### 2021-01-07 install.Rmd
Initial commit

---
title: "Installation of Required Software"
author: "Aaron Peikert"
date: "05/01/2021"
output:
  github_document:
    number_sections: true
repro:
  packages:
    - here
  scripts:
    - R/link.R
---

```{r setup, include=FALSE}
source(here::here("R", "link.R"))
```

# Why you want to use package managers

Installing software can be a painful experience, especially if you have to sift through the documentation for multiple platforms.
A package manager helps you to install, update and remove software on your computer, automating large parts of the process for you.
Most importantly, the installation process is the same for every software, not unlike the appstore on your phone.
Unfortunately though, it is not as pleasantly designed, but is used through the terminal.
If the thought of using a terminal lets you hesitate, rest assured it is easier than you might think.

# Windows 10

## How to open a terminal on Windows 10?

1. Press Windows 10 key + X.
2. Click on "Windows Powershell" or "Windows Powershell (Admin)" (if you are using Chocolately)

## How to install a package manager for Windows 10?

`r link("https://chocolatey.org/", "Chocolately")` is a great package manager for Windows 10.
You should follow these installation instructions: `r link("https://chocolatey.org/docs/installation")`, but if you are in a hurry:

1. Press Windows key + X, click on "Windows Powershell (Admin)".
2. Paste: `Set-ExecutionPolicy Bypass -Scope Process -Force; [System.Net.ServicePointManager]::SecurityProtocol = [System.Net.ServicePointManager]::SecurityProtocol -bor 3072; iex ((New-Object System.Net.WebClient).DownloadString('https://chocolatey.org/install.ps1'))`
3. Press enter.
4. Close the PowerShell

## How to install R + Rstudio with a package manager?

Since you probably already have R and RStudio you may skip this step.

1. Open a new terminal (Press Windows key + X, click on "Windows Powershell (Admin)").
2. Paste `choco install -y r r.studio`
3. Press enter

## How to install Git, Make, and Docker with a package manager?

Please note that Docker does not work on Windows 10 Home, you need Windows Pro or Education.

1. Open a new terminal (Press Windows key + X, click on "Windows Powershell (Admin)").
2. Paste `choco install -y git make docker`
3. Press enter.

# Mac OS

## How to open a terminal on Mac OS?

1. Press Command + Space.
2. Type "Terminal".
3. Press enter.

## How to install a package manager for Mac OS?

`r link("https://brew.sh", "Homebrew")` is a great package manager for Mac OS. To install it:

1. Open a new terminal (Press Command + Space, type "Terminal", press enter).
2. Paste `xcode-select --install` and press enter to install xcode.
3. Paste: `/bin/bash -c "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/HEAD/install.sh)"`.
4. Press enter.
5. Close the terminal.

## How to install R + Rstudio with a package manager?

Since you probably already have R and RStudio you may skip this step.

1. Open a new terminal (Press Command + Space, type "Terminal", press enter).
2. Paste `brew install r` and press enter to install R.
3. Paste `brew install --cask rstudio` and press enter to install RStudio.

## How to install Git, Make, and Docker with a package manager?

1. Open a new terminal (Press Command + Space, type "Terminal", press enter).
2. Paste `brew install git make docker`.
3. Press enter.

```{r link_index, results='asis', echo=FALSE}
link_index()
```

### 2021-01-07 R/link.R

@@ -1,0 +1,18 @@
+link <- function(link, linktext = link, footnote = link != linktext, file = here::here("links.txt")){
+  markdown <- paste0("[", linktext, "](", link, ")")
+  if(footnote){
+    ref <- gsub("[^a-zA-Z]*", "", linktext)
+    markdown <- paste0(markdown, "[^", ref, "]", collapse = "")
+    cat(paste0("[^", ref, "]: ", link, "\n"), file = file, append = TRUE)
+  }
+  return(markdown)
+}
+
+link_index <- function(file = here::here("links.txt")){
+  if(file.exists("links.txt")){
+    links <- readLines(file)
+    cat(links, sep = "\n")
+    file.remove(file)
+    invisible()
+  } else return(invisible())
+}

### 2021-01-07 repro_log.R

@@ -1,0 +1,44 @@
+# I try to log what I am doing
+if(!requireNamespace("remotes"))install.packages("remotes")
+remotes::install_github("aaronpeikert/repro")
+repro::use_repro_template()
+# ISSUE: https://github.com/aaronpeikert/repro/issues/53
+# manually add install.Rmd, delete examples
+# configure Docker and Make using repro:
+repro::check_make()
+repro::check_docker()
+repro::automate()
+# remind git add correct files
+
+# add git:
+repro::check_git()
+usethis::use_git()
+
+# add a licence:
+usethis::use_ccby_license()
+git2r::add("." ,"LICENSE.md")
+git2r::commit(".", "usethis::use_ccby_license()")
+
+# add a readme:
+usethis::use_readme_rmd()
+# manually edit readme
+# add readme to makefile:
+repro::automate()
+# ISSUE: https://github.com/aaronpeikert/repro/issues/46
+system("make docker")
+system("make -B DOCKER=TRUE")
+# ISSUE: https://github.com/aaronpeikert/repro/issues/54
+# changed R version to 4.0.3 in Dockerfile
+
+# add a code of conduct
+usethis::use_code_of_conduct()
+git2r::add("." ,"README.Rmd")
+git2r::add("." ,"CODE_OF_CONDUCT.md")
+system("make DOCKER=TRUE")
+git2r::add("." ,"README.md")
+git2r::commit(".", "usethis::use_code_of_conduct()")
+
+# use github
+repro::check_github()
+# ISSUE: https://github.com/aaronpeikert/repro/issues/55
+usethis::use_github()

### 2021-01-07 README.Rmd
add README & build documents

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# A quick tour through the `repro`-package

WIP

### 2021-01-07 repro_log.R
update repro_log.R

@@ -14,11 +14,6 @@ repro::automate()
 repro::check_git()
 usethis::use_git()
 
-# add a licence:
-usethis::use_ccby_license()
-git2r::add("." ,"LICENSE.md")
-git2r::commit(".", "usethis::use_ccby_license()")
-
 # add a readme:
 usethis::use_readme_rmd()
 # manually edit readme
@@ -29,9 +24,23 @@ system("make docker")
 system("make -B DOCKER=TRUE")
 # ISSUE: https://github.com/aaronpeikert/repro/issues/54
 # changed R version to 4.0.3 in Dockerfile
+git2r::add(".", c("Dockerfile", "Makefile", "README.Rmd", "README.md", "install.md", ".repro/Dockerfile_base", ".repro/Makefile_Rmds"))
+git2r::commit(".", "add README & build documents")
 
-# add a code of conduct
+# ignore htmls:
+usethis::use_git_ignore("*.html")
+git2r::add(".", c(".gitignore"))
+git2r::commit(".", "ignore htmls")
+
+
+# add a licence:
+usethis::use_ccby_license()
+git2r::add("." ,"LICENSE.md")
+git2r::commit(".", "usethis::use_ccby_license()")
+
+# add a code of conduct:
 usethis::use_code_of_conduct()
+# manually edit README.Rmd
 git2r::add("." ,"README.Rmd")
 git2r::add("." ,"CODE_OF_CONDUCT.md")
 system("make DOCKER=TRUE")
@@ -42,3 +51,13 @@ git2r::commit(".", "usethis::use_code_of_conduct()")
 repro::check_github()
 # ISSUE: https://github.com/aaronpeikert/repro/issues/55
 usethis::use_github()
+
+# add autogenerated not so important files:
+git2r::add("." , c(".Rbuildignore", "repro-tutorial.Rproj"))
+git2r::commit(".", "add autogenerated not so important files")
+git2r::push()
+
+# update repro_log.R:
+git2r::add("." , c("repro_log.R"))
+git2r::commit(".", "update repro_log.R")
+git2r::push()

### 2021-01-07 repro_log.R
remove reminder

@@ -8,7 +8,6 @@ repro::use_repro_template()
 repro::check_make()
 repro::check_docker()
 repro::automate()
-# remind git add correct files
 
 # add git:
 repro::check_git()
@@ -47,7 +46,7 @@ system("make DOCKER=TRUE")
 git2r::add("." ,"README.md")
 git2r::commit(".", "usethis::use_code_of_conduct()")
 
-# use github
+# use github:
 repro::check_github()
 # ISSUE: https://github.com/aaronpeikert/repro/issues/55
 usethis::use_github()

### 2021-01-12 README.Rmd
usethis::use_code_of_conduct()
add 'concept'-badge

@@
 # A quick tour through the `repro`-package
+<!-- badges: start -->
+[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
+<!-- badges: end -->
 
 WIP
+
+## Code of Conduct
+
+Please note that the repro-tutorial project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
+By contributing to this project, you agree to abide by its terms.

### 2021-01-12 repro_log.R
use seperate branch for planning

@@ -60,3 +60,8 @@ git2r::push()
 git2r::add("." , c("repro_log.R"))
 git2r::commit(".", "update repro_log.R")
 git2r::push()
+
+# switch to a new branch to plan the content:
+git2r::branch_create(git2r::commits()[[1]] ,"plan")
+git2r::checkout(branch = "plan")
+file.create("manuscript.Rmd")

### 2021-01-12 manuscript.Rmd
add draft outline of manuscript

---
title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
output: pdf_document
repro:
  packages:
    - usethis
---

<!-- the HTML comments, like this one, are meta comments, mainly describing the intent --->
<!-- each sentence below a heading summarizes what I want to say there --->
<!-- "Hands-on:" means concrete practical application, they roughly proceed from easy/familiar to hard/unfamiliar-->

# Why you should care about reproducibility

<!-- define reproducibility -->

## Productivity

<!-- lure the reader in -->

Adherence to best practices promises to increase productivity for individual researchers and the community as a whole.

## Metascientific Arguments

<!-- present reproducibility as something that fits right into the researchers' objectives-->

Reproducibility is caused by openness, enables replication, and all three together warrant trust in science.

# The lifecycle of an open research project

<!-- many small steps archive reproducibility -->

## Setup

<!-- keep it brief, let `repro` do the work -->

We assume that you have already installed R and RStudio (if not check the installation guide).
Additionally, you'll need the [`repro`-package](https://github.com/aaronpeikert/repro):

```r
if(!requireNamespace("remotes"))install.packages("remotes")
remotes::install_github("aaronpeikert/repro")
```

## Planing

<!-- researcher should begin early to enjoy most benefits -->

Organize files in one place.

<!-- Hands-on: Introducing RProjects-->

Keep notes.

<!-- Hands on: Introduce Markdown -->

Record changes.

<!-- Hands-on: Introduce Git -->

<!-- This is maybe the place to introduce the example-->
<!-- something roughly as complex as a bachelor thesis, like:

Hypothesis: Machivallianism is higher in male persons
Sample: Adults from western countries
Analytic strategy: Multigroup CFA + Partial Measurement Invariance
Data from: https://openpsychometrics.org/_rawdata/ either SD3 or MACH-IV
-->

```{r}
library("repro")
check_git()
```

### Simulation (optional)

<!-- impress with neatness → sample size planning, preregistration and analysis in one 😎-->
<!-- but explain that good science is often not as neat, strive for the ideal -->

Simulating data helps prevent unpleasant surprises and lets you double-check that the data you expect to gather and your analysis fit together.

<!--Hands-on: build simple functions tailored for your analysis-->

### Preregestration (optional)

A preregistration is a tool that may help to increase the credibility of empirical results.

<!--Hands on: Introduce RMarkdown-->
<!--Hands on: gitignore resulting .html/pdf-->
<!-- A preregistration needs to be public -->

## Prepare public release (optional)

<!--Hands on: Add a readme-->
```r
usethis::use_readme_rmd()
```

<!--Hands on: Add a license-->
```r
usethis::use_cc0_license()
```

<!--Hands on: Add a code of conduct-->
```r
usethis::use_code_of_conduct()
```

## Collaboration

### GitHub

GitHub/GitLab are great as a shared information basis: sharing and editing files, discussing and distributing tasks via issues and documenting the project via wikis even before data collection starts.

<!--Hands-on: Introduce Github-->

```{r}
check_github()
```

```r
usethis::use_github()
```

<!--Hands on: Introduce GitHub Issues-->
<!--Hands on: Introduce GitHub Projects-->

### Internal Dependencies

A well defined "entry point" that automatically reproduces everything eases collaboration.

<!--I am not sure this is the right place, should we introduce this earlier? -->

<!--Hands on: Introduce Make-->
<!--Hands on: Add targets for README.Rmd, preregistration.Rmd -->

``` r
automate_make()
```

### External Dependencies

A considerable number of external dependencies may hinder collaboration and long term reproducibility, but automated solutions can "bundle" and fix dependencies.

<!-- Hands on: Introduce Docker -->

``` r
automate_docker()
```

<-- after this everything is optional? -->

#### renv (alternative)

#### Singularity (optional)

## Analyzing Data

<!-- Hands on: add data via automate() -->
<!-- Hands on: automatically download data via manually editing the Makefile -->

### Data Integrety (optional)

<!-- Hands-on: check changed hash -->

### Anonymising Data (optional)

<!-- Should we add a section on synthetic data generation? -->
<!-- Hands on: Create synthetic data -->

## Releasing a Preprint (optional)

<!-- Hands on: add commit hash to results-->
<!-- Hands on: git tag + github release-->

## Peer Review (optional)

## Post Publication Review (optional)

## Archiving

<!-- Hands on: explain how to archive data-->
<!-- Hands on: explain how to archive docker images-->

### 2021-01-12 repro_log.R
housekeeping

@@ -65,3 +65,5 @@ git2r::push()
 git2r::branch_create(git2r::commits()[[1]] ,"plan")
 git2r::checkout(branch = "plan")
 file.create("manuscript.Rmd")
+
+usethis::use_git_ignore("*.pdf")

### 2021-01-13 manuscript.Rmd
load repro earlier

@@
     - usethis
+    - repro
 ---
@@
 remotes::install_github("aaronpeikert/repro")
+library("repro")
 ```
@@
 ```{r}
-library("repro")
 check_git()

### 2021-01-22 manuscript.Rmd
should we add a checklist @brandmaier?
add introduction

@@
     - repro
+    - here
 ---
 
-<!-- the HTML comments, like this one, are meta comments, mainly describing the intent --->
+<<!-- the HTML comments, like this one, are meta comments, mainly describing the intent --->
 <!-- each sentence below a heading summarizes what I want to say there --->
 <!-- "Hands-on:" means concrete practical application, they roughly proceed from easy/familiar to hard/unfamiliar-->
+
+```{r setup, include=FALSE}
+library(repro)
+source(here::here("R", "link.R"))
+```
 
@@
 <!-- define reproducibility -->
+<!-- is a direct quote from my master thesis -->
+<!-- I wouldn't want to cite it. @brandmaier, would you please rewrite? -->
+A reproducible scientific product allows other researchers to obtain the same results from the same dataset in a way that enables substantive criticism and therefore facilitates replication.
+<!-- lure the reader in -->
+This notion of independent verifiability has long been part of "good scientific practice", but has not reached its full potential to benefit the scientific community.
+<!-- expand: who is calling it a requirement, maybe metascientific arguments-->
+Reproducing an empirical study is often tedious and frustrating work for the researcher.
+However, if reproduction would require mere minutes instead of hours, it could greatly facilitate collaboration within and across research projects.
+<!-- Guiding principle: spending machine compute time instead of human research time -->
+To archive this boost, we must place the burden of reproducing something upon computers instead of human researchers.
+A scientific document should therefore be understood by researchers but reproduced by computers.
 
-## Productivity
+This tutorial walks through the creation of such a document that can be reproduced automatically.
+Though reproducibility is but a small part of the research process, it can serve as the mortar of open science building blocks (e.g. preregistration, open data, postpublication review).
+Therefore, we show the whole lifecycle of an open research project, where some parts are essential to archive reproducibility.
+Still, the reader can safely skip sections marked as optional, when unfit for their research question or methods.
+However, any empirical data analysis will suffer from the following threats to reproducibility:
+<!-- Which problems are solved by which tool (like in Andreas talk)-->
+<!--stolen from: https://brandmaier.github.io/reproducible-data-analysis-materials/KULeuvenQuantPsy2020.html#11 -->
 
-<!-- lure the reader in -->
+1. Multiple versions of scripts/data (e.g., dataset has changed over time, i.e., was further cleaned or extended)
+2. Copy&paste errors (e.g., inconsistency between reported result and reproduced result)
+3. unclear which scripts should be executed in which order
+4. Broken software dependencies (e.g., analysis broken after update, missing package)
 
-Adherence to best practices promises to increase productivity for individual researchers and the community as a whole.
+It is common to rely on the craftiness of the researcher to debug these problems, but we could prevent them altogether with:
 
-## Metascientific Arguments
+1. Version control
+2. Dynamic document creation
+3. Dependency tracking
+4. Software management
 
-<!-- present reproducibility as something that fits right into the researchers' objectives-->
+We deem each concept necessary to archive longterm reproducibility.
+However, the tools you use to archive them are a mere matter of taste and project requirements.
+This tutorial will follow the recommendations of @peikertMissing and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking and Docker for software management.
+Their interplay is shown in Figure \@ref(fig:schematic).
+At several points we suggest other viable alternatives.
 
-Reproducibility is caused by openness, enables replication, and all three together warrant trust in science.
+```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (git), dependency management (Make), containerization (Docker), and dynamic document generation (R Markdown).
+Git tracks changes to the project over time.
+Make manages dependencies among the files.
+Docker provides a container, in which the final report is built using dynamic document generation in R Markdown.
+Reproduced from @peikertMissing  "}
+# file gets downloaded in Makefile
+knitr::include_graphics("images/nutshell.svg", auto_pdf = TRUE)
+```
 
@@
 
-### Data Integrety (optional)
+### Data Integrity (optional)
 
@@
 <!-- Hands on: explain how to archive docker images-->
+
+## Checklist
+
+<!-- Should we design a checklist? -->`

### 2021-01-22 manuscript.Rmd
add minimal setup

@@
 
-# The lifecycle of an open research project
-
-<!-- many small steps archive reproducibility -->
-
-## Setup
+# Setup
 
@@
 
-We assume that you have already installed R and RStudio (if not check the installation guide).
-Additionally, you'll need the [`repro`-package](https://github.com/aaronpeikert/repro):
+We assume that you have already installed R and RStudio (if not check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/install.md", "installation guide")`).
+Additionally, you'll need the [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:
+
+[^repropackage]: https://github.com/aaronpeikert/repro
 
@@
 ```
+
+The [`repro`-package](https://github.com/aaronpeikert/repro) contains several helpers to work with reproducibility tools.
+For example, it can check if you computer is set up to use the tools we rely on:
+
+```{r}
+check_git()
+```
+
+```{r}
+check_make()
+```
+
+```{r}
+check_docker()
+```
+
+If these commands detect that something is not installed or set up, they will give you detailed instructions (tailored to your operating system) on how to remedy the situation.
+Follow the instructions, then rerun the command, till it tells you to not worry.
 
@@
 
-```{r}
-check_github()
-```
-
 ```r

### 2021-01-22 manuscript.Rmd
add introduction to git & markdown in planning

@@
 <!-- researcher should begin early to enjoy most benefits -->
+It can be tempting to leave considerations of reproducibility till the very last when you finish a project.
+<!-- weird sentence structure, I'll leave it as a placeholder-->
+However, reproducibility is invaluable for collaboration---including with the version of yourself from a few weeks ago.
+We therefore suggest to start early, ideally from the moment you decide to pursue a project to gradually incorporate tools to help you archive reproducibility.
+The first step is to organize all files in one folder.
+This way you never loose anything and you can easily share the files.
+<!-- Hands-on: Introducing RProjects-->
+This concept of a project folder is facilitated by RStudio's RProjects.
+To create one, click on:
 
-Organize files in one place.
+> File → New Project... → New Directory → New Project
 
-<!-- Hands-on: Introducing RProjects-->
+<!--Keep notes.-->
+<!-- Hands on: Introduce Markdown -->
+A great tool to capture ideas or meeting notes is Markdown.
+To create a Markdown document click on:
 
-Keep notes.
+> File → New File → Markdown File
 
-<!-- Hands on: Introduce Markdown -->
+Markdown is simply text, where some characters let you add a minimal amount of structure:
 
-Record changes.
+```{r, results='asis', echo=FALSE}
+# in case we want to reuse this example later
+markdown_example <- "
+# Header
+Normal text.
+Important **word**.
 
+To do list:
+
+* do research
+* do more research
+* spend time with family
+"
+cat("``` markdown", markdown_example, "```", sep = "")
+```
+
+We will later see how this simplicity enables us to create many different document formats (Word, PDF, HTML) and incorporate code (R, Python, Julia).
+Another advantage of Markdown is, that you can easily version control it with Git.
+You may be familiar with Microsoft Words' "Track Changes"-Feature and version control is similar in its basic idea, but more powerful in its features.
+Instead of tracking only one file, you can keep track of the whole project directory and instead of loosing all changes when you agree to them, the whole history is preserved transparently.
+For example this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/master", "iterations")`.
+To activate Git in a given project you can call:
 <!-- Hands-on: Introduce Git -->
 
+```r
+# package::function → use function from package without library(package)
+usethis::use_git()
+```
+
+Recording each change you made, helps you to iterate more quickly because you know that you can go back to previous versions easily.
+For example you can write down a rough project idea, knowing your collaborators and you will iterate and improve.
 <!-- This is maybe the place to introduce the example-->
-<!-- something roughly as complex as a bachelor thesis, like:
+First you could create a new file named `idea.md` that looks like this:
 
-Hypothesis: Machivallianism is higher in male persons
-Sample: Adults from western countries
-Analytic strategy: Multigroup CFA + Partial Measurement Invariance
-Data from: https://openpsychometrics.org/_rawdata/ either SD3 or MACH-IV
--->
+```markdown
+# Hypothesis
 
-```{r}
-check_git()
+Machivallianism is higher in male persons
+
+# Analytic Strategy
+
+t-test
+
+# Sample
+
+We could use openpsychometrics.org data?
 ```
+
+And you could commit it in Git:
+
+```r
+git2r::add(".", "idea.md")
+git2r::commit(".", "add a first concept")
+```
+
+But then you realize that the new literature suggests that there could be a bias in measurement:
+
+``` markdown
+# Hypothesis
+
+Machivallianism is higher in male persons
+
+# Analytic Strategy
+
+Multigroup CFA + Measurement Invariance
+
+# Sample
+
+We could use openpsychometrics.org data?
+```
+
+Commit this version:
+
+```r
+git2r::add(".", "idea.md")
+git2r::commit(".", "exclude bias in measurement as possible confounder")
+```
+
+And Git can tell you anytime what has changed.
+
+<!--insert screenshot-->
 

### 2021-01-22 manuscript.Rmd
add autolinked footnotes
remove typo

@@
 
-<<!-- the HTML comments, like this one, are meta comments, mainly describing the intent --->
+<!-- the HTML comments, like this one, are meta comments, mainly describing the intent --->
 <!-- each sentence below a heading summarizes what I want to say there --->
@@
 <!-- Should we design a checklist? -->`
+
+```{r, results='asis', echo=FALSE}
+link_index()
+```

### 2021-01-22 manuscript.Rmd
check with grammerly

@@
 
-1. Multiple versions of scripts/data (e.g., dataset has changed over time, i.e., was further cleaned or extended)
+1. Multiple versions of scripts/data (e.g., the dataset has changed over time, i.e., was further cleaned or extended)
 2. Copy&paste errors (e.g., inconsistency between reported result and reproduced result)
 3. unclear which scripts should be executed in which order
-4. Broken software dependencies (e.g., analysis broken after update, missing package)
+4. Broken software dependencies (e.g., analysis broken after an update, missing package)
 
@@
 Their interplay is shown in Figure \@ref(fig:schematic).
-At several points we suggest other viable alternatives.
+At several points, we suggest other viable alternatives.
 
-```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (git), dependency management (Make), containerization (Docker), and dynamic document generation (R Markdown).
+```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency management (Make), containerization (Docker), and dynamic document generation (R Markdown).
 Git tracks changes to the project over time.
 Make manages dependencies among the files.
-Docker provides a container, in which the final report is built using dynamic document generation in R Markdown.
+Docker provides a container in which the final report is built using dynamic document generation in R Markdown.
 Reproduced from @peikertMissing  "}
@@
 The [`repro`-package](https://github.com/aaronpeikert/repro) contains several helpers to work with reproducibility tools.
-For example, it can check if you computer is set up to use the tools we rely on:
+For example, it can check if your computer is set up to use the tools we rely on:
 
@@
 
-If these commands detect that something is not installed or set up, they will give you detailed instructions (tailored to your operating system) on how to remedy the situation.
-Follow the instructions, then rerun the command, till it tells you to not worry.
+If these commands detect that something is not installed or set up, they will give you detailed instructions (tailored to your operating system) to remedy the situation.
+Follow the instructions, then rerun the command, until it tells you not to worry.
 
@@
 However, reproducibility is invaluable for collaboration---including with the version of yourself from a few weeks ago.
-We therefore suggest to start early, ideally from the moment you decide to pursue a project to gradually incorporate tools to help you archive reproducibility.
+Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you archive reproducibility gradually.
 The first step is to organize all files in one folder.
-This way you never loose anything and you can easily share the files.
+This way, you never lose anything, and you can easily share the files.
 <!-- Hands-on: Introducing RProjects-->
-This concept of a project folder is facilitated by RStudio's RProjects.
+RStudio's RProjects facilitate this concept of a project folder.
 To create one, click on:
@@
 We will later see how this simplicity enables us to create many different document formats (Word, PDF, HTML) and incorporate code (R, Python, Julia).
-Another advantage of Markdown is, that you can easily version control it with Git.
-You may be familiar with Microsoft Words' "Track Changes"-Feature and version control is similar in its basic idea, but more powerful in its features.
-Instead of tracking only one file, you can keep track of the whole project directory and instead of loosing all changes when you agree to them, the whole history is preserved transparently.
+Another advantage of Markdown is that you can easily version-control it with Git.
+You may be familiar with Microsoft Words' "Track Changes"-Feature and version control is similar in its basic idea but more potent in its features.
+Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
 For example this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/master", "iterations")`.
-To activate Git in a given project you can call:
+To activate Git in a given project, you can call:
 <!-- Hands-on: Introduce Git -->
@@
 
-Recording each change you made, helps you to iterate more quickly because you know that you can go back to previous versions easily.
-For example you can write down a rough project idea, knowing your collaborators and you will iterate and improve.
+Recording each change you made helps you to iterate more quickly because you know that you can effortlessly go back to previous versions.
+For example, you can write down a rough project idea, knowing your collaborators and you will iterate and improve.
 <!-- This is maybe the place to introduce the example-->
-First you could create a new file named `idea.md` that looks like this:
+First, you could create a new markdown file named `idea.md` that looks like this:
 
@@
 
-Machivallianism is higher in male persons
+Machiavellianism is higher in male persons.
 
@@
 
-We could use openpsychometrics.org data?
+Could we use openpsychometrics.org data?
 ```
@@
 And you could commit it in Git:
+<!--I am leaning towards doing everything in R, not sure though-->
 
@@
 
-<-- after this everything is optional? -->
+<!-- after this everything is optional? -->
 

### 2021-01-22 manuscript.Rmd
add references

@@
     - here
+csl: apa7.csl
+bibliography: references.bib
 ---
@@
 However, the tools you use to archive them are a mere matter of taste and project requirements.
-This tutorial will follow the recommendations of @peikertMissing and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking and Docker for software management.
+This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking and Docker for software management.
 Their interplay is shown in Figure \@ref(fig:schematic).
@@
 Docker provides a container in which the final report is built using dynamic document generation in R Markdown.
-Reproduced from @peikertMissing  "}
+Reproduced from @Peikert2019.
+"}
 # file gets downloaded in Makefile
@@
 
-## Planing
+# Planing
 
@@
 ```r
-# package::function → use function from package without library(package)
+# `package::function()` → use function from package without `library(package)`
 usethis::use_git()
@@
 <!--I am leaning towards doing everything in R, not sure though-->
+<!--For Git we maybe should use RStudio's Git pane-->
 
@@
 
+We recommend to attend an online lecture or workshop to learn the ins and outs of Git.
+
 ### Simulation (optional)
@@
 
-Simulating data helps prevent unpleasant surprises and lets you double-check that the data you expect to gather and your analysis fit together.
+There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
+A step to prevent unpleasant surprises is to simulate data.
+Simulated data lets you verify that the data you expect to gather and the analysis that you plan fit together.
 
@@
 
-<!-- Should we design a checklist? -->`
+<!-- Should we design a checklist? -->
 
@@
 ```
+
+# References

### 2021-01-25 manuscript.Rmd
switch to bookdown, so that figure references work

@@
 title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
-output: pdf_document
+output: bookdown::pdf_document2
 repro:

### 2021-01-25 manuscript.Rmd
revise some phrasing

@@
 We deem each concept necessary to archive longterm reproducibility.
-However, the tools you use to archive them are a mere matter of taste and project requirements.
+However, which tools you use to implement this concepts are mere matter of taste and project requirements.
 This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking and Docker for software management.
@@
 <!-- researcher should begin early to enjoy most benefits -->
-It can be tempting to leave considerations of reproducibility till the very last when you finish a project.
+It can be tempting to leave considerations of reproducibility for when you finish a project.
 <!-- weird sentence structure, I'll leave it as a placeholder-->
@@
 
-We will later see how this simplicity enables us to create many different document formats (Word, PDF, HTML) and incorporate code (R, Python, Julia).
+We will later see how this simplicity enables us to create many different document formats (i.e. Word, PDF, HTML) and incorporate code (i.e. R, Python, Julia).
 Another advantage of Markdown is that you can easily version-control it with Git.

### 2021-01-25 manuscript.Rmd
add screenshot of git change

@@
 
-And Git can tell you anytime what has changed.
+And Git can show you what has changed, like in Figure \@ref(fig:idea-change).
 
-<!--insert screenshot-->
-
+```{r idea-change, echo=FALSE, fig.cap="A screenshot of how a change tracked in Git is represented by GitHub.  "}
+# unfortunately css selectors won't work for pdf so we'll use png instead
+idea_change <- here::here("images", "idea-change.png")
+webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/5b3f4641ff542158184f85d458880c35e8f09f3c?diff=split",
+                  file = idea_change,
+                  selector = "div#diff-2f8beaa39e5c98706d8abb3361a8383776f8df8a195a21c0e6eeb25e5aa48f45", # only select file change
+                  zoom = 2)# higher resolution
+```
+<!--I am troubled by how we should recomend to learn git-->
 We recommend to attend an online lecture or workshop to learn the ins and outs of Git.

### 2021-01-26 manuscript.Rmd
divide taking and including screenshot

--- 2021-01-25 manuscript.Rmd
@@
 
-And Git can tell you anytime what has changed.
+And Git can show you what has changed, like in Figure \@ref(fig:idea-change).
 
-<!--insert screenshot-->
+```{r idea-change-screenshot, include=FALSE}
+# unfortunately css selectors won't work for pdf so we'll use png instead
+idea_change <- here::here("images", "idea-change.png")
+webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/5b3f4641ff542158184f85d458880c35e8f09f3c?diff=split",
+                  file = idea_change,
+                  selector = "div#diff-2f8beaa39e5c98706d8abb3361a8383776f8df8a195a21c0e6eeb25e5aa48f45", # only select file change
+                  zoom = 2)# higher resolution
+```
 
+```{r idea-change, echo=FALSE, fig.cap="A screenshot of how a change tracked in Git is represented by GitHub.  ", out.width='100%'}
+knitr::include_graphics(idea_change)
+```
+<!--I am troubled by how we should recomend to learn git-->
 We recommend to attend an online lecture or workshop to learn the ins and outs of Git.

### 2021-01-26 manuscript.Rmd
start working on simulation

@@
 We deem each concept necessary to archive longterm reproducibility.
-However, which tools you use to implement this concepts are mere matter of taste and project requirements.
-This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking and Docker for software management.
+However, which tools you use to implement these concepts is a matter of taste and project requirements.
+This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
 Their interplay is shown in Figure \@ref(fig:schematic).
@@
 
-```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency management (Make), containerization (Docker), and dynamic document generation (R Markdown).
+```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency management (Make), software management (Docker), and dynamic document generation (R Markdown).
 Git tracks changes to the project over time.
@@
 ```r
+# -- type this on the R console --
 if(!requireNamespace("remotes"))install.packages("remotes")
@@
 ```{r}
+# -- type this on the R console --
 check_git()
@@
 ```{r}
+# -- type this on the R console --
 check_make()
@@
 ```{r}
+# -- type this on the R console --
 check_docker()
@@
 It can be tempting to leave considerations of reproducibility for when you finish a project.
-<!-- weird sentence structure, I'll leave it as a placeholder-->
+<!-- weird sentence, I'll leave it as a placeholder-->
 However, reproducibility is invaluable for collaboration---including with the version of yourself from a few weeks ago.
@@
 markdown_example <- "
+<!--this is a Markdown file -->
 # Header
@@
 ```markdown
+<!--this is a Markdown file -->
+
 # Hypothesis
@@
 ``` markdown
+<!--this is a Markdown file -->
+
 # Hypothesis
@@
 ```r
+# -- type this on the R console --
 git2r::add(".", "idea.md")
@@
 ```
+
 <!--I am troubled by how we should recomend to learn git-->
@@
 
+```{r, include=FALSE}
+source(here::here("R", "simulation.R"))
+```
+
 There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
@@
 Simulated data lets you verify that the data you expect to gather and the analysis that you plan fit together.
+Of course one has to make a lot of educated guesses.
+Generally there are of course hundreds of different equal possibilities, but we just care about the worst reasonable settings.
+Sticking with our example about machiavilism, we simulate multivariate normal data representing the machiavilism subscale of the short dark triad (SD3).
+As a starting point we can use already publishes information.
+@Jones2013 report loadings seen in Table \@ref(tab:loadings) and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
+
+```{r loadings, echo=FALSE}
+knitr::kable(data.frame(Item = 1:9,
+                        Loading = loadings_jones_paulhus),
+             caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
+```
+
+It is probably better to assume values that are more pessimistic.
+Instead of taking these values as they are we will therefore assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.1.
+These assumption can be simply translated into `lavaan` model syntax:
+
+```{r, results='asis', echo=FALSE}
+model_truth <- paste(measurement(
+  "MACH",
+  paste0("x", 1:9),
+  loadings = round(loadings_jones_paulhus * 0.7, 2)
+),
+intercepts("MACH", list(0, 0.1)),
+collapse = "\n", sep = "\n")
+cat("``` r",
+    model_truth,
+    "```",
+    sep = "\n")
+model <- paste(measurement(
+  "MACH",
+  paste0("x", 1:9)),
+  intercepts("MACH", list(0, "NA")),
+  sep = "\n")
+```
+
+From which we can simulate data and fit a model to it:
+
+```{r, results='hide'}
+data <- lavaan::simulateData(model_truth, "cfa", sample.nobs = c(20000, 20000), meanstructure = TRUE, standardized = TRUE)
+fit <- lavaan::cfa(model, data = data, group = "group", meanstructure = TRUE, std.lv = TRUE, group.equal = c("loadings", "intercepts")) 
+lavaan::summary(fit)
+```
 

### 2021-01-26 R/lavaan_helper.R

@@ -1,0 +1,27 @@
+intercepts <- function(variables, intercepts){
+  stopifnot(is.character(variables))
+  stringr::str_c(variables, " ~ ", modifier(intercepts), "1")
+}
+
+modifier <- function(values){
+  if (missing(values)) {
+    values <- ""
+  } else if (is.list(values)) {
+    values <- do.call(stringr::str_c, c(values, sep = ", "))
+    values <- stringr::str_c("c(", values, ")*")
+  } else {
+    values <- stringr::str_c(values, "*")
+  }
+  return(values)
+}
+
+measurement <- function(latent, variables, loadings){
+  stopifnot(is.character(latent),
+            is.character(variables),
+            length(latent) == 1)
+  loadings <- modifier(loadings)
+  variables <- stringr::str_c(loadings, variables, collapse = " + ")
+  stringr::str_c(latent, " =~ ", variables)
+}
+
+

### 2021-01-26 R/simulation.R

@@ -1,0 +1,15 @@
+source(here::here("R", "lavaan_helper.R"))
+generate <- function(model, n){
+  
+}
+
+analyze <- function(model, data){
+  
+}
+
+summarise <- function(fit){
+  
+}
+
+loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
+cohend_jones_paulhus <- c(24, 29, 35)/100

### 2021-01-26 manuscript.Rmd
cite lavaan

@@
 csl: apa7.csl
-bibliography: references.bib
+bibliography: ["references.bib", "packages.bib"]
 ---
@@
 Instead of taking these values as they are we will therefore assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.1.
-These assumption can be simply translated into `lavaan` model syntax:
+These assumption can be simply translated into [`lavaan` [@lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
+
+[^lavaan]: https://lavaan.ugent.be
 

### 2021-01-26 manuscript.Rmd
don't screenshot all the time

@@
 idea_change <- here::here("images", "idea-change.png")
+if(!file.exists(idea_change)){
   webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/5b3f4641ff542158184f85d458880c35e8f09f3c?diff=split",
@@
                   zoom = 2)# higher resolution
+}
 ```

### 2021-01-26 manuscript.Rmd
anontate code chunks with programming language

--- 2021-01-26 manuscript.Rmd
@@
 We deem each concept necessary to archive longterm reproducibility.
-However, which tools you use to implement these concepts is a matter of taste and project requirements.
-This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
+However, which tools you use to implement this concepts are mere matter of taste and project requirements.
+This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking and Docker for software management.
 Their interplay is shown in Figure \@ref(fig:schematic).
@@
 
-```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency management (Make), software management (Docker), and dynamic document generation (R Markdown).
+```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency management (Make), containerization (Docker), and dynamic document generation (R Markdown).
 Git tracks changes to the project over time.
@@
 It can be tempting to leave considerations of reproducibility for when you finish a project.
-<!-- weird sentence, I'll leave it as a placeholder-->
+<!-- weird sentence structure, I'll leave it as a placeholder-->
 However, reproducibility is invaluable for collaboration---including with the version of yourself from a few weeks ago.
@@
 ```
-
 <!--I am troubled by how we should recomend to learn git-->
@@
 
-```{r, include=FALSE}
-source(here::here("R", "simulation.R"))
-```
-
 There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
@@
 Simulated data lets you verify that the data you expect to gather and the analysis that you plan fit together.
-Of course one has to make a lot of educated guesses.
-Generally there are of course hundreds of different equal possibilities, but we just care about the worst reasonable settings.
-Sticking with our example about machiavilism, we simulate multivariate normal data representing the machiavilism subscale of the short dark triad (SD3).
-As a starting point we can use already publishes information.
-@Jones2013 report loadings seen in Table \@ref(tab:loadings) and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
-
-```{r loadings, echo=FALSE}
-knitr::kable(data.frame(Item = 1:9,
-                        Loading = loadings_jones_paulhus),
-             caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
-```
-
-It is probably better to assume values that are more pessimistic.
-Instead of taking these values as they are we will therefore assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.1.
-These assumption can be simply translated into `lavaan` model syntax:
-
-```{r, results='asis', echo=FALSE}
-model_truth <- paste(measurement(
-  "MACH",
-  paste0("x", 1:9),
-  loadings = round(loadings_jones_paulhus * 0.7, 2)
-),
-intercepts("MACH", list(0, 0.1)),
-collapse = "\n", sep = "\n")
-cat("``` r",
-    model_truth,
-    "```",
-    sep = "\n")
-model <- paste(measurement(
-  "MACH",
-  paste0("x", 1:9)),
-  intercepts("MACH", list(0, "NA")),
-  sep = "\n")
-```
-
-From which we can simulate data and fit a model to it:
-
-```{r, results='hide'}
-data <- lavaan::simulateData(model_truth, "cfa", sample.nobs = c(20000, 20000), meanstructure = TRUE, standardized = TRUE)
-fit <- lavaan::cfa(model, data = data, group = "group", meanstructure = TRUE, std.lv = TRUE, group.equal = c("loadings", "intercepts")) 
-lavaan::summary(fit)
-```
 

### 2021-01-26 manuscript.Rmd
explain some ways to simulate more realistic data

--- 2021-01-26 manuscript.Rmd
@@
 Simulated data lets you verify that the data you expect to gather and the analysis that you plan fit together.
-Of course one has to make a lot of educated guesses.
-Generally there are of course hundreds of different equal possibilities, but we just care about the worst reasonable settings.
-Sticking with our example about machiavilism, we simulate multivariate normal data representing the machiavilism subscale of the short dark triad (SD3).
+Of course one has to make a lot of educated guesses to simulate data that closely matches your expectation.
+These guesses strongly depend on the study you are planning and there cannot be a one fits all solution.
+However, the analytic strategy often gives you a hint about how to simulate data.
+In the machiavilism example the analytic strategy is to use structural equation modeling (SEM).
+To simulate data for SEM one simply has to specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
+The model parameters should mimic the machiavilism subscale of the short dark triad (SD3).
 As a starting point we can use already publishes information.
@@
 
-From which we can simulate data and fit a model to it:
+From which we can simulate data:
 
 ```{r, results='hide'}
-data <- lavaan::simulateData(model_truth, "cfa", sample.nobs = c(20000, 20000), meanstructure = TRUE, standardized = TRUE)
-fit <- lavaan::cfa(model, data = data, group = "group", meanstructure = TRUE, std.lv = TRUE, group.equal = c("loadings", "intercepts")) 
-lavaan::summary(fit)
+simulated <- lavaan::simulateData(
+  model_truth,
+  sample.nobs = c(1000, 1000),
+  meanstructure = TRUE,
+  standardized = TRUE
+)
 ```
 
+And then fit the statistical model:
+
+```{r}
+fit <-
+  lavaan::cfa(
+    model,
+    data = simulated,
+    group = "group",
+    meanstructure = TRUE,
+    std.lv = TRUE
+  ) 
+```
+
+But wait `lavaan` warns us that this model may not be identified.
+We overlooked a crucial assumption of this model, that is measurement invariance.
+This assumption is also baked into our simulation of the data.
+It is probably more realistic to violate this assumption and plan our analysis accordingly.
+We therefore simulate the data so that some of the items parameter differ among groups and will then alter the analysis to assume partial measurement invariance.
+While reconsidering our assumptions it is also unlikely that the data is normally distributed.
+Not only would the real items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, therefore suggesting some skewness to the items.
+@Rhemtulla2012 conclude that using Likert-like items is acceptable when they employ five or more categories.
+We therefore just continue simulating continues data.
+However, adding a little skewness to the items is easy and helps make the simulation more realistically.
+Anyhow, remember that what we consider realistically for our case could be unfit for your case.
 <!--Hands-on: build simple functions tailored for your analysis-->

### 2021-01-26 manuscript.Rmd
check spelling & grammar

@@
 
-### Simulation (optional)
+## Simulation (optional)
 
@@
 A step to prevent unpleasant surprises is to simulate data.
-Simulated data lets you verify that the data you expect to gather and the analysis that you plan fit together.
-Of course one has to make a lot of educated guesses to simulate data that closely matches your expectation.
-These guesses strongly depend on the study you are planning and there cannot be a one fits all solution.
+Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
+Of course, one has to make many educated guesses to simulate data that closely matches your expectation.
+These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
 However, the analytic strategy often gives you a hint about how to simulate data.
-In the machiavilism example the analytic strategy is to use structural equation modeling (SEM).
-To simulate data for SEM one simply has to specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
-The model parameters should mimic the machiavilism subscale of the short dark triad (SD3).
-As a starting point we can use already publishes information.
+In the Machiavellism example, the analytic strategy is to use structural equation modelling (SEM).
+To simulate data for SEM, one has to specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
+The model parameters should mimic the Machiavellism subscale of the short dark triad (SD3).
+As a starting point, we can use already publishes information.
 @Jones2013 report loadings seen in Table \@ref(tab:loadings) and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
@@
 
-It is probably better to assume values that are more pessimistic.
-Instead of taking these values as they are we will therefore assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.1.
+It is probably better to assume more pessimistic values.
+Therefore, we will assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.1.
 These assumption can be simply translated into [`lavaan` [@lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
@@
 It is probably more realistic to violate this assumption and plan our analysis accordingly.
-We therefore simulate the data so that some of the items parameter differ among groups and will then alter the analysis to assume partial measurement invariance.
-While reconsidering our assumptions it is also unlikely that the data is normally distributed.
-Not only would the real items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, therefore suggesting some skewness to the items.
+Therefore, we simulate the data so that some of the items parameter differ among groups and then alter the analysis to assume partial measurement invariance.
+While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
+Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
 @Rhemtulla2012 conclude that using Likert-like items is acceptable when they employ five or more categories.
-We therefore just continue simulating continues data.
+Hence, we continue simulating continues data.
 However, adding a little skewness to the items is easy and helps make the simulation more realistically.

### 2021-01-27 manuscript.Rmd
explain why pessimistic
fix spelling

--- 2021-01-26 manuscript.Rmd
@@
 We deem each concept necessary to archive longterm reproducibility.
-However, which tools you use to implement this concepts are mere matter of taste and project requirements.
-This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking and Docker for software management.
+However, which tools you use to implement these concepts are mere matter of taste and project requirements.
+This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
 Their interplay is shown in Figure \@ref(fig:schematic).

### 2021-01-27 manuscript.Rmd
write more lavaan helper

--- 2021-01-26 manuscript.Rmd
@@
     - here
+  scripts:
+    - R/lavaan_helper.R
+    - R/simulation.R
+    - R/link.R
 csl: apa7.csl
@@
 library(repro)
+automate_load_scripts()
 source(here::here("R", "link.R"))
@@
 It is probably better to assume more pessimistic values.
+This way, we make sure that our analytic strategy is viable even in non-optimal circumstances.
 Therefore, we will assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.1.
@@
 ```{r, results='asis', echo=FALSE}
-model_truth <- paste(measurement(
+model_truth <- combine(
+  measurement(
     "MACH",
-  paste0("x", 1:9),
+    items("x", 9),
     loadings = round(loadings_jones_paulhus * 0.7, 2)
     ),
-intercepts("MACH", list(0, 0.1)),
-collapse = "\n", sep = "\n")
+  intercepts("MACH", list(0, 0.1))
+  )
 cat("``` r",
@@
     sep = "\n")
-model <- paste(measurement(
+model <- combine(measurement(
   "MACH",
-  paste0("x", 1:9)),
-  intercepts("MACH", list(0, "NA")),
-  sep = "\n")
+  items("x", 9)),
+  intercepts("MACH", list(0, "NA")))
 ```

### 2021-01-27 R/lavaan_helper.R

@@ -1,3 +1,7 @@
+items <- function(item, n){
+  stringr::str_c(item, seq_len(n))
+}
+
 intercepts <- function(variables, intercepts){
   stopifnot(is.character(variables))
   stringr::str_c(variables, " ~ ", modifier(intercepts), "1")
@@ -24,4 +28,20 @@ measurement <- function(latent, variables, loadings){
   stringr::str_c(latent, " =~ ", variables)
 }
 
+varying_modifer <- function(modifier, n, rel_change, abs_change){
+  if(missing(n))n <- ifelse(missing(abs_change), length(rel_change), length(abs_change))
+  pos <- sample(seq_along(modifier), n)
+  varied <- modifier
+  if(missing(abs_change)){
+    varied[pos] <- modifier[pos] * (1 + rel_change)
+    return(list(modifier, varied))
+  } else {
+    varied[pos] <- modifier[pos] + abs_change
+    return(list(modifier, varied))
+  }
+}
 
+combine <- function(...)stringr::str_c(..., sep = "\n", collapse = "\n")
+shuffle <- function(x)sample(x, length(x))
+pad <- function(x, replace, length)c(x, rep(replace, length))
+pad_zero <- function(x, length)pad(x, 0, length)

### 2021-01-27 manuscript.Rmd
add more realistic simulation function

@@
 Hence, we continue simulating continues data.
-However, adding a little skewness to the items is easy and helps make the simulation more realistically.
-Anyhow, remember that what we consider realistically for our case could be unfit for your case.
+However, adding a little skewness to the items is easy and helps make the simulation more realistic.
+Anyhow, remember that what we consider realistic for our case could be unfit for your case.
+If you are satisfied with your assumptions it helps to write a function that expresses them, like this one:
 <!--Hands-on: build simple functions tailored for your analysis-->
+
+```{r generate}
+```
+
+Note however that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/R/simulation.R", "GitHub")`.
+
+With its help we can simulate 500 cases with a mean difference between groups of 0.1, where the intercepts of three random items vary by .3 and three loadings vary by 30% and five random items have a skew.
+
+```{r}
+simulated <- generate(n = 500,
+                      diff = 0.1,
+                      vary_intercepts = rep(0.3, 3),
+                      vary_loadings = rep(0.3, 3),
+                      skewness = rep(5, -0.5))
+```
+
+```{r}
+fit <-
+  lavaan::cfa(
+    model,
+    data = simulated,
+    group = "group",
+    meanstructure = TRUE,
+    std.lv = TRUE,
+    group.equal = c("intercepts", "loadings")
+  ) 
+```
 

### 2021-01-27 R/lavaan_helper.R

@@ -4,7 +4,7 @@ items <- function(item, n){
 
 intercepts <- function(variables, intercepts){
   stopifnot(is.character(variables))
-  stringr::str_c(variables, " ~ ", modifier(intercepts), "1")
+  stringr::str_c(variables, " ~ ", modifier(intercepts), "1", collapse = "\n")
 }
 
 modifier <- function(values){
@@ -20,7 +20,7 @@ modifier <- function(values){
 }
 
 measurement <- function(latent, variables, loadings){
-  stopifnot(is.character(latent),
+stopifnot(is.character(latent),
             is.character(variables),
             length(latent) == 1)
   loadings <- modifier(loadings)

### 2021-01-27 R/simulation.R

@@ -1,15 +1,28 @@
 source(here::here("R", "lavaan_helper.R"))
-generate <- function(model, n){
-  
+loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
+cohend_jones_paulhus <- c(24, 29, 35)/100
+#----generate----
+generate <- function(n, diff, vary_intercepts, vary_loadings, skewness, ...){
+  latent <- "MACH"
+  items <- items("x", 9)
+  model <-
+    combine(
+      measurement(
+        latent,
+        items,
+        varying_modifer(loadings_jones_paulhus, rel_change = vary_loadings)
+      ),
+      intercepts(latent, list(0, diff)),
+      intercepts(items, varying_modifer(rep(0, 9), 3, abs_change = vary_intercepts))
+    )
+  skewness <- shuffle(pad_zero(skewness, length(items)))
+  lavaan::simulateData(model, skewness = skewness, sample.nobs = c(n/2, n/2), ...)
 }
-
+#----analyze----
 analyze <- function(model, data){
   
 }
-
+#----summarise----
 summarise <- function(fit){
   
 }
-
-loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
-cohend_jones_paulhus <- c(24, 29, 35)/100

### 2021-02-01 manuscript.Rmd
adept wording

--- 2021-01-27 manuscript.Rmd
@@
 We deem each concept necessary to archive longterm reproducibility.
-However, which tools you use to implement these concepts are mere matter of taste and project requirements.
+However, which tools you use to implement these concepts is a matter of taste and project requirements.
 This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.

### 2021-02-11 manuscript.Rmd
work on simulation; unfinished!

--- 2021-01-27 manuscript.Rmd
@@
 
+Note, that this certainly results in different numbers each time we run it.
+To consistently get the same results it is good practice to use a seed.
+The function `seeded_generate()` will always generate the same pseudo-random numbers, when given the same seed.
+
 ```{r}
-fit <-
-  lavaan::cfa(
-    model,
-    data = simulated,
-    group = "group",
-    meanstructure = TRUE,
-    std.lv = TRUE,
-    group.equal = c("intercepts", "loadings")
+seeded_generate <-  function(seed, ...)withr::with_seed(seed, generate(...))
+simulated <- seeded_generate(
+  seed = 1234,
+  n = 100000,
+  diff = 0.1,
+  vary_intercepts = rep(0.3, 3),
+  vary_loadings = rep(0.3, 3),
+  skewness = rep(5,-0.5)
 )
+```
+
+```{r}
+lavaan::summary(analyze(simulated))
 ```

### 2021-02-11 R/lavaan_helper.R

@@ -41,6 +41,13 @@ varying_modifer <- function(modifier, n, rel_change, abs_change){
   }
 }
 
+covariances <- function(x, y, covariances){
+  stopifnot(is.character(x), is.character(y))
+  stringr::str_c(x, " ~~ ", modifier(covariances), y, collapse = "\n")
+}
+
+variances <- function(x, variances)covariances(x, x, covariances = variances)
+
 combine <- function(...)stringr::str_c(..., sep = "\n", collapse = "\n")
 shuffle <- function(x)sample(x, length(x))
 pad <- function(x, replace, length)c(x, rep(replace, length))

### 2021-02-11 R/simulation.R

@@ -1,6 +1,19 @@
 source(here::here("R", "lavaan_helper.R"))
 loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
 cohend_jones_paulhus <- c(24, 29, 35)/100
+#----model-truth----
+model_truth <- combine(
+  measurement(
+    "MACH",
+    items("x", 9),
+    loadings = round(loadings_jones_paulhus * 0.7, 2)
+  ),
+  intercepts("MACH", list(0, 0.1)),
+  variances("MACH", list(1, 1))
+)
+#----model----
+
+
 #----generate----
 generate <- function(n, diff, vary_intercepts, vary_loadings, skewness, ...){
   latent <- "MACH"
@@ -13,14 +26,34 @@ generate <- function(n, diff, vary_intercepts, vary_loadings, skewness, ...){
         varying_modifer(loadings_jones_paulhus, rel_change = vary_loadings)
       ),
       intercepts(latent, list(0, diff)),
-      intercepts(items, varying_modifer(rep(0, 9), 3, abs_change = vary_intercepts))
+      intercepts(items, varying_modifer(rep(0, 9), 3, abs_change = vary_intercepts)),
+      variances(latent, list(1, 1))
     )
   skewness <- shuffle(pad_zero(skewness, length(items)))
-  lavaan::simulateData(model, skewness = skewness, sample.nobs = c(n/2, n/2), ...)
+  lavaan::simulateData(model, skewness = skewness, sample.nobs = c(n/2, n/2), standardized = TRUE, ...)
 }
 #----analyze----
-analyze <- function(model, data){
-  
+analyze <- function(simulated){
+  latent <- "MACH"
+  items <- items("x", 9)
+  model <-
+    combine(
+      measurement(
+        latent,
+        items
+      ),
+      intercepts(latent, list(0, "NA")),
+      variances(latent, list(1, 1))
+    )
+  browser()
+    lavaan::cfa(
+      model,
+      data = simulated,
+      group = "group",
+      meanstructure = TRUE,
+      std.lv = TRUE,
+      group.equal = c("intercepts", "loadings")
+    )
 }
 #----summarise----
 summarise <- function(fit){

### 2021-02-11 manuscript.Rmd
add part about readme

--- 2021-02-01 manuscript.Rmd
@@
 
-<!--I am troubled by how we should recomend to learn git-->
+<!--I am troubled by how we should recommend to learn git-->
 We recommend to attend an online lecture or workshop to learn the ins and outs of Git.
@@
 
+One of the greatest benefits of open research is that other people can freely reuse your materials.
+To use your materials, they first have to find them and learn how to use them.
+Interested users find out about your project via the readme.
+The readme is the first thing they see when they visit your Project.
+It should explain to them what your project is about and how to reproduce your project.
+You could for example include the abstract of your (planned) paper here.
+
+To add a readme to a repository run:
+
 <!--Hands on: Add a readme-->
@@
 
+This will create a file with the name `README.Rmd`.
+When you knit this file, it in turn creates a `README.md`.
+That means you can include R code or plots etc. within the readme.
+The `README.md` file will look pleasant in the webbrowser, just like in Figure \@ref(fig:idea-change).
+
+```{r readme-screenshot, include=FALSE}
+# unfortunately css selectors won't work for pdf so we'll use png instead
+readme <- here::here("images", "readme.png")
+webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/",
+                  file = readme,
+                  selector = "div#readme", # only select file change
+                  zoom = 1)# higher resolution
+```
+
+```{r readme, echo=FALSE, fig.cap="A screenshot of how a change tracked in Git is represented by GitHub.  ", out.width='100%'}
+knitr::include_graphics(readme)
+```
+
 <!--Hands on: Add a license-->
 ```r
-usethis::use_cc0_license()
+usethis::use_ccby_license()
 ```

### 2021-02-11 manuscript.Rmd
add part about licences

@@
 
+Even when other researchers are excited to use your code or materials, they are usually legally inhibited from to doing so.
+That is because in many countries (i.e. the USA) the creator of something has exclusive right to use their creation.
+A license allows other people to use what you have created.
+<!---exact copy from our previous manuscript, do we need to change? -->
+In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
+Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
+These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
+A great resource to choose a license is [choosealicense.com](https://choosealicense.com), however, no resource, including our recommendation, replaces legal advice.
+<!-- plagiary end -->
+It is important to get written consent from collaborators when you change/add the license.
+However, if they contribute to the project, after you have added a license, they give their consent by contributing.
+To add a license to your project simply run:
+
 <!--Hands on: Add a license-->

### 2021-02-12 manuscript.Rmd
add section on contributor covenant

@@
 That means you can include R code or plots etc. within the readme.
-The `README.md` file will look pleasant in the webbrowser, just like in Figure \@ref(fig:idea-change).
+The `README.md` file will look pleasant in the browser, just like in Figure \@ref(fig:idea-change).
 
@@
                   file = readme,
-                  selector = "div#readme", # only select file change
+                  selector = "div#readme",
                   zoom = 1)# higher resolution
@@
 It is important to get written consent from collaborators when you change/add the license.
-However, if they contribute to the project, after you have added a license, they give their consent by contributing.
-To add a license to your project simply run:
+However, if their first contribution occurs after you have added a license, they give their consent by contributing.
+To add a CC-BY license to your project simply run (there are similar commands for other licenses):
 
@@
 ```
+
+We strongly suggest that you strive for respectful interactions with anyone who might want to contribute.
+In fact, we hope it comes naturally to you.
+To communicate that contributors can expect a welcoming and respectful environment, you can add the `r link( "https://www.contributor-covenant.org", "Contributor Covenant")`.
+The Contributor Covenant is the code of conduct used by many open source project.
+Contributing to an open project is daunting and a public pledge to value any contributions (even when you do not incorporate them) can help others to take the first step.
+
+To add the Contributor Covenant to your project use:
 

### 2021-02-12 manuscript.Rmd
restructure headings

@@
 csl: apa7.csl
+numbered_headings: true
+toc_depth: 5
 bibliography: references.bib
@@
 
-### Preregestration (optional)
+## Preregestration (optional)
 
@@
 
-### Internal Dependencies
+# Analyzing Data
+
+## Internal Dependencies
 
@@
 
-### External Dependencies
+## External Dependencies
 
@@
 
-#### renv (alternative)
+### renv (alternative)
 
-#### Singularity (optional)
-
-## Analyzing Data
-
+### Singularity (optional)
 <!-- Hands on: add data via automate() -->
@@
 
-### Data Integrity (optional)
+## Data Integrity (optional)
 
@@
 
-### Anonymising Data (optional)
+## Anonymising Data (optional)
 
@@
 
-## Peer Review (optional)
+# Peer Review (optional)
 
-## Post Publication Review (optional)
+# Post Publication Review (optional)
 
-## Archiving
+# Archiving
 
@@
 
-## Checklist
+# Checklist
 

### 2021-02-12 manuscript.Rmd
adept readme screenshot

@@
                   file = readme,
+                  vwidth = 1920, # height & width are not resolution
+                  vheight = 1080, # but what website thinks the browser size is
                   selector = "div#readme",
-                  zoom = 1)# higher resolution
+                  zoom = 2) # higher resolution
 ```

### 2021-02-12 manuscript.Rmd
explain how to learn git

@@
 <!--I am troubled by how we should recommend to learn git-->
-We recommend to attend an online lecture or workshop to learn the ins and outs of Git.
+Git is can be challenging to learn, because we are used to learn alone and Git is best learned with others.
+We recommend to find a partner to attend an online lecture or workshop to learn the ins and outs of Git.
 
@@
 
-One of the greatest benefits of open research is that other people can freely reuse your materials.
+One of the greatest benefits of open research is that other people can freely reuse materials.
 To use your materials, they first have to find them and learn how to use them.

### 2021-02-15 manuscript.Rmd
minor changes

@@
 Therefore, we show the whole lifecycle of an open research project, where some parts are essential to archive reproducibility.
+All code chunks are meant for the reader to try out for themselves.
 Still, the reader can safely skip sections marked as optional, when unfit for their research question or methods.
@@
 However, which tools you use to implement these concepts is a matter of taste and project requirements.
-This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
+This tutorial will follow the recommendations of @Peikert2019 and utilize Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
 Their interplay is shown in Figure \@ref(fig:schematic).
@@
 In fact, we hope it comes naturally to you.
-To communicate that contributors can expect a welcoming and respectful environment, you can add the `r link( "https://www.contributor-covenant.org", "Contributor Covenant")`.
+To communicate that contributors can expect a welcoming and respectful environment, you can add the `r link( "https://www.contributor-covenant.org", "Contributor Covenant")` to your repository.
 The Contributor Covenant is the code of conduct used by many open source project.

### 2021-02-15 manuscript.Rmd
add newline after header
fix figure ref

@@
 # Header
+
 Normal text.
@@
 That means you can include R code or plots etc. within the readme.
-The `README.md` file will look pleasant in the browser, just like in Figure \@ref(fig:idea-change).
+The `README.md` file will look pleasant in the browser, just like in Figure \@ref(fig:readme).
 

### 2021-02-15 manuscript.Rmd
summarise preparations for publishing
fix typo

@@
 <!--I am troubled by how we should recommend to learn git-->
-Git is can be challenging to learn, because we are used to learn alone and Git is best learned with others.
+Git can be challenging to learn, because we are used to learn alone and Git is best learned with others.
 We recommend to find a partner to attend an online lecture or workshop to learn the ins and outs of Git.
@@
 To communicate that contributors can expect a welcoming and respectful environment, you can add the `r link( "https://www.contributor-covenant.org", "Contributor Covenant")` to your repository.
-The Contributor Covenant is the code of conduct used by many open source project.
+The Contributor Covenant is a code of conduct used by many open source project.
 Contributing to an open project is daunting and a public pledge to value any contributions (even when you do not incorporate them) can help others to take the first step.
@@
 ```
+
+If you have considered what you want to tell the public about the project with a readme, have chosen a license and thought about how to make collaboration welcoming, you may publish your Git repository.
 

### 2021-02-15 manuscript.Rmd
replace fig caption placeholder
add introduction to collabortation

@@
 
-```{r readme, echo=FALSE, fig.cap="A screenshot of how a change tracked in Git is represented by GitHub.  ", out.width='100%'}
+```{r readme, echo=FALSE, fig.cap="A screenshot of how a README.md is rendered/presented on GitHub.  ", out.width='100%'}
 knitr::include_graphics(readme)
@@
 If you have considered what you want to tell the public about the project with a readme, have chosen a license and thought about how to make collaboration welcoming, you may publish your Git repository.
+See the section on [GitHub] on how to publish your project.
 
 ## Collaboration
+
+Often, reproducibility is reduced to the merits it has for the reader.
+However, we argue that reproducibility is vital for collaboration, not only in the sense that the consumer of a scientific product build upon it, but also that the creators work together.
+The time that you invest into ensuring reproducibility, is directed at ensuring effective collaboration.
+Therefore, reproducibility is at its core about collaboration.
 

### 2021-02-19 manuscript.Rmd
add webshots to docker: - install google chrome in docker - install webshot2

@@
     - usethis
-    - repro
+    - aaronpeikert/repro@7bfaf98
     - here
+    - rstudio/webshot2@f62e743
 csl: apa7.csl

### 2021-02-23 manuscript.Rmd
add section on github

@@
 <!--I am leaning towards doing everything in R, not sure though-->
-<!--For Git we maybe should use RStudio's Git pane-->
+<!--For Git we maybe should use RStudio's Git pane?-->
 
@@
 
-And Git can show you what has changed, like in Figure \@ref(fig:idea-change).
+And GitHub can show you what has changed, like in Figure \@ref(fig:idea-change).
+GitHub is in its essence a Webinterface to Git and you may learn more about how to use it under the section on [GitHub].
 
@@
 Even when other researchers are excited to use your code or materials, they are usually legally inhibited from to doing so.
-That is because in many countries (i.e. the USA) the creator of something has exclusive right to use their creation.
+That is because in many countries (i.e. the USA, European Union) the creator of something has exclusive right to use their creation.
 A license allows other people to use what you have created.
@@
 
-GitHub/GitLab are great as a shared information basis: sharing and editing files, discussing and distributing tasks via issues and documenting the project via wikis even before data collection starts.
+On your local machine, you use Git to track changes.
+However, to share these changes with others, you need to use an online service like GitHub.
+GitHub is not the only solution; there is also GitLab or BitBucket, which are as feature-rich as GitHub.
+GitLab has the advantage that your institution can host it themselves; therefore, you are not relying on commercial service providers.
+We focus on GitHub because it is the most popular service.
+To upload your changes to GitHub, you need to create an account under https://github.com/join.
+As a researcher, you are eligible for the `r link("https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount", "Researcher/Educator discount").
+The following command creates a new GitHub repository under your GitHub account, beware this is public:
+<!--Hands-on: Introduce Github-->
 
-<!--Hands-on: Introduce Github-->
+```r
+usethis::use_github()
+```
+
+To create a private repository (only you and people you invite have access), use:
 

### 2021-02-23 manuscript.Rmd
add links
revise style

@@
 <!-- lure the reader in -->
-This notion of independent verifiability has long been part of "good scientific practice", but has not reached its full potential to benefit the scientific community.
+The research community has long accepted reproducibility as part of "good scientific practice", but has not has not benefited much from this self-imposed requirement.
 <!-- expand: who is calling it a requirement, maybe metascientific arguments-->
@@
 However, which tools you use to implement these concepts is a matter of taste and project requirements.
-This tutorial will follow the recommendations of @Peikert2019 and utilize Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
+This tutorial follows the recommendations of @Peikert2019 and utilize Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
 Their interplay is shown in Figure \@ref(fig:schematic).
@@
 <!-- researcher should begin early to enjoy most benefits -->
-It can be tempting to leave considerations of reproducibility for when you finish a project.
-<!-- weird sentence, I'll leave it as a placeholder-->
-However, reproducibility is invaluable for collaboration---including with the version of yourself from a few weeks ago.
+Reproducibility is invaluable for effective collaboration---including with your "past self".
+While you might be tempted to consider reproducibility only when you about to finish a project, upfront investments pay of through your increased productivity.
 Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you archive reproducibility gradually.
@@
 Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
-For example this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/master", "iterations")`.
+For example, this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/master", "iterations")`.
 To activate Git in a given project, you can call:
@@
 And GitHub can show you what has changed, like in Figure \@ref(fig:idea-change).
-GitHub is in its essence a Webinterface to Git and you may learn more about how to use it under the section on [GitHub].
+GitHub is in its essence a webinterface to Git and you may learn more about how to use it under the section on [GitHub].
 
@@
 Interested users find out about your project via the readme.
-The readme is the first thing they see when they visit your Project.
+The readme is the first thing they see when they visit your project on [GitHub].
 It should explain to them what your project is about and how to reproduce your project.
@@
 
-This will create a file with the name `README.Rmd`.
+This command will create a file with the name `README.Rmd`.
+<!--knit/rmarkdown will be explained in section above-->
 When you knit this file, it in turn creates a `README.md`.
@@
 On your local machine, you use Git to track changes.
-However, to share these changes with others, you need to use an online service like GitHub.
-GitHub is not the only solution; there is also GitLab or BitBucket, which are as feature-rich as GitHub.
+Online services like GitHub allow you to share these changes with others.
+GitHub is not the only solution; there is also [GitLab](gitlab.com) or [BitBucket](https://bitbucket.org), which are as feature-rich as GitHub.
 GitLab has the advantage that your institution can host it themselves; therefore, you are not relying on commercial service providers.
@@
 ```r
-usethis::use_github()
+usethis::use_github(private = TRUE)
 ```

### 2021-02-24 manuscript.Rmd
add missing backtick

@@
 title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
-output: bookdown::pdf_document2
+output:
+  bookdown::pdf_document2:
+    keep_tex: true
 repro:
@@
 To communicate that contributors can expect a welcoming and respectful environment, you can add the `r link( "https://www.contributor-covenant.org", "Contributor Covenant")` to your repository.
+Contributing to an open project is daunting and a public pledge to value any contributions (even when you do not incorporate them) can help others to take the first step.
 The Contributor Covenant is a code of conduct used by many open source project.
-Contributing to an open project is daunting and a public pledge to value any contributions (even when you do not incorporate them) can help others to take the first step.
 
@@
 ```
+
+However, there is some controversy around how to handle violations of a code of conduct.
+For that or any other reason, you may want to adapt the Contributor Covenant, to fit the needs of you project (which you are allowed to to because it is published with a CC-BY license).
 
@@
 To upload your changes to GitHub, you need to create an account under https://github.com/join.
-As a researcher, you are eligible for the `r link("https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount", "Researcher/Educator discount").
+As a researcher, you are eligible for the `r link("https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount", "Researcher/Educator discount")`.
 The following command creates a new GitHub repository under your GitHub account, beware this is public:

### 2021-02-24 manuscript.Rmd
Revert "add missing backtick"
This reverts commit 1a3ddd3a3960909c8b5e8f9ce641fbef61763c31.
add missing backtick

@@
 title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
-output:
-  bookdown::pdf_document2:
-    keep_tex: true
+output: bookdown::pdf_document2
 repro:

### 2021-02-24 manuscript.Rmd
add (R) Markdown, Git as header for cross references

@@
 
-```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency management (Make), software management (Docker), and dynamic document generation (R Markdown).
+```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency tracking (Make), software management (Docker), and dynamic document generation (R Markdown).
 Git tracks changes to the project over time.
@@
 
+## Markdown
+
 <!--Keep notes.-->
@@
 We will later see how this simplicity enables us to create many different document formats (i.e. Word, PDF, HTML) and incorporate code (i.e. R, Python, Julia).
+
+## Git
+
 Another advantage of Markdown is that you can easily version-control it with Git.
@@
 A preregistration is a tool that may help to increase the credibility of empirical results.
+
+### R Markdown
 

### 2021-02-24 manuscript.Rmd
explain what is still missing
close #3

@@
 However, which tools you use to implement these concepts is a matter of taste and project requirements.
-This tutorial follows the recommendations of @Peikert2019 and utilize Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
+This tutorial follows the recommendations of @Peikert2019 and utilize Git for version control, R Markdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
 Their interplay is shown in Figure \@ref(fig:schematic).
@@
 <!-- This is maybe the place to introduce the example-->
-First, you could create a new markdown file named `idea.md` that looks like this:
+First, you could create a new Markdown file named `idea.md` that looks like this:
 
@@
 
-<!--Hands on: Introduce RMarkdown-->
+<!--Hands on: Introduce R Markdown-->
 <!--Hands on: gitignore resulting .html/pdf-->
@@
 
+While you already know how to write dynamic documents with [Markdown] and [R Markdown], and track changes with [Git] and [GitHub], two components to archive longterm reproducibility are still missing.
+
 ## Internal Dependencies

### 2021-03-05 manuscript.Rmd
add alternative codes of conduct

@@
 However, there is some controversy around how to handle violations of a code of conduct.
-For that or any other reason, you may want to adapt the Contributor Covenant, to fit the needs of you project (which you are allowed to to because it is published with a CC-BY license).
+For that or any other reason, you may want to adapt the Contributor Covenant, to fit the needs of you project (which you are allowed to do because it is published with a CC-BY license).
+You can also consider alternative Codes of Conduct, like the `r link("https://opensource.microsoft.com/codeofconduct/", "Microsoft Open Source Code of Conduct")` or the `r link("https://berlincodeofconduct.org", "Berlin Code of Conduct")`.
 
@@
 GitLab has the advantage that your institution can host it themselves; therefore, you are not relying on commercial service providers.
-We focus on GitHub because it is the most popular service.
+Anyhow, we focus on GitHub because it is the most popular service.
 To upload your changes to GitHub, you need to create an account under https://github.com/join.
@@
 
-A well defined "entry point" that automatically reproduces everything eases collaboration.
+<!--I am not sure this is the right place, should we introduce this earlier? -->
 
-<!--I am not sure this is the right place, should we introduce this earlier? -->
+First dependency management, which means that you capture the interrelations between the components of you analysis.
+You cannot run an analysis without data, you cannot render a dynamic document without an analysis, and so forth.
+Understanding these relations is often key to reproducing a project.
+A reader of your reproducible manuscript should be able to derive how they are to reproduce it (ideally by a statement in your [README])
+The "components" of an analysis are often loosely mapped onto files.
+For a simple project it might be enough to have a single R Markdown.
+However, while it is easy to reproduce this approach quickly gets unwieldy.
 

### 2021-04-06 manuscript.Rmd
fix typo

--- 2021-02-24 manuscript.Rmd
@@
     - usethis
-    - aaronpeikert/repro@7bfaf98
+    - repro
     - here
-    - rstudio/webshot2@f62e743
+  scripts:
+    - R/lavaan_helper.R
+    - R/simulation.R
+    - R/link.R
 csl: apa7.csl
-numbered_headings: true
-toc_depth: 5
-bibliography: references.bib
+bibliography: ["references.bib", "packages.bib"]
 ---
@@
 <!-- lure the reader in -->
-The research community has long accepted reproducibility as part of "good scientific practice", but has not has not benefited much from this self-imposed requirement.
+This notion of independent verifiability has long been part of "good scientific practice", but has not reached its full potential to benefit the scientific community.
 <!-- expand: who is calling it a requirement, maybe metascientific arguments-->
@@
 Therefore, we show the whole lifecycle of an open research project, where some parts are essential to archive reproducibility.
-All code chunks are meant for the reader to try out for themselves.
 Still, the reader can safely skip sections marked as optional, when unfit for their research question or methods.
@@
 However, which tools you use to implement these concepts is a matter of taste and project requirements.
-This tutorial follows the recommendations of @Peikert2019 and utilize Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
+This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
 Their interplay is shown in Figure \@ref(fig:schematic).
@@
 
-```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency tracking (Make), software management (Docker), and dynamic document generation (R Markdown).
+```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency management (Make), software management (Docker), and dynamic document generation (R Markdown).
 Git tracks changes to the project over time.
@@
 <!-- researcher should begin early to enjoy most benefits -->
-Reproducibility is invaluable for effective collaboration---including with your "past self".
-While you might be tempted to consider reproducibility only when you about to finish a project, upfront investments pay of through your increased productivity.
+It can be tempting to leave considerations of reproducibility for when you finish a project.
+<!-- weird sentence, I'll leave it as a placeholder-->
+However, reproducibility is invaluable for collaboration---including with the version of yourself from a few weeks ago.
 Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you archive reproducibility gradually.
@@
 > File → New Project... → New Directory → New Project
-
-## Markdown
 
@@
 # Header
-
 Normal text.
@@
 We will later see how this simplicity enables us to create many different document formats (i.e. Word, PDF, HTML) and incorporate code (i.e. R, Python, Julia).
-
-## Git
-
 Another advantage of Markdown is that you can easily version-control it with Git.
@@
 Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
-For example, this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/master", "iterations")`.
+For example this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/master", "iterations")`.
 To activate Git in a given project, you can call:
@@
 <!--I am leaning towards doing everything in R, not sure though-->
-<!--For Git we maybe should use RStudio's Git pane?-->
+<!--For Git we maybe should use RStudio's Git pane-->
 
@@
 
-And GitHub can show you what has changed, like in Figure \@ref(fig:idea-change).
-GitHub is in its essence a webinterface to Git and you may learn more about how to use it under the section on [GitHub].
+And Git can show you what has changed, like in Figure \@ref(fig:idea-change).
 
@@
 
-<!--I am troubled by how we should recommend to learn git-->
-Git can be challenging to learn, because we are used to learn alone and Git is best learned with others.
-We recommend to find a partner to attend an online lecture or workshop to learn the ins and outs of Git.
+<!--I am troubled by how we should recomend to learn git-->
+We recommend to attend an online lecture or workshop to learn the ins and outs of Git.
 
@@
 
+```{r, include=FALSE}
+source(here::here("R", "simulation.R"))
+```
+
 There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
 A step to prevent unpleasant surprises is to simulate data.
-Simulated data lets you verify that the data you expect to gather and the analysis that you plan fit together.
+Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
+Of course, one has to make many educated guesses to simulate data that closely matches your expectation.
+These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
+However, the analytic strategy often gives you a hint about how to simulate data.
+In the Machiavellism example, the analytic strategy is to use structural equation modelling (SEM).
+To simulate data for SEM, one has to specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
+The model parameters should mimic the Machiavellism subscale of the short dark triad (SD3).
+As a starting point, we can use already published information.
+@Jones2013 report loadings seen in Table \@ref(tab:loadings) and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
 
+```{r loadings, echo=FALSE}
+knitr::kable(data.frame(Item = 1:9,
+                        Loading = loadings_jones_paulhus),
+             caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
+```
+
+It is probably better to assume more pessimistic values.
+This way, we make sure that our analytic strategy is viable even in non-optimal circumstances.
+Therefore, we will assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.1.
+These assumption can be simply translated into [`lavaan` [@lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
+
+[^lavaan]: https://lavaan.ugent.be
+
+```{r, results='asis', echo=FALSE}
+model_truth <- combine(
+  measurement(
+    "MACH",
+    items("x", 9),
+    loadings = round(loadings_jones_paulhus * 0.7, 2)
+    ),
+  intercepts("MACH", list(0, 0.1))
+  )
+cat("``` r",
+    model_truth,
+    "```",
+    sep = "\n")
+model <- combine(measurement(
+  "MACH",
+  items("x", 9)),
+  intercepts("MACH", list(0, "NA")))
+```
+
+From which we can simulate data:
+
+```{r, results='hide'}
+simulated <- lavaan::simulateData(
+  model_truth,
+  sample.nobs = c(1000, 1000),
+  meanstructure = TRUE,
+  standardized = TRUE
+)
+```
+
+And then fit the statistical model:
+
+```{r}
+fit <-
+  lavaan::cfa(
+    model,
+    data = simulated,
+    group = "group",
+    meanstructure = TRUE,
+    std.lv = TRUE
+  ) 
+```
+
+But wait `lavaan` warns us that this model may not be identified.
+We overlooked a crucial assumption of this model, that is measurement invariance.
+This assumption is also baked into our simulation of the data.
+It is probably more realistic to violate this assumption and plan our analysis accordingly.
+Therefore, we simulate the data so that some of the items parameter differ among groups and then alter the analysis to assume partial measurement invariance.
+While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
+Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
+@Rhemtulla2012 conclude that using Likert-like items is acceptable when they employ five or more categories.
+Hence, we continue simulating continues data.
+However, adding a little skewness to the items is easy and helps make the simulation more realistic.
+Anyhow, remember that what we consider realistic for our case could be unfit for your case.
+If you are satisfied with your assumptions it helps to write a function that expresses them, like this one:
 <!--Hands-on: build simple functions tailored for your analysis-->
 
-## Preregestration (optional)
+```{r generate}
+```
+
+Note however that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/R/simulation.R", "GitHub")`.
+
+With its help we can simulate 500 cases with a mean difference between groups of 0.1, where the intercepts of three random items vary by .3 and three loadings vary by 30% and five random items have a skew.
+
+```{r}
+simulated <- generate(n = 500,
+                      diff = 0.1,
+                      vary_intercepts = rep(0.3, 3),
+                      vary_loadings = rep(0.3, 3),
+                      skewness = rep(5, -0.5))
+```
+
+Note, that this certainly results in different numbers each time we run it.
+To consistently get the same results it is good practice to use a seed.
+The function `seeded_generate()` will always generate the same pseudo-random numbers, when given the same seed.
+
+```{r}
+seeded_generate <-  function(seed, ...)withr::with_seed(seed, generate(...))
+simulated <- seeded_generate(
+  seed = 1234,
+  n = 100000,
+  diff = 0.1,
+  vary_intercepts = rep(0.3, 3),
+  vary_loadings = rep(0.3, 3),
+  skewness = rep(5,-0.5)
+)
+```
+
+```{r}
+lavaan::summary(analyze(simulated))
+```
+
+### Preregestration (optional)
 
 A preregistration is a tool that may help to increase the credibility of empirical results.
-
-### R Markdown
 
@@
 
-One of the greatest benefits of open research is that other people can freely reuse materials.
-To use your materials, they first have to find them and learn how to use them.
-Interested users find out about your project via the readme.
-The readme is the first thing they see when they visit your project on [GitHub].
-It should explain to them what your project is about and how to reproduce your project.
-You could for example include the abstract of your (planned) paper here.
-
-To add a readme to a repository run:
-
 <!--Hands on: Add a readme-->
@@
 
-This command will create a file with the name `README.Rmd`.
-<!--knit/rmarkdown will be explained in section above-->
-When you knit this file, it in turn creates a `README.md`.
-That means you can include R code or plots etc. within the readme.
-The `README.md` file will look pleasant in the browser, just like in Figure \@ref(fig:readme).
-
-```{r readme-screenshot, include=FALSE}
-# unfortunately css selectors won't work for pdf so we'll use png instead
-readme <- here::here("images", "readme.png")
-webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/",
-                  file = readme,
-                  vwidth = 1920, # height & width are not resolution
-                  vheight = 1080, # but what website thinks the browser size is
-                  selector = "div#readme",
-                  zoom = 2) # higher resolution
-```
-
-```{r readme, echo=FALSE, fig.cap="A screenshot of how a README.md is rendered/presented on GitHub.  ", out.width='100%'}
-knitr::include_graphics(readme)
-```
-
-Even when other researchers are excited to use your code or materials, they are usually legally inhibited from to doing so.
-That is because in many countries (i.e. the USA, European Union) the creator of something has exclusive right to use their creation.
-A license allows other people to use what you have created.
-<!---exact copy from our previous manuscript, do we need to change? -->
-In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
-Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
-These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
-A great resource to choose a license is [choosealicense.com](https://choosealicense.com), however, no resource, including our recommendation, replaces legal advice.
-<!-- plagiary end -->
-It is important to get written consent from collaborators when you change/add the license.
-However, if their first contribution occurs after you have added a license, they give their consent by contributing.
-To add a CC-BY license to your project simply run (there are similar commands for other licenses):
-
 <!--Hands on: Add a license-->
 ```r
-usethis::use_ccby_license()
+usethis::use_cc0_license()
 ```
-
-We strongly suggest that you strive for respectful interactions with anyone who might want to contribute.
-In fact, we hope it comes naturally to you.
-To communicate that contributors can expect a welcoming and respectful environment, you can add the `r link( "https://www.contributor-covenant.org", "Contributor Covenant")` to your repository.
-Contributing to an open project is daunting and a public pledge to value any contributions (even when you do not incorporate them) can help others to take the first step.
-The Contributor Covenant is a code of conduct used by many open source project.
-
-To add the Contributor Covenant to your project use:
 
@@
 
-However, there is some controversy around how to handle violations of a code of conduct.
-For that or any other reason, you may want to adapt the Contributor Covenant, to fit the needs of you project (which you are allowed to to because it is published with a CC-BY license).
-
-If you have considered what you want to tell the public about the project with a readme, have chosen a license and thought about how to make collaboration welcoming, you may publish your Git repository.
-See the section on [GitHub] on how to publish your project.
-
 ## Collaboration
-
-Often, reproducibility is reduced to the merits it has for the reader.
-However, we argue that reproducibility is vital for collaboration, not only in the sense that the consumer of a scientific product build upon it, but also that the creators work together.
-The time that you invest into ensuring reproducibility, is directed at ensuring effective collaboration.
-Therefore, reproducibility is at its core about collaboration.
 
@@
 
-On your local machine, you use Git to track changes.
-Online services like GitHub allow you to share these changes with others.
-GitHub is not the only solution; there is also [GitLab](gitlab.com) or [BitBucket](https://bitbucket.org), which are as feature-rich as GitHub.
-GitLab has the advantage that your institution can host it themselves; therefore, you are not relying on commercial service providers.
-We focus on GitHub because it is the most popular service.
-To upload your changes to GitHub, you need to create an account under https://github.com/join.
-As a researcher, you are eligible for the `r link("https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount", "Researcher/Educator discount")`.
-The following command creates a new GitHub repository under your GitHub account, beware this is public:
+GitHub/GitLab are great as a shared information basis: sharing and editing files, discussing and distributing tasks via issues and documenting the project via wikis even before data collection starts.
+
 <!--Hands-on: Introduce Github-->
@@
 
-To create a private repository (only you and people you invite have access), use:
-
-```r
-usethis::use_github(private = TRUE)
-```
-
 <!--Hands on: Introduce GitHub Issues-->
@@
 
-# Analyzing Data
-
-## Internal Dependencies
+### Internal Dependencies
 
@@
 
-## External Dependencies
+### External Dependencies
 
@@
 
-### renv (alternative)
+#### renv (alternative)
 
-### Singularity (optional)
+#### Singularity (optional)
+
+## Analyzing Data
+
 <!-- Hands on: add data via automate() -->
@@
 
-## Data Integrity (optional)
+### Data Integrity (optional)
 
@@
 
-## Anonymising Data (optional)
+### Anonymising Data (optional)
 
@@
 
-# Peer Review (optional)
+## Peer Review (optional)
 
-# Post Publication Review (optional)
+## Post Publication Review (optional)
 
-# Archiving
+## Archiving
 
@@
 
-# Checklist
+## Checklist
 

### 2021-04-06 R/simulation.R
add basic data and models for simulation

@@ -1,4 +1,5 @@
 source(here::here("R", "lavaan_helper.R"))
+library(lavaan)
 loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
 cohend_jones_paulhus <- c(24, 29, 35)/100
 #----model-truth----
@@ -11,10 +12,37 @@ model_truth <- combine(
   intercepts("MACH", list(0, 0.1)),
   variances("MACH", list(1, 1))
 )
-#----model----
+#----models----
+model_same <- 
+  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+  MACH ~ c(a,a)*1
+  x1 ~ 0*1"
+
+model_differ <- 
+  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+  MACH ~ c(a,b)*1
+  x1 ~ 0*1"
 
 
 #----generate----
+n = 3000 
+data <- lavaan::simulateData(
+  model_truth,
+  sample.nobs = c(n/2, n/2), 
+  standardized = TRUE)
+
+#----analyze----
+res_same <- cfa(model_same, data, group = "group", group.equal = c("loadings", "intercepts"))
+
+summary(res_same)
+
+res_differ <- cfa(model_differ, data, group = "group", group.equal = c("loadings", "intercepts"))
+
+summary(res_differ)
+
+
+#----some functions----
+
 generate <- function(n, diff, vary_intercepts, vary_loadings, skewness, ...){
   latent <- "MACH"
   items <- items("x", 9)
@@ -32,7 +60,7 @@ generate <- function(n, diff, vary_intercepts, vary_loadings, skewness, ...){
   skewness <- shuffle(pad_zero(skewness, length(items)))
   lavaan::simulateData(model, skewness = skewness, sample.nobs = c(n/2, n/2), standardized = TRUE, ...)
 }
-#----analyze----
+
 analyze <- function(simulated){
   latent <- "MACH"
   items <- items("x", 9)

### 2021-04-06 R/simulation.R
add function to simulate data and first power plot

@@ -9,7 +9,7 @@ model_truth <- combine(
     items("x", 9),
     loadings = round(loadings_jones_paulhus * 0.7, 2)
   ),
-  intercepts("MACH", list(0, 0.1)),
+  intercepts("MACH", list(0, 0.5)),
   variances("MACH", list(1, 1))
 )
 #----models----
@@ -25,7 +25,7 @@ model_differ <-
 
 
 #----generate----
-n = 3000 
+n = 100 
 data <- lavaan::simulateData(
   model_truth,
   sample.nobs = c(n/2, n/2), 
@@ -40,9 +40,56 @@ res_differ <- cfa(model_differ, data, group = "group", group.equal = c("loadings
 
 summary(res_differ)
 
+fitMeasures(res_same, "BIC")
+fitMeasures(res_differ, "BIC")
+
+generate_data <- function(n_obs, n_sim, model_truth, model_1, model_2, ...){
+  res_n <- list()
+  for (i in 1:length(n_obs)){
+    res <- c()
+    for (j in 1:n_sim){
+      data <- lavaan::simulateData(
+        model_truth,
+        sample.nobs = c(n_obs[i]/2, n_obs[i]/2), 
+        standardized = TRUE)
+      
+      res_1 <- cfa(model_1, data, group = "group", ...)
+      res_2 <- cfa(model_2, data, group = "group", ...)
+      if(lavInspect(res_1, "converged") & lavInspect(res_2, "converged"))
+        res[j] <- (fitMeasures(res_1, "BIC") - fitMeasures(res_2, "BIC")) > 0
+      else
+        res[j] <- NA
+    }
+    res_n[[i]] <- res
+  }
+  return(res_n)
+}
+
+n_obs = c(50, 100, 200, 300, 500, 1000)
+
+res <- generate_data(
+  n_obs, 
+  100, 
+  model_truth, 
+  model_same, 
+  model_differ, 
+  group.equal = c("loadings", "intercepts"))
+
+res_data <- data.frame(n = n_obs)
+
+res_data$result = res
+res_data$p <- purrr::map_dbl(res_data$result, mean, na.rm = TRUE)
+res_data$n_conv <- purrr::map_dbl(res_data$result, ~sum(is.na(.x)))
+
+res_data$n_conv
+
+library(ggplot2)
+
+ggplot(res_data, aes(x = n, y = p)) + geom_point() + geom_line() + 
+  scale_x_continuous(breaks = n_obs)
+  
 
[…]

### 2021-04-06 R/simulation.R
add aic, p_value etc. to power simulation

@@ -9,85 +9,144 @@ model_truth <- combine(
     items("x", 9),
     loadings = round(loadings_jones_paulhus * 0.7, 2)
   ),
-  intercepts("MACH", list(0, 0.5)),
+  intercepts("MACH", list(0, 0.2)),
   variances("MACH", list(1, 1))
 )
 #----models----
 model_same <- 
   "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(a,a)*1
-  x1 ~ 0*1"
+  MACH ~ c(0,0)*1
+  MACH ~~ 1*MACH"
 
 model_differ <- 
   "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(a,b)*1
-  x1 ~ 0*1"
+  MACH ~ c(0,NA)*1
+  MACH ~~ 1*MACH"
 
 
 #----generate----
-n = 100 
+n = 500
 data <- lavaan::simulateData(
   model_truth,
   sample.nobs = c(n/2, n/2), 
   standardized = TRUE)
 
 #----analyze----
-res_same <- cfa(model_same, data, group = "group", group.equal = c("loadings", "intercepts"))
+res_same <- cfa(
+  model_same, 
+  data, group = "group", 
+  group.equal = c("loadings", "intercepts"), 
+  std.lv= TRUE)
 
 summary(res_same)
 
-res_differ <- cfa(model_differ, data, group = "group", group.equal = c("loadings", "intercepts"))
+res_differ <- cfa(
+  model_differ, 
+  data, 
+  group = "group", 
+  group.equal = c("loadings", "intercepts"), 
+  std.lv= TRUE)
 
 summary(res_differ)
 
 fitMeasures(res_same, "BIC")
 fitMeasures(res_differ, "BIC")
 
+lavaan::anova(res_same, res_differ)$`Pr(>Chisq)`[2]
+
+dplyr::filter(parameterEstimates(res_differ), lhs == "MACH", op == "~1", group == 2)$est
+dplyr::filter(parameterEstimates(res_differ), lhs == "MACH", op == "~1", group == 2)$se
+dplyr::filter(parameterEstimates(res_differ), lhs == "MACH", op == "~1", group == 2)$pvalue
+
+
 generate_data <- function(n_obs, n_sim, model_truth, model_1, model_2, ...){
-  res_n <- list()
[…]

### 2021-04-07 R/simulation.R
compare free variance in one group

@@ -16,12 +16,18 @@ model_truth <- combine(
 model_same <- 
   "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
   MACH ~ c(0,0)*1
-  MACH ~~ 1*MACH"
+  MACH ~~ c(1, NA)*MACH"
+# MACH ~~ 1*MACH
+# mean(res_data$dec_p)
+# MACH ~~ c(1, NA)*MACH
+# [1] 0.466
+# mean(res_data$dec_p)
+# [1] 0.464
 
 model_differ <- 
   "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
   MACH ~ c(0,NA)*1
-  MACH ~~ 1*MACH"
+  MACH ~~ c(1, NA)*MACH"
 
 
 #----generate----
@@ -115,11 +121,11 @@ generate_data <- function(n_obs, n_sim, model_truth, model_1, model_2, ...){
     est = est_n))
 }
 
-n_obs = c(25, 50, 100, 200, 300, 500)
+n_obs = c(300)
 
 res <- generate_data(
   n_obs, 
-  20, 
+  1000, 
   model_truth, 
   model_same, 
   model_differ, 
@@ -143,7 +149,6 @@ res_data %>% group_by(n) %>%
   pivot_longer(c(bic, aic, p_aov), names_to = "metric", values_to = "power") %>% 
   ggplot(aes(x = n, y = power, color = metric)) + geom_point() + geom_line() +
   theme_minimal() + scale_x_continuous(breaks = n_obs)
-
 
 res_data %>% ggplot(aes(x = est, color = factor(n))) + geom_density() +
   theme_minimal()

### 2021-04-07 R/simulation.R
tidy up

@@ -1,7 +1,11 @@
 source(here::here("R", "lavaan_helper.R"))
+library(tidyverse)
 library(lavaan)
+
+#----jones_paulhus----
 loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
 cohend_jones_paulhus <- c(24, 29, 35)/100
+
 #----model-truth----
 model_truth <- combine(
   measurement(
@@ -12,6 +16,7 @@ model_truth <- combine(
   intercepts("MACH", list(0, 0.2)),
   variances("MACH", list(1, 1))
 )
+
 #----models----
 model_same <- 
   "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
@@ -26,12 +31,12 @@ model_same <-
 
 model_differ <- 
   "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0,NA)*1
+  MACH ~ c(0, NA)*1 + c(zero, diff)*1
   MACH ~~ c(1, NA)*MACH"
 
 
 #----generate----
-n = 500
+n = 5000
 data <- lavaan::simulateData(
   model_truth,
   sample.nobs = c(n/2, n/2), 
@@ -43,158 +48,98 @@ res_same <- cfa(
   data, group = "group", 
   group.equal = c("loadings", "intercepts"), 
   std.lv= TRUE)
-
 summary(res_same)
-
 res_differ <- cfa(
   model_differ, 
   data, 
   group = "group", 
   group.equal = c("loadings", "intercepts"), 
   std.lv= TRUE)
-
 summary(res_differ)
 
-fitMeasures(res_same, "BIC")
-fitMeasures(res_differ, "BIC")
-
-lavaan::anova(res_same, res_differ)$`Pr(>Chisq)`[2]
-
-dplyr::filter(parameterEstimates(res_differ), lhs == "MACH", op == "~1", group == 2)$est
-dplyr::filter(parameterEstimates(res_differ), lhs == "MACH", op == "~1", group == 2)$se
-dplyr::filter(parameterEstimates(res_differ), lhs == "MACH", op == "~1", group == 2)$pvalue
-
-
-generate_data <- function(n_obs, n_sim, model_truth, model_1, model_2, ...){
-  bic_n <- list()
-  aic_n <- list()
-  p_aov_n <- list()
-  se_n <- list()
[…]

### 2021-04-08 R/lavaan_helper.R
compare three model types, add std

@@ -1,3 +1,5 @@
+source(here::here("R", "standardizedSolution.R"))
+
 items <- function(item, n){
   stringr::str_c(item, seq_len(n))
 }

### 2021-04-08 R/simulation.R

@@ -21,22 +21,35 @@ model_truth <- combine(
 model_same <- 
   "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
   MACH ~ c(0,0)*1
-  MACH ~~ c(1, NA)*MACH"
-# MACH ~~ 1*MACH
-# mean(res_data$dec_p)
-# MACH ~~ c(1, NA)*MACH
-# [1] 0.466
-# mean(res_data$dec_p)
-# [1] 0.464
+  MACH ~~ c(1, 1)*MACH"
 
 model_differ <- 
   "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
   MACH ~ c(0, NA)*1 + c(zero, diff)*1
+  MACH ~~ c(1, 1)*MACH"
+
+model_same2 <- 
+  "MACH =~ 1*x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+  MACH ~ c(0,0)*1
+  MACH ~~ c(NA, NA)*MACH"
+
+model_differ2 <- 
+  "MACH =~ 1*x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+  MACH ~ c(0, NA)*1 + c(zero, diff)*1
+  MACH ~~ c(NA, NA)*MACH"
+
+model_same3 <- 
+  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+  MACH ~ c(0,0)*1
   MACH ~~ c(1, NA)*MACH"
 
+model_differ3 <- 
+  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+  MACH ~ c(0, NA)*1 + c(zero, diff)*1
+  MACH ~~ c(1, NA)*MACH"
 
 #----generate----
-n = 5000
+n = 500000
 data <- lavaan::simulateData(
   model_truth,
   sample.nobs = c(n/2, n/2), 
@@ -78,16 +91,28 @@ get_lrt_p <- function(fits){
   if(!all_converged(fits))return(NA)
   lavaan::anova(fits$same, fits$differ)$`Pr(>Chisq)`[2]
 }
-get_parameter <- function(fits, what, label){
+get_parameter <- function(fits, what, label, how){
   if(!all_converged(fits))return(NA)
-  pars <- parameterestimates(fits$differ)
[…]

### 2021-04-09 R/simulation.R
use future for multicore/hpc

@@ -1,6 +1,16 @@
 source(here::here("R", "lavaan_helper.R"))
 library(tidyverse)
 library(lavaan)
+library(furrr)
+# if you have access to a hpc envir specify this in R/hpc.R
+# if not local multicore is used
+hpc_config <- here::here("R", "hpc.R")
+if(fs::file_exists(hpc_config)){
+  source(hpc_config)
+} else {
+  plan(list(transparent,
+       tweak(multisession, workers = 4L)))
+}
 
 #----jones_paulhus----
 loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
@@ -122,8 +132,21 @@ generate_data_ <- function(n_obs, truth, same, differ, extract_fns, ...){
   as_tibble(map(extract_fns, exec, fits))
 }
 
-generate_data <- function(n_sim, setup){
-  map(seq_len(n_sim), ~mutate(setup, results = pmap(setup, generate_data_)))
+generate_data <- function(n_sim, setup, .furrr_options = furrr_options()){
+  future_map(seq_len(n_sim),
+             ~ mutate(
+               setup,
+               results = future_pmap(setup, generate_data_, .options = .furrr_options)
+             ),
+             .options = .furrr_options)
+}
+
+ls_funs <- function(pos = parent.frame(), ...){
+  everything <- ls(pos = pos, ...)
+  names(everything) <- everything
+  out <- names(keep(map(everything, get), is.function))
+  names(out) <- out
+  out
 }
 
 #----sim----
@@ -134,7 +157,7 @@ extract_fns <- list(bic = get_bic,
                     estimate = get_estimate,
                     std_delta_p = get_std_delta_p,
                     std_estimate = get_std_estimate)
-n_obs <- c(1,5, 100)*100
+n_obs <- c(1:100)*100
 setup <- expand_grid(n_obs = n_obs,
                      truth = model_truth,
                      same = model_same,
@@ -144,18 +167,19 @@ setup <- expand_grid(n_obs = n_obs,
                      auto.fix.first = FALSE)
 setup2 <- mutate(setup, same = model_same2, differ = model_differ2)
 setup3 <- mutate(setup, same = model_same3, differ = model_differ3)
-n_sim <- 2
-res_raw <-
[…]

### 2021-04-09 R/simulation.R
optimize nesting and size of results

@@ -11,6 +11,8 @@ if(fs::file_exists(hpc_config)){
   plan(list(transparent,
        tweak(multisession, workers = 4L)))
 }
+# debug:
+#   plan(transparent)
 
 #----jones_paulhus----
 loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
@@ -59,7 +61,7 @@ model_differ3 <-
   MACH ~~ c(1, NA)*MACH"
 
 #----generate----
-n = 500000
+n = 50000
 data <- lavaan::simulateData(
   model_truth,
   sample.nobs = c(n/2, n/2), 
@@ -107,7 +109,6 @@ get_parameter <- function(fits, what, label, how){
   which <- which(pars$label == label)
   if(length(which) == 0L)stop("No parameter labeled '", label, "'.")
   out <- pars[which, what]
-  if(is.null(out))browser()
   out
 }
 
@@ -133,10 +134,11 @@ generate_data_ <- function(n_obs, truth, same, differ, extract_fns, ...){
 }
 
 generate_data <- function(n_sim, setup, .furrr_options = furrr_options()){
+  without_id <- select(setup, -matches("^id\\d*$"))
   future_map(seq_len(n_sim),
              ~ mutate(
-               setup,
-               results = future_pmap(setup, generate_data_, .options = .furrr_options)
+               select(setup, -extract_fns),
+               results = future_pmap(without_id, generate_data_, .options = .furrr_options)
              ),
              .options = .furrr_options)
 }
@@ -157,6 +159,9 @@ extract_fns <- list(bic = get_bic,
                     estimate = get_estimate,
                     std_delta_p = get_std_delta_p,
                     std_estimate = get_std_estimate)
+#quick test
+#n_obs <- c(100, 10000)
+
 n_obs <- c(1:100)*100
 setup <- expand_grid(n_obs = n_obs,
                      truth = model_truth,
@@ -167,21 +172,33 @@ setup <- expand_grid(n_obs = n_obs,
                      auto.fix.first = FALSE)
 setup2 <- mutate(setup, same = model_same2, differ = model_differ2)
 setup3 <- mutate(setup, same = model_same3, differ = model_differ3)
-n_sim <- 1000
+n_sim <- 100
[…]

### 2021-04-09 R/modind.R
add function to free parameters by modindices

@@ -1,0 +1,44 @@
+stitch <- function(pars){
+  with(pars, str_c(lhs, op, rhs))
+}
+
+competing_models_mod <- function(data, same, differ, ...){
+  pars <- c()
+  res_mod <- set_free_mod(differ, data, pars, op = "=~", ...)
+  res_mod <- set_free_mod(differ, data, res_mod$pars, op = "~1", ...)
+  
+  fits <- list(same = cfa(same, data, group.partial = res_mod$pars, ...),
+               differ = res_mod$differ)
+  
+}
+
+set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...){
+  
+  differ_fit <- cfa(differ, data, group.partial = pars, ...)
+  
+  depth <- depth - 1
+  if(depth < 0){return(list(pars = pars, differ = differ_fit))}
+  
+  parest <- parameterEstimates(differ_fit)
+  testscores <- lavTestScore(differ_fit)
+  sorted_pars <- testscores$uni %>% 
+    rename(label = lhs) %>% 
+    select(-c(op, rhs, df)) %>% 
+    left_join(parest, by = c("label")) %>% 
+    filter(group == 1, .data$op %in% !!op) %>% 
+    arrange(desc(X2))
+  sorted_pars <- sorted_pars[1,]
+  
+  if(pull(sorted_pars, X2) > qchisq(0.95, 1)){
+    pars <- append(pars, stitch(sorted_pars))
+    set_free_mod(differ, data, pars, depth, op = op, ...)
+  }else{
+    return(list(pars = pars, differ = differ_fit))
+  }
+}
+
+res <- competing_models_mod(data, model_same3, model_differ3, 
+                            group.equal = c("loadings", "intercepts"),
+                            auto.fix.first = FALSE,
+                            std.lv = TRUE,
+                            group = "group")

### 2021-04-12 R/simulation.R
revamp plots

@@ -1,4 +1,5 @@
 source(here::here("R", "lavaan_helper.R"))
+source(here::here("R", "modind.R"))
 library(tidyverse)
 library(lavaan)
 library(furrr)
@@ -69,13 +70,13 @@ data <- lavaan::simulateData(
 
 #----analyze----
 res_same <- cfa(
-  model_same, 
+  model_same3, 
   data, group = "group", 
   group.equal = c("loadings", "intercepts"), 
   std.lv= TRUE)
 summary(res_same)
 res_differ <- cfa(
-  model_differ, 
+  model_differ3, 
   data, 
   group = "group", 
   group.equal = c("loadings", "intercepts"), 
@@ -160,7 +161,7 @@ extract_fns <- list(bic = get_bic,
                     std_delta_p = get_std_delta_p,
                     std_estimate = get_std_estimate)
 #quick test
-#n_obs <- c(100, 10000)
+#n_obs <- c(10000)
 
 n_obs <- c(1:100)*100
 setup <- expand_grid(n_obs = n_obs,
@@ -172,7 +173,7 @@ setup <- expand_grid(n_obs = n_obs,
                      auto.fix.first = FALSE)
 setup2 <- mutate(setup, same = model_same2, differ = model_differ2)
 setup3 <- mutate(setup, same = model_same3, differ = model_differ3)
-n_sim <- 100
+n_sim <- 1000
 #quick test
 #n_sim <- 2
 #setup_all <- mutate(setup3, id = "model3")
@@ -193,7 +194,7 @@ res_raw %<-%
                   seed = TRUE,
                   packages = c("furrr", "lavaan", "tidyverse")
                 ))
-res_raw
+write_rds(res_raw, "res_raw.rds")
 
 res <- res_raw  %>%
   map(bind_rows) %>% 
@@ -225,17 +226,11 @@ res %>%
   NULL
 
 res %>%
-  ggplot(aes(x = estimate, color = factor(n_obs))) +
-  geom_density() +
-  facet_wrap(~id, ncol = 1) +
-  theme_minimal()
-
-res %>%
   ggplot(aes(x = std_estimate, group = factor(n_obs), color = log10(n_obs))) +
   geom_density() +
   facet_wrap(~id, ncol = 1) +
   theme_minimal() +
-  scale_color_viridis_c(option = "magma")
+  scale_color_viridis_c()
   
 
[…]

### 2021-04-12 R/simulation.R
only use model3, remove others

@@ -34,29 +34,9 @@ model_truth <- combine(
 model_same <- 
   "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
   MACH ~ c(0,0)*1
-  MACH ~~ c(1, 1)*MACH"
-
-model_differ <- 
-  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0, NA)*1 + c(zero, diff)*1
-  MACH ~~ c(1, 1)*MACH"
-
-model_same2 <- 
-  "MACH =~ 1*x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0,0)*1
-  MACH ~~ c(NA, NA)*MACH"
-
-model_differ2 <- 
-  "MACH =~ 1*x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0, NA)*1 + c(zero, diff)*1
-  MACH ~~ c(NA, NA)*MACH"
-
-model_same3 <- 
-  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0,0)*1
   MACH ~~ c(1, NA)*MACH"
 
-model_differ3 <- 
+model_differ <- 
   "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
   MACH ~ c(0, NA)*1 + c(zero, diff)*1
   MACH ~~ c(1, NA)*MACH"
@@ -163,7 +143,7 @@ extract_fns <- list(bic = get_bic,
 #quick test
 #n_obs <- c(10000)
 
-n_obs <- c(1:100)*100
+n_obs <- seq(100, 10000, 100)
 setup <- expand_grid(n_obs = n_obs,
                      truth = model_truth,
                      same = model_same,
@@ -171,24 +151,18 @@ setup <- expand_grid(n_obs = n_obs,
                      extract_fns = list(extract_fns),
                      group.equal = list(c("loadings", "intercepts")),
                      auto.fix.first = FALSE)
-setup2 <- mutate(setup, same = model_same2, differ = model_differ2)
-setup3 <- mutate(setup, same = model_same3, differ = model_differ3)
-n_sim <- 1000
+n_sim <- 100
 #quick test
 #n_sim <- 2
[…]

### 2021-04-12 R/modind.R
get modind to work

@@ -2,20 +2,32 @@ stitch <- function(pars){
   with(pars, str_c(lhs, op, rhs))
 }
 
-competing_models_mod <- function(data, same, differ, ...){
-  pars <- c()
-  res_mod <- set_free_mod(differ, data, pars, op = "=~", ...)
-  res_mod <- set_free_mod(differ, data, res_mod$pars, op = "~1", ...)
-  
-  fits <- list(same = cfa(same, data, group.partial = res_mod$pars, ...),
+filter_dots <- function(..., fun, what = names(formals(fun))){
+  dots <- list(...)
+  if(any(names(dots) == ""))stop("filter_dots does not allow unnamed args!")
+  exists <- setdiff(what, setdiff(what, names(dots)))
+  dots[exists]
+}
+
+safe_cfa <- function(...){
+  valid_args <-
+    c(names(formals(lavaan::lavaanify)),
+      names(formals(lavaan::lavaan)),
+      names(lavOptions()))
+  dots <- filter_dots(..., fun = lavaan::cfa, what = valid_args)
+  do.call("cfa", dots, envir = getNamespace("lavaan"))
+}
+
+competing_models_mod <- function(data, same, differ, mod_load = 0L, mod_int = 0L, pars = "", ...){
+  res_mod <- set_free_mod(differ, data, pars, op = "=~", depth = mod_load, ...)
+  res_mod <- set_free_mod(differ, data, res_mod$pars, op = "~1", depth = mod_int, ...)
+  fits <- list(same = safe_cfa(model = same, data = data, group.partial = res_mod$pars, ...),
                differ = res_mod$differ)
   
 }
 
 set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...){
-  
-  differ_fit <- cfa(differ, data, group.partial = pars, ...)
-  
+  differ_fit <- safe_cfa(model = differ, data = data, group.partial = pars, ...)
   depth <- depth - 1
   if(depth < 0){return(list(pars = pars, differ = differ_fit))}
   
@@ -36,9 +48,3 @@ set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...)
     return(list(pars = pars, differ = differ_fit))
   }
 }
-
-res <- competing_models_mod(data, model_same3, model_differ3, 
-                            group.equal = c("loadings", "intercepts"),
-                            auto.fix.first = FALSE,
-                            std.lv = TRUE,
-                            group = "group")

### 2021-04-12 R/simulation.R

@@ -50,13 +50,13 @@ data <- lavaan::simulateData(
 
 #----analyze----
 res_same <- cfa(
-  model_same3, 
+  model_same, 
   data, group = "group", 
   group.equal = c("loadings", "intercepts"), 
   std.lv= TRUE)
 summary(res_same)
 res_differ <- cfa(
-  model_differ3, 
+  model_differ, 
   data, 
   group = "group", 
   group.equal = c("loadings", "intercepts"), 
@@ -110,7 +110,7 @@ generate_data_ <- function(n_obs, truth, same, differ, extract_fns, ...){
     truth,
     sample.nobs = rep(n_obs, 2), 
     standardized = TRUE)
-  fits <- competing_models(data, same, differ, group = "group", ...)
+  fits <- competing_models_mod(data, same, differ, group = "group", ...)
   as_tibble(map(extract_fns, exec, fits))
 }
 
@@ -148,10 +148,12 @@ setup <- expand_grid(n_obs = n_obs,
                      truth = model_truth,
                      same = model_same,
                      differ = model_differ,
+                     mod_load = c(0, 3),
+                     mod_int = c(0, 3),
                      extract_fns = list(extract_fns),
                      group.equal = list(c("loadings", "intercepts")),
                      auto.fix.first = FALSE)
-n_sim <- 100
+n_sim <- 1000
 #quick test
 #n_sim <- 2
 
@@ -173,7 +175,6 @@ write_rds(res_raw, "res_raw.rds")
 res <- res_raw  %>%
   map(bind_rows) %>% 
   bind_rows() %>% 
-  select(n_obs, results) %>% 
   unnest(results)
 
 res <- mutate(
@@ -193,9 +194,9 @@ interval <- function(x, alpha = c(0.05)){
 }
 
 res_interval <- res %>%
-  select(n_obs, std_estimate) %>% 
+  select(n_obs, std_estimate, mod_load, mod_int) %>% 
   filter(!is.na(std_estimate)) %>% 
-  group_by(n_obs) %>% 
+  group_by(n_obs, mod_load, mod_int) %>% 
   summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
             .groups = "drop") %>% 
   unnest(c(interval)) %>% 
@@ -211,6 +212,8 @@ plot_interval <- res_interval %>%
 plot_interval
 ggsave("interval.png")
 
[…]

### 2021-04-12 preregistration.Rmd
rearrange simulation files

---
title: "Preregistration"
author: "Aaron Peikert"
date: "4/12/2021"
output: html_document
params:
  nsim: 16
  nmin: 100
  nmax: 10000
  nstep: 1000
  seed: 1234
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lavaan)
library(furrr)
source(here::here("R", "simulation.R"))
# if you have access to a hpc envir specify this in R/hpc.R
# if not we use local multicore, consider reduce nsim; increase nstep in ↑YAML↑

hpc_config <- here::here("R", "hpc.R")
if(fs::file_exists(hpc_config)){
  source(hpc_config)
} else {
  plan(list(transparent,
       tweak(multisession, workers = 4L)))
}
# debug:
#   plan(transparent)
```

```{r}
loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
cohend_jones_paulhus <- c(24, 29, 35)/100
```

```{r models}
model_truth <- combine(
  measurement(
    "MACH",
    items("x", 9),
    loadings = round(loadings_jones_paulhus * 0.7, 2)
  ),
  intercepts("MACH", list(0, 0.2)),
  variances("MACH", list(1, 1))
)

model_same <- 
  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
  MACH ~ c(0,0)*1
  MACH ~~ c(1, NA)*MACH"

model_differ <- 
  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
  MACH ~ c(0, NA)*1 + c(zero, diff)*1
  MACH ~~ c(1, NA)*MACH"
```

```{r sim}
extract_fns <- list(bic = get_bic, 
                    aic = get_aic,
                    lrt_p = get_lrt_p,
                    delta_p = get_delta_p,
                    estimate = get_estimate,
                    std_delta_p = get_std_delta_p,
                    std_estimate = get_std_estimate)

n_obs <- seq(params$nmin, params$nmax, params$nstep)
n_sim <- params$nsim

setup <- expand_grid(n_obs = n_obs,
                     truth = model_truth,
                     same = model_same,
                     differ = model_differ,
                     mod_load = c(3),
                     mod_int = c(3),
                     extract_fns = list(extract_fns),
                     group.equal = list(c("loadings", "intercepts")),
                     auto.fix.first = FALSE)
to_export <- ls_funs() %>% map(get)
to_export <- c(list(setup = setup), to_export)
RNGkind("L'Ecuyer-CMRG")
set.seed(params$seed)
res_raw %<-% 
  generate_data(n_sim,
                setup,
                furrr_options(
                  globals = to_export,
                  seed = TRUE,
                  packages = c("furrr", "lavaan", "tidyverse")
                ))
```

```{r write-out, include=FALSE}
invisible(res_raw)
write_rds(res_raw, here::here("res_raw.rds"))
```

```{r checksum}
# compute checksum using md5
checksum <- digest::digest(res_raw, "md5")
if(checksum != "856fed767a7e0c3efcb5ac76359f7707"){
  warning("Mismatch between original and current simulations!")
}
```

```{r}
res <- res_raw  %>%
  map(bind_rows) %>% 
  bind_rows() %>% 
  unnest(results)

res <- mutate(
  res,
  dec_bic = bic < 0,
  dec_aic = aic < 0,
  dec_aic2 = aic < -2,
  dec_lrt_p = lrt_p < 0.05,
  dec_delta_p = delta_p < 0.05)

res_interval <- res %>%
  select(n_obs, std_estimate, mod_load, mod_int) %>% 
  filter(!is.na(std_estimate)) %>% 
  group_by(n_obs, mod_load, mod_int) %>% 
  summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
            .groups = "drop") %>% 
  unnest(c(interval)) %>% 
  mutate(width = upper - lower)
```

```{r}
plot_interval <- res_interval %>% 
  ggplot(aes(n_obs, ymin = lower, ymax = upper, fill = alpha, group = alpha)) + 
  geom_ribbon() +
  scale_fill_viridis_c() +
  scale_y_continuous(breaks = seq(-0.2, 0.7, .1)) +
  theme_minimal()

plot_interval
```

```{r}
plot_width <- res_interval %>% 
  ggplot(aes(n_obs, width, color = alpha, group = alpha)) +
  geom_line() +
  scale_color_viridis_c() +
  theme_minimal() +
  scale_y_log10(breaks = seq(0, 0.8, .1)) +
  theme(axis.text.x = element_text(angle = -90)) +
  NULL

plot_width
```

### 2021-04-12 R/lavaan_helper.R

@@ -1,5 +1,9 @@
 source(here::here("R", "standardizedSolution.R"))
 
+stitch <- function(pars){
+  with(pars, str_c(lhs, op, rhs))
+}
+
 items <- function(item, n){
   stringr::str_c(item, seq_len(n))
 }

### 2021-04-12 R/modind.R

@@ -1,50 +1,0 @@
-stitch <- function(pars){
-  with(pars, str_c(lhs, op, rhs))
-}
-
-filter_dots <- function(..., fun, what = names(formals(fun))){
-  dots <- list(...)
-  if(any(names(dots) == ""))stop("filter_dots does not allow unnamed args!")
-  exists <- setdiff(what, setdiff(what, names(dots)))
-  dots[exists]
-}
-
-safe_cfa <- function(...){
-  valid_args <-
-    c(names(formals(lavaan::lavaanify)),
-      names(formals(lavaan::lavaan)),
-      names(lavOptions()))
-  dots <- filter_dots(..., fun = lavaan::cfa, what = valid_args)
-  do.call("cfa", dots, envir = getNamespace("lavaan"))
-}
-
-competing_models_mod <- function(data, same, differ, mod_load = 0L, mod_int = 0L, pars = "", ...){
-  res_mod <- set_free_mod(differ, data, pars, op = "=~", depth = mod_load, ...)
-  res_mod <- set_free_mod(differ, data, res_mod$pars, op = "~1", depth = mod_int, ...)
-  fits <- list(same = safe_cfa(model = same, data = data, group.partial = res_mod$pars, ...),
-               differ = res_mod$differ)
-  
-}
-
-set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...){
-  differ_fit <- safe_cfa(model = differ, data = data, group.partial = pars, ...)
-  depth <- depth - 1
-  if(depth < 0){return(list(pars = pars, differ = differ_fit))}
-  
-  parest <- parameterEstimates(differ_fit)
-  testscores <- lavTestScore(differ_fit)
-  sorted_pars <- testscores$uni %>% 
-    rename(label = lhs) %>% 
-    select(-c(op, rhs, df)) %>% 
-    left_join(parest, by = c("label")) %>% 
-    filter(group == 1, .data$op %in% !!op) %>% 
-    arrange(desc(X2))
-  sorted_pars <- sorted_pars[1,]
-  
-  if(pull(sorted_pars, X2) > qchisq(0.95, 1)){
-    pars <- append(pars, stitch(sorted_pars))
-    set_free_mod(differ, data, pars, depth, op = op, ...)
-  }else{
-    return(list(pars = pars, differ = differ_fit))
-  }
-}

### 2021-04-12 R/simulation.R

@@ -1,72 +1,56 @@
 source(here::here("R", "lavaan_helper.R"))
-source(here::here("R", "modind.R"))
-library(tidyverse)
-library(lavaan)
-library(furrr)
-# if you have access to a hpc envir specify this in R/hpc.R
-# if not local multicore is used
-hpc_config <- here::here("R", "hpc.R")
-if(fs::file_exists(hpc_config)){
-  source(hpc_config)
-} else {
-  plan(list(transparent,
-       tweak(multisession, workers = 4L)))
-}
-# debug:
-#   plan(transparent)
-
-#----jones_paulhus----
-loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
-cohend_jones_paulhus <- c(24, 29, 35)/100
-
-#----model-truth----
-model_truth <- combine(
-  measurement(
-    "MACH",
-    items("x", 9),
-    loadings = round(loadings_jones_paulhus * 0.7, 2)
-  ),
-  intercepts("MACH", list(0, 0.2)),
-  variances("MACH", list(1, 1))
-)
-
-#----models----
-model_same <- 
-  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0,0)*1
-  MACH ~~ c(1, NA)*MACH"
-
-model_differ <- 
-  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0, NA)*1 + c(zero, diff)*1
-  MACH ~~ c(1, NA)*MACH"
-
-#----generate----
-n = 50000
-data <- lavaan::simulateData(
-  model_truth,
-  sample.nobs = c(n/2, n/2), 
-  standardized = TRUE)
-
-#----analyze----
-res_same <- cfa(
-  model_same, 
-  data, group = "group", 
-  group.equal = c("loadings", "intercepts"), 
-  std.lv= TRUE)
-summary(res_same)
-res_differ <- cfa(
-  model_differ, 
-  data, 
-  group = "group", 
-  group.equal = c("loadings", "intercepts"), 
-  std.lv= TRUE)
-summary(res_differ)
 
 #----sim-functions----
 competing_models <- function(data, same, differ, ...){
   list(same = cfa(same, data, ...),
        differ = cfa(differ, data, ...))
+}
+
+filter_dots <- function(..., fun, what = names(formals(fun))){
[…]

### 2021-04-12 preregistration.Rmd
fix seed

@@
                   globals = to_export,
-                  seed = TRUE,
+                  seed = params$seed,
                   packages = c("furrr", "lavaan", "tidyverse")
@@
 checksum <- digest::digest(res_raw, "md5")
-if(checksum != "856fed767a7e0c3efcb5ac76359f7707"){
+if(checksum != "e3b116b36c2a267d0e848e692b76fc1a"){
   warning("Mismatch between original and current simulations!")

### 2021-04-12 R/simulation.R

@@ -99,6 +99,7 @@ generate_data_setup_ <- function(n_obs, truth, same, differ, extract_fns, ...){
 }
 
 generate_data_setup <- function(setup, .furrr_options = furrr_options()){
+  .furrr_options$seed <- TRUE
   mutate(select(setup, -extract_fns),
     results = future_pmap(setup, generate_data_setup_, .options = .furrr_options))
 }

### 2021-04-12 preregistration.Rmd
show checksum

@@
 params:
-  nsim: 16
+  nsim: 1000
   nmin: 100
   nmax: 10000
-  nstep: 1000
+  nstep: 100
   seed: 1234
@@
 if(checksum != "e3b116b36c2a267d0e848e692b76fc1a"){
-  warning("Mismatch between original and current simulations!")
+  warning("Mismatch between original and current simulations!
+          Checksum now is:\n    ", checksum)
 }

### 2021-04-12 manuscript.Rmd

--- 2021-03-05 manuscript.Rmd
@@
     - rstudio/webshot2@f62e743
+  scripts:
+    - R/lavaan_helper.R
+    - R/simulation.R
+    - R/link.R
 csl: apa7.csl
 numbered_headings: true
-toc_depth: 5
-bibliography: references.bib
+bibliography: ["references.bib", "packages.bib"]
 ---
@@
 
+```{r, include=FALSE}
+source(here::here("R", "simulation.R"))
+```
+
 There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
 A step to prevent unpleasant surprises is to simulate data.
-Simulated data lets you verify that the data you expect to gather and the analysis that you plan fit together.
+Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
+Of course, one has to make many educated guesses to simulate data that closely matches your expectation.
+These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
+However, the analytic strategy often gives you a hint about how to simulate data.
+In the Machiavellism example, the analytic strategy is to use structural equation modelling (SEM).
+To simulate data for SEM, one has to specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
+The model parameters should mimic the Machiavellism subscale of the short dark triad (SD3).
+As a starting point, we can use already published information.
+@Jones2013 report loadings seen in Table \@ref(tab:loadings) and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
+
+```{r loadings, echo=FALSE}
+knitr::kable(data.frame(Item = 1:9,
+                        Loading = loadings_jones_paulhus),
+             caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
+```
+
+It is probably better to assume more pessimistic values.
+This way, we make sure that our analytic strategy is viable even in non-optimal circumstances.
+Therefore, we will assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.1.
+These assumption can be simply translated into [`lavaan` [@lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
+
+[^lavaan]: https://lavaan.ugent.be
+
+```{r, results='asis', echo=FALSE}
+model_truth <- combine(
+  measurement(
+    "MACH",
+    items("x", 9),
+    loadings = round(loadings_jones_paulhus * 0.7, 2)
+    ),
+  intercepts("MACH", list(0, 0.1))
+  )
+cat("``` r",
+    model_truth,
+    "```",
+    sep = "\n")
+model <- combine(measurement(
+  "MACH",
+  items("x", 9)),
+  intercepts("MACH", list(0, "NA")))
+```
+
+From which we can simulate data:
+
+```{r, results='hide'}
+simulated <- lavaan::simulateData(
+  model_truth,
+  sample.nobs = c(1000, 1000),
+  meanstructure = TRUE,
+  standardized = TRUE
+)
+```
+
+And then fit the statistical model:
+
+```{r}
+fit <-
+  lavaan::cfa(
+    model,
+    data = simulated,
+    group = "group",
+    meanstructure = TRUE,
+    std.lv = TRUE
+  ) 
+```
+
+But wait `lavaan` warns us that this model may not be identified.
+We overlooked a crucial assumption of this model, that is measurement invariance.
+This assumption is also baked into our simulation of the data.
+It is probably more realistic to violate this assumption and plan our analysis accordingly.
+Therefore, we simulate the data so that some of the items parameter differ among groups and then alter the analysis to assume partial measurement invariance.
+While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
+Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
+@Rhemtulla2012 conclude that using Likert-like items is acceptable when they employ five or more categories.
+Hence, we continue simulating continues data.
+However, adding a little skewness to the items is easy and helps make the simulation more realistic.
+Anyhow, remember that what we consider realistic for our case could be unfit for your case.
+If you are satisfied with your assumptions it helps to write a function that expresses them, like this one:
+<!--Hands-on: build simple functions tailored for your analysis-->
+
+```{r generate}
+```
+
+Note however that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/R/simulation.R", "GitHub")`.
+
+With its help we can simulate 500 cases with a mean difference between groups of 0.1, where the intercepts of three random items vary by .3 and three loadings vary by 30% and five random items have a skew.
+
+```{r}
+simulated <- generate(n = 500,
+                      diff = 0.1,
+                      vary_intercepts = rep(0.3, 3),
+                      vary_loadings = rep(0.3, 3),
+                      skewness = rep(5, -0.5))
+```
+
+Note, that this certainly results in different numbers each time we run it.
+To consistently get the same results it is good practice to use a seed.
+The function `seeded_generate()` will always generate the same pseudo-random numbers, when given the same seed.
+
+```{r}
+seeded_generate <-  function(seed, ...)withr::with_seed(seed, generate(...))
+simulated <- seeded_generate(
+  seed = 1234,
+  n = 100000,
+  diff = 0.1,
+  vary_intercepts = rep(0.3, 3),
+  vary_loadings = rep(0.3, 3),
+  skewness = rep(5,-0.5)
+)
+```
+
+```{r}
+lavaan::summary(analyze(simulated))
+```
+
+### Preregestration (optional)
 

### 2021-04-13 manuscript.Rmd
change wording here and there

@@
 <!-- expand: who is calling it a requirement, maybe metascientific arguments-->
-Reproducing an empirical study is often tedious and frustrating work for the researcher.
+Reproducing an empirical study is often too tedious and frustrating work for the researchers to benefit them.
 However, if reproduction would require mere minutes instead of hours, it could greatly facilitate collaboration within and across research projects.
@@
 Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you archive reproducibility gradually.
-The first step is to organize all files in one folder.
+The first step is to bundle all required files in one folder.
 This way, you never lose anything, and you can easily share the files.
@@
 
-## Preregestration (optional)
-
 A preregistration is a tool that may help to increase the credibility of empirical results.
@@
 As a researcher, you are eligible for the `r link("https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount", "Researcher/Educator discount")`.
-The following command creates a new GitHub repository under your GitHub account, beware this is public:
+The following command creates a new GitHub repository under your GitHub account (beware this is public):
 <!--Hands-on: Introduce Github-->
@@
 
+Only from this moment on, files are uploaded to the internet and only those which you have explicitly added and committed in Git.
+
 <!--Hands on: Introduce GitHub Issues-->
@@
 
-While you already know how to write dynamic documents with [Markdown] and [R Markdown], and track changes with [Git] and [GitHub], two components to archive longterm reproducibility are still missing.
+While you already know how to write dynamic documents with [Markdown] and [R Markdown], and track changes with [Git] and [GitHub], two components to archive longterm reproducibility are still missing, namely internal and external dependencies.
 
@@
 
-First dependency management, which means that you capture the interrelations between the components of you analysis.
+First dependency management, which means that you capture the interrelations between the components of your analysis.
 You cannot run an analysis without data, you cannot render a dynamic document without an analysis, and so forth.
 Understanding these relations is often key to reproducing a project.
-A reader of your reproducible manuscript should be able to derive how they are to reproduce it (ideally by a statement in your [README])
+A reader of your reproducible manuscript should be able to derive how they are to reproduce it (ideally by a statement in your [README]).
 The "components" of an analysis are often loosely mapped onto files.

### 2021-04-13 manuscript.Rmd
add some rambling about internal dependencies

@@
 library(repro)
+library(tidyverse)
 automate_load_scripts()
@@
 Therefore, we will assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.1.
-These assumption can be simply translated into [`lavaan` [@lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
+These assumption can be simply translated into [`lavaan` [@R-lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
 
@@
 For a simple project it might be enough to have a single R Markdown.
-However, while it is easy to reproduce this approach quickly gets unwieldy.
+However, while it is easy to understand how to reproduce the end result, this approach quickly gets unwieldy.
+To better navigate the data analysis, a researcher often splits it across several scripts.
+For example, they create a script to load the data and another to run the analysis.
+This modularisation facilitates reuse, because the loading script may be used by other projects building on the same data, while the analysis script may be employed on other data and so forth.
+The downside is that reproducibility is complicated, because instead of reproducing everything in one step, multiple steps are required.
+
+Of course there a several ways to solve this problem.
+One could build an R package to separate analysis and data or use code externalization in the RMarkdown to facilitating reuse or build a "master" script that combines all steps.
+We found Make to be particularly useful for two reasons.
+First, it is language agnostic, therefore fusing steps that are carried out in different programming languages.
+Second, it does implement a simple, yet powerful caching mechanism of results which is welcome in computing intensive analysis often found in neuroimaging or machine learning.
+
+Make is build around the concept of computational recipes.
+Like cooking recipes, the dish you want to create is at the top, followed by the necessary ingredients with the procedure at last.
+In Make the dish is usually a file and is called "target".
+The ingredients are called "requirements" and are usually also files.
+The procedures are formulated as shell commands.
+One (hypothetical) example is this:
+
+```
+spaghetti_arrabiata.pdf: spaghetti_arrabiata.Rmd arrabiata_sauce.csv pasta.csv
+  Rscript -e 'rmarkdown::render("spagetti_arrabiata.Rmd")'
+```
+
+An important property is that each requirement can be itself be a target, like here:
+
+```
+spaghetti_arrabiata.pdf: spagetti_arrabiata.Rmd arrabiata_sauce.csv pasta.csv
+  Rscript -e 'rmarkdown::render("spaghetti_arrabiata.Rmd")'
+
+pasta.csv: cook_pasta.R
+  Rscript -e 'source("cook_pasta.R")'
+
+arrabiata_sauce.csv: cook_sauce.R canned_tomatoes.csv
+  Rscript -e 'source("cook_sauce.R")'
+```
+
+Lets suppose you want to eat spaghetti arrabiata and you still have delicious arrabiata sauce in the fridge.
+Would you cook the sauce again?
+No, you would only cook the pasta.
+Make behaves the same.
+If a file already exists it does not recreate it unnecessarily.
+However, if a requirement is newer than its target, make recreates the target to reflect the updated dependency.
+Therefore, when you update the data file, the analysis is run again.
+Because Make knows the whole dependency tree, its caching is smarter than many alternatives.
+
+One alternative that is even smarter than Make is the R package [`targets` [@targets2021]](https://books.ropensci.org/targets/)[^targets].
+`targets` can infer more about your project than Make, because it analysis the R code you write to build the dependency tree.
+For this inference to work well you have to write everything in R and you have to write everything strictly functional.
+
+[^targets]: https://books.ropensci.org/targets/
 
@@
 <!--Hands on: Add targets for README.Rmd, preregistration.Rmd -->
+
+The `repro` package can automatically write Makefiles for simple scenarios.
+For this purpose, `repro` utilizes the YAML metadata:
+
+```{r}
+x <- readLines(here::here("preregistration.Rmd"))
+yaml_ind <- stringr::str_which(x, "^---[[:space:]]*")
+stripped <- stringr::str_c(x[seq(yaml_ind[[1]], yaml_ind[[2]])],
+                  collapse = "\n")
+cat("```\n", stripped, "\n```", sep = "")
+```
 
@@
 # References
+
+```{r, include=FALSE}
+knitr::write_bib(c(.packages(), "bookdown", "lavaan", "knitr", "targets"), "packages.bib")
+```

### 2021-04-14 R/simulation.R
safetify simulation

@@ -30,11 +30,14 @@ competing_models_mod <- function(data, same, differ, mod_load = 0L, mod_int = 0L
   
 }
 
+safe_competing_models <- safely(competing_models_mod,
+                                otherwise = list(same = NA, differ = NA))
+
 set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...){
   differ_fit <- safe_cfa(model = differ, data = data, group.partial = pars, ...)
   depth <- depth - 1
   if(depth < 0){return(list(pars = pars, differ = differ_fit))}
-  
+  if(!is_converged(differ_fit)){return(list(pars = pars, differ = differ_fit))}
   parest <- parameterEstimates(differ_fit)
   testscores <- lavTestScore(differ_fit)
   sorted_pars <- testscores$uni %>% 
@@ -53,7 +56,14 @@ set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...)
   }
 }
 
-is_converged <- function(fit)lavInspect(fit, "converged")
+safe_set_free_mod <- safely(set_free_mod,
+                            otherwise = list(same = NA, differ = NA))
+
+is_converged_ <- function(fit)lavInspect(fit, "converged")
+is_converged <- function(fit){
+  if(inherits(fit, "lavaan"))return(is_converged_(fit))
+  else return(FALSE)
+}
 all_converged <- function(fits)all(purrr::map_lgl(fits, is_converged))
 
 get_fit <- function(fits, measure){
@@ -89,24 +99,28 @@ get_std_delta_p <- function(fits, label = "diff"){
 get_std_estimate <- function(fits, label = "diff"){
   get_parameter(fits, "est.std", label, how = standardizedSolution)
 }
-generate_data_setup_ <- function(n_obs, truth, same, differ, extract_fns, ...){
+generate_data <- function(n_obs, truth, same, differ, extract_fns, ...){
   data <- lavaan::simulateData(
     truth,
     sample.nobs = rep(n_obs, 2), 
     standardized = TRUE)
   fits <- competing_models_mod(data, same, differ, group = "group", ...)
-  as_tibble(map(extract_fns, exec, fits))
+  save_fns <- map(extract_fns, purrr::safely, otherwise = NA)
+  results <- map(save_fns, exec, fits)
+  as_tibble(map(results, "result"))
 }
 
-generate_data_setup <- function(setup, .furrr_options = furrr_options()){
[…]

### 2021-04-14 preregistration.Rmd
add smoothed plots

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, error = TRUE)
 
@@
 library(furrr)
+library(slider)
 source(here::here("R", "simulation.R"))
 # if you have access to a hpc envir specify this in R/hpc.R
-# if not we use local multicore, consider reduce nsim; increase nstep in ↑YAML↑
+# if not we use local multicore
+# to speed up consider reduce nsim; increase nstep in ↑YAML↑
 
@@
 set.seed(params$seed)
-res_raw %<-% 
-  generate_data(n_sim,
+res_raw %<-% generate_data_setup(n_sim,
                 setup,
@@
 
-```{r write-out, include=FALSE}
-invisible(res_raw)
+```{r write-out, error=TRUE, include=FALSE}
 write_rds(res_raw, here::here("res_raw.rds"))
@@
 checksum <- digest::digest(res_raw, "md5")
-if(checksum != "e3b116b36c2a267d0e848e692b76fc1a"){
+if(checksum != "4203854848ad0cc5d2f848c44b70815d"){
   warning("Mismatch between original and current simulations!
-          Checksum now is:\n    ", checksum)
+Checksum now is:\n    '", checksum, "'")
 }
@@
   bind_rows() %>% 
+  mutate(results = map(results, "result")) %>% 
   unnest(results)
@@
   unnest(c(interval)) %>% 
-  mutate(width = upper - lower)
+  mutate(width = upper - lower) %>% 
+  group_by(alpha) %>% 
+  arrange(n_obs) %>% 
+  mutate(across(c(width, upper, lower), list(smooth = ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE)))) %>% 
+  ungroup()
 ```
@@
 plot_interval
+
+plot_interval %+% aes(ymin = lower_smooth, ymax = upper_smooth)
 ```
@@
   NULL
-
 plot_width
+plot_width %+% aes(y = width_smooth)
 ```
+```{r}
+res_interval %>% 
+  ggplot(aes(n_obs, width, color = alpha, group = alpha)) +
+  geom_smooth(se = FALSE, span = .1, size = .5) +
+  scale_color_viridis_c() +
+  theme_minimal() +
+  scale_y_log10(breaks = seq(0, 0.8, .1)) +
+  theme(axis.text.x = element_text(angle = -90)) +
+  NULL
+```

### 2021-04-14 manuscript.Rmd
add rambling about preregistration

@@
 
-A preregistration is a tool that may help to increase the credibility of empirical results.
+A preregistration serves as a tool that may help to increase the credibility of empirical results.
+Without getting lost in the meta scientific arguments for preregistration, we want to note two advantages.
+First, a preregistration is an opportunity to carefully plan a study.
+Gathering empirical data often requires enormous resources and better planning increases the chances that these efforts are worthwhile.
+For us a good preregistration is the same as a good plan; the only difference is that a preregistration is made public beforehand.
+Some [citation!] argue that preregistration can not fulfill its promise to reduce the false discovery rate of science, if we allow to deviate from them.
+While such argument has some merits, we believe another advantage of preregistration is transparency.
+If researchers deviate from their preregistration, they probably have good reason to do so, which means they have learned something from the process of conducting the study.
+A preregistration, together with version control allows their readers to inform themselfs about these lessons learned.
 

### 2021-04-14 preregistration.Rmd
add classic power plot

@@
 library(slider)
+library(patchwork)
 source(here::here("R", "simulation.R"))
@@
   summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
-            .groups = "drop") %>% 
+            .groups = "drop", median = median(std_estimate)) %>% 
   unnest(c(interval)) %>% 
-  mutate(width = upper - lower) %>% 
+  mutate(width = upper - lower, perc = 1 - alpha) %>% 
   group_by(alpha) %>% 
   arrange(n_obs) %>% 
-  mutate(across(c(width, upper, lower), list(smooth = ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE)))) %>% 
-  ungroup()
+  mutate(across(c(width, upper, lower),
+                ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE))) %>% 
+  ungroup() %>% 
+  arrange(perc)
+```
+
+```{r}
+minn <- filter(res_interval, perc == .8, width < .2) %>% 
+  filter(n_obs == min(n_obs))
 ```
@@
 plot_interval <- res_interval %>%
-  ggplot(aes(n_obs, ymin = lower, ymax = upper, fill = alpha, group = alpha)) + 
+  ggplot(aes(
+    n_obs,
+    ymin = lower,
+    ymax = upper,
+    fill = perc,
+    group = alpha
+  )) +
   geom_ribbon() +
-  scale_fill_viridis_c() +
-  scale_y_continuous(breaks = seq(-0.2, 0.7, .1)) +
-  theme_minimal()
-
-plot_interval
-
-plot_interval %+% aes(ymin = lower_smooth, ymax = upper_smooth)
+  geom_line(aes(y = median), color = "white") +
+  geom_hline(yintercept = .2, linetype = "dotted") +
+  scale_fill_viridis_c(
+    begin = 1,
+    end = 0,
+    breaks  = 1 - seq(0.05, .2, 0.05),
+    labels = . %>% scales::percent(accuracy = 1)
+  ) +
+  scale_y_continuous(n.breaks = 10) +
+  scale_x_continuous(breaks = seq(0, params$nmax, params$nstep * 10)) +
+  theme_minimal() +
+  labs(x = "N per Group", y = "Standardized Effect Size") +
+  theme(legend.title = element_blank()) +
+  NULL
 ```
@@
 plot_width <- res_interval %>%
-  ggplot(aes(n_obs, width, color = alpha, group = alpha)) +
+  ggplot(aes(n_obs, width, color = perc, group = perc)) +
+  geom_line() +
+  geom_point(
+    data = minn,
+    size = 3,
+    color = "black",
+    shape = 3
+  ) +
+  scale_color_viridis_c(
+    begin = 1,
+    end = 0,
+    breaks = 1 - seq(0.05, .2, 0.05),
+    labels = . %>% scales::percent(accuracy = 1)
+  ) +
+  theme_minimal() +
+  scale_y_log10(breaks = seq(0, 0.8, .1)) +
+  scale_x_continuous(breaks = seq(0, params$nmax, params$nstep * 10)) +
+  labs(x = "N per Group", y = "Width of quantile\n(Standardized Effect Size)") +
+  theme(legend.title = element_blank()) +
+  NULL
+```
+
+```{r}
+plot_interval +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + 
+  plot_width +
+  plot_layout(nrow = 2, guides = 'collect') +
+  plot_annotation(tag_levels = 'A')
+ggsave("plot.svg", device = svglite::svglite, width = 7, height = 4)
+```
+
+Plot A depicts the range of effectsizes we observed in the simulation, where the color encodes the density.
+The light yellow interval shows where 80% of the estimates land.
+The dotted line represents the simulated true effect size; the white line shows the median of observed effect sizes.
+
+Plot B represents the width of the interval shown in Plot A on a logarithmic scale.
+The cross shows the minimal required sample size so that 80% of the estimates land in a range of +/- .1.
+
+```{r}
+alpha <- seq(0.05, .2, 0.01)
+res %>% 
+  group_by(n_obs) %>% 
+  summarise(power = list(tibble(alpha = alpha,
+            power = map_dbl(alpha/2, ~mean(lrt_p < .x, na.rm = TRUE))))) %>% 
+  unnest(power) %>% 
+  ggplot(aes(n_obs, power, color = alpha, group = alpha)) +
   geom_line() +
   scale_color_viridis_c() +
-  theme_minimal() +
-  scale_y_log10(breaks = seq(0, 0.8, .1)) +
-  theme(axis.text.x = element_text(angle = -90)) +
-  NULL
-plot_width
-plot_width %+% aes(y = width_smooth)
+  theme_minimal()
 ```
-```{r}
-res_interval %>% 
-  ggplot(aes(n_obs, width, color = alpha, group = alpha)) +
-  geom_smooth(se = FALSE, span = .1, size = .5) +
-  scale_color_viridis_c() +
-  theme_minimal() +
-  scale_y_log10(breaks = seq(0, 0.8, .1)) +
-  theme(axis.text.x = element_text(angle = -90)) +
-  NULL
-```

### 2021-04-14 preregistration.Rmd
Revert "add classic power plot"
This reverts commit 7472b19564739c8b49785d141a24c96deec2cfcb.

@@
 The cross shows the minimal required sample size so that 80% of the estimates land in a range of +/- .1.
-
-```{r}
-alpha <- seq(0.05, .2, 0.01)
-res %>% 
-  group_by(n_obs) %>% 
-  summarise(power = list(tibble(alpha = alpha,
-            power = map_dbl(alpha/2, ~mean(lrt_p < .x, na.rm = TRUE))))) %>% 
-  unnest(power) %>% 
-  ggplot(aes(n_obs, power, color = alpha, group = alpha)) +
-  geom_line() +
-  scale_color_viridis_c() +
-  theme_minimal()
-```

### 2021-04-14 preregistration.Rmd
summarise results

@@
 ```{r models}
+true_diff <- 0.2
 model_truth <- combine(
@@
   ),
-  intercepts("MACH", list(0, 0.2)),
+  intercepts("MACH", list(0, true_diff)),
   variances("MACH", list(1, 1))
@@
 ```{r}
-minn <- filter(res_interval, perc == .8, width < .2) %>% 
+choosen_perc <- .8
+choosen_width <- .2
+minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
   filter(n_obs == min(n_obs))
@@
   geom_line(aes(y = median), color = "white") +
-  geom_hline(yintercept = .2, linetype = "dotted") +
+  geom_hline(yintercept = true_diff, linetype = "dotted") +
   scale_fill_viridis_c(
@@
 
-Plot A depicts the range of effectsizes we observed in the simulation, where the color encodes the density.
+Plot A depicts the range of effect sizes we observed in the simulation, where the colour encodes the density.
 The light yellow interval shows where 80% of the estimates land.
-The dotted line represents the simulated true effect size; the white line shows the median of observed effect sizes.
+The dotted line represents the simulated true effect size; the white line shows the median observed effect sizes.
 
 Plot B represents the width of the interval shown in Plot A on a logarithmic scale.
-The cross shows the minimal required sample size so that 80% of the estimates land in a range of +/- .1.
+The cross shows the minimum required sample size so that 80% of the estimates land in a range of +/- .1.
+
+```{r}
+nmax <- filter(res_interval, n_obs == max(n_obs))
+bias <- (slice(nmax, 1)$median - true_diff)/true_diff
+```
+
+We can therefore conclude, that when we want to plan a study where we expect that `r scales::percent(choosen_perc)` of the estimates fall into a range of +/- `r choosen_width/2`, we need a minimum sample size of `r minn$n_obs` per group.
+The simulation also shows that we overestimate the effect size by `r scales::percent(bias)` (true: `r true_diff`, median of observed: `round(slice(nmax, 1)$median, 2)`) with a sample size per group of `r slice(nmax, 1)$n_obs`.

### 2021-04-15 R/hpc.R
add tardis access

@@ -1,0 +1,15 @@
+library("future.batchtools")
+# ssh keys login
+#ssh-copy-id -i /home/aaron/.ssh/id_rsa peikert@tardis.mpib-berlin.mpg.de
+tardis <- parallelly::makeClusterPSOCK("tardis.mpib-berlin.mpg.de",
+                port='random', user="peikert",
+                rscript=c("/opt/software/R/4.0.3/bin/Rscript"),
+                homogeneous = TRUE)
+ncpus <- 10
+plan(list(tweak(cluster, workers=tardis),
+          tweak(batchtools_slurm, workers = 16,
+                resources=list(ncpus=ncpus,
+                               memory='700m',
+                               walltime=6600,
+                               partition=c('gpu'))),
+          tweak(multicore, workers=ncpus)))

### 2021-04-15 R/hpc.R
use singularity on cluster

@@ -3,11 +3,13 @@ library("future.batchtools")
 #ssh-copy-id -i /home/aaron/.ssh/id_rsa peikert@tardis.mpib-berlin.mpg.de
 tardis <- parallelly::makeClusterPSOCK("tardis.mpib-berlin.mpg.de",
                 port='random', user="peikert",
-                rscript=c("/opt/software/R/4.0.3/bin/Rscript"),
+                rscript=c("/opt/software/R/4.0.3/bin/Rscript"), 
                 homogeneous = TRUE)
 ncpus <- 10
 plan(list(tweak(cluster, workers=tardis),
-          tweak(batchtools_slurm, workers = 16,
+          tweak(batchtools_slurm,
+                workers = 16,
+                template = ".batchtools.slurm.singularity.tmpl",
                 resources=list(ncpus=ncpus,
                                memory='700m',
                                walltime=6600,

### 2021-04-15 preregistration.Rmd
containerise preregistration/simulation

@@
   nmin: 100
-  nmax: 10000
+  nmax: 1000
   nstep: 100
   seed: 1234
+repro:
+  packages:
+    - tidyverse
+    - here
+    - lavaan
+    - furrr
+    - future.batchtools
+    - slider
+    - patchwork
+    - svglite
+  scripts:
+    - R/simulation.R
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, error = TRUE)
-
 library(tidyverse)
@@
 set.seed(params$seed)
+seed <- params$seed
 res_raw %<-% generate_data_setup(n_sim,
@@
                   globals = to_export,
-                  seed = params$seed,
+                  seed = seed,
                   packages = c("furrr", "lavaan", "tidyverse")
@@
 ```{r write-out, error=TRUE, include=FALSE}
+invisible(res_raw)
 write_rds(res_raw, here::here("res_raw.rds"))
@@
 The simulation also shows that we overestimate the effect size by `r scales::percent(bias)` (true: `r true_diff`, median of observed: `round(slice(nmax, 1)$median, 2)`) with a sample size per group of `r slice(nmax, 1)$n_obs`.
+
+# Session Info
+
+```{r}
+if(fs::file_exists(hpc_config)){
+  list(local = sessioninfo::session_info(),
+     login_node = value(future(sessioninfo::session_info())),
+     worker = value(future(value(future(sessioninfo::session_info())))))
+} else {
+  sessioninfo::session_info()
+}
+```

### 2021-04-16 manuscript.Rmd
streamline internal dependency

@@
 If researchers deviate from their preregistration, they probably have good reason to do so, which means they have learned something from the process of conducting the study.
-A preregistration, together with version control allows their readers to inform themselfs about these lessons learned.
+A preregistration, together with version control allows their readers to inform themselves about these lessons learned.
 
@@
 
-First dependency management, which means that you capture the interrelations between the components of your analysis.
+A reader of your reproducible manuscript should be able to derive how they are to reproduce it (ideally by a statement in your [README]).
+To be able to do that, it is often key to understand how different components of the project relate to each other.
 You cannot run an analysis without data, you cannot render a dynamic document without an analysis, and so forth.
-Understanding these relations is often key to reproducing a project.
-A reader of your reproducible manuscript should be able to derive how they are to reproduce it (ideally by a statement in your [README]).
+Dependency management means that you capture the interrelations between the components of your analysis.
 The "components" of an analysis are often loosely mapped onto files.

### 2021-04-16 manuscript.Rmd
update section on simulation to reflect the actual simulation

@@
 
+```{r, include=FALSE}
+loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
+cohend_jones_paulhus <- c(24, 29, 35)/100
+```
+
 There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
@@
 Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
+No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly, than simply thinking about it.
+If you identify unrealistic assumptions you can either violate these assumptions and see if the analytic strategy is robust or modify the analytical strategy to address them.
+We will show an example using structural equation modeling (SEM) for our Machiavellism example.
+However, the general process is applicable to any statistical model, but of course the simulation here is specific to our hypothetical research question.
 Of course, one has to make many educated guesses to simulate data that closely matches your expectation.
 These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
-However, the analytic strategy often gives you a hint about how to simulate data.
-In the Machiavellism example, the analytic strategy is to use structural equation modelling (SEM).
+The analytic strategy (i.e. SEM) often gives you a hint about how to simulate data.
+
+If you are unfamiliar with SEM, you may skim over the next sections.
 To simulate data for SEM, one has to specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
@@
 This way, we make sure that our analytic strategy is viable even in non-optimal circumstances.
-Therefore, we will assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.1.
+Therefore, we will assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.2.
 These assumption can be simply translated into [`lavaan` [@R-lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
@@
     ),
-  intercepts("MACH", list(0, 0.1))
+  intercepts("MACH", list(0, 0.2))
   )
@@
 This assumption is also baked into our simulation of the data.
-It is probably more realistic to violate this assumption and plan our analysis accordingly.
-Therefore, we simulate the data so that some of the items parameter differ among groups and then alter the analysis to assume partial measurement invariance.
+There is a good chance that this assumption is violated.
+We therefore adept our analysis to plan for this contingency.
+After fitting the model with strict measurement invariance, we free equality constrains first on loadings and then on intercepts based on the Lagrange Multiplier test.
+The model is iteratively refit and improved till at most three loadings or three intercepts are freed or the Lagrange Multiplier test is not significant for the constrains on loadings or intercepts.
 While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
@@
 @Rhemtulla2012 conclude that using Likert-like items is acceptable when they employ five or more categories.
-Hence, we continue simulating continues data.
-However, adding a little skewness to the items is easy and helps make the simulation more realistic.
+For simplicity we do not address these concerns (but doing so would certainly improve the simulation).
 Anyhow, remember that what we consider realistic for our case could be unfit for your case.
@@
 
-```{r generate}
+```{r}
+generate_data
 ```
 
-Note however that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/R/simulation.R", "GitHub")`.
+Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/R/simulation.R", "GitHub")`.
 
-With its help we can simulate 500 cases with a mean difference between groups of 0.1, where the intercepts of three random items vary by .3 and three loadings vary by 30% and five random items have a skew.
+With its help we can simulate 500 cases and fit our analysis.
 
 ```{r}
-simulated <- generate(n = 500,
-                      diff = 0.1,
-                      vary_intercepts = rep(0.3, 3),
-                      vary_loadings = rep(0.3, 3),
-                      skewness = rep(5, -0.5))
+true_diff <- 0.2
+model_truth <- combine(
+  measurement(
+    "MACH",
+    items("x", 9),
+    loadings = round(loadings_jones_paulhus * 0.7, 2)
+  ),
+  intercepts("MACH", list(0, true_diff)),
+  variances("MACH", list(1, 1))
+)
+
+model_same <- 
+  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+  MACH ~ c(0,0)*1
+  MACH ~~ c(1, NA)*MACH"
+
+model_differ <- 
+  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+  MACH ~ c(0, NA)*1 + c(zero, diff)*1
+  MACH ~~ c(1, NA)*MACH"
+generate_data(500,
+              model_truth,
+              model_same,
+              model_differ,
+              list(std_estimate = get_std_estimate),
+              group.equal = c("loadings", "intercepts"),
+              mod_load = 3,
+              mod_int = 3,
+              std.lv = TRUE)
 ```
@@
 Note, that this certainly results in different numbers each time we run it.
-To consistently get the same results it is good practice to use a seed.
-The function `seeded_generate()` will always generate the same pseudo-random numbers, when given the same seed.
 
 ```{r}
-seeded_generate <-  function(seed, ...)withr::with_seed(seed, generate(...))
-simulated <- seeded_generate(
-  seed = 1234,
-  n = 100000,
-  diff = 0.1,
-  vary_intercepts = rep(0.3, 3),
-  vary_loadings = rep(0.3, 3),
-  skewness = rep(5,-0.5)
-)
+generate_data(500,
+              model_truth,
+              model_same,
+              model_differ,
+              list(std_estimate = get_std_estimate),
+              group.equal = c("loadings", "intercepts"),
+              mod_load = 3,
+              mod_int = 3,
+              std.lv = TRUE)
+```
+
+To consistently get the same results it is good practice to use a seed with `set.seed(1235)`.
+It is a little more complicated when you have parallelized your code, like we did, but the future package handles this for us:
+
+```{r}
+set.seed(1235)
+runif(1)
+set.seed(1235)
+runif(1)
 ```
@@
 ```{r}
-lavaan::summary(analyze(simulated))
+future::value(future::future(runif(1), seed = 1235))
+future::value(future::future(runif(1), seed = 1235))
 ```

### 2021-04-16 R/simulation.R

@@ -17,7 +17,7 @@ safe_cfa <- function(...){
   valid_args <-
     c(names(formals(lavaan::lavaanify)),
       names(formals(lavaan::lavaan)),
-      names(lavOptions()))
+      names(lavaan::lavOptions()))
   dots <- filter_dots(..., fun = lavaan::cfa, what = valid_args)
   do.call("cfa", dots, envir = getNamespace("lavaan"))
 }
@@ -38,8 +38,8 @@ set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...)
   depth <- depth - 1
   if(depth < 0){return(list(pars = pars, differ = differ_fit))}
   if(!is_converged(differ_fit)){return(list(pars = pars, differ = differ_fit))}
-  parest <- parameterEstimates(differ_fit)
-  testscores <- lavTestScore(differ_fit)
+  parest <- lavaan::parameterEstimates(differ_fit)
+  testscores <- lavaan::lavTestScore(differ_fit)
   sorted_pars <- testscores$uni %>% 
     rename(label = lhs) %>% 
     select(-c(op, rhs, df)) %>% 
@@ -59,7 +59,7 @@ set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...)
 safe_set_free_mod <- safely(set_free_mod,
                             otherwise = list(same = NA, differ = NA))
 
-is_converged_ <- function(fit)lavInspect(fit, "converged")
+is_converged_ <- function(fit)lavaan::lavInspect(fit, "converged")
 is_converged <- function(fit){
   if(inherits(fit, "lavaan"))return(is_converged_(fit))
   else return(FALSE)

### 2021-04-16 R/standardizedSolution.R

@@ -45,7 +45,7 @@ standardizedSolution <-
     }
     
     if(is.null(partable)) {
-      PARTABLE <- inspect(object, "list")
+      PARTABLE <- lavaan::inspect(object, "list")
     } else {
       PARTABLE <- partable
     }

### 2021-04-16 manuscript.Rmd
update R + packages

@@
     - usethis
+    - git2r
     - aaronpeikert/repro@7bfaf98
@@
     - rstudio/webshot2@f62e743
+    - targets
   scripts:
@@
 However, the general process is applicable to any statistical model, but of course the simulation here is specific to our hypothetical research question.
-Of course, one has to make many educated guesses to simulate data that closely matches your expectation.
+One has to make many educated guesses to simulate data that closely matches your expectation.
 These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
@@
 To simulate data for SEM, one has to specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
-The model parameters should mimic the Machiavellism subscale of the short dark triad (SD3).
+The model parameters should express out expectations for the Machiavellism subscale of the short dark triad (SD3).
 As a starting point, we can use already published information.

### 2021-04-16 preregistration.Rmd
update R + packages

@@
 params:
-  nsim: 1000
+  nsim: 2
   nmin: 100
-  nmax: 1000
+  nmax: 300
   nstep: 100

### 2021-04-20 R/hpc.R
use absolut path for template

@@ -9,7 +9,7 @@ ncpus <- 10
 plan(list(tweak(cluster, workers=tardis),
           tweak(batchtools_slurm,
                 workers = 16,
-                template = ".batchtools.slurm.singularity.tmpl",
+                template = "/home/mpib/peikert/repro-tutorial/.batchtools.slurm.singularity.tmpl",
                 resources=list(ncpus=ncpus,
                                memory='700m',
                                walltime=6600,

### 2021-04-20 R/hpc.R
try out non singularity use

@@ -9,7 +9,7 @@ ncpus <- 10
 plan(list(tweak(cluster, workers=tardis),
           tweak(batchtools_slurm,
                 workers = 16,
-                template = "/home/mpib/peikert/repro-tutorial/.batchtools.slurm.singularity.tmpl",
+                template = "/home/mpib/peikert/repro-tutorial/.batchtools.slurm.tmpl",
                 resources=list(ncpus=ncpus,
                                memory='700m',
                                walltime=6600,

### 2021-04-20 manuscript.Rmd
preregistration as code

@@
 
+Often a preregistration is guided by a template, which formalizes the research questions and the planned analysis.
+An alternative to such protocol is to submit preregistration as code.
+Therefore, instead of describing the analysis that you plan to turn into code and run, you write the code beforehand.
+This is obviously challenging to do, because it is so concrete.
+
 ### R Markdown

### 2021-04-21 manuscript.Rmd
preregistration as code

--- 2021-04-16 manuscript.Rmd
@@
 For us a good preregistration is the same as a good plan; the only difference is that a preregistration is made public beforehand.
+
+The second advantage is transparency.
+A preregistration offers the interested reader more insight into the research process than a paper alone could.
 Some [citation!] argue that preregistration can not fulfill its promise to reduce the false discovery rate of science, if we allow to deviate from them.
-While such argument has some merits, we believe another advantage of preregistration is transparency.
+While such argument has some merits, we believe that transparency outweighs these concerns.
 If researchers deviate from their preregistration, they probably have good reason to do so, which means they have learned something from the process of conducting the study.
 A preregistration, together with version control allows their readers to inform themselves about these lessons learned.
+
+The liklihood that you have to diverge from your preregistration increases the more specific and concrete the preregistration is.
+To nudge researcher to be specific enough, there are many preregistrations template.
+These templates formalize somewhat the research questions and the planned analysis.
+An alternative to such protocol is to submit preregistration as code.
+Therefore, instead of describing the analysis that you plan to turn into code, you skip a step and write the code beforehand.
+This is obviously challenging to do, because it leaves no place to hide in the additional layer of abstraction.
+It may well be that for any given project this is impossible, without deviating from the preregistration.
+However, preregistered and version controlled code lets anyone see what was changed very concretely.
+
+One way to ensure that the code for the analysis works (at least in theory) is to simulate data.
+A data simulation allows you to verify that the code runs error free and lets you form more realistic expectation about the correctness and precision of the estimates.
+We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
+This showed that we have to gather at least `missing_rcode_simulation` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `missing_rcode_simulation`%.
+If you went to the trouble to simulate data, we argue that you get a preregistration for free.
 
@@
 <!--Hands on: Introduce R Markdown-->
+A preregistration in code of course describes sufficiently how you want to analyze the data, but ideally you also describe the why.
+Literate programming combines code and text, enabling you to explain your reasoning.
+As @knuthCWEBSystemStructured states: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
+We already learned about [Markdown], to write text.
+R Markdown allows you to include R code (in fact arbitrary computer code; see @riedererChapter15Other).
+To create a R Markdown document click on:
+
+> File → New File → Markdown File
+
+The code is separated from the text by three backtics, like here:
+
+````md
+```{r}`r ''`
+set.seed(1235)
+runif(1)
+```
+````
+
+When you include code into a R Markdown this code is run and its results are included in the rendered document.
+With R Markdown, you can write whole articles (like this one).
+You could in principle even write large parts of the manuscript as part of your preregistration, i.e. the introduction, theoretical background, the method section and parts of the discussion.
+In fact, if you have simulated data, you could even include fake results.
+Such dynamic document is arguably the most concrete preregistration possible.
+One advantage is that writing the manuscript and writing the preregistration are no longer two distinct tasks.
+Therefore a preregistration can be thought of as simply rearranging tasks that you have to do anyway.
+
 <!--Hands on: gitignore resulting .html/pdf-->

### 2021-04-21 manuscript.Rmd
remove citation for likert like vs continues

@@
 Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
-@Rhemtulla2012 conclude that using Likert-like items is acceptable when they employ five or more categories.
-For simplicity we do not address these concerns (but doing so would certainly improve the simulation).
+For simplicity we do not simulate these likely data problems (but doing so would certainly improve the simulation).
 Anyhow, remember that what we consider realistic for our case could be unfit for your case.
@@
 To consistently get the same results it is good practice to use a seed with `set.seed(1235)`.
-It is a little more complicated when you have parallelized your code, like we did, but the future package handles this for us:
+It is a little more complicated to archive uncorrelted randome numbers when you have parallelized your code, like we did, but the future package handles this for us:
 

### 2021-04-22 manuscript.Rmd
github is not a preregistration plattform

@@
 The liklihood that you have to diverge from your preregistration increases the more specific and concrete the preregistration is.
-To nudge researcher to be specific enough, there are many preregistrations template.
+To nudge researcher to still be specific enough, there are many preregistrations templates.
 These templates formalize somewhat the research questions and the planned analysis.
@@
 Therefore, instead of describing the analysis that you plan to turn into code, you skip a step and write the code beforehand.
-This is obviously challenging to do, because it leaves no place to hide in the additional layer of abstraction.
+This is obviously challenging to do, because it leaves no place to hide behind the additional layer of abstraction.
 It may well be that for any given project this is impossible, without deviating from the preregistration.
@@
 
-One way to ensure that the code for the analysis works (at least in theory) is to simulate data.
+One way to ensure that you can expect your analysis to work with real data is to simulate data.
 A data simulation allows you to verify that the code runs error free and lets you form more realistic expectation about the correctness and precision of the estimates.
@@
 This showed that we have to gather at least `missing_rcode_simulation` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `missing_rcode_simulation`%.
-If you went to the trouble to simulate data, we argue that you get a preregistration for free.
+If you went to the trouble to simulate data, we argue that you get a preregistration almost for free.
 
@@
 <!-- A preregistration needs to be public -->
+
+### Preregistration on GitHub
+
+An important aspect of a preregistration is that you have to make it public.
+Of course if you commit the preregistration to Git and upload it into a public repository it is also public.
+However, simply using GitHub as a preregistration site is not ideal.
+First, you can delete your GitHub repository whenever you want.
+Second, while Git records meticulously when you made which change, committing and publishing are two different steps.
+Therefore, you can choose to upload to GitHub after conducting the study, while still making it look like you have made the preregistration public before conducting the study. 
+Third, any person with some experience in Git can retroactively change the history.
+These downsides can be alleviated by using zenodo.org to give a digital object identifier (DOI) to your preregistration, because zenondo.org saves an unchangeable copy of your repository.
+You can learn more about this in [Creating a public release].
 
@@
 <!--Hands on: Introduce GitHub Projects-->
+
+## Creating a public release
 

### 2021-04-22 manuscript.Rmd
several minor text changes

@@
 The readme is the first thing they see when they visit your project on [GitHub].
-It should explain to them what your project is about and how to reproduce your project.
+It should explain to them what your project is about and how to reproduce it.
 You could for example include the abstract of your (planned) paper here.
@@
 The "components" of an analysis are often loosely mapped onto files.
-For a simple project it might be enough to have a single R Markdown.
+For a simple project it might be enough to have a single R Markdown file.
 However, while it is easy to understand how to reproduce the end result, this approach quickly gets unwieldy.
@@
 
-Lets suppose you want to eat spaghetti arrabiata and you still have delicious arrabiata sauce in the fridge.
-Would you cook the sauce again?
-No, you would only cook the pasta.
+When you cook something and you have a neccesary ingriedient already at hand, would you throw it away just to prepare it again?
+Unlikely.
 Make behaves the same.
@@
 ```
+
+This command will go through all R Markdown files in your repository and infer a Makefile from them:
 

### 2021-04-26 manuscript.Rmd
add section on docker

@@
 
+```
+---
+title: "Preregistration"
+author: "Aaron Peikert"
+date: "4/12/2021"
+output: html_document
+repro:
+  scripts:
+    - R/simulation.R
+---
+```
+
 This command will go through all R Markdown files in your repository and infer a Makefile from them:
@@
 
-A considerable number of external dependencies may hinder collaboration and long term reproducibility, but automated solutions can "bundle" and fix dependencies.
+Modern data analysis is only feasible because we rely on much external software.
+One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
+The key problem is the fragility of a software setup.
+Many components have to fit perfectly together and each change e.g. an update, risks that balance.
+Updates are, of course, very much necessary.
+Another problem is that the software environment is also variable and what works on one machine, might not work on any other.
+Docker circumvents such issues.
+Instead of delivering instructions on how to recreate some required components of an analysis, Docker delivers all software that was used in an analysis, down to the operating system.
+Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
+Such snapshot of the software stack is called "image".
+Of course packaging all needed software in such image, requires considerable amounts of storage.
+However, two strategies help to keep the storage requirements manageable.
+First, there is a community that maintains images for special purposes.
+For example, using R [rocker-citation].
+Such images only contain what is absolutely necessary for the specific purpose.
+Users can then install whatever they additionally need on top of these purpose built images.
+The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
+It does include R, RStudio, LaTeX as well as many R packages like the tidyverse (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
+Second, instead of saving a binary image of the software, one can also save a textual description of the programmatic steps to install the software in a `Dockerfile`.
+Such files are very small (the Dockerfile for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
+The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:
+
+```
+---
+title: "Preregistration"
+author: "Aaron Peikert"
+date: "4/12/2021"
+output: html_document
+repro:
+  scripts:
+    - R/simulation.R
+  packages:
+    - tidyverse
+    - here
+    - lavaan
+    - furrr
+---
+```
+
+These packages together with the R version that you currently used are then translated into an `Dockerfile`.
 

### 2021-04-26 manuscript.Rmd
check with grammerly

@@
 <!-- lure the reader in -->
-The research community has long accepted reproducibility as part of "good scientific practice", but has not has not benefited much from this self-imposed requirement.
+The research community has long accepted reproducibility as part of "good scientific practice" but has not benefited much from this self-imposed requirement.
 <!-- expand: who is calling it a requirement, maybe metascientific arguments-->
@@
 This tutorial walks through the creation of such a document that can be reproduced automatically.
-Though reproducibility is but a small part of the research process, it can serve as the mortar of open science building blocks (e.g. preregistration, open data, postpublication review).
+Though reproducibility is a small part of the research process, it can serve as the mortar of open science building blocks (e.g. preregistration, open data, postpublication review).
 Therefore, we show the whole lifecycle of an open research project, where some parts are essential to archive reproducibility.
 All code chunks are meant for the reader to try out for themselves.
-Still, the reader can safely skip sections marked as optional, when unfit for their research question or methods.
+Still, the reader can safely skip sections marked as optional when unfit for their research question or methods.
 However, any empirical data analysis will suffer from the following threats to reproducibility:
@@
 
-We deem each concept necessary to archive longterm reproducibility.
+We deem each concept necessary to archive long-term reproducibility.
 However, which tools you use to implement these concepts is a matter of taste and project requirements.
@@
 
-We assume that you have already installed R and RStudio (if not check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/install.md", "installation guide")`).
+We assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/install.md", "installation guide")`).
 Additionally, you'll need the [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:
@@
 If these commands detect that something is not installed or set up, they will give you detailed instructions (tailored to your operating system) to remedy the situation.
-Follow the instructions, then rerun the command, until it tells you not to worry.
+Follow the instructions, then rerun the command until it tells you not to worry.
 
@@
 Reproducibility is invaluable for effective collaboration---including with your "past self".
-While you might be tempted to consider reproducibility only when you about to finish a project, upfront investments pay of through your increased productivity.
+While you might be tempted to consider reproducibility only when you about to finish a project, upfront investments pay off through your increased productivity.
 Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you archive reproducibility gradually.
@@
 A great tool to capture ideas or meeting notes is Markdown.
-To create a Markdown document click on:
+To create a Markdown document, click on:
 
@@
 Another advantage of Markdown is that you can easily version-control it with Git.
-You may be familiar with Microsoft Words' "Track Changes"-Feature and version control is similar in its basic idea but more potent in its features.
+You may be familiar with Microsoft Words' "Track Changes"-Feature and version control is similar in their basic idea but more potent in their features.
 Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
@@
 And GitHub can show you what has changed, like in Figure \@ref(fig:idea-change).
-GitHub is in its essence a webinterface to Git and you may learn more about how to use it under the section on [GitHub].
+GitHub is, in its essence, a web interface to Git, and you may learn more about how to use it under the section on [GitHub].
 
@@
 
-<!--I am troubled by how we should recommend to learn git-->
-Git can be challenging to learn, because we are used to learn alone and Git is best learned with others.
-We recommend to find a partner to attend an online lecture or workshop to learn the ins and outs of Git.
+<!--I am troubled by how we should recommend to learn Git-->
+Git can be challenging to learn because we are used to learning alone, and Git is best learned with others.
+We recommend finding a partner to attend an online lecture or workshop to learn the ins and outs of Git.
 
@@
 Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
-No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly, than simply thinking about it.
-If you identify unrealistic assumptions you can either violate these assumptions and see if the analytic strategy is robust or modify the analytical strategy to address them.
-We will show an example using structural equation modeling (SEM) for our Machiavellism example.
-However, the general process is applicable to any statistical model, but of course the simulation here is specific to our hypothetical research question.
+No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about it.
+If you identify unrealistic assumptions, you can either violate these assumptions and see if the analytic strategy is robust or modify the analytical approach to address them.
+We will show an example using structural equation modelling (SEM) for our Machiavellism example.
+However, the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
 One has to make many educated guesses to simulate data that closely matches your expectation.
@@
 To simulate data for SEM, one has to specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
-The model parameters should express out expectations for the Machiavellism subscale of the short dark triad (SD3).
+The model parameters should express our expectations for the Machiavellism subscale of the short dark triad (SD3).
 As a starting point, we can use already published information.
@@
 
-But wait `lavaan` warns us that this model may not be identified.
-We overlooked a crucial assumption of this model, that is measurement invariance.
+But wait, `lavaan` warns us that this model may not be identified.
+We overlooked a crucial assumption of this model, that is, measurement invariance.
 This assumption is also baked into our simulation of the data.
 There is a good chance that this assumption is violated.
-We therefore adept our analysis to plan for this contingency.
-After fitting the model with strict measurement invariance, we free equality constrains first on loadings and then on intercepts based on the Lagrange Multiplier test.
-The model is iteratively refit and improved till at most three loadings or three intercepts are freed or the Lagrange Multiplier test is not significant for the constrains on loadings or intercepts.
+We, therefore, adept our analysis to plan for this contingency.
+After fitting the model with strict measurement invariance, we free equality constraints first on loadings and then on intercepts based on the Lagrange Multiplier test.
+The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
 While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
 Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
-For simplicity we do not simulate these likely data problems (but doing so would certainly improve the simulation).
+For simplicity, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).
 Anyhow, remember that what we consider realistic for our case could be unfit for your case.
-If you are satisfied with your assumptions it helps to write a function that expresses them, like this one:
+If you are satisfied with your assumptions, it helps to write a function that expresses them, like this one:
 <!--Hands-on: build simple functions tailored for your analysis-->
@@
 
-To consistently get the same results it is good practice to use a seed with `set.seed(1235)`.
-It is a little more complicated to archive uncorrelted randome numbers when you have parallelized your code, like we did, but the future package handles this for us:
+To consistently get the same results, it is good practice to set a seed with `set.seed(1235)`.
+It is a little more complicated to archive uncorrelated random numbers when you have parallelized your code like we did, but the future package handles this for us:
 
@@
 
-A preregistration serves as a tool that may help to increase the credibility of empirical results.
-Without getting lost in the meta scientific arguments for preregistration, we want to note two advantages.
-First, a preregistration is an opportunity to carefully plan a study.
-Gathering empirical data often requires enormous resources and better planning increases the chances that these efforts are worthwhile.
-For us a good preregistration is the same as a good plan; the only difference is that a preregistration is made public beforehand.
+Preregistration serves as a tool that may help to increase the credibility of empirical results.
+Without getting lost in the meta-scientific arguments for preregistration, we want to note two advantages.
+First, a preregistration is an opportunity to plan a study carefully.
+Gathering empirical data requires many resources, and better planning increases the chances that these efforts are worthwhile.
+For us, a good preregistration is the same as a good plan; the only difference is that a preregistration is made public beforehand.
 
@@
 A preregistration offers the interested reader more insight into the research process than a paper alone could.
-Some [citation!] argue that preregistration can not fulfill its promise to reduce the false discovery rate of science, if we allow to deviate from them.
-While such argument has some merits, we believe that transparency outweighs these concerns.
+Some [citation!] argue that preregistration can not fulfil its promise to reduce the false discovery rate of science if we allow deviating from them.
+While such an argument has some merits, we believe that transparency outweighs these concerns.
 If researchers deviate from their preregistration, they probably have good reason to do so, which means they have learned something from the process of conducting the study.
-A preregistration, together with version control allows their readers to inform themselves about these lessons learned.
+A preregistration, together with version control, allows their readers to inform themselves about these lessons learned.
 
-The liklihood that you have to diverge from your preregistration increases the more specific and concrete the preregistration is.
-To nudge researcher to still be specific enough, there are many preregistrations templates.
+The likelihood that you have to diverge from your preregistration increases the more specific and concrete the preregistration is.
+To nudge the researcher to still be specific enough, there are many preregistrations templates.
 These templates formalize somewhat the research questions and the planned analysis.
 An alternative to such protocol is to submit preregistration as code.
-Therefore, instead of describing the analysis that you plan to turn into code, you skip a step and write the code beforehand.
-This is obviously challenging to do, because it leaves no place to hide behind the additional layer of abstraction.
-It may well be that for any given project this is impossible, without deviating from the preregistration.
-However, preregistered and version controlled code lets anyone see what was changed very concretely.
+Therefore, instead of describing the analysis, you plan to turn into code, you skip a step and write the code beforehand.
+Such preregistration as code is challenging because it leaves no place to hide behind the additional layer of abstraction that verbal descriptions provide.
+It may well be that for any given project, this is impossible without deviating from the preregistration.
+However, preregistered and version-controlled code lets anyone see what was changed very concretely.
 
 One way to ensure that you can expect your analysis to work with real data is to simulate data.
-A data simulation allows you to verify that the code runs error free and lets you form more realistic expectation about the correctness and precision of the estimates.
+A data simulation allows you to verify that the code runs error-free and lets you form more realistic expectations about the estimates' correctness and precision.
 We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
-This showed that we have to gather at least `missing_rcode_simulation` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `missing_rcode_simulation`%.
+The simulation showed that we have to gather at least `missing_rcode_simulation` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `missing_rcode_simulation`%.
 If you went to the trouble to simulate data, we argue that you get a preregistration almost for free.
@@
 <!--Hands on: Introduce R Markdown-->
-A preregistration in code of course describes sufficiently how you want to analyze the data, but ideally you also describe the why.
+A preregistration in code describes sufficiently how you want to analyze the data, but ideally, you also convey the why.
 Literate programming combines code and text, enabling you to explain your reasoning.
@@
 We already learned about [Markdown], to write text.
-R Markdown allows you to include R code (in fact arbitrary computer code; see @riedererChapter15Other).
-To create a R Markdown document click on:
+R Markdown allows you to include R code (in fact, arbitrary computer code; see @riedererChapter15Other).
+To create an R Markdown document, click on:
 
@@
 
-The code is separated from the text by three backtics, like here:
+The code is separated from the text by three backticks, like here:
 
@@
 
-When you include code into a R Markdown this code is run and its results are included in the rendered document.
+When you include code into an R Markdown, this code is run, and its results are included in the rendered document.
 With R Markdown, you can write whole articles (like this one).
-You could in principle even write large parts of the manuscript as part of your preregistration, i.e. the introduction, theoretical background, the method section and parts of the discussion.
-In fact, if you have simulated data, you could even include fake results.
-Such dynamic document is arguably the most concrete preregistration possible.
+You could, in principle, even write large parts of the manuscript as part of your preregistration, i.e. the introduction, theoretical background, the method section and parts of the discussion.
+If you have simulated data, you could even include fake results.
+Such a dynamic document is arguably the most concrete preregistration possible.
 One advantage is that writing the manuscript and writing the preregistration are no longer two distinct tasks.
@@
 
-An important aspect of a preregistration is that you have to make it public.
-Of course if you commit the preregistration to Git and upload it into a public repository it is also public.
+An essential aspect of a preregistration is that you have to make it public.
+Of course, if you commit the preregistration to Git and upload it into a public repository, it is also public.
 However, simply using GitHub as a preregistration site is not ideal.
 First, you can delete your GitHub repository whenever you want.
-Second, while Git records meticulously when you made which change, committing and publishing are two different steps.
-Therefore, you can choose to upload to GitHub after conducting the study, while still making it look like you have made the preregistration public before conducting the study. 
+Second, while Git meticulously records when you made which change, committing and publishing are two different steps.
+Therefore, you can choose to upload to GitHub after conducting the study while still making it look like you have made the preregistration public before conducting the study. 
 Third, any person with some experience in Git can retroactively change the history.
-These downsides can be alleviated by using zenodo.org to give a digital object identifier (DOI) to your preregistration, because zenondo.org saves an unchangeable copy of your repository.
+You can alleviate these downsides by using zenodo.org to give a digital object identifier (DOI) to your preregistration because zenondo.org saves an unchangeable copy of your repository.
 You can learn more about this in [Creating a public release].
@@
 
-One of the greatest benefits of open research is that other people can freely reuse materials.
+One of the most significant benefits of open research is that other people can freely reuse materials.
 To use your materials, they first have to find them and learn how to use them.
@@
 It should explain to them what your project is about and how to reproduce it.
-You could for example include the abstract of your (planned) paper here.
+You could, for example, include the abstract of your (planned) paper here.
 
@@
 <!--knit/rmarkdown will be explained in section above-->
-When you knit this file, it in turn creates a `README.md`.
-That means you can include R code or plots etc. within the readme.
+When you knit this file, it, in turn, creates a `README.md`.
+That means you can include R code or plots etc., within the readme.
 The `README.md` file will look pleasant in the browser, just like in Figure \@ref(fig:readme).
@@
 
-Even when other researchers are excited to use your code or materials, they are usually legally inhibited from to doing so.
-That is because in many countries (i.e. the USA, European Union) the creator of something has exclusive right to use their creation.
+Even when other researchers are excited to use your code or materials, they are usually legally inhibited from doing so.
+In many countries (i.e. the USA, European Union), the creator of something has exclusive right to use their creation.
 A license allows other people to use what you have created.
 <!---exact copy from our previous manuscript, do we need to change? -->
-In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
+In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media.
+In contrast, scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
 These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
-A great resource to choose a license is [choosealicense.com](https://choosealicense.com), however, no resource, including our recommendation, replaces legal advice.
+A great resource to choose a license is [choosealicense.com](https://choosealicense.com).
+However, no resource, including our recommendation, replaces legal advice.
 <!-- plagiary end -->
 It is important to get written consent from collaborators when you change/add the license.
-However, if their first contribution occurs after you have added a license, they give their consent by contributing.
-To add a CC-BY license to your project simply run (there are similar commands for other licenses):
+However, if their first contribution occurs after adding a license, they give their consent by contributing.
+To add a CC-BY license to your project, simply run (there are similar commands for other licenses):
 
@@
 We strongly suggest that you strive for respectful interactions with anyone who might want to contribute.
-In fact, we hope it comes naturally to you.
+We hope it comes naturally to you.
 To communicate that contributors can expect a welcoming and respectful environment, you can add the `r link( "https://www.contributor-covenant.org", "Contributor Covenant")` to your repository.
-Contributing to an open project is daunting and a public pledge to value any contributions (even when you do not incorporate them) can help others to take the first step.
+Contributing to an open project is daunting, and a public pledge to value any contributions (even when you do not incorporate them) can help others to take the first step.
 The Contributor Covenant is a code of conduct used by many open source project.
@@
 However, there is some controversy around how to handle violations of a code of conduct.
-For that or any other reason, you may want to adapt the Contributor Covenant, to fit the needs of you project (which you are allowed to do because it is published with a CC-BY license).
+For that or any other reason, you may want to change the Contributor Covenant to fit your project's needs (which you are allowed to do because it is published with a CC-BY license).
 You can also consider alternative Codes of Conduct, like the `r link("https://opensource.microsoft.com/codeofconduct/", "Microsoft Open Source Code of Conduct")` or the `r link("https://berlincodeofconduct.org", "Berlin Code of Conduct")`.
@@
 Often, reproducibility is reduced to the merits it has for the reader.
-However, we argue that reproducibility is vital for collaboration, not only in the sense that the consumer of a scientific product build upon it, but also that the creators work together.
-The time that you invest into ensuring reproducibility, is directed at ensuring effective collaboration.
+However, we argue that reproducibility is vital for collaboration, not only in the sense that the consumer of a scientific product builds upon it but also that the creators work together.
+The time that you invest into ensuring reproducibility is directed at ensuring effective collaboration.
 Therefore, reproducibility is at its core about collaboration.
@@
 
-Only from this moment on, files are uploaded to the internet and only those which you have explicitly added and committed in Git.
+Only from this moment on files are uploaded to the internet and only those you have explicitly added and committed in Git.
 
@@
 
-While you already know how to write dynamic documents with [Markdown] and [R Markdown], and track changes with [Git] and [GitHub], two components to archive longterm reproducibility are still missing, namely internal and external dependencies.
+While you already know how to write dynamic documents with [Markdown] and [R Markdown], and track changes with [Git] and [GitHub], two components to archive long-term reproducibility are still missing, namely internal and external dependencies.
 
@@
 A reader of your reproducible manuscript should be able to derive how they are to reproduce it (ideally by a statement in your [README]).
-To be able to do that, it is often key to understand how different components of the project relate to each other.
-You cannot run an analysis without data, you cannot render a dynamic document without an analysis, and so forth.
+To do that, it is often crucial to understand how different project components relate to each other.
+You cannot run an analysis without data; you cannot render a dynamic document without an analysis, and so forth.
 Dependency management means that you capture the interrelations between the components of your analysis.
 The "components" of an analysis are often loosely mapped onto files.
-For a simple project it might be enough to have a single R Markdown file.
+For a simple project, it might be enough to have a single R Markdown file.
 However, while it is easy to understand how to reproduce the end result, this approach quickly gets unwieldy.
@@
 For example, they create a script to load the data and another to run the analysis.
-This modularisation facilitates reuse, because the loading script may be used by other projects building on the same data, while the analysis script may be employed on other data and so forth.
-The downside is that reproducibility is complicated, because instead of reproducing everything in one step, multiple steps are required.
+This modularisation facilitates reuse because the loading script may be used by other projects building on the same data, while the analysis script may be employed on different data and so forth.
+The downside is that reproducibility is complicated because instead of reproducing everything in one step, multiple steps are required.
 
-Of course there a several ways to solve this problem.
-One could build an R package to separate analysis and data or use code externalization in the RMarkdown to facilitating reuse or build a "master" script that combines all steps.
+Of course, there a several ways to solve this problem.
+One could build an R package to separate analysis and data or use code externalization in the RMarkdown to facilitating reuse or create a "master" script that combines all steps.
 We found Make to be particularly useful for two reasons.
 First, it is language agnostic, therefore fusing steps that are carried out in different programming languages.
-Second, it does implement a simple, yet powerful caching mechanism of results which is welcome in computing intensive analysis often found in neuroimaging or machine learning.
+Second, it does implement a simple yet powerful caching mechanism of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning.
 
@@
 
-When you cook something and you have a neccesary ingriedient already at hand, would you throw it away just to prepare it again?
+When you cook something, and you have the necessary ingredients already at hand, would you throw it away just to prepare it again?
 Unlikely.
 Make behaves the same.
-If a file already exists it does not recreate it unnecessarily.
+If a file already exists, it does not recreate it unnecessarily.
 However, if a requirement is newer than its target, make recreates the target to reflect the updated dependency.
-Therefore, when you update the data file, the analysis is run again.
+Therefore, when you update the data file, the analysis is rerun.
 Because Make knows the whole dependency tree, its caching is smarter than many alternatives.
@@
 `targets` can infer more about your project than Make, because it analysis the R code you write to build the dependency tree.
-For this inference to work well you have to write everything in R and you have to write everything strictly functional.
+For this inference to work well, you have to write everything in R, and you have to write everything strictly functional.
 
@@
 One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
-The key problem is the fragility of a software setup.
-Many components have to fit perfectly together and each change e.g. an update, risks that balance.
+The critical problem is the fragility of a software setup.
+Many components have to fit perfectly together, and each change, e.g. an update, risks that balance.
 Updates are, of course, very much necessary.
-Another problem is that the software environment is also variable and what works on one machine, might not work on any other.
+Another problem is that the software environment is also variable and what works on one machine might not work on any other.
 Docker circumvents such issues.
-Instead of delivering instructions on how to recreate some required components of an analysis, Docker delivers all software that was used in an analysis, down to the operating system.
+Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, down to the operating system.
 Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
-Such snapshot of the software stack is called "image".
-Of course packaging all needed software in such image, requires considerable amounts of storage.
+Such a snapshot of the software stack is called "image".
+Of course, packaging all needed software in such an image requires considerable amounts of storage.
 However, two strategies help to keep the storage requirements manageable.
-First, there is a community that maintains images for special purposes.
+First, there is a community that maintains images for particular purposes.
 For example, using R [rocker-citation].
-Such images only contain what is absolutely necessary for the specific purpose.
-Users can then install whatever they additionally need on top of these purpose built images.
+Such images only contain what is necessary for a specific purpose.
+Users can then install whatever they additionally need on top of these purpose-built images.
 The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
@@
 Second, instead of saving a binary image of the software, one can also save a textual description of the programmatic steps to install the software in a `Dockerfile`.
-Such files are very small (the Dockerfile for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
+Such files are tiny (the Dockerfile for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
 The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:
@@
 
-These packages together with the R version that you currently used are then translated into an `Dockerfile`.
+With the R version that you currently use, this specification is then translated into a `Dockerfile`.
 

### 2021-04-26 manuscript.Rmd
some changes after reading

@@
 Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
-No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about it.
+No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about them.
 If you identify unrealistic assumptions, you can either violate these assumptions and see if the analytic strategy is robust or modify the analytical approach to address them.
-We will show an example using structural equation modelling (SEM) for our Machiavellism example.
+We will show an example using structural equation modeling (SEM) for our Machiavellism example.
 However, the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
@@
 If you are unfamiliar with SEM, you may skim over the next sections.
-To simulate data for SEM, one has to specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
+To simulate data for SEM, one could specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
 The model parameters should express our expectations for the Machiavellism subscale of the short dark triad (SD3).
@@
 Anyhow, remember that what we consider realistic for our case could be unfit for your case.
+
 If you are satisfied with your assumptions, it helps to write a function that expresses them, like this one:
@@
 On your local machine, you use Git to track changes.
-Online services like GitHub allow you to share these changes with others.
+Online services like GitHub allow you to share these changes with others and collaborate with them.
 GitHub is not the only solution; there is also [GitLab](gitlab.com) or [BitBucket](https://bitbucket.org), which are as feature-rich as GitHub.
@@
 For this purpose, `repro` utilizes the YAML metadata:
-
-```{r}
-x <- readLines(here::here("preregistration.Rmd"))
-yaml_ind <- stringr::str_which(x, "^---[[:space:]]*")
-stripped <- stringr::str_c(x[seq(yaml_ind[[1]], yaml_ind[[2]])],
-                  collapse = "\n")
-cat("```\n", stripped, "\n```", sep = "")
-```
 
@@
 ```{r, include=FALSE}
-knitr::write_bib(c(.packages(), "bookdown", "lavaan", "knitr", "targets"), "packages.bib")
+knitr::write_bib(c(.packages(), "bookdown", "lavaan", "knitr", "targets", "renv"), "packages.bib")
 ```

### 2021-04-26 manuscript.Rmd
add renv

@@
 
+Docker can be a bit unfriendly, because it requires a complicated mental model and some knowledge about how to install software on Linux.
+An R-focused alternative is the R package [`renv`](https://rstudio.github.io/renv/articles/renv.html)[^renv][@R-renv].
+`renv` isolates all R packages used in a project and records there version as well as the source from where they where installed.
+This mechanism is faster and simpler to use than Docker, at the expense of only tracking R packages.
+`renv` therefore is already of great use for reproducibility.
+However, even when your project relies solely on R, external dependencies still change the outcome of computations, like the operating system, the compilers used, the BLAS system in use and so forth.
+Most importantly the R version itself has to match.
+
+[^renv]: https://rstudio.github.io/renv/articles/renv.html
+
 ### Singularity (optional)
+
 <!-- Hands on: add data via automate() -->

### 2021-04-26 manuscript.Rmd
add singularity

@@
 
+One usecase in which Docker sometimes can't be used is in high performance computing (HPC).
+The simulation we did is computationally expensive enough to warrant the use of a HPC cluster.
+To run the simulation on a cluster we require the exact same software, not only on one or two computers, but potentially hundreds.
+Docker is a natural solution for this problem and is the defacto standard for cloud computing.
+However, Docker requires admin rights, while this is no issue on cloud computing infrastructure it is an issue in classical HPC environments.
+Singularity was designed to run without admin rights, while maintaining full compatibility with Docker images.
+If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
+Anyway, from within the container i.e. for your analysis it does not make any difference.
+
 <!-- Hands on: add data via automate() -->

### 2021-04-26 preregistration.Rmd
externalise hash

@@
     - R/simulation.R
+    - R/hash.R
 ---
@@
 
-```{r checksum}
-# compute checksum using md5
-checksum <- digest::digest(res_raw, "md5")
-if(checksum != "4203854848ad0cc5d2f848c44b70815d"){
-  warning("Mismatch between original and current simulations!
-Checksum now is:\n    '", checksum, "'")
-}
+```{r hash}
 ```

### 2021-04-26 R/hash.R

@@ -1,0 +1,6 @@
+#----hash----
+# compute hash using md5
+hash <- digest::digest(res_raw, "md5")
+if(hash != "4203854848ad0cc5d2f848c44b70815d"){
+  warning("Mismatch between original and current simulations!\nHash now is:\n    '", hash, "'")
+}

### 2021-04-26 manuscript.Rmd
add section on hashing/data integrity

@@
 
+To reproduce an analysis, you need to have access to the data.
+There are several reasons which may prohibit you from making your data available, like legal or regulatory or privacy concerns.
+Often you find yourself in a situation where you can make the data only availabe on request.
+A possibility for irreproducibility is that the data has changed in the meantime.
+To at least check that the data was not changed you can provide a so called "hash".
+This is a cryptographic fingerprint.
+If the data changes, so does its fingerprint or hash.
+The user can therefore compare if the data they have received is the same that was used in the original publication, as long as the original publication does contain the hash of the data.
+
+To verify the data integrity you can include a line of code after reading the data that looks like this:
+
+```{r}
+# create a dummy data.frame with two columns
+x <- data.frame(VAR1 = c(1, 2, 3, 4), VAR2 = c(0, 4, 6, 9))
+# compute hash using md5
+hash <- digest::digest(x, "md5")
+if(hash != "4203854848ad0cc5d2f848c44b70815d"){
+  warning("Mismatch between original and current simulations!\nHash now is:\n    '", hash, "'")
+}
+```
+
+We did the same for our simulation to verify that the simulation is really reproducible[^simhash].
+
+[^simhash]: https://github.com/aaronpeikert/repro-tutorial/blob/master/R/hash.R
+
 ## Anonymising Data (optional)

### 2021-04-26 manuscript.Rmd
merge sections on data checking and synthetic creation

@@
 
-## Data Integrity (optional)
+## If you can not make the data public (optional)
 
@@
 
-We did the same for our simulation to verify that the simulation is really reproducible[^simhash].
+We employed the same mechanism for our simulation to verify that the simulation is really reproducible[^simhash].
 
@@
 
-## Anonymising Data (optional)
-
-<!-- Should we add a section on synthetic data generation? -->
-<!-- Hands on: Create synthetic data -->
+An compromise between making the data publicly available and keeping it closed is to publish synthetic data.
+Such approach tries to recreate the marginal distribution and bivariate relationships in simulated data, e.g. with the R package `synthpop`[@nowokSynthpopBespokeCreation2016].
+The results of such data should generally not be interpreted.
+However, synthetic data provides an opportunity for researcher to test if a project is in principle reproducible, even though they have no access to the data.
 

### 2021-04-28 manuscript.Rmd
it should be dependency tracking #4

@@
 You cannot run an analysis without data; you cannot render a dynamic document without an analysis, and so forth.
-Dependency management means that you capture the interrelations between the components of your analysis.
+Dependency tracking means that you capture the interrelations between the components of your analysis.
 The "components" of an analysis are often loosely mapped onto files.

### 2021-04-30 manuscript.Rmd
small changes

@@
 
-Even when other researchers are excited to use your code or materials, they are usually legally inhibited from doing so.
+Even when other researchers are excited to use your published code or materials, they are usually legally inhibited from doing so.
 In many countries (i.e. the USA, European Union), the creator of something has exclusive right to use their creation.
@@
 
-This command will go through all R Markdown files in your repository and infer a Makefile from them:
+This command will go through all R Markdown files, and their YAML metadata, in your repository and infer a Makefile from them:
 
@@
 The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
-It does include R, RStudio, LaTeX as well as many R packages like the tidyverse (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
+It does include Ubuntu (a Linux operating system) R, RStudio, LaTeX as well as many R packages like the tidyverse (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
 Second, instead of saving a binary image of the software, one can also save a textual description of the programmatic steps to install the software in a `Dockerfile`.
-Such files are tiny (the Dockerfile for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
+Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
 The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:
@@
 `renv` therefore is already of great use for reproducibility.
-However, even when your project relies solely on R, external dependencies still change the outcome of computations, like the operating system, the compilers used, the BLAS system in use and so forth.
+However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS system in use and so forth.
 Most importantly the R version itself has to match.
@@
 
-One usecase in which Docker sometimes can't be used is in high performance computing (HPC).
+One usecase in which Docker sometimes can not be used is in high performance computing (HPC).
 The simulation we did is computationally expensive enough to warrant the use of a HPC cluster.
@@
 hash <- digest::digest(x, "md5")
-if(hash != "4203854848ad0cc5d2f848c44b70815d"){
-  warning("Mismatch between original and current simulations!\nHash now is:\n    '", hash, "'")
+if(hash != "5ba412f5a26f43842971dd74954fcdeb"){
+  warning("Mismatch between original and current data!\nHash now is:\n    '", hash, "'")
 }
@@
 ```{r, include=FALSE}
-knitr::write_bib(c(.packages(), "bookdown", "lavaan", "knitr", "targets", "renv"), "packages.bib")
+knitr::write_bib(c(.packages(), "bookdown", "lavaan", "knitr", "targets", "renv"),
+                 "packages.bib")
 ```

### 2021-04-30 manuscript.Rmd
code is R code by default
rearrange headings

@@
 All code chunks are meant for the reader to try out for themselves.
+If not declared otherwise the code is meant to run in R (version `r with(R.version, paste0(major, ".", minor))`).
 Still, the reader can safely skip sections marked as optional when unfit for their research question or methods.
@@
 
+### Preregistration as Code
+
 The likelihood that you have to diverge from your preregistration increases the more specific and concrete the preregistration is.
@@
 If you went to the trouble to simulate data, we argue that you get a preregistration almost for free.
+A preregistration in code describes sufficiently how you want to analyze the data, but ideally, you also convey the why.
+Dynamic document generation combines code and text, enabling you to explain your reasoning.
+See the section on [R Markdown] for an example of a dynamic document generation paradigm.
 
@@
 <!--Hands on: Introduce R Markdown-->
-A preregistration in code describes sufficiently how you want to analyze the data, but ideally, you also convey the why.
-Literate programming combines code and text, enabling you to explain your reasoning.
 As @knuthCWEBSystemStructured states: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
@@
 
+### README (optional)
+
 One of the most significant benefits of open research is that other people can freely reuse materials.
@@
 
+### License
+
 Even when other researchers are excited to use your published code or materials, they are usually legally inhibited from doing so.
@@
 ```
+
+### Code of Conduct
 
@@
 <!--Hands on: Introduce GitHub Projects-->
-
-## Creating a public release
 
@@
 
-## Releasing a Preprint (optional)
+# Creating a public release
 
+<!-- public releases like preregistration, preprint -->
 <!-- Hands on: add commit hash to results-->
 <!-- Hands on: git tag + github release-->
-
-# Peer Review (optional)
-
-# Post Publication Review (optional)
-
-# Archiving
-
 <!-- Hands on: explain how to archive data-->

### 2021-04-30 manuscript.Rmd
update docker

@@
     - targets
+    - renv
   scripts:

### 2021-04-30 manuscript.Rmd
restructure the archiving section #15

@@
 
-# Creating a public release
+# Archive Milestones
+
+The open nature of the here proposed workflow does blur the lines between published and unpublished.
+However, even when your Git repository is public, you may want to signify important stages of your project, like preregistration, preprint, submission and acceptance to the journal.
+These milestones deserve to be archived.
+A project is only then properly archived when it is stored at least at two independent storage locations and reproducible from all locations.
+Therefore, you should make extra sure that the project at these points is reproducible.
+Ideally, you ask a coworker or collaborator to reproduce the project independently from you.
+The sections [Creating a DOI] and [Creating a Release] describe how you can easily create three different openly accessible longterm archived copies of your project.
+
+## Creating a DOI
+
+## Creating a Release
+
+GitHub has a "release" function, which creates a link to a specific version of your project, where everything is available as a zip file.
+You have the opportunity to describe the release in the release notes and to upload other additional files (e.g. the rendered pdf).
+
+## Archiving Docker
+
+## Archiving Data
 

### 2021-05-01 manuscript.Rmd
close #15
add zenodo

@@
 
-# Archive Milestones
+# Archiving and Dissemination
 
@@
 Therefore, you should make extra sure that the project at these points is reproducible.
-Ideally, you ask a coworker or collaborator to reproduce the project independently from you.
+To ensure reproducible, you could ask a coworker or collaborator to reproduce the project independently from you.
 The sections [Creating a DOI] and [Creating a Release] describe how you can easily create three different openly accessible longterm archived copies of your project.
@@
 ## Creating a DOI
+
+zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides digital object identifiers (DOI) for these archives.
+While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
+Please note that you can only link public GitHub repositories to zenodo.
+You may log into zenodo.org through your GitHub account.
+
+> Log into zenodo.org → Account → GitHub[^zenodo-github]
+
+[^zenodo-github]: https://zenodo.org/account/settings/github/
+
+After you have linked a GitHub repository you trigger the archivation by creating a GitHub release, as described in the next section.
 

### 2021-05-01 manuscript.Rmd
include hash in manuscript

@@
 
+When you disseminate the paper across more locations than just GitHub it can be ambiguous which version of code and data have created the document (e.g. the manuscript.pdf).
+To create a clear link between the repository version and some end result you can automatically include the hash of the commit.
+Either you use Git from within R:
+
+```{r}
+system("git rev-parse --short HEAD")
+```
+
+Or you can use the `gert` package:
+
+```{r}
+strtrim(gert::git_commit_id(), 7)
+```
+
 ## Creating a DOI
 
+<!-- public releases like preregistration, preprint -->
 zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides digital object identifiers (DOI) for these archives.
@@
 
-<!-- public releases like preregistration, preprint -->
 <!-- Hands on: add commit hash to results-->

### 2021-05-03 manuscript.Rmd
actually print hash
add introduction to "prepare public release"

@@
 
+An open research project usually entails that you publish content to the public web, without an intermediate publisher.
+Since you are directly responsible for such content, you have to prepare a few things that are usually done by the publisher for you.
+
 ### README (optional)
@@
 ```{r}
-system("git rev-parse --short HEAD")
+system2("git", c("rev-parse",  "--short",  "HEAD"), stdout = TRUE)
 ```
@@
 
-> Log into zenodo.org → Account → GitHub[^zenodo-github]
+> Log into zenodo.org → Account → [GitHub[^zenodo-github]](https://zenodo.org/account/settings/github/)
 

### 2021-05-03 manuscript.Rmd
small changes after reading

@@
 All code chunks are meant for the reader to try out for themselves.
-If not declared otherwise the code is meant to run in R (version `r with(R.version, paste0(major, ".", minor))`).
+If not declared otherwise the code is meant to run in R (tested with version `r with(R.version, paste0(major, ".", minor))`).
 Still, the reader can safely skip sections marked as optional when unfit for their research question or methods.
@@
 We, therefore, adept our analysis to plan for this contingency.
-After fitting the model with strict measurement invariance, we free equality constraints first on loadings and then on intercepts based on the Lagrange Multiplier test.
+After fitting the model with strict measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
 The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
@@
 We already learned about [Markdown], to write text.
-R Markdown allows you to include R code (in fact, arbitrary computer code; see @riedererChapter15Other).
+R Markdown allows you to include R code (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown.
 To create an R Markdown document, click on:
 
-> File → New File → Markdown File
+> File → New File → R Markdown...
 
@@
 When you include code into an R Markdown, this code is run, and its results are included in the rendered document.
-With R Markdown, you can write whole articles (like this one).
+With R Markdown, you can write whole articles (`r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/manuscript.Rmd", "like the one you are reading")`).
 You could, in principle, even write large parts of the manuscript as part of your preregistration, i.e. the introduction, theoretical background, the method section and parts of the discussion.
@@
 The results of such data should generally not be interpreted.
-However, synthetic data provides an opportunity for researcher to test if a project is in principle reproducible, even though they have no access to the data.
+However, synthetic data provides an opportunity for readers to test if a project is in principle reproducible, even though they have no access to the data.
 

### 2021-05-03 manuscript.Rmd
add placeholder for GitHub collaboration
explain the basics of issues #20

@@
 
+### GitHub Issues
+
 <!--Hands on: Introduce GitHub Issues-->
+Effective collaboration needs effective organization.
+GitHub offers a ticket system called "Issues".
+Here you and your collaborators can discuss and distribute tasks.
+
+To take a look at the issue page of your project, you may use:
+
+``` r
+usethis::browse_github_issues()
+```
+
+```{r issue-screenshot, include=FALSE}
+# unfortunately css selectors won't work for pdf so we'll use png instead
+issue <- here::here("images", "issue.png")
+if(!file.exists(issue)){
+  webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/issues/20",
+                  file = issue,
+                  selector = "div#repo-content-pjax-container", # only select file change
+                  zoom = 2)# higher resolution
+}
+```
+
+```{r issue, echo=FALSE, fig.cap="A screenshot of an example issue.  ", out.width='100%'}
+knitr::include_graphics(issue)
+```
+
+One advantage of using the issue system of GitHub is that it enables you to link changes you made to their tasks by including the issue number (i.e. #20) in the commit message:
+
+``` r
+# do not run this
+gert::git_add(c("manuscript.Rmd", "references.bib"))
+gert::git_commit("explain the basics of issues #20")
+```
+
+While it is a valuable device for working with direct collaborators, it is also great to facilitate review [@rougierSustainableComputationalScience2017].
+Even if a journal does not support this, it enables post-publication review.
+Readers of your manuscript can ask questions or point out how to improve it after you have published it.
+
+### GitHub Pull Requests
+
+<!--Hands on: Introduce GitHub PR and review-->
+
+### GitHub Project Management
+
 <!--Hands on: Introduce GitHub Projects-->

### 2021-05-03 manuscript.Rmd
add section on pr

@@
 ``` r
-# do not run this
+# do not run:
 gert::git_add(c("manuscript.Rmd", "references.bib"))
@@
 
+Pull requests allow anyone to contribute to a project, even though they do not have write access, while the authors still retain all control.
+If someone wants to contribute, but they have no permission to write to a repository, they can create a copy of the project, called "fork", where they can push their changes.
+They then request that you (or anyone with write permissions) "pull" these changes.
+Any new pull request can be seen on GitHub:
+
+``` r
+usethis::browse_github_pulls()
+```
+
+Before you incorporate changes from someone else, you can review and change them.
+This review step can also be useful for collaboration, because you can invite a collaborator to take a look at changes you want to do.
+It is therefore sometimes useful to create a pull request even though you have permission to push, because it invites your collaborators to comment or change your proposal.
+
+You can initiate a new pull request with:
+
+``` r
+usethis::pr_init("myrevision")
+```
+
+Then you make and commit your changes and open a new pull request with:
+
+``` r
+usethis::pr_push()
+```
+
+After you or someone else has merged the pull request on GitHub, you can wrap it up with:
+
+```r
+usethis::pr_finish()
+```
+
+This process is especially useful if you want the feedback of a more senior researcher, because they do not need to know how to use Git but have a convenient web interface to provide comments or changes.
+
 ### GitHub Project Management
@@
 <!--Hands on: Introduce GitHub Projects-->
+<!--I love it, but maybe we should remove it-->
+While issues and pull requests already provide much structure to organize your work, GitHub projects allow you to categorize and prioritize issues and pull request, while informing you on the progress you made.
 

### 2021-05-04 manuscript.Rmd
small changes after reading

@@
 Another advantage of Markdown is that you can easily version-control it with Git.
-You may be familiar with Microsoft Words' "Track Changes"-Feature and version control is similar in their basic idea but more potent in their features.
+You may be familiar with Microsoft Words' "Track Changes"-Feature.
+Version control is similar in its basic idea but more potent in its features.
 Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
@@
 
+You see the linkage in Figure \@ref(fig:issue).
+
 While it is a valuable device for working with direct collaborators, it is also great to facilitate review [@rougierSustainableComputationalScience2017].
@@
 If someone wants to contribute, but they have no permission to write to a repository, they can create a copy of the project, called "fork", where they can push their changes.
-They then request that you (or anyone with write permissions) "pull" these changes.
-Any new pull request can be seen on GitHub:
+For example if you found an error in this manuscript, you can not simply change it, because you have no write access.
+However, you can create a fork in your account, where you have write permission:
+
+``` r
+usethis::create_from_github("aaronpeikert/repro-tutorial", fork = TRUE)
+```
+In this copy you then make any changes you want.
+You could then request that we (or anyone with write permissions) "pull" these changes.
+You can view new pull request for any project with:
 
@@
 
-Before you incorporate changes from someone else, you can review and change them.
+Before incorporating changes from someone else, they can be reviewed and adepted.
 This review step can also be useful for collaboration, because you can invite a collaborator to take a look at changes you want to do.

### 2021-05-04 manuscript.Rmd
explaint how to reproduce a project close #22 related #25

@@
 
+## Reproduce a Project
+
+A reproducible project should not only be theoretically reproducible, but it should be simple to reproduce.
+Ideally, you have a statement in the README that contains a few steps to reproduce the project, where most steps are automatic and standardized.
+This statement should also note the required software.
+E.g. you can download any git repository with these commands:
+
+``` bash
+# this should be run in the terminal
+git clone https://github.com/aaronpeikert/repro-tutorial.git
+cd repro-tutorial
+```
+
+Projects that combine Make and Docker then can usually be reproduced with:
+
+``` bash
+# this should be run in the terminal
+make docker &&
+make -B DOCKER=TRUE 
+```
+
+<!-- maybe we add here the real data? #25 -->
+
 ## If you can not make the data public (optional)

### 2021-05-04 manuscript.Rmd
explain interplay of Docker and Make close #13

@@
 
-This command will go through all R Markdown files, and their YAML metadata, in your repository and infer a Makefile from them:
+This command will go through all R Markdown files, and their YAML metadata, in your repository and infer a `Makefile` from them:
 
@@
 ```
+
+The generated `Makefile` automatically contains targets for all R Markdown documents.
+To, for example, generate the `README.md` from the `README.Rmd` you can run:
+
+``` bash
+# this should be run in the terminal
+make README.md
+```
+
+It is common practice to have a target all, that encompasses all important files of the project as the first target.
+If you do not specify a target, Make defaults to the first target, which is usually `all`.
+`repro` reminds you of that and suggests which file you might want to add to the target `all`.
+After you have done so, you can then run in the terminal:
+
+``` bash
+# this should be run in the terminal
+make
+```
+
+And it will fully reproduce the project, if you have all required software installed.
 
@@
 Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
+For archiving, you therefore have to store the image, as explained in [Archiving Docker].
 The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:
@@
 
-With the R version that you currently use, this specification is then translated into a `Dockerfile`.
+With the R version that you currently use, this specification is then translated into a `Dockerfile`:
 
@@
 automate_docker()
+```
+
+From the `Dockerfile` you can then build an image.
+While you can do so manually we prefer to use Make, see [Internal Dependencies].
+The `Makefile` build by `repro` does contain a recipe for creating the `Dockerfile`.
+
+``` bash
+make docker
+```
+
+The `Makefile` from `repro` does also include a mechanism to send commands conditionally through the Docker container.
+This mechanism is triggered by `DOCKER=TRUE` in the make call:
+
+``` bash
+# this should be run in the terminal
+make DOCKER=TRUE
 ```
@@
 Anyway, from within the container i.e. for your analysis it does not make any difference.
+Like with Docker, the default `Makefile` of `repro` has a mechanism to automatically send all commands to Singularity.
+It is triggered by `SINGULARITY=TRUE`:
+
+``` bash
+make SINGULARITY=TRUE
+```
 

### 2021-05-04 manuscript.Rmd
remove space in code chunk header closes #29

@@
 "
-cat("``` markdown", markdown_example, "```", sep = "")
+cat("```markdown", markdown_example, "```", sep = "")
 ```
@@
 
-``` markdown
+```markdown
 <!--this is a Markdown file -->
@@
   )
-cat("``` r",
+cat("```r",
     model_truth,
@@
 
-``` r
+```r
 usethis::browse_github_issues()
@@
 
-``` r
+```r
 # do not run:
@@
 
-``` r
+```r
 usethis::create_from_github("aaronpeikert/repro-tutorial", fork = TRUE)
@@
 
-``` r
+```r
 usethis::browse_github_pulls()
@@
 
-``` r
+```r
 usethis::pr_init("myrevision")
@@
 
-``` r
+```r
 usethis::pr_push()
@@
 
-``` r
+```r
 automate_make()
@@
 
-``` bash
+```bash
 # this should be run in the terminal
@@
 
-``` r
+```r
 automate_docker()
@@
 
-``` bash
+```bash
 make docker
@@
 
-``` bash
+```bash
 make SINGULARITY=TRUE

### 2021-05-04 manuscript.Rmd
added some text to the introduction

@@
 
-<!-- define reproducibility -->
+A cornerstone of scientific progress is the possibility to critically assess the validity of peer scientists’ scientific conclusions.
+Scientific claims should not be credible because of their originators' authority but by the transparency and replicability of their supporting evidence [@aac4716].
+Enabling this assessment requires a systematic and rigorous approach to research.
+<!-- define replicability and then... 
+ define reproducibility -->
 <!-- is a direct quote from my master thesis -->
@@
 <!-- lure the reader in -->
-The research community has long accepted reproducibility as part of "good scientific practice" but has not benefited much from this self-imposed requirement.
 <!-- expand: who is calling it a requirement, maybe metascientific arguments-->
-Reproducing an empirical study is often too tedious and frustrating work for the researchers to benefit them.
-However, if reproduction would require mere minutes instead of hours, it could greatly facilitate collaboration within and across research projects.
+In fact, the the research community has long accepted reproducibility as an important element of “good scientific practice.
+<!-- examples missing --> However, ensuring reproducibility of computational analyses is still tedious and difficult, and there are still no widely accepted standards that can guide researchers when setting up their data analyses.
+Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.
+For example, @hardwicke2018 attempted to replicate open materials in the journal Cognition.
+Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
+For 13 articles, at least one outcome could not be reproduced — even with the original authors' assistance.
+@obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
+The authors could execute only 31 of the scripts without error and were able to reproduce the results of only 21 articles (within a reasonable period of time).
+Thus, there is a dire need to establish standards for reproducible workflows in order to move current research from just being transparent to becoming a cumulative computational knowledge acquisition process.
+
+From our personal experience, the lack of a proper incentive structure and the dreary perspective of having to spend several extra hours if not days to properly set of a code repository to become reproducible, makes many scholars skip this extra step.
+If reproduction would require merely minutes instead of hours, it could greatly facilitate collaboration within and across research projects.
 <!-- Guiding principle: spending machine compute time instead of human research time -->

### 2021-05-04 manuscript.Rmd
each sentance on its own line for better diffs
added some text to the introduction (#31)
* added some text to the introduction

* each sentance on its own line for better diffs

* correct quotes

Co-authored-by: Aaron Peikert <aaron.peikert@posteo.de>

@@
 <!-- expand: who is calling it a requirement, maybe metascientific arguments-->
-In fact, the the research community has long accepted reproducibility as an important element of “good scientific practice.
-<!-- examples missing --> However, ensuring reproducibility of computational analyses is still tedious and difficult, and there are still no widely accepted standards that can guide researchers when setting up their data analyses.
+In fact, the the research community has long accepted reproducibility as an important element of "good scientific practice." <!-- examples missing -->
+However, ensuring reproducibility of computational analyses is still tedious and difficult, and there are still no widely accepted standards that can guide researchers when setting up their data analyses.
 Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.

### 2021-05-04 manuscript.Rmd
explain pull and push close #21
explain .gitignore closes #9
add a reference for learning Git
closes #23

@@
 <!--I am troubled by how we should recommend to learn Git-->
-Git can be challenging to learn because we are used to learning alone, and Git is best learned with others.
-We recommend finding a partner to attend an online lecture or workshop to learn the ins and outs of Git.
+Git can be challenging to learn, because it requires a rather complex mental model.
+However, there is a subsets of commands which we present here and in the section on [GitHub] (commit, push, pull) which enable you to utilize Git for simple version control.
+There are many tutorials on learning Git and @vuorreCuratingResearchAssets2018 provide a good starting point.
+However Git is designed for collaboration, we therefore recommend finding a partner to attend an online lecture or workshop to learn the ins and outs of Git.
 
@@
 Only from this moment on files are uploaded to the internet and only those you have explicitly added and committed in Git.
+
+After you have initially uploaded your repository to GitHub, you need to regularly synchronize your local copy and GitHub.
+Uploading commits to GitHub is called "push":
+
+```r
+gert::git_push()
+```
+
+Downloading changes, e.g. from your collaborators is called "pull".
+
+```r
+gert::git_pull()
+```
 

### 2021-05-04 manuscript.Rmd
add reference for Docker #24

@@
 
+You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manuel` to install software beyond R packages.
+Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manuel`.
+We recomend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
+
 <!-- after this everything is optional? -->

### 2021-05-04 manuscript.Rmd
more changes to intro

--- 2021-05-04 manuscript.Rmd
@@
 Scientific claims should not be credible because of their originators' authority but by the transparency and replicability of their supporting evidence [@aac4716].
-Enabling this assessment requires a systematic and rigorous approach to research.
+Here, _replicability_ refers to the degree to which original studies and their derivatives yield consistent results.
+In other words, a succesful replication study finds a consistent effect with an original study in a new sample of data.
+Enabling the assessment of scientific conclusions requires a systematic and rigorous approach to research.
+Researchers across various fields have identified a _crisis_ of non-replicability of findings [e.g.
+@aac4716]- <!--citation needed, e.g.
+[@aac4716]-->
+
 <!-- define replicability and then... 
  define reproducibility -->
-<!-- is a direct quote from my master thesis -->
-<!-- I wouldn't want to cite it. @brandmaier, would you please rewrite? -->
-A reproducible scientific product allows other researchers to obtain the same results from the same dataset in a way that enables substantive criticism and therefore facilitates replication.
-<!-- lure the reader in -->
-<!-- expand: who is calling it a requirement, maybe metascientific arguments-->
-In fact, the the research community has long accepted reproducibility as an important element of "good scientific practice." <!-- examples missing -->
+Even worse, researchers might often not even be able to achieve reproducibility -- defined as the ability to obtain identical results from the _same_ data with the _same_ computer code.
+Over the past decades, computational statistical methods have radically changed the ability of researchers from all disciplines to process and analyze data with complex statistical models.
+With these advances came challenges that contribute to broader concerns of irreproducibility in the published literature, among them the lack of exact and unambiguous documentation of computational methods and each computational step taken.
+With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves.
+Only with access to both data and procedures can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.
+Only if both the data and code are shared, can we obtain a reproducible scientific product, facilitating assessment of its validity and later reuse.
+
+In fact, the research community has long accepted reproducibility as an important element of "good scientific practice." <!-- examples missing -->
 However, ensuring reproducibility of computational analyses is still tedious and difficult, and there are still no widely accepted standards that can guide researchers when setting up their data analyses.
@@
 The authors could execute only 31 of the scripts without error and were able to reproduce the results of only 21 articles (within a reasonable period of time).
-Thus, there is a dire need to establish standards for reproducible workflows in order to move current research from just being transparent to becoming a cumulative computational knowledge acquisition process.
+Thus, there is a dire need to establish standards for reproducible workflows for scientific rigor.
 
@@
 <!-- Guiding principle: spending machine compute time instead of human research time -->
-To archive this boost, we must place the burden of reproducing something upon computers instead of human researchers.
-A scientific document should therefore be understood by researchers but reproduced by computers.
-
-This tutorial walks through the creation of such a document that can be reproduced automatically.
-Though reproducibility is a small part of the research process, it can serve as the mortar of open science building blocks (e.g. preregistration, open data, postpublication review).
-Therefore, we show the whole lifecycle of an open research project, where some parts are essential to archive reproducibility.
-All code chunks are meant for the reader to try out for themselves.
-If not declared otherwise the code is meant to run in R (tested with version `r with(R.version, paste0(major, ".", minor))`).
-Still, the reader can safely skip sections marked as optional when unfit for their research question or methods.
-However, any empirical data analysis will suffer from the following threats to reproducibility:
+To achieve this boost in productivity, we must place the burden of reproducibility upon computers instead of the researchers.
+To this end, we demand that a scientific product (no matter whether this is a preliminary report or a full research article) should both be understandable by researchers and automatically reproducible by computers with minimal constraints to the researchers' own workflows.
+
 <!-- Which problems are solved by which tool (like in Andreas talk)-->
 <!--stolen from: https://brandmaier.github.io/reproducible-data-analysis-materials/KULeuvenQuantPsy2020.html#11 -->
-
-1. Multiple versions of scripts/data (e.g., the dataset has changed over time, i.e., was further cleaned or extended)
-2. Copy&paste errors (e.g., inconsistency between reported result and reproduced result)
-3. unclear which scripts should be executed in which order
-4. Broken software dependencies (e.g., analysis broken after an update, missing package)
-
-It is common to rely on the craftiness of the researcher to debug these problems, but we could prevent them altogether with:
+We identified the following common four causes for non-reproducibility:
+
+1.	Multiple inconsistent versions of code and data; for example, the dataset has changed over time because outliers were removed at a later stage, or the code has changed because a bug was removed at some point in time;
+2.	Copy-and-paste errors; for example, results are manually copied from a computing environment into a text processor, causing inconsistencies between the reported result and the reproduced result;
+.3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;
+4.	Broken dependencies; for example, a given analysis depends on a software package that might not be available on a different computer or no longer exists.
+
+We have eveloped a workflow that leverages established tools and practices from software engineering to achieve long-term computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:
 
@@
 
-We deem each concept necessary to archive long-term reproducibility.
 However, which tools you use to implement these concepts is a matter of taste and project requirements.
-This tutorial follows the recommendations of @Peikert2019 and utilize Git for version control, R Markdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
+This tutorial follows the recommendations of @Peikert2019 and use Git for version control, R Markdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
 Their interplay is shown in Figure \@ref(fig:schematic).
@@
 
+This tutorial walks through the creation of a reproducible document that can be reproduced automatically.
+Though reproducibility is a small part of the research process, it can serve as the mortar of open science building blocks (e.g.
+preregistration, open data, postpublication review) -- but not that even non-open science will benefit from ensuring reproducibility because good scientific practice demands that every scientist is able to reproduce their own results later in time.
+Therefore, we show the whole lifecycle of an open research project, where some parts are essential to archive reproducibility.
+All code chunks are meant for the reader to try out for themselves.
+If not declared otherwise the code is meant to run in R (tested with version `r with(R.version, paste0(major, ".", minor))`).
+Still, the reader can safely skip sections marked as optional when unfit for their research question or methods.
+
 # Setup
@@
 
-We assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/install.md", "installation guide")`).
-Additionally, you'll need the [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:
+This tutorial is targeted at scholars with at least intermediate R knowledge and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/install.md", "installation guide")`).
+To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:
 
@@
 
-The [`repro`-package](https://github.com/aaronpeikert/repro) contains several helpers to work with reproducibility tools.
-For example, it can check if your computer is set up to use the tools we rely on:
+The [`repro`-package](https://github.com/aaronpeikert/repro) contains several helpers to work with the suggested tools for reproducibility.
+For example, the package can check if your computer has all the required software packages installed and ready.
 
@@
 check_git()
-```
-
-```{r}
-# -- type this on the R console --
 check_make()
-```
-
-```{r}
-# -- type this on the R console --
 check_docker()
@@
 
-If these commands detect that something is not installed or set up, they will give you detailed instructions (tailored to your operating system) to remedy the situation.
-Follow the instructions, then rerun the command until it tells you not to worry.
-
-# Planing
+If these commands detect that something is not installed or set up properly, they will alert you and give you detailed instructions (tailored to your operating system) to set up your computer.
+Please make sure that all these commands pass before you continue with the tutorial.
+
+# Planning
 
@@
 While you might be tempted to consider reproducibility only when you about to finish a project, upfront investments pay off through your increased productivity.
-Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you archive reproducibility gradually.
+Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you achieve reproducibility gradually.
 The first step is to bundle all required files in one folder.

### 2021-05-04 manuscript.Rmd
reformat

@@
 Scientific claims should not be credible because of their originators' authority but by the transparency and replicability of their supporting evidence [@aac4716].
-Here, _replicability_ refers to the degree to which original studies and their derivatives yield consistent results.
-In other words, a succesful replication study finds a consistent effect with an original study in a new sample of data.
+Here, *replicability* refers to the degree to which original studies and their derivatives yield consistent results.In other words, a succesful replication study finds a consistent effect with an original study in a new sample of data.
 Enabling the assessment of scientific conclusions requires a systematic and rigorous approach to research.
-Researchers across various fields have identified a _crisis_ of non-replicability of findings [e.g.
-@aac4716]- <!--citation needed, e.g.
-[@aac4716]-->
+Researchers across various fields have identified a *crisis* of non-replicability of findings [e.g. @aac4716]- <!--citation needed, e.g. [@aac4716]-->
 
@@
  define reproducibility -->
-Even worse, researchers might often not even be able to achieve reproducibility -- defined as the ability to obtain identical results from the _same_ data with the _same_ computer code.
+Even worse, researchers might often not even be able to achieve reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code.
 Over the past decades, computational statistical methods have radically changed the ability of researchers from all disciplines to process and analyze data with complex statistical models.
@@
 Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
-For 13 articles, at least one outcome could not be reproduced — even with the original authors' assistance.
+For 13 articles, at least one outcome could not be reproduced —-- even with the original authors' assistance.
 @obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
@@
 Though reproducibility is a small part of the research process, it can serve as the mortar of open science building blocks (e.g.
-preregistration, open data, postpublication review) -- but not that even non-open science will benefit from ensuring reproducibility because good scientific practice demands that every scientist is able to reproduce their own results later in time.
+preregistration, open data, postpublication review) --- but not that even non-open science will benefit from ensuring reproducibility because good scientific practice demands that every scientist is able to reproduce their own results later in time.
 Therefore, we show the whole lifecycle of an open research project, where some parts are essential to archive reproducibility.

### 2021-05-04 manuscript.Rmd
add missing rocker citation close #24
master -> main closes #33

@@
 
-This tutorial is targeted at scholars with at least intermediate R knowledge and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/install.md", "installation guide")`).
+This tutorial is targeted at scholars with at least intermediate R knowledge and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md", "installation guide")`).
 To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:
@@
 Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
-For example, this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/master", "iterations")`.
+For example, this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/main", "iterations")`.
 To activate Git in a given project, you can call:
@@
 
-Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/R/simulation.R", "GitHub")`.
+Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R", "GitHub")`.
 
@@
 When you include code into an R Markdown, this code is run, and its results are included in the rendered document.
-With R Markdown, you can write whole articles (`r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/manuscript.Rmd", "like the one you are reading")`).
+With R Markdown, you can write whole articles (`r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/manuscript.Rmd", "like the one you are reading")`).
 You could, in principle, even write large parts of the manuscript as part of your preregistration, i.e. the introduction, theoretical background, the method section and parts of the discussion.
@@
 Of course, there a several ways to solve this problem.
-One could build an R package to separate analysis and data or use code externalization in the RMarkdown to facilitating reuse or create a "master" script that combines all steps.
+One could build an R package to separate analysis and data or use code externalization in the RMarkdown to facilitating reuse or create a "main" script that combines all steps.
 We found Make to be particularly useful for two reasons.
@@
 First, there is a community that maintains images for particular purposes.
-For example, using R [rocker-citation].
+For example, using R [@boettigerIntroductionRockerDocker2017].
 Such images only contain what is necessary for a specific purpose.
@@
 Second, instead of saving a binary image of the software, one can also save a textual description of the programmatic steps to install the software in a `Dockerfile`.
-Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
+Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
 For archiving, you therefore have to store the image, as explained in [Archiving Docker].
@@
 
-[^simhash]: https://github.com/aaronpeikert/repro-tutorial/blob/master/R/hash.R
+[^simhash]: https://github.com/aaronpeikert/repro-tutorial/blob/main/R/hash.R
 

### 2021-05-04 manuscript.Rmd
git2r -> gert
closes #16

@@
     - usethis
-    - git2r
+    - gert
     - aaronpeikert/repro@7bfaf98
@@
 Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
-For example, this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/main", "iterations")`.
+For example, this paper has been going through `r system2("git", c("rev-list",  "--count",  "HEAD"), stdout = TRUE)` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/main", "iterations")`.
 To activate Git in a given project, you can call:
@@
 
-And you could commit it in Git:
+And you could commit it in Git.
+For convenience, we use the [`gert`-package](https://github.com/r-lib/gert) [@R-gert] to interface with Git, but it does not make a difference:
 <!--I am leaning towards doing everything in R, not sure though-->
@@
 ```r
-git2r::add(".", "idea.md")
-git2r::commit(".", "add a first concept")
+gert::git_add("idea.md")
+gert::git_add("add a first concept")
 ```
@@
 # -- type this on the R console --
-git2r::add(".", "idea.md")
-git2r::commit(".", "exclude bias in measurement as possible confounder")
+gert::git_add("idea.md")
+gert::commit("exclude bias in measurement as possible confounder")
 ```
@@
 ```{r, include=FALSE}
-knitr::write_bib(c(.packages(), "bookdown", "lavaan", "knitr", "targets", "renv"),
+knitr::write_bib(c(.packages(), "gert", "bookdown", "lavaan", "knitr", "targets", "renv"),
                  "packages.bib")

### 2021-05-04 README.Rmd
add reproduction statement to readme closes #27

--- 2021-01-07 README.Rmd
@@
 
-# A quick tour through the `repro`-package
+# A Hitchhiker's Guide to Reproducible Research
+<!-- badges: start -->
+[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
+<!-- badges: end -->
 
 WIP
+
+## How to reproduce this manuscript
+
+To reproduce this project Git, Make, and Docker is required (see [the installation guide](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)).
+
+Open the terminal, download the repository, and enter the directory:
+
+```bash
+git clone https://github.com/aaronpeikert/repro-tutorial.git
+cd repro-tutorial
+```
+
+Then build the Docker image, and run Make:
+
+```bash
+make docker &&
+make -B DOCKER=TRUE 
+```
+
+## Code of Conduct
+
+Please note that the repro-tutorial project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).
+By contributing to this project, you agree to abide by its terms.

### 2021-05-04 manuscript.Rmd
add missing newline
setup, only once per machine closes #6
mention our reproduction statement close #28 related #12
some changes in the example git_add(). I think there was an error in l.220. It should be git_commit() instea of git_add(), right?

@@
 Please make sure that all these commands pass before you continue with the tutorial.
+This process is only required once per computer and does not need to be repeated per project.
 
@@
 
-Could we use openpsychometrics.org data?
+Can we use openpsychometrics.org data?
 ```
@@
 And you could commit it in Git.
-For convenience, we use the [`gert`-package](https://github.com/r-lib/gert) [@R-gert] to interface with Git, but it does not make a difference:
+For convenience, we use the [`gert`-package](https://github.com/r-lib/gert) [@R-gert], which wraps most git commands into R functions.
+Alternatively, we suggest to either work directly in the terminal (for example, RStudio offers a command shell with various options accessible in the Menu Tools -> Terminal), or work using the Git pane in RStudio that offers buttons for basic operations.
 <!--I am leaning towards doing everything in R, not sure though-->
@@
 
+To commit your first changes to the project, start with staging all files to be included in the commit using `git_add()`.
+Here, we stage our idea collection in the file `idea.md` and then commit it to our local repository with a meaningful commit message that will later help us keep track of all the changes over time.
+
 ```r
 gert::git_add("idea.md")
-gert::git_add("add a first concept")
+gert::git_commit("add a first concept")
 ```
@@
 This statement should also note the required software.
+Figure \@ref(fig:readme) shows `r link("https://github.com/aaronpeikert/repro-tutorial#how-to-reproduce-this-manuscript", "the reproduction statement of this project")`.
 E.g. you can download any git repository with these commands:

### 2021-05-04 manuscript.Rmd
some more suggestions re modeling

@@
 
-t-test
+Independent t-test
 
@@
 
-But then you realize that the new literature suggests that there could be a bias in measurement:
+Now assume that after discussing with your colleagues, you settle onto a final analysis strategy using latent variable models with tests for measurement invariance because power is greater and you want to exclude bias in measurement as possible confounder <!-- Aaron, reference needed-->.
+This results in changes to your notes as follows:
 
@@
 
-Multigroup CFA + Measurement Invariance
+Multigroup Confirmatory Factor Analysis
+First, establish strong measurement invariance using a CFI cutoff-criterion
+Second, test for differences in latent factor means of Machiavellianism using a $\chi^2$ test
 
@@
 
-Commit this version:
+Again, stage the changes to this file and commit like before using `git_add()` and `git_commit()` or use the following shorthand to stage all new and modified files and then commit:
 
@@
 # -- type this on the R console --
-gert::git_add("idea.md")
-gert::commit("exclude bias in measurement as possible confounder")
+gert::commit_all("switch to CFA to exclude bias in measurement as possible confounder")
 ```
@@
 
-It is probably better to assume more pessimistic values.
-This way, we make sure that our analytic strategy is viable even in non-optimal circumstances.
-Therefore, we will assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.2.
+Typically, it is recommended to choose conservative values during study design planning  [@brandmaier2018precision].
+This way, we make sure that our analytic strategy is viable even in non-optimal circumstances, for example if the measurement instrument works less well then expected or the population characteristics are more different than expected.
+As a heuristic guess, we assume that the factor loadings are 30% lower than reported and that the standardized mean difference is only 0.2.
 These assumption can be simply translated into [`lavaan` [@R-lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
@@
 
-From which we can simulate data:
+Structural Equation Models are generative models that not only fit models but lend themselves to data simulation. This is particularly easy in lavaan syntax using the function `simulateData`.
 

### 2021-05-04 manuscript.Rmd
some more changes and a lot of "AP" call-outs with tasks for Aaron

@@
 There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
-A step to prevent unpleasant surprises is to simulate data.
+A step to prevent unpleasant surprises is to simulate data.<!-- AP: this paragraph deserves some references.
+I assume those can be found in the Bayesian literature.
+But are there classic SEM literature references talking of the value of simulation.
+Please make a lit research -->
 Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
@@
 
-If you are unfamiliar with SEM, you may skim over the next sections.
-To simulate data for SEM, one could specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
-The model parameters should express our expectations for the Machiavellism subscale of the short dark triad (SD3).
+If you are unfamiliar with the specifics of SEM, we you may skip the next sections and continue in section XXX <!--AP: fill in reference.-->.
+To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.
+
+The selected model parameters should express our expectations for the Machiavellism subscale of the short dark triad (SD3).
+<!-- AP: reference needed; explain the scale.
+What is the construct that we are measuring here?
+-->
 As a starting point, we can use already published information.
-@Jones2013 report loadings seen in Table \@ref(tab:loadings) and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
+@Jones2013 report a set of factor loadings (see Table \@ref(tab:loadings)) <!--AP what was the reference population?
+what was the sample size?
+--> and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
+<!--AP why not on the latent level?
+They report factor loadings after all?
+Please explain-->
 
@@
 
-Typically, it is recommended to choose conservative values during study design planning  [@brandmaier2018precision].
+Typically, it is recommended to choose conservative values in the study planning phase [@brandmaier2018precision].
 This way, we make sure that our analytic strategy is viable even in non-optimal circumstances, for example if the measurement instrument works less well then expected or the population characteristics are more different than expected.
-As a heuristic guess, we assume that the factor loadings are 30% lower than reported and that the standardized mean difference is only 0.2.
+As a heuristic guess, we assume that the factor loadings are 30% lower than those reported by @Jones2013 and that the standardized mean difference is only 0.2.
 These assumption can be simply translated into [`lavaan` [@R-lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
@@
 
-Structural Equation Models are generative models that not only fit models but lend themselves to data simulation. This is particularly easy in lavaan syntax using the function `simulateData`.
+Data simulation is particularly easy in lavaan syntax using the function `simulateData`.
+More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem <!--AP please add reference, and possibly other packages-->.
+The following code simulates data from the population model `model_truth` with 1.000 cases in each group.
 
@@
 
-But wait, `lavaan` warns us that this model may not be identified.
+Now, `lavaan` warns us that this model may not be identified.
 We overlooked a crucial assumption of this model, that is, measurement invariance.
@@
 There is a good chance that this assumption is violated.
-We, therefore, adept our analysis to plan for this contingency.
+We, therefore, adapt our analysis to plan for this contingency.
 After fitting the model with strict measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
@@
 
-To consistently get the same results, it is good practice to set a seed with `set.seed(1235)`.
-It is a little more complicated to archive uncorrelated random numbers when you have parallelized your code like we did, but the future package handles this for us:
-
+Strict reproducibility is at stake with any code that is based on the computer's pseudo random number generator (PRNG) because random numbers will almost always differ with each execution of the code.
+To consistently get the same results while preserving the pseudo-random property of the generated numbers, it is good practice to set a random seed that sets the PRNG into a restorable state, in R code: `set.seed(1235)`.
+It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster), but the `future` <!--AP add citation --> package handles this for us:
+<!--AB I don't understand how this solves the problem yet - please explain in more detail -->
 ```{r}
@@
 
-Preregistration serves as a tool that may help to increase the credibility of empirical results.
-Without getting lost in the meta-scientific arguments for preregistration, we want to note two advantages.
+Preregistration serves as a tool that may help to increase the credibility of empirical results. <!-- AP citation needed-->
+Without getting into too much detail of the meta-scientific arguments for preregistration, we want to note two advantages. <!-- AP can we add a citation to someone who has the arguments?-->
 First, a preregistration is an opportunity to plan a study carefully.
@@
 
-One of the most significant benefits of open research is that other people can freely reuse materials.
+One of the most significant benefits of open research is that other people can freely reuse materials, ideally including everything from stimuli, over computerized experiments, to data analysis code.
 To use your materials, they first have to find them and learn how to use them.
 Interested users find out about your project via the readme.
-The readme is the first thing they see when they visit your project on [GitHub].
-It should explain to them what your project is about and how to reproduce it.
-You could, for example, include the abstract of your (planned) paper here.
+Historically\footnote{Apparently, the use of readme files can be traced back to at least 1974 <!--AP reference needed, I guess-->}, a readme file is the default place to look for human readable information in a computer archive
+the first thing they see when they visit your project on [GitHub].
+It is supposed to explain to visitors what your project is about and how to reproduce it.
+You could, for example, include the abstract of your (planned) paper here.<!-- AP: shouldn't this be a function in repro? to auto-insert the abstract?-->
 
@@
 
-Even when other researchers are excited to use your published code or materials, they are usually legally inhibited from doing so.
-In many countries (i.e. the USA, European Union), the creator of something has exclusive right to use their creation.
-A license allows other people to use what you have created.
+Note that publishing data, code, or other materials is not sufficient to allow other users to use it, modify it, or redistribute it.
+For example, according to German copyright law, a repository without a license must be assumed to be protected by copyright <!-- AP citation needed -->.
+In many countries (i.e.
+the USA, European Union), a content creator has exclusive right to use their creation.
+A license is the legal authorization of other people to use what you have created.
 <!---exact copy from our previous manuscript, do we need to change? -->
-In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media.
+In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media (such as stimulus materials if you created them).
 In contrast, scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
@@
 ```
+
+<!-- AP: Can we show how to set MIT license for code an CCBY license for the manuscript? Can we make a repro function that creates such a double-license? How to do it actually? Two license files? -->
 

### 2021-05-05 manuscript.Rmd
add more details to release. Spelling out your thoughts in more detail

@@
 
-GitHub has a "release" function, which creates a link to a specific version of your project, where everything is available as a zip file.
-You have the opportunity to describe the release in the release notes and to upload other additional files (e.g. the rendered pdf).
+In software development, a release is the distribution of a piece of computer software indicating a certain maturity in development and level of completeness.
+GitHub has a "release" function, which creates a link to a specific snapshot of your project that then is available as a single compressed file.
+We believe that the idea of software releases can be leveraged for singling out important versions in a research project.
+We found it useful to label the following stages of development as release:
+- Preregistration (repository contains the final data analysis plan; also see the Section on PAC)
+- Completion of data collection (repository contains all study data)
+- Preprint (repository exactly reproduces a preprint that is hosted externally)
+- Submission or Resubmission (repository exactly reproduces a given version (re)submitted to a journal)
+- In Press (repository exactly reproduces the final manuscript as accepted by a scientific journal)
+
+You have the opportunity to describe the release in more detail in the accompanying release notes and to add extra files (e.g., complementing the Markdown source code of a journal article with a rendered PDF document).
 

### 2021-05-05 manuscript.Rmd
some more changes re closed data

@@
 
-## If you can not make the data public (optional)
+## Closed Data (optional)
 
 <!-- Hands-on: check changed hash -->
-
-To reproduce an analysis, you need to have access to the data.
-There are several reasons which may prohibit you from making your data available, like legal or regulatory or privacy concerns.
-Often you find yourself in a situation where you can make the data only availabe on request.
-A possibility for irreproducibility is that the data has changed in the meantime.
-To at least check that the data was not changed you can provide a so called "hash".
-This is a cryptographic fingerprint.
+A prerequisite to reproducibility is the availability of code and data <!-- AP: can you add the reference to this tight reproducibility requirements that you told me about?-->.
+We use containers to guarantee that all code components are available and work as expected.
+Ideally, all data is also stored in the repository.
+In practice, however, there are various reasons, which may prohibit you from making your data available, for example legal or regulatory or privacy concerns.
+You may find yourself in a situation where you can make the data only available on request. <!-- AP please cite the SHARE data and explain as an example what one has to do to ontain these data -->.
+A major threat to reproducibility is that the data has changed in the meantime or multiple versions are circulating and it is not exactly clear what version was used originally.
+To at least check that the data was not changed, we suggest to provide a so called "hash", which is a cryptographic fingerprint.
 If the data changes, so does its fingerprint or hash.
@@
 ## Creating a DOI
+<!-- AP add references; why are DOIs helpful-->
 

### 2021-05-05 manuscript.Rmd
add statement on non-reproducibility if lack of structure and docs

@@
 
-A reproducible project should not only be theoretically reproducible, but it should be simple to reproduce.
+A reproducible project should not only be reproducible in theory, but it should be simple to reproduce.
+Typically, it is argued that results should be reproducible by a Ph.D.
+student from the same discipline who had experience with the statistical software used [@obels2020].
+Even if all code and data are available in a repository, if reproducibility takes too long because of a lack of structure and documentation, it must be regarded as non-reproducible [@obels2020].
+
 Ideally, you have a statement in the README that contains a few steps to reproduce the project, where most steps are automatic and standardized.

### 2021-05-05 manuscript.Rmd
more on singularity. also changed headline

@@
 
-### Singularity (optional)
-
-One usecase in which Docker sometimes can not be used is in high performance computing (HPC).
-The simulation we did is computationally expensive enough to warrant the use of a HPC cluster.
-To run the simulation on a cluster we require the exact same software, not only on one or two computers, but potentially hundreds.
-Docker is a natural solution for this problem and is the defacto standard for cloud computing.
-However, Docker requires admin rights, while this is no issue on cloud computing infrastructure it is an issue in classical HPC environments.
-Singularity was designed to run without admin rights, while maintaining full compatibility with Docker images.
+### High-performance computing environments (optional)
+
+The `repro` package is particularly well-suited for computationally intensive data analyses and simulations that  are to be run on distributed computing environments.
+By using containers, `repro` ensures identical execution across platforms and thus guarantees that a data analysis or simulation that was designed and tested on a local computer will be executed in the identical manner on a remote machine or various remote machines across an entire computing cluster.
+Thereby, the repro package makes massive parallel computing in heterogeneous computing environments less error-prone while ensuring the full documentation of all computational steps taken — even if some or all of them were run in parallel on distributed platforms.
+<!-- AP can we add examples, are there open repositores eg from neuroscience that have massive datasets published or massive simulations?
+-->
+As default, `repro` uses Docker, which is the defacto standard for cloud computing, as container solution.
+However, Docker requires administrator rights for execution. Whereas this is no issue on most modern cloud computing infrastructures, it is often an issue in classic HPC environments.
+Singularity <!--AP add reference --> is an alternative solution. It was designed to run without administratot rights, while maintaining full compatibility with Docker images.
 If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
-Anyway, from within the container i.e. for your analysis it does not make any difference.
-Like with Docker, the default `Makefile` of `repro` has a mechanism to automatically send all commands to Singularity.
-It is triggered by `SINGULARITY=TRUE`:
+Importantly, this choice does not affect how code is executed inside the container.
+Like with Docker, the default `Makefile` generated by `repro` comes with a mechanism to automatically send all commands to Singularity.
+To reproduce an analysis with Singularity, simple add the parameter `SINGULARITY=TRUE` as follows:
 

### 2021-05-05 manuscript.Rmd
more on external dependencies

@@
 One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
+The dependency on a single R package may easily create a dependency tree of more than hundred direct or indirect dependencies [@epskamp2019].
 The critical problem is the fragility of a software setup.
-Many components have to fit perfectly together, and each change, e.g. an update, risks that balance.
-Updates are, of course, very much necessary.
+Many components have to fit perfectly together, and each change, e.g.
+an update, risks that balance .
+At the same time, software updates are very much necessary to remove errors or improve methodology [@epskamp2019].
 Another problem is that the software environment is also variable and what works on one machine might not work on any other.
-Docker circumvents such issues.
-Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, down to the operating system.
+Containers directly address such issues.
+Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself.
+The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
+<!--AP can we add an example here with the non-reproducibility of an eigen problem solver - I think we discussed at some point in time when WOlfgang Viechtbauer sent us a link to a paper -->.
 Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
-Such a snapshot of the software stack is called "image".
-Of course, packaging all needed software in such an image requires considerable amounts of storage.
-However, two strategies help to keep the storage requirements manageable.
+Such a snapshot of the software stack is called an "image".
+Packaging all needed software in such an image requires considerable amounts of storage space.
+There are two major strategies helping to keep the storage requirements reasonable.
 First, there is a community that maintains images for particular purposes.

### 2021-05-05 manuscript.Rmd
more on dependencies

@@
 First, there is a community that maintains images for particular purposes.
-For example, using R [@boettigerIntroductionRockerDocker2017].
-Such images only contain what is necessary for a specific purpose.
-Users can then install whatever they additionally need on top of these purpose-built images.
+For example, there are ready-made images for R based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
+Users can then install whatever they additionally need on top of these pre-made images.
 The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
-It does include Ubuntu (a Linux operating system) R, RStudio, LaTeX as well as many R packages like the tidyverse (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
-Second, instead of saving a binary image of the software, one can also save a textual description of the programmatic steps to install the software in a `Dockerfile`.
-Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
-For archiving, you therefore have to store the image, as explained in [Archiving Docker].
+It does include Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse <!-- AP add citations to all packages, please--> (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
+Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
+This recipe is a textual description of all commands that need to be executed to recreate the image.
+Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
+However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
+For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
 The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:

### 2021-05-05 manuscript.Rmd
Johnes2013 is the construction of SD3

--- 2021-05-05 manuscript.Rmd
@@
 
-The selected model parameters should express our expectations for the Machiavellism subscale of the short dark triad (SD3).
-<!-- AP: reference needed; explain the scale.
-What is the construct that we are measuring here?
--->
+The selected model parameters should express our expectations for the Machiavellism subscale of the short dark triad [SD3, @Jones2013].
+
 As a starting point, we can use already published information.
-@Jones2013 report a set of factor loadings (see Table \@ref(tab:loadings)) <!--AP what was the reference population?
-what was the sample size?
---> and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
-<!--AP why not on the latent level?
-They report factor loadings after all?
-Please explain-->
+@Jones2013 report a set of factor loadings from an Exploratory Structural Equation Model [@asparouhovExploratoryStructuralEquation2009].
+The estimates for the construction sample, which was recruited from Amazon’s Mechanical Turk (MTurk) and consists of 489 adults from Canada and the United States, can be seen in Table \@ref(tab:loadings))
+They report a gender difference (at the manifest level) depending on the study between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
 

### 2021-05-05 manuscript.Rmd
Johnes2013 is the construction of SD3

--- 2021-05-05 manuscript.Rmd
@@
 <!--Hands on: Introduce GitHub PR and review-->
-
-Pull requests allow anyone to contribute to a project, even though they do not have write access, while the authors still retain all control.
+There are various ways, in which other scholars may interact with your published repository.
+For example, in our exposition of the workflow for reproducibility [@Peikert2019], once we had published our preprint, scholars previously unknown to us started to send us spelling corrections.
+When multiple persons start modifying a project in parallel, keeping track of the edits and how they are supposed to be merged is crucial.
+Clearly, allowing everyone to edit the project is difficult <!--AP are there papers who analyzed Wikipedia and the way concurrent changes and discussions are resolved - is this relevant at all?-->
+
+_Pull requests_ allow anyone to contribute to a project, even though they were not granted write access, while the authors still retain full control over all changes.
 If someone wants to contribute, but they have no permission to write to a repository, they can create a copy of the project, called "fork", where they can push their changes.
@@
 
-### GitHub Project Management
-
-<!--Hands on: Introduce GitHub Projects-->
-<!--I love it, but maybe we should remove it-->
-While issues and pull requests already provide much structure to organize your work, GitHub projects allow you to categorize and prioritize issues and pull request, while informing you on the progress you made.
-
 # Analyzing Data
 
-While you already know how to write dynamic documents with [Markdown] and [R Markdown], and track changes with [Git] and [GitHub], two components to archive long-term reproducibility are still missing, namely internal and external dependencies.
+While you already know how to write dynamic documents with [Markdown] and [R Markdown], and track changes with [Git] and [GitHub], two components to achieve long-term reproducibility are still missing, namely handling internal and external dependencies.
 
@@
 
-Modern data analysis is only feasible because we rely on much external software.
+Modern data analysis is only feasible because researchers rely on much external software. <!-- AP: can we give examples of cool applications in psychology -->
 One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
@@
 For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
-The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:
+With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`
 
@@
 
-With the R version that you currently use, this specification is then translated into a `Dockerfile`:
-
+With the R version that you currently use, this specification is then translated into a `Dockerfile` that makes sure that the container image has all these packages installed:
+<!--AP: can you say something about ensuring a given version of the R package; can the header be automatically generated from a given R script?-->
 <!-- Hands on: Introduce Docker -->
@@
 
-To verify the data integrity you can include a line of code after reading the data that looks like this:
+To verify the data integrity, you can include a line of code after reading the data that looks like this:
+<!-- AP: shouldn't this be wrapped into a repro function and also stored in the YAML header??? -->
 

### 2021-05-05 manuscript.Rmd
some more on PR

@@
 
-The selected model parameters should express our expectations for the Machiavellism subscale of the short dark triad [SD3, @Jones2013].
-
+The selected model parameters should express our expectations for the Machiavellism subscale of the short dark triad (SD3).
+<!-- AP: reference needed; explain the scale.
+What is the construct that we are measuring here?
+-->
 As a starting point, we can use already published information.
-@Jones2013 report a set of factor loadings from an Exploratory Structural Equation Model [@asparouhovExploratoryStructuralEquation2009].
-The estimates for the construction sample, which was recruited from Amazon’s Mechanical Turk (MTurk) and consists of 489 adults from Canada and the United States, can be seen in Table \@ref(tab:loadings))
-They report a gender difference (at the manifest level) depending on the study between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
+@Jones2013 report a set of factor loadings (see Table \@ref(tab:loadings)) <!--AP what was the reference population?
+what was the sample size?
+--> and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
+<!--AP why not on the latent level?
+They report factor loadings after all?
+Please explain-->
 
@@
 This process is especially useful if you want the feedback of a more senior researcher, because they do not need to know how to use Git but have a convenient web interface to provide comments or changes.
+
+### GitHub Project Management
+
+<!--Hands on: Introduce GitHub Projects-->
+<!--I love it, but maybe we should remove it-->
+While issues and pull requests already provide much structure to organize your work, GitHub projects allow you to categorize and prioritize issues and pull request, while informing you on the progress you made.
 
@@
 
-To verify the data integrity, you can include a line of code after reading the data that looks like this:
-<!-- AP: shouldn't this be wrapped into a repro function and also stored in the YAML header??? -->
+To verify the data integrity you can include a line of code after reading the data that looks like this:
 

### 2021-05-05 manuscript.Rmd
removed PR management

@@
 This process is especially useful if you want the feedback of a more senior researcher, because they do not need to know how to use Git but have a convenient web interface to provide comments or changes.
-
-### GitHub Project Management
-
-<!--Hands on: Introduce GitHub Projects-->
-<!--I love it, but maybe we should remove it-->
-While issues and pull requests already provide much structure to organize your work, GitHub projects allow you to categorize and prioritize issues and pull request, while informing you on the progress you made.
 

### 2021-05-05 manuscript.Rmd
cite simsem and simstudy closes #37

@@
 Data simulation is particularly easy in lavaan syntax using the function `simulateData`.
-More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem <!--AP please add reference, and possibly other packages-->.
+More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem [@simsem], or simstudy [@goldfeldSimstudyIlluminatingResearch2020].
 The following code simulates data from the population model `model_truth` with 1.000 cases in each group.

### 2021-05-05 manuscript.Rmd
add broader references for simulation as technical and conceptual check

@@
 There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
-A step to prevent unpleasant surprises is to simulate data.<!-- AP: this paragraph deserves some references.
-I assume those can be found in the Bayesian literature.
-But are there classic SEM literature references talking of the value of simulation.
-Please make a lit research -->
+A step to prevent unpleasant surprises is to simulate data.
+The machine learning community has begun to use simulated data to test models and their implementation [@braiekTestingMachineLearning2020].
+Beyond checking the technical side of the model, simulation is also beneficial on a conceptual leven [@axelrodAdvancingArtSimulation1997].
 Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.

### 2021-05-05 manuscript.Rmd
explain advantages of DOI #38

@@
 ## Creating a DOI
-<!-- AP add references; why are DOIs helpful-->
-
+
+The approach we advocate for distributes the research project, e.g. across GitHub, preregistration platforms, the publishers site, and more.
+Digital Object Identifiers (DOI) tie all these locations together under one link [@derisiWhatWhysDOIs2003].
+Not only can a DOI resolve to several locations, several parts of a research project can get their own DOI besides the DOI of the publication.
+This is useful to link or cite a specific part of a research project, such as the preregistration, an illustration in the paper or the data that was gathered.
+The DOI can also contain information about the license of the object it refers to.
 <!-- public releases like preregistration, preprint -->
-zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides digital object identifiers (DOI) for these archives.
+zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides doi for these archives.
 While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.

### 2021-05-05 manuscript.Rmd
explain why paralell RNGs are hard

@@
 To consistently get the same results while preserving the pseudo-random property of the generated numbers, it is good practice to set a random seed that sets the PRNG into a restorable state, in R code: `set.seed(1235)`.
-It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster), but the `future` <!--AP add citation --> package handles this for us:
-<!--AB I don't understand how this solves the problem yet - please explain in more detail -->
+It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
+The standard random number generator (RNG) produces high quality random numbers, but has is not safe to use in parrallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
+The problem is that concurrent random number streams are correlated, because they repeat after some time.
+Therefore, if you set a seed for one worker and another seed for another worker, you could end up with the same random numbers but with a phase shift.
+@lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
+While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the [`future`-package](futureverse.org)[^futureverse] to handles this for us across all workers:
+
 ```{r}
@@
 ```
+
+[^futureverse]: https://www.futureverse.org
 

### 2021-05-06 manuscript.Rmd
add reference for the oldest available readme

@@
 Interested users find out about your project via the readme.
-Historically\footnote{Apparently, the use of readme files can be traced back to at least 1974 <!--AP reference needed, I guess-->}, a readme file is the default place to look for human readable information in a computer archive
+Historically[^readme-age], a readme file is the default place to look for human readable information in a computer archive 
 the first thing they see when they visit your project on [GitHub].
@@
 You could, for example, include the abstract of your (planned) paper here.<!-- AP: shouldn't this be a function in repro? to auto-insert the abstract?-->
+
+[^readme-age]: Apparently, the use of readme files can be traced back to at least 1974 as part of UCI LISP in this file: http://pdp-10.trailing-edge.com/decuslib10-04/01/43,50322/read.me.html
 

### 2021-05-06 manuscript.Rmd
cite another reference for copyright

@@
 Note that publishing data, code, or other materials is not sufficient to allow other users to use it, modify it, or redistribute it.
-For example, according to German copyright law, a repository without a license must be assumed to be protected by copyright <!-- AP citation needed -->.
+For example, according to German copyright law, a repository without a license must be assumed to be protected by copyright ([UrhG[^urhg]](https://www.gesetze-im-internet.de/englisch_urhg/englisch_urhg.html), Section 1, 7, 11).
 In many countries (i.e.
-the USA, European Union), a content creator has exclusive right to use their creation.
+the USA, European Union), a content creator has exclusive right to use their creation[@stoddenLegalFrameworkReproducible2009].
 A license is the legal authorization of other people to use what you have created.
@@
 In contrast, scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
-Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
+Copyright does often not apply to data, we therefore recommend to share data with a CC0 license to clarify that you do not hold copyright to it.
+Giving data another license than CC0 is often just confusing, because a license is invalid if there is no copyright.
+All licenses assure maximal freedom for future users while requiring the attribution, if possible, of the original authors in derivative work.
 These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].

### 2021-05-07 manuscript.Rmd
cite tidyverse

@@
 The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
-It does include Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse <!-- AP add citations to all packages, please--> (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
+It does include Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
 Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
@@
 ```{r, include=FALSE}
-knitr::write_bib(c(.packages(), "gert", "bookdown", "lavaan", "knitr", "targets", "renv"),
+knitr::write_bib(c(.packages(), "gert", "bookdown", "lavaan", "knitr", "targets", "renv", "tidyverse"),
                  "packages.bib")

### 2021-05-07 manuscript.Rmd
describe a possible solution for multiple licences

@@
 However, if their first contribution occurs after adding a license, they give their consent by contributing.
-To add a CC-BY license to your project, simply run (there are similar commands for other licenses):
+To add a CC-BY license to your project in general, simply run (there are similar commands for other licenses):
 
@@
 
-<!-- AP: Can we show how to set MIT license for code an CCBY license for the manuscript? Can we make a repro function that creates such a double-license? How to do it actually? Two license files? -->
+You can add other licenses for specific directories (e.g. copy `LICENSE.md` from `usethis::use_cc0_license()` into the `data` folder or the MIT License into the `R` folder).
 

### 2021-05-07 manuscript.Rmd
explain that the package versions are indeed fixed

--- 2021-05-07 manuscript.Rmd
@@
 title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
+author: Aaron Peikert and Andreas M. Brandmaier 
 output: bookdown::pdf_document2
@@
 
-This tutorial walks through the creation of a reproducible document that can be reproduced automatically.
-Though reproducibility is a small part of the research process, it can serve as the mortar of open science building blocks (e.g.
-preregistration, open data, postpublication review) --- but not that even non-open science will benefit from ensuring reproducibility because good scientific practice demands that every scientist is able to reproduce their own results later in time.
-Therefore, we show the whole lifecycle of an open research project, where some parts are essential to archive reproducibility.
-All code chunks are meant for the reader to try out for themselves.
-If not declared otherwise the code is meant to run in R (tested with version `r with(R.version, paste0(major, ".", minor))`).
-Still, the reader can safely skip sections marked as optional when unfit for their research question or methods.
+This tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
+We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.
+We continue with the creation of a concrete data analysis plan, which ideally is already implemented in code, and a dynamic report document that can be preregistered.
+Once the empirical data has been collected, the results can be calculated and appear directly in the dynamically generated document, where ideally only the discussion section needs to be added before the whole archive can be permanently stored as a fully reproducible digital artifact using a unique DOI in a persistent location.
+
+We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.
+We explicitly encourage readers to try out all R commands for themselves.
+Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).
+
+Here, we present a workflow that has proven useful for our way of working with a strong focus on Open Science.
+However, not all steps are necessary to achieve reproducibility.
+We have marked those steps that can be omitted as "optional" in the section headings.
 
@@
 
-If these commands detect that something is not installed or set up properly, they will alert you and give you detailed instructions (tailored to your operating system) to set up your computer.
-Please make sure that all these commands pass before you continue with the tutorial.
+If these commands detect that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
+Please make sure that all three commands pass before you continue with the tutorial.
 This process is only required once per computer and does not need to be repeated per project.
@@
 
-With the R version that you currently use, this specification is then translated into a `Dockerfile` that makes sure that the container image has all these packages installed:
-<!--AP: can you say something about ensuring a given version of the R package; can the header be automatically generated from a given R script?-->
+With the R version that you currently use, this specification is then translated into a `Dockerfile` that makes sure that the container image has all these packages installed.
+The `Dockerfile` also contains the current date and fixes packages versions to the version of the package that were available on that day on the [Comprehensive R Archive Network (CRAN)](https://cran.r-project.org)[^cran] by relying on the [CRAN timemachine of Microsoft R Application Network (MRAN)](https://mran.microsoft.com/timemachine)[^mran].
 <!-- Hands on: Introduce Docker -->
@@
 ```
+
+[^cran]: https://cran.r-project.org
+[^mran]: https://mran.microsoft.com/timemachine
 

### 2021-05-07 manuscript.Rmd
explain that the package versions are indeed fixed

@@
 title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
-author: Aaron Peikert and Andreas M. Brandmaier 
 output: bookdown::pdf_document2
@@
 
-If these commands detect that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
-Please make sure that all three commands pass before you continue with the tutorial.
+If these commands detect that something is not installed or set up properly, they will alert you and give you detailed instructions (tailored to your operating system) to set up your computer.
+Please make sure that all these commands pass before you continue with the tutorial.
 This process is only required once per computer and does not need to be repeated per project.

### 2021-05-07 manuscript.Rmd
some amendments to the intro

@@
 
-With the R version that you currently use, this specification is then translated into a `Dockerfile` that makes sure that the container image has all these packages installed.
-The `Dockerfile` also contains the current date and fixes packages versions to the version of the package that were available on that day on the [Comprehensive R Archive Network (CRAN)](https://cran.r-project.org)[^cran] by relying on the [CRAN timemachine of Microsoft R Application Network (MRAN)](https://mran.microsoft.com/timemachine)[^mran].
+With the R version that you currently use, this specification is then translated into a `Dockerfile` that makes sure that the container image has all these packages installed:
+<!--AP: can you say something about ensuring a given version of the R package; can the header be automatically generated from a given R script?-->
 <!-- Hands on: Introduce Docker -->
@@
 ```
-
-[^cran]: https://cran.r-project.org
-[^mran]: https://mran.microsoft.com/timemachine
 

### 2021-05-07 manuscript.Rmd
cite some big data sets

@@
 Thereby, the repro package makes massive parallel computing in heterogeneous computing environments less error-prone while ensuring the full documentation of all computational steps taken — even if some or all of them were run in parallel on distributed platforms.
-<!-- AP can we add examples, are there open repositores eg from neuroscience that have massive datasets published or massive simulations?
--->
-As default, `repro` uses Docker, which is the defacto standard for cloud computing, as container solution.
+Alone the sheer data size often warrants distributed computing, e.g.
+the European Bioinformatics Institute stores more than 120 Petabyte of publicly accessible data [@cookEuropeanBioinformaticsInstitute2018], single FMRI studies quickly amount to hundreds of Gigabyte[^openneuro] and even purely behavioral data like from the moral machine experiment [>10GB, @awadMoralMachineExperiment2018] can become to big to handle .
+
+[^openneuro]: https://openneuro.org/public/datasets
+
+As default, `repro` uses Docker, which is the defacto standard for cloud computing, as a container solution.
 However, Docker requires administrator rights for execution. Whereas this is no issue on most modern cloud computing infrastructures, it is often an issue in classic HPC environments.

### 2021-05-07 manuscript.Rmd
more verbose explanations of what is happening here....

--- 2021-05-07 manuscript.Rmd
@@
 
-We have eveloped a workflow that leverages established tools and practices from software engineering to achieve long-term computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:
+We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:
 
@@
 
-However, which tools you use to implement these concepts is a matter of taste and project requirements.
-This tutorial follows the recommendations of @Peikert2019 and use Git for version control, R Markdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
-Their interplay is shown in Figure \@ref(fig:schematic).
+This tutorial follows the recommendations of @Peikert2019 and uses _Git_ for version control, _R Markdown_ for dynamic document creation, _Make_ for dependency tracking, and _Docker_ for software management.
+The interplay of these four components is shown in Figure \@ref(fig:schematic). 
+However, which tools you use to implement these concepts is a matter of taste and project requirements, and other choices are possible depending on the context and the degree of reproducibility one aims for [e.g., @vanlissa2020worcs; @turingway2021turing].
 At several points, we suggest other viable alternatives.
@@
 
+<!-- AP is there a single command check() that does all of this?-->
 If these commands detect that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
@@
 
-# Planning
+# Organizing Files
 
@@
 Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you achieve reproducibility gradually.
-The first step is to bundle all required files in one folder.
-This way, you never lose anything, and you can easily share the files.
+The first step towards reproducibility is to bundle all required files in one folder.
+This way, you never lose anything, and you can easily share the files without missing anything.
 <!-- Hands-on: Introducing RProjects-->
-RStudio's RProjects facilitate this concept of a project folder.
+RStudio's _R Projects_ facilitate this concept of a project folder.
 To create one, click on:
@@
 
-## Markdown
+<!-- AP: clearly content is missing. What about the here package. What are the benefits of an R package? What if I have external data on a network drive ..... ->
+
+## Dynamic Documents
 
@@
 A great tool to capture ideas or meeting notes is Markdown.
+Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
+This way, Markup does not need any specialized software for editting, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, or a Powerpoint presentation <!--AP: What is Markdown?
+Why Markdown?
+Why not MS Word?
+Why should I care?
+What is so great about it?-->
 To create a Markdown document, click on:

### 2021-05-07 manuscript.Rmd
more verbose explanations of what is happening here....

@@
 Thereby, the repro package makes massive parallel computing in heterogeneous computing environments less error-prone while ensuring the full documentation of all computational steps taken — even if some or all of them were run in parallel on distributed platforms.
-Alone the sheer data size often warrants distributed computing, e.g.
-the European Bioinformatics Institute stores more than 120 Petabyte of publicly accessible data [@cookEuropeanBioinformaticsInstitute2018], single FMRI studies quickly amount to hundreds of Gigabyte[^openneuro] and even purely behavioral data like from the moral machine experiment [>10GB, @awadMoralMachineExperiment2018] can become to big to handle .
-
-[^openneuro]: https://openneuro.org/public/datasets
-
-As default, `repro` uses Docker, which is the defacto standard for cloud computing, as a container solution.
+<!-- AP can we add examples, are there open repositores eg from neuroscience that have massive datasets published or massive simulations?
+-->
+As default, `repro` uses Docker, which is the defacto standard for cloud computing, as container solution.
 However, Docker requires administrator rights for execution. Whereas this is no issue on most modern cloud computing infrastructures, it is often an issue in classic HPC environments.

### 2021-05-07 manuscript.Rmd
more on Markdown

--- 2021-05-07 manuscript.Rmd
@@
 <!--Keep notes.-->
-<!-- Hands on: Introduce Markdown -->
  A great tool to capture ideas or meeting notes is Markdown.
 Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
-This way, Markup does not need any specialized software for editting, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, or a Powerpoint presentation <!--AP: What is Markdown?
+This way, Markup does not need any specialized software for editting, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format.
+We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations -- and even your CV.
+<!--AP: What is Markdown?
 Why Markdown?
@@
 Why should I care?
-What is so great about it?-->
+What is so great about it?
+This was missing.
+Do you want to add anything?
+Please add citation to how you can create a CV with Markdown -->
 To create a Markdown document, click on:
@@
 
-Another advantage of Markdown is that you can easily version-control it with Git.
+Another advantage of Markdown is that because of its alikeness to plain text, you can easily version-control it with Git and leverage Gits features for tracking, displaying, and merging sequential and concurrent changes.
 You may be familiar with Microsoft Words' "Track Changes"-Feature.

### 2021-05-07 manuscript.Rmd
added parts from the PAC abstract for FGME

@@
 
-## Preregestration (optional)
+## Preregistration (optional)
 
@@
 
-Preregistration serves as a tool that may help to increase the credibility of empirical results. <!-- AP citation needed-->
+Preregistration is the specification of a research plan in advance of data analysis. The goal of preregistration is transparency and a clear-cut divide into exploration and confirmation.
+Thus, it serves as a tool that may help to increase the credibility of empirical results. <!-- AP citation needed-->
 Without getting into too much detail of the meta-scientific arguments for preregistration, we want to note two advantages. <!-- AP can we add a citation to someone who has the arguments?-->
@@
 Some [citation!] argue that preregistration can not fulfil its promise to reduce the false discovery rate of science if we allow deviating from them.
-While such an argument has some merits, we believe that transparency outweighs these concerns.
-If researchers deviate from their preregistration, they probably have good reason to do so, which means they have learned something from the process of conducting the study.
+While such an argument has some merits, we believe that often transparency outweighs these concerns.
+Deviation from a preregistration is still possible.
+If researchers deviate from their preregistration, they probably have good reason to do so, which means they have learned something from the process of conducting the study, and any deviation will simply be explicitly marked as such.
 A preregistration, together with version control, allows their readers to inform themselves about these lessons learned.
@@
 
-The likelihood that you have to diverge from your preregistration increases the more specific and concrete the preregistration is.
-To nudge the researcher to still be specific enough, there are many preregistrations templates.
-These templates formalize somewhat the research questions and the planned analysis.
-An alternative to such protocol is to submit preregistration as code.
-Therefore, instead of describing the analysis, you plan to turn into code, you skip a step and write the code beforehand.
-Such preregistration as code is challenging because it leaves no place to hide behind the additional layer of abstraction that verbal descriptions provide.
-It may well be that for any given project, this is impossible without deviating from the preregistration.
-However, preregistered and version-controlled code lets anyone see what was changed very concretely.
+Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration <!--AP citation needed-->.
+Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code <!--AP are the classic computer science literature who talks about this ambiguity...
+theoretical CS?
+NLP research or such?-->.
+Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity.
+PAC essentially reverses the process of empirical research.
+Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results --- initially based on simulated data.
+Then, the source code of this manuscript serves as an unambiguous description of the analysis plan and is published as a preregistration.
+
+PAC works well within this reproducible framework.
+First, the preregistration becomes a part of the reproducible repository.
+It is part of the set of version-controlled files and provides transparency about when the preregistration file, data files, and code files were changed over the course of a research project.
+If code was version-controlled and tagged (see Section XXX) before data collection, all deviations from the preregistration or any post-hoc additions that were performed after data collection are transparent and are visible and comparable in the version history.
 

### 2021-05-07 manuscript.Rmd
add summary

@@
 A data simulation allows you to verify that the code runs error-free and lets you form more realistic expectations about the estimates' correctness and precision.
+<!--AP shouldn't this be in the simulation part?-->
 We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
@@
 
+### Data Documentation
+
+<!--AP cite WORCS and some arguments for documenting your data. Also, cite the codebook package -->
+
 ### Code of Conduct
@@
 One (hypothetical) example is this:
+
+<!-- AP: this is a nice example but why is this an abstract example now? I thought that 
+we walk through one(!) example. I think this really should be the concrete Makefile from
+the analysis that we are talking about -->
 
@@
 
+Note that both this tutorial as well as our article on the reproducible workflow [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories <!--AP add links -->.
 <!-- maybe we add here the real data? #25 -->
@@
 
+<!--AP: make sure to cite WORCS and describe what WORCS is proposing here-->
 <!-- Hands-on: check changed hash -->
@@
 
-# Archiving and Dissemination
+# Archival and Dissemination
 
@@
 
+# Summary
+
+Here, we provided a step-by-step tutorial with what we think are best practices for creating a fully reproducible repository for data analysis projects that center on the statistical programming language R.
+The workflow avoids common threats to reproducibility, such as missing files, missing software dependencies, side-effects through changing dependent software packages, changes in the programming language, or even at the level of the operating system itself.
+<!-- AP: add more.
+Summarize.
+Talk about the benefits.
+Talk about the limitations.
+What are the threats to reproducibility.
+Make the case that adopting this workflow is easy and can even be done for existing repositories.
+Also, wasn't an important point of repro that it itself is not a dependency?
+I think this needs to be strengthened and probably already said in the intro-->
+
 # References

### 2021-05-07 manuscript.Rmd
skip sem to "Randome Number Generation"

@@
 
-<!-- AP is there a single command check() that does all of this?-->
 If these commands detect that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
@@
 
-<!-- AP: clearly content is missing. What about the here package. What are the benefits of an R package? What if I have external data on a network drive ..... ->
-
 ## Dynamic Documents
@@
 We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations -- and even your CV.
-<!--AP: What is Markdown?
-Why Markdown?
-Why not MS Word?
-Why should I care?
-What is so great about it?
-This was missing.
-Do you want to add anything?
-Please add citation to how you can create a CV with Markdown -->
 To create a Markdown document, click on:
@@
 
-If you are unfamiliar with the specifics of SEM, we you may skip the next sections and continue in section XXX <!--AP: fill in reference.-->.
+If you are unfamiliar with the specifics of SEM, we you may skip the next sections and continue in section ["Randome Number Generation"](#prng).
 To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.
@@
 ```
+
+### Randome Number Generation (optional) {#prng}
 

### 2021-05-07 manuscript.Rmd
more to discussion

@@
 Here, we provided a step-by-step tutorial with what we think are best practices for creating a fully reproducible repository for data analysis projects that center on the statistical programming language R.
-The workflow avoids common threats to reproducibility, such as missing files, missing software dependencies, side-effects through changing dependent software packages, changes in the programming language, or even at the level of the operating system itself.
+The workflow avoids common threats to reproducibility, such as copy-and-paste errors, missing files, missing software dependencies, or side-effects through changing dependent software packages, changes in the programming language, or even at the level of the operating system itself.
+However, the workflow as proposed by @Peikert2019 requires mastering up to four different software tools and their specification languages, which are typically not taught in the curriculum of social and behavioral sciences but rather belong to computer science and software development.
+Currently, this is a serious obstacle in adopting the workflow for researchers who strive for reproducibility but have little training in computer science and software development.  
+The repro package, which we introduced here, supports scholars in creating reproducible workflows.
+It abstracts tedious and error-prone tasks such as the creation of a Makefile or the specification of a container.
+By extending the YAML specification that is used for storing meta information in R Markdown files, which are becoming increasingly popular among R users, repro helps users to implement @Peikert2019's workflow with minimal effort.
+Most importantly, repro is implemented such that the package itself is not needed to reproduce a workflow that was generated according to the workflow.
+Importantly, this means that users that adopt the workflow do not have to rely on the future availability and maintenance of the repro package. 
 <!-- AP: add more.

### 2021-05-07 manuscript.Rmd
added ref

@@
 ```
-
+<!--AP emphasize more how cool it is that this stuff is automatically generated by repro!-->
 With the R version that you currently use, this specification is then translated into a `Dockerfile` that makes sure that the container image has all these packages installed.
@@
 This mechanism is faster and simpler to use than Docker, at the expense of only tracking R packages.
-`renv` therefore is already of great use for reproducibility.
-However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS system in use and so forth.
+`renv` therefore is already of great use for reproducibility [@vanlissa2020worcs].
+However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS system <!--AP reference, also explain what this is...
+most people will not know--> in use and so forth.
 Most importantly the R version itself has to match. 
+For example, @Peikert2019 illustrate how the generation of random numbers across different R versions is a threat to reproducibility.
 
@@
 <!-- AP: shouldn't this be wrapped into a repro function and also stored in the YAML header??? -->
-
+<!-- AP: can you demonstrate this with the actual openpsychometric data instead?-->
 ```{r}

### 2021-05-09 manuscript.Rmd
put each sentance on its own line #50

--- 2021-05-07 manuscript.Rmd
@@
  Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
-This way, Markup does not need any specialized software for editting, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format.
-We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations -- and even your CV.
+ This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+ R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format
+ We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations---and even your CV.
 To create a Markdown document, click on:
@@
 
-Preregistration is the specification of a research plan in advance of data analysis. The goal of preregistration is transparency and a clear-cut divide into exploration and confirmation.
-Thus, it serves as a tool that may help to increase the credibility of empirical results. <!-- AP citation needed-->
-Without getting into too much detail of the meta-scientific arguments for preregistration, we want to note two advantages. <!-- AP can we add a citation to someone who has the arguments?-->
+Preregistration is the specification of a research plan in advance of data analysis.
+The goal of preregistration is transparency and a clear-cut divide into exploration and confirmation.
+Thus, it serves as a tool that may help to increase the credibility of empirical results.
+Without getting into too much detail of the meta-scientific arguments for preregistration, we want to note two advantages.
 First, a preregistration is an opportunity to plan a study carefully.
@@
 
-Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration <!--AP citation needed-->.
-Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code <!--AP are the classic computer science literature who talks about this ambiguity...
-theoretical CS?
-NLP research or such?-->.
+Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration. 
+Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code.
 Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity.
@@
 A data simulation allows you to verify that the code runs error-free and lets you form more realistic expectations about the estimates' correctness and precision.
-<!--AP shouldn't this be in the simulation part?-->
 We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
@@
 It is supposed to explain to visitors what your project is about and how to reproduce it.
-You could, for example, include the abstract of your (planned) paper here.<!-- AP: shouldn't this be a function in repro? to auto-insert the abstract?-->
+You could, for example, include the abstract of your (planned) paper here.
 
@@
 For example, according to German copyright law, a repository without a license must be assumed to be protected by copyright ([UrhG[^urhg]](https://www.gesetze-im-internet.de/englisch_urhg/englisch_urhg.html), Section 1, 7, 11).
-In many countries (i.e.
-the USA, European Union), a content creator has exclusive right to use their creation[@stoddenLegalFrameworkReproducible2009].
+In many countries (i.e. the USA, European Union), a content creator has exclusive right to use their creation[@stoddenLegalFrameworkReproducible2009].
 A license is the legal authorization of other people to use what you have created.
@@
 
-<!-- AP: Can we show how to set MIT license for code an CCBY license for the manuscript? Can we make a repro function that creates such a double-license? How to do it actually? Two license files? -->
-
 ### Data Documentation
-
-<!--AP cite WORCS and some arguments for documenting your data. Also, cite the codebook package -->
 
@@
 When multiple persons start modifying a project in parallel, keeping track of the edits and how they are supposed to be merged is crucial.
-Clearly, allowing everyone to edit the project is difficult <!--AP are there papers who analyzed Wikipedia and the way concurrent changes and discussions are resolved - is this relevant at all?-->
+Clearly, allowing everyone to edit the project is difficult.
 
@@
 
-<!-- AP: this is a nice example but why is this an abstract example now? I thought that 
-we walk through one(!) example. I think this really should be the concrete Makefile from
-the analysis that we are talking about -->
-
 ```
@@
 
-Modern data analysis is only feasible because researchers rely on much external software. <!-- AP: can we give examples of cool applications in psychology -->
+Modern data analysis is only feasible because researchers rely on much external software.
 One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
@@
 The critical problem is the fragility of a software setup.
-Many components have to fit perfectly together, and each change, e.g.
-an update, risks that balance .
+Many components have to fit perfectly together, and each change, e.g. an update, risks that balance .
 At the same time, software updates are very much necessary to remove errors or improve methodology [@epskamp2019].
@@
 The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
-<!--AP can we add an example here with the non-reproducibility of an eigen problem solver - I think we discussed at some point in time when WOlfgang Viechtbauer sent us a link to a paper -->.
 Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
@@
 This mechanism is faster and simpler to use than Docker, at the expense of only tracking R packages.
-`renv` therefore is already of great use for reproducibility.
+`renv` therefore is already of great use for reproducibility [@vanlissa2020worcs].
 However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS system in use and so forth.
 Most importantly the R version itself has to match.
+For example, @Peikert2019 illustrate how the generation of random numbers across different R versions is a threat to reproducibility.
 
@@
 As default, `repro` uses Docker, which is the defacto standard for cloud computing, as a container solution.
-However, Docker requires administrator rights for execution. Whereas this is no issue on most modern cloud computing infrastructures, it is often an issue in classic HPC environments.
-Singularity <!--AP add reference --> is an alternative solution. It was designed to run without administratot rights, while maintaining full compatibility with Docker images.
+However, Docker requires administrator rights for execution.
+Whereas this is no issue on most modern cloud computing infrastructures, it is often an issue in classic HPC environments.
+Singularity is an alternative solution.
+It was designed to run without administratot rights, while maintaining full compatibility with Docker images.
 If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
@@
 A reproducible project should not only be reproducible in theory, but it should be simple to reproduce.
-Typically, it is argued that results should be reproducible by a Ph.D.
-student from the same discipline who had experience with the statistical software used [@obels2020].
+Typically, it is argued that results should be reproducible by a Ph.D. student from the same discipline who had experience with the statistical software used [@obels2020].
 Even if all code and data are available in a repository, if reproducibility takes too long because of a lack of structure and documentation, it must be regarded as non-reproducible [@obels2020].
@@
 
-Note that both this tutorial as well as our article on the reproducible workflow [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories <!--AP add links -->.
+Note that both this tutorial as well as our article on the reproducible workflow [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories.
 <!-- maybe we add here the real data? #25 -->
@@
 
-<!--AP: make sure to cite WORCS and describe what WORCS is proposing here-->
 <!-- Hands-on: check changed hash -->
-A prerequisite to reproducibility is the availability of code and data <!-- AP: can you add the reference to this tight reproducibility requirements that you told me about?-->.
+A prerequisite to reproducibility is the availability of code and data.
 We use containers to guarantee that all code components are available and work as expected.
@@
 In practice, however, there are various reasons, which may prohibit you from making your data available, for example legal or regulatory or privacy concerns.
-You may find yourself in a situation where you can make the data only available on request. <!-- AP please cite the SHARE data and explain as an example what one has to do to ontain these data -->.
+You may find yourself in a situation where you can make the data only available on request.
 A major threat to reproducibility is that the data has changed in the meantime or multiple versions are circulating and it is not exactly clear what version was used originally.
@@
 To verify the data integrity, you can include a line of code after reading the data that looks like this:
-<!-- AP: shouldn't this be wrapped into a repro function and also stored in the YAML header??? -->
 
@@
 Importantly, this means that users that adopt the workflow do not have to rely on the future availability and maintenance of the repro package.
-<!-- AP: add more.
-Summarize.
-Talk about the benefits.
-Talk about the limitations.
-What are the threats to reproducibility.
-Make the case that adopting this workflow is easy and can even be done for existing repositories.
-Also, wasn't an important point of repro that it itself is not a dependency?
-I think this needs to be strengthened and probably already said in the intro-->
 

### 2021-05-10 manuscript.Rmd
recomend to put each sentence on a newline closes #50
some spelling errors and clarifications

@@
 
-If these commands detect that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
+If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
 Please make sure that all three commands pass before you continue with the tutorial.
-This process is only required once per computer and does not need to be repeated per project.
+This process is only required once per computer and does not need to be repeated for each project.
 
@@
 RStudio's _R Projects_ facilitate this concept of a project folder.
+The `repro` package offers an R Studio template, which sets up a file structure that follows best practices and conventions.<!--AP this is from your thesis - can you please add citations as to which practices and conventions these are?-->
 To create one, click on:
@@
 And GitHub can show you what has changed, like in Figure \@ref(fig:idea-change).
+This detection of what has changed works on a line-by-line basis.
+We recommend therefore to keep each sentence in a Markdown on its own line to generate clean difference views.
 GitHub is, in its essence, a web interface to Git, and you may learn more about how to use it under the section on [GitHub].
@@
 Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manuel`.
-We recomend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
+We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
 

### 2021-05-10 preregistration.Rmd
simplify simulation

@@
 source(here::here("R", "simulation.R"))
+repro::automate_load_scripts()
 # if you have access to a hpc envir specify this in R/hpc.R
@@
 
-```{r models}
-true_diff <- 0.2
-model_truth <- combine(
-  measurement(
-    "MACH",
-    items("x", 9),
-    loadings = round(loadings_jones_paulhus * 0.7, 2)
-  ),
-  intercepts("MACH", list(0, true_diff)),
-  variances("MACH", list(1, 1))
-)
-
-model_same <- 
-  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0,0)*1
-  MACH ~~ c(1, NA)*MACH"
-
-model_differ <- 
-  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0, NA)*1 + c(zero, diff)*1
-  MACH ~~ c(1, NA)*MACH"
-```
-
 ```{r sim}
-extract_fns <- list(bic = get_bic, 
-                    aic = get_aic,
-                    lrt_p = get_lrt_p,
-                    delta_p = get_delta_p,
-                    estimate = get_estimate,
-                    std_delta_p = get_std_delta_p,
-                    std_estimate = get_std_estimate)
-
 n_obs <- seq(params$nmin, params$nmax, params$nstep)
 n_sim <- params$nsim
-
-setup <- expand_grid(n_obs = n_obs,
-                     truth = model_truth,
-                     same = model_same,
-                     differ = model_differ,
-                     mod_load = c(3),
-                     mod_int = c(3),
-                     extract_fns = list(extract_fns),
-                     group.equal = list(c("loadings", "intercepts")),
-                     auto.fix.first = FALSE)
+# n_sim * length(n_obs)
+# [1] 100000
 to_export <- ls_funs() %>% map(get)
-to_export <- c(list(setup = setup), to_export)
-RNGkind("L'Ecuyer-CMRG")
-set.seed(params$seed)
-seed <- params$seed
-res_raw %<-% generate_data_setup(n_sim,
-                setup,
+to_export <-
+  c(to_export, list(
+    n_obs = seq(params$nmin, params$nmax, params$nstep),
+    n_sim = params$nsim
+  ))
+res_raw %<-% simulation_study(n_obs, n_sim,
                 furrr_options(
@@
 ```{r}
-res <- res_raw  %>%
-  map(bind_rows) %>% 
-  bind_rows() %>% 
-  mutate(results = map(results, "result")) %>% 
-  unnest(results)
-
-res <- mutate(
-  res,
-  dec_bic = bic < 0,
-  dec_aic = aic < 0,
-  dec_aic2 = aic < -2,
-  dec_lrt_p = lrt_p < 0.05,
-  dec_delta_p = delta_p < 0.05)
-
+res <- res_raw %>% mutate(n_obs = as.numeric(n_obs))
 res_interval <- res %>%
-  select(n_obs, std_estimate, mod_load, mod_int) %>% 
+  select(n_obs, std_estimate) %>% 
   filter(!is.na(std_estimate)) %>% 
-  group_by(n_obs, mod_load, mod_int) %>% 
+  group_by(n_obs) %>% 
   summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
@@
 choosen_width <- .2
+true_diff <- .2
 minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
@@
 ```{r}
-plot_interval +
+simulation_plot <- plot_interval +
   theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + 
@@
   plot_annotation(tag_levels = 'A')
-ggsave("plot.svg", device = svglite::svglite, width = 7, height = 4)
+simulation_plot
+write_rds(simulation_plot, "simulation_plot.rds")
+ggsave("simulation.svg", device = svglite::svglite, width = 7, height = 4)
 ```

### 2021-05-10 R/simulation.R

@@ -1,11 +1,6 @@
-source(here::here("R", "lavaan_helper.R"))
+source(here::here("R", "standardizedSolution.R"))
 
 #----sim-functions----
-competing_models <- function(data, same, differ, ...){
-  list(same = cfa(same, data, ...),
-       differ = cfa(differ, data, ...))
-}
-
 filter_dots <- function(..., fun, what = names(formals(fun))){
   dots <- list(...)
   if(any(names(dots) == ""))stop("filter_dots does not allow unnamed args!")
@@ -22,16 +17,17 @@ safe_cfa <- function(...){
   do.call("cfa", dots, envir = getNamespace("lavaan"))
 }
 
-competing_models_mod <- function(data, same, differ, mod_load = 0L, mod_int = 0L, pars = "", ...){
+stitch <- function(pars){
+  with(pars, str_c(lhs, op, rhs))
+}
+
+partial_measurement_invariance <- function(data, same, differ, mod_load = 0L, mod_int = 0L, pars = "", ...){
   res_mod <- set_free_mod(differ, data, pars, op = "=~", depth = mod_load, ...)
   res_mod <- set_free_mod(differ, data, res_mod$pars, op = "~1", depth = mod_int, ...)
   fits <- list(same = safe_cfa(model = same, data = data, group.partial = res_mod$pars, ...),
                differ = res_mod$differ)
   
 }
-
-safe_competing_models <- safely(competing_models_mod,
-                                otherwise = list(same = NA, differ = NA))
 
 set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...){
   differ_fit <- safe_cfa(model = differ, data = data, group.partial = pars, ...)
@@ -56,9 +52,6 @@ set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...)
   }
 }
 
-safe_set_free_mod <- safely(set_free_mod,
-                            otherwise = list(same = NA, differ = NA))
-
 is_converged_ <- function(fit)lavaan::lavInspect(fit, "converged")
 is_converged <- function(fit){
   if(inherits(fit, "lavaan"))return(is_converged_(fit))
@@ -68,12 +61,13 @@ all_converged <- function(fits)all(purrr::map_lgl(fits, is_converged))
 
 get_fit <- function(fits, measure){
   stopifnot(length(fits) == 2L, length(measure) == 1L)
[…]

### 2021-05-10 manuscript.Rmd
revamp simulation still WIP

--- 2021-05-09 manuscript.Rmd
@@
   packages:
+    - tidyverse
     - usethis
@@
     - renv
+    - slider
+    - patchwork
   scripts:
-    - R/lavaan_helper.R
     - R/simulation.R
+    - R/simulation_funs.R
     - R/link.R
@@
 library(repro)
-library(tidyverse)
+automate_load_packages()
 automate_load_scripts()
@@
 
-Multigroup Confirmatory Factor Analysis
-First, establish strong measurement invariance using a CFI cutoff-criterion
-Second, test for differences in latent factor means of Machiavellianism using a $\chi^2$ test
+Multigroup Confirmatory Factor Analysis with equal loadings and intercepts
 
@@
 ```{r, include=FALSE}
-source(here::here("R", "simulation.R"))
+source(here::here("R", "simulation_funs.R"))
 ```
@@
 The machine learning community has begun to use simulated data to test models and their implementation [@braiekTestingMachineLearning2020].
-Beyond checking the technical side of the model, simulation is also beneficial on a conceptual leven [@axelrodAdvancingArtSimulation1997].
+Beyond checking the technical side of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
 Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
@@
 ```{r loadings, echo=FALSE}
-knitr::kable(data.frame(Item = 1:9,
-                        Loading = loadings_jones_paulhus),
+knitr::kable(t(data.frame(Item = as.character(1:9),
+                        Loading = loadings_jones_paulhus)),
              caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
@@
 
+```{r}
+model_truth <-
+"MACH =~ 0.27*x1 + 0.22*x2 + 0.28*x3 + 0.36*x4 + 0.41*x5 + 0.5*x6 + 0.43*x7 + 0.32*x8 + 0.36*x9
+MACH ~ c(0, 0.2)*1
+MACH ~~ c(1, 1)*MACH"
+```
+
 [^lavaan]: https://lavaan.ugent.be
-
-```{r, results='asis', echo=FALSE}
-model_truth <- combine(
-  measurement(
-    "MACH",
-    items("x", 9),
-    loadings = round(loadings_jones_paulhus * 0.7, 2)
-    ),
-  intercepts("MACH", list(0, 0.2))
-  )
-cat("```r",
-    model_truth,
-    "```",
-    sep = "\n")
-model <- combine(measurement(
-  "MACH",
-  items("x", 9)),
-  intercepts("MACH", list(0, "NA")))
-```
 
@@
 
-```{r, results='hide'}
+```{r}
 simulated <- lavaan::simulateData(
@@
 ```{r}
+model <- "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+MACH ~~ c(1, NA)*MACH
+MACH ~ c(0, NA)*1 + c(zero, diff)*1"
 fit <-
@@
 We, therefore, adapt our analysis to plan for this contingency.
-After fitting the model with strict measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
+After fitting the model with strong measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
 The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
@@
 
-If you are satisfied with your assumptions, it helps to write a function that expresses them, like this one:
+If you are satisfied with your assumptions for the data simulation, it helps to write a function that expresses them, like this one:
 <!--Hands-on: build simple functions tailored for your analysis-->
@@
 
+And to build another function that captures your planned analysis.
 Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R", "GitHub")`.
 
-With its help we can simulate 500 cases and fit our analysis.
-
-```{r}
-true_diff <- 0.2
-model_truth <- combine(
-  measurement(
-    "MACH",
-    items("x", 9),
-    loadings = round(loadings_jones_paulhus * 0.7, 2)
-  ),
-  intercepts("MACH", list(0, true_diff)),
-  variances("MACH", list(1, 1))
-)
-
-model_same <- 
-  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0,0)*1
-  MACH ~~ c(1, NA)*MACH"
-
-model_differ <- 
-  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0, NA)*1 + c(zero, diff)*1
-  MACH ~~ c(1, NA)*MACH"
-generate_data(500,
-              model_truth,
-              model_same,
-              model_differ,
-              list(std_estimate = get_std_estimate),
-              group.equal = c("loadings", "intercepts"),
-              mod_load = 3,
-              mod_int = 3,
-              std.lv = TRUE)
-```
-
-Note, that this certainly results in different numbers each time we run it.
-
-```{r}
-generate_data(500,
-              model_truth,
-              model_same,
-              model_differ,
-              list(std_estimate = get_std_estimate),
-              group.equal = c("loadings", "intercepts"),
-              mod_load = 3,
-              mod_int = 3,
-              std.lv = TRUE)
-```
+```{r}
+planed_analysis
+extract_results
+```
+
+With its help we can simulate 1000 cases and fit our analysis.
+
+```{r}
+generate_data(1000) %>% 
+  planed_analysis() %>% 
+  extract_results()
+```
+
+Note, that this certainly results in different numbers each time we run it (see ["Randome Number Generation"](#prng)).
+Of course it is already helpful to simulate a single dataset to test if the analysis works in principle, but repeating the analysis several times and vary parameters lets you verify more properties of your analysis.
+First, you can check if the planned analysis recovers the simulated values (`std_estimate` should be close to the simulated mean difference of 0.2).
+Second, it enables you to plan an appropriate sample size because you are able to quantify the sampling variability.
+
+To complete a simulation study we need a function that repeats the analysis for different conditions (here only for different n):
+
+```{r}
+simulation_study
+```
+
+```{r}
+res <- read_csv(here::here("results.csv"), col_types = "ddddd")
+```
+
+```{r}
+res_interval <- res %>%
+  select(n_obs, std_estimate) %>% 
+  filter(!is.na(std_estimate)) %>% 
+  group_by(n_obs) %>% 
+  summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
+            .groups = "drop", median = median(std_estimate)) %>% 
+  unnest(c(interval)) %>% 
+  mutate(width = upper - lower, perc = 1 - alpha) %>% 
+  group_by(alpha) %>% 
+  arrange(n_obs) %>% 
+  mutate(across(c(width, upper, lower),
+                ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE))) %>% 
+  ungroup() %>% 
+  arrange(perc)
+```
+
+```{r}
+choosen_perc <- .8
+choosen_width <- .2
+true_diff <- .2
+minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
+  filter(n_obs == min(n_obs))
+sim_n <- pull(res, n_obs)
+sim_nmax <- max(sim_n)
+sim_n_unique <- unique(sim_n)
+sim_nstep <- sim_n_unique[2] - sim_n_unique[1]
+```
+
+```{r}
+plot_interval <- res_interval %>%
+  ggplot(aes(
+    n_obs,
+    ymin = lower,
+    ymax = upper,
+    fill = perc,
+    group = alpha
+  )) +
+  geom_ribbon() +
+  geom_line(aes(y = median), color = "white") +
+  geom_hline(yintercept = true_diff, linetype = "dotted") +
+  scale_fill_viridis_c(
+    begin = 1,
+    end = 0,
+    breaks  = 1 - seq(0.05, .2, 0.05),
+    labels = . %>% scales::percent(accuracy = 1)
+  ) +
+  scale_y_continuous(n.breaks = 10) +
+  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
+  theme_minimal() +
+  labs(x = "N per Group", y = "Standardized Effect Size") +
+  theme(legend.title = element_blank()) +
+  NULL
+```
+
+```{r}
+plot_width <- res_interval %>%
+  ggplot(aes(n_obs, width, color = perc, group = perc)) +
+  geom_line() +
+  geom_point(
+    data = minn,
+    size = 3,
+    color = "black",
+    shape = 3
+  ) +
+  scale_color_viridis_c(
+    begin = 1,
+    end = 0,
+    breaks = 1 - seq(0.05, .2, 0.05),
+    labels = . %>% scales::percent(accuracy = 1)
+  ) +
+  theme_minimal() +
+  scale_y_log10(breaks = seq(0, 0.8, .1)) +
+  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
+  labs(x = "N per Group", y = "Width of quantile\n(Standardized Effect Size)") +
+  theme(legend.title = element_blank()) +
+  NULL
+```
+
+```{r}
+simulation_plot <- plot_interval +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + 
+  plot_width +
+  plot_layout(nrow = 2, guides = 'collect') +
+  plot_annotation(tag_levels = 'A')
+```
+
+```{r}
+simulation_plot
+```
+
+Plot A depicts the range of effect sizes we observed in the simulation, where the colour encodes the density.
+The light yellow interval shows where 80% of the estimates land.
+The dotted line represents the simulated true effect size; the white line shows the median observed effect sizes.
+
+Plot B represents the width of the interval shown in Plot A on a logarithmic scale.
+The cross shows the minimum required sample size so that 80% of the estimates land in a range of +/- .1.
 

### 2021-05-10 R/hash.R
revamp simulation still WIP

@@ -1,6 +1,6 @@
 #----hash----
 # compute hash using md5
 hash <- digest::digest(res_raw, "md5")
-if(hash != "4203854848ad0cc5d2f848c44b70815d"){
+if(hash != "cd14a91f680aef9355d8b7fe99e7f20f"){
   warning("Mismatch between original and current simulations!\nHash now is:\n    '", hash, "'")
 }

### 2021-05-10 R/simulation_funs.R

@@ -1,0 +1,178 @@
+source(here::here("R", "standardizedSolution.R"))
+
+#----sim-functions----
+filter_dots <- function(..., fun, what = names(formals(fun))){
+  dots <- list(...)
+  if(any(names(dots) == ""))stop("filter_dots does not allow unnamed args!")
+  exists <- setdiff(what, setdiff(what, names(dots)))
+  dots[exists]
+}
+
+safe_cfa <- function(...){
+  valid_args <-
+    c(names(formals(lavaan::lavaanify)),
+      names(formals(lavaan::lavaan)),
+      names(lavaan::lavOptions()))
+  dots <- filter_dots(..., fun = lavaan::cfa, what = valid_args)
+  do.call("cfa", dots, envir = getNamespace("lavaan"))
+}
+
+stitch <- function(pars){
+  with(pars, str_c(lhs, op, rhs))
+}
+
+partial_measurement_invariance <- function(data, same, differ, mod_load = 0L, mod_int = 0L, pars = "", ...){
+  res_mod <- set_free_mod(differ, data, pars, op = "=~", depth = mod_load, ...)
+  res_mod <- set_free_mod(differ, data, res_mod$pars, op = "~1", depth = mod_int, ...)
+  fits <- list(same = safe_cfa(model = same, data = data, group.partial = res_mod$pars, ...),
+               differ = res_mod$differ)
+  
+}
+
+set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...){
+  differ_fit <- safe_cfa(model = differ, data = data, group.partial = pars, ...)
+  depth <- depth - 1
+  if(depth < 0){return(list(pars = pars, differ = differ_fit))}
+  if(!is_converged(differ_fit)){return(list(pars = pars, differ = differ_fit))}
+  parest <- lavaan::parameterEstimates(differ_fit)
+  testscores <- lavaan::lavTestScore(differ_fit)
+  sorted_pars <- testscores$uni %>% 
+    rename(label = lhs) %>% 
+    select(-c(op, rhs, df)) %>% 
+    left_join(parest, by = c("label")) %>% 
+    filter(group == 1, .data$op %in% !!op) %>% 
+    arrange(desc(X2))
+  sorted_pars <- sorted_pars[1,]
+  
+  if(pull(sorted_pars, X2) > qchisq(0.95, 1)){
+    pars <- append(pars, stitch(sorted_pars))
[…]

### 2021-05-10 manuscript.Rmd
revamp simulation still WIP

@@
 Scientific claims should not be credible because of their originators' authority but by the transparency and replicability of their supporting evidence [@aac4716].
-Here, *replicability* refers to the degree to which original studies and their derivatives yield consistent results.In other words, a succesful replication study finds a consistent effect with an original study in a new sample of data.
+Here, *replicability* refers to the degree to which original studies and their derivatives yield consistent results.
+In other words, a successful replication study finds a consistent effect with an original study in a new sample of data.
 Enabling the assessment of scientific conclusions requires a systematic and rigorous approach to research.
-Researchers across various fields have identified a *crisis* of non-replicability of findings [e.g. @aac4716]- <!--citation needed, e.g. [@aac4716]-->
+Researchers across various fields have identified a *crisis* of non-replicability of findings [e.g. @aac4716]. <!--citation needed, e.g. [@aac4716]-->
 
@@
 With these advances came challenges that contribute to broader concerns of irreproducibility in the published literature, among them the lack of exact and unambiguous documentation of computational methods and each computational step taken.
-With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves.
+With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
 Only with access to both data and procedures can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.
-Only if both the data and code are shared, can we obtain a reproducible scientific product, facilitating assessment of its validity and later reuse.
+With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
 
@@
 2.	Copy-and-paste errors; for example, results are manually copied from a computing environment into a text processor, causing inconsistencies between the reported result and the reproduced result;
-.3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;
+3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;
 4.	Broken dependencies; for example, a given analysis depends on a software package that might not be available on a different computer or no longer exists.
@@
 This tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
+This paper was designed to help researchers to practically apply the TOP guidelines [except for number 8.
+which does not apply to single research projects; @nosekPromotingOpenResearch2015; @vanlissa2020worcs].
 We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.
@@
 
-# Organizing Files
+# Planning Phase
 
@@
 
-## Dynamic Documents
+## Markdown
 
@@
 `repro` reminds you of that and suggests which file you might want to add to the target `all`.
+You can generally change the `Makefile` anyway you want, e.g. to add targets which are not supported by `repro`.
 After you have done so, you can then run in the terminal:
@@
 ```
-
-You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manuel` to install software beyond R packages.
-Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manuel`.
-We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
+The function `automate()` combines `automate_docker()` & `automate_make()`.
+
+You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manual` to install software beyond R packages.
+Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manual`.
+We recomend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
 
@@
 
+Because you probably want to reproduce your own project each time something changes, you will often want to reproduce evertyhting.
+The `repro`-package contains a function designed for that purpose:
+
+```{r}
+repro::rerun()
+```
+
+This function  automatically figures out how to reproduce a project  and also recognizes other employed reproducibility tools (like `renv` and `singularity`), but can of course only guess the appropriate way to reproduce a project.
+
 Note that both this tutorial as well as our article on the reproducible workflow [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories.
@@
 We use containers to guarantee that all code components are available and work as expected.
-Ideally, all data is also stored in the repository.
-In practice, however, there are various reasons, which may prohibit you from making your data available, for example legal or regulatory or privacy concerns.
+Ideally, all data is also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs], but this is not always feasible.
+Sometimes it is just the required storage capacity that exceeds what Git can handle (Git gets slow around more than 100mb).
+In such case we strongly recommend using [DataLad [@DataladDatalad2021]](http://handbook.datalad.org/en/latest/)[^datalad] or the below procedure to guarantee that the data has not changed.
+In practice, however, there are various reasons, which may prohibit you from making your data publicly available, for example legal or regulatory or privacy concerns.
 You may find yourself in a situation where you can make the data only available on request.
@@
 To verify the data integrity, you can include a line of code after reading the data that looks like this:
+
+[^datalad]: http://handbook.datalad.org/en/latest/
 
@@
 By extending the YAML specification that is used for storing meta information in R Markdown files, which are becoming increasingly popular among R users, repro helps users to implement @Peikert2019's workflow with minimal effort.
-Most importantly, repro is implemented such that the package itself is not needed to reproduce a workflow that was generated according to the workflow.
-Importantly, this means that users that adopt the workflow do not have to rely on the future availability and maintenance of the repro package.
+Most importantly, `repro`-is implemented such that the package itself is not needed to reproduce a workflow that was generated according to the workflow.
+Importantly, this means that users that adopt the workflow do not have to rely on the future availability and maintenance of the `repro`-package.
 

### 2021-05-10 R/simulation.R

@@ -1,178 +1,49 @@
-source(here::here("R", "standardizedSolution.R"))
+library(tidyverse)
+library(here)
+library(lavaan)
+library(furrr)
+source(here("R", "simulation_funs.R"))
+# if you have access to a hpc envir specify this in R/hpc.R
+# if not we use local multicore
+# to speed up consider reduce nsim; increase nstep ↓↓↓
 
-#----sim-functions----
-filter_dots <- function(..., fun, what = names(formals(fun))){
-  dots <- list(...)
-  if(any(names(dots) == ""))stop("filter_dots does not allow unnamed args!")
-  exists <- setdiff(what, setdiff(what, names(dots)))
-  dots[exists]
+hpc_config <- here::here("R", "hpc.R")
+if(fs::file_exists(hpc_config)){
+  source(hpc_config)
+} else {
+  plan(list(transparent,
+            tweak(multisession, workers = 4L)))
 }
+# debug:
+#   plan(transparent)
 
-safe_cfa <- function(...){
-  valid_args <-
-    c(names(formals(lavaan::lavaanify)),
-      names(formals(lavaan::lavaan)),
-      names(lavaan::lavOptions()))
-  dots <- filter_dots(..., fun = lavaan::cfa, what = valid_args)
-  do.call("cfa", dots, envir = getNamespace("lavaan"))
-}
+nmin <- 100
+nmax <- 300
+nstep <- 100
+nsim <- 2
+seed <- 1234
+n_obs <- seq(nmin, nmax, nstep)
 
-stitch <- function(pars){
-  with(pars, str_c(lhs, op, rhs))
-}
+to_export <- ls_funs() %>% map(get)
+to_export <-
+  c(to_export, list(
+    n_obs = seq(nmin, nmax, nstep),
+    n_sim = nsim
+  ))
+res_raw %<-% simulation_study(n_obs, n_sim,
+                              furrr_options(
+                                globals = to_export,
+                                seed = seed,
+                                packages = c("furrr", "lavaan", "tidyverse")
+                              ))
+invisible(res_raw)
+readr::write_csv(res_raw, "results.csv")
+readr::write_rds(res_raw, "results.csv")
 
-partial_measurement_invariance <- function(data, same, differ, mod_load = 0L, mod_int = 0L, pars = "", ...){
[…]

### 2021-05-11 R/simulation.R
update simulation, add as make dependency

@@ -1,10 +1,13 @@
-library(tidyverse)
-library(here)
-library(lavaan)
-library(furrr)
+suppressPackageStartupMessages(library(tidyverse))
+suppressPackageStartupMessages(library(here))
+suppressPackageStartupMessages(library(lavaan))
+suppressPackageStartupMessages(library(furrr))
+
 source(here("R", "simulation_funs.R"))
 # if you have access to a hpc envir specify this in R/hpc.R
-# if not we use local multicore
+# see https://github.com/aaronpeikert/repro-tutorial/blob/hpc/R/hpc.R
+# or `git checkout hpc R/hpc.R`
+# if no hpc is availible we use local multicore with all available cores
 # to speed up consider reduce nsim; increase nstep ↓↓↓
 
 hpc_config <- here::here("R", "hpc.R")
@@ -12,33 +15,38 @@ if(fs::file_exists(hpc_config)){
   source(hpc_config)
 } else {
   plan(list(transparent,
-            tweak(multisession, workers = 4L)))
+            tweak(multisession)))
 }
 # debug:
 #   plan(transparent)
 
 nmin <- 100
-nmax <- 300
+nmax <- 10000
 nstep <- 100
-nsim <- 2
+n_sim <- 1000
 seed <- 1234
 n_obs <- seq(nmin, nmax, nstep)
+
+options(scipen = 999)
+message("This may take a while! You fit approximately ", n_sim*length(n_obs)*5*2, " SEMs.")
 
 to_export <- ls_funs() %>% map(get)
 to_export <-
   c(to_export, list(
     n_obs = seq(nmin, nmax, nstep),
-    n_sim = nsim
+    n_sim = n_sim
   ))
 res_raw %<-% simulation_study(n_obs, n_sim,
                               furrr_options(
                                 globals = to_export,
                                 seed = seed,
-                                packages = c("furrr", "lavaan", "tidyverse")
+                                packages = c("furrr", "lavaan", "tidyverse"),
+                                scheduling = 10
                               ))
 invisible(res_raw)
-readr::write_csv(res_raw, "results.csv")
-readr::write_rds(res_raw, "results.csv")
+fs::dir_create("data")
+readr::write_csv(res_raw, here::here("data", "simulation_results.csv"))
+readr::write_rds(res_raw, here::here("data", "simulation_results.rds"))
 
 if(fs::file_exists(hpc_config)){
   list(local = sessioninfo::session_info(),

### 2021-05-11 manuscript.Rmd
update simulation, add as make dependency
cite worcs for data #54

--- 2021-05-10 manuscript.Rmd
@@
 
-If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
+If these commands detect that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
 Please make sure that all three commands pass before you continue with the tutorial.
-This process is only required once per computer and does not need to be repeated for each project.
+This process is only required once per computer and does not need to be repeated per project.
 
@@
 RStudio's _R Projects_ facilitate this concept of a project folder.
-The `repro` package offers an R Studio template, which sets up a file structure that follows best practices and conventions.<!--AP this is from your thesis - can you please add citations as to which practices and conventions these are?-->
 To create one, click on:
@@
 Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manuel`.
-We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
+We recomend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
 
@@
 We use containers to guarantee that all code components are available and work as expected.
-Ideally, all data is also stored in the repository.
+Ideally, all data is also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs].
 In practice, however, there are various reasons, which may prohibit you from making your data available, for example legal or regulatory or privacy concerns.

### 2021-05-11 manuscript.Rmd
cite datalad

@@
 We use containers to guarantee that all code components are available and work as expected.
-Ideally, all data is also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs].
-In practice, however, there are various reasons, which may prohibit you from making your data available, for example legal or regulatory or privacy concerns.
+Ideally, all data is also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs], but this is not always feasible.
+Sometimes it is just the required storage capacity that exceeds what Git can handle (Git gets slow around more than 100mb).
+In such case we strongly recommend using [DataLad [@DataladDatalad2021]](http://handbook.datalad.org/en/latest/)[^datalad] or the below procedure to garunteee that the data has not changed.
+In practice, however, there are various reasons, which may prohibit you from making your data publically available, for example legal or regulatory or privacy concerns.
 You may find yourself in a situation where you can make the data only available on request.

### 2021-05-11 manuscript.Rmd
note top guidelines #54

@@
 Sometimes it is just the required storage capacity that exceeds what Git can handle (Git gets slow around more than 100mb).
-In such case we strongly recommend using [DataLad [@DataladDatalad2021]](http://handbook.datalad.org/en/latest/)[^datalad] or the below procedure to garunteee that the data has not changed.
-In practice, however, there are various reasons, which may prohibit you from making your data publically available, for example legal or regulatory or privacy concerns.
+In such case we strongly recommend using [DataLad [@DataladDatalad2021]](http://handbook.datalad.org/en/latest/)[^datalad] or the below procedure to guarantee that the data has not changed.
+In practice, however, there are various reasons, which may prohibit you from making your data publicly available, for example legal or regulatory or privacy concerns.
 You may find yourself in a situation where you can make the data only available on request.

### 2021-05-11 manuscript.Rmd
fix typo

@@
 
-You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manuel` to install software beyond R packages.
-Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manuel`.
+You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manual` to install software beyond R packages.
+Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manual`.
 We recomend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.

### 2021-05-11 manuscript.Rmd
mention rerun closes #67

@@
 
+The `repro` also contains a function that automatically figures out how to reproduce a project with:
+
+```{r}
+repro::rerun()
+```
+
+This function also recognizes other employed reproducibility tools (like `renv` and `singularity`), but can of course only guess the appropriate way to reproduce a project.
+
 Note that both this tutorial as well as our article on the reproducible workflow [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories.

### 2021-05-11 manuscript.Rmd
talk about automate
explain why you need `rerun()`

@@
 
+The function `automate()` combines `automate_docker()` & `automate_make()`.
+
 <!-- after this everything is optional? -->
@@
 
-The `repro` also contains a function that automatically figures out how to reproduce a project with:
+Because you probably want to reproduce your own project each time something changes, you will often want to reproduce evertyhting.
+The `repro`-package contains a function designed for that purpose:
 
@@
 
-This function also recognizes other employed reproducibility tools (like `renv` and `singularity`), but can of course only guess the appropriate way to reproduce a project.
+This function  automatically figures out how to reproduce a project  and also recognizes other employed reproducibility tools (like `renv` and `singularity`), but can of course only guess the appropriate way to reproduce a project.
 

### 2021-05-11 manuscript.Rmd
actually include fake results

@@
 
+And to build another function that captures your planned analysis.
 Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R", "GitHub")`.
 
-With its help we can simulate 500 cases and fit our analysis.
-
-```{r}
-true_diff <- 0.2
-model_truth <- combine(
-  measurement(
-    "MACH",
-    items("x", 9),
-    loadings = round(loadings_jones_paulhus * 0.7, 2)
-  ),
-  intercepts("MACH", list(0, true_diff)),
-  variances("MACH", list(1, 1))
-)
-
-model_same <- 
-  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0,0)*1
-  MACH ~~ c(1, NA)*MACH"
-
-model_differ <- 
-  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-  MACH ~ c(0, NA)*1 + c(zero, diff)*1
-  MACH ~~ c(1, NA)*MACH"
-generate_data(500,
-              model_truth,
-              model_same,
-              model_differ,
-              list(std_estimate = get_std_estimate),
-              group.equal = c("loadings", "intercepts"),
-              mod_load = 3,
-              mod_int = 3,
-              std.lv = TRUE)
-```
-
-Note, that this certainly results in different numbers each time we run it.
-
-```{r}
-generate_data(500,
-              model_truth,
-              model_same,
-              model_differ,
-              list(std_estimate = get_std_estimate),
-              group.equal = c("loadings", "intercepts"),
-              mod_load = 3,
-              mod_int = 3,
-              std.lv = TRUE)
-```
+```{r}
+planed_analysis
+extract_results
+```
+
+With its help we can simulate 1000 cases and fit our analysis.
+
+```{r}
+generate_data(1000) %>% 
+  planed_analysis() %>% 
+  extract_results()
+```
+
+Note, that this certainly results in different numbers each time we run it (see ["Randome Number Generation"](#prng)).
+Of course it is already helpful to simulate a single dataset to test if the analysis works in principle, but repeating the analysis several times and vary parameters lets you verify more properties of your analysis.
+First, you can check if the planned analysis recovers the simulated values (`std_estimate` should be close to the simulated mean difference of 0.2).
+Second, it enables you to plan an appropriate sample size because you are able to quantify the sampling variability.
+
+To complete a simulation study we need a function that repeats the analysis for different conditions (here only for different n):
+
+```{r}
+simulation_study
+```
+
+```{r}
+res <- read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddd")
+```
+
+```{r}
+res_interval <- res %>%
+  select(n_obs, std_estimate) %>% 
+  filter(!is.na(std_estimate)) %>% 
+  group_by(n_obs) %>% 
+  summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
+            .groups = "drop", median = median(std_estimate)) %>% 
+  unnest(c(interval)) %>% 
+  mutate(width = upper - lower, perc = 1 - alpha) %>% 
+  group_by(alpha) %>% 
+  arrange(n_obs) %>% 
+  mutate(across(c(width, upper, lower),
+                ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE))) %>% 
+  ungroup() %>% 
+  arrange(perc)
+```
+
+```{r}
+choosen_perc <- .8
+choosen_width <- .2
+true_diff <- .2
+minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
+  filter(n_obs == min(n_obs))
+sim_n <- pull(res, n_obs)
+sim_nmax <- max(sim_n)
+sim_n_unique <- unique(sim_n)
+sim_nstep <- sim_n_unique[2] - sim_n_unique[1]
+```
+
+```{r}
+plot_interval <- res_interval %>%
+  ggplot(aes(
+    n_obs,
+    ymin = lower,
+    ymax = upper,
+    fill = perc,
+    group = alpha
+  )) +
+  geom_ribbon() +
+  geom_line(aes(y = median), color = "white") +
+  geom_hline(yintercept = true_diff, linetype = "dotted") +
+  scale_fill_viridis_c(
+    begin = 1,
+    end = 0,
+    breaks  = 1 - seq(0.05, .2, 0.05),
+    labels = . %>% scales::percent(accuracy = 1)
+  ) +
+  scale_y_continuous(n.breaks = 10) +
+  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
+  theme_minimal() +
+  labs(x = "N per Group", y = "Standardized Effect Size") +
+  theme(legend.title = element_blank()) +
+  NULL
+```
+
+```{r}
+plot_width <- res_interval %>%
+  ggplot(aes(n_obs, width, color = perc, group = perc)) +
+  geom_line() +
+  geom_point(
+    data = minn,
+    size = 3,
+    color = "black",
+    shape = 3
+  ) +
+  scale_color_viridis_c(
+    begin = 1,
+    end = 0,
+    breaks = 1 - seq(0.05, .2, 0.05),
+    labels = . %>% scales::percent(accuracy = 1)
+  ) +
+  theme_minimal() +
+  scale_y_log10(breaks = seq(0, 0.8, .1)) +
+  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
+  labs(x = "N per Group", y = "Width of quantile\n(Standardized Effect Size)") +
+  theme(legend.title = element_blank()) +
+  NULL
+```
+
+```{r}
+simulation_plot <- plot_interval +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + 
+  plot_width +
+  plot_layout(nrow = 2, guides = 'collect') +
+  plot_annotation(tag_levels = 'A')
+```
+
+```{r}
+simulation_plot
+```
+
+Plot A depicts the range of effect sizes we observed in the simulation, where the colour encodes the density.
+The light yellow interval shows where 80% of the estimates land.
+The dotted line represents the simulated true effect size; the white line shows the median observed effect sizes.
+
+Plot B represents the width of the interval shown in Plot A on a logarithmic scale.
+The cross shows the minimum required sample size so that 80% of the estimates land in a range of +/- .1.
 
@@
 ```
+The function `automate()` combines `automate_docker()` & `automate_make()`.
 
@@
 We recomend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
-
-The function `automate()` combines `automate_docker()` & `automate_make()`.
 
@@
 This function automatically figures out how to reproduce a project and also recognizes other employed reproducibility tools (like `renv` and `singularity`), but can of course only guess the appropriate way to reproduce a project.
+
+If you have written a [Preregistration as Code] you can swap the data file and reproduce the project.
+In fact, before the`r link("https://github.com/aaronpeikert/repro-tutorial/releases/tag/ v0.0.1-preregistration", "preregistration")` were the results below on simulated data, only after we preregistered the whole manuscript, they were based on actual data from https://openpsychometrics.org/_rawdata/.
+
+```{r}
+read_csv(here("data", "sd3.csv"), col_types = "dddddddddd") %>% 
+  planed_analysis() %>% 
+  extract_results()
+```
 

### 2021-05-11 manuscript.Rmd
explain when exactly we plan to use real data

@@
 If you have written a [Preregistration as Code] you can swap the data file and reproduce the project.
-In fact, before the`r link("https://github.com/aaronpeikert/repro-tutorial/releases/tag/ v0.0.1-preregistration", "preregistration")` were the results below on simulated data, only after we preregistered the whole manuscript, they were based on actual data from https://openpsychometrics.org/_rawdata/.
+In fact, till after we preregistered the study `r link("https://github.com/aaronpeikert/repro-tutorial/releases/tag/ v0.0.1-preregistration", "preregistration")` were the results below based on simulated data.
+Only after we preregistered the whole manuscript and upon submission, they were based on actual data from https://openpsychometrics.org/_rawdata/.
 
@@
 Note that both this tutorial as well as our article on the reproducible workflow [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories.
-<!-- maybe we add here the real data? #25 -->
 

### 2021-05-12 manuscript.Rmd
hide code for plot

@@
 
-```{r}
+```{r, include=FALSE}
 res <- read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddd")
@@
 
-```{r}
+```{r, include=FALSE}
 res_interval <- res %>%
@@
 
-```{r}
+```{r, include=FALSE}
 choosen_perc <- .8
@@
 
-```{r}
+```{r, include=FALSE}
 plot_interval <- res_interval %>%
@@
 
-```{r}
+```{r, include=FALSE}
 plot_width <- res_interval %>%
@@
 
-```{r}
+```{r, include=FALSE}
 simulation_plot <- plot_interval +
@@
 
-```{r}
-simulation_plot
-```
-
-Plot A depicts the range of effect sizes we observed in the simulation, where the colour encodes the density.
+```{r, simulation-plot, echo=FALSE, fig.cap="*Plot A* depicts the range of effect sizes we observed in the simulation, where the colour encodes the density.
 The light yellow interval shows where 80% of the estimates land.
 The dotted line represents the simulated true effect size; the white line shows the median observed effect sizes.
-
-Plot B represents the width of the interval shown in Plot A on a logarithmic scale.
+*Plot B* represents the width of the interval shown in Plot A on a logarithmic scale.
 The cross shows the minimum required sample size so that 80% of the estimates land in a range of +/- .1.
+"}
+simulation_plot
+```
 

### 2021-05-12 manuscript.Rmd
remove duplicated sentence
show required sample size and bias of sim in PAC

@@
 Only with access to both data and procedures can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.
-With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
 
@@
 It is part of the set of version-controlled files and provides transparency about when the preregistration file, data files, and code files were changed over the course of a research project.
-If code was version-controlled and tagged (see Section XXX) before data collection, all deviations from the preregistration or any post-hoc additions that were performed after data collection are transparent and are visible and comparable in the version history.
+If code was version-controlled and tagged (see Section [Git]) before data collection, all deviations from the preregistration or any post-hoc additions that were performed after data collection are transparent and are visible and comparable in the version history.
 
@@
 We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
-The simulation showed that we have to gather at least `missing_rcode_simulation` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `missing_rcode_simulation`%.
+The simulation showed that we have to gather at least `r minn$n_obs` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `r round((est_diff - true_diff)*100/true_diff, 2)`%.
+These results are also visualized in Figure \@ref(fig:simulation-plot).
 If you went to the trouble to simulate data, we argue that you get a preregistration almost for free.

### 2021-05-12 manuscript.Rmd
small changes here and there

@@
 This paper was designed to help researchers to practically apply the TOP guidelines [except for number 8.
-which does not apply to single research projects; @nosekPromotingOpenResearch2015; @vanlissa2020worcs].
+which does not apply to single research projects, @nosekPromotingOpenResearch2015; @vanlissa2020worcs].
 We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.
@@
  This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
- R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format
 We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations---and even your CV.
@@
 For convenience, we use the [`gert`-package](https://github.com/r-lib/gert) [@R-gert], which wraps most git commands into R functions.
-Alternatively, we suggest to either work directly in the terminal (for example, RStudio offers a command shell with various options accessible in the Menu Tools -> Terminal), or work using the Git pane in RStudio that offers buttons for basic operations.
+Alternatively, we suggest to either work directly in the terminal (for example, RStudio offers a command shell with various options accessible in the Menu Tools → Terminal), or work using the Git pane in RStudio that offers buttons for basic operations.
 <!--I am leaning towards doing everything in R, not sure though-->
@@
 More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem [@simsem], or simstudy [@goldfeldSimstudyIlluminatingResearch2020].
-The following code simulates data from the population model `model_truth` with 1.000 cases in each group.
+The following code simulates data from the population model `model_truth` with 1000 cases in each group.
 
@@
 true_diff <- .2
+est_diff <- res %>% filter(n_obs == max(n_obs)) %>% pull(std_estimate) %>% mean()
 minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
@@
 It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
-The standard random number generator (RNG) produces high quality random numbers, but has is not safe to use in parrallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
+The standard random number generator (RNG) produces high quality random numbers, but is not safe to use in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
 The problem is that concurrent random number streams are correlated, because they repeat after some time.
@@
 @lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
-While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the [`future`-package](futureverse.org)[^futureverse] to handles this for us across all workers:
+While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers:
 
@@
 PAC essentially reverses the process of empirical research.
-Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results --- initially based on simulated data.
+Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results---initially based on simulated data.
 Then, the source code of this manuscript serves as an unambiguous description of the analysis plan and is published as a preregistration.
@@
 R Markdown allows you to include R code (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown.
+R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format
 To create an R Markdown document, click on:

### 2021-05-12 manuscript.Rmd
closes #52

@@
 Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
+Beyond ensuring that the code runs without technical error and it lets you form more realistic expectations about the estimates' correctness and precision.
 No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about them.
@@
 PAC works well within this reproducible framework.
+
 First, the preregistration becomes a part of the reproducible repository.
@@
 One way to ensure that you can expect your analysis to work with real data is to simulate data.
-A data simulation allows you to verify that the code runs error-free and lets you form more realistic expectations about the estimates' correctness and precision.
 We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
@@
 See the section on [R Markdown] for an example of a dynamic document generation paradigm.
+In fact, this manuscript and therefore the here presented Machiavialismus example were preregistered as a full manuscript, and you can see the results on real data in section [Reproduce a Project].
 

### 2021-05-12 R/simulation.R
run simulation only when no simulation data is there

@@ -36,17 +36,20 @@ to_export <-
     n_obs = seq(nmin, nmax, nstep),
     n_sim = n_sim
   ))
-res_raw %<-% simulation_study(n_obs, n_sim,
-                              furrr_options(
-                                globals = to_export,
-                                seed = seed,
-                                packages = c("furrr", "lavaan", "tidyverse"),
-                                scheduling = 10
-                              ))
-invisible(res_raw)
-fs::dir_create("data")
-readr::write_csv(res_raw, here::here("data", "simulation_results.csv"))
-readr::write_rds(res_raw, here::here("data", "simulation_results.rds"))
+out_file <- here::here("data", "simulation_results.csv")
+if(!fs::file_exists(out_file)){
+  res_raw %<-% simulation_study(n_obs, n_sim,
+                                furrr_options(
+                                  globals = to_export,
+                                  seed = seed,
+                                  packages = c("furrr", "lavaan", "tidyverse"),
+                                  scheduling = 10
+                                ))
+  invisible(res_raw)
+  fs::dir_create("data")
+  readr::write_csv(res_raw, here::here("data", "simulation_results.csv"))
+  readr::write_rds(res_raw, here::here("data", "simulation_results.rds"))
+}
 
 if(fs::file_exists(hpc_config)){
   list(local = sessioninfo::session_info(),

### 2021-05-12 R/simulation.R
set seed for small examples in manuscript

@@ -28,7 +28,6 @@ seed <- 1234
 n_obs <- seq(nmin, nmax, nstep)
 
 options(scipen = 999)
-message("This may take a while! You fit approximately ", n_sim*length(n_obs)*5*2, " SEMs.")
 
 to_export <- ls_funs() %>% map(get)
 to_export <-
@@ -38,6 +37,7 @@ to_export <-
   ))
 out_file <- here::here("data", "simulation_results.csv")
 if(!fs::file_exists(out_file)){
+  message("This may take a while! You fit approximately ", n_sim*length(n_obs)*5*2, " SEMs.")
   res_raw %<-% simulation_study(n_obs, n_sim,
                                 furrr_options(
                                   globals = to_export,

### 2021-05-12 manuscript.Rmd
set seed for small examples in manuscript
fix typo
correct spelling of Machiavellism

@@
 
-Machiavellianism is higher in male persons.
+Machiavellism is higher in male persons.
 
@@
 
-If you are unfamiliar with the specifics of SEM, we you may skip the next sections and continue in section ["Randome Number Generation"](#prng).
+If you are unfamiliar with the specifics of SEM, you may briefly skim the next sections and continue in section ["Randome Number Generation"](#prng).
 To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.
@@
 The following code simulates data from the population model `model_truth` with 1000 cases in each group.
+
+```{r, include=FALSE}
+set.seed(1234)
+```
 
@@
 See the section on [R Markdown] for an example of a dynamic document generation paradigm.
-In fact, this manuscript and therefore the here presented Machiavialismus example were preregistered as a full manuscript, and you can see the results on real data in section [Reproduce a Project].
+In fact, this manuscript and therefore the here presented Machiavellism example were preregistered as a full manuscript, and you can see the results on real data in section [Reproduce a Project].
 

### 2021-05-13 manuscript.Rmd
simulation lets you plan your sample size

@@
 
-To complete a simulation study we need a function that repeats the analysis for different conditions (here only for different n):
+To complete a such a  Monte-Carlo simulation study we need a function that repeats the analysis for different conditions (here only for different n):
 
@@
 ```
+
+Such Monte-Carlo simulation can help you decide between several tradeoffs.
+One historically important tradeoff is between sample size and power (the probability to detect an effect of given size).
+More recently there has been growing interest in other resource relevant parameters as well [!citation], such as number of measurement occasions, reliability of measures etc.
+To show that there can also be alternative approaches for sample size planning beyond null hypothesis testing, we choose to visualize the density distribution of the parameter of interest (latent mean difference by gender of Machiavellism) instead of power.
+Such process still lets you formulate expectations for the precision of the analysis, e.g.
+with a sample size of 1000 per group 80% of the simulated standardized effect size fall into a range of +/- .1, but without necessarily rely on null hypothesis testing.
 

### 2021-05-13 manuscript.Rmd
add links to github repo closes #61

@@
 And to build another function that captures your planned analysis.
-Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R", "GitHub")`.
+Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R", "GitHub")`.
 
@@
 with a sample size of 1000 per group 80% of the simulated standardized effect size fall into a range of +/- .1, but without necessarily rely on null hypothesis testing.
+For simplicity we do not show the code for the simulation/graphic, but it is available in the accompanying [GitHub repository](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo]:.
+
+[^gh-repo]: https://github.com/aaronpeikert/repro-tutorial
 
@@
 
-Note that both this tutorial as well as our article on the reproducible workflow [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories.
-
+Note that both [this tutorial](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo2] as well as our article on the [reproducible workflow](https://github.com/aaronpeikert/reproducible-research)[^gh-repo3] [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories.
+
+[^gh-repo2]: https://github.com/aaronpeikert/repro-tutorial
+[^gh-repo3]: https://github.com/aaronpeikert/reproducible-research
 ## Closed Data (optional)
@@
 
-> Log into zenodo.org → Account → [GitHub[^zenodo-github]](https://zenodo.org/account/settings/github/)
+> Log into zenodo.org → Account → [GitHub](https://zenodo.org/account/settings/github/)[^zenodo-github]
 

### 2021-05-13 R/simulation_funs.R
explizitly set all loadings free closes #70

@@ -96,7 +96,8 @@ generate_data <- function(n_obs){
 }
 
 planed_analysis <- function(data){
-  model <- "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+  model <- "MACH =~  c(NA, NA)*x1 + c(NA, NA)*x2 + c(NA, NA)*x3 + c(NA, NA)*x4 +
+  c(NA, NA)*x5 + c(NA, NA)*x6 + c(NA, NA)*x7 + c(NA, NA)*x8 + c(NA, NA)*x9
    MACH ~~ c(1, NA)*MACH"
   # intercept is set to zero
   same <- paste0(model, "
@@ -112,8 +113,7 @@ planed_analysis <- function(data){
     mod_load = 3L,
     mod_int = 3L,
     group = "group",
-    group.equal = c("loadings", "intercepts"),
-    auto.fix.first = FALSE
+    group.equal = c("loadings", "intercepts")
   )
 }
 

### 2021-05-13 manuscript.Rmd
beauty corrections for the code

@@
     - patchwork
+    - knitr
   scripts:
@@
 source(here::here("R", "link.R"))
+```
+
+```{r}
+# define a ürint method for objects of the class data.frame
+knit_print.data.frame = function(x, ...) {
+  res = paste(c('', '', kable(x)), collapse = '\n')
+  asis_output(res)
+}
+# register the method
+registerS3method("knit_print", "data.frame", knit_print.data.frame)
 ```
@@
 model_truth <-
-"MACH =~ 0.27*x1 + 0.22*x2 + 0.28*x3 + 0.36*x4 + 0.41*x5 + 0.5*x6 + 0.43*x7 + 0.32*x8 + 0.36*x9
+"MACH =~ 0.27*x1 + 0.22*x2 + 0.28*x3 + 0.36*x4 + 0.41*x5 + 0.5*x6 + 0.43*x7 +
+0.32*x8 + 0.36*x9
 MACH ~ c(0, 0.2)*1
@@
 
-```{r}
-generate_data
+```{r generate_data}
 ```
@@
 
-```{r}
-planed_analysis
-extract_results
+```{r planed_analysis}
+```
+
+```{r extract_results}
 ```
@@
 
-```{r}
-simulation_study
+```{r simulation_study}
 ```

### 2021-05-13 R/simulation_funs.R

@@ -85,8 +85,10 @@ get_std_estimate <- function(fits, label = "diff"){
   get_parameter(fits, "est.std", label, how = standardizedSolution)
 }
 
+#----generate_data----
 generate_data <- function(n_obs){
-  truth <- "MACH =~ 0.27*x1 + 0.22*x2 + 0.28*x3 + 0.36*x4 + 0.41*x5 + 0.5*x6 + 0.43*x7 + 0.32*x8 + 0.36*x9
+  truth <- "MACH =~ 0.27*x1 + 0.22*x2 + 0.28*x3 + 0.36*x4 + 0.41*x5 + 0.5*x6 +
+  0.43*x7 + 0.32*x8 + 0.36*x9
   MACH ~ c(0, 0.2)*1
   MACH ~~ c(1, 1)*MACH"
   lavaan::simulateData(
@@ -95,10 +97,13 @@ generate_data <- function(n_obs){
     standardized = TRUE)
 }
 
+#----planed_analysis----
 planed_analysis <- function(data){
+  # force loadings to be freely estimated
   model <- "MACH =~  c(NA, NA)*x1 + c(NA, NA)*x2 + c(NA, NA)*x3 + c(NA, NA)*x4 +
   c(NA, NA)*x5 + c(NA, NA)*x6 + c(NA, NA)*x7 + c(NA, NA)*x8 + c(NA, NA)*x9
-   MACH ~~ c(1, NA)*MACH"
+  # model is identified via latent variance set to 1
+  MACH ~~ c(1, NA)*MACH"
   # intercept is set to zero
   same <- paste0(model, "
   MACH ~ c(0,0)*1")
@@ -117,6 +122,7 @@ planed_analysis <- function(data){
   )
 }
 
+#----extract_results----
 extract_results <- function(fits) {
   # extract_fns are helper that extract variables of interest, like p value etc.
   extract_fns <-
@@ -133,6 +139,7 @@ extract_results <- function(fits) {
   as_tibble(map(results, "result"))
 }
 
+#----simulation_study_funs----
 simulation <- function(n_obs){
   fits <- purrr::safely(planed_analysis)(generate_data(n_obs))$result
   extract_results(fits)
@@ -149,6 +156,7 @@ simulation_study_ <-
                    .options = .options)
   }
 
+#----simulation_study----
 simulation_study <- function(n_obs, n_sim, .options = furrr::furrr_options()){
   # `map` is equivalent to lapply or for-loop
[…]

### 2021-05-13 manuscript.Rmd
small changes after reading

@@
 A step to prevent unpleasant surprises is to simulate data.
-The machine learning community has begun to use simulated data to test models and their implementation [@braiekTestingMachineLearning2020].
+The machine learning community has begun to routinely simulate data to test models and their implementation [@braiekTestingMachineLearning2020].
 Beyond checking the technical side of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
@@
 [^gh-repo3]: https://github.com/aaronpeikert/reproducible-research
+
 ## Closed Data (optional)

### 2021-05-13 manuscript.Rmd
add use_github()
fix list

@@
 
+```{r, eval=FALSE}
+check_github()
+```
+
 If these commands detect that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
@@
 We found it useful to label the following stages of development as release:
-- Preregistration (repository contains the final data analysis plan; also see the Section on PAC)
-- Completion of data collection (repository contains all study data)
-- Preprint (repository exactly reproduces a preprint that is hosted externally)
-- Submission or Resubmission (repository exactly reproduces a given version (re)submitted to a journal)
-- In Press (repository exactly reproduces the final manuscript as accepted by a scientific journal)
+
+* Preregistration (repository contains the final data analysis plan; also see the Section on PAC)
+* Completion of data collection (repository contains all study data)
+* Preprint (repository exactly reproduces a preprint that is hosted externally)
+* Submission or Resubmission (repository exactly reproduces a given version (re)submitted to a journal)
+* In Press (repository exactly reproduces the final manuscript as accepted by a scientific journal)
 

### 2021-05-13 R/hash.R
use hash hash is incorrect, because simulation is running

@@ -1,6 +1,6 @@
 #----hash----
 # compute hash using md5
-hash <- digest::digest(res_raw, "md5")
-if(hash != "cd14a91f680aef9355d8b7fe99e7f20f"){
+hash <- digest::digest(res, "md5")
+if(hash != "e9e0dfe4b4753c1e00bb02e4205d8772"){
   warning("Mismatch between original and current simulations!\nHash now is:\n    '", hash, "'")
 }

### 2021-05-13 manuscript.Rmd
use hash hash is incorrect, because simulation is running
cite worcs for synthethic data
forgot to clean up after removing preregistration.Rmd

@@
     - knitr
+    - pander
+    - lavaan
+    - furrr
+    - future.batchtools
   scripts:
@@
 res <- read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddd")
+```
+
+```{r hash}
 ```
@@
 
-An compromise between making the data publicly available and keeping it closed is to publish synthetic data.
+An compromise between making the data publicly available and keeping it closed is to publish synthetic data [@vanlissa2020worcs].
 Such approach tries to recreate the marginal distribution and bivariate relationships in simulated data, e.g. with the R package `synthpop`[@nowokSynthpopBespokeCreation2016].

### 2021-05-13 manuscript.Rmd
add overview table

@@
 
+We have marked those steps that can be omitted as "optional" in the section headings.
+Table \@ref(tab:overview) shows all steps required to to create and open and reproducible research project as they are presented in the remainder of the manuscript.
+
+```{r overview, echo=FALSE}
+overview <-
+  read_csv(
+    here::here("data", "overview.csv"),
+    col_types = "cc",
+    col_names = c("Phase", "Steps"),
+    trim_ws = FALSE
+  )
+pander::pander(
+  overview,
+  split.tables = Inf,
+  split.cells = Inf,
+  missing = "",
+  keep.line.breaks = TRUE,
+  style = "grid",
+  justify = 'left',
+  caption = "Overview over all steps to create and open and reproducible research project."
+)
+```
+
 # Setup

### 2021-05-13 manuscript.Rmd
don't delete .tex so make is happy

--- 2021-05-13 manuscript.Rmd
@@
 author: Aaron Peikert and Andreas M. Brandmaier 
-output: bookdown::pdf_document2
+output:
+  bookdown::pdf_document2:
+    keep_tex: TRUE
 repro:
@@
 
+We have marked those steps that can be omitted as "optional" in the section headings.
+Table \@ref(tab:overview) shows all steps required to to create and open and reproducible research project as they are presented in the remainder of the manuscript.
+
+```{r overview, echo=FALSE}
+overview <-
+  read_csv(
+    here::here("data", "overview.csv"),
+    col_types = "cc",
+    col_names = c("Phase", "Steps"),
+    trim_ws = FALSE
+  )
+pander::pander(
+  overview,
+  split.tables = Inf,
+  split.cells = Inf,
+  missing = "",
+  keep.line.breaks = TRUE,
+  style = "grid",
+  justify = 'left',
+  caption = "\\label{tab:overview} Overview over all steps to create and open and reproducible research project."
+)
+```
+
 # Setup

### 2021-05-13 manuscript.Rmd
invite casper to collaborate
update idea-change closes #66

@@
 if(!file.exists(idea_change)){
-  webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/5b3f4641ff542158184f85d458880c35e8f09f3c?diff=split",
+  webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/76f5d4e78b3d3bf22e84e9137756bc17767dff3b?diff=split",
                   file = idea_change,

### 2021-05-13 manuscript.Rmd
fix gert command
cite some reporting packages closes #74

@@
 # -- type this on the R console --
-gert::commit_all("switch to CFA to exclude bias in measurement as possible confounder")
+gert::git_commit_all("switch to CFA to exclude bias in measurement as possible confounder")
 ```
@@
 If you have simulated data, you could even include fake results.
+Packages like [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf)[^stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html)[^apaTables] [citation] and [`papaja`](http://frederikaust.com/papaja_man/)[^papaja] help you to create dynamically generated professional looking results.
+The package [`report`](https://github.com/easystats/report)[^report] [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
 Such a dynamic document is arguably the most concrete preregistration possible.
@@
 
+[^report]: https://github.com/easystats/report
+[^stargazer]: https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf
+[^apaTables]: https://dstanley4.github.io/apaTables/articles/apaTables.html
+[^papaja]: http://frederikaust.com/papaja_man/
 <!--Hands on: gitignore resulting .html/pdf-->

### 2021-05-13 manuscript.Rmd
actually cite packages closes #75

@@
 If you have simulated data, you could even include fake results.
-Packages like [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf)[^stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html)[^apaTables] [citation] and [`papaja`](http://frederikaust.com/papaja_man/)[^papaja] help you to create dynamically generated professional looking results.
+Packages like [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf)[^stargazer] [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html)[^apaTables] [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/)[^papaja] [@papaja] help you to create dynamically generated professional looking results.
 The package [`report`](https://github.com/easystats/report)[^report] [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).

### 2021-05-15 manuscript.Rmd
Add citations.

@@
  This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations---and even your CV.
+We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations---and even your CV [@vitae].
 To create a Markdown document, click on:
@@
 
-Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration. 
+Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration [@osfstandard; @socpsytemp; @howtoreplicate]. 
 Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code.
@@
 ### Data Documentation
+
+[@codebook]
 
@@
 Whereas this is no issue on most modern cloud computing infrastructures, it is often an issue in classic HPC environments.
-Singularity is an alternative solution.
+Singularity is an alternative solution [@singularity].
 It was designed to run without administratot rights, while maintaining full compatibility with Docker images.
@@
 <!-- Hands-on: check changed hash -->
-A prerequisite to reproducibility is the availability of code and data.
+A prerequisite to reproducibility is the availability of code and data [@akhlaghi2020towards].
 We use containers to guarantee that all code components are available and work as expected.

### 2021-05-15 manuscript.Rmd
fix #53

@@
 Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration [@osfstandard; @socpsytemp; @howtoreplicate]. 
-Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code.
+Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code [@sep-ambiguity].
 Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity.

### 2021-05-17 manuscript.Rmd
changes

--- 2021-05-13 manuscript.Rmd
@@
 author: Aaron Peikert and Andreas M. Brandmaier 
-output:
-  bookdown::pdf_document2:
-    keep_tex: TRUE
+output: bookdown::pdf_document2
 repro:
   packages:
-    - tidyverse
     - usethis
@@
     - renv
-    - slider
-    - patchwork
-    - knitr
-    - pander
-    - lavaan
-    - furrr
-    - future.batchtools
   scripts:
+    - R/lavaan_helper.R
     - R/simulation.R
-    - R/simulation_funs.R
     - R/link.R
@@
 library(repro)
-automate_load_packages()
+library(tidyverse)
 automate_load_scripts()
@@
 
-```{r}
-# define a ürint method for objects of the class data.frame
-knit_print.data.frame = function(x, ...) {
-  res = paste(c('', '', kable(x)), collapse = '\n')
-  asis_output(res)
-}
-# register the method
-registerS3method("knit_print", "data.frame", knit_print.data.frame)
-```
-
 # Why you should care about reproducibility
@@
 Scientific claims should not be credible because of their originators' authority but by the transparency and replicability of their supporting evidence [@aac4716].
-Here, *replicability* refers to the degree to which original studies and their derivatives yield consistent results.
-In other words, a successful replication study finds a consistent effect with an original study in a new sample of data.
+Here, *replicability* refers to the degree to which original studies and their derivatives yield consistent results.In other words, a succesful replication study finds a consistent effect with an original study in a new sample of data.
 Enabling the assessment of scientific conclusions requires a systematic and rigorous approach to research.
-Researchers across various fields have identified a *crisis* of non-replicability of findings [e.g. @aac4716]. <!--citation needed, e.g. [@aac4716]-->
+Researchers across various fields have identified a *crisis* of non-replicability of findings [e.g. @aac4716]- <!--citation needed, e.g. [@aac4716]-->
 
@@
 With these advances came challenges that contribute to broader concerns of irreproducibility in the published literature, among them the lack of exact and unambiguous documentation of computational methods and each computational step taken.
-With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
+With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves.
 Only with access to both data and procedures can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.
+Only if both the data and code are shared, can we obtain a reproducible scientific product, facilitating assessment of its validity and later reuse.
 
@@
 2.	Copy-and-paste errors; for example, results are manually copied from a computing environment into a text processor, causing inconsistencies between the reported result and the reproduced result;
-3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;
+.3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;
 4.	Broken dependencies; for example, a given analysis depends on a software package that might not be available on a different computer or no longer exists.
@@
 This tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
-This paper was designed to help researchers to practically apply the TOP guidelines [except for number 8.
-which does not apply to single research projects, @nosekPromotingOpenResearch2015; @vanlissa2020worcs].
 We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.
@@
 We have marked those steps that can be omitted as "optional" in the section headings.
-
-We have marked those steps that can be omitted as "optional" in the section headings.
-Table \@ref(tab:overview) shows all steps required to to create and open and reproducible research project as they are presented in the remainder of the manuscript.
-
-```{r overview, echo=FALSE}
-overview <-
-  read_csv(
-    here::here("data", "overview.csv"),
-    col_types = "cc",
-    col_names = c("Phase", "Steps"),
-    trim_ws = FALSE
-  )
-pander::pander(
-  overview,
-  split.tables = Inf,
-  split.cells = Inf,
-  missing = "",
-  keep.line.breaks = TRUE,
-  style = "grid",
-  justify = 'left',
-  caption = "\\label{tab:overview} Overview over all steps to create and open and reproducible research project."
-)
-```
 
@@
 
-```{r, eval=FALSE}
-check_github()
-```
-
-If these commands detect that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
+If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
 Please make sure that all three commands pass before you continue with the tutorial.
-This process is only required once per computer and does not need to be repeated per project.
-
-# Planning Phase
+This process is only required once per computer and does not need to be repeated for each project.
+
+# Organizing Files
 
@@
 
-## Markdown
+## Dynamic Documents
 
@@
  This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+ R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format
  We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations---and even your CV.
@@
 
-Machiavellism is higher in male persons.
+Machiavellianism is higher in male persons.
 
@@
 For convenience, we use the [`gert`-package](https://github.com/r-lib/gert) [@R-gert], which wraps most git commands into R functions.
-Alternatively, we suggest to either work directly in the terminal (for example, RStudio offers a command shell with various options accessible in the Menu Tools → Terminal), or work using the Git pane in RStudio that offers buttons for basic operations.
+Alternatively, we suggest to either work directly in the terminal (for example, RStudio offers a command shell with various options accessible in the Menu Tools -> Terminal), or work using the Git pane in RStudio that offers buttons for basic operations.
 <!--I am leaning towards doing everything in R, not sure though-->
@@
 
-Multigroup Confirmatory Factor Analysis with equal loadings and intercepts
+Multigroup Confirmatory Factor Analysis
+First, establish strong measurement invariance using a CFI cutoff-criterion
+Second, test for differences in latent factor means of Machiavellianism using a $\chi^2$ test
 
@@
 # -- type this on the R console --
-gert::git_commit_all("switch to CFA to exclude bias in measurement as possible confounder")
+gert::commit_all("switch to CFA to exclude bias in measurement as possible confounder")
 ```
@@
 if(!file.exists(idea_change)){
-  webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/76f5d4e78b3d3bf22e84e9137756bc17767dff3b?diff=split",
+  webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/5b3f4641ff542158184f85d458880c35e8f09f3c?diff=split",
                   file = idea_change,
@@
 <!--I am troubled by how we should recommend to learn Git-->
-Git can be challenging to learn, because it requires a rather complex mental model.
-However, there is a subsets of commands which we present here and in the section on [GitHub] (commit, push, pull) which enable you to utilize Git for simple version control.
+Git can be challenging to learn, because it requires a rather complex mental model of maintaining potentially differing versions between different repositories (typically one central repository and one local repository per each author while each repository may have different branches).
+However, there is a subsets of commands which we present here and in the section on [GitHub] (commit, push, pull) which enable you to use Git for simple version control.
 There are many tutorials on learning Git and @vuorreCuratingResearchAssets2018 provide a good starting point.
@@
 ```{r, include=FALSE}
-source(here::here("R", "simulation_funs.R"))
+source(here::here("R", "simulation.R"))
 ```
@@
 A step to prevent unpleasant surprises is to simulate data.
-The machine learning community has begun to routinely simulate data to test models and their implementation [@braiekTestingMachineLearning2020].
-Beyond checking the technical side of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
+The machine learning community has begun to use simulated data to test models and their implementation [@braiekTestingMachineLearning2020].
+Beyond checking the technical side of the model, simulation is also beneficial on a conceptual leven [@axelrodAdvancingArtSimulation1997].
 Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
-Beyond ensuring that the code runs without technical error and it lets you form more realistic expectations about the estimates' correctness and precision.
 No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about them.
@@
 
-If you are unfamiliar with the specifics of SEM, you may briefly skim the next sections and continue in section ["Randome Number Generation"](#prng).
+If you are unfamiliar with the specifics of SEM, we you may skip the next sections and continue in section ["Randome Number Generation"](#prng).
 To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.
@@
 ```{r loadings, echo=FALSE}
-knitr::kable(t(data.frame(Item = as.character(1:9),
-                        Loading = loadings_jones_paulhus)),
+knitr::kable(data.frame(Item = 1:9,
+                        Loading = loadings_jones_paulhus),
              caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
@@
 
-```{r}
-model_truth <-
-"MACH =~ 0.27*x1 + 0.22*x2 + 0.28*x3 + 0.36*x4 + 0.41*x5 + 0.5*x6 + 0.43*x7 +
-0.32*x8 + 0.36*x9
-MACH ~ c(0, 0.2)*1
-MACH ~~ c(1, 1)*MACH"
-```
-
 [^lavaan]: https://lavaan.ugent.be
+
+```{r, results='asis', echo=FALSE}
+model_truth <- combine(
+  measurement(
+    "MACH",
+    items("x", 9),
+    loadings = round(loadings_jones_paulhus * 0.7, 2)
+    ),
+  intercepts("MACH", list(0, 0.2))
+  )
+cat("```r",
+    model_truth,
+    "```",
+    sep = "\n")
+model <- combine(measurement(
+  "MACH",
+  items("x", 9)),
+  intercepts("MACH", list(0, "NA")))
+```
 
@@
 More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem [@simsem], or simstudy [@goldfeldSimstudyIlluminatingResearch2020].
-The following code simulates data from the population model `model_truth` with 1000 cases in each group.
-
-```{r, include=FALSE}
-set.seed(1234)
-```
-
-```{r}
+The following code simulates data from the population model `model_truth` with 1.000 cases in each group.
+
+```{r, results='hide'}
 simulated <- lavaan::simulateData(
@@
 ```{r}
-model <- "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-MACH ~~ c(1, NA)*MACH
-MACH ~ c(0, NA)*1 + c(zero, diff)*1"
 fit <-
@@
 We, therefore, adapt our analysis to plan for this contingency.
-After fitting the model with strong measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
+After fitting the model with strict measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
 The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
@@
 
-If you are satisfied with your assumptions for the data simulation, it helps to write a function that expresses them, like this one:
+If you are satisfied with your assumptions, it helps to write a function that expresses them, like this one:
 <!--Hands-on: build simple functions tailored for your analysis-->
 
-```{r generate_data}
-```
-
-And to build another function that captures your planned analysis.
-Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R", "GitHub")`.
-
-```{r planed_analysis}
-```
-
-```{r extract_results}
-```
-
-With its help we can simulate 1000 cases and fit our analysis.
-
 ```{r}
-generate_data(1000) %>% 
-  planed_analysis() %>% 
-  extract_results()
-```
-
-Note, that this certainly results in different numbers each time we run it (see ["Randome Number Generation"](#prng)).
-Of course it is already helpful to simulate a single dataset to test if the analysis works in principle, but repeating the analysis several times and vary parameters lets you verify more properties of your analysis.
-First, you can check if the planned analysis recovers the simulated values (`std_estimate` should be close to the simulated mean difference of 0.2).
-Second, it enables you to plan an appropriate sample size because you are able to quantify the sampling variability.
-
-To complete a such a  Monte-Carlo simulation study we need a function that repeats the analysis for different conditions (here only for different n):
-
-```{r simulation_study}
-```
-
-Such Monte-Carlo simulation can help you decide between several tradeoffs.
-One historically important tradeoff is between sample size and power (the probability to detect an effect of given size).
-More recently there has been growing interest in other resource relevant parameters as well [!citation], such as number of measurement occasions, reliability of measures etc.
-To show that there can also be alternative approaches for sample size planning beyond null hypothesis testing, we choose to visualize the density distribution of the parameter of interest (latent mean difference by gender of Machiavellism) instead of power.
-Such process still lets you formulate expectations for the precision of the analysis, e.g.
-with a sample size of 1000 per group 80% of the simulated standardized effect size fall into a range of +/- .1, but without necessarily rely on null hypothesis testing.
-For simplicity we do not show the code for the simulation/graphic, but it is available in the accompanying [GitHub repository](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo]:.
-
-[^gh-repo]: https://github.com/aaronpeikert/repro-tutorial
-
-```{r, include=FALSE}
-res <- read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddd")
-```
-
-```{r hash}
-```
-
-```{r, include=FALSE}
-res_interval <- res %>%
-  select(n_obs, std_estimate) %>% 
-  filter(!is.na(std_estimate)) %>% 
-  group_by(n_obs) %>% 
-  summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
-            .groups = "drop", median = median(std_estimate)) %>% 
-  unnest(c(interval)) %>% 
-  mutate(width = upper - lower, perc = 1 - alpha) %>% 
-  group_by(alpha) %>% 
-  arrange(n_obs) %>% 
-  mutate(across(c(width, upper, lower),
-                ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE))) %>% 
-  ungroup() %>% 
-  arrange(perc)
-```
-
-```{r, include=FALSE}
-choosen_perc <- .8
-choosen_width <- .2
-true_diff <- .2
-est_diff <- res %>% filter(n_obs == max(n_obs)) %>% pull(std_estimate) %>% mean()
-minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
-  filter(n_obs == min(n_obs))
-sim_n <- pull(res, n_obs)
-sim_nmax <- max(sim_n)
-sim_n_unique <- unique(sim_n)
-sim_nstep <- sim_n_unique[2] - sim_n_unique[1]
-```
-
-```{r, include=FALSE}
-plot_interval <- res_interval %>%
-  ggplot(aes(
-    n_obs,
-    ymin = lower,
-    ymax = upper,
-    fill = perc,
-    group = alpha
-  )) +
-  geom_ribbon() +
-  geom_line(aes(y = median), color = "white") +
-  geom_hline(yintercept = true_diff, linetype = "dotted") +
-  scale_fill_viridis_c(
-    begin = 1,
-    end = 0,
-    breaks  = 1 - seq(0.05, .2, 0.05),
-    labels = . %>% scales::percent(accuracy = 1)
-  ) +
-  scale_y_continuous(n.breaks = 10) +
-  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
-  theme_minimal() +
-  labs(x = "N per Group", y = "Standardized Effect Size") +
-  theme(legend.title = element_blank()) +
-  NULL
-```
-
-```{r, include=FALSE}
-plot_width <- res_interval %>%
-  ggplot(aes(n_obs, width, color = perc, group = perc)) +
-  geom_line() +
-  geom_point(
-    data = minn,
-    size = 3,
-    color = "black",
-    shape = 3
-  ) +
-  scale_color_viridis_c(
-    begin = 1,
-    end = 0,
-    breaks = 1 - seq(0.05, .2, 0.05),
-    labels = . %>% scales::percent(accuracy = 1)
-  ) +
-  theme_minimal() +
-  scale_y_log10(breaks = seq(0, 0.8, .1)) +
-  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
-  labs(x = "N per Group", y = "Width of quantile\n(Standardized Effect Size)") +
-  theme(legend.title = element_blank()) +
-  NULL
-```
-
-```{r, include=FALSE}
-simulation_plot <- plot_interval +
-  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + 
-  plot_width +
-  plot_layout(nrow = 2, guides = 'collect') +
-  plot_annotation(tag_levels = 'A')
-```
-
-```{r, simulation-plot, echo=FALSE, fig.cap="*Plot A* depicts the range of effect sizes we observed in the simulation, where the colour encodes the density.
-The light yellow interval shows where 80% of the estimates land.
-The dotted line represents the simulated true effect size; the white line shows the median observed effect sizes.
-*Plot B* represents the width of the interval shown in Plot A on a logarithmic scale.
-The cross shows the minimum required sample size so that 80% of the estimates land in a range of +/- .1.
-"}
-simulation_plot
+generate_data
+```
+
+Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R", "GitHub")`.
+
+With its help we can simulate 500 cases and fit our analysis.
+
+```{r}
+true_diff <- 0.2
+model_truth <- combine(
+  measurement(
+    "MACH",
+    items("x", 9),
+    loadings = round(loadings_jones_paulhus * 0.7, 2)
+  ),
+  intercepts("MACH", list(0, true_diff)),
+  variances("MACH", list(1, 1))
+)
+
+model_same <- 
+  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+  MACH ~ c(0,0)*1
+  MACH ~~ c(1, NA)*MACH"
+
+model_differ <- 
+  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
+  MACH ~ c(0, NA)*1 + c(zero, diff)*1
+  MACH ~~ c(1, NA)*MACH"
+generate_data(500,
+              model_truth,
+              model_same,
+              model_differ,
+              list(std_estimate = get_std_estimate),
+              group.equal = c("loadings", "intercepts"),
+              mod_load = 3,
+              mod_int = 3,
+              std.lv = TRUE)
+```
+
+Note, that this certainly results in different numbers each time we run it.
+
+```{r}
+generate_data(500,
+              model_truth,
+              model_same,
+              model_differ,
+              list(std_estimate = get_std_estimate),
+              group.equal = c("loadings", "intercepts"),
+              mod_load = 3,
+              mod_int = 3,
+              std.lv = TRUE)
 ```
@@
 It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
-The standard random number generator (RNG) produces high quality random numbers, but is not safe to use in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
+The standard random number generator (RNG) produces high quality random numbers, but has is not safe to use in parrallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
 The problem is that concurrent random number streams are correlated, because they repeat after some time.
@@
 @lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
-While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers:
+While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the [`future`-package](futureverse.org)[^futureverse] to handles this for us across all workers:
 
@@
 PAC essentially reverses the process of empirical research.
-Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results---initially based on simulated data.
+Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results --- initially based on simulated data.
 Then, the source code of this manuscript serves as an unambiguous description of the analysis plan and is published as a preregistration.
@@
 PAC works well within this reproducible framework.
-
 First, the preregistration becomes a part of the reproducible repository.
 It is part of the set of version-controlled files and provides transparency about when the preregistration file, data files, and code files were changed over the course of a research project.
-If code was version-controlled and tagged (see Section [Git]) before data collection, all deviations from the preregistration or any post-hoc additions that were performed after data collection are transparent and are visible and comparable in the version history.
+If code was version-controlled and tagged (see Section XXX) before data collection, all deviations from the preregistration or any post-hoc additions that were performed after data collection are transparent and are visible and comparable in the version history.
 
 One way to ensure that you can expect your analysis to work with real data is to simulate data.
+A data simulation allows you to verify that the code runs error-free and lets you form more realistic expectations about the estimates' correctness and precision.
 We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
-The simulation showed that we have to gather at least `r minn$n_obs` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `r round((est_diff - true_diff)*100/true_diff, 2)`%.
-These results are also visualized in Figure \@ref(fig:simulation-plot).
+The simulation showed that we have to gather at least `missing_rcode_simulation` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `missing_rcode_simulation`%.
 If you went to the trouble to simulate data, we argue that you get a preregistration almost for free.
@@
 See the section on [R Markdown] for an example of a dynamic document generation paradigm.
-In fact, this manuscript and therefore the here presented Machiavellism example were preregistered as a full manuscript, and you can see the results on real data in section [Reproduce a Project].
 
@@
 As @knuthCWEBSystemStructured states: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
-We already learned about [Markdown], to write text.
-R Markdown allows you to include R code (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown.
-R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format
+Above, we have already suggested [Markdown] as a universal and portable text format. 
+R Markdown extends regular Markdown by allowing useres you to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
+Upon rendering of the document, the code blocks are executed and their output is dynamically inserted into the document.
+This allows the creation of conditional text blocks, statistical results, or figures that are guarantueed to be uptodate because they are created everytime anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
 To create an R Markdown document, click on:
@@
 You could, in principle, even write large parts of the manuscript as part of your preregistration, i.e. the introduction, theoretical background, the method section and parts of the discussion.
-If you have simulated data, you could even include fake results.
-Packages like [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf)[^stargazer] [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html)[^apaTables] [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/)[^papaja] [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report)[^report] [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
+If you have simulated data, you could even already generate all figures and statistical results based on fake data, which will later be replaced by the real data.
 Such a dynamic document is arguably the most concrete preregistration possible.
@@
 
-[^report]: https://github.com/easystats/report
-[^stargazer]: https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf
-[^apaTables]: https://dstanley4.github.io/apaTables/articles/apaTables.html
-[^papaja]: http://frederikaust.com/papaja_man/
 <!--Hands on: gitignore resulting .html/pdf-->
@@
 `repro` reminds you of that and suggests which file you might want to add to the target `all`.
-You can generally change the `Makefile` anyway you want, e.g. to add targets which are not supported by `repro`.
 After you have done so, you can then run in the terminal:
@@
 ```
-The function `automate()` combines `automate_docker()` & `automate_make()`.
-
-You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manual` to install software beyond R packages.
-Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manual`.
-We recomend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
+
+You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manuel` to install software beyond R packages.
+Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manuel`.
+We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
 
@@
 
-Because you probably want to reproduce your own project each time something changes, you will often want to reproduce evertyhting.
-The `repro`-package contains a function designed for that purpose:
-
-```{r}
-repro::rerun()
-```
-
-This function automatically figures out how to reproduce a project and also recognizes other employed reproducibility tools (like `renv` and `singularity`), but can of course only guess the appropriate way to reproduce a project.
-
-If you have written a [Preregistration as Code] you can swap the data file and reproduce the project.
-In fact, till after we preregistered the study `r link("https://github.com/aaronpeikert/repro-tutorial/releases/tag/ v0.0.1-preregistration", "preregistration")` were the results below based on simulated data.
-Only after we preregistered the whole manuscript and upon submission, they were based on actual data from https://openpsychometrics.org/_rawdata/.
-
-```{r}
-read_csv(here("data", "sd3.csv"), col_types = "dddddddddd") %>% 
-  planed_analysis() %>% 
-  extract_results()
-```
-
-Note that both [this tutorial](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo2] as well as our article on the [reproducible workflow](https://github.com/aaronpeikert/reproducible-research)[^gh-repo3] [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories.
-
-[^gh-repo2]: https://github.com/aaronpeikert/repro-tutorial
-[^gh-repo3]: https://github.com/aaronpeikert/reproducible-research
+Note that both this tutorial as well as our article on the reproducible workflow [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories.
+<!-- maybe we add here the real data? #25 -->
 
@@
 We use containers to guarantee that all code components are available and work as expected.
-Ideally, all data is also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs], but this is not always feasible.
-Sometimes it is just the required storage capacity that exceeds what Git can handle (Git gets slow around more than 100mb).
-In such case we strongly recommend using [DataLad [@DataladDatalad2021]](http://handbook.datalad.org/en/latest/)[^datalad] or the below procedure to guarantee that the data has not changed.
-In practice, however, there are various reasons, which may prohibit you from making your data publicly available, for example legal or regulatory or privacy concerns.
+Ideally, all data is also stored in the repository.
+In practice, however, there are various reasons, which may prohibit you from making your data available, for example legal or regulatory or privacy concerns.
 You may find yourself in a situation where you can make the data only available on request.
@@
 To verify the data integrity, you can include a line of code after reading the data that looks like this:
-
-[^datalad]: http://handbook.datalad.org/en/latest/
 
@@
 
-An compromise between making the data publicly available and keeping it closed is to publish synthetic data [@vanlissa2020worcs].
+An compromise between making the data publicly available and keeping it closed is to publish synthetic data.
 Such approach tries to recreate the marginal distribution and bivariate relationships in simulated data, e.g. with the R package `synthpop`[@nowokSynthpopBespokeCreation2016].
@@
 
-> Log into zenodo.org → Account → [GitHub](https://zenodo.org/account/settings/github/)[^zenodo-github]
+> Log into zenodo.org → Account → [GitHub[^zenodo-github]](https://zenodo.org/account/settings/github/)
 
@@
 We found it useful to label the following stages of development as release:
-
-* Preregistration (repository contains the final data analysis plan; also see the Section on PAC)
-* Completion of data collection (repository contains all study data)
-* Preprint (repository exactly reproduces a preprint that is hosted externally)
-* Submission or Resubmission (repository exactly reproduces a given version (re)submitted to a journal)
-* In Press (repository exactly reproduces the final manuscript as accepted by a scientific journal)
+- Preregistration (repository contains the final data analysis plan; also see the Section on PAC)
+- Completion of data collection (repository contains all study data)
+- Preprint (repository exactly reproduces a preprint that is hosted externally)
+- Submission or Resubmission (repository exactly reproduces a given version (re)submitted to a journal)
+- In Press (repository exactly reproduces the final manuscript as accepted by a scientific journal)
 
@@
 By extending the YAML specification that is used for storing meta information in R Markdown files, which are becoming increasingly popular among R users, repro helps users to implement @Peikert2019's workflow with minimal effort.
-Most importantly, `repro`-is implemented such that the package itself is not needed to reproduce a workflow that was generated according to the workflow.
-Importantly, this means that users that adopt the workflow do not have to rely on the future availability and maintenance of the `repro`-package.
+Most importantly, repro is implemented such that the package itself is not needed to reproduce a workflow that was generated according to the workflow.
+Importantly, this means that users that adopt the workflow do not have to rely on the future availability and maintenance of the repro package.
 

### 2021-05-17 manuscript.Rmd

--- 2021-05-15 manuscript.Rmd
@@
  This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations---and even your CV [@vitae].
+We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations---and even your CV.
 To create a Markdown document, click on:
@@
 
-Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration [@osfstandard; @socpsytemp; @howtoreplicate]. 
+Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration. 
 Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code.
@@
 As @knuthCWEBSystemStructured states: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
-We already learned about [Markdown], to write text.
-R Markdown allows you to include R code (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown.
-R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format
+
+Above, we have already suggested [Markdown] as a universal and portable text format. 
+R Markdown extends regular Markdown by allowing useres you to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
+Upon rendering of the document, the code blocks are executed and their output is dynamically inserted into the document.
+This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
+
 To create an R Markdown document, click on:
@@
 ### Data Documentation
-
-[@codebook]
 
@@
 Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manual`.
-We recomend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
+We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
 
@@
 Whereas this is no issue on most modern cloud computing infrastructures, it is often an issue in classic HPC environments.
-Singularity is an alternative solution [@singularity].
+Singularity is an alternative solution.
 It was designed to run without administratot rights, while maintaining full compatibility with Docker images.
@@
 <!-- Hands-on: check changed hash -->
-A prerequisite to reproducibility is the availability of code and data [@akhlaghi2020towards].
+A prerequisite to reproducibility is the availability of code and data.
 We use containers to guarantee that all code components are available and work as expected.

### 2021-05-17 manuscript.Rmd
more on simulations

@@
 
-Another advantage of Markdown is that because of its alikeness to plain text, you can easily version-control it with Git and leverage Gits features for tracking, displaying, and merging sequential and concurrent changes.
+Another advantage of Markdown is that because of its likeness to plain text, you can easily version-control it with Git and leverage Gits features for tracking, displaying, and merging sequential and concurrent changes.
 You may be familiar with Microsoft Words' "Track Changes"-Feature.
@@
 The machine learning community has begun to routinely simulate data to test models and their implementation [@braiekTestingMachineLearning2020].
-Beyond checking the technical side of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
-Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
-Beyond ensuring that the code runs without technical error and it lets you form more realistic expectations about the estimates' correctness and precision.
-No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about them.
-If you identify unrealistic assumptions, you can either violate these assumptions and see if the analytic strategy is robust or modify the analytical approach to address them.
+Other than checking technical and statistical aspects of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
+With simulated data you can check if the expected data and the planned analysis match.
+First of all, simulated data can be used as an initial check that the code for the planned analysis runs without error. 
+This is particularly important if you want to preregister pieces of your analysis before data were collected.
+For example, we can use simulated data to check whether a given data analysis outputs parameter estimates when we simulate missing data values.
+Second, Monte Carlo simulations allow us to form realistic expectations about the unbiasedness and precision of parameter estimates and it can help with sample size planning to obtain a sensible study design.
+For a variety of standard models, there are analytic strategies to evaluate precision or statistical power, however, Monte Carlo simulation offers more flexibility in terms of mode realistic modeling conditions (e.g., missing data or non-normality).
+For example, @maccallum1999sample used Monte Carlo simulations to investigate optimal sample size for exploratory factor analysis, or @curran1996robustness studied the robustness of the $\chi^2$ test statistic under non-normal data.
+No matter which analytic strategy you plan, it will come with assumptions about the data generating process under investigation. 
+A simulation forces us to deal with these assumptions and their consequences explicitly.
+If you identify unrealistic assumptions, you can either deliberately violate these assumptions and, by simulation, check if the analytic strategy is robust or modify the modeling approach to explicitly address them.
 We will show an example using structural equation modeling (SEM) for our Machiavellism example.
-However, the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
-One has to make many educated guesses to simulate data that closely matches your expectation.
-These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
-The analytic strategy (i.e. SEM) often gives you a hint about how to simulate data.
-
-If you are unfamiliar with the specifics of SEM, you may briefly skim the next sections and continue in section ["Randome Number Generation"](#prng).
+Note that the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
+Typically, you have to make a number of educated guesses to simulate data that closely matches your expectation.
+These guesses strongly depend on the study you are planning, and there cannot be a one-fits-all solution.
+Since we typically use parametric statistical models (general linear model or SEM) or hypothesis tests that can be expressed as Gaussian models, we can usually use these models directly to simulate data.
+
+If you are unfamiliar with the specifics of SEM, you may briefly skim the next sections and continue in section ["Random Number Generation"](#prng).
 To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.
@@
 The estimates for the construction sample, which was recruited from Amazon’s Mechanical Turk (MTurk) and consists of 489 adults from Canada and the United States, can be seen in Table \@ref(tab:loadings))
-They report a gender difference (at the manifest level) depending on the study between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
+They report a gender difference (at the manifest level), which depending on the study ranges between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
 
@@
 We overlooked a crucial assumption of this model, that is, measurement invariance.
-This assumption is also baked into our simulation of the data.
+This assumption has also been incorporated into our simulation of the data.
 There is a good chance that this assumption is violated.
-We, therefore, adapt our analysis to plan for this contingency.
-After fitting the model with strong measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
+We therefore adjust our analysis plan to include tests for measurement invariance.
+We propose the following procedure. 
+After fitting the model with strong measurement invariance <!--AP: citation needed-->, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
 The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
-While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
-Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
-For simplicity, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).
-Anyhow, remember that what we consider realistic for our case could be unfit for your case.
+While reconsidering our modeling assumptions, it is also unlikely that the data are perfectly normally distributed.
+First, we have rating scale items and thus our observations are categorical.
+Second, @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
+For simplicity of our illustration, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).
 
@@
 
-Note, that this certainly results in different numbers each time we run it (see ["Randome Number Generation"](#prng)).
+Note, that this certainly results in different numbers each time we run it (see ["Random Number Generation"](#prng)).
 Of course it is already helpful to simulate a single dataset to test if the analysis works in principle, but repeating the analysis several times and vary parameters lets you verify more properties of your analysis.
@@
 
-### Randome Number Generation (optional) {#prng}
+### Random Number Generation (optional) {#prng}
 
@@
 To consistently get the same results while preserving the pseudo-random property of the generated numbers, it is good practice to set a random seed that sets the PRNG into a restorable state, in R code: `set.seed(1235)`.
+Note however that the consistency of a given seed is not necessarily maintained across different R versions because the default algorithm for generating pseudo random numbers may change (see @Peikert2019 for a demonstration with different R versions).
 It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).

### 2021-05-18 manuscript.Rmd
added abstract removed toc

@@
     keep_tex: TRUE
+    toc: FALSE
 repro:
@@
 bibliography: ["references.bib", "packages.bib"]
+abstract:
+  Computational reproducibility is the ability to obtain identical results from the *same* data with the *same* computer code.
+It is a building block for transparent and cumulative science because it enables other researchers, on other computers and later in time, to reproduce and thus understand how results came about while, at the same time, it avoids a variety of errors that may lead to erroneous reporting of statistical and computational results.
+In this tutorial, we demonstrate how the R package `repro` supports researchers in creating fully computationally reproducible research projects with a focus on current Open Science practices; starting from sketching the research proposal, over preregistering an analysis as executable code with simulated data, until the creation and dissemination of the final scientific report.
 ---
@@
 
-```{r}
+```{r echo=FALSE}
 # define a ürint method for objects of the class data.frame
@@
 
-The `repro` package is particularly well-suited for computationally intensive data analyses and simulations that  are to be run on distributed computing environments.
-By using containers, `repro` ensures identical execution across platforms and thus guarantees that a data analysis or simulation that was designed and tested on a local computer will be executed in the identical manner on a remote machine or various remote machines across an entire computing cluster.
+The `repro` package is particularly well-suited for computationally intensive data analyses and simulations that  are to run on distributed computing environments.
+Through the use of containers, `repro` ensures identical execution across platforms, guaranteeing that a given data analysis or simulation designed and tested on a local machine will execute identically on a remote machine or multiple different remote machines across an entire computing cluster.
 Thereby, the repro package makes massive parallel computing in heterogeneous computing environments less error-prone while ensuring the full documentation of all computational steps taken — even if some or all of them were run in parallel on distributed platforms.
-Alone the sheer data size often warrants distributed computing, e.g.
-the European Bioinformatics Institute stores more than 120 Petabyte of publicly accessible data [@cookEuropeanBioinformaticsInstitute2018], single FMRI studies quickly amount to hundreds of Gigabyte[^openneuro] and even purely behavioral data like from the moral machine experiment [>10GB, @awadMoralMachineExperiment2018] can become to big to handle .
+The sheer volume of data often necessitates distributed computing, e.g.
+the European Bioinformatics Institute stores more than 120 Petabyte of publicly accessible data [@cookEuropeanBioinformaticsInstitute2018], single FMRI studies quickly amount to hundreds of Gigabyte[^openneuro] and even purely behavioral data like those from the moral machine experiment [>10GB, @awadMoralMachineExperiment2018] can become too large to be handled by a single local computer.
 
@@
 
-As default, `repro` uses Docker, which is the defacto standard for cloud computing, as a container solution.
+As default container solution, `repro` uses Docker, which is a widely adopted standard for cloud computing.
 However, Docker requires administrator rights for execution.
@@
 Singularity is an alternative solution.
-It was designed to run without administratot rights, while maintaining full compatibility with Docker images.
+It was designed to run without administrator rights, while maintaining full compatibility with Docker images.
 If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
@@
 
-## Reproduce a Project
+## Reproducing a Project
 
@@
 Figure \@ref(fig:readme) shows `r link("https://github.com/aaronpeikert/repro-tutorial#how-to-reproduce-this-manuscript", "the reproduction statement of this project")`.
-E.g. you can download any git repository with these commands:
+For example, you can download any git repository with these commands:
 
@@
 
-Projects that combine Make and Docker then can usually be reproduced with:
+Projects that combine Make and Docker and adhere to the standard proposed by @Peikert2019 can be reproduced with:
 
@@
 
-Because you probably want to reproduce your own project each time something changes, you will often want to reproduce evertyhting.
+Because you probably want to reproduce your own project each time something changes, you will often want to reproduce everything.
 The `repro`-package contains a function designed for that purpose:
@@
 
-If you have written a [Preregistration as Code] you can swap the data file and reproduce the project.
-In fact, till after we preregistered the study `r link("https://github.com/aaronpeikert/repro-tutorial/releases/tag/ v0.0.1-preregistration", "preregistration")` were the results below based on simulated data.
+If you have written a [Preregistration as Code] you can add the real data file to the project and reproduce the entire project with this single command.
+In fact, until after we preregistered the study `r link("https://github.com/aaronpeikert/repro-tutorial/releases/tag/ v0.0.1-preregistration", "preregistration")` were the results below based on simulated data.
 Only after we preregistered the whole manuscript and upon submission, they were based on actual data from https://openpsychometrics.org/_rawdata/.
@@
 <!-- Hands-on: check changed hash -->
-A prerequisite to reproducibility is the availability of code and data.
+A prerequisite to reproducibility is the permanent availability of both code and data.
 We use containers to guarantee that all code components are available and work as expected.
-Ideally, all data is also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs], but this is not always feasible.
+Ideally, all data are also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs], but this is not always feasible.
 Sometimes it is just the required storage capacity that exceeds what Git can handle (Git gets slow around more than 100mb).
@@
 
-To verify the data integrity, you can include a line of code after reading the data that looks like this:
+To verify data integrity, you can include a line of code after reading the data that looks like this:
 
@@
 
-We employed the same mechanism for our simulation to verify that the simulation is really reproducible[^simhash].
+We employed the same mechanism for our simulation to verify that the simulation is indeed reproducible[^simhash].
 
@@
 
-An compromise between making the data publicly available and keeping it closed is to publish synthetic data [@vanlissa2020worcs].
-Such approach tries to recreate the marginal distribution and bivariate relationships in simulated data, e.g. with the R package `synthpop`[@nowokSynthpopBespokeCreation2016].
-The results of such data should generally not be interpreted.
-However, synthetic data provides an opportunity for readers to test if a project is in principle reproducible, even though they have no access to the data.
+A compromise between making the data publicly available and keeping it closed is the publication of synthetic data [@vanlissa2020worcs].
+Synthetic data are an attempt to recreate the marginal distribution and bivariate relationships in simulated data.
+The R package `synthpop`[@nowokSynthpopBespokeCreation2016] provides functions to generate synthetic data .
+Of course, the results of such synthetic data should generally not be interpreted but are rather meant to illustrate the workings and the structure of the output of a given data analysis.
+Also, synthetic data provides an opportunity for readers to test if a project is in principle reproducible, even though they have no access to the data.
 

### 2021-05-18 manuscript.Rmd
some modifications to rerun()

@@
 
-Ideally, you have a statement in the README that contains a few steps to reproduce the project, where most steps are automatic and standardized.
-This statement should also note the required software.
+Ideally, you have a statement in the README that contains a short description of how to reproduce the project, whereas most steps are automatic and standardized.
+This statement should also list all software requirements.
+<!--AP: I think this Figure is superfluous. Please remove. All the information is in the following code blocks.-->
 Figure \@ref(fig:readme) shows `r link("https://github.com/aaronpeikert/repro-tutorial#how-to-reproduce-this-manuscript", "the reproduction statement of this project")`.
@@
 
-Projects that combine Make and Docker and adhere to the standard proposed by @Peikert2019 can be reproduced with:
+Projects that combine Make and Docker and adhere to the standard proposed by @Peikert2019 can then be reproduced with:
 
@@
 
-Because you probably want to reproduce your own project each time something changes, you will often want to reproduce everything.
-The `repro`-package contains a function designed for that purpose:
+Typically there is a need to repeatedly reproduce a project during the development phase (e.g., on every change of  the analysis cod or, every change to the dataset).
+The `repro`-package contains a function to reproduce all targets:
 
@@
 
-This function automatically figures out how to reproduce a project and also recognizes other employed reproducibility tools (like `renv` and `singularity`), but can of course only guess the appropriate way to reproduce a project.
-
-If you have written a [Preregistration as Code] you can add the real data file to the project and reproduce the entire project with this single command.
+This function automatically figures out how to reproduce an entire project and also recognizes non-default reproducibility tools (like `renv` and `singularity`). 
+Of course, this automatic derivation of tools and steps is limited to the packages we have discussed here and the program may not always succeed if researchers deviate from our suggestions.
+
+If you have written a [Preregistration as Code] you can add the real data file to the project and (re)produce the entire project with this single command.
 In fact, until after we preregistered the study `r link("https://github.com/aaronpeikert/repro-tutorial/releases/tag/ v0.0.1-preregistration", "preregistration")` were the results below based on simulated data.

### 2021-05-19 manuscript.Rmd
Add citations. (#76)
* Add citations.

* fix #53

* remove faulty coded chars and abstracts

* add counter paper
#49

* place citations

Co-authored-by: Aaron Peikert <aaron.peikert@posteo.de>

--- 2021-05-17 manuscript.Rmd
@@
 
-Another advantage of Markdown is that because of its likeness to plain text, you can easily version-control it with Git and leverage Gits features for tracking, displaying, and merging sequential and concurrent changes.
+Another advantage of Markdown is that because of its alikeness to plain text, you can easily version-control it with Git and leverage Gits features for tracking, displaying, and merging sequential and concurrent changes.
 You may be familiar with Microsoft Words' "Track Changes"-Feature.
@@
 
-Now assume that after discussing with your colleagues, you settle onto a final analysis strategy using latent variable models with tests for measurement invariance because power is greater and you want to exclude bias in measurement as possible confounder <!-- Aaron, reference needed-->.
+Now assume that after discussing with your colleagues, you settle onto a final analysis strategy using latent variable models with tests for measurement invariance because power is greater and you want to exclude bias in measurement as possible confounder [@flake2020].
 This results in changes to your notes as follows:
@@
 <!--I am troubled by how we should recommend to learn Git-->
-Git can be challenging to learn, because it requires a rather complex mental model of maintaining potentially differing versions between different repositories (typically one central repository and one local repository per each author while each repository may have different branches).
-However, there is a subsets of commands which we present here and in the section on [GitHub] (commit, push, pull) which enable you to use Git for simple version control.
+Git can be challenging to learn, because it requires a rather complex mental model.
+However, there is a subsets of commands which we present here and in the section on [GitHub] (commit, push, pull) which enable you to utilize Git for simple version control.
 There are many tutorials on learning Git and @vuorreCuratingResearchAssets2018 provide a good starting point.
@@
 The machine learning community has begun to routinely simulate data to test models and their implementation [@braiekTestingMachineLearning2020].
-Other than checking technical and statistical aspects of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
-With simulated data you can check if the expected data and the planned analysis match.
-First of all, simulated data can be used as an initial check that the code for the planned analysis runs without error. 
-This is particularly important if you want to preregister pieces of your analysis before data were collected.
-For example, we can use simulated data to check whether a given data analysis outputs parameter estimates when we simulate missing data values.
-Second, Monte Carlo simulations allow us to form realistic expectations about the unbiasedness and precision of parameter estimates and it can help with sample size planning to obtain a sensible study design.
-For a variety of standard models, there are analytic strategies to evaluate precision or statistical power, however, Monte Carlo simulation offers more flexibility in terms of mode realistic modeling conditions (e.g., missing data or non-normality).
-For example, @maccallum1999sample used Monte Carlo simulations to investigate optimal sample size for exploratory factor analysis, or @curran1996robustness studied the robustness of the $\chi^2$ test statistic under non-normal data.
-No matter which analytic strategy you plan, it will come with assumptions about the data generating process under investigation. 
-A simulation forces us to deal with these assumptions and their consequences explicitly.
-If you identify unrealistic assumptions, you can either deliberately violate these assumptions and, by simulation, check if the analytic strategy is robust or modify the modeling approach to explicitly address them.
+Beyond checking the technical side of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
+Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
+Beyond ensuring that the code runs without technical error and it lets you form more realistic expectations about the estimates' correctness and precision.
+No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about them.
+If you identify unrealistic assumptions, you can either violate these assumptions and see if the analytic strategy is robust or modify the analytical approach to address them.
 We will show an example using structural equation modeling (SEM) for our Machiavellism example.
-Note that the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
-Typically, you have to make a number of educated guesses to simulate data that closely matches your expectation.
-These guesses strongly depend on the study you are planning, and there cannot be a one-fits-all solution.
-Since we typically use parametric statistical models (general linear model or SEM) or hypothesis tests that can be expressed as Gaussian models, we can usually use these models directly to simulate data.
-
-If you are unfamiliar with the specifics of SEM, you may briefly skim the next sections and continue in section ["Random Number Generation"](#prng).
+However, the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
+One has to make many educated guesses to simulate data that closely matches your expectation.
+These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
+The analytic strategy (i.e. SEM) often gives you a hint about how to simulate data.
+
+If you are unfamiliar with the specifics of SEM, you may briefly skim the next sections and continue in section ["Randome Number Generation"](#prng).
 To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.
@@
 The estimates for the construction sample, which was recruited from Amazon’s Mechanical Turk (MTurk) and consists of 489 adults from Canada and the United States, can be seen in Table \@ref(tab:loadings))
-They report a gender difference (at the manifest level), which depending on the study ranges between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
+They report a gender difference (at the manifest level) depending on the study between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
 
@@
 Data simulation is particularly easy in lavaan syntax using the function `simulateData`.
-More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem [@simsem], or simstudy [@goldfeldSimstudyIlluminatingResearch2020].
+More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem [@simsem], faux [@faux], or simstudy [@goldfeldSimstudyIlluminatingResearch2020].
 The following code simulates data from the population model `model_truth` with 1000 cases in each group.
@@
 We overlooked a crucial assumption of this model, that is, measurement invariance.
-This assumption has also been incorporated into our simulation of the data.
+This assumption is also baked into our simulation of the data.
 There is a good chance that this assumption is violated.
-We therefore adjust our analysis plan to include tests for measurement invariance.
-We propose the following procedure. 
-After fitting the model with strong measurement invariance <!--AP: citation needed-->, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
+We, therefore, adapt our analysis to plan for this contingency.
+After fitting the model with strong measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
 The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
-While reconsidering our modeling assumptions, it is also unlikely that the data are perfectly normally distributed.
-First, we have rating scale items and thus our observations are categorical.
-Second, @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
-For simplicity of our illustration, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).
+While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
+Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
+For simplicity, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).
+Anyhow, remember that what we consider realistic for our case could be unfit for your case.
 
@@
 
-Note, that this certainly results in different numbers each time we run it (see ["Random Number Generation"](#prng)).
+Note, that this certainly results in different numbers each time we run it (see ["Randome Number Generation"](#prng)).
 Of course it is already helpful to simulate a single dataset to test if the analysis works in principle, but repeating the analysis several times and vary parameters lets you verify more properties of your analysis.
@@
 
-### Random Number Generation (optional) {#prng}
+### Randome Number Generation (optional) {#prng}
 
@@
 To consistently get the same results while preserving the pseudo-random property of the generated numbers, it is good practice to set a random seed that sets the PRNG into a restorable state, in R code: `set.seed(1235)`.
-Note however that the consistency of a given seed is not necessarily maintained across different R versions because the default algorithm for generating pseudo random numbers may change (see @Peikert2019 for a demonstration with different R versions).
 It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
@@
 A preregistration offers the interested reader more insight into the research process than a paper alone could.
-Some [citation!] argue that preregistration can not fulfil its promise to reduce the false discovery rate of science if we allow deviating from them.
-While such an argument has some merits, we believe that often transparency outweighs these concerns.
+@szollosi_is_2020 argue that preregistration can not fulfil its promise to reduce the false discovery rate of science if we allow deviating from them.
+While such an argument has some merits, we believe that often transparency outweighs these concerns [@nosekPreregistrationHardWorthwhile2019].
 Deviation from a preregistration is still possible.
@@
 
-Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration. 
-Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code.
+Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration [@osfstandard; @socpsytemp; @howtoreplicate]. 
+Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code [@sep-ambiguity].
 Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity.
@@
 As @knuthCWEBSystemStructured states: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
-
-Above, we have already suggested [Markdown] as a universal and portable text format. 
-R Markdown extends regular Markdown by allowing useres you to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
-Upon rendering of the document, the code blocks are executed and their output is dynamically inserted into the document.
-This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
-
+We already learned about [Markdown], to write text.
+R Markdown allows you to include R code (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown.
+R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format
 To create an R Markdown document, click on:
@@
 Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manual`.
-We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
+We recomend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
 
@@
 <!-- Hands-on: check changed hash -->
-A prerequisite to reproducibility is the availability of code and data.
+A prerequisite to reproducibility is the availability of code and data [@akhlaghi2020towards].
 We use containers to guarantee that all code components are available and work as expected.
-Ideally, all data is also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs], but this is not always feasible.
+Ideally, all data is also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs] and well documented [@codebook], but this is not always feasible.
 Sometimes it is just the required storage capacity that exceeds what Git can handle (Git gets slow around more than 100mb).
@@
 In practice, however, there are various reasons, which may prohibit you from making your data publicly available, for example legal or regulatory or privacy concerns.
-You may find yourself in a situation where you can make the data only available on request.
+You may find yourself in a situation where you can make the data only available on request [e.g. @share].
 A major threat to reproducibility is that the data has changed in the meantime or multiple versions are circulating and it is not exactly clear what version was used originally.

### 2021-05-19 manuscript.Rmd
cite rmarkdown formats closes #85

@@
     - future.batchtools
+    - rticles
   scripts:
@@
 ```{r}
-# define a ürint method for objects of the class data.frame
+# define a print method for objects of the class data.frame
 knit_print.data.frame = function(x, ...) {
@@
  This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations---and even your CV [@vitae].
+We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
 To create a Markdown document, click on:
@@
 ```{r, include=FALSE}
-knitr::write_bib(c(.packages(), "gert", "bookdown", "lavaan", "knitr", "targets", "renv", "tidyverse"),
+knitr::write_bib(c(.packages(), "rticles", "gert", "bookdown", "lavaan", "knitr", "targets", "renv", "tidyverse"),
                  "packages.bib")

### 2021-05-19 manuscript.Rmd
add section on Archiving Docker #82

@@
 
+A valid point of criticism is that this workflow introduces a bulk of software solutions and services that may not be available for long.
+Superficially it relies on R, RStudio, Pandoc, Git, Make, Docker, Homebrew (OS X only) and Chocolaty (Windows only) as software and on GitHub, DockerHub, MRAN and Ubuntu Package Archive as services.
+These many breakpoints can be reduced to one service and one software.
+That is because all required software is bundled into the container, whose image can be stored.
+Hence, one needs a storage provider which saves the container image and the other files associated with the project.
+We already recommend to store all project files redundantly (your computer, on GitHub and zenodo.org).
+Building the Docker image from the Dockerfile still requires DockerHub, MRAN, and Ubuntu Package Archive to be online.
+We therefore suggest that you include an already build image into the releases.
+The Makefile generated by `repro` does include a target to facilitate this.
+To write a compressed image to disk use:
+
+```bash
+# this should be run in the terminal
+make save-docker
+```
+
+Unfortunately, this image can only be used when Docker is installed.
+To remove any software dependency including Docker from the archived project, one can convert the Docker image to a full system image.
+Such system image can subsequently be booted on any personal computer without relying on any software or service and therefore grantees reproducibility for decades to come.
+Currently, `repro` does not facilitate this conversion, but we hope to include this feature soon.
+
 ## Archiving Data

### 2021-05-19 manuscript.Rmd
remove empty sections closes #82
state size of docker only when image was saved closes #81

@@
 
+```{r, include=FALSE}
+container_size <- as.numeric(fs::file_size("reprotutorial.sif"))/1024^3
+```
+
 Modern data analysis is only feasible because researchers rely on much external software.
@@
 First, there is a community that maintains images for particular purposes.
-For example, there are ready-made images for R based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
+For example, there are ready-made images for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
 Users can then install whatever they additionally need on top of these pre-made images.
-The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
-It does include Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
+`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
+Our image includes Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
 Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
@@
 
-## Archiving Data
-
 <!-- Hands on: add commit hash to results-->
@@
 <!-- Hands on: explain how to archive docker images-->
-
-# Checklist
-
-<!-- Should we design a checklist? -->
 

### 2021-05-19 manuscript.Rmd
cache screenshot of readme

@@
 readme <- here::here("images", "readme.png")
-webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/",
+if (!file.exists(readme)) {
+  webshot2::webshot(
+    "https://github.com/aaronpeikert/repro-tutorial/",
     file = readme,
-                  vwidth = 1920, # height & width are not resolution
-                  vheight = 1080, # but what website thinks the browser size is
+    vwidth = 1920,
+    # height & width are not resolution
+    vheight = 1080,
+    # but what website thinks the browser size is
     selector = "div#readme",
-                  zoom = 2) # higher resolution
+    zoom = 2
+  ) # higher resolution
+}
 ```

### 2021-05-19 manuscript.Rmd
describe the example research question earlier

@@
 Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).
+To guide the reader through all steps of a research project, we will use a exemplary research question:
+
+> "Is there a mean difference in the personality trait 'Machiavellism' between genders?"
+
+This example serves only didactic purposes and not to derive substantive claims.
+Nevertheless, everything else is kept as close to a real research question as possible and we have preregistered the analysis without looking at the data beforehand and use actual data subsequently.
 

### 2021-05-19 manuscript.Rmd
remove dublicated sentence

--- 2021-05-19 manuscript.Rmd
@@
 
+```{r echo=FALSE}
 ```{r}
@@
 
-We have marked those steps that can be omitted as "optional" in the section headings.
 Table \@ref(tab:overview) shows all steps required to to create and open and reproducible research project as they are presented in the remainder of the manuscript.
@@
 Singularity is an alternative solution [@singularity].
-It was designed to run without administratot rights, while maintaining full compatibility with Docker images.
+It was designed to run without administrator rights, while maintaining full compatibility with Docker images.
+
 If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
@@
 
-Ideally, you have a statement in the README that contains a few steps to reproduce the project, where most steps are automatic and standardized.
-This statement should also note the required software.
+Ideally, you have a statement in the README that contains a short description of how to reproduce the project, whereas most steps are automatic and standardized.
+This statement should also list all software requirements.
+<!--AP: I think this Figure is superfluous. Please remove. All the information is in the following code blocks.-->
 Figure \@ref(fig:readme) shows `r link("https://github.com/aaronpeikert/repro-tutorial#how-to-reproduce-this-manuscript", "the reproduction statement of this project")`.
-E.g. you can download any git repository with these commands:
+For example, you can download any git repository with these commands:
 
@@
 
-Projects that combine Make and Docker then can usually be reproduced with:
+Projects that combine Make and Docker and adhere to the standard proposed by @Peikert2019 can then be reproduced with:
 
@@
 
-Because you probably want to reproduce your own project each time something changes, you will often want to reproduce evertyhting.
-The `repro`-package contains a function designed for that purpose:
+Typically there is a need to repeatedly reproduce a project during the development phase (e.g., on every change of  the analysis cod or, every change to the dataset).
+The `repro`-package contains a function to reproduce all targets:
 
@@
 
-This function automatically figures out how to reproduce a project and also recognizes other employed reproducibility tools (like `renv` and `singularity`), but can of course only guess the appropriate way to reproduce a project.
-
-If you have written a [Preregistration as Code] you can swap the data file and reproduce the project.
-In fact, till after we preregistered the study `r link("https://github.com/aaronpeikert/repro-tutorial/releases/tag/ v0.0.1-preregistration", "preregistration")` were the results below based on simulated data.
+This function automatically figures out how to reproduce an entire project and also recognizes non-default reproducibility tools (like `renv` and `singularity`). 
+Of course, this automatic derivation of tools and steps is limited to the packages we have discussed here and the program may not always succeed if researchers deviate from our suggestions.
+
+If you have written a [Preregistration as Code] you can add the real data file to the project and (re)produce the entire project with this single command.
+In fact, until after we preregistered the study `r link("https://github.com/aaronpeikert/repro-tutorial/releases/tag/ v0.0.1-preregistration", "preregistration")` were the results below based on simulated data.
 Only after we preregistered the whole manuscript and upon submission, they were based on actual data from https://openpsychometrics.org/_rawdata/.
@@
 <!-- Hands-on: check changed hash -->
-A prerequisite to reproducibility is the availability of code and data [@akhlaghi2020towards].
+A prerequisite to reproducibility is the permanent availability of both code and data [@akhlaghi2020towards].
 We use containers to guarantee that all code components are available and work as expected.
-Ideally, all data is also stored in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs] and well documented [@codebook], but this is not always feasible.
+Ideally, all data are also stored well documented [@codebook] in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs], but this is not always feasible.
+
 Sometimes it is just the required storage capacity that exceeds what Git can handle (Git gets slow around more than 100mb).

### 2021-05-19 manuscript.Rmd
first attempt to rewrite simulation section

@@
 
-There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
-A step to prevent unpleasant surprises is to simulate data.
-The machine learning community has begun to routinely simulate data to test models and their implementation [@braiekTestingMachineLearning2020].
-Other than checking technical and statistical aspects of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
-With simulated data you can check if the expected data and the planned analysis match.
-First of all, simulated data can be used as an initial check that the code for the planned analysis runs without error. 
-This is particularly important if you want to preregister pieces of your analysis before data were collected.
-For example, we can use simulated data to check whether a given data analysis outputs parameter estimates when we simulate missing data values.
-Second, Monte Carlo simulations allow us to form realistic expectations about the unbiasedness and precision of parameter estimates and it can help with sample size planning to obtain a sensible study design.
-For a variety of standard models, there are analytic strategies to evaluate precision or statistical power, however, Monte Carlo simulation offers more flexibility in terms of mode realistic modeling conditions (e.g., missing data or non-normality).
+Monte Carlo simulations are a crucial tool to investigate properties of a statistical model and they are routinely employed to develop or evaluate statistical models.
 For example, @maccallum1999sample used Monte Carlo simulations to investigate optimal sample size for exploratory factor analysis, or @curran1996robustness studied the robustness of the $\chi^2$ test statistic under non-normal data.
+
+While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
+First, you have to address and could plan for data analytic problems on a conceptual level before you made the considerable effort to acquire data [@axelrodAdvancingArtSimulation1997].
+Second, if you only simulate a single dataset, you can perform a simple technical check, that is the analysis software outputs parameter estimates [@braiekTestingMachineLearning2020], e.g.
+when the simulated data contains missings.
+Third, if you perform a Monte Carlo simulation, you may evaluate how the data analysis copes with expected data complications, e.g.
+non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
+Fourth, you can plan your resources efficiently, e.g. determine required sample size to achieve certain precision or power [@brandmaier2018precision].
+
+A Monte Carlo simulation affords you an ideal planning opportunity.
+We realize that this comes at an expensive but believe the reduction in the risk that the data is not sufficient to answer the research questions offsets the required effort.
+This advantage is especially pronounced when you plan to preregister your study.
+In fact, we argue that a Monte Carlo simulation can combine the analysis, sample size planning and preregistration (see [Preregistration as Code].
+
 No matter which analytic strategy you plan, it will come with assumptions about the data generating process under investigation. 
@@
 If you identify unrealistic assumptions, you can either deliberately violate these assumptions and, by simulation, check if the analytic strategy is robust or modify the modeling approach to explicitly address them.
-We will show an example using structural equation modeling (SEM) for our Machiavellism example.
-Note that the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
+
 Typically, you have to make a number of educated guesses to simulate data that closely matches your expectation.
@@
 Since we typically use parametric statistical models (general linear model or SEM) or hypothesis tests that can be expressed as Gaussian models, we can usually use these models directly to simulate data.
+We will show an example using structural equation modeling (SEM) for our Machiavellism example.
+Note that the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
 
@@
 
-And then fit the statistical model:
-
-```{r}
-model <- "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
-MACH ~~ c(1, NA)*MACH
-MACH ~ c(0, NA)*1 + c(zero, diff)*1"
-fit <-
-  lavaan::cfa(
-    model,
-    data = simulated,
-    group = "group",
-    meanstructure = TRUE,
-    std.lv = TRUE
-  ) 
-```
-
-Now, `lavaan` warns us that this model may not be identified.
-We overlooked a crucial assumption of this model, that is, measurement invariance.
-This assumption has also been incorporated into our simulation of the data.
-There is a good chance that this assumption is violated.
-We therefore adjust our analysis plan to include tests for measurement invariance.
-We propose the following procedure. 
-After fitting the model with strong measurement invariance <!--AP: citation needed-->, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
+We propose the following procedure for analyzing the data and dealing with possible violations of strong measurement invariance. 
+After fitting the model with strong measurement invariance [@flake2020], we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
 The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.

### 2021-05-20 manuscript.Rmd
extend simulation and PAC section

@@
 Thus, it serves as a tool that may help to increase the credibility of empirical results.
-Without getting into too much detail of the meta-scientific arguments for preregistration, we want to note two advantages.
+Without getting into too much detail of the meta-scientific arguments for preregistration, we want to note two advantages [@nosekPreregistrationHardWorthwhile2019].
 First, a preregistration is an opportunity to plan a study carefully.
@@
 PAC essentially reverses the process of empirical research.
-Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results---initially based on simulated data.
+Before researchers gather data, they write a reproducible, dynamically generated manuscript (see [R Markdown]) including the sections Introduction, Methods, and Results---initially based on simulated data.
 Then, the source code of this manuscript serves as an unambiguous description of the analysis plan and is published as a preregistration.
+One particular advantage of this approach is that you can precisely define if-else conditions under which you choose a different analytic strategy to address possible complications (e.g.
+to establish partial measurement invariance, when full measurement invariance is not attainable).
+This approach allows for more flexibility than a traditional template based preregistration but is still well defined.
 
@@
 
-One way to ensure that you can expect your analysis to work with real data is to simulate data.
+One way to ensure that your analysis to works with real data is to simulate data.
 We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
@@
 If you went to the trouble to simulate data, we argue that you get a preregistration almost for free.
+
 A preregistration in code describes sufficiently how you want to analyze the data, but ideally, you also convey the why.

### 2021-05-20 manuscript.Rmd
explain blas

--- 2021-05-19 manuscript.Rmd
@@
 `renv` therefore is already of great use for reproducibility [@vanlissa2020worcs].
-However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS system in use and so forth.
+However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS (Basic Linear Algebra Subprograms, responsible for carrying out many computations like matrix multiplication needed in the linear regression) system in use and so forth.
 Most importantly the R version itself has to match.

### 2021-05-20 manuscript.Rmd
fix typo

@@
 ```{r echo=FALSE}
-```{r}
 # define a print method for objects of the class data.frame

### 2021-05-20 manuscript.Rmd
close #69

@@
 RStudio's _R Projects_ facilitate this concept of a project folder.
-The `repro` package offers an R Studio template, which sets up a file structure that follows best practices and conventions.<!--AP this is from your thesis - can you please add citations as to which practices and conventions these are?-->
+The `repro` package offers an R Studio template, which sets up a file structure that follows best practices and conventions for placing scripts and data [@rextensions].
 To create one, click on:

### 2021-05-20 manuscript.Rmd
cite FAIR principles close #18

@@
 We use containers to guarantee that all code components are available and work as expected.
-Ideally, all data are also stored well documented [@codebook] in the repository [for a more comprehensive argument for publishing data see @vanlissa2020worcs], but this is not always feasible.
+Ideally, all data are also stored well documented [@codebook] in the repository following the FAIR principles [@wilkinsonFAIRGuidingPrinciples2016].
+We believe that sharing data is highly beneficial for the research community [for a more comprehensive argument for publishing data see @vanlissa2020worcs], but this is not always feasible.
 
@@
 ```{r, include=FALSE}
-knitr::write_bib(c(.packages(), "rticles", "gert", "bookdown", "lavaan", "knitr", "targets", "renv", "tidyverse"),
+knitr::write_bib(c(.packages(), "here", "rticles", "gert", "bookdown", "lavaan", "knitr", "targets", "renv", "tidyverse"),
                  "packages.bib")

### 2021-05-20 manuscript.Rmd
move abstract to own file

@@
 bibliography: ["references.bib", "packages.bib"]
-abstract:
-  Computational reproducibility is the ability to obtain identical results from the *same* data with the *same* computer code.
-It is a building block for transparent and cumulative science because it enables other researchers, on other computers and later in time, to reproduce and thus understand how results came about while, at the same time, it avoids a variety of errors that may lead to erroneous reporting of statistical and computational results.
-In this tutorial, we demonstrate how the R package `repro` supports researchers in creating fully computationally reproducible research projects with a focus on current Open Science practices; starting from sketching the research proposal, over preregistering an analysis as executable code with simulated data, until the creation and dissemination of the final scientific report.
+abstract: "`r tryCatch(trimws(readr::read_file('abstract.Rmd')))`"
 ---

### 2021-05-20 abstract.Rmd
each sentance on newline

Computational reproducibility is the ability to obtain identical results from the *same* data with the *same* computer code.
It is a building block for transparent and cumulative science because it enables other researchers, on other computers and later in time, to reproduce and thus understand how results came about while, at the same time, it avoids a variety of errors that may lead to erroneous reporting of statistical and computational results.
In this tutorial, we demonstrate how the R package `repro` supports researchers in creating fully computationally reproducible research projects with a focus on current Open Science practices; starting from sketching the research proposal, over preregistering an analysis as executable code with simulated data, until the creation and dissemination of the final scientific report.

### 2021-05-20 manuscript.Rmd
remove readme figure

@@
 That means you can include R code or plots etc., within the readme.
-The `README.md` file will look pleasant in the browser, just like in Figure \@ref(fig:readme).
-
-```{r readme-screenshot, include=FALSE}
-# unfortunately css selectors won't work for pdf so we'll use png instead
-readme <- here::here("images", "readme.png")
-if (!file.exists(readme)) {
-  webshot2::webshot(
-    "https://github.com/aaronpeikert/repro-tutorial/",
-    file = readme,
-    vwidth = 1920,
-    # height & width are not resolution
-    vheight = 1080,
-    # but what website thinks the browser size is
-    selector = "div#readme",
-    zoom = 2
-  ) # higher resolution
-}
-```
-
-```{r readme, echo=FALSE, fig.cap="A screenshot of how a README.md is rendered/presented on GitHub.  ", out.width='100%'}
-knitr::include_graphics(readme)
-```
+The `README.md` file will look pleasant in the browser on `r link("https://github.com/aaronpeikert/repro-tutorial#readme", "GitHub")`.
 
@@
 This statement should also list all software requirements.
-<!--AP: I think this Figure is superfluous. Please remove. All the information is in the following code blocks.-->
-Figure \@ref(fig:readme) shows `r link("https://github.com/aaronpeikert/repro-tutorial#how-to-reproduce-this-manuscript", "the reproduction statement of this project")`.
 For example, you can download any git repository with these commands:

### 2021-05-20 manuscript.Rmd
cite here package closes #46

@@
 RStudio's _R Projects_ facilitate this concept of a project folder.
+We recomend the [`here`](https://github.com/r-lib/here)[^here] package [@R-here] to refer to files, because it works consistently across different operating systems, and between R scripts and [R Markdown].
 The `repro` package offers an R Studio template, which sets up a file structure that follows best practices and conventions for placing scripts and data [@rextensions].
 To create one, click on:
+
+[^here]: https://github.com/r-lib/here
 
@@
 That means you can include R code or plots etc., within the readme.
-The `README.md` file will look pleasant in the browser on `r link("https://github.com/aaronpeikert/repro-tutorial#readme", "GitHub")`.
+The `README.md` file will look pleasant in the browser on [GitHub](https://github.com/aaronpeikert/repro-tutorial#readme).
 

### 2021-05-20 manuscript.Rmd
added simulation references

@@
 
-Such Monte-Carlo simulation can help you decide between several tradeoffs.
-One historically important tradeoff is between sample size and power (the probability to detect an effect of given size).
-More recently there has been growing interest in other resource relevant parameters as well [!citation], such as number of measurement occasions, reliability of measures etc.
-To show that there can also be alternative approaches for sample size planning beyond null hypothesis testing, we choose to visualize the density distribution of the parameter of interest (latent mean difference by gender of Machiavellism) instead of power.
-Such process still lets you formulate expectations for the precision of the analysis, e.g.
-with a sample size of 1000 per group 80% of the simulated standardized effect size fall into a range of +/- .1, but without necessarily rely on null hypothesis testing.
-For simplicity we do not show the code for the simulation/graphic, but it is available in the accompanying [GitHub repository](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo]:.
+Such Monte-Carlo simulation can help to decide between several trade-offs.
+Historically, one important trade-off is between sample size and power (the probability to detect an effect of given size if it exists).
+However, the question can be put more broadly and it is possible to consider various other trade-offs of relevant study design parameters other than sample size, such as number of measurement occasions [e.g., @brandmaier2015lifespan], reliability of measures [e.g., @brandmaier2018precision], or parameters related to the structure of planned-missing data design [@brandmaier2020optimal].
+To show that there can also be alternative approaches for sample size planning other than those that are restricted to null hypothesis testing, we choose to visualize the precision of estimation as the density distribution of the parameter of interest (latent mean difference by gender of Machiavellism) instead of statistical power.
+Such process still allow us to formulate expectations for the expected precision of the chosen analysis.
+For example, with a sample size of 1.000 per group, 80% of the simulated standardized effect sizes fall into a range of +/- .1 standardized difference points.
+This claim is useful and informative irrespective of whether we embark on null hypothesis testing later on or not.
+The R code to reproduce the simulation and the figure can be found in the accompanying [GitHub repository](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo]:.
 

### 2021-05-20 manuscript.Rmd
Update manuscript.Rmd

--- 2021-05-20 manuscript.Rmd
@@
 Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).
-To guide the reader through all steps of a research project, we will use a exemplary research question:
-
-> "Is there a mean difference in the personality trait 'Machiavellism' between genders?"
-
-This example serves only didactic purposes and not to derive substantive claims.
-Nevertheless, everything else is kept as close to a real research question as possible and we have preregistered the analysis without looking at the data beforehand and use actual data subsequently.
 
@@
 
-Such Monte-Carlo simulation can help you decide between several tradeoffs.
-One historically important tradeoff is between sample size and power (the probability to detect an effect of given size).
-More recently there has been growing interest in other resource relevant parameters as well [!citation], such as number of measurement occasions, reliability of measures etc.
-To show that there can also be alternative approaches for sample size planning beyond null hypothesis testing, we choose to visualize the density distribution of the parameter of interest (latent mean difference by gender of Machiavellism) instead of power.
-Such process still lets you formulate expectations for the precision of the analysis, e.g.
-with a sample size of 1000 per group 80% of the simulated standardized effect size fall into a range of +/- .1, but without necessarily rely on null hypothesis testing.
-For simplicity we do not show the code for the simulation/graphic, but it is available in the accompanying [GitHub repository](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo]:.
+Such Monte-Carlo simulation can help to decide between several trade-offs.
+Historically, one important trade-off is between sample size and power (the probability to detect an effect of given size if it exists).
+However, the question can be put more broadly and it is possible to consider various other trade-offs of relevant study design parameters other than sample size, such as number of measurement occasions [e.g., @brandmaier2015lifespan], reliability of measures [e.g., @brandmaier2018precision], or parameters related to the structure of planned-missing data design [@brandmaier2020optimal].
+To show that there can also be alternative approaches for sample size planning other than those that are restricted to null hypothesis testing, we choose to visualize the precision of estimation as the density distribution of the parameter of interest (latent mean difference by gender of Machiavellism) instead of statistical power.
+Such process still allow us to formulate expectations for the expected precision of the chosen analysis.
+For example, with a sample size of 1,000 per group, 80% of the simulated standardized effect sizes fall into a range of +/- .1 standardized difference points.
+This claim is useful and informative irrespective of whether we embark on null hypothesis testing later on or not.
+The R code to reproduce the simulation and the figure can be found in the accompanying [GitHub repository](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo]:.
 

### 2021-05-20 manuscript.Rmd
First round of comments, back after my meeting

@@
 A cornerstone of scientific progress is the possibility to critically assess the validity of peer scientists’ scientific conclusions.
+<!-- CJ More fundamental: Reproducibility is a cornerstone of the scientific METHOD.
+I think it's more important to emphasize this, rather than critical assessment of others' work.
+See Karl Popper, The Logic of Scientific Discovery, Routledge, London, 1992, p.
+66.
+-->
 Scientific claims should not be credible because of their originators' authority but by the transparency and replicability of their supporting evidence [@aac4716].
+<!-- CJ This is a red herring; nobody said that an argument to authority should be the foundation of scientific evidence.
+I would suggest focusing on replicability as an essential component of the scientific method.
+-->
 Here, *replicability* refers to the degree to which original studies and their derivatives yield consistent results.
@@
 Even worse, researchers might often not even be able to achieve reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code.
+<!-- CJ I would introduce the topic of reproducibility differently.
+"Even worse" is editorializing; not appropriate for a scientific paper.
+Instead, consider saying: More fundamental even than replication across datasets is the reproducibility of findings within a single dataset.
+-->
+When analyses are not properly documented, small mistakes, or circumstances outside the researcher's control (e.g., a corrupted file, a coauthor leaving academia) may render them non-reproducible. 
 Over the past decades, computational statistical methods have radically changed the ability of researchers from all disciplines to process and analyze data with complex statistical models.
+<!-- CJ I think it's more important to focus on the methods that have been developed to ensure reproducibility!)-->
 With these advances came challenges that contribute to broader concerns of irreproducibility in the published literature, among them the lack of exact and unambiguous documentation of computational methods and each computational step taken.
-With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
-Only with access to both data and procedures can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.
-
-In fact, the research community has long accepted reproducibility as an important element of "good scientific practice." <!-- examples missing -->
-However, ensuring reproducibility of computational analyses is still tedious and difficult, and there are still no widely accepted standards that can guide researchers when setting up their data analyses.
-Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.
+With computationally intensive data analyses <!-- CJ The complexity of the analyses is orthogonal to the necessity to make them reproducible.
+-->, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
+Only with access to both data and procedures <!-- CJ and software environment --> can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.
+
+In fact, the research community has long accepted reproducibility as an important element of "good scientific practice." <!-- CJ: This is the crucial point.
+I would start the paper with this.
+See my reference to Popper in the opening sentence.
+-->
+However, ensuring reproducibility of computational analyses is still tedious and difficult <!-- CJ I would not editorialize this.
+Moreover, this sounds discouraging.
+I would focus on what you have done to make reproducibility more attainable.
+-->, and there are still no widely accepted standards that can guide researchers when setting up their data analyses.
+<!-- CJ These are two separate points, each important enough to have their own sentence, with some further clarification.
+-->
+Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories <!-- CJ you have not introduced any open repositories yet.
+--> do not provide sufficient information to reproduce relevant computational and statistical results.
 For example, @hardwicke2018 attempted to replicate open materials in the journal Cognition.

### 2021-05-20 manuscript.Rmd
clarify why future is a good solution

--- 2021-05-20 manuscript.Rmd
@@
 ```{r}
+# reproducible but unsafe in parallel code
 set.seed(1235)
@@
 ```{r}
+# reproducible and safe in parallel code
 future::value(future::future(runif(1), seed = 1235))

### 2021-05-20 manuscript.Rmd
various changes - more explanations :)

--- 2021-05-20 manuscript.Rmd
@@
 
-Before incorporating changes from someone else, they can be reviewed and adepted.
-This review step can also be useful for collaboration, because you can invite a collaborator to take a look at changes you want to do.
-It is therefore sometimes useful to create a pull request even though you have permission to push, because it invites your collaborators to comment or change your proposal.
+Before incorporating changes from someone else, they can be reviewed and adapted.
+This review step is particularly useful for remote collaboration, because collaborators can be explicitly pointed to changes and asked to approve them.
+Therefore, it is sometimes useful to create a pull request even though you have permission to push, because it invites your collaborators to comment on, change, or approve proposed changes.
 
@@
 
-This process is especially useful if you want the feedback of a more senior researcher, because they do not need to know how to use Git but have a convenient web interface to provide comments or changes.
-
-# Analyzing Data
-
-While you already know how to write dynamic documents with [Markdown] and [R Markdown], and track changes with [Git] and [GitHub], two components to achieve long-term reproducibility are still missing, namely handling internal and external dependencies.
+This process can be executed via Github's web interface and can work well with collaborators who do not actually want to bother with installing and running the entire code base of a project.
+
+# The Data Analysis Phase
+
+So far, we have covered the generation of dynamic documents with [Markdown] and [R Markdown], and tracking changes with [Git] and [GitHub].
+These are two of the four components [see @Peikert2019] to achieve long-term reproducibility.
+To complete computational reroducibility, researchers need to properly handle dependency relations between their files (internal dependencies) and all dependencies on external software, most importantly dependend R packages (external dependencies).
+In the following, we will show how the `repro` package supports researchers in mapping all dependencies completely and unambiguously.
 
@@
 
-A reader of your reproducible manuscript should be able to derive how they are to reproduce it (ideally by a statement in your [README]).
-To do that, it is often crucial to understand how different project components relate to each other.
-You cannot run an analysis without data; you cannot render a dynamic document without an analysis, and so forth.
-Dependency tracking means that you capture the interrelations between the components of your analysis.
-The "components" of an analysis are often loosely mapped onto files.
-For a simple project, it might be enough to have a single R Markdown file.
-However, while it is easy to understand how to reproduce the end result, this approach quickly gets unwieldy.
-To better navigate the data analysis, a researcher often splits it across several scripts.
-For example, they create a script to load the data and another to run the analysis.
-This modularisation facilitates reuse because the loading script may be used by other projects building on the same data, while the analysis script may be employed on different data and so forth.
-The downside is that reproducibility is complicated because instead of reproducing everything in one step, multiple steps are required.
-
-Of course, there a several ways to solve this problem.
-One could build an R package to separate analysis and data or use code externalization in the RMarkdown to facilitating reuse or create a "main" script that combines all steps.
-We found Make to be particularly useful for two reasons.
-First, it is language agnostic, therefore fusing steps that are carried out in different programming languages.
+A reader of a reproducible research project should be able to easily derive how it can be reproduced (ideally by a statement in your [README]).
+To do that, it is often crucial to understand how different project components relate to each other and depend on each other.
+For example, you cannot run an analysis without data, or you cannot render a statistical report without having the results from the statistical analysis that is reported.
+Dependency tracking means that you capture and map the interrelations between all components of your analysis.
+The "components" of an analysis are typically all involved files including raw data files, intermediate files (such as a preprocessed data file), and final results (such as a rendered manuscript in PDF format).
+With larger projects, there is typically a higher degree of modularization because code fragments are written for re-use in other projects (or are already re-used).
+Also, more complicated empirical research projects may have different hypotheses that are answered with different data sets and different statistical models, such that it often makes sense to split the different analyses into different files.
+Reproducibility then demands the careful documentation and mapping of all necessary steps in the correct order.
+
+Make is a build automation tool for software projects that typically are challenged by maintaing and resolving a variety of dependencies in the correct order when compiling a piece of software-
+We found _Make_ to be particularly useful for our purposes for two reasons.
+First, it is language-agnostic, therefore enabling us to map steps not only taken within R but also across different programming languages or other external programs.
 Second, it does implement a simple yet powerful caching mechanism of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning. 
+That is, Make determines automatically which branches in the dependency graph needs updating when files have changed and does not necessarily recompute or rebuild everything.
 
@@
 
-When you cook something, and you have the necessary ingredients already at hand, would you throw it away just to prepare it again?
-Unlikely.
-Make behaves the same.
-If a file already exists, it does not recreate it unnecessarily.
-However, if a requirement is newer than its target, make recreates the target to reflect the updated dependency.
-Therefore, when you update the data file, the analysis is rerun.
-Because Make knows the whole dependency tree, its caching is smarter than many alternatives.
-
-One alternative that is even smarter than Make is the R package [`targets` [@targets2021]](https://books.ropensci.org/targets/)[^targets].
-`targets` can infer more about your project than Make, because it analysis the R code you write to build the dependency tree.
+If a target file already exists, Make only recreates it if necessary.
+If a dependency is newer than its target, make recreates the target to reflect the updated dependency.
+In our case, when we (re)build a manuscript and some dependent data file has changed, the data analysis is automatically rerun.
+This avoids mistakes in the reporting that are due to reports and analyses being out-of-sync.
+
+<!--
+AP: I suggest to remove this simply because it distracts from our primary tool and we need to shorten
+One alternative that is even smarter (in terms of automatically inferring dependencies) than Make is the R package [`targets` [@targets2021]](https://books.ropensci.org/targets/)[^targets].
+`targets` can infer more about your project than Make, because it analyzes the R code you write to build the dependency tree.
 For this inference to work well, you have to write everything in R, and you have to write everything strictly functional.
@@
 [^targets]: https://books.ropensci.org/targets/
+-->
 
@@
 
-The `repro` package can automatically write Makefiles for simple scenarios.
-For this purpose, `repro` utilizes the YAML metadata:
+Writing Makefiles manually is cumbersome for larger projects. 
+Therefore, the `repro` package provides means to automatically write Makefiles for a given R project.
+To this end, `repro` extends the YAML metadata header of R Markdown files.
+`repro` introduces a new first-level tag `repro:` with second-level tags (`scripts:`, `packages:`, and `data:`) that take lists of dependent files.
+For example, if an R Markdown document depends on a further file with R code, the header would look like this:
 
@@
 
-This command will go through all R Markdown files, and their YAML metadata, in your repository and infer a `Makefile` from them:
+To automatically generate the Makefile, `repro` provides a command that goes through all R Markdown files in the project, parses their YAML metadata, and infers a `Makefile` from them:
 
@@
 
-The generated `Makefile` automatically contains targets for all R Markdown documents.
-To, for example, generate the `README.md` from the `README.Rmd` you can run:
+The generated `Makefile` automatically contains separate targets for each R Markdown documents.
+The names of the Makefile targets are identical to the Markdown file names.
+For example, to generate the Markdown file `README.md` from the R Markdown `README.Rmd`, you can run:
 
@@
 
-It is common practice to have a target all, that encompasses all important files of the project as the first target.
-If you do not specify a target, Make defaults to the first target, which is usually `all`.
-`repro` reminds you of that and suggests which file you might want to add to the target `all`.
-You can generally change the `Makefile` anyway you want, e.g. to add targets which are not supported by `repro`.
-After you have done so, you can then run in the terminal:
+It is common practice to name the very first target in a Makefile "all" and have it include all important files of the project.
+If you do not specify a target when calling Make, Make defaults to this first target.
+`repro` reminds you of that common practice and suggests which file you might want to add to the target `all`.
+Thus, if researchers who try to reproduce a given project and are unaware of its potentially various contents, will always build all reproducible artifacts of a repository when calling Make without further arguments:
 
@@
 
-And it will fully reproduce the project, if you have all required software installed.
+Running this command fully reproduces the entire project -- but only if you have all required software installed.
+Next, we will describe how `repro` helpts to manage all software dependencies automatically.
 
@@
 Modern data analysis is only feasible because researchers rely on much external software. 
-One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
 The dependency on a single R package may easily create a dependency tree of more than hundred direct or indirect dependencies [@epskamp2019].
+External dependencies are unavoidable but, at the same time, a major threat to long-term reproducibility.
 The critical problem is the fragility of a software setup.
-Many components have to fit perfectly together, and each change, e.g. an update, risks that balance .
+Many components have to fit perfectly together, and each change, e.g. a software update or bug fixe, risks that balance.
 At the same time, software updates are very much necessary to remove errors or improve methodology [@epskamp2019].
@@
 
-Docker can be a bit unfriendly, because it requires a complicated mental model and some knowledge about how to install software on Linux.
+Setting up and running Docker is not always painless, because it requires a complicated mental model of the underlying virtual computer and, specifically, knowledge about Linux environments.
 An R-focused alternative is the R package [`renv`](https://rstudio.github.io/renv/articles/renv.html)[^renv][@R-renv].

### 2021-05-20 manuscript.Rmd
minor edits

@@
 From the `Dockerfile` you can then build an image.
-While you can do so manually we prefer to use Make, see [Internal Dependencies].
-The `Makefile` build by `repro` does contain a recipe for creating the `Dockerfile`.
+While you can do so manually, we suggest to again use Make ( see [Internal Dependencies]).
+The `Makefile` built by `repro` does contain a recipe for creating the `Dockerfile`.
 
@@
 
-The `Makefile` from `repro` does also include a mechanism to send commands conditionally through the Docker container.
-This mechanism is triggered by `DOCKER=TRUE` in the make call:
+Importantly, the `Makefile` created by `repro` includes a mechanism to execute commands in the Docker container instead of locally.
+This mechanism is triggered by the argument `DOCKER=TRUE` appended to any make call. 
+The following call builds all targets inside the container instead of in the local computing environment:
 
@@
 ```
-The function `automate()` combines `automate_docker()` & `automate_make()`.
-
-You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manual` to install software beyond R packages.
-Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manual`.
+
+Last, to create or update the container and all Make targets, `repro` provides a single function `automate()` that calls `automate_docker()` and `automate_make()`.
+
+You can also manually edit the `Dockerfile` by changing the file `.repro/Dockerfile_manual` to install software other than R packages.
+Remember to run `repro::automate()` every time you made edits to `.repro/Dockerfile_manual`.
 We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.

### 2021-05-20 manuscript.Rmd
more on docker

@@
 
-The open nature of the here proposed workflow does blur the lines between published and unpublished.
-However, even when your Git repository is public, you may want to signify important stages of your project, like preregistration, preprint, submission and acceptance to the journal.
-These milestones deserve to be archived.
-A project is only then properly archived when it is stored at least at two independent storage locations and reproducible from all locations.
-Therefore, you should make extra sure that the project at these points is reproducible.
-To ensure reproducible, you could ask a coworker or collaborator to reproduce the project independently from you.
+In a public repository, you may want to mark important stages of your project, like preregistration, preprint, submission and acceptance at a scientific journal.
+We suggest to archive these milestones individually.
+A reproducible project is only then properly archived when it is stored at least at two independent storage locations and reproducible from all locations. <!--AP citation? -->
+To really make sure that a given project is reproducible, one should aim for an explicit test of reproduction before the project is published. 
+We recommend to check out and reproduce the project in a new directory in a "fresh" environment (in particular, a new R session should be used for this). 
+For maximal validitiy of this test, we suggest to ask a coworker or collaborator to reproduce the project independently from you.
 The sections [Creating a DOI] and [Creating a Release] describe how you can easily create three different openly accessible longterm archived copies of your project.
 
-When you disseminate the paper across more locations than just GitHub it can be ambiguous which version of code and data have created the document (e.g. the manuscript.pdf).
-To create a clear link between the repository version and some end result you can automatically include the hash of the commit.
-Either you use Git from within R:
+When you disseminate multiple versions of a document (e.g., at different stages of a review process), it seems good practice to specify the exact version of code and data used to create a given document (e.g.
+the final type-set version of a journal article).
+To create a clear link between the repository version and a given end result you can automatically include the unique hash identifying the project state in the git version history.
+<!--Either you use Git from within R:
 
@@
 ```
-
-Or you can use the `gert` package:
+-->
+To this end, we use the `gert` package:
 
@@
 
-The approach we advocate for distributes the research project, e.g. across GitHub, preregistration platforms, the publishers site, and more.
+Typical open research project will have their contents available from multiple sites, e.g. from GitHub, preregistration platforms, preprint servers, or publisher websites. 
+Under these circumstances it is important to provide unique links 
 Digital Object Identifiers (DOI) tie all these locations together under one link [@derisiWhatWhysDOIs2003].
@@
 This is useful to link or cite a specific part of a research project, such as the preregistration, an illustration in the paper or the data that was gathered.
-The DOI can also contain information about the license of the object it refers to.
+The DOI can also contain information about the license of its target.
 <!-- public releases like preregistration, preprint -->
-zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides doi for these archives.
+zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides DOIs for these archives.
 While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
@@
 
-After you have linked a GitHub repository you trigger the archivation by creating a GitHub release, as described in the next section.
+After you have linked a GitHub repository, you trigger the archival by creating a GitHub release, as described in the next section.
 
@@
 
-A valid point of criticism is that this workflow introduces a bulk of software solutions and services that may not be available for long.
-Superficially it relies on R, RStudio, Pandoc, Git, Make, Docker, Homebrew (OS X only) and Chocolaty (Windows only) as software and on GitHub, DockerHub, MRAN and Ubuntu Package Archive as services.
-These many breakpoints can be reduced to one service and one software.
+A central point of criticism is that the proposed workflow introduces a bulk of software solutions and services that may not be available in the long run.
+First and foremost, it relies on R, RStudio, Pandoc, Git, Make, Docker, Homebrew (OS X only) and Chocolaty (Windows only) as software and on GitHub, DockerHub, MRAN and Ubuntu Package Archive as services.
+These many breakpoints can be reduced to a single service and a single piece of software.
 That is because all required software is bundled into the container, whose image can be stored.
 Hence, one needs a storage provider which saves the container image and the other files associated with the project.
-We already recommend to store all project files redundantly (your computer, on GitHub and zenodo.org).
+We already recommend to store all project files redundantly (your computer, a remote git server such as GitHub, and a publicly funded, open-science-compatible service provider such as zenodo.org).
 Building the Docker image from the Dockerfile still requires DockerHub, MRAN, and Ubuntu Package Archive to be online.
-We therefore suggest that you include an already build image into the releases.
-The Makefile generated by `repro` does include a target to facilitate this.
+We therefore suggest that you include an already built image into the releases.
+The Makefile generated by `repro` does include a target to do so.
 To write a compressed image to disk use:
@@
 
-Unfortunately, this image can only be used when Docker is installed.
-To remove any software dependency including Docker from the archived project, one can convert the Docker image to a full system image.
-Such system image can subsequently be booted on any personal computer without relying on any software or service and therefore grantees reproducibility for decades to come.
-Currently, `repro` does not facilitate this conversion, but we hope to include this feature soon.
+This image allows anyone with a computer equipped with Docker to reproduce the analysis (if the repository contains all code and data and has no other external dependencies).
+Last, one can convert the Docker image to a full bootable system image.
+Such a system image can subsequently be booted on any personal computer without relying on any software or service and therefore grantees reproducibility for decades to come.
+Currently, `repro` does not yet facilitate this conversion, but we hope to include this feature soon.
 

### 2021-05-20 manuscript.Rmd
changes to random numbers trying to be shorter

@@
 @lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
-While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers:
-
+While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers.
+
+<!--
 ```{r}
@@
 ```
+-->
+By wrapping code that uses the internal random number generator (here, the command to create a random number from the uniform distribution, `runif(1)`) with the `future` approach, we guarantee that the random numbers will differ across parallel executions of that code, will achieve a high quality of randomness, but still will be deterministic and thus reproducible:
 
 ```{r}
-future::value(future::future(runif(1), seed = 1235))
 future::value(future::future(runif(1), seed = 1235))

### 2021-05-20 manuscript.Rmd
added Bayesian prior predictive checks reference

@@
 Reproducibility is invaluable for effective collaboration---including with your "past self".
-While you might be tempted to consider reproducibility only when you about to finish a project, upfront investments pay off through your increased productivity.
+While you may be tempted to consider reproducibility only when you are close to completing a project, we believe that upfront investment pays off in your increased productivity.
 Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you achieve reproducibility gradually.
@@
 <!--Keep notes.-->
- A great tool to capture ideas or meeting notes is Markdown.
+ A great tool to capture ideas or meeting notes and eventually complete entire journal articles is Markdown.
  Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
@@
 For example, @maccallum1999sample used Monte Carlo simulations to investigate optimal sample size for exploratory factor analysis, or @curran1996robustness studied the robustness of the $\chi^2$ test statistic under non-normal data.
+In Bayesian data analysis, prior predictive checks generate data from a model according to the prior distribution in order to asses whether a prior is appropriate [@gabry2019visualization].
 

### 2021-05-20 manuscript.Rmd
Second round of comments; more after dinner

--- 2021-05-20 manuscript.Rmd
@@
 Only with access to both data and procedures can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.
+<!-- CJ this final claim is questionable, at least before introducing the concept of containerization.
+Just using data and procedures, you cannot guarantee reproducibility on any computer.-->
 
@@
 Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.
-For example, @hardwicke2018 attempted to replicate open materials in the journal Cognition.
+For example, @hardwicke2018 attempted to replicate studies ostensibly adhering to the standards of open science in the journal Cognition.<!-- CJ How do you replicate open materials?
+At least in the social sciences "open materials" is taken to mean that you have used a permissive license on a newly developed questionnaire or other research material.
+-->
 Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
@@
 The authors could execute only 31 of the scripts without error and were able to reproduce the results of only 21 articles (within a reasonable period of time).
-Thus, there is a dire need to establish standards for reproducible workflows for scientific rigor.
-
-From our personal experience, the lack of a proper incentive structure and the dreary perspective of having to spend several extra hours if not days to properly set of a code repository to become reproducible, makes many scholars skip this extra step.
-If reproduction would require merely minutes instead of hours, it could greatly facilitate collaboration within and across research projects.
+Thus, there is a dire need to establish standards for reproducible workflows for scientific rigor.<!-- CJ It does not automatically follow that the problem you describe necessitates reproducible workflows.
+I would procede as follows: 1) use this example to conclude the existence of a problem (even "open" studies do not reproduce), 2) present workflows as a solution.
+-->
+
+From our personal experience,<!--CJ This undermines the credibility of the argument.
+This is not an opinion piece--> the lack of a proper incentive structure and the dreary perspective of having to spend several extra hours if not days to properly set of a code repository to become reproducible, makes many scholars skip this extra step.
+<!-- CJ I am not sure this is the primary reason.
+Indeed, incentives are important - but in that context, perhaps reference DORA?
+I think you were just on a good track, arguing that even with the best intentions, researchers often fail to make their work reproducible.
+Your workflow will ensure that their next attempt is succesful.-->
+If reproduction would require merely minutes instead of hours,<!-- CJ Now we're back to ease-of-use, not incentives.-->
+it could greatly facilitate collaboration within and across research projects.
+<!-- CJ This is a different, third argument: In order to collaborate effectively, researchers should be looking at exactly the same data, code and results.
+I suggest to organize these arguments, and address them each exhaustively in turn.
+-->
 <!-- Guiding principle: spending machine compute time instead of human research time -->
-To achieve this boost in productivity, we must place the burden of reproducibility upon computers instead of the researchers.
-To this end, we demand that a scientific product (no matter whether this is a preliminary report or a full research article) should both be understandable by researchers and automatically reproducible by computers with minimal constraints to the researchers' own workflows.
+To achieve this boost in productivity, we must place the burden of reproducibility upon computers instead of the researchers.<!-- CJ I understand what you mean, but it's not yet communicated effectively because you have not yet made explicit that reproducible projects will indeed facilitate collaboration, and increase opportunities for collaboration (e.g., with strangers through github), and will therefore lead to a boost in productivity.
+Also - ask yourself if this boost in productivity is convincing to individual researchers, your target audience.
+It seems like this would be more convincing to management and policy makers.-->
+To this end, we demand <!-- CJ It's you creating the product, so.. you demand this of yourself? Or do you mean that you demand this of your readership? Either way, strange choice of words.-->
+that a scientific product (no matter whether this is a preliminary report or a full research article) should both be understandable by researchers and automatically reproducible by computers with minimal constraints to the researchers' own workflows.
+<!-- CJ This is ambiguous.
+The point of "understandable by researchers and automatically reproducible by computers" requires clarification.
+"Minimal constraints to the workflow" is also not clear.
+Aren't you providing readers with a workflow?
+-->
 
@@
 <!--stolen from: https://brandmaier.github.io/reproducible-data-analysis-materials/KULeuvenQuantPsy2020.html#11 -->
-We identified the following common four causes for non-reproducibility:
+We identified the following common four causes for non-reproducibility: <!-- CJ How did you identify these causes? Is this exhaustive?-->
 
@@
 4.	Broken dependencies; for example, a given analysis depends on a software package that might not be available on a different computer or no longer exists.
-
-We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:
+<!-- CJ 5.
+Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
+We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph.
+The opening paragraph should at least:
+1) Introduce the topic area
+2) Introduce the problem/knowledge gap
+3) Introduce how you will fill this knowledge gap-->
 
@@
 This tutorial follows the recommendations of @Peikert2019 and uses _Git_ for version control, _R Markdown_ for dynamic document creation, _Make_ for dependency tracking, and _Docker_ for software management.
+<!-- CJ None of these terms have been defined yet-->
 The interplay of these four components is shown in Figure \@ref(fig:schematic).
-However, which tools you use to implement these concepts is a matter of taste and project requirements, and other choices are possible depending on the context and the degree of reproducibility one aims for [e.g., @vanlissa2020worcs; @turingway2021turing].
+However, which tools you use to implement these concepts is a matter of taste and project requirements, and other choices are possible depending on the context and the degree of reproducibility one aims for [e.g., @vanlissa2020worcs; @turingway2021turing].<!-- CJ But readers do not yet know what these concepts are, nor what tools exist.
+At this point in the writing, I think you need to take them by the hand and explain.
+After doing that, you can let them know that there are alternatives.-->
 At several points, we suggest other viable alternatives.
@@
 This paper was designed to help researchers to practically apply the TOP guidelines [except for number 8.
-which does not apply to single research projects, @nosekPromotingOpenResearch2015; @vanlissa2020worcs].
-We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.
+which does not apply to single research projects, @nosekPromotingOpenResearch2015; @vanlissa2020worcs].<!-- CJ Is this really the case?
+Seems like there is little attention to the TOP guidelines, and the present paper is much more widely applicable.
+Maybe make the general point that workflows can help address formal guidelines for transparency, like TOP and FAIR - and consider citing WORCS for more information.
+Then explain that your workflow is flexible enough to go beyond-->
+We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.<!-- CJ Is this the description of the workflow?-->
 We continue with the creation of a concrete data analysis plan, which ideally is already implemented in code, and a dynamic report document that can be preregistered.
-Once the empirical data has been collected, the results can be calculated and appear directly in the dynamically generated document, where ideally only the discussion section needs to be added before the whole archive can be permanently stored as a fully reproducible digital artifact using a unique DOI in a persistent location.
+Once the empirical data has been collected, the results can be calculated and appear directly in the dynamically generated document, where ideally only the discussion section needs to be added before the whole archive can be permanently stored as a fully reproducible digital artifact in a persistent location, and be assigned a DOI to increase findability.<!--CJ All DOIs are unique, and it is not stored using a DOI-->
 
@@
 
-Here, we present a workflow that has proven useful for our way of working with a strong focus on Open Science.
+Here, we present a workflow that has proven useful for our way of working with a strong focus on Open Science.<!-- CJ What do you mean here? -->
 However, not all steps are necessary to achieve reproducibility.
@@
 
-This tutorial is targeted at scholars with at least intermediate R knowledge and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md", "installation guide")`).
-To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:
+This tutorial is targeted at scholars with at least intermediate R knowledge<!--CJ This sounds like something that belongs in the beginning of the introduction; it's not part of setup--> and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md", "installation guide")`).
+To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]: <!-- CJ This also seems like something that should be mentioned in the intro; once people are setting up their systems, they should havebeen told what they're setting them up for-->
 
@@
 if(!requireNamespace("remotes"))install.packages("remotes")
-remotes::install_github("aaronpeikert/repro")
+remotes::install_github("aaronpeikert/repro") # CJ Suggest publishing (at least aminimal working version) of the repro package to CRAN before publishing this paper, to reduce one layer of complexity for users
 library("repro")
@@
 The [`repro`-package](https://github.com/aaronpeikert/repro) contains several helpers to work with the suggested tools for reproducibility.
-For example, the package can check if your computer has all the required software packages installed and ready.
+For example, the package can check if your computer has all the required software packages installed and ready. <!-- CJ Does repro assume that these have already been installed  by hand? -->
 
@@
 
-If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
+If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.<!-- CJ I think you should explain your philosophy that the package provides additional tutorial instructions depending on the user's situation before diving into the actual tutorial.
+I foresee that people are going to be puzzled when the tutorial steps present them with additional mini tutorials.
+Explain in the introduction what the problem is (users need to install andfamiliarize themselves with many different software solutions) and how you solved it (you wrote a package that tells users what dependencies they need to install and how -->
 Please make sure that all three commands pass before you continue with the tutorial.
@@
 
-# Planning Phase
+# Planning Phase <!-- CJ This reminds me of the WORCS phases, which is fine, but it's the only header that does remind me of these phases.
+You've not introduced a breakdown of the research process into phases anywhere.
+Moreover, the text below does not really relate to "planning"; it goes back to reasons for reproducibility, and then an explanation of technical tools (Rmarkdown and Git), neither of which relate to planning!
+-->
 
@@
 
-Before incorporating changes from someone else, they can be reviewed and adapted.
-This review step is particularly useful for remote collaboration, because collaborators can be explicitly pointed to changes and asked to approve them.
-Therefore, it is sometimes useful to create a pull request even though you have permission to push, because it invites your collaborators to comment on, change, or approve proposed changes.
+Before incorporating changes from someone else, they can be reviewed and adepted.
+This review step can also be useful for collaboration, because you can invite a collaborator to take a look at changes you want to do.
+It is therefore sometimes useful to create a pull request even though you have permission to push, because it invites your collaborators to comment or change your proposal.
 
@@
 
-This process can be executed via Github's web interface and can work well with collaborators who do not actually want to bother with installing and running the entire code base of a project.
-
-# The Data Analysis Phase
-
-So far, we have covered the generation of dynamic documents with [Markdown] and [R Markdown], and tracking changes with [Git] and [GitHub].
-These are two of the four components [see @Peikert2019] to achieve long-term reproducibility.
-To complete computational reroducibility, researchers need to properly handle dependency relations between their files (internal dependencies) and all dependencies on external software, most importantly dependend R packages (external dependencies).
-In the following, we will show how the `repro` package supports researchers in mapping all dependencies completely and unambiguously.
+This process is especially useful if you want the feedback of a more senior researcher, because they do not need to know how to use Git but have a convenient web interface to provide comments or changes.
+
+# Analyzing Data
+
+While you already know how to write dynamic documents with [Markdown] and [R Markdown], and track changes with [Git] and [GitHub], two components to achieve long-term reproducibility are still missing, namely handling internal and external dependencies.
 
@@
 
-A reader of a reproducible research project should be able to easily derive how it can be reproduced (ideally by a statement in your [README]).
-To do that, it is often crucial to understand how different project components relate to each other and depend on each other.
-For example, you cannot run an analysis without data, or you cannot render a statistical report without having the results from the statistical analysis that is reported.
-Dependency tracking means that you capture and map the interrelations between all components of your analysis.
-The "components" of an analysis are typically all involved files including raw data files, intermediate files (such as a preprocessed data file), and final results (such as a rendered manuscript in PDF format).
-With larger projects, there is typically a higher degree of modularization because code fragments are written for re-use in other projects (or are already re-used).
-Also, more complicated empirical research projects may have different hypotheses that are answered with different data sets and different statistical models, such that it often makes sense to split the different analyses into different files.
-Reproducibility then demands the careful documentation and mapping of all necessary steps in the correct order.
-
-Make is a build automation tool for software projects that typically are challenged by maintaing and resolving a variety of dependencies in the correct order when compiling a piece of software-
-We found _Make_ to be particularly useful for our purposes for two reasons.
-First, it is language-agnostic, therefore enabling us to map steps not only taken within R but also across different programming languages or other external programs.
+A reader of your reproducible manuscript should be able to derive how they are to reproduce it (ideally by a statement in your [README]).
+To do that, it is often crucial to understand how different project components relate to each other.
+You cannot run an analysis without data; you cannot render a dynamic document without an analysis, and so forth.
+Dependency tracking means that you capture the interrelations between the components of your analysis.
+The "components" of an analysis are often loosely mapped onto files.
+For a simple project, it might be enough to have a single R Markdown file.
+However, while it is easy to understand how to reproduce the end result, this approach quickly gets unwieldy.
+To better navigate the data analysis, a researcher often splits it across several scripts.
+For example, they create a script to load the data and another to run the analysis.
+This modularisation facilitates reuse because the loading script may be used by other projects building on the same data, while the analysis script may be employed on different data and so forth.
+The downside is that reproducibility is complicated because instead of reproducing everything in one step, multiple steps are required.
+
+Of course, there a several ways to solve this problem.
+One could build an R package to separate analysis and data or use code externalization in the RMarkdown to facilitating reuse or create a "main" script that combines all steps.
+We found Make to be particularly useful for two reasons.
+First, it is language agnostic, therefore fusing steps that are carried out in different programming languages.
 Second, it does implement a simple yet powerful caching mechanism of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning.
-That is, Make determines automatically which branches in the dependency graph needs updating when files have changed and does not necessarily recompute or rebuild everything.
 
@@
 
-If a target file already exists, Make only recreates it if necessary.
-If a dependency is newer than its target, make recreates the target to reflect the updated dependency.
-In our case, when we (re)build a manuscript and some dependent data file has changed, the data analysis is automatically rerun.
-This avoids mistakes in the reporting that are due to reports and analyses being out-of-sync.
-
-<!--
-AP: I suggest to remove this simply because it distracts from our primary tool and we need to shorten
-One alternative that is even smarter (in terms of automatically inferring dependencies) than Make is the R package [`targets` [@targets2021]](https://books.ropensci.org/targets/)[^targets].
-`targets` can infer more about your project than Make, because it analyzes the R code you write to build the dependency tree.
+When you cook something, and you have the necessary ingredients already at hand, would you throw it away just to prepare it again?
+Unlikely.
+Make behaves the same.
+If a file already exists, it does not recreate it unnecessarily.
+However, if a requirement is newer than its target, make recreates the target to reflect the updated dependency.
+Therefore, when you update the data file, the analysis is rerun.
+Because Make knows the whole dependency tree, its caching is smarter than many alternatives.
+
+One alternative that is even smarter than Make is the R package [`targets` [@targets2021]](https://books.ropensci.org/targets/)[^targets].
+`targets` can infer more about your project than Make, because it analysis the R code you write to build the dependency tree.
 For this inference to work well, you have to write everything in R, and you have to write everything strictly functional.
@@
 [^targets]: https://books.ropensci.org/targets/
--->
 
@@
 
-Writing Makefiles manually is cumbersome for larger projects. 
-Therefore, the `repro` package provides means to automatically write Makefiles for a given R project.
-To this end, `repro` extends the YAML metadata header of R Markdown files.
-`repro` introduces a new first-level tag `repro:` with second-level tags (`scripts:`, `packages:`, and `data:`) that take lists of dependent files.
-For example, if an R Markdown document depends on a further file with R code, the header would look like this:
+The `repro` package can automatically write Makefiles for simple scenarios.
+For this purpose, `repro` utilizes the YAML metadata:
 
@@
 
-To automatically generate the Makefile, `repro` provides a command that goes through all R Markdown files in the project, parses their YAML metadata, and infers a `Makefile` from them:
+This command will go through all R Markdown files, and their YAML metadata, in your repository and infer a `Makefile` from them:
 
@@
 
-The generated `Makefile` automatically contains separate targets for each R Markdown documents.
-The names of the Makefile targets are identical to the Markdown file names.
-For example, to generate the Markdown file `README.md` from the R Markdown `README.Rmd`, you can run:
+The generated `Makefile` automatically contains targets for all R Markdown documents.
+To, for example, generate the `README.md` from the `README.Rmd` you can run:
 
@@
 
-It is common practice to name the very first target in a Makefile "all" and have it include all important files of the project.
-If you do not specify a target when calling Make, Make defaults to this first target.
-`repro` reminds you of that common practice and suggests which file you might want to add to the target `all`.
-Thus, if researchers who try to reproduce a given project and are unaware of its potentially various contents, will always build all reproducible artifacts of a repository when calling Make without further arguments:
+It is common practice to have a target all, that encompasses all important files of the project as the first target.
+If you do not specify a target, Make defaults to the first target, which is usually `all`.
+`repro` reminds you of that and suggests which file you might want to add to the target `all`.
+You can generally change the `Makefile` anyway you want, e.g. to add targets which are not supported by `repro`.
+After you have done so, you can then run in the terminal:
 
@@
 
-Running this command fully reproduces the entire project -- but only if you have all required software installed.
-Next, we will describe how `repro` helpts to manage all software dependencies automatically.
+And it will fully reproduce the project, if you have all required software installed.
 
@@
 Modern data analysis is only feasible because researchers rely on much external software.
+One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
 The dependency on a single R package may easily create a dependency tree of more than hundred direct or indirect dependencies [@epskamp2019].
-External dependencies are unavoidable but, at the same time, a major threat to long-term reproducibility.
 The critical problem is the fragility of a software setup.
-Many components have to fit perfectly together, and each change, e.g. a software update or bug fixe, risks that balance.
+Many components have to fit perfectly together, and each change, e.g. an update, risks that balance .
 At the same time, software updates are very much necessary to remove errors or improve methodology [@epskamp2019].
@@
 From the `Dockerfile` you can then build an image.
-While you can do so manually, we suggest to again use Make ( see [Internal Dependencies]).
-The `Makefile` built by `repro` does contain a recipe for creating the `Dockerfile`.
+While you can do so manually we prefer to use Make, see [Internal Dependencies].
+The `Makefile` build by `repro` does contain a recipe for creating the `Dockerfile`.
 
@@
 
-Importantly, the `Makefile` created by `repro` includes a mechanism to execute commands in the Docker container instead of locally.
-This mechanism is triggered by the argument `DOCKER=TRUE` appended to any make call. 
-The following call builds all targets inside the container instead of in the local computing environment:
+The `Makefile` from `repro` does also include a mechanism to send commands conditionally through the Docker container.
+This mechanism is triggered by `DOCKER=TRUE` in the make call:
 
@@
 ```
-
-Last, to create or update the container and all Make targets, `repro` provides a single function `automate()` that calls `automate_docker()` and `automate_make()`.
-
-You can also manually edit the `Dockerfile` by changing the file `.repro/Dockerfile_manual` to install software other than R packages.
-Remember to run `repro::automate()` every time you made edits to `.repro/Dockerfile_manual`.
+The function `automate()` combines `automate_docker()` & `automate_make()`.
+
+You can also manually edit the `Dockerfile` by amending the file `.repro/Dockerfile_manual` to install software beyond R packages.
+Remember to run `repro::automate()` after you made edits to `.repro/Dockerfile_manual`.
 We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
@@
 
-Setting up and running Docker is not always painless, because it requires a complicated mental model of the underlying virtual computer and, specifically, knowledge about Linux environments.
+Docker can be a bit unfriendly, because it requires a complicated mental model and some knowledge about how to install software on Linux.
 An R-focused alternative is the R package [`renv`](https://rstudio.github.io/renv/articles/renv.html)[^renv][@R-renv].
@@
 
-In a public repository, you may want to mark important stages of your project, like preregistration, preprint, submission and acceptance at a scientific journal.
-We suggest to archive these milestones individually.
-A reproducible project is only then properly archived when it is stored at least at two independent storage locations and reproducible from all locations. <!--AP citation? -->
-To really make sure that a given project is reproducible, one should aim for an explicit test of reproduction before the project is published. 
-We recommend to check out and reproduce the project in a new directory in a "fresh" environment (in particular, a new R session should be used for this). 
-For maximal validitiy of this test, we suggest to ask a coworker or collaborator to reproduce the project independently from you.
+The open nature of the here proposed workflow does blur the lines between published and unpublished.
+However, even when your Git repository is public, you may want to signify important stages of your project, like preregistration, preprint, submission and acceptance to the journal.
+These milestones deserve to be archived.
+A project is only then properly archived when it is stored at least at two independent storage locations and reproducible from all locations.
+Therefore, you should make extra sure that the project at these points is reproducible.
+To ensure reproducible, you could ask a coworker or collaborator to reproduce the project independently from you.
 The sections [Creating a DOI] and [Creating a Release] describe how you can easily create three different openly accessible longterm archived copies of your project.
 
-When you disseminate multiple versions of a document (e.g., at different stages of a review process), it seems good practice to specify the exact version of code and data used to create a given document (e.g.
-the final type-set version of a journal article).
-To create a clear link between the repository version and a given end result you can automatically include the unique hash identifying the project state in the git version history.
-<!--Either you use Git from within R:
+When you disseminate the paper across more locations than just GitHub it can be ambiguous which version of code and data have created the document (e.g. the manuscript.pdf).
+To create a clear link between the repository version and some end result you can automatically include the hash of the commit.
+Either you use Git from within R:
 
@@
 ```
--->
-To this end, we use the `gert` package:
+
+Or you can use the `gert` package:
 
@@
 
-Typical open research project will have their contents available from multiple sites, e.g. from GitHub, preregistration platforms, preprint servers, or publisher websites. 
-Under these circumstances it is important to provide unique links 
+The approach we advocate for distributes the research project, e.g. across GitHub, preregistration platforms, the publishers site, and more.
 Digital Object Identifiers (DOI) tie all these locations together under one link [@derisiWhatWhysDOIs2003].
@@
 This is useful to link or cite a specific part of a research project, such as the preregistration, an illustration in the paper or the data that was gathered.
-The DOI can also contain information about the license of its target.
+The DOI can also contain information about the license of the object it refers to.
 <!-- public releases like preregistration, preprint -->
-zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides DOIs for these archives.
+zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides doi for these archives.
 While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
@@
 
-After you have linked a GitHub repository, you trigger the archival by creating a GitHub release, as described in the next section.
+After you have linked a GitHub repository you trigger the archivation by creating a GitHub release, as described in the next section.
 
@@
 
-A central point of criticism is that the proposed workflow introduces a bulk of software solutions and services that may not be available in the long run.
-First and foremost, it relies on R, RStudio, Pandoc, Git, Make, Docker, Homebrew (OS X only) and Chocolaty (Windows only) as software and on GitHub, DockerHub, MRAN and Ubuntu Package Archive as services.
-These many breakpoints can be reduced to a single service and a single piece of software.
+A valid point of criticism is that this workflow introduces a bulk of software solutions and services that may not be available for long.
+Superficially it relies on R, RStudio, Pandoc, Git, Make, Docker, Homebrew (OS X only) and Chocolaty (Windows only) as software and on GitHub, DockerHub, MRAN and Ubuntu Package Archive as services.
+These many breakpoints can be reduced to one service and one software.
 That is because all required software is bundled into the container, whose image can be stored.
 Hence, one needs a storage provider which saves the container image and the other files associated with the project.
-We already recommend to store all project files redundantly (your computer, a remote git server such as GitHub, and a publicly funded, open-science-compatible service provider such as zenodo.org).
+We already recommend to store all project files redundantly (your computer, on GitHub and zenodo.org).
 Building the Docker image from the Dockerfile still requires DockerHub, MRAN, and Ubuntu Package Archive to be online.
-We therefore suggest that you include an already built image into the releases.
-The Makefile generated by `repro` does include a target to do so.
+We therefore suggest that you include an already build image into the releases.
+The Makefile generated by `repro` does include a target to facilitate this.
 To write a compressed image to disk use:
@@
 
-This image allows anyone with a computer equipped with Docker to reproduce the analysis (if the repository contains all code and data and has no other external dependencies).
-Last, one can convert the Docker image to a full bootable system image.
-Such a system image can subsequently be booted on any personal computer without relying on any software or service and therefore grantees reproducibility for decades to come.
-Currently, `repro` does not yet facilitate this conversion, but we hope to include this feature soon.
+Unfortunately, this image can only be used when Docker is installed.
+To remove any software dependency including Docker from the archived project, one can convert the Docker image to a full system image.
+Such system image can subsequently be booted on any personal computer without relying on any software or service and therefore grantees reproducibility for decades to come.
+Currently, `repro` does not facilitate this conversion, but we hope to include this feature soon.
 

### 2021-05-20 manuscript.Rmd
Final changes

@@
 Only with access to both data and procedures can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.
-<!-- CJ this final claim is questionable, at least before introducing the concept of containerization.
-Just using data and procedures, you cannot guarantee reproducibility on any computer.-->
 
@@
 Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.
-For example, @hardwicke2018 attempted to replicate studies ostensibly adhering to the standards of open science in the journal Cognition.<!-- CJ How do you replicate open materials?
-At least in the social sciences "open materials" is taken to mean that you have used a permissive license on a newly developed questionnaire or other research material.
--->
+For example, @hardwicke2018 attempted to replicate open materials in the journal Cognition.
 Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
@@
 The authors could execute only 31 of the scripts without error and were able to reproduce the results of only 21 articles (within a reasonable period of time).
-Thus, there is a dire need to establish standards for reproducible workflows for scientific rigor.<!-- CJ It does not automatically follow that the problem you describe necessitates reproducible workflows.
-I would procede as follows: 1) use this example to conclude the existence of a problem (even "open" studies do not reproduce), 2) present workflows as a solution.
--->
-
-From our personal experience,<!--CJ This undermines the credibility of the argument.
-This is not an opinion piece--> the lack of a proper incentive structure and the dreary perspective of having to spend several extra hours if not days to properly set of a code repository to become reproducible, makes many scholars skip this extra step.
-<!-- CJ I am not sure this is the primary reason.
-Indeed, incentives are important - but in that context, perhaps reference DORA?
-I think you were just on a good track, arguing that even with the best intentions, researchers often fail to make their work reproducible.
-Your workflow will ensure that their next attempt is succesful.-->
-If reproduction would require merely minutes instead of hours,<!-- CJ Now we're back to ease-of-use, not incentives.-->
-it could greatly facilitate collaboration within and across research projects.
-<!-- CJ This is a different, third argument: In order to collaborate effectively, researchers should be looking at exactly the same data, code and results.
-I suggest to organize these arguments, and address them each exhaustively in turn.
--->
+Thus, there is a dire need to establish standards for reproducible workflows for scientific rigor.
+
+From our personal experience, the lack of a proper incentive structure and the dreary perspective of having to spend several extra hours if not days to properly set of a code repository to become reproducible, makes many scholars skip this extra step.
+If reproduction would require merely minutes instead of hours, it could greatly facilitate collaboration within and across research projects.
 <!-- Guiding principle: spending machine compute time instead of human research time -->
-To achieve this boost in productivity, we must place the burden of reproducibility upon computers instead of the researchers.<!-- CJ I understand what you mean, but it's not yet communicated effectively because you have not yet made explicit that reproducible projects will indeed facilitate collaboration, and increase opportunities for collaboration (e.g., with strangers through github), and will therefore lead to a boost in productivity.
-Also - ask yourself if this boost in productivity is convincing to individual researchers, your target audience.
-It seems like this would be more convincing to management and policy makers.-->
-To this end, we demand <!-- CJ It's you creating the product, so.. you demand this of yourself? Or do you mean that you demand this of your readership? Either way, strange choice of words.-->
-that a scientific product (no matter whether this is a preliminary report or a full research article) should both be understandable by researchers and automatically reproducible by computers with minimal constraints to the researchers' own workflows.
-<!-- CJ This is ambiguous.
-The point of "understandable by researchers and automatically reproducible by computers" requires clarification.
-"Minimal constraints to the workflow" is also not clear.
-Aren't you providing readers with a workflow?
--->
+To achieve this boost in productivity, we must place the burden of reproducibility upon computers instead of the researchers.
+To this end, we demand that a scientific product (no matter whether this is a preliminary report or a full research article) should both be understandable by researchers and automatically reproducible by computers with minimal constraints to the researchers' own workflows.
 
@@
 <!--stolen from: https://brandmaier.github.io/reproducible-data-analysis-materials/KULeuvenQuantPsy2020.html#11 -->
-We identified the following common four causes for non-reproducibility: <!-- CJ How did you identify these causes? Is this exhaustive?-->
+We identified the following common four causes for non-reproducibility:
 
@@
 4.	Broken dependencies; for example, a given analysis depends on a software package that might not be available on a different computer or no longer exists.
-<!-- CJ 5.
-Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
-We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph.
-The opening paragraph should at least:
-1) Introduce the topic area
-2) Introduce the problem/knowledge gap
-3) Introduce how you will fill this knowledge gap-->
+
+We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:
 
@@
 This tutorial follows the recommendations of @Peikert2019 and uses _Git_ for version control, _R Markdown_ for dynamic document creation, _Make_ for dependency tracking, and _Docker_ for software management.
-<!-- CJ None of these terms have been defined yet-->
 The interplay of these four components is shown in Figure \@ref(fig:schematic).
-However, which tools you use to implement these concepts is a matter of taste and project requirements, and other choices are possible depending on the context and the degree of reproducibility one aims for [e.g., @vanlissa2020worcs; @turingway2021turing].<!-- CJ But readers do not yet know what these concepts are, nor what tools exist.
-At this point in the writing, I think you need to take them by the hand and explain.
-After doing that, you can let them know that there are alternatives.-->
+However, which tools you use to implement these concepts is a matter of taste and project requirements, and other choices are possible depending on the context and the degree of reproducibility one aims for [e.g., @vanlissa2020worcs; @turingway2021turing].
 At several points, we suggest other viable alternatives.
@@
 This paper was designed to help researchers to practically apply the TOP guidelines [except for number 8.
-which does not apply to single research projects, @nosekPromotingOpenResearch2015; @vanlissa2020worcs].<!-- CJ Is this really the case?
-Seems like there is little attention to the TOP guidelines, and the present paper is much more widely applicable.
-Maybe make the general point that workflows can help address formal guidelines for transparency, like TOP and FAIR - and consider citing WORCS for more information.
-Then explain that your workflow is flexible enough to go beyond-->
-We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.<!-- CJ Is this the description of the workflow?-->
+which does not apply to single research projects, @nosekPromotingOpenResearch2015; @vanlissa2020worcs].
+We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.
 We continue with the creation of a concrete data analysis plan, which ideally is already implemented in code, and a dynamic report document that can be preregistered.
-Once the empirical data has been collected, the results can be calculated and appear directly in the dynamically generated document, where ideally only the discussion section needs to be added before the whole archive can be permanently stored as a fully reproducible digital artifact in a persistent location, and be assigned a DOI to increase findability.<!--CJ All DOIs are unique, and it is not stored using a DOI-->
+Once the empirical data has been collected, the results can be calculated and appear directly in the dynamically generated document, where ideally only the discussion section needs to be added before the whole archive can be permanently stored as a fully reproducible digital artifact using a unique DOI in a persistent location.
 
@@
 
-Here, we present a workflow that has proven useful for our way of working with a strong focus on Open Science.<!-- CJ What do you mean here? -->
+Here, we present a workflow that has proven useful for our way of working with a strong focus on Open Science.
 However, not all steps are necessary to achieve reproducibility.
@@
 
-This tutorial is targeted at scholars with at least intermediate R knowledge<!--CJ This sounds like something that belongs in the beginning of the introduction; it's not part of setup--> and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md", "installation guide")`).
-To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]: <!-- CJ This also seems like something that should be mentioned in the intro; once people are setting up their systems, they should havebeen told what they're setting them up for-->
+This tutorial is targeted at scholars with at least intermediate R knowledge and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md", "installation guide")`).
+To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:
 
@@
 if(!requireNamespace("remotes"))install.packages("remotes")
-remotes::install_github("aaronpeikert/repro") # CJ Suggest publishing (at least aminimal working version) of the repro package to CRAN before publishing this paper, to reduce one layer of complexity for users
+remotes::install_github("aaronpeikert/repro")
 library("repro")
@@
 The [`repro`-package](https://github.com/aaronpeikert/repro) contains several helpers to work with the suggested tools for reproducibility.
-For example, the package can check if your computer has all the required software packages installed and ready. <!-- CJ Does repro assume that these have already been installed  by hand? -->
+For example, the package can check if your computer has all the required software packages installed and ready.
 
@@
 
-If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.<!-- CJ I think you should explain your philosophy that the package provides additional tutorial instructions depending on the user's situation before diving into the actual tutorial.
-I foresee that people are going to be puzzled when the tutorial steps present them with additional mini tutorials.
-Explain in the introduction what the problem is (users need to install andfamiliarize themselves with many different software solutions) and how you solved it (you wrote a package that tells users what dependencies they need to install and how -->
+If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
 Please make sure that all three commands pass before you continue with the tutorial.
@@
 
-# Planning Phase <!-- CJ This reminds me of the WORCS phases, which is fine, but it's the only header that does remind me of these phases.
-You've not introduced a breakdown of the research process into phases anywhere.
-Moreover, the text below does not really relate to "planning"; it goes back to reasons for reproducibility, and then an explanation of technical tools (Rmarkdown and Git), neither of which relate to planning!
--->
+# Planning Phase
 
@@
 While you might be tempted to consider reproducibility only when you about to finish a project, upfront investments pay off through your increased productivity.
-Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you achieve reproducibility gradually.
+Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you achieve reproducibility gradually.<!--CJ The three preceding sentences belong in the introduction; this tutorial part should be more applied, less conceptual-->
 The first step towards reproducibility is to bundle all required files in one folder.
@@
  Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
- This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+ This way, Markup<!-- CJ Up or down?--> does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
 We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
@@
 
-We will later see how this simplicity enables us to create many different document formats (i.e. Word, PDF, HTML) and incorporate code (i.e. R, Python, Julia).
+We will later see how this simplicity enables us to create many different document formats (i.e. Word, PDF, HTML) and incorporate code in different languages (i.e. R, Python, Julia).
 
@@
 
-Another advantage of Markdown is that because of its likeness to plain text, you can easily version-control it with Git and leverage Gits features for tracking, displaying, and merging sequential and concurrent changes.
+Another advantage of Markdown is that because of its likeness to plain text<!-- CJ This will not make sense to someone who does not yet know how git works and why plain text is preferred, i.e., your target audience-->,
+you can easily version-control it with Git and leverage Gits features for tracking, displaying, and merging sequential and concurrent changes.
 You may be familiar with Microsoft Words' "Track Changes"-Feature.
-Version control is similar in its basic idea but more potent in its features.
+Version control <!-- CJ You're now introducing version control after using it in a sentence (preceding).
+That's too late.
+Moreover, as the paragraph is labeled "Git", I feel that you should explain git, not version control.
+Also...
+shouldn't tools be introduced in the introduction, instead of the tutorial?
+The tutorial should show users how to use tools that have been explained to them in the introduction.-->
+  is similar in its basic idea but more potent in its features.
 Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
 For example, this paper has been going through `r system2("git", c("rev-list",  "--count",  "HEAD"), stdout = TRUE)` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/main", "iterations")`.
+<!-- CJ Weird flex, readers will only see the number and won't know that it was dynamically generated from a git call.
+They also won't understand how that number conveys anything about the benefits of git.-->
 To activate Git in a given project, you can call:
@@
 
-And you could commit it in Git.
-For convenience, we use the [`gert`-package](https://github.com/r-lib/gert) [@R-gert], which wraps most git commands into R functions.
+And you could commit it in Git.<!-- CJ Do people know what a commit is yet?-->
+For convenience, we use the [`gert`-package](https://github.com/r-lib/gert) [@R-gert], which wraps most git commands into R functions.<!-- CJ Actually, I believe gert is an independent git client - but not 100% sure-->
 Alternatively, we suggest to either work directly in the terminal (for example, RStudio offers a command shell with various options accessible in the Menu Tools → Terminal), or work using the Git pane in RStudio that offers buttons for basic operations.
@@
 <!--For Git we maybe should use RStudio's Git pane?-->
-
-To commit your first changes to the project, start with staging all files to be included in the commit using `git_add()`.
+<!-- CJ No preference yet.
+I think command line R is convenient, but the R git pane gives a nice overview of new and changed files etc.
+That would be very instructive to new users.
+Maybe you can demo how the command line relates to the git pane?
+I.e., the git pane will be updated when using gert::git_add()-->
+To commit your first changes to the project, start with staging all files <!-- CJ do they know what staging is yet? Do they need to know?--> to be included in the commit using `git_add()`.
 Here, we stage our idea collection in the file `idea.md` and then commit it to our local repository with a meaningful commit message that will later help us keep track of all the changes over time.
@@
 
-Now assume that after discussing with your colleagues, you settle onto a final analysis strategy using latent variable models with tests for measurement invariance because power is greater and you want to exclude bias in measurement as possible confounder [@flake2020].
+Now assume that after discussing with your colleagues, you decide on a different strategy for the final analysis <!-- CJ I missed what the original plan was-->: using latent variable models with tests for measurement invariance <!--CK consider dropping, it's not essential and confusing to readers who don't know SEM: because power is greater and you want to exclude bias in measurement as possible confounder--> [@flake2020].
 This results in changes to your notes as follows:
@@
 
-And GitHub can show you what has changed, like in Figure \@ref(fig:idea-change).
-This detection of what has changed works on a line-by-line basis.
+And GitHub can show you what has changed, like in Figure \@ref(fig:idea-change). <!-- CJ Is the project linked to github yet? Have the changes been pushed? Do readers know what github is yet? -->
+This detection of what has changed works on a line-by-line basis. <!-- CJ this should have been explained when you first introduced git -->
 We recommend therefore to keep each sentence in a Markdown on its own line to generate clean difference views.
-GitHub is, in its essence, a web interface to Git, and you may learn more about how to use it under the section on [GitHub].
+GitHub is, in its essence, a web interface to Git, and you may learn more about how to use it under the section on [GitHub]. <!-- CJ This should have been introduced earlier-->
 
@@
 <!--I am troubled by how we should recommend to learn Git-->
+<!-- CJ I think you should avoid teaching people git.
+You're not teaching them R or how to use RStudio either.
+Instead, teach readers only the absolute minimum they need to use it.
+If possible, consider importing the git_update() function into repro, because that will save a lot of explaining in the paper.
+Do provide a link to a good resource for people who want to learn more; I believe github has tutorial pages.-->
+<!-- CJ Also, I feel like most of what is said about git below here should have been in the introduction.
+Some other snippets (e.g., explaining commit, push and pull) should have been presented before you tell users to make commits.
+It comes too late here-->
 Git can be challenging to learn, because it requires a rather complex mental model of maintaining potentially differing versions between different repositories (typically one central repository and one local repository per each author while each repository may have different branches).
@@
 There are many tutorials on learning Git and @vuorreCuratingResearchAssets2018 provide a good starting point.
-However Git is designed for collaboration, we therefore recommend finding a partner to attend an online lecture or workshop to learn the ins and outs of Git.
+However Git is designed for collaboration, we therefore recommend finding a partner to attend an online lecture or workshop to learn the ins and outs of Git.<!-- CJ I think this is a good sentiment; however, I would (also) provide a link to a good tutorial, and emphasize that all they need to know to get started will be covered in this tutorial, otherwise people get frustrated that they have to go read another tutorial halfway through your tutorial.-->
 
@@
 
-Monte Carlo simulations are a crucial tool to investigate properties of a statistical model and they are routinely employed to develop or evaluate statistical models.
+Monte Carlo simulations <!-- CJ Explain what this is!
+Or consider using plain language instead.--> are a crucial tool to investigate properties of a statistical model and they are routinely employed to develop or evaluate statistical models.
 For example, @maccallum1999sample used Monte Carlo simulations to investigate optimal sample size for exploratory factor analysis, or @curran1996robustness studied the robustness of the $\chi^2$ test statistic under non-normal data.
 
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
+While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.<!-- CJ I feel that this is a better opening paragraph, because it introduces the concept of simulating data!
+However, consider that some readers may have no idea of what that means, and might require even more detail.-->
 First, you have to address and could plan for data analytic problems on a conceptual level before you made the considerable effort to acquire data [@axelrodAdvancingArtSimulation1997].
@@
 
-A Monte Carlo simulation affords you an ideal planning opportunity.
-We realize that this comes at an expensive but believe the reduction in the risk that the data is not sufficient to answer the research questions offsets the required effort.
+A Monte Carlo simulation affords you an ideal planning opportunity.<!-- CJ Put the other monte carlo stuff here, and explain what it is-->
+We realize that this comes at an expensive but believe the reduction in the risk that the data is not sufficient to answer the research questions offsets the required effort.<!-- CJ What are you trying to say in this sentence?
+-->
 This advantage is especially pronounced when you plan to preregister your study.
-In fact, we argue that a Monte Carlo simulation can combine the analysis, sample size planning and preregistration (see [Preregistration as Code].
+In fact, a Monte Carlo simulation can help plan the analysis in detail, form the basis of power calculation and sample size planning, and can be preregistered integrally (see [Preregistration as Code].<!-- CJ Should this preregistration as code have been introduced earlier?
+There, you could discuss the notion of simulating data-->
 
@@
 As a heuristic guess, we assume that the factor loadings are 30% lower than those reported by @Jones2013 and that the standardized mean difference is only 0.2.
-These assumption can be simply translated into [`lavaan` [@R-lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
+These assumption can be simply<!--CJ I wouldn't use this word here.
+People might find this all quite complicated, and could be disparaged if they read that this should be simple--> translated into [`lavaan` [@R-lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
 
@@
 The following code simulates data from the population model `model_truth` with 1000 cases in each group.
+<!-- CJ I'm left wondering whether the example shouldn't simply be a t-test with observed variables.
+Everybody knows linear models.
+Lavaan is quite a specialized piece of software, and you might lose a lot of readers when the examples relate to SEM and require understanding of factor loadings, model implied covariance matrices, etc -->
 
@@
 We propose the following procedure for analyzing the data and dealing with possible violations of strong measurement invariance.
+<!-- CJ Again, mweasurement invariance is probably too technical for an introductory example in a paper about workflows, not SEM-->
 After fitting the model with strong measurement invariance [@flake2020], we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
+<!-- CJ This is really getting too complex for an example -->
 The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
@@
 ```
+<!-- CJ I would just use an example as simple as something like this:
+x <- rnorm(100)
+y <- .2 + 2.3 * x + rnorm(100)
+That's already going to be a big step for many, but hopefully manageable. After grokking that, they can get back to the true point of the paper: Workflows.
+-->
 
@@
 Note however that the consistency of a given seed is not necessarily maintained across different R versions because the default algorithm for generating pseudo random numbers may change (see @Peikert2019 for a demonstration with different R versions).
-It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
+It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).<!-- CJ This is probably too specialized for a paper about workflows.
+You have not introduced parallelization, so it doesn't make sense to talk about special provisions for paralellized code.
+-->
 The standard random number generator (RNG) produces high quality random numbers, but is not safe to use in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
@@
 @lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
-While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers:
+While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers:<!-- CJ These are many technical terms that apply only to parallelized code; a concept you have not introduced and one that seems out of scope-->
 
 ```{r}
+# reproducible but unsafe in parallel code
 set.seed(1235)
@@
 The goal of preregistration is transparency and a clear-cut divide into exploration and confirmation.
-Thus, it serves as a tool that may help to increase the credibility of empirical results.
-Without getting into too much detail of the meta-scientific arguments for preregistration, we want to note two advantages [@nosekPreregistrationHardWorthwhile2019].
+Thus, it serves as a tool that may help to increase the credibility of empirical results.<!-- CJ For this sentence to make sense, you have to briefly mention the problem with non-preregistered studies: Undisclosed exploration, HARKing, and the file drawer effect-->
+Without getting into too much detail of the meta-scientific arguments for preregistration <!-- CJ I think you do need to mention some of those arguments.-->, we want to note two advantages [@nosekPreregistrationHardWorthwhile2019].
 First, a preregistration is an opportunity to plan a study carefully.
 Gathering empirical data requires many resources, and better planning increases the chances that these efforts are worthwhile.
-For us, a good preregistration is the same as a good plan; the only difference is that a preregistration is made public beforehand.
+A preregistration is essentially a detailed research plan; the only difference is that a preregistration is made public <!-- CJ Not necessarily!
+prereg can be deposited in a private or embargoed repository.
+The crucial point is that it is time-stamped and indelible, so that future auditors may verify correspondence between the prereg and final analysis.
+--> beforehand.
 
 The second advantage is transparency.
-A preregistration offers the interested reader more insight into the research process than a paper alone could.
+A preregistration offers the interested reader more insight into the research process than a paper alone could.<!-- CJ Yes, but is that so important? -->
 @szollosi_is_2020 argue that preregistration can not fulfil its promise to reduce the false discovery rate of science if we allow deviating from them.
 While such an argument has some merits, we believe that often transparency outweighs these concerns [@nosekPreregistrationHardWorthwhile2019].
-Deviation from a preregistration is still possible.
-If researchers deviate from their preregistration, they probably have good reason to do so, which means they have learned something from the process of conducting the study, and any deviation will simply be explicitly marked as such.
+<!-- CJ Not clear how transparency outweighs (?) concerns about inflated false positive rate.-->
+Researchers may have good reason to deviate from a preregistration; for instance, if real data are found to violate assumptions (e.g., about distributions, or measurement levels) that were made at the time of study planning.
+To account for such unforeseen circumstances, there is some room to deviate from a preregistration - provided the deviation is clearly disclosed.
 A preregistration, together with version control, allows their readers to inform themselves about these lessons learned.
+<!-- CJ I would emphasize here that version control functions as a research log / diary, which shows how a project progressed from idea to prereg to preliminary analysis to corrected analyses with some deviations from the original plan.
+If the project log is straigfhtforward, with some disclosed deviations from prereg - the research is credible.
+If the project log is convoluted as f***, then it's not credible.-->
 
@@
 Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration [@osfstandard; @socpsytemp; @howtoreplicate]. 
-Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code [@sep-ambiguity].
+All of these templates use natural language to describe the analysis plan.
+The crucial limitation of any natural-language representation of an analysis plan is that it leaves room for ambiguity when translating the plan into code [@sep-ambiguity].
+Preregistration as code (PAC) addresses this limitation, by... <!-- CJ Provide definition first.-->
 Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity.
@@
 to establish partial measurement invariance, when full measurement invariance is not attainable).
-This approach allows for more flexibility than a traditional template based preregistration but is still well defined.
+<!-- CJ This is a great idea, but I feel like you should first explain the general concept of PAC in more detail before presenting specific benefits, like forking paths.-->
+This approach allows for more flexibility <!-- CJ This is confusing, because you just said that traditional templates allowed for too much flexibility--> than a traditional template based preregistration but is still well defined.
 
@@
 One way to ensure that your analysis to works with real data is to simulate data.
+<!-- CJ I'm having flashbacks to the previous paragraph about simulating data.
+Haven't you said this already?
+Merge all sections about the same topic-->
 We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
@@
 Above, we have already suggested [Markdown] as a universal and portable text format.
+<!-- CJ I feel like these sections should be merged.
+It does not make sense to introduce markdown first, and then several pages down, introduce R Markdown.-->
 R Markdown extends regular Markdown by allowing useres you to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
@@
 You could, in principle, even write large parts of the manuscript as part of your preregistration, i.e. the introduction, theoretical background, the method section and parts of the discussion.
-If you have simulated data, you could even include fake results.
+If you have simulated data, you could even analyze them to include dummy results.
 Packages like [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf)[^stargazer] [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html)[^apaTables] [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/)[^papaja] [@papaja] help you to create dynamically generated professional looking results.
@@
 An essential aspect of a preregistration is that you have to make it public.
-Of course, if you commit the preregistration to Git and upload it into a public repository, it is also public.
+<!-- CJ This requires a reference.
+I believe I found one that defined prereg as an "indelible, time-stamped record".
+That's useful-->
+Of course, if you commit the preregistration to Git and upload it into a public repository, it is also public. <!-- CJ You should credit my paper for this, as I'm the first to make this argument.-->
 However, simply using GitHub as a preregistration site is not ideal.
@@
 You can alleviate these downsides by using zenodo.org to give a digital object identifier (DOI) to your preregistration because zenondo.org saves an unchangeable copy of your repository.
-You can learn more about this in [Creating a public release].
+You can learn more about this in [Creating a public release].<!-- CJ That's a great idea! -->
 
@@
 
-### README (optional)
+### README (optional) <!-- CJ I think this is too much information about the readme file. -->
 
@@
 <!---exact copy from our previous manuscript, do we need to change? -->
-In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media (such as stimulus materials if you created them).
+In our experience <!-- CJ This sentence should not be used in relation to legal matters.
+Refer to a trusted legal authority, not personal experience.
+-->, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media (such as stimulus materials if you created them).
 In contrast, scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
-Copyright does often not apply to data, we therefore recommend to share data with a CC0 license to clarify that you do not hold copyright to it.
+Copyright does often not apply to data <!-- CJ Reference, also this is a very important point that might require some elaboration.-->, we therefore recommend to share data with a CC0 license to clarify that you do not hold copyright to it.
 Giving data another license than CC0 is often just confusing, because a license is invalid if there is no copyright.
@@
 It is important to get written consent from collaborators when you change/add the license.
-However, if their first contribution occurs after adding a license, they give their consent by contributing.
+However, if their first contribution occurs after adding a license, they give their consent by contributing.<!-- CJ Is this true by default, or should it be included in your contributor guidelines or someting?-->
 To add a CC-BY license to your project, simply run (there are similar commands for other licenses):
@@
 
-### Code of Conduct
+### Code of Conduct <!-- CJ I have mixed feelings about this.
+Considering that CoC is a recent development in software communities, and that it has caused substantial debate (and some casualties) in those communities, and considering that it is not currently "a thing" in scientific communities, and considering that it does not directly relate to reproducible workflows - I sincerely doubt whether this has any place in the present tutorialn.
+I would recommend sticking with solutions that solve issues experienced by scientists, not by programmers.-->
 
@@
 
-## Collaboration
+## Collaboration <!-- CJ This should be in the introduction - in fact, there was already a remark about it. Merge those sections -->
 
@@
 
-### GitHub
+### GitHub <!-- CJ merge this with your earlier github section -->
 
@@
 
-### GitHub Issues
+### GitHub Issues <!-- CJ This is another example of something not essential for reproducibility; consider leavcing it out of the paper. Definitely not worth it's own header! -->
 
@@
 
-### GitHub Pull Requests
+### GitHub Pull Requests <!-- CJ Same, not relevant -->
 
@@
 
-<!--I am not sure this is the right place, should we introduce this earlier? -->
+<!--I am not sure this is the right place, should we introduce this earlier? --> <!-- CJ Yes-->
 
@@
 For example, they create a script to load the data and another to run the analysis.
-This modularisation facilitates reuse because the loading script may be used by other projects building on the same data, while the analysis script may be employed on different data and so forth.
+This modularisation <!-- CJ Are you using UK or US spelling?
+Modularization?
+--> facilitates reuse because the loading script may be used by other projects building on the same data, while the analysis script may be employed on different data and so forth.
 The downside is that reproducibility is complicated because instead of reproducing everything in one step, multiple steps are required.
@@
 Of course, there a several ways to solve this problem.
-One could build an R package to separate analysis and data or use code externalization in the RMarkdown to facilitating reuse or create a "main" script that combines all steps.
-We found Make to be particularly useful for two reasons.
-First, it is language agnostic, therefore fusing steps that are carried out in different programming languages.
-Second, it does implement a simple yet powerful caching mechanism of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning.
-
-Make is build around the concept of computational recipes.
+One could build an R package to separate analysis <!-- CJ This sentence is going to lose all readers who are not R package developers, i.e., the entire target readership--> and data or use code externalization <!-- CJ Now you're even losing me ;) --> in the RMarkdown to facilitating reuse or create a "main" script that combines all steps.
+We found Make to be particularly useful for two reasons. <!-- CJ Unless I'm mistaken, Make has not yet been introduced? -->
+First, it is language agnostic, therefore fusing steps that are carried out in different programming languages. <!-- CJ This is confusing, because you have not yet explained what make is / does -->
+Second, it does implement a simple yet powerful caching mechanism <!-- CJ Eh??
+What is this --> of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning.
+
+Make is build around the concept of computational recipes.<!-- CJ This is better. Start with this (and add one sentence first with a definition of Make-->
 Like cooking recipes, the dish you want to create is at the top, followed by the necessary ingredients with the procedure at last.
@@
 The ingredients are called "requirements" and are usually also files.
-The procedures are formulated as shell commands.
+The procedures are formulated as shell commands. <!-- CJ This might shock you, but many readers won't know what shell commands are. -->
 One (hypothetical) example is this:
@@
 
-Modern data analysis is only feasible because researchers rely on much external software.
+Modern data analysis is only feasible because researchers rely on much external software. <!-- CJ This is kind of a sweeping statement... what does it mean?-->
 One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.

### 2021-05-21 manuscript.Rmd

@@
 With these advances came challenges that contribute to broader concerns of irreproducibility in the published literature, among them the lack of exact and unambiguous documentation of computational methods and each computational step taken.
+
 With computationally intensive data analyses <!-- CJ The complexity of the analyses is orthogonal to the necessity to make them reproducible.
 -->, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
-Only with access to both data and procedures <!-- CJ and software environment --> can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.
+Only with access to both data and procedures <!-- CJ and software environment --> can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.<!-- CJ this final claim is questionable, at least before introducing the concept of containerization.
+Just using data and procedures, you cannot guarantee reproducibility on any computer.-->
 
@@
 For example, @hardwicke2018 attempted to replicate open materials in the journal Cognition.
+<!-- CJ How do you replicate open materials?
+At least in the social sciences "open materials" is taken to mean that you have used a permissive license on a newly developed questionnaire or other research material.
+-->
+
 Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.

### 2021-05-21 manuscript.Rmd

@@
 Reproducibility is invaluable for effective collaboration---including with your "past self".
-While you may be tempted to consider reproducibility only when you are close to completing a project, we believe that upfront investment pays off in your increased productivity.
-Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you achieve reproducibility gradually.
+
+While you might be tempted to consider reproducibility only when you about to finish a project, upfront investments pay off through your increased productivity.
+Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you achieve reproducibility gradually.<!--CJ The three preceding sentences belong in the introduction; this tutorial part should be more applied, less conceptual-->
+
 The first step towards reproducibility is to bundle all required files in one folder.
@@
 
-Monte Carlo simulations are a crucial tool to investigate properties of a statistical model and they are routinely employed to develop or evaluate statistical models.
+Monte Carlo simulations <!-- CJ Explain what this is!
+Or consider using plain language instead.--> are a crucial tool to investigate properties of a statistical model and they are routinely employed to develop or evaluate statistical models.
 For example, @maccallum1999sample used Monte Carlo simulations to investigate optimal sample size for exploratory factor analysis, or @curran1996robustness studied the robustness of the $\chi^2$ test statistic under non-normal data.
-In Bayesian data analysis, prior predictive checks generate data from a model according to the prior distribution in order to asses whether a prior is appropriate [@gabry2019visualization].
-
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
+
+While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.<!-- CJ I feel that this is a better opening paragraph, because it introduces the concept of simulating data!
+However, consider that some readers may have no idea of what that means, and might require even more detail.-->
 First, you have to address and could plan for data analytic problems on a conceptual level before you made the considerable effort to acquire data [@axelrodAdvancingArtSimulation1997].
@@
 @lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
-While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers.
-
-<!--
+
+While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers:<!-- CJ These are many technical terms that apply only to parallelized code; a concept you have not introduced and one that seems out of scope-->
+
 ```{r}
@@
 ```
--->
+
 By wrapping code that uses the internal random number generator (here, the command to create a random number from the uniform distribution, `runif(1)`) with the `future` approach, we guarantee that the random numbers will differ across parallel executions of that code, will achieve a high quality of randomness, but still will be deterministic and thus reproducible:
@@
 
-<!--I am not sure this is the right place, should we introduce this earlier? -->
-
-A reader of a reproducible research project should be able to easily derive how it can be reproduced (ideally by a statement in your [README]).
-To do that, it is often crucial to understand how different project components relate to each other and depend on each other.
-For example, you cannot run an analysis without data, or you cannot render a statistical report without having the results from the statistical analysis that is reported.
-Dependency tracking means that you capture and map the interrelations between all components of your analysis.
-The "components" of an analysis are typically all involved files including raw data files, intermediate files (such as a preprocessed data file), and final results (such as a rendered manuscript in PDF format).
-With larger projects, there is typically a higher degree of modularization because code fragments are written for re-use in other projects (or are already re-used).
-Also, more complicated empirical research projects may have different hypotheses that are answered with different data sets and different statistical models, such that it often makes sense to split the different analyses into different files.
-Reproducibility then demands the careful documentation and mapping of all necessary steps in the correct order.
-
-Make is a build automation tool for software projects that typically are challenged by maintaing and resolving a variety of dependencies in the correct order when compiling a piece of software-
-We found _Make_ to be particularly useful for our purposes for two reasons.
-First, it is language-agnostic, therefore enabling us to map steps not only taken within R but also across different programming languages or other external programs.
-Second, it does implement a simple yet powerful caching mechanism of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning. 
-That is, Make determines automatically which branches in the dependency graph needs updating when files have changed and does not necessarily recompute or rebuild everything.
-
-Make is build around the concept of computational recipes.
+<!--I am not sure this is the right place, should we introduce this earlier? --> <!-- CJ Yes-->
+
+A reader of your reproducible manuscript should be able to derive how they are to reproduce it (ideally by a statement in your [README]).
+To do that, it is often crucial to understand how different project components relate to each other.
+You cannot run an analysis without data; you cannot render a dynamic document without an analysis, and so forth.
+Dependency tracking means that you capture the interrelations between the components of your analysis.
+The "components" of an analysis are often loosely mapped onto files.
+For a simple project, it might be enough to have a single R Markdown file.
+However, while it is easy to understand how to reproduce the end result, this approach quickly gets unwieldy.
+To better navigate the data analysis, a researcher often splits it across several scripts.
+For example, they create a script to load the data and another to run the analysis.
+This modularisation <!-- CJ Are you using UK or US spelling?
+Modularization?
+--> facilitates reuse because the loading script may be used by other projects building on the same data, while the analysis script may be employed on different data and so forth.
+The downside is that reproducibility is complicated because instead of reproducing everything in one step, multiple steps are required.
+
+Of course, there a several ways to solve this problem.
+One could build an R package to separate analysis <!-- CJ This sentence is going to lose all readers who are not R package developers, i.e., the entire target readership--> and data or use code externalization <!-- CJ Now you're even losing me ;) --> in the RMarkdown to facilitating reuse or create a "main" script that combines all steps.
+We found Make to be particularly useful for two reasons. <!-- CJ Unless I'm mistaken, Make has not yet been introduced? -->
+First, it is language agnostic, therefore fusing steps that are carried out in different programming languages. <!-- CJ This is confusing, because you have not yet explained what make is / does -->
+Second, it does implement a simple yet powerful caching mechanism <!-- CJ Eh??
+What is this --> of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning.
+
+Make is build around the concept of computational recipes.<!-- CJ This is better. Start with this (and add one sentence first with a definition of Make-->
 Like cooking recipes, the dish you want to create is at the top, followed by the necessary ingredients with the procedure at last.
@@
 
-Modern data analysis is only feasible because researchers rely on much external software. 
+Modern data analysis is only feasible because researchers rely on much external software. <!-- CJ This is kind of a sweeping statement... what does it mean?-->
+One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
 The dependency on a single R package may easily create a dependency tree of more than hundred direct or indirect dependencies [@epskamp2019].

### 2021-05-25 manuscript.Rmd
beginn rewrite, add introduction

--- 2021-05-21 manuscript.Rmd
@@
 
-# Why you should care about reproducibility
+# Introduction
 
-A cornerstone of scientific progress is the possibility to critically assess the validity of peer scientists’ scientific conclusions.
-<!-- CJ More fundamental: Reproducibility is a cornerstone of the scientific METHOD.
-I think it's more important to emphasize this, rather than critical assessment of others' work.
-See Karl Popper, The Logic of Scientific Discovery, Routledge, London, 1992, p.
-66.
--->
-Scientific claims should not be credible because of their originators' authority but by the transparency and replicability of their supporting evidence [@aac4716].
-<!-- CJ This is a red herring; nobody said that an argument to authority should be the foundation of scientific evidence.
-I would suggest focusing on replicability as an essential component of the scientific method.
--->
-Here, *replicability* refers to the degree to which original studies and their derivatives yield consistent results.
-In other words, a successful replication study finds a consistent effect with an original study in a new sample of data.
-Enabling the assessment of scientific conclusions requires a systematic and rigorous approach to research.
-Researchers across various fields have identified a *crisis* of non-replicability of findings [e.g. @aac4716]. <!--citation needed, e.g. [@aac4716]-->
+The papers aim is twofold.
+First, we want to introduce the reader to a workflow that ensures reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code. 
+The research community has long accepted reproducibility as integral to "good scientific practice."
+Often, however, projects can not be reproduced by other research teams, despite the best efforts by the creators.
+We present four reasons that lead to irreproducibility, explain four solutions from the software engineering community that eliminate these and show the researchers how the R package `repro` helps them to implement the solutions.
+The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
 
-<!-- define replicability and then... 
- define reproducibility -->
-Even worse, researchers might often not even be able to achieve reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code.
-<!-- CJ I would introduce the topic of reproducibility differently.
-"Even worse" is editorializing; not appropriate for a scientific paper.
-Instead, consider saying: More fundamental even than replication across datasets is the reproducibility of findings within a single dataset.
--->
-When analyses are not properly documented, small mistakes, or circumstances outside the researcher's control (e.g., a corrupted file, a coauthor leaving academia) may render them non-reproducible. 
-Over the past decades, computational statistical methods have radically changed the ability of researchers from all disciplines to process and analyze data with complex statistical models.
-<!-- CJ I think it's more important to focus on the methods that have been developed to ensure reproducibility!)-->
-With these advances came challenges that contribute to broader concerns of irreproducibility in the published literature, among them the lack of exact and unambiguous documentation of computational methods and each computational step taken.
-With computationally intensive data analyses <!-- CJ The complexity of the analyses is orthogonal to the necessity to make them reproducible.
--->, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
-Only with access to both data and procedures <!-- CJ and software environment --> can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.
-
-In fact, the research community has long accepted reproducibility as an important element of "good scientific practice." <!-- CJ: This is the crucial point.
-I would start the paper with this.
-See my reference to Popper in the opening sentence.
--->
-However, ensuring reproducibility of computational analyses is still tedious and difficult <!-- CJ I would not editorialize this.
-Moreover, this sounds discouraging.
-I would focus on what you have done to make reproducibility more attainable.
--->, and there are still no widely accepted standards that can guide researchers when setting up their data analyses.
-<!-- CJ These are two separate points, each important enough to have their own sentence, with some further clarification.
--->
-Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories <!-- CJ you have not introduced any open repositories yet.
---> do not provide sufficient information to reproduce relevant computational and statistical results.
-For example, @hardwicke2018 attempted to replicate open materials in the journal Cognition.
+This tutorial is aimed at researchers who are willing to make code, data, and materials available and should help them to ensure reproducibility of relevant statistical results.
+While the willingness to share materials is required for reproducing a project it is not sufficient.
+For example, @hardwicke2018 attempted to reproduce results from open materials in the journal Cognition. 
 Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
-For 13 articles, at least one outcome could not be reproduced —-- even with the original authors' assistance.
+For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
 @obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
 The authors could execute only 31 of the scripts without error and were able to reproduce the results of only 21 articles (within a reasonable period of time).
-Thus, there is a dire need to establish standards for reproducible workflows for scientific rigor.
 
-From our personal experience, the lack of a proper incentive structure and the dreary perspective of having to spend several extra hours if not days to properly set of a code repository to become reproducible, makes many scholars skip this extra step.
-If reproduction would require merely minutes instead of hours, it could greatly facilitate collaboration within and across research projects.
-<!-- Guiding principle: spending machine compute time instead of human research time -->
-To achieve this boost in productivity, we must place the burden of reproducibility upon computers instead of the researchers.
-To this end, we demand that a scientific product (no matter whether this is a preliminary report or a full research article) should both be understandable by researchers and automatically reproducible by computers with minimal constraints to the researchers' own workflows.
+Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.
+These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
+To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use .
+A rigorous standard increases the likelihood of any individual project to be actually reproducible.
+An easy to use standard on the other hand is more likely to be employed across research projects.
+The remainder of this paper is structured along these lines.
+The first part introduces theoretical concepts and software solutions that together form a rigorous reproducibility standard.
+The second part is a tutorial that shows how these software solutions can be easily employed by researchers with the help of the R package `repro`.
 
-<!-- Which problems are solved by which tool (like in Andreas talk)-->
-<!--stolen from: https://brandmaier.github.io/reproducible-data-analysis-materials/KULeuvenQuantPsy2020.html#11 -->
-We identified the following common four causes for non-reproducibility:
+# Concepts and Software Solutions
 
-1.	Multiple inconsistent versions of code and data; for example, the dataset has changed over time because outliers were removed at a later stage, or the code has changed because a bug was removed at some point in time;
-2.	Copy-and-paste errors; for example, results are manually copied from a computing environment into a text processor, causing inconsistencies between the reported result and the reproduced result;
-3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;
-4.	Broken dependencies; for example, a given analysis depends on a software package that might not be available on a different computer or no longer exists.
+<!-- could andreas write this? --->
 
-We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:
+# Tutorial
 
-1. Version control
-2. Dynamic document creation
-3. Dependency tracking
-4. Software management
+# Preregistration as Code
 
-This tutorial follows the recommendations of @Peikert2019 and uses _Git_ for version control, _R Markdown_ for dynamic document creation, _Make_ for dependency tracking, and _Docker_ for software management.
-The interplay of these four components is shown in Figure \@ref(fig:schematic).
-However, which tools you use to implement these concepts is a matter of taste and project requirements, and other choices are possible depending on the context and the degree of reproducibility one aims for [e.g., @vanlissa2020worcs; @turingway2021turing].
-At several points, we suggest other viable alternatives.
+We argue that a preregistration as code (PAC) is an excellent planning tool that offers several benefits over traditional preregistration.
+For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data.
 
-```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency tracking (Make), software management (Docker), and dynamic document generation (R Markdown).
-Git tracks changes to the project over time.
-Make manages dependencies among the files.
-Docker provides a container in which the final report is built using dynamic document generation in R Markdown.
-Reproduced from @Peikert2019.
-"}
-# file gets downloaded in Makefile
-knitr::include_graphics("images/nutshell.svg", auto_pdf = TRUE)
-```
+We consider three criteria when planning a study.
+First, the plan should be comprehensive, e.g. state the research question and describe the study design and analysis.
+Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand. 
+Third, the plan should be efficient; therefore, it only employs the necessary amount of resources.
 
-This tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
-This paper was designed to help researchers to practically apply the TOP guidelines [except for number 8.
-which does not apply to single research projects, @nosekPromotingOpenResearch2015; @vanlissa2020worcs].
-We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.
-We continue with the creation of a concrete data analysis plan, which ideally is already implemented in code, and a dynamic report document that can be preregistered.
-Once the empirical data has been collected, the results can be calculated and appear directly in the dynamically generated document, where ideally only the discussion section needs to be added before the whole archive can be permanently stored as a fully reproducible digital artifact using a unique DOI in a persistent location.
+To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript.
+The introduction and theoretical background provide the basis for the studies design and can be written before gathering data.
+We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
+To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
+The researchers can include a "faked " results section based on the simulated data, employing dynamic document generation.
+Repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to assure efficiency.
+For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study.
 
-We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.
-We explicitly encourage readers to try out all R commands for themselves.
-Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).
-To guide the reader through all steps of a research project, we will use a exemplary research question:
+Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [https://www.pnas.org/content/115/11/2600].
+Preregistration increases the credibility of empirical results in three ways [preregistration is hard and worthwhile, Nosek].
 
-> "Is there a mean difference in the personality trait 'Machiavellism' between genders?"
+> First, preregistration of analysis plans makes clear which analyses were planned a priori and which were conducted post hoc.
+This improves calibration of uncertainty for unplanned analyses, and diagnosticity of statistical inferences for planned analyses.
+[...] Doing so has the benefit of strengthening statistical inferences as compared with unplanned analyses.
+[...] Second, preregistration enables detection of questionable research practices such as selective outcome reporting (John et al.
+2012) or Hypothesizing After the Results are Known (HARKing; Kerr, 1998).
+Third, preregistration of studies can reduce the impact of publication bias—particularly the prioritization of publishing positive over negative results—by making all planned studies discoverable whether or not they are ultimately published.
 
-This example serves only didactic purposes and not to derive substantive claims.
-Nevertheless, everything else is kept as close to a real research question as possible and we have preregistered the analysis without looking at the data beforehand and use actual data subsequently.
-
-Here, we present a workflow that has proven useful for our way of working with a strong focus on Open Science.
-However, not all steps are necessary to achieve reproducibility.
-We have marked those steps that can be omitted as "optional" in the section headings.
-
-Table \@ref(tab:overview) shows all steps required to to create and open and reproducible research project as they are presented in the remainder of the manuscript.
-
-```{r overview, echo=FALSE}
-overview <-
-  read_csv(
-    here::here("data", "overview.csv"),
-    col_types = "cc",
-    col_names = c("Phase", "Steps"),
-    trim_ws = FALSE
-  )
-pander::pander(
-  overview,
-  split.tables = Inf,
-  split.cells = Inf,
-  missing = "",
-  keep.line.breaks = TRUE,
-  style = "grid",
-  justify = 'left',
-  caption = "\\label{tab:overview} Overview over all steps to create and open and reproducible research project."
-)
-```
-
-# Setup
-
-<!-- keep it brief, let `repro` do the work -->
-
-This tutorial is targeted at scholars with at least intermediate R knowledge and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md", "installation guide")`).
-To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:
-
-[^repropackage]: https://github.com/aaronpeikert/repro
-
-```r
-# -- type this on the R console --
-if(!requireNamespace("remotes"))install.packages("remotes")
-remotes::install_github("aaronpeikert/repro")
-library("repro")
-```
-
-The [`repro`-package](https://github.com/aaronpeikert/repro) contains several helpers to work with the suggested tools for reproducibility.
-For example, the package can check if your computer has all the required software packages installed and ready.
-
-```{r}
-# -- type this on the R console --
-check_git()
-check_make()
-check_docker()
-```
-
-```{r, eval=FALSE}
-check_github()
-```
-
-If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
-Please make sure that all three commands pass before you continue with the tutorial.
-This process is only required once per computer and does not need to be repeated for each project.
-
-# Planning Phase
-
-<!-- researcher should begin early to enjoy most benefits -->
-Reproducibility is invaluable for effective collaboration---including with your "past self".
-While you may be tempted to consider reproducibility only when you are close to completing a project, we believe that upfront investment pays off in your increased productivity.
-Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you achieve reproducibility gradually.
-The first step towards reproducibility is to bundle all required files in one folder.
-This way, you never lose anything, and you can easily share the files without missing anything.
-<!-- Hands-on: Introducing RProjects-->
-RStudio's _R Projects_ facilitate this concept of a project folder.
-We recomend the [`here`](https://github.com/r-lib/here)[^here] package [@R-here] to refer to files, because it works consistently across different operating systems, and between R scripts and [R Markdown].
-The `repro` package offers an R Studio template, which sets up a file structure that follows best practices and conventions for placing scripts and data [@rextensions].
-To create one, click on:
-
-[^here]: https://github.com/r-lib/here
-
-> File → New Project... → New Directory → New Project
-
-## Markdown
-
-<!--Keep notes.-->
- A great tool to capture ideas or meeting notes and eventually complete entire journal articles is Markdown.
- Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
- This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
-To create a Markdown document, click on:
-
-> File → New File → Markdown File
-
-Markdown is simply text, where some characters let you add a minimal amount of structure:
-
-```{r, results='asis', echo=FALSE}
-# in case we want to reuse this example later
-markdown_example <- "
-<!--this is a Markdown file -->
-# Header
-
-Normal text.
-Important **word**.
-
-To do list:
-
-* do research
-* do more research
-* spend time with family
-"
-cat("```markdown", markdown_example, "```", sep = "")
-```
-
-We will later see how this simplicity enables us to create many different document formats (i.e. Word, PDF, HTML) and incorporate code (i.e. R, Python, Julia).
-
-## Git
-
-Another advantage of Markdown is that because of its likeness to plain text, you can easily version-control it with Git and leverage Gits features for tracking, displaying, and merging sequential and concurrent changes.
-You may be familiar with Microsoft Words' "Track Changes"-Feature.
-Version control is similar in its basic idea but more potent in its features.
-Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
-For example, this paper has been going through `r system2("git", c("rev-list",  "--count",  "HEAD"), stdout = TRUE)` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/main", "iterations")`.
-To activate Git in a given project, you can call:
-<!-- Hands-on: Introduce Git -->
-
-```r
-# `package::function()` → use function from package without `library(package)`
-usethis::use_git()
-```
-
-Recording each change you made helps you to iterate more quickly because you know that you can effortlessly go back to previous versions.
-For example, you can write down a rough project idea, knowing your collaborators and you will iterate and improve.
-<!-- This is maybe the place to introduce the example-->
-First, you could create a new Markdown file named `idea.md` that looks like this:
-
-```markdown
-<!--this is a Markdown file -->
-
-# Hypothesis
-
-Machiavellism is higher in male persons.
-
-# Analytic Strategy
-
-Independent t-test
-
-# Sample
-
-Can we use openpsychometrics.org data?
-```
-
-And you could commit it in Git.
-For convenience, we use the [`gert`-package](https://github.com/r-lib/gert) [@R-gert], which wraps most git commands into R functions.
-Alternatively, we suggest to either work directly in the terminal (for example, RStudio offers a command shell with various options accessible in the Menu Tools → Terminal), or work using the Git pane in RStudio that offers buttons for basic operations.
-<!--I am leaning towards doing everything in R, not sure though-->
-<!--For Git we maybe should use RStudio's Git pane?-->
-
-To commit your first changes to the project, start with staging all files to be included in the commit using `git_add()`.
-Here, we stage our idea collection in the file `idea.md` and then commit it to our local repository with a meaningful commit message that will later help us keep track of all the changes over time.
-
-```r
-gert::git_add("idea.md")
-gert::git_commit("add a first concept")
-```
-
-Now assume that after discussing with your colleagues, you settle onto a final analysis strategy using latent variable models with tests for measurement invariance because power is greater and you want to exclude bias in measurement as possible confounder [@flake2020].
-This results in changes to your notes as follows:
-
-```markdown
-<!--this is a Markdown file -->
-
-# Hypothesis
-
-Machiavellism is higher in male persons.
-
-# Analytic Strategy
-
-Multigroup Confirmatory Factor Analysis with equal loadings and intercepts
-
-# Sample
-
-Can we use openpsychometrics.org data? 
-```
-
-Again, stage the changes to this file and commit like before using `git_add()` and `git_commit()` or use the following shorthand to stage all new and modified files and then commit:
-
-```r
-# -- type this on the R console --
-gert::git_commit_all("switch to CFA to exclude bias in measurement as possible confounder")
-```
-
-And GitHub can show you what has changed, like in Figure \@ref(fig:idea-change).
-This detection of what has changed works on a line-by-line basis.
-We recommend therefore to keep each sentence in a Markdown on its own line to generate clean difference views.
-GitHub is, in its essence, a web interface to Git, and you may learn more about how to use it under the section on [GitHub].
-
-```{r idea-change-screenshot, include=FALSE}
-# unfortunately css selectors won't work for pdf so we'll use png instead
-idea_change <- here::here("images", "idea-change.png")
-if(!file.exists(idea_change)){
-  webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/76f5d4e78b3d3bf22e84e9137756bc17767dff3b?diff=split",
-                  file = idea_change,
-                  selector = "div#diff-2f8beaa39e5c98706d8abb3361a8383776f8df8a195a21c0e6eeb25e5aa48f45", # only select file change
-                  zoom = 2)# higher resolution
-}
-```
-
-```{r idea-change, echo=FALSE, fig.cap="A screenshot of how a change tracked in Git is represented by GitHub.  ", out.width='100%'}
-knitr::include_graphics(idea_change)
-```
-
-<!--I am troubled by how we should recommend to learn Git-->
-Git can be challenging to learn, because it requires a rather complex mental model of maintaining potentially differing versions between different repositories (typically one central repository and one local repository per each author while each repository may have different branches).
-However, there is a subsets of commands which we present here and in the section on [GitHub] (commit, push, pull) which enable you to use Git for simple version control.
-There are many tutorials on learning Git and @vuorreCuratingResearchAssets2018 provide a good starting point.
-However Git is designed for collaboration, we therefore recommend finding a partner to attend an online lecture or workshop to learn the ins and outs of Git.
-
-## Simulation (optional)
-
-<!-- impress with neatness → sample size planning, preregistration and analysis in one 😎-->
-<!-- but explain that good science is often not as neat, strive for the ideal -->
-
-```{r, include=FALSE}
-source(here::here("R", "simulation_funs.R"))
-```
-
-```{r, include=FALSE}
-loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
-cohend_jones_paulhus <- c(24, 29, 35)/100
-```
-
-Monte Carlo simulations are a crucial tool to investigate properties of a statistical model and they are routinely employed to develop or evaluate statistical models.
-For example, @maccallum1999sample used Monte Carlo simulations to investigate optimal sample size for exploratory factor analysis, or @curran1996robustness studied the robustness of the $\chi^2$ test statistic under non-normal data.
-In Bayesian data analysis, prior predictive checks generate data from a model according to the prior distribution in order to asses whether a prior is appropriate [@gabry2019visualization].
-
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-First, you have to address and could plan for data analytic problems on a conceptual level before you made the considerable effort to acquire data [@axelrodAdvancingArtSimulation1997].
-Second, if you only simulate a single dataset, you can perform a simple technical check, that is the analysis software outputs parameter estimates [@braiekTestingMachineLearning2020], e.g.
-when the simulated data contains missings.
-Third, if you perform a Monte Carlo simulation, you may evaluate how the data analysis copes with expected data complications, e.g.
-non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
-Fourth, you can plan your resources efficiently, e.g. determine required sample size to achieve certain precision or power [@brandmaier2018precision].
-
-A Monte Carlo simulation affords you an ideal planning opportunity.
-We realize that this comes at an expensive but believe the reduction in the risk that the data is not sufficient to answer the research questions offsets the required effort.
-This advantage is especially pronounced when you plan to preregister your study.
-In fact, we argue that a Monte Carlo simulation can combine the analysis, sample size planning and preregistration (see [Preregistration as Code].
-
-No matter which analytic strategy you plan, it will come with assumptions about the data generating process under investigation. 
-A simulation forces us to deal with these assumptions and their consequences explicitly.
-If you identify unrealistic assumptions, you can either deliberately violate these assumptions and, by simulation, check if the analytic strategy is robust or modify the modeling approach to explicitly address them.
-
-Typically, you have to make a number of educated guesses to simulate data that closely matches your expectation.
-These guesses strongly depend on the study you are planning, and there cannot be a one-fits-all solution.
-Since we typically use parametric statistical models (general linear model or SEM) or hypothesis tests that can be expressed as Gaussian models, we can usually use these models directly to simulate data.
-We will show an example using structural equation modeling (SEM) for our Machiavellism example.
-Note that the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
-
-If you are unfamiliar with the specifics of SEM, you may briefly skim the next sections and continue in section ["Random Number Generation"](#prng).
-To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.
-
-The selected model parameters should express our expectations for the Machiavellism subscale of the short dark triad [SD3, @Jones2013].
-
-As a starting point, we can use already published information.
-@Jones2013 report a set of factor loadings from an Exploratory Structural Equation Model [@asparouhovExploratoryStructuralEquation2009].
-The estimates for the construction sample, which was recruited from Amazon’s Mechanical Turk (MTurk) and consists of 489 adults from Canada and the United States, can be seen in Table \@ref(tab:loadings))
-They report a gender difference (at the manifest level), which depending on the study ranges between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.
-
-```{r loadings, echo=FALSE}
-knitr::kable(t(data.frame(Item = as.character(1:9),
-                        Loading = loadings_jones_paulhus)),
-             caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
-```
-
-Typically, it is recommended to choose conservative values in the study planning phase [@brandmaier2018precision].
-This way, we make sure that our analytic strategy is viable even in non-optimal circumstances, for example if the measurement instrument works less well then expected or the population characteristics are more different than expected.
-As a heuristic guess, we assume that the factor loadings are 30% lower than those reported by @Jones2013 and that the standardized mean difference is only 0.2.
-These assumption can be simply translated into [`lavaan` [@R-lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:
-
-```{r}
-model_truth <-
-"MACH =~ 0.27*x1 + 0.22*x2 + 0.28*x3 + 0.36*x4 + 0.41*x5 + 0.5*x6 + 0.43*x7 +
-0.32*x8 + 0.36*x9
-MACH ~ c(0, 0.2)*1
-MACH ~~ c(1, 1)*MACH"
-```
-
-[^lavaan]: https://lavaan.ugent.be
-
-Data simulation is particularly easy in lavaan syntax using the function `simulateData`.
-More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem [@simsem], faux [@faux], or simstudy [@goldfeldSimstudyIlluminatingResearch2020].
-The following code simulates data from the population model `model_truth` with 1000 cases in each group.
-
-```{r, include=FALSE}
-set.seed(1234)
-```
-
-```{r}
-simulated <- lavaan::simulateData(
-  model_truth,
-  sample.nobs = c(1000, 1000),
-  meanstructure = TRUE,
-  standardized = TRUE
-)
-```
-
-We propose the following procedure for analyzing the data and dealing with possible violations of strong measurement invariance. 
-After fitting the model with strong measurement invariance [@flake2020], we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
-The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
-While reconsidering our modeling assumptions, it is also unlikely that the data are perfectly normally distributed.
-First, we have rating scale items and thus our observations are categorical.
-Second, @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
-For simplicity of our illustration, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).
-
-If you are satisfied with your assumptions for the data simulation, it helps to write a function that expresses them, like this one:
-<!--Hands-on: build simple functions tailored for your analysis-->
-
-```{r generate_data}
-```
-
-And to build another function that captures your planned analysis.
-Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R", "GitHub")`.
-
-```{r planed_analysis}
-```
-
-```{r extract_results}
-```
-
-With its help we can simulate 1000 cases and fit our analysis.
-
-```{r}
-generate_data(1000) %>% 
-  planed_analysis() %>% 
-  extract_results()
-```
-
-Note, that this certainly results in different numbers each time we run it (see ["Random Number Generation"](#prng)).
-Of course it is already helpful to simulate a single dataset to test if the analysis works in principle, but repeating the analysis several times and vary parameters lets you verify more properties of your analysis.
-First, you can check if the planned analysis recovers the simulated values (`std_estimate` should be close to the simulated mean difference of 0.2).
-Second, it enables you to plan an appropriate sample size because you are able to quantify the sampling variability.
-
-To complete a such a  Monte-Carlo simulation study we need a function that repeats the analysis for different conditions (here only for different n):
-
-```{r simulation_study}
-```
-
-Such Monte-Carlo simulation can help to decide between several trade-offs.
-Historically, one important trade-off is between sample size and power (the probability to detect an effect of given size if it exists).
-However, the question can be put more broadly and it is possible to consider various other trade-offs of relevant study design parameters other than sample size, such as number of measurement occasions [e.g., @brandmaier2015lifespan], reliability of measures [e.g., @brandmaier2018precision], or parameters related to the structure of planned-missing data design [@brandmaier2020optimal].
-To show that there can also be alternative approaches for sample size planning other than those that are restricted to null hypothesis testing, we choose to visualize the precision of estimation as the density distribution of the parameter of interest (latent mean difference by gender of Machiavellism) instead of statistical power.
-Such process still allow us to formulate expectations for the expected precision of the chosen analysis.
-For example, with a sample size of 1,000 per group, 80% of the simulated standardized effect sizes fall into a range of +/- .1 standardized difference points.
-This claim is useful and informative irrespective of whether we embark on null hypothesis testing later on or not.
-The R code to reproduce the simulation and the figure can be found in the accompanying [GitHub repository](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo]:.
-
-[^gh-repo]: https://github.com/aaronpeikert/repro-tutorial
-
-```{r, include=FALSE}
-res <- read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddd")
-```
-
-```{r hash}
-```
-
-```{r, include=FALSE}
-res_interval <- res %>%
-  select(n_obs, std_estimate) %>% 
-  filter(!is.na(std_estimate)) %>% 
-  group_by(n_obs) %>% 
-  summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
-            .groups = "drop", median = median(std_estimate)) %>% 
-  unnest(c(interval)) %>% 
-  mutate(width = upper - lower, perc = 1 - alpha) %>% 
-  group_by(alpha) %>% 
-  arrange(n_obs) %>% 
-  mutate(across(c(width, upper, lower),
-                ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE))) %>% 
-  ungroup() %>% 
-  arrange(perc)
-```
-
-```{r, include=FALSE}
-choosen_perc <- .8
-choosen_width <- .2
-true_diff <- .2
-est_diff <- res %>% filter(n_obs == max(n_obs)) %>% pull(std_estimate) %>% mean()
-minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
-  filter(n_obs == min(n_obs))
-sim_n <- pull(res, n_obs)
-sim_nmax <- max(sim_n)
-sim_n_unique <- unique(sim_n)
-sim_nstep <- sim_n_unique[2] - sim_n_unique[1]
-```
-
-```{r, include=FALSE}
-plot_interval <- res_interval %>%
-  ggplot(aes(
-    n_obs,
-    ymin = lower,
-    ymax = upper,
-    fill = perc,
-    group = alpha
-  )) +
-  geom_ribbon() +
-  geom_line(aes(y = median), color = "white") +
-  geom_hline(yintercept = true_diff, linetype = "dotted") +
-  scale_fill_viridis_c(
-    begin = 1,
-    end = 0,
-    breaks  = 1 - seq(0.05, .2, 0.05),
-    labels = . %>% scales::percent(accuracy = 1)
-  ) +
-  scale_y_continuous(n.breaks = 10) +
-  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
-  theme_minimal() +
-  labs(x = "N per Group", y = "Standardized Effect Size") +
-  theme(legend.title = element_blank()) +
-  NULL
-```
-
-```{r, include=FALSE}
-plot_width <- res_interval %>%
-  ggplot(aes(n_obs, width, color = perc, group = perc)) +
-  geom_line() +
-  geom_point(
-    data = minn,
-    size = 3,
-    color = "black",
-    shape = 3
-  ) +
-  scale_color_viridis_c(
-    begin = 1,
-    end = 0,
-    breaks = 1 - seq(0.05, .2, 0.05),
-    labels = . %>% scales::percent(accuracy = 1)
-  ) +
-  theme_minimal() +
-  scale_y_log10(breaks = seq(0, 0.8, .1)) +
-  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
-  labs(x = "N per Group", y = "Width of quantile\n(Standardized Effect Size)") +
-  theme(legend.title = element_blank()) +
-  NULL
-```
-
-```{r, include=FALSE}
-simulation_plot <- plot_interval +
-  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + 
-  plot_width +
-  plot_layout(nrow = 2, guides = 'collect') +
-  plot_annotation(tag_levels = 'A')
-```
-
-```{r, simulation-plot, echo=FALSE, fig.cap="*Plot A* depicts the range of effect sizes we observed in the simulation, where the colour encodes the density.
-The light yellow interval shows where 80% of the estimates land.
-The dotted line represents the simulated true effect size; the white line shows the median observed effect sizes.
-*Plot B* represents the width of the interval shown in Plot A on a logarithmic scale.
-The cross shows the minimum required sample size so that 80% of the estimates land in a range of +/- .1.
-"}
-simulation_plot
-```
-
-### Random Number Generation (optional) {#prng}
-
-Strict reproducibility is at stake with any code that is based on the computer's pseudo random number generator (PRNG) because random numbers will almost always differ with each execution of the code.
-To consistently get the same results while preserving the pseudo-random property of the generated numbers, it is good practice to set a random seed that sets the PRNG into a restorable state, in R code: `set.seed(1235)`.
-Note however that the consistency of a given seed is not necessarily maintained across different R versions because the default algorithm for generating pseudo random numbers may change (see @Peikert2019 for a demonstration with different R versions).
-It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
-The standard random number generator (RNG) produces high quality random numbers, but is not safe to use in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
-The problem is that concurrent random number streams are correlated, because they repeat after some time.
-Therefore, if you set a seed for one worker and another seed for another worker, you could end up with the same random numbers but with a phase shift.
-@lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
-While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers.
-
-<!--
-```{r}
-# reproducible but unsafe in parallel code
-set.seed(1235)
-runif(1)
-set.seed(1235)
-runif(1)
-```
--->
-By wrapping code that uses the internal random number generator (here, the command to create a random number from the uniform distribution, `runif(1)`) with the `future` approach, we guarantee that the random numbers will differ across parallel executions of that code, will achieve a high quality of randomness, but still will be deterministic and thus reproducible:
-
-```{r}
-# reproducible and safe in parallel code
-future::value(future::future(runif(1), seed = 1235))
-```
-
-[^futureverse]: https://www.futureverse.org
-
-## Preregistration (optional)
-
-<!--Hands-on: build simple functions tailored for your analysis-->
-
-Preregistration is the specification of a research plan in advance of data analysis.
-The goal of preregistration is transparency and a clear-cut divide into exploration and confirmation.
-Thus, it serves as a tool that may help to increase the credibility of empirical results.
-Without getting into too much detail of the meta-scientific arguments for preregistration, we want to note two advantages [@nosekPreregistrationHardWorthwhile2019].
-First, a preregistration is an opportunity to plan a study carefully.
-Gathering empirical data requires many resources, and better planning increases the chances that these efforts are worthwhile.
-For us, a good preregistration is the same as a good plan; the only difference is that a preregistration is made public beforehand.
-
-The second advantage is transparency.
-A preregistration offers the interested reader more insight into the research process than a paper alone could.
-@szollosi_is_2020 argue that preregistration can not fulfil its promise to reduce the false discovery rate of science if we allow deviating from them.
-While such an argument has some merits, we believe that often transparency outweighs these concerns [@nosekPreregistrationHardWorthwhile2019].
-Deviation from a preregistration is still possible.
-If researchers deviate from their preregistration, they probably have good reason to do so, which means they have learned something from the process of conducting the study, and any deviation will simply be explicitly marked as such.
-A preregistration, together with version control, allows their readers to inform themselves about these lessons learned.
-
-### Preregistration as Code
-
-Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration [@osfstandard; @socpsytemp; @howtoreplicate]. 
-Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code [@sep-ambiguity].
-Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity.
-PAC essentially reverses the process of empirical research.
-Before researchers gather data, they write a reproducible, dynamically generated manuscript (see [R Markdown]) including the sections Introduction, Methods, and Results---initially based on simulated data.
-Then, the source code of this manuscript serves as an unambiguous description of the analysis plan and is published as a preregistration.
-One particular advantage of this approach is that you can precisely define if-else conditions under which you choose a different analytic strategy to address possible complications (e.g.
-to establish partial measurement invariance, when full measurement invariance is not attainable).
-This approach allows for more flexibility than a traditional template based preregistration but is still well defined.
-
-PAC works well within this reproducible framework.
-
-First, the preregistration becomes a part of the reproducible repository.
-It is part of the set of version-controlled files and provides transparency about when the preregistration file, data files, and code files were changed over the course of a research project.
-If code was version-controlled and tagged (see Section [Git]) before data collection, all deviations from the preregistration or any post-hoc additions that were performed after data collection are transparent and are visible and comparable in the version history.
-
-One way to ensure that your analysis to works with real data is to simulate data.
-We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
-The simulation showed that we have to gather at least `r minn$n_obs` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `r round((est_diff - true_diff)*100/true_diff, 2)`%.
-These results are also visualized in Figure \@ref(fig:simulation-plot).
-If you went to the trouble to simulate data, we argue that you get a preregistration almost for free.
-
-A preregistration in code describes sufficiently how you want to analyze the data, but ideally, you also convey the why.
-Dynamic document generation combines code and text, enabling you to explain your reasoning.
-See the section on [R Markdown] for an example of a dynamic document generation paradigm.
-In fact, this manuscript and therefore the here presented Machiavellism example were preregistered as a full manuscript, and you can see the results on real data in section [Reproduce a Project].
-
-### R Markdown
-
-<!--Hands on: Introduce R Markdown-->
-As @knuthCWEBSystemStructured states: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
-
-Above, we have already suggested [Markdown] as a universal and portable text format. 
-R Markdown extends regular Markdown by allowing useres you to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
-Upon rendering of the document, the code blocks are executed and their output is dynamically inserted into the document.
-This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
-
-To create an R Markdown document, click on:
-
-> File → New File → R Markdown...
-
-The code is separated from the text by three backticks, like here:
-
-````md
-```{r}`r ''`
-set.seed(1235)
-runif(1)
-```
-````
-
-When you include code into an R Markdown, this code is run, and its results are included in the rendered document.
-With R Markdown, you can write whole articles (`r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/manuscript.Rmd", "like the one you are reading")`).
-You could, in principle, even write large parts of the manuscript as part of your preregistration, i.e. the introduction, theoretical background, the method section and parts of the discussion.
-If you have simulated data, you could even include fake results.
-Packages like [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf)[^stargazer] [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html)[^apaTables] [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/)[^papaja] [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report)[^report] [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
-Such a dynamic document is arguably the most concrete preregistration possible.
-One advantage is that writing the manuscript and writing the preregistration are no longer two distinct tasks.
-Therefore a preregistration can be thought of as simply rearranging tasks that you have to do anyway.
-
-[^report]: https://github.com/easystats/report
-[^stargazer]: https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf
-[^apaTables]: https://dstanley4.github.io/apaTables/articles/apaTables.html
-[^papaja]: http://frederikaust.com/papaja_man/
-<!--Hands on: gitignore resulting .html/pdf-->
-<!-- A preregistration needs to be public -->
-
-### Preregistration on GitHub
-
-An essential aspect of a preregistration is that you have to make it public.
-Of course, if you commit the preregistration to Git and upload it into a public repository, it is also public.
-However, simply using GitHub as a preregistration site is not ideal.
-First, you can delete your GitHub repository whenever you want.
-Second, while Git meticulously records when you made which change, committing and publishing are two different steps.
-Therefore, you can choose to upload to GitHub after conducting the study while still making it look like you have made the preregistration public before conducting the study.
-Third, any person with some experience in Git can retroactively change the history.
-You can alleviate these downsides by using zenodo.org to give a digital object identifier (DOI) to your preregistration because zenondo.org saves an unchangeable copy of your repository.
-You can learn more about this in [Creating a public release].
-
-## Prepare public release (optional)
-
-An open research project usually entails that you publish content to the public web, without an intermediate publisher.
-Since you are directly responsible for such content, you have to prepare a few things that are usually done by the publisher for you.
-
-### README (optional)
-
-One of the most significant benefits of open research is that other people can freely reuse materials, ideally including everything from stimuli, over computerized experiments, to data analysis code.
-To use your materials, they first have to find them and learn how to use them.
-Interested users find out about your project via the readme.
-Historically[^readme-age], a readme file is the default place to look for human readable information in a computer archive 
-the first thing they see when they visit your project on [GitHub].
-It is supposed to explain to visitors what your project is about and how to reproduce it.
-You could, for example, include the abstract of your (planned) paper here.
-
-[^readme-age]: Apparently, the use of readme files can be traced back to at least 1974 as part of UCI LISP in this file: http://pdp-10.trailing-edge.com/decuslib10-04/01/43,50322/read.me.html
-
-To add a readme to a repository run:
-
-<!--Hands on: Add a readme-->
-```r
-usethis::use_readme_rmd()
-```
-
-This command will create a file with the name `README.Rmd`.
-<!--knit/rmarkdown will be explained in section above-->
-When you knit this file, it, in turn, creates a `README.md`.
-That means you can include R code or plots etc., within the readme.
-The `README.md` file will look pleasant in the browser on [GitHub](https://github.com/aaronpeikert/repro-tutorial#readme).
-
-### License
-
-Note that publishing data, code, or other materials is not sufficient to allow other users to use it, modify it, or redistribute it.
-For example, according to German copyright law, a repository without a license must be assumed to be protected by copyright ([UrhG[^urhg]](https://www.gesetze-im-internet.de/englisch_urhg/englisch_urhg.html), Section 1, 7, 11).
-In many countries (i.e. the USA, European Union), a content creator has exclusive right to use their creation[@stoddenLegalFrameworkReproducible2009].
-A license is the legal authorization of other people to use what you have created.
-<!---exact copy from our previous manuscript, do we need to change? -->
-In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media (such as stimulus materials if you created them).
-In contrast, scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
-Copyright does often not apply to data, we therefore recommend to share data with a CC0 license to clarify that you do not hold copyright to it.
-Giving data another license than CC0 is often just confusing, because a license is invalid if there is no copyright.
-All licenses assure maximal freedom for future users while requiring the attribution, if possible, of the original authors in derivative work.
-These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
-A great resource to choose a license is [choosealicense.com](https://choosealicense.com).
-However, no resource, including our recommendation, replaces legal advice.
-<!-- plagiary end -->
-It is important to get written consent from collaborators when you change/add the license.
-However, if their first contribution occurs after adding a license, they give their consent by contributing.
-To add a CC-BY license to your project, simply run (there are similar commands for other licenses):
-
-<!--Hands on: Add a license-->
-```r
-usethis::use_ccby_license()
-```
-
-### Data Documentation
-
-[@codebook]
-
-### Code of Conduct
-
-We strongly suggest that you strive for respectful interactions with anyone who might want to contribute.
-We hope it comes naturally to you.
-To communicate that contributors can expect a welcoming and respectful environment, you can add the `r link( "https://www.contributor-covenant.org", "Contributor Covenant")` to your repository.
-Contributing to an open project is daunting, and a public pledge to value any contributions (even when you do not incorporate them) can help others to take the first step.
-The Contributor Covenant is a code of conduct used by many open source project.
-
-To add the Contributor Covenant to your project use:
-
-<!--Hands on: Add a code of conduct-->
-```r
-usethis::use_code_of_conduct()
-```
-
-However, there is some controversy around how to handle violations of a code of conduct.
-For that or any other reason, you may want to change the Contributor Covenant to fit your project's needs (which you are allowed to do because it is published with a CC-BY license).
-You can also consider alternative Codes of Conduct, like the `r link("https://opensource.microsoft.com/codeofconduct/", "Microsoft Open Source Code of Conduct")` or the `r link("https://berlincodeofconduct.org", "Berlin Code of Conduct")`.
-
-If you have considered what you want to tell the public about the project with a readme, have chosen a license and thought about how to make collaboration welcoming, you may publish your Git repository.
-See the section on [GitHub] on how to publish your project.
-
-## Collaboration
-
-Often, reproducibility is reduced to the merits it has for the reader.
-However, we argue that reproducibility is vital for collaboration, not only in the sense that the consumer of a scientific product builds upon it but also that the creators work together.
-The time that you invest into ensuring reproducibility is directed at ensuring effective collaboration.
-Therefore, reproducibility is at its core about collaboration.
-
-### GitHub
-
-On your local machine, you use Git to track changes.
-Online services like GitHub allow you to share these changes with others and collaborate with them.
-GitHub is not the only solution; there is also [GitLab](gitlab.com) or [BitBucket](https://bitbucket.org), which are as feature-rich as GitHub.
-GitLab has the advantage that your institution can host it themselves; therefore, you are not relying on commercial service providers.
-Anyhow, we focus on GitHub because it is the most popular service.
-To upload your changes to GitHub, you need to create an account under https://github.com/join.
-As a researcher, you are eligible for the `r link("https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount", "Researcher/Educator discount")`.
-The following command creates a new GitHub repository under your GitHub account (beware this is public):
-<!--Hands-on: Introduce Github-->
-
-```r
-usethis::use_github()
-```
-
-To create a private repository (only you and people you invite have access), use:
-
-```r
-usethis::use_github(private = TRUE)
-```
-
-Only from this moment on files are uploaded to the internet and only those you have explicitly added and committed in Git.
-
-After you have initially uploaded your repository to GitHub, you need to regularly synchronize your local copy and GitHub.
-Uploading commits to GitHub is called "push":
-
-```r
-gert::git_push()
-```
-
-Downloading changes, e.g. from your collaborators is called "pull".
-
-```r
-gert::git_pull()
-```
-
-### GitHub Issues
-
-<!--Hands on: Introduce GitHub Issues-->
-Effective collaboration needs effective organization.
-GitHub offers a ticket system called "Issues".
-Here you and your collaborators can discuss and distribute tasks.
-
-To take a look at the issue page of your project, you may use:
-
-```r
-usethis::browse_github_issues()
-```
-
-```{r issue-screenshot, include=FALSE}
-# unfortunately css selectors won't work for pdf so we'll use png instead
-issue <- here::here("images", "issue.png")
-if(!file.exists(issue)){
-  webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/issues/20",
-                  file = issue,
-                  selector = "div#repo-content-pjax-container", # only select file change
-                  zoom = 2)# higher resolution
-}
-```
-
-```{r issue, echo=FALSE, fig.cap="A screenshot of an example issue.  ", out.width='100%'}
-knitr::include_graphics(issue)
-```
-
-One advantage of using the issue system of GitHub is that it enables you to link changes you made to their tasks by including the issue number (i.e. #20) in the commit message:
-
-```r
-# do not run:
-gert::git_add(c("manuscript.Rmd", "references.bib"))
-gert::git_commit("explain the basics of issues #20")
-```
-
-You see the linkage in Figure \@ref(fig:issue).
-
-While it is a valuable device for working with direct collaborators, it is also great to facilitate review [@rougierSustainableComputationalScience2017].
-Even if a journal does not support this, it enables post-publication review.
-Readers of your manuscript can ask questions or point out how to improve it after you have published it.
-
-### GitHub Pull Requests
-
-<!--Hands on: Introduce GitHub PR and review-->
-There are various ways, in which other scholars may interact with your published repository.
-For example, in our exposition of the workflow for reproducibility [@Peikert2019], once we had published our preprint, scholars previously unknown to us started to send us spelling corrections.
-When multiple persons start modifying a project in parallel, keeping track of the edits and how they are supposed to be merged is crucial.
-Clearly, allowing everyone to edit the project is difficult.
-
-_Pull requests_ allow anyone to contribute to a project, even though they were not granted write access, while the authors still retain full control over all changes.
-If someone wants to contribute, but they have no permission to write to a repository, they can create a copy of the project, called "fork", where they can push their changes.
-For example if you found an error in this manuscript, you can not simply change it, because you have no write access.
-However, you can create a fork in your account, where you have write permission:
-
-```r
-usethis::create_from_github("aaronpeikert/repro-tutorial", fork = TRUE)
-```
-In this copy you then make any changes you want.
-You could then request that we (or anyone with write permissions) "pull" these changes.
-You can view new pull request for any project with:
-
-```r
-usethis::browse_github_pulls()
-```
-
-Before incorporating changes from someone else, they can be reviewed and adapted.
-This review step is particularly useful for remote collaboration, because collaborators can be explicitly pointed to changes and asked to approve them.
-Therefore, it is sometimes useful to create a pull request even though you have permission to push, because it invites your collaborators to comment on, change, or approve proposed changes.
-
-You can initiate a new pull request with:
-
-```r
-usethis::pr_init("myrevision")
-```
-
-Then you make and commit your changes and open a new pull request with:
-
-```r
-usethis::pr_push()
-```
-
-After you or someone else has merged the pull request on GitHub, you can wrap it up with:
-
-```r
-usethis::pr_finish()
-```
-
-This process can be executed via Github's web interface and can work well with collaborators who do not actually want to bother with installing and running the entire code base of a project.
-
-# The Data Analysis Phase
-
-So far, we have covered the generation of dynamic documents with [Markdown] and [R Markdown], and tracking changes with [Git] and [GitHub].
-These are two of the four components [see @Peikert2019] to achieve long-term reproducibility.
-To complete computational reroducibility, researchers need to properly handle dependency relations between their files (internal dependencies) and all dependencies on external software, most importantly dependend R packages (external dependencies).
-In the following, we will show how the `repro` package supports researchers in mapping all dependencies completely and unambiguously.
-
-## Internal Dependencies
-
-<!--I am not sure this is the right place, should we introduce this earlier? -->
-
-A reader of a reproducible research project should be able to easily derive how it can be reproduced (ideally by a statement in your [README]).
-To do that, it is often crucial to understand how different project components relate to each other and depend on each other.
-For example, you cannot run an analysis without data, or you cannot render a statistical report without having the results from the statistical analysis that is reported.
-Dependency tracking means that you capture and map the interrelations between all components of your analysis.
-The "components" of an analysis are typically all involved files including raw data files, intermediate files (such as a preprocessed data file), and final results (such as a rendered manuscript in PDF format).
-With larger projects, there is typically a higher degree of modularization because code fragments are written for re-use in other projects (or are already re-used).
-Also, more complicated empirical research projects may have different hypotheses that are answered with different data sets and different statistical models, such that it often makes sense to split the different analyses into different files.
-Reproducibility then demands the careful documentation and mapping of all necessary steps in the correct order.
-
-Make is a build automation tool for software projects that typically are challenged by maintaing and resolving a variety of dependencies in the correct order when compiling a piece of software-
-We found _Make_ to be particularly useful for our purposes for two reasons.
-First, it is language-agnostic, therefore enabling us to map steps not only taken within R but also across different programming languages or other external programs.
-Second, it does implement a simple yet powerful caching mechanism of results which is welcome in computing-intensive analysis often found in neuroimaging or machine learning. 
-That is, Make determines automatically which branches in the dependency graph needs updating when files have changed and does not necessarily recompute or rebuild everything.
-
-Make is build around the concept of computational recipes.
-Like cooking recipes, the dish you want to create is at the top, followed by the necessary ingredients with the procedure at last.
-In Make the dish is usually a file and is called "target".
-The ingredients are called "requirements" and are usually also files.
-The procedures are formulated as shell commands.
-One (hypothetical) example is this:
-
-```
-spaghetti_arrabiata.pdf: spaghetti_arrabiata.Rmd arrabiata_sauce.csv pasta.csv
-  Rscript -e 'rmarkdown::render("spagetti_arrabiata.Rmd")'
-```
-
-An important property is that each requirement can be itself be a target, like here:
-
-```
-spaghetti_arrabiata.pdf: spagetti_arrabiata.Rmd arrabiata_sauce.csv pasta.csv
-  Rscript -e 'rmarkdown::render("spaghetti_arrabiata.Rmd")'
-
-pasta.csv: cook_pasta.R
-  Rscript -e 'source("cook_pasta.R")'
-
-arrabiata_sauce.csv: cook_sauce.R canned_tomatoes.csv
-  Rscript -e 'source("cook_sauce.R")'
-```
-
-If a target file already exists, Make only recreates it if necessary.
-If a dependency is newer than its target, make recreates the target to reflect the updated dependency.
-In our case, when we (re)build a manuscript and some dependent data file has changed, the data analysis is automatically rerun.
-This avoids mistakes in the reporting that are due to reports and analyses being out-of-sync.
-
-<!--
-AP: I suggest to remove this simply because it distracts from our primary tool and we need to shorten
-One alternative that is even smarter (in terms of automatically inferring dependencies) than Make is the R package [`targets` [@targets2021]](https://books.ropensci.org/targets/)[^targets].
-`targets` can infer more about your project than Make, because it analyzes the R code you write to build the dependency tree.
-For this inference to work well, you have to write everything in R, and you have to write everything strictly functional.
-
-[^targets]: https://books.ropensci.org/targets/
--->
-
-<!--Hands on: Introduce Make-->
-<!--Hands on: Add targets for README.Rmd, preregistration.Rmd -->
-
-Writing Makefiles manually is cumbersome for larger projects. 
-Therefore, the `repro` package provides means to automatically write Makefiles for a given R project.
-To this end, `repro` extends the YAML metadata header of R Markdown files.
-`repro` introduces a new first-level tag `repro:` with second-level tags (`scripts:`, `packages:`, and `data:`) that take lists of dependent files.
-For example, if an R Markdown document depends on a further file with R code, the header would look like this:
-
-```
----
-title: "Preregistration"
-author: "Aaron Peikert"
-date: "4/12/2021"
-output: html_document
-repro:
-  scripts:
-    - R/simulation.R
----
-```
-
-To automatically generate the Makefile, `repro` provides a command that goes through all R Markdown files in the project, parses their YAML metadata, and infers a `Makefile` from them:
-
-```r
-automate_make()
-```
-
-The generated `Makefile` automatically contains separate targets for each R Markdown documents.
-The names of the Makefile targets are identical to the Markdown file names.
-For example, to generate the Markdown file `README.md` from the R Markdown `README.Rmd`, you can run:
-
-```bash
-# this should be run in the terminal
-make README.md
-```
-
-It is common practice to name the very first target in a Makefile "all" and have it include all important files of the project.
-If you do not specify a target when calling Make, Make defaults to this first target.
-`repro` reminds you of that common practice and suggests which file you might want to add to the target `all`.
-Thus, if researchers who try to reproduce a given project and are unaware of its potentially various contents, will always build all reproducible artifacts of a repository when calling Make without further arguments:
-
-```bash
-# this should be run in the terminal
-make
-```
-
-Running this command fully reproduces the entire project -- but only if you have all required software installed.
-Next, we will describe how `repro` helpts to manage all software dependencies automatically.
-
-## External Dependencies
-
-```{r, include=FALSE}
-container_size <- as.numeric(fs::file_size("reprotutorial.sif"))/1024^3
-```
-
-Modern data analysis is only feasible because researchers rely on much external software. 
-The dependency on a single R package may easily create a dependency tree of more than hundred direct or indirect dependencies [@epskamp2019].
-External dependencies are unavoidable but, at the same time, a major threat to long-term reproducibility.
-The critical problem is the fragility of a software setup.
-Many components have to fit perfectly together, and each change, e.g. a software update or bug fixe, risks that balance.
-At the same time, software updates are very much necessary to remove errors or improve methodology [@epskamp2019].
-Another problem is that the software environment is also variable and what works on one machine might not work on any other.
-Containers directly address such issues.
-Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself.
-The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
-Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
-Such a snapshot of the software stack is called an "image".
-Packaging all needed software in such an image requires considerable amounts of storage space.
-There are two major strategies helping to keep the storage requirements reasonable.
-First, there is a community that maintains images for particular purposes.
-For example, there are ready-made images for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
-Users can then install whatever they additionally need on top of these pre-made images.
-`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
-Our image includes Ubuntu, R, RStudio, LaTeX as well as many R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
-Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
-This recipe is a textual description of all commands that need to be executed to recreate the image.
-Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
-However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
-For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
-With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`
-
-```
----
-title: "Preregistration"
-author: "Aaron Peikert"
-date: "4/12/2021"
-output: html_document
-repro:
-  scripts:
-    - R/simulation.R
-  packages:
-    - tidyverse
-    - here
-    - lavaan
-    - furrr
----
-```
-<!--AP emphasize more how cool it is that this stuff is automatically generated by repro!-->
-With the R version that you currently use, this specification is then translated into a `Dockerfile` that makes sure that the container image has all these packages installed.
-The `Dockerfile` also contains the current date and fixes packages versions to the version of the package that were available on that day on the [Comprehensive R Archive Network (CRAN)](https://cran.r-project.org)[^cran] by relying on the [CRAN timemachine of Microsoft R Application Network (MRAN)](https://mran.microsoft.com/timemachine)[^mran].
-<!-- Hands on: Introduce Docker -->
-
-```r
-automate_docker()
-```
-
-[^cran]: https://cran.r-project.org
-[^mran]: https://mran.microsoft.com/timemachine
-
-From the `Dockerfile` you can then build an image.
-While you can do so manually, we suggest to again use Make ( see [Internal Dependencies]).
-The `Makefile` built by `repro` does contain a recipe for creating the `Dockerfile`.
-
-```bash
-make docker
-```
-
-Importantly, the `Makefile` created by `repro` includes a mechanism to execute commands in the Docker container instead of locally.
-This mechanism is triggered by the argument `DOCKER=TRUE` appended to any make call. 
-The following call builds all targets inside the container instead of in the local computing environment:
-
-```bash
-# this should be run in the terminal
-make DOCKER=TRUE
-```
-
-Last, to create or update the container and all Make targets, `repro` provides a single function `automate()` that calls `automate_docker()` and `automate_make()`.
-
-You can also manually edit the `Dockerfile` by changing the file `.repro/Dockerfile_manual` to install software other than R packages.
-Remember to run `repro::automate()` every time you made edits to `.repro/Dockerfile_manual`.
-We recommend @wiebelsLeveragingContainersReproducible2021 as a starting point for researcher who want to better understand how they can employ Docker for research.
-
-<!-- after this everything is optional? -->
-
-### renv (alternative)
-
-Setting up and running Docker is not always painless, because it requires a complicated mental model of the underlying virtual computer and, specifically, knowledge about Linux environments.
-An R-focused alternative is the R package [`renv`](https://rstudio.github.io/renv/articles/renv.html)[^renv][@R-renv].
-`renv` isolates all R packages used in a project and records there version as well as the source from where they where installed.
-This mechanism is faster and simpler to use than Docker, at the expense of only tracking R packages.
-`renv` therefore is already of great use for reproducibility [@vanlissa2020worcs].
-However, even when your project relies solely on R, external dependencies can still change the outcome of computations, like the operating system, the compilers used, the BLAS (Basic Linear Algebra Subprograms, responsible for carrying out many computations like matrix multiplication needed in the linear regression) system in use and so forth.
-Most importantly the R version itself has to match.
-For example, @Peikert2019 illustrate how the generation of random numbers across different R versions is a threat to reproducibility.
-
-[^renv]: https://rstudio.github.io/renv/articles/renv.html
-
-### High-performance computing environments (optional)
-
-The `repro` package is particularly well-suited for computationally intensive data analyses and simulations that  are to run on distributed computing environments.
-Through the use of containers, `repro` ensures identical execution across platforms, guaranteeing that a given data analysis or simulation designed and tested on a local machine will execute identically on a remote machine or multiple different remote machines across an entire computing cluster.
-Thereby, the repro package makes massive parallel computing in heterogeneous computing environments less error-prone while ensuring the full documentation of all computational steps taken — even if some or all of them were run in parallel on distributed platforms.
-The sheer volume of data often necessitates distributed computing, e.g.
-the European Bioinformatics Institute stores more than 120 Petabyte of publicly accessible data [@cookEuropeanBioinformaticsInstitute2018], single FMRI studies quickly amount to hundreds of Gigabyte[^openneuro] and even purely behavioral data like those from the moral machine experiment [>10GB, @awadMoralMachineExperiment2018] can become too large to be handled by a single local computer.
-
-[^openneuro]: https://openneuro.org/public/datasets
-
-As default container solution, `repro` uses Docker, which is a widely adopted standard for cloud computing.
-However, Docker requires administrator rights for execution.
-Whereas this is no issue on most modern cloud computing infrastructures, it is often an issue in classic HPC environments.
-Singularity is an alternative solution [@singularity].
-It was designed to run without administrator rights, while maintaining full compatibility with Docker images.
-
-If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
-Importantly, this choice does not affect how code is executed inside the container.
-Like with Docker, the default `Makefile` generated by `repro` comes with a mechanism to automatically send all commands to Singularity.
-To reproduce an analysis with Singularity, simple add the parameter `SINGULARITY=TRUE` as follows:
-
-```bash
-make SINGULARITY=TRUE
-```
-
-<!-- Hands on: add data via automate() -->
-<!-- Hands on: automatically download data via manually editing the Makefile -->
-
-## Reproducing a Project
-
-A reproducible project should not only be reproducible in theory, but it should be simple to reproduce.
-Typically, it is argued that results should be reproducible by a Ph.D. student from the same discipline who had experience with the statistical software used [@obels2020].
-Even if all code and data are available in a repository, if reproducibility takes too long because of a lack of structure and documentation, it must be regarded as non-reproducible [@obels2020].
-
-Ideally, you have a statement in the README that contains a short description of how to reproduce the project, whereas most steps are automatic and standardized.
-This statement should also list all software requirements.
-For example, you can download any git repository with these commands:
-
-```bash
-# this should be run in the terminal
-git clone https://github.com/aaronpeikert/repro-tutorial.git
-cd repro-tutorial
-```
-
-Projects that combine Make and Docker and adhere to the standard proposed by @Peikert2019 can then be reproduced with:
-
-```bash
-# this should be run in the terminal
-make docker &&
-make -B DOCKER=TRUE 
-```
-
-Typically there is a need to repeatedly reproduce a project during the development phase (e.g., on every change of  the analysis cod or, every change to the dataset).
-The `repro`-package contains a function to reproduce all targets:
-
-```{r}
-repro::rerun()
-```
-
-This function automatically figures out how to reproduce an entire project and also recognizes non-default reproducibility tools (like `renv` and `singularity`). 
-Of course, this automatic derivation of tools and steps is limited to the packages we have discussed here and the program may not always succeed if researchers deviate from our suggestions.
-
-If you have written a [Preregistration as Code] you can add the real data file to the project and (re)produce the entire project with this single command.
-In fact, until after we preregistered the study `r link("https://github.com/aaronpeikert/repro-tutorial/releases/tag/ v0.0.1-preregistration", "preregistration")` were the results below based on simulated data.
-Only after we preregistered the whole manuscript and upon submission, they were based on actual data from https://openpsychometrics.org/_rawdata/.
-
-```{r}
-read_csv(here("data", "sd3.csv"), col_types = "dddddddddd") %>% 
-  planed_analysis() %>% 
-  extract_results()
-```
-
-Note that both [this tutorial](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo2] as well as our article on the [reproducible workflow](https://github.com/aaronpeikert/reproducible-research)[^gh-repo3] [@Peikert2019] were created using this workflow and can be reproduced from their source code repositories.
-
-[^gh-repo2]: https://github.com/aaronpeikert/repro-tutorial
-[^gh-repo3]: https://github.com/aaronpeikert/reproducible-research
-
-## Closed Data (optional)
-
-<!-- Hands-on: check changed hash -->
-A prerequisite to reproducibility is the permanent availability of both code and data [@akhlaghi2020towards].
-We use containers to guarantee that all code components are available and work as expected.
-Ideally, all data are also stored well documented [@codebook] in the repository following the FAIR principles [@wilkinsonFAIRGuidingPrinciples2016].
-We believe that sharing data is highly beneficial for the research community [for a more comprehensive argument for publishing data see @vanlissa2020worcs], but this is not always feasible.
-
-Sometimes it is just the required storage capacity that exceeds what Git can handle (Git gets slow around more than 100mb).
-In such case we strongly recommend using [DataLad [@DataladDatalad2021]](http://handbook.datalad.org/en/latest/)[^datalad] or the below procedure to guarantee that the data has not changed.
-In practice, however, there are various reasons, which may prohibit you from making your data publicly available, for example legal or regulatory or privacy concerns.
-You may find yourself in a situation where you can make the data only available on request [e.g. @share].
-A major threat to reproducibility is that the data has changed in the meantime or multiple versions are circulating and it is not exactly clear what version was used originally.
-To at least check that the data was not changed, we suggest to provide a so called "hash", which is a cryptographic fingerprint.
-If the data changes, so does its fingerprint or hash.
-The user can therefore compare if the data they have received is the same that was used in the original publication, as long as the original publication does contain the hash of the data.
-
-To verify data integrity, you can include a line of code after reading the data that looks like this:
-
-[^datalad]: http://handbook.datalad.org/en/latest/
-
-```{r}
-# create a dummy data.frame with two columns
-x <- data.frame(VAR1 = c(1, 2, 3, 4), VAR2 = c(0, 4, 6, 9))
-# compute hash using md5
-hash <- digest::digest(x, "md5")
-if(hash != "5ba412f5a26f43842971dd74954fcdeb"){
-  warning("Mismatch between original and current data!\nHash now is:\n    '", hash, "'")
-}
-```
-
-We employed the same mechanism for our simulation to verify that the simulation is indeed reproducible[^simhash].
-
-[^simhash]: https://github.com/aaronpeikert/repro-tutorial/blob/main/R/hash.R
-
-A compromise between making the data publicly available and keeping it closed is the publication of synthetic data [@vanlissa2020worcs].
-Synthetic data are an attempt to recreate the marginal distribution and bivariate relationships in simulated data.
-The R package `synthpop`[@nowokSynthpopBespokeCreation2016] provides functions to generate synthetic data .
-Of course, the results of such synthetic data should generally not be interpreted but are rather meant to illustrate the workings and the structure of the output of a given data analysis.
-Also, synthetic data provides an opportunity for readers to test if a project is in principle reproducible, even though they have no access to the data.
-
-# Archival and Dissemination
-
-In a public repository, you may want to mark important stages of your project, like preregistration, preprint, submission and acceptance at a scientific journal.
-We suggest to archive these milestones individually.
-A reproducible project is only then properly archived when it is stored at least at two independent storage locations and reproducible from all locations. <!--AP citation? -->
-To really make sure that a given project is reproducible, one should aim for an explicit test of reproduction before the project is published. 
-We recommend to check out and reproduce the project in a new directory in a "fresh" environment (in particular, a new R session should be used for this). 
-For maximal validitiy of this test, we suggest to ask a coworker or collaborator to reproduce the project independently from you.
-The sections [Creating a DOI] and [Creating a Release] describe how you can easily create three different openly accessible longterm archived copies of your project.
-
-When you disseminate multiple versions of a document (e.g., at different stages of a review process), it seems good practice to specify the exact version of code and data used to create a given document (e.g.
-the final type-set version of a journal article).
-To create a clear link between the repository version and a given end result you can automatically include the unique hash identifying the project state in the git version history.
-<!--Either you use Git from within R:
-
-```{r}
-system2("git", c("rev-parse",  "--short",  "HEAD"), stdout = TRUE)
-```
--->
-To this end, we use the `gert` package:
-
-```{r}
-strtrim(gert::git_commit_id(), 7)
-```
-
-## Creating a DOI
-
-Typical open research project will have their contents available from multiple sites, e.g. from GitHub, preregistration platforms, preprint servers, or publisher websites. 
-Under these circumstances it is important to provide unique links 
-Digital Object Identifiers (DOI) tie all these locations together under one link [@derisiWhatWhysDOIs2003].
-Not only can a DOI resolve to several locations, several parts of a research project can get their own DOI besides the DOI of the publication.
-This is useful to link or cite a specific part of a research project, such as the preregistration, an illustration in the paper or the data that was gathered.
-The DOI can also contain information about the license of its target.
-<!-- public releases like preregistration, preprint -->
-zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides DOIs for these archives.
-While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
-Please note that you can only link public GitHub repositories to zenodo.
-You may log into zenodo.org through your GitHub account.
-
-> Log into zenodo.org → Account → [GitHub](https://zenodo.org/account/settings/github/)[^zenodo-github]
-
-[^zenodo-github]: https://zenodo.org/account/settings/github/
-
-After you have linked a GitHub repository, you trigger the archival by creating a GitHub release, as described in the next section.
-
-## Creating a Release
-
-In software development, a release is the distribution of a piece of computer software indicating a certain maturity in development and level of completeness.
-GitHub has a "release" function, which creates a link to a specific snapshot of your project that then is available as a single compressed file.
-We believe that the idea of software releases can be leveraged for singling out important versions in a research project.
-We found it useful to label the following stages of development as release:
-
-* Preregistration (repository contains the final data analysis plan; also see the Section on PAC)
-* Completion of data collection (repository contains all study data)
-* Preprint (repository exactly reproduces a preprint that is hosted externally)
-* Submission or Resubmission (repository exactly reproduces a given version (re)submitted to a journal)
-* In Press (repository exactly reproduces the final manuscript as accepted by a scientific journal)
-
-You have the opportunity to describe the release in more detail in the accompanying release notes and to add extra files (e.g., complementing the Markdown source code of a journal article with a rendered PDF document).
-
-## Archiving Docker
-
-A central point of criticism is that the proposed workflow introduces a bulk of software solutions and services that may not be available in the long run.
-First and foremost, it relies on R, RStudio, Pandoc, Git, Make, Docker, Homebrew (OS X only) and Chocolaty (Windows only) as software and on GitHub, DockerHub, MRAN and Ubuntu Package Archive as services.
-These many breakpoints can be reduced to a single service and a single piece of software.
-That is because all required software is bundled into the container, whose image can be stored.
-Hence, one needs a storage provider which saves the container image and the other files associated with the project.
-We already recommend to store all project files redundantly (your computer, a remote git server such as GitHub, and a publicly funded, open-science-compatible service provider such as zenodo.org).
-Building the Docker image from the Dockerfile still requires DockerHub, MRAN, and Ubuntu Package Archive to be online.
-We therefore suggest that you include an already built image into the releases.
-The Makefile generated by `repro` does include a target to do so.
-To write a compressed image to disk use:
-
-```bash
-# this should be run in the terminal
-make save-docker
-```
-
-This image allows anyone with a computer equipped with Docker to reproduce the analysis (if the repository contains all code and data and has no other external dependencies).
-Last, one can convert the Docker image to a full bootable system image.
-Such a system image can subsequently be booted on any personal computer without relying on any software or service and therefore grantees reproducibility for decades to come.
-Currently, `repro` does not yet facilitate this conversion, but we hope to include this feature soon.
-
-<!-- Hands on: add commit hash to results-->
-<!-- Hands on: git tag + github release-->
-<!-- Hands on: explain how to archive data-->
-<!-- Hands on: explain how to archive docker images-->
-
-```{r, results='asis', echo=FALSE}
-link_index()
-```
-
-# Summary
-
-Here, we provided a step-by-step tutorial with what we think are best practices for creating a fully reproducible repository for data analysis projects that center on the statistical programming language R.
-The workflow avoids common threats to reproducibility, such as copy-and-paste errors, missing files, missing software dependencies, or side-effects through changing dependent software packages, changes in the programming language, or even at the level of the operating system itself.
-However, the workflow as proposed by @Peikert2019 requires mastering up to four different software tools and their specification languages, which are typically not taught in the curriculum of social and behavioral sciences but rather belong to computer science and software development.
-Currently, this is a serious obstacle in adopting the workflow for researchers who strive for reproducibility but have little training in computer science and software development.
-The repro package, which we introduced here, supports scholars in creating reproducible workflows.
-It abstracts tedious and error-prone tasks such as the creation of a Makefile or the specification of a container.
-By extending the YAML specification that is used for storing meta information in R Markdown files, which are becoming increasingly popular among R users, repro helps users to implement @Peikert2019's workflow with minimal effort.
-Most importantly, `repro`-is implemented such that the package itself is not needed to reproduce a workflow that was generated according to the workflow.
-Importantly, this means that users that adopt the workflow do not have to rely on the future availability and maintenance of the `repro`-package.
+A preregistration as code offers four advantages over classical preregistration.
+First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
+Second, despite being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
+Researchers can, for example, formulate conditions under which they prefer one analysis over the other, i.e.
+if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
+Third, deviations from the preregistration are more explicit because they are reflected in changes to the code a property facilitated by version control.
+Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
 

### 2021-05-25 manuscript.Rmd
first half draft of tools section

@@
 
-The papers aim is twofold.
-First, we want to introduce the reader to a workflow that ensures reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code. 
+Increasingly, scientists openly share their data, materials, and analysis code.
+Sharing these digital artifacts increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results.
+However, making a research project reproducible comes with a variety of challenges and there is a dire need for standards of reproducibility.
+
+This paper has two aims.
+First, we want to introduce the reader to a workflow [see @Peikert2019 for details] that ensures reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code.
 The research community has long accepted reproducibility as integral to "good scientific practice."
 Often, however, projects can not be reproduced by other research teams, despite the best efforts by the creators.
-We present four reasons that lead to irreproducibility, explain four solutions from the software engineering community that eliminate these and show the researchers how the R package `repro` helps them to implement the solutions.
+We present four common threats to reproducibility and explain how we can address these with a combination of four solutions from the software engineering community.
+Specifically, we demonstrate how the R package `repro` supports researchers in implementing this solution.
 The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
@@
 This tutorial is aimed at researchers who are willing to make code, data, and materials available and should help them to ensure reproducibility of relevant statistical results.
-While the willingness to share materials is required for reproducing a project it is not sufficient.
+While the willingness to share materials is required for reproducing a project, it is not sufficient.
 For example, @hardwicke2018 attempted to reproduce results from open materials in the journal Cognition. 
@@
 
+After introducing the conceptual building blocks, this tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
+We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.
+We explicitly encourage readers to try out all R commands for themselves.
+Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).
+To guide the reader through all steps of a research project, we will use a exemplary research question:
+
+> "Is there a mean difference in the personality trait 'Machiavellism' between genders?"
+
+This example serves only didactic purposes and not to derive substantive claims.
+Nevertheless, everything else is kept as close to a real research question as possible.
+
 # Concepts and Software Solutions
@@
 <!-- could andreas write this? --->
+We identified the following common threat to reproducibility: <!-- CJ How did you identify these causes? Is this exhaustive?-->
+
+1.	The existence of multiple inconsistent versions of code, data, or both; for example, the dataset could have changed over time because outliers were removed at a later stage, or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
+2.	Copy-and-paste; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
+3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order; 
+4.	Code dependencies; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist; or, a different version of the same software may produce different results.
+<!-- CJ 5.
+Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
+
+We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term and cross-platform computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph.
+The opening paragraph should at least:
+1) Introduce the topic area
+2) Introduce the problem/knowledge gap
+3) Introduce how you will fill this knowledge gap-->
+
+1. Version control
+2. Dynamic document creation
+3. Dependency tracking
+4. Software management
+
+In the remainder of this section, we will briefly explain why each of these four building blocks is needed and how it plays a role in guaranteeing reproducibility.
+To resolve ambiguity across multiple versions of code and data that may be come into existence during the development phase of a research project, we recommend the use of a version control system.
+A version control system
+tracks changes to all project-related files (e.g., materials, data, and code) over time.
+At a later stage, single files or the entire project can be compared to or reverted to an earlier version.
+Additional features such as branches allow parallel versions to exist that can be merged later.
+Specific snapshots of a repository can be marked, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
+Using version control protects against loss of files and also supports remote collaboration with semi-automated tools for tracking and merging changes.
+
+Second, we rely on dynamic document generation.
+The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps.
+Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer." This paradigm has gained popularity with the Sweave approach [@leisch2002sweave] that allowed embedding the data analysis into a TeX document, which then could be dynamically rendered into a final manuscript.
+As of recently, R Markdown has added to the success of this idea.
+Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
+This way, Markdown does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
+R Markdown extends regular Markdown by allowing useres you to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
+Upon rendering of the document, the code blocks are executed and their output is dynamically inserted into the document.
+This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
+
+<!-- TODO Andreas: Internal and externdal dependencies -->
+
+```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency tracking (Make), software management (Docker), and dynamic document generation (R Markdown).
+Git tracks changes to the project over time.
+Make manages dependencies among the files.
+Docker provides a container in which the final report is built using dynamic document generation in R Markdown.
+Reproduced from @Peikert2019.
+"}
+# file gets downloaded in Makefile
+knitr::include_graphics("images/nutshell.svg", auto_pdf = TRUE)
+```
 

### 2021-05-26 manuscript.Rmd
add newlines

@@
 These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
-To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use .
+To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use.
 A rigorous standard increases the likelihood of any individual project to be actually reproducible.
@@
 To resolve ambiguity across multiple versions of code and data that may be come into existence during the development phase of a research project, we recommend the use of a version control system.
-A version control system
-tracks changes to all project-related files (e.g., materials, data, and code) over time.
+A version control system tracks changes to all project-related files (e.g., materials, data, and code) over time.
 At a later stage, single files or the entire project can be compared to or reverted to an earlier version.
@@
 The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps.
-Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer." This paradigm has gained popularity with the Sweave approach [@leisch2002sweave] that allowed embedding the data analysis into a TeX document, which then could be dynamically rendered into a final manuscript.
+Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
+This paradigm has gained popularity with the Sweave approach [@leisch2002sweave] that allowed embedding the data analysis into a TeX document, which then could be dynamically rendered into a final manuscript.
 As of recently, R Markdown has added to the success of this idea.
@@
 We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
-R Markdown extends regular Markdown by allowing useres you to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
+R Markdown extends regular Markdown by allowing users to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
 Upon rendering of the document, the code blocks are executed and their output is dynamically inserted into the document.

### 2021-05-26 manuscript.Rmd
add metascientific motivation for reproducibility

@@
 First, we want to introduce the reader to a workflow [see @Peikert2019 for details] that ensures reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code.
-The research community has long accepted reproducibility as integral to "good scientific practice."
+The research community has long accepted reproducibility as integral to "good scientific practice", because empirical research hinges on its objectivity meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
+22][^popper]
 Often, however, projects can not be reproduced by other research teams, despite the best efforts by the creators.
@@
 The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
+
+[^popper]: In fact @popperLogicScientificDiscovery2002's famous criterion of demarcation is build around "inter-subjective testing", a concept which he later generalized to inter-subjective criticism (Footnote *¹, p.
+22).
+Somewhat confusingly @popperLogicScientificDiscovery2002 uses the term "Reproducibility" for what we would call "Replication".
 

### 2021-05-26 manuscript.Rmd
stress that *others* should be able to reproduce

@@
 This paper has two aims.
-First, we want to introduce the reader to a workflow [see @Peikert2019 for details] that ensures reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code.
+First, we want to introduce the reader to a workflow [see @Peikert2019 for details] that ensures reproducibility---defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code.
 The research community has long accepted reproducibility as integral to "good scientific practice", because empirical research hinges on its objectivity meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.

### 2021-05-28 manuscript.Rmd
new text for conceptual section

@@
 
-In the remainder of this section, we will briefly explain why each of these four building blocks is needed and how it plays a role in guaranteeing reproducibility.
-To resolve ambiguity across multiple versions of code and data that may be come into existence during the development phase of a research project, we recommend the use of a version control system.
+The remainder of this section briefly explains why each of these four building blocks is needed and the role they play in ensuring reproducibility.
+
+First, to resolve ambiguity across multiple versions of code and data that may be come into existence during the development phase of a research project, we recommend the use of a version control system.
 A version control system tracks changes to all project-related files (e.g., materials, data, and code) over time.
@@
 This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
+
+One would hope that sharing a literate statistical analysis would allow others (including one own's future self) to simply download and run (ie., reproduce) a given analysis. 
+However, as @ argue, in most instances this approach fails. 
+In practice, dependencies on external software (such as system libraries or components of the programming language) are frequently unmentioned and the components or packages are not included.
+Troubleshooting  issues specific to a certain programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
+One solution to this problem is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
+"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silver2017].
+Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself.
+The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
+Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
+Such a snapshot of the software stack is called an "image".
+Packaging all needed software in such an image requires considerable amounts of storage space.
+There are two major strategies helping to keep the storage requirements reasonable.
+First, there is a community that maintains images for particular purposes.
+For example, there are ready-made images for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
+Users can then install whatever they need in addition to what is provided by these pre-compiled images.
+`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
+Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
+Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
+This recipe is a textual description of all commands that need to be executed to recreate the image.
+Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
+However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
+For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
+With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`
+
+The fourth and last concept we rely on is 
+In software projects, tools like the make utility, are used to automatically determine, which pieces of a program need to be recompiled when the source code or dependent libraries are changed and also to issue the commands how to build the final product.
+In essence, developers provide a single _Makefile_, which can be thought of as a collection of computational recipes.
+Similar to a collection of cooking recipes, we can have multiple products (_targets_ in Make speak) with different ingredients (_requirements_) and different steps of preparation (_recipe_).
+In the context of a scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results, such as preprocessed data files, simulation results, analysis results).
+To make this more concrete, consider the following hypothetical example, in which a research project contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
+A Makefile could then look like this:
+
+```
+manuscript.pdf: manuscript.Rmd simulation_results.csv 
+  Rscript -e 'rmarkdown::render("manuscript.Rmd")'
+
+simulation_results.csv: simulate.R
+  Rscript -e 'source("simulate.R")'
+```
+
+There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).
+Each target is followed by a colon and a list of requirements.
+If a requirement is  newer than the target, the recipe will be executed to rebuild the target.
+If a requirement does not exist, Make seeks to find a recipe to build the requirement first before building the target.
+Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would make sure that the file `simulation_results.csv` exists; if not, it would find that this is also a target and issue the command in l.5 to run the simulation first before rendering the manuscript.
+This way, we ensure that the simulation is always run before the manuscript is built and that the manuscript is rebuilt if the simulation code was changed.
+<!-- AP: I was just wondering...
+is this true?
+Is the caching mechanism transitive?-->
+
+To summarize, the workflow by @Peikert2019 requires four components, dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), containerization (using Docker).
+While R Markdown and git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which represents a considerable hurdle in the acceptance and implementation of the workflow.
+To remove this hurdle, we have developed a new R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
+In the remainder, we will walk you through a complete research project, from sketching the very first idea over preregistration, data analysis, and submission of a preprint while focusing on how the `repro`-package enables users to maintain reproducibility of all steps.
 

### 2021-05-28 manuscript.Rmd
added figure ref

@@
 
-<!-- could andreas write this? --->
 We identified the following common threat to reproducibility: <!-- CJ How did you identify these causes? Is this exhaustive?-->
@@
 
-To summarize, the workflow by @Peikert2019 requires four components, dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), containerization (using Docker).
+To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref{schematic}), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), containerization (using Docker).
 While R Markdown and git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which represents a considerable hurdle in the acceptance and implementation of the workflow.
@@
 In the remainder, we will walk you through a complete research project, from sketching the very first idea over preregistration, data analysis, and submission of a preprint while focusing on how the `repro`-package enables users to maintain reproducibility of all steps.
-
-<!-- TODO Andreas: Internal and externdal dependencies -->
 

### 2021-05-31 R/simulation.R
backup

@@ -1,60 +1,73 @@
-suppressPackageStartupMessages(library(tidyverse))
-suppressPackageStartupMessages(library(here))
-suppressPackageStartupMessages(library(lavaan))
-suppressPackageStartupMessages(library(furrr))
-
-source(here("R", "simulation_funs.R"))
-# if you have access to a hpc envir specify this in R/hpc.R
-# see https://github.com/aaronpeikert/repro-tutorial/blob/hpc/R/hpc.R
-# or `git checkout hpc R/hpc.R`
-# if no hpc is availible we use local multicore with all available cores
-# to speed up consider reduce nsim; increase nstep ↓↓↓
-
-hpc_config <- here::here("R", "hpc.R")
-if(fs::file_exists(hpc_config)){
-  source(hpc_config)
-} else {
-  plan(list(transparent,
-            tweak(multisession)))
-}
-# debug:
-#   plan(transparent)
-
-nmin <- 100
-nmax <- 10000
-nstep <- 100
-n_sim <- 1000
-seed <- 1234
-n_obs <- seq(nmin, nmax, nstep)
-
-options(scipen = 999)
-
-to_export <- ls_funs() %>% map(get)
-to_export <-
-  c(to_export, list(
-    n_obs = seq(nmin, nmax, nstep),
-    n_sim = n_sim
-  ))
-out_file <- here::here("data", "simulation_results.csv")
-if(!fs::file_exists(out_file)){
-  message("This may take a while! You fit approximately ", n_sim*length(n_obs)*5*2, " SEMs.")
-  res_raw %<-% simulation_study(n_obs, n_sim,
-                                furrr_options(
-                                  globals = to_export,
-                                  seed = seed,
-                                  packages = c("furrr", "lavaan", "tidyverse"),
-                                  scheduling = 10
-                                ))
-  invisible(res_raw)
-  fs::dir_create("data")
-  readr::write_csv(res_raw, here::here("data", "simulation_results.csv"))
-  readr::write_rds(res_raw, here::here("data", "simulation_results.rds"))
+with_lecuyer <- function(seed, code) {
+  withr::with_seed(
+    seed,
+    code,
+    .rng_kind = "L'Ecuyer-CMRG"
+  )
 }
 
-if(fs::file_exists(hpc_config)){
-  list(local = sessioninfo::session_info(),
[…]

### 2021-05-31 R/simulation_funs.R

@@ -1,188 +1,0 @@
-source(here::here("R", "standardizedSolution.R"))
-
-#----sim-functions----
-filter_dots <- function(..., fun, what = names(formals(fun))){
-  dots <- list(...)
-  if(any(names(dots) == ""))stop("filter_dots does not allow unnamed args!")
-  exists <- setdiff(what, setdiff(what, names(dots)))
-  dots[exists]
-}
-
-safe_cfa <- function(...){
-  valid_args <-
-    c(names(formals(lavaan::lavaanify)),
-      names(formals(lavaan::lavaan)),
-      names(lavaan::lavOptions()))
-  dots <- filter_dots(..., fun = lavaan::cfa, what = valid_args)
-  do.call("cfa", dots, envir = getNamespace("lavaan"))
-}
-
-stitch <- function(pars){
-  with(pars, str_c(lhs, op, rhs))
-}
-
-partial_measurement_invariance <- function(data, same, differ, mod_load = 0L, mod_int = 0L, pars = "", ...){
-  res_mod <- set_free_mod(differ, data, pars, op = "=~", depth = mod_load, ...)
-  res_mod <- set_free_mod(differ, data, res_mod$pars, op = "~1", depth = mod_int, ...)
-  fits <- list(same = safe_cfa(model = same, data = data, group.partial = res_mod$pars, ...),
-               differ = res_mod$differ)
-  
-}
-
-set_free_mod <- function(differ, data, pars, depth = 3, op = c("~1", "=~"), ...){
-  differ_fit <- safe_cfa(model = differ, data = data, group.partial = pars, ...)
-  depth <- depth - 1
-  if(depth < 0){return(list(pars = pars, differ = differ_fit))}
-  if(!is_converged(differ_fit)){return(list(pars = pars, differ = differ_fit))}
-  parest <- lavaan::parameterEstimates(differ_fit)
-  testscores <- lavaan::lavTestScore(differ_fit)
-  sorted_pars <- testscores$uni %>% 
-    rename(label = lhs) %>% 
-    select(-c(op, rhs, df)) %>% 
-    left_join(parest, by = c("label")) %>% 
-    filter(group == 1, .data$op %in% !!op) %>% 
-    arrange(desc(X2))
-  sorted_pars <- sorted_pars[1,]
-  
-  if(pull(sorted_pars, X2) > qchisq(0.95, 1)){
-    pars <- append(pars, stitch(sorted_pars))
[…]

### 2021-05-31 R/standardizedSolution.R

@@ -1,171 +1,0 @@
-# https://github.com/aaronpeikert/lavaan/commit/a528028973c9b9ffc4afbde3c4a663a9e9a9be23
-standardizedSolution <-
-  standardizedsolution <- function(object,
-                                   type = "std.all",
-                                   se = TRUE,
-                                   zstat = TRUE,
-                                   pvalue = TRUE,
-                                   ci = TRUE,
-                                   level = 0.95,
-                                   cov.std = TRUE,
-                                   remove.eq = TRUE,
-                                   remove.ineq = TRUE,
-                                   remove.def = FALSE,
-                                   partable = NULL,
-                                   GLIST = NULL,
-                                   est   = NULL,
-                                   output = "data.frame") {
-    
-    stopifnot(type %in% c("std.all", "std.lv", "std.nox"))
-    
-    # check output= argument
-    output <- tolower(output)
-    if(output %in% c("data.frame", "table")) {
-      output <- "data.frame"
-    } else if(output %in% c("text", "pretty")) {
-      output <- "text"
-    } else {
-      stop("lavaan ERROR: output must be ", sQuote("data.frame"),
-           " or ", sQuote("text"))
-    }
-    
-    # no zstat + pvalue if estimator is Bayes
-    if(object@Options$estimator == "Bayes") {
-      zstat <- pvalue <- FALSE
-    }
-    
-    # no se if class is not lavaan
-    # using class() -- can't use inherits(), as this includes blavaan
-    if(class(object)[1L] != "lavaan") {
-      if(missing(se) || !se) {
-        se <- FALSE
-        zstat <- FALSE
-        pvalue <- FALSE
-      }
-    }
-    
-    if(is.null(partable)) {
-      PARTABLE <- lavaan::inspect(object, "list")
-    } else {
-      PARTABLE <- partable
-    }
-    LIST <- PARTABLE[,c("lhs", "op", "rhs", "exo")]
-    if(!is.null(PARTABLE$group)) {
-      LIST$group <- PARTABLE$group
-    }
-    if(!is.null(PARTABLE$block)) {
-      LIST$block <- PARTABLE$block
-    }
-    if (sum(nchar(PARTABLE$label)) != 0L) {
-      LIST$label <- PARTABLE$label
-    }
-    
[…]

### 2021-05-31 manuscript.Rmd
newlines
cite statcheck

@@
 A rigorous standard increases the likelihood of any individual project to be actually reproducible.
-An easy to use standard on the other hand is more likely to be employed across research projects.
+An easy to use standard on the other hand is more likely to be employed across many research projects.
 The remainder of this paper is structured along these lines.
@@
 Second, we rely on dynamic document generation.
-The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps.
+The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps which are conducted in two different programs.
+In a word processor the researcher writes the text and in another software they conduct the analysis.
+Both steps are linked by manually copying and pasting results from one software to the other; a process that often produces in inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
+
 Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."

### 2021-05-31 manuscript.Rmd
fix typo

@@
 In a word processor the researcher writes the text and in another software they conduct the analysis.
-Both steps are linked by manually copying and pasting results from one software to the other; a process that often produces in inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
+Both steps are linked by manually copying and pasting results from one software to the other; a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
 

### 2021-06-01 manuscript.Rmd
reinclude chunk for container size
explain what CLI is and how we make it easy closes #112

@@
 At a later stage, single files or the entire project can be compared to or reverted to an earlier version.
-Additional features such as branches allow parallel versions to exist that can be merged later.
+Additional features, such as branches, facilitate collaboration.
 Specific snapshots of a repository can be marked, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
@@
 As of recently, R Markdown has added to the success of this idea.
+R Markdown uses Markdown for writing the text, and R (or other programming languages) for the analysis.
 Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
@@
 Troubleshooting  issues specific to a certain programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
+
+```{r, include=FALSE}
+container_size <- as.numeric(fs::file_size("reprotutorial.sif"))/1024^3
+```
+
 One solution to this problem is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
@@
 
+A major hurdle for adoption of the before mentioned tools for researchers is that they are typically used through a command line interface (CLI).
+The CLI is vastly more user friendly than its predecessor, the punch card, but many users (including the authors) prefer a graphical user interface (GUI) over text based interactions for most tasks.
+A GUI has the obvious disadvantages for reproducibility, that it requires a user who clicks on the correct buttons in the correct order.
+Depending on the number of steps this is not only much effort but also error prone.
+This tutorial seeks a compromise.
+We use a GUI (and similar helpers) where we can, to create a reproducible workflow but the reproduction of such workflow does not rely on manual steps.
+At several points we rely on or have self developed a CLI that wraps a more complicated (and powerful) CLI.
+These wrapper use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in laymans terms.
+We hope this make reproducibility tools more accessible by enabling untrained users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
+These alternative tools are merely an assistance system and as users get more comfortable they can use the underlying tools directly to solve more complex problems.
+
+One example for such assistance are the check functions from `repro`.
+
+Two properties of CLI tools make them useful but at the same time difficult to learn.
+Many CLI tools are decades old, and are therefore stable, widely used, and widely supported.
+Because of their age they seem archaic and are difficult to use when compared to more modern computer tools.
+These tools are also more powerful than many GUI counterparts, because of their flexibility---many options to change their behavior and composabibility---the output of one, can be the input of another.
+
 # Preregistration as Code
@@
 if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
-Third, deviations from the preregistration are more explicit because they are reflected in changes to the code a property facilitated by version control.
+Third, deviations from the preregistration are more explicit because they are reflected in changes to the codes a property facilitated by version control.
 Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.

### 2021-06-01 manuscript.Rmd
fix missing citation

@@
 One solution to this problem is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
-"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silver2017].
+"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
 Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself.

### 2021-06-01 manuscript.Rmd
realign introducion closes #110

@@
 However, as @ argue, in most instances this approach fails.
-In practice, dependencies on external software (such as system libraries or components of the programming language) are frequently unmentioned and the components or packages are not included.
+In practice, dependencies between project files (the raw data was preprocessed) or on external software (such as system libraries or components of the programming language) are frequently unmentioned.
+
+To automatically resolve dependencies between project files we rely on dependency tracking.
+In essence, developers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and in the end a final product.
+Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipe_).
+In the context of a scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results, such as preprocessed data files, simulation results, analysis results).
+
+We recommend to use Make for dependency tracking because it is language independent.
+To make this more concrete, consider the following hypothetical example, in which a research project contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
+A Makefile could then look like this:
+
+```
+manuscript.pdf: manuscript.Rmd simulation_results.csv 
+  Rscript -e 'rmarkdown::render("manuscript.Rmd")'
+
+simulation_results.csv: simulate.R
+  Rscript -e 'source("simulate.R")'
+```
+There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).
+Each target is followed by a colon and a list of requirements.
+If a requirement is  newer than the target, the recipe will be executed to rebuild the target.
+If a requirement does not exist, Make seeks to find a recipe to build the requirement first before building the target.
+Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would make sure that the file `simulation_results.csv` exists; if not, it would find that this is also a target and issue the command in l.5 to run the simulation first before rendering the manuscript.
+This way, we ensure that the simulation is always run before the manuscript is built and that the manuscript is rebuilt if the simulation code was changed.
+
 Troubleshooting  issues specific to a certain programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
@@
 
-One solution to this problem is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
+One solution to problem of exteranal software dependencies is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
 "By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
@@
 With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`
-
-The fourth and last concept we rely on is 
-In software projects, tools like the make utility, are used to automatically determine, which pieces of a program need to be recompiled when the source code or dependent libraries are changed and also to issue the commands how to build the final product.
-In essence, developers provide a single _Makefile_, which can be thought of as a collection of computational recipes.
-Similar to a collection of cooking recipes, we can have multiple products (_targets_ in Make speak) with different ingredients (_requirements_) and different steps of preparation (_recipe_).
-In the context of a scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results, such as preprocessed data files, simulation results, analysis results).
-To make this more concrete, consider the following hypothetical example, in which a research project contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
-A Makefile could then look like this:
-
-```
-manuscript.pdf: manuscript.Rmd simulation_results.csv 
-  Rscript -e 'rmarkdown::render("manuscript.Rmd")'
-
-simulation_results.csv: simulate.R
-  Rscript -e 'source("simulate.R")'
-```
-
-There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).
-Each target is followed by a colon and a list of requirements.
-If a requirement is  newer than the target, the recipe will be executed to rebuild the target.
-If a requirement does not exist, Make seeks to find a recipe to build the requirement first before building the target.
-Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would make sure that the file `simulation_results.csv` exists; if not, it would find that this is also a target and issue the command in l.5 to run the simulation first before rendering the manuscript.
-This way, we ensure that the simulation is always run before the manuscript is built and that the manuscript is rebuilt if the simulation code was changed.
-<!-- AP: I was just wondering...
-is this true?
-Is the caching mechanism transitive?-->
 

### 2021-06-02 manuscript.Rmd
add setup + begin of Git

@@
 At several points we rely on or have self developed a CLI that wraps a more complicated (and powerful) CLI.
-These wrapper use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in laymans terms.
+These wrapper use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in laymans terms and are accessible from within R.
 We hope this make reproducibility tools more accessible by enabling untrained users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
 These alternative tools are merely an assistance system and as users get more comfortable they can use the underlying tools directly to solve more complex problems.
+Please install the package [`gert`](https://github.com/r-lib/gert)[@R-gert] which wraps Git commands and [`repro`](https://github.com/aaronpeikert)[@R-repro] so they can act as assistance systems while you follow the tutorial:
+
+```{r, eval=FALSE}
+install.packages("gert")
+# repro is not on CRAN yet, install from GitHub
+if(!requireNamespace("remotes"))install.packages("remotes")
+remotes::install_github("aaronpeikert/repro")
+```
 
 One example for such assistance are the check functions from `repro`.
+They verify that you indeed have installed and setup the required software for this workflow:
+
+```{r}
+# `package::function()` → use function from package without `library(package)`
+repro::check_git()
+repro::check_make()
+repro::check_docker()
+```
+
+They detect the users system state and explain what this means to the user.
+Sometimes they ask the user to do something, e.g. the following happens if you were a windows user who does not have installed Git:
+
+```{r, include=FALSE}
+opts <- options()
+options(repro.os = "windows",
+        repro.git = FALSE)
+```
+
+```{r}
+check_git()
+```
+
+```{r, include=FALSE}
+options(opts)
+```
+
+Before you continue we ask you to run the above commands to check Git, Make, and Docker.
+If necessary follow the instructions of repro to install them.
+
+The first concept we want to practically apply is version control with Git.
+Git tracks files within one project folder.
+We start by creating such project folder with RStudio:
+
+> File → New Project... → New Directory → Example Repro Template
+
+This creates a project containing an example analysis, but you may use any other template or existing R Project.
+
+To prevent Git from accidentally committing files or you do not want to track you can use the `.gitignore`, for example:
+
+```{r}
+usethis::use_git_ignore("private.md")
+```
+
+We activate Git in any project with:
+
+```{r, eval=FALSE}
+usethis::use_git()
+```
 
@@
 ```{r, include=FALSE}
-knitr::write_bib(c(.packages(), "here", "rticles", "gert", "bookdown", "lavaan", "knitr", "targets", "renv", "tidyverse"),
+knitr::write_bib(c(.packages(), "repro", "here", "rticles", "gert", "bookdown", "lavaan", "knitr", "targets", "renv", "tidyverse"),
                  "packages.bib")

### 2021-06-04 manuscript.Rmd
update git tutorial section

@@
 This tutorial seeks a compromise.
-We use a GUI (and similar helpers) where we can, to create a reproducible workflow but the reproduction of such workflow does not rely on manual steps.
+We use simpler alternatives including a GUI where we can, to create a reproducible workflow but the reproduction of such workflow does not rely on manual steps.
 At several points we rely on or have self developed a CLI that wraps a more complicated (and powerful) CLI.
@@
 
-To prevent Git from accidentally committing files or you do not want to track you can use the `.gitignore`, for example:
+You can prevent Git from accidentally committing files which you do not want to track with the `.gitignore`, for example:
 
@@
 
-We activate Git in any project with:
+Now the file `private.md` will not be committed to Git and therefore will not be made public.
+
+After you have ignored sensitive files (like not yet anonymized data) you can activate Git with:
 

### 2021-06-04 manuscript.Rmd
extend git introduction

@@
 First, to resolve ambiguity across multiple versions of code and data that may be come into existence during the development phase of a research project, we recommend the use of a version control system.
+Version control allows a clear link between which results were generated by which verion of code and data.
+
 A version control system tracks changes to all project-related files (e.g., materials, data, and code) over time.
 At a later stage, single files or the entire project can be compared to or reverted to an earlier version.
-Additional features, such as branches, facilitate collaboration.
-Specific snapshots of a repository can be marked, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
-Using version control protects against loss of files and also supports remote collaboration with semi-automated tools for tracking and merging changes.
+A version controlled project is makes the loss of files unlikely and also supports remote collaboration.
+
+We recommend to use Git for version control, because of its widespread adoption in the R community.
+Git is build around snapshots which represent the entire project state at a given point in time.
+Those snapshots are called "commits".
+Each file that was changed has to be added to a commit and each commit requires a commit message that succinctly describes the changes made.
+Each commit "knows" its ancestor and they build a timeline together.
+The conglomerate of commits (i.e. the version controlled project) is called repository.
+In Git specific snapshots of a repository can be tagged, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
+Git has additional features beyond basic version control, such as branches to facilitate collaboration.
+To collaborate on a Git repository, the repository has to be uploaded somewhere.
+We recommend GitHub to host Git repositories because of its popularity among R users.
+GitHub has many tools that ease project management and collaboration and these tools provide much value in our everyday work but they are not central to archiving reproducibility.
 
@@
 Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
-This paradigm has gained popularity with the Sweave approach [@leisch2002sweave] that allowed embedding the data analysis into a TeX document, which then could be dynamically rendered into a final manuscript.
-As of recently, R Markdown has added to the success of this idea.
 R Markdown uses Markdown for writing the text, and R (or other programming languages) for the analysis.
@@
 ```
+
 There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).

### 2021-06-04 manuscript.Rmd
transition from Git + RMarkdown to Make + Docker

@@
 One would hope that sharing a literate statistical analysis would allow others (including one own's future self) to simply download and run (ie., reproduce) a given analysis.
-However, as @ argue, in most instances this approach fails.
-In practice, dependencies between project files (the raw data was preprocessed) or on external software (such as system libraries or components of the programming language) are frequently unmentioned.
+However, while using version control and dynamic document generation are becoming more common @ argue in most instances this approach fails to garuntee reproducibility.
+In practice, dependencies between project files (e.g.
+the raw data was preprocessed) or on external software (such as system libraries or components of the programming language) are frequently unmentioned.
 
@@
 
-One solution to problem of exteranal software dependencies is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
+One solution to problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
 "By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].

### 2021-06-04 manuscript.Rmd
run introduction through grammerly

@@
 Increasingly, scientists openly share their data, materials, and analysis code.
-Sharing these digital artifacts increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results.
-However, making a research project reproducible comes with a variety of challenges and there is a dire need for standards of reproducibility.
+Sharing these digital artefacts increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results.
+However, making a research project reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
 
@@
 First, we want to introduce the reader to a workflow [see @Peikert2019 for details] that ensures reproducibility---defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code.
-The research community has long accepted reproducibility as integral to "good scientific practice", because empirical research hinges on its objectivity meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
+The research community has long accepted reproducibility as integral to "good scientific practice" because empirical research hinges on its objectivity meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
 22][^popper]
@@
 
-[^popper]: In fact @popperLogicScientificDiscovery2002's famous criterion of demarcation is build around "inter-subjective testing", a concept which he later generalized to inter-subjective criticism (Footnote *¹, p.
+[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalised to inter-subjective criticism (Footnote *¹, p.
 22).
-Somewhat confusingly @popperLogicScientificDiscovery2002 uses the term "Reproducibility" for what we would call "Replication".
+Somewhat confusingly, @popperLogicScientificDiscovery2002 uses the term "Reproducibility" for what we would call "Replication".
 
-This tutorial is aimed at researchers who are willing to make code, data, and materials available and should help them to ensure reproducibility of relevant statistical results.
+This tutorial aims at researchers who are willing to make code, data, and materials available and should help them ensure the reproducibility of relevant statistical results.
 While the willingness to share materials is required for reproducing a project, it is not sufficient.
@@
 @obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
-The authors could execute only 31 of the scripts without error and were able to reproduce the results of only 21 articles (within a reasonable period of time).
+The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
 
@@
 A rigorous standard increases the likelihood of any individual project to be actually reproducible.
-An easy to use standard on the other hand is more likely to be employed across many research projects.
+An easy to use standard, on the other hand, is more likely to be employed across many research projects.
 The remainder of this paper is structured along these lines.
@@
 
-After introducing the conceptual building blocks, this tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
+After introducing the conceptual building blocks, this tutorial guides the reader through each step of an open research project, emphasising guaranteeing reproducibility from the very first step on.
 We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.
@@
 Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).
-To guide the reader through all steps of a research project, we will use a exemplary research question:
+To guide the reader through all steps of a research project, we will use an exemplary research question:
 

### 2021-06-04 manuscript.Rmd
grammerly on concepts section

@@
 
-The remainder of this section briefly explains why each of these four building blocks is needed and the role they play in ensuring reproducibility.
+The remainder of this section briefly explains why each of these four building blocks is needed and their role in ensuring reproducibility.
 
-First, to resolve ambiguity across multiple versions of code and data that may be come into existence during the development phase of a research project, we recommend the use of a version control system.
-Version control allows a clear link between which results were generated by which verion of code and data.
+First, to resolve ambiguity across multiple versions of code and data that come into existence during the development phase of a research project, we recommend using a version control system.
+Version control allows a clear link between which results were generated by which version of code and data.
 
@@
 At a later stage, single files or the entire project can be compared to or reverted to an earlier version.
-A version controlled project is makes the loss of files unlikely and also supports remote collaboration.
+A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
 
-We recommend to use Git for version control, because of its widespread adoption in the R community.
-Git is build around snapshots which represent the entire project state at a given point in time.
+We recommend using Git for version control, because of its widespread adoption in the R community.
+Git is built around snapshots that represent the entire project state at a given point in time.
 Those snapshots are called "commits".
-Each file that was changed has to be added to a commit and each commit requires a commit message that succinctly describes the changes made.
-Each commit "knows" its ancestor and they build a timeline together.
-The conglomerate of commits (i.e. the version controlled project) is called repository.
-In Git specific snapshots of a repository can be tagged, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
+Each file that was changed has to be added to a commit, and each commit requires a commit message that succinctly describes the changes made.
+Each commit "knows" its ancestor, and they form a timeline together.
+The conglomerate of commits (i.e. the version-controlled project) is called a repository.
+In Git, specific snapshots of a repository can be tagged, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
 Git has additional features beyond basic version control, such as branches to facilitate collaboration.
-To collaborate on a Git repository, the repository has to be uploaded somewhere.
+To collaborate via Git,<!--dangling modify, cant help it--> the repository has to be uploaded somewhere.
 We recommend GitHub to host Git repositories because of its popularity among R users.
-GitHub has many tools that ease project management and collaboration and these tools provide much value in our everyday work but they are not central to archiving reproducibility.
+GitHub has many tools that ease project management and collaboration, and these tools provide much value in our everyday work, but they are not central to archiving reproducibility.
 
 Second, we rely on dynamic document generation.
-The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps which are conducted in two different programs.
-In a word processor the researcher writes the text and in another software they conduct the analysis.
-Both steps are linked by manually copying and pasting results from one software to the other; a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
+The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
+In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
+Both steps are linked by manually copying and pasting results from one software to the other, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
 
-Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
-R Markdown uses Markdown for writing the text, and R (or other programming languages) for the analysis.
-Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
-This way, Markdown does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
+Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured, who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
+R Markdown uses Markdown for writing the text and R (or other programming languages) for the analysis.
+Markdown is a lightweight text format in plain text with minimal reserved symbols for formatting instructions.
+This way, Markdown does not need any specialized software for editing.
+It is highly fool-proof (unlike, for example, LaTeX), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+We suggest using Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
 R Markdown extends regular Markdown by allowing users to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
-Upon rendering of the document, the code blocks are executed and their output is dynamically inserted into the document.
+Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
 This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
@@
 One would hope that sharing a literate statistical analysis would allow others (including one own's future self) to simply download and run (ie., reproduce) a given analysis.
-However, while using version control and dynamic document generation are becoming more common @ argue in most instances this approach fails to garuntee reproducibility.
+However, while version control and dynamic document generation are becoming more common @, this approach fails to guarantee reproducibility in most instances.
 In practice, dependencies between project files (e.g.
@@
 
-To automatically resolve dependencies between project files we rely on dependency tracking.
-In essence, developers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and in the end a final product.
+To automatically resolve dependencies between project files, we rely on dependency tracking.
+In essence, developers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
 Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipe_).
-In the context of a scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results, such as preprocessed data files, simulation results, analysis results).
+In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results, such as preprocessed data files, simulation results, analysis results).
 
-We recommend to use Make for dependency tracking because it is language independent.
+We recommend using Make for dependency tracking because it is language independent.
 To make this more concrete, consider the following hypothetical example, in which a research project contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
@@
 If a requirement is newer than the target, the recipe will be executed to rebuild the target.
-If a requirement does not exist, Make seeks to find a recipe to build the requirement first before building the target.
+If a requirement does not exist, Make uses a recipe to build the requirement first before building the target.
 Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would make sure that the file `simulation_results.csv` exists; if not, it would find that this is also a target and issue the command in l.5 to run the simulation first before rendering the manuscript.
@@
 
-Troubleshooting  issues specific to a certain programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
+Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
 
@@
 
-One solution to problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
+One solution to the problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
 "By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
@@
 The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
-Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
+Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
 Such a snapshot of the software stack is called an "image".
 Packaging all needed software in such an image requires considerable amounts of storage space.
-There are two major strategies helping to keep the storage requirements reasonable.
+Two major strategies are helping to keep the storage requirements reasonable.
 First, there is a community that maintains images for particular purposes.
-For example, there are ready-made images for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
+For example, there are ready-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
 Users can then install whatever they need in addition to what is provided by these pre-compiled images.
@@
 For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
-With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`
+With the `repro`-package, adding R packages dependencies is as simple as adding them to the header under the tag `packages:`
 

### 2021-06-07 manuscript.Rmd
reformat code chunk

@@
 ```{r, include=FALSE}
-knitr::write_bib(c(.packages(), "repro", "here", "rticles", "gert", "bookdown", "lavaan", "knitr", "targets", "renv", "tidyverse"),
-                 "packages.bib")
+knitr::write_bib(
+  c(
+    .packages(),
+    "repro",
+    "here",
+    "rticles",
+    "gert",
+    "bookdown",
+    "lavaan",
+    "knitr",
+    "targets",
+    "renv",
+    "tidyverse"
+  ),
+  "packages.bib"
+)
 ```

### 2021-06-08 abstract.Rmd
Revert "each sentance on newline"
This reverts commit b5e77d9ef35d43f6cf9a43d2bcb77f5e4218d61b.
include originator as beneficiary

@@
 Computational reproducibility is the ability to obtain identical results from the *same* data with the *same* computer code.
-It is a building block for transparent and cumulative science because it enables other researchers, on other computers and later in time, to reproduce and thus understand how results came about while, at the same time, it avoids a variety of errors that may lead to erroneous reporting of statistical and computational results.
+It is a building block for transparent and cumulative science because it enables the originator and other researchers, on other computers and later in time, to reproduce and thus understand how results came about while, at the same time, it avoids a variety of errors that may lead to erroneous reporting of statistical and computational results.
 In this tutorial, we demonstrate how the R package `repro` supports researchers in creating fully computationally reproducible research projects with a focus on current Open Science practices; starting from sketching the research proposal, over preregistering an analysis as executable code with simulated data, until the creation and dissemination of the final scientific report.

### 2021-06-08 manuscript.Rmd
Concepts and Software Solutions is taking shape

@@
 
-First, to resolve ambiguity across multiple versions of code and data that come into existence during the development phase of a research project, we recommend using a version control system.
+First, to resolve ambiguity across multiple versions of code and data that come into existence during the development phase of a research project, we recommend using a **version control** system.
 Version control allows a clear link between which results were generated by which version of code and data.
@@
 
-We recommend using Git for version control, because of its widespread adoption in the R community.
+We recommend using **Git** for version control, because of its widespread adoption in the R community.
 Git is built around snapshots that represent the entire project state at a given point in time.
@@
 
-Second, we rely on dynamic document generation.
+Second, we rely on **dynamic document generation**.
 The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
@@
 Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured, who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
-R Markdown uses Markdown for writing the text and R (or other programming languages) for the analysis.
+**R Markdown** uses Markdown for writing the text and R (or other programming languages) for the analysis.
 Markdown is a lightweight text format in plain text with minimal reserved symbols for formatting instructions.
@@
 
-To automatically resolve dependencies between project files, we rely on dependency tracking.
+To automatically resolve dependencies between project files, we rely on **dependency tracking**.
 In essence, developers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
 Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipe_).
-In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results, such as preprocessed data files, simulation results, analysis results).
+In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, analysis results).
 
-We recommend using Make for dependency tracking because it is language independent.
+We recommend using **Make** for dependency tracking because it is language independent.
 To make this more concrete, consider the following hypothetical example, in which a research project contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
@@
 
+A version controlled dynamic document with dependency tracking still relies on external software.
 Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
@@
 
-One solution to the problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
+One solution to the problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
+One comprehensive approach to software management is containerization.
 "By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
-Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself.
+Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
 The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
@@
 For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
-With the `repro`-package, adding R packages dependencies is as simple as adding them to the header under the tag `packages:`
 
@@
 
-```{r}
+```{r, eval=FALSE}
 usethis::use_git_ignore("private.md")

### 2021-06-08 manuscript.Rmd
some buity corrections

@@
 author: Aaron Peikert and Andreas M. Brandmaier 
-output:
-  bookdown::pdf_document2:
-    keep_tex: TRUE
-    toc: FALSE
+output: bookdown::pdf_document2
 repro:
@@
 
-1.	The existence of multiple inconsistent versions of code, data, or both; for example, the dataset could have changed over time because outliers were removed at a later stage, or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
-2.	Copy-and-paste; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order; 
-4.	Code dependencies; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist; or, a different version of the same software may produce different results.
+1.	**Multiple inconsistent versions of code, data, or both**; for example, the dataset could have changed over time because outliers were removed at a later stage, or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
+2.	**Copy-and-paste**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
+3.	**Ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order; 
+4.	**External code dependencies**; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist; or, a different version of the same software may produce different results.
 <!-- CJ 5.
@@
 It is highly fool-proof (unlike, for example, LaTeX), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
+We suggest using Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your CV [@vitae].
 R Markdown extends regular Markdown by allowing users to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
@@
 In essence, developers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
-Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipe_).
+Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
 In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, analysis results).
@@
 
-```
+```{bash, eval=FALSE}
 manuscript.pdf: manuscript.Rmd simulation_results.csv 
@@
 For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
+
+\@ref{schematic}
 

### 2021-06-09 manuscript.Rmd

@@
 ---
-title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
-author: Aaron Peikert and Andreas M. Brandmaier 
-output: bookdown::pdf_document2
+title: A Hitchhiker's Guide to Reproducible Research
+author:
+  - name: Aaron Peikert
+    affil: 1, *
+    orcid: 0000-0001-7813-818X
+  - name: Caspar J. Van Lissa
+    affil: 2, 3
+    orcid: 0000-0002-0808-5024
+  - name: Andreas M. Brandmaier
+    affil: 1, 4
+    orcid: 0000-0001-8765-6982
+affiliation:
+  - num: 1
+    address: |
+      Center for Lifespan Psychology---Max Planck Institute for Human Development
+      Lentzeallee 94, 14195 Berlin, Germany
+  - num: 2
+    address: |
+      Department of Methodology & Statistics---Utrecht University faculty of Social and Behavioral Sciences, Utrecht, Netherlands
+  - num: 3
+    address: |
+      Open Science Community Utrecht, Utrecht, Netherlands
+  - num: 4
+    address: |
+      Max Planck UCL Centre for Computational Psychiatry and Ageing Research
+      Berlin, Germany and London, UK
+# firstnote to eighthnote
+correspondence: |
+  peikert@mpib-berlin.mpg.de
+journal: psych
+type: tutorial
+status: submit
+bibliography: temp.bib
+appendix: appendix.tex
+simplesummary: |
+  A Simple summary goes here.
+abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
+keywords: |
+  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
+  to the article, yet reasonably common within the subject discipline.).
+acknowledgement: |
+  All sources of funding of the study should be disclosed. Please clearly 
+  indicate grants that you have received in support of your research work. 
+  Clearly state if you received funds for covering the costs to publish in open 
+  access.
+authorcontributions: |
+  For research articles with several authors, a short paragraph specifying their 
+  individual contributions must be provided. The following statements should be 
+  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
+  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
+  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
+  limited to those who have contributed substantially to the work reported.
+conflictsofinterest: |
+  The authors declare no conflict of interest.
+abbreviations:
+  - short: GUI
+    long: Graphical User Interface
+  - short: CLI
+    long: Comand Line Interface
 repro:
@@
     - R/link.R
-csl: apa7.csl
-numbered_headings: true
-bibliography: ["references.bib", "packages.bib"]
-abstract: "`r tryCatch(trimws(readr::read_file('abstract.Rmd')))`"
+output:
+  bookdown::markdown_document2:
+    base_format: rticles::mdpi_article
+header-includes:
+   - \usepackage[labelformat=empty]{caption}
 ---
-
-<!-- the HTML comments, like this one, are meta comments, mainly describing the intent --->
-<!-- each sentence below a heading summarizes what I want to say there --->
-<!-- "Hands-on:" means concrete practical application, they roughly proceed from easy/familiar to hard/unfamiliar-->
 
@@
 
-[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalised to inter-subjective criticism (Footnote *¹, p.
+[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalised to inter-subjective criticism (Footnote *1, p.
 22).
@@
 Version control allows a clear link between which results were generated by which version of code and data.
-
 A version control system tracks changes to all project-related files (e.g., materials, data, and code) over time.
@@
 
-\@ref{schematic}
-
-To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref{schematic}), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), containerization (using Docker).
+To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), containerization (using Docker).
 While R Markdown and git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which represents a considerable hurdle in the acceptance and implementation of the workflow.
@@
 # file gets downloaded in Makefile
-knitr::include_graphics("images/nutshell.svg", auto_pdf = TRUE)
+knitr::include_graphics(here::here("images/nutshell.svg"), auto_pdf = TRUE)
 ```
@@
   ),
-  "packages.bib"
+  here("packages.bib")
 )
+cat(c(readLines(here::here("packages.bib")), "\n", readLines(here::here("references.bib"))), sep = "\n", file = "temp.bib")
 ```

### 2021-06-09 manuscript.Rmd
install repro from r-universe closes #113

@@
 install.packages("gert")
-# repro is not on CRAN yet, install from GitHub
-if(!requireNamespace("remotes"))install.packages("remotes")
-remotes::install_github("aaronpeikert/repro")
+# repro is not on CRAN yet
+install.packages('repro', repos = 'https://aaronpeikert.r-universe.dev')
 ```

### 2021-06-09 manuscript.Rmd
include "R" in title closes #94
minor changes

@@
 ---
-title: A Hitchhiker's Guide to Reproducible Research
+title: A Hitchhiker's Guide to Reproducible Research in R
 author:
@@
 
-This tutorial aims at researchers who are willing to make code, data, and materials available and should help them ensure the reproducibility of relevant statistical results.
+This tutorial aims at researchers who are willing to make all code, data, and materials---publicly or on request---available and should help them ensure the reproducibility of relevant statistical results.
 While the willingness to share materials is required for reproducing a project, it is not sufficient.

### 2021-06-09 R/simulation.R
rename variables for clarity

@@ -37,12 +37,13 @@ planned_analysis <- function(data){
   x <- rowMeans(dplyr::select(data, dplyr::starts_with("x")))
   y <- as.factor(data$group)
   skew <- moments::skewness(x)
-  rank <- abs(skew) > 1
-  if(rank){
+  # skewness cutoff
+  use_rank <- abs(skew) > 1
+  if(use_rank){
     x <- rank(x)
   }
   test <- t.test(x ~ y)
-  list(test = test, skew = skew, rank = rank, n = length(x))
+  list(test = test, skew = skew, use_rank = use_rank, n = length(x))
 }
 
 extract_results <- function(analysis){

### 2021-06-09 manuscript.Rmd
remove orphan section on cli

@@
 
-Two properties of CLI tools make them useful but at the same time difficult to learn.
-Many CLI tools are decades old, and are therefore stable, widely used, and widely supported.
-Because of their age they seem archaic and are difficult to use when compared to more modern computer tools.
-These tools are also more powerful than many GUI counterparts, because of their flexibility---many options to change their behavior and composabibility---the output of one, can be the input of another.
-
 # Preregistration as Code

### 2021-06-14 manuscript.Rmd
explain the tutorial character of repro

@@
 
-Before you continue we ask you to run the above commands to check Git, Make, and Docker.
+The messages emitted by `repro` and other packages we present try to take the user by the hand to help them solve problems.
+This tutorial describes steps that we expect apply to all users, the messages from the packages fill the gap to your specific situation and extend the tutorial in this way.
+Before you continue, we ask you to run the above commands to check Git, Make, and Docker.
 If necessary follow the instructions of repro to install them.

### 2021-06-14 manuscript.Rmd
what is rmarkdown metadata and how does repro use it

@@
 
+If you made changes to a file you have to stage and then commit these changes.
+The easiest way is to use the RStudio Git pane.
+You click on the empty box next to the file you want to stage and when a tick appears the file is staged.
+After you have staged all files you want click on the commit button, explain curtly why you made those changes in a commit message and then click on commit.
+
+The files you created and the changes you made still have not left your computer.
+To backup and share the files, you can publicly upload those changes to GitHub with:
+
+```{r, eval=FALSE}
+usethis::use_github()
+```
+
+Depending on your computer configuration this may ask to setup a secure connection to GitHub.
+
+If you are not yet ready to share you project publicly use:
+
+```{r, eval=FALSE}
+usethis::use_github(private = TRUE)
+```
+
+Now that you have created a project that is under version control we turn to dynamic document generation.
+Three aspects are important to understand a dynamic document:
+
+1. Text
+2. Code
+3. Metadata
+
+The text of an R Markdown is written in Markdown.
+The following example shows how to signify a header, a bold word and a list in Markdown:
+
+```markdown
+<!--this is a Markdown file -->
+# Header
+
+Normal text.
+Important **word** in bold.
+
+To do list:
+
+* do research
+* do more research
+* spend time with family
+```
+
+Code is separated by three backtics and the programming language in curly braces like here:
+
+````markdown
+This is normal text, written in Markdown.
+
+`r ''````{r}
+# this is R code
+1 + 1
+```
+````
+
+Now that you know how to write text and R code in an RMarkdown the only thing you need to know about is metadata, like the title or the output format.
+
+Metadata are placed at the beginning of the document and are seperated by three dashes; the following example is a full markdown document where the header is in lines 1-6.
+
+````markdown
+---
+title: A Hitchhiker's Guide to Reproducible Research in R
+author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
+abstract: A paper about recalculating the same thing more than once.
+output: html_document
+---
+
+# Introduction
+
+Important for reproducibility:
+
+1. Version Control
+2. Dynamic Document Creation
+3. Dependency Tracking
+4. Software Management
+
+`r ''````{r}
+# this is R code
+t.test(extra ~ group, data = sleep)
+```
+````
+
+Most of the metadata fields are self explanatory (like the author field) but any package may use the metadata for their purposes.
+Often, for example, does a specific output format require more metadata i.e. the `rticles`-package for authoring scientific manuscripts.
+RMarkdown was extended to hundreds of document formats which are often highly configurable by their metadata.
+The `repro` package uses the metadata, not to change the output, but to make the document more reproducible, simply by listing resources, i.e., scripts, data, and packages, your document depends on.
+
+```markdown
+---
+title: A Hitchhiker's Guide to Reproducible Research in R
+author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
+abstract: A paper about recalculating the same thing more than once.
+output: html_document
+repro:
+  scripts:
+    - R/load.R
+  data:
+    - data/mtcars.csv
+  packages:
+    - tidyverse
+    - usethis
+    - gert
+---
+```
+
+These information make internal dependencies to other files in the project and external dependencies to R packages explicit.
+`repro` can use these information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
+In day-to-day work the simplest solution is to use the function `repro::automate()` which both infers a Makefile and a Dockerfile from all RMarkdown files (all files with the ending `.Rmd`).
+Everytime you changed the `repro`-metadata, change the output format, or add a new RMarkdown, the Makefile and Dockerfile are potentially outdated and you should run `repro::automate()` to keep them up to date.
+The function `repro::automate()` does not overwrite manual changes and there is no harm in running it too often.
+
 # Preregistration as Code
@@
 We argue that a preregistration as code (PAC) is an excellent planning tool that offers several benefits over traditional preregistration.
-For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data.
+For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data using simulated data.
 

### 2021-06-15 manuscript.Rmd
explain how reproduction works with repro

@@
 
-These information make internal dependencies to other files in the project and external dependencies to R packages explicit.
-`repro` can use these information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
-In day-to-day work the simplest solution is to use the function `repro::automate()` which both infers a Makefile and a Dockerfile from all RMarkdown files (all files with the ending `.Rmd`).
-Everytime you changed the `repro`-metadata, change the output format, or add a new RMarkdown, the Makefile and Dockerfile are potentially outdated and you should run `repro::automate()` to keep them up to date.
-The function `repro::automate()` does not overwrite manual changes and there is no harm in running it too often.
+This information makes internal dependencies to other files in the project and external dependencies to R packages explicit.
+`repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
+In day-to-day work the simplest solution is to use the function `repro::automate()` which both infers a Makefile and a Dockerfile from all RMarkdown  (all files with the ending `.Rmd`) and there respective repro metadata.
+Every time you change the `repro`-metadata, change the output format, or add a new RMarkdown, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
+The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
+Besides, the Makefile and the Dockerfile repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
+
+When you or people who received your project from GitHub want to reproduce the project, they can ask repro to explain how this should be done:
+
+```{r}
+repro::rerun()
+```
+
+However, users do not have to have repro installed, in fact, they do not even need R installed. 
+The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.
+They can be pasted into a terminal and will reproduce the project in many cases.
+Some projects, however, are not as easy to make reproducible.
+As explained above, repro is merely a simplified interface.
+These simplifications are possible because of two restrictions.
+First, the project can only rely on R packages for the Dockerfile and second, it can have no multilayered preprocessing for the Makefile.
+Lifting these restrictions requires the user to interact more directly with Make or Docker.
+Hence, it is necessary to understand what these commands do to satisfy more complicated requirements because then you need to understand how repro structures its interaction with Make and Docker.
+
+Let us dissect the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`.
+The command `make docker` will trigger Make to build the target `docker`.
+The recipe for this target builds an image from your Dockerfile.
+The `&&` combines two commands in one line.
+The call `make` without any explicit target will build the Make target `all`, where the flag `-B` means that Make will consider all prerequisites as outdated and hence will rebuild everything.
+repro constructs Make targets so that if you supply `DOCKER=TRUE` they are executed within the Docker image of the project.
 

### 2021-06-15 manuscript.Rmd
check with grammerly

@@
 
-A major hurdle for adoption of the before mentioned tools for researchers is that they are typically used through a command line interface (CLI).
-The CLI is vastly more user friendly than its predecessor, the punch card, but many users (including the authors) prefer a graphical user interface (GUI) over text based interactions for most tasks.
-A GUI has the obvious disadvantages for reproducibility, that it requires a user who clicks on the correct buttons in the correct order.
-Depending on the number of steps this is not only much effort but also error prone.
+A significant hurdle for adopting the before-mentioned tools for researchers is that they are typically used through a command-line interface (CLI).
+The CLI is vastly more user friendly than its predecessor, the punch card, but many users (including the authors) prefer a graphical user interface (GUI) over text-based interactions for most tasks.
+A GUI has the apparent disadvantages for reproducibility that it requires a user who clicks on the correct buttons in the correct order.
+Depending on the number of steps, this is not only much effort but also error-prone.
 This tutorial seeks a compromise.
-We use simpler alternatives including a GUI where we can, to create a reproducible workflow but the reproduction of such workflow does not rely on manual steps.
-At several points we rely on or have self developed a CLI that wraps a more complicated (and powerful) CLI.
-These wrapper use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in laymans terms and are accessible from within R.
-We hope this make reproducibility tools more accessible by enabling untrained users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
-These alternative tools are merely an assistance system and as users get more comfortable they can use the underlying tools directly to solve more complex problems.
+We use simpler alternatives, including a GUI where we can, to create a reproducible workflow, but the reproduction of such workflow does not rely on manual steps.
+At several points, we rely on or have self-developed a CLI that wraps a more complicated (and powerful) CLI.
+These wrappers use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
+We hope this makes reproducibility tools more accessible by enabling untrained users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
+These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
 Please install the package [`gert`](https://github.com/r-lib/gert)[@R-gert] which wraps Git commands and [`repro`](https://github.com/aaronpeikert)[@R-repro] so they can act as assistance systems while you follow the tutorial:
@@
 
-They detect the users system state and explain what this means to the user.
+They detect the users' system state and explain what this means to the user.
 Sometimes they ask the user to do something, e.g. the following happens if you were a windows user who does not have installed Git:
@@
 The messages emitted by `repro` and other packages we present try to take the user by the hand to help them solve problems.
-This tutorial describes steps that we expect apply to all users, the messages from the packages fill the gap to your specific situation and extend the tutorial in this way.
+This tutorial describes steps that we expect to apply to all users; the messages from the packages fill the gap to your specific situation and extend the tutorial in this way.
 Before you continue, we ask you to run the above commands to check Git, Make, and Docker.
-If necessary follow the instructions of repro to install them.
+If necessary, follow the instructions of repro to install them.
 
-The first concept we want to practically apply is version control with Git.
+The first concept we want to apply practically is version control with Git.
 Git tracks files within one project folder.
@@
 
-You can prevent Git from accidentally committing files which you do not want to track with the `.gitignore`, for example:
+You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore`, for example:
 
@@
 
-Now the file `private.md` will not be committed to Git and therefore will not be made public.
+Now the file `private.md` will not be committed to Git and will not be made public.
 
-After you have ignored sensitive files (like not yet anonymized data) you can activate Git with:
+After you have ignored sensitive files (like not yet anonymized data), you can activate Git with:
 
@@
 
-If you made changes to a file you have to stage and then commit these changes.
+If you made changes to a file, you have to stage and then commit these changes.
 The easiest way is to use the RStudio Git pane.
-You click on the empty box next to the file you want to stage and when a tick appears the file is staged.
-After you have staged all files you want click on the commit button, explain curtly why you made those changes in a commit message and then click on commit.
+You click on the empty box next to the file you want to stage, and when a tick appears, the file is staged.
+After you have staged all files you want, click on the commit button, explain curtly why you made those changes in a commit message and then click on commit.
 
 The files you created and the changes you made still have not left your computer.
-To backup and share the files, you can publicly upload those changes to GitHub with:
+To back up and share the files, you can publicly upload those changes to GitHub with:
 
@@
 
-Depending on your computer configuration this may ask to setup a secure connection to GitHub.
+Depending on your computer configuration, this may ask to set up a secure connection to GitHub.
 
-If you are not yet ready to share you project publicly use:
+If you are not yet ready to share your project publicly use:
 
@@
 
-Now that you have created a project that is under version control we turn to dynamic document generation.
-Three aspects are important to understand a dynamic document:
+Now that you have created a project that is under version control, we turn to dynamic document generation.
+Three aspects are essential to understand a dynamic document:
 
@@
 
-Code is separated by three backtics and the programming language in curly braces like here:
+Code is separated by three backticks (and the programming language in curly braces) like here:
 
@@
 
-Now that you know how to write text and R code in an RMarkdown the only thing you need to know about is metadata, like the title or the output format.
+Now that you know how to write text and R code in an RMarkdown, you need to know about metadata, like the title or the output format.
 
-Metadata are placed at the beginning of the document and are seperated by three dashes; the following example is a full markdown document where the header is in lines 1-6.
+Metadata are placed at the beginning of the document and are separated by three dashes; the following example is a full markdown document where the header is in lines 1-6.
 
@@
 
-Most of the metadata fields are self explanatory (like the author field) but any package may use the metadata for their purposes.
-Often, for example, does a specific output format require more metadata i.e. the `rticles`-package for authoring scientific manuscripts.
+Most metadata fields are self-explanatory (like the author field), but any package may use the metadata for their purposes.
+Often, for example, does a specific output format require more metadata, i.e. the `rticles`-package for authoring scientific manuscripts.
 RMarkdown was extended to hundreds of document formats which are often highly configurable by their metadata.

### 2021-06-15 manuscript.Rmd
reconsider wording

@@
 
-You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore`, for example:
+You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore` file.
+You can add something to the `.gitignore` file directly or with this command:
 
@@
 
-Depending on your computer configuration, this may ask to set up a secure connection to GitHub.
+Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow its suggestions.
 
@@
 Now that you know how to write text and R code in an RMarkdown, you need to know about metadata, like the title or the output format.
-
 Metadata are placed at the beginning of the document and are separated by three dashes; the following example is a full markdown document where the header is in lines 1-6.
@@
 RMarkdown was extended to hundreds of document formats which are often highly configurable by their metadata.
-The `repro` package uses the metadata, not to change the output, but to make the document more reproducible, simply by listing resources, i.e., scripts, data, and packages, your document depends on.
+The `repro` package uses the metadata, not to change the output, but to make the document more reproducible, simply by listing resources, i.e., scripts, data, and packages, your document depends on, in the following format:
 
@@
 The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
-Besides, the Makefile and the Dockerfile repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
+Besides the main Makefile and the Dockerfile repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
 
@@
 Lifting these restrictions requires the user to interact more directly with Make or Docker.
-Hence, it is necessary to understand what these commands do to satisfy more complicated requirements because then you need to understand how repro structures its interaction with Make and Docker.
+To satisfy more complicated requirements users need to understand how repro utilizes Make and Docker internally.
 
-Let us dissect the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`.
+Let us dissect the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two steps.
+First it virtually recreates the software environment and then it executes computational recipes in the virtual software environment.
+The first step is represented by the command `make docker`.
 The command `make docker` will trigger Make to build the target `docker`.
 The recipe for this target builds an image from your Dockerfile.
-The `&&` combines two commands in one line.
-The call `make` without any explicit target will build the Make target `all`, where the flag `-B` means that Make will consider all prerequisites as outdated and hence will rebuild everything.
-repro constructs Make targets so that if you supply `DOCKER=TRUE` they are executed within the Docker image of the project.
+The `&&` only runs the second command if the first was successful.
+The second step is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy properties.
+First, a call to `make` without any explicit target will build the Make target `all`.
+Second, the flag `-B` means that Make will consider all prerequisites as outdated and hence will rebuild everything.
+Third, repro constructs Make targets so that if you supply `DOCKER=TRUE` they are executed within the Docker image of the project.
 

### 2021-06-15 manuscript.Rmd
address andreas' comments

@@
 Increasingly, scientists openly share their data, materials, and analysis code.
-Sharing these digital artefacts increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results.
+Sharing these digital artefacts increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
 However, making a research project reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
@@
 Specifically, we demonstrate how the R package `repro` supports researchers in implementing this solution.
-The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
+The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
 
@@
 
-Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.
 These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
+These open repositories do not routinely provide sufficient to reproduce relevant computational and statistical results.
+We argue that the research community could benefit more from the increased transparency across scholarly disciplines (that is, data and methods are increasingly openly shared), if more projects could be reproduced.
 To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use.
@@
 
-> "Is there a mean difference in the personality trait 'Machiavellism' between genders?"
+> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"
 
@@
 To automatically resolve dependencies between project files, we rely on **dependency tracking**.
-In essence, developers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
+In essence, researchers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
 Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
@@
 A significant hurdle for adopting the before-mentioned tools for researchers is that they are typically used through a command-line interface (CLI).
-The CLI is vastly more user friendly than its predecessor, the punch card, but many users (including the authors) prefer a graphical user interface (GUI) over text-based interactions for most tasks.
+The CLI predates and coexists with graphical user interfaces (GUIs) which many users (including the authors) prefer over the text-based interactions of a CLI for most tasks.
 A GUI has the apparent disadvantages for reproducibility that it requires a user who clicks on the correct buttons in the correct order.

### 2021-06-15 manuscript.Rmd
number code chunk lines closes #117

@@
    - \usepackage[labelformat=empty]{caption}
+   - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, numbers=left}
 ---

### 2021-06-16 manuscript.Rmd
Opening paragraph edits
A few more comments and some edits to show how I think the clarity of the writing can be improved. Comments on the remainder of the manuscript will follow

@@
 Increasingly, scientists openly share their data, materials, and analysis code.
-Sharing these digital artefacts increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
-However, making a research project reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
+Sharing these digital artefacts <!-- consider using the term 'research output' instead, as digital artefacts might be unnecessarily technical-->increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible<!--Does this last statement apply to data and materials?
+Reproducible materials?-->.
+<!-- Suggest adding sentence: The technology is now available to make all artefacts related to a research project available in such a way that the project can be reproduced with minimal effort. --> 
+Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
+22][^popper]
+In contemporary research, reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
+Unfortunately, despite increasing commitment to open science practices, many projects can not yet be reproduced by other research teams.
+This is because making a research project reproducible <!-- how does a research project relate to data/materials/code/artefacts?--> comes with a variety of challenges, and there is a dire need for reproducibility standards.
 
 This paper has two aims.
-First, we want to introduce the reader to a workflow [see @Peikert2019 for details] that ensures reproducibility---defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code.
-The research community has long accepted reproducibility as integral to "good scientific practice" because empirical research hinges on its objectivity meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
-22][^popper]
-Often, however, projects can not be reproduced by other research teams, despite the best efforts by the creators.
-We present four common threats to reproducibility and explain how we can address these with a combination of four solutions from the software engineering community.
-Specifically, we demonstrate how the R package `repro` supports researchers in implementing this solution.
-The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
+First, we want to introduce the reader to a workflow that ensures reproducibility [see @Peikert2019 for details]. 
+<!--Second?
+This is really a different point than introducing the workflow-->We present four common threats to reproducibility and explain how we can address these with a combination of four solutions from the software engineering community.
+<!-- Third? -->Specifically, we demonstrate how the R package `repro` supports researchers in implementing this solution.
+<!-- Not sure if this belongs in the present paper.
+If it does, then I wouldn't introduce it as a separate goal.
+Instead, I would say something like: The present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community.
+We show how these solutions address several common threats to reproducibility, and enable novel solutions further enhancing the open science workflow, such as 'preregistration as code'.
+-->The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
+This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project, and are willing to make relevant code, data, and materials available; whether publicly or on request.
 
@@
 
-This tutorial aims at researchers who are willing to make all code, data, and materials---publicly or on request---available and should help them ensure the reproducibility of relevant statistical results.
-While the willingness to share materials is required for reproducing a project, it is not sufficient.
+While the willingness to share materials is required for reproducing a project, it is not sufficient.<!-- This sounds like the start of a new paragraph-->
 For example, @hardwicke2018 attempted to reproduce results from open materials in the journal Cognition. 
@@
 The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
+These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
+The results further illustrate that open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
 
-These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
-These open repositories do not routinely provide sufficient to reproduce relevant computational and statistical results.
-We argue that the research community could benefit more from the increased transparency across scholarly disciplines (that is, data and methods are increasingly openly shared), if more projects could be reproduced.
-To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use.
+We argue that the benefits associated with transparency increase if research can additionally be reproduced.
+This paper tutorial demonstrates how to make research projects reproducible, while striking a balance between rigor and ease-of-use.
 A rigorous standard increases the likelihood of any individual project to be actually reproducible.
-An easy to use standard, on the other hand, is more likely to be employed across many research projects.
+An easy to use standard, on the other hand, is more likely to be adopted.
 The remainder of this paper is structured along these lines.
@@
 
-After introducing the conceptual building blocks, this tutorial guides the reader through each step of an open research project, emphasising guaranteeing reproducibility from the very first step on.
+After introducing the conceptual building blocks, this tutorial guides the reader through each step of an open research project, <!--What does this mean?
+Isn't the whole paper about reproducibility?-->emphasising guaranteeing reproducibility from the very first step on.
 We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.

### 2021-06-24 manuscript.Rmd
replace research artefacts with research output keep reproducible materials, makes sense for me

@@
 Increasingly, scientists openly share their data, materials, and analysis code.
-Sharing these digital artefacts <!-- consider using the term 'research output' instead, as digital artefacts might be unnecessarily technical-->increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible<!--Does this last statement apply to data and materials?
-Reproducible materials?-->.
-<!-- Suggest adding sentence: The technology is now available to make all artefacts related to a research project available in such a way that the project can be reproduced with minimal effort. --> 
+Sharing the digital research output increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
+The technology is now available to make all artefacts related to a research project available so that the project can be reproduced with minimal effort.
 Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.

### 2021-06-24 manuscript.Rmd
research project = code, data, materials

@@
 Unfortunately, despite increasing commitment to open science practices, many projects can not yet be reproduced by other research teams.
-This is because making a research project reproducible <!-- how does a research project relate to data/materials/code/artefacts?--> comes with a variety of challenges, and there is a dire need for reproducibility standards.
+This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
 

### 2021-06-24 manuscript.Rmd
restructure the two aims of the paper

@@
 
-This paper has two aims.
-First, we want to introduce the reader to a workflow that ensures reproducibility [see @Peikert2019 for details]. 
-<!--Second?
-This is really a different point than introducing the workflow-->We present four common threats to reproducibility and explain how we can address these with a combination of four solutions from the software engineering community.
-<!-- Third? -->Specifically, we demonstrate how the R package `repro` supports researchers in implementing this solution.
-<!-- Not sure if this belongs in the present paper.
-If it does, then I wouldn't introduce it as a separate goal.
-Instead, I would say something like: The present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community.
-We show how these solutions address several common threats to reproducibility, and enable novel solutions further enhancing the open science workflow, such as 'preregistration as code'.
--->The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
-This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project, and are willing to make relevant code, data, and materials available; whether publicly or on request.
+This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
+We pursue two aims.
+First, the present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community that eliminate four common threats to reproducibility [see @Peikert2019 for details].
+Specifically, we demonstrate how the R package `repro` supports researchers in implementing these solutions.
+The second objective is to discuss opportunities of such workflow to improve not only how research outputs are created and recreated but also how research itself is conducted.
+One such opportunity is an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
+This dynamic document which closely resembles an actual journal article but with "fake" results and without discussion of results, then serves as preregistration.
 
@@
 
-While the willingness to share materials is required for reproducing a project, it is not sufficient.<!-- This sounds like the start of a new paragraph-->
+While the willingness to share materials is required for reproducing a project, it is not sufficient.
 For example, @hardwicke2018 attempted to reproduce results from open materials in the journal Cognition. 
@@
 
-After introducing the conceptual building blocks, this tutorial guides the reader through each step of an open research project, <!--What does this mean?
-Isn't the whole paper about reproducibility?-->emphasising guaranteeing reproducibility from the very first step on.
 We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.

### 2021-06-28 manuscript.Rmd
explain that PAC is not new

@@
 Increasingly, scientists openly share their data, materials, and analysis code.
-Sharing the digital research output increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
-The technology is now available to make all artefacts related to a research project available so that the project can be reproduced with minimal effort.
-Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
+Sharing these digital artefacts increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
+However, making a research project reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
+
+This paper has two aims.
+First, we want to introduce the reader to a workflow [see @Peikert2019 for details] that ensures reproducibility---defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code.
+The research community has long accepted reproducibility as integral to "good scientific practice" because empirical research hinges on its objectivity meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
 22][^popper]
-In contemporary research, reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
-Unfortunately, despite increasing commitment to open science practices, many projects can not yet be reproduced by other research teams.
-This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
-
-This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
-We pursue two aims.
-First, the present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community that eliminate four common threats to reproducibility [see @Peikert2019 for details].
-Specifically, we demonstrate how the R package `repro` supports researchers in implementing these solutions.
-The second objective is to discuss opportunities of such workflow to improve not only how research outputs are created and recreated but also how research itself is conducted.
-One such opportunity is an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
-This dynamic document which closely resembles an actual journal article but with "fake" results and without discussion of results, then serves as preregistration.
+Often, however, projects can not be reproduced by other research teams, despite the best efforts by the creators.
+We present four common threats to reproducibility and explain how we can address these with a combination of four solutions from the software engineering community.
+Specifically, we demonstrate how the R package `repro` supports researchers in implementing this solution.
+The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
 
@@
 
+This tutorial aims at researchers who are willing to make all code, data, and materials---publicly or on request---available and should help them ensure the reproducibility of relevant statistical results.
 While the willingness to share materials is required for reproducing a project, it is not sufficient.
@@
 The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
+
 These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
-The results further illustrate that open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
-
-We argue that the benefits associated with transparency increase if research can additionally be reproduced.
-This paper tutorial demonstrates how to make research projects reproducible, while striking a balance between rigor and ease-of-use.
+These open repositories do not routinely provide sufficient to reproduce relevant computational and statistical results.
+We argue that the research community could benefit more from the increased transparency across scholarly disciplines (that is, data and methods are increasingly openly shared), if more projects could be reproduced.
+To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use.
 A rigorous standard increases the likelihood of any individual project to be actually reproducible.
-An easy to use standard, on the other hand, is more likely to be adopted.
+An easy to use standard, on the other hand, is more likely to be employed across many research projects.
 The remainder of this paper is structured along these lines.
@@
 
+After introducing the conceptual building blocks, this tutorial guides the reader through each step of an open research project, emphasising guaranteeing reproducibility from the very first step on.
 We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.
@@
 They can be pasted into a terminal and will reproduce the project in many cases.
+
 Some projects, however, are not as easy to make reproducible.
@@
 We argue that a preregistration as code (PAC) is an excellent planning tool that offers several benefits over traditional preregistration.
-For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data using simulated data.
+For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data, were the results are initially based on simulated data.
+
+The idea of submitting code as part of a preregistration is not new.
+A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods and in an informal literature research (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
+
+We discuss PAC as an example of how a robust reproducibility standard has the potential to benefit research projects throughout the whole research process, starting at the beginning, with planning a study.
 

### 2021-06-28 manuscript.Rmd
cita apa

@@
 
-To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript.
+To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
 The introduction and theoretical background provide the basis for the studies design and can be written before gathering data.
@@
 To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
-The researchers can include a "faked " results section based on the simulated data, employing dynamic document generation.
+The researchers can include a "faked" results section based on the simulated data, employing dynamic document generation.
 Repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to assure efficiency.

### 2021-06-28 manuscript.Rmd
include benefits of simulation from earlier draft

@@
 The researchers can include a "faked" results section based on the simulated data, employing dynamic document generation.
+
 Repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to assure efficiency.
 For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study.
+While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
+First, you have to address and could plan for data analytic problems on a conceptual level before you made the considerable effort to acquire data [@axelrodAdvancingArtSimulation1997].
+Second, if you only simulate a single dataset, you can perform a simple technical check, that is the analysis software outputs parameter estimates [@braiekTestingMachineLearning2020], e.g.
+when the simulated data contains missings.
+Third, if you perform a Monte Carlo simulation, you may evaluate how the data analysis copes with expected data complications, e.g.
+non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
+Fourth, you can plan your resources efficiently, e.g. determine required sample size to achieve certain precision or power [@brandmaier2018precision].
 

### 2021-06-28 manuscript.Rmd
cite potential applications for "reusability"

@@
 To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
-The introduction and theoretical background provide the basis for the studies design and can be written before gathering data.
+The introduction and theoretical background provide the basis for the studies' design and can be written before gathering data.
 We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
@@
 
-A preregistration as code offers four advantages over classical preregistration.
+A PAC offers four advantages over classical preregistration.
 First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
@@
 if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
-Third, deviations from the preregistration are more explicit because they are reflected in changes to the codes a property facilitated by version control.
+Third, deviations from the preregistration are more explicit because they are reflected in changes to the code a property facilitated by version control.
 Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
+
+Why are PAC and reproducibility related?
+A reproducible script is supposed to work on the same data at a later point in time.
+PAC requires a script to work on different data at a later point in time, a property we call "reusability".
+This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] so it is build-for-reuse [@al-badareenReusableSoftwareComponents2010].
+Reusability promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
 

### 2021-06-29 manuscript.Rmd
move sentance
preregistrations are not always reproducible

@@
 For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data, were the results are initially based on simulated data.
+We discuss PAC as an example of how a robust reproducibility standard has the potential to benefit research projects throughout the whole research process, starting at the beginning, with planning a study.
 
@@
 A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods and in an informal literature research (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
-
-We discuss PAC as an example of how a robust reproducibility standard has the potential to benefit research projects throughout the whole research process, starting at the beginning, with planning a study.
 
@@
 Why are PAC and reproducibility related?
+One the one hand, preregistrations are plagued by nonreproducibility like any other research outcome.
+@bakkerRecommendationsPreregistrationsInternal2020 found from the 210 preregistrations they investigated 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to theoretically reproduce the power analysis.
+Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
+@steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
+
+On the other hand have PACs greater requirements for reproducibility than a simple analysis script.
 A reproducible script is supposed to work on the same data at a later point in time.
@@
 This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] so it is build-for-reuse [@al-badareenReusableSoftwareComponents2010].
-Reusability promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
+Reusability promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
 

### 2021-06-29 manuscript.Rmd
fix small mistakes

@@
 These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
-These open repositories do not routinely provide sufficient to reproduce relevant computational and statistical results.
+These open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
 We argue that the research community could benefit more from the increased transparency across scholarly disciplines (that is, data and methods are increasingly openly shared), if more projects could be reproduced.
@@
 This recipe is a textual description of all commands that need to be executed to recreate the image.
-Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
+Such files are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an sufficently describe how to recreate a given container image.
 However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
@@
 To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), containerization (using Docker).
-While R Markdown and git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which represents a considerable hurdle in the acceptance and implementation of the workflow.
+While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which represents a considerable hurdle in the acceptance and implementation of the workflow.
 To remove this hurdle, we have developed a new R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
@@
 
-Now that you know how to write text and R code in an RMarkdown, you need to know about metadata, like the title or the output format.
+Now that you know how to write text and R code in an R Markdown, you need to know about metadata, like the title or the output format.
 Metadata are placed at the beginning of the document and are separated by three dashes; the following example is a full markdown document where the header is in lines 1-6.
@@
 Often, for example, does a specific output format require more metadata, i.e. the `rticles`-package for authoring scientific manuscripts.
-RMarkdown was extended to hundreds of document formats which are often highly configurable by their metadata.
+R Markdown was extended to hundreds of document formats which are often highly configurable by their metadata.
 The `repro` package uses the metadata, not to change the output, but to make the document more reproducible, simply by listing resources, i.e., scripts, data, and packages, your document depends on, in the following format:
@@
 `repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
-In day-to-day work the simplest solution is to use the function `repro::automate()` which both infers a Makefile and a Dockerfile from all RMarkdown  (all files with the ending `.Rmd`) and there respective repro metadata.
-Every time you change the `repro`-metadata, change the output format, or add a new RMarkdown, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
+In day-to-day work the simplest solution is to use the function `repro::automate()` which both infers a Makefile and a Dockerfile from all R Markdown  (all files with the ending `.Rmd`) and there respective repro metadata.
+Every time you change the `repro`-metadata, change the output format, or add a new R Markdown, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
 The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
@@
 
-Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [https://www.pnas.org/content/115/11/2600].
-Preregistration increases the credibility of empirical results in three ways [preregistration is hard and worthwhile, Nosek].
+Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
+Preregistration increases the credibility of empirical results in three ways, @nosekPreregistrationHardWorthwhile2019:
 

### 2021-06-29 manuscript.Rmd

--- 2021-06-28 manuscript.Rmd
@@
 Increasingly, scientists openly share their data, materials, and analysis code.
-Sharing these digital artefacts increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
-However, making a research project reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
+Sharing the digital research output increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
+The technology is now available to make all artefacts related to a research project available so that the project can be reproduced with minimal effort.
+Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
+22][^popper]
+In contemporary research, reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
+Unfortunately, despite increasing commitment to open science practices, many projects can not yet be reproduced by other research teams.
+This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
 
-This paper has two aims.
-First, we want to introduce the reader to a workflow [see @Peikert2019 for details] that ensures reproducibility---defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code.
-The research community has long accepted reproducibility as integral to "good scientific practice" because empirical research hinges on its objectivity meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
-22][^popper]
-Often, however, projects can not be reproduced by other research teams, despite the best efforts by the creators.
-We present four common threats to reproducibility and explain how we can address these with a combination of four solutions from the software engineering community.
-Specifically, we demonstrate how the R package `repro` supports researchers in implementing this solution.
-The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
+This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
+We pursue two aims.
+First, the present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community that eliminate four common threats to reproducibility [see @Peikert2019 for details].
+Specifically, we demonstrate how the R package `repro` supports researchers in implementing these solutions.
+The second objective is to discuss opportunities of such workflow to improve not only how research outputs are created and recreated but also how research itself is conducted.
+One such opportunity is an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
+This dynamic document closely resembles an actual journal article but with "fake" results and without discussion of results, then serves as preregistration.
 
@@
 
-This tutorial aims at researchers who are willing to make all code, data, and materials---publicly or on request---available and should help them ensure the reproducibility of relevant statistical results.
 While the willingness to share materials is required for reproducing a project, it is not sufficient.
@@
 
-After introducing the conceptual building blocks, this tutorial guides the reader through each step of an open research project, emphasising guaranteeing reproducibility from the very first step on.
 We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.
@@
 They can be pasted into a terminal and will reproduce the project in many cases.
-
 Some projects, however, are not as easy to make reproducible.
@@
 We argue that a preregistration as code (PAC) is an excellent planning tool that offers several benefits over traditional preregistration.
-For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data, were the results are initially based on simulated data.
-
-The idea of submitting code as part of a preregistration is not new.
-A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods and in an informal literature research (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
-
-We discuss PAC as an example of how a robust reproducibility standard has the potential to benefit research projects throughout the whole research process, starting at the beginning, with planning a study.
+For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data using simulated data.
 
@@
 
-To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
-The introduction and theoretical background provide the basis for the studies' design and can be written before gathering data.
+To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript.
+The introduction and theoretical background provide the basis for the studies design and can be written before gathering data.
 We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
 To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
-The researchers can include a "faked" results section based on the simulated data, employing dynamic document generation.
-
+The researchers can include a "faked " results section based on the simulated data, employing dynamic document generation.
 Repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to assure efficiency.
 For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study.
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-First, you have to address and could plan for data analytic problems on a conceptual level before you made the considerable effort to acquire data [@axelrodAdvancingArtSimulation1997].
-Second, if you only simulate a single dataset, you can perform a simple technical check, that is the analysis software outputs parameter estimates [@braiekTestingMachineLearning2020], e.g.
-when the simulated data contains missings.
-Third, if you perform a Monte Carlo simulation, you may evaluate how the data analysis copes with expected data complications, e.g.
-non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
-Fourth, you can plan your resources efficiently, e.g. determine required sample size to achieve certain precision or power [@brandmaier2018precision].
 
@@
 
-A PAC offers four advantages over classical preregistration.
+A preregistration as code offers four advantages over classical preregistration.
 First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
@@
 if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
-Third, deviations from the preregistration are more explicit because they are reflected in changes to the code a property facilitated by version control.
+Third, deviations from the preregistration are more explicit because they are reflected in changes to the codes a property facilitated by version control.
 Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
-
-Why are PAC and reproducibility related?
-A reproducible script is supposed to work on the same data at a later point in time.
-PAC requires a script to work on different data at a later point in time, a property we call "reusability".
-This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] so it is build-for-reuse [@al-badareenReusableSoftwareComponents2010].
-Reusability promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
 

### 2021-06-29 manuscript.Rmd
explain how to install external software in Docker

@@
 The `&&` only runs the second command if the first was successful.
-The second step is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy properties.
+Therefore, the computational steps are only executed, when the software environment is set up.
+The second step and actual reproduction is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy properties.
 First, a call to `make` without any explicit target will build the Make target `all`.
@@
 Third, repro constructs Make targets so that if you supply `DOCKER=TRUE` they are executed within the Docker image of the project.
+
+This interplay between Docker and Make is so complicated because it resembles a chicken or egg problem.
+We have computational steps (Make) which depend on the software environment (Docker) for which we again have computational steps that create it.
+Users only require a deeper understanding when they either want to have more complex computational recipes than rendering an R Markdown or require other software than R packages.
+
+Users can have full control over the software installed within the image of the project.
+`repro` creates three Dockerfiles inside the `.repro` directory.
+The file `.repro/Dockerfile_base` contains information about the base image on which all the rest is build.
+By default we rely on the "verse" images provided by the Rocker project.
+These contain (among other) the packages `tidyverse`, `rmarkdown` and a Latex installation, which make them ideal for writing scientific manuscripts.
+Users can choose which R version they want by changing the version number in line 1 to the desired R version number.
+By default the R version corresponds to the version on which `repro::automate()` was called the first time.
+The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed, by default this date is set to the date on which `repro::automate()` was called the first time.
+
+```{r, results='asis', echo=FALSE}
+cat(
+  "```bash",
+  readLines(here::here(".repro", "Dockerfile_base"))[1:3],
+  "```",
+  sep = "\n"
+)
+```
+
+The file `.repro/Dockerfile_manual` is by default empty and is supposed to allow dependencies outside of R, like system libraries or external software.
+The most convinient way to install software on Ubuntu is through its package manager `apt`.
+This snipped installs python:
+
+```bash
+RUN apt-get update && apt-get install -y python3
+```
 

### 2021-06-29 manuscript.Rmd
explain the autogenerated Dockerfile

@@
 
+The the last Docker, `.repro/Dockerfile_packages`, should not be manually changed.
+`repro::automate()` gathers all packages from all `.Rmd` files and figures out if they should be installed from CRAN or GitHub.
+`repro::automate()` also simply combines the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the project.
+The seperation into several Dockerfiles is only to provide a clear seperation between automatically generated parts and manually altered parts of the Dockerfile.
+However, Docker requires in the end the one Dockerfile that `repro::automate()` creates from them.
+
 # Preregistration as Code

### 2021-06-30 R/simulation.R
dont mess around with seeds

@@ -1,24 +1,4 @@
-with_lecuyer <- function(seed, code) {
-  withr::with_seed(
-    seed,
-    code,
-    .rng_kind = "L'Ecuyer-CMRG"
-  )
-}
-
-generate_seeds <- function(k, seed) {
-  with_lecuyer(
-    seed,
-    purrr::accumulate(seq_len(k), function(x, y) # discard y
-      parallel::nextRNGStream(x), .init = .Random.seed)
-  )
-}
-
-error <- function(i){
-  as.numeric(scale(rnorm(i), scale = FALSE)) # only centered
-}
-
-simulate_data_ <- function(n, df, d, i){
+simulate_data <- function(n, df, d, i){
   stopifnot(n %% 2L == 0L)
   group <- rep(c(0L, 1L), each = n/2)
   rand <-  matrix(rchisq(n * i, df/i), ncol = i) # sum of n chisq has df of n*df
@@ -26,11 +6,6 @@ simulate_data_ <- function(n, df, d, i){
   colnames(effect) <- stringr::str_c("x", seq_len(i))
   dplyr::mutate(tibble::as_tibble(effect),
                 group = group)
-}
-
-simulate_data <- function(seed, n, df, d, i = 9, ...){
-  stopifnot(length(list(...)) == 0L)
-  with_lecuyer(seed, simulate_data_(n, df, d, i))
 }
 
 planned_analysis <- function(data){
@@ -59,16 +34,14 @@ simulation_study <- function(setup){
 }
 
 setup <- tidyr::expand_grid(
-  seed = generate_seeds(10, 1002),
   n = seq(10, 1000, 10),
   df = 8, # skew = sqrt(8/df)
-  d = seq(0, .5, 0.1)
-)
+  d = seq(0, .5, 0.1),
+  i = 10)
 
 res <- simulation_study(setup)
 
 test <- res %>% 
-  select(-seed) %>% 
   group_by(across(-results)) %>% 
   unnest_wider(results) %>% 
   summarise(power = mean(p_value < 0.05), .groups = "drop")

### 2021-06-30 R/simulation.R
only use base R in the fundamental functions

@@ -3,13 +3,14 @@ simulate_data <- function(n, df, d, i){
   group <- rep(c(0L, 1L), each = n/2)
   rand <-  matrix(rchisq(n * i, df/i), ncol = i) # sum of n chisq has df of n*df
   effect <- rand + d * group
-  colnames(effect) <- stringr::str_c("x", seq_len(i))
-  dplyr::mutate(tibble::as_tibble(effect),
-                group = group)
+  colnames(effect) <- paste0("x", seq_len(i))
+  effect <- as.data.frame(effect)
+  effect$group <- group
+  return(effect)
 }
 
 planned_analysis <- function(data){
-  x <- rowMeans(dplyr::select(data, dplyr::starts_with("x")))
+  x <- rowMeans(data["group" != names(data)])
   y <- as.factor(data$group)
   skew <- moments::skewness(x)
   # skewness cutoff
@@ -17,6 +18,7 @@ planned_analysis <- function(data){
   if(use_rank){
     x <- rank(x)
   }
+  x <- scale(x)
   test <- t.test(x ~ y)
   list(test = test, skew = skew, use_rank = use_rank, n = length(x))
 }

### 2021-07-02 R/simulation.R
verify that simulation works

@@ -2,23 +2,24 @@ simulate_data <- function(n, df, d, i){
   stopifnot(n %% 2L == 0L)
   group <- rep(c(0L, 1L), each = n/2)
   rand <-  matrix(rchisq(n * i, df/i), ncol = i) # sum of n chisq has df of n*df
-  effect <- rand + d * group
+  d_scaled <- d * sqrt(2*df)/i # scale d to express sd units (var(chisq) = 2df)
+  #browser()
+  effect <- rand + d_scaled * group
   colnames(effect) <- paste0("x", seq_len(i))
   effect <- as.data.frame(effect)
   effect$group <- group
   return(effect)
 }
 
-planned_analysis <- function(data){
+planned_analysis <- function(data, use_rank = "skew"){
   x <- rowMeans(data["group" != names(data)])
   y <- as.factor(data$group)
   skew <- moments::skewness(x)
   # skewness cutoff
-  use_rank <- abs(skew) > 1
+  if(use_rank == "skew")use_rank <- abs(skew) > 1
   if(use_rank){
     x <- rank(x)
   }
-  x <- scale(x)
   test <- t.test(x ~ y)
   list(test = test, skew = skew, use_rank = use_rank, n = length(x))
 }
@@ -27,6 +28,27 @@ extract_results <- function(analysis){
   list(p_value = analysis$test$p.value,
        skew = analysis$skew)
 }
+
+t2d <- function(t, n1, n2)t * sqrt(((n1 + n2)/(n1 * n2)) * (n1 + n2)/(n1 + n2 -2))
+parameter_recovery <- function(n, df, d, i, rank = FALSE){
+  t2d(planned_analysis(simulate_data(n, df, d, i), rank)$test$statistic, n/2, n/2)
+}
+
+test <- simulate_data(1000000, 100, 0.1, 10)
+m1 <- filter(test, group == 1) %>% 
+  select(-group )%>% 
+  rowSums() %>% 
+  mean()
+m0 <- filter(test, group == 0) %>% 
+  select(-group )%>% 
+  rowSums() %>% 
+  mean()
+sd <- test %>% rowSums() %>% sd()
+
+(m1 - m0)/sd
+
+t2d(planned_analysis(test)$test$statistic, 1000000/2, 1000000/2)
+parameter_recovery(1000000, 100, 0.1, 10)
 
 simulation_study <- function(setup){
   all_steps <- purrr::compose(extract_results, planned_analysis, simulate_data)

### 2021-07-02 R/simulation.R
scale simulation

@@ -1,6 +1,6 @@
 simulate_data <- function(n, df, d, i){
   stopifnot(n %% 2L == 0L)
-  group <- rep(c(0L, 1L), each = n/2)
+  group <- rep(c(1L, 0L), each = n/2)
   rand <-  matrix(rchisq(n * i, df/i), ncol = i) # sum of n chisq has df of n*df
   d_scaled <- d * sqrt(2*df)/i # scale d to express sd units (var(chisq) = 2df)
   #browser()
@@ -11,12 +11,12 @@ simulate_data <- function(n, df, d, i){
   return(effect)
 }
 
-planned_analysis <- function(data, use_rank = "skew"){
+planned_analysis <- function(data, use_rank = "skew", skew_cutoff = 1){
   x <- rowMeans(data["group" != names(data)])
   y <- as.factor(data$group)
   skew <- moments::skewness(x)
   # skewness cutoff
-  if(use_rank == "skew")use_rank <- abs(skew) > 1
+  if(use_rank == "skew")use_rank <- abs(skew) > skew_cutoff
   if(use_rank){
     x <- rank(x)
   }
@@ -24,48 +24,33 @@ planned_analysis <- function(data, use_rank = "skew"){
   list(test = test, skew = skew, use_rank = use_rank, n = length(x))
 }
 
-extract_results <- function(analysis){
-  list(p_value = analysis$test$p.value,
-       skew = analysis$skew)
+#t2d <- function(t, n1, n2)t * sqrt(((n1 + n2)/(n1 * n2)) * (n1 + n2)/(n1 + n2 -2))
+t2d <- function(test){
+  unname(2 * test$statistic/sqrt(test$parameter))
 }
-
-t2d <- function(t, n1, n2)t * sqrt(((n1 + n2)/(n1 * n2)) * (n1 + n2)/(n1 + n2 -2))
 parameter_recovery <- function(n, df, d, i, rank = FALSE){
   t2d(planned_analysis(simulate_data(n, df, d, i), rank)$test$statistic, n/2, n/2)
 }
 
-test <- simulate_data(1000000, 100, 0.1, 10)
-m1 <- filter(test, group == 1) %>% 
-  select(-group )%>% 
-  rowSums() %>% 
-  mean()
-m0 <- filter(test, group == 0) %>% 
-  select(-group )%>% 
-  rowSums() %>% 
-  mean()
-sd <- test %>% rowSums() %>% sd()
+extract_results <- function(analysis){
+  list(cohend = t2d(analysis$test),
[…]

### 2021-07-02 R/simulation.R
oops switch y <-> x

@@ -12,16 +12,16 @@ simulate_data <- function(n, df, d, i){
 }
 
 planned_analysis <- function(data, use_rank = "skew", skew_cutoff = 1){
-  x <- rowMeans(data["group" != names(data)])
-  y <- as.factor(data$group)
-  skew <- moments::skewness(x)
+  y <- rowMeans(data["group" != names(data)])
+  x <- as.factor(data$group)
+  skew <- moments::skewness(y)
   # skewness cutoff
   if(use_rank == "skew")use_rank <- abs(skew) > skew_cutoff
   if(use_rank){
-    x <- rank(x)
+    y <- rank(y)
   }
-  test <- t.test(x ~ y)
-  list(test = test, skew = skew, use_rank = use_rank, n = length(x))
+  test <- t.test(y ~ x)
+  list(test = test, skew = skew, use_rank = use_rank, n = length(y))
 }
 
 #t2d <- function(t, n1, n2)t * sqrt(((n1 + n2)/(n1 * n2)) * (n1 + n2)/(n1 + n2 -2))

### 2021-07-04 R/simulation.R
split again in simulation functions and script

@@ -1,56 +1,36 @@
-simulate_data <- function(n, df, d, i){
-  stopifnot(n %% 2L == 0L)
-  group <- rep(c(1L, 0L), each = n/2)
-  rand <-  matrix(rchisq(n * i, df/i), ncol = i) # sum of n chisq has df of n*df
-  d_scaled <- d * sqrt(2*df)/i # scale d to express sd units (var(chisq) = 2df)
-  #browser()
-  effect <- rand + d_scaled * group
-  colnames(effect) <- paste0("x", seq_len(i))
-  effect <- as.data.frame(effect)
-  effect$group <- group
-  return(effect)
-}
+library(tidyverse)
+source(here::here("R", "simulation_funs.R"))
+# if you have access to a hpc envir specify this in R/hpc.R
+# see https://github.com/aaronpeikert/repro-tutorial/blob/hpc/R/hpc.R
+# or `git checkout hpc R/hpc.R`
+# if no hpc is availible we use local multicore with all available cores
+# to speed up consider reduce nsim; increase nstep ↓↓↓
 
-planned_analysis <- function(data, use_rank = "skew", skew_cutoff = 1){
-  y <- rowMeans(data["group" != names(data)])
-  x <- as.factor(data$group)
-  skew <- moments::skewness(y)
-  # skewness cutoff
-  if(use_rank == "skew")use_rank <- abs(skew) > skew_cutoff
-  if(use_rank){
-    y <- rank(y)
-  }
-  test <- t.test(y ~ x)
-  list(test = test, skew = skew, use_rank = use_rank, n = length(y))
+hpc_config <- here::here("R", "hpc.R")
+if(fs::file_exists(hpc_config)){
+  source(hpc_config)
+} else {
+  plan(list(transparent,
+            tweak(multisession)))
 }
+trans_seq <- function(min, max, length, trans = function(x)x^2){
+  stopifnot(is.function(trans))
+  lin <- seq(min, max, length.out = length)
+  transformed <- trans(lin)
+  scaled <- (transformed - min(transformed) + min)/max(transformed)*max
+  scaled
+}
+setup <- tidyr::expand_grid(
+  n = c(10, seq(100, 1000, 100)),
+  df = 8, # skew = sqrt(8/df)
+  d = seq(0, .5, 0.05),
+  i = 10)
+res_raw %<-% simulation_study(setup, 10000, 1235)
+res <- res_raw %>% 
+  group_by(across(-results)) %>% 
[…]

### 2021-07-04 R/simulation_funs.R

@@ -1,0 +1,52 @@
+#----simulate_data----
+simulate_data <- function(n, df, d, i){
+  stopifnot(n %% 2L == 0L)
+  group <- rep(c(1L, 0L), each = n/2)
+  rand <-  matrix(rchisq(n * i, df/i), ncol = i) # sum of n chisq has df of n*df
+  d_scaled <- d * sqrt(2*df)/i # scale d to express sd units (var(chisq) = 2df)
+  #browser()
+  effect <- rand + d_scaled * group
+  colnames(effect) <- paste0("x", seq_len(i))
+  effect <- as.data.frame(effect)
+  effect$group <- group
+  return(effect)
+}
+#----planned_analysis----
+planned_analysis <- function(data, use_rank = "skew", skew_cutoff = 1){
+  y <- rowMeans(data["group" != names(data)])
+  x <- as.factor(data$group)
+  skew <- moments::skewness(y)
+  # skewness cutoff
+  if(use_rank == "skew")use_rank <- abs(skew) > skew_cutoff
+  if(use_rank){
+    y <- rank(y)
+  }
+  test <- t.test(y ~ x)
+  list(test = test, skew = skew, use_rank = use_rank, n = length(y))
+}
+#----extract_funs----
+#t2d <- function(t, n1, n2)t * sqrt(((n1 + n2)/(n1 * n2)) * (n1 + n2)/(n1 + n2 -2))
+t2d <- function(test){
+  unname(2 * test$statistic/sqrt(test$parameter))
+}
+parameter_recovery <- function(n, df, d, i, rank = FALSE){
+  t2d(planned_analysis(simulate_data(n, df, d, i), rank)$test$statistic, n/2, n/2)
+}
+
+extract_results <- function(analysis){
+  list(cohend = t2d(analysis$test),
+       p_value = analysis$test$p.value,
+       skew = analysis$skew)
+}
+
+#----simulation_study----
+simulation_study_ <- function(setup){
+  all_steps <- purrr::compose(extract_results, planned_analysis, simulate_data)
+  out <- dplyr::mutate(setup, results = furrr::future_pmap(setup, all_steps, .options = furrr::furrr_options(seed = TRUE)))
+  #tidyr::unnest(out, results)
+  out
+}
+
+simulation_study <- function(setup, k, seed = NULL){
+  furrr::future_map_dfr(seq_len(k), function(irrelevant)simulation_study_(setup), .options = furrr::furrr_options(seed = seed))
+}

### 2021-07-04 manuscript.Rmd
add packages

--- 2021-06-29 manuscript.Rmd
@@
 ---
-title: A Hitchhiker's Guide to Reproducible Research in R
-author:
-  - name: Aaron Peikert
-    affil: 1, *
-    orcid: 0000-0001-7813-818X
-  - name: Caspar J. Van Lissa
-    affil: 2, 3
-    orcid: 0000-0002-0808-5024
-  - name: Andreas M. Brandmaier
-    affil: 1, 4
-    orcid: 0000-0001-8765-6982
-affiliation:
-  - num: 1
-    address: |
-      Center for Lifespan Psychology---Max Planck Institute for Human Development
-      Lentzeallee 94, 14195 Berlin, Germany
-  - num: 2
-    address: |
-      Department of Methodology & Statistics---Utrecht University faculty of Social and Behavioral Sciences, Utrecht, Netherlands
-  - num: 3
-    address: |
-      Open Science Community Utrecht, Utrecht, Netherlands
-  - num: 4
-    address: |
-      Max Planck UCL Centre for Computational Psychiatry and Ageing Research
-      Berlin, Germany and London, UK
-# firstnote to eighthnote
-correspondence: |
-  peikert@mpib-berlin.mpg.de
-journal: psych
-type: tutorial
-status: submit
-bibliography: temp.bib
-appendix: appendix.tex
-simplesummary: |
-  A Simple summary goes here.
-abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
-keywords: |
-  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
-  to the article, yet reasonably common within the subject discipline.).
-acknowledgement: |
-  All sources of funding of the study should be disclosed. Please clearly 
-  indicate grants that you have received in support of your research work. 
-  Clearly state if you received funds for covering the costs to publish in open 
-  access.
-authorcontributions: |
-  For research articles with several authors, a short paragraph specifying their 
-  individual contributions must be provided. The following statements should be 
-  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
-  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
-  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
-  limited to those who have contributed substantially to the work reported.
-conflictsofinterest: |
-  The authors declare no conflict of interest.
-abbreviations:
-  - short: GUI
-    long: Graphical User Interface
-  - short: CLI
-    long: Comand Line Interface
+title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
+author: Aaron Peikert and Andreas M. Brandmaier 
+output:
+  bookdown::pdf_document2:
+    keep_tex: TRUE
+    toc: FALSE
 repro:
@@
     - rticles
+    - moments
+    - report
   scripts:
@@
     - R/link.R
-output:
-  bookdown::markdown_document2:
-    base_format: rticles::mdpi_article
-header-includes:
-   - \usepackage[labelformat=empty]{caption}
-   - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, numbers=left}
+csl: apa7.csl
+numbered_headings: true
+bibliography: ["references.bib", "packages.bib"]
+abstract: "`r tryCatch(trimws(readr::read_file('abstract.Rmd')))`"
 ---
+
+<!-- the HTML comments, like this one, are meta comments, mainly describing the intent --->
+<!-- each sentence below a heading summarizes what I want to say there --->
+<!-- "Hands-on:" means concrete practical application, they roughly proceed from easy/familiar to hard/unfamiliar-->
 
@@
 Increasingly, scientists openly share their data, materials, and analysis code.
-Sharing the digital research output increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
-The technology is now available to make all artefacts related to a research project available so that the project can be reproduced with minimal effort.
-Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
+Sharing these digital artifacts increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results.
+However, making a research project reproducible comes with a variety of challenges and there is a dire need for standards of reproducibility.
+
+This paper has two aims.
+First, we want to introduce the reader to a workflow [see @Peikert2019 for details] that ensures reproducibility---defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code.
+The research community has long accepted reproducibility as integral to "good scientific practice", because empirical research hinges on its objectivity meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
 22][^popper]
-In contemporary research, reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
-Unfortunately, despite increasing commitment to open science practices, many projects can not yet be reproduced by other research teams.
-This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
+Often, however, projects can not be reproduced by other research teams, despite the best efforts by the creators.
+We present four common threats to reproducibility and explain how we can address these with a combination of four solutions from the software engineering community.
+Specifically, we demonstrate how the R package `repro` supports researchers in implementing this solution.
+The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
 
-This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
-We pursue two aims.
-First, the present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community that eliminate four common threats to reproducibility [see @Peikert2019 for details].
-Specifically, we demonstrate how the R package `repro` supports researchers in implementing these solutions.
-The second objective is to discuss opportunities of such workflow to improve not only how research outputs are created and recreated but also how research itself is conducted.
-One such opportunity is an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
-This dynamic document closely resembles an actual journal article but with "fake" results and without discussion of results, then serves as preregistration.
+[^popper]: In fact @popperLogicScientificDiscovery2002's famous criterion of demarcation is build around "inter-subjective testing", a concept which he later generalized to inter-subjective criticism (Footnote *¹, p.
+22).
+Somewhat confusingly @popperLogicScientificDiscovery2002 uses the term "Reproducibility" for what we would call "Replication".
 
-[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalised to inter-subjective criticism (Footnote *1, p.
-22).
-Somewhat confusingly, @popperLogicScientificDiscovery2002 uses the term "Reproducibility" for what we would call "Replication".
-
+This tutorial is aimed at researchers who are willing to make code, data, and materials available and should help them to ensure reproducibility of relevant statistical results.
 While the willingness to share materials is required for reproducing a project, it is not sufficient.
@@
 @obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
-The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
+The authors could execute only 31 of the scripts without error and were able to reproduce the results of only 21 articles (within a reasonable period of time).
+
+Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.
 These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
-The results further illustrate that open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
-
-We argue that the benefits associated with transparency increase if research can additionally be reproduced.
-This paper tutorial demonstrates how to make research projects reproducible, while striking a balance between rigor and ease-of-use.
+To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use.
 A rigorous standard increases the likelihood of any individual project to be actually reproducible.
-An easy to use standard, on the other hand, is more likely to be adopted.
+An easy to use standard on the other hand is more likely to be employed across research projects.
 The remainder of this paper is structured along these lines.
@@
 
+After introducing the conceptual building blocks, this tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
 We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.
@@
 Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).
-To guide the reader through all steps of a research project, we will use an exemplary research question:
+To guide the reader through all steps of a research project, we will use a exemplary research question:
 
-> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"
+> "Is there a mean difference in the personality trait 'Machiavellism' between genders?"
 
@@
 
+<!-- could andreas write this? --->
 We identified the following common threat to reproducibility: <!-- CJ How did you identify these causes? Is this exhaustive?-->
 
-1.	**Multiple inconsistent versions of code, data, or both**; for example, the dataset could have changed over time because outliers were removed at a later stage, or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
-2.	**Copy-and-paste**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	**Ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order; 
-4.	**External code dependencies**; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist; or, a different version of the same software may produce different results.
+1.	The existence of multiple inconsistent versions of code, data, or both; for example, the dataset could have changed over time because outliers were removed at a later stage, or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
+2.	Copy-and-paste; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
+3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order; 
+4.	Code dependencies; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist; or, a different version of the same software may produce different results.
 <!-- CJ 5.
@@
 
-The remainder of this section briefly explains why each of these four building blocks is needed and their role in ensuring reproducibility.
-
-First, to resolve ambiguity across multiple versions of code and data that come into existence during the development phase of a research project, we recommend using a **version control** system.
-Version control allows a clear link between which results were generated by which version of code and data.
+In the remainder of this section, we will briefly explain why each of these four building blocks is needed and how it plays a role in guaranteeing reproducibility.
+To resolve ambiguity across multiple versions of code and data that may be come into existence during the development phase of a research project, we recommend the use of a version control system.
 A version control system tracks changes to all project-related files (e.g., materials, data, and code) over time.
 At a later stage, single files or the entire project can be compared to or reverted to an earlier version.
-A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
+Additional features such as branches allow parallel versions to exist that can be merged later.
+Specific snapshots of a repository can be marked, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
+Using version control protects against loss of files and also supports remote collaboration with semi-automated tools for tracking and merging changes.
 
-We recommend using **Git** for version control, because of its widespread adoption in the R community.
-Git is built around snapshots that represent the entire project state at a given point in time.
-Those snapshots are called "commits".
-Each file that was changed has to be added to a commit, and each commit requires a commit message that succinctly describes the changes made.
-Each commit "knows" its ancestor, and they form a timeline together.
-The conglomerate of commits (i.e. the version-controlled project) is called a repository.
-In Git, specific snapshots of a repository can be tagged, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
-Git has additional features beyond basic version control, such as branches to facilitate collaboration.
-To collaborate via Git,<!--dangling modify, cant help it--> the repository has to be uploaded somewhere.
-We recommend GitHub to host Git repositories because of its popularity among R users.
-GitHub has many tools that ease project management and collaboration, and these tools provide much value in our everyday work, but they are not central to archiving reproducibility.
-
-Second, we rely on **dynamic document generation**.
-The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
-In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
-Both steps are linked by manually copying and pasting results from one software to the other, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
-
-Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured, who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
-**R Markdown** uses Markdown for writing the text and R (or other programming languages) for the analysis.
-Markdown is a lightweight text format in plain text with minimal reserved symbols for formatting instructions.
-This way, Markdown does not need any specialized software for editing.
-It is highly fool-proof (unlike, for example, LaTeX), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your CV [@vitae].
+Second, we rely on dynamic document generation.
+The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps.
+Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
+This paradigm has gained popularity with the Sweave approach [@leisch2002sweave] that allowed embedding the data analysis into a TeX document, which then could be dynamically rendered into a final manuscript.
+As of recently, R Markdown has added to the success of this idea.
+Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
+This way, Markdown does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
 R Markdown extends regular Markdown by allowing users to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
-Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
+Upon rendering of the document, the code blocks are executed and their output is dynamically inserted into the document.
 This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
 
-One would hope that sharing a literate statistical analysis would allow others (including one own's future self) to simply download and run (ie., reproduce) a given analysis.
-However, while version control and dynamic document generation are becoming more common @, this approach fails to guarantee reproducibility in most instances.
-In practice, dependencies between project files (e.g.
-the raw data was preprocessed) or on external software (such as system libraries or components of the programming language) are frequently unmentioned.
-
-To automatically resolve dependencies between project files, we rely on **dependency tracking**.
-In essence, researchers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
-Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
-In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, analysis results).
-
-We recommend using **Make** for dependency tracking because it is language independent.
-To make this more concrete, consider the following hypothetical example, in which a research project contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
-A Makefile could then look like this:
-
-```{bash, eval=FALSE}
-manuscript.pdf: manuscript.Rmd simulation_results.csv 
-  Rscript -e 'rmarkdown::render("manuscript.Rmd")'
-
-simulation_results.csv: simulate.R
-  Rscript -e 'source("simulate.R")'
-```
-
-There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).
-Each target is followed by a colon and a list of requirements.
-If a requirement is newer than the target, the recipe will be executed to rebuild the target.
-If a requirement does not exist, Make uses a recipe to build the requirement first before building the target.
-Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would make sure that the file `simulation_results.csv` exists; if not, it would find that this is also a target and issue the command in l.5 to run the simulation first before rendering the manuscript.
-This way, we ensure that the simulation is always run before the manuscript is built and that the manuscript is rebuilt if the simulation code was changed.
-
-A version controlled dynamic document with dependency tracking still relies on external software.
-Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
-
-```{r, include=FALSE}
-container_size <- as.numeric(fs::file_size("reprotutorial.sif"))/1024^3
-```
-
-One solution to the problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
-One comprehensive approach to software management is containerization.
-"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
-Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
-The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
-Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
-Such a snapshot of the software stack is called an "image".
-Packaging all needed software in such an image requires considerable amounts of storage space.
-Two major strategies are helping to keep the storage requirements reasonable.
-First, there is a community that maintains images for particular purposes.
-For example, there are ready-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
-Users can then install whatever they need in addition to what is provided by these pre-compiled images.
-`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
-Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
-Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
-This recipe is a textual description of all commands that need to be executed to recreate the image.
-Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
-However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
-For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
-
-To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), containerization (using Docker).
-While R Markdown and git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which represents a considerable hurdle in the acceptance and implementation of the workflow.
-To remove this hurdle, we have developed a new R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
-In the remainder, we will walk you through a complete research project, from sketching the very first idea over preregistration, data analysis, and submission of a preprint while focusing on how the `repro`-package enables users to maintain reproducibility of all steps.
+<!-- TODO Andreas: Internal and externdal dependencies -->
 
@@
 # file gets downloaded in Makefile
-knitr::include_graphics(here::here("images/nutshell.svg"), auto_pdf = TRUE)
+knitr::include_graphics("images/nutshell.svg", auto_pdf = TRUE)
 ```
@@
 
-A significant hurdle for adopting the before-mentioned tools for researchers is that they are typically used through a command-line interface (CLI).
-The CLI predates and coexists with graphical user interfaces (GUIs) which many users (including the authors) prefer over the text-based interactions of a CLI for most tasks.
-A GUI has the apparent disadvantages for reproducibility that it requires a user who clicks on the correct buttons in the correct order.
-Depending on the number of steps, this is not only much effort but also error-prone.
-This tutorial seeks a compromise.
-We use simpler alternatives, including a GUI where we can, to create a reproducible workflow, but the reproduction of such workflow does not rely on manual steps.
-At several points, we rely on or have self-developed a CLI that wraps a more complicated (and powerful) CLI.
-These wrappers use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
-We hope this makes reproducibility tools more accessible by enabling untrained users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
-These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
-Please install the package [`gert`](https://github.com/r-lib/gert)[@R-gert] which wraps Git commands and [`repro`](https://github.com/aaronpeikert)[@R-repro] so they can act as assistance systems while you follow the tutorial:
-
-```{r, eval=FALSE}
-install.packages("gert")
-# repro is not on CRAN yet
-install.packages('repro', repos = 'https://aaronpeikert.r-universe.dev')
-```
-
-One example for such assistance are the check functions from `repro`.
-They verify that you indeed have installed and setup the required software for this workflow:
-
-```{r}
-# `package::function()` → use function from package without `library(package)`
-repro::check_git()
-repro::check_make()
-repro::check_docker()
-```
-
-They detect the users' system state and explain what this means to the user.
-Sometimes they ask the user to do something, e.g. the following happens if you were a windows user who does not have installed Git:
-
-```{r, include=FALSE}
-opts <- options()
-options(repro.os = "windows",
-        repro.git = FALSE)
-```
-
-```{r}
-check_git()
-```
-
-```{r, include=FALSE}
-options(opts)
-```
-
-The messages emitted by `repro` and other packages we present try to take the user by the hand to help them solve problems.
-This tutorial describes steps that we expect to apply to all users; the messages from the packages fill the gap to your specific situation and extend the tutorial in this way.
-Before you continue, we ask you to run the above commands to check Git, Make, and Docker.
-If necessary, follow the instructions of repro to install them.
-
-The first concept we want to apply practically is version control with Git.
-Git tracks files within one project folder.
-We start by creating such project folder with RStudio:
-
-> File → New Project... → New Directory → Example Repro Template
-
-This creates a project containing an example analysis, but you may use any other template or existing R Project.
-
-You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore` file.
-You can add something to the `.gitignore` file directly or with this command:
-
-```{r, eval=FALSE}
-usethis::use_git_ignore("private.md")
-```
-
-Now the file `private.md` will not be committed to Git and will not be made public.
-
-After you have ignored sensitive files (like not yet anonymized data), you can activate Git with:
-
-```{r, eval=FALSE}
-usethis::use_git()
-```
-
-If you made changes to a file, you have to stage and then commit these changes.
-The easiest way is to use the RStudio Git pane.
-You click on the empty box next to the file you want to stage, and when a tick appears, the file is staged.
-After you have staged all files you want, click on the commit button, explain curtly why you made those changes in a commit message and then click on commit.
-
-The files you created and the changes you made still have not left your computer.
-To back up and share the files, you can publicly upload those changes to GitHub with:
-
-```{r, eval=FALSE}
-usethis::use_github()
-```
-
-Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow its suggestions.
-
-If you are not yet ready to share your project publicly use:
-
-```{r, eval=FALSE}
-usethis::use_github(private = TRUE)
-```
-
-Now that you have created a project that is under version control, we turn to dynamic document generation.
-Three aspects are essential to understand a dynamic document:
-
-1. Text
-2. Code
-3. Metadata
-
-The text of an R Markdown is written in Markdown.
-The following example shows how to signify a header, a bold word and a list in Markdown:
-
-```markdown
-<!--this is a Markdown file -->
-# Header
-
-Normal text.
-Important **word** in bold.
-
-To do list:
-
-* do research
-* do more research
-* spend time with family
-```
-
-Code is separated by three backticks (and the programming language in curly braces) like here:
-
-````markdown
-This is normal text, written in Markdown.
-
-`r ''````{r}
-# this is R code
-1 + 1
-```
-````
-
-Now that you know how to write text and R code in an RMarkdown, you need to know about metadata, like the title or the output format.
-Metadata are placed at the beginning of the document and are separated by three dashes; the following example is a full markdown document where the header is in lines 1-6.
-
-````markdown
----
-title: A Hitchhiker's Guide to Reproducible Research in R
-author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
-abstract: A paper about recalculating the same thing more than once.
-output: html_document
----
-
-# Introduction
-
-Important for reproducibility:
-
-1. Version Control
-2. Dynamic Document Creation
-3. Dependency Tracking
-4. Software Management
-
-`r ''````{r}
-# this is R code
-t.test(extra ~ group, data = sleep)
-```
-````
-
-Most metadata fields are self-explanatory (like the author field), but any package may use the metadata for their purposes.
-Often, for example, does a specific output format require more metadata, i.e. the `rticles`-package for authoring scientific manuscripts.
-RMarkdown was extended to hundreds of document formats which are often highly configurable by their metadata.
-The `repro` package uses the metadata, not to change the output, but to make the document more reproducible, simply by listing resources, i.e., scripts, data, and packages, your document depends on, in the following format:
-
-```markdown
----
-title: A Hitchhiker's Guide to Reproducible Research in R
-author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
-abstract: A paper about recalculating the same thing more than once.
-output: html_document
-repro:
-  scripts:
-    - R/load.R
-  data:
-    - data/mtcars.csv
-  packages:
-    - tidyverse
-    - usethis
-    - gert
----
-```
-
-This information makes internal dependencies to other files in the project and external dependencies to R packages explicit.
-`repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
-In day-to-day work the simplest solution is to use the function `repro::automate()` which both infers a Makefile and a Dockerfile from all RMarkdown  (all files with the ending `.Rmd`) and there respective repro metadata.
-Every time you change the `repro`-metadata, change the output format, or add a new RMarkdown, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
-The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
-Besides the main Makefile and the Dockerfile repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
-
-When you or people who received your project from GitHub want to reproduce the project, they can ask repro to explain how this should be done:
-
-```{r}
-repro::rerun()
-```
-
-However, users do not have to have repro installed, in fact, they do not even need R installed. 
-The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.
-They can be pasted into a terminal and will reproduce the project in many cases.
-Some projects, however, are not as easy to make reproducible.
-As explained above, repro is merely a simplified interface.
-These simplifications are possible because of two restrictions.
-First, the project can only rely on R packages for the Dockerfile and second, it can have no multilayered preprocessing for the Makefile.
-Lifting these restrictions requires the user to interact more directly with Make or Docker.
-To satisfy more complicated requirements users need to understand how repro utilizes Make and Docker internally.
-
-Let us dissect the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two steps.
-First it virtually recreates the software environment and then it executes computational recipes in the virtual software environment.
-The first step is represented by the command `make docker`.
-The command `make docker` will trigger Make to build the target `docker`.
-The recipe for this target builds an image from your Dockerfile.
-The `&&` only runs the second command if the first was successful.
-The second step is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy properties.
-First, a call to `make` without any explicit target will build the Make target `all`.
-Second, the flag `-B` means that Make will consider all prerequisites as outdated and hence will rebuild everything.
-Third, repro constructs Make targets so that if you supply `DOCKER=TRUE` they are executed within the Docker image of the project.
-
 # Preregistration as Code
@@
 We argue that a preregistration as code (PAC) is an excellent planning tool that offers several benefits over traditional preregistration.
-For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data using simulated data.
+For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data.
 
@@
 if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
-Third, deviations from the preregistration are more explicit because they are reflected in changes to the codes a property facilitated by version control.
+Third, deviations from the preregistration are more explicit because they are reflected in changes to the code a property facilitated by version control.
 Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
@@
 ```{r, include=FALSE}
-knitr::write_bib(
-  c(
-    .packages(),
-    "repro",
-    "here",
-    "rticles",
-    "gert",
-    "bookdown",
-    "lavaan",
-    "knitr",
-    "targets",
-    "renv",
-    "tidyverse"
-  ),
-  here("packages.bib")
-)
-cat(c(readLines(here::here("packages.bib")), "\n", readLines(here::here("references.bib"))), sep = "\n", file = "temp.bib")
+knitr::write_bib(c(.packages(), "here", "rticles", "gert", "bookdown", "lavaan", "knitr", "targets", "renv", "tidyverse"),
+                 "packages.bib")
 ```

### 2021-07-04 manuscript.Rmd
describe example a bit

@@
 
+# Preregistration as Code --- an example
+
+At the heart of a PAC lies the planned analysis.
+We suggest writing a function that captures the full process of testing a research question beginning from the dataset.
+This is a simple example:
+
+```{r planned_analysis}
+```
+
+All variables of the dataset except the grouping variable are averaged.
+Depending on how skewed the dataset is we apply a rank transformation to the average and then we calculate a t-test with Welch correction in the hope of creating a more robust analysis for non-normal data.
+At last the function returns an object with all relevant information.
+
+To use this function one needs a compatible dataset.
+At the time of preregistration there is of course no actual data availible.
+We therefore simulate a dataset.
+
+```{r simulate_data}
+```
+
+We decided to draw non-normal data to test how well our procedure performed under non-optimal situations.
+Specifically, we draw from a $\Chi^2$-distribution, where in one group we add difference that conforms to a standardized effect size.
+This simulated dataset fits perfectly into our planned analysis:
+
+```{r}
+set.seed(112358) # fix state of RNG
+simulate_data(n = 400, # n/2 per group
+              df = 8, # Chisq with skewness == 1
+              d = .2, # Cohen's D
+              i = 10) %>% # number of items (should not effect outcome)
+  planned_analysis() %>% # `%>%` = output of 1. function becomes argument 2.
+  .$test %>% # extract the t.test object
+  report::report_text() # use the `report`-package
+```
+
 # References

### 2021-07-05 R/simulation.R
write csv out

@@ -29,7 +29,7 @@ res_raw %<-% simulation_study(setup, 10000, 1235)
 res <- res_raw %>% 
   group_by(across(-results)) %>% 
   unnest_wider(results)
-
+write_csv(res, here::here("data", "simulation_results.csv"))
 res %>%
   summarise(power = mean(p_value < 0.025)) %>% 
   ggplot(aes(n, power, color = d, group = d)) +

### 2021-07-05 R/simulation.R
summarise results before saving to disk

@@ -13,24 +13,26 @@ if(fs::file_exists(hpc_config)){
   plan(list(transparent,
             tweak(multisession)))
 }
-trans_seq <- function(min, max, length, trans = function(x)x^2){
-  stopifnot(is.function(trans))
-  lin <- seq(min, max, length.out = length)
-  transformed <- trans(lin)
-  scaled <- (transformed - min(transformed) + min)/max(transformed)*max
-  scaled
-}
+
 setup <- tidyr::expand_grid(
   n = c(10, seq(100, 1000, 100)),
   df = 8, # skew = sqrt(8/df)
   d = seq(0, .5, 0.05),
   i = 10)
-res_raw %<-% simulation_study(setup, 10000, 1235)
+res_raw %<-% simulation_study(setup, 100, 1235)
+
 res <- res_raw %>% 
   group_by(across(-results)) %>% 
-  unnest_wider(results)
+  unnest_wider(results) %>% 
+  summarise(power = mean(p_value < 0.025),
+            cohend = mean(cohend),
+            cohend_sd = sd(cohend),
+            .groups = "drop")
+fs::dir_create(here::here("data"))
 write_csv(res, here::here("data", "simulation_results.csv"))
-res %>%
-  summarise(power = mean(p_value < 0.025)) %>% 
+
+res %>% 
   ggplot(aes(n, power, color = d, group = d)) +
-  geom_line()
+  geom_line() + 
+  theme_minimal() +
+  NULL

### 2021-07-05 manuscript.Rmd
move introduction of hypothetical research question

@@
 Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).
-To guide the reader through all steps of a research project, we will use an exemplary research question:
-
-> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"
-
-This example serves only didactic purposes and not to derive substantive claims.
-Nevertheless, everything else is kept as close to a real research question as possible.
 
@@
 
+To guide the reader through all steps of a research project, we will use an exemplary research question:
+
+> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"
+
+This example serves only didactic purposes and not to derive substantive claims.
+Nevertheless, everything else is kept as close to a real research question as possible.
+
 At the heart of a PAC lies the planned analysis.
-We suggest writing a function that captures the full process of testing a research question beginning from the dataset.
-This is a simple example:
+We suggest writing a function that captures the full process of testing a research question that receives a dataset as input.
+For the hypothetical research question, we could for example plan to conduct a simple Student's t-test with Welch's correction an the average of machiavillism items between the gender groups.
+As an example of data dependend decisions we included a rank transformation if the skewness greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+This example is suboptimal with regard to several criteria (untested assumption of measurement invariance, underestimation of effectsize in presence of measurement error, overestimation of effectsize for highly skewed distributions), we merely choose it for didactic purposes because similar procedures a widely used (@horton reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005).
+
+[^caution-mww]: The Mann–Whitney–Wilcoxon is **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
+The interested reader can use the provided code for the simulation to verify that the t.test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
 
@@
 We decided to draw non-normal data to test how well our procedure performed under non-optimal situations.
-Specifically, we draw from a $\Chi^2$-distribution, where in one group we add difference that conforms to a standardized effect size.
+Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a standardized effect size.
 This simulated dataset fits perfectly into our planned analysis:

### 2021-07-05 manuscript.Rmd
add citation for analytic procedure

@@
 We suggest writing a function that captures the full process of testing a research question that receives a dataset as input.
-For the hypothetical research question, we could for example plan to conduct a simple Student's t-test with Welch's correction an the average of machiavillism items between the gender groups.
-As an example of data dependend decisions we included a rank transformation if the skewness greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
-This example is suboptimal with regard to several criteria (untested assumption of measurement invariance, underestimation of effectsize in presence of measurement error, overestimation of effectsize for highly skewed distributions), we merely choose it for didactic purposes because similar procedures a widely used (@horton reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005).
+For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavillism items of the Short Dard Triad (SD3) between the gender groups.
+As an example of a data dependent decisions we included a rank transformation if the skewness greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+This analyitical strategie is suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions[^caution-mww] [@stonehouseRobustnessTestsCombined1998]); we merely choose it for didactic purposes because similar procedures a widely used (@horton reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005).
 

### 2021-07-05 manuscript.Rmd
add more references

--- 2021-07-05 manuscript.Rmd
@@
 
-To guide the reader through all steps of a research project, we will use an exemplary research question:
+To guide the reader through a preregistration as code (PAC), we will use an exemplary research question:
 
@@
 We suggest writing a function that captures the full process of testing a research question that receives a dataset as input.
-For the hypothetical research question, we could for example plan to conduct a simple Student's t-test with Welch's correction an the average of machiavillism items between the gender groups.
-As an example of data dependend decisions we included a rank transformation if the skewness greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
-This example is suboptimal with regard to several criteria (untested assumption of measurement invariance, underestimation of effectsize in presence of measurement error, overestimation of effectsize for highly skewed distributions), we merely choose it for didactic purposes because similar procedures a widely used (@horton reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005).
+For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavillism items of the Short Dard Triad [SD3, @Jones2013] between the gender groups.
+As an example of a data dependent decisions we included a rank transformation if the skewness greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+This analyitical strategie is suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]); we merely choose it for didactic purposes because similar procedures a widely used [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
 
@@
 
-All variables of the dataset except the grouping variable are averaged.
-Depending on how skewed the dataset is we apply a rank transformation to the average and then we calculate a t-test with Welch correction in the hope of creating a more robust analysis for non-normal data.
-At last the function returns an object with all relevant information.
-
 To use this function one needs a compatible dataset.
-At the time of preregistration there is of course no actual data availible.
+At the time of preregistration there is of course no actual data available.
 We therefore simulate a dataset.

### 2021-07-05 manuscript.Rmd
dont use pipe in example

@@
 set.seed(112358) # fix state of RNG
-simulate_data(n = 400, # n/2 per group
-              df = 8, # Chisq with skewness == 1
+report::report_text(
+  planned_analysis(
+    simulate_data(
+      n = 400,# Sample size
+      df = 8, # Chisq with DF = 8 has skewness = 1
       d = .2, # Cohen's D
-              i = 10) %>% # number of items (should not effect outcome)
-  planned_analysis() %>% # `%>%` = output of 1. function becomes argument 2.
-  .$test %>% # extract the t.test object
-  report::report_text() # use the `report`-package
+      i = 10  # Number of items (should not effect anything)
+    )
+  )$test
+)
 ```

### 2021-07-05 R/simulation.R
increase simulation size

@@ -19,13 +19,13 @@ setup <- tidyr::expand_grid(
   df = 8, # skew = sqrt(8/df)
   d = seq(0, .5, 0.05),
   i = 10)
-res_raw %<-% simulation_study(setup, 100, 1235)
+res_raw %<-% simulation_study(setup, 10000, 1235)
 
 res <- res_raw %>% 
   group_by(across(-results)) %>% 
   unnest_wider(results) %>% 
   summarise(power = mean(p_value < 0.025),
-            cohend = mean(cohend),
+            cohend_mean = mean(cohend),
             cohend_sd = sd(cohend),
             .groups = "drop")
 fs::dir_create(here::here("data"))

### 2021-07-05 manuscript.Rmd
add simulation_results.csv as dependency

@@
     - R/link.R
+  data:
+    - data/simulation_results.csv
 output:
@@
 [^caution-mww]: The Mann–Whitney–Wilcoxon is **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
-The interested reader can use the provided code for the simulation to verify that the t.test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
+The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
 
@@
 
-```{r}
+***
+
+These results are based on simulated data.
+
+```{r, results='asis'}
 set.seed(112358) # fix state of RNG
@@
 )
+```
+
+***
+
+```{r, echo=FALSE}
+res <-
+  read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddddd")
+choosen_power <- .8
+choosen_d <- .2
+minn <- filter(res, d == choosen_d, power > choosen_power) %>% 
+  filter(n == min(n))
+```
+
+To determine a minimal sample size for this study we repeat this simulation across several combination of effect sizes and sample sizes.
+One standard recomendations for choosing a sample size is to have an expected power of 80% based on the expected effect size.
+@Jones2013 found in three studies gender differences of .24, .29, and .35 for their machiavilism subscale.
+If we want to detect an effect of size .2 with an 80% probability, our simulation shows that we require at least `r minn$n` participants.
+
+```{r, echo=FALSE}
+res %>%
+  ggplot(aes(n, power, color = d, group = d)) +
+  geom_line() +
+  geom_point(
+    data = minn,
+    size = 3,
+    color = "black",
+    shape = 3
+  ) +
+  theme_minimal() +
+  scale_color_viridis_c() +
+  theme(legend.title = element_text()) +
+  labs(y = "Power\n(Proportion of significant results)",
+       x = "Sample size",
+       color = "Cohen's D") +
+  NULL
 ```

### 2021-07-05 manuscript.Rmd
add a draft of a simple summary

@@
 simplesummary: |
-  A Simple summary goes here.
+  People trust science because other experts can vet it.
+  For example, suppose another person than the author of an article can recreate the same results on another computer.
+  In that case, they can more easily trust that the results are correct.
+  If anyone can recreate an article, it is called "reproducible".
+  We explain common technical problems (e.g.
+making sure that: 1.
+nothing has changed, 2.
+there are no copy-and-paste errors, 3.
+all steps are in the correct order, and 4.
+that the necessary software is installed) of reproducibility and show technical solutions.
+  However, because these solutions are complicated to use, we have developed some helpers to make them more accessible.
+  We then discuss why reproducibility also helps researchers to be more productive and produce higher-quality research.
+  One example is that with reproducibility, researchers can plan studies and their analyses with great precision.
+  If they publish such a plan, this is called "preregistration".
+  We urge researchers to formulate their data analysis for a preregistration as computer code because computer code is more precise than describing it with words.
+  They then should write all other parts of the preregistration with the same standards as a scientific article.
+  We call this preregistration as code.
+  Such a precise plan is beneficial for science because researchers generally trust results more if they are not explained after the fact.
 abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"

### 2021-07-05 manuscript.Rmd
add funding statement

--- 2021-07-05 manuscript.Rmd
@@
 acknowledgement: |
-  All sources of funding of the study should be disclosed. Please clearly 
-  indicate grants that you have received in support of your research work. 
-  Clearly state if you received funds for covering the costs to publish in open 
-  access.
+  The authors received no financial support for the research, authorship, and/or publication of this article.
 authorcontributions: |

### 2021-07-05 manuscript.Rmd
describe which author did what

@@
 authorcontributions: |
-  For research articles with several authors, a short paragraph specifying their 
-  individual contributions must be provided. The following statements should be 
-  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
-  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
-  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
-  limited to those who have contributed substantially to the work reported.
+  Aaron Peikert wrote the initial draft of the introduction, the tutorial and the preregistration of code.
+  Andreas Brandmaier wrote the summary of the workflow and provided extensive feedback.
+  Caspar J. Van Lissa proposed a more accessible structure of the paper, contributed ideas and, provided critical feedback.
 conflictsofinterest: |

### 2021-07-05 manuscript.Rmd
note Maximilians contribution
close #127

--- 2021-07-05 manuscript.Rmd
@@
 
-Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [https://www.pnas.org/content/115/11/2600].
-Preregistration increases the credibility of empirical results in three ways [preregistration is hard and worthwhile, Nosek].
+Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
+Preregistration increases the credibility of empirical results in three ways @nosekPreregistrationHardWorthwhile2019:
 

### 2021-07-05 manuscript.Rmd
closes #120
fix typo

@@
 For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
+A more comprehensive overview over how to employ containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
 

### 2021-07-05 manuscript.Rmd
Update manuscript.Rmd
I feel the original author contributions were too detailed - readers are likely not interested in a level of detail pertaining to the individual sections of the manuscript. I think it's important to state that AP took the lead in writing and that AB and CvL helped revising the draft.

--- 2021-07-05 manuscript.Rmd
@@
 authorcontributions: |
-  Aaron Peikert wrote the initial draft of the introduction, the tutorial and the preregistration of code.
-  Andreas Brandmaier wrote the summary of the workflow and provided extensive feedback.
-  Caspar J. Van Lissa proposed a more accessible structure of the paper, contributed ideas and, provided critical feedback.
+  Aaron Peikert took the lead in writing and provided the initial draft of the manuscript.
+  Andreas Brandmaier and Caspar J. Van Lissa contributed further ideas, critical feedback, and revisions of the original manuscript.
+  Furthermore, we would like to thank Maximilian Stefan Ernst (not an author) for his contributions to the code for the simulation study.
 conflictsofinterest: |

### 2021-07-05 manuscript.Rmd
add Authot Contribution (#129)
* describe which author did what

* note Maximilians contribution

* Update manuscript.Rmd

I feel the original author contributions were too detailed - readers are likely not interested in a level of detail pertaining to the individual sections of the manuscript. I think it's important to state that AP took the lead in writing and that AB and CvL helped revising the draft.

Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>

--- 2021-07-05 manuscript.Rmd
@@
 authorcontributions: |
-  For research articles with several authors, a short paragraph specifying their 
-  individual contributions must be provided. The following statements should be 
-  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
-  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
-  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
-  limited to those who have contributed substantially to the work reported.
+  Aaron Peikert took the lead in writing and provided the initial draft of the manuscript.
+  Andreas Brandmaier and Caspar J. Van Lissa contributed further ideas, critical feedback, and revisions of the original manuscript.
+  Furthermore, we would like to thank Maximilian Stefan Ernst (not an author) for his contributions to the code for the simulation study.
 conflictsofinterest: |

### 2021-07-05 manuscript.Rmd
add inline r code

@@
 @Jones2013 found in three studies gender differences of .24, .29, and .35 for their machiavilism subscale.
-If we want to detect an effect of size .2 with an 80% probability, our simulation shows that we require at least `r minn$n` participants.
+If we want to detect an effect of size `r choosen_d` with an `r choosen_power*100`% probability, our simulation shows that we require at least `r minn$n` participants.
 

### 2021-07-06 manuscript.Rmd
get changes from 3a7f563 back

@@
 
+# Preregistration as Code
+
 The file `.repro/Dockerfile_manual` is by default empty and is supposed to allow dependencies outside of R, like system libraries or external software.
@@
 
-To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript.
-The introduction and theoretical background provide the basis for the studies design and can be written before gathering data.
+To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
+The introduction and theoretical background provide the basis for the studies' design and can be written before gathering data.
 We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
@@
 if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
-Third, deviations from the preregistration are more explicit because they are reflected in changes to the codes a property facilitated by version control.
+Third, deviations from the preregistration are more explicit because they are reflected in changes to the code, a property facilitated by version control.
 Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
+
+Why are PAC and reproducibility related?
+A reproducible script is supposed to work on the same data at a later point in time.
+PAC requires a script to work on different data at a later point in time, a property we call "reusability".
+This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] so it is build-for-reuse [@al-badareenReusableSoftwareComponents2010].
+Reusability promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
 

### 2021-07-06 manuscript.Rmd
bring back 03336ac

@@
 The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
-These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
-The results further illustrate that open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
+These failed attempts to reproduce highlight the need for widely accepted reproducibility standards, because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
 
@@
 Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
-Preregistration increases the credibility of empirical results in three ways @nosekPreregistrationHardWorthwhile2019:
+Preregistration increases the credibility of empirical results in three ways, @nosekPreregistrationHardWorthwhile2019:
 

### 2021-07-06 manuscript.Rmd
change how r code is presented

@@
 
-These results are based on simulated data.
-
 ```{r, results='asis'}
 set.seed(112358) # fix state of RNG
-report::report_text(
+results <- 
   planned_analysis(
@@
     )
-  )$test
   )
+```
+
+These results are based on simulated data:
+
+```{r}
+report::report(results)
 ```

### 2021-07-06 R/simulation_funs.R

@@ -4,7 +4,6 @@ simulate_data <- function(n, df, d, i){
   group <- rep(c(1L, 0L), each = n/2)
   rand <-  matrix(rchisq(n * i, df/i), ncol = i) # sum of n chisq has df of n*df
   d_scaled <- d * sqrt(2*df)/i # scale d to express sd units (var(chisq) = 2df)
-  #browser()
   effect <- rand + d_scaled * group
   colnames(effect) <- paste0("x", seq_len(i))
   effect <- as.data.frame(effect)

### 2021-07-06 manuscript.Rmd
close #125
close #119

@@
 The file `.repro/Dockerfile_base` contains information about the base image on which all the rest is build.
-By default we rely on the "verse" images provided by the Rocker project.
+By default we rely on the "verse" images provided by the Rocker project [@boettigerIntroductionRockerDocker2017].
 These contain (among other) the packages `tidyverse`, `rmarkdown` and a Latex installation, which make them ideal for writing scientific manuscripts.
@@
 Fourth, you can plan your resources efficiently, e.g. determine required sample size to achieve certain precision or power [@brandmaier2018precision].
+@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.
 

### 2021-07-06 manuscript.Rmd
fix tiny error
reorder PAC

@@
 
-The idea of submitting code as part of a preregistration is not new.
-A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods and in an informal literature research (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
-
 We consider three criteria when planning a study.
@@
 
-A PAC offers four advantages over classical preregistration.
+The idea of submitting code as part of a preregistration is not new.
+A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods and in an informal literature research (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
+
+Why are PAC and reproducibility related?
+Preregistrations are plagued by nonreproducibility like any other research outcome.
+@bakkerRecommendationsPreregistrationsInternal2020 found from the 210 preregistrations they investigated 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to theoretically reproduce the power analysis.
+Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
+@steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
+
+We believe that pairing PAC with the here presented workflow offers four advantages over classical preregistration.
 First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
@@
 
-Why are PAC and reproducibility related?
-One the one hand, preregistrations are plagued by nonreproducibility like any other research outcome.
-@bakkerRecommendationsPreregistrationsInternal2020 found from the 210 preregistrations they investigated 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to theoretically reproduce the power analysis.
-Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
-@steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
-
-On the other hand have PACs greater requirements for reproducibility than a simple analysis script.
-A reproducible script is supposed to work on the same data at a later point in time.
-PAC requires a script to work on different data at a later point in time, a property we call "reusability".
+At the same time require PACs more than "just" reproducibility.
+In contrast to a reproducible script, that should work on the same data, a PAC must function on different data.
+We call this property "reusability".
 This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] so it is build-for-reuse [@al-badareenReusableSoftwareComponents2010].
 Reusability promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
+Reusability is naturally desirable in multi lab studies (studies with more than one research site).
+PAC could offers additional
 
@@
 ```{r}
-report::report(results)
+report::report(results$test)
 ```

### 2021-07-06 manuscript.Rmd
close #122
remove section
results asis

@@
 
-# Preregistration as Code
-
 The file `.repro/Dockerfile_manual` is by default empty and is supposed to allow dependencies outside of R, like system libraries or external software.
@@
 First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
+Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
 Second, despite being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
@@
 
-```{r, results='asis'}
+```{r}
 set.seed(112358) # fix state of RNG
@@
 
-```{r}
+```{r, results='asis'}
 report::report(results$test)

### 2021-07-06 manuscript.Rmd
closes #123

@@
 
-We believe that pairing PAC with the here presented workflow offers four advantages over classical preregistration.
+We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
 First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
@@
 if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
-Third, deviations from the preregistration are more explicit because they are reflected in changes to the code, a property facilitated by version control.
-Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
+Third, a PAC more comprehensive is by design because e.g. missing steps are obvious when the code is tested on simulated data.
+A due to its high structure PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
+Forth, deviations from the preregistration are more explicit because they are reflected in changes to the code, a property facilitated by version control.
+Fiveth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
 
@@
 Reusability promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
-Reusability is naturally desirable in multi lab studies (studies with more than one research site).
-PAC could offers additional
 

### 2021-07-06 manuscript.Rmd
remove gert, its not needed for the tutorial

@@
 These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
-Please install the package [`gert`](https://github.com/r-lib/gert)[@R-gert] which wraps Git commands and [`repro`](https://github.com/aaronpeikert)[@R-repro] so they can act as assistance systems while you follow the tutorial:
+Please install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so it can act as assistance systems while you follow the tutorial:
 
 ```{r, eval=FALSE}
-install.packages("gert")
 # repro is not on CRAN yet
@@
 
-However, users do not have to have repro installed, in fact, they do not even need R installed. 
+However, users do not have to have repro installed to reproduce a project, in fact, they do not even need R installed. 
 The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.

### 2021-07-06 manuscript.Rmd
better explain where software is installed

@@
 The file `.repro/Dockerfile_manual` is by default empty and is supposed to allow dependencies outside of R, like system libraries or external software.
-The most convinient way to install software on Ubuntu is through its package manager `apt`.
-This snipped installs python:
+Remember, the software environment of the Docker image is completely separate from software installed on your computer.
+This separation is excellent for reproducibility, but requires you to install software on an operating system that may not be familiar to you.
+The images supplied by [@boettigerIntroductionRockerDocker2017] are based on Ubuntu, and hence software has to be installed in a way that is compatible with Ubuntu.
+The most convenient way to install software on Ubuntu is through its package manager `apt`.
+This snipped installs python, added to `.repro/Dockerfile_manual` you get an image that has python available:
 

### 2021-07-06 manuscript.Rmd
caution to not use t.test or wilcox test

@@
 
-At the same time require PACs more than "just" reproducibility.
-In contrast to a reproducible script, that should work on the same data, a PAC must function on different data.
+PACs come with a challange though, in contrast to a reproducible script, that should work on the same data, a PAC must function on different data.
 We call this property "reusability".
@@
 As an example of a data dependent decisions we included a rank transformation if the skewness greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
-This analyitical strategie is suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]); we merely choose it for didactic purposes because similar procedures a widely used [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
 
@@
 
+This function is one possible translation of the verbal descriptions:
+
 ```{r planned_analysis}
 ```
+
+One point of note about the analytical strategy and the function implementing it is that there are several translations into code that would fit the descriptions but would produce different results.
+One example would be using the function `wilcox.test` instead of the combinations of `rank` and `t.test`.
+Both are valid implementation of Mann–Whitney–Wilcoxon test but the first assumes equal variance while the second applies Welch's correction and hence is robust even with unequal variances [@zimmermanRankTransformationsPower1993].
+Mentioning every such minute implementation detail is would result in overly verbose preregistrations but does change the outcome.
+
+We also want to caution researchers to not use the here shown procedure.
+We merely choose it for didactic purposes because it may be familiar to many researchers because the t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
+This analytical strategie is suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]);
 

### 2021-07-06 manuscript.Rmd
include more application of automation closes #124

@@
 This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] so it is build-for-reuse [@al-badareenReusableSoftwareComponents2010].
-Reusability promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
+The reproducible workflow we present here is heavily automated and hence promotes reusibility.
+Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations employ automation, like machine readible hypothethis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
+Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
 

### 2021-07-06 manuscript.Rmd
suggest Visual Editor for R Markdown closes #118

@@
 
+This type of markup for formatting, instead of the more visual approach of e.g. Microsoft Word ©, is compatible with many output formats and is more easily version controlled.
+However, researchers might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor with still creates an R Markdown and has all of its benefits.
+The Visual Editor has the additional benefit of promoting best practices, e.g. each sentence on a newline for better version control or automated citations, cross references and more.
+
+Now that you are familiar with writing text we turn our attention to including code into the text.
 Code is separated by three backticks (and the programming language in curly braces) like here:

### 2021-07-06 manuscript.Rmd
add all abbrevations closes #116

@@
 abbreviations:
+  - short: PAC
+    long: Preregistration as 
+  - short: Gb
+    long: Gigabyte
+  - short: Kb
+    long: Kilobyte
   - short: GUI
@@
     long: Comand Line Interface
+  - short: CRAN
+    long: Comprehensive R Archive Network
 repro:
@@
 It is highly fool-proof (unlike, for example, LaTeX), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your CV [@vitae].
+We suggest using Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your résumé [@vitae].
 R Markdown extends regular Markdown by allowing users to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.

### 2021-07-06 manuscript.Rmd
remove appendix

@@
 bibliography: temp.bib
-appendix: appendix.tex
 simplesummary: |

### 2021-07-07 manuscript.Rmd
small changes after reading

@@
 In contemporary research, reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
-Unfortunately, despite increasing commitment to open science practices, many projects can not yet be reproduced by other research teams.
+Unfortunately, despite increasing commitment to open science practices, many projects can not yet be reproduced by other research teams [@obels2020].
 This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
@@
 One such opportunity is an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
-This dynamic document closely resembles an actual journal article but with "fake" results and without discussion of results, then serves as preregistration.
+This dynamic document closely resembles an actual journal article but with "fake" results and without discussion of results and serves as preregistration.
 
@@
 We suggest using Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your résumé [@vitae].
-R Markdown extends regular Markdown by allowing users to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
+R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
 Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.

### 2021-07-07 manuscript.Rmd
I give up finding a reference for this

@@
 One would hope that sharing a literate statistical analysis would allow others (including one own's future self) to simply download and run (ie., reproduce) a given analysis.
-However, while version control and dynamic document generation are becoming more common @, this approach fails to guarantee reproducibility in most instances.
+However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in most instances [@Peikert2019].
 In practice, dependencies between project files (e.g.

### 2021-07-07 manuscript.Rmd
cite further materials for Git closes #133
explain how to extend Make

@@
 Git has additional features beyond basic version control, such as branches to facilitate collaboration.
+@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git.
+@bryanExcuseMeYou2018 provides additional information on how to track R Markdown documents.
 To collaborate via Git,<!--dangling modify, cant help it--> the repository has to be uploaded somewhere.
@@
 ```{bash, eval=FALSE}
-manuscript.pdf: manuscript.Rmd simulation_results.csv 
+manuscript.pdf: manuscript.Rmd simulated_data.csv
   Rscript -e 'rmarkdown::render("manuscript.Rmd")'
 
-simulation_results.csv: simulate.R
+simulated_data.csv: simulate.R
   Rscript -e 'source("simulate.R")'
@@
 This way, we ensure that the simulation is always run before the manuscript is built and that the manuscript is rebuilt if the simulation code was changed.
-
-A version controlled dynamic document with dependency tracking still relies on external software.
-Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
+Make therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
 
@@
 
+A version controlled dynamic document with dependency tracking still relies on external software.
+Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
 One solution to the problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
@@
 Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
+Researchers therefore have the opportunity to use the exact same software that was used by the original authors.
 Such a snapshot of the software stack is called an "image".
@@
 This recipe is a textual description of all commands that need to be executed to recreate the image.
-Such files are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an sufficently describe how to recreate a given container image.
+Such files are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and sufficiently describe how to recreate a given container image.
 However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
-For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
+For archiving, we therefore recommend that a full image is always kept.
 A more comprehensive overview over how to employ containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
@@
 To remove this hurdle, we have developed a new R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
-In the remainder, we will walk you through a complete research project, from sketching the very first idea over preregistration, data analysis, and submission of a preprint while focusing on how the `repro`-package enables users to maintain reproducibility of all steps.
+In the remainder, we will walk you through a the creation of a reproducible research project aided by the package `repro`.
 
@@
 They detect the users' system state and explain what this means to the user.
-Sometimes they ask the user to do something, e.g. the following happens if you were a windows user who does not have installed Git:
+Sometimes they ask the user to do something, e.g. the following happens if you were a Windows user who does not have installed Git:
 
@@
 This type of markup for formatting, instead of the more visual approach of e.g. Microsoft Word ©, is compatible with many output formats and is more easily version controlled.
-However, researchers might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor with still creates an R Markdown and has all of its benefits.
+However, researchers might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown and has all of its benefits.
 The Visual Editor has the additional benefit of promoting best practices, e.g. each sentence on a newline for better version control or automated citations, cross references and more.
@@
 The most convenient way to install software on Ubuntu is through its package manager `apt`.
-This snipped installs python, added to `.repro/Dockerfile_manual` you get an image that has python available:
+The following snipped installs Python.
+If added to `.repro/Dockerfile_manual` the Docker image will have Python installed.
 
@@
 
-The the last Docker, `.repro/Dockerfile_packages`, should not be manually changed.
+The last Dockerfile, `.repro/Dockerfile_packages`, should not be manually changed.
 `repro::automate()` gathers all packages from all `.Rmd` files and figures out if they should be installed from CRAN or GitHub.
-`repro::automate()` also simply combines the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the project.
+
 The seperation into several Dockerfiles is only to provide a clear seperation between automatically generated parts and manually altered parts of the Dockerfile.
-However, Docker requires in the end the one Dockerfile that `repro::automate()` creates from them.
+However, Docker requires single Dockerfile so `repro::automate()` simply combines the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the
+
+Users of `repro` can not only build arbitrarily complex software environments they can also implement complex file dependencies.
+The standard `repro` metadata only makes sure that all dependencies are availible but does not allow you to specify a recipe for them in the metadata.
+If you can formulate the creation of dependencies in terms of computational steps you should include these in the Makefile.
+The Makefile that `repro` creates is only an template and you are free to change it.
+However, you want to make sure to not remove the following two lines:
+
+```bash
+include .repro/Makefile_Rmds
+include .repro/Makefile_Docker
+```
+
+The file `.repro/Makefile_Rmds` contains the automatically generated targets from `repro::automate()` for the R Markdown files.
+This file should not be altered manually.
+If you want to adept a target than simply provide an alternative target in the main `Makefile`.
+Targets in the main `Makefile` take precedent.
+
+The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but seldomly should alter.
+This Makefile coordinates the interplay between Make and Docker and contains targets for building (with `make docker`) and saving (with `make save-docker`) the Docker image.
+Additionally it provides facilities to execute commands within the container.
+If you write computational recipe for a target it will be by default evaluated using the locally installed software.
+To instead evaluate commends inside the Docker image you should wrap them in `$(RUN1) command $(RUN2)`, like in this example:
+
+```bash
+simulated_data.csv: R/simulate.R
+  $(RUN1) Rscript -e 'source("R/simulate.R")' $(RUN1)
+```
+
+If users use
+
+```bash
+make simulated_data.csv
+```
+
+this is simply translated to:
+
+```{r, echo=FALSE, results='asis'}
+cat("```bash",
+  stringi::stri_wrap(system2("make", " --dry-run simulated_data.csv", stdout = TRUE)),
+  "```", sep = "\n")
+```
+
+But if users use
+
+```bash
+make DOCKER=TRUE simulated_data.csv
+```
+
+it is evaluated within the Docker container:
+
+```{r, echo=FALSE, results='asis'}
+cat("```bash",
+  stringi::stri_wrap(system2("make", " --dry-run DOCKER=TRUE simulated_data.csv", stdout = TRUE)),
+  "```", sep = "\n")
+```
 
@@
 The reproducible workflow we present here is heavily automated and hence promotes reusibility.
-Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations employ automation, like machine readible hypothethis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
+Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readible hypothethis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
 Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
+
+Therefore reproducibility facilitates traditional good scientific practices and provides the fundament for promising innovations.
 

### 2021-07-11 manuscript.Rmd
change headings and add summary to advanced section
a first set of suggestions covering lines 1--374

@@
 
-Increasingly, scientists openly share their data, materials, and analysis code.
-Sharing the digital research output increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
-The technology is now available to make all artefacts related to a research project available so that the project can be reproduced with minimal effort.
+Increasingly, scientists make their data, materials, and analysis code openly available.
+Sharing all these types of digital research output increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
+We now have the technology to make all artefacts related to a research project available so that a research project can be reproduced with minimal effort.
 Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
 22][^popper]
-In contemporary research, reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
-Unfortunately, despite increasing commitment to open science practices, many projects can not yet be reproduced by other research teams [@obels2020].
+Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
+Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
 This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
@@
 Specifically, we demonstrate how the R package `repro` supports researchers in implementing these solutions.
-The second objective is to discuss opportunities of such workflow to improve not only how research outputs are created and recreated but also how research itself is conducted.
-One such opportunity is an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
-This dynamic document closely resembles an actual journal article but with "fake" results and without discussion of results and serves as preregistration.
+The second objective is to discuss opportunities of such workflows to improve not only how research outputs are created and recreated but also how research itself is conducted.
+One such opportunity is an alternative preregistration <!--AP preregistration is not explained yet; citation?--> scheme that builds upon a reproducible workflow, the so-called _preregistration as code_ (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results before they gather empirical data.
+This dynamic document closely resembles an actual journal article but with "fake" results from simulated data that serve to generate placeholders for figures, tables, and statistics. 
+This document serves as unambiguous preregistration of the planned analysis and how the results are going to be reported.
+The only parts that are missing from this first draft are the discussion of results, which will completed once the real data were gathered.
 
@@
 
-While the willingness to share materials is required for reproducing a project, it is not sufficient.
+While the willingness to share materials is required for reproducing a project, it is not sufficient to achieve it.
 For example, @hardwicke2018 attempted to reproduce results from open materials in the journal Cognition. 
@@
 
-We argue that the benefits associated with transparency increase if research can additionally be reproduced.
-This paper tutorial demonstrates how to make research projects reproducible, while striking a balance between rigor and ease-of-use.
-A rigorous standard increases the likelihood of any individual project to be actually reproducible.
-An easy to use standard, on the other hand, is more likely to be adopted.
+We argue that the benefits of open science increase if research is not only transparent but reproducible.
+To this end, this tutorial demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
+A rigorous standard increases the likelihood of any individual project to be actually reproducible and remain reproducible as long as possible.
+An easy-to-use standard, on the other hand, is less error-prone, produces less overhead and is thus more likely to be adopted.
+To us, it makes sense to strive for a broad adoption of easy-to-use practices and then gradually move into more rigorous standards as researchers become more proficient with the tools involved.
 The remainder of this paper is structured along these lines.
@@
 
-We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.
+We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along on their own computers.
 We explicitly encourage readers to try out all R commands for themselves.
@@
 
-We identified the following common threat to reproducibility: <!-- CJ How did you identify these causes? Is this exhaustive?-->
+From our own experience with various research projects, we identified the following common threats to reproducibility: <!-- CJ How did you identify these causes? Is this exhaustive?-->
 
-1.	**Multiple inconsistent versions of code, data, or both**; for example, the dataset could have changed over time because outliers were removed at a later stage, or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
-2.	**Copy-and-paste**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	**Ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order; 
-4.	**External code dependencies**; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist; or, a different version of the same software may produce different results.
+1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
+2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
+3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly)
+4.	**External code dependencies**; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
 <!-- CJ 5.
 Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
+<!--AB Caspar's 5th point is important.
+In my mind, this best fits to point #3, which adressed an insufficient level of coding/automation/documentation of the necessary steps taken.
+Can we merge this into one new point #3?
+I will commit a suggested change of point #3 here (see above);.-->
 
@@
 A version control system tracks changes to all project-related files (e.g., materials, data, and code) over time.
-At a later stage, single files or the entire project can be compared to or reverted to an earlier version.
+At a later stage, single files or the entire project can be compared to or reverted to any earlier version.
 A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
@@
 Git is built around snapshots that represent the entire project state at a given point in time.
-Those snapshots are called "commits".
-Each file that was changed has to be added to a commit, and each commit requires a commit message that succinctly describes the changes made.
+Those snapshots are called "commits" and work like a "save" action.
+Changes to files are collected in a _commit_, and each commit has a message that succinctly describes the changes made by the author.
 Each commit "knows" its ancestor, and they form a timeline together.
-The conglomerate of commits (i.e. the version-controlled project) is called a repository.
+The entirety of commits (i.e. the version-controlled project) is called a repository.
 In Git, specific snapshots of a repository can be tagged, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
@@
 We recommend GitHub to host Git repositories because of its popularity among R users.
-GitHub has many tools that ease project management and collaboration, and these tools provide much value in our everyday work, but they are not central to archiving reproducibility.
+GitHub has many tools that ease project management and collaboration, and these tools provide much value in our everyday work, but they are not central to achieving reproducibility.
 
@@
 In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
-Both steps are linked by manually copying and pasting results from one software to the other, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
+Typically, results are manually copyied and pasted from one software to the other, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
 
-Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured, who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
+Dynamic document generation integrates both steps.
+This paradigm can be traced back to @knuthCWEBSystemStructured who suggested that a computer prorgram should be interspersed with explanations of its logic in a natural language.
+<!--AB I removed the quotation because it is not really clear to me how it matches what we say about R Markdown - I am not sure whether my changes make things better, though-->
 **R Markdown** uses Markdown for writing the text and R (or other programming languages) for the analysis.
-Markdown is a lightweight text format in plain text with minimal reserved symbols for formatting instructions.
+Markdown is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
 This way, Markdown does not need any specialized software for editing.
-It is highly fool-proof (unlike, for example, LaTeX), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your résumé [@vitae].
+It is highly fool-proof (unlike, for example, LaTeX <!--AB citation needed!-->), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your résumé [@vitae].
 R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
@@
 However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in most instances [@Peikert2019].
-In practice, dependencies between project files (e.g.
-the raw data was preprocessed) or on external software (such as system libraries or components of the programming language) are frequently unmentioned.
+In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambigously documented.
 
@@
 We recommend using **Make** for dependency tracking because it is language independent.
-To make this more concrete, consider the following hypothetical example, in which a research project contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
+To make the utility of Make more concrete, consider the following hypothetical example, in which a research project contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
 A Makefile could then look like this:
@@
 
-A version controlled dynamic document with dependency tracking still relies on external software.
-Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
-One solution to the problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
+A version controlled dynamic document with dependency tracking may still rely on external software.
+Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
+One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
 One comprehensive approach to software management is containerization.
 "By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
+<!--AB: direct quotes need a line number!
+See APA style guide-->
 Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
@@
 Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
-Researchers therefore have the opportunity to use the exact same software that was used by the original authors.
+Researchers therefore have the opportunity to use the exact same software setup that was used by the original authors.
 Such a snapshot of the software stack is called an "image".
-Packaging all needed software in such an image requires considerable amounts of storage space.
-Two major strategies are helping to keep the storage requirements reasonable.
-First, there is a community that maintains images for particular purposes.
-For example, there are ready-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
+Packaging all required software in such an image requires considerable amounts of storage space.
+Two major strategies help to keep the storage requirements reasonable.
+First, there is a community that maintains pre-made images for particular purposes.
+For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
 Users can then install whatever they need in addition to what is provided by these pre-compiled images.
-`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
-Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
-Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
+`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
+Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
+Second, instead of saving a binary image of the software environment, one can also save a container recipe, a so-called  `Dockerfile`.
 This recipe is a textual description of all commands that need to be executed to recreate the image.
-Such files are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and sufficiently describe how to recreate a given container image.
-However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
-For archiving, we therefore recommend that a full image is always kept.
-A more comprehensive overview over how to employ containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
+Such files are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb) and sufficiently describes how to recreate a given container image.
+However, Dockerfiles rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
+For proper archival, we therefore recommend to keep a full image in addition to the Dockerfile.
+A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
 
-To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), containerization (using Docker).
-While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which represents a considerable hurdle in the acceptance and implementation of the workflow.
+To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
+While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
 To remove this hurdle, we have developed a new R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
-In the remainder, we will walk you through a the creation of a reproducible research project aided by the package `repro`.
+In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.
 
@@
 
-A significant hurdle for adopting the before-mentioned tools for researchers is that they are typically used through a command-line interface (CLI).
-The CLI predates and coexists with graphical user interfaces (GUIs) which many users (including the authors) prefer over the text-based interactions of a CLI for most tasks.
-A GUI has the apparent disadvantages for reproducibility that it requires a user who clicks on the correct buttons in the correct order.
-Depending on the number of steps, this is not only much effort but also error-prone.
+A major barrier to adoption of the above tools for researchers is that they are typically used via a command line interface (CLI).
+The CLI predates and coexists with graphical user interfaces (GUIs), which many users (including the authors) prefer over the text-based interactions of a CLI for most tasks.
+A GUI has the apparent disadvantage for reproducibility that it requires a user who clicks on the correct elements (such as buttons or menu items) in the correct order.<!--AB not sure about the entire paragraph about CLI vs GUI here -  can we shorten this?-->
+Depending on the number of steps taken, this is not only much effort but also error-prone.
 This tutorial seeks a compromise.
-We use simpler alternatives, including a GUI where we can, to create a reproducible workflow, but the reproduction of such workflow does not rely on manual steps.
-At several points, we rely on or have self-developed a CLI that wraps a more complicated (and powerful) CLI.
+We use simple GUI actions to create a reproducible workflow template, but the reproduction of such workflow does not rely on manual steps.
+<!--AB I modified this sentence - but I am not sure what we are saying here with the "manual steps"-->
+At several points, we rely on or have developed a CLI that wraps a more complicated (and powerful) CLI.
 These wrappers use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
-We hope this makes reproducibility tools more accessible by enabling untrained users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
+We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
 These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
-Please install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so it can act as assistance systems while you follow the tutorial:
+
+Before we can start creating reproducible research projects, we need a computer with a working R Studio installation. 
+Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so that it can act as an assistance systems while you follow the tutorial:
 
@@
 
-One example for such assistance are the check functions from `repro`.
-They verify that you indeed have installed and setup the required software for this workflow:
+An example of the assistance that `repro` provides are the check functions.
+They verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
 
 ```{r}
-# `package::function()` → use function from package without `library(package)`
 repro::check_git()
@@
 
-They detect the users' system state and explain what this means to the user.
-Sometimes they ask the user to do something, e.g. the following happens if you were a Windows user who does not have installed Git:
+They detect the users' system state and explain if further action is needed.
+Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:
 
@@
 The messages emitted by `repro` and other packages we present try to take the user by the hand to help them solve problems.
-This tutorial describes steps that we expect to apply to all users; the messages from the packages fill the gap to your specific situation and extend the tutorial in this way.
+This tutorial describes steps that we expect to apply to all users; the messages from the packages are adjusted to your specific situation (that is, operating system and the specific problem of missing software) and complement the tutorial in this way.
 Before you continue, we ask you to run the above commands to check Git, Make, and Docker.
-If necessary, follow the instructions of repro to install them.
+If necessary, follow the instructions of repro to install them until the check functions are all fine.
 
-The first concept we want to apply practically is version control with Git.
+The first concept we want to introduce is version control with Git.
 Git tracks files within one project folder.
-We start by creating such project folder with RStudio:
+We start by creating such a project folder with RStudio by clicking the menu item:
 
@@
 
-If you made changes to a file, you have to stage and then commit these changes.
-The easiest way is to use the RStudio Git pane.
-You click on the empty box next to the file you want to stage, and when a tick appears, the file is staged.
-After you have staged all files you want, click on the commit button, explain curtly why you made those changes in a commit message and then click on commit.
+If you made changes to a file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
+The easiest way to do this is to use the RStudio Git pane.
+Click on the empty box next to the file you want to stage, and when a tick appears, which means that the file is staged.
+After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
 
 The files you created and the changes you made still have not left your computer.
-To back up and share the files, you can publicly upload those changes to GitHub with:
+All snapshots are stored in a local repository in your project folder.
+To back up and share the files, you can publicly upload those changes to a public repository.
+To synchronize your local repository with a remote repository from the GitHub service, issue the following command:
 
@@
 
-Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow its suggestions.
+Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the command line.
 
-If you are not yet ready to share your project publicly use:
+If you are not yet ready to share your project publicly but you want to synchronize your local repository to a remote repository (for example, as a simple backup service), use:
 
@@
 They can be pasted into a terminal and will reproduce the project in many cases.
+
+# Advanced Tutorial
 
@@
 
-# Preregistration as Code
+To summarise, `repro` provides untrained researchers with an easy to use solution for dependency tracking (in form of Make) and software management (using Docker).
+Users with more advanced requirements can manually change all aspects while still profiting from can mix `repro`'s automations.
+
+# Introducing Preregistration as Code
 
@@
 
-# Preregistration as Code --- an example
+# Preregistration as Code --- a Tutorial
 

### 2021-07-12 manuscript.Rmd
some more suggestions until l.514

@@
 
-This creates a project containing an example analysis, but you may use any other template or existing R Project.
+This creates a project containing an example analysis, but you may use any other template or existing R Project to turn into a reproducible research project using the functionality of `repro`.
 
@@
 
-After you have ignored sensitive files (like not yet anonymized data), you can activate Git with:
+After you have excluded sensitive files (like not yet anonymized data files), you can activate Git with:
 
@@
 
-If you made changes to a file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
+Whenever you made changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
 The easiest way to do this is to use the RStudio Git pane.
@@
 
-Now that you have created a project that is under version control, we turn to dynamic document generation.
-Three aspects are essential to understand a dynamic document:
+Now that you have created a project that is under version control, we continue with adding dynamic document generation.
+Adynamic document has three elements:
 
-1. Text
-2. Code
+1. Text in a natural language
+2. Executable code 
 3. Metadata
 
-The text of an R Markdown is written in Markdown.
-The following example shows how to signify a header, a bold word and a list in Markdown:
+The text of an R Markdown is written in Markdown. <!--AB citation needed for further information-->
+The following Markdown example serves to illustrate the Markdown syntax. 
+It shows how to create a heading, a word in bold font, and a list of several items in Markdown:
 
@@
 <!--this is a Markdown file -->
-# Header
+# Heading (level 1)
 
@@
 Important **word** in bold.
+
+## Subheading (level 2)
 
@@
 * do more research
-* spend time with family
+* spend time with family and friends
 ```
 
-This type of markup for formatting, instead of the more visual approach of e.g. Microsoft Word ©, is compatible with many output formats and is more easily version controlled.
-However, researchers might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown and has all of its benefits.
-The Visual Editor has the additional benefit of promoting best practices, e.g. each sentence on a newline for better version control or automated citations, cross references and more.
+This type of markup for formatting, instead of the more visual approach of e.g. Microsoft Word, is compatible with many output formats and is easier to put under version control.
+Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
+The visual editor has the additional benefit of promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions).
+Among other features, it eases the generation of citations and cross references.
 
-Now that you are familiar with writing text we turn our attention to including code into the text.
+Now that you are familiar with the basics of writing text in Markdown, we turn our attention to including code into the text.
 Code is separated by three backticks (and the programming language in curly braces) like here:
@@
 Now that you know how to write text and R code in an R Markdown, you need to know about metadata, like the title or the output format.
-Metadata are placed at the beginning of the document and are separated by three dashes; the following example is a full markdown document where the header is in lines 1-6.
+Metadata are placed at the beginning of the document and are separated by three dashes from the document body; the following example is a full markdown document where the header is in lines 1--6.
 
@@
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
-abstract: A paper about recalculating the same thing more than once.
+abstract: A tutorial about doing the same thing more than once.
 output: html_document
@@
 
-Most metadata fields are self-explanatory (like the author field), but any package may use the metadata for their purposes.
-Often, for example, does a specific output format require more metadata, i.e. the `rticles`-package for authoring scientific manuscripts.
-R Markdown was extended to hundreds of document formats which are often highly configurable by their metadata.
-The `repro` package uses the metadata, not to change the output, but to make the document more reproducible, simply by listing resources, i.e., scripts, data, and packages, your document depends on, in the following format:
+Most metadata fields are self-explanatory (like the author field). 
+Sometimes, R package may extend the metadata fields for their own purposes.
+Certain output formats may require specific  metadata. 
+For example, the `rticles`-package that offers templates for authoring scientific manuscripts comes with extra fields that are described in the package documentation and offer extra options for flexible formatting of documents.
+R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
+The `repro` package uses the metadata for hints to make the research project reproducible.
+Researchers should list all dependent resources of their research project in the metadata, that is, R scripts, data files, and external packages in the following format (see everything below the line *repro:*):
 
@@
 `repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
-In day-to-day work the simplest solution is to use the function `repro::automate()` which both infers a Makefile and a Dockerfile from all R Markdown  (all files with the ending `.Rmd`) and there respective repro metadata.
-Every time you change the `repro`-metadata, change the output format, or add a new R Markdown, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
+In everyday work, the simplest solution is to now call the function `repro::automate()` which both creates a Makefile and a Dockerfile from all R Markdown files in the project (all files with the ending `.Rmd`) based on the respective repro metadata entries.
+Every time you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
 The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
-Besides the main Makefile and the Dockerfile repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
+Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
+<!--AB: why is this addition not automated?
+Feels strange given that this is our recommended best practice?!-->
 
-When you or people who received your project from GitHub want to reproduce the project, they can ask repro to explain how this should be done:
+When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:
 
@@
 ```
+<!--AB: I still think this should be repro::reproduce() please change!-->
+Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
+These projects will remain reproducible irrespective of whether `repro` will be maintained and remain accessible in future.
+Users do not have to have repro installed to reproduce a project, in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
+The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.
+The commands can be pasted into a terminal and will reproduce the project in many cases.
+<!--AB: I said this before.
+I think it really should be possible to execute this command from within R.
+Never underestimate the troubles people can have with the terminal.
+I think we discussed this before.
+Please add a function repro::rerun(execute=TRUE) or -- if we rename it -- repro::reproduce(execute=TRUE).
+-->
 
-However, users do not have to have repro installed to reproduce a project, in fact, they do not even need R installed. 
-The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.
-They can be pasted into a terminal and will reproduce the project in many cases.
-
-# Advanced Tutorial
+# Advanced Features
 
 Some projects, however, are not as easy to make reproducible.
-As explained above, repro is merely a simplified interface.
+As explained above, repro is merely a simplified interface that guarantuees a high level of reproducibility for simple projects.
 These simplifications are possible because of two restrictions.
-First, the project can only rely on R packages for the Dockerfile and second, it can have no multilayered preprocessing for the Makefile.
+First, the project can only rely on R packages for the Dockerfile and second, it can have no multilayered preprocessing for the Makefile.<!--AB: please explain both points in more detail-->
 Lifting these restrictions requires the user to interact more directly with Make or Docker.
@@
 Let us dissect the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two steps.
-First it virtually recreates the software environment and then it executes computational recipes in the virtual software environment.
+First it recreates the virtual software environment (Docker) and then it executes computational recipes in the virtual software environment (Make).
 The first step is represented by the command `make docker`.
@@
 The `&&` only runs the second command if the first was successful.
-Therefore, the computational steps are only executed, when the software environment is set up.
-The second step and actual reproduction is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy properties.
+Therefore, the computational steps are only executed when the software environment is set up.
+The second step and actual reproduction is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy parts.
 First, a call to `make` without any explicit target will build the Make target `all`.
-Second, the flag `-B` means that Make will consider all prerequisites as outdated and hence will rebuild everything.
+Second, the flag `-B` means that Make will consider all dependencies as outdated and hence will rebuild everything.
 Third, repro constructs Make targets so that if you supply `DOCKER=TRUE` they are executed within the Docker image of the project.
 
-This interplay between Docker and Make is so complicated because it resembles a chicken or egg problem.
+This interplay between Docker and Make is relatively complicated because it resembles a chicken or egg problem.
 We have computational steps (Make) which depend on the software environment (Docker) for which we again have computational steps that create it.

### 2021-07-12 manuscript.Rmd
added some lines wrt Caspar's point #5 later on

@@
 
-# Tutorial
+# Creating reproducible research projects
 
@@
 `repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
+These two files together form the basis for consistency within a research project and consistency across different systems.
+At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.
+This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
+Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).
 In everyday work, the simplest solution is to now call the function `repro::automate()` which both creates a Makefile and a Dockerfile from all R Markdown files in the project (all files with the ending `.Rmd`) based on the respective repro metadata entries.

### 2021-07-12 manuscript.Rmd
suggestions until PAC

@@
 The CLI predates and coexists with graphical user interfaces (GUIs), which many users (including the authors) prefer over the text-based interactions of a CLI for most tasks.
-A GUI has the apparent disadvantage for reproducibility that it requires a user who clicks on the correct elements (such as buttons or menu items) in the correct order.<!--AB not sure about the entire paragraph about CLI vs GUI here -  can we shorten this?-->
+A GUI has the apparent disadvantage for reproducibility that it requires a user who clicks on the correct elements (such as buttons or menu items) in the correct order.<!--AB not sure about the entire paragraph about CLI vs GUI here -  can we shorten this entire paragraph?-->
 Depending on the number of steps taken, this is not only much effort but also error-prone.
@@
 
-Let us dissect the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two steps.
+Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two processing steps.
 First it recreates the virtual software environment (Docker) and then it executes computational recipes in the virtual software environment (Make).
-The first step is represented by the command `make docker`.
-The command `make docker` will trigger Make to build the target `docker`.
-The recipe for this target builds an image from your Dockerfile.
-The `&&` only runs the second command if the first was successful.
+The first step is done by the command `make docker`.
+The command `make docker` will trigger Make to build the target called `docker`.
+The recipe for this target builds an image from the Dockerfile of the repository.
+The `&&` concatenates both commands and only runs the second command if the first was successful.
 Therefore, the computational steps are only executed when the software environment is set up.
-The second step and actual reproduction is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy parts.
+The second step executes the actual reproduction and is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy parts.
 First, a call to `make` without any explicit target will build the Make target `all`.
@@
 We have computational steps (Make) which depend on the software environment (Docker) for which we again have computational steps that create it.
-Users only require a deeper understanding when they either want to have more complex computational recipes than rendering an R Markdown or require other software than R packages.
+Users only require a deeper understanding of this interdependence when they either want to have more complex computational recipes than rendering an R Markdown or require other software than R packages.
 
@@
 `repro` creates three Dockerfiles inside the `.repro` directory.
-The file `.repro/Dockerfile_base` contains information about the base image on which all the rest is build.
+The file `.repro/Dockerfile_base` contains information about the base image on which all the remaining software is installed.
 By default we rely on the "verse" images provided by the Rocker project [@boettigerIntroductionRockerDocker2017].
-These contain (among other) the packages `tidyverse`, `rmarkdown` and a Latex installation, which make them ideal for writing scientific manuscripts.
-Users can choose which R version they want by changing the version number in line 1 to the desired R version number.
-By default the R version corresponds to the version on which `repro::automate()` was called the first time.
+These contain (among other things) the packages `tidyverse`, `rmarkdown` and a complete LaTeX installation, which makes these images ideal for the creation of scientific manuscripts.
+Users can choose which R version they want to have inside the container by changing the version number in line 1 to the desired R version number.
+By default the R version corresponds to the locally installed <!--AB: please check - did I get this correctly?
+It is the local version?-->version on which `repro::automate()` was called the first time.
 The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed, by default this date is set to the date on which `repro::automate()` was called the first time.
+This way, the call to the automate function virtually freezes the R environment to the state it was called the first time inside the container. 
+Below, you see the Docker base file we used to create this manuscript:
 
@@
 
-The file `.repro/Dockerfile_manual` is by default empty and is supposed to allow dependencies outside of R, like system libraries or external software.
+The file `.repro/Dockerfile_manual` is empty by default and can be used to add further dependencies outside of R, like system libraries or external software. 
 Remember, the software environment of the Docker image is completely separate from software installed on your computer.
@@
 
-The last Dockerfile, `.repro/Dockerfile_packages`, should not be manually changed.
-`repro::automate()` gathers all packages from all `.Rmd` files and figures out if they should be installed from CRAN or GitHub.
+The last Dockerfile, `.repro/Dockerfile_packages`, is automatically generated from the metadata provided and must not be manually changed.
+Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub.
+<!-- does this also figure out the specific version?
+If so, please say so-->
 
-The seperation into several Dockerfiles is only to provide a clear seperation between automatically generated parts and manually altered parts of the Dockerfile.
-However, Docker requires single Dockerfile so `repro::automate()` simply combines the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the
+The separation into several Dockerfiles is a design choice to provide a clear separation between automatically generated parts based on metadata provided and manually altered parts of the Dockerfile.
+Docker eventually requires a single Dockerfile to run, so `repro::automate()` simply concatenates the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the R project.
 
-Users of `repro` can not only build arbitrarily complex software environments they can also implement complex file dependencies.
-The standard `repro` metadata only makes sure that all dependencies are availible but does not allow you to specify a recipe for them in the metadata.
-If you can formulate the creation of dependencies in terms of computational steps you should include these in the Makefile.
+With this approach, users of `repro` can not only build complex software environments but they can also implement complex file dependencies.
+The standard `repro` metadata only makes sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
+If you can formulate the creation of dependencies in terms of computational steps you should include these in the Makefile.<!--AB I do not understand this sentence-->
 The Makefile that `repro` creates is only an template and you are free to change it.
-However, you want to make sure to not remove the following two lines:
+However, make sure to never remove the following two lines:
 
@@
 This file should not be altered manually.
-If you want to adept a target than simply provide an alternative target in the main `Makefile`.
+If you want to adept<!--AB strange word?! change--> a target than simply provide an alternative target in the main `Makefile`.
 Targets in the main `Makefile` take precedent.
 
-The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but seldomly should alter.
+The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but usually should not modify.
 This Makefile coordinates the interplay between Make and Docker and contains targets for building (with `make docker`) and saving (with `make save-docker`) the Docker image.
-Additionally it provides facilities to execute commands within the container.
+Additionally, it provides facilities to execute commands within the container.
 If you write computational recipe for a target it will be by default evaluated using the locally installed software.
-To instead evaluate commends inside the Docker image you should wrap them in `$(RUN1) command $(RUN2)`, like in this example:
+To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example:<!--AB please explain this hypothetical example in more detail, so that we can follow along-->
 
@@
 
-If users use
+If users execute this in the terminal
 
@@
 
-To summarise, `repro` provides untrained researchers with an easy to use solution for dependency tracking (in form of Make) and software management (using Docker).
-Users with more advanced requirements can manually change all aspects while still profiting from can mix `repro`'s automations.
+To summarise, `repro` provides researchers with an easy to use solution for dependency tracking (in form of Make) and software management (using Docker) without the necessity to learn both languages.
+Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automations.
 

### 2021-07-13 manuscript.Rmd

@@
 The second objective is to discuss opportunities of such workflows to improve not only how research outputs are created and recreated but also how research itself is conducted.
-One such opportunity is an alternative preregistration <!--AP preregistration is not explained yet; citation?--> scheme that builds upon a reproducible workflow, the so-called _preregistration as code_ (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results before they gather empirical data.
+One such opportunity is an alternative preregistration [@NosekRevolution2018] scheme that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results before they gather empirical data.
 This dynamic document closely resembles an actual journal article but with "fake" results from simulated data that serve to generate placeholders for figures, tables, and statistics. 
 This document serves as unambiguous preregistration of the planned analysis and how the results are going to be reported.
-The only parts that are missing from this first draft are the discussion of results, which will completed once the real data were gathered.
+The only parts that are missing from this first draft are the discussion of results, which the authors add after they gathered empirical data.
 
@@
 An easy-to-use standard, on the other hand, is less error-prone, produces less overhead and is thus more likely to be adopted.
-To us, it makes sense to strive for a broad adoption of easy-to-use practices and then gradually move into more rigorous standards as researchers become more proficient with the tools involved.
+To us, it makes sense to strive for a broad adoption of easy-to-use practices and then gradually move into more complex workflows as researchers become more proficient with the tools involved.
 The remainder of this paper is structured along these lines.
@@
 
-We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along on their own computers.
+We have structured the tutorial with a *learning by doing* approach in mind, such that readers can follow along on their own computers.
 We explicitly encourage readers to try out all R commands for themselves.
@@
 2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly)
+3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
 4.	**External code dependencies**; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
@@
 ```{r}
+# `package::function()` → use function from package without `library(package)`
 repro::check_git()
@@
 
-Whenever you made changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
+For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
 The easiest way to do this is to use the RStudio Git pane.
@@
 
-Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the command line.
+Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the R console.
 
@@
 The text of an R Markdown is written in Markdown. <!--AB citation needed for further information-->
-The following Markdown example serves to illustrate the Markdown syntax. 
+The following example serves to illustrate the Markdown syntax. 
 It shows how to create a heading, a word in bold font, and a list of several items in Markdown:
@@
 By default we rely on the "verse" images provided by the Rocker project [@boettigerIntroductionRockerDocker2017].
-These contain (among other things) the packages `tidyverse`, `rmarkdown` and a complete LaTeX installation, which makes these images ideal for the creation of scientific manuscripts.
+These contain (among other software) the packages `tidyverse`, `rmarkdown` and a complete LaTeX installation, which makes these images ideal for the creation of scientific manuscripts.
 Users can choose which R version they want to have inside the container by changing the version number in line 1 to the desired R version number.
-By default the R version corresponds to the locally installed <!--AB: please check - did I get this correctly?
-It is the local version?-->version on which `repro::automate()` was called the first time.
+By default the R version corresponds to the locally installed version on which `repro::automate()` was called the first time.
 The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed, by default this date is set to the date on which `repro::automate()` was called the first time.
@@
 The last Dockerfile, `.repro/Dockerfile_packages`, is automatically generated from the metadata provided and must not be manually changed.
-Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub.
-<!-- does this also figure out the specific version?
-If so, please say so-->
+Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.
 
@@
 This file should not be altered manually.
-If you want to adept<!--AB strange word?! change--> a target than simply provide an alternative target in the main `Makefile`.
+If you are not satisfied with the automatically generated target than simply provide an alternative target in the main `Makefile`.
 Targets in the main `Makefile` take precedent.

### 2021-07-13 manuscript.Rmd
more comments until l.712

@@
 
-# Introducing Preregistration as Code
+# Preregistration as Code
 
-We argue that a preregistration as code (PAC) is an excellent planning tool that offers several benefits over traditional preregistration.
-For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data, were the results are initially based on simulated data.
-We discuss PAC as an example of how a robust reproducibility standard has the potential to benefit research projects throughout the whole research process, starting at the beginning, with planning a study.
-
-We consider three criteria when planning a study.
-First, the plan should be comprehensive, e.g. state the research question and describe the study design and analysis.
-Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
-Third, the plan should be efficient; therefore, it only employs the necessary amount of resources.
-
-To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
-The introduction and theoretical background provide the basis for the studies' design and can be written before gathering data.
-We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
-To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
-The researchers can include a "faked" results section based on the simulated data, employing dynamic document generation.
-
-Repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to assure efficiency.
-For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study.
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-First, you have to address and could plan for data analytic problems on a conceptual level before you made the considerable effort to acquire data [@axelrodAdvancingArtSimulation1997].
-Second, if you only simulate a single dataset, you can perform a simple technical check, that is the analysis software outputs parameter estimates [@braiekTestingMachineLearning2020], e.g.
-when the simulated data contains missings.
-Third, if you perform a Monte Carlo simulation, you may evaluate how the data analysis copes with expected data complications, e.g.
-non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
-Fourth, you can plan your resources efficiently, e.g. determine required sample size to achieve certain precision or power [@brandmaier2018precision].
-@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.
+<!--AB this is a too steep intro.
+please add two lines on what traditional preregistration is and why we would like to do it (including references to classic papers on pregistration and its benefits.
+What are the problems (ambiguity)?
+This should be said before we offer a solution.
+I moved some parts from below to here-->
 
@@
 Third, preregistration of studies can reduce the impact of publication bias—particularly the prioritization of publishing positive over negative results—by making all planned studies discoverable whether or not they are ultimately published.
+<!--AB you need to add the line numbers to the citation-->
+
+Here we propose to use reproducible research projects with the full analysis code written in advance to data collection as comprehensive and unambiguous form of preregistration.
+We term this approach _preregistration as code (PAC)_.
+We argue that a PAC is an excellent planning tool that offers several benefits over traditional preregistration.
+For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
+In this document, the results are initially based on simulated data that mimicks the expected output.
+We discuss PAC as an example of how a robust reproducibility standard has the potential to benefit research projects throughout the whole research process, starting at the beginning, that is, when planning a study.
+
+## Benefits of PAC for study planning
+
+We consider three quality criteria to be important for juding a study plan.<!--AB: criteria for what?-->
+First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
+Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
+Third, the plan should be efficient; the plan should only require the necessary amount of resources.
+
+To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
+The introduction and theoretical background provide the basis for the studies' design and can be written before gathering data.
+We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
+To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
+The researchers can include a temporary "faked" results section based on the simulated data, employing dynamic document generation.
+This allows for an initial check whether the analysis actually runs without failure and helps to see whether the planned reporting (e.g., the selected statistics, figures, and tables) provides sufficient detail.
+
+While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
+First, you have to think about and address data analytic problems on a conceptual level before you make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
+Second, with simulated data you can perform a simple technical check of whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
+Third, repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to ensure efficiency.
+For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
+@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.
+You may also evaluate how the data analysis copes with expected data complications, e.g.
+non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
 
 The idea of submitting code as part of a preregistration is not new.
-A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods and in an informal literature research (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
+A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
+In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
 
-Why are PAC and reproducibility related?
-Preregistrations are plagued by nonreproducibility like any other research outcome.
-@bakkerRecommendationsPreregistrationsInternal2020 found from the 210 preregistrations they investigated 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to theoretically reproduce the power analysis.
+## Why are PAC and reproducibility related? <!--AB maybe turn this into a subheading? Then, please think what other subsections under PAC deserve a subheading - I added some suggestions-->
+Preregistrations are plagued by non-reproducibility like any other research outcome.
+@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to theoretically reproduce the power analysis.
 Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
-@steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
+For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
+
+## Advantages of PAC over traditional preregistration
 
 We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
+First, a PAC removes any ambiguity regarding the translation of an natural language analysis plan into code.
 Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
-Second, despite being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
-Researchers can, for example, formulate conditions under which they prefer one analysis over the other, i.e.
-if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
-Third, a PAC more comprehensive is by design because e.g. missing steps are obvious when the code is tested on simulated data.
-A due to its high structure PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
-Forth, deviations from the preregistration are more explicit because they are reflected in changes to the code, a property facilitated by version control.
-Fiveth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
+Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
+Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
+For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
+Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
+We believe it is easier with PAC to detect missing steps or ambiguous decisions.
+Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
+Forth, deviations from the preregistration are more explicit because they are reflected in changes to the code, which are managed and tracked with version control.
+Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.
 
-PACs come with a challange though, in contrast to a reproducible script, that should work on the same data, a PAC must function on different data.
-We call this property "reusability".
-This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] so it is build-for-reuse [@al-badareenReusableSoftwareComponents2010].
-The reproducible workflow we present here is heavily automated and hence promotes reusibility.
-Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readible hypothethis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
+PACs come with a challenge though.
+In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
+We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
+This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
+The reproducible workflow we present here is heavily automated and hence promotes reusability.
+Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
+Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
 
-Therefore reproducibility facilitates traditional good scientific practices and provides the fundament for promising innovations.
+Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.
 
@@
 
-To guide the reader through a preregistration as code (PAC), we will use an exemplary research question:
+To guide the reader through a preregistration as code (PAC), we will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
 
@@
 At the heart of a PAC lies the planned analysis.
-We suggest writing a function that captures the full process of testing a research question that receives a dataset as input.
-For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavillism items of the Short Dard Triad [SD3, @Jones2013] between the gender groups.
-As an example of a data dependent decisions we included a rank transformation if the skewness greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+We suggest writing a main analysis function that captures the full process of testing a research question.
+This function receives a dataset as input and performs all statistical processing. <!--AB say anything about its output? is the Markdown the output?-->
+For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
+As an example of a data-dependent decision, we included a rank transformation if the skewness was greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
 
-[^caution-mww]: The Mann–Whitney–Wilcoxon is **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
+[^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
 The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.

### 2021-07-14 preregistration.Rmd
add draft for preregistration

@@
 ---
-title: "Preregistration"
-author: "Aaron Peikert"
-date: "4/12/2021"
-output: html_document
-params:
-  nsim: 2
-  nmin: 100
-  nmax: 300
-  nstep: 100
-  seed: 1234
-repro:
-  packages:
-    - tidyverse
-    - here
-    - lavaan
-    - furrr
-    - future.batchtools
-    - slider
-    - patchwork
-    - svglite
-  scripts:
-    - R/simulation.R
-    - R/hash.R
+title: "A Preregistered Comparison of Male and Female Machiavellianism"
+date: "7/10/2021"
+author:
+  - name: Aaron Peikert
+    affil: 1, *
+    orcid: 0000-0001-7813-818X
+  - name: Caspar J. Van Lissa
+    affil: 2, 3
+    orcid: 0000-0002-0808-5024
+  - name: Andreas M. Brandmaier
+    affil: 1, 4
+    orcid: 0000-0001-8765-6982
+affiliation:
+  - num: 1
+    address: |
+      Center for Lifespan Psychology---Max Planck Institute for Human Development
+      Lentzeallee 94, 14195 Berlin, Germany
+  - num: 2
+    address: |
+      Department of Methodology & Statistics---Utrecht University faculty of Social and Behavioral Sciences, Utrecht, Netherlands
+  - num: 3
+    address: |
+      Open Science Community Utrecht, Utrecht, Netherlands
+  - num: 4
+    address: |
+      Max Planck UCL Centre for Computational Psychiatry and Ageing Research
+      Berlin, Germany and London, UK
+# firstnote to eighthnote
+correspondence: |
+  peikert@mpib-berlin.mpg.de
+journal: psych
+type: tutorial
+status: submit
+bibliography: preregistration.bib
+abstract: |
+  We test for group differences in machiavellism between self identified males and females in the MACH-IV in a well powered preregistered study.
+output:
+  bookdown::markdown_document2:
+    base_format: rticles::mdpi_article
+header-includes:
+  -   \def\baselinestretch{0.98}
 ---
+
+\setlength{\bibsep}{0cm}
+\titlespacing{\section}{.1cm}{.1cm}{.1cm}
 
@@
 library(tidyverse)
-library(lavaan)
-library(furrr)
-library(slider)
-library(patchwork)
-source(here::here("R", "simulation.R"))
-repro::automate_load_scripts()
-# if you have access to a hpc envir specify this in R/hpc.R
-# if not we use local multicore
-# to speed up consider reduce nsim; increase nstep in ↑YAML↑
-
-hpc_config <- here::here("R", "hpc.R")
-if(fs::file_exists(hpc_config)){
-  source(hpc_config)
-} else {
-  plan(list(transparent,
-       tweak(multisession, workers = 4L)))
-}
-# debug:
-#   plan(transparent)
+library(report)
+source(here::here("R", "simulation_funs.R"))
 ```
 
-```{r}
-loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
-cohend_jones_paulhus <- c(24, 29, 35)/100
+Theoretical Background
+======================
+
+There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self identified males and females.
+@muris2017 note that studies focusing on gender differences in machiavellianism are sparse.
+We therefore want to add to this body of literature.
+
+Method
+======
+
+```{r, echo=FALSE}
+res <-
+  readr::read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddddd")
+choosen_power <- .8
+choosen_d <- .2
+minn <- filter(res, d == choosen_d, power > choosen_power) %>% 
+  filter(n == min(n))
 ```
 
-```{r sim}
-n_obs <- seq(params$nmin, params$nmax, params$nstep)
-n_sim <- params$nsim
-# n_sim * length(n_obs)
-# [1] 100000
-to_export <- ls_funs() %>% map(get)
-to_export <-
-  c(to_export, list(
-    n_obs = seq(params$nmin, params$nmax, params$nstep),
-    n_sim = params$nsim
-  ))
-res_raw %<-% simulation_study(n_obs, n_sim,
-                furrr_options(
-                  globals = to_export,
-                  seed = seed,
-                  packages = c("furrr", "lavaan", "tidyverse")
-                ))
+We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study [@simmons2012].
+We use data available from [openpsychometrics.org](https://openpsychometrics.org/_rawdata/) from the online version of the MACH-IV [@christie1970]and included participants that have responded to at least one machiavellianism item and reported their gender as either "male" or "female".
+We conduct a t-test on the average of machiavellianism items, if the skew of this avarage is below 1.0, else we conduct a Mann--Whitney--Wilcoxon test between the gender groups.
+A simulation we conducted indicated that with a sample size of `r minn$n` for an alpha of .05 (two-sided) we archive at least `r choosen_power*100`% power assuming an effect size of `r choosen_d`.
+
+Results
+=======
+
+```{r, echo=FALSE, results='asis'}
+real_data <- here::here("data", "data.csv")
+simulated <- !fs::file_exists(real_data)
+if(simulated){
+  cat("\\textcolor{red}{This document and the results presented herein are based on simulated data and should not be interpreted.}")
+  set.seed(1234)
+  mach <- simulate_data(900, 8, 0.3, 10)
+} else {
+  mach <- readr::read_csv(real_data)
+}
 ```
 
-```{r write-out, error=TRUE, include=FALSE}
-invisible(res_raw)
-write_rds(res_raw, here::here("res_raw.rds"))
+```{r, echo=FALSE, results='asis'}
+analysis <- planned_analysis(mach)
+report <- function(analysis) {
+  params <- report::report_parameters(analysis$test)
+  table <- attributes(params)$table
+  model <- report::report_model(analysis$test, table = table)
+  if (analysis$use_rank)
+    model <-
+    str_replace(model,
+                fixed("Welch Two Sample t-test"),
+                "Mann--Whitney--Wilcoxon test")
+  cat("The ", model, " suggests that the effect is ", params, sep = "")
+}
 ```
 
-```{r hash}
+```{r, echo=FALSE, results='asis'}
 ```
 
-```{r}
-res <- res_raw %>% mutate(n_obs = as.numeric(n_obs))
-res_interval <- res %>%
-  select(n_obs, std_estimate) %>% 
-  filter(!is.na(std_estimate)) %>% 
-  group_by(n_obs) %>% 
-  summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
-            .groups = "drop", median = median(std_estimate)) %>% 
-  unnest(c(interval)) %>% 
-  mutate(width = upper - lower, perc = 1 - alpha) %>% 
-  group_by(alpha) %>% 
-  arrange(n_obs) %>% 
-  mutate(across(c(width, upper, lower),
-                ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE))) %>% 
-  ungroup() %>% 
-  arrange(perc)
+Discussion
+==========
+
+```{r, echo=FALSE, results='asis'}
+if(simulated){
+  cat("\\textcolor{red}{This document was created before we have gathered any data, we therefore have nothing to discuss.}")
+}
 ```
 
-```{r}
-choosen_perc <- .8
-choosen_width <- .2
-true_diff <- .2
-minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
-  filter(n_obs == min(n_obs))
-```
-
-```{r}
-plot_interval <- res_interval %>%
-  ggplot(aes(
-    n_obs,
-    ymin = lower,
-    ymax = upper,
-    fill = perc,
-    group = alpha
-  )) +
-  geom_ribbon() +
-  geom_line(aes(y = median), color = "white") +
-  geom_hline(yintercept = true_diff, linetype = "dotted") +
-  scale_fill_viridis_c(
-    begin = 1,
-    end = 0,
-    breaks  = 1 - seq(0.05, .2, 0.05),
-    labels = . %>% scales::percent(accuracy = 1)
-  ) +
-  scale_y_continuous(n.breaks = 10) +
-  scale_x_continuous(breaks = seq(0, params$nmax, params$nstep * 10)) +
-  theme_minimal() +
-  labs(x = "N per Group", y = "Standardized Effect Size") +
-  theme(legend.title = element_blank()) +
-  NULL
-```
-
-```{r}
-plot_width <- res_interval %>%
-  ggplot(aes(n_obs, width, color = perc, group = perc)) +
-  geom_line() +
-  geom_point(
-    data = minn,
-    size = 3,
-    color = "black",
-    shape = 3
-  ) +
-  scale_color_viridis_c(
-    begin = 1,
-    end = 0,
-    breaks = 1 - seq(0.05, .2, 0.05),
-    labels = . %>% scales::percent(accuracy = 1)
-  ) +
-  theme_minimal() +
-  scale_y_log10(breaks = seq(0, 0.8, .1)) +
-  scale_x_continuous(breaks = seq(0, params$nmax, params$nstep * 10)) +
-  labs(x = "N per Group", y = "Width of quantile\n(Standardized Effect Size)") +
-  theme(legend.title = element_blank()) +
-  NULL
-```
-
-```{r}
-simulation_plot <- plot_interval +
-  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + 
-  plot_width +
-  plot_layout(nrow = 2, guides = 'collect') +
-  plot_annotation(tag_levels = 'A')
-simulation_plot
-write_rds(simulation_plot, "simulation_plot.rds")
-ggsave("simulation.svg", device = svglite::svglite, width = 7, height = 4)
-```
-
-Plot A depicts the range of effect sizes we observed in the simulation, where the colour encodes the density.
-The light yellow interval shows where 80% of the estimates land.
-The dotted line represents the simulated true effect size; the white line shows the median observed effect sizes.
-
-Plot B represents the width of the interval shown in Plot A on a logarithmic scale.
-The cross shows the minimum required sample size so that 80% of the estimates land in a range of +/- .1.
-
-```{r}
-nmax <- filter(res_interval, n_obs == max(n_obs))
-bias <- (slice(nmax, 1)$median - true_diff)/true_diff
-```
-
-We can therefore conclude, that when we want to plan a study where we expect that `r scales::percent(choosen_perc)` of the estimates fall into a range of +/- `r choosen_width/2`, we need a minimum sample size of `r minn$n_obs` per group.
-The simulation also shows that we overestimate the effect size by `r scales::percent(bias)` (true: `r true_diff`, median of observed: `round(slice(nmax, 1)$median, 2)`) with a sample size per group of `r slice(nmax, 1)$n_obs`.
-
-# Session Info
-
-```{r}
-if(fs::file_exists(hpc_config)){
-  list(local = sessioninfo::session_info(),
-     login_node = value(future(sessioninfo::session_info())),
-     worker = value(future(value(future(sessioninfo::session_info())))))
-} else {
-  sessioninfo::session_info()
-}
-```
+References
+==========

### 2021-07-14 manuscript.Rmd
strike section CLI vs GUI closes #139

--- 2021-07-13 manuscript.Rmd
@@
     long: Kilobyte
-  - short: GUI
-    long: Graphical User Interface
-  - short: CLI
-    long: Comand Line Interface
   - short: CRAN
@@
 
-A major barrier to adoption of the above tools for researchers is that they are typically used via a command line interface (CLI).
-The CLI predates and coexists with graphical user interfaces (GUIs), which many users (including the authors) prefer over the text-based interactions of a CLI for most tasks.
-A GUI has the apparent disadvantage for reproducibility that it requires a user who clicks on the correct elements (such as buttons or menu items) in the correct order.<!--AB not sure about the entire paragraph about CLI vs GUI here -  can we shorten this entire paragraph?-->
-Depending on the number of steps taken, this is not only much effort but also error-prone.
-This tutorial seeks a compromise.
-We use simple GUI actions to create a reproducible workflow template, but the reproduction of such workflow does not rely on manual steps.
-<!--AB I modified this sentence - but I am not sure what we are saying here with the "manual steps"-->
-At several points, we rely on or have developed a CLI that wraps a more complicated (and powerful) CLI.
-These wrappers use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
+The above mentioned tools require much expertise that is beyond the traditional curriculum of researchers.
+To aid researcher adopting these tools we rely on or have developed helpers that wrap more complicated (and powerful) tools.
+These wrappers use standard tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
 We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].

### 2021-07-14 manuscript.Rmd
better guide readers through the advanced section closes #145 closes #147

@@
 Some projects, however, are not as easy to make reproducible.
-As explained above, repro is merely a simplified interface that guarantuees a high level of reproducibility for simple projects.
+As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
 These simplifications are possible because of two restrictions.
-First, the project can only rely on R packages for the Dockerfile and second, it can have no multilayered preprocessing for the Makefile.<!--AB: please explain both points in more detail-->
+Users who ask themselves either "How can I install software that is not an R package in the Docker image?" or "How can I express complex dependencies between files, e.g.
+raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
+All other users may safely skip this section and may return to it if they become limited by restriction.
+
+The metadata the `repro::automate()` function relies on can only express dependencies R packages for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
+The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
+Other software that users might need like other programming languages, not already installed latex packages etc. have to be added manually.
+We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g.
+for system libraries (via `apt` the Ubuntu package manager), Latex packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
+The second limitation is about dependencies.
+Make can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
+If B is missing in this example, Make would know how to recreate it from C and D.
+These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
+Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D therefore leaving out important information or express the dependencies directly within the Makefile.
+
+The following section explains how to overcome these limitation while still relying on the automatons afforded by `repro`.
 Lifting these restrictions requires the user to interact more directly with Make or Docker.
@@
 The standard `repro` metadata only makes sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
-If you can formulate the creation of dependencies in terms of computational steps you should include these in the Makefile.<!--AB I do not understand this sentence-->
+If you can formulate the creation of dependencies in terms of computational steps, e.g.
+the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the Makefile.
 The Makefile that `repro` creates is only an template and you are free to change it.

### 2021-07-14 manuscript.Rmd
detail what has changed for Make in the advanced section closes #146

@@
 If you write computational recipe for a target it will be by default evaluated using the locally installed software.
-To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example:<!--AB please explain this hypothetical example in more detail, so that we can follow along-->
-
+To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first Make example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
+2:
 ```bash
@@
 
-If users execute this in the terminal
+If users execute this in the terminal:
 
@@
 ```
-
-this is simply translated to:
+It behaves exactly as the in the first Make example, the script `R/simulate.R` is run using the locally installed R.
+Because this translates simply to:
 
@@
 
-it is evaluated within the Docker container:
+it is evaluated within the Docker container using the software within it and not the locally installed R version:
 

### 2021-07-14 manuscript.Rmd
code comunicates closes #136

@@
 Dynamic document generation integrates both steps.
-This paradigm can be traced back to @knuthCWEBSystemStructured who suggested that a computer prorgram should be interspersed with explanations of its logic in a natural language.
-<!--AB I removed the quotation because it is not really clear to me how it matches what we say about R Markdown - I am not sure whether my changes make things better, though-->
+Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allows interested readers deeper understanding [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
 **R Markdown** uses Markdown for writing the text and R (or other programming languages) for the analysis.

### 2021-07-14 manuscript.Rmd
closes #141

@@
 This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
+@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.
 
@@
 
-The text of an R Markdown is written in Markdown. <!--AB citation needed for further information-->
+The text of an R Markdown is written in Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
 The following example serves to illustrate the Markdown syntax. 

### 2021-07-14 manuscript.Rmd
final comments

--- 2021-07-14 manuscript.Rmd
@@
     long: Kilobyte
+  - short: GUI
+    long: Graphical User Interface
+  - short: CLI
+    long: Comand Line Interface
   - short: CRAN
@@
 
-The above mentioned tools require much expertise that is beyond the traditional curriculum of researchers.
-To aid researcher adopting these tools we rely on or have developed helpers that wrap more complicated (and powerful) tools.
-These wrappers use standard tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
+A major barrier to adoption of the above tools for researchers is that they are typically used via a command line interface (CLI).
+The CLI predates and coexists with graphical user interfaces (GUIs), which many users (including the authors) prefer over the text-based interactions of a CLI for most tasks.
+A GUI has the apparent disadvantage for reproducibility that it requires a user who clicks on the correct elements (such as buttons or menu items) in the correct order.<!--AB not sure about the entire paragraph about CLI vs GUI here -  can we shorten this entire paragraph?-->
+Depending on the number of steps taken, this is not only much effort but also error-prone.
+This tutorial seeks a compromise.
+We use simple GUI actions to create a reproducible workflow template, but the reproduction of such workflow does not rely on manual steps.
+<!--AB I modified this sentence - but I am not sure what we are saying here with the "manual steps"-->
+At several points, we rely on or have developed a CLI that wraps a more complicated (and powerful) CLI.
+These wrappers use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
 We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
@@
 Some projects, however, are not as easy to make reproducible.
-As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
+As explained above, repro is merely a simplified interface that guarantuees a high level of reproducibility for simple projects.
 These simplifications are possible because of two restrictions.
-Users who ask themselves either "How can I install software that is not an R package in the Docker image?" or "How can I express complex dependencies between files, e.g.
-raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
-All other users may safely skip this section and may return to it if they become limited by restriction.
-
-The metadata the `repro::automate()` function relies on can only express dependencies R packages for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
-The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
-Other software that users might need like other programming languages, not already installed latex packages etc. have to be added manually.
-We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g.
-for system libraries (via `apt` the Ubuntu package manager), Latex packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
-The second limitation is about dependencies.
-Make can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
-If B is missing in this example, Make would know how to recreate it from C and D.
-These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
-Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D therefore leaving out important information or express the dependencies directly within the Makefile.
-
-The following section explains how to overcome these limitation while still relying on the automatons afforded by `repro`.
+First, the project can only rely on R packages for the Dockerfile and second, it can have no multilayered preprocessing for the Makefile.<!--AB: please explain both points in more detail-->
 Lifting these restrictions requires the user to interact more directly with Make or Docker.
@@
 The standard `repro` metadata only makes sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
-If you can formulate the creation of dependencies in terms of computational steps, e.g.
-the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the Makefile.
+If you can formulate the creation of dependencies in terms of computational steps you should include these in the Makefile.<!--AB I do not understand this sentence-->
 The Makefile that `repro` creates is only an template and you are free to change it.
@@
 If you write computational recipe for a target it will be by default evaluated using the locally installed software.
-To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first Make example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
-2:
+To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example:<!--AB please explain this hypothetical example in more detail, so that we can follow along-->
+
 ```bash
@@
 
-If users execute this in the terminal:
+If users execute this in the terminal
 
@@
 ```
-It behaves exactly as the in the first Make example, the script `R/simulate.R` is run using the locally installed R.
-Because this translates simply to:
+
+this is simply translated to:
 
@@
 
-it is evaluated within the Docker container using the software within it and not the locally installed R version:
+it is evaluated within the Docker container:
 
@@
 At the heart of a PAC lies the planned analysis.
-We suggest writing a function that captures the full process of testing a research question that receives a dataset as input.
-For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavillism items of the Short Dard Triad [SD3, @Jones2013] between the gender groups.
-As an example of a data dependent decisions we included a rank transformation if the skewness greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+We suggest writing a main analysis function that captures the full process of testing a research question.
+This function receives a dataset as input and performs all statistical processing. <!--AB say anything about its output? is the Markdown the output?-->
+For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
+As an example of a data-dependent decision, we included a rank transformation if the skewness was greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
 
-[^caution-mww]: The Mann–Whitney–Wilcoxon is **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
+[^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
 The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
 
-This function is one possible translation of the verbal descriptions:
+This function is one possible translation of the verbal description above:
 
@@
 
-One point of note about the analytical strategy and the function implementing it is that there are several translations into code that would fit the descriptions but would produce different results.
+Importantly, natural language descriptions do not map unambiguously onto computer code implementations. 
+The same textual descriptions may be implemented differently by different persons depending on their statistical and programming knowledge but also based on their own expectations and beliefs.
 One example would be using the function `wilcox.test` instead of the combinations of `rank` and `t.test`.
-Both are valid implementation of Mann–Whitney–Wilcoxon test but the first assumes equal variance while the second applies Welch's correction and hence is robust even with unequal variances [@zimmermanRankTransformationsPower1993].
-Mentioning every such minute implementation detail is would result in overly verbose preregistrations but does change the outcome.
+Both are valid implementation of the Mann–Whitney–Wilcoxon test but the first assumes equal variance whereas the second applies Welch's correction and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
+Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
+Still, these details can make a difference and provide unwanted researcher degrees-of-freedom.
 
-We also want to caution researchers to not use the here shown procedure.
-We merely choose it for didactic purposes because it may be familiar to many researchers because the t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
-This analytical strategie is suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]);
+Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
+The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
+The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
+<!--AB can we move this to a different place?
+Maybe this would actually better fit to the discussion/limitation section!-->
 
-To use this function one needs a compatible dataset.
+To use this function one needs a compatible dataset. <!--AB what is "this" function? What does "compatible" mean?-->
 At the time of preregistration there is of course no actual data available.
-We therefore simulate a dataset.
+We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
 
@@
 ```
+<!--AB it is odd that the return variable is called effect because it is a dataset-->
 
-We decided to draw non-normal data to test how well our procedure performed under non-optimal situations.
-Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a standardized effect size.
-This simulated dataset fits perfectly into our planned analysis:
+Here, we decided to randomly draw non-normal data to test how well our procedure performed under non-optimal situations.
+Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a given standardized effect size (a Cohen's d of .2).
+This simulated dataset fits perfectly into our planned analysis with a sample size of 400 and 10 items: <!--AB in what way? please explain-->
 
@@
 
-These results are based on simulated data:
+Using the `report` package, we can obtain a preview of how our reporting will look like based on simulated data:
 
-```{r, results='asis'}
+```{r, results='asis', eval=FALSE}
 report::report(results$test)
@@
 
+This call to the `report()` function will yield the following automatically generated output:
+
+```{r, results='asis', echo=FALSE}
+report::report(results$test)
+```
+<!--AB it is difficult to discern the automatically generated text and out manuscript text even though there is the horizontal line. Not sure how to improve this.-->
 ***
@@
 If we want to detect an effect of size `r choosen_d` with an `r choosen_power*100`% probability, our simulation shows that we require at least `r minn$n` participants.
+
+<!--AB please add a figure number and a proper citation to the figure. Also, tell people where they can find the code to run the power analysis themselves-->
 
@@
 ```
+
+<!--AB: add some text explaining what to once the final data is available-->
+
+# Discussion
+
+<!--AB: TODO add a quick summary -->
+
+# Limitations
+
+<!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->
+
+# Outlook
+
+<!--AB: two lines of what the future of the package, of such standards, and prereigstration is.
+Maybe say something that preregistration is not always possible or wanted but that reproducibility is almost always a good idea?-->
 

### 2021-07-14 manuscript.Rmd
added two more ideas

@@
 
-<!--AB: add some text explaining what to once the final data is available-->
+<!--AB: add some text explaining what to once the final data is available.
+What parts have to be replaced in the code and in the project structure; also recommend tagging the different versions with releases.
+Do we have functions in repro to do this?
+Can we have something like repro::release("preprint") or repro::release("preregistration") -- that would be nice!
+-->
 

### 2021-07-14 manuscript.Rmd
moved t test part to limitations

@@
 
-Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
-The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
-The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-<!--AB can we move this to a different place?
-Maybe this would actually better fit to the discussion/limitation section!-->
-
+<!--AB here is kind of a break in the text-->
 To use this function one needs a compatible dataset. <!--AB what is "this" function? What does "compatible" mean?-->
@@
 We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
+Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
+The following code simulates a dataset with one observed variable in two groups.
 
@@
 
-Using the `report` package, we can obtain a preview of how our reporting will look like based on simulated data:
+Next, we add dynamic code blocks to our Markdown report that implement the planned reporting of the final results.
+A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures.
+<!--AB add citations to some of the packages papaja?
+stargazer?
+and more...-->
+Here we use the `report` package to create a preview of how our reporting will look like based on simulated data:<!--AB add citation to the report package-->
 
@@
 <!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->
+Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
+The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
+The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
+<!--AB can we move this to a different place?
+Maybe this would actually better fit to the discussion/limitation section!-->
 

### 2021-07-14 manuscript.Rmd
remove comment #159

@@
 
-<!--AB: add some text explaining what to once the final data is available.
-What parts have to be replaced in the code and in the project structure; also recommend tagging the different versions with releases.
-Do we have functions in repro to do this?
-Can we have something like repro::release("preprint") or repro::release("preregistration") -- that would be nice!
--->
-
 # Discussion

### 2021-07-15 manuscript.Rmd
recommend inclusion of hash closes #101

@@
     - gert
-    - aaronpeikert/repro@7bfaf98
+    - aaronpeikert/repro@ba1dda75
     - here
@@
 ````
+
+The results of your code are than dynamically inserted into the document.
+Dynamically means that whenever you turn the R Markdown into its output format the code is executed again and if you changed something the results are updated.
+The output document is than static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
+This static document is somewhat decoupled from the R Markdown and the rest of the code that did create it.
+It is quite possible that several slightly different versions are therefore circulating.
+We therefore recomend to connect the static document to the version control system by dynamically including the identifier of the identifier of the Git commit, called hash, in the document.
+The `repro` package comes with the function `repro::current_hash()` for that purpose.
+This document was created from the commit with the hash ``r repro::current_hash()``.
 

### 2021-07-15 manuscript.Rmd
rerun() -> reproduce() closes #143

@@
     - gert
-    - aaronpeikert/repro@ba1dda75
+    - aaronpeikert/repro@368b897d
     - here
@@
 ```{r}
-repro::rerun()
+repro::reproduce()
 ```
-<!--AB: I still think this should be repro::reproduce() please change!-->
+
 Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.

### 2021-07-15 manuscript.Rmd
how to avoid the terminal for repro::reproduce() draft for #144

@@
     - gert
-    - aaronpeikert/repro@368b897d
+    - aaronpeikert/repro@c3cae1f
     - here
@@
 
+If you uncomfortable with using the terminal directly, you can sent the command to the terminal from within R:
+
+```{r, eval=FALSE}
+system(reproduce())
+```
+
 Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.

### 2021-07-15 manuscript.Rmd
add pagenumber for direct quote closes #138

@@
 One comprehensive approach to software management is containerization.
-"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
-<!--AB: direct quotes need a line number!
-See APA style guide-->
+"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p.
+174].
+<!--APA requires no line number, but pagenumber or section heading-->
+
 Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
@@
 Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
-Preregistration increases the credibility of empirical results in three ways, @nosekPreregistrationHardWorthwhile2019:
+Preregistration increases the credibility of empirical results in three ways, @nosekPreregistrationHardWorthwhile2019, p. 2:
 
@@
 Third, preregistration of studies can reduce the impact of publication bias—particularly the prioritization of publishing positive over negative results—by making all planned studies discoverable whether or not they are ultimately published.
-<!--AB you need to add the line numbers to the citation-->
+<!--APA requires no line number, but pagenumber or section heading-->
 

### 2021-07-15 manuscript.Rmd
what do we consider while planning

@@
 
-We consider three quality criteria to be important for juding a study plan.<!--AB: criteria for what?-->
+We consider three criteria when planning a study.
 First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.

### 2021-07-15 manuscript.Rmd
cite LaTeX closes #137

@@
 This way, Markdown does not need any specialized software for editing.
-It is highly fool-proof (unlike, for example, LaTeX <!--AB citation needed!-->), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+It is highly fool-proof [unlike, for example, LaTeX @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
 We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your résumé [@vitae].
@@
 The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
-Other software that users might need like other programming languages, not already installed latex packages etc. have to be added manually.
+Other software that users might need like other programming languages, not already installed LaTeX packages etc. have to be added manually.
 We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g.
-for system libraries (via `apt` the Ubuntu package manager), Latex packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
+for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
 The second limitation is about dependencies.

### 2021-07-15 manuscript.Rmd
bring back all the formatting packages closes #154

@@
 Next, we add dynamic code blocks to our Markdown report that implement the planned reporting of the final results.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures.
-<!--AB add citations to some of the packages papaja?
-stargazer?
-and more...-->
+A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
+Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
+The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
 Here we use the `report` package to create a preview of how our reporting will look like based on simulated data:<!--AB add citation to the report package-->

### 2021-07-17 manuscript.Rmd
rewrite introduction to PAC closes #151 adress some points in #149

@@
 
-<!--AB this is a too steep intro.
-please add two lines on what traditional preregistration is and why we would like to do it (including references to classic papers on pregistration and its benefits.
-What are the problems (ambiguity)?
-This should be said before we offer a solution.
-I moved some parts from below to here-->
+Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
+Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
+For an a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
+Preregistrations, however, have a problem.
+Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
+The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates, like the ones supplied by [The Open Science Framework](https://osf.io/).
+The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
+But these templates can not remove researchers degrees of freedoms completely, because it is impossible to describe every minute detail of an analysis for all but the simplest analysis with natural language.
 
-Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
-Preregistration increases the credibility of empirical results in three ways, @nosekPreregistrationHardWorthwhile2019, p. 2:
+Our proposed solution is simple.
+Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
+We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
+PAC has the potential to all but eliminate researchers degrees of freedom.
 
-> First, preregistration of analysis plans makes clear which analyses were planned a priori and which were conducted post hoc.
-This improves calibration of uncertainty for unplanned analyses, and diagnosticity of statistical inferences for planned analyses.
-[...] Doing so has the benefit of strengthening statistical inferences as compared with unplanned analyses.
-[...] Second, preregistration enables detection of questionable research practices such as selective outcome reporting (John et al.
-2012) or Hypothesizing After the Results are Known (HARKing; Kerr, 1998).
-Third, preregistration of studies can reduce the impact of publication bias—particularly the prioritization of publishing positive over negative results—by making all planned studies discoverable whether or not they are ultimately published.
-<!--APA requires no line number, but pagenumber or section heading-->
+The idea of submitting code as part of a preregistration is not new.
+A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
+In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
 
-Here we propose to use reproducible research projects with the full analysis code written in advance to data collection as comprehensive and unambiguous form of preregistration.
-We term this approach _preregistration as code (PAC)_.
-We argue that a PAC is an excellent planning tool that offers several benefits over traditional preregistration.
+We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.
+
+## Why are PAC and reproducibility related?
+
+First, preregistrations are plagued by non-reproducibility like any other research outcome.
+@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to theoretically reproduce the power analysis (they did not try to reproduce these in practice).
+Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
+For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
+
+Second, although it has been almost a decade since @steegenMeasuringCrowdAgain2014 have preregistered their code, the idea of publishing code before analyzing data has gained not much traction.
+This could change when combining the here proposed workflow and PAC.
+We want to extend PAC from preregistering code to preregistering *the whole manuscript as dynamic document*.
+The idea is as follows.
 For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
-In this document, the results are initially based on simulated data that mimicks the expected output.
-We discuss PAC as an example of how a robust reproducibility standard has the potential to benefit research projects throughout the whole research process, starting at the beginning, that is, when planning a study.
+In this document, the results are initially based on simulated data that mimics the expected output.
+
+Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in how to write a scientific manuscript.
+It follows that for preregistration they should stick to widely accepted standards of scientific writing [e.g. @apa7].
+We have already explained the need for reproducibility standards in the process of writing such manuscript and it is therefore only natural to extend these arguments to preregistration.
+
+## Advantages of PAC over traditional preregistration
+
+We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
+First, a PAC removes any ambiguity regarding the translation of an natural language analysis plan into code.
+Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
+Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
+For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
+Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
+We believe it is easier with PAC to detect missing steps or ambiguous decisions.
+Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
+Forth, deviations from the preregistration are more explicit because they are reflected in changes to the code, which are managed and tracked with version control.
+Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.
 
@@
 
+Leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
 We consider three criteria when planning a study.
@@
 non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
-
-The idea of submitting code as part of a preregistration is not new.
-A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
-In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
-
-## Why are PAC and reproducibility related? <!--AB maybe turn this into a subheading? Then, please think what other subsections under PAC deserve a subheading - I added some suggestions-->
-Preregistrations are plagued by non-reproducibility like any other research outcome.
-@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to theoretically reproduce the power analysis.
-Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
-For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
-
-## Advantages of PAC over traditional preregistration
-
-We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, a PAC removes any ambiguity regarding the translation of an natural language analysis plan into code.
-Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
-Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
-Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
-For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
-Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
-We believe it is easier with PAC to detect missing steps or ambiguous decisions.
-Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
-Forth, deviations from the preregistration are more explicit because they are reflected in changes to the code, which are managed and tracked with version control.
-Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.
-
-PACs come with a challenge though.
-In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
-We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
-This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
-The reproducible workflow we present here is heavily automated and hence promotes reusability.
-Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
-
-Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.
 
@@
 
+PACs come with a challenge though.
+In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
+We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
+This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
+The reproducible workflow we present here is heavily automated and hence promotes reusability.
+Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
+Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
+
+Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.
+
 <!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->

### 2021-07-17 manuscript.Rmd
typo
some remarks

@@
 The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
-But these templates can not remove researchers degrees of freedoms completely, because it is impossible to describe every minute detail of an analysis for all but the simplest analysis with natural language.
+But these templates can not remove researchers degrees of freedoms completely, because it is impossible to describe every detail of an analysis for all but the simplest analysis with natural language.
+# Hier vielleicht schon ein Beispiel?
 
-Our proposed solution is simple.
-Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
+Our proposed solution is simple: Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
 We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
-PAC has the potential to all but eliminate researchers degrees of freedom.
+PAC has the potential to all but eliminate researchers degrees of freedom. # Den Satz verstehe ich nicht
 
@@
 
-We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.
+We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community. # Sehr gut
 
@@
 First, preregistrations are plagued by non-reproducibility like any other research outcome.
-@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to theoretically reproduce the power analysis (they did not try to reproduce these in practice).
+@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to reproduce the power analysis.
 Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
@@
 This could change when combining the here proposed workflow and PAC.
-We want to extend PAC from preregistering code to preregistering *the whole manuscript as dynamic document*.
+We want to extend PAC from preregistering code to preregistering *the whole manuscript as a dynamic document*.
 The idea is as follows.
@@
 We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, a PAC removes any ambiguity regarding the translation of an natural language analysis plan into code.
-Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
+First, a PAC removes any ambiguity regarding the translation of an analysis plan into code. # I would remove the term "natural language" everywhere
+Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions, formulated as code. # Every data-depended decision that can be preregistered can be formulated as code
 Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
@@
 Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
-Forth, deviations from the preregistration are more explicit because they are reflected in changes to the code, which are managed and tracked with version control.
+Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
 Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.
@@
 Leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
-We consider three criteria when planning a study.
+To use PAC for study planing, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
+To that end, they should borrow from widely employed standards for writing an empirical manuscript [e.g.
+@apa7].
+Then they should translate their method section into executable computer code and simulate data that has the same structure as the data they expect to obtain from their experiments.
+
+We consider three domains where this procedure can help planning a study.
 First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
@@
 Third, the plan should be efficient; the plan should only require the necessary amount of resources.
+
+### Comprehensiveness
+
+### Efficiency
+
+### Effectiveness
+
+# everything that is written here should be sorted under the specific point
+The introduction and theoretical background provide the basis for the studies' design.
+We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
+To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
+The researchers can include a temporary "faked" results section based on the simulated data, employing dynamic document generation.
+This allows for an initial check whether the analysis actually runs without failure and helps to see whether the planned reporting (e.g., the selected statistics, figures, and tables) provides sufficient detail.
 

### 2021-07-18 manuscript.Rmd
natural language -> verbal closes #162

@@
 The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
-But these templates can not remove researchers degrees of freedoms completely, because it is impossible to describe every detail of an analysis for all but the simplest analysis with natural language.
-# Hier vielleicht schon ein Beispiel?
+But these templates can not remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
 
-Our proposed solution is simple: Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
+Our proposed solution is simple:
+Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
 We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
-PAC has the potential to all but eliminate researchers degrees of freedom. # Den Satz verstehe ich nicht
+PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration afford.
 
@@
 
-We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community. # Sehr gut
+We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.
 
@@
 Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in how to write a scientific manuscript.
-It follows that for preregistration they should stick to widely accepted standards of scientific writing [e.g. @apa7].
+It follows that for preregistrations they should stick to widely accepted standards of scientific writing [e.g. @apa7].
 We have already explained the need for reproducibility standards in the process of writing such manuscript and it is therefore only natural to extend these arguments to preregistration.
@@
 We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, a PAC removes any ambiguity regarding the translation of an analysis plan into code. # I would remove the term "natural language" everywhere
-Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions, formulated as code. # Every data-depended decision that can be preregistered can be formulated as code
+First, a PAC removes any ambiguity regarding the translation of an verbal analysis plan into code.
+Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
 Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
@@
 To use PAC for study planing, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
-To that end, they should borrow from widely employed standards for writing an empirical manuscript [e.g.
-@apa7].
+To that end, they should borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
 Then they should translate their method section into executable computer code and simulate data that has the same structure as the data they expect to obtain from their experiments.
@@
 
-# everything that is written here should be sorted under the specific point
 The introduction and theoretical background provide the basis for the studies' design.

### 2021-07-18 manuscript.Rmd
rewrite advantage for study planning

@@
 Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For an a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
+For a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
 Preregistrations, however, have a problem.
 Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
-The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates, like the ones supplied by [The Open Science Framework](https://osf.io/).
+The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
 The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
@@
 We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
-PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration afford.
+PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration despite the best efforts afford.
 
@@
 We want to extend PAC from preregistering code to preregistering *the whole manuscript as a dynamic document*.
-The idea is as follows.
+The idea is as follows:
 For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
@@
 
-Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in how to write a scientific manuscript.
+Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
 It follows that for preregistrations they should stick to widely accepted standards of scientific writing [e.g. @apa7].
@@
 
-### Efficiency
+Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
+One particular advantage is the strong foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
+Nevertheless it can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section.
+
+Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
+While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
+Simulating data forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
+It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
+Writing incomplete code is technically impossible or at least researchers would notice if they e.g. forgot to aggregate the data etc.
+For this technical check researcher do not necessarily have to simulated data, they could also rely on synthetic data.
+Synthetic data means that they apply measures to obfuscate true relations in the data, e.g. shuffle the main dependent variable etc.
 
@@
 
-The introduction and theoretical background provide the basis for the studies' design.
-We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
-To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
-The researchers can include a temporary "faked" results section based on the simulated data, employing dynamic document generation.
-This allows for an initial check whether the analysis actually runs without failure and helps to see whether the planned reporting (e.g., the selected statistics, figures, and tables) provides sufficient detail.
+With simulated data researchers can check if the analysis is technically complete and can also recover the true parameters of the simulation.
+When an analysis fails to recover the correct values in idealized conditions with big samples researchers should not expect this analysis to work under realistic conditions.
+Because the simulation captures the researchers expectations about the real world conditions it is the best way to assess the effectiveness of the analysis without actually gathering the data.
 
-To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
-The introduction and theoretical background provide the basis for the studies' design and can be written before gathering data.
-We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
-To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
-The researchers can include a temporary "faked" results section based on the simulated data, employing dynamic document generation.
-This allows for an initial check whether the analysis actually runs without failure and helps to see whether the planned reporting (e.g., the selected statistics, figures, and tables) provides sufficient detail.
+### Efficiency
 
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-First, you have to think about and address data analytic problems on a conceptual level before you make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
-Second, with simulated data you can perform a simple technical check of whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
-Third, repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to ensure efficiency.
-For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
+A single simulated or synthesized dataset already helps to ensure that the plan is viable under certain cirumstances.
+However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
+Monte Carlo simulation lets researcher navigate several trade offs regarding their resources.
+In the simplest case they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
 @bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.

### 2021-07-18 manuscript.Rmd
remove synthethic data
add a conceptual overview for PAC

@@
 Writing incomplete code is technically impossible or at least researchers would notice if they e.g. forgot to aggregate the data etc.
-For this technical check researcher do not necessarily have to simulated data, they could also rely on synthetic data.
-Synthetic data means that they apply measures to obfuscate true relations in the data, e.g. shuffle the main dependent variable etc.
 
@@
 
-A single simulated or synthesized dataset already helps to ensure that the plan is viable under certain cirumstances.
+A single simulated or synthesized dataset already helps to ensure that the plan is viable under certain circumstances.
 However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
@@
 non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
+
+## What you need to know before writing a PAC
+
+To write a PAC researchers should start by drafting a manuscript in form of an R Markdown.
+When they have completed the method section, they should begin writing code.
+We suggest writing three functions for each hypothesis, one for the planned analysis, one for the simulating the data and one for reporting the results.
+At the heart of a PAC lies the planned analysis so researchers should first draft a function called `planned_analysis()` that captures the description in the method section.
+This function should receive the data and derive relevant results from it.
+
+```{r}
+planned_analyis <- function(data){
+  # 1. preprocess e.g. with `rowMeans(data)`
+  # 2. conduct analysis e.g. with `t.test()`
+  # 3. `return(results)`
+}
+```
+
+Because it is difficult to draft such function without interactively test it, researcher should than proceed to write a function for simulating data.
+
+```{r}
+simulate_data <- function(n, effect_size){
+  # 1. warn users that the results are "fake"
+  # 2. draw `n` samples with `effect_size`
+  # 3. format and return in expected data format
+}
+```
+
+As soon as they have written `planned_anaylsis()` and `simulate_data()` they can iteratively improve both functions, e.g.
+till `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
+The goal is that the output of `simulate_data()` works as input to the function `planned_analysis()`.
+
+This process can be tedious and complicated.
+If it proofs infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
+Of course researcher must not analyze the data before they write the preregistration but they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
+
+When the researchers are satisfied with the function `planned_analysis()` they can think about how they would like to report the results of the analysis, via tables, plots and text.
+The implementation of this reporting should be in the function `report_analysis()`.
+
+```{r}
+report_results <- function(results){
+  # 1. create markdown tables from results
+  # 2. conditionally interpret results e.g. if(p < .025)"Result is significant."
+  # (optional) visualize results
+  # 3. return results section formatted in markdown
+}
+```
+
+These function again should be accept the output of `planned_analysis()` as input.
+Ideally, these three functions can be composed to create a "fake" results section, e.g.
+when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section formatted in Markdown.
 

### 2021-07-19 manuscript.Rmd
Comments to introduction

--- 2021-07-18 manuscript.Rmd
@@
 
-Increasingly, scientists make their data, materials, and analysis code openly available.
-Sharing all these types of digital research output increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
-We now have the technology to make all artefacts related to a research project available so that a research project can be reproduced with minimal effort.
+Scientists increasingly strive to make research data, materials, and analysis code openly available.
+Sharing these types of digital research artifacts has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
+To some extent these potential benefits are contingent on whether these digital research artifacts are reproducible. <!--CJ Define reproducibility here-->
+Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
 Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
 22][^popper]
-Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
-This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
+This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.<!-- CJ Are we introducing such standards?
+Otherwise, that comment belongs in the Discussion, as a recommendation for future work.
+Also, if there are no standards - then how do we know repro meets the required goals?-->
+<!-- CJ This is superfluous; the technology has existed for a while: We now have the technology to make all artefacts related to a research project available so that a research project can be reproduced with minimal effort.-->
 
@@
 We pursue two aims.
-First, the present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community that eliminate four common threats to reproducibility [see @Peikert2019 for details].
-Specifically, we demonstrate how the R package `repro` supports researchers in implementing these solutions.
-The second objective is to discuss opportunities of such workflows to improve not only how research outputs are created and recreated but also how research itself is conducted.
-One such opportunity is an alternative preregistration [@NosekRevolution2018] scheme that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results before they gather empirical data.
-This dynamic document closely resembles an actual journal article but with "fake" results from simulated data that serve to generate placeholders for figures, tables, and statistics. 
-This document serves as unambiguous preregistration of the planned analysis and how the results are going to be reported.
-The only parts that are missing from this first draft are the discussion of results, which the authors add after they gathered empirical data.
+First, the present paper introduces the reader to a workflow that maximizes the potential reproducibility of a research project.<!--CJ: Start new sentence.
+Ensures reproducibility...by eliminating threats to reproducibility is an awkward sentence construction-->
+This workflow builds on several technical solutions originating in the software engineering community. 
+We discuss how the workflow addresses four common threats to reproducibility [see @Peikert2019 for details].
+We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
+The second objective is to discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
+One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
+PAC involves preregistering the intended analysis code and the bulk of the manuscript as dynamic document, including the Introduction, Methods, and Results sections<!--CJ Preregistration may occur at different stages, including after data collection:, before they gather empirical data.-->
+The resulting dynamic document closely resembles the final manuscript, but uses preliminary ("mock") Results, based on the analysis of simulated data, as a placeholder for figures, tables, and statistics.
+<!-- CJ This is a repetition: PAC serves as unambiguous preregistration of the planned analysis and how the results are going to be reported.-->
+After the empirical data are analyzed, the Results are updated, and the Discussion is written.
 
@@
 
-While the willingness to share materials is required for reproducing a project, it is not sufficient to achieve it.
-For example, @hardwicke2018 attempted to reproduce results from open materials in the journal Cognition. 
-Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
+Scientific organizations and funding bodies increasingly demand transparent sharing of research artifacts, and researchers are increasingly willing to do so.
+However, although the sharing of such artifacts is a necessary condition for reproducibility, it is not a sufficient one.
+This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
+Out of 35 published articles with open research artifacts, results of 22 articles could be reproduced but in 11 of these cases, further assistance from the original authors was required.
 For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
-@obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
+Another study of 62 Registered Reports found that only 41 had data available, and 37 had analysis scripts available [@obels2020].
 The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
 These failed attempts to reproduce highlight the need for widely accepted reproducibility standards, because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
+If research artifacts are available but not reproducible, their added value is limited.
 
-We argue that the benefits of open science increase if research is not only transparent but reproducible.
-To this end, this tutorial demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
-A rigorous standard increases the likelihood of any individual project to be actually reproducible and remain reproducible as long as possible.
-An easy-to-use standard, on the other hand, is less error-prone, produces less overhead and is thus more likely to be adopted.
-To us, it makes sense to strive for a broad adoption of easy-to-use practices and then gradually move into more complex workflows as researchers become more proficient with the tools involved.
-The remainder of this paper is structured along these lines.
-The first part introduces theoretical concepts and software solutions that together form a rigorous reproducibility standard.
-The second part is a tutorial that shows how these software solutions can be easily employed by researchers with the help of the R package `repro`.
+This tutorial therefore demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
+A rigorous standard increases the likelihood that a project will remain reproducible as long as possible.
+An easy-to-use standard, on the other hand, is more likely to be adopted.
+Our approach is to promote broad adoption of easy-to-use practices, i.e., a "low threshold", and ensure that these practices are compatible with more complex rigorous solutions, i.e., a "high ceiling".
+As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
+The first part of this tutorial outlines a reproducibility standard, and introduces the underlying theoretical concepts and software solutions.
+The second part is a tutorial that demonstrates how this standard can be easily adopted with the help of the R package `repro`.
 

### 2021-07-19 manuscript.Rmd
Comments to tools section

--- 2021-07-18 manuscript.Rmd
@@
 
-# Concepts and Software Solutions
+# Concepts and Software Solutions<!--CJ Some inconsistencies:
+1) From the end of the previous paragraph, I would expect a description of the "rigorous reproducibility standard" here.
+2) The paragraph is labeled "Concepts and Software SOlutions, but it begins with a list of threats to reproducibility. Threats are not concepts nor software solutions.
+3) After the list of threats, we get a description of the "rigorous reproducibility standard".
+That is what I would expect based on the preceding paragraph.
+Note that the "rigorous reproducibility standard" is also not a concept, nor a software solution.
+Consider re-naming this section.-->
 
-From our own experience with various research projects, we identified the following common threats to reproducibility: <!-- CJ How did you identify these causes? Is this exhaustive?-->
+From our own experience with various research projects, we identified the following common threats to reproducibility: <!-- CJ This still requires attention!
+How did you identify these causes?
+Is this exhaustive?-->
 
@@
 3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
-4.	**External code dependencies**; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
+4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
 <!-- CJ 5.
@@
 1. Version control
-2. Dynamic document creation
+2. Dynamic document generation
 3. Dependency tracking
@@
 
-First, to resolve ambiguity across multiple versions of code and data that come into existence during the development phase of a research project, we recommend using a **version control** system.
+**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the development phase of a research project.
 Version control allows a clear link between which results were generated by which version of code and data.
-A version control system tracks changes to all project-related files (e.g., materials, data, and code) over time.
-At a later stage, single files or the entire project can be compared to or reverted to any earlier version.
+We recommend using **Git** for version control, because of its widespread adoption in the R community.
+Git tracks changes to all project-related files (e.g., materials, data, and code) over time.
+At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
 A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
-
-We recommend using **Git** for version control, because of its widespread adoption in the R community.
-Git is built around snapshots that represent the entire project state at a given point in time.
+Git is built around snapshots that represent the project state at a given point in time.
 Those snapshots are called "commits" and work like a "save" action.
-Changes to files are collected in a _commit_, and each commit has a message that succinctly describes the changes made by the author.
+Changes to files are collected in a _commit_.
+Ideally, each commit has a message that succinctly describes these changes.
+It is good practice to make commits for concrete milestones, e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3".
+This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
 Each commit "knows" its ancestor, and they form a timeline together.
 The entirety of commits (i.e. the version-controlled project) is called a repository.
-In Git, specific snapshots of a repository can be tagged, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
+In Git, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
 Git has additional features beyond basic version control, such as branches to facilitate collaboration.
-@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git.
+@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git for research.
 @bryanExcuseMeYou2018 provides additional information on how to track R Markdown documents.
-To collaborate via Git,<!--dangling modify, cant help it--> the repository has to be uploaded somewhere.
+Collaborating via Git requires the repository to be uploaded somewhere.
 We recommend GitHub to host Git repositories because of its popularity among R users.
@@
 
-One would hope that sharing a literate statistical analysis would allow others (including one own's future self) to simply download and run (ie., reproduce) a given analysis.
-However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in most instances [@Peikert2019].
+One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
+However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in most instances [@Peikert2019].<!--CJ This is a very strong claim.
+Consider rephrasing-->
 In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambigously documented.
@@
 We recommend using **Make** for dependency tracking because it is language independent.
-To make the utility of Make more concrete, consider the following hypothetical example, in which a research project contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
-A Makefile could then look like this:
+The following hypothesical example illustrates the utility of Make.
+Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
+A Makefile for this project could look like this:
 
@@
 If a requirement does not exist, Make uses a recipe to build the requirement first before building the target.
-Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would make sure that the file `simulation_results.csv` exists; if not, it would find that this is also a target and issue the command in l.5 to run the simulation first before rendering the manuscript.
-This way, we ensure that the simulation is always run before the manuscript is built and that the manuscript is rebuilt if the simulation code was changed.
+Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
+This ensures that the simulation is always run before the manuscript is built, and that the manuscript is rebuilt if the simulation code was changed.
 Make therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
@@
 One comprehensive approach to software management is containerization.
-"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p.
+"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p.
 174].
@@
 
-Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
+**Docker** is an example of containerization. It provides all software used in an analysis, including the operating system itself.
 The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
@@
 Two major strategies help to keep the storage requirements reasonable.
-First, there is a community that maintains pre-made images for particular purposes.
+One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
 For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
@@
 `r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
-Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
-Second, instead of saving a binary image of the software environment, one can also save a container recipe, a so-called  `Dockerfile`.
-This recipe is a textual description of all commands that need to be executed to recreate the image.
-Such files are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb) and sufficiently describes how to recreate a given container image.
-However, Dockerfiles rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
-For proper archival, we therefore recommend to keep a full image in addition to the Dockerfile.
+Our image <!--CJ What is our image?
+I don't think you introduced this-->includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse <!--CJ Yuck-->[@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
+A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
+Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
+However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
+For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the Dockerfile.
 A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
 
-To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
-While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
-To remove this hurdle, we have developed a new R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
+To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).<!-- CJ: Please briefly mention the similarity / differences from worcs.
+Perhaps here would be a good place.
+"This approach is similar to ..., and enhances it by ..."-->
+While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
+To remove this hurdle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
 In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.
@@
 Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
+For an a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
 Preregistrations, however, have a problem.
 Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
-The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
+The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates, like the ones supplied by [The Open Science Framework](https://osf.io/).
 The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
@@
 We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
-PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration despite the best efforts afford.
+PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration afford.
 
@@
 We want to extend PAC from preregistering code to preregistering *the whole manuscript as a dynamic document*.
-The idea is as follows:
+The idea is as follows.
 For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
@@
 
-Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
+Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in how to write a scientific manuscript.
 It follows that for preregistrations they should stick to widely accepted standards of scientific writing [e.g. @apa7].
@@
 
-Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
-One particular advantage is the strong foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
-Nevertheless it can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section.
-
-Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-Simulating data forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
-It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
-Writing incomplete code is technically impossible or at least researchers would notice if they e.g. forgot to aggregate the data etc.
-For this technical check researcher do not necessarily have to simulated data, they could also rely on synthetic data.
-Synthetic data means that they apply measures to obfuscate true relations in the data, e.g. shuffle the main dependent variable etc.
+### Efficiency
 
@@
 
-With simulated data researchers can check if the analysis is technically complete and can also recover the true parameters of the simulation.
-When an analysis fails to recover the correct values in idealized conditions with big samples researchers should not expect this analysis to work under realistic conditions.
-Because the simulation captures the researchers expectations about the real world conditions it is the best way to assess the effectiveness of the analysis without actually gathering the data.
+The introduction and theoretical background provide the basis for the studies' design.
+We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
+To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
+The researchers can include a temporary "faked" results section based on the simulated data, employing dynamic document generation.
+This allows for an initial check whether the analysis actually runs without failure and helps to see whether the planned reporting (e.g., the selected statistics, figures, and tables) provides sufficient detail.
 
-### Efficiency
+To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
+The introduction and theoretical background provide the basis for the studies' design and can be written before gathering data.
+We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
+To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
+The researchers can include a temporary "faked" results section based on the simulated data, employing dynamic document generation.
+This allows for an initial check whether the analysis actually runs without failure and helps to see whether the planned reporting (e.g., the selected statistics, figures, and tables) provides sufficient detail.
 
-A single simulated or synthesized dataset already helps to ensure that the plan is viable under certain cirumstances.
-However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
-Monte Carlo simulation lets researcher navigate several trade offs regarding their resources.
-In the simplest case they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
+While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
+First, you have to think about and address data analytic problems on a conceptual level before you make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
+Second, with simulated data you can perform a simple technical check of whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
+Third, repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to ensure efficiency.
+For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
 @bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.

### 2021-07-19 manuscript.Rmd
move sections around

--- 2021-07-18 manuscript.Rmd
@@
 These function again should be accept the output of `planned_analysis()` as input.
+The output of this function should be formatted in Markdown.
+A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
+Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
+The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
+
 Ideally, these three functions can be composed to create a "fake" results section, e.g.
-when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section formatted in Markdown.
+when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section.
+
+To summarize researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
+After they gathered the actual data, they replace the simulated data, run `planned_analysis()` and write the discussion.
 

### 2021-07-20 R/simulation_funs.R
planned_analysis handles missingness

@@ -10,9 +10,12 @@ simulate_data <- function(n, df, d, i){
   effect$group <- group
   return(effect)
 }
+
 #----planned_analysis----
 planned_analysis <- function(data, use_rank = "skew", skew_cutoff = 1){
-  y <- rowMeans(data["group" != names(data)])
+  y <- rowMeans(data["group" != names(data)], na.rm = TRUE)
+  data <- data[!is.na(y),]
+  y <- y[!is.na(y)]
   x <- as.factor(data$group)
   skew <- moments::skewness(y)
   # skewness cutoff
@@ -23,6 +26,7 @@ planned_analysis <- function(data, use_rank = "skew", skew_cutoff = 1){
   test <- t.test(y ~ x)
   list(test = test, skew = skew, use_rank = use_rank, n = length(y))
 }
+
 #----extract_funs----
 #t2d <- function(t, n1, n2)t * sqrt(((n1 + n2)/(n1 * n2)) * (n1 + n2)/(n1 + n2 -2))
 t2d <- function(test){

### 2021-07-20 R/simulation_funs.R
add report_analysis

@@ -27,6 +27,19 @@ planned_analysis <- function(data, use_rank = "skew", skew_cutoff = 1){
   list(test = test, skew = skew, use_rank = use_rank, n = length(y))
 }
 
+#----report_analysis----
+report_analysis <- function(analysis) {
+  params <- report::report_parameters(analysis$test)
+  table <- attributes(params)$table
+  model <- report::report_model(analysis$test, table = table)
+  if (analysis$use_rank)
+    model <-
+    stringr::str_replace(model,
+                fixed("Welch Two Sample t-test"),
+                "Mann--Whitney--Wilcoxon test")
+  stringr::str_c("The ", model, " suggests that the effect is ", params, sep = "")
+}
+
 #----extract_funs----
 #t2d <- function(t, n1, n2)t * sqrt(((n1 + n2)/(n1 * n2)) * (n1 + n2)/(n1 + n2 -2))
 t2d <- function(test){

### 2021-07-20 R/simulation_funs.R
actually print results by default

@@ -28,7 +28,7 @@ planned_analysis <- function(data, use_rank = "skew", skew_cutoff = 1){
 }
 
 #----report_analysis----
-report_analysis <- function(analysis) {
+report_analysis <- function(analysis, cat = TRUE) {
   params <- report::report_parameters(analysis$test)
   table <- attributes(params)$table
   model <- report::report_model(analysis$test, table = table)
@@ -37,7 +37,13 @@ report_analysis <- function(analysis) {
     stringr::str_replace(model,
                 fixed("Welch Two Sample t-test"),
                 "Mann--Whitney--Wilcoxon test")
-  stringr::str_c("The ", model, " suggests that the effect is ", params, sep = "")
+  out <- stringr::str_c("The ", model, " suggests that the effect is ", params, sep = "")
+  if(cat){
+    cat(out)
+    return(invisible(out))
+  } else {
+    return(out)
+  }
 }
 
 #----extract_funs----

### 2021-07-20 manuscript.Rmd
dont describe example, directly include it

@@
 source(here::here("R", "link.R"))
+source(here::here("R", "simulation_funs.R"))
 ```
@@
 Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
+For a priori planned analyses, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
 Preregistrations, however, have a problem.
@@
 The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
-The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
+The structure these templates provide can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
 But these templates can not remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
@@
 We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, a PAC removes any ambiguity regarding the translation of an verbal analysis plan into code.
+First, a PAC removes any ambiguity regarding the translation of a verbal analysis plan into code.
 Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
@@
 Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
-Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.
+Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
 
@@
 
-With simulated data researchers can check if the analysis is technically complete and can also recover the true parameters of the simulation.
-When an analysis fails to recover the correct values in idealized conditions with big samples researchers should not expect this analysis to work under realistic conditions.
-Because the simulation captures the researchers expectations about the real world conditions it is the best way to assess the effectiveness of the analysis without actually gathering the data.
+With simulated data researchers can check if the analysis recover the known parameters of the simulation.
+When an analysis fails to recover the correct values in idealized conditions with big samples researchers should not expect this analysis to be effective under realistic conditions.
+Because the simulation captures the researchers expectations about the real world it is the best way to assess the effectiveness of the analysis without actually gathering the data.
 
@@
 
-A single simulated or synthesized dataset already helps to ensure that the plan is viable under certain circumstances.
+A single simulated dataset already helps to ensure that the plan is viable under certain circumstances.
 However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
@@
 @bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.
-You may also evaluate how the data analysis copes with expected data complications, e.g.
-non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
+Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
 
@@
 We suggest writing three functions for each hypothesis, one for the planned analysis, one for the simulating the data and one for reporting the results.
-At the heart of a PAC lies the planned analysis so researchers should first draft a function called `planned_analysis()` that captures the description in the method section.
+At the heart of a PAC lies the planned analysis so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
 This function should receive the data and derive relevant results from it.
 
-```{r}
+```{r, eval=FALSE}
 planned_analyis <- function(data){
@@
 
-```{r}
+```{r, eval=FALSE}
 simulate_data <- function(n, effect_size){
@@
 
-As soon as they have written `planned_anaylsis()` and `simulate_data()` they can iteratively improve both functions, e.g.
+As soon as they have written a first draft of `planned_anaylsis()` and `simulate_data()` they can interactively improve both functions, e.g.
 till `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
@@
 
-```{r}
+```{r, eval=FALSE}
 report_results <- function(results){
@@
 
-These function again should be accept the output of `planned_analysis()` as input.
+This function should again accept the output of `planned_analysis()` as input.
 The output of this function should be formatted in Markdown.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
+A variety of packages automatically generate well formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
 Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
@@
 Ideally, these three functions can be composed to create a "fake" results section, e.g.
-when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section.
-
-To summarize researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
-After they gathered the actual data, they replace the simulated data, run `planned_analysis()` and write the discussion.
-
-# Preregistration as Code --- a Tutorial
-
-To guide the reader through a preregistration as code (PAC), we will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
+when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` they output a full results section.
+
+To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
+After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()`) and write the discussion.
+
+## An example of Preregistration as Code
+
+We have made the argument that PAC has great advantages over classical preregistration and have outlined how to implement PAC, but can PAC actually work?
+To answer this question and to help researchers implementing PAC themselves, we have conducted a PAC that can serve as an example.
+We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
 
@@
 
-This example serves only didactic purposes and not to derive substantive claims.
-Nevertheless, everything else is kept as close to a real research question as possible.
-
-At the heart of a PAC lies the planned analysis.
-We suggest writing a main analysis function that captures the full process of testing a research question.
-This function receives a dataset as input and performs all statistical processing. <!--AB say anything about its output? is the Markdown the output?-->
-For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
-As an example of a data-dependent decision, we included a rank transformation if the skewness was greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+Again, we propose a preregistration format that closely resembles a classical journal article but that utilizes simulated data and dynamic document generation to serve as preregistration and final manuscript.
+The full example is available in the (https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
+In this section we show excerpts of this example formatted in light grey and provide some commentary.
+
+```{r, echo=FALSE}
+show <- function(x, wrap = TRUE, cat = TRUE, fence = TRUE){
+  stopifnot(length(x) == 1L)
+  if(wrap)x <- str_c(str_wrap(str_split(x, "\\n\\n")[[1]]), collapse = "\n\n")
+  if(!isFALSE(fence)){
+    if(isTRUE(fence))fence <- "markdown"
+    fence <- str_c("````", fence)
+    x <-
+      str_c(fence,
+            x,
+            "````",
+            sep = "\n",
+            collapse = "\n")
+  }
+  if(cat){
+    cat(x)
+    return(invisible(x))
+  } else {
+    return(x)
+  }
+}
+show_prereg <-
+  function(section,
+           file = "preregistration.Rmd",
+           wrap = TRUE,
+           cat = TRUE,
+           fence = TRUE) {
+    prereg <- readr::read_lines(here::here(file))
+    start <- stringr::str_c("<!--begin-", section, "-->")
+    end <- stringr::str_c("<!--end-", section, "-->")
+    start_pos <- which(prereg == start)
+    end_pos <- which(prereg == end)
+    if (length(start_pos) != 1L || length(end_pos) != 1L)
+      stop("No section '", section, "' found.")
+    excerpt <-
+      str_c(prereg[(start_pos + 1):(end_pos - 1)], collapse = "\n")
+    show(excerpt, wrap = wrap, cat = cat, fence = fence)
+  }
+```
+
+As usual researchers state why they are interested in their research question in the section "Introduction" and provide the necessary information and literature to understand the rest of the paper.
+This example is drastically shortened for illustration purposes:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("theory")
+```
+
+After researchers have provided the research question they proceed to explain how they want to study it.
+For simplicity, we will use already published data that we have not yet analyzed:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("method")
+```
+
+We choose the following statistical procedure because many researchers may be familiar with it[^caution-mww]:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("ttest")
+```
 
 [^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
-The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
-
-This function is one possible translation of the verbal description above:
+The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
+
+We than translated this methods section into the following `planned_analysis()` function:
 
@@
 
-Importantly, natural language descriptions do not map unambiguously onto computer code implementations. 
+This function illustrates two advantage of PAC.
+First, a PAC can easily include data dependent decisions creating different branches of analysis under different conditions.
+Second, it highlights how difficult it is to precisely describe a statistical analysis verbally.
 The same textual descriptions may be implemented differently by different persons depending on their statistical and programming knowledge but also based on their own expectations and beliefs.
@@
 
-<!--AB here is kind of a break in the text-->
-To use this function one needs a compatible dataset. <!--AB what is "this" function? What does "compatible" mean?-->
-At the time of preregistration there is of course no actual data available.
-We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
-Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
-The following code simulates a dataset with one observed variable in two groups.
-
-```{r simulate_data}
-```
-<!--AB it is odd that the return variable is called effect because it is a dataset-->
-
-Here, we decided to randomly draw non-normal data to test how well our procedure performed under non-optimal situations.
-Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a given standardized effect size (a Cohen's d of .2).
-This simulated dataset fits perfectly into our planned analysis with a sample size of 400 and 10 items: <!--AB in what way? please explain-->
-
-***
-
-```{r}
-set.seed(112358) # fix state of RNG
-results <- 
-  planned_analysis(
-    simulate_data(
-      n = 400,# Sample size
-      df = 8, # Chisq with DF = 8 has skewness = 1
-      d = .2, # Cohen's D
-      i = 10  # Number of items (should not effect anything)
-    )
-  )
-```
-
-Next, we add dynamic code blocks to our Markdown report that implement the planned reporting of the final results.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
-Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
-Here we use the `report` package to create a preview of how our reporting will look like based on simulated data:<!--AB add citation to the report package-->
-
-```{r, results='asis', eval=FALSE}
-report::report(results$test)
-```
-
-This call to the `report()` function will yield the following automatically generated output:
-
-```{r, results='asis', echo=FALSE}
-report::report(results$test)
-```
-<!--AB it is difficult to discern the automatically generated text and out manuscript text even though there is the horizontal line. Not sure how to improve this.-->
-***
+The function `planned_analysis()` together with the function `simulate_data()` (not shown here) can be used to justify the planned sample size.
+To that end `simulate_data()` is repeatedly called with increased sample size and the proportion of significant results (power) is recorded.
+The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
+The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R).
+
+Monte Carlo simulation are of course not only applicable for this analysis method and and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
 
@@
 
-To determine a minimal sample size for this study we repeat this simulation across several combination of effect sizes and sample sizes.
-One standard recomendations for choosing a sample size is to have an expected power of 80% based on the expected effect size.
-@Jones2013 found in three studies gender differences of .24, .29, and .35 for their machiavilism subscale.
-If we want to detect an effect of size `r choosen_d` with an `r choosen_power*100`% probability, our simulation shows that we require at least `r minn$n` participants.
-
-<!--AB please add a figure number and a proper citation to the figure. Also, tell people where they can find the code to run the power analysis themselves-->
-
-```{r, echo=FALSE}
+```{r, include=FALSE}
+fig_caption <- str_c("Results of simulation for the power analysis. The cross indicates the sample size that archives ", choosen_power*100, "% assuming a Cohen's D of ", choosen_d, ".")
+```
+
+```{r power, echo=FALSE, fig.cap=fig_caption}
 res %>%
@@
 
+```{r, echo=FALSE, results='asis'}
+show_prereg("power-code", wrap = FALSE)
+```
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("power")
+```
+
+For the results section we implemented a mechanism that only uses simulated data when the actual data is not yet available.
+This mechanism also warns readers if these results are based on simulated data.
+As soon as the actual data are available, the simulated data is no longer used and the results represent the actual empirical results of the study.
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("results", wrap = FALSE)
+```
+
+The last thing we need to do is to report our results with the combination of `planned_analysis()` and `report_analysis()`.
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("report", wrap = FALSE)
+```
+
+This is an example of how the results could be reported (based on simulated data):
+
+```{r, echo=FALSE, results='asis'}
+set.seed(1234)
+show(report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)), cat = FALSE), fence = "")
+```
+
+The interested reader can find the example rendered with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
+
 # Discussion
@@
 
-PACs come with a challenge though.
+PACs come with a challenge.
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
@@
 Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
+Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
 
@@
 
-<!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->
-Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
+Please note that the statistical procedure in the example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for addressing your research question.
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
 The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-<!--AB can we move this to a different place?
-Maybe this would actually better fit to the discussion/limitation section!-->
+
+We also suggest that version control with Git and GitHub fulfills the criteria of a preregistration as long as you state your research plan comprehensively.
+This view is not shared by everyone [@szollosi_is_2020] because with some effort and malice intend the record of Git/GitHub can be altered.
+This concern can be addressed by using zenodo.org to give a digital object identifier (DOI) to your preregistration (see [this guide](https://guides.github.com/activities/citable-code/)) because zenondo.org saves an unchangeable copy of your repository.
 

### 2021-07-20 manuscript.Rmd
dont describe example, directly include it

@@
 
-Increasingly, scientists make their data, materials, and analysis code openly available.
-Sharing all these types of digital research output increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
-We now have the technology to make all artefacts related to a research project available so that a research project can be reproduced with minimal effort.
+Scientists increasingly strive to make research data, materials, and analysis code openly available.
+Sharing these types of digital research artifacts has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
+To some extent these potential benefits are contingent on whether these digital research artifacts are reproducible. 
+Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
 Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
 22][^popper]
-Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
 This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
+We now have the technology to make all artefacts related to a research project available so that a research project can be reproduced with minimal effort.
 
@@
 We pursue two aims.
-First, the present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community that eliminate four common threats to reproducibility [see @Peikert2019 for details].
-Specifically, we demonstrate how the R package `repro` supports researchers in implementing these solutions.
-The second objective is to discuss opportunities of such workflows to improve not only how research outputs are created and recreated but also how research itself is conducted.
-One such opportunity is an alternative preregistration [@NosekRevolution2018] scheme that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results before they gather empirical data.
-This dynamic document closely resembles an actual journal article but with "fake" results from simulated data that serve to generate placeholders for figures, tables, and statistics. 
-This document serves as unambiguous preregistration of the planned analysis and how the results are going to be reported.
-The only parts that are missing from this first draft are the discussion of results, which the authors add after they gathered empirical data.
-
-[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalised to inter-subjective criticism (Footnote *1, p.
+First, the present paper introduces the reader to a workflow that maximizes the potential reproducibility of a research project.
+This workflow builds on several technical solutions originating in the software engineering community. 
+We discuss how these solution address four common threats to reproducibility [see @Peikert2019 for details] and form a coherent workflow.
+We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
+The second objective is to discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
+One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
+PAC involves preregistering the intended analysis code and the bulk of the manuscript as dynamic document, including the Introduction, Methods, and Results sections.
+The resulting dynamic document closely resembles the final manuscript, but uses preliminary ("mock") results, based on the analysis of simulated data, as a placeholder for figures, tables, and statistics.
+After the empirical data are analyzed, the Results are updated, and the Discussion is written.
+
+[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalized to inter-subjective criticism (Footnote *1, p.
 22).
@@
 
-While the willingness to share materials is required for reproducing a project, it is not sufficient to achieve it.
-For example, @hardwicke2018 attempted to reproduce results from open materials in the journal Cognition. 
-Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
+Scientific organizations and funding bodies increasingly demand transparent sharing of research artifacts, and researchers are increasingly willing to do so.
+However, although the sharing of such artifacts is a necessary condition for reproducibility, it is not a sufficient one.
+This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
+Out of 35 published articles with open research artifacts, results of 22 articles could be reproduced but in 11 of these cases, further assistance from the original authors was required.
 For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
-@obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
+Another study of 62 Registered Reports found that only 41 had data available, and 37 had analysis scripts available [@obels2020].
 The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
 These failed attempts to reproduce highlight the need for widely accepted reproducibility standards, because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
-
-We argue that the benefits of open science increase if research is not only transparent but reproducible.
-To this end, this tutorial demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
-A rigorous standard increases the likelihood of any individual project to be actually reproducible and remain reproducible as long as possible.
-An easy-to-use standard, on the other hand, is less error-prone, produces less overhead and is thus more likely to be adopted.
-To us, it makes sense to strive for a broad adoption of easy-to-use practices and then gradually move into more complex workflows as researchers become more proficient with the tools involved.
-The remainder of this paper is structured along these lines.
-The first part introduces theoretical concepts and software solutions that together form a rigorous reproducibility standard.
-The second part is a tutorial that shows how these software solutions can be easily employed by researchers with the help of the R package `repro`.
+If research artifacts are available but not reproducible, their added value is limited.
+
+This tutorial therefore demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
+A rigorous standard increases the likelihood that a project will remain reproducible as long as possible.
+An easy-to-use standard, on the other hand, is more likely to be adopted.
+Our approach is to promote broad adoption of easy-to-use practices, i.e., a "low threshold", and ensure that these practices are compatible with more complex rigorous solutions, i.e., a "high ceiling".
+As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
+The first part of this tutorial outlines a reproducibility standard, and introduces the underlying theoretical concepts and software solutions.
+The second part is a tutorial that demonstrates how this standard can be easily adopted with the help of the R package `repro`.
 
@@
 
-# Concepts and Software Solutions
-
-From our own experience with various research projects, we identified the following common threats to reproducibility: <!-- CJ How did you identify these causes? Is this exhaustive?-->
+# Concepts and Software Solutions<!--CJ Some inconsistencies:
+1) From the end of the previous paragraph, I would expect a description of the "rigorous reproducibility standard" here.
+2) The paragraph is labeled "Concepts and Software SOlutions, but it begins with a list of threats to reproducibility. Threats are not concepts nor software solutions.
+3) After the list of threats, we get a description of the "rigorous reproducibility standard".
+That is what I would expect based on the preceding paragraph.
+Note that the "rigorous reproducibility standard" is also not a concept, nor a software solution.
+Consider re-naming this section.-->
+
+From our own experience with various research projects, we identified the following common threats to reproducibility: <!-- CJ This still requires attention!
+How did you identify these causes?
+Is this exhaustive?-->
 
@@
 
-First, to resolve ambiguity across multiple versions of code and data that come into existence during the development phase of a research project, we recommend using a **version control** system.
+**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the development phase of a research project.
 Version control allows a clear link between which results were generated by which version of code and data.
-A version control system tracks changes to all project-related files (e.g., materials, data, and code) over time.
-At a later stage, single files or the entire project can be compared to or reverted to any earlier version.
+We recommend using **Git** for version control, because of its widespread adoption in the R community.
+
+Git tracks changes to all project-related files (e.g., materials, data, and code) over time.
+At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
 A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
-
-We recommend using **Git** for version control, because of its widespread adoption in the R community.
-Git is built around snapshots that represent the entire project state at a given point in time.
+Git is built around snapshots that represent the project state at a given point in time.
 Those snapshots are called "commits" and work like a "save" action.
-Changes to files are collected in a _commit_, and each commit has a message that succinctly describes the changes made by the author.
+Changes to files are collected in a _commit_.
+Ideally, each commit has a message that succinctly describes these changes.
+It is good practice to make commits for concrete milestones, e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3".
+This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
 Each commit "knows" its ancestor, and they form a timeline together.
 The entirety of commits (i.e. the version-controlled project) is called a repository.
-In Git, specific snapshots of a repository can be tagged, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
+In Git, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
 Git has additional features beyond basic version control, such as branches to facilitate collaboration.
-@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git.
+@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git for research.
 @bryanExcuseMeYou2018 provides additional information on how to track R Markdown documents.
-To collaborate via Git,<!--dangling modify, cant help it--> the repository has to be uploaded somewhere.
+Collaborating via Git requires the repository to be uploaded somewhere.
 We recommend GitHub to host Git repositories because of its popularity among R users.
@@
 
-One would hope that sharing a literate statistical analysis would allow others (including one own's future self) to simply download and run (ie., reproduce) a given analysis.
-However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in most instances [@Peikert2019].
+One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
+However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
 In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambigously documented.
@@
 One comprehensive approach to software management is containerization.
-"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p.
+"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p.
 174].
-<!--APA requires no line number, but pagenumber or section heading-->
-
-Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
+
+**Docker** is an example of containerization.
+It provides all software used in an analysis, including the operating system itself.
 The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
@@
 `r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
-Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
-Second, instead of saving a binary image of the software environment, one can also save a container recipe, a so-called  `Dockerfile`.
-This recipe is a textual description of all commands that need to be executed to recreate the image.
-Such files are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb) and sufficiently describes how to recreate a given container image.
-However, Dockerfiles rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
-For proper archival, we therefore recommend to keep a full image in addition to the Dockerfile.
+The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
+A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
+Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
+However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
+For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the Dockerfile.
 A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
@@
 To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
-While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
-To remove this hurdle, we have developed a new R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
+While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
+To remove this hurdle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
 In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.

### 2021-07-20 preregistration.Rmd
dont describe example, directly include it

@@
 
-Theoretical Background
-======================
+<!--begin-theory-->
+# Theoretical Background
 
@@
 We therefore want to add to this body of literature.
-
-Method
-======
+<!--end-theory-->
 
 ```{r, echo=FALSE}
-res <-
+simulation_results <-
   readr::read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddddd")
+```
+
+<!--begin-method-->
+# Method
+
+We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study [@simmons2012].
+We use data available from [openpsychometrics.org](https://openpsychometrics.org/_rawdata/) from the online version of the MACH-IV[@christie1970] and included participants that have responded to at least one machiavellianism item and reported their gender as either "male" or "female".
+<!--end-method-->
+
+<!--begin-ttest-->
+We conduct a Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items between the binary-coded gender groups.
+If the skew of this average is greater than 1.0 we conduct a supposedly more robust Mann--Whitney--Wilcoxon test [@Wilcoxon1945] instead.
+<!--end-ttest-->
+<!--begin-power-code-->
+```{r}
 choosen_power <- .8
 choosen_d <- .2
-minn <- filter(res, d == choosen_d, power > choosen_power) %>% 
-  filter(n == min(n))
+minn <-
+  filter(filter(simulation_results, d == choosen_d, power > choosen_power),
+         n == min(n))$n
 ```
+<!--end-power-code-->
+<!--begin-power-->
+A simulation we conducted indicated that with a sample size of
+`r minn` for an alpha of .05 (two-sided) we archive at least `r choosen_power*100`% power assuming an effect size of `r choosen_d`.
+<!--end-power-->
 
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study [@simmons2012].
-We use data available from [openpsychometrics.org](https://openpsychometrics.org/_rawdata/) from the online version of the MACH-IV [@christie1970]and included participants that have responded to at least one machiavellianism item and reported their gender as either "male" or "female".
-We conduct a t-test on the average of machiavellianism items, if the skew of this avarage is below 1.0, else we conduct a Mann--Whitney--Wilcoxon test between the gender groups.
-A simulation we conducted indicated that with a sample size of `r minn$n` for an alpha of .05 (two-sided) we archive at least `r choosen_power*100`% power assuming an effect size of `r choosen_d`.
-
-Results
-=======
+<!--begin-results-->
+# Results
 
@@
 if(simulated){
-  cat("\\textcolor{red}{This document and the results presented herein are based on simulated data and should not be interpreted.}")
+  cat("\\textcolor{red}{This document and the results presented herein are based
+      on simulated data and should not be interpreted.}")
   set.seed(1234)
@@
 ```
+<!--end-results-->
 
+<!--begin-report-->
 ```{r, echo=FALSE, results='asis'}
-analysis <- planned_analysis(mach)
-report <- function(analysis) {
-  params <- report::report_parameters(analysis$test)
-  table <- attributes(params)$table
-  model <- report::report_model(analysis$test, table = table)
-  if (analysis$use_rank)
-    model <-
-    str_replace(model,
-                fixed("Welch Two Sample t-test"),
-                "Mann--Whitney--Wilcoxon test")
-  cat("The ", model, " suggests that the effect is ", params, sep = "")
-}
+report_analysis(planned_analysis(mach))
 ```
+<!--end-report-->
 
-```{r, echo=FALSE, results='asis'}
-```
+# Discussion
 
-Discussion
-==========
+\textcolor{red}{This document was created before we have gathered any data, we therefore have nothing to discuss.}
 
-```{r, echo=FALSE, results='asis'}
-if(simulated){
-  cat("\\textcolor{red}{This document was created before we have gathered any data, we therefore have nothing to discuss.}")
-}
-```
-
-References
-==========
+\newpage
+# References

### 2021-07-20 manuscript.Rmd

--- 2021-07-19 manuscript.Rmd
@@
 We pursue two aims.
-First, the present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community that eliminate four common threats to reproducibility [see @Peikert2019 for details].
-Specifically, we demonstrate how the R package `repro` supports researchers in implementing these solutions.
-The second objective is to discuss opportunities of such workflows to improve not only how research outputs are created and recreated but also how research itself is conducted.
-One such opportunity is an alternative preregistration [@NosekRevolution2018] scheme that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results before they gather empirical data.
-This dynamic document closely resembles an actual journal article but with "fake" results from simulated data that serve to generate placeholders for figures, tables, and statistics. 
-This document serves as unambiguous preregistration of the planned analysis and how the results are going to be reported.
-The only parts that are missing from this first draft are the discussion of results, which the authors add after they gathered empirical data.
+First, the present paper introduces the reader to a workflow that maximizes the potential reproducibility of a research project.
+This workflow builds on several technical solutions originating in the software engineering community. 
+We discuss how these solution address four common threats to reproducibility [see @Peikert2019 for details] and form a coherent workflow.
+We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
+The second objective is to discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
+One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
+PAC involves preregistering the intended analysis code and the bulk of the manuscript as dynamic document, including the Introduction, Methods, and Results sections.
+The resulting dynamic document closely resembles the final manuscript, but uses preliminary ("mock") results, based on the analysis of simulated data, as a placeholder for figures, tables, and statistics.
+After the empirical data are analyzed, the Results are updated, and the Discussion is written.
 
-[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalised to inter-subjective criticism (Footnote *1, p.
+[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalized to inter-subjective criticism (Footnote *1, p.
 22).

### 2021-07-20 manuscript.Rmd
address some comments from #165

@@
 We recommend using **Git** for version control, because of its widespread adoption in the R community.
+
 Git tracks changes to all project-related files (e.g., materials, data, and code) over time.
@@
 174].
-<!--APA requires no line number, but pagenumber or section heading-->
 
-**Docker** is an example of containerization. It provides all software used in an analysis, including the operating system itself.
+**Docker** is an example of containerization.
+It provides all software used in an analysis, including the operating system itself.
 The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
@@
 `r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
-Our image <!--CJ What is our image?
-I don't think you introduced this-->includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse <!--CJ Yuck-->[@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
+The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse <!--CJ Yuck-->[@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
 A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.

### 2021-07-20 manuscript.Rmd
soften phrasing a tiny bit #168

@@
 One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
-However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in most instances [@Peikert2019].<!--CJ This is a very strong claim.
-Consider rephrasing-->
+However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
 In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambigously documented.

### 2021-07-21 manuscript.Rmd
turn caspers comments into issues

@@
 `r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
-The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse <!--CJ Yuck-->[@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
+The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
 A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
@@
 
-To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).<!-- CJ: Please briefly mention the similarity / differences from worcs.
-Perhaps here would be a good place.
-"This approach is similar to ..., and enhances it by ..."-->
+To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
 While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.

### 2021-07-21 manuscript.Rmd
grammerly

@@
 Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
+For a priori planned analyses, preregistration should reduce "researcher degrees of freedom" therefore, the number of decisions researcher can make that could unduly influence the results.
 Preregistrations, however, have a problem.
-Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
-The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
-The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
-But these templates can not remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
+Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
+The current best practice to ensure that the preregistration is comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
+The structure these templates provide can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
+Nevertheless, these templates can not remove researchers' degrees of freedom because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
 
@@
 Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
-We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
-PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration despite the best efforts afford.
+We term this writing and publishing code before data collection *preregistration as code (PAC)*.
+PAC has the potential to all but eliminates researchers degrees of freedom that standard preregistration despite the best efforts afford.
 
@@
 A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
-In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
+In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`), we found close to a dozen published articles that included some form of script as part of their preregistration.
 
@@
 First, preregistrations are plagued by non-reproducibility like any other research outcome.
-@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to reproduce the power analysis.
-Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
+@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis, but from these, only 34 (20%) provided enough information to reproduce the power analysis.
+Even researchers who have gone to great lengths in submitting a script in their preregistration sometimes inexplicably fail to reproduce their results.
 For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
@@
 
-Leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
-To use PAC for study planing, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
+When leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
+To use PAC for study planning, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
 To that end, they should borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
@@
 Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
-One particular advantage is the strong foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
-Nevertheless it can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section.
+One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
+Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
 
 Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-Simulating data forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
+While relatively uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
+When simulating data, researchers have to think about and address data-analytic problems on a conceptual level before making an effort to acquire data [@axelrodAdvancingArtSimulation1997].
 It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
-Writing incomplete code is technically impossible or at least researchers would notice if they e.g. forgot to aggregate the data etc.
+Writing incomplete code is technically impossible, or at least researchers would notice if they, e.g. forgot to aggregate the data, and so forth.
 
@@
 
-With simulated data researchers can check if the analysis is technically complete and can also recover the true parameters of the simulation.
-When an analysis fails to recover the correct values in idealized conditions with big samples researchers should not expect this analysis to work under realistic conditions.
-Because the simulation captures the researchers expectations about the real world conditions it is the best way to assess the effectiveness of the analysis without actually gathering the data.
+With simulated data, researchers can check if the analysis recovers the known parameters of the simulation.
+When an analysis fails to recover the correct values in idealized conditions with big samples, researchers should not expect this analysis to be effective under realistic conditions.
+Because the simulation captures the researchers' expectations about the real world, it is the best way to assess the effectiveness of the analysis without actually gathering the data.
 
@@
 
-A single simulated or synthesized dataset already helps to ensure that the plan is viable under certain circumstances.
+A single simulated dataset already helps to ensure that the plan is viable under certain circumstances.
 However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
-Monte Carlo simulation lets researcher navigate several trade offs regarding their resources.
-In the simplest case they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
+Monte Carlo simulation lets researchers navigate several trade-offs regarding their resources.
+In the simplest case, they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
 @bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.
-You may also evaluate how the data analysis copes with expected data complications, e.g.
-non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
+Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
 
@@
 
-To write a PAC researchers should start by drafting a manuscript in form of an R Markdown.
+To write a PAC, researchers should start by drafting a manuscript in the form of an R Markdown.
 When they have completed the method section, they should begin writing code.
-We suggest writing three functions for each hypothesis, one for the planned analysis, one for the simulating the data and one for reporting the results.
-At the heart of a PAC lies the planned analysis so researchers should first draft a function called `planned_analysis()` that captures the description in the method section.
+We suggest writing three functions for each hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
+At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
 This function should receive the data and derive relevant results from it.
 
-```{r}
+```{r, eval=FALSE}
 planned_analyis <- function(data){
@@
 
-Because it is difficult to draft such function without interactively test it, researcher should than proceed to write a function for simulating data.
-
-```{r}
+Because it is difficult to draft such a function without interactively testing it, the researcher should then write a function for simulating data.
+
+```{r, eval=FALSE}
 simulate_data <- function(n, effect_size){
-  # 1. warn users that the results are "fake"
+  # 1. warn users that the results are mocked
   # 2. draw `n` samples with `effect_size`
@@
 This process can be tedious and complicated.
-If it proofs infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
-Of course researcher must not analyze the data before they write the preregistration but they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
-
-When the researchers are satisfied with the function `planned_analysis()` they can think about how they would like to report the results of the analysis, via tables, plots and text.
+If it proves infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
+Of course, researchers must not analyze the data before they write the preregistration.
+However, they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
+
+When the researchers are satisfied with the function `planned_analysis()`, they can think about how they would like to report the analysis results via tables, plots, and text.
 The implementation of this reporting should be in the function `report_analysis()`.
 
-```{r}
+```{r, eval=FALSE}
 report_results <- function(results){
@@
 
-These function again should be accept the output of `planned_analysis()` as input.
+This function should again accept the output of `planned_analysis()` as input.
 The output of this function should be formatted in Markdown.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
+Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
 Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
-
-Ideally, these three functions can be composed to create a "fake" results section, e.g.
-when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section.
-
-To summarize researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
-After they gathered the actual data, they replace the simulated data, run `planned_analysis()` and write the discussion.
-
-# Preregistration as Code --- a Tutorial
-
-To guide the reader through a preregistration as code (PAC), we will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
-
-> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"
-
-This example serves only didactic purposes and not to derive substantive claims.
-Nevertheless, everything else is kept as close to a real research question as possible.
-
-At the heart of a PAC lies the planned analysis.
-We suggest writing a main analysis function that captures the full process of testing a research question.
-This function receives a dataset as input and performs all statistical processing. <!--AB say anything about its output? is the Markdown the output?-->
-For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
-As an example of a data-dependent decision, we included a rank transformation if the skewness was greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it generates not only tables but also a straightforward interpretation of the effects as actual prose (e.g.
+it quantifies the size of an effect).
+
+Ideally, these three functions can be composed to create a mock results section, e.g.
+when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` they output a full results section.
+
+To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
+After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()`) and write the discussion.
+
+## An example of Preregistration as Code
+
+We have made the argument that PAC has significant advantages over classical preregistration and have outlined how to implement PAC, but can PAC work in practice?
+To answer this question and help researchers implementing PAC themselves, we have conducted a PAC that can serve as an example.
+We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
+
+> "Is there a mean difference in the personality trait 'Machiavellism' between self-identified females and males?"
+
+Again, we propose a preregistration format that closely resembles a classical journal article, but that utilizes simulated data and dynamic document generation to serve simultaneously as preregistration and final manuscript.
+The complete example is available in the (https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
+In this section, we show excerpts of this example formatted in light grey and provide some commentary.
+
+```{r, echo=FALSE}
+show <- function(x, wrap = TRUE, cat = TRUE, fence = TRUE){
+  stopifnot(length(x) == 1L)
+  if(wrap)x <- str_c(str_wrap(str_split(x, "\\n\\n")[[1]]), collapse = "\n\n")
+  if(!isFALSE(fence)){
+    if(isTRUE(fence))fence <- "markdown"
+    fence <- str_c("````", fence)
+    x <-
+      str_c(fence,
+            x,
+            "````",
+            sep = "\n",
+            collapse = "\n")
+  }
+  if(cat){
+    cat(x)
+    return(invisible(x))
+  } else {
+    return(x)
+  }
+}
+show_prereg <-
+  function(section,
+           file = "preregistration.Rmd",
+           wrap = TRUE,
+           cat = TRUE,
+           fence = TRUE) {
+    prereg <- readr::read_lines(here::here(file))
+    start <- stringr::str_c("<!--begin-", section, "-->")
+    end <- stringr::str_c("<!--end-", section, "-->")
+    start_pos <- which(prereg == start)
+    end_pos <- which(prereg == end)
+    if (length(start_pos) != 1L || length(end_pos) != 1L)
+      stop("No section '", section, "' found.")
+    excerpt <-
+      str_c(prereg[(start_pos + 1):(end_pos - 1)], collapse = "\n")
+    show(excerpt, wrap = wrap, cat = cat, fence = fence)
+  }
+```
+
+As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary information and literature to understand the rest of the paper.
+This example is drastically shortened for illustration purposes:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("theory")
+```
+
+After researchers have provided the research question, they proceed to explain how they want to study it.
+For simplicity, we will use already published data that we have not yet analyzed:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("method")
+```
+
+We choose the following statistical procedure because many researchers may be familiar with it[^caution-mww]:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("ttest")
+```
 
 [^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
-The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
-
-This function is one possible translation of the verbal description above:
+The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
+
+We then translated this methods section into the following `planned_analysis()` function:
 
@@
 
-Importantly, natural language descriptions do not map unambiguously onto computer code implementations. 
-The same textual descriptions may be implemented differently by different persons depending on their statistical and programming knowledge but also based on their own expectations and beliefs.
+This function illustrates two advantages of PAC.
+First, a PAC can easily include data-dependent decisions creating different branches of analysis under different conditions.
+Second, it highlights how difficult it is to describe a statistical analysis precisely.
+The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and their expectations and beliefs.
 One example would be using the function `wilcox.test` instead of the combinations of `rank` and `t.test`.
-Both are valid implementation of the Mann–Whitney–Wilcoxon test but the first assumes equal variance whereas the second applies Welch's correction and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
+Both are valid implementations of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
+In contrast, the second applies Welch's correction and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
 Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
-Still, these details can make a difference and provide unwanted researcher degrees-of-freedom.
-
-<!--AB here is kind of a break in the text-->
-To use this function one needs a compatible dataset. <!--AB what is "this" function? What does "compatible" mean?-->
-At the time of preregistration there is of course no actual data available.
-We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
-Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
-The following code simulates a dataset with one observed variable in two groups.
-
-```{r simulate_data}
-```
-<!--AB it is odd that the return variable is called effect because it is a dataset-->
-
-Here, we decided to randomly draw non-normal data to test how well our procedure performed under non-optimal situations.
-Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a given standardized effect size (a Cohen's d of .2).
-This simulated dataset fits perfectly into our planned analysis with a sample size of 400 and 10 items: <!--AB in what way? please explain-->
-
-***
-
-```{r}
-set.seed(112358) # fix state of RNG
-results <- 
-  planned_analysis(
-    simulate_data(
-      n = 400,# Sample size
-      df = 8, # Chisq with DF = 8 has skewness = 1
-      d = .2, # Cohen's D
-      i = 10  # Number of items (should not effect anything)
-    )
-  )
-```
-
-Next, we add dynamic code blocks to our Markdown report that implement the planned reporting of the final results.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
-Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
-Here we use the `report` package to create a preview of how our reporting will look like based on simulated data:<!--AB add citation to the report package-->
-
-```{r, results='asis', eval=FALSE}
-report::report(results$test)
-```
-
-This call to the `report()` function will yield the following automatically generated output:
-
-```{r, results='asis', echo=FALSE}
-report::report(results$test)
-```
-<!--AB it is difficult to discern the automatically generated text and out manuscript text even though there is the horizontal line. Not sure how to improve this.-->
-***
+Still, these details can make a difference and provide unwanted researchers' degrees of freedom.
+
+The function `planned_analysis()` together with the function `simulate_data()` (not shown here) can be used to justify the planned sample size.
+To that end, `simulate_data()` is repeatedly called with increased sample size and the proportion of significant results (power) is recorded.
+The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
+The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R).
+
+Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
 
@@
 
+```{r, echo=FALSE, results='asis'}
+show_prereg("power-code", wrap = FALSE)
+```
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("power")
+```
+
+We implemented a mechanism that only uses simulated data when the actual data is not yet available for the results section.
+This mechanism also warns readers if these results are based on simulated data.
+As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study.
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("results", wrap = FALSE)
+```
+
+The last thing we need to do is to report our results with the combination of `planned_analysis()` and `report_analysis()`.
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("report", wrap = FALSE)
+```
+
+This is an example of how the results could be reported (based on simulated data):
+
+```{r, echo=FALSE, results='asis'}
+set.seed(1234)
+show(report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)), cat = FALSE), fence = "")
+```
+
+The interested reader can find the example rendered with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
+
 # Discussion

### 2021-07-21 preregistration.Rmd
grammerly

@@
 
-There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self identified males and females.
+There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self-identified males and females.
 @muris2017 note that studies focusing on gender differences in machiavellianism are sparse.
@@
 <!--begin-ttest-->
-We conduct a Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items between the binary-coded gender groups.
+We conduct a Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] of the average of machiavellism items between the binary-coded gender groups.
 If the skew of this average is greater than 1.0 we conduct a supposedly more robust Mann--Whitney--Wilcoxon test [@Wilcoxon1945] instead.
@@
 
-\textcolor{red}{This document was created before we have gathered any data, we therefore have nothing to discuss.}
+\textcolor{red}{This document was created before we have gathered any data, we, therefore, have nothing to discuss.}
 

### 2021-07-22 manuscript.Rmd
PAC example

@@
 
-Please note that the statistical procedure in the example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for addressing your research question.
+Please note that the statistical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for addressing your research question.
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].

### 2021-07-22 manuscript.Rmd
add first draft of discussion

--- 2021-07-21 manuscript.Rmd
@@
 
-<!--AB: TODO add a quick summary -->
+We have presented a workflow that rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to ensure reproducibility.
+This workflow is straightforward to implement with the `repro`-package, which provides a low threshold for adoption without limiting the users' ability to implement complex projects and is, therefore, a "high ceiling" solution.
+The `repro`-package builds upon and extends established best practices, i.e. R Markdown and Git, in two critical ways.
+First, it provides the required software for reproduction by automatically generating Dockerfiles.
+Second, by generating Makefiles the `repro`-package establishes a unified mechanism for reproducing research projects.
+
+We than proceeded to show one example of how such rigorous and automated reproducibility workflow may enable other innovations, e.g. preregistration as code.
+Complementing a preregistration with code is not unheard of but writing a manuscript with code as an integral dynamic element with mock results based on simulated data offers several advantages over classical preregistration.
+PAC lets consumers of a research document (e.g. editors, reviewers, readers) clearly distinguish how analyses were planned and how they were conducted via version control.
+On the researchers' side, PAC assures that the analysis can effectively answer the research question, i.e.
+that the analysis can recover the parameters of interest in simulation and lets them employ their resources efficiently by supporting study design decisions.
 
 # Limitations
+
+`repro`'s main goal is to increase the user friendliness of reproducibility tools for researchers.
+While we have invested much effort to ensure this goal, the package is still in early development and undoubtedly has rough edges and even bugs that went undetected.
+We hope that despite the early development stage `repro` can support researchers to work more reproducible.
 
@@
 
-<!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->
 Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
@@
 The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-<!--AB can we move this to a different place?
-Maybe this would actually better fit to the discussion/limitation section!-->
 

### 2021-07-22 manuscript.Rmd
First comments on tutorial

@@
 
-The above mentioned tools require much expertise that is beyond the traditional curriculum of researchers.
-To aid researcher adopting these tools we rely on or have developed helpers that wrap more complicated (and powerful) tools.
-These wrappers use standard tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
-We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
-These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
-
-Before we can start creating reproducible research projects, we need a computer with a working R Studio installation. 
-Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so that it can act as an assistance systems while you follow the tutorial:
+One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be time-intensive.
+To lower the threshold to adoption, the R-package `repro` introduces helper functions that wrap more complicated (and powerful) tools.
+These so-called wrappers guide end-users in the use of reproducibility <!--CJ Can't say "standard tools" after just saying that these tools are not yet widely adopted -->tools, provide feedback about what the computer does, and suggest what the user should do in layman's terms.
+We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
+These alternative tools<!-- CJ: Are they wrappers, or alternative tools--> are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
+
+This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
+It describes basic steps that we expect to apply to all users.
+Your specific situation might involve additional, more specialized steps.
+After completing the tutorial, you should be able to customize your workflow accordingly.<!--CJ: Cut/pasted from below; the purpose of the tutorial should be explained beforehand, not in the middle-->
+The first step is to install the required software.
+Start R Studio<!--CJ I am concerned that new users may not install R if you only mention RStudio.--> and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
+It will assist you while you follow the tutorial.
 
@@
 ```
+<!--
+This command fails for me:
+> install.packages('repro', repos = 'https://aaronpeikert.r-universe.dev')
+Installing package into ‘C:/Git_Repositories/veni_forest/renv/library/R-4.1/x86_64-w64-mingw32’
+(as ‘lib’ is unspecified)
+Warning: dependencies ‘usethis’, ‘fs’, ‘gh’ are not available
+trying URL 'https://aaronpeikert.r-universe.dev/src/contrib/repro_0.1.0.zip'
+Error in download.file(url, destfile, method, mode = "wb", ...) : 
+  cannot open URL 'https://aaronpeikert.r-universe.dev/src/contrib/repro_0.1.0.zip'
+In addition: Warning message:
+In download.file(url, destfile, method, mode = "wb", ...) :
+  cannot open URL 'https://aaronpeikert.r-universe.dev/src/contrib/repro_0.1.0.zip': HTTP status was '404 Not Found'
+Warning in download.packages(pkgs, destdir = tmpd, available = available,  :
+  download of package ‘repro’ failed
+
+But this one succeeds; it is what is suggested by R-universe:
+> options(repos = c(
++     aaronpeikert = 'https://aaronpeikert.r-universe.dev',
++     CRAN = 'https://cloud.r-project.org'))
+> 
+> # Install some packages
+> install.packages('repro')
+-->
 
 An example of the assistance that `repro` provides are the check functions.
-They verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
+These verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
 
@@
 ```
-
-They detect the users' system state and explain if further action is needed.
+<!-- CJ Be mindful of small spelling mistakes in the status messages, e.g., infos instead of info.
+Also: Not clear what different icons mean before the status messages.
+Also: Do you want to use informal language?
+E.g.: V Git is installed, don't worry!
+or: V Git is installed.-->
+These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
+<!-- CJ: This is so flipping important!!!
+We also need something like this for the summer school, to check if users have Mplus and if it can communicate with RStudio.
+I would emphasize this simple / valuable feature earlier, in the abstract and intro.-->
 Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:
@@
 
-The messages emitted by `repro` and other packages we present try to take the user by the hand to help them solve problems.
-This tutorial describes steps that we expect to apply to all users; the messages from the packages are adjusted to your specific situation (that is, operating system and the specific problem of missing software) and complement the tutorial in this way.
-Before you continue, we ask you to run the above commands to check Git, Make, and Docker.
-If necessary, follow the instructions of repro to install them until the check functions are all fine.
-
-The first concept we want to introduce is version control with Git.
+The messages emitted by `repro` and other packages we present <!-- CJ: This either needs more elaboration, or remove it entirely.
+It is currently confusing; what is the difference between "emitting" and "presenting"?
+I understand that you introduce some new functions, and import some existing ones.
+I would remove that distinction and just cite some of the packages you rely on.-->try to take the user by the hand to help them solve problems.
+These messages are adjusted to your specific operating system and installed dependencies.
+Before you continue, we ask you to run the above commands to check Git, Make, and Docker - both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
+If necessary, follow any instructions presented until all checks pass.
+A shortcut to test whether you are ready to continue is:
+
+```{r}
+check_tutorial()
+```
+
+## Implementing version control
+
+Now that your system is set up for this tutorial, we will introduce you to version control with Git.
 Git tracks files within one project folder.
@@
 
-You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore` file.
+Git tracks any files that you manually add to the repository.
+To make sure you do not accidentally add files that you do not wish to share (e.g., privacy sensitive data),
+you can list specific files that you do not want to track in the `.gitignore` file.
 You can add something to the `.gitignore` file directly or with this command:
@@
 ```
-
+<!--CJ: Also worcs::git_ignore()-->
 Now the file `private.md` will not be committed to Git and will not be made public.
 
-After you have excluded sensitive files (like not yet anonymized data files), you can activate Git with:
+Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
+When you are ready, you can begin tracking your files using Git by running:
 
@@
 For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
-The easiest way to do this is to use the RStudio Git pane.
-Click on the empty box next to the file you want to stage, and when a tick appears, which means that the file is staged.
+One way to do this is through the visual user interface in the RStudio Git pane.
+Click on the empty box next to the file you want to stage. A check mark appears to indicate that the file is staged.
 After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
+This stores a snapshot of the current state of the project.
 
@@
 All snapshots are stored in a local repository in your project folder.
-To back up and share the files, you can publicly upload those changes to a public repository.
-To synchronize your local repository with a remote repository from the GitHub service, issue the following command:
+To back up and/or share the files online, you can push your local repository to a remote repository.
+Specifically, to synchronize your local repository with a remote repository on GitHub, issue the following command:
 
@@
 
-If you are not yet ready to share your project publicly but you want to synchronize your local repository to a remote repository (for example, as a simple backup service), use:
+If you are not yet ready to share your project publicly<!--CJ: Note that you are conflating "online" with "public".
+People can push their repo to a private remote repository.
+Consider discussing this distinction first before showing the two commands.--> but you want to synchronize your local repository to a remote repository (for example, as a simple backup service), use:
 
@@
 
-Now that you have created a project that is under version control, we continue with adding dynamic document generation.
-Adynamic document has three elements:
-
-1. Text in a natural language
-2. Executable code 
-3. Metadata
-
-The text of an R Markdown is written in Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
+Once the connection with GitHub (or another remote repository, like GitLab or BitBucket) is set up, you can continue staging changes, committing them, and pushing your commits to the remote repository.
+One convenience function that does all this in a single step is:
+
+```{r}
+worcs::git_update("Your commit message")
+```
+
+This is essentially a 'quick save' function that stages all changes to files not listed in `.gitIgnore`, commits them, and pushes the commit.
+
+## Dynamic document generation
+
+Now that you have created a version controlled project, we will proceed with dynamic document generation.
+A dynamic document has three elements:
+
+1. Text in natural language (prose; e.g., a scientific paper or presentation)
+2. Executable code (e.g., analyses)
+3. Metadata (?)
+
+R Markdown is a type of dynamic document well-suited to the RStudio user interface.
+The natural language text in an R Markdown file can be formatted using Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
+The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
 The following example serves to illustrate the Markdown syntax. 
@@
 
-This type of markup for formatting, instead of the more visual approach of e.g. Microsoft Word, is compatible with many output formats and is easier to put under version control.
+Using markup for formatting, instead of the visual approach used in, e.g., Microsoft Word, is compatible with many output formats and is easier to put under version control.
 Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
@@
 
-We have presented a workflow that rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to ensure reproducibility.
-This workflow is straightforward to implement with the `repro`-package, which provides a low threshold for adoption without limiting the users' ability to implement complex projects and is, therefore, a "high ceiling" solution.
-The `repro`-package builds upon and extends established best practices, i.e. R Markdown and Git, in two critical ways.
-First, it provides the required software for reproduction by automatically generating Dockerfiles.
-Second, by generating Makefiles the `repro`-package establishes a unified mechanism for reproducing research projects.
-
-We than proceeded to show one example of how such rigorous and automated reproducibility workflow may enable other innovations, e.g. preregistration as code.
-Complementing a preregistration with code is not unheard of but writing a manuscript with code as an integral dynamic element with mock results based on simulated data offers several advantages over classical preregistration.
-PAC lets consumers of a research document (e.g. editors, reviewers, readers) clearly distinguish how analyses were planned and how they were conducted via version control.
-On the researchers' side, PAC assures that the analysis can effectively answer the research question, i.e.
-that the analysis can recover the parameters of interest in simulation and lets them employ their resources efficiently by supporting study design decisions.
+<!--AB: TODO add a quick summary -->
 
 # Limitations
-
-`repro`'s main goal is to increase the user friendliness of reproducibility tools for researchers.
-While we have invested much effort to ensure this goal, the package is still in early development and undoubtedly has rough edges and even bugs that went undetected.
-We hope that despite the early development stage `repro` can support researchers to work more reproducible.
 
@@
 
+<!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->
 Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
@@
 The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
+<!--AB can we move this to a different place?
+Maybe this would actually better fit to the discussion/limitation section!-->
 

### 2021-07-22 manuscript.Rmd
More comments to tutorial

--- 2021-07-22 manuscript.Rmd
@@
 The following example serves to illustrate the Markdown syntax. 
-It shows how to create a heading, a word in bold font, and a list of several items in Markdown:
+It shows how to create a heading, a word in bold font, and a list of several items in Markdown:<!-- CJ Consider suggesting a specific template, like papaja, that has such an example by default - so users only need to edit the existing template.
+BTW - is there a check_papaja()?
+I want one!-->
 
@@
 
-This type of markup for formatting, instead of the more visual approach of e.g. Microsoft Word, is compatible with many output formats and is easier to put under version control.
-Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
-The visual editor has the additional benefit of promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions).
-Among other features, it eases the generation of citations and cross references.
-
-Now that you are familiar with the basics of writing text in Markdown, we turn our attention to including code into the text.
+One advantage of this type of markup for formatting is that it can be rendered to many different output formats - both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g.
+specific journal requirements.
+Another advantage is that, because R Markdown files are in plain text, they are more suitable for version control using Git.
+Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass <!--CJ Add picture-->in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
+The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
+and easing the generation of citations and references to Tables and Figures.
+
+Now that you are familiar with the basics of Markdown formatting, we turn our attention to including code into the text.
 Code is separated by three backticks (and the programming language in curly braces) like here:
@@
 ````
-
-The results of your code are than dynamically inserted into the document.
-Dynamically means that whenever you turn the R Markdown into its output format the code is executed again and if you changed something the results are updated.
-The output document is than static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
-This static document is somewhat decoupled from the R Markdown and the rest of the code that did create it.
-It is quite possible that several slightly different versions are therefore circulating.
-We therefore recomend to connect the static document to the version control system by dynamically including the identifier of the identifier of the Git commit, called hash, in the document.
-The `repro` package comes with the function `repro::current_hash()` for that purpose.
+The hotkey [Control]+[Alt]+[i] inserts a block of code in the file.
+The results of code enclosed in such backticks will be dynamically inserted into the document (depending on specific settings).
+This means that, whenever you render the R Markdown to its intended output format, the code will be executed and the results updated.
+The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
+
+Once the R Markdown file has been rendered to a static document, the resulting file is decoupled from the R Markdown and the code that created it.
+This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
+To avoid ambiguity, we therefore recommend referencing the identifier of the Git commit at the time of rendering in the static document.
+Simply put, a static document should link to the version of the code that was used to create it.
+The `repro` package comes with the function `repro::current_hash()` for this purpose.
 This document was created from the commit with the hash ``r repro::current_hash()``.
 
-Now that you know how to write text and R code in an R Markdown, you need to know about metadata, like the title or the output format.
+Now that you know how to write text and R code in an R Markdown, you need to know about metadata (also called: YAML front matter).
+These metadate contain information about the document, like the title and the output format.
 Metadata are placed at the beginning of the document and are separated by three dashes from the document body; the following example is a full markdown document where the header is in lines 1--6.
@@
 title: A Hitchhiker's Guide to Reproducible Research in R
-author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
+author: Aaron Peikert, Caspar J. van Lissa and Andreas M. Brandmaier
 abstract: A tutorial about doing the same thing more than once.
@@
 
-Most metadata fields are self-explanatory (like the author field). 
-Sometimes, R package may extend the metadata fields for their own purposes.
-Certain output formats may require specific  metadata. 
-For example, the `rticles`-package that offers templates for authoring scientific manuscripts comes with extra fields that are described in the package documentation and offer extra options for flexible formatting of documents.
-R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
-The `repro` package uses the metadata for hints to make the research project reproducible.
-Researchers should list all dependent resources of their research project in the metadata, that is, R scripts, data files, and external packages in the following format (see everything below the line *repro:*):
+Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
+However, R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
+Some R packages add additional metadata fields.
+The `repro` package adds fields to the metadata to help make the research project reproducible.
+These fields are used to list all dependencies of the research project.
+This includes R scripts, data files, and external packages.
+The format is as follows (see everything below the line *repro:*):
 
@@
 
-This information makes internal dependencies to other files in the project and external dependencies to R packages explicit.
-`repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
-These two files together form the basis for consistency within a research project and consistency across different systems.
-At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.
+This information makes explicitly declares the project's dependencies on files and R packages.
+`repro` uses this information to infer a Makefile<!--CJ Has not been introduced yet--> for the dependencies on other files and a Dockerfile<!--CJ Has not been introduced yet--> that includes all required packages.
+These two files together form the basis for consistency within a research project and consistency across different systems.<!--CJ Consider explaining this in the theoretical intro-->
+At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.<!--CJ This comes way to late.
+Shouldbe in the opening paragraph or something.-->
 This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
-Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).
-In everyday work, the simplest solution is to now call the function `repro::automate()` which both creates a Makefile and a Dockerfile from all R Markdown files in the project (all files with the ending `.Rmd`) based on the respective repro metadata entries.
-Every time you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
-The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
+Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).<!--CJ These points describe the formal definition of reproducibility used in this paper, and the goal of the workflow.
+That needs to be explained in the introduction.
+If you only find out what you are working toward halfway through the tutorial - that's too late!-->
+
+The function `repro::automate()` converts the metadata from all R Markdown files in the project (all files with the ending `.Rmd`) to a Makefile and a Dockerfile.
+Together, these files allow users (including your future self) to reproduce every step in the analysis automatically.
+It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, to keep the Makefile and Dockerfile up to date.
+The function `repro::automate()` does not overwrite manual changes<!--CJ What does that mean?-->, and there is no harm in running it too often.<!-- CJ Can that not be automated?
+Or at least provide a suggestion when to run it?-->
 Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
 <!--AB: why is this addition not automated?
-Feels strange given that this is our recommended best practice?!-->
-
+Feels strange given that this is our recommended best practice?!
+CJ: Agreed.-->
+
+<!-- CJ What i'm missing here is something to wrap up the tutorial. Do users now have a project that could be reproduced? The tutorial should not end until that is the case.-->
+
+## Reproducing a project
+
+<!--CJ Suggestion: Instead of explaining reproduction right in the middle of the tutorial for setting up a project, add a later section called "reproducing a project", and provide an example project for users to reproduce.-->
 When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:
@@
 ```
-
+<!-- CJ: You didn't say anything about using the terminal....
+or will that be explained by messages printed by the code above?
+Either way, I suggest making the fully automatic version the default instead of making users interact with system().-->
 If you uncomfortable with using the terminal directly, you can sent the command to the terminal from within R:
@@
 
-Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
+Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.<!--CJ This belongs in the intro, not the tutorial.
+It is part of explaining the benefits of repro.-->
 These projects will remain reproducible irrespective of whether `repro` will be maintained and remain accessible in future.
 Users do not have to have repro installed to reproduce a project, in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
-The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.
+<!-- CJ This could go into the "reproducing a project" section.
+BTW, it is important to be clear about what *does* need to be installed.
+For example, I don't find it objectionable to have to install R.
+But do users need to install docker?-->
+The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.<!--CJ This really belongs in a "reproducing a project" section"-->
 The commands can be pasted into a terminal and will reproduce the project in many cases.
@@
 Please add a function repro::rerun(execute=TRUE) or -- if we rename it -- repro::reproduce(execute=TRUE).
--->
+CJ: I completely agree with this, too.-->
 
@@
 
-Some projects, however, are not as easy to make reproducible.
-As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
-These simplifications are possible because of two restrictions.
-Users who ask themselves either "How can I install software that is not an R package in the Docker image?" or "How can I express complex dependencies between files, e.g.
-raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
-All other users may safely skip this section and may return to it if they become limited by restriction.
+Some projects, however, are not as easy to make reproducible.<!--CJ Many might not find the above 'easy'. Consider focusing on 'custom workflows' instead of calling the above easy.-->
+As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.<!--CJ This is confusing.
+How can an interface guarantee reproducibility?
+Does repro guarantee reproducibility?
+If so, how?
+Why is this only suitable for simple projects?
+What makes a project simple?-->
+These simplifications<!--CJ Which simplifications?--> are possible because of two restrictions.
+Users who ask themselves either "How can I install software dependencies outside of R in the Docker image?" or "How can I express complex dependencies between files, e.g.
+raw data turned into preprocessed data required by analysis?"<!--CJ Only if the preprocessing happens outside of R, right?
+Isn't that covered by the previous point?
+How about instead: Users that need to work on a access restricted system, that's a very common hurdle I hear--> need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
+Other users may safely skip this section, or return to it if they encounter such challenges.
 

### 2021-07-26 manuscript.Rmd
add intro to outlook

--- 2021-07-22 manuscript.Rmd
@@
 
-<!--AB: two lines of what the future of the package, of such standards, and prereigstration is.
-Maybe say something that preregistration is not always possible or wanted but that reproducibility is almost always a good idea?-->
+Open science practices are a continually evolving field were technical innovations foster changes in research practice.
+Open data is much more widespread use thanks to online storage facilities, preregistration are possible because there are preregistration platforms and so forth.
+We hope that similarly, fully automatic reproduction increases scientific rigor and efficiency.
 

### 2021-07-26 preregistration.Rmd
address andreas comments

@@
 ---
-title: "A Preregistered Comparison of Male and Female Machiavellianism"
+title: "Preregistered as Code: A Comparison of Male and Female Machiavellianism"
 date: "7/10/2021"
@@
 abstract: |
-  We test for group differences in machiavellism between self identified males and females in the MACH-IV in a well powered preregistered study.
+  We test for group differences in machiavellism between self-identified males and females in the MACH-IV in a well powered preregistered study.
+keywords: |
+  preregistration as code, power simulation, machiavellism, gender
 output:
@@
   - \def\baselinestretch{0.98}
+  - \usepackage{draftwatermark}
 ---
@@
 \titlespacing{\section}{.1cm}{.1cm}{.1cm}
+\SetWatermarkText{PAC}
 
@@
 
-There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self-identified males and females.
-@muris2017 note that studies focusing on gender differences in machiavellianism are sparse.
-We therefore want to add to this body of literature.
+There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self-identified males and females [@muris2017], but studies focusing solely on machiavialism are sparse.
+Machiavellianism describes a personality dimension characterized by a cynical disregard of morals in the pursuit of one's own interest, e.g. through manipulation.
+We aim to replicate the finding that males tend to have higher machiavellianism scores [@muris2017].
 <!--end-theory-->
+This research question serves as a testbed for a preregistration scheme we call preregistration as code (PAC) to test if PAC is feasible under realistic conditions.
 
@@
 <!--begin-power-code-->
-```{r}
+```{r, echo=FALSE}
 choosen_power <- .8
@@
 A simulation we conducted indicated that with a sample size of
-`r minn` for an alpha of .05 (two-sided) we archive at least `r choosen_power*100`% power assuming an effect size of `r choosen_d`.
+`r minn` for an alpha of .05 (two-sided) we achieve at least `r choosen_power*100`% power assuming a standardized effect size of d=`r choosen_d`.
 <!--end-power-->

### 2021-07-26 manuscript.Rmd
pregistering explorative analyses

@@
 Complementing a preregistration with code is not unheard of but writing a manuscript with code as an integral dynamic element with mock results based on simulated data offers several advantages over classical preregistration.
-PAC lets consumers of a research document (e.g. editors, reviewers, readers) clearly distinguish how analyses were planned and how they were conducted via version control.
-On the researchers' side, PAC assures that the analysis can effectively answer the research question, i.e.
+First, PAC is more precise by design without becoming overly verbose and can thus more effectively minimize researcher degrees of freedom.
+Second, PAC lets consumers of a research document (e.g.
+editors, reviewers, readers) clearly distinguish how analyses were planned and how they were conducted via version control because preregistration and final manuscript stem from the same dynamic document.
+Third, on the researchers' side, PAC assures that the analysis can effectively answer the research question, i.e.
 that the analysis can recover the parameters of interest in simulation and lets them employ their resources efficiently by supporting study design decisions.
+Fourth, a PAC can incorporate data-dependent decisions.
+Researchers can therefore plan for possible data analytic problems without deviation from their preregistration.
+However, researchers can also preregister explorative analyses, i.e. explorative factor analysis or machine learning methods but precisely calibrate how much exploration is allowed.
+When communicating a finding as explorative, it may make a difference if tens or hundreds of variables were considered.
 
@@
 
-PACs come with a challenge though.
+We realize that PACs are challenging.
+Not only is every research question or project is suitable for preregistration but PAC requires more programming expertise than a usual analysis.
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.

### 2021-07-26 R/simulation_funs.R
use informative variable names for output

@@ -1,30 +1,30 @@
 #----simulate_data----
 simulate_data <- function(n, df, d, i){
   stopifnot(n %% 2L == 0L)
-  group <- rep(c(1L, 0L), each = n/2)
+  gender <- rep(c(0L, 1L), each = n/2)
   rand <-  matrix(rchisq(n * i, df/i), ncol = i) # sum of n chisq has df of n*df
   d_scaled <- d * sqrt(2*df)/i # scale d to express sd units (var(chisq) = 2df)
-  effect <- rand + d_scaled * group
-  colnames(effect) <- paste0("x", seq_len(i))
+  effect <- rand + d_scaled * gender
+  colnames(effect) <- paste0("mach", seq_len(i))
   effect <- as.data.frame(effect)
-  effect$group <- group
+  effect$gender <- factor(gender, levels = c(1L, 0L), labels = c("male", "female"))
   return(effect)
 }
 
 #----planned_analysis----
 planned_analysis <- function(data, use_rank = "skew", skew_cutoff = 1){
-  y <- rowMeans(data["group" != names(data)], na.rm = TRUE)
-  data <- data[!is.na(y),]
-  y <- y[!is.na(y)]
-  x <- as.factor(data$group)
-  skew <- moments::skewness(y)
+  machiavellianism <- rowMeans(data["gender" != names(data)], na.rm = TRUE)
+  data <- data[!is.na(machiavellianism),]
+  machiavellianism <- machiavellianism[!is.na(machiavellianism)]
+  gender <- as.factor(data$gender)
+  skew <- moments::skewness(machiavellianism)
   # skewness cutoff
   if(use_rank == "skew")use_rank <- abs(skew) > skew_cutoff
   if(use_rank){
-    y <- rank(y)
+    machiavellianism <- rank(machiavellianism)
   }
-  test <- t.test(y ~ x)
-  list(test = test, skew = skew, use_rank = use_rank, n = length(y))
+  test <- t.test(machiavellianism ~ gender)
+  list(test = test, skew = skew, use_rank = use_rank, n = length(gender))
 }
 
 #----report_analysis----

### 2021-07-27 preregistration.Rmd
cite definition
change title

@@
 ---
-title: "Preregistered as Code: A Comparison of Male and Female Machiavellianism"
+title: "Preregistered as Code: Gender differences in Machiavellianism"
 date: "7/10/2021"
@@
 There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self-identified males and females [@muris2017], but studies focusing solely on machiavialism are sparse.
-Machiavellianism describes a personality dimension characterized by a cynical disregard of morals in the pursuit of one's own interest, e.g. through manipulation.
+Machiavellianism describes a personality dimension characterized by a cynical disregard of morals in the pursuit of one's own interest, e.g. through manipulation [@christie1970].
 We aim to replicate the finding that males tend to have higher machiavellianism scores [@muris2017].

### 2021-07-27 preregistration.Rmd
address andreas comments

@@
 abstract: |
-  We test for group differences in machiavellism between self-identified males and females in the MACH-IV in a well powered preregistered study.
+  We test for group differences in machiavellianism between self-identified males and females in the MACH-IV in a well powered preregistered study.
 keywords: |
-  preregistration as code, power simulation, machiavellism, gender
+  preregistration as code, power simulation, machiavellianism, gender
 output:
@@
 
-There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self-identified males and females [@muris2017], but studies focusing solely on machiavialism are sparse.
 Machiavellianism describes a personality dimension characterized by a cynical disregard of morals in the pursuit of one's own interest, e.g. through manipulation [@christie1970].
+There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self-identified males and females [@muris2017] but only few studies focus solely on machiavellianism
 We aim to replicate the finding that males tend to have higher machiavellianism scores [@muris2017].
 <!--end-theory-->
-This research question serves as a testbed for a preregistration scheme we call preregistration as code (PAC) to test if PAC is feasible under realistic conditions.
+This research question serves as a testbed for preregistration as code (PAC) to examine if PAC is feasible under realistic conditions.
+In the PAC paradigm, all analysis code is written before real data are gathered or accessed the first time.
 
@@
 <!--begin-ttest-->
-We conduct a Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] of the average of machiavellism items between the binary-coded gender groups.
+We conduct a Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] of the average of machiavellianism items between the binary-coded gender groups.
 If the skew of this average is greater than 1.0 we conduct a supposedly more robust Mann--Whitney--Wilcoxon test [@Wilcoxon1945] instead.
@@
 if(simulated){
-  cat("\\textcolor{red}{This document and the results presented herein are based
-      on simulated data and should not be interpreted.}")
+  cat("\\textcolor{red}{The results are based on simulated data and must not be interpreted. They only serve to illustrate the result of the preregistered code.}")
   set.seed(1234)
@@
 
-\textcolor{red}{This document was created before we have gathered any data, we, therefore, have nothing to discuss.}
+\textcolor{red}{This document was created before we have gathered gathered data. The discussion will only be added to the final report.}
 

### 2021-07-27 preregistration.Rmd
add full stop and linebreaks

@@
 Machiavellianism describes a personality dimension characterized by a cynical disregard of morals in the pursuit of one's own interest, e.g. through manipulation [@christie1970].
-There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self-identified males and females [@muris2017] but only few studies focus solely on machiavellianism
+There is extensive literature reporting differences in the dark triad (narcissism, machiavellianism, and psychopathy) between self-identified males and females [@muris2017] but only few studies focus solely on machiavellianism.
 We aim to replicate the finding that males tend to have higher machiavellianism scores [@muris2017].
@@
 if(simulated){
-  cat("\\textcolor{red}{The results are based on simulated data and must not be interpreted. They only serve to illustrate the result of the preregistered code.}")
+  cat("\\textcolor{red}{The results are based on simulated data and must not be
+      interpreted. They only serve to illustrate the result of the preregistered
+      code.}")
   set.seed(1234)

### 2021-07-27 manuscript.Rmd
fix link

--- 2021-07-26 manuscript.Rmd
@@
 
-These function again should be accept the output of `planned_analysis()` as input.
+This function should again accept the output of `planned_analysis()` as input.
 The output of this function should be formatted in Markdown.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
+Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
 Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
-
-Ideally, these three functions can be composed to create a "fake" results section, e.g.
-when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section.
-
-To summarize researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
-After they gathered the actual data, they replace the simulated data, run `planned_analysis()` and write the discussion.
-
-# Preregistration as Code --- a Tutorial
-
-To guide the reader through a preregistration as code (PAC), we will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
-
-> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"
-
-This example serves only didactic purposes and not to derive substantive claims.
-Nevertheless, everything else is kept as close to a real research question as possible.
-
-At the heart of a PAC lies the planned analysis.
-We suggest writing a main analysis function that captures the full process of testing a research question.
-This function receives a dataset as input and performs all statistical processing. <!--AB say anything about its output? is the Markdown the output?-->
-For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
-As an example of a data-dependent decision, we included a rank transformation if the skewness was greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it generates not only tables but also a straightforward interpretation of the effects as actual prose (e.g.
+it quantifies the size of an effect).
+
+Ideally, these three functions can be composed to create a mock results section, e.g.
+when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` they output a full results section.
+
+To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
+After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()`) and write the discussion.
+
+## An example of Preregistration as Code
+
+We have made the argument that PAC has significant advantages over classical preregistration and have outlined how to implement PAC, but can PAC work in practice?
+To answer this question and help researchers implementing PAC themselves, we have conducted a PAC that can serve as an example.
+We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
+
+> "Is there a mean difference in the personality trait 'Machiavellism' between self-identified females and males?"
+
+Again, we propose a preregistration format that closely resembles a classical journal article, but that utilizes simulated data and dynamic document generation to serve simultaneously as preregistration and final manuscript.
+The complete example is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
+In this section, we show excerpts of this example formatted in monospace and provide some commentary.
+
+```{r, echo=FALSE}
+show <- function(x, wrap = TRUE, cat = TRUE, fence = TRUE){
+  stopifnot(length(x) == 1L)
+  if(wrap)x <- str_c(str_wrap(str_split(x, "\\n\\n")[[1]]), collapse = "\n\n")
+  if(!isFALSE(fence)){
+    if(isTRUE(fence))fence <- "markdown"
+    fence <- str_c("````", fence)
+    x <-
+      str_c(fence,
+            x,
+            "````",
+            sep = "\n",
+            collapse = "\n")
+  }
+  if(cat){
+    cat(x)
+    return(invisible(x))
+  } else {
+    return(x)
+  }
+}
+show_prereg <-
+  function(section,
+           file = "preregistration.Rmd",
+           wrap = TRUE,
+           cat = TRUE,
+           fence = TRUE) {
+    prereg <- readr::read_lines(here::here(file))
+    start <- stringr::str_c("<!--begin-", section, "-->")
+    end <- stringr::str_c("<!--end-", section, "-->")
+    start_pos <- which(prereg == start)
+    end_pos <- which(prereg == end)
+    if (length(start_pos) != 1L || length(end_pos) != 1L)
+      stop("No section '", section, "' found.")
+    excerpt <-
+      str_c(prereg[(start_pos + 1):(end_pos - 1)], collapse = "\n")
+    show(excerpt, wrap = wrap, cat = cat, fence = fence)
+  }
+```
+
+As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary information and literature to understand the rest of the paper.
+This example is drastically shortened for illustration purposes:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("theory")
+```
+
+After researchers have provided the research question, they proceed to explain how they want to study it.
+For simplicity, we will use already published data that we have not yet analyzed:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("method")
+```
+
+We choose the following statistical procedure because many researchers may be familiar with it[^caution-mww]:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("ttest")
+```
 
 [^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
-The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
-
-This function is one possible translation of the verbal description above:
+The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
+
+We then translated this methods section into the following `planned_analysis()` function:
 
@@
 
+```{r, echo=FALSE, results='asis'}
+show_prereg("power-code", wrap = FALSE)
+```
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("power")
+```
+
+We implemented a mechanism that only uses simulated data when the actual data is not yet available for the results section.
+This mechanism also warns readers if these results are based on simulated data.
+As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study.
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("results", wrap = FALSE)
+```
+
+The last thing we need to do is to report our results with the combination of `planned_analysis()` and `report_analysis()`.
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("report", wrap = FALSE)
+```
+
+This is an example of how the results could be reported (based on simulated data):
+
+```{r, echo=FALSE, results='asis'}
+set.seed(1234)
+show(report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)), cat = FALSE), fence = "")
+```
+
+The interested reader can find the example rendered with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
+
 # Discussion
 
-We have presented a workflow that rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to ensure reproducibility.
-This workflow is straightforward to implement with the `repro`-package, which provides a low threshold for adoption without limiting the users' ability to implement complex projects and is, therefore, a "high ceiling" solution.
-The `repro`-package builds upon and extends established best practices, i.e. R Markdown and Git, in two critical ways.
-First, it provides the required software for reproduction by automatically generating Dockerfiles.
-Second, by generating Makefiles the `repro`-package establishes a unified mechanism for reproducing research projects.
-
-We than proceeded to show one example of how such rigorous and automated reproducibility workflow may enable other innovations, e.g. preregistration as code.
-Complementing a preregistration with code is not unheard of but writing a manuscript with code as an integral dynamic element with mock results based on simulated data offers several advantages over classical preregistration.
-PAC lets consumers of a research document (e.g. editors, reviewers, readers) clearly distinguish how analyses were planned and how they were conducted via version control.
-On the researchers' side, PAC assures that the analysis can effectively answer the research question, i.e.
-that the analysis can recover the parameters of interest in simulation and lets them employ their resources efficiently by supporting study design decisions.
+<!--AB: TODO add a quick summary -->
 
@@
 
-`repro`'s main goal is to increase the user friendliness of reproducibility tools for researchers.
-While we have invested much effort to ensure this goal, the package is still in early development and undoubtedly has rough edges and even bugs that went undetected.
-We hope that despite the early development stage `repro` can support researchers to work more reproducible.
-
-PACs come with a challenge though.
+PACs come with a challenge.
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
@@
 
-Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
+Please note that the statistical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for addressing your research question.
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
@@
 
+We also suggest that version control with Git and GitHub fulfills the criteria of a preregistration as long as you state your research plan comprehensively.
+This view is not shared by everyone [@szollosi_is_2020] because with some effort and malice intend the record of Git/GitHub can be altered.
+This concern can be addressed by using zenodo.org to give a digital object identifier (DOI) to your preregistration (see [this guide](https://guides.github.com/activities/citable-code/)) because zenondo.org saves an unchangeable copy of your repository.
+
 # Outlook
 
-Open science practices are a continually evolving field were technical innovations foster changes in research practice.
-Open data is much more widespread use thanks to online storage facilities, preregistration are possible because there are preregistration platforms and so forth.
-We hope that similarly, fully automatic reproduction increases scientific rigor and efficiency.
+<!--AB: two lines of what the future of the package, of such standards, and prereigstration is.
+Maybe say something that preregistration is not always possible or wanted but that reproducibility is almost always a good idea?-->
 

### 2021-07-27 manuscript.Rmd
add outlook

--- 2021-07-26 manuscript.Rmd
@@
 
-We than proceeded to show one example of how such rigorous and automated reproducibility workflow may enable other innovations, e.g. preregistration as code.
+We then proceeded to show one example of how such rigorous and automated reproducibility workflow may enable other innovations, e.g. preregistration as code.
 Complementing a preregistration with code is not unheard of but writing a manuscript with code as an integral dynamic element with mock results based on simulated data offers several advantages over classical preregistration.
@@
 
-`repro`'s main goal is to increase the user friendliness of reproducibility tools for researchers.
+`repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
 While we have invested much effort to ensure this goal, the package is still in early development and undoubtedly has rough edges and even bugs that went undetected.
-We hope that despite the early development stage `repro` can support researchers to work more reproducible.
+We hope that despite the early development stage, `repro` can support researchers to work more reproducible.
 
 We realize that PACs are challenging.
-Not only is every research question or project is suitable for preregistration but PAC requires more programming expertise than a usual analysis.
+Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove unwieldy for a PAC.
+Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not preregistered research but as a compliment.
+We are convinced that many reasons justify changing the preregistered code, but we are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
+
+Another challenge is that PAC requires more programming expertise than a usual analysis.
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
 We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
-This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
+This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
 The reproducible workflow we present here is heavily automated and hence promotes reusability.
-Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
+Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
+Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
 
@@
 
-Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
+Please note that the analytical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes, and it is not necessarily optimal for addressing your research question.
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
-The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
+The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
 
@@
 
-Open science practices are a continually evolving field were technical innovations foster changes in research practice.
-Open data is much more widespread use thanks to online storage facilities, preregistration are possible because there are preregistration platforms and so forth.
-We hope that similarly, fully automatic reproduction increases scientific rigor and efficiency.
+Open science practices are a continually evolving field where technical innovations foster changes in research practice.
+Open data is much more widespread thanks to online storage facilities; preregistration are possible because there are preregistration platforms and so forth.
+Similarly, we hope that fully automatic reproduction, e.g. with `repro` as technical innovation, increase scientific rigour and efficiency.
+
+This ideal of a fully automatic reproduction of research projects meets in practice a wide range of demands on how user-friendly and how powerful the software solutions are.
+Some may find Make too complicated or that Docker requires too much storage.
+Yet others may find that they require other programming languages or scale their computation across hundreds of computers, e.g. via high-performance computing clusters or cloud computing.
+
+`repro` was designed modularly to meet such demands.
+At the moment, `repro` only supports the combination of RMarkdown, Git, Make, and Docker.
+However, there are alternatives for each of these elements that may fit better into an individual research project.
+RMarkdown could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer` or `officedown` to accommodate a wider arrange of journal submission standards.
+Instead of using formal version control with Git, `repro` could automatically save snapshots for increasing user-friendliness.
+Make could be replaced by the more R-centred alternative `targets` for more convenience.
+Docker could be combined with `renv` to control the package versions precisely (our approach fixes the date, `renv` the exact package version).
+Alternatively, Docker could be replaced by the more lightweight `renv` if no dependencies outside of R are considered crucial.
+Docker does not satisfy the requirements of many HPC environments, but Singularity was designed to avoid this limitation while still being compatible with Docker images.
+
+`repro`'s modular structure allows such alternative workflow, though they have not been implemented.
+Depending on the demand by users, we will implement some of these alternative workflows in `repro`.
 

### 2021-07-27 manuscript.Rmd
add outlook polish discussion

@@
 Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not preregistered research but as a compliment.
-We are convinced that many reasons justify changing the preregistered code, but we are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
+Some may see the rigidness of PAC as to much of a good thing.
+However, we are convinced that many reasons justify changing the preregistered code, but we are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
 
@@
 However, there are alternatives for each of these elements that may fit better into an individual research project.
-RMarkdown could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer` or `officedown` to accommodate a wider arrange of journal submission standards.
+RMarkdown could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer`[@officer] or `officedown`[@officedown] to accommodate a wider arrange of journal submission standards.
 Instead of using formal version control with Git, `repro` could automatically save snapshots for increasing user-friendliness.
 Make could be replaced by the more R-centred alternative `targets` for more convenience.
-Docker could be combined with `renv` to control the package versions precisely (our approach fixes the date, `renv` the exact package version).
+Docker could be combined with `renv` [@R-renv] to control the package versions precisely (our approach fixes the date, `renv` the exact package version).
 Alternatively, Docker could be replaced by the more lightweight `renv` if no dependencies outside of R are considered crucial.

### 2021-07-27 manuscript.Rmd
Grammar until line 300

--- 2021-07-27 manuscript.Rmd
@@
 Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
-Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
-22][^popper]
+The credibility of empirical research hinges on its objectivity.
+Objectivity in this context means "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
+22][^popper].
+Therefore, reproducibility has long been considered integral to empirical research.
+<!---
+You tend to use quite long sentences with many subclauses.
+While this has a long tradition in the german language, it is considered bad style in english.
+Also it often leads to minor mistakes like dangling modifiers.
+-->
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
@@
 We pursue two aims.
-First, the present paper introduces the reader to a workflow that maximizes the potential reproducibility of a research project.
+First, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
 This workflow builds on several technical solutions originating in the software engineering community. 
@@
 We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
-The second objective is to discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
+Second, we discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
 One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
-PAC involves preregistering the intended analysis code and the bulk of the manuscript as dynamic document, including the Introduction, Methods, and Results sections.
-The resulting dynamic document closely resembles the final manuscript, but uses preliminary ("mock") results, based on the analysis of simulated data, as a placeholder for figures, tables, and statistics.
-After the empirical data are analyzed, the Results are updated, and the Discussion is written.
+PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including the Introduction, Methods, and Results sections.
+The resulting dynamic document closely resembles the final manuscript, but uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
+Once the empirical data are analyzed, the Results are updated, and the Discussion is written.
 
@@
 Scientific organizations and funding bodies increasingly demand transparent sharing of research artifacts, and researchers are increasingly willing to do so.
+<!---
+I am not shure about the word artifact in general. Is it common terminology?
+-->
 However, although the sharing of such artifacts is a necessary condition for reproducibility, it is not a sufficient one.
@@
 An easy-to-use standard, on the other hand, is more likely to be adopted.
-Our approach is to promote broad adoption of easy-to-use practices, i.e., a "low threshold", and ensure that these practices are compatible with more complex rigorous solutions, i.e., a "high ceiling".
+Our approach is to promote broad adoption of easy-to-use practices (a "low threshold"), and ensure that these practices are compatible with more complex rigorous solutions (a "high ceiling").
 As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
@@
 
-1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
+1.	**Multiple inconsistent versions of code, data, or both**; for example
+  - the data set could have changed over time because outliers were removed at a later stage or an item was later recoded
+  - the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
+It may then be unclear which version of code and data was used to produce some reported set of results.
 2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
@@
 
-We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term and cross-platform computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph.
+We have developed a workflow that achieves long-term and cross-platform computational reproducibility of scientific data analyses.
+It leverages established tools and practices from software engineering and rests on four pillars that address the aforementioned causes of non-reproducibility: [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph.
 The opening paragraph should at least:
@@
 We recommend GitHub to host Git repositories because of its popularity among R users.
-GitHub has many tools that ease project management and collaboration, and these tools provide much value in our everyday work, but they are not central to achieving reproducibility.
+GitHub has many tools that ease project management and collaboration, and these tools are valuable in scientists everyday work, but they are not central to achieving reproducibility.
 
@@
 In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
-Typically, results are manually copyied and pasted from one software to the other, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
+Typically, results are manually copied and pasted from one software another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
 
@@
 It is highly fool-proof [unlike, for example, LaTeX @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your résumé [@vitae].
+We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
 R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
 Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
-This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
+This allows the creation of formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
 @xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.
@@
 However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
-In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambigously documented.
+In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
 
 To automatically resolve dependencies between project files, we rely on **dependency tracking**.
-In essence, researchers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
+In essence, researchers provide a collection of computational instructions (_recipes_) that describe how ingredients are processed to create intermediate products and, in the end, a final product.
 Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
-In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, analysis results).
+In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
 
 We recommend using **Make** for dependency tracking because it is language independent.
-The following hypothesical example illustrates the utility of Make.
+The following hypothetical example illustrates the utility of Make.
 Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
@@
 Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
-One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
+One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via **software management**.
 One comprehensive approach to software management is containerization.
@@
 It provides all software used in an analysis, including the operating system itself.
-The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
+(The operating system level is important because some functionality may be passed to software layers beneath the programming language, for example random number generators and linear algebra libraries.)
 Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
@@
 The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
-A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
+A second strategy is to save a container recipe, a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
 Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
 However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
-For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the Dockerfile.
+For proper archival, we therefore recommend storing a complete image of the software environment besides the Dockerfile.
 A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
 
-To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
+To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)); dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
 While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.

### 2021-07-27 manuscript.Rmd
rest of grammar

@@
 
-The above mentioned tools require much expertise that is beyond the traditional curriculum of researchers.
-To aid researcher adopting these tools we rely on or have developed helpers that wrap more complicated (and powerful) tools.
-These wrappers use standard tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
-We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
+The above mentioned tools require expertise that is beyond the traditional curriculum of researchers.
+To aid researchers adopting these tools we rely on and have developed helpers that wrap more complicated (and powerful) tools.
+These helpers wrap standard tools but provide feedback about what the computer does and what the user should do in layman's terms and are accessible from within R.
 These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
@@
 Before we can start creating reproducible research projects, we need a computer with a working R Studio installation. 
-Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so that it can act as an assistance systems while you follow the tutorial:
+Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so it can assist you following the tutorial:
 
@@
 An example of the assistance that `repro` provides are the check functions.
-They verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
+They verify that you indeed have installed and setup the required pieces of software for this workflow (we use the double-colon operator to access functions from the `repro` namespace to make clear which functions are native `repro` functions):
 
@@
 They detect the users' system state and explain if further action is needed.
-Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:
+For example, the following happens if you were a Windows user who does not have Git installed:
 
@@
 You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore` file.
-You can add something to the `.gitignore` file directly or with this command:
+You can edit the `.gitignore` file directly or with this command:
 
@@
 Click on the empty box next to the file you want to stage, and when a tick appears, which means that the file is staged.
-After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
+After you have staged all files you would like to, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
+<!---
+Maybe first click on the commit button and then stage, because in the bigger window you can browse the changes.
+-->
 
@@
 Now that you have created a project that is under version control, we continue with adding dynamic document generation.
-Adynamic document has three elements:
-
-1. Text in a natural language
+A dynamic document has three elements:
+
+1. Text
 2. Executable code
@@
 
-The text of an R Markdown is written in Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
+The text of an R Markdown is formatted by Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
 The following example serves to illustrate the Markdown syntax.
@@
 Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
-The visual editor has the additional benefit of promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions).
+The visual editor has the additional benefit of promoting best practices (for example, each sentence should be written on a new line, which eases tracking changes across versions).
 Among other features, it eases the generation of citations and cross references.
@@
 Now that you are familiar with the basics of writing text in Markdown, we turn our attention to including code into the text.
-Code is separated by three backticks (and the programming language in curly braces) like here:
+Code is separated by three backticks (and the programming language in curly braces):
 
@@
 
-The results of your code are than dynamically inserted into the document.
+The results of your code are dynamically inserted into the document.
 Dynamically means that whenever you turn the R Markdown into its output format the code is executed again and if you changed something the results are updated.
-The output document is than static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
+<!---
+you did not explain how to knit and to choose an output format
+-->
+The output document is then static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
 This static document is somewhat decoupled from the R Markdown and the rest of the code that did create it.
 It is quite possible that several slightly different versions are therefore circulating.
-We therefore recomend to connect the static document to the version control system by dynamically including the identifier of the identifier of the Git commit, called hash, in the document.
+We therefore recommend to connect the static document to the version control system by dynamically including the identifier of the Git commit, called hash, in the document.
 The `repro` package comes with the function `repro::current_hash()` for that purpose.
@@
 Now that you know how to write text and R code in an R Markdown, you need to know about metadata, like the title or the output format.
-Metadata are placed at the beginning of the document and are separated by three dashes from the document body; the following example is a full markdown document where the header is in lines 1--6.
+Metadata are placed at the beginning of the document and are separated by three dashes from the document body.
+The following example is a full markdown document where the the metadata (the "header") is in lines 1--6.
 
@@
 Most metadata fields are self-explanatory (like the author field). 
-Sometimes, R package may extend the metadata fields for their own purposes.
+Sometimes, R packages may extend the metadata fields for their own purposes.
 Certain output formats may require specific  metadata. 
-For example, the `rticles`-package that offers templates for authoring scientific manuscripts comes with extra fields that are described in the package documentation and offer extra options for flexible formatting of documents.
+For example, the `rticles`-package, which offers templates for authoring scientific manuscripts, comes with extra fields that are described in the package documentation and offer extra options for flexible formatting of documents.
 R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
 The `repro` package uses the metadata for hints to make the research project reproducible.
-Researchers should list all dependent resources of their research project in the metadata, that is, R scripts, data files, and external packages in the following format (see everything below the line *repro:*):
+Researchers should list all dependent resources of their research project in the metadata, that is, R scripts, data files, and external packages, in the following format (see everything below the line *repro:*):
 
@@
 These two files together form the basis for consistency within a research project and consistency across different systems.
-At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.
+<!---
+What is consistency?
+-->
+At this point, we would like to emphasize that a prerequisite for reproducibility is that every step taken in an analysis can be **automatically** reproduced.
 This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
 Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).
-In everyday work, the simplest solution is to now call the function `repro::automate()` which both creates a Makefile and a Dockerfile from all R Markdown files in the project (all files with the ending `.Rmd`) based on the respective repro metadata entries.
+
+In everyday work, the simplest solution is to now call the function `repro::automate()` which creates a Makefile and a Dockerfile from all R Markdown files in the project based on the respective repro metadata entries.
 Every time you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
@@
 
-When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:
+When you or someone else who received your project would like to reproduce it, they can ask repro to explain how this should be done:
 
@@
 
-If you uncomfortable with using the terminal directly, you can sent the command to the terminal from within R:
+If you feel uncomfortable using the terminal directly, you can sent the command to the terminal from within R:
 
@@
 
-Some projects, however, are not as easy to make reproducible.
+Some projects, however, are not so easy to make reproducible.
 As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
@@
 raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
-All other users may safely skip this section and may return to it if they become limited by restriction.
-
-The metadata the `repro::automate()` function relies on can only express dependencies R packages for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
+All other users may safely skip this section and return if they need to.
+
+The metadata the `repro::automate()` function relies on can only express R packages as dependencies for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile.
 The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
-Other software that users might need like other programming languages, not already installed LaTeX packages etc. have to be added manually.
-We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g.
-for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
+Other software that users might need like other programming languages and not preinstalled LaTeX packages have to be added manually.
+We plan to add support for commonly used ways to install software beyond R packages, e.g.
+for system libraries (via the Ubuntu package manager `apt`), LaTeX packages (via the Tex Live package manager `tlmgr`), and python packages (via the python package manager `pip`).
 The second limitation is about dependencies.
@@
 These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
-Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D therefore leaving out important information or express the dependencies directly within the Makefile.
-
-The following section explains how to overcome these limitation while still relying on the automatons afforded by `repro`.
-Lifting these restrictions requires the user to interact more directly with Make or Docker.
+Users therefore have to either "flatten" the dependency structure by simply stating that A depends on B, C and D, thereby leaving out important information, or express the dependencies directly within the Makefile.
+
+The following section explains how to overcome these limitations while still relying on the automations provided by `repro`.
+Lifting these limitations requires the user to interact more directly with Make or Docker.
 To satisfy more complicated requirements users need to understand how repro utilizes Make and Docker internally.
-
-Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two processing steps.
+<!-- I am pretty shure you are saying that users should satisfy more complicated requirements ^^ -->
+
+Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`. It consists of two processing steps:
 First it recreates the virtual software environment (Docker) and then it executes computational recipes in the virtual software environment (Make).
-The first step is done by the command `make docker`.
-The command `make docker` will trigger Make to build the target called `docker`.
+The first step is done by the command `make docker`, it will trigger Make to build the target called `docker`.
 The recipe for this target builds an image from the Dockerfile of the repository.
@@
 By default the R version corresponds to the locally installed version on which `repro::automate()` was called the first time.
-The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed, by default this date is set to the date on which `repro::automate()` was called the first time.
+The build date is used to install packages in the version available on CRAN on this specific date and can also be changed, by default it is set to the date on which `repro::automate()` was called the first time.
 This way, the call to the automate function virtually freezes the R environment to the state it was called the first time inside the container. 
@@
 The most convenient way to install software on Ubuntu is through its package manager `apt`.
-The following snipped installs Python.
-If added to `.repro/Dockerfile_manual` the Docker image will have Python installed.
+If the following snipped is added to `.repro/Dockerfile_manual` the Docker image will have Python installed.
 
@@
 the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the Makefile.
-The Makefile that `repro` creates is only an template and you are free to change it.
+The Makefile that `repro` creates is only a template and you are free to change it.
 However, make sure to never remove the following two lines:
@@
 This file should not be altered manually.
-If you are not satisfied with the automatically generated target than simply provide an alternative target in the main `Makefile`.
+If you are not satisfied with the automatically generated target simply provide an alternative target in the main `Makefile`.
 Targets in the main `Makefile` take precedent.
@@
 Additionally, it provides facilities to execute commands within the container.
-If you write computational recipe for a target it will be by default evaluated using the locally installed software.
+If you write a computational recipe for a target it will be by default evaluated using the locally installed software.
 To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first Make example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
@@
 Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For a priori planned analyses, preregistration should reduce "researcher degrees of freedom" therefore, the number of decisions researcher can make that could unduly influence the results.
-Preregistrations, however, have a problem.
+For a priori planned analyses, preregistration should reduce "researcher degrees of freedom", the number of decisions researchers can make that could influence the results.
+However, some problems described in the literature make it questionable whether preregistration can achieve its goals.
 Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
 The current best practice to ensure that the preregistration is comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
-The structure these templates provide can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
-Nevertheless, these templates can not remove researchers' degrees of freedom because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
+The structure these templates provide can more effectively reduce researcher degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
+Nevertheless, these templates can not remove all researcher degrees of freedom because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
 
 Our proposed solution is simple:
-Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
-We term this writing and publishing code before data collection *preregistration as code (PAC)*.
-PAC has the potential to all but eliminates researchers degrees of freedom that standard preregistration despite the best efforts afford.
+Instead of describing the analysis, you write and publish the code to conduct the analysis before you gather data.
+<!-- I am not shure if it is good to directly talk to the reader in this form -->
+We term this *preregistration as code (PAC)*.
+PAC has the potential to eliminate all researcher degrees of freedom, even those standard preregistration can not eliminate.
 
@@
 
-We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.
+We believe, however, that combining ideas from reproducibility and preregistration can turn this idea into a tool with tangible benefits for the research community.
 
@@
 Second, although it has been almost a decade since @steegenMeasuringCrowdAgain2014 have preregistered their code, the idea of publishing code before analyzing data has gained not much traction.
+<!-- This is not a reason why PAC and reproducibility are related -->
 This could change when combining the here proposed workflow and PAC.
@@
 It follows that for preregistrations they should stick to widely accepted standards of scientific writing [e.g. @apa7].
-We have already explained the need for reproducibility standards in the process of writing such manuscript and it is therefore only natural to extend these arguments to preregistration.
+We have already explained the need for reproducibility standards in the process of writing such manuscripts and it is therefore only natural to extend these arguments to preregistration.
 
@@
 Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
-For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
+For example, if distributional assumptions are not met, the code may branch out to employ robust methods or an analysis may perform automated variable selection mechanisms before running the final model.
 Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
-We believe it is easier with PAC to detect missing steps or ambiguous decisions.
-Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
+We believe PAC reveals missing steps or ambiguous decisions. 
+It can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
 Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
@@
 
-When leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
+Leaving meta-scientific benefits aside, a preregistration is also a tool for study planning.
 To use PAC for study planning, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
@@
 Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
-Third, the plan should be efficient; the plan should only require the necessary amount of resources.
+Third, the plan should be efficient; it should only require the necessary amount of resources.
 
@@
 Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
+<!-- I dont understand this sentence -->
 One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
@@
 It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
-Writing incomplete code is technically impossible, or at least researchers would notice if they, e.g. forgot to aggregate the data, and so forth.
+Writing incomplete code is technically impossible, or at least researchers would notice if something is omitted (e.g. if they forgot to aggregate the data).
 
@@
 Because the simulation captures the researchers' expectations about the real world, it is the best way to assess the effectiveness of the analysis without actually gathering the data.
+<!-- I would maybe say something about why the simulated data captures researchers expectations.
+I think for most people data simulation is a novel topic, and even more so the idea that complicated assumptions about the real world (skewness, factor structure of a test, violated measurement invariance) can be incorporated into the simulation -->
 
@@
   # 2. conduct analysis e.g. with `t.test()`
-  # 3. `return(results)`
+  # 3. return the results
 }
@@
 ```{r, eval=FALSE}
-simulate_data <- function(n, effect_size){
+simulate_data <- function(n, effect_size, measurement_error, etc.){
   # 1. warn users that the results are mocked
   # 2. draw `n` samples with `effect_size`
-  # 3. format and return in expected data format
+  # 3. Further assumptions can be passed to this function, like measurement error, skewness, etc.
+  # 4. format and return in expected data format
 }
@@
 The implementation of this reporting should be in the function `report_analysis()`.
+<!-- Inconsistent naming of the function report_analysis() !! -->
 
@@
   # 2. conditionally interpret results e.g. if(p < .025)"Result is significant."
-  # (optional) visualize results
-  # 3. return results section formatted in markdown
+  # 3. visualize results
+  # 4. return results section formatted in markdown
 }
@@
 
-To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
-After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()`) and write the discussion.
+To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration.
+After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data) and write the discussion.
 
@@
 
-We have made the argument that PAC has significant advantages over classical preregistration and have outlined how to implement PAC, but can PAC work in practice?
-To answer this question and help researchers implementing PAC themselves, we have conducted a PAC that can serve as an example.
+We have made the argument that PAC has significant advantages over classical preregistration and have outlined how to implement it, but some may be sceptical if it is feasible.
+To address this concern and help researchers implementing PAC themselves, we have conducted a PAC that can serve as an example.
 We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
@@
 
-Again, we propose a preregistration format that closely resembles a classical journal article, but that utilizes simulated data and dynamic document generation to serve simultaneously as preregistration and final manuscript.
+Again, we propose a preregistration format that closely resembles a classical journal article, but utilizes simulated data and dynamic document generation to serve simultaneously as preregistration and final manuscript.
 The complete example is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
@@
 This function illustrates two advantages of PAC.
-First, a PAC can easily include data-dependent decisions creating different branches of analysis under different conditions.
+First, a PAC can easily include data-dependent decisions creating different analysis branches under different conditions.
 Second, it highlights how difficult it is to describe a statistical analysis precisely.
@@
 
-The interested reader can find the example rendered with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
+The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
+<!-- I would show the rendered PDF (as long as it is shorter than one page) -->
 
@@
 
-PACs come with a challenge.
+PAC comes with a challenge.
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
@@
 Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
+Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
 
@@
 
-Please note that the statistical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for addressing your research question.
+Please note that the statistical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and is not necessarily optimal for addressing your research question.
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
 The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
+<!-- nobody cares about limitations of the example analysis, I would delete all of this -->
 

### 2021-07-28 manuscript.Rmd
review max's edits

@@
 Therefore, reproducibility has long been considered integral to empirical research.
-<!---
-You tend to use quite long sentences with many subclauses.
-While this has a long tradition in the german language, it is considered bad style in english.
-Also it often leads to minor mistakes like dangling modifiers.
--->
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
@@
 Scientific organizations and funding bodies increasingly demand transparent sharing of research artifacts, and researchers are increasingly willing to do so.
-<!---
-I am not shure about the word artifact in general. Is it common terminology?
--->
+
 However, although the sharing of such artifacts is a necessary condition for reproducibility, it is not a sufficient one.
@@
 In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
-Typically, results are manually copied and pasted from one software another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
+Typically, results are manually copied and pasted from one software to another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
 
@@
 Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
-This allows the creation of formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
+This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
 @xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.
@@
 To automatically resolve dependencies between project files, we rely on **dependency tracking**.
-In essence, researchers provide a collection of computational instructions (_recipes_) that describe how ingredients are processed to create intermediate products and, in the end, a final product.
+In essence, researchers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
 Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
@@
 It provides all software used in an analysis, including the operating system itself.
-(The operating system level is important because some functionality may be passed to software layers beneath the programming language, for example random number generators and linear algebra libraries.)
+The operating system level is important because some functionality may be passed to software layers beneath the programming language, for example random number generators and linear algebra libraries.
 Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
@@
 These helpers wrap standard tools but provide feedback about what the computer does and what the user should do in layman's terms and are accessible from within R.
+We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
 These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
@@
 They detect the users' system state and explain if further action is needed.
-For example, the following happens if you were a Windows user who does not have Git installed:
+Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:
 
@@
 After you have staged all files you would like to, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
-<!---
-Maybe first click on the commit button and then stage, because in the bigger window you can browse the changes.
--->
 
@@
 Dynamically means that whenever you turn the R Markdown into its output format the code is executed again and if you changed something the results are updated.
-<!---
-you did not explain how to knit and to choose an output format
--->
 The output document is then static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
@@
 These two files together form the basis for consistency within a research project and consistency across different systems.
-<!---
-What is consistency?
--->
 At this point, we would like to emphasize that a prerequisite for reproducibility is that every step taken in an analysis can be **automatically** reproduced.
@@
 Lifting these limitations requires the user to interact more directly with Make or Docker.
-To satisfy more complicated requirements users need to understand how repro utilizes Make and Docker internally.
-<!-- I am pretty shure you are saying that users should satisfy more complicated requirements ^^ -->
-
-Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`. It consists of two processing steps:
+To satisfy more complicated requirements useres may have, they need to understand how `repro` utilizes Make and Docker internally.
+
+Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`.
+It consists of two processing steps:
 First it recreates the virtual software environment (Docker) and then it executes computational recipes in the virtual software environment (Make).
@@
 Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For a priori planned analyses, preregistration should reduce "researcher degrees of freedom", the number of decisions researchers can make that could influence the results.
-However, some problems described in the literature make it questionable whether preregistration can achieve its goals.
+For a priori planned analyses, preregistration should reduce "researcher degrees of freedom" therefore, the number of decisions researcher can make that could unduly influence the results.
+Preregistrations, however, have a problem.
 Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
@@
 Instead of describing the analysis, you write and publish the code to conduct the analysis before you gather data.
-<!-- I am not shure if it is good to directly talk to the reader in this form -->
 We term this *preregistration as code (PAC)*.
@@
 Second, although it has been almost a decade since @steegenMeasuringCrowdAgain2014 have preregistered their code, the idea of publishing code before analyzing data has gained not much traction.
-<!-- This is not a reason why PAC and reproducibility are related -->
 This could change when combining the here proposed workflow and PAC.
@@
 For example, if distributional assumptions are not met, the code may branch out to employ robust methods or an analysis may perform automated variable selection mechanisms before running the final model.
-Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
-We believe PAC reveals missing steps or ambiguous decisions. 
+Third, a PAC is more comprehensive by design because its completeness can be checked with simulated data.
+Using simulated data makes it straightforward to detect missing steps or ambiguous decisions.
+Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
 It can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
@@
 
-Leaving meta-scientific benefits aside, a preregistration is also a tool for study planning.
+When leaving meta-scientific benefits aside, a preregistration is also a tool for planning a study.
 To use PAC for study planning, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
@@
 Because the simulation captures the researchers' expectations about the real world, it is the best way to assess the effectiveness of the analysis without actually gathering the data.
-<!-- I would maybe say something about why the simulated data captures researchers expectations.
-I think for most people data simulation is a novel topic, and even more so the idea that complicated assumptions about the real world (skewness, factor structure of a test, violated measurement invariance) can be incorporated into the simulation -->
 
@@
 ```{r, eval=FALSE}
-simulate_data <- function(n, effect_size, measurement_error, etc.){
+simulate_data <- function(n, effect_size){
   # 1. warn users that the results are mocked
   # 2. draw `n` samples with `effect_size`
-  # 3. Further assumptions can be passed to this function, like measurement error, skewness, etc.
-  # 4. format and return in expected data format
+  # 3. format and return in expected data format
 }
@@
 The implementation of this reporting should be in the function `report_analysis()`.
-<!-- Inconsistent naming of the function report_analysis() !! -->
 
 ```{r, eval=FALSE}
-report_results <- function(results){
+report_analysis <- function(results){
   # 1. create markdown tables from results
@@
 The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
-<!-- I would show the rendered PDF (as long as it is shorter than one page) -->
 
@@
 The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-<!-- nobody cares about limitations of the example analysis, I would delete all of this -->
 

### 2021-07-28 manuscript.Rmd
fix botched merge

--- 2021-07-27 manuscript.Rmd
@@
 source(here::here("R", "link.R"))
-source(here::here("R", "simulation_funs.R"))
 ```
@@
 Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
-The credibility of empirical research hinges on its objectivity.
-Objectivity in this context means "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
-22][^popper].
-Therefore, reproducibility has long been considered integral to empirical research.
-<!---
-You tend to use quite long sentences with many subclauses.
-While this has a long tradition in the german language, it is considered bad style in english.
-Also it often leads to minor mistakes like dangling modifiers.
--->
+Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
+22][^popper]
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
@@
 We pursue two aims.
-First, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
+First, the present paper introduces the reader to a workflow that maximizes the potential reproducibility of a research project.
 This workflow builds on several technical solutions originating in the software engineering community. 
@@
 We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
-Second, we discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
+The second objective is to discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
 One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
-PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including the Introduction, Methods, and Results sections.
-The resulting dynamic document closely resembles the final manuscript, but uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
-Once the empirical data are analyzed, the Results are updated, and the Discussion is written.
+PAC involves preregistering the intended analysis code and the bulk of the manuscript as dynamic document, including the Introduction, Methods, and Results sections.
+The resulting dynamic document closely resembles the final manuscript, but uses preliminary ("mock") results, based on the analysis of simulated data, as a placeholder for figures, tables, and statistics.
+After the empirical data are analyzed, the Results are updated, and the Discussion is written.
 
@@
 Scientific organizations and funding bodies increasingly demand transparent sharing of research artifacts, and researchers are increasingly willing to do so.
-<!---
-I am not shure about the word artifact in general. Is it common terminology?
--->
 However, although the sharing of such artifacts is a necessary condition for reproducibility, it is not a sufficient one.
@@
 An easy-to-use standard, on the other hand, is more likely to be adopted.
-Our approach is to promote broad adoption of easy-to-use practices (a "low threshold"), and ensure that these practices are compatible with more complex rigorous solutions (a "high ceiling").
+Our approach is to promote broad adoption of easy-to-use practices, i.e., a "low threshold", and ensure that these practices are compatible with more complex rigorous solutions, i.e., a "high ceiling".
 As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
@@
 
-1.	**Multiple inconsistent versions of code, data, or both**; for example
-  - the data set could have changed over time because outliers were removed at a later stage or an item was later recoded
-  - the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
-It may then be unclear which version of code and data was used to produce some reported set of results.
+1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
 2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
@@
 
-We have developed a workflow that achieves long-term and cross-platform computational reproducibility of scientific data analyses.
-It leverages established tools and practices from software engineering and rests on four pillars that address the aforementioned causes of non-reproducibility: [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph.
+We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term and cross-platform computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph.
 The opening paragraph should at least:
@@
 We recommend GitHub to host Git repositories because of its popularity among R users.
-GitHub has many tools that ease project management and collaboration, and these tools are valuable in scientists everyday work, but they are not central to achieving reproducibility.
+GitHub has many tools that ease project management and collaboration, and these tools provide much value in our everyday work, but they are not central to achieving reproducibility.
 
@@
 In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
-Typically, results are manually copied and pasted from one software another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
+Typically, results are manually copyied and pasted from one software to the other, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
 
@@
 It is highly fool-proof [unlike, for example, LaTeX @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
+We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your résumé [@vitae].
 R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
 Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
-This allows the creation of formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
+This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
 @xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.
@@
 However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
-In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
+In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambigously documented.
 
 To automatically resolve dependencies between project files, we rely on **dependency tracking**.
-In essence, researchers provide a collection of computational instructions (_recipes_) that describe how ingredients are processed to create intermediate products and, in the end, a final product.
+In essence, researchers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
 Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
-In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
+In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, analysis results).
 
 We recommend using **Make** for dependency tracking because it is language independent.
-The following hypothetical example illustrates the utility of Make.
+The following hypothesical example illustrates the utility of Make.
 Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
@@
 Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
-One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via **software management**.
+One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
 One comprehensive approach to software management is containerization.
@@
 It provides all software used in an analysis, including the operating system itself.
-(The operating system level is important because some functionality may be passed to software layers beneath the programming language, for example random number generators and linear algebra libraries.)
+The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
 Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
@@
 The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
-A second strategy is to save a container recipe, a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
+A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
 Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
 However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
-For proper archival, we therefore recommend storing a complete image of the software environment besides the Dockerfile.
+For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the Dockerfile.
 A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
 
-To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)); dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
+To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
 While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
@@
 
-The above mentioned tools require much expertise that is beyond the traditional curriculum of researchers.
-To aid researcher adopting these tools we rely on or have developed helpers that wrap more complicated (and powerful) tools.
-These wrappers use standard tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
-We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
-These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
-
-Before we can start creating reproducible research projects, we need a computer with a working R Studio installation. 
-Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so that it can act as an assistance systems while you follow the tutorial:
+One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be time-intensive.
+To lower the threshold to adoption, the R-package `repro` introduces helper functions that wrap more complicated (and powerful) tools.
+These so-called wrappers guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user should do in layman's terms.
+We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
+These wrappers are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
+
+This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
+It describes basic steps that we expect to apply to all users.
+Your specific situation might involve additional, more specialized steps.
+After completing the tutorial, you should be able to customize your workflow accordingly.
+The first step is to install the required software.
+Start R Studio<!--CJ I am concerned that new users may not install R if you only mention RStudio.--> and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
+It will assist you while you follow the tutorial.
 
@@
 # repro is not on CRAN yet
-install.packages('repro', repos = 'https://aaronpeikert.r-universe.dev')
+options(
+  repos = c(aaronpeikert = 'https://aaronpeikert.r-universe.dev',
+            CRAN = 'https://cloud.r-project.org')
+)
+install.packages('repro')
 ```
@@
 An example of the assistance that `repro` provides are the check functions.
-They verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
+These verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
 
@@
 
-They detect the users' system state and explain if further action is needed.
+These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
 Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:
@@
 
-The messages emitted by `repro` and other packages we present try to take the user by the hand to help them solve problems.
-This tutorial describes steps that we expect to apply to all users; the messages from the packages are adjusted to your specific situation (that is, operating system and the specific problem of missing software) and complement the tutorial in this way.
-Before you continue, we ask you to run the above commands to check Git, Make, and Docker.
-If necessary, follow the instructions of repro to install them until the check functions are all fine.
-
-The first concept we want to introduce is version control with Git.
+The messages from `repro` try to take the user by the hand to help them solve problems.
+These messages are adjusted to your specific operating system and installed dependencies.
+Before you continue, we ask you to run the above commands to check Git, Make, and Docker - both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
+If necessary, follow any instructions presented until all checks pass.
+
+## Implementing version control
+
+Now that your system is set up for this tutorial, we will introduce you to version control with Git.
 Git tracks files within one project folder.
@@
 ```
-
 Now the file `private.md` will not be committed to Git and will not be made public.
 
-After you have excluded sensitive files (like not yet anonymized data files), you can activate Git with:
+Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
+When you are ready, you can begin tracking your files using Git by running:
 
@@
 All snapshots are stored in a local repository in your project folder.
-To back up and share the files, you can publicly upload those changes to a public repository.
-To synchronize your local repository with a remote repository from the GitHub service, issue the following command:
+To back up and/or share the files online, you can push your local repository to a remote repository.
+While you can choose any Git service (like GitLab or BitBucket), we will use GitHub in this tutorial.
+Before you upload your project to Git, you need to decide if you would like the project publically accessable (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
+To upload the project publically to GitHub use:
 
@@
 
-Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the R console.
-
-If you are not yet ready to share your project publicly but you want to synchronize your local repository to a remote repository (for example, as a simple backup service), use:
+To upload it privately:
 
@@
 
-Now that you have created a project that is under version control, we continue with adding dynamic document generation.
-Adynamic document has three elements:
-
-1. Text in a natural language
-2. Executable code 
-3. Metadata
-
-The text of an R Markdown is written in Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
+Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the R console.
+
+## Dynamic document generation
+
+Now that you have created a version controlled project, we will proceed with dynamic document generation.
+A dynamic document has three elements:
+
+1. Text in natural language (prose; e.g., a scientific paper or presentation)
+2. Executable code (e.g., analyses)
+3. Metadata (e.g. title, authors, document format)
+
+R Markdown is a type of dynamic document well-suited to the RStudio user interface.
+The natural language text in an R Markdown file can be formatted using Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
+The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
 The following example serves to illustrate the Markdown syntax. 
@@
 Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For a priori planned analyses, preregistration should reduce "researcher degrees of freedom" therefore, the number of decisions researcher can make that could unduly influence the results.
+For a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
 Preregistrations, however, have a problem.
-Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
-The current best practice to ensure that the preregistration is comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
-The structure these templates provide can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
-Nevertheless, these templates can not remove researchers' degrees of freedom because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
+Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
+The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
+The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
+But these templates can not remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
 
@@
 Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
-We term this writing and publishing code before data collection *preregistration as code (PAC)*.
-PAC has the potential to all but eliminates researchers degrees of freedom that standard preregistration despite the best efforts afford.
+We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
+PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration despite the best efforts afford.
 
@@
 A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
-In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`), we found close to a dozen published articles that included some form of script as part of their preregistration.
+In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
 
@@
 First, preregistrations are plagued by non-reproducibility like any other research outcome.
-@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis, but from these, only 34 (20%) provided enough information to reproduce the power analysis.
-Even researchers who have gone to great lengths in submitting a script in their preregistration sometimes inexplicably fail to reproduce their results.
+@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to reproduce the power analysis.
+Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
 For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
@@
 We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, a PAC removes any ambiguity regarding the translation of a verbal analysis plan into code.
+First, a PAC removes any ambiguity regarding the translation of an verbal analysis plan into code.
 Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
@@
 Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
-Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
+Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.
 
@@
 
-When leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
-To use PAC for study planning, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
+Leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
+To use PAC for study planing, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
 To that end, they should borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
@@
 Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
-One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
-Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
+One particular advantage is the strong foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
+Nevertheless it can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section.
 
 Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
-While relatively uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-When simulating data, researchers have to think about and address data-analytic problems on a conceptual level before making an effort to acquire data [@axelrodAdvancingArtSimulation1997].
+While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
+Simulating data forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
 It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
-Writing incomplete code is technically impossible, or at least researchers would notice if they, e.g. forgot to aggregate the data, and so forth.
+Writing incomplete code is technically impossible or at least researchers would notice if they e.g. forgot to aggregate the data etc.
 
@@
 
-With simulated data, researchers can check if the analysis recovers the known parameters of the simulation.
-When an analysis fails to recover the correct values in idealized conditions with big samples, researchers should not expect this analysis to be effective under realistic conditions.
-Because the simulation captures the researchers' expectations about the real world, it is the best way to assess the effectiveness of the analysis without actually gathering the data.
+With simulated data researchers can check if the analysis is technically complete and can also recover the true parameters of the simulation.
+When an analysis fails to recover the correct values in idealized conditions with big samples researchers should not expect this analysis to work under realistic conditions.
+Because the simulation captures the researchers expectations about the real world conditions it is the best way to assess the effectiveness of the analysis without actually gathering the data.
 
@@
 
-A single simulated dataset already helps to ensure that the plan is viable under certain circumstances.
+A single simulated or synthesized dataset already helps to ensure that the plan is viable under certain circumstances.
 However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
-Monte Carlo simulation lets researchers navigate several trade-offs regarding their resources.
-In the simplest case, they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
+Monte Carlo simulation lets researcher navigate several trade offs regarding their resources.
+In the simplest case they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
 @bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.
-Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
+You may also evaluate how the data analysis copes with expected data complications, e.g.
+non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
 
@@
 
-To write a PAC, researchers should start by drafting a manuscript in the form of an R Markdown.
+To write a PAC researchers should start by drafting a manuscript in form of an R Markdown.
 When they have completed the method section, they should begin writing code.
-We suggest writing three functions for each hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
-At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
+We suggest writing three functions for each hypothesis, one for the planned analysis, one for the simulating the data and one for reporting the results.
+At the heart of a PAC lies the planned analysis so researchers should first draft a function called `planned_analysis()` that captures the description in the method section.
 This function should receive the data and derive relevant results from it.
 
-```{r, eval=FALSE}
+```{r}
 planned_analyis <- function(data){
@@
 
-Because it is difficult to draft such a function without interactively testing it, the researcher should then write a function for simulating data.
-
-```{r, eval=FALSE}
+Because it is difficult to draft such function without interactively test it, researcher should than proceed to write a function for simulating data.
+
+```{r}
 simulate_data <- function(n, effect_size){
-  # 1. warn users that the results are mocked
+  # 1. warn users that the results are "fake"
   # 2. draw `n` samples with `effect_size`
@@
 
-As soon as they have written a first draft of `planned_anaylsis()` and `simulate_data()` they can interactively improve both functions, e.g.
+As soon as they have written `planned_anaylsis()` and `simulate_data()` they can iteratively improve both functions, e.g.
 till `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
@@
 This process can be tedious and complicated.
-If it proves infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
-Of course, researchers must not analyze the data before they write the preregistration.
-However, they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
-
-When the researchers are satisfied with the function `planned_analysis()`, they can think about how they would like to report the analysis results via tables, plots, and text.
+If it proofs infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
+Of course researcher must not analyze the data before they write the preregistration but they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
+
+When the researchers are satisfied with the function `planned_analysis()` they can think about how they would like to report the results of the analysis, via tables, plots and text.
 The implementation of this reporting should be in the function `report_analysis()`.
 
-```{r, eval=FALSE}
+```{r}
 report_results <- function(results){
@@
 
-This function should again accept the output of `planned_analysis()` as input.
+These function again should be accept the output of `planned_analysis()` as input.
 The output of this function should be formatted in Markdown.
-Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
+A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
 Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it generates not only tables but also a straightforward interpretation of the effects as actual prose (e.g.
-it quantifies the size of an effect).
-
-Ideally, these three functions can be composed to create a mock results section, e.g.
-when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` they output a full results section.
-
-To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
-After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()`) and write the discussion.
-
-## An example of Preregistration as Code
-
-We have made the argument that PAC has significant advantages over classical preregistration and have outlined how to implement PAC, but can PAC work in practice?
-To answer this question and help researchers implementing PAC themselves, we have conducted a PAC that can serve as an example.
-We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
-
-> "Is there a mean difference in the personality trait 'Machiavellism' between self-identified females and males?"
-
-Again, we propose a preregistration format that closely resembles a classical journal article, but that utilizes simulated data and dynamic document generation to serve simultaneously as preregistration and final manuscript.
-The complete example is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
-In this section, we show excerpts of this example formatted in monospace and provide some commentary.
-
-```{r, echo=FALSE}
-show <- function(x, wrap = TRUE, cat = TRUE, fence = TRUE){
-  stopifnot(length(x) == 1L)
-  if(wrap)x <- str_c(str_wrap(str_split(x, "\\n\\n")[[1]]), collapse = "\n\n")
-  if(!isFALSE(fence)){
-    if(isTRUE(fence))fence <- "markdown"
-    fence <- str_c("````", fence)
-    x <-
-      str_c(fence,
-            x,
-            "````",
-            sep = "\n",
-            collapse = "\n")
-  }
-  if(cat){
-    cat(x)
-    return(invisible(x))
-  } else {
-    return(x)
-  }
-}
-show_prereg <-
-  function(section,
-           file = "preregistration.Rmd",
-           wrap = TRUE,
-           cat = TRUE,
-           fence = TRUE) {
-    prereg <- readr::read_lines(here::here(file))
-    start <- stringr::str_c("<!--begin-", section, "-->")
-    end <- stringr::str_c("<!--end-", section, "-->")
-    start_pos <- which(prereg == start)
-    end_pos <- which(prereg == end)
-    if (length(start_pos) != 1L || length(end_pos) != 1L)
-      stop("No section '", section, "' found.")
-    excerpt <-
-      str_c(prereg[(start_pos + 1):(end_pos - 1)], collapse = "\n")
-    show(excerpt, wrap = wrap, cat = cat, fence = fence)
-  }
-```
-
-As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary information and literature to understand the rest of the paper.
-This example is drastically shortened for illustration purposes:
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("theory")
-```
-
-After researchers have provided the research question, they proceed to explain how they want to study it.
-For simplicity, we will use already published data that we have not yet analyzed:
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("method")
-```
-
-We choose the following statistical procedure because many researchers may be familiar with it[^caution-mww]:
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("ttest")
-```
+The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
+
+Ideally, these three functions can be composed to create a "fake" results section, e.g.
+when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section.
+
+To summarize researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
+After they gathered the actual data, they replace the simulated data, run `planned_analysis()` and write the discussion.
+
+# Preregistration as Code --- a Tutorial
+
+To guide the reader through a preregistration as code (PAC), we will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
+
+> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"
+
+This example serves only didactic purposes and not to derive substantive claims.
+Nevertheless, everything else is kept as close to a real research question as possible.
+
+At the heart of a PAC lies the planned analysis.
+We suggest writing a main analysis function that captures the full process of testing a research question.
+This function receives a dataset as input and performs all statistical processing. <!--AB say anything about its output? is the Markdown the output?-->
+For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
+As an example of a data-dependent decision, we included a rank transformation if the skewness was greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
 
 [^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
-The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
-
-We then translated this methods section into the following `planned_analysis()` function:
+The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
+
+This function is one possible translation of the verbal description above:
 
@@
 
-This function illustrates two advantages of PAC.
-First, a PAC can easily include data-dependent decisions creating different branches of analysis under different conditions.
-Second, it highlights how difficult it is to describe a statistical analysis precisely.
-The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and their expectations and beliefs.
+Importantly, natural language descriptions do not map unambiguously onto computer code implementations. 
+The same textual descriptions may be implemented differently by different persons depending on their statistical and programming knowledge but also based on their own expectations and beliefs.
 One example would be using the function `wilcox.test` instead of the combinations of `rank` and `t.test`.
-Both are valid implementations of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
-In contrast, the second applies Welch's correction and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
+Both are valid implementation of the Mann–Whitney–Wilcoxon test but the first assumes equal variance whereas the second applies Welch's correction and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
 Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
-Still, these details can make a difference and provide unwanted researchers' degrees of freedom.
-
-The function `planned_analysis()` together with the function `simulate_data()` (not shown here) can be used to justify the planned sample size.
-To that end, `simulate_data()` is repeatedly called with increased sample size and the proportion of significant results (power) is recorded.
-The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
-The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R).
-
-Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
+Still, these details can make a difference and provide unwanted researcher degrees-of-freedom.
+
+<!--AB here is kind of a break in the text-->
+To use this function one needs a compatible dataset. <!--AB what is "this" function? What does "compatible" mean?-->
+At the time of preregistration there is of course no actual data available.
+We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
+Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
+The following code simulates a dataset with one observed variable in two groups.
+
+```{r simulate_data}
+```
+<!--AB it is odd that the return variable is called effect because it is a dataset-->
+
+Here, we decided to randomly draw non-normal data to test how well our procedure performed under non-optimal situations.
+Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a given standardized effect size (a Cohen's d of .2).
+This simulated dataset fits perfectly into our planned analysis with a sample size of 400 and 10 items: <!--AB in what way? please explain-->
+
+***
+
+```{r}
+set.seed(112358) # fix state of RNG
+results <- 
+  planned_analysis(
+    simulate_data(
+      n = 400,# Sample size
+      df = 8, # Chisq with DF = 8 has skewness = 1
+      d = .2, # Cohen's D
+      i = 10  # Number of items (should not effect anything)
+    )
+  )
+```
+
+Next, we add dynamic code blocks to our Markdown report that implement the planned reporting of the final results.
+A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
+Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
+The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
+Here we use the `report` package to create a preview of how our reporting will look like based on simulated data:<!--AB add citation to the report package-->
+
+```{r, results='asis', eval=FALSE}
+report::report(results$test)
+```
+
+This call to the `report()` function will yield the following automatically generated output:
+
+```{r, results='asis', echo=FALSE}
+report::report(results$test)
+```
+<!--AB it is difficult to discern the automatically generated text and out manuscript text even though there is the horizontal line. Not sure how to improve this.-->
+***
 
@@
 
-```{r, include=FALSE}
-fig_caption <- str_c("Results of simulation for the power analysis. The cross indicates the sample size that archives ", choosen_power*100, "% assuming a Cohen's D of ", choosen_d, ".")
-```
-
-```{r power, echo=FALSE, fig.cap=fig_caption}
+To determine a minimal sample size for this study we repeat this simulation across several combination of effect sizes and sample sizes.
+One standard recomendations for choosing a sample size is to have an expected power of 80% based on the expected effect size.
+@Jones2013 found in three studies gender differences of .24, .29, and .35 for their machiavilism subscale.
+If we want to detect an effect of size `r choosen_d` with an `r choosen_power*100`% probability, our simulation shows that we require at least `r minn$n` participants.
+
+<!--AB please add a figure number and a proper citation to the figure. Also, tell people where they can find the code to run the power analysis themselves-->
+
+```{r, echo=FALSE}
 res %>%
@@
 
-```{r, echo=FALSE, results='asis'}
-show_prereg("power-code", wrap = FALSE)
-```
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("power")
-```
-
-We implemented a mechanism that only uses simulated data when the actual data is not yet available for the results section.
-This mechanism also warns readers if these results are based on simulated data.
-As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study.
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("results", wrap = FALSE)
-```
-
-The last thing we need to do is to report our results with the combination of `planned_analysis()` and `report_analysis()`.
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("report", wrap = FALSE)
-```
-
-This is an example of how the results could be reported (based on simulated data):
-
-```{r, echo=FALSE, results='asis'}
-set.seed(1234)
-show(report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)), cat = FALSE), fence = "")
-```
-
-The interested reader can find the example rendered with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
-
 # Discussion
@@
 
-PACs come with a challenge.
+PACs come with a challenge though.
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
@@
 Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
+Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
 
@@
 
-Please note that the statistical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for addressing your research question.
+<!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->
+Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
 The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-
-We also suggest that version control with Git and GitHub fulfills the criteria of a preregistration as long as you state your research plan comprehensively.
-This view is not shared by everyone [@szollosi_is_2020] because with some effort and malice intend the record of Git/GitHub can be altered.
-This concern can be addressed by using zenodo.org to give a digital object identifier (DOI) to your preregistration (see [this guide](https://guides.github.com/activities/citable-code/)) because zenondo.org saves an unchangeable copy of your repository.
+<!--AB can we move this to a different place?
+Maybe this would actually better fit to the discussion/limitation section!-->
 

### 2021-07-28 manuscript.Rmd

@@
 
-Using markup for formatting, instead of the visual approach used in, e.g., Microsoft Word, is compatible with many output formats and is easier to put under version control.
-Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
-The visual editor has the additional benefit of promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions).
-Among other features, it eases the generation of citations and cross references.
-
-Now that you are familiar with the basics of writing text in Markdown, we turn our attention to including code into the text.
+One advantage of this type of markup for formatting is that it can be rendered to many different output formats - both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g.
+specific journal requirements.
+Another advantage is that, because R Markdown files are in plain text, they are more suitable for version control using Git.
+Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass <!--CJ Add picture-->in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
+The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
+and easing the generation of citations and references to Tables and Figures.
+
+Now that you are familiar with the basics of Markdown formatting, we turn our attention to including code into the text.
 Code is separated by three backticks (and the programming language in curly braces) like here:
@@
 
-This information makes internal dependencies to other files in the project and external dependencies to R packages explicit.
-`repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
-These two files together form the basis for consistency within a research project and consistency across different systems.
-At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.
+This information makes explicitly declares the project's dependencies on files and R packages.
+`repro` uses this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
+These two files together form the basis for consistency within a research project and consistency across different systems.<!--CJ Consider explaining this in the theoretical intro-->
+At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.<!--CJ This comes way to late.
+Shouldbe in the opening paragraph or something.-->
 This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
-Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).
-In everyday work, the simplest solution is to now call the function `repro::automate()` which both creates a Makefile and a Dockerfile from all R Markdown files in the project (all files with the ending `.Rmd`) based on the respective repro metadata entries.
-Every time you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
-The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
+Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).<!--CJ These points describe the formal definition of reproducibility used in this paper, and the goal of the workflow.
+That needs to be explained in the introduction.
+If you only find out what you are working toward halfway through the tutorial - that's too late!-->
+
+The function `repro::automate()` converts the metadata from all R Markdown files in the project (all files with the ending `.Rmd`) to a Makefile and a Dockerfile.
+Together, these files allow users (including your future self) to reproduce every step in the analysis automatically.
+It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, to keep the Makefile and Dockerfile up to date.
+The function `repro::automate()` does not overwrite manual changes<!--CJ What does that mean?-->, and there is no harm in running it too often.<!-- CJ Can that not be automated?
+Or at least provide a suggestion when to run it?-->
 Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
 <!--AB: why is this addition not automated?
-Feels strange given that this is our recommended best practice?!-->
-
+Feels strange given that this is our recommended best practice?!
+CJ: Agreed.-->
+
+<!-- CJ What i'm missing here is something to wrap up the tutorial. Do users now have a project that could be reproduced? The tutorial should not end until that is the case.-->
+
+## Reproducing a project
+
+<!--CJ Suggestion: Instead of explaining reproduction right in the middle of the tutorial for setting up a project, add a later section called "reproducing a project", and provide an example project for users to reproduce.-->
 When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:
@@
 
-Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
+Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.<!--CJ This belongs in the intro, not the tutorial.
+It is part of explaining the benefits of repro.-->
 These projects will remain reproducible irrespective of whether `repro` will be maintained and remain accessible in future.
 Users do not have to have repro installed to reproduce a project, in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
-The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.
+The only required software for reproducing a project is Docker, assuming users know how to build a Docker image and run Make within the container.
+However, if they have installed Make additionally to Docker, they do not need to know how to use Docker and can simply rely on the two Make commands `make docker` and `make -B DOCKER=TRUE`.
 The commands can be pasted into a terminal and will reproduce the project in many cases.
-<!--AB: I said this before.
-I think it really should be possible to execute this command from within R.
-Never underestimate the troubles people can have with the terminal.
-I think we discussed this before.
-Please add a function repro::rerun(execute=TRUE) or -- if we rename it -- repro::reproduce(execute=TRUE).
--->
 
@@
 
-Some projects, however, are not as easy to make reproducible.
-As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
-These simplifications are possible because of two restrictions.
-Users who ask themselves either "How can I install software that is not an R package in the Docker image?" or "How can I express complex dependencies between files, e.g.
-raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
-All other users may safely skip this section and may return to it if they become limited by restriction.
+As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.
+The simplification of Docker and Make are possible because of two restrictions.
+Users who ask themselves either "How can I install software dependencies outside of R in the Docker image?" or "How can I express complex dependencies between files, e.g.
+hundreds of data files are preprocessed and combined?") need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
+Other users may safely skip this section, or return to it if they encounter such challenges.
 

### 2021-07-28 manuscript.Rmd
useres can skip advanced features

--- 2021-07-28 manuscript.Rmd
@@
 
-The above mentioned tools require expertise that is beyond the traditional curriculum of researchers.
-To aid researchers adopting these tools we rely on and have developed helpers that wrap more complicated (and powerful) tools.
-These helpers wrap standard tools but provide feedback about what the computer does and what the user should do in layman's terms and are accessible from within R.
-We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
-These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
-
-Before we can start creating reproducible research projects, we need a computer with a working R Studio installation. 
-Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so it can assist you following the tutorial:
+One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be time-intensive.
+To lower the threshold to adoption, the R-package `repro` introduces helper functions that wrap more complicated (and powerful) tools.
+These so-called wrappers guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user should do in layman's terms.
+We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
+These wrappers are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
+
+This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
+It describes basic steps that we expect to apply to all users.
+Your specific situation might involve additional, more specialized steps.
+After completing the tutorial, you should be able to customize your workflow accordingly.
+The first step is to install the required software.
+Start R Studio<!--CJ I am concerned that new users may not install R if you only mention RStudio.--> and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
+It will assist you while you follow the tutorial.
+>>>>>>> main
 
@@
 An example of the assistance that `repro` provides are the check functions.
-They verify that you indeed have installed and setup the required pieces of software for this workflow (we use the double-colon operator to access functions from the `repro` namespace to make clear which functions are native `repro` functions):
+
+These verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
 
@@
 
-You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore` file.
-You can edit the `.gitignore` file directly or with this command:
+Git tracks any files that you manually add to the repository.
+To make sure you do not accidentally add files that you do not wish to share (e.g., privacy sensitive data),
+you can list specific files that you do not want to track in the `.gitignore` file.
+You can add something to the `.gitignore` file directly or with this command:
 
@@
 
-After you have excluded sensitive files (like not yet anonymized data files), you can activate Git with:
+Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
+When you are ready, you can begin tracking your files using Git by running:
 
@@
 For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
-The easiest way to do this is to use the RStudio Git pane.
-Click on the empty box next to the file you want to stage, and when a tick appears, which means that the file is staged.
-After you have staged all files you would like to, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
+
+One way to do this is through the visual user interface in the RStudio Git pane.
+Click on the empty box next to the file you want to stage.
+A check mark appears which indicates that the file is staged.
+After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
+This stores a snapshot of the current state of the project.
 
@@
 
-Now that you have created a project that is under version control, we continue with adding dynamic document generation.
+Depending on your computer configuration, this may ask to set up a secure connection to GitHub.
+In this case first follow the suggestions shown on the R console.
+
+## Dynamic document generation
+
+Now that you have created a version controlled project, we will proceed with dynamic document generation.
 A dynamic document has three elements:
 
-1. Text
-2. Executable code
-3. Metadata
-
+1. Text (prose; e.g., a scientific paper or presentation)
+2. Executable code (e.g., analyses)
+3. Metadata (e.g. title, authors, document format)
+
+R Markdown is a type of dynamic document well-suited to the RStudio user interface.
 The text of an R Markdown is formatted by Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
+The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
 The following example serves to illustrate the Markdown syntax.
+
 It shows how to create a heading, a word in bold font, and a list of several items in Markdown:
@@
 
-This type of markup for formatting, instead of the more visual approach of e.g. Microsoft Word, is compatible with many output formats and is easier to put under version control.
+One advantage of this type of markup for formatting is that it can be rendered to many different output formats - both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g.
+specific journal requirements.
+Another advantage is that, because R Markdown files are in plain text, they are more suitable for version control using Git.
 Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
-The visual editor has the additional benefit of promoting best practices (for example, each sentence should be written on a new line, which eases tracking changes across versions).
-Among other features, it eases the generation of citations and cross references.
-
-Now that you are familiar with the basics of writing text in Markdown, we turn our attention to including code into the text.
-Code is separated by three backticks (and the programming language in curly braces):
+The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
+and easing the generation of citations and references to Tables and Figures.
+
+Now that you are familiar with the basics of Markdown formatting, we turn our attention to including code into the text.
+Code is separated by three backticks (and the programming language in curly braces) like here:
 
@@
 
-The results of your code are dynamically inserted into the document.
-Dynamically means that whenever you turn the R Markdown into its output format the code is executed again and if you changed something the results are updated.
-The output document is then static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
-This static document is somewhat decoupled from the R Markdown and the rest of the code that did create it.
-It is quite possible that several slightly different versions are therefore circulating.
-We therefore recommend to connect the static document to the version control system by dynamically including the identifier of the Git commit, called hash, in the document.
-The `repro` package comes with the function `repro::current_hash()` for that purpose.
+The hotkey [Control]+[Alt]+[i] inserts a block of code in the file.
+The results of code enclosed in such backticks will be dynamically inserted into the document (depending on specific settings).
+This means that, whenever you render the R Markdown to its intended output format, the code will be executed and the results updated.
+The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
+
+Once the R Markdown file has been rendered to a static document, the resulting file is decoupled from the R Markdown and the code that created it.
+This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
+To avoid ambiguity, we therefore recommend referencing the identifier of the Git commit at the time of rendering in the static document.
+Simply put, a static document should link to the version of the code that was used to create it.
+The `repro` package comes with the function `repro::current_hash()` for this purpose.
 This document was created from the commit with the hash ``r repro::current_hash()``.
 
-Now that you know how to write text and R code in an R Markdown, you need to know about metadata, like the title or the output format.
+Now that you know how to write text and R code in an R Markdown, you need to know about metadata (also called: YAML front matter).
+These metadate contain information about the document, like the title and the output format.
 Metadata are placed at the beginning of the document and are separated by three dashes from the document body.
-The following example is a full markdown document where the the metadata (the "header") is in lines 1--6.
+The following example is a full markdown document where the metadata (the "YAML front matter") is in lines 1--6.
 
@@
 
-Most metadata fields are self-explanatory (like the author field). 
-Sometimes, R packages may extend the metadata fields for their own purposes.
-Certain output formats may require specific  metadata. 
-For example, the `rticles`-package, which offers templates for authoring scientific manuscripts, comes with extra fields that are described in the package documentation and offer extra options for flexible formatting of documents.
-R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
-The `repro` package uses the metadata for hints to make the research project reproducible.
-Researchers should list all dependent resources of their research project in the metadata, that is, R scripts, data files, and external packages, in the following format (see everything below the line *repro:*):
+Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
+However, R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
+Some R packages add additional metadata fields.
+The `repro` package adds fields to the metadata to help make the research project reproducible.
+These fields are used to list all dependencies of the research project.
+This includes R scripts, data files, and external packages.
+The format is as follows (see everything below the line *repro:*):
 
@@
 
-This information makes internal dependencies to other files in the project and external dependencies to R packages explicit.
-`repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
-These two files together form the basis for consistency within a research project and consistency across different systems.
-At this point, we would like to emphasize that a prerequisite for reproducibility is that every step taken in an analysis can be **automatically** reproduced.
+This information makes explicitly declares the project's dependencies on files and R packages.
+`repro` uses this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
+These two files together form the basis for consistency within a research project and consistency across different systems.<!--CJ Consider explaining this in the theoretical intro-->
+At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.<!--CJ This comes way to late.
+Shouldbe in the opening paragraph or something.-->
 This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
-Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).
-
-In everyday work, the simplest solution is to now call the function `repro::automate()` which creates a Makefile and a Dockerfile from all R Markdown files in the project based on the respective repro metadata entries.
-Every time you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
-The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
+Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).<!--CJ These points describe the formal definition of reproducibility used in this paper, and the goal of the workflow.
+That needs to be explained in the introduction.
+If you only find out what you are working toward halfway through the tutorial - that's too late!-->
+
+The function `repro::automate()` converts the metadata from all R Markdown files in the project (all files with the ending `.Rmd`) to a Makefile and a Dockerfile.
+Together, these files allow users (including your future self) to reproduce every step in the analysis automatically.
+It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, to keep the Makefile and Dockerfile up to date.
+The function `repro::automate()` does not overwrite manual changes<!--CJ What does that mean?-->, and there is no harm in running it too often.<!-- CJ Can that not be automated?
+Or at least provide a suggestion when to run it?-->
 Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
 <!--AB: why is this addition not automated?
-Feels strange given that this is our recommended best practice?!-->
-
-When you or someone else who received your project would like to reproduce it, they can ask repro to explain how this should be done:
+Feels strange given that this is our recommended best practice?!
+CJ: Agreed.-->
+
+<!-- CJ What i'm missing here is something to wrap up the tutorial. Do users now have a project that could be reproduced? The tutorial should not end until that is the case.-->
+
+## Reproducing a project
+
+<!--CJ Suggestion: Instead of explaining reproduction right in the middle of the tutorial for setting up a project, add a later section called "reproducing a project", and provide an example project for users to reproduce.-->
+When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:
 
@@
 
-Some projects, however, are not so easy to make reproducible.
-As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
-These simplifications are possible because of two restrictions.
-Users who ask themselves either "How can I install software that is not an R package in the Docker image?" or "How can I express complex dependencies between files, e.g.
-raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
-All other users may safely skip this section and return if they need to.
+This section is only for advanced users who want to overcome some limitations of `repro`.
+If you read this paper the first time you probably want to skip this section.
+As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.
+The simplification of Docker and Make are possible because of two restrictions.
+Users who ask themselves either "How can I install software dependencies outside of R in the Docker image?" or "How can I express complex dependencies between files, e.g.
+hundreds of data files are preprocessed and combined?") need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
+Other users may safely skip this section, or return to it if they encounter such challenges.
 

### 2021-07-28 manuscript.Rmd
Comments on abstract
Half the abstract focused on PAC; that does not accurately reflect the content of the paper. I've attempted to address that. Still needs a powerful concluding message!

--- 2021-07-28 manuscript.Rmd
@@
 bibliography: temp.bib
+appendix: appendix.tex
 simplesummary: |
-  A Simple summary goes here.
+  Reproducibility has long been considered integral to the scientific method.
+  Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
+  Technological advancements now enable scientists to achieve a much higher standard of reproducibility;
+  one in which any individual can be granted access to a digital research repository, and reproduce the analyses from raw data to final manuscript with a single command.
+  This method has far-reaching implications for scientific archiving,
+  reproducibility and replication,
+  scientific productivity,
+  and the credibility and reliability of scientific findings.
+  One obstacle preventing the widespread adoption of this method is that the underlying technological advancements
+  are complicated to use.
+  This paper introduces `repro`; an R-package that guides researchers in the installation and use of the tools required for computational reproducibility.
+  It includes a tutorial on how to make a research project reproducible with the aid of `repro`.
+  Finally, it introduces novel applications of reproducibility tools for scientific research;
+  including the preregistration of study plans as reproducible computer code.
+  This circumvents the shortcomings of ambiguous preregistrations that allow for researcher degrees of freedom,
+  because computer code describes study plans more precisely than prose.
+  <!-- CJ Add concluding sentence(s) -->
 abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
@@
 acknowledgement: |
-  The authors received no financial support for the research, authorship, and/or publication of this article.
+  All sources of funding of the study should be disclosed. Please clearly 
+  indicate grants that you have received in support of your research work. 
+  Clearly state if you received funds for covering the costs to publish in open 
+  access.
 authorcontributions: |
-  Aaron Peikert took the lead in writing and provided the initial draft of the manuscript.
-  Andreas Brandmaier and Caspar J. Van Lissa contributed further ideas, critical feedback, and revisions of the original manuscript.
-  Furthermore, we would like to thank Maximilian Stefan Ernst (not an author) for his contributions to the code for the simulation study.
+  For research articles with several authors, a short paragraph specifying their 
+  individual contributions must be provided. The following statements should be 
+  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
+  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
+  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
+  limited to those who have contributed substantially to the work reported.
 conflictsofinterest: |
@@
 abbreviations:
-  - short: PAC
-    long: Preregistration as 
-  - short: Gb
-    long: Gigabyte
-  - short: Kb
-    long: Kilobyte
-  - short: CRAN
-    long: Comprehensive R Archive Network
+  - short: GUI
+    long: Graphical User Interface
+  - short: CLI
+    long: Comand Line Interface
 repro:
@@
     - gert
-    - aaronpeikert/repro@c3cae1f
+    - aaronpeikert/repro@7bfaf98
     - here
@@
 
-Scientists increasingly strive to make research data, materials, and analysis code openly available.
-Sharing these types of digital research artifacts has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
-To some extent these potential benefits are contingent on whether these digital research artifacts are reproducible. 
-Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
+Increasingly, scientists openly share their data, materials, and analysis code.
+Sharing the digital research output increases scientific efficiency by enabling researchers to learn from each other, reuse materials, or review published research results, but only if these artifacts are reproducible.
+The technology is now available to make all artefacts related to a research project available so that the project can be reproduced with minimal effort.
 Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
 22][^popper]
-Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
+In contemporary research, reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
+Unfortunately, despite increasing commitment to open science practices, many projects can not yet be reproduced by other research teams.
 This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
-We now have the technology to make all artefacts related to a research project available so that a research project can be reproduced with minimal effort.
 
@@
 We pursue two aims.
-First, the present paper introduces the reader to a workflow that maximizes the potential reproducibility of a research project.
-This workflow builds on several technical solutions originating in the software engineering community. 
-We discuss how these solution address four common threats to reproducibility [see @Peikert2019 for details] and form a coherent workflow.
-We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
-The second objective is to discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
-One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
-PAC involves preregistering the intended analysis code and the bulk of the manuscript as dynamic document, including the Introduction, Methods, and Results sections.
-The resulting dynamic document closely resembles the final manuscript, but uses preliminary ("mock") results, based on the analysis of simulated data, as a placeholder for figures, tables, and statistics.
-After the empirical data are analyzed, the Results are updated, and the Discussion is written.
+First, the present paper introduces the reader to a workflow that ensures reproducibility by building on several technical solutions originating in the software engineering community that eliminate four common threats to reproducibility [see @Peikert2019 for details].
+Specifically, we demonstrate how the R package `repro` supports researchers in implementing these solutions.
+The second objective is to discuss opportunities of such workflow to improve not only how research outputs are created and recreated but also how research itself is conducted.
+One such opportunity is an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the complete analysis as code and the bulk of the manuscript as dynamic document, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.
+This dynamic document closely resembles an actual journal article but with "fake" results and without discussion of results, then serves as preregistration.
 
-[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalized to inter-subjective criticism (Footnote *1, p.
+[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalised to inter-subjective criticism (Footnote *1, p.
 22).
@@
 
-Scientific organizations and funding bodies increasingly demand transparent sharing of research artifacts, and researchers are increasingly willing to do so.
-However, although the sharing of such artifacts is a necessary condition for reproducibility, it is not a sufficient one.
-This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
-Out of 35 published articles with open research artifacts, results of 22 articles could be reproduced but in 11 of these cases, further assistance from the original authors was required.
+While the willingness to share materials is required for reproducing a project, it is not sufficient.
+For example, @hardwicke2018 attempted to reproduce results from open materials in the journal Cognition. 
+Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
 For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
-Another study of 62 Registered Reports found that only 41 had data available, and 37 had analysis scripts available [@obels2020].
+@obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
 The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
-These failed attempts to reproduce highlight the need for widely accepted reproducibility standards, because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
-If research artifacts are available but not reproducible, their added value is limited.
+These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
+The results further illustrate that open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
 
-This tutorial therefore demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
-A rigorous standard increases the likelihood that a project will remain reproducible as long as possible.
-An easy-to-use standard, on the other hand, is more likely to be adopted.
-Our approach is to promote broad adoption of easy-to-use practices, i.e., a "low threshold", and ensure that these practices are compatible with more complex rigorous solutions, i.e., a "high ceiling".
-As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
-The first part of this tutorial outlines a reproducibility standard, and introduces the underlying theoretical concepts and software solutions.
-The second part is a tutorial that demonstrates how this standard can be easily adopted with the help of the R package `repro`.
+We argue that the benefits associated with transparency increase if research can additionally be reproduced.
+This paper tutorial demonstrates how to make research projects reproducible, while striking a balance between rigor and ease-of-use.
+A rigorous standard increases the likelihood of any individual project to be actually reproducible.
+An easy to use standard, on the other hand, is more likely to be adopted.
+The remainder of this paper is structured along these lines.
+The first part introduces theoretical concepts and software solutions that together form a rigorous reproducibility standard.
+The second part is a tutorial that shows how these software solutions can be easily employed by researchers with the help of the R package `repro`.
 
-We have structured the tutorial with a *learning by doing* approach in mind, such that readers can follow along on their own computers.
+We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.
 We explicitly encourage readers to try out all R commands for themselves.
@@
 
-# Concepts and Software Solutions<!--CJ Some inconsistencies:
-1) From the end of the previous paragraph, I would expect a description of the "rigorous reproducibility standard" here.
-2) The paragraph is labeled "Concepts and Software SOlutions, but it begins with a list of threats to reproducibility. Threats are not concepts nor software solutions.
-3) After the list of threats, we get a description of the "rigorous reproducibility standard".
-That is what I would expect based on the preceding paragraph.
-Note that the "rigorous reproducibility standard" is also not a concept, nor a software solution.
-Consider re-naming this section.-->
+# Concepts and Software Solutions
 
-From our own experience with various research projects, we identified the following common threats to reproducibility: <!-- CJ This still requires attention!
-How did you identify these causes?
-Is this exhaustive?-->
+We identified the following common threat to reproducibility: <!-- CJ How did you identify these causes? Is this exhaustive?-->
 
-1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
-2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
-4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
+1.	**Multiple inconsistent versions of code, data, or both**; for example, the dataset could have changed over time because outliers were removed at a later stage, or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
+2.	**Copy-and-paste**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
+3.	**Ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order; 
+4.	**External code dependencies**; for example, a given analysis may depend on a specific version of a specific software package that might not be available on a different computer or no longer exist; or, a different version of the same software may produce different results.
 <!-- CJ 5.
 Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
-<!--AB Caspar's 5th point is important.
-In my mind, this best fits to point #3, which adressed an insufficient level of coding/automation/documentation of the necessary steps taken.
-Can we merge this into one new point #3?
-I will commit a suggested change of point #3 here (see above);.-->
 
@@
 1. Version control
-2. Dynamic document generation
+2. Dynamic document creation
 3. Dependency tracking
@@
 
-**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the development phase of a research project.
+First, to resolve ambiguity across multiple versions of code and data that come into existence during the development phase of a research project, we recommend using a **version control** system.
 Version control allows a clear link between which results were generated by which version of code and data.
+A version control system tracks changes to all project-related files (e.g., materials, data, and code) over time.
+At a later stage, single files or the entire project can be compared to or reverted to an earlier version.
+A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
+
 We recommend using **Git** for version control, because of its widespread adoption in the R community.
-
-Git tracks changes to all project-related files (e.g., materials, data, and code) over time.
-At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
-A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
-Git is built around snapshots that represent the project state at a given point in time.
-Those snapshots are called "commits" and work like a "save" action.
-Changes to files are collected in a _commit_.
-Ideally, each commit has a message that succinctly describes these changes.
-It is good practice to make commits for concrete milestones, e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3".
-This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
+Git is built around snapshots that represent the entire project state at a given point in time.
+Those snapshots are called "commits".
+Each file that was changed has to be added to a commit, and each commit requires a commit message that succinctly describes the changes made.
 Each commit "knows" its ancestor, and they form a timeline together.
-The entirety of commits (i.e. the version-controlled project) is called a repository.
-In Git, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
+The conglomerate of commits (i.e. the version-controlled project) is called a repository.
+In Git, specific snapshots of a repository can be tagged, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
 Git has additional features beyond basic version control, such as branches to facilitate collaboration.
-@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git for research.
-@bryanExcuseMeYou2018 provides additional information on how to track R Markdown documents.
-Collaborating via Git requires the repository to be uploaded somewhere.
+To collaborate via Git,<!--dangling modify, cant help it--> the repository has to be uploaded somewhere.
 We recommend GitHub to host Git repositories because of its popularity among R users.
-GitHub has many tools that ease project management and collaboration, and these tools provide much value in our everyday work, but they are not central to achieving reproducibility.
+GitHub has many tools that ease project management and collaboration, and these tools provide much value in our everyday work, but they are not central to archiving reproducibility.
 
@@
 In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
-Typically, results are manually copyied and pasted from one software to the other, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
+Both steps are linked by manually copying and pasting results from one software to the other, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
 
-Dynamic document generation integrates both steps.
-Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allows interested readers deeper understanding [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
+Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured, who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
 **R Markdown** uses Markdown for writing the text and R (or other programming languages) for the analysis.
-Markdown is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
+Markdown is a lightweight text format in plain text with minimal reserved symbols for formatting instructions.
 This way, Markdown does not need any specialized software for editing.
-It is highly fool-proof [unlike, for example, LaTeX @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your résumé [@vitae].
-R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
+It is highly fool-proof (unlike, for example, LaTeX), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+We suggest using Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your CV [@vitae].
+R Markdown extends regular Markdown by allowing users to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
 Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
 This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
-@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.
 
-One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
-However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
-In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambigously documented.
+One would hope that sharing a literate statistical analysis would allow others (including one own's future self) to simply download and run (ie., reproduce) a given analysis.
+However, while version control and dynamic document generation are becoming more common @, this approach fails to guarantee reproducibility in most instances.
+In practice, dependencies between project files (e.g.
+the raw data was preprocessed) or on external software (such as system libraries or components of the programming language) are frequently unmentioned.
 
@@
 We recommend using **Make** for dependency tracking because it is language independent.
-The following hypothesical example illustrates the utility of Make.
-Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
-A Makefile for this project could look like this:
+To make this more concrete, consider the following hypothetical example, in which a research project contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
+A Makefile could then look like this:
 
 ```{bash, eval=FALSE}
-manuscript.pdf: manuscript.Rmd simulated_data.csv
+manuscript.pdf: manuscript.Rmd simulation_results.csv 
   Rscript -e 'rmarkdown::render("manuscript.Rmd")'
 
-simulated_data.csv: simulate.R
+simulation_results.csv: simulate.R
   Rscript -e 'source("simulate.R")'
@@
 If a requirement does not exist, Make uses a recipe to build the requirement first before building the target.
-Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
-This ensures that the simulation is always run before the manuscript is built, and that the manuscript is rebuilt if the simulation code was changed.
-Make therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
+Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would make sure that the file `simulation_results.csv` exists; if not, it would find that this is also a target and issue the command in l.5 to run the simulation first before rendering the manuscript.
+This way, we ensure that the simulation is always run before the manuscript is built and that the manuscript is rebuilt if the simulation code was changed.
+
+A version controlled dynamic document with dependency tracking still relies on external software.
+Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
 
@@
 
-A version controlled dynamic document with dependency tracking may still rely on external software.
-Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
-One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
+One solution to the problem of external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
 One comprehensive approach to software management is containerization.
-"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p.
-174].
-
-**Docker** is an example of containerization.
-It provides all software used in an analysis, including the operating system itself.
+"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silverSoftwareSimplified2017].
+Instead of leaving it to the user which software should be installed how on their computer, **Docker** provides all software used in an analysis, including the operating system itself.
 The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
 Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
-Researchers therefore have the opportunity to use the exact same software setup that was used by the original authors.
 Such a snapshot of the software stack is called an "image".
-Packaging all required software in such an image requires considerable amounts of storage space.
-Two major strategies help to keep the storage requirements reasonable.
-One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
-For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
+Packaging all needed software in such an image requires considerable amounts of storage space.
+Two major strategies are helping to keep the storage requirements reasonable.
+First, there is a community that maintains images for particular purposes.
+For example, there are ready-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
 Users can then install whatever they need in addition to what is provided by these pre-compiled images.
-`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
-The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
-A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
-Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
-However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
-For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the Dockerfile.
-A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
+`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
+Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
+Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
+This recipe is a textual description of all commands that need to be executed to recreate the image.
+Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
+However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
+For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
 
-To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
-While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
-To remove this hurdle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
-In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.
+To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), containerization (using Docker).
+While R Markdown and git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside studies of information technology, which represents a considerable hurdle in the acceptance and implementation of the workflow.
+To remove this hurdle, we have developed a new R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
+In the remainder, we will walk you through a complete research project, from sketching the very first idea over preregistration, data analysis, and submission of a preprint while focusing on how the `repro`-package enables users to maintain reproducibility of all steps.
 
@@
 
-# Creating reproducible research projects
+# Tutorial
 
-One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be time-intensive.
-To lower the threshold to adoption, the R-package `repro` introduces helper functions that wrap more complicated (and powerful) tools.
-These so-called wrappers guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user should do in layman's terms.
-We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
-These wrappers are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
-
-This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
-It describes basic steps that we expect to apply to all users.
-Your specific situation might involve additional, more specialized steps.
-After completing the tutorial, you should be able to customize your workflow accordingly.
-The first step is to install the required software.
-Start R Studio<!--CJ I am concerned that new users may not install R if you only mention RStudio.--> and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
-It will assist you while you follow the tutorial.
+A significant hurdle for adopting the before-mentioned tools for researchers is that they are typically used through a command-line interface (CLI).
+The CLI predates and coexists with graphical user interfaces (GUIs) which many users (including the authors) prefer over the text-based interactions of a CLI for most tasks.
+A GUI has the apparent disadvantages for reproducibility that it requires a user who clicks on the correct buttons in the correct order.
+Depending on the number of steps, this is not only much effort but also error-prone.
+This tutorial seeks a compromise.
+We use simpler alternatives, including a GUI where we can, to create a reproducible workflow, but the reproduction of such workflow does not rely on manual steps.
+At several points, we rely on or have self-developed a CLI that wraps a more complicated (and powerful) CLI.
+These wrappers use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
+We hope this makes reproducibility tools more accessible by enabling untrained users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
+These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
+Please install the package [`gert`](https://github.com/r-lib/gert)[@R-gert] which wraps Git commands and [`repro`](https://github.com/aaronpeikert)[@R-repro] so they can act as assistance systems while you follow the tutorial:
 
 ```{r, eval=FALSE}
+install.packages("gert")
 # repro is not on CRAN yet
-options(
-  repos = c(aaronpeikert = 'https://aaronpeikert.r-universe.dev',
-            CRAN = 'https://cloud.r-project.org')
-)
-install.packages('repro')
+install.packages('repro', repos = 'https://aaronpeikert.r-universe.dev')
 ```
 
-An example of the assistance that `repro` provides are the check functions.
-These verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
+One example for such assistance are the check functions from `repro`.
+They verify that you indeed have installed and setup the required software for this workflow:
 
@@
 
-These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
-Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:
+They detect the users' system state and explain what this means to the user.
+Sometimes they ask the user to do something, e.g. the following happens if you were a windows user who does not have installed Git:
 
@@
 
-The messages from `repro` try to take the user by the hand to help them solve problems.
-These messages are adjusted to your specific operating system and installed dependencies.
-Before you continue, we ask you to run the above commands to check Git, Make, and Docker - both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
-If necessary, follow any instructions presented until all checks pass.
+The messages emitted by `repro` and other packages we present try to take the user by the hand to help them solve problems.
+This tutorial describes steps that we expect to apply to all users; the messages from the packages fill the gap to your specific situation and extend the tutorial in this way.
+Before you continue, we ask you to run the above commands to check Git, Make, and Docker.
+If necessary, follow the instructions of repro to install them.
 
-## Implementing version control
-
-Now that your system is set up for this tutorial, we will introduce you to version control with Git.
+The first concept we want to apply practically is version control with Git.
 Git tracks files within one project folder.
-We start by creating such a project folder with RStudio by clicking the menu item:
+We start by creating such project folder with RStudio:
 
@@
 
-This creates a project containing an example analysis, but you may use any other template or existing R Project to turn into a reproducible research project using the functionality of `repro`.
+This creates a project containing an example analysis, but you may use any other template or existing R Project.
 
-Git tracks any files that you manually add to the repository.
-To make sure you do not accidentally add files that you do not wish to share (e.g., privacy sensitive data),
-you can list specific files that you do not want to track in the `.gitignore` file.
+You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore` file.
 You can add something to the `.gitignore` file directly or with this command:
@@
 ```
+
 Now the file `private.md` will not be committed to Git and will not be made public.
 
-Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
-When you are ready, you can begin tracking your files using Git by running:
+After you have ignored sensitive files (like not yet anonymized data), you can activate Git with:
 
@@
 
-For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
-One way to do this is through the visual user interface in the RStudio Git pane.
-Click on the empty box next to the file you want to stage. A check mark appears to indicate that the file is staged.
-After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
-This stores a snapshot of the current state of the project.
+If you made changes to a file, you have to stage and then commit these changes.
+The easiest way is to use the RStudio Git pane.
+You click on the empty box next to the file you want to stage, and when a tick appears, the file is staged.
+After you have staged all files you want, click on the commit button, explain curtly why you made those changes in a commit message and then click on commit.
 
 The files you created and the changes you made still have not left your computer.
-All snapshots are stored in a local repository in your project folder.
-To back up and/or share the files online, you can push your local repository to a remote repository.
-While you can choose any Git service (like GitLab or BitBucket), we will use GitHub in this tutorial.
-Before you upload your project to Git, you need to decide if you would like the project publically accessable (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
-To upload the project publically to GitHub use:
+To back up and share the files, you can publicly upload those changes to GitHub with:
 
@@
 
-To upload it privately:
+Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow its suggestions.
+
+If you are not yet ready to share your project publicly use:
 
@@
 
-Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the R console.
+Now that you have created a project that is under version control, we turn to dynamic document generation.
+Three aspects are essential to understand a dynamic document:
 
-## Dynamic document generation
+1. Text
+2. Code
+3. Metadata
 
-Now that you have created a version controlled project, we will proceed with dynamic document generation.
-A dynamic document has three elements:
-
-1. Text in natural language (prose; e.g., a scientific paper or presentation)
-2. Executable code (e.g., analyses)
-3. Metadata (e.g. title, authors, document format)
-
-R Markdown is a type of dynamic document well-suited to the RStudio user interface.
-The natural language text in an R Markdown file can be formatted using Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
-The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
-The following example serves to illustrate the Markdown syntax. 
-It shows how to create a heading, a word in bold font, and a list of several items in Markdown:
+The text of an R Markdown is written in Markdown.
+The following example shows how to signify a header, a bold word and a list in Markdown:
 
@@
 <!--this is a Markdown file -->
-# Heading (level 1)
+# Header
 
@@
 Important **word** in bold.
-
-## Subheading (level 2)
 
@@
 * do more research
-* spend time with family and friends
+* spend time with family
 ```
 
-Using markup for formatting, instead of the visual approach used in, e.g., Microsoft Word, is compatible with many output formats and is easier to put under version control.
-Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
-The visual editor has the additional benefit of promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions).
-Among other features, it eases the generation of citations and cross references.
-
-Now that you are familiar with the basics of writing text in Markdown, we turn our attention to including code into the text.
 Code is separated by three backticks (and the programming language in curly braces) like here:
@@
 
-The results of your code are than dynamically inserted into the document.
-Dynamically means that whenever you turn the R Markdown into its output format the code is executed again and if you changed something the results are updated.
-The output document is than static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
-This static document is somewhat decoupled from the R Markdown and the rest of the code that did create it.
-It is quite possible that several slightly different versions are therefore circulating.
-We therefore recomend to connect the static document to the version control system by dynamically including the identifier of the identifier of the Git commit, called hash, in the document.
-The `repro` package comes with the function `repro::current_hash()` for that purpose.
-This document was created from the commit with the hash ``r repro::current_hash()``.
-
-Now that you know how to write text and R code in an R Markdown, you need to know about metadata, like the title or the output format.
-Metadata are placed at the beginning of the document and are separated by three dashes from the document body; the following example is a full markdown document where the header is in lines 1--6.
+Now that you know how to write text and R code in an RMarkdown, you need to know about metadata, like the title or the output format.
+Metadata are placed at the beginning of the document and are separated by three dashes; the following example is a full markdown document where the header is in lines 1-6.
 
@@
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
-abstract: A tutorial about doing the same thing more than once.
+abstract: A paper about recalculating the same thing more than once.
 output: html_document
@@
 
-Most metadata fields are self-explanatory (like the author field). 
-Sometimes, R package may extend the metadata fields for their own purposes.
-Certain output formats may require specific  metadata. 
-For example, the `rticles`-package that offers templates for authoring scientific manuscripts comes with extra fields that are described in the package documentation and offer extra options for flexible formatting of documents.
-R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
-The `repro` package uses the metadata for hints to make the research project reproducible.
-Researchers should list all dependent resources of their research project in the metadata, that is, R scripts, data files, and external packages in the following format (see everything below the line *repro:*):
+Most metadata fields are self-explanatory (like the author field), but any package may use the metadata for their purposes.
+Often, for example, does a specific output format require more metadata, i.e. the `rticles`-package for authoring scientific manuscripts.
+RMarkdown was extended to hundreds of document formats which are often highly configurable by their metadata.
+The `repro` package uses the metadata, not to change the output, but to make the document more reproducible, simply by listing resources, i.e., scripts, data, and packages, your document depends on, in the following format:
 
@@
 `repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
-These two files together form the basis for consistency within a research project and consistency across different systems.
-At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.
-This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
-Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).
-In everyday work, the simplest solution is to now call the function `repro::automate()` which both creates a Makefile and a Dockerfile from all R Markdown files in the project (all files with the ending `.Rmd`) based on the respective repro metadata entries.
-Every time you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
+In day-to-day work the simplest solution is to use the function `repro::automate()` which both infers a Makefile and a Dockerfile from all RMarkdown  (all files with the ending `.Rmd`) and there respective repro metadata.
+Every time you change the `repro`-metadata, change the output format, or add a new RMarkdown, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
 The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
-Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
-<!--AB: why is this addition not automated?
-Feels strange given that this is our recommended best practice?!-->
+Besides the main Makefile and the Dockerfile repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
 
-When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:
+When you or people who received your project from GitHub want to reproduce the project, they can ask repro to explain how this should be done:
 
 ```{r}
-repro::reproduce()
+repro::rerun()
 ```
 
-If you uncomfortable with using the terminal directly, you can sent the command to the terminal from within R:
-
-```{r, eval=FALSE}
-system(reproduce())
-```
-
-Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
-These projects will remain reproducible irrespective of whether `repro` will be maintained and remain accessible in future.
-Users do not have to have repro installed to reproduce a project, in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
+However, users do not have to have repro installed, in fact, they do not even need R installed. 
 The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.
-The commands can be pasted into a terminal and will reproduce the project in many cases.
-<!--AB: I said this before.
-I think it really should be possible to execute this command from within R.
-Never underestimate the troubles people can have with the terminal.
-I think we discussed this before.
-Please add a function repro::rerun(execute=TRUE) or -- if we rename it -- repro::reproduce(execute=TRUE).
--->
-
-# Advanced Features
-
+They can be pasted into a terminal and will reproduce the project in many cases.
 Some projects, however, are not as easy to make reproducible.
-As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
+As explained above, repro is merely a simplified interface.
 These simplifications are possible because of two restrictions.
-Users who ask themselves either "How can I install software that is not an R package in the Docker image?" or "How can I express complex dependencies between files, e.g.
-raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
-All other users may safely skip this section and may return to it if they become limited by restriction.
-
-The metadata the `repro::automate()` function relies on can only express dependencies R packages for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
-The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
-Other software that users might need like other programming languages, not already installed LaTeX packages etc. have to be added manually.
-We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g.
-for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
-The second limitation is about dependencies.
-Make can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
-If B is missing in this example, Make would know how to recreate it from C and D.
-These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
-Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D therefore leaving out important information or express the dependencies directly within the Makefile.
-
-The following section explains how to overcome these limitation while still relying on the automatons afforded by `repro`.
+First, the project can only rely on R packages for the Dockerfile and second, it can have no multilayered preprocessing for the Makefile.
 Lifting these restrictions requires the user to interact more directly with Make or Docker.
@@
 
-Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two processing steps.
-First it recreates the virtual software environment (Docker) and then it executes computational recipes in the virtual software environment (Make).
-The first step is done by the command `make docker`.
-The command `make docker` will trigger Make to build the target called `docker`.
-The recipe for this target builds an image from the Dockerfile of the repository.
-The `&&` concatenates both commands and only runs the second command if the first was successful.
-Therefore, the computational steps are only executed when the software environment is set up.
-The second step executes the actual reproduction and is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy parts.
+Let us dissect the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two steps.
+First it virtually recreates the software environment and then it executes computational recipes in the virtual software environment.
+The first step is represented by the command `make docker`.
+The command `make docker` will trigger Make to build the target `docker`.
+The recipe for this target builds an image from your Dockerfile.
+The `&&` only runs the second command if the first was successful.
+Therefore, the computational steps are only executed, when the software environment is set up.
+The second step and actual reproduction is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy properties.
 First, a call to `make` without any explicit target will build the Make target `all`.
-Second, the flag `-B` means that Make will consider all dependencies as outdated and hence will rebuild everything.
+Second, the flag `-B` means that Make will consider all prerequisites as outdated and hence will rebuild everything.
 Third, repro constructs Make targets so that if you supply `DOCKER=TRUE` they are executed within the Docker image of the project.
 
-This interplay between Docker and Make is relatively complicated because it resembles a chicken or egg problem.
+This interplay between Docker and Make is so complicated because it resembles a chicken or egg problem.
 We have computational steps (Make) which depend on the software environment (Docker) for which we again have computational steps that create it.
-Users only require a deeper understanding of this interdependence when they either want to have more complex computational recipes than rendering an R Markdown or require other software than R packages.
+Users only require a deeper understanding when they either want to have more complex computational recipes than rendering an R Markdown or require other software than R packages.
 
@@
 `repro` creates three Dockerfiles inside the `.repro` directory.
-The file `.repro/Dockerfile_base` contains information about the base image on which all the remaining software is installed.
-By default we rely on the "verse" images provided by the Rocker project [@boettigerIntroductionRockerDocker2017].
-These contain (among other software) the packages `tidyverse`, `rmarkdown` and a complete LaTeX installation, which makes these images ideal for the creation of scientific manuscripts.
-Users can choose which R version they want to have inside the container by changing the version number in line 1 to the desired R version number.
-By default the R version corresponds to the locally installed version on which `repro::automate()` was called the first time.
+The file `.repro/Dockerfile_base` contains information about the base image on which all the rest is build.
+By default we rely on the "verse" images provided by the Rocker project.
+These contain (among other) the packages `tidyverse`, `rmarkdown` and a Latex installation, which make them ideal for writing scientific manuscripts.
+Users can choose which R version they want by changing the version number in line 1 to the desired R version number.
+By default the R version corresponds to the version on which `repro::automate()` was called the first time.
 The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed, by default this date is set to the date on which `repro::automate()` was called the first time.
-This way, the call to the automate function virtually freezes the R environment to the state it was called the first time inside the container. 
-Below, you see the Docker base file we used to create this manuscript:
 
@@
 
-The file `.repro/Dockerfile_manual` is empty by default and can be used to add further dependencies outside of R, like system libraries or external software. 
-Remember, the software environment of the Docker image is completely separate from software installed on your computer.
-This separation is excellent for reproducibility, but requires you to install software on an operating system that may not be familiar to you.
-The images supplied by [@boettigerIntroductionRockerDocker2017] are based on Ubuntu, and hence software has to be installed in a way that is compatible with Ubuntu.
-The most convenient way to install software on Ubuntu is through its package manager `apt`.
-The following snipped installs Python.
-If added to `.repro/Dockerfile_manual` the Docker image will have Python installed.
+The file `.repro/Dockerfile_manual` is by default empty and is supposed to allow dependencies outside of R, like system libraries or external software.
+The most convinient way to install software on Ubuntu is through its package manager `apt`.
+This snipped installs python:
 
@@
 
-The last Dockerfile, `.repro/Dockerfile_packages`, is automatically generated from the metadata provided and must not be manually changed.
-Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.
-
-The separation into several Dockerfiles is a design choice to provide a clear separation between automatically generated parts based on metadata provided and manually altered parts of the Dockerfile.
-Docker eventually requires a single Dockerfile to run, so `repro::automate()` simply concatenates the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the R project.
-
-With this approach, users of `repro` can not only build complex software environments but they can also implement complex file dependencies.
-The standard `repro` metadata only makes sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
-If you can formulate the creation of dependencies in terms of computational steps, e.g.
-the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the Makefile.
-The Makefile that `repro` creates is only an template and you are free to change it.
-However, make sure to never remove the following two lines:
-
-```bash
-include .repro/Makefile_Rmds
-include .repro/Makefile_Docker
-```
-
-The file `.repro/Makefile_Rmds` contains the automatically generated targets from `repro::automate()` for the R Markdown files.
-This file should not be altered manually.
-If you are not satisfied with the automatically generated target than simply provide an alternative target in the main `Makefile`.
-Targets in the main `Makefile` take precedent.
-
-The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but usually should not modify.
-This Makefile coordinates the interplay between Make and Docker and contains targets for building (with `make docker`) and saving (with `make save-docker`) the Docker image.
-Additionally, it provides facilities to execute commands within the container.
-If you write computational recipe for a target it will be by default evaluated using the locally installed software.
-To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first Make example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
-2:
-```bash
-simulated_data.csv: R/simulate.R
-  $(RUN1) Rscript -e 'source("R/simulate.R")' $(RUN1)
-```
-
-If users execute this in the terminal:
-
-```bash
-make simulated_data.csv
-```
-It behaves exactly as the in the first Make example, the script `R/simulate.R` is run using the locally installed R.
-Because this translates simply to:
-
-```{r, echo=FALSE, results='asis'}
-cat("```bash",
-  stringi::stri_wrap(system2("make", " --dry-run simulated_data.csv", stdout = TRUE)),
-  "```", sep = "\n")
-```
-
-But if users use
-
-```bash
-make DOCKER=TRUE simulated_data.csv
-```
-
-it is evaluated within the Docker container using the software within it and not the locally installed R version:
-
-```{r, echo=FALSE, results='asis'}
-cat("```bash",
-  stringi::stri_wrap(system2("make", " --dry-run DOCKER=TRUE simulated_data.csv", stdout = TRUE)),
-  "```", sep = "\n")
-```
-
-To summarise, `repro` provides researchers with an easy to use solution for dependency tracking (in form of Make) and software management (using Docker) without the necessity to learn both languages.
-Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automations.
+The the last Docker, `.repro/Dockerfile_packages`, should not be manually changed.
+`repro::automate()` gathers all packages from all `.Rmd` files and figures out if they should be installed from CRAN or GitHub.
+`repro::automate()` also simply combines the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the project.
+The seperation into several Dockerfiles is only to provide a clear seperation between automatically generated parts and manually altered parts of the Dockerfile.
+However, Docker requires in the end the one Dockerfile that `repro::automate()` creates from them.
 
@@
 
-Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
-Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
-Preregistrations, however, have a problem.
-Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
-The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
-The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
-But these templates can not remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
+We argue that a preregistration as code (PAC) is an excellent planning tool that offers several benefits over traditional preregistration.
+For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data using simulated data.
 
-Our proposed solution is simple:
-Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
-We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
-PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration despite the best efforts afford.
+We consider three criteria when planning a study.
+First, the plan should be comprehensive, e.g. state the research question and describe the study design and analysis.
+Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
+Third, the plan should be efficient; therefore, it only employs the necessary amount of resources.
 
-The idea of submitting code as part of a preregistration is not new.
-A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
-In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
+To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript.
+The introduction and theoretical background provide the basis for the studies design and can be written before gathering data.
+We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
+To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
+The researchers can include a "faked " results section based on the simulated data, employing dynamic document generation.
+Repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to assure efficiency.
+For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study.
 
-We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.
+Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [https://www.pnas.org/content/115/11/2600].
+Preregistration increases the credibility of empirical results in three ways [preregistration is hard and worthwhile, Nosek].
 
-## Why are PAC and reproducibility related?
+> First, preregistration of analysis plans makes clear which analyses were planned a priori and which were conducted post hoc.
+This improves calibration of uncertainty for unplanned analyses, and diagnosticity of statistical inferences for planned analyses.
+[...] Doing so has the benefit of strengthening statistical inferences as compared with unplanned analyses.
+[...] Second, preregistration enables detection of questionable research practices such as selective outcome reporting (John et al.
+2012) or Hypothesizing After the Results are Known (HARKing; Kerr, 1998).
+Third, preregistration of studies can reduce the impact of publication bias—particularly the prioritization of publishing positive over negative results—by making all planned studies discoverable whether or not they are ultimately published.
 
-First, preregistrations are plagued by non-reproducibility like any other research outcome.
-@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to reproduce the power analysis.
-Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
-For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
+A preregistration as code offers four advantages over classical preregistration.
+First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
+Second, despite being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
+Researchers can, for example, formulate conditions under which they prefer one analysis over the other, i.e.
+if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
+Third, deviations from the preregistration are more explicit because they are reflected in changes to the codes a property facilitated by version control.
+Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
 
-Second, although it has been almost a decade since @steegenMeasuringCrowdAgain2014 have preregistered their code, the idea of publishing code before analyzing data has gained not much traction.
-This could change when combining the here proposed workflow and PAC.
-We want to extend PAC from preregistering code to preregistering *the whole manuscript as a dynamic document*.
-The idea is as follows:
-For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
-In this document, the results are initially based on simulated data that mimics the expected output.
+# Preregistration as Code --- an example
 
-Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
-It follows that for preregistrations they should stick to widely accepted standards of scientific writing [e.g. @apa7].
-We have already explained the need for reproducibility standards in the process of writing such manuscript and it is therefore only natural to extend these arguments to preregistration.
-
-## Advantages of PAC over traditional preregistration
-
-We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, a PAC removes any ambiguity regarding the translation of an verbal analysis plan into code.
-Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
-Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
-For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
-Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
-We believe it is easier with PAC to detect missing steps or ambiguous decisions.
-Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
-Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
-Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.
-
-## Benefits of PAC for study planning
-
-Leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
-To use PAC for study planing, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
-To that end, they should borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
-Then they should translate their method section into executable computer code and simulate data that has the same structure as the data they expect to obtain from their experiments.
-
-We consider three domains where this procedure can help planning a study.
-First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
-Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
-Third, the plan should be efficient; the plan should only require the necessary amount of resources.
-
-### Comprehensiveness
-
-Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
-One particular advantage is the strong foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
-Nevertheless it can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section.
-
-Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-Simulating data forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
-It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
-Writing incomplete code is technically impossible or at least researchers would notice if they e.g. forgot to aggregate the data etc.
-
-### Effectiveness
-
-With simulated data researchers can check if the analysis is technically complete and can also recover the true parameters of the simulation.
-When an analysis fails to recover the correct values in idealized conditions with big samples researchers should not expect this analysis to work under realistic conditions.
-Because the simulation captures the researchers expectations about the real world conditions it is the best way to assess the effectiveness of the analysis without actually gathering the data.
-
-### Efficiency
-
-A single simulated or synthesized dataset already helps to ensure that the plan is viable under certain circumstances.
-However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
-Monte Carlo simulation lets researcher navigate several trade offs regarding their resources.
-In the simplest case they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
-@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.
-You may also evaluate how the data analysis copes with expected data complications, e.g.
-non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
-
-## What you need to know before writing a PAC
-
-To write a PAC researchers should start by drafting a manuscript in form of an R Markdown.
-When they have completed the method section, they should begin writing code.
-We suggest writing three functions for each hypothesis, one for the planned analysis, one for the simulating the data and one for reporting the results.
-At the heart of a PAC lies the planned analysis so researchers should first draft a function called `planned_analysis()` that captures the description in the method section.
-This function should receive the data and derive relevant results from it.
-
-```{r}
-planned_analyis <- function(data){
-  # 1. preprocess e.g. with `rowMeans(data)`
-  # 2. conduct analysis e.g. with `t.test()`
-  # 3. `return(results)`
-}
-```
-
-Because it is difficult to draft such function without interactively test it, researcher should than proceed to write a function for simulating data.
-
-```{r}
-simulate_data <- function(n, effect_size){
-  # 1. warn users that the results are "fake"
-  # 2. draw `n` samples with `effect_size`
-  # 3. format and return in expected data format
-}
-```
-
-As soon as they have written `planned_anaylsis()` and `simulate_data()` they can iteratively improve both functions, e.g.
-till `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
-The goal is that the output of `simulate_data()` works as input to the function `planned_analysis()`.
-
-This process can be tedious and complicated.
-If it proofs infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
-Of course researcher must not analyze the data before they write the preregistration but they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
-
-When the researchers are satisfied with the function `planned_analysis()` they can think about how they would like to report the results of the analysis, via tables, plots and text.
-The implementation of this reporting should be in the function `report_analysis()`.
-
-```{r}
-report_results <- function(results){
-  # 1. create markdown tables from results
-  # 2. conditionally interpret results e.g. if(p < .025)"Result is significant."
-  # (optional) visualize results
-  # 3. return results section formatted in markdown
-}
-```
-
-These function again should be accept the output of `planned_analysis()` as input.
-The output of this function should be formatted in Markdown.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
-Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
-
-Ideally, these three functions can be composed to create a "fake" results section, e.g.
-when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section.
-
-To summarize researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
-After they gathered the actual data, they replace the simulated data, run `planned_analysis()` and write the discussion.
-
-# Preregistration as Code --- a Tutorial
-
-To guide the reader through a preregistration as code (PAC), we will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
+To guide the reader through a preregistration as code (PAC), we will use an exemplary research question:
 
@@
 At the heart of a PAC lies the planned analysis.
-We suggest writing a main analysis function that captures the full process of testing a research question.
-This function receives a dataset as input and performs all statistical processing. <!--AB say anything about its output? is the Markdown the output?-->
-For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
-As an example of a data-dependent decision, we included a rank transformation if the skewness was greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+We suggest writing a function that captures the full process of testing a research question that receives a dataset as input.
+For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavillism items of the Short Dard Triad [SD3, @Jones2013] between the gender groups.
+As an example of a data dependent decisions we included a rank transformation if the skewness greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+This analyitical strategie is suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]); we merely choose it for didactic purposes because similar procedures a widely used [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
 
-[^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
+[^caution-mww]: The Mann–Whitney–Wilcoxon is **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
 The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
-
-This function is one possible translation of the verbal description above:
 
@@
 
-Importantly, natural language descriptions do not map unambiguously onto computer code implementations. 
-The same textual descriptions may be implemented differently by different persons depending on their statistical and programming knowledge but also based on their own expectations and beliefs.
-One example would be using the function `wilcox.test` instead of the combinations of `rank` and `t.test`.
-Both are valid implementation of the Mann–Whitney–Wilcoxon test but the first assumes equal variance whereas the second applies Welch's correction and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
-Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
-Still, these details can make a difference and provide unwanted researcher degrees-of-freedom.
-
-<!--AB here is kind of a break in the text-->
-To use this function one needs a compatible dataset. <!--AB what is "this" function? What does "compatible" mean?-->
+To use this function one needs a compatible dataset.
 At the time of preregistration there is of course no actual data available.
-We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
-Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
-The following code simulates a dataset with one observed variable in two groups.
+We therefore simulate a dataset.
 
@@
 ```
-<!--AB it is odd that the return variable is called effect because it is a dataset-->
 
-Here, we decided to randomly draw non-normal data to test how well our procedure performed under non-optimal situations.
-Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a given standardized effect size (a Cohen's d of .2).
-This simulated dataset fits perfectly into our planned analysis with a sample size of 400 and 10 items: <!--AB in what way? please explain-->
+We decided to draw non-normal data to test how well our procedure performed under non-optimal situations.
+Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a standardized effect size.
+This simulated dataset fits perfectly into our planned analysis:
 
@@
 
-```{r}
+These results are based on simulated data.
+
+```{r, results='asis'}
 set.seed(112358) # fix state of RNG
-results <- 
+report::report_text(
   planned_analysis(
@@
     )
+  )$test
 )
@@
 
-Next, we add dynamic code blocks to our Markdown report that implement the planned reporting of the final results.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
-Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
-Here we use the `report` package to create a preview of how our reporting will look like based on simulated data:<!--AB add citation to the report package-->
-
-```{r, results='asis', eval=FALSE}
-report::report(results$test)
-```
-
-This call to the `report()` function will yield the following automatically generated output:
-
-```{r, results='asis', echo=FALSE}
-report::report(results$test)
-```
-<!--AB it is difficult to discern the automatically generated text and out manuscript text even though there is the horizontal line. Not sure how to improve this.-->
 ***
@@
 @Jones2013 found in three studies gender differences of .24, .29, and .35 for their machiavilism subscale.
-If we want to detect an effect of size `r choosen_d` with an `r choosen_power*100`% probability, our simulation shows that we require at least `r minn$n` participants.
-
-<!--AB please add a figure number and a proper citation to the figure. Also, tell people where they can find the code to run the power analysis themselves-->
+If we want to detect an effect of size .2 with an 80% probability, our simulation shows that we require at least `r minn$n` participants.
 
@@
 ```
-
-# Discussion
-
-<!--AB: TODO add a quick summary -->
-
-# Limitations
-
-PACs come with a challenge though.
-In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
-We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
-This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
-The reproducible workflow we present here is heavily automated and hence promotes reusability.
-Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
-
-Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.
-
-<!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->
-Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
-The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
-The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-<!--AB can we move this to a different place?
-Maybe this would actually better fit to the discussion/limitation section!-->
-
-# Outlook
-
-<!--AB: two lines of what the future of the package, of such standards, and prereigstration is.
-Maybe say something that preregistration is not always possible or wanted but that reproducibility is almost always a good idea?-->
 

### 2021-07-28 manuscript.Rmd
Update manuscript.Rmd
added more to discussion + TODOs for AP

--- 2021-07-28 manuscript.Rmd
@@
 
-<!--AB: TODO add a quick summary -->
-
-# Limitations
-
-PAC comes with a challenge.
-In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
+## Summary
+
+We demonstrated how the R package 'repro' supports researchers in creating reproducible repositories and reproducible manuscripts.
+These are important building blocks for transparent and cumulative science
+because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
+This is made possible because the repro approach fully documents all software dependencies from the exact version of specific R packages down to the entire operating system layer.
+Using a virtualization approach, the entire software environment can be recreated on a different computer without interfering with locally installed software.
+Building upon reproducible manuscripts, we suggest to use reproducible PAC as a means to create and publish unambiguous pre-registrations in the form of code.
+Ideally, a PAC contains all analysis code and a reproducible manuscript that already renders the entire reporting of statistical and computational results.
+These results are based on simulated data, such that there is no danger of peeking at the real data while writing the scripts.
+This way, all researcher's degrees of freedom are eliminated.
+Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
+All deviations from the pre-registration can be documented with version control systems that produce accurate documentations of the differences in the analysis code.
+
+## Limitations of reproducibility
+
+<!--AB: please add what possible problems are there that endanger reproducibility; what are the assumptions under which this all works... -->
+Reproducible repositories created with the proposed workflow will work as long as operating systems with support for Make, Docker, and git are available.
+If development of git should be stopped one day, we believe that there will be other version control systems that will offer similar features and that can be substituted for git.
+If Make is not available on a future system, the Makefile itself is written in a way that one can manually trace the dependencies and execute commands in the right order.
+If Docker is not available on a future system, reproducing the comoutational environment will be more difficult, but the Dockerfile will still serve as unambiguous documentation of how the original system was set up and may help future users to create a software environment as close as possible to this environment.
+<!--AB: please finalize; say something to the effect of Dockerfiles vs full containers (again)?-->
+<!--AB: are there any specific limitations of what the package can/cannot do?-->
+
+## Limitations of PAC
+
+PAC comes with challenges.
+In contrast to a reproducible script that needs to work on a single, given dataset, a PAC must work across different data sets that are yet unknown at the time the script is created.
 We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
-This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
-The reproducible workflow we present here is heavily automated and hence promotes reusability.
+This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _built-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
+The reproducible workflow we present here is heavily automated and hence promotes reusability; however, there is an increased demand on the researchers' side to write and maintain code that is flexible enough to work with different inputs.
 Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
 Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
-
-Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.
+Another challenge are analyses that do not have linear execution but branch out.
+If such conditional analysis steps exist (such as those we demonstrated before with the test of mean differences), scholars using PAC need to write code that may be never executed.
+Typical examples of conditional analysis steps could be an exploratory factor analysis that is only executed if the theory-driven model does not fit well, or a group comparison is only executed if measurement invariance in a latent factor model is established.
+Writing code that may be never executed puts additional burden on the researcher.
+However, we hope that with the principle of reusability in mind, open and reproduce PAC elements may be reused by others and in turn save time to future developers.
+
+We also suggest that version control with Git and GitHub fulfills the criteria of a preregistration as long as you state your research plan comprehensively.
+This view is not shared by everyone [@szollosi_is_2020] because with some effort and malice intend the record of Git/GitHub can be altered.
+This concern can be addressed by using zenodo.org to give a digital object identifier (DOI) to your preregistration (see [this guide](https://guides.github.com/activities/citable-code/)) because zenondo.org saves an unchangeable copy of your repository.
+
+## Limitations of the test for mean differences
 
@@
 
-We also suggest that version control with Git and GitHub fulfills the criteria of a preregistration as long as you state your research plan comprehensively.
-This view is not shared by everyone [@szollosi_is_2020] because with some effort and malice intend the record of Git/GitHub can be altered.
-This concern can be addressed by using zenodo.org to give a digital object identifier (DOI) to your preregistration (see [this guide](https://guides.github.com/activities/citable-code/)) because zenondo.org saves an unchangeable copy of your repository.
-
-# Outlook
-
-<!--AB: two lines of what the future of the package, of such standards, and prereigstration is.
+## Outlook
+
+<!--AB: TODO for AP: two/three lines of what the future of the package, of such standards, and prereigstration is.
 Maybe say something that preregistration is not always possible or wanted but that reproducibility is almost always a good idea?-->

### 2021-07-28 manuscript.Rmd
comments AB

--- 2021-07-28 manuscript.Rmd
@@
 
-Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
-Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For a priori planned analyses, preregistration should reduce "researcher degrees of freedom" therefore, the number of decisions researcher can make that could unduly influence the results.
-Preregistrations, however, have a problem.
+Defining the research questions and planning data analysis before observing the research outcomes is referred to _preregistration_ [@NosekRevolution2018].
+Preregistration increases the credibility of empirical results by separating a-priori planned analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
+For a priori planned analyses, preregistration is supposed to reduce so-called "researcher degrees of freedom", and therefore, the number of decisions researcher can make that could unduly influence the results.
+Preregistrations, however, often do not achieve the level of unambiguity that are required to effectively zero the researcher degrees of freedom.
 Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
@@
 We term this *preregistration as code (PAC)*.
-PAC has the potential to eliminate all researcher degrees of freedom, even those standard preregistration can not eliminate.
+A reproducible PAC has the potential to eliminate all researcher degrees of freedom, even those that standard preregistration can not eliminate.
 
@@
 @bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis, but from these, only 34 (20%) provided enough information to reproduce the power analysis.
-Even researchers who have gone to great lengths in submitting a script in their preregistration sometimes inexplicably fail to reproduce their results.
+Even researchers who have gone to great lengths in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
 For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
@@
 The idea is as follows:
-For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
-In this document, the results are initially based on simulated data that mimics the expected output.
-
-Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
-It follows that for preregistrations they should stick to widely accepted standards of scientific writing [e.g. @apa7].
+For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather or acquire empirical data.
+In this document, the results are initially based on simulated data that merely serves the purpose to mimic the expected output. 
+For example, this simulated data allows to write code that already produces figures or tables.
+
+Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific report, ie., researchers do not have to learn specifics of a given preregistration template.
+It follows that for preregistrations they should simply stick to widely accepted standards of scientific writing [e.g. @apa7].
 We have already explained the need for reproducibility standards in the process of writing such manuscripts and it is therefore only natural to extend these arguments to preregistration.
@@
 
-We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
+We believe that pairing PAC with the workflow presented here offers five advantages over classical preregistration.
 First, a PAC removes any ambiguity regarding the translation of a verbal analysis plan into code.
@@
 Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
-Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
+Fifth, the preregistration is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
 
@@
 When they have completed the method section, they should begin writing code.
-We suggest writing three functions for each hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
+We suggest writing three functions for each hypothesis tested, one for the planned analysis, one for simulating the data and one for reporting the results.
 At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
@@
 Because it is difficult to draft such a function without interactively testing it, the researcher should then write a function for simulating data.
+The follow code illustrates the signature of a simple simulation function:
 
@@
 As soon as they have written a first draft of `planned_anaylsis()` and `simulate_data()` they can interactively improve both functions, e.g.
-till `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
+until `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
 The goal is that the output of `simulate_data()` works as input to the function `planned_analysis()`.
 
-This process can be tedious and complicated.
-If it proves infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
-Of course, researchers must not analyze the data before they write the preregistration.
-However, they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
+If it is infeasible to simulate realistic data, researcher can delay their preregistration until they have gathered a given percentage of the data.
+Of course, researchers must not analyze these data before they write the preregistration.
+However, they can obfuscate these data by shuffling the dependent variable or otherwise createing synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
 
@@
 The implementation of this reporting should be in the function `report_analysis()`.
+<!--AB hier ist mir unklar, wie das genau geht. baue ich hier alle Figures und Tables und Textstücke und benutze die dann erst später? Bitte genauer-->
 
@@
 The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it generates not only tables but also a straightforward interpretation of the effects as actual prose (e.g.
-it quantifies the size of an effect).
+it verbally quantifies the size of an effect).
 
@@
 
-We have made the argument that PAC has significant advantages over classical preregistration and have outlined how to implement it, but some may be sceptical if it is feasible.
-To address this concern and help researchers implementing PAC themselves, we have conducted a PAC that can serve as an example.
+We argued that PAC has several advantages over classical preregistration and have outlined how to implement it.
+To illustrate how PAC works in practice and help researchers implementing PAC themselves, we have conducted a PAC that serves as a guiding example.
 We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
@@
 
-Again, we propose a preregistration format that closely resembles a classical journal article, but utilizes simulated data and dynamic document generation to serve simultaneously as preregistration and final manuscript.
-The complete example is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
-In this section, we show excerpts of this example formatted in monospace and provide some commentary.
+Again, we propose a preregistration format that closely resembles a classical journal article, but uses simulated data and dynamic document generation to create a document that starts out as a preregistration and eventually becomes the final report.
+The complete preregistration source is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
+In this section, we show code excerpts of this preregistration (formatted in monospace) and explain the rationale behind them.
 
@@
 
-As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary information and literature to understand the rest of the paper.
-This example is drastically shortened for illustration purposes:
+As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary background information and literature to understand the context and purpose of the research question.
+This example is drastically shortened for purposes of illustration:
 
@@
 
-After researchers have provided the research question, they proceed to explain how they want to study it.
+After researchers have provided the research question, they typically proceed to explain how they want to study it.
 For simplicity, we will use already published data that we have not yet analyzed:
@@
 
-We choose the following statistical procedure because many researchers may be familiar with it[^caution-mww]:
+We choose the following statistical procedure because many researchers are familiar with it[^caution-mww]:
 
@@
 The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and their expectations and beliefs.
-One example would be using the function `wilcox.test` instead of the combinations of `rank` and `t.test`.
-Both are valid implementations of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
-In contrast, the second applies Welch's correction and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
+One example would be using the function `wilcox.test` instead of the combinations of the functions `rank` and `t.test`.
+Either of them are valid implementations of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
+In contrast, the second applies Welch's correction by default and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
 Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
@@
 We implemented a mechanism that only uses simulated data when the actual data is not yet available for the results section.
-This mechanism also warns readers if these results are based on simulated data.
-As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study.
+This mechanism also warns readers if these results are based on simulated data. <!--AB how does it warn?-->
+As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study. <!--AB how exactly does the script know?-->
 
@@
 
-The last thing we need to do is to report our results with the combination of `planned_analysis()` and `report_analysis()`.
+The last thing we need to do is to report our results with the combination of the functions `planned_analysis()` and `report_analysis()`.
 

### 2021-07-29 manuscript.Rmd

@@
 Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
-<!-- I dont understand this sentence -->
 One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
@@
 When they have completed the method section, they should begin writing code.
-We suggest writing three functions for each hypothesis tested, one for the planned analysis, one for simulating the data and one for reporting the results.
+We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
 At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
@@
 As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary background information and literature to understand the context and purpose of the research question.
-This example is drastically shortened for purposes of illustration:
+This example is drastically shortened for illustration purposes:
 
@@
 
-We implemented a mechanism that only uses simulated data when the actual data is not yet available for the results section.
-This mechanism also warns readers if these results are based on simulated data. <!--AB how does it warn?-->
-As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study. <!--AB how exactly does the script know?-->
+We implemented a mechanism that only uses simulated data when the actual data is not yet available  (in this example if the file `data/data.csv` does not exist)for the results section.
+This mechanism also warns readers if these results are based on simulated data.
+The warning is colored red to avoid any confusion between mock and actual results.
+As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study.
 

### 2021-07-29 manuscript.Rmd
idea: preregister explorative analysis

--- 2021-07-28 manuscript.Rmd
@@
 simplesummary: |
-  A Simple summary goes here.
+  Reproducibility has long been considered integral to the scientific method.
+  Something is called reproducible when independend researcher arrive at the same results from the same data.
+  Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
+  Technological advancements now enable scientists to achieve a much higher standard;
+  one in which any individual can be granted access to a digital research repository, and reproduce the analyses from raw data to final manuscript with a single command.
+  This method has far-reaching implications for scientific archiving,
+  reproducibility and replication,
+  scientific productivity,
+  and the credibility and reliability of scientific findings.
+  One obstacle preventing the widespread adoption of this method is that the underlying technological advancements
+  are complicated to use.
+  This paper introduces `repro`; an R-package that guides researchers in the installation and use of the tools required for computational reproducibility.
+  It includes a tutorial on how to make a research project reproducible with the aid of `repro`.
+  Finally, it introduces novel applications of reproducibility tools for scientific research;
+  including the preregistration of study plans as reproducible computer code.
+  This circumvents the shortcomings of ambiguous preregistrations that allow for researcher degrees of freedom,
+  because computer code describes study plans more precisely than prose.
+  Reproducibility made convenient with automation has hence a wide range of applications to accelerate scientific progress.
 abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
@@
 Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
-The credibility of empirical research hinges on its objectivity.
-Objectivity in this context means "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
-22][^popper].
-Therefore, reproducibility has long been considered integral to empirical research.
+Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
+22][^popper]
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
@@
 Scientific organizations and funding bodies increasingly demand transparent sharing of research artifacts, and researchers are increasingly willing to do so.
-
 However, although the sharing of such artifacts is a necessary condition for reproducibility, it is not a sufficient one.
@@
 In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
-Typically, results are manually copied and pasted from one software to another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
+Typically, results are manually copyied and pasted from one software to the other, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
 
@@
 It is highly fool-proof [unlike, for example, LaTeX @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
+We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your résumé [@vitae].
 R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
 Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
-This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
+This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
 @xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.
@@
 However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
-In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
+In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambigously documented.
 
@@
 Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
-In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
+In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, analysis results).
 
 We recommend using **Make** for dependency tracking because it is language independent.
-The following hypothetical example illustrates the utility of Make.
+The following hypothesical example illustrates the utility of Make.
 Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
@@
 It provides all software used in an analysis, including the operating system itself.
-The operating system level is important because some functionality may be passed to software layers beneath the programming language, for example random number generators and linear algebra libraries.
+The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
 Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
@@
 It will assist you while you follow the tutorial.
->>>>>>> main
 
@@
 An example of the assistance that `repro` provides are the check functions.
-
 These verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
@@
 ```
-
 Now the file `private.md` will not be committed to Git and will not be made public.
@@
 For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
-
 One way to do this is through the visual user interface in the RStudio Git pane.
-Click on the empty box next to the file you want to stage.
-A check mark appears which indicates that the file is staged.
+Click on the empty box next to the file you want to stage. A check mark appears to indicate that the file is staged.
 After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
@@
 
-Depending on your computer configuration, this may ask to set up a secure connection to GitHub.
-In this case first follow the suggestions shown on the R console.
+Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the R console.
 
@@
 
-1. Text (prose; e.g., a scientific paper or presentation)
+1. Text in natural language (prose; e.g., a scientific paper or presentation)
 2. Executable code (e.g., analyses)
@@
 R Markdown is a type of dynamic document well-suited to the RStudio user interface.
-The text of an R Markdown is formatted by Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
+The natural language text in an R Markdown file can be formatted using Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
 The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
 The following example serves to illustrate the Markdown syntax. 
-
 It shows how to create a heading, a word in bold font, and a list of several items in Markdown:
@@
 Another advantage is that, because R Markdown files are in plain text, they are more suitable for version control using Git.
-Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
+Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass <!--CJ Add picture-->in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
 The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
@@
 ````
-
 The hotkey [Control]+[Alt]+[i] inserts a block of code in the file.
@@
 These metadate contain information about the document, like the title and the output format.
-Metadata are placed at the beginning of the document and are separated by three dashes from the document body.
-The following example is a full markdown document where the metadata (the "YAML front matter") is in lines 1--6.
+Metadata are placed at the beginning of the document and are separated by three dashes from the document body; the following example is a full markdown document where the header is in lines 1--6.
 
@@
 
-If you feel uncomfortable using the terminal directly, you can sent the command to the terminal from within R:
+If you uncomfortable with using the terminal directly, you can sent the command to the terminal from within R:
 
@@
 
-The metadata the `repro::automate()` function relies on can only express R packages as dependencies for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile.
+The metadata the `repro::automate()` function relies on can only express dependencies R packages for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
 The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
-Other software that users might need like other programming languages and not preinstalled LaTeX packages have to be added manually.
-We plan to add support for commonly used ways to install software beyond R packages, e.g.
-for system libraries (via the Ubuntu package manager `apt`), LaTeX packages (via the Tex Live package manager `tlmgr`), and python packages (via the python package manager `pip`).
+Other software that users might need like other programming languages, not already installed LaTeX packages etc. have to be added manually.
+We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g.
+for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
 The second limitation is about dependencies.
@@
 These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
-Users therefore have to either "flatten" the dependency structure by simply stating that A depends on B, C and D, thereby leaving out important information, or express the dependencies directly within the Makefile.
-
-The following section explains how to overcome these limitations while still relying on the automations provided by `repro`.
-Lifting these limitations requires the user to interact more directly with Make or Docker.
-To satisfy more complicated requirements useres may have, they need to understand how `repro` utilizes Make and Docker internally.
-
-Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`.
-It consists of two processing steps:
+Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D therefore leaving out important information or express the dependencies directly within the Makefile.
+
+The following section explains how to overcome these limitation while still relying on the automatons afforded by `repro`.
+Lifting these restrictions requires the user to interact more directly with Make or Docker.
+To satisfy more complicated requirements users need to understand how repro utilizes Make and Docker internally.
+
+Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two processing steps.
 First it recreates the virtual software environment (Docker) and then it executes computational recipes in the virtual software environment (Make).
-The first step is done by the command `make docker`, it will trigger Make to build the target called `docker`.
+The first step is done by the command `make docker`.
+The command `make docker` will trigger Make to build the target called `docker`.
 The recipe for this target builds an image from the Dockerfile of the repository.
@@
 By default the R version corresponds to the locally installed version on which `repro::automate()` was called the first time.
-The build date is used to install packages in the version available on CRAN on this specific date and can also be changed, by default it is set to the date on which `repro::automate()` was called the first time.
+The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed, by default this date is set to the date on which `repro::automate()` was called the first time.
 This way, the call to the automate function virtually freezes the R environment to the state it was called the first time inside the container. 
@@
 The most convenient way to install software on Ubuntu is through its package manager `apt`.
-If the following snipped is added to `.repro/Dockerfile_manual` the Docker image will have Python installed.
+The following snipped installs Python.
+If added to `.repro/Dockerfile_manual` the Docker image will have Python installed.
 
@@
 the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the Makefile.
-The Makefile that `repro` creates is only a template and you are free to change it.
+The Makefile that `repro` creates is only an template and you are free to change it.
 However, make sure to never remove the following two lines:
@@
 This file should not be altered manually.
-If you are not satisfied with the automatically generated target simply provide an alternative target in the main `Makefile`.
+If you are not satisfied with the automatically generated target than simply provide an alternative target in the main `Makefile`.
 Targets in the main `Makefile` take precedent.
@@
 Additionally, it provides facilities to execute commands within the container.
-If you write a computational recipe for a target it will be by default evaluated using the locally installed software.
+If you write computational recipe for a target it will be by default evaluated using the locally installed software.
 To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first Make example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
@@
 
-Defining the research questions and planning data analysis before observing the research outcomes is referred to _preregistration_ [@NosekRevolution2018].
-Preregistration increases the credibility of empirical results by separating a-priori planned analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
-For a priori planned analyses, preregistration is supposed to reduce so-called "researcher degrees of freedom", and therefore, the number of decisions researcher can make that could unduly influence the results.
-Preregistrations, however, often do not achieve the level of unambiguity that are required to effectively zero the researcher degrees of freedom.
-Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
-The current best practice to ensure that the preregistration is comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
-The structure these templates provide can more effectively reduce researcher degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
-Nevertheless, these templates can not remove all researcher degrees of freedom because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
+Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
+Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
+For a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
+Preregistrations, however, have a problem.
+Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
+The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
+The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
+But these templates can not remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
 
 Our proposed solution is simple:
-Instead of describing the analysis, you write and publish the code to conduct the analysis before you gather data.
-We term this *preregistration as code (PAC)*.
-A reproducible PAC has the potential to eliminate all researcher degrees of freedom, even those that standard preregistration can not eliminate.
+Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
+We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
+PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration despite the best efforts afford.
 
@@
 A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
-In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`), we found close to a dozen published articles that included some form of script as part of their preregistration.
-
-We believe, however, that combining ideas from reproducibility and preregistration can turn this idea into a tool with tangible benefits for the research community.
+In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
+
+We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.
 
@@
 First, preregistrations are plagued by non-reproducibility like any other research outcome.
-@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis, but from these, only 34 (20%) provided enough information to reproduce the power analysis.
-Even researchers who have gone to great lengths in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
+@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to reproduce the power analysis.
+Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
 For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
@@
 The idea is as follows:
-For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather or acquire empirical data.
-In this document, the results are initially based on simulated data that merely serves the purpose to mimic the expected output. 
-For example, this simulated data allows to write code that already produces figures or tables.
-
-Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific report, ie., researchers do not have to learn specifics of a given preregistration template.
-It follows that for preregistrations they should simply stick to widely accepted standards of scientific writing [e.g. @apa7].
-We have already explained the need for reproducibility standards in the process of writing such manuscripts and it is therefore only natural to extend these arguments to preregistration.
+For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
+In this document, the results are initially based on simulated data that mimics the expected output.
+
+Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
+It follows that for preregistrations they should stick to widely accepted standards of scientific writing [e.g. @apa7].
+We have already explained the need for reproducibility standards in the process of writing such manuscript and it is therefore only natural to extend these arguments to preregistration.
 
@@
 
-We believe that pairing PAC with the workflow presented here offers five advantages over classical preregistration.
-First, a PAC removes any ambiguity regarding the translation of a verbal analysis plan into code.
+We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
+First, a PAC removes any ambiguity regarding the translation of an verbal analysis plan into code.
 Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
 Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
-For example, if distributional assumptions are not met, the code may branch out to employ robust methods or an analysis may perform automated variable selection mechanisms before running the final model.
-Third, a PAC is more comprehensive by design because its completeness can be checked with simulated data.
-Using simulated data makes it straightforward to detect missing steps or ambiguous decisions.
+For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
+Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
+With PAC researchers can precisely calibrate how much they want to explore.
+When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
+Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
+We believe it is easier with PAC to detect missing steps or ambiguous decisions.
 Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
-It can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
 Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
-Fifth, the preregistration is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
+Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.
 
@@
 
-When leaving meta-scientific benefits aside, a preregistration is also a tool for planning a study.
-To use PAC for study planning, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
+Leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
+To use PAC for study planing, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
 To that end, they should borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
@@
 Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
-Third, the plan should be efficient; it should only require the necessary amount of resources.
+Third, the plan should be efficient; the plan should only require the necessary amount of resources.
 
@@
 Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
-<!-- I dont understand this sentence -->
-One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
-Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
+One particular advantage is the strong foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
+Nevertheless it can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section.
 
 Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
-While relatively uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-When simulating data, researchers have to think about and address data-analytic problems on a conceptual level before making an effort to acquire data [@axelrodAdvancingArtSimulation1997].
+While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
+Simulating data forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
 It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
-Writing incomplete code is technically impossible, or at least researchers would notice if something is omitted (e.g. if they forgot to aggregate the data).
+Writing incomplete code is technically impossible or at least researchers would notice if they e.g. forgot to aggregate the data etc.
 
@@
 
-To write a PAC, researchers should start by drafting a manuscript in the form of an R Markdown.
+To write a PAC researchers should start by drafting a manuscript in form of an R Markdown.
 When they have completed the method section, they should begin writing code.
-We suggest writing three functions for each hypothesis tested, one for the planned analysis, one for simulating the data and one for reporting the results.
-At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
+We suggest writing three functions for each hypothesis, one for the planned analysis, one for the simulating the data and one for reporting the results.
+At the heart of a PAC lies the planned analysis so researchers should first draft a function called `planned_analysis()` that captures the description in the method section.
 This function should receive the data and derive relevant results from it.
 
-```{r, eval=FALSE}
+```{r}
 planned_analyis <- function(data){
@@
   # 2. conduct analysis e.g. with `t.test()`
-  # 3. return the results
+  # 3. `return(results)`
 }
@@
 
-Because it is difficult to draft such a function without interactively testing it, the researcher should then write a function for simulating data.
-The follow code illustrates the signature of a simple simulation function:
-
-```{r, eval=FALSE}
+Because it is difficult to draft such function without interactively test it, researcher should than proceed to write a function for simulating data.
+
+```{r}
 simulate_data <- function(n, effect_size){
-  # 1. warn users that the results are mocked
+  # 1. warn users that the results are "fake"
   # 2. draw `n` samples with `effect_size`
@@
 
-As soon as they have written a first draft of `planned_anaylsis()` and `simulate_data()` they can interactively improve both functions, e.g.
-until `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
+As soon as they have written `planned_anaylsis()` and `simulate_data()` they can iteratively improve both functions, e.g.
+till `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
 The goal is that the output of `simulate_data()` works as input to the function `planned_analysis()`.
 
-If it is infeasible to simulate realistic data, researcher can delay their preregistration until they have gathered a given percentage of the data.
-Of course, researchers must not analyze these data before they write the preregistration.
-However, they can obfuscate these data by shuffling the dependent variable or otherwise createing synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
-
-When the researchers are satisfied with the function `planned_analysis()`, they can think about how they would like to report the analysis results via tables, plots, and text.
+This process can be tedious and complicated.
+If it proofs infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
+Of course researcher must not analyze the data before they write the preregistration but they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
+
+When the researchers are satisfied with the function `planned_analysis()` they can think about how they would like to report the results of the analysis, via tables, plots and text.
 The implementation of this reporting should be in the function `report_analysis()`.
-<!--AB hier ist mir unklar, wie das genau geht. baue ich hier alle Figures und Tables und Textstücke und benutze die dann erst später? Bitte genauer-->
-
-```{r, eval=FALSE}
-report_analysis <- function(results){
+
+```{r}
+report_results <- function(results){
   # 1. create markdown tables from results
   # 2. conditionally interpret results e.g. if(p < .025)"Result is significant."
-  # 3. visualize results
-  # 4. return results section formatted in markdown
+  # (optional) visualize results
+  # 3. return results section formatted in markdown
 }
@@
 
-This function should again accept the output of `planned_analysis()` as input.
+These function again should be accept the output of `planned_analysis()` as input.
 The output of this function should be formatted in Markdown.
-Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
+A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
 Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it generates not only tables but also a straightforward interpretation of the effects as actual prose (e.g.
-it verbally quantifies the size of an effect).
-
-Ideally, these three functions can be composed to create a mock results section, e.g.
-when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` they output a full results section.
-
-To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration.
-After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data) and write the discussion.
-
-## An example of Preregistration as Code
-
-We argued that PAC has several advantages over classical preregistration and have outlined how to implement it.
-To illustrate how PAC works in practice and help researchers implementing PAC themselves, we have conducted a PAC that serves as a guiding example.
-We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
-
-> "Is there a mean difference in the personality trait 'Machiavellism' between self-identified females and males?"
-
-Again, we propose a preregistration format that closely resembles a classical journal article, but uses simulated data and dynamic document generation to create a document that starts out as a preregistration and eventually becomes the final report.
-The complete preregistration source is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
-In this section, we show code excerpts of this preregistration (formatted in monospace) and explain the rationale behind them.
-
-```{r, echo=FALSE}
-show <- function(x, wrap = TRUE, cat = TRUE, fence = TRUE){
-  stopifnot(length(x) == 1L)
-  if(wrap)x <- str_c(str_wrap(str_split(x, "\\n\\n")[[1]]), collapse = "\n\n")
-  if(!isFALSE(fence)){
-    if(isTRUE(fence))fence <- "markdown"
-    fence <- str_c("````", fence)
-    x <-
-      str_c(fence,
-            x,
-            "````",
-            sep = "\n",
-            collapse = "\n")
-  }
-  if(cat){
-    cat(x)
-    return(invisible(x))
-  } else {
-    return(x)
-  }
-}
-show_prereg <-
-  function(section,
-           file = "preregistration.Rmd",
-           wrap = TRUE,
-           cat = TRUE,
-           fence = TRUE) {
-    prereg <- readr::read_lines(here::here(file))
-    start <- stringr::str_c("<!--begin-", section, "-->")
-    end <- stringr::str_c("<!--end-", section, "-->")
-    start_pos <- which(prereg == start)
-    end_pos <- which(prereg == end)
-    if (length(start_pos) != 1L || length(end_pos) != 1L)
-      stop("No section '", section, "' found.")
-    excerpt <-
-      str_c(prereg[(start_pos + 1):(end_pos - 1)], collapse = "\n")
-    show(excerpt, wrap = wrap, cat = cat, fence = fence)
-  }
-```
-
-As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary background information and literature to understand the context and purpose of the research question.
-This example is drastically shortened for purposes of illustration:
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("theory")
-```
-
-After researchers have provided the research question, they typically proceed to explain how they want to study it.
-For simplicity, we will use already published data that we have not yet analyzed:
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("method")
-```
-
-We choose the following statistical procedure because many researchers are familiar with it[^caution-mww]:
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("ttest")
-```
+The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
+
+Ideally, these three functions can be composed to create a "fake" results section, e.g.
+when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section.
+
+To summarize researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
+After they gathered the actual data, they replace the simulated data, run `planned_analysis()` and write the discussion.
+
+# Preregistration as Code --- a Tutorial
+
+To guide the reader through a preregistration as code (PAC), we will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
+
+> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"
+
+This example serves only didactic purposes and not to derive substantive claims.
+Nevertheless, everything else is kept as close to a real research question as possible.
+
+At the heart of a PAC lies the planned analysis.
+We suggest writing a main analysis function that captures the full process of testing a research question.
+This function receives a dataset as input and performs all statistical processing. <!--AB say anything about its output? is the Markdown the output?-->
+For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
+As an example of a data-dependent decision, we included a rank transformation if the skewness was greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
 
 [^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
-The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
-
-We then translated this methods section into the following `planned_analysis()` function:
+The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
+
+This function is one possible translation of the verbal description above:
 
@@
 
-This function illustrates two advantages of PAC.
-First, a PAC can easily include data-dependent decisions creating different analysis branches under different conditions.
-Second, it highlights how difficult it is to describe a statistical analysis precisely.
-The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and their expectations and beliefs.
-One example would be using the function `wilcox.test` instead of the combinations of the functions `rank` and `t.test`.
-Either of them are valid implementations of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
-In contrast, the second applies Welch's correction by default and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
+Importantly, natural language descriptions do not map unambiguously onto computer code implementations. 
+The same textual descriptions may be implemented differently by different persons depending on their statistical and programming knowledge but also based on their own expectations and beliefs.
+One example would be using the function `wilcox.test` instead of the combinations of `rank` and `t.test`.
+Both are valid implementation of the Mann–Whitney–Wilcoxon test but the first assumes equal variance whereas the second applies Welch's correction and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
 Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
-Still, these details can make a difference and provide unwanted researchers' degrees of freedom.
-
-The function `planned_analysis()` together with the function `simulate_data()` (not shown here) can be used to justify the planned sample size.
-To that end, `simulate_data()` is repeatedly called with increased sample size and the proportion of significant results (power) is recorded. 
-The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
-The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R).
-
-Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
+Still, these details can make a difference and provide unwanted researcher degrees-of-freedom.
+
+<!--AB here is kind of a break in the text-->
+To use this function one needs a compatible dataset. <!--AB what is "this" function? What does "compatible" mean?-->
+At the time of preregistration there is of course no actual data available.
+We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
+Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
+The following code simulates a dataset with one observed variable in two groups.
+
+```{r simulate_data}
+```
+<!--AB it is odd that the return variable is called effect because it is a dataset-->
+
+Here, we decided to randomly draw non-normal data to test how well our procedure performed under non-optimal situations.
+Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a given standardized effect size (a Cohen's d of .2).
+This simulated dataset fits perfectly into our planned analysis with a sample size of 400 and 10 items: <!--AB in what way? please explain-->
+
+***
+
+```{r}
+set.seed(112358) # fix state of RNG
+results <- 
+  planned_analysis(
+    simulate_data(
+      n = 400,# Sample size
+      df = 8, # Chisq with DF = 8 has skewness = 1
+      d = .2, # Cohen's D
+      i = 10  # Number of items (should not effect anything)
+    )
+  )
+```
+
+Next, we add dynamic code blocks to our Markdown report that implement the planned reporting of the final results.
+A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
+Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
+The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
+Here we use the `report` package to create a preview of how our reporting will look like based on simulated data:<!--AB add citation to the report package-->
+
+```{r, results='asis', eval=FALSE}
+report::report(results$test)
+```
+
+This call to the `report()` function will yield the following automatically generated output:
+
+```{r, results='asis', echo=FALSE}
+report::report(results$test)
+```
+<!--AB it is difficult to discern the automatically generated text and out manuscript text even though there is the horizontal line. Not sure how to improve this.-->
+***
 
@@
 
-```{r, echo=FALSE, results='asis'}
-show_prereg("power-code", wrap = FALSE)
-```
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("power")
-```
-
-We implemented a mechanism that only uses simulated data when the actual data is not yet available for the results section.
-This mechanism also warns readers if these results are based on simulated data. <!--AB how does it warn?-->
-As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study. <!--AB how exactly does the script know?-->
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("results", wrap = FALSE)
-```
-
-The last thing we need to do is to report our results with the combination of the functions `planned_analysis()` and `report_analysis()`.
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("report", wrap = FALSE)
-```
-
-This is an example of how the results could be reported (based on simulated data):
-
-```{r, echo=FALSE, results='asis'}
-set.seed(1234)
-show(report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)), cat = FALSE), fence = "")
-```
-
-The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
-
 # Discussion
@@
 
-PAC comes with a challenge.
+PACs come with a challenge though.
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
@@
 Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
+Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
 
@@
 
-Please note that the statistical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and is not necessarily optimal for addressing your research question.
+<!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->
+Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
 The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-
-We also suggest that version control with Git and GitHub fulfills the criteria of a preregistration as long as you state your research plan comprehensively.
-This view is not shared by everyone [@szollosi_is_2020] because with some effort and malice intend the record of Git/GitHub can be altered.
-This concern can be addressed by using zenodo.org to give a digital object identifier (DOI) to your preregistration (see [this guide](https://guides.github.com/activities/citable-code/)) because zenondo.org saves an unchangeable copy of your repository.
+<!--AB can we move this to a different place?
+Maybe this would actually better fit to the discussion/limitation section!-->
 

### 2021-07-29 manuscript.Rmd
start incorporating Andreas' draft discussion

@@
 simplesummary: |
-  Reproducibility has long been considered integral to the scientific method.
-  Something is called reproducible when independend researcher arrive at the same results from the same data.
-  Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
-  Technological advancements now enable scientists to achieve a much higher standard;
-  one in which any individual can be granted access to a digital research repository, and reproduce the analyses from raw data to final manuscript with a single command.
-  This method has far-reaching implications for scientific archiving,
-  reproducibility and replication,
-  scientific productivity,
-  and the credibility and reliability of scientific findings.
-  One obstacle preventing the widespread adoption of this method is that the underlying technological advancements
-  are complicated to use.
-  This paper introduces `repro`; an R-package that guides researchers in the installation and use of the tools required for computational reproducibility.
-  It includes a tutorial on how to make a research project reproducible with the aid of `repro`.
-  Finally, it introduces novel applications of reproducibility tools for scientific research;
-  including the preregistration of study plans as reproducible computer code.
-  This circumvents the shortcomings of ambiguous preregistrations that allow for researcher degrees of freedom,
-  because computer code describes study plans more precisely than prose.
-  Reproducibility made convenient with automation has hence a wide range of applications to accelerate scientific progress.
+  A Simple summary goes here.
 abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
@@
 
-One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be time-intensive.
-To lower the threshold to adoption, the R-package `repro` introduces helper functions that wrap more complicated (and powerful) tools.
-These so-called wrappers guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user should do in layman's terms.
-We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
-These wrappers are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
-
-This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
-It describes basic steps that we expect to apply to all users.
-Your specific situation might involve additional, more specialized steps.
-After completing the tutorial, you should be able to customize your workflow accordingly.
-The first step is to install the required software.
-Start R Studio<!--CJ I am concerned that new users may not install R if you only mention RStudio.--> and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
-It will assist you while you follow the tutorial.
+The above mentioned tools require much expertise that is beyond the traditional curriculum of researchers.
+To aid researcher adopting these tools we rely on or have developed helpers that wrap more complicated (and powerful) tools.
+These wrappers use standard tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
+We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
+These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
+
+Before we can start creating reproducible research projects, we need a computer with a working R Studio installation. 
+Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so that it can act as an assistance systems while you follow the tutorial:
 
@@
 # repro is not on CRAN yet
-options(
-  repos = c(aaronpeikert = 'https://aaronpeikert.r-universe.dev',
-            CRAN = 'https://cloud.r-project.org')
-)
-install.packages('repro')
+install.packages('repro', repos = 'https://aaronpeikert.r-universe.dev')
 ```
@@
 An example of the assistance that `repro` provides are the check functions.
-These verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
+They verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
 
@@
 
-These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
+They detect the users' system state and explain if further action is needed.
 Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:
@@
 
-The messages from `repro` try to take the user by the hand to help them solve problems.
-These messages are adjusted to your specific operating system and installed dependencies.
-Before you continue, we ask you to run the above commands to check Git, Make, and Docker - both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
-If necessary, follow any instructions presented until all checks pass.
-
-## Implementing version control
-
-Now that your system is set up for this tutorial, we will introduce you to version control with Git.
+The messages emitted by `repro` and other packages we present try to take the user by the hand to help them solve problems.
+This tutorial describes steps that we expect to apply to all users; the messages from the packages are adjusted to your specific situation (that is, operating system and the specific problem of missing software) and complement the tutorial in this way.
+Before you continue, we ask you to run the above commands to check Git, Make, and Docker.
+If necessary, follow the instructions of repro to install them until the check functions are all fine.
+
+The first concept we want to introduce is version control with Git.
 Git tracks files within one project folder.
@@
 
-Git tracks any files that you manually add to the repository.
-To make sure you do not accidentally add files that you do not wish to share (e.g., privacy sensitive data),
-you can list specific files that you do not want to track in the `.gitignore` file.
+You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore` file.
 You can add something to the `.gitignore` file directly or with this command:
@@
 ```
+
 Now the file `private.md` will not be committed to Git and will not be made public.
 
-Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
-When you are ready, you can begin tracking your files using Git by running:
+After you have excluded sensitive files (like not yet anonymized data files), you can activate Git with:
 
@@
 For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
-One way to do this is through the visual user interface in the RStudio Git pane.
-Click on the empty box next to the file you want to stage. A check mark appears to indicate that the file is staged.
+The easiest way to do this is to use the RStudio Git pane.
+Click on the empty box next to the file you want to stage, and when a tick appears, which means that the file is staged.
 After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
-This stores a snapshot of the current state of the project.
 
@@
 All snapshots are stored in a local repository in your project folder.
-To back up and/or share the files online, you can push your local repository to a remote repository.
-While you can choose any Git service (like GitLab or BitBucket), we will use GitHub in this tutorial.
-Before you upload your project to Git, you need to decide if you would like the project publically accessable (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
-To upload the project publically to GitHub use:
+To back up and share the files, you can publicly upload those changes to a public repository.
+To synchronize your local repository with a remote repository from the GitHub service, issue the following command:
 
@@
 
-To upload it privately:
+Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the R console.
+
+If you are not yet ready to share your project publicly but you want to synchronize your local repository to a remote repository (for example, as a simple backup service), use:
 
@@
 
-Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the R console.
-
-## Dynamic document generation
-
-Now that you have created a version controlled project, we will proceed with dynamic document generation.
-A dynamic document has three elements:
-
-1. Text in natural language (prose; e.g., a scientific paper or presentation)
-2. Executable code (e.g., analyses)
-3. Metadata (e.g. title, authors, document format)
-
-R Markdown is a type of dynamic document well-suited to the RStudio user interface.
-The natural language text in an R Markdown file can be formatted using Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
-The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
+Now that you have created a project that is under version control, we continue with adding dynamic document generation.
+Adynamic document has three elements:
+
+1. Text in a natural language
+2. Executable code 
+3. Metadata
+
+The text of an R Markdown is written in Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
 The following example serves to illustrate the Markdown syntax. 
@@
 
-One advantage of this type of markup for formatting is that it can be rendered to many different output formats - both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g.
-specific journal requirements.
-Another advantage is that, because R Markdown files are in plain text, they are more suitable for version control using Git.
-Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass <!--CJ Add picture-->in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
-The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
-and easing the generation of citations and references to Tables and Figures.
-
-Now that you are familiar with the basics of Markdown formatting, we turn our attention to including code into the text.
+This type of markup for formatting, instead of the more visual approach of e.g. Microsoft Word, is compatible with many output formats and is easier to put under version control.
+Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
+The visual editor has the additional benefit of promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions).
+Among other features, it eases the generation of citations and cross references.
+
+Now that you are familiar with the basics of writing text in Markdown, we turn our attention to including code into the text.
 Code is separated by three backticks (and the programming language in curly braces) like here:
@@
 ````
-The hotkey [Control]+[Alt]+[i] inserts a block of code in the file.
-The results of code enclosed in such backticks will be dynamically inserted into the document (depending on specific settings).
-This means that, whenever you render the R Markdown to its intended output format, the code will be executed and the results updated.
-The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
-
-Once the R Markdown file has been rendered to a static document, the resulting file is decoupled from the R Markdown and the code that created it.
-This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
-To avoid ambiguity, we therefore recommend referencing the identifier of the Git commit at the time of rendering in the static document.
-Simply put, a static document should link to the version of the code that was used to create it.
-The `repro` package comes with the function `repro::current_hash()` for this purpose.
+
+The results of your code are than dynamically inserted into the document.
+Dynamically means that whenever you turn the R Markdown into its output format the code is executed again and if you changed something the results are updated.
+The output document is than static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
+This static document is somewhat decoupled from the R Markdown and the rest of the code that did create it.
+It is quite possible that several slightly different versions are therefore circulating.
+We therefore recomend to connect the static document to the version control system by dynamically including the identifier of the identifier of the Git commit, called hash, in the document.
+The `repro` package comes with the function `repro::current_hash()` for that purpose.
 This document was created from the commit with the hash ``r repro::current_hash()``.
 
-Now that you know how to write text and R code in an R Markdown, you need to know about metadata (also called: YAML front matter).
-These metadate contain information about the document, like the title and the output format.
+Now that you know how to write text and R code in an R Markdown, you need to know about metadata, like the title or the output format.
 Metadata are placed at the beginning of the document and are separated by three dashes from the document body; the following example is a full markdown document where the header is in lines 1--6.
@@
 title: A Hitchhiker's Guide to Reproducible Research in R
-author: Aaron Peikert, Caspar J. van Lissa and Andreas M. Brandmaier
+author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
 abstract: A tutorial about doing the same thing more than once.
@@
 
-Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
-However, R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
-Some R packages add additional metadata fields.
-The `repro` package adds fields to the metadata to help make the research project reproducible.
-These fields are used to list all dependencies of the research project.
-This includes R scripts, data files, and external packages.
-The format is as follows (see everything below the line *repro:*):
+Most metadata fields are self-explanatory (like the author field). 
+Sometimes, R package may extend the metadata fields for their own purposes.
+Certain output formats may require specific  metadata. 
+For example, the `rticles`-package that offers templates for authoring scientific manuscripts comes with extra fields that are described in the package documentation and offer extra options for flexible formatting of documents.
+R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
+The `repro` package uses the metadata for hints to make the research project reproducible.
+Researchers should list all dependent resources of their research project in the metadata, that is, R scripts, data files, and external packages in the following format (see everything below the line *repro:*):
 
@@
 
-This information makes explicitly declares the project's dependencies on files and R packages.
-`repro` uses this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
-These two files together form the basis for consistency within a research project and consistency across different systems.<!--CJ Consider explaining this in the theoretical intro-->
-At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.<!--CJ This comes way to late.
-Shouldbe in the opening paragraph or something.-->
+This information makes internal dependencies to other files in the project and external dependencies to R packages explicit.
+`repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
+These two files together form the basis for consistency within a research project and consistency across different systems.
+At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.
 This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
-Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).<!--CJ These points describe the formal definition of reproducibility used in this paper, and the goal of the workflow.
-That needs to be explained in the introduction.
-If you only find out what you are working toward halfway through the tutorial - that's too late!-->
-
-The function `repro::automate()` converts the metadata from all R Markdown files in the project (all files with the ending `.Rmd`) to a Makefile and a Dockerfile.
-Together, these files allow users (including your future self) to reproduce every step in the analysis automatically.
-It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, to keep the Makefile and Dockerfile up to date.
-The function `repro::automate()` does not overwrite manual changes<!--CJ What does that mean?-->, and there is no harm in running it too often.<!-- CJ Can that not be automated?
-Or at least provide a suggestion when to run it?-->
+Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).
+In everyday work, the simplest solution is to now call the function `repro::automate()` which both creates a Makefile and a Dockerfile from all R Markdown files in the project (all files with the ending `.Rmd`) based on the respective repro metadata entries.
+Every time you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
+The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
 Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
 <!--AB: why is this addition not automated?
-Feels strange given that this is our recommended best practice?!
-CJ: Agreed.-->
-
-<!-- CJ What i'm missing here is something to wrap up the tutorial. Do users now have a project that could be reproduced? The tutorial should not end until that is the case.-->
-
-## Reproducing a project
-
-<!--CJ Suggestion: Instead of explaining reproduction right in the middle of the tutorial for setting up a project, add a later section called "reproducing a project", and provide an example project for users to reproduce.-->
+Feels strange given that this is our recommended best practice?!-->
+
 When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:
@@
 
-Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.<!--CJ This belongs in the intro, not the tutorial.
-It is part of explaining the benefits of repro.-->
+Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
 These projects will remain reproducible irrespective of whether `repro` will be maintained and remain accessible in future.
 Users do not have to have repro installed to reproduce a project, in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
-The only required software for reproducing a project is Docker, assuming users know how to build a Docker image and run Make within the container.
-However, if they have installed Make additionally to Docker, they do not need to know how to use Docker and can simply rely on the two Make commands `make docker` and `make -B DOCKER=TRUE`.
+The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.
 The commands can be pasted into a terminal and will reproduce the project in many cases.
+<!--AB: I said this before.
+I think it really should be possible to execute this command from within R.
+Never underestimate the troubles people can have with the terminal.
+I think we discussed this before.
+Please add a function repro::rerun(execute=TRUE) or -- if we rename it -- repro::reproduce(execute=TRUE).
+-->
 
@@
 
-This section is only for advanced users who want to overcome some limitations of `repro`.
-If you read this paper the first time you probably want to skip this section.
-As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.
-The simplification of Docker and Make are possible because of two restrictions.
-Users who ask themselves either "How can I install software dependencies outside of R in the Docker image?" or "How can I express complex dependencies between files, e.g.
-hundreds of data files are preprocessed and combined?") need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
-Other users may safely skip this section, or return to it if they encounter such challenges.
+Some projects, however, are not as easy to make reproducible.
+As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
+These simplifications are possible because of two restrictions.
+Users who ask themselves either "How can I install software that is not an R package in the Docker image?" or "How can I express complex dependencies between files, e.g.
+raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
+All other users may safely skip this section and may return to it if they become limited by restriction.
 
@@
 For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
-Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
-With PAC researchers can precisely calibrate how much they want to explore.
-When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
 Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
@@
 
-<!--AB: TODO add a quick summary -->
+# Discussion
+
+## Summary
+
+We demonstrated how the R package `repro` supports researchers in creating reproducible repositories and reproducible manuscripts.
+These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
+The workflow we present here rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to all but guarantee reproducibility.
+
+We then proceeded to show preregistration as code (PAC) as one example of how such rigorous and automated reproducibility workflow may enable other innovations.
+In a PAC researcher write all analysis code and a reproducible manuscript that already renders the entire reporting of statistical and computational results before they gather data.
+These results are based on simulated data, such that there is no danger of peeking at the real data while writing the scripts.
+This way, all researcher's degrees of freedom are eliminated.
+Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
+Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
+To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything is under version control.
+The likelihood of those deviations, while unavoidable, is decreased by using simulated data to verify the suitability of the analysis plan.
+This analysis plan may contain data depended decisions to further decrease the likelihood that researcher are unable stick to their preregistration
+
+## Limitations of repro
+
+`repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
+While we have invested much effort to ensure this goal, the package is still in early development and undoubtedly has rough edges and even bugs that went undetected.
+We hope that despite the early development stage, `repro` can support researchers to work more reproducible.
+
+The workflow we present here introduces several software solutions which in itself presents a danger to reproducibility.
+To profit from automatic and convenient reproduction researcher must install Git, Make, and Docker.
+However, Git and Make are itself included in the Docker image created by `repro`.
+Researchers can therefore employ the Docker image manually to download the Git repository and execute Make for full reproduction.
+In other words the only requirement for reproduction and therefore Achilles' heel for this workflow is Docker.
+The Docker approach has two vulnerabilities.
+First, and more important, the Docker image for the project has to remain available.
+The Dockerfile on the other hand is insufficient because it relies on to many service providers (e.g. Microsoft R Application Network, Ubuntu Package Archive).
+The Docker image can be saved to disk and treated like any other file that requires careful archivation.
+If the existence of the Docker image is guaranteed, researchers still require software to run the image.
+To that end they can either rely on Docker itself or Docker-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
+The only way to remove the reliance on such external software is to turn the Docker image into a full operating system that subsequently can be installed and run on almost any modern computer.
+This process is technically possible and would guarantee reproducibility for decades.
+However, this process requires much technical knowledge and is currently not supported by `repro`.
+In any case and without such precautions do the R Markdown, the Makefile and the Dockerfile provide information that allow researchers to manually trace the computational steps and recreate the computational environment.
 
@@
 
-PACs come with a challenge though.
+We realize that PACs are challenging.
+Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove unwieldy for a PAC.
+Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not preregistered research but as a compliment.
+Some may see the rigidness of PAC as to much of a good thing.
+However, we are convinced that many reasons justify changing the preregistered code, but we are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
+
+Another challenge is that PAC requires more programming expertise than a usual analysis.
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
 We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
-This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
+This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
 The reproducible workflow we present here is heavily automated and hence promotes reusability.
-Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
+Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
+Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
 
@@
 
-<!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->
-Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
+Please note that the analytical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes, and it is not necessarily optimal for addressing your research question.
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
-The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-<!--AB can we move this to a different place?
-Maybe this would actually better fit to the discussion/limitation section!-->
+The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
 
@@
 
-<!--AB: two lines of what the future of the package, of such standards, and prereigstration is.
-Maybe say something that preregistration is not always possible or wanted but that reproducibility is almost always a good idea?-->
+Open science practices are a continually evolving field where technical innovations foster changes in research practice.
+Open data is much more widespread thanks to online storage facilities; preregistration are possible because there are preregistration platforms and so forth.
+Similarly, we hope that fully automatic reproduction, e.g. with `repro` as technical innovation, increase scientific rigour and efficiency.
+
+This ideal of a fully automatic reproduction of research projects meets in practice a wide range of demands on how user-friendly and how powerful the software solutions are.
+Some may find Make too complicated or that Docker requires too much storage.
+Yet others may find that they require other programming languages or scale their computation across hundreds of computers, e.g. via high-performance computing clusters or cloud computing.
+
+`repro` was designed modularly to meet such demands.
+At the moment, `repro` only supports the combination of RMarkdown, Git, Make, and Docker.
+However, there are alternatives for each of these elements that may fit better into an individual research project.
+RMarkdown could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer`[@officer] or `officedown`[@officedown] to accommodate a wider arrange of journal submission standards.
+Instead of using formal version control with Git, `repro` could automatically save snapshots for increasing user-friendliness.
+Make could be replaced by the more R-centred alternative `targets` for more convenience.
+Docker could be combined with `renv` [@R-renv] to control the package versions precisely (our approach fixes the date, `renv` the exact package version).
+Alternatively, Docker could be replaced by the more lightweight `renv` if no dependencies outside of R are considered crucial.
+Docker does not satisfy the requirements of many HPC environments, but Singularity was designed to avoid this limitation while still being compatible with Docker images.
+
+`repro`'s modular structure allows such alternative workflow, though they have not been implemented.
+Depending on the demand by users, we will implement some of these alternative workflows in `repro`.
 

### 2021-07-30 manuscript.Rmd
fix typo
revise discussion

@@
 
-# Discussion
-
 ## Summary
@@
 In a PAC researcher write all analysis code and a reproducible manuscript that already renders the entire reporting of statistical and computational results before they gather data.
-These results are based on simulated data, such that there is no danger of peeking at the real data while writing the scripts.
+These results are based on simulated data, so that there is no danger of peeking at the real data while writing the scripts.
 This way, all researcher's degrees of freedom are eliminated.
@@
 Researchers can therefore employ the Docker image manually to download the Git repository and execute Make for full reproduction.
-In other words the only requirement for reproduction and therefore Achilles' heel for this workflow is Docker.
+In other words the only hard requirement for reproduction and therefore Achilles' heel for this workflow is Docker.
 The Docker approach has two vulnerabilities.
 First, and more important, the Docker image for the project has to remain available.
-The Dockerfile on the other hand is insufficient because it relies on to many service providers (e.g. Microsoft R Application Network, Ubuntu Package Archive).
+The Dockerfile (the plain text description to build a Docker image) on the other hand is insufficient because it relies on to many service providers (e.g.
+Microsoft R Application Network, Ubuntu Package Archive).
 The Docker image can be saved to disk and treated like any other file that requires careful archivation.
-If the existence of the Docker image is guaranteed, researchers still require software to run the image.
+However, if the existence of the Docker image is guaranteed, researchers still require software to run the image.
 To that end they can either rely on Docker itself or Docker-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
 The only way to remove the reliance on such external software is to turn the Docker image into a full operating system that subsequently can be installed and run on almost any modern computer.
-This process is technically possible and would guarantee reproducibility for decades.
+This process is technically possible and would guarantee reproducibility for decades without any software dependency.
 However, this process requires much technical knowledge and is currently not supported by `repro`.
 In any case and without such precautions do the R Markdown, the Makefile and the Dockerfile provide information that allow researchers to manually trace the computational steps and recreate the computational environment.
+The Makefile, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are, for whatever reason, unable to run Make.
+Similarly, even when Docker does not work, the Dockerfile still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
 
@@
 Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove unwieldy for a PAC.
-Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not preregistered research but as a compliment.
-Some may see the rigidness of PAC as to much of a good thing.
-However, we are convinced that many reasons justify changing the preregistered code, but we are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
+Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not preregistered research but as a complement.
+The rigidness of PAC is only than to much of a good thing, if one holds the view that researchers can for no reason deviate from their preregistration.
+However, we are convinced that many reasons justify changing the preregistered code, but are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
+PAC enables just that.
+Researcher can change their plans and possibly defend their decision, e.g. in the discussion, while readers of the manuscript can clearly see what was changed through version control.
 
@@
 Instead of using formal version control with Git, `repro` could automatically save snapshots for increasing user-friendliness.
-Make could be replaced by the more R-centred alternative `targets` for more convenience.
+Make could be replaced by the more R-centered alternative `targets` for more convenience.
 Docker could be combined with `renv` [@R-renv] to control the package versions precisely (our approach fixes the date, `renv` the exact package version).

### 2021-07-30 manuscript.Rmd
add definition of reproducibility

--- 2021-07-29 manuscript.Rmd
@@
 bibliography: temp.bib
+appendix: appendix.tex
 simplesummary: |
@@
   because computer code describes study plans more precisely than prose.
-  Reproducibility made convenient with automation has hence a wide range of applications to accelerate scientific progress.
+  <!-- CJ Add concluding sentence(s) -->
 abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
@@
 * do more research
-* spend time with family and friends
+* spend time with family
 ```
 
-One advantage of this type of markup for formatting is that it can be rendered to many different output formats - both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g.
-specific journal requirements.
-Another advantage is that, because R Markdown files are in plain text, they are more suitable for version control using Git.
-Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass <!--CJ Add picture-->in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
-The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
-and easing the generation of citations and references to Tables and Figures.
-
-Now that you are familiar with the basics of Markdown formatting, we turn our attention to including code into the text.
 Code is separated by three backticks (and the programming language in curly braces) like here:
@@
 ````
-The hotkey [Control]+[Alt]+[i] inserts a block of code in the file.
-The results of code enclosed in such backticks will be dynamically inserted into the document (depending on specific settings).
-This means that, whenever you render the R Markdown to its intended output format, the code will be executed and the results updated.
-The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
 
-Once the R Markdown file has been rendered to a static document, the resulting file is decoupled from the R Markdown and the code that created it.
-This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
-To avoid ambiguity, we therefore recommend referencing the identifier of the Git commit at the time of rendering in the static document.
-Simply put, a static document should link to the version of the code that was used to create it.
-The `repro` package comes with the function `repro::current_hash()` for this purpose.
-This document was created from the commit with the hash ``r repro::current_hash()``.
-
-Now that you know how to write text and R code in an R Markdown, you need to know about metadata (also called: YAML front matter).
-These metadate contain information about the document, like the title and the output format.
-Metadata are placed at the beginning of the document and are separated by three dashes from the document body; the following example is a full markdown document where the header is in lines 1--6.
+Now that you know how to write text and R code in an RMarkdown, you need to know about metadata, like the title or the output format.
+Metadata are placed at the beginning of the document and are separated by three dashes; the following example is a full markdown document where the header is in lines 1-6.
 
@@
 title: A Hitchhiker's Guide to Reproducible Research in R
-author: Aaron Peikert, Caspar J. van Lissa and Andreas M. Brandmaier
-abstract: A tutorial about doing the same thing more than once.
+author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
+abstract: A paper about recalculating the same thing more than once.
 output: html_document
@@
 
-Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
-However, R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
-Some R packages add additional metadata fields.
-The `repro` package adds fields to the metadata to help make the research project reproducible.
-These fields are used to list all dependencies of the research project.
-This includes R scripts, data files, and external packages.
-The format is as follows (see everything below the line *repro:*):
+Most metadata fields are self-explanatory (like the author field), but any package may use the metadata for their purposes.
+Often, for example, does a specific output format require more metadata, i.e. the `rticles`-package for authoring scientific manuscripts.
+RMarkdown was extended to hundreds of document formats which are often highly configurable by their metadata.
+The `repro` package uses the metadata, not to change the output, but to make the document more reproducible, simply by listing resources, i.e., scripts, data, and packages, your document depends on, in the following format:
 
@@
 
-This information makes explicitly declares the project's dependencies on files and R packages.
-`repro` uses this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
-These two files together form the basis for consistency within a research project and consistency across different systems.<!--CJ Consider explaining this in the theoretical intro-->
-At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.<!--CJ This comes way to late.
-Shouldbe in the opening paragraph or something.-->
-This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
-Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).<!--CJ These points describe the formal definition of reproducibility used in this paper, and the goal of the workflow.
-That needs to be explained in the introduction.
-If you only find out what you are working toward halfway through the tutorial - that's too late!-->
+This information makes internal dependencies to other files in the project and external dependencies to R packages explicit.
+`repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
+In day-to-day work the simplest solution is to use the function `repro::automate()` which both infers a Makefile and a Dockerfile from all RMarkdown  (all files with the ending `.Rmd`) and there respective repro metadata.
+Every time you change the `repro`-metadata, change the output format, or add a new RMarkdown, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
+The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
+Besides the main Makefile and the Dockerfile repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
 
-The function `repro::automate()` converts the metadata from all R Markdown files in the project (all files with the ending `.Rmd`) to a Makefile and a Dockerfile.
-Together, these files allow users (including your future self) to reproduce every step in the analysis automatically.
-It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, to keep the Makefile and Dockerfile up to date.
-The function `repro::automate()` does not overwrite manual changes<!--CJ What does that mean?-->, and there is no harm in running it too often.<!-- CJ Can that not be automated?
-Or at least provide a suggestion when to run it?-->
-Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
-<!--AB: why is this addition not automated?
-Feels strange given that this is our recommended best practice?!
-CJ: Agreed.-->
-
-<!-- CJ What i'm missing here is something to wrap up the tutorial. Do users now have a project that could be reproduced? The tutorial should not end until that is the case.-->
-
-## Reproducing a project
-
-<!--CJ Suggestion: Instead of explaining reproduction right in the middle of the tutorial for setting up a project, add a later section called "reproducing a project", and provide an example project for users to reproduce.-->
-When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:
+When you or people who received your project from GitHub want to reproduce the project, they can ask repro to explain how this should be done:
 
 ```{r}
-repro::reproduce()
+repro::rerun()
 ```
 
-If you uncomfortable with using the terminal directly, you can sent the command to the terminal from within R:
-
-```{r, eval=FALSE}
-system(reproduce())
-```
-
-Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.<!--CJ This belongs in the intro, not the tutorial.
-It is part of explaining the benefits of repro.-->
-These projects will remain reproducible irrespective of whether `repro` will be maintained and remain accessible in future.
-Users do not have to have repro installed to reproduce a project, in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
-The only required software for reproducing a project is Docker, assuming users know how to build a Docker image and run Make within the container.
-However, if they have installed Make additionally to Docker, they do not need to know how to use Docker and can simply rely on the two Make commands `make docker` and `make -B DOCKER=TRUE`.
-The commands can be pasted into a terminal and will reproduce the project in many cases.
-
-# Advanced Features
-
-This section is only for advanced users who want to overcome some limitations of `repro`.
-If you read this paper the first time you probably want to skip this section.
-As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.
-The simplification of Docker and Make are possible because of two restrictions.
-Users who ask themselves either "How can I install software dependencies outside of R in the Docker image?" or "How can I express complex dependencies between files, e.g.
-hundreds of data files are preprocessed and combined?") need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
-Other users may safely skip this section, or return to it if they encounter such challenges.
-
-The metadata the `repro::automate()` function relies on can only express dependencies R packages for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
-The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
-Other software that users might need like other programming languages, not already installed LaTeX packages etc. have to be added manually.
-We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g.
-for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
-The second limitation is about dependencies.
-Make can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
-If B is missing in this example, Make would know how to recreate it from C and D.
-These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
-Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D therefore leaving out important information or express the dependencies directly within the Makefile.
-
-The following section explains how to overcome these limitation while still relying on the automatons afforded by `repro`.
+However, users do not have to have repro installed, in fact, they do not even need R installed. 
+The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.
+They can be pasted into a terminal and will reproduce the project in many cases.
+Some projects, however, are not as easy to make reproducible.
+As explained above, repro is merely a simplified interface.
+These simplifications are possible because of two restrictions.
+First, the project can only rely on R packages for the Dockerfile and second, it can have no multilayered preprocessing for the Makefile.
 Lifting these restrictions requires the user to interact more directly with Make or Docker.
@@
 
-Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
-Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
-Preregistrations, however, have a problem.
-Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
-The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
-The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
-But these templates can not remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
+We argue that a preregistration as code (PAC) is an excellent planning tool that offers several benefits over traditional preregistration.
+For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data using simulated data.
 
-Our proposed solution is simple:
-Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
-We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
-PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration despite the best efforts afford.
+We consider three criteria when planning a study.
+First, the plan should be comprehensive, e.g. state the research question and describe the study design and analysis.
+Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
+Third, the plan should be efficient; therefore, it only employs the necessary amount of resources.
 
-The idea of submitting code as part of a preregistration is not new.
-A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
-In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
+To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript.
+The introduction and theoretical background provide the basis for the studies design and can be written before gathering data.
+We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
+To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
+The researchers can include a "faked " results section based on the simulated data, employing dynamic document generation.
+Repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to assure efficiency.
+For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study.
 
-We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.
+Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [https://www.pnas.org/content/115/11/2600].
+Preregistration increases the credibility of empirical results in three ways [preregistration is hard and worthwhile, Nosek].
 
-## Why are PAC and reproducibility related?
+> First, preregistration of analysis plans makes clear which analyses were planned a priori and which were conducted post hoc.
+This improves calibration of uncertainty for unplanned analyses, and diagnosticity of statistical inferences for planned analyses.
+[...] Doing so has the benefit of strengthening statistical inferences as compared with unplanned analyses.
+[...] Second, preregistration enables detection of questionable research practices such as selective outcome reporting (John et al.
+2012) or Hypothesizing After the Results are Known (HARKing; Kerr, 1998).
+Third, preregistration of studies can reduce the impact of publication bias—particularly the prioritization of publishing positive over negative results—by making all planned studies discoverable whether or not they are ultimately published.
 
-First, preregistrations are plagued by non-reproducibility like any other research outcome.
-@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to reproduce the power analysis.
-Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
-For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
+A preregistration as code offers four advantages over classical preregistration.
+First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
+Second, despite being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
+Researchers can, for example, formulate conditions under which they prefer one analysis over the other, i.e.
+if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
+Third, deviations from the preregistration are more explicit because they are reflected in changes to the codes a property facilitated by version control.
+Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
 
-Second, although it has been almost a decade since @steegenMeasuringCrowdAgain2014 have preregistered their code, the idea of publishing code before analyzing data has gained not much traction.
-This could change when combining the here proposed workflow and PAC.
-We want to extend PAC from preregistering code to preregistering *the whole manuscript as a dynamic document*.
-The idea is as follows:
-For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
-In this document, the results are initially based on simulated data that mimics the expected output.
+# Preregistration as Code --- an example
 
-Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
-It follows that for preregistrations they should stick to widely accepted standards of scientific writing [e.g. @apa7].
-We have already explained the need for reproducibility standards in the process of writing such manuscript and it is therefore only natural to extend these arguments to preregistration.
-
-## Advantages of PAC over traditional preregistration
-
-We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, a PAC removes any ambiguity regarding the translation of an verbal analysis plan into code.
-Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
-Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
-For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
-Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
-With PAC researchers can precisely calibrate how much they want to explore.
-When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
-Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
-We believe it is easier with PAC to detect missing steps or ambiguous decisions.
-Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
-Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
-Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.
-
-## Benefits of PAC for study planning
-
-Leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
-To use PAC for study planing, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
-To that end, they should borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
-Then they should translate their method section into executable computer code and simulate data that has the same structure as the data they expect to obtain from their experiments.
-
-We consider three domains where this procedure can help planning a study.
-First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
-Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
-Third, the plan should be efficient; the plan should only require the necessary amount of resources.
-
-### Comprehensiveness
-
-Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
-One particular advantage is the strong foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
-Nevertheless it can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section.
-
-Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-Simulating data forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
-It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
-Writing incomplete code is technically impossible or at least researchers would notice if they e.g. forgot to aggregate the data etc.
-
-### Effectiveness
-
-With simulated data researchers can check if the analysis is technically complete and can also recover the true parameters of the simulation.
-When an analysis fails to recover the correct values in idealized conditions with big samples researchers should not expect this analysis to work under realistic conditions.
-Because the simulation captures the researchers expectations about the real world conditions it is the best way to assess the effectiveness of the analysis without actually gathering the data.
-
-### Efficiency
-
-A single simulated or synthesized dataset already helps to ensure that the plan is viable under certain circumstances.
-However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
-Monte Carlo simulation lets researcher navigate several trade offs regarding their resources.
-In the simplest case they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
-@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.
-You may also evaluate how the data analysis copes with expected data complications, e.g.
-non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
-
-## What you need to know before writing a PAC
-
-To write a PAC researchers should start by drafting a manuscript in form of an R Markdown.
-When they have completed the method section, they should begin writing code.
-We suggest writing three functions for each hypothesis, one for the planned analysis, one for the simulating the data and one for reporting the results.
-At the heart of a PAC lies the planned analysis so researchers should first draft a function called `planned_analysis()` that captures the description in the method section.
-This function should receive the data and derive relevant results from it.
-
-```{r}
-planned_analyis <- function(data){
-  # 1. preprocess e.g. with `rowMeans(data)`
-  # 2. conduct analysis e.g. with `t.test()`
-  # 3. `return(results)`
-}
-```
-
-Because it is difficult to draft such function without interactively test it, researcher should than proceed to write a function for simulating data.
-
-```{r}
-simulate_data <- function(n, effect_size){
-  # 1. warn users that the results are "fake"
-  # 2. draw `n` samples with `effect_size`
-  # 3. format and return in expected data format
-}
-```
-
-As soon as they have written `planned_anaylsis()` and `simulate_data()` they can iteratively improve both functions, e.g.
-till `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
-The goal is that the output of `simulate_data()` works as input to the function `planned_analysis()`.
-
-This process can be tedious and complicated.
-If it proofs infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
-Of course researcher must not analyze the data before they write the preregistration but they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
-
-When the researchers are satisfied with the function `planned_analysis()` they can think about how they would like to report the results of the analysis, via tables, plots and text.
-The implementation of this reporting should be in the function `report_analysis()`.
-
-```{r}
-report_results <- function(results){
-  # 1. create markdown tables from results
-  # 2. conditionally interpret results e.g. if(p < .025)"Result is significant."
-  # (optional) visualize results
-  # 3. return results section formatted in markdown
-}
-```
-
-These function again should be accept the output of `planned_analysis()` as input.
-The output of this function should be formatted in Markdown.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
-Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
-
-Ideally, these three functions can be composed to create a "fake" results section, e.g.
-when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section.
-
-To summarize researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
-After they gathered the actual data, they replace the simulated data, run `planned_analysis()` and write the discussion.
-
-# Preregistration as Code --- a Tutorial
-
-To guide the reader through a preregistration as code (PAC), we will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
+To guide the reader through a preregistration as code (PAC), we will use an exemplary research question:
 

### 2021-07-30 manuscript.Rmd
add concluding sentance
add a draft of a simple summary (#128)
* add a draft of a simple summary

* Comments on abstract

Half the abstract focused on PAC; that does not accurately reflect the content of the paper. I've attempted to address that. Still needs a powerful concluding message!

* Update manuscript.Rmd

* add definition of reproducibility

* add concluding sentance

Co-authored-by: C. J. van Lissa <c.j.vanlissa@uu.nl>

--- 2021-07-29 manuscript.Rmd
@@
 
-Some projects, however, are not as easy to make reproducible.
-As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
-These simplifications are possible because of two restrictions.
-Users who ask themselves either "How can I install software that is not an R package in the Docker image?" or "How can I express complex dependencies between files, e.g.
-raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
-All other users may safely skip this section and may return to it if they become limited by restriction.
+This section is only for advanced users who want to overcome some limitations of `repro`.
+If you read this paper the first time you probably want to skip this section.
+As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.
+The simplification of Docker and Make are possible because of two restrictions.
+Users who ask themselves either "How can I install software dependencies outside of R in the Docker image?" or "How can I express complex dependencies between files, e.g.
+hundreds of data files are preprocessed and combined?") need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
+Other users may safely skip this section, or return to it if they encounter such challenges.
 
@@
 
-# Discussion
-
-## Summary
-
-We demonstrated how the R package `repro` supports researchers in creating reproducible repositories and reproducible manuscripts.
-These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
-The workflow we present here rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to all but guarantee reproducibility.
-
-We then proceeded to show preregistration as code (PAC) as one example of how such rigorous and automated reproducibility workflow may enable other innovations.
-In a PAC researcher write all analysis code and a reproducible manuscript that already renders the entire reporting of statistical and computational results before they gather data.
-These results are based on simulated data, such that there is no danger of peeking at the real data while writing the scripts.
-This way, all researcher's degrees of freedom are eliminated.
-Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
-Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
-To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything is under version control.
-The likelihood of those deviations, while unavoidable, is decreased by using simulated data to verify the suitability of the analysis plan.
-This analysis plan may contain data depended decisions to further decrease the likelihood that researcher are unable stick to their preregistration
-
-## Limitations of repro
-
-`repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
-While we have invested much effort to ensure this goal, the package is still in early development and undoubtedly has rough edges and even bugs that went undetected.
-We hope that despite the early development stage, `repro` can support researchers to work more reproducible.
-
-The workflow we present here introduces several software solutions which in itself presents a danger to reproducibility.
-To profit from automatic and convenient reproduction researcher must install Git, Make, and Docker.
-However, Git and Make are itself included in the Docker image created by `repro`.
-Researchers can therefore employ the Docker image manually to download the Git repository and execute Make for full reproduction.
-In other words the only requirement for reproduction and therefore Achilles' heel for this workflow is Docker.
-The Docker approach has two vulnerabilities.
-First, and more important, the Docker image for the project has to remain available.
-The Dockerfile on the other hand is insufficient because it relies on to many service providers (e.g. Microsoft R Application Network, Ubuntu Package Archive).
-The Docker image can be saved to disk and treated like any other file that requires careful archivation.
-If the existence of the Docker image is guaranteed, researchers still require software to run the image.
-To that end they can either rely on Docker itself or Docker-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
-The only way to remove the reliance on such external software is to turn the Docker image into a full operating system that subsequently can be installed and run on almost any modern computer.
-This process is technically possible and would guarantee reproducibility for decades.
-However, this process requires much technical knowledge and is currently not supported by `repro`.
-In any case and without such precautions do the R Markdown, the Makefile and the Dockerfile provide information that allow researchers to manually trace the computational steps and recreate the computational environment.
+<!--AB: TODO add a quick summary -->
 
@@
 
-We realize that PACs are challenging.
-Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove unwieldy for a PAC.
-Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not preregistered research but as a compliment.
-Some may see the rigidness of PAC as to much of a good thing.
-However, we are convinced that many reasons justify changing the preregistered code, but we are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
-
-Another challenge is that PAC requires more programming expertise than a usual analysis.
+PACs come with a challenge though.
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
 We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
-This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
+This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
 The reproducible workflow we present here is heavily automated and hence promotes reusability.
-Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
+Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
+Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
 
@@
 
-Please note that the analytical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes, and it is not necessarily optimal for addressing your research question.
+<!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->
+Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
-The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
+The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
+<!--AB can we move this to a different place?
+Maybe this would actually better fit to the discussion/limitation section!-->
 
@@
 
-Open science practices are a continually evolving field where technical innovations foster changes in research practice.
-Open data is much more widespread thanks to online storage facilities; preregistration are possible because there are preregistration platforms and so forth.
-Similarly, we hope that fully automatic reproduction, e.g. with `repro` as technical innovation, increase scientific rigour and efficiency.
-
-This ideal of a fully automatic reproduction of research projects meets in practice a wide range of demands on how user-friendly and how powerful the software solutions are.
-Some may find Make too complicated or that Docker requires too much storage.
-Yet others may find that they require other programming languages or scale their computation across hundreds of computers, e.g. via high-performance computing clusters or cloud computing.
-
-`repro` was designed modularly to meet such demands.
-At the moment, `repro` only supports the combination of RMarkdown, Git, Make, and Docker.
-However, there are alternatives for each of these elements that may fit better into an individual research project.
-RMarkdown could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer`[@officer] or `officedown`[@officedown] to accommodate a wider arrange of journal submission standards.
-Instead of using formal version control with Git, `repro` could automatically save snapshots for increasing user-friendliness.
-Make could be replaced by the more R-centred alternative `targets` for more convenience.
-Docker could be combined with `renv` [@R-renv] to control the package versions precisely (our approach fixes the date, `renv` the exact package version).
-Alternatively, Docker could be replaced by the more lightweight `renv` if no dependencies outside of R are considered crucial.
-Docker does not satisfy the requirements of many HPC environments, but Singularity was designed to avoid this limitation while still being compatible with Docker images.
-
-`repro`'s modular structure allows such alternative workflow, though they have not been implemented.
-Depending on the demand by users, we will implement some of these alternative workflows in `repro`.
+<!--AB: two lines of what the future of the package, of such standards, and prereigstration is.
+Maybe say something that preregistration is not always possible or wanted but that reproducibility is almost always a good idea?-->
 

### 2021-07-30 manuscript.Rmd
artifact -> digital research product closes #180

@@
 Scientists increasingly strive to make research data, materials, and analysis code openly available.
-Sharing these types of digital research artifacts has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
-To some extent these potential benefits are contingent on whether these digital research artifacts are reproducible. 
+Sharing these types of digital research products has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
+To some extent these potential benefits are contingent on whether these digital research products are reproducible. 
 Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
@@
 This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
-We now have the technology to make all artefacts related to a research project available so that a research project can be reproduced with minimal effort.
+We now have the technology to make all digital products related to a research project available so that a research project can be reproduced with minimal effort.
 
@@
 
-Scientific organizations and funding bodies increasingly demand transparent sharing of research artifacts, and researchers are increasingly willing to do so.
-However, although the sharing of such artifacts is a necessary condition for reproducibility, it is not a sufficient one.
+Scientific organizations and funding bodies increasingly demand transparent sharing of digital research products, and researchers are increasingly willing to do so.
+However, although the sharing of such digital research products is a necessary condition for reproducibility, it is not a sufficient one.
 This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
-Out of 35 published articles with open research artifacts, results of 22 articles could be reproduced but in 11 of these cases, further assistance from the original authors was required.
+Out of 35 published articles with open code and data, results of 22 articles could be reproduced but in 11 of these cases, further assistance from the original authors was required.
 For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
@@
 These failed attempts to reproduce highlight the need for widely accepted reproducibility standards, because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
-If research artifacts are available but not reproducible, their added value is limited.
+If digital research products are available but not reproducible, their added value is limited.
 
@@
 For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
+Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
+With PAC researchers can precisely calibrate how much they want to explore.
+When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
 Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.

### 2021-07-30 abstract.Rmd
add PAC to abstract closes #131

@@
 It is a building block for transparent and cumulative science because it enables the originator and other researchers, on other computers and later in time, to reproduce and thus understand how results came about while, at the same time, it avoids a variety of errors that may lead to erroneous reporting of statistical and computational results.
-In this tutorial, we demonstrate how the R package `repro` supports researchers in creating fully computationally reproducible research projects with a focus on current Open Science practices; starting from sketching the research proposal, over preregistering an analysis as executable code with simulated data, until the creation and dissemination of the final scientific report.
+In this tutorial, we demonstrate how the R package `repro` supports researchers in creating fully computationally reproducible research projects with tools from the software engineering community.
+Building upon this notion of fully automated reproducibility we present several application including the preregistration of research plans with code (Preregistration as Code, PAC).
+PAC eschews all ambiguity of traditional preregistration and offers several other advantages.

### 2021-07-30 manuscript.Rmd

@@
 
-<!--AB: TODO add a quick summary -->
+## Summary
+
+We demonstrated how the R package `repro` supports researchers in creating reproducible repositories and reproducible manuscripts.
+These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
+The workflow we present here rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to all but guarantee reproducibility.
+
+We then proceeded to show preregistration as code (PAC) as one example of how such rigorous and automated reproducibility workflow may enable other innovations.
+In a PAC researcher write all analysis code and a reproducible manuscript that already renders the entire reporting of statistical and computational results before they gather data.
+These results are based on simulated data, so that there is no danger of peeking at the real data while writing the scripts.
+This way, all researcher's degrees of freedom are eliminated.
+Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
+Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
+To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything is under version control.
+The likelihood of those deviations, while unavoidable, is decreased by using simulated data to verify the suitability of the analysis plan.
+This analysis plan may contain data depended decisions to further decrease the likelihood that researcher are unable stick to their preregistration
+
+## Limitations of repro
+
+`repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
+While we have invested much effort to ensure this goal, the package is still in early development and undoubtedly has rough edges and even bugs that went undetected.
+We hope that despite the early development stage, `repro` can support researchers to work more reproducible.
+
+The workflow we present here introduces several software solutions which in itself presents a danger to reproducibility.
+To profit from automatic and convenient reproduction researcher must install Git, Make, and Docker.
+However, Git and Make are itself included in the Docker image created by `repro`.
+Researchers can therefore employ the Docker image manually to download the Git repository and execute Make for full reproduction.
+In other words the only hard requirement for reproduction and therefore Achilles' heel for this workflow is Docker.
+The Docker approach has two vulnerabilities.
+First, and more important, the Docker image for the project has to remain available.
+The Dockerfile (the plain text description to build a Docker image) on the other hand is insufficient because it relies on to many service providers (e.g.
+Microsoft R Application Network, Ubuntu Package Archive).
+The Docker image can be saved to disk and treated like any other file that requires careful archivation.
+However, if the existence of the Docker image is guaranteed, researchers still require software to run the image.
+To that end they can either rely on Docker itself or Docker-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
+The only way to remove the reliance on such external software is to turn the Docker image into a full operating system that subsequently can be installed and run on almost any modern computer.
+This process is technically possible and would guarantee reproducibility for decades without any software dependency.
+However, this process requires much technical knowledge and is currently not supported by `repro`.
+In any case and without such precautions do the R Markdown, the Makefile and the Dockerfile provide information that allow researchers to manually trace the computational steps and recreate the computational environment.
+The Makefile, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are, for whatever reason, unable to run Make.
+Similarly, even when Docker does not work, the Dockerfile still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
 
@@
 
-PACs come with a challenge though.
+We realize that PACs are challenging.
+Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove unwieldy for a PAC.
+Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not preregistered research but as a complement.
+The rigidness of PAC is only than to much of a good thing, if one holds the view that researchers can for no reason deviate from their preregistration.
+However, we are convinced that many reasons justify changing the preregistered code, but are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
+PAC enables just that.
+Researcher can change their plans and possibly defend their decision, e.g. in the discussion, while readers of the manuscript can clearly see what was changed through version control.
+
+Another challenge is that PAC requires more programming expertise than a usual analysis.
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
 We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
-This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
+This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
 The reproducible workflow we present here is heavily automated and hence promotes reusability.
-Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
+Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
+Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
 
@@
 
-<!--AB: two lines of what the future of the package, of such standards, and prereigstration is.
-Maybe say something that preregistration is not always possible or wanted but that reproducibility is almost always a good idea?-->
+Open science practices are a continually evolving field where technical innovations foster changes in research practice.
+Open data is much more widespread thanks to online storage facilities; preregistration are possible because there are preregistration platforms and so forth.
+Similarly, we hope that fully automatic reproduction, e.g. with `repro` as technical innovation, increase scientific rigour and efficiency.
+
+This ideal of a fully automatic reproduction of research projects meets in practice a wide range of demands on how user-friendly and how powerful the software solutions are.
+Some may find Make too complicated or that Docker requires too much storage.
+Yet others may find that they require other programming languages or scale their computation across hundreds of computers, e.g. via high-performance computing clusters or cloud computing.
+
+`repro` was designed modularly to meet such demands.
+At the moment, `repro` only supports the combination of RMarkdown, Git, Make, and Docker.
+However, there are alternatives for each of these elements that may fit better into an individual research project.
+RMarkdown could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer`[@officer] or `officedown`[@officedown] to accommodate a wider arrange of journal submission standards.
+Instead of using formal version control with Git, `repro` could automatically save snapshots for increasing user-friendliness.
+Make could be replaced by the more R-centered alternative `targets` for more convenience.
+Docker could be combined with `renv` [@R-renv] to control the package versions precisely (our approach fixes the date, `renv` the exact package version).
+Alternatively, Docker could be replaced by the more lightweight `renv` if no dependencies outside of R are considered crucial.
+Docker does not satisfy the requirements of many HPC environments, but Singularity was designed to avoid this limitation while still being compatible with Docker images.
+
+`repro`'s modular structure allows such alternative workflow, though they have not been implemented.
+Depending on the demand by users, we will implement some of these alternative workflows in `repro`.
 

### 2021-07-30 manuscript.Rmd
don't use bookdown because automate() does not inference the correct output for base_format
repro now acknowledges more file dependencies
fix dependency structure for temp.bib

@@
     - data/simulation_results.csv
+  images:
+    - images/nutshell.svg
+  files:
+    - references.bib
 output:
-  bookdown::markdown_document2:
-    base_format: rticles::mdpi_article
+  rticles::mdpi_article
 header-includes:
-   - \usepackage[labelformat=empty]{caption}
    - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, numbers=left}
@@
 # References
-
-```{r, include=FALSE}
-knitr::write_bib(
-  c(
-    .packages(),
-    "repro",
-    "here",
-    "rticles",
-    "gert",
-    "bookdown",
-    "lavaan",
-    "knitr",
-    "targets",
-    "renv",
-    "tidyverse"
-  ),
-  here("packages.bib")
-)
-cat(c(readLines(here::here("packages.bib")), "\n", readLines(here::here("references.bib"))), sep = "\n", file = "temp.bib")
-```

### 2021-07-30 manuscript.Rmd
the pdf version is the actual dependency
rename second section

@@
   images:
-    - images/nutshell.svg
+    - images/nutshell.pdf
   files:
@@
 As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
-The first part of this tutorial outlines a reproducibility standard, and introduces the underlying theoretical concepts and software solutions.
+The first part of this tutorial outlines threats to reproducibility and how the here presented reproducibility standard addresses these threats.
 The second part is a tutorial that demonstrates how this standard can be easily adopted with the help of the R package `repro`.
@@
 
-# Concepts and Software Solutions<!--CJ Some inconsistencies:
-1) From the end of the previous paragraph, I would expect a description of the "rigorous reproducibility standard" here.
-2) The paragraph is labeled "Concepts and Software SOlutions, but it begins with a list of threats to reproducibility. Threats are not concepts nor software solutions.
-3) After the list of threats, we get a description of the "rigorous reproducibility standard".
-That is what I would expect based on the preceding paragraph.
-Note that the "rigorous reproducibility standard" is also not a concept, nor a software solution.
-Consider re-naming this section.-->
-
-From our own experience with various research projects, we identified the following common threats to reproducibility: <!-- CJ This still requires attention!
-How did you identify these causes?
-Is this exhaustive?-->
+# Threats to reproducibility and appropriate remedies
+
+From our own experience with various research projects, we identified the following common threats to reproducibility:
 

### 2021-07-30 manuscript.Rmd
reference install.md #140

@@
 The first step is to install the required software.
-Start R Studio<!--CJ I am concerned that new users may not install R if you only mention RStudio.--> and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
+We assume that you have installed R and RStudio already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert)[@R-repro].
+In case you either have not installed R and RStudio or are unsure if they are up to date you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers how to install all software necessary for this tutorial in three steps.
+Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
 It will assist you while you follow the tutorial.

### 2021-07-30 manuscript.Rmd

@@
 Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
-Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
-22][^popper]
+The credibility of empirical research hinges on its objectivity.
+Objectivity in this context means "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
+22][^popper].
+Therefore, reproducibility has long been considered integral to empirical research.
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
@@
 However, although the sharing of such digital research products is a necessary condition for reproducibility, it is not a sufficient one.
+
 This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
@@
 In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
-Typically, results are manually copyied and pasted from one software to the other, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
+Typically, results are manually copied and pasted from one software to another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
 
@@
 It is highly fool-proof [unlike, for example, LaTeX @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles], and presentations[@revealjs] or even your résumé [@vitae].
+We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
 R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
 Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
-This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
+This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
 @xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.
@@
 However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
-In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambigously documented.
+In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
 
@@
 Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
-In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, analysis results).
+In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
 
 We recommend using **Make** for dependency tracking because it is language independent.
-The following hypothesical example illustrates the utility of Make.
+The following hypothetical example illustrates the utility of Make.
 Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
@@
 It provides all software used in an analysis, including the operating system itself.
-The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
+The operating system level is important because some functionality may be passed to software layers beneath the programming language, for example random number generators and linear algebra libraries.
 Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
@@
 It will assist you while you follow the tutorial.
+>>>>>>> main
 
@@
 An example of the assistance that `repro` provides are the check functions.
+
 These verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
@@
 ```
+
 Now the file `private.md` will not be committed to Git and will not be made public.
@@
 For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
+
 One way to do this is through the visual user interface in the RStudio Git pane.
-Click on the empty box next to the file you want to stage. A check mark appears to indicate that the file is staged.
+Click on the empty box next to the file you want to stage.
+A check mark appears which indicates that the file is staged.
 After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
@@
 
-Depending on your computer configuration, this may ask to set up a secure connection to GitHub. In this case first follow the suggestions shown on the R console.
+Depending on your computer configuration, this may ask to set up a secure connection to GitHub.
+In this case first follow the suggestions shown on the R console.
 
@@
 
-1. Text in natural language (prose; e.g., a scientific paper or presentation)
+1. Text (prose; e.g., a scientific paper or presentation)
 2. Executable code (e.g., analyses)
@@
 R Markdown is a type of dynamic document well-suited to the RStudio user interface.
-The natural language text in an R Markdown file can be formatted using Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
+The text of an R Markdown is formatted by Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
 The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
 The following example serves to illustrate the Markdown syntax.
+
 It shows how to create a heading, a word in bold font, and a list of several items in Markdown:
@@
 Another advantage is that, because R Markdown files are in plain text, they are more suitable for version control using Git.
-Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass <!--CJ Add picture-->in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
+Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
 The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
@@
 ````
+
 The hotkey [Control]+[Alt]+[i] inserts a block of code in the file.
@@
 These metadate contain information about the document, like the title and the output format.
-Metadata are placed at the beginning of the document and are separated by three dashes from the document body; the following example is a full markdown document where the header is in lines 1--6.
+Metadata are placed at the beginning of the document and are separated by three dashes from the document body.
+The following example is a full markdown document where the metadata (the "YAML front matter") is in lines 1--6.
 
@@
 
-The metadata the `repro::automate()` function relies on can only express dependencies R packages for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
+The metadata the `repro::automate()` function relies on can only express R packages as dependencies for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile.
 The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
-Other software that users might need like other programming languages, not already installed LaTeX packages etc. have to be added manually.
-We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g.
-for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
+Other software that users might need like other programming languages and not preinstalled LaTeX packages have to be added manually.
+We plan to add support for commonly used ways to install software beyond R packages, e.g.
+for system libraries (via the Ubuntu package manager `apt`), LaTeX packages (via the Tex Live package manager `tlmgr`), and python packages (via the python package manager `pip`).
 The second limitation is about dependencies.
@@
 These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
-Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D therefore leaving out important information or express the dependencies directly within the Makefile.
-
-The following section explains how to overcome these limitation while still relying on the automatons afforded by `repro`.
-Lifting these restrictions requires the user to interact more directly with Make or Docker.
-To satisfy more complicated requirements users need to understand how repro utilizes Make and Docker internally.
-
-Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two processing steps.
+Users therefore have to either "flatten" the dependency structure by simply stating that A depends on B, C and D, thereby leaving out important information, or express the dependencies directly within the Makefile.
+
+The following section explains how to overcome these limitations while still relying on the automations provided by `repro`.
+Lifting these limitations requires the user to interact more directly with Make or Docker.
+To satisfy more complicated requirements useres may have, they need to understand how `repro` utilizes Make and Docker internally.
+
+Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`.
+It consists of two processing steps:
 First it recreates the virtual software environment (Docker) and then it executes computational recipes in the virtual software environment (Make).
-The first step is done by the command `make docker`.
-The command `make docker` will trigger Make to build the target called `docker`.
+The first step is done by the command `make docker`, it will trigger Make to build the target called `docker`.
 The recipe for this target builds an image from the Dockerfile of the repository.
@@
 
-Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
-Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
-Preregistrations, however, have a problem.
-Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
-The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
-The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
-But these templates can not remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
+Defining the research questions and planning data analysis before observing the research outcomes is referred to _preregistration_ [@NosekRevolution2018].
+Preregistration increases the credibility of empirical results by separating a-priori planned analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
+For a priori planned analyses, preregistration is supposed to reduce so-called "researcher degrees of freedom", and therefore, the number of decisions researcher can make that could unduly influence the results.
+Preregistrations, however, often do not achieve the level of unambiguity that are required to effectively zero the researcher degrees of freedom.
+Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
+The current best practice to ensure that the preregistration is comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
+The structure these templates provide can more effectively reduce researcher degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
+Nevertheless, these templates can not remove all researcher degrees of freedom because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
 
 Our proposed solution is simple:
-Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
-We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
-PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration despite the best efforts afford.
+Instead of describing the analysis, you write and publish the code to conduct the analysis before you gather data.
+We term this *preregistration as code (PAC)*.
+A reproducible PAC has the potential to eliminate all researcher degrees of freedom, even those that standard preregistration can not eliminate.
 
@@
 A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
-In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
-
-We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.
+In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`), we found close to a dozen published articles that included some form of script as part of their preregistration.
+
+We believe, however, that combining ideas from reproducibility and preregistration can turn this idea into a tool with tangible benefits for the research community.
 
@@
 First, preregistrations are plagued by non-reproducibility like any other research outcome.
-@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to reproduce the power analysis.
-Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
+@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis, but from these, only 34 (20%) provided enough information to reproduce the power analysis.
+Even researchers who have gone to great lengths in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
 For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
@@
 The idea is as follows:
-For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
-In this document, the results are initially based on simulated data that mimics the expected output.
-
-Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
-It follows that for preregistrations they should stick to widely accepted standards of scientific writing [e.g. @apa7].
-We have already explained the need for reproducibility standards in the process of writing such manuscript and it is therefore only natural to extend these arguments to preregistration.
+For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather or acquire empirical data.
+In this document, the results are initially based on simulated data that merely serves the purpose to mimic the expected output. 
+For example, this simulated data allows to write code that already produces figures or tables.
+
+Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific report, ie., researchers do not have to learn specifics of a given preregistration template.
+It follows that for preregistrations they should simply stick to widely accepted standards of scientific writing [e.g. @apa7].
+We have already explained the need for reproducibility standards in the process of writing such manuscripts and it is therefore only natural to extend these arguments to preregistration.
 
@@
 
-We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, a PAC removes any ambiguity regarding the translation of an verbal analysis plan into code.
+We believe that pairing PAC with the workflow presented here offers five advantages over classical preregistration.
+First, a PAC removes any ambiguity regarding the translation of a verbal analysis plan into code.
 Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
@@
 Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
-We believe it is easier with PAC to detect missing steps or ambiguous decisions.
+Using simulated data makes it straightforward to detect missing steps or ambiguous decisions.
 Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
 Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
-Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.
+Fifth, the preregistration is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
 
@@
 
-Leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
-To use PAC for study planing, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
+When leaving meta-scientific benefits aside, a preregistration is also a tool for planning a study.
+To use PAC for study planning, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
 To that end, they should borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
@@
 Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
-One particular advantage is the strong foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
-Nevertheless it can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section.
+One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
+Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
 
 Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-Simulating data forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
+While relatively uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
+When simulating data, researchers have to think about and address data-analytic problems on a conceptual level before making an effort to acquire data [@axelrodAdvancingArtSimulation1997].
 It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
-Writing incomplete code is technically impossible or at least researchers would notice if they e.g. forgot to aggregate the data etc.
+Writing incomplete code is technically impossible, or at least researchers would notice if something is omitted (e.g. if they forgot to aggregate the data).
 
@@
 
-To write a PAC researchers should start by drafting a manuscript in form of an R Markdown.
+To write a PAC, researchers should start by drafting a manuscript in the form of an R Markdown.
 When they have completed the method section, they should begin writing code.
-We suggest writing three functions for each hypothesis, one for the planned analysis, one for the simulating the data and one for reporting the results.
-At the heart of a PAC lies the planned analysis so researchers should first draft a function called `planned_analysis()` that captures the description in the method section.
+We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
+At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
 This function should receive the data and derive relevant results from it.
 
-```{r}
+```{r, eval=FALSE}
 planned_analyis <- function(data){
@@
 
-Because it is difficult to draft such function without interactively test it, researcher should than proceed to write a function for simulating data.
-
-```{r}
+Because it is difficult to draft such a function without interactively testing it, the researcher should then write a function for simulating data.
+The follow code illustrates the signature of a simple simulation function:
+
+```{r, eval=FALSE}
 simulate_data <- function(n, effect_size){
-  # 1. warn users that the results are "fake"
+  # 1. warn users that the results are mocked
   # 2. draw `n` samples with `effect_size`
@@
 
-As soon as they have written `planned_anaylsis()` and `simulate_data()` they can iteratively improve both functions, e.g.
-till `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
+As soon as they have written a first draft of `planned_anaylsis()` and `simulate_data()` they can interactively improve both functions, e.g.
+until `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
 The goal is that the output of `simulate_data()` works as input to the function `planned_analysis()`.
 
-This process can be tedious and complicated.
-If it proofs infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
-Of course researcher must not analyze the data before they write the preregistration but they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
-
-When the researchers are satisfied with the function `planned_analysis()` they can think about how they would like to report the results of the analysis, via tables, plots and text.
+If it is infeasible to simulate realistic data, researcher can delay their preregistration until they have gathered a given percentage of the data.
+Of course, researchers must not analyze these data before they write the preregistration.
+However, they can obfuscate these data by shuffling the dependent variable or otherwise createing synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
+
+When the researchers are satisfied with the function `planned_analysis()`, they can think about how they would like to report the analysis results via tables, plots, and text.
 The implementation of this reporting should be in the function `report_analysis()`.
-
-```{r}
-report_results <- function(results){
+<!--AB hier ist mir unklar, wie das genau geht. baue ich hier alle Figures und Tables und Textstücke und benutze die dann erst später? Bitte genauer-->
+
+```{r, eval=FALSE}
+report_analysis <- function(results){
   # 1. create markdown tables from results
   # 2. conditionally interpret results e.g. if(p < .025)"Result is significant."
-  # (optional) visualize results
-  # 3. return results section formatted in markdown
+  # 3. visualize results
+  # 4. return results section formatted in markdown
 }
@@
 
-These function again should be accept the output of `planned_analysis()` as input.
+This function should again accept the output of `planned_analysis()` as input.
 The output of this function should be formatted in Markdown.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
+Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
 Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
-
-Ideally, these three functions can be composed to create a "fake" results section, e.g.
-when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` outputs a results section.
-
-To summarize researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
-After they gathered the actual data, they replace the simulated data, run `planned_analysis()` and write the discussion.
-
-# Preregistration as Code --- a Tutorial
-
-To guide the reader through a preregistration as code (PAC), we will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
-
-> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"
-
-This example serves only didactic purposes and not to derive substantive claims.
-Nevertheless, everything else is kept as close to a real research question as possible.
-
-At the heart of a PAC lies the planned analysis.
-We suggest writing a main analysis function that captures the full process of testing a research question.
-This function receives a dataset as input and performs all statistical processing. <!--AB say anything about its output? is the Markdown the output?-->
-For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
-As an example of a data-dependent decision, we included a rank transformation if the skewness was greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it generates not only tables but also a straightforward interpretation of the effects as actual prose (e.g.
+it verbally quantifies the size of an effect).
+
+Ideally, these three functions can be composed to create a mock results section, e.g.
+when composed to `report_analysis(planned_analysis(simulate_data()))` or `simulate_data() %>% planned_analysis() %>% report_analysis()` they output a full results section.
+
+To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration.
+After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data) and write the discussion.
+
+## An example of Preregistration as Code
+
+We argued that PAC has several advantages over classical preregistration and have outlined how to implement it.
+To illustrate how PAC works in practice and help researchers implementing PAC themselves, we have conducted a PAC that serves as a guiding example.
+We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
+
+> "Is there a mean difference in the personality trait 'Machiavellism' between self-identified females and males?"
+
+Again, we propose a preregistration format that closely resembles a classical journal article, but uses simulated data and dynamic document generation to create a document that starts out as a preregistration and eventually becomes the final report.
+The complete preregistration source is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
+In this section, we show code excerpts of this preregistration (formatted in monospace) and explain the rationale behind them.
+
+```{r, echo=FALSE}
+show <- function(x, wrap = TRUE, cat = TRUE, fence = TRUE){
+  stopifnot(length(x) == 1L)
+  if(wrap)x <- str_c(str_wrap(str_split(x, "\\n\\n")[[1]]), collapse = "\n\n")
+  if(!isFALSE(fence)){
+    if(isTRUE(fence))fence <- "markdown"
+    fence <- str_c("````", fence)
+    x <-
+      str_c(fence,
+            x,
+            "````",
+            sep = "\n",
+            collapse = "\n")
+  }
+  if(cat){
+    cat(x)
+    return(invisible(x))
+  } else {
+    return(x)
+  }
+}
+show_prereg <-
+  function(section,
+           file = "preregistration.Rmd",
+           wrap = TRUE,
+           cat = TRUE,
+           fence = TRUE) {
+    prereg <- readr::read_lines(here::here(file))
+    start <- stringr::str_c("<!--begin-", section, "-->")
+    end <- stringr::str_c("<!--end-", section, "-->")
+    start_pos <- which(prereg == start)
+    end_pos <- which(prereg == end)
+    if (length(start_pos) != 1L || length(end_pos) != 1L)
+      stop("No section '", section, "' found.")
+    excerpt <-
+      str_c(prereg[(start_pos + 1):(end_pos - 1)], collapse = "\n")
+    show(excerpt, wrap = wrap, cat = cat, fence = fence)
+  }
+```
+
+As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary background information and literature to understand the context and purpose of the research question.
+This example is drastically shortened for illustration purposes:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("theory")
+```
+
+After researchers have provided the research question, they typically proceed to explain how they want to study it.
+For simplicity, we will use already published data that we have not yet analyzed:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("method")
+```
+
+We choose the following statistical procedure because many researchers are familiar with it[^caution-mww]:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("ttest")
+```
 
 [^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
-The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
-
-This function is one possible translation of the verbal description above:
+The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
+
+We then translated this methods section into the following `planned_analysis()` function:
 
@@
 
-Importantly, natural language descriptions do not map unambiguously onto computer code implementations. 
-The same textual descriptions may be implemented differently by different persons depending on their statistical and programming knowledge but also based on their own expectations and beliefs.
-One example would be using the function `wilcox.test` instead of the combinations of `rank` and `t.test`.
-Both are valid implementation of the Mann–Whitney–Wilcoxon test but the first assumes equal variance whereas the second applies Welch's correction and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
+This function illustrates two advantages of PAC.
+First, a PAC can easily include data-dependent decisions creating different analysis branches under different conditions.
+Second, it highlights how difficult it is to describe a statistical analysis precisely.
+The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and their expectations and beliefs.
+One example would be using the function `wilcox.test` instead of the combinations of the functions `rank` and `t.test`.
+Either of them are valid implementations of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
+In contrast, the second applies Welch's correction by default and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
 Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
-Still, these details can make a difference and provide unwanted researcher degrees-of-freedom.
-
-<!--AB here is kind of a break in the text-->
-To use this function one needs a compatible dataset. <!--AB what is "this" function? What does "compatible" mean?-->
-At the time of preregistration there is of course no actual data available.
-We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
-Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
-The following code simulates a dataset with one observed variable in two groups.
-
-```{r simulate_data}
-```
-<!--AB it is odd that the return variable is called effect because it is a dataset-->
-
-Here, we decided to randomly draw non-normal data to test how well our procedure performed under non-optimal situations.
-Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a given standardized effect size (a Cohen's d of .2).
-This simulated dataset fits perfectly into our planned analysis with a sample size of 400 and 10 items: <!--AB in what way? please explain-->
-
-***
-
-```{r}
-set.seed(112358) # fix state of RNG
-results <- 
-  planned_analysis(
-    simulate_data(
-      n = 400,# Sample size
-      df = 8, # Chisq with DF = 8 has skewness = 1
-      d = .2, # Cohen's D
-      i = 10  # Number of items (should not effect anything)
-    )
-  )
-```
-
-Next, we add dynamic code blocks to our Markdown report that implement the planned reporting of the final results.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
-Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
-Here we use the `report` package to create a preview of how our reporting will look like based on simulated data:<!--AB add citation to the report package-->
-
-```{r, results='asis', eval=FALSE}
-report::report(results$test)
-```
-
-This call to the `report()` function will yield the following automatically generated output:
-
-```{r, results='asis', echo=FALSE}
-report::report(results$test)
-```
-<!--AB it is difficult to discern the automatically generated text and out manuscript text even though there is the horizontal line. Not sure how to improve this.-->
-***
+Still, these details can make a difference and provide unwanted researchers' degrees of freedom.
+
+The function `planned_analysis()` together with the function `simulate_data()` (not shown here) can be used to justify the planned sample size.
+To that end, `simulate_data()` is repeatedly called with increased sample size and the proportion of significant results (power) is recorded.
+The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
+The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R).
+
+Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
 
@@
 ```
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("power-code", wrap = FALSE)
+```
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("power")
+```
+
+We implemented a mechanism that only uses simulated data when the actual data is not yet available  (in this example if the file `data/data.csv` does not exist)for the results section.
+This mechanism also warns readers if these results are based on simulated data.
+The warning is colored red to avoid any confusion between mock and actual results.
+As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study.
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("results", wrap = FALSE)
+```
+
+The last thing we need to do is to report our results with the combination of the functions `planned_analysis()` and `report_analysis()`.
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("report", wrap = FALSE)
+```
+
+This is an example of how the results could be reported (based on simulated data):
+
+```{r, echo=FALSE, results='asis'}
+set.seed(1234)
+show(report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)), cat = FALSE), fence = "")
+```
+
+The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
 
@@
 Another challenge is that PAC requires more programming expertise than a usual analysis.
+
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
 We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
-This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
+This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
 The reproducible workflow we present here is heavily automated and hence promotes reusability.
+
 Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
+Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
 
@@
 
-Please note that the analytical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes, and it is not necessarily optimal for addressing your research question.
+Please note that the statistical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and is not necessarily optimal for addressing your research question.
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].

### 2021-07-31 manuscript.Rmd
Comments to remainder of the paper

--- 2021-07-30 manuscript.Rmd
@@
 First, the present paper introduces the reader to a workflow that maximizes the potential reproducibility of a research project.
+The end goal of this workflow is a project that can be completely, automatically, and unambiguously reproduced.<!--CJ From tutorial-->
+This means that any manual steps, such as copy and pasting results, manual modification of data (e.g., manually removing observations or manually recoding variables),
+are replaced with reproducible code. 
+The result of this workflow will be a portable research project containing all digital research artefacts,
+in which all steps required to produce the (published) manuscript from the raw (and ideally, open) data are well-documented
+and automatically reproducible with a single command.<!--CJ From tutorial-->
+
 This workflow builds on several technical solutions originating in the software engineering community. 
@@
 
-**Docker** is an example of containerization.
-It provides all software used in an analysis, including the operating system itself.
-The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
-Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
-Researchers therefore have the opportunity to use the exact same software setup that was used by the original authors.
-Such a snapshot of the software stack is called an "image".
+**Docker** is a popular protocol for containerization.
+It manages software dependencies by constructing a virtual software environment independent of the host software environment.
+These so-called "Docker images" function like a virtual computer  (i.e., a sort of "sand box").
+A Docker image should contain *all* software dependencies used in an analysis - not just R-packages, but also R and Rstudio, and even the computer's operating system.
+This is important because some of its functionality may impact the workings of higher-order software, such as calls to random number generators, linear algebra libraries and such within R.<!-- CJ From tutorial-->
+
+Note that the software environment of the Docker image is completely separate from software installed on your computer.<!-- CJ This paragraph is also relevant for users that are not advanced!!!
+Moved it here, from the tutorial-->
+This separation is excellent for reproducibility, but take some getting used to.
+For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the Docker image.
+Each dependency that you want to use within the Docker image must be explicitly added as a dependency.
+Furthermore, using Docker mayrequire you to install software on an operating system that may not be familiar to you.
+The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.
+
 Packaging all required software in such an image requires considerable amounts of storage space.
@@
 To remove this hurdle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
-In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.
+Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
+These projects will remain reproducible irrespective of whether `repro` remains accessible in future.
+Users do not have to have repro installed to reproduce a project;
+in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
+In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.<!-- CJ From tutorial-->
 
@@
 title: A Hitchhiker's Guide to Reproducible Research in R
-author: Aaron Peikert, Caspar J. van Lissa and Andreas M. Brandmaier
-abstract: A tutorial about doing the same thing more than once.
+author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
+abstract: A tutorial about doing the same thing more than once.<!-- CJ I actually really like this title, and would consider it for the main manuscript.
+E.g., "Reprodicible research: A tutorial about.."-->
 output: html_document
@@
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
-abstract: A paper about recalculating the same thing more than once.
+abstract: A paper about recalculating the same thing more than once.<!-- CJ Is it correct that this is different from the example above?-->
 output: html_document
@@
 
-This information makes explicitly declares the project's dependencies on files and R packages.
-`repro` uses this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
+This information makes explicitly declares the project's dependencies on files and R packages.<!-- CJ How about version numbers and sources?
+Would it make sense to declare a dependency on renv and include a .renv file?-->
+`repro` uses this information to construct a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
 These two files together form the basis for consistency within a research project and consistency across different systems.<!--CJ Consider explaining this in the theoretical intro-->
-At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.<!--CJ This comes way to late.
-Shouldbe in the opening paragraph or something.-->
-This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
-Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).<!--CJ These points describe the formal definition of reproducibility used in this paper, and the goal of the workflow.
-That needs to be explained in the introduction.
-If you only find out what you are working toward halfway through the tutorial - that's too late!-->
-
 The function `repro::automate()` converts the metadata from all R Markdown files in the project (all files with the ending `.Rmd`) to a Makefile and a Dockerfile.
@@
 
-Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.<!--CJ This belongs in the intro, not the tutorial.
-It is part of explaining the benefits of repro.-->
-These projects will remain reproducible irrespective of whether `repro` will be maintained and remain accessible in future.
-Users do not have to have repro installed to reproduce a project, in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
-The only required software for reproducing a project is Docker, assuming users know how to build a Docker image and run Make within the container.
+The only software required for reproducing a project is Docker, assuming users know how to build a Docker image and run Make within the container.
 However, if they have installed Make additionally to Docker, they do not need to know how to use Docker and can simply rely on the two Make commands `make docker` and `make -B DOCKER=TRUE`.
 The commands can be pasted into a terminal and will reproduce the project in many cases.
+<!-- CJ This paragraph needs to be elaborated further-->
 
@@
 
-This section is only for advanced users who want to overcome some limitations of `repro`.
+This section is only for advanced users who want to overcome some limitations of `repro`.<!-- CJ Consider sticking this in a vignette, which would allow it to be updated more regularly-->
 If you read this paper the first time you probably want to skip this section.
@@
 `repro` creates three Dockerfiles inside the `.repro` directory.
-The file `.repro/Dockerfile_base` contains information about the base image on which all the remaining software is installed.
+Two Dockerfiles are automatically generated.
+The first is `.repro/Dockerfile_base`.
+It contains information about the base image on which all the remaining software is installed.
 By default we rely on the "verse" images provided by the Rocker project [@boettigerIntroductionRockerDocker2017].
@@
 
-The file `.repro/Dockerfile_manual` is empty by default and can be used to add further dependencies outside of R, like system libraries or external software. 
-Remember, the software environment of the Docker image is completely separate from software installed on your computer.
-This separation is excellent for reproducibility, but requires you to install software on an operating system that may not be familiar to you.
-The images supplied by [@boettigerIntroductionRockerDocker2017] are based on Ubuntu, and hence software has to be installed in a way that is compatible with Ubuntu.
+The second automatically generated Dockerfile is `.repro/Dockerfile_packages`.
+Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.<!-- CJ Suggestion: Make a directory called "dockerfiles", and name the files "image", "external_dependencies", and "rpackages"-->
+
+Finally, there is one manually edited Dockerfile: `.repro/Dockerfile_manual`.
+<!-- CJ I would expect this file to contain a "manual" based on the title.
+Consider if there is a better name, eg: .repro/external_dependencies-->
+It is blank by default, and can be used to add further dependencies outside of R, like system libraries or external software.
+
+<!--Note that the software environment of the Docker image is completely separate from software installed on your computer. CJ This paragraph is also relevant for users that are not advanced!!!
+This separation is excellent for reproducibility, but take some getting used to.
+For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the Docker image.
+Each dependency that you want to use within the Docker image must be explicitly added as a dependency.-->
+Using Docker mayrequire you to install software on an operating system that may not be familiar to you.
+The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.
 The most convenient way to install software on Ubuntu is through its package manager `apt`.
-The following snipped installs Python.
+The following snipped installs Python.<!--CJ You have not explained why one might want to install python.
+I think this is too specific for the general paper; put in a vignette and make a separate vignette for "Adding external dependencies",addressing python and others-->
 If added to `.repro/Dockerfile_manual` the Docker image will have Python installed.
@@
 
-The last Dockerfile, `.repro/Dockerfile_packages`, is automatically generated from the metadata provided and must not be manually changed.
-Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.
-
-The separation into several Dockerfiles is a design choice to provide a clear separation between automatically generated parts based on metadata provided and manually altered parts of the Dockerfile.
-Docker eventually requires a single Dockerfile to run, so `repro::automate()` simply concatenates the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the R project.
+<!--
+Docker eventually requires a single Dockerfile to run, so `repro::automate()` simply concatenates the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the R project.CJ Too much technical detail about the implementation.
+This section promised to help advanced users enhance the functionality of repro - but what functionality are they gaining by knowing that repro concatenates these files?
+-->
 
@@
 
-The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but usually should not modify.
+The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but usually should not modify.<!-- CJ Under what conditions could you modify it? -->
 This Makefile coordinates the interplay between Make and Docker and contains targets for building (with `make docker`) and saving (with `make save-docker`) the Docker image.
 Additionally, it provides facilities to execute commands within the container.
-If you write computational recipe for a target it will be by default evaluated using the locally installed software.
+If you write a computational recipe<--CJ this term has not been defined--> for a target it will be by default evaluated using the locally installed software.
 To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first Make example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
@@
 Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automations.
+<!-- CJ I would only leave one paragraph for advanced users, stating that they can manually change the Dockerfiles and makefiles following the vignette on the repro website.
+Skip the details about implementation in this paper!!!
+That's not what a tutorial is for-->
 
@@
 Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
-Preregistration increases the credibility of empirical results by separating a priori planned analyses from those conducted unplanned and post hoc.
-For a priori planned analysis, preregistration should reduce "researcher degrees of freedom" therefore the number of decisions researcher can make that could unduly influence the results.
-Preregistrations, however, have a problem.
-Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis the results vary considerably [@silberzahnManyAnalystsOne2018].
+Preregistration increases the credibility of empirical results by separating theory-driven planned analyses from unplanned exploratory analyses.
+Preregistration seeks to reduce "researcher degrees of freedom" [@reference] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could influence the results.
+However, there are several shortcomings to preregistration.
+One shortcoming is that written study plans are often multi-interpretable.
+Even when several researchers set out to investigate the same hypothesis using the same data, their approaches vary considerably [@silberzahnManyAnalystsOne2018].
 The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
-The structure these templates provides can more effectively reduce researchers degrees of freedom than an unstructured approach [@bakkerEnsuringQualitySpecificity2020].
-But these templates can not remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
-
-Our proposed solution is simple:
-Instead of describing the analysis, you write the code to conduct the analysis before you gather data.
-We term this approach of writing and publishing code before data collection *preregistration as code (PAC)*.
-PAC has the potential to all but eliminate researchers degrees of freedom that standard preregistration despite the best efforts afford.
+These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
+However, templates cannot remove researchers degrees of freedoms completely,
+because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
+A second limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
+To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
+Even when researchers have gone to great length in preregistering an analysis script,
+they sometimes inexplicably fail to reproduce their own results.
+For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
+A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
+For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables, or more parameters than observed data.
+Conversely, a model may be mis-specified in a major way; for example by omitting direct effects when the research question is about mediation, thus leading to a model with unacceptable fit.
+Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.
+
+The workflow described in this paper facilitates a rigorous and innovative solution to this problem:
+Instead of describing the analysis in prose, researchers include the code required to conduct the analysis in the preregistration.
+We term this approach of writing and publishing code at the preregistration stage *preregistration as code (PAC)*.
+PAC has the potential to eliminate researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
+Moreover, it removes the necessity to write two separate documents: A preregistration, and a manuscript.
+For a PAC, researchers can write a reproducible, dynamically generated draft of their intended manuscript at the preregistration stage.
+This includes the Introduction, Methods, and Results sections.
+These results are initially based on simulated data with the same structure as the data they expect to obtain from their experiments.
+For guidance on how to simulate data, see [@ref].
+Once the preregistration is submitted and real data have been collected or made available,
+the document can be reproduced with a single command, thus updating the Results to their final version.
 
@@
 In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
-
-We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.
-
-## Why are PAC and reproducibility related?
-
-First, preregistrations are plagued by non-reproducibility like any other research outcome.
-@bakkerRecommendationsPreregistrationsInternal2020 found that, among the 210 preregistrations they investigated, 174 (67%) included a formal power analysis but from these only 34 (20%) provided enough information to reproduce the power analysis.
-Even researcher who have gone to great length in submitting a script in their preregistration sometimes inexplicably fail to reproduce their own results.
-For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
-
-Second, although it has been almost a decade since @steegenMeasuringCrowdAgain2014 have preregistered their code, the idea of publishing code before analyzing data has gained not much traction.
-This could change when combining the here proposed workflow and PAC.
-We want to extend PAC from preregistering code to preregistering *the whole manuscript as a dynamic document*.
-The idea is as follows:
-For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
-In this document, the results are initially based on simulated data that mimics the expected output.
-
-Writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
-It follows that for preregistrations they should stick to widely accepted standards of scientific writing [e.g. @apa7].
-We have already explained the need for reproducibility standards in the process of writing such manuscript and it is therefore only natural to extend these arguments to preregistration.
+<!-- CJ This is not convincing:
+We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.-->
+But although the notion of preregistering code has been around for a while [cf. @steegenMeasuringCrowdAgain2014],
+it has not gained much traction - perhaps because, until now, this has constituted an extra non-standard step in the research process.
+This tutorial integrates the preregistration of code into the research workflow
+by encouraging researchers to preregister the whole manuscript as a reproducible dynamic document.
 
@@
 We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, a PAC removes any ambiguity regarding the translation of an verbal analysis plan into code.
-Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
-Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
+First, in PAC, the preregistration is simply a first draft of the final manuscript,
+thus sparing authors from writing (and editors and reviewers from evaluating) two separate documents.
+Relatedly, writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
+By contrast, filling out a preregistration template is a novel type of scientific writing for many.
+Second, the PAC is a tool for study planning.
+When all steps to be executed are documented clearly, the execution of the study will proceed more efficiently than when every step is planned in medias res.
+Third, a PAC removes any ambiguity regarding the translation of verbal analysis plans into code.
+Fourth, despite its rigor, PAC does accommodate data-dependent decisions if these can be formulated as code.
+Researchers can, for example, formulate conditions (`if() {} else {}`) under which they prefer one analysis type over the other.
 For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
-Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
-With PAC researchers can precisely calibrate how much they want to explore.
-When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
-Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
-We believe it is easier with PAC to detect missing steps or ambiguous decisions.
-Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
-Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
-Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.
-
-## Benefits of PAC for study planning
-
-Leaving meta-scientific benefits aside, a preregistration is a tool for planning a study.
-To use PAC for study planing, we suggest that researchers write a dynamically generated manuscript including the sections Introduction, Methods, and Results (or similar) _before_ they gather empirical data.
-To that end, they should borrow from widely employed standards for writing an empirical manuscript [e.g. @apa7].
-Then they should translate their method section into executable computer code and simulate data that has the same structure as the data they expect to obtain from their experiments.
-
-We consider three domains where this procedure can help planning a study.
+Fifth, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
+Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
+PAC therefore minimizes researchers degrees of freedom more effectively than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
+Sixth, deviations from the preregistration are clearly documented, because they are reflected in changes to the code which are managed and tracked with version control.
+One could even make a commit labeled "Necessary deviations from the preregistration after data collection",
+and reference it in the manuscript.
+
+Note that preregistration, or PAC, does not preclude exploratory analyses.
+Such analyses should merely be explicitly declared as such in the manuscript.
+
+## What you need to know before writing a PAC
+
+When writing a PAC, researchers ought to strive to meet three goals:
 First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
@@
 
-Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
+Describing the preregistration in the form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
 One particular advantage is the strong foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
-Nevertheless it can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section.
+It can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section [@ref provide specific good examples].
 
 Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
-While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
-Simulating data forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
+Simulating data offers various benefits in the planning phase of a study.
+It forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
 It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
@@
 
-With simulated data researchers can check if the analysis is technically complete and can also recover the true parameters of the simulation.
-When an analysis fails to recover the correct values in idealized conditions with big samples researchers should not expect this analysis to work under realistic conditions.
-Because the simulation captures the researchers expectations about the real world conditions it is the best way to assess the effectiveness of the analysis without actually gathering the data.
+When the comprehensiveness criterion is met, researchers can check if the analysis is effective at recovering the true parameters of the simulation.
+If an analysis fails to recover the correct values in idealized (simulated) conditions,
+it would be unreasonable to expect this analysis to work under realistic conditions.
+Because the simulation describes the researchers expectations about real world conditions,
+it is the best way to assess the effectiveness of the analysis without actually gathering the data.
+If a researcher experiences difficulty in specifying such expectations,
+that may be an indication that further literature research, or exploratory pilot studies, are required.
 
@@
 
-## What you need to know before writing a PAC
+### The writing process
 
@@
 As soon as they have written `planned_anaylsis()` and `simulate_data()` they can iteratively improve both functions, e.g.
-till `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
+until `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
 The goal is that the output of `simulate_data()` works as input to the function `planned_analysis()`.
 
-This process can be tedious and complicated.
-If it proofs infeasible to simulate data, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
+#### Alternatives to simulated data
+
+If simulating data proves to be too challenging, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
 Of course researcher must not analyze the data before they write the preregistration but they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
@@
 These function again should be accept the output of `planned_analysis()` as input.
+
+<!-- CJ: The text below is about reporting standards; it's useful, but not directly linked to the text above-->
 The output of this function should be formatted in Markdown.
@@
 The same textual descriptions may be implemented differently by different persons depending on their statistical and programming knowledge but also based on their own expectations and beliefs.
-One example would be using the function `wilcox.test` instead of the combinations of `rank` and `t.test`.
-Both are valid implementation of the Mann–Whitney–Wilcoxon test but the first assumes equal variance whereas the second applies Welch's correction and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
-Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
-Still, these details can make a difference and provide unwanted researcher degrees-of-freedom.
+The present example illustrates this ambiguity, 
+because the functions `wilcox.test` and the combinations of `rank` and `t.test` are both valid implementation of the Mann–Whitney–Wilcoxon test.
+However, whereas the former assumes equal variance whereas the latter is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
+Mentioning every such detail in a written preregistration plan is almost impossible.
+Still, these details impact the results and provide unwanted researcher degrees-of-freedom.
 
@@
 To use this function one needs a compatible dataset. <!--AB what is "this" function? What does "compatible" mean?-->
-At the time of preregistration there is of course no actual data available.
+At the time of preregistration, researchers should remain insulated from real data.
 We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
@@
 
+<!-- CJ Repetition from above!!!-->
 Next, we add dynamic code blocks to our Markdown report that implement the planned reporting of the final results.
@@
 ```
+<!-- CJ This section also needs a concluding paragraph to wrap it up and present a take-home message, before moving on to Discussion-->
 
@@
 
-We demonstrated how the R package `repro` supports researchers in creating reproducible repositories and reproducible manuscripts.
+This paper demonstrated how the R package `repro` supports researchers in creating reproducible repositories and reproducible manuscripts.
 These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
 The workflow we present here rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to all but guarantee reproducibility.
-
-We then proceeded to show preregistration as code (PAC) as one example of how such rigorous and automated reproducibility workflow may enable other innovations.
-In a PAC researcher write all analysis code and a reproducible manuscript that already renders the entire reporting of statistical and computational results before they gather data.
+We first demonstrated how to create a reproducible research project.
+Then, we illustrated how such a project can be reproduced - either by the original author and/or collaborators, or by a third party.
+We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations,
+such as the preregistration as code (PAC).
+A PAC incorporates all planned analysis code into a reproducible manuscript that already renders the expected results.
 These results are based on simulated data, so that there is no danger of peeking at the real data while writing the scripts.
+The entire manuscript, including planned analyses, is preregistered.
 This way, all researcher's degrees of freedom are eliminated.
@@
 `repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
-While we have invested much effort to ensure this goal, the package is still in early development and undoubtedly has rough edges and even bugs that went undetected.
-We hope that despite the early development stage, `repro` can support researchers to work more reproducible.
-
+Despite our efforts to ensure this goal, the package is still in development and undoubtedly has limitations.
 The workflow we present here introduces several software solutions which in itself presents a danger to reproducibility.
-To profit from automatic and convenient reproduction researcher must install Git, Make, and Docker.
+To benefit from automatic and convenient reproduction researcher must install Git, Make, and Docker.
 However, Git and Make are itself included in the Docker image created by `repro`.
@@
 The only way to remove the reliance on such external software is to turn the Docker image into a full operating system that subsequently can be installed and run on almost any modern computer.
-This process is technically possible and would guarantee reproducibility for decades without any software dependency.
-However, this process requires much technical knowledge and is currently not supported by `repro`.
+This process is technically possible and would guarantee reproducibility for decades<!--CJ It's hard to know this for sure--> without any software dependency.
+However, this process requires much technical knowledge and is currently not facilitated by `repro`.
 In any case and without such precautions do the R Markdown, the Makefile and the Dockerfile provide information that allow researchers to manually trace the computational steps and recreate the computational environment.
 The Makefile, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are, for whatever reason, unable to run Make.
-Similarly, even when Docker does not work, the Dockerfile still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
+Hypothetically, even if Docker were unavaible, the Dockerfile still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
 
@@
 
-We realize that PACs are challenging.
+We realize that PACs are challenging.<!--CJ Why a separate limitations section for PACs???
+I don't think this is a very strong section, and if PACs are just an example of innovations afforded by repro, then it does not warrant a separate Limitations section.
+Discuss only the MOST IMPORTANT points below in the PAC section instead-->
 Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove unwieldy for a PAC.
 Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not preregistered research but as a complement.
-The rigidness of PAC is only than to much of a good thing, if one holds the view that researchers can for no reason deviate from their preregistration.
+The rigidity of PAC is only than to much of a good thing, if one holds the view that researchers can for no reason deviate from their preregistration.
 However, we are convinced that many reasons justify changing the preregistered code, but are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
@@
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
-We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
+We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".<!--CJ This property is relevant for the whole workflow, so it needs to go into the general discussion-->
 This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
@@
 Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.
-
+<!-- CJ: Toooooo detailed!!!-->
 Please note that the analytical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes, and it is not necessarily optimal for addressing your research question.

### 2021-08-03 manuscript.Rmd
review Casper's changes

@@
 First, the present paper introduces the reader to a workflow that maximizes the potential reproducibility of a research project.
-The end goal of this workflow is a project that can be completely, automatically, and unambiguously reproduced.<!--CJ From tutorial-->
+The end goal of this workflow is a project that can be completely, automatically, and unambiguously reproduced.
 This means that any manual steps, such as copy and pasting results, manual modification of data (e.g., manually removing observations or manually recoding variables),
@@
 in which all steps required to produce the (published) manuscript from the raw (and ideally, open) data are well-documented
-and automatically reproducible with a single command.<!--CJ From tutorial-->
+and automatically reproducible with a single command.
 
@@
 4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
-<!-- CJ 5.
-Missing steps: Some of the steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., copy-pasting results from one program to another; reverse coding items on the fly).-->
-<!--AB Caspar's 5th point is important.
-In my mind, this best fits to point #3, which adressed an insufficient level of coding/automation/documentation of the necessary steps taken.
-Can we merge this into one new point #3?
-I will commit a suggested change of point #3 here (see above);.-->
-
-We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term and cross-platform computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]: <!-- CJ In fact, this is the central message of your paper: "We have developed this new workflow, based on four pillars." I would suggest mentioning this very clearly in the opening paragraph.
-The opening paragraph should at least:
-1) Introduce the topic area
-2) Introduce the problem/knowledge gap
-3) Introduce how you will fill this knowledge gap-->
+
+We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term and cross-platform computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:
 
@@
 
-**Docker** is a popular protocol for containerization.
+**Docker** is a popular tool for containerization.
 It manages software dependencies by constructing a virtual software environment independent of the host software environment.
@@
 A Docker image should contain *all* software dependencies used in an analysis - not just R-packages, but also R and Rstudio, and even the computer's operating system.
-This is important because some of its functionality may impact the workings of higher-order software, such as calls to random number generators, linear algebra libraries and such within R.<!-- CJ From tutorial-->
-
-Note that the software environment of the Docker image is completely separate from software installed on your computer.<!-- CJ This paragraph is also relevant for users that are not advanced!!!
-Moved it here, from the tutorial-->
+This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators, linear algebra libraries and such within R.
+
+Note that the software environment of the Docker image is completely separate from software installed on your computer.
 This separation is excellent for reproducibility, but take some getting used to.
@@
 in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
-In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.<!-- CJ From tutorial-->
+In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.
 
@@
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
-abstract: A tutorial about doing the same thing more than once.<!-- CJ I actually really like this title, and would consider it for the main manuscript.
-E.g., "Reprodicible research: A tutorial about.."-->
+abstract: A tutorial about doing the same thing more than once.
 output: html_document
@@
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
-abstract: A paper about recalculating the same thing more than once.<!-- CJ Is it correct that this is different from the example above?-->
+abstract: A tutorial about doing the same thing more than once.
 output: html_document
@@
 
-This information makes explicitly declares the project's dependencies on files and R packages.<!-- CJ How about version numbers and sources?
-Would it make sense to declare a dependency on renv and include a .renv file?-->
+This information makes explicitly declares the project's dependencies on files and R packages.
 `repro` uses this information to construct a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
-These two files together form the basis for consistency within a research project and consistency across different systems.<!--CJ Consider explaining this in the theoretical intro-->
+These two files together form the basis for consistency within a research project and consistency across different systems.
 The function `repro::automate()` converts the metadata from all R Markdown files in the project (all files with the ending `.Rmd`) to a Makefile and a Dockerfile.
@@
 It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, to keep the Makefile and Dockerfile up to date.
-The function `repro::automate()` does not overwrite manual changes<!--CJ What does that mean?-->, and there is no harm in running it too often.<!-- CJ Can that not be automated?
-Or at least provide a suggestion when to run it?-->
+There is no harm in running it too often.
 Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
-<!--AB: why is this addition not automated?
-Feels strange given that this is our recommended best practice?!
-CJ: Agreed.-->
-
-<!-- CJ What i'm missing here is something to wrap up the tutorial. Do users now have a project that could be reproduced? The tutorial should not end until that is the case.-->
 
@@
 
-<!--CJ Suggestion: Instead of explaining reproduction right in the middle of the tutorial for setting up a project, add a later section called "reproducing a project", and provide an example project for users to reproduce.-->
 When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:
@@
 The commands can be pasted into a terminal and will reproduce the project in many cases.
-<!-- CJ This paragraph needs to be elaborated further-->
 
@@
 
-This section is only for advanced users who want to overcome some limitations of `repro`.<!-- CJ Consider sticking this in a vignette, which would allow it to be updated more regularly-->
+This section is only for advanced users who want to overcome some limitations of `repro`.
 If you read this paper the first time you probably want to skip this section.
@@
 The second automatically generated Dockerfile is `.repro/Dockerfile_packages`.
-Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.<!-- CJ Suggestion: Make a directory called "dockerfiles", and name the files "image", "external_dependencies", and "rpackages"-->
+Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.
 
 Finally, there is one manually edited Dockerfile: `.repro/Dockerfile_manual`.
-<!-- CJ I would expect this file to contain a "manual" based on the title.
-Consider if there is a better name, eg: .repro/external_dependencies-->
 It is blank by default, and can be used to add further dependencies outside of R, like system libraries or external software.
 
-<!--Note that the software environment of the Docker image is completely separate from software installed on your computer. CJ This paragraph is also relevant for users that are not advanced!!!
-This separation is excellent for reproducibility, but take some getting used to.
-For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the Docker image.
-Each dependency that you want to use within the Docker image must be explicitly added as a dependency.-->
-Using Docker mayrequire you to install software on an operating system that may not be familiar to you.
+Using Docker may require you to install software on an operating system that may not be familiar to you.
 The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.
 The most convenient way to install software on Ubuntu is through its package manager `apt`.
-The following snipped installs Python.<!--CJ You have not explained why one might want to install python.
-I think this is too specific for the general paper; put in a vignette and make a separate vignette for "Adding external dependencies",addressing python and others-->
+The following snipped, for example, installs Python but works the same for any other software users might want to install.
 If added to `.repro/Dockerfile_manual` the Docker image will have Python installed.
@@
 
-<!--
-Docker eventually requires a single Dockerfile to run, so `repro::automate()` simply concatenates the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the R project.CJ Too much technical detail about the implementation.
-This section promised to help advanced users enhance the functionality of repro - but what functionality are they gaining by knowing that repro concatenates these files?
--->
+Docker eventually requires a single Dockerfile to run, so `repro::automate()` simply concatenates the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the R project.
 
@@
 
-The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but usually should not modify.<!-- CJ Under what conditions could you modify it? -->
+The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but usually should not modify.
 This Makefile coordinates the interplay between Make and Docker and contains targets for building (with `make docker`) and saving (with `make save-docker`) the Docker image.
 Additionally, it provides facilities to execute commands within the container.
-If you write a computational recipe<--CJ this term has not been defined--> for a target it will be by default evaluated using the locally installed software.
+If you write a computational recipe for a target it will be by default evaluated using the locally installed software.
 To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first Make example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
@@
 Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automations.
-<!-- CJ I would only leave one paragraph for advanced users, stating that they can manually change the Dockerfiles and makefiles following the vignette on the repro website.
-Skip the details about implementation in this paper!!!
-That's not what a tutorial is for-->
 
@@
 In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
-<!-- CJ This is not convincing:
-We believe, however, that combining ideas from reproducibility and preregistration can turn this "neat" idea into a tool with tangible benefits for the research community.-->
 But although the notion of preregistering code has been around for a while [cf. @steegenMeasuringCrowdAgain2014],
-it has not gained much traction - perhaps because, until now, this has constituted an extra non-standard step in the research process.
-This tutorial integrates the preregistration of code into the research workflow
-by encouraging researchers to preregister the whole manuscript as a reproducible dynamic document.
+it has not gained much traction---perhaps because, until now, this has constituted an extra non-standard step in the research process.
+This tutorial integrates the preregistration of code into the reproducible research workflow
+by encouraging researchers to preregister the whole manuscript as a dynamic document.
 
@@
 
-These function again should be accept the output of `planned_analysis()` as input.
-
-<!-- CJ: The text below is about reporting standards; it's useful, but not directly linked to the text above-->
+These function again should accept the output of `planned_analysis()` as input.
+
+The goal of this function is to describe the results to readers automatically.
+This way, the preregistration specifies not only the computation  also how the computation's results are reported.
 The output of this function should be formatted in Markdown.
@@
 We suggest writing a main analysis function that captures the full process of testing a research question.
-This function receives a dataset as input and performs all statistical processing. <!--AB say anything about its output? is the Markdown the output?-->
+This function receives a dataset as input and performs all statistical processing.
 For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
@@
 
-<!--AB here is kind of a break in the text-->
-To use this function one needs a compatible dataset. <!--AB what is "this" function? What does "compatible" mean?-->
-At the time of preregistration, researchers should remain insulated from real data.
+It is difficult to write code when it is not clear how exactly the input should look like, however, at the time of preregistration, researchers should remain insulated from real data.
 We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
@@
 ```
-<!--AB it is odd that the return variable is called effect because it is a dataset-->
 
@@
 Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a given standardized effect size (a Cohen's d of .2).
-This simulated dataset fits perfectly into our planned analysis with a sample size of 400 and 10 items: <!--AB in what way? please explain-->
+This simulated dataset fits perfectly into our planned analysis with a sample size of 400 and 10 items:
 
@@
 If we want to detect an effect of size `r choosen_d` with an `r choosen_power*100`% probability, our simulation shows that we require at least `r minn$n` participants.
-
-<!--AB please add a figure number and a proper citation to the figure. Also, tell people where they can find the code to run the power analysis themselves-->
 
@@
 ```
-<!-- CJ This section also needs a concluding paragraph to wrap it up and present a take-home message, before moving on to Discussion-->
 
@@
 The only way to remove the reliance on such external software is to turn the Docker image into a full operating system that subsequently can be installed and run on almost any modern computer.
-This process is technically possible and would guarantee reproducibility for decades<!--CJ It's hard to know this for sure--> without any software dependency.
+This process is technically possible and would guarantee reproducibility for decades without any software dependency, assuming hardware that conforms to the x86 instruction set architecture continues to be available.
 However, this process requires much technical knowledge and is currently not facilitated by `repro`.
@@
 
-We realize that PACs are challenging.<!--CJ Why a separate limitations section for PACs???
-I don't think this is a very strong section, and if PACs are just an example of innovations afforded by repro, then it does not warrant a separate Limitations section.
-Discuss only the MOST IMPORTANT points below in the PAC section instead-->
+We realize that PACs are challenging.
 Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove unwieldy for a PAC.
@@
 In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
-We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".<!--CJ This property is relevant for the whole workflow, so it needs to go into the general discussion-->
+We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
 This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
@@
 Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->
+Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
 
 Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.
-<!-- CJ: Toooooo detailed!!!-->
 Please note that the analytical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes, and it is not necessarily optimal for addressing your research question.

### 2021-08-03 manuscript.Rmd

@@
 We pursue two aims.
-First, the present paper introduces the reader to a workflow that maximizes the potential reproducibility of a research project.
-The end goal of this workflow is a project that can be completely, automatically, and unambiguously reproduced.
-This means that any manual steps, such as copy and pasting results, manual modification of data (e.g., manually removing observations or manually recoding variables),
-are replaced with reproducible code. 
-The result of this workflow will be a portable research project containing all digital research artefacts,
-in which all steps required to produce the (published) manuscript from the raw (and ideally, open) data are well-documented
-and automatically reproducible with a single command.
+
+First, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
 
@@
 
-We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term and cross-platform computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:
+We have developed a workflow that achieves long-term and cross-platform computational reproducibility of scientific data analyses.
+It leverages established tools and practices from software engineering and rests on four pillars that address the aforementioned causes of non-reproducibility: [@Peikert2019]:
 
@@
 One way to do this is through the visual user interface in the RStudio Git pane.
-Click on the empty box next to the file you want to stage. A check mark appears to indicate that the file is staged.
+Click on the empty box next to the file you want to stage.
+A check mark appears to indicate that the file is staged.
 After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
@@
 R Markdown is a type of dynamic document well-suited to the RStudio user interface.
-The natural language text in an R Markdown file can be formatted using Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
+The text of an R Markdown is formatted by Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
 The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
@@
 
-The metadata the `repro::automate()` function relies on can only express dependencies R packages for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
+The metadata the `repro::automate()` function relies on can only express R packages as dependencies for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile.
 The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
@@
 These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
-Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D therefore leaving out important information or express the dependencies directly within the Makefile.
+Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D, therefore leaving out important information or express the dependencies directly within the Makefile.
 
@@
 The most convenient way to install software on Ubuntu is through its package manager `apt`.
-The following snipped, for example, installs Python but works the same for any other software users might want to install.
-If added to `.repro/Dockerfile_manual` the Docker image will have Python installed.
+
+If the following snipped is added to `.repro/Dockerfile_manual` the Docker image will have Python installed.
 
@@
 
-Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
-Preregistration increases the credibility of empirical results by separating theory-driven planned analyses from unplanned exploratory analyses.
+Defining the research questions and planning data analysis before observing the research outcomes is referred to as *preregistration* [@NosekRevolution2018].
+Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
 Preregistration seeks to reduce "researcher degrees of freedom" [@reference] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could influence the results.
-However, there are several shortcomings to preregistration.
-One shortcoming is that written study plans are often multi-interpretable.
-Even when several researchers set out to investigate the same hypothesis using the same data, their approaches vary considerably [@silberzahnManyAnalystsOne2018].
+Preregistrations, however, have several shortcomings.
+One shortcoming is that written study plans are often multi-interpretable and often do not achieve the level of unambiguity that are required to effectively zero the researcher degrees of freedom.
+Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
 The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
 These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
-However, templates cannot remove researchers degrees of freedoms completely,
-because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
+However, templates cannot remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
 A second limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
@@
 We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, in PAC, the preregistration is simply a first draft of the final manuscript,
-thus sparing authors from writing (and editors and reviewers from evaluating) two separate documents.
+First, in PAC, is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
 Relatedly, writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
@@
 Third, a PAC removes any ambiguity regarding the translation of verbal analysis plans into code.
+A PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
+Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
+PAC therefore minimizes researchers degrees of freedom more effectively than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
 Fourth, despite its rigor, PAC does accommodate data-dependent decisions if these can be formulated as code.
@@
 For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
-Fifth, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
-Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
-PAC therefore minimizes researchers degrees of freedom more effectively than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
-Sixth, deviations from the preregistration are clearly documented, because they are reflected in changes to the code which are managed and tracked with version control.
+Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
+With PAC researchers can precisely calibrate how much they want to explore.
+When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
+Fifth, deviations from the preregistration are clearly documented, because they are reflected in changes to the code which are managed and tracked with version control.
 One could even make a commit labeled "Necessary deviations from the preregistration after data collection",
@@
 
-When writing a PAC, researchers ought to strive to meet three goals:
+When planning a study and hence when writing a PAC, researchers ought to strive to meet three goals:
 First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
@@
 
-Describing the preregistration in the form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
-One particular advantage is the strong foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
-It can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section [@ref provide specific good examples].
+Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
+One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
+Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
 
@@
 It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
-Writing incomplete code is technically impossible or at least researchers would notice if they e.g. forgot to aggregate the data etc.
+Writing incomplete code is technically impossible, or at least researchers would notice if something is omitted (e.g. if they forgot to aggregate the data).
 
@@
 
-To write a PAC researchers should start by drafting a manuscript in form of an R Markdown.
+To write a PAC, researchers should start by drafting a manuscript in the form of an R Markdown.
 When they have completed the method section, they should begin writing code.
-We suggest writing three functions for each hypothesis, one for the planned analysis, one for the simulating the data and one for reporting the results.
-At the heart of a PAC lies the planned analysis so researchers should first draft a function called `planned_analysis()` that captures the description in the method section.
+We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
+At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
 This function should receive the data and derive relevant results from it.
@@
 
-Because it is difficult to draft such function without interactively test it, researcher should than proceed to write a function for simulating data.
+It is difficult to write code when it is not clear how exactly the input should look like, however, at the time of preregistration, researchers should remain insulated from real data.
+We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
+Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
+The following code simulates a dataset with one observed variable in two groups.
 
@@
 
-When the researchers are satisfied with the function `planned_analysis()` they can think about how they would like to report the results of the analysis, via tables, plots and text.
+When the researchers are satisfied with the function `planned_analysis()`, they can think about how they would like to report the analysis results via tables, plots, and text.
 The implementation of this reporting should be in the function `report_analysis()`.
 
-```{r}
-report_results <- function(results){
+```{r, eval=FALSE}
+report_analysis <- function(results){
   # 1. create markdown tables from results
@@
 
-These function again should accept the output of `planned_analysis()` as input.
-
-The goal of this function is to describe the results to readers automatically.
-This way, the preregistration specifies not only the computation  also how the computation's results are reported.
+This function should again accept the output of `planned_analysis()` as input.
 The output of this function should be formatted in Markdown.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
+The idea is to automatically generate the full results section from the analysis.
+This way, the preregistration specifies not only the computation but also how the computation's results are reported.
+Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R objects to help with this objective.
 Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
+The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it generates not only tables but also a straightforward interpretation of the effects as actual prose (e.g.
+it verbally quantifies the size of an effect).
 
@@
 
-To summarize researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
-After they gathered the actual data, they replace the simulated data, run `planned_analysis()` and write the discussion.
+To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration.
+After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data) and write the discussion.
 
@@
 
-To guide the reader through a preregistration as code (PAC), we will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
-
-> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"
-
-This example serves only didactic purposes and not to derive substantive claims.
-Nevertheless, everything else is kept as close to a real research question as possible.
-
-At the heart of a PAC lies the planned analysis.
-We suggest writing a main analysis function that captures the full process of testing a research question.
-This function receives a dataset as input and performs all statistical processing.
-For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
-As an example of a data-dependent decision, we included a rank transformation if the skewness was greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]
+We argued that PAC has several advantages over classical preregistration and have outlined how to implement it.
+To illustrate how PAC works in practice and help researchers implementing PAC themselves, we have conducted a PAC that serves as a guiding example.
+We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
+
+> "Is there a mean difference in the personality trait 'Machiavellism' between self-identified females and males?"
+
+Again, we propose a preregistration format that closely resembles a classical journal article, but uses simulated data and dynamic document generation to create a document that starts out as a preregistration and eventually becomes the final report.
+The complete preregistration source is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
+In this section, we show code excerpts of this preregistration (formatted in monospace) and explain the rationale behind them.
+
+```{r, echo=FALSE}
+show <- function(x, wrap = TRUE, cat = TRUE, fence = TRUE){
+  stopifnot(length(x) == 1L)
+  if(wrap)x <- str_c(str_wrap(str_split(x, "\\n\\n")[[1]]), collapse = "\n\n")
+  if(!isFALSE(fence)){
+    if(isTRUE(fence))fence <- "markdown"
+    fence <- str_c("````", fence)
+    x <-
+      str_c(fence,
+            x,
+            "````",
+            sep = "\n",
+            collapse = "\n")
+  }
+  if(cat){
+    cat(x)
+    return(invisible(x))
+  } else {
+    return(x)
+  }
+}
+show_prereg <-
+  function(section,
+           file = "preregistration.Rmd",
+           wrap = TRUE,
+           cat = TRUE,
+           fence = TRUE) {
+    prereg <- readr::read_lines(here::here(file))
+    start <- stringr::str_c("<!--begin-", section, "-->")
+    end <- stringr::str_c("<!--end-", section, "-->")
+    start_pos <- which(prereg == start)
+    end_pos <- which(prereg == end)
+    if (length(start_pos) != 1L || length(end_pos) != 1L)
+      stop("No section '", section, "' found.")
+    excerpt <-
+      str_c(prereg[(start_pos + 1):(end_pos - 1)], collapse = "\n")
+    show(excerpt, wrap = wrap, cat = cat, fence = fence)
+  }
+```
+
+As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary background information and literature to understand the context and purpose of the research question.
+This example is drastically shortened for illustration purposes:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("theory")
+```
+
+After researchers have provided the research question, they typically proceed to explain how they want to study it.
+For simplicity, we will use already published data that we have not yet analyzed:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("method")
+```
+
+We choose the following statistical procedure because many researchers are familiar with it[^caution-mww]:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("ttest")
+```
 
 [^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
-The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
-
-This function is one possible translation of the verbal description above:
+The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
+
+We then translated this methods section into the following `planned_analysis()` function:
 
@@
 
-Importantly, natural language descriptions do not map unambiguously onto computer code implementations. 
-The same textual descriptions may be implemented differently by different persons depending on their statistical and programming knowledge but also based on their own expectations and beliefs.
-The present example illustrates this ambiguity, 
-because the functions `wilcox.test` and the combinations of `rank` and `t.test` are both valid implementation of the Mann–Whitney–Wilcoxon test.
-However, whereas the former assumes equal variance whereas the latter is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
-Mentioning every such detail in a written preregistration plan is almost impossible.
-Still, these details impact the results and provide unwanted researcher degrees-of-freedom.
-
-It is difficult to write code when it is not clear how exactly the input should look like, however, at the time of preregistration, researchers should remain insulated from real data.
-We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
-Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
-The following code simulates a dataset with one observed variable in two groups.
-
-```{r simulate_data}
-```
-
-Here, we decided to randomly draw non-normal data to test how well our procedure performed under non-optimal situations.
-Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a given standardized effect size (a Cohen's d of .2).
-This simulated dataset fits perfectly into our planned analysis with a sample size of 400 and 10 items:
-
-***
-
-```{r}
-set.seed(112358) # fix state of RNG
-results <- 
-  planned_analysis(
-    simulate_data(
-      n = 400,# Sample size
-      df = 8, # Chisq with DF = 8 has skewness = 1
-      d = .2, # Cohen's D
-      i = 10  # Number of items (should not effect anything)
-    )
-  )
-```
-
-<!-- CJ Repetition from above!!!-->
-Next, we add dynamic code blocks to our Markdown report that implement the planned reporting of the final results.
-A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
-Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
-Here we use the `report` package to create a preview of how our reporting will look like based on simulated data:<!--AB add citation to the report package-->
-
-```{r, results='asis', eval=FALSE}
-report::report(results$test)
-```
-
-This call to the `report()` function will yield the following automatically generated output:
-
-```{r, results='asis', echo=FALSE}
-report::report(results$test)
-```
-<!--AB it is difficult to discern the automatically generated text and out manuscript text even though there is the horizontal line. Not sure how to improve this.-->
-***
+This function illustrates two advantages of PAC.
+First, a PAC can easily include data-dependent decisions creating different analysis branches under different conditions.
+Second, it highlights how difficult it is to describe a statistical analysis precisely.
+The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and their expectations and beliefs.
+One example would be using the function `wilcox.test` instead of the combinations of the functions `rank` and `t.test`.
+Either of them are valid implementations of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
+In contrast, the second applies Welch's correction by default and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
+Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
+Still, these details can make a difference and provide unwanted researchers' degrees of freedom.
+
+The function `planned_analysis()` together with the function `simulate_data()` (not shown here) can be used to justify the planned sample size.
+To that end, `simulate_data()` is repeatedly called with increased sample size and the proportion of significant results (power) is recorded.
+The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
+The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R).
+
+Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
 
@@
 
-To determine a minimal sample size for this study we repeat this simulation across several combination of effect sizes and sample sizes.
-One standard recomendations for choosing a sample size is to have an expected power of 80% based on the expected effect size.
-@Jones2013 found in three studies gender differences of .24, .29, and .35 for their machiavilism subscale.
-If we want to detect an effect of size `r choosen_d` with an `r choosen_power*100`% probability, our simulation shows that we require at least `r minn$n` participants.
-
-```{r, echo=FALSE}
+```{r, include=FALSE}
+fig_caption <- str_c("Results of simulation for the power analysis. The cross indicates the sample size that archives ", choosen_power*100, "% assuming a Cohen's D of ", choosen_d, ".")
+```
+
+```{r power, echo=FALSE, fig.cap=fig_caption}
 res %>%
@@
 The reproducible workflow we present here is heavily automated and hence promotes reusability.
+
 Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
@@
 Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.
-Please note that the analytical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes, and it is not necessarily optimal for addressing your research question.
+
+Please note that the statistical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and is not necessarily optimal for addressing your research question.
+
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].

### 2021-08-04 manuscript.Rmd
incorparoate Parthena's comments

@@
   It includes a tutorial on how to make a research project reproducible with the aid of `repro`.
-  Finally, it introduces novel applications of reproducibility tools for scientific research;
-  including the preregistration of study plans as reproducible computer code.
-  This circumvents the shortcomings of ambiguous preregistrations that allow for researcher degrees of freedom,
-  because computer code describes study plans more precisely than prose.
+  Finally, it introduces novel applications of reproducibility tools for scientific research, notably it includes the preregistration of study plans as reproducible computer code.
+  Since computer code describes study plans more precisely than prose, this circumvents the shortcomings of ambiguous preregistrations that give researchers undesired degrees of freedom.
   Reproducibility made convenient with automation has hence a wide range of applications to accelerate scientific progress.
@@
 
-Scientists increasingly strive to make research data, materials, and analysis code openly available.
+Scientists increasingly strive to make research data, materials, and analysis code openly available i.e. within the Open Science movement.
 Sharing these types of digital research products has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
@@
 Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
-The credibility of empirical research hinges on its objectivity.
-Objectivity in this context means "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p.
-22][^popper].
+The credibility of empirical research results hingeson their objectivity.
+Objectivity in this context means "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
+Only something reproducible can be tested and understood by anyone.
 Therefore, reproducibility has long been considered integral to empirical research.
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
-This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
-We now have the technology to make all digital products related to a research project available so that a research project can be reproduced with minimal effort.
+This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (multiple versions of code and data, copy-and-paste errors, missing software, ambiguous computation), and there is a dire need for reproducibility standards.
+With the technological advancement, it is now possible to make all digital products related to a research project available so that a research project can be reproduced with minimal effort.
 
@@
 This workflow builds on several technical solutions originating in the software engineering community. 
-We discuss how these solution address four common threats to reproducibility [see @Peikert2019 for details] and form a coherent workflow.
+We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they form a coherent workflow.
 We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
-Second, we discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
-One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
+Second, we discuss opportunities that the reproducible workflow affords to improve not only how research outputs are created and recreated, but also how research itself is conducted.
+One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
 PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including the Introduction, Methods, and Results sections.
-The resulting dynamic document closely resembles the final manuscript, but uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
-Once the empirical data are analyzed, the Results are updated, and the Discussion is written.
+The resulting dynamic document closely resembles the final manuscript, but temporarily uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
+Once the empirical data are analyzed, the Results are updated automatically, and the Discussion can be written.
 
@@
 The first part of this tutorial outlines threats to reproducibility and how the here presented reproducibility standard addresses these threats.
-The second part is a tutorial that demonstrates how this standard can be easily adopted with the help of the R package `repro`.
 
@@
 It may then be unclear which version of code and data was used to produce some reported set of results.
-2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
+2.	**Copy-and-paste errors**; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
+3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
 4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
@@
 
-1. Version control
-2. Dynamic document generation
-3. Dependency tracking
-4. Software management
+1. *Version control*
+2. *Dynamic document generation*
+3. *Dependency tracking*
+4. *Software management*
 
@@
 Version control allows a clear link between which results were generated by which version of code and data.
+This is important for reproducibility because only if it is obvious which version of data and code produced which results they can be said to be reproducible.
 We recommend using **Git** for version control, because of its widespread adoption in the R community.
@@
 `r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
-The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
+The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like tidyverse [@tidyverse] and all its dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
 A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.

### 2021-08-04 manuscript.Rmd
alternative title suggestion

@@
 ---
-title: A Hitchhiker's Guide to Reproducible Research in R
+title: "Reproducible research: A tutorial about doing the same thing more than once in R"
 author:
@@
 ---
-title: A Hitchhiker's Guide to Reproducible Research in R
+title: "Reproducible research: A tutorial about doing the same thing more than once in R"
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
-abstract: A tutorial about doing the same thing more than once.
+abstract: A Hitchhiker's Guide to Reproducible Research in R
 output: html_document
@@
 ---
-title: A Hitchhiker's Guide to Reproducible Research in R
+title: "Reproducible research: A tutorial about doing the same thing more than once in R"
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
-abstract: A tutorial about doing the same thing more than once.
 output: html_document

### 2021-08-04 manuscript.Rmd
should have been bold

@@
 
-1. *Version control*
-2. *Dynamic document generation*
-3. *Dependency tracking*
-4. *Software management*
+1. **Version control**
+2. **Dynamic document generation**
+3. **Dependency tracking**
+4. **Software management**
 
@@
 Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
-Preregistration seeks to reduce "researcher degrees of freedom" [@reference] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could influence the results.
+Preregistration seeks to reduce "researcher degrees of freedom" [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could influence the results.
 Preregistrations, however, have several shortcomings.

### 2021-08-04 manuscript.Rmd
alternative title suggestion (#198)

@@
 ---
-title: "Reproducible research: A tutorial about doing the same thing more than once in R"
+title: "Reproducible Research: A Tutorial on how to do the same thing more than once"
 author:
@@
 ---
-title: "Reproducible research: A tutorial about doing the same thing more than once in R"
+title: "Reproducible Research: A Tutorial on how to do the same thing more than once"
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier

### 2021-08-04 manuscript.Rmd
forgot "R"

@@
 ---
-title: "Reproducible Research: A Tutorial on how to do the same thing more than once"
+title: "Reproducible Research: A Tutorial on how to do the same thing more than once in R"
 author:
@@
 ---
-title: "Reproducible Research: A Tutorial on how to do the same thing more than once"
+title: "Reproducible Research: A Tutorial on how to do the same thing more than once in R"
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier

### 2021-08-04 manuscript.Rmd
add abbrevation WORCS
remove "fail to guarantee" closes #168

@@
   - short: PAC
-    long: Preregistration as 
+    long: Preregistration as Code
   - short: Gb
@@
     long: Comprehensive R Archive Network
+  - short: WORCS
+    long: Workflow for Open Reproducible Code in Science
 repro:
@@
 One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
-However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
+While version control and dynamic document generation are becoming more common, we have argued that two more components are required and that each component alone is unlikely to guarantee reproducibility [@Peikert2019; @vanlissa2020worcs].
 In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.

### 2021-08-04 manuscript.Rmd
note that WORCS does not contain Make
compare renv and docker in light of worcs

@@
 Make therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
+Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] does not contain this element of dependency tracking explicitly but its strict structure of only containing one definite R Markdown still makes dependencies between files unambiguous.
 
@@
 A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
+Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] relies for software management on the R package `renv`[@R-renv] which is more lightweight and easier to use on the one hand, but not as comprehensive on the other hand because it only snapshots the R packages instead of all software used.
 

### 2021-08-04 manuscript.Rmd
cite worcs as a source of PAC
include setup in reproducing a project closes #187

--- 2021-08-04 manuscript.Rmd
@@
 
-When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:
+When you or someone else who received your project want to reproduce it, they first have to install the required software, which is Make, and Docker.
+Remember, you can use the `check_*`-functions to test if these are installed:
+
+```{r}
+check_make()
+check_docker()
+```
+
+When they are setup, they can ask `repro` to explain how they should use Make and Docker to reproduce the project (or you could explain it to them):
 

### 2021-08-04 manuscript.Rmd
add summary as code

@@
 
+## Summary
+
+1. Install the `repro` package:
+
+```{r, eval=FALSE}
+options(
+  repos = c(aaronpeikert = 'https://aaronpeikert.r-universe.dev',
+            CRAN = 'https://cloud.r-project.org')
+)
+install.packages('repro')
+```
+
+2. Check the required software:
+
+```{r}
+repro::check_git()
+repro::check_make()
+repro::check_docker()
+```
+
+3. Create an R project or use an existing one. Do not forget to add `repro` metadata (i.e. packages, scripts, data).
+
+```yaml
+repro:
+  scripts:
+    - R/load.R
+  data:
+    - data/mtcars.csv
+  packages:
+    - tidyverse
+```
+
+The example `repro` project already has this metadata:
+
+```{r,eval=FALSE}
+repro::use_repro_template("/some/folder")
+```
+
+4. Let `repro` generate Docker- and Makfile:
+
+```{r, eval=FALSE}
+repro::automate()
+```
+
+5. Enjoy automated reproducibility:
+
+```{r}
+repro::reproduce()
+```
+
 # Advanced Features

### 2021-08-04 install.Rmd
add rtools
update required dependencies and advice windows users closes #140

@@
 1. Open a new terminal (Press Windows key + X, click on "Windows Powershell (Admin)").
-2. Paste `choco install -y r r.studio`
+2. Paste `choco install -y r r.studio rtools`
 3. Press enter
@@
 
-Please note that Docker does not work on Windows 10 Home, you need Windows Pro or Education.
+Please note that Docker does may not work on Windows 10 Home, because Docker requires a feature called virtualization which you may have to enable.
+In that case follow this [guide](https://docs.docker.com/docker-for-windows/troubleshoot/#virtualization).
+This feature is usually enabled in Windows Pro or Education.
+On Windows Home, you also require the Windows Subsystem for Linux (WSL) as well as the WSL Ubuntu extension.
 
-1. Open a new terminal (Press Windows key + X, click on "Windows Powershell (Admin)").
+1.  Open a new terminal (Press Windows key + X, click on “Windows
+    Powershell (Admin)”).
 2.  Paste `choco install -y git make docker`
 3.  Press enter.
+
+On Windows Home and systems without Hyper V:
+
+1. Open a new terminal (Press Windows key + X, click on “Windows
+    Powershell (Admin)”).
+2. Paste `choco install -y git make docker wsl`
+3. Press enter.
+4. Restart.
+5. Open a new terminal (Press Windows key + X, click on “Windows 
+    Powershell (Admin)”).
+6. Paste `choco install -y wsl-ubuntu-2004`
+7. Press enter.
 

### 2021-08-04 manuscript.Rmd
add keywords
update required dependencies and advice windows users closes #140

--- 2021-08-04 manuscript.Rmd
@@
 These so-called "Docker images" function like a virtual computer  (i.e., a sort of "sand box").
-A Docker image should contain *all* software dependencies used in an analysis - not just R-packages, but also R and Rstudio, and even the computer's operating system.
+A Docker image should contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the computer's operating system.
 This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators, linear algebra libraries and such within R.
@@
 The first step is to install the required software.
-We assume that you have installed R and RStudio already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert)[@R-repro].
+We assume that you have installed R [@R-base, version `r with(R.version, paste0(major, ".", minor))`] and RStudio [@rstudio, version 1.4] already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert)[@R-repro].
 In case you either have not installed R and RStudio or are unsure if they are up to date you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers how to install all software necessary for this tutorial in three steps.
+The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing Docker.
 Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 

### 2021-08-05 manuscript.Rmd
abstract and intro

@@
 ---
-title: "Reproducible Research: A Tutorial on how to do the same thing more than once in R"
+title: "Reproducible Research in R: A tutorial on how to do the same thing more than once"
 author:
@@
   Reproducibility has long been considered integral to the scientific method.
-  Something is called reproducible when independend researcher arrive at the same results from the same data.
+  Something is called reproducible when an independent person obtains the same results from the same data.
   Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
-  Technological advancements now enable scientists to achieve a much higher standard;
-  one in which any individual can be granted access to a digital research repository, and reproduce the analyses from raw data to final manuscript with a single command.
+  Technological advancements now enable scientists to achieve a more comprehensive standard;
+  one in which any individual can be granted access to a digital research repository, and reproduce the analyses from the raw data to the final report including all relevant statistical analyses with a single command.
   This method has far-reaching implications for scientific archiving,
@@
   are complicated to use.
-  This paper introduces `repro`; an R-package that guides researchers in the installation and use of the tools required for computational reproducibility.
-  It includes a tutorial on how to make a research project reproducible with the aid of `repro`.
-  Finally, it introduces novel applications of reproducibility tools for scientific research, notably it includes the preregistration of study plans as reproducible computer code.
-  Since computer code describes study plans more precisely than prose, this circumvents the shortcomings of ambiguous preregistrations that give researchers undesired degrees of freedom.
-  Reproducibility made convenient with automation has hence a wide range of applications to accelerate scientific progress.
+  This paper introduces `repro`, an R-package, which guides researchers in the installation and use of the tools required for making a research project reproducible.
+  Finally, we suggest the use of the proposed tools for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
+  Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions of those analyses, which merely complement the PAC as a more readable summary.
+  PAC circumvents the shortcomings of ambiguous preregistrations that may give researchers undesired degrees of freedom.
+  <Hence, reproducibility made convenient with automation has a wide range of applications to accelerate scientific progress.
 abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
@@
 
-Scientists increasingly strive to make research data, materials, and analysis code openly available i.e. within the Open Science movement.
+Scientists increasingly strive to make research data, materials, and analysis code openly available <!-- AB "i.e." does not work in this context.
+I just removed this because it's irrelevant whether this happens within the movement or not?!-->
 Sharing these types of digital research products has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
@@
 Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
-The credibility of empirical research results hingeson their objectivity.
+The credibility of empirical research results hinges on their objectivity.
 Objectivity in this context means "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
@@
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
-This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (multiple versions of code and data, copy-and-paste errors, missing software, ambiguous computation), and there is a dire need for reproducibility standards.
+This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (e.g., missing software or ambiguous documentation of all steps taken <!-- AB the previous list was difficult to understand without more context; these two points are hopefully better to understand--> ), and there is a dire need for reproducibility standards.
 With the technological advancement, it is now possible to make all digital products related to a research project available so that a research project can be reproduced with minimal effort.
@@
 We pursue two aims.
-
 First, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
-
-This workflow builds on several technical solutions originating in the software engineering community. 
-We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they form a coherent workflow.
+Second, we introduce a computational preregistration scheme based on reproducible research projects.
+<!-- AB Absätze können nicht nur einen Satz haben! Außerdem fehlte das "Second"-->
+
+<!--AB "This" geht nicht am Absatzanfang--> The proposed workflow builds on several technical solutions originating from software engineering. 
+We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they interact to form a coherent workflow.
 We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
@@
 One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
-PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including the Introduction, Methods, and Results sections.
+PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including typical sections like Introduction, Methods, and Results.
 The resulting dynamic document closely resembles the final manuscript, but temporarily uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
-Once the empirical data are analyzed, the Results are updated automatically, and the Discussion can be written.
+With these placeholders, authors preregister the exact presentation of their results without having seen real data yet.
+Once the empirical data are analyzed, the Results are then updated automatically, and the Discussion can be written to finalize the report.
 
@@
 
-This tutorial therefore demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
+Therefore, this tutorial demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
 A rigorous standard increases the likelihood that a project will remain reproducible as long as possible.
@@
 As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
-The first part of this tutorial outlines threats to reproducibility and how the here presented reproducibility standard addresses these threats.
+<!--The first part... dieser Satz wirkte verloren. Habe ich gestrichen-->
 

### 2021-08-05 manuscript.Rmd
until advanced features

@@
 
-1.	**Multiple inconsistent versions of code, data, or both**; for example
-  - the data set could have changed over time because outliers were removed at a later stage or an item was later recoded
-  - the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
+1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time <!-- the bullet points looked really strange in the formatted document-->
 It may then be unclear which version of code and data was used to produce some reported set of results.
 2.	**Copy-and-paste errors**; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
+3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another)
 4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
@@
 
-1. **Version control**
-2. **Dynamic document generation**
-3. **Dependency tracking**
-4. **Software management**
-
-The remainder of this section briefly explains why each of these four building blocks is needed and their role in ensuring reproducibility.
-
-**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the development phase of a research project.
+<!--why boldface? looks odd and it is easy to confuse with section headings-->
+1. Version control
+2. Dynamic document generation
+3. Dependency tracking
+4. Software management
+
+The remainder of this section briefly explains why each of these four building blocks is needed and details their role in ensuring reproducibility.
+
+**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime (and in particular, in the development phase) of a research project.
 Version control allows a clear link between which results were generated by which version of code and data.
@@
 Ideally, each commit has a message that succinctly describes these changes.
-It is good practice to make commits for concrete milestones, e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3".
+It is good practice to make commits for concrete milestones (e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3").
 This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
-Each commit "knows" its ancestor, and they form a timeline together.
+Each commit refers back to its ancestor, and they form a timeline together.
 The entirety of commits (i.e. the version-controlled project) is called a repository.
@@
 Dynamic document generation integrates both steps.
-Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allows interested readers deeper understanding [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
-**R Markdown** uses Markdown for writing the text and R (or other programming languages) for the analysis.
+Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allowing interested readers to gain a deeper understanding of the contents [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
+**R Markdown** uses Markdown for formatting text and R (or other programming languages) for writing the statistical analysis.
 Markdown is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
@@
 However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
-In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
+In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or dependencies on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
 
@@
 Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
-This ensures that the simulation is always run before the manuscript is built, and that the manuscript is rebuilt if the simulation code was changed.
+This ensures that the simulated data are present before the manuscript is built, and that the simulation is rerun and the manuscript is rebuilt if the simulation code was changed.
 Make therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
@@
 A Docker image should contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the computer's operating system.
-This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators, linear algebra libraries and such within R.
+This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.
 
 Note that the software environment of the Docker image is completely separate from software installed on your computer.
-This separation is excellent for reproducibility, but take some getting used to.
+This separation is excellent for reproducibility, but takes some getting used to.
 For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the Docker image.
@@
 To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
-While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
-To remove this hurdle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
+While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
+To remove this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
 Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
@@
 
+<!--AB according to our own license in the original paper, we need to state the license under which the image was released. Please complement this information.-->
 ```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency tracking (Make), software management (Docker), and dynamic document generation (R Markdown).
@@
 
-One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be time-intensive.
-To lower the threshold to adoption, the R-package `repro` introduces helper functions that wrap more complicated (and powerful) tools.
-These so-called wrappers guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user should do in layman's terms.
+One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be quite time-intensive.
+To lower the threshold to adoption, the R-package `repro` introduces helper functions that wraps and hides the complexity of the necessary more complicated (and powerful) tools.
+The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user is supposed to do next.
 We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
@@
 This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
-It describes basic steps that we expect to apply to all users.
-Your specific situation might involve additional, more specialized steps.
+It describes basic steps that we expect to be representative of a small-scale psychological research project. 
+Of course, your specific situation might involve additional, more specialized steps. 
 After completing the tutorial, you should be able to customize your workflow accordingly.
@@
 
-This information makes explicitly declares the project's dependencies on files and R packages.
 `repro` uses this information to construct a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
@@
 This section is only for advanced users who want to overcome some limitations of `repro`.
-If you read this paper the first time you probably want to skip this section.
+If you read this paper the first time you probably want to skip this section and continue reading at the section "Preregistration as Code".
 As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.

### 2021-08-05 manuscript.Rmd
until right before discussion

@@
 Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
-Preregistration seeks to reduce "researcher degrees of freedom" [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could influence the results.
+Preregistration seeks to reduce "researcher degrees of freedom" [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could bias the reported results.
 Preregistrations, however, have several shortcomings.
@@
 A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
-For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables, or more parameters than observed data.
+For example, a verbal description of a statistical model may be unidentified, e.g.
+if it includes reciprocal paths between variables, or more parameters than observed data.
+<!-- AB note though that we still cannot exclude the possibility of empirical underidentification with this approach.
+-->
 Conversely, a model may be mis-specified in a major way; for example by omitting direct effects when the research question is about mediation, thus leading to a model with unacceptable fit.
@@
 
-The workflow described in this paper facilitates a rigorous and innovative solution to this problem:
+The workflow described in this paper facilitates a rigorous solution to this problem: <!--AB is it really innovative?
+I guess since people have preregistered code before, it's not entirely new...
+even though the reproducibility aspect is innovative-->
 Instead of describing the analysis in prose, researchers include the code required to conduct the analysis in the preregistration.
@@
 PAC has the potential to eliminate researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
-Moreover, it removes the necessity to write two separate documents: A preregistration, and a manuscript.
+Moreover, it removes the necessity to write two separate documents: A preregistration _and_ a manuscript.
 For a PAC, researchers can write a reproducible, dynamically generated draft of their intended manuscript at the preregistration stage.
-This includes the Introduction, Methods, and Results sections.
+This already includes most typical sections, such as Introduction, Methods, and Results.
 These results are initially based on simulated data with the same structure as the data they expect to obtain from their experiments.
@@
 We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, in PAC, is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
+First, the PAC is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
 Relatedly, writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
@@
 Fourth, despite its rigor, PAC does accommodate data-dependent decisions if these can be formulated as code.
-Researchers can, for example, formulate conditions (`if() {} else {}`) under which they prefer one analysis type over the other.
+Researchers can, for example, formulate conditions (e.g., in the form of if-else-blocks) under which they prefer one analysis type over the other.
 For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
@@
 Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
-One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
+One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the design of the study.
 Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
@@
 @bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.
+<!--AB I am not sure what you want to say here.
+Is this saying that researchers usually already have a good hunch and they should not be afraid?
+I think this needs more explanation -- or remove the sentence-->
 Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
@@
 When they have completed the method section, they should begin writing code.
-We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
+For basic research projects, we suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
+<!-- AB not sure what this means.
+Why do I have to write 3 functions for each planned hypothesis?
+Isn't it one for each hypothesis?
+Please check -->
 At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
@@
 
-We argued that PAC has several advantages over classical preregistration and have outlined how to implement it.
+We argued that PAC has several advantages over classic <!--AB please check different meanings of classic VS classical--> preregistration and have outlined how to implement it.
 To illustrate how PAC works in practice and help researchers implementing PAC themselves, we have conducted a PAC that serves as a guiding example.
@@
 
-Again, we propose a preregistration format that closely resembles a classical journal article, but uses simulated data and dynamic document generation to create a document that starts out as a preregistration and eventually becomes the final report.
+Again, we propose a preregistration format that closely resembles a classic journal article, but uses simulated data and dynamic document generation to create a document that starts out as a preregistration and eventually becomes the final report.
 The complete preregistration source is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
@@
 As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary background information and literature to understand the context and purpose of the research question.
+<!--AB do we explain anywhere that the [@XXX] are bibtex references?
+This may be not clear to everyone-->
 This example is drastically shortened for illustration purposes:
@@
 
+<!-- AB I am not happy with the next sentence. Isn't it the other way round? I thought we basically write the analysis first and then translate the analysis into a method section?-->
 We then translated this methods section into the following `planned_analysis()` function:
@@
 
+<!--AB this code will not be understood by most of the readers?
+Can we remove this?
+Instead keep the results text below but explain what the variables choosen_power and minn mean and that they were either set in the simulation or simulation results #necessary-->
 ```{r, echo=FALSE}

### 2021-08-05 manuscript.Rmd
discussion

@@
 
-This paper demonstrated how the R package `repro` supports researchers in creating reproducible repositories and reproducible manuscripts.
+This paper demonstrated how the R package `repro` supports researchers in creating reproducible research projects including reproducible manuscripts.
 These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
-The workflow we present here rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to all but guarantee reproducibility.
+The workflow we present here rests on four software solutions to guarantee reproducibility: version control, dynamic document generation, dependency tracking, and software management.
 We first demonstrated how to create a reproducible research project.
@@
 To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything is under version control.
-The likelihood of those deviations, while unavoidable, is decreased by using simulated data to verify the suitability of the analysis plan.
-This analysis plan may contain data depended decisions to further decrease the likelihood that researcher are unable stick to their preregistration
+The likelihood of those deviations, while never unavoidable, is decreased by using simulated data to verify the suitability of the proposed analysis plan.
+This analysis plan may contain data-dependent decisions to increase flexibility of research plans.
 
@@
 Despite our efforts to ensure this goal, the package is still in development and undoubtedly has limitations.
-The workflow we present here introduces several software solutions which in itself presents a danger to reproducibility.
-To benefit from automatic and convenient reproduction researcher must install Git, Make, and Docker.
+The workflow we present here introduces several software solutions, which in itself present a threat to fully automated reproducibility.
+To benefit from automatic and convenient reproduction, researcher must have Git, Make, and Docker.
 However, Git and Make are itself included in the Docker image created by `repro`.
 Researchers can therefore employ the Docker image manually to download the Git repository and execute Make for full reproduction.
-In other words the only hard requirement for reproduction and therefore Achilles' heel for this workflow is Docker.
+In other words the only hard requirement for reproduction and therefore the Achilles' heel for this workflow is Docker.
 The Docker approach has two vulnerabilities.
@@
 Microsoft R Application Network, Ubuntu Package Archive).
-The Docker image can be saved to disk and treated like any other file that requires careful archivation.
-However, if the existence of the Docker image is guaranteed, researchers still require software to run the image.
+The Docker image can be saved to disk and treated like any other file that requires careful archival.
+However, even if the existence of the Docker image is guaranteed, future researchers still require software to run the image.
 To that end they can either rely on Docker itself or Docker-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
@@
 The Makefile, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are, for whatever reason, unable to run Make.
-Hypothetically, even if Docker were unavaible, the Dockerfile still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
+Hypothetically, even if Docker were unavailabe one day, the Dockerfile still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
 
@@
 We realize that PACs are challenging.
-Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove unwieldy for a PAC.
-Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not preregistered research but as a complement.
-The rigidity of PAC is only than to much of a good thing, if one holds the view that researchers can for no reason deviate from their preregistration.
-However, we are convinced that many reasons justify changing the preregistered code, but are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
+Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove impractical for a PAC.
+Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not-preregistered research but as a complementary approach.
+We are convinced that many reasons justify changing the preregistered code, but are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
 PAC enables just that.
@@
 Another challenge is that PAC requires more programming expertise than a usual analysis.
-In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
+In contrast to a reproducible script, which should work on the same data, a PAC must work on different data of similar structure.
 We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
@@
 The reproducible workflow we present here is heavily automated and hence promotes reusability.
+<!--AB is it correct to claim the following?
+If yes, there is probably research on this?: Furthermore, adhering to principles of reusability typically removes errors in the code and thus increases the likelihood that the statistical analysis is correct-->
 
@@
 
+<!--AB this paragraph does not really fit here. Either have a subsection heading or, better, move it back up again-->
 The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
@@
 
-`repro` was designed modularly to meet such demands.
+`repro` was designed modularly to meet many such demands.
 At the moment, `repro` only supports the combination of RMarkdown, Git, Make, and Docker.
@@
 `repro`'s modular structure allows such alternative workflow, though they have not been implemented.
-Depending on the demand by users, we will implement some of these alternative workflows in `repro`.
+Depending on the demand by users, we will implement some of these alternative workflows in `repro` and hope for a broad adoption of computational reproducibility in the near future.
 

### 2021-08-05 manuscript.Rmd
Update manuscript.Rmd
just removed `workflow`

--- 2021-08-05 manuscript.Rmd
@@
 ---
-title: "Reproducible Research in R: A tutorial on how to do the same thing more than once"
+title: "Reproducible Research: A Tutorial on how to do the same thing more than once in R"
 author:
@@
   Reproducibility has long been considered integral to the scientific method.
-  Something is called reproducible when an independent person obtains the same results from the same data.
+  Something is called reproducible when independend researcher arrive at the same results from the same data.
   Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
-  Technological advancements now enable scientists to achieve a more comprehensive standard;
-  one in which any individual can be granted access to a digital research repository, and reproduce the analyses from the raw data to the final report including all relevant statistical analyses with a single command.
+  Technological advancements now enable scientists to achieve a much higher standard;
+  one in which any individual can be granted access to a digital research repository, and reproduce the analyses from raw data to final manuscript with a single command.
   This method has far-reaching implications for scientific archiving,
@@
   are complicated to use.
-  This paper introduces `repro`, an R-package, which guides researchers in the installation and use of the tools required for making a research project reproducible.
-  Finally, we suggest the use of the proposed tools for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
-  Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions of those analyses, which merely complement the PAC as a more readable summary.
-  PAC circumvents the shortcomings of ambiguous preregistrations that may give researchers undesired degrees of freedom.
-  <Hence, reproducibility made convenient with automation has a wide range of applications to accelerate scientific progress.
+  This paper introduces `repro`; an R-package that guides researchers in the installation and use of the tools required for computational reproducibility.
+  It includes a tutorial on how to make a research project reproducible with the aid of `repro`.
+  Finally, it introduces novel applications of reproducibility tools for scientific research, notably it includes the preregistration of study plans as reproducible computer code.
+  Since computer code describes study plans more precisely than prose, this circumvents the shortcomings of ambiguous preregistrations that give researchers undesired degrees of freedom.
+  Reproducibility made convenient with automation has hence a wide range of applications to accelerate scientific progress.
 abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
 keywords: |
-  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
-  to the article, yet reasonably common within the subject discipline.).
+  open science; compuational reproducibility; preregistration; R; R Markdown; Make; GitHub; Docker
 acknowledgement: |
@@
 
-Scientists increasingly strive to make research data, materials, and analysis code openly available <!-- AB "i.e." does not work in this context.
-I just removed this because it's irrelevant whether this happens within the movement or not?!-->
+Scientists increasingly strive to make research data, materials, and analysis code openly available i.e. within the Open Science movement.
 Sharing these types of digital research products has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
@@
 Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
-The credibility of empirical research results hinges on their objectivity.
+The credibility of empirical research results hingeson their objectivity.
 Objectivity in this context means "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
@@
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
-This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (e.g., missing software or ambiguous documentation of all steps taken <!-- AB the previous list was difficult to understand without more context; these two points are hopefully better to understand--> ), and there is a dire need for reproducibility standards.
+This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (multiple versions of code and data, copy-and-paste errors, missing software, ambiguous computation), and there is a dire need for reproducibility standards.
 With the technological advancement, it is now possible to make all digital products related to a research project available so that a research project can be reproduced with minimal effort.
@@
 We pursue two aims.
+
 First, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
-Second, we introduce a computational preregistration scheme based on reproducible research projects.
-<!-- AB Absätze können nicht nur einen Satz haben! Außerdem fehlte das "Second"-->
-
-<!--AB "This" geht nicht am Absatzanfang--> The proposed workflow builds on several technical solutions originating from software engineering. 
-We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they interact to form a coherent workflow.
+
+This workflow builds on several technical solutions originating in the software engineering community. 
+We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they form a coherent workflow.
 We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
@@
 One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
-PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including typical sections like Introduction, Methods, and Results.
+PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including the Introduction, Methods, and Results sections.
 The resulting dynamic document closely resembles the final manuscript, but temporarily uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
-With these placeholders, authors preregister the exact presentation of their results without having seen real data yet.
-Once the empirical data are analyzed, the Results are then updated automatically, and the Discussion can be written to finalize the report.
+Once the empirical data are analyzed, the Results are updated automatically, and the Discussion can be written.
 
@@
 
-Therefore, this tutorial demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
+This tutorial therefore demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
 A rigorous standard increases the likelihood that a project will remain reproducible as long as possible.
@@
 As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
-<!--The first part... dieser Satz wirkte verloren. Habe ich gestrichen-->
+The first part of this tutorial outlines threats to reproducibility and how the here presented reproducibility standard addresses these threats.
 
@@
 
-1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time <!-- the bullet points looked really strange in the formatted document-->
+1.	**Multiple inconsistent versions of code, data, or both**; for example
+  - the data set could have changed over time because outliers were removed at a later stage or an item was later recoded
+  - the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
 It may then be unclear which version of code and data was used to produce some reported set of results.
 2.	**Copy-and-paste errors**; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another)
+3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
 4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
@@
 
-<!--why boldface? looks odd and it is easy to confuse with section headings-->
-1. Version control
-2. Dynamic document generation
-3. Dependency tracking
-4. Software management
-
-The remainder of this section briefly explains why each of these four building blocks is needed and details their role in ensuring reproducibility.
-
-**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime (and in particular, in the development phase) of a research project.
+1. **Version control**
+2. **Dynamic document generation**
+3. **Dependency tracking**
+4. **Software management**
+
+The remainder of this section briefly explains why each of these four building blocks is needed and their role in ensuring reproducibility.
+
+**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the development phase of a research project.
 Version control allows a clear link between which results were generated by which version of code and data.
@@
 Ideally, each commit has a message that succinctly describes these changes.
-It is good practice to make commits for concrete milestones (e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3").
+It is good practice to make commits for concrete milestones, e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3".
 This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
-Each commit refers back to its ancestor, and they form a timeline together.
+Each commit "knows" its ancestor, and they form a timeline together.
 The entirety of commits (i.e. the version-controlled project) is called a repository.
@@
 Dynamic document generation integrates both steps.
-Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allowing interested readers to gain a deeper understanding of the contents [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
-**R Markdown** uses Markdown for formatting text and R (or other programming languages) for writing the statistical analysis.
+Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allows interested readers deeper understanding [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
+**R Markdown** uses Markdown for writing the text and R (or other programming languages) for the analysis.
 Markdown is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
@@
 However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
-In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or dependencies on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
+In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
 
@@
 Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
-This ensures that the simulated data are present before the manuscript is built, and that the simulation is rerun and the manuscript is rebuilt if the simulation code was changed.
+This ensures that the simulation is always run before the manuscript is built, and that the manuscript is rebuilt if the simulation code was changed.
 Make therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
@@
 These so-called "Docker images" function like a virtual computer  (i.e., a sort of "sand box").
-A Docker image should contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the computer's operating system.
-This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.
+A Docker image should contain *all* software dependencies used in an analysis - not just R-packages, but also R and Rstudio, and even the computer's operating system.
+This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators, linear algebra libraries and such within R.
 
 Note that the software environment of the Docker image is completely separate from software installed on your computer.
-This separation is excellent for reproducibility, but takes some getting used to.
+This separation is excellent for reproducibility, but take some getting used to.
 For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the Docker image.
@@
 To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
-While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
-To remove this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
+While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable hurdle in the acceptance and implementation of the workflow.
+To remove this hurdle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
 Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
@@
 
-<!--AB according to our own license in the original paper, we need to state the license under which the image was released. Please complement this information.-->
 ```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency tracking (Make), software management (Docker), and dynamic document generation (R Markdown).
@@
 
-One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be quite time-intensive.
-To lower the threshold to adoption, the R-package `repro` introduces helper functions that wraps and hides the complexity of the necessary more complicated (and powerful) tools.
-The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user is supposed to do next.
+One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be time-intensive.
+To lower the threshold to adoption, the R-package `repro` introduces helper functions that wrap more complicated (and powerful) tools.
+These so-called wrappers guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user should do in layman's terms.
 We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
@@
 This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
-It describes basic steps that we expect to be representative of a small-scale psychological research project. 
-Of course, your specific situation might involve additional, more specialized steps. 
+It describes basic steps that we expect to apply to all users.
+Your specific situation might involve additional, more specialized steps.
 After completing the tutorial, you should be able to customize your workflow accordingly.
 The first step is to install the required software.
-We assume that you have installed R [@R-base, version `r with(R.version, paste0(major, ".", minor))`] and RStudio [@rstudio, version 1.4] already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert)[@R-repro].
+We assume that you have installed R and RStudio already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert)[@R-repro].
 In case you either have not installed R and RStudio or are unsure if they are up to date you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers how to install all software necessary for this tutorial in three steps.
-The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing Docker.
 Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
@@
 
+This information makes explicitly declares the project's dependencies on files and R packages.
 `repro` uses this information to construct a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
@@
 This section is only for advanced users who want to overcome some limitations of `repro`.
-If you read this paper the first time you probably want to skip this section and continue reading at the section "Preregistration as Code".
+If you read this paper the first time you probably want to skip this section.
 As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.

### 2021-08-05 preregistration.Rmd
run automate to include prereg in Makefile

@@
   preregistration as code, power simulation, machiavellianism, gender
-output:
-  bookdown::markdown_document2:
-    base_format: rticles::mdpi_article
+output: rticles::mdpi_article
 header-includes:

### 2021-08-05 manuscript.Rmd

@@
   are complicated to use.
-  This paper introduces `repro`; an R-package that guides researchers in the installation and use of the tools required for computational reproducibility.
-  It includes a tutorial on how to make a research project reproducible with the aid of `repro`.
-  Finally, it introduces novel applications of reproducibility tools for scientific research, notably it includes the preregistration of study plans as reproducible computer code.
-  Since computer code describes study plans more precisely than prose, this circumvents the shortcomings of ambiguous preregistrations that give researchers undesired degrees of freedom.
-  Reproducibility made convenient with automation has hence a wide range of applications to accelerate scientific progress.
+  This paper introduces `repro`, an R-package, which guides researchers in the installation and use of the tools required for making a research project reproducible.
+  Finally, we suggest the use of the proposed tools for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
+  Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions of those analyses, which merely complement the PAC as a more readable summary.
+  PAC circumvents the shortcomings of ambiguous preregistrations that may give researchers undesired degrees of freedom.
+  Hence, reproducibility made convenient with automation has a wide range of applications to accelerate scientific progress.
 abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
@@
 
-Scientists increasingly strive to make research data, materials, and analysis code openly available i.e. within the Open Science movement.
+Scientists increasingly strive to make research data, materials, and analysis code openly available.
 Sharing these types of digital research products has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
@@
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
-This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (multiple versions of code and data, copy-and-paste errors, missing software, ambiguous computation), and there is a dire need for reproducibility standards.
+This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (e.g., missing software or ambiguous documentation of all steps taken), and there is a dire need for reproducibility standards.
 With the technological advancement, it is now possible to make all digital products related to a research project available so that a research project can be reproduced with minimal effort.
@@
 We pursue two aims.
-
 First, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
-
-This workflow builds on several technical solutions originating in the software engineering community. 
-We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they form a coherent workflow.
+Second, we introduce a computational preregistration scheme based on reproducible research projects.
+
+The proposed workflow builds on several technical solutions originating from software engineering. 
+We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they interact to form a coherent workflow.
 We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
@@
 As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
-The first part of this tutorial outlines threats to reproducibility and how the here presented reproducibility standard addresses these threats.
 
@@
 
-1.	**Multiple inconsistent versions of code, data, or both**; for example
-  - the data set could have changed over time because outliers were removed at a later stage or an item was later recoded
-  - the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
+1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
 It may then be unclear which version of code and data was used to produce some reported set of results.
 2.	**Copy-and-paste errors**; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
+3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another)
 4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
@@
 
-1. **Version control**
-2. **Dynamic document generation**
-3. **Dependency tracking**
-4. **Software management**
-
-The remainder of this section briefly explains why each of these four building blocks is needed and their role in ensuring reproducibility.
-
-**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the development phase of a research project.
+1. Version control
+2. Dynamic document generation
+3. Dependency tracking
+4. Software management
+
+The remainder of this section briefly explains why each of these four building blocks is needed and details their role in ensuring reproducibility.
+
+**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime (and in particular, in the development phase) of a research project.
 Version control allows a clear link between which results were generated by which version of code and data.
@@
 
-This information makes explicitly declares the project's dependencies on files and R packages.
+This information makes explicit what dependencies (in form of files and R packages) a project relies on .
 `repro` uses this information to construct a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
@@
 Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.
-
-The workflow described in this paper facilitates a rigorous and innovative solution to this problem:
+Of course this approach is not failsafe and researchers could still encounter problems after gathering data (ie.
+empierical underidentification) in which case they still have to deviate from their preregistration.
+
+The workflow described in this paper facilitates a rigorous solution to this problem:
 Instead of describing the analysis in prose, researchers include the code required to conduct the analysis in the preregistration.
@@
 PAC has the potential to eliminate researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
-Moreover, it removes the necessity to write two separate documents: A preregistration, and a manuscript.
+Moreover, it removes the necessity to write two separate documents:
+A preregistration *and* a manuscript.
 For a PAC, researchers can write a reproducible, dynamically generated draft of their intended manuscript at the preregistration stage.
-This includes the Introduction, Methods, and Results sections.
+This already includes most typical sections, such as Introduction, Methods, and Results.
 These results are initially based on simulated data with the same structure as the data they expect to obtain from their experiments.
@@
 In the simplest case, they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
-@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not necessarily increase the planned sample size.
+@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not increase the planned sample size on avaerage but could help to invest resources more efficiently.
 Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
@@
 We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
+We capture these distinct tasks in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
+The simulate and report function can of course be reused, if appropriate, across multiple hypotheses.
 At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
@@
 
-We argued that PAC has several advantages over classical preregistration and have outlined how to implement it.
+We argued that PAC has several advantages over classic preregistration and have outlined how to implement it.
 To illustrate how PAC works in practice and help researchers implementing PAC themselves, we have conducted a PAC that serves as a guiding example.
@@
 
-We then translated this methods section into the following `planned_analysis()` function:
+The methods section is the translation of the following `planned_analysis()` function:
 
@@
 
-This paper demonstrated how the R package `repro` supports researchers in creating reproducible repositories and reproducible manuscripts.
+This paper demonstrated how the R package `repro` supports researchers in creating reproducible research projects including reproducible manuscripts.
 These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
@@
 The reproducible workflow we present here is heavily automated and hence promotes reusability.
+Furthermore, adhering to principles of reusability typically removes errors in the code and thus increases the likelihood that the statistical analysis is correct
 

### 2021-08-05 manuscript.Rmd

@@
 One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
-However, while version control and dynamic document generation are becoming more common, this approach fails to guarantee reproducibility in many instances [@Peikert2019].
-In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or dependencies on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
+While version control and dynamic document generation are becoming more common, we have argued that two more components are required and that each component alone is unlikely to guarantee reproducibility [@Peikert2019; @vanlissa2020worcs].
+In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
 

### 2021-08-05 manuscript.Rmd
add ref for simulating data closes #199
shorten title in example

@@
 ---
-title: "Reproducible Research in R: A tutorial on how to do the same thing more than once"
+title: "A tutorial on how to do the same thing more than once"
 author:
@@
 ---
-title: "Reproducible Research: A Tutorial on how to do the same thing more than once in R"
+title: "A Tutorial on how to do the same thing more than once"
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
@@
 ---
-title: "Reproducible Research: A Tutorial on how to do the same thing more than once in R"
+title: "Reproducible Research: A Tutorial on how to do the same thing more than once"
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
@@
 These results are initially based on simulated data with the same structure as the data they expect to obtain from their experiments.
-For guidance on how to simulate data, see [@ref].
+For guidance on how to simulate data, see @paxtonMonteCarloExperiments2001 and @skrondalDesignAnalysisMonte2000.
 Once the preregistration is submitted and real data have been collected or made available,

### 2021-08-05 manuscript.Rmd
use bookdown again closes #205 reintroduces #10
R Studio -> RStudio closes #206

@@
 output:
-  rticles::mdpi_article
+  bookdown::markdown_document2:
+    base_format: rticles::mdpi_article
 header-includes:
@@
 `r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
-The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like tidyverse [@tidyverse] and all its dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
+The image for this project includes Ubuntu, R, RStudio, LaTeX as well as a variety of R packages like tidyverse [@tidyverse] and all its dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
 A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
@@
 To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
-While R Markdown and Git are well integrated into the R environment through R Studio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
+While R Markdown and Git are well integrated into the R environment through RStudio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
 To remove this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
@@
 The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing Docker.
-Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
+Start RStudio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
 It will assist you while you follow the tutorial.
@@
 Another advantage is that, because R Markdown files are in plain text, they are more suitable for version control using Git.
-Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
+Some users might find it easier to activate the "Visual Editor" of RStudio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
 The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)

### 2021-08-06 manuscript.Rmd
state licence for image closes #209
m-dash

@@
 Docker provides a container in which the final report is built using dynamic document generation in R Markdown.
-Reproduced from @Peikert2019.
+Reproduced from @Peikert2019 licensed under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0).
 "}
@@
 These messages are adjusted to your specific operating system and installed dependencies.
-Before you continue, we ask you to run the above commands to check Git, Make, and Docker - both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
+Before you continue, we ask you to run the above commands to check Git, Make, and Docker---both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
 If necessary, follow any instructions presented until all checks pass.
@@
 
-One advantage of this type of markup for formatting is that it can be rendered to many different output formats - both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g.
+One advantage of this type of markup for formatting is that it can be rendered to many different output formats---both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g.
 specific journal requirements.

### 2021-08-06 manuscript.Rmd
move this discussion to the footnote closes #192

@@
 
-[^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes.
+[^caution-mww]: The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
+The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
 The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
@@
 
-The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
-The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-
 # Outlook

### 2021-08-06 manuscript.Rmd
typo

@@
 ---
-title: "A Tutorial on how to do the same thing more than once"
+title: "A tutorial on how to do the same thing more than once"
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
@@
 ---
-title: "Reproducible Research: A Tutorial on how to do the same thing more than once"
+title: "A tutorial on how to do the same thing more than once"
 author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier

### 2021-08-06 manuscript.Rmd
link hash closes #207
update repro closes #211

@@
     - gert
-    - aaronpeikert/repro@c3cae1f
+    - aaronpeikert/repro@27e7d22
     - here
@@
 The `repro` package comes with the function `repro::current_hash()` for this purpose.
-This document was created from the commit with the hash ``r repro::current_hash()``.
+This document was created from the commit with the hash [``r repro::current_hash()``](`r stringr::str_c("https://github.com/aaronpeikert/repro-tutorial/tree/", repro::current_hash(40))`).
 
@@
 The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R).
-
-Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
+The next snippet shows how we included the results dynamically into the preregistration (the origin of the R-variables `minn`, `choosen_power`, and `choosen_d` is not shown).
 
@@
 ```{r, echo=FALSE, results='asis'}
-show_prereg("power-code", wrap = FALSE)
-```
-
-```{r, echo=FALSE, results='asis'}
 show_prereg("power")
 ```
+
+Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
 

### 2021-08-06 preregistration.Rmd
update repro closes #211

@@
 <!--end-ttest-->
-<!--begin-power-code-->
 ```{r, echo=FALSE}
@@
 ```
-<!--end-power-code-->
 <!--begin-power-->

### 2021-08-06 manuscript.Rmd
closes #188

@@
 2.	**Copy-and-paste errors**; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another)
+3.	**Undocumented or ambiguous order of computation**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another)
 4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.

### 2021-08-06 manuscript.Rmd
change formatting of software and concepts closes #197 closes #204

@@
 
-1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
+1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
 It may then be unclear which version of code and data was used to produce some reported set of results.
-2.	**Copy-and-paste errors**; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	**Undocumented or ambiguous order of computation**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another)
-4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
+2.	*Copy-and-paste errors*; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
+3.	*Undocumented or ambiguous order of computation*; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another)
+4.	*Ambiguous dependencies*; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
 
@@
 
-**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime (and in particular, in the development phase) of a research project.
+*Version control* prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime (and in particular, in the development phase) of a research project.
 Version control allows a clear link between which results were generated by which version of code and data.
 This is important for reproducibility because only if it is obvious which version of data and code produced which results they can be said to be reproducible.
-We recommend using **Git** for version control, because of its widespread adoption in the R community.
-
-Git tracks changes to all project-related files (e.g., materials, data, and code) over time.
+We recommend using *Git* for version control, because of its widespread adoption in the R community.
+
+`Git` tracks changes to all project-related files (e.g., materials, data, and code) over time.
 At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
 A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
-Git is built around snapshots that represent the project state at a given point in time.
+`Git` is built around snapshots that represent the project state at a given point in time.
 Those snapshots are called "commits" and work like a "save" action.
@@
 The entirety of commits (i.e. the version-controlled project) is called a repository.
-In Git, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
-Git has additional features beyond basic version control, such as branches to facilitate collaboration.
-@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git for research.
-@bryanExcuseMeYou2018 provides additional information on how to track R Markdown documents.
-Collaborating via Git requires the repository to be uploaded somewhere.
-We recommend GitHub to host Git repositories because of its popularity among R users.
+In `Git`, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
+`Git` has additional features beyond basic version control, such as branches to facilitate collaboration.
+@vuorreCuratingResearchAssets2018 provide a more extensive treatment how `Git` functions and how to use `Git` for research.
+@bryanExcuseMeYou2018 provides additional information on how to track `R Markdown` documents.
+Collaborating via `Git` requires the repository to be uploaded somewhere.
+We recommend GitHub to host `Git` repositories because of its popularity among R users.
 GitHub has many tools that ease project management and collaboration, and these tools are valuable in scientists everyday work, but they are not central to achieving reproducibility.
 
-Second, we rely on **dynamic document generation**.
+Second, we rely on *dynamic document generation*.
 The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
@@
 Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allowing interested readers to gain a deeper understanding of the contents [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
-**R Markdown** uses Markdown for formatting text and R (or other programming languages) for writing the statistical analysis.
-Markdown is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
+*R Markdown* uses `Markdown` for formatting text and R (or other programming languages) for writing the statistical analysis.
+`Markdown` is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
 This way, Markdown does not need any specialized software for editing.
-It is highly fool-proof [unlike, for example, LaTeX @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using Markdown for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
-R Markdown extends regular Markdown by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a Markdown document.
+It is highly fool-proof [unlike, for example, `LaTeX` @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+We suggest using `Markdown` for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
+`R Markdown` extends regular `Markdown` by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a `Markdown` document.
 Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
 This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
-@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of R Markdown.
+@xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of `R Markdown`.
 
@@
 
-To automatically resolve dependencies between project files, we rely on **dependency tracking**.
+To automatically resolve dependencies between project files, we rely on *dependency tracking*.
 In essence, researchers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
@@
 
-We recommend using **Make** for dependency tracking because it is language independent.
-The following hypothetical example illustrates the utility of Make.
-Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in R Markdown (`manuscript.Rmd`).
-A Makefile for this project could look like this:
+We recommend using *Make* for dependency tracking because it is language independent.
+The following hypothetical example illustrates the utility of `Make`.
+Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in `R Markdown` (`manuscript.Rmd`).
+A `Makefile` for this project could look like this:
 
@@
 If a requirement is newer than the target, the recipe will be executed to rebuild the target.
-If a requirement does not exist, Make uses a recipe to build the requirement first before building the target.
-Here, if one were to build the final `manuscript.pdf` by rendering the R Markdown with the command shown in l.2, Make would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
+If a requirement does not exist, `Make` uses a recipe to build the requirement first before building the target.
+Here, if one were to build the final `manuscript.pdf` by rendering the `R Markdown` with the command shown in l.2, `Make` would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
 This ensures that the simulated data are present before the manuscript is built, and that the simulation is rerun and the manuscript is rebuilt if the simulation code was changed.
-Make therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
-Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] does not contain this element of dependency tracking explicitly but its strict structure of only containing one definite R Markdown still makes dependencies between files unambiguous.
+`Make` therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
+Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] does not contain this element of dependency tracking explicitly but its strict structure of only containing one definite `R Markdown` still makes dependencies between files unambiguous.
 
@@
 Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
-One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via **software management**.
+One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via *software management*.
 One comprehensive approach to software management is containerization.
@@
 
-**Docker** is a popular tool for containerization.
+*`Docker`* is a popular tool for containerization.
 It manages software dependencies by constructing a virtual software environment independent of the host software environment.
-These so-called "Docker images" function like a virtual computer  (i.e., a sort of "sand box").
-A Docker image should contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the computer's operating system.
+These so-called "`Docker` images" function like a virtual computer  (i.e., a sort of "sand box").
+A `Docker` image should contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the computer's operating system.
 This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.
@@
 This separation is excellent for reproducibility, but takes some getting used to.
-For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the Docker image.
-Each dependency that you want to use within the Docker image must be explicitly added as a dependency.
-Furthermore, using Docker mayrequire you to install software on an operating system that may not be familiar to you.
+For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the `Docker` image.
+Each dependency that you want to use within the `Docker` image must be explicitly added as a dependency.
+Furthermore, using `Docker` may require you to install software on an operating system that may not be familiar to you.
 The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.
@@
 A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
-Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
+`Dockerfiles` are tiny (the [`Dockerfile`](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
 However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
-For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the Dockerfile.
+For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the `Dockerfile`.
 A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
@@
 
-To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using R Markdown), version control (using Git), internal dependency management (using Make), and containerization (using Docker).
-While R Markdown and Git are well integrated into the R environment through RStudio, Make and Docker require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
+To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using `R Markdown`), version control (using `Git`), internal dependency management (using `Make`), and containerization (using `Docker`).
+While `R Markdown` and `Git` are well integrated into the R environment through RStudio, `Make` and `Docker` require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
 To remove this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
@@
 
-```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency tracking (Make), software management (Docker), and dynamic document generation (R Markdown).
-Git tracks changes to the project over time.
-Make manages dependencies among the files.
-Docker provides a container in which the final report is built using dynamic document generation in R Markdown.
+```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (`Git`), dependency tracking (`Make`), software management (`Docker`), and dynamic document generation (`R Markdown`).
+`Git` tracks changes to the project over time.
+`Make` manages dependencies among the files.
+`Docker` provides a container in which the final report is built using dynamic document generation in `R Markdown`.
 Reproduced from @Peikert2019 licensed under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0).
@@
 In case you either have not installed R and RStudio or are unsure if they are up to date you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers how to install all software necessary for this tutorial in three steps.
-The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing Docker.
+The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing `Docker`.
 Start RStudio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
@@
 These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
-Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:
+Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have `Git` installed:
 
@@
 These messages are adjusted to your specific operating system and installed dependencies.
-Before you continue, we ask you to run the above commands to check Git, Make, and Docker---both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
+Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
 If necessary, follow any instructions presented until all checks pass.
@@
 
-Now that your system is set up for this tutorial, we will introduce you to version control with Git.
-Git tracks files within one project folder.
+Now that your system is set up for this tutorial, we will introduce you to version control with `Git`.
+`Git` tracks files within one project folder.
 We start by creating such a project folder with RStudio by clicking the menu item:
@@
 
-Git tracks any files that you manually add to the repository.
+`Git` tracks any files that you manually add to the repository.
 To make sure you do not accidentally add files that you do not wish to share (e.g., privacy sensitive data),
@@
 ```
-Now the file `private.md` will not be committed to Git and will not be made public.
+Now the file `private.md` will not be committed to `Git` and will not be made public.
 
 Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
-When you are ready, you can begin tracking your files using Git by running:
+When you are ready, you can begin tracking your files using `Git` by running:
 
@@
 
-For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
-One way to do this is through the visual user interface in the RStudio Git pane.
+For `Git` to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
+One way to do this is through the visual user interface in the RStudio `Git` pane.
 Click on the empty box next to the file you want to stage.
@@
 To back up and/or share the files online, you can push your local repository to a remote repository.
-While you can choose any Git service (like GitLab or BitBucket), we will use GitHub in this tutorial.
-Before you upload your project to Git, you need to decide if you would like the project publically accessable (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
+While you can choose any `Git` service (like GitLab or BitBucket), we will use GitHub in this tutorial.
+Before you upload your project to `Git`, you need to decide if you would like the project publically accessable (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
 To upload the project publically to GitHub use:
@@
 
-R Markdown is a type of dynamic document well-suited to the RStudio user interface.
-The text of an R Markdown is formatted by Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
+`R Markdown` is a type of dynamic document well-suited to the RStudio user interface.
+The text of an `R Markdown` is formatted by `Markdown` (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
 The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
-The following example serves to illustrate the Markdown syntax. 
-It shows how to create a heading, a word in bold font, and a list of several items in Markdown:
+The following example serves to illustrate the `Markdown` syntax. 
+It shows how to create a heading, a word in bold font, and a list of several items in `Markdown`:
 
@@
 specific journal requirements.
-Another advantage is that, because R Markdown files are in plain text, they are more suitable for version control using Git.
-Some users might find it easier to activate the "Visual Editor" of RStudio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
+Another advantage is that, because `R Markdown` files are in plain text, they are more suitable for version control using `Git`.
+Some users might find it easier to activate the "Visual Editor" of RStudio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the `R Markdown` document) which features more graphical elements like a traditional word processor but still creates an `R Markdown` underneath with all of its flexibility.
 The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
@@
 
-Now that you are familiar with the basics of Markdown formatting, we turn our attention to including code into the text.
+Now that you are familiar with the basics of `Markdown` formatting, we turn our attention to including code into the text.
 Code is separated by three backticks (and the programming language in curly braces) like here:
@@
 The results of code enclosed in such backticks will be dynamically inserted into the document (depending on specific settings).
-This means that, whenever you render the R Markdown to its intended output format, the code will be executed and the results updated.
+This means that, whenever you render the `R Markdown` to its intended output format, the code will be executed and the results updated.
 The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
 
-Once the R Markdown file has been rendered to a static document, the resulting file is decoupled from the R Markdown and the code that created it.
+Once the `R Markdown` file has been rendered to a static document, the resulting file is decoupled from the `R Markdown` and the code that created it.
 This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
-To avoid ambiguity, we therefore recommend referencing the identifier of the Git commit at the time of rendering in the static document.
+To avoid ambiguity, we therefore recommend referencing the identifier of the `Git` commit at the time of rendering in the static document.
 Simply put, a static document should link to the version of the code that was used to create it.
@@
 
-Now that you know how to write text and R code in an R Markdown, you need to know about metadata (also called: YAML front matter).
+Now that you know how to write text and R code in an `R Markdown`, you need to know about metadata (also called: YAML front matter).
 These metadate contain information about the document, like the title and the output format.
@@
 
-1. Version Control
-2. Dynamic Document Creation
-3. Dependency Tracking
-4. Software Management
+1. *Version Control*
+2. *Dynamic Document Creation*
+3. *Dependency Tracking*
+4. *Software Management*
 
@@
 Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
-However, R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
+However, `R Markdown` has been extended to hundreds of document formats which are often highly configurable by their metadata.
 Some R packages add additional metadata fields.
@@
 This information makes explicit what dependencies (in form of files and R packages) a project relies on .
-`repro` uses this information to construct a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
+`repro` uses this information to construct a `Makefile` for the dependencies on other files and a `Dockerfile` that includes all required packages.
 These two files together form the basis for consistency within a research project and consistency across different systems.
-The function `repro::automate()` converts the metadata from all R Markdown files in the project (all files with the ending `.Rmd`) to a Makefile and a Dockerfile.
+The function `repro::automate()` converts the metadata from all `R Markdown` files in the project (all files with the ending `.Rmd`) to a `Makefile` and a `Dockerfile`.
 Together, these files allow users (including your future self) to reproduce every step in the analysis automatically.
-It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, to keep the Makefile and Dockerfile up to date.
+It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new `R Markdown` file to the project, to keep the `Makefile` and `Dockerfile` up to date.
 There is no harm in running it too often.
-Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git.
+Other than the `Makefile` and the `Dockerfile`, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to `Git`.
 
@@
 
-When you or someone else who received your project want to reproduce it, they first have to install the required software, which is Make, and Docker.
+When you or someone else who received your project want to reproduce it, they first have to install the required software, which is `Make`, and `Docker`.
 Remember, you can use the `check_*`-functions to test if these are installed:
@@
 
-When they are setup, they can ask `repro` to explain how they should use Make and Docker to reproduce the project (or you could explain it to them):
+When they are setup, they can ask `repro` to explain how they should use `Make` and `Docker` to reproduce the project (or you could explain it to them):
 
@@
 
-The only software required for reproducing a project is Docker, assuming users know how to build a Docker image and run Make within the container.
-However, if they have installed Make additionally to Docker, they do not need to know how to use Docker and can simply rely on the two Make commands `make docker` and `make -B DOCKER=TRUE`.
+The only software required for reproducing a project is `Docker`, assuming users know how to build a `Docker` image and run `Make` within the container.
+However, if they have installed `Make` additionally to `Docker`, they do not need to know how to use `Docker` and can simply rely on the two `Make` commands `make docker` and `make -B DOCKER=TRUE`.
 The commands can be pasted into a terminal and will reproduce the project in many cases.
@@
 
-4. Let `repro` generate Docker- and Makfile:
+4. Let `repro` generate `Docker`- and `Makefile`:
 
@@
 As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.
-The simplification of Docker and Make are possible because of two restrictions.
-Users who ask themselves either "How can I install software dependencies outside of R in the Docker image?" or "How can I express complex dependencies between files, e.g.
+The simplification of `Docker` and `Make` are possible because of two restrictions.
+Users who ask themselves either "How can I install software dependencies outside of R in the `Docker` image?" or "How can I express complex dependencies between files, e.g.
 hundreds of data files are preprocessed and combined?") need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
@@
 
-The metadata the `repro::automate()` function relies on can only express R packages as dependencies for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile.
-The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
+The metadata the `repro::automate()` function relies on can only express R packages as dependencies for the `Dockerfile` and only trivial dependencies (in the form of "file must exist") for the `Makefile`.
+The first limitation means that users must rely on software that is either already provided by the base `Dockerimage` "rocker/verse" or the R packages they list in the metadata.
 Other software that users might need like other programming languages, not already installed LaTeX packages etc. have to be added manually.
@@
 The second limitation is about dependencies.
-Make can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
-If B is missing in this example, Make would know how to recreate it from C and D.
+`Make` can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
+If B is missing in this example, `Make` would know how to recreate it from C and D.
 These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
-Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D, therefore leaving out important information or express the dependencies directly within the Makefile.
+Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D, therefore leaving out important information or express the dependencies directly within the `Makefile`.
 
 The following section explains how to overcome these limitation while still relying on the automatons afforded by `repro`.
-Lifting these restrictions requires the user to interact more directly with Make or Docker.
-To satisfy more complicated requirements users need to understand how repro utilizes Make and Docker internally.
+Lifting these restrictions requires the user to interact more directly with `Make` or `Docker`.
+To satisfy more complicated requirements users need to understand how repro utilizes `Make` and `Docker` internally.
 
 Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two processing steps.
-First it recreates the virtual software environment (Docker) and then it executes computational recipes in the virtual software environment (Make).
+First it recreates the virtual software environment (`Docker`) and then it executes computational recipes in the virtual software environment (`Make`).
 The first step is done by the command `make docker`.
-The command `make docker` will trigger Make to build the target called `docker`.
-The recipe for this target builds an image from the Dockerfile of the repository.
+The command `make docker` will trigger `Make` to build the target called `docker`.
+The recipe for this target builds an image from the `Dockerfile` of the repository.
 The `&&` concatenates both commands and only runs the second command if the first was successful.
 Therefore, the computational steps are only executed when the software environment is set up.
-The second step executes the actual reproduction and is again a call to Make in the form of `make -B DOCKER=TRUE` with three noteworthy parts.
-First, a call to `make` without any explicit target will build the Make target `all`.
-Second, the flag `-B` means that Make will consider all dependencies as outdated and hence will rebuild everything.
-Third, repro constructs Make targets so that if you supply `DOCKER=TRUE` they are executed within the Docker image of the project.
-
-This interplay between Docker and Make is relatively complicated because it resembles a chicken or egg problem.
-We have computational steps (Make) which depend on the software environment (Docker) for which we again have computational steps that create it.
-Users only require a deeper understanding of this interdependence when they either want to have more complex computational recipes than rendering an R Markdown or require other software than R packages.
+The second step executes the actual reproduction and is again a call to `Make` in the form of `make -B DOCKER=TRUE` with three noteworthy parts.
+First, a call to `make` without any explicit target will build the `Make` target `all`.
+Second, the flag `-B` means that `Make` will consider all dependencies as outdated and hence will rebuild everything.
+Third, repro constructs `Make` targets so that if you supply `DOCKER=TRUE` they are executed within the `Docker` image of the project.
+
+This interplay between `Docker` and `Make` is relatively complicated because it resembles a chicken or egg problem.
+We have computational steps (`Make`) which depend on the software environment (`Docker`) for which we again have computational steps that create it.
+Users only require a deeper understanding of this interdependence when they either want to have more complex computational recipes than rendering an `R Markdown` or require other software than R packages.
 
 Users can have full control over the software installed within the image of the project.
-`repro` creates three Dockerfiles inside the `.repro` directory.
-Two Dockerfiles are automatically generated.
+`repro` creates three `Dockerfiles` inside the `.repro` directory.
+Two `Dockerfiles` are automatically generated.
 The first is `.repro/Dockerfile_base`.
@@
 This way, the call to the automate function virtually freezes the R environment to the state it was called the first time inside the container. 
-Below, you see the Docker base file we used to create this manuscript:
+Below, you see the `Docker` base file we used to create this manuscript:
 
@@
 
-The second automatically generated Dockerfile is `.repro/Dockerfile_packages`.
+The second automatically generated `Dockerfile` is `.repro/Dockerfile_packages`.
 Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.
 
-Finally, there is one manually edited Dockerfile: `.repro/Dockerfile_manual`.
+Finally, there is one manually edited `Dockerfile`: `.repro/Dockerfile_manual`.
 It is blank by default, and can be used to add further dependencies outside of R, like system libraries or external software.
 
-Using Docker may require you to install software on an operating system that may not be familiar to you.
+Using `Docker` may require you to install software on an operating system that may not be familiar to you.
 The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.
@@
 
-If the following snipped is added to `.repro/Dockerfile_manual` the Docker image will have Python installed.
+If the following snipped is added to `.repro/Dockerfile_manual` the `Docker` image will have Python installed.
 
@@
 
-Docker eventually requires a single Dockerfile to run, so `repro::automate()` simply concatenates the three Dockerfiles and saves the result into the main `Dockerfile` at the top level of the R project.
+`Docker` eventually requires a single `Dockerfile` to run, so `repro::automate()` simply concatenates the three `Dockerfiles` and saves the result into the main `Dockerfile` at the top level of the R project.
 
@@
 If you can formulate the creation of dependencies in terms of computational steps, e.g.
-the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the Makefile.
-The Makefile that `repro` creates is only a template and you are free to change it.
+the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the `Makefile`.
+The `Makefile` that `repro` creates is only a template and you are free to change it.
 However, make sure to never remove the following two lines:
@@
 
-The file `.repro/Makefile_Rmds` contains the automatically generated targets from `repro::automate()` for the R Markdown files.
+The file `.repro/Makefile_Rmds` contains the automatically generated targets from `repro::automate()` for the `R Markdown` files.
 This file should not be altered manually.
@@
 The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but usually should not modify.
-This Makefile coordinates the interplay between Make and Docker and contains targets for building (with `make docker`) and saving (with `make save-docker`) the Docker image.
+This `Makefile` coordinates the interplay between `Make` and `Docker` and contains targets for building (with `make docker`) and saving (with `make save-docker`) the `Docker` image.
 Additionally, it provides facilities to execute commands within the container.
 If you write a computational recipe for a target it will be by default evaluated using the locally installed software.
-To instead evaluate commands inside the Docker image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first Make example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
+To instead evaluate commands inside the `Docker` image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first `Make` example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
 2:
@@
 ```
-It behaves exactly as the in the first Make example, the script `R/simulate.R` is run using the locally installed R.
+It behaves exactly as the in the first `Make` example, the script `R/simulate.R` is run using the locally installed R.
 Because this translates simply to:
@@
 
-it is evaluated within the Docker container using the software within it and not the locally installed R version:
+it is evaluated within the `Docker` container using the software within it and not the locally installed R version:
 
@@
 
-To summarise, `repro` provides researchers with an easy to use solution for dependency tracking (in form of Make) and software management (using Docker) without the necessity to learn both languages.
+To summarise, `repro` provides researchers with an easy to use solution for dependency tracking (in form of `Make`) and software management (using `Docker`) without the necessity to learn both languages.
 Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automations.
@@
 
-To write a PAC, researchers should start by drafting a manuscript in the form of an R Markdown.
+To write a PAC, researchers should start by drafting a manuscript in the form of an `R Markdown`.
 When they have completed the method section, they should begin writing code.
@@
 The workflow we present here introduces several software solutions, which in itself present a threat to fully automated reproducibility.
-To benefit from automatic and convenient reproduction, researcher must have Git, Make, and Docker.
-However, Git and Make are itself included in the Docker image created by `repro`.
-Researchers can therefore employ the Docker image manually to download the Git repository and execute Make for full reproduction.
-In other words the only hard requirement for reproduction and therefore the Achilles' heel for this workflow is Docker.
-The Docker approach has two vulnerabilities.
-First, and more important, the Docker image for the project has to remain available.
-The Dockerfile (the plain text description to build a Docker image) on the other hand is insufficient because it relies on to many service providers (e.g.
+To benefit from automatic and convenient reproduction, researcher must have `Git`, `Make`, and `Docker`.
+However, `Git` and `Make` are itself included in the `Docker` image created by `repro`.
+Researchers can therefore employ the `Docker` image manually to download the `Git` repository and execute `Make` for full reproduction.
+In other words the only hard requirement for reproduction and therefore the Achilles' heel for this workflow is `Docker`.
+The `Docker` approach has two vulnerabilities.
+First, and more important, the `Docker` image for the project has to remain available.
+The `Dockerfile` (the plain text description to build a `Docker` image) on the other hand is insufficient because it relies on to many service providers (e.g.
 Microsoft R Application Network, Ubuntu Package Archive).
-The Docker image can be saved to disk and treated like any other file that requires careful archival.
-However, even if the existence of the Docker image is guaranteed, future researchers still require software to run the image.
-To that end they can either rely on Docker itself or Docker-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
-The only way to remove the reliance on such external software is to turn the Docker image into a full operating system that subsequently can be installed and run on almost any modern computer.
+The `Docker` image can be saved to disk and treated like any other file that requires careful archival.
+However, even if the existence of the `Docker` image is guaranteed, future researchers still require software to run the image.
+To that end they can either rely on `Docker` itself or `Docker`-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
+The only way to remove the reliance on such external software is to turn the `Docker` image into a full operating system that subsequently can be installed and run on almost any modern computer.
 This process is technically possible and would guarantee reproducibility for decades without any software dependency, assuming hardware that conforms to the x86 instruction set architecture continues to be available.
 However, this process requires much technical knowledge and is currently not facilitated by `repro`.
-In any case and without such precautions do the R Markdown, the Makefile and the Dockerfile provide information that allow researchers to manually trace the computational steps and recreate the computational environment.
-The Makefile, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are, for whatever reason, unable to run Make.
-Hypothetically, even if Docker were unavailabe one day, the Dockerfile still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
+In any case and without such precautions do the `R Markdown`, the `Makefile` and the `Dockerfile` provide information that allow researchers to manually trace the computational steps and recreate the computational environment.
+The `Makefile`, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are, for whatever reason, unable to run `Make`.
+Hypothetically, even if `Docker` were unavailable one day, the `Dockerfile` still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
 
@@
 This ideal of a fully automatic reproduction of research projects meets in practice a wide range of demands on how user-friendly and how powerful the software solutions are.
-Some may find Make too complicated or that Docker requires too much storage.
+Some may find `Make` too complicated or that `Docker` requires too much storage.
 Yet others may find that they require other programming languages or scale their computation across hundreds of computers, e.g. via high-performance computing clusters or cloud computing.
@@
 `repro` was designed modularly to meet many such demands.
-At the moment, `repro` only supports the combination of RMarkdown, Git, Make, and Docker.
+At the moment, `repro` only supports the combination of `R Markdown`, `Git`, `Make`, and `Docker`.
 However, there are alternatives for each of these elements that may fit better into an individual research project.
-RMarkdown could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer`[@officer] or `officedown`[@officedown] to accommodate a wider arrange of journal submission standards.
-Instead of using formal version control with Git, `repro` could automatically save snapshots for increasing user-friendliness.
-Make could be replaced by the more R-centered alternative `targets` for more convenience.
-Docker could be combined with `renv` [@R-renv] to control the package versions precisely (our approach fixes the date, `renv` the exact package version).
-Alternatively, Docker could be replaced by the more lightweight `renv` if no dependencies outside of R are considered crucial.
-Docker does not satisfy the requirements of many HPC environments, but Singularity was designed to avoid this limitation while still being compatible with Docker images.
+`R Markdown` could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer`[@officer] or `officedown`[@officedown] to accommodate a wider arrange of journal submission standards.
+Instead of using formal version control with `Git`, `repro` could automatically save snapshots for increasing user-friendliness.
+`Make` could be replaced by the more R-centered alternative `targets` for more convenience.
+`Docker` could be combined with `renv` [@R-renv] to control the package versions precisely (our approach fixes the date, `renv` the exact package version).
+Alternatively, `Docker` could be replaced by the more lightweight `renv` if no dependencies outside of R are considered crucial.
+`Docker` does not satisfy the requirements of many HPC environments, but Singularity was designed to avoid this limitation while still being compatible with `Docker` images.
 

### 2021-08-06 manuscript.Rmd
fix titles

@@
 ---
-title: "A tutorial on how to do the same thing more than once"
+title: "Reproducible Research in R: A tutorial on how to do the same thing more than once"
 author:
@@
 
-## Dynamic document generation
+## Using dynamic document generation
 
@@
 ````
+
+## Manage software and file dependencies
 
@@
 
-Another challenge is that PAC requires more programming expertise than a usual analysis.
-In contrast to a reproducible script, which should work on the same data, a PAC must work on different data of similar structure.
-We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
-This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
-The reproducible workflow we present here is heavily automated and hence promotes reusability.
-Furthermore, adhering to principles of reusability typically removes errors in the code and thus increases the likelihood that the statistical analysis is correct
-
 Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].

### 2021-08-06 manuscript.Rmd
restructure discussion (headings made no sense) closes #190 closes #191

@@
 
-We implemented a mechanism that only uses simulated data when the actual data is not yet available  (in this example if the file `data/data.csv` does not exist)for the results section.
+We implemented a mechanism that only uses simulated data when the actual data is not yet available (in this example if the file `data/data.csv` does not exist) for the results section.
 This mechanism also warns readers if these results are based on simulated data.
@@
 
+We realize that PACs are challenging.
+Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove impractical for a PAC.
+Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not-preregistered research but as a complementary approach.
+We are convinced that many reasons justify changing the preregistered code, but are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
+PAC enables just that.
+Researcher can change their plans and possibly defend their decision, e.g. in the discussion, while readers of the manuscript can clearly see what was changed through version control.
+
 # Discussion
+
+Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
+Automated research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
+
+Central to these innovations is a property we call "reusability" which we hope this workflow promotes.
+Reusable code can run on different inputs from a similar context and produce valid outputs.
+This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
+The reproducible workflow we present here is heavily automated and hence promotes reusability.
+Furthermore, adhering to principles of reusability typically removes errors in the code and thus increases the likelihood that the statistical analysis is correct.
+Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.
 
@@
 
-## Limitations of repro
+## Limitations
 
@@
 
-# Limitations
-
-We realize that PACs are challenging.
-Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove impractical for a PAC.
-Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not-preregistered research but as a complementary approach.
-We are convinced that many reasons justify changing the preregistered code, but are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
-PAC enables just that.
-Researcher can change their plans and possibly defend their decision, e.g. in the discussion, while readers of the manuscript can clearly see what was changed through version control.
-
-Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
-
-Therefore reproducibility facilitates traditional good scientific practices and provides the foundation for promising innovations.
-
-Please note that the statistical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes and is not necessarily optimal for addressing your research question.
-
-# Outlook
+## Outlook
 

### 2021-08-06 manuscript.Rmd
highlight Docker- and Makefile closes #183

@@
 We recommend using *Make* for dependency tracking because it is language independent.
-The following hypothetical example illustrates the utility of `Make`.
+The following hypothetical example illustrates the utility of `Make` and a suitable `Makefile`.
 Consider a research project that contains a script to simulate data (`simulate.R`) and a scientific report of the simulation results written in `R Markdown` (`manuscript.Rmd`).
@@
 These so-called "`Docker` images" function like a virtual computer  (i.e., a sort of "sand box").
-A `Docker` image should contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the computer's operating system.
+A `Docker` image does contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the computer's operating system.
 This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.
@@
 The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.
+
+There are two ways to build a `Docker` image.
+First users can manually install whatever software they like from within the virtual environment.
+Such manually build environment can still be ported to all computers which support `Docker`.
+We, however, prefer the second way of building images automatically from a textual description called `Dockerfile`.
+This enables users to quickly change the software environment e.g. update to another R or package version without being "stuck" with a unmaintainable because manually build `Docker` image.
 

### 2021-08-06 manuscript.Rmd
put automation at the center of the argument closes #184 closes #177

@@
 This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (e.g., missing software or ambiguous documentation of all steps taken), and there is a dire need for reproducibility standards.
-With the technological advancement, it is now possible to make all digital products related to a research project available so that a research project can be reproduced with minimal effort.
-
-This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
+With the technological advancement, it is now possible to make all digital products related to a research project available so that a research project can be automatically reproduced with minimal effort.
+
 We pursue two aims.
-First, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
-Second, we introduce a computational preregistration scheme based on reproducible research projects.
-
-The proposed workflow builds on several technical solutions originating from software engineering. 
+First, we want to introduce researchers to a notion of automated reproducibility that requires, apart from the setup of software, no manual steps.
+Second, we discuss opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated, but also how research itself is conducted.
+With regard to the first goal, we present a tutorial that is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
+In this tutorial, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
+The proposed workflow builds on several technical solutions originating from software engineering to automate reproduction. 
 We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they interact to form a coherent workflow.
-We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
-Second, we discuss opportunities that the reproducible workflow affords to improve not only how research outputs are created and recreated, but also how research itself is conducted.
-One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
+We further demonstrate how the R package `repro` supports researchers in setup the required software and in adopting this workflow.
+In pursuit of the second aim we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
 PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including typical sections like Introduction, Methods, and Results.

### 2021-08-06 manuscript.Rmd
differentiate software dependencies

@@
 3.	*Undocumented or ambiguous order of computation*; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another)
-4.	*Ambiguous dependencies*; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
+4.	*Ambiguous software dependencies*; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
 

### 2021-08-06 manuscript.Rmd
typo

@@
 We have developed a workflow that achieves long-term and cross-platform computational reproducibility of scientific data analyses.
-It leverages established tools and practices from software engineering and rests on four pillars that address the aforementioned causes of non-reproducibility: [@Peikert2019]:
+It leverages established tools and practices from software engineering and rests on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:
 

### 2021-08-06 manuscript.Rmd
highlight the concept of computational recipes closes #189

@@
 To automatically resolve dependencies between project files, we rely on *dependency tracking*.
-In essence, researchers provide a collection of computational recipes that describe how ingredients are processed to create intermediate products and, in the end, a final product.
+In essence, researchers provide a collection of *computational recipes*.
+A computational recipe describes how inputs are processed to create a specific output in a way that is automatically executable.
+The concept of computational recipes is central to our understanding of reproducibility because it enables a unified way to automatically reproduce a project.
 Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).

### 2021-08-06 manuscript.Rmd
explain why standards for scientific writing ensure comprehensive study plans closes #181

@@
 
-Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
+Standards for scientific writing are designed to result in understandable and comprehensive research articles and can serve the same purpose for writing a preregistration.
+Standards for scientific writing [e.g. @apa7] require that research articles are described in such detail that any researcher that is familiar with the field is able to repeat a given study.
+They provide, hence, an excellent vehicle for planning a study.
 One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
 Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
+If the PAC is submitted as a registered report [an article format in which the preregistration is reviewed and the article accepted "in principle", @nosekRegisteredReports2014; @chambersWhatNextRegistered2019] a journal article might be more familiar to the reviewers that ensure a comprehensive study plan.
 

### 2021-08-06 manuscript.Rmd
elaborate how assumptions are simulated closes #182

@@
 it is the best way to assess the effectiveness of the analysis without actually gathering the data.
-If a researcher experiences difficulty in specifying such expectations,
-that may be an indication that further literature research, or exploratory pilot studies, are required.
+A simulation of data requires the researchers to translate their expectations about the underlying process into concrete computational terms, e.g.
+assumptions about which distribution does the dependent variable follow, or how do the independent variables influence each other and the dependent variable.
+If a researcher experiences difficulty in specifying such expectations, that may be an indication that further literature research, or exploratory pilot studies, are required.
 

### 2021-08-06 manuscript.Rmd
show citation in markdown example closes #210
dont use svg version

@@
 # file gets downloaded in Makefile
-knitr::include_graphics(here::here("images/nutshell.svg"), auto_pdf = TRUE)
+knitr::include_graphics(here::here("images/nutshell.pdf"))
 ```
@@
 The following example serves to illustrate the `Markdown` syntax. 
-It shows how to create a heading, a word in bold font, and a list of several items in `Markdown`:
+It shows how to create a heading, a word in bold font, a citation and a list of several items in `Markdown`:
 
@@
 Important **word** in bold.
+A citation: @einstein1935 did important research on this topic.
 

### 2021-08-08 manuscript.Rmd
last changes after reading before prereg

@@
 
-1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
+1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time.
 It may then be unclear which version of code and data was used to produce some reported set of results.
-2.	*Copy-and-paste errors*; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
-3.	*Undocumented or ambiguous order of computation*; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another)
+2.	*Copy-and-paste errors*; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result.
+3.	*Undocumented or ambiguous order of computation*; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another).
 4.	*Ambiguous software dependencies*; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
@@
 The remainder of this section briefly explains why each of these four building blocks is needed and details their role in ensuring reproducibility.
+A more extensive treatment of these tools is given in @Peikert2019.
 
@@
 `Markdown` is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
-This way, Markdown does not need any specialized software for editing.
+This way, `Markdown` does not need any specialized software for editing.
 It is highly fool-proof [unlike, for example, `LaTeX` @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
@@
 These so-called "`Docker` images" function like a virtual computer  (i.e., a sort of "sand box").
-A `Docker` image does contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the computer's operating system.
+A `Docker` image does contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the operating system.
 This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.
@@
 Furthermore, using `Docker` may require you to install software on an operating system that may not be familiar to you.
-The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.
+The images supplied by @boettigerIntroductionRockerDocker2017, for example, are based on the Ubuntu operating system.
 
@@
 A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
-Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] relies for software management on the R package `renv`[@R-renv] which is more lightweight and easier to use on the one hand, but not as comprehensive on the other hand because it only snapshots the R packages instead of all software used.
-
-To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using `R Markdown`), version control (using `Git`), internal dependency management (using `Make`), and containerization (using `Docker`).
+Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] relies for software management on the R package `renv`[@R-renv] which is more lightweight and easier to use than `Docker` on the one hand, but not as comprehensive on the other hand because it only snapshots the R packages instead of all software used.
+
+To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using `R Markdown`), version control (using `Git`), dependency tracking (using `Make`), and Software management (using `Docker`).
 While `R Markdown` and `Git` are well integrated into the R environment through RStudio, `Make` and `Docker` require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
@@
 These projects will remain reproducible irrespective of whether `repro` remains accessible in future.
-Users do not have to have repro installed to reproduce a project;
+Users do not have to have `repro` installed to reproduce a project;
 in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
@@
 # file gets downloaded in Makefile
-knitr::include_graphics(here::here("images/nutshell.pdf"))
+knitr::include_graphics(here::here("images/nutshell.svg"), auto_pdf = TRUE)
 ```
@@
 The `repro` package comes with the function `repro::current_hash()` for this purpose.
-This document was created from the commit with the hash [``r repro::current_hash()``](`r stringr::str_c("https://github.com/aaronpeikert/repro-tutorial/tree/", repro::current_hash(40))`).
+This document was created from the commit with the hash [``r repro::current_hash()`` (view on GitHub)](`r stringr::str_c("https://github.com/aaronpeikert/repro-tutorial/tree/", repro::current_hash(40))`).
 
@@
 
-The following section explains how to overcome these limitation while still relying on the automatons afforded by `repro`.
+The following section explains how to overcome these limitation while still relying on the automations afforded by `repro`.
 Lifting these restrictions requires the user to interact more directly with `Make` or `Docker`.
@@
 
-If the following snipped is added to `.repro/Dockerfile_manual` the `Docker` image will have Python installed.
+If the following snipped is added to `.repro/Dockerfile_manual` the `Docker` image will have, for example, Python installed.
+Other software is installed identically only the software name is exchanged.
 
@@
 ```bash
-make simulated_data.csv
+make data/simulation_results.csv
 ```
@@
 cat("```bash",
-  stringi::stri_wrap(system2("make", " --dry-run simulated_data.csv", stdout = TRUE)),
+  stringi::stri_wrap(system2("make", " --dry-run data/simulation_results.csv", stdout = TRUE)),
   "```", sep = "\n")
@@
 ```bash
-make DOCKER=TRUE simulated_data.csv
+make DOCKER=TRUE data/simulation_results.csv
 ```
@@
 
-To summarise, `repro` provides researchers with an easy to use solution for dependency tracking (in form of `Make`) and software management (using `Docker`) without the necessity to learn both languages.
+To summarise, `repro` provides researchers with an easy to use solution for dependency tracking (in form of `Make`) and software management (using `Docker`) without the necessity to learn both tools.
 Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automations.
@@
 Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.
-Of course this approach is not failsafe and researchers could still encounter problems after gathering data (ie.
-empierical underidentification) in which case they still have to deviate from their preregistration.
 
@@
 To write a PAC, researchers should start by drafting a manuscript in the form of an `R Markdown`.
-When they have completed the method section, they should begin writing code.
+When they have have outlined the studies design, they should begin writing code.
 We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
@@
 The simulate and report function can of course be reused, if appropriate, across multiple hypotheses.
-At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
-This function should receive the data and derive relevant results from it.
+At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` from which they later derive the method section.
+This function should receive the data and compute all relevant results from it.
 
@@
 Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
-The following code simulates a dataset with one observed variable in two groups.
 

### 2021-08-08 README.Rmd
use cache when reproduce this manuscript
repro::automate()
change status to active

@@
 output: github_document
+repro:
+  images:
+    - images/nutshell.svg
 ---
@@
 <!-- badges: start -->
-[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
+[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
 <!-- badges: end -->
-
-WIP
 
@@
 
+## Abstract
+
+```{r abstract, child = 'abstract.Rmd'}
+```
+
+```{r, echo=FALSE}
+knitr::include_graphics("images/nutshell.svg")
+```
+
 ## Code of Conduct

### 2021-08-09 preregistration.Rmd
last changes after reading before prereg
Necessary deviations from the preregistration

@@
 
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study [@simmons2012].
+We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study [cf.
+@simmons2012].
 We use data available from [openpsychometrics.org](https://openpsychometrics.org/_rawdata/) from the online version of the MACH-IV[@christie1970] and included participants that have responded to at least one machiavellianism item and reported their gender as either "male" or "female".
@@
 
-```{r, echo=FALSE, results='asis'}
+```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
 real_data <- here::here("data", "data.csv")
@@
 } else {
-  mach <- readr::read_csv(real_data)
+  mach <- readr::read_delim(real_data, delim = "\t")
+  # only keep MACH items + gender
+  mach <- dplyr::select(mach, dplyr::matches("^Q\\d+A$"), gender)
+  # code gender according to codebook (3 would be other)
+  mach <-
+    dplyr::mutate(mach, gender = factor(
+      gender,
+      levels = 1:2,
+      labels = c("male", "female")
+    ))
 }
@@
 
-\textcolor{red}{This document was created before we have gathered gathered data. The discussion will only be added to the final report.}
+We created this document to illustrate Preregistration as Code.
+We therefore do not discuss the results.
 

### 2021-08-09 preregistration.Rmd
add discussion to preregistration

@@
 
-We created this document to illustrate Preregistration as Code.
-We therefore do not discuss the results.
+This document only serves to illustrate Preregistration as Code.
+We, therefore, do not discuss the results.
+After we have acquired the data, we realized that we had to change the code for reading the data (see commit [6556a93](https://github.com/aaronpeikert/repro-tutorial/commit/6556a9395fcdd600b5b0c5358f92a2c6635ae360)).
+We do not believe that these changes influence the results substantively.
 

### 2021-08-09 README.Rmd
add link to zenodo for preregistration closes #196

@@
 [![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
+
+Preregistration as Code: [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.5170740.svg)](https://doi.org/10.5281/zenodo.5170740)
+
+Results after Preregistration: [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.5171678.svg)](https://doi.org/10.5281/zenodo.5171678)
+
 <!-- badges: end -->

### 2021-08-09 manuscript.Rmd
add links to prereg and its changes closes #213

@@
 
-The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
+The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/releases/tag/compav0.0.2.0-results).
 

### 2021-08-19 README.Rmd
add altmetric

--- 2021-08-08 README.Rmd
@@
 <!-- badges: start -->
-[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
+<script type='text/javascript' src='https://d1bxh8uas1mnw7.cloudfront.net/assets/embed.js'></script>
+
+<table>
+<tbody>
+  <tr>
+    <td><div data-badge-popover="right" data-badge-type="1" data-doi="10.31234/osf.io/fwxs4" data-hide-no-mentions="true" class="altmetric-embed"></div></td>
+    <td>Preregistration as Code: <a href="https://doi.org/10.5281/zenodo.5170740" target="_blank" rel="noopener noreferrer">[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.5170740.svg)](https://doi.org/10.5281/zenodo.5170740)</a></td>
+  </tr>
+  <tr>
+    <td><a href="https://www.repostatus.org/#active" target="_blank" rel="noopener noreferrer">[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)</a></td>
+    <td>Results after Preregistration: <a href="https://doi.org/10.5281/zenodo.5171678">[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.5171678.svg)](https://doi.org/10.5281/zenodo.5171678)</a></td>
+  </tr>
+</tbody>
+</table>
+
 <!-- badges: end -->

### 2021-08-19 README.Rmd
link preprint
remove table

--- 2021-08-09 README.Rmd
@@
 <!-- badges: start -->
+
+<script type='text/javascript' src='https://d1bxh8uas1mnw7.cloudfront.net/assets/embed.js'></script>
+
+<div data-badge-popover="right" data-badge-type="1" data-doi="10.31234/osf.io/fwxs4" data-hide-no-mentions="true" class="altmetric-embed"></div>
+
 [![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
@@
 Results after Preregistration: [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.5171678.svg)](https://doi.org/10.5281/zenodo.5171678)
+
+[📃Preprint📃](https://www.doi.org/10.31234/osf.io/fwxs4)
 

### 2021-09-17 preregistration.Rmd
Necessary deviations from the preregistration - reversed items

@@
 } else {
-  mach <- readr::read_delim(real_data, delim = "\t")
+  mach <- readr::read_delim(real_data, delim = "\t", na = c("", "NA", "NULL"))
   # only keep MACH items + gender
@@
     ))
+  # some items are reversed, see https://core.ac.uk/download/pdf/38810542.pdf
+  reversed_nr <- c(1, 15, 2, 12, 4, 11, 14, 19)
+  reversed <- stringr::str_c("Q", reversed_nr, "A")
+  mach <- dplyr::mutate(mach, dplyr::across(one_of(reversed), ~ 6 - .x))
 }

### 2021-09-17 preregistration.Rmd
log changes to preregistration

@@
 We, therefore, do not discuss the results.
-After we have acquired the data, we realized that we had to change the code for reading the data (see commit [6556a93](https://github.com/aaronpeikert/repro-tutorial/commit/6556a9395fcdd600b5b0c5358f92a2c6635ae360)).
+After we have acquired the data, we realized that we had to change the code for reading the data, including recoding gender, missing values and reversed items (see commit [6556a93](https://github.com/aaronpeikert/repro-tutorial/commit/6556a9395fcdd600b5b0c5358f92a2c6635ae360) and commit [9f7ab21](https://github.com/aaronpeikert/repro-tutorial/commit/9f7ab212dfaf84a0398752a4b80cf14c71000d00)).
 We do not believe that these changes influence the results substantively.

### 2021-09-17 manuscript.Rmd
explain our position on deviating from the prereg closes #219 closes #237

--- 2021-08-08 manuscript.Rmd
@@
 However, templates cannot remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
-A second limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
+This ambiguity causes a second problem, that is comparing the preregistration and the resulting publication to decide if and how researcher deviated from the preregistration.
+This task is difficult because it impossible to decide without additional information if the analysis actually was carried out differently or just described differently.
+Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare to different texts which may be worded differently or describe the same thing in varying level of detail.
+A third limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
 To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
@@
 and reference it in the manuscript.
+The convenient comparison of preregistration and publication with PAC is particularly noteworthy.
+Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
+We, however, oppose this view.
+The problem is not deviating from the preregistration, it is not making these changes transparent.
+In fact, we expect that only a fraction of PACs will require no changes after actual data was collected.
+With PAC, researcher can reflect any deviations with ease and should summarize and justify the changes for their reader[^summarize-changes].
+The reader is also in a position to inspects the changes and can therefore make an informed judgment.
+
+[^summarize-changes]: In section [Preregistration as Code --- a Tutorial] we have conducted an actual PAC and summarize the changes we made to the preregistered code in the discussion.
 
 Note that preregistration, or PAC, does not preclude exploratory analyses.
-Such analyses should merely be explicitly declared as such in the manuscript.
+Again such analyses should merely be explicitly declared as such in the manuscript to allow appropriate judgment of the results.
 
@@
 
-The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](this link does not work) and can view the changes we made since preregistering it on this [GitHub page](this link does not work).
+Of course it is likely that the preregistered code has to be adapted to the actual data.
+Changing the preregistration must be discussed.
+This is our summary of what we changed:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("discuss-changes")
+```
+
+But readers can inspect and judge the changes for themselves on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
+
+The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
 

### 2021-09-17 manuscript.Rmd
add multicollinearity example

@@
 
+Take the following example: A preregistration intended to include "working memory" and "fluid intelligence" as predictors for reaction time in an experiment but after gathering data both constructs exhibit high collinearity.
+The authors decide that for their purpose these variables are interchangable and remove one.
+For them this change simply does not effect the core of their theory but was an auxiliary assumption and they might be right.
+Another researcher, however, who e.g. hoped to incorporate their results into a theory about intelligence, might disagree about the importance of this assumption.
+
 [^summarize-changes]: In section [Preregistration as Code --- a Tutorial] we have conducted an actual PAC and summarize the changes we made to the preregistered code in the discussion.

### 2021-09-20 manuscript.Rmd
improve wording

--- 2021-09-17 manuscript.Rmd
@@
 When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
-Fifth, deviations from the preregistration are clearly documented, because they are reflected in changes to the code which are managed and tracked with version control.
-One could even make a commit labeled "Necessary deviations from the preregistration after data collection",
+Fifth, deviations from the preregistration are clearly documented because they are reflected in changes to the code, which are managed and tracked with version control.
+One could even make a commit labelled "Necessary deviations from the preregistration after data collection",
 and reference it in the manuscript.
@@
 We, however, oppose this view.
-The problem is not deviating from the preregistration, it is not making these changes transparent.
-In fact, we expect that only a fraction of PACs will require no changes after actual data was collected.
-With PAC, researcher can reflect any deviations with ease and should summarize and justify the changes for their reader[^summarize-changes].
-The reader is also in a position to inspects the changes and can therefore make an informed judgment.
+The problem is not deviating from the preregistration; it is not making these changes transparent.
+In fact, we expect that most PACs will require changes after the researcher collects data.
+With PAC, the authors can reflect any deviations with ease and should summarize and justify the changes for their readers [^summarize-changes].
+The reader is also in a position to inspects the changes and can therefore make an informed judgment about weighing the evidence presented by the authors.
+
+Take the following example: A preregistration intended to include "working memory" and "fluid intelligence" as predictors for reaction time in an experiment.
+However, after fitting the model on the actual data both constructs exhibit high collinearity.
+The authors decide that for their purpose, these variables are interchangeable and remove one.
+For them, this change simply does not affect their theory's core but was an auxiliary assumption, and they might be right.
+Another researcher, however, who, e.g.
+hoped to incorporate their results into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
 
@@
 
-Note that preregistration, or PAC, does not preclude exploratory analyses.
-Again such analyses should merely be explicitly declared as such in the manuscript to allow appropriate judgment of the results.
+Similarly, preregistration, or PAC, does not preclude exploratory analyses but should demark confirmatory results.
+Again such exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results.
 

### 2021-09-20 manuscript.Rmd
supress "your are inside docker" closes #232

--- 2021-09-17 manuscript.Rmd
@@
 
+```{r, include=FALSE}
+opts <- options()
+options(repro.docker = TRUE,
+        repro.docker.running = FALSE)
+```
+
 ```{r}
@@
 repro::check_docker()
+```
+
+```{r, include=FALSE}
+options(opts)
 ```
@@
 However, templates cannot remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
-This ambiguity causes a second problem, that is comparing the preregistration and the resulting publication to decide if and how researcher deviated from the preregistration.
-This task is difficult because it impossible to decide without additional information if the analysis actually was carried out differently or just described differently.
-Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare to different texts which may be worded differently or describe the same thing in varying level of detail.
-A third limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
+A second limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
 To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
@@
 and reference it in the manuscript.
-The convenient comparison of preregistration and publication with PAC is particularly noteworthy.
-Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
-We, however, oppose this view.
-The problem is not deviating from the preregistration, it is not making these changes transparent.
-In fact, we expect that only a fraction of PACs will require no changes after actual data was collected.
-With PAC, researcher can reflect any deviations with ease and should summarize and justify the changes for their reader[^summarize-changes].
-The reader is also in a position to inspects the changes and can therefore make an informed judgment.
-
-[^summarize-changes]: In section [Preregistration as Code --- a Tutorial] we have conducted an actual PAC and summarize the changes we made to the preregistered code in the discussion.
 
 Note that preregistration, or PAC, does not preclude exploratory analyses.
-Again such analyses should merely be explicitly declared as such in the manuscript to allow appropriate judgment of the results.
+Such analyses should merely be explicitly declared as such in the manuscript.
 
@@
 
-Of course it is likely that the preregistered code has to be adapted to the actual data.
-Changing the preregistration must be discussed.
-This is our summary of what we changed:
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("discuss-changes")
-```
-
-But readers can inspect and judge the changes for themselves on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
-
-The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
+The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/releases/tag/compav0.0.2.0-results).
 

### 2021-09-20 manuscript.Rmd
closes #225
cite usethis appropriatly

@@
 In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
+One noteworthy property of computational recipes is that they require a textual representation and the ability of a program to make sense of this textual representation.
+A workflow that includes, e.g. renaming variable names by hand in a graphical spreadsheet application is therefore incompatible with automated reproducibility.
 
@@
 The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user is supposed to do next.
+The `repro` package extends the [`usethis`](https://usethis.r-lib.org) [@usethis] package.
+While the `usethis` package aims to simplifies the development of R packages, `repro` contains reproducibility specific utilities.
 We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].

### 2021-09-20 manuscript.Rmd
address reviewer Is minor issues

@@
 We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they interact to form a coherent workflow.
-We further demonstrate how the R package `repro` supports researchers in setup the required software and in adopting this workflow.
+We further demonstrate how the R package `repro` supports researchers in setting up the required software and in adopting this workflow.
 In pursuit of the second aim we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
@@
 `Git` is built around snapshots that represent the project state at a given point in time.
-Those snapshots are called "commits" and work like a "save" action.
-Changes to files are collected in a _commit_.
+Those snapshots are called *commits* and work like a "save" action.
 Ideally, each commit has a message that succinctly describes these changes.
@@
 One comprehensive approach to software management is containerization.
-"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p.
+The central idea is that by
+"packaging the key elements of the computational  environment  needed  to  run  the  desired  software the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p.
 174].
@@
 Furthermore, using `Docker` may require you to install software on an operating system that may not be familiar to you.
-The images supplied by @boettigerIntroductionRockerDocker2017, for example, are based on the Ubuntu operating system.
+The images supplied by @boettigerIntroductionRockerDocker2017, for example, are based on Linux.
 
@@
 One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
-For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
+For example, there are pre-made images that only include what is necessary for R, based on Ubuntu containers [@boettigerIntroductionRockerDocker2017].
 Users can then install whatever they need in addition to what is provided by these pre-compiled images.
@@
 The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
+The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
 
@@
 The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
-The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R).
+The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R).
 The next snippet shows how we included the results dynamically into the preregistration (the origin of the R-variables `minn`, `choosen_power`, and `choosen_d` is not shown).

### 2021-09-21 manuscript.Rmd
add an explicit discussion of principle limitations closes #223

@@
 
+We have to distinguish two kinds of limitations.
+First, limitations by design that are unlikely to be changed in the near future.
+This workflow does inherit these both from the software it does rely on and the fundamental design principles this software shares with the workflow and `repro`.
+Second, limitations that we hope development effort of the open source community and ,with regard to `repro`, us will ease.
+
+As a principle, this workflow is incompatible with steps that cannot be automated.
+This means many researcher will have to search for alternative software that allows for automation and it is at odds with some ingrained habits of researchers to mix and match manual and automated steps in data analysis.
+This overarching principle manifests itself as several technical but nevertheless important limitations.
+`Git` can track any files (files bigger than 1GB only with extensions) but can only compare text files (with endings like, `.txt`, `.csv`, `.R`,  or `.Rmd`), not binary files (with endings like `.docx`, `.exe` or `.zip`).
+With `R Markdown` tables and graphics are dynamically generated from code but cannot be crafted by hand.
+`Make` can orchestrate any software that is programmable but fails to automate software that will only obey a user clicking at the right buttons.
+`Docker` can ship any software that runs on linux and can be automatically installed but that precludes many commercial or closed source software.
+
+This move away from software that has served researchers well for decades is understandably difficult and presents us with a conundrum.
+On the one hand, we firmly believe that automated reproducibility makes research more productive and collaboration easier but on the other hand we expect that researcher invest considerable time in learning these tools and than must "force" their collaborators to do the same.
+Two arguments reconcile this apparent paradox.
+First, this change will not happen all at once.
+Automated reproducibility is an ideal that we believe has many advantages but it is not an all or nothing decision.
+Researchers can pick up one skill at a time and than help their fellow collaborators to do the same.
+Second, the upfront investment is one time (and efforts such as `repro` are underway to reduce it) while the complexity and demands of research projects are steadily increasing.
+
 `repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.

### 2021-09-21 manuscript.Rmd
grammerly

--- 2021-09-20 manuscript.Rmd
@@
 
+We have to distinguish two kinds of limitations.
+First, some limitations are by design and are unlikely to be changed soon.
+Our workflow does inherit these from the software it relies on and the fundamental design principles these share with the workflow and `repro`.
+Second, there are other limitations that we hope development effort of the open-source community and, concerning `repro`, us will ease.
+
+As a principle, this workflow is incompatible with steps that cannot be automated.
+This principle is at odds with some ingrained habits of researchers to mix and match manual and automated steps in data analysis.
+To allow for automation, many researchers will have to search for alternative software.  
+The automation friendly software we present here has several technical but critical limitations.
+`Git` can track any files (files bigger than 1GB only with extensions) but can only compare text files (with endings like, `.txt`, `.csv`, `.R`,  or `.Rmd`), not binary files (with endings like `.docx`, `.exe` or `.zip`).
+With `R Markdown`, tables and graphics are dynamically generated from code but cannot be crafted by hand.
+`Make` can orchestrate any programmable software but fails to automate software that will only obey a user clicking at the right buttons.
+`Docker` can ship software that runs on Linux and can be automatically installed, but that requirement precludes many commercial or closed source software.
+
+This move away from software that has served researchers well for decades is understandably difficult and presents us with a conundrum.
+On the one hand, we firmly believe that automated reproducibility makes research more productive and collaboration easier.
+But, on the other hand, we expect researchers to invest considerable time in learning new tools and which they then must "force" on their collaborators.
+Two arguments reconcile this apparent paradox.
+First, this change will not happen all at once.
+Automated reproducibility is an ideal that we believe has many advantages, but it is not an all or nothing decision.
+Researchers can pick up one skill at a time and then help their fellow collaborators do the same.
+Second, the upfront investment is one time (and efforts such as `repro` are underway to reduce it) while the complexity and demands of research projects are steadily increasing.
+
 `repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.

### 2021-09-22 manuscript.Rmd
readd zenodo.org closes #245

--- 2021-09-20 manuscript.Rmd
@@
 
+As a last step researcher should create an archived and unchangeable version of the project as the preregistration.
+zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides digital object identifiers (DOI) for these archives.
+While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
+Please note that you can only link public GitHub repositories to zenodo.
+You may log into zenodo.org through your GitHub account.
+
+> Log into zenodo.org → Account → GitHub[^zenodo-github]
+[^zenodo-github]: https://zenodo.org/account/settings/github/
+
+After you have linked a GitHub repository you trigger the archivation by creating a GitHub release.
+To create GitHub release, navigate to GitHub:
+
+```{r}
+browse_github()
+```
+
+And click on Releases → Draft a new release.
+Here you can add all relevant binary files but at least a rendered version of the manuscript and the docker image.
+
 We realize that PACs are challenging.
@@
 Despite our efforts to ensure this goal, the package is still in development and undoubtedly has limitations.
-The workflow we present here introduces several software solutions, which in itself present a threat to fully automated reproducibility.
+The workflow we present here introduces several software solutions, which in itself present a threat to long-term reproducibility.
 To benefit from automatic and convenient reproduction, researcher must have `Git`, `Make`, and `Docker`.
@@
 The `Docker` approach has two vulnerabilities.
-First, and more important, the `Docker` image for the project has to remain available.
-The `Dockerfile` (the plain text description to build a `Docker` image) on the other hand is insufficient because it relies on to many service providers (e.g.
+First, and more important, the `Docker` image for the project and the `Git` repository have to remain available.
+The `Dockerfile` (the plain text description to build a `Docker` image) as opposed to the image is insufficient because it relies on too many service providers (e.g.
 Microsoft R Application Network, Ubuntu Package Archive).
-The `Docker` image can be saved to disk and treated like any other file that requires careful archival.
-However, even if the existence of the `Docker` image is guaranteed, future researchers still require software to run the image.
+We recommend to archive the `Git` repository and the `Docker` image with zenodo.org, a non-profit long-term storage for scientific data.
+The necessary steps for archival on zenodo.org are described in at the end of section [Preregistration as Code --- a Tutorial].
+The second vulnerability is that even if the existence of the `Docker` image and `Git` repository is guaranteed, future researchers still require software to run the image.
 To that end they can either rely on `Docker` itself or `Docker`-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).

### 2021-09-22 manuscript.Rmd
clarify that system libraries don't change results
clarify the advantages of Dockerfiles

@@
 A `Docker` image does contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the operating system.
+All differences in computational results that could be caused by software are hence eliminated.
 This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.
@@
 We, however, prefer the second way of building images automatically from a textual description called `Dockerfile`.
-This enables users to quickly change the software environment e.g. update to another R or package version without being "stuck" with a unmaintainable because manually build `Docker` image.
+Because the `Dockerfile` clearly describes which software is installed how, users can quickly change the software environment e.g.
+update to another R or package version without being "stuck" with an unmaintainable because manually build `Docker` image.
 

### 2021-09-22 manuscript.Rmd
add missing `::`
closes #246
fix typo
fix some typos

@@
 ```{r}
-check_git()
+repro::check_git()
 ```
@@
 Now that you know how to write text and R code in an `R Markdown`, you need to know about metadata (also called: YAML front matter).
-These metadate contain information about the document, like the title and the output format.
+These metadata contain information about the document, like the title and the output format.
 Metadata are placed at the beginning of the document and are separated by three dashes from the document body.
@@
 ```{r}
-check_make()
-check_docker()
+repro::check_make()
+repro::check_docker()
 ```
@@
 ```{r, eval=FALSE}
-system(reproduce())
+system(repro::reproduce())
 ```
@@
 
-The following section explains how to overcome these limitation while still relying on the automations afforded by `repro`.
+The following section explains how to overcome these limitations while still relying on the automations afforded by `repro`.
 Lifting these restrictions requires the user to interact more directly with `Make` or `Docker`.
@@
 
-If the following snipped is added to `.repro/Dockerfile_manual` the `Docker` image will have, for example, Python installed.
+If the following snippet is added to `.repro/Dockerfile_manual` the `Docker` image will have, for example, Python installed.
 Other software is installed identically only the software name is exchanged.
@@
 A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
-In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
+In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did include some form of script as part of their preregistration.
 But although the notion of preregistering code has been around for a while [cf. @steegenMeasuringCrowdAgain2014],
@@
 ```{r}
-browse_github()
+usethis::browse_github()
 ```
@@
 Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
-Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
+Because the same dynamic document serves as preregistration and final manuscript, researchers, reviewers, and editors only have to revise one document.
 To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything is under version control.

### 2021-09-22 manuscript.Rmd
there were two last steps

@@
 If simulating data proves to be too challenging, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
-Of course researcher must not analyze the data before they write the preregistration but they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
+In order to do so researcher must walk a fine but utmost important line.
+Of course researcher must not analyze the data before they write the preregistration.
+Nevertheless, they may obfuscate the real data by shuffling all variables or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
+Preregistration after data collection is common when reanalyzing data from another research group [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
+We argue that it is still an eligible preregistration.
+Since the data is obfuscated, almost no results are revealed and the researchers can not employ their researcher degrees of freedom effectively to influence the results.
+Shuffling leaves only univariate information, so this procedure is only applicable for bi-or multivariate hypotheses.
+Researcher still can access the information about means or proportions (e.g.
+how many participants belonging to group "A" are in the dataset) but are unbiased for all relations between variables (e.g.
+members of group "A" have a greater mean in variable "Z").
 
@@
 
-As a last step researcher should create an archived and unchangeable version of the project as the preregistration.
+After the researcher is satisfied with their draft preregistration, they should create an archived and unchangeable version of the project that serves as the preregistration.
 zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides digital object identifiers (DOI) for these archives.

### 2021-09-23 manuscript.Rmd
explain preregistration after data
suggest organization for multiple tests

@@
 We capture these distinct tasks in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
+The functions shown here are empty (they do not contain executable code) and the interested reader may find function filled with some life in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.
+
 The simulate and report function can of course be reused, if appropriate, across multiple hypotheses.
@@
 
+This example preregistration covers a single study with a single hypothesis.
+To organize studies with multi hypotheses we suggest multiple (possibly numbered in accord with the hypotheses, e.g. 1.2, 2.3 etc.) `planned_analysis()` and `report_analysis()` functions.
+Steps that apply to multiple analyses should be extracted into their own function e.g. `discard_outliers()` or `extract_scores()`.
+Preregistrations that cover multiple distinct data sources may employ multiple `simulate_data()` functions.
+These are merely suggestions, researcher who are more comfortable with another programming style will surely find an appropriate system.
+
 The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/releases/tag/compav0.0.2.0-results).

### 2021-09-23 R/simulation_funs.R
comment simulate_data

@@ -1,12 +1,17 @@
 #----simulate_data----
 simulate_data <- function(n, df, d, i){
+  # groups have same size
   stopifnot(n %% 2L == 0L)
   gender <- rep(c(0L, 1L), each = n/2)
+  # use Chi^2 distributed random data for some skew
   rand <-  matrix(rchisq(n * i, df/i), ncol = i) # sum of n chisq has df of n*df
   d_scaled <- d * sqrt(2*df)/i # scale d to express sd units (var(chisq) = 2df)
+  # add difference only to group coded with `1`
   effect <- rand + d_scaled * gender
+  # name items
   colnames(effect) <- paste0("mach", seq_len(i))
   effect <- as.data.frame(effect)
+  # recode gender from numeric to factor
   effect$gender <- factor(gender, levels = c(1L, 0L), labels = c("male", "female"))
   return(effect)
 }

### 2021-09-23 R/simulation_funs.R
comment planned_analysis()

@@ -18,17 +18,23 @@ simulate_data <- function(n, df, d, i){
 
 #----planned_analysis----
 planned_analysis <- function(data, use_rank = "skew", skew_cutoff = 1){
+  # average over all variable supplied, except gender
   machiavellianism <- rowMeans(data["gender" != names(data)], na.rm = TRUE)
+  # discard rows that only contain NAs
   data <- data[!is.na(machiavellianism),]
   machiavellianism <- machiavellianism[!is.na(machiavellianism)]
+  # assure gender is factor
   gender <- as.factor(data$gender)
+  # note skewness and decide t.test vs wilcox based on it
   skew <- moments::skewness(machiavellianism)
   # skewness cutoff
   if(use_rank == "skew")use_rank <- abs(skew) > skew_cutoff
   if(use_rank){
+    # t.test + rank = wilcox test
     machiavellianism <- rank(machiavellianism)
   }
   test <- t.test(machiavellianism ~ gender)
+  # return a bunch of information
   list(test = test, skew = skew, use_rank = use_rank, n = length(gender))
 }
 

### 2021-09-23 manuscript.Rmd
move to dedicated section
note that these are not repro functions

@@
 We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
-We capture these distinct tasks in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
+For each study these tasks have to specifically coded.
+While capture the tasks here in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
 The functions shown here are empty (they do not contain executable code) and the interested reader may find function filled with some life in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.
 
-The simulate and report function can of course be reused, if appropriate, across multiple hypotheses.
 At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` from which they later derive the method section.

### 2021-09-23 manuscript.Rmd
cite anoter good resource for monte carlo closes #199
PAC as a source of truth

--- 2021-09-22 manuscript.Rmd
@@
 We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they interact to form a coherent workflow.
-We further demonstrate how the R package `repro` supports researchers in setting up the required software and in adopting this workflow.
+We further demonstrate how the R package `repro` supports researchers in setup the required software and in adopting this workflow.
 In pursuit of the second aim we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
@@
 `Git` is built around snapshots that represent the project state at a given point in time.
-Those snapshots are called *commits* and work like a "save" action.
+Those snapshots are called "commits" and work like a "save" action.
+Changes to files are collected in a _commit_.
 Ideally, each commit has a message that succinctly describes these changes.
@@
 In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
-One noteworthy property of computational recipes is that they require a textual representation and the ability of a program to make sense of this textual representation.
-A workflow that includes, e.g. renaming variable names by hand in a graphical spreadsheet application is therefore incompatible with automated reproducibility.
 
@@
 One comprehensive approach to software management is containerization.
-The central idea is that by
-"packaging the key elements of the computational  environment  needed  to  run  the  desired  software the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p.
+"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p.
 174].
@@
 A `Docker` image does contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the operating system.
-All differences in computational results that could be caused by software are hence eliminated.
 This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.
@@
 Furthermore, using `Docker` may require you to install software on an operating system that may not be familiar to you.
-The images supplied by @boettigerIntroductionRockerDocker2017, for example, are based on Linux.
+The images supplied by @boettigerIntroductionRockerDocker2017, for example, are based on the Ubuntu operating system.
 
@@
 We, however, prefer the second way of building images automatically from a textual description called `Dockerfile`.
-Because the `Dockerfile` clearly describes which software is installed how, users can quickly change the software environment e.g.
-update to another R or package version without being "stuck" with an unmaintainable because manually build `Docker` image.
+This enables users to quickly change the software environment e.g. update to another R or package version without being "stuck" with a unmaintainable because manually build `Docker` image.
 
@@
 One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
-For example, there are pre-made images that only include what is necessary for R, based on Ubuntu containers [@boettigerIntroductionRockerDocker2017].
+For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
 Users can then install whatever they need in addition to what is provided by these pre-compiled images.
@@
 The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user is supposed to do next.
-The `repro` package extends the [`usethis`](https://usethis.r-lib.org) [@usethis] package.
-While the `usethis` package aims to simplifies the development of R packages, `repro` contains reproducibility specific utilities.
 We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
@@
 
-```{r, include=FALSE}
-opts <- options()
-options(repro.docker = TRUE,
-        repro.docker.running = FALSE)
-```
-
 ```{r}
@@
 
-```{r, include=FALSE}
-options(opts)
-```
-
 These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
@@
 ```{r}
-repro::check_git()
+check_git()
 ```
@@
 Now that you know how to write text and R code in an `R Markdown`, you need to know about metadata (also called: YAML front matter).
-These metadata contain information about the document, like the title and the output format.
+These metadate contain information about the document, like the title and the output format.
 Metadata are placed at the beginning of the document and are separated by three dashes from the document body.
@@
 ```{r}
-repro::check_make()
-repro::check_docker()
+check_make()
+check_docker()
 ```
@@
 ```{r, eval=FALSE}
-system(repro::reproduce())
+system(reproduce())
 ```
@@
 
-The following section explains how to overcome these limitations while still relying on the automations afforded by `repro`.
+The following section explains how to overcome these limitation while still relying on the automations afforded by `repro`.
 Lifting these restrictions requires the user to interact more directly with `Make` or `Docker`.
@@
 
-If the following snippet is added to `.repro/Dockerfile_manual` the `Docker` image will have, for example, Python installed.
+If the following snipped is added to `.repro/Dockerfile_manual` the `Docker` image will have, for example, Python installed.
 Other software is installed identically only the software name is exchanged.
@@
 A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
-In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did include some form of script as part of their preregistration.
+In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did included some form of script as part of their preregistration.
 But although the notion of preregistering code has been around for a while [cf. @steegenMeasuringCrowdAgain2014],
@@
 When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
-Fifth, deviations from the preregistration are clearly documented, because they are reflected in changes to the code which are managed and tracked with version control.
-One could even make a commit labeled "Necessary deviations from the preregistration after data collection",
+Fifth, deviations from the preregistration are clearly documented because they are reflected in changes to the code, which are managed and tracked with version control.
+One could even make a commit labelled "Necessary deviations from the preregistration after data collection",
 and reference it in the manuscript.
-
-Note that preregistration, or PAC, does not preclude exploratory analyses.
-Such analyses should merely be explicitly declared as such in the manuscript.
+The convenient comparison of preregistration and publication with PAC is particularly noteworthy.
+Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
+We, however, oppose this view.
+The problem is not deviating from the preregistration; it is not making these changes transparent.
+In fact, we expect that most PACs will require changes after the researcher collects data.
+With PAC, the authors can reflect any deviations with ease and should summarize and justify the changes for their readers [^summarize-changes].
+The reader is also in a position to inspects the changes and can therefore make an informed judgment about weighing the evidence presented by the authors.
+
+Take the following example: A preregistration intended to include "working memory" and "fluid intelligence" as predictors for reaction time in an experiment.
+However, after fitting the model on the actual data both constructs exhibit high collinearity.
+The authors decide that for their purpose, these variables are interchangeable and remove one.
+For them, this change simply does not affect their theory's core but was an auxiliary assumption, and they might be right.
+Another researcher, however, who, e.g.
+hoped to incorporate their results into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
+
+[^summarize-changes]: In section [Preregistration as Code --- a Tutorial] we have conducted an actual PAC and summarize the changes we made to the preregistered code in the discussion.
+
+Similarly, preregistration, or PAC, does not preclude exploratory analyses but should demark confirmatory results.
+Again such exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results.
 
@@
 If simulating data proves to be too challenging, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
-In order to do so researcher must walk a fine but utmost important line.
-Of course researcher must not analyze the data before they write the preregistration.
-Nevertheless, they may obfuscate the real data by shuffling all variables or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
-Preregistration after data collection is common when reanalyzing data from another research group [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
-We argue that it is still an eligible preregistration.
-Since the data is obfuscated, almost no results are revealed and the researchers can not employ their researcher degrees of freedom effectively to influence the results.
-Shuffling leaves only univariate information, so this procedure is only applicable for bi-or multivariate hypotheses.
-Researcher still can access the information about means or proportions (e.g.
-how many participants belonging to group "A" are in the dataset) but are unbiased for all relations between variables (e.g.
-members of group "A" have a greater mean in variable "Z").
+Of course researcher must not analyze the data before they write the preregistration but they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
 
@@
 The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
+The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
 
@@
 The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
-The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R).
+The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R).
 The next snippet shows how we included the results dynamically into the preregistration (the origin of the R-variables `minn`, `choosen_power`, and `choosen_d` is not shown).
@@
 
-The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/releases/tag/compav0.0.2.0-results).
-
-After the researcher is satisfied with their draft preregistration, they should create an archived and unchangeable version of the project that serves as the preregistration.
-zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides digital object identifiers (DOI) for these archives.
-While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
-Please note that you can only link public GitHub repositories to zenodo.
-You may log into zenodo.org through your GitHub account.
-
-> Log into zenodo.org → Account → GitHub[^zenodo-github]
-[^zenodo-github]: https://zenodo.org/account/settings/github/
-
-After you have linked a GitHub repository you trigger the archivation by creating a GitHub release.
-To create GitHub release, navigate to GitHub:
-
-```{r}
-usethis::browse_github()
-```
-
-And click on Releases → Draft a new release.
-Here you can add all relevant binary files but at least a rendered version of the manuscript and the docker image.
+Of course it is likely that the preregistered code has to be adapted to the actual data.
+Changing the preregistration must be discussed.
+This is our summary of what we changed:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("discuss-changes")
+```
+
+But readers can inspect and judge the changes for themselves on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
+
+The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
 
@@
 Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
-Because the same dynamic document serves as preregistration and final manuscript, researchers, reviewers, and editors only have to revise one document.
-To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything is under version control.
+Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
+To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything there is a single "source of truth"---dynamic manuscript including code---under version control.
+
 The likelihood of those deviations, while never unavoidable, is decreased by using simulated data to verify the suitability of the proposed analysis plan.
@@
 Despite our efforts to ensure this goal, the package is still in development and undoubtedly has limitations.
-The workflow we present here introduces several software solutions, which in itself present a threat to long-term reproducibility.
+The workflow we present here introduces several software solutions, which in itself present a threat to fully automated reproducibility.
 To benefit from automatic and convenient reproduction, researcher must have `Git`, `Make`, and `Docker`.
@@
 The `Docker` approach has two vulnerabilities.
-First, and more important, the `Docker` image for the project and the `Git` repository have to remain available.
-The `Dockerfile` (the plain text description to build a `Docker` image) as opposed to the image is insufficient because it relies on too many service providers (e.g.
+First, and more important, the `Docker` image for the project has to remain available.
+The `Dockerfile` (the plain text description to build a `Docker` image) on the other hand is insufficient because it relies on to many service providers (e.g.
 Microsoft R Application Network, Ubuntu Package Archive).
-We recommend to archive the `Git` repository and the `Docker` image with zenodo.org, a non-profit long-term storage for scientific data.
-The necessary steps for archival on zenodo.org are described in at the end of section [Preregistration as Code --- a Tutorial].
-The second vulnerability is that even if the existence of the `Docker` image and `Git` repository is guaranteed, future researchers still require software to run the image.
+The `Docker` image can be saved to disk and treated like any other file that requires careful archival.
+However, even if the existence of the `Docker` image is guaranteed, future researchers still require software to run the image.
 To that end they can either rely on `Docker` itself or `Docker`-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).

### 2021-09-23 manuscript.Rmd
turn this into a stand alone section

@@
 Fifth, deviations from the preregistration are clearly documented because they are reflected in changes to the code, which are managed and tracked with version control.
-One could even make a commit labelled "Necessary deviations from the preregistration after data collection",
-and reference it in the manuscript.
+
+## Deviating from the preregistration and exploration
+
 The convenient comparison of preregistration and publication with PAC is particularly noteworthy.
+Because there is a single "source of truth", the dynamic document including code, under version control, changes can be inspected by reader and authors alike.
+To make the comparison easier, we highly recommend that the authors review all changes to summarize and justify them for their reader [^summarize-changes].
+The reader is also in a position to inspects the changes and can therefore make an informed judgment about accepting or rejecting the justifications.
+
 Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
@@
 In fact, we expect that most PACs will require changes after the researcher collects data.
-With PAC, the authors can reflect any deviations with ease and should summarize and justify the changes for their readers [^summarize-changes].
-The reader is also in a position to inspects the changes and can therefore make an informed judgment about weighing the evidence presented by the authors.
+And often deviations provide opportunity to learn from the unexpected.
+Another reason why transparency is important in this regard, is that an empirical test of an theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].
 
@@
 Another researcher, however, who, e.g.
-hoped to incorporate their results into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
+hopes to incorporate their results into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
+In other words, "one man's trash is another man's treasure".
 
@@
 
-Similarly, preregistration, or PAC, does not preclude exploratory analyses but should demark confirmatory results.
+Transparency of deviations is also preferable to avoiding them by vagueness in the first place.
+A vague preregistration must have reduced diagnosticity for demarcating exploration and conformation.
+A reader may falsely conclude that a result is explorative because the publication is more specific than the preregistration, therefore commiting an alpha error (believing there is a deviation when in fact there is none).
+At the same time, a researcher may overlook a deviation---the goal of intentional vagueness---therefore committing a beta error.
+At its most extreme, the preregistration "We will do something with some data." has lost all ability to differentiate between exploration and conformation, which was what we set out to do with preregistrations.
+
+Another common misunderstanding is that, preregistration, and therefore PAC, preclude exploratory analyses.
 Again such exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results.
+In fact, PAC allows to preregister exploratory statistical models.
+Such preregistration communicates clearly how much exploration was going on, an information that is crucial for assessing the uncertainty of such results.
 

### 2021-09-23 manuscript.Rmd
distinguish exploration

@@
 Because there is a single "source of truth", the dynamic document including code, under version control, changes can be inspected by reader and authors alike.
-To make the comparison easier, we highly recommend that the authors review all changes to summarize and justify them for their reader [^summarize-changes].
+To make the comparison easier, we highly recommend that the authors review all changes to summarize and justify them for their reader e.g. in the discussion [^summarize-changes].
 The reader is also in a position to inspects the changes and can therefore make an informed judgment about accepting or rejecting the justifications.
@@
 In fact, we expect that most PACs will require changes after the researcher collects data.
-And often deviations provide opportunity to learn from the unexpected.
-Another reason why transparency is important in this regard, is that an empirical test of an theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].
+Often deviations provide opportunity to learn from the unexpected.
+Transparency is important in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].
 
@@
 Another researcher, however, who, e.g.
-hopes to incorporate their results into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
+hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
 In other words, "one man's trash is another man's treasure".
@@
 Another common misunderstanding is that, preregistration, and therefore PAC, preclude exploratory analyses.
-Again such exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results.
-In fact, PAC allows to preregister exploratory statistical models.
-Such preregistration communicates clearly how much exploration was going on, an information that is crucial for assessing the uncertainty of such results.
+We must differentiate here between two kinds of exploration.
+First, the traditional meaning of ad hoc statistical procedures and post hoc explanation of the results.
+Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results.
+The second kind of exploration is that through procedural well defined exploration in the form of exploratory statistical models that are common in machine learning.
+Somewhat ironically, PAC enables preregistration of such procedures and offers the merit to communicate precisely how much exploration was done; an information that is crucial for assessing the uncertainty of such "preregisterable" but exploratory results.
 

### 2021-09-23 manuscript.Rmd
grammerly

@@
 The convenient comparison of preregistration and publication with PAC is particularly noteworthy.
-Because there is a single "source of truth", the dynamic document including code, under version control, changes can be inspected by reader and authors alike.
-To make the comparison easier, we highly recommend that the authors review all changes to summarize and justify them for their reader e.g. in the discussion [^summarize-changes].
+Because there is a single "source of truth", the dynamic document, including code, under version control, changes can be inspected by readers and authors alike.
+To make the comparison more accessible, we highly recommend that the authors review all changes to summarize and justify them for their reader, e.g. in the discussion [^summarize-changes].
 The reader is also in a position to inspects the changes and can therefore make an informed judgment about accepting or rejecting the justifications.
@@
 In fact, we expect that most PACs will require changes after the researcher collects data.
-Often deviations provide opportunity to learn from the unexpected.
-Transparency is important in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].
+Often deviations provide an opportunity to learn from the unexpected.
+Transparency is essential in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].
 
 Take the following example: A preregistration intended to include "working memory" and "fluid intelligence" as predictors for reaction time in an experiment.
-However, after fitting the model on the actual data both constructs exhibit high collinearity.
+However, after fitting the model on the actual data, both constructs exhibit high collinearity.
 The authors decide that for their purpose, these variables are interchangeable and remove one.
@@
 A vague preregistration must have reduced diagnosticity for demarcating exploration and conformation.
-A reader may falsely conclude that a result is explorative because the publication is more specific than the preregistration, therefore commiting an alpha error (believing there is a deviation when in fact there is none).
+A reader may falsely conclude that a result is explorative because the publication is more specific than the preregistration, therefore committing an alpha error (believing there is a deviation when, in fact, there is none).
 At the same time, a researcher may overlook a deviation---the goal of intentional vagueness---therefore committing a beta error.
@@
 
-Another common misunderstanding is that, preregistration, and therefore PAC, preclude exploratory analyses.
-We must differentiate here between two kinds of exploration.
-First, the traditional meaning of ad hoc statistical procedures and post hoc explanation of the results.
+Another common misunderstanding is that preregistration, and therefore PAC, preclude exploratory analyses.
+We must differentiate here between two kinds of exploration, and neither of them is inhibited.
+The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
 Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results.
-The second kind of exploration is that through procedural well defined exploration in the form of exploratory statistical models that are common in machine learning.
-Somewhat ironically, PAC enables preregistration of such procedures and offers the merit to communicate precisely how much exploration was done; an information that is crucial for assessing the uncertainty of such "preregisterable" but exploratory results.
+The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning.
+These models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
+We recommend PAC for exploratory statistical models without any irony.
+The merit of preregistration in such cases is to communicate precisely how much exploration was done, information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.
 

### 2021-09-23 manuscript.Rmd
more precise wording

@@
 The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning.
-These models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
-We recommend PAC for exploratory statistical models without any irony.
-The merit of preregistration in such cases is to communicate precisely how much exploration was done, information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.
+These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
+We recommend PAC for those exploratory statistical models without any irony.
+The merit of preregistration in such cases is to communicate precisely how much exploration was done, a piece of information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.
 

### 2021-09-23 abstract.Rmd

@@
 In this tutorial, we demonstrate how the R package `repro` supports researchers in creating fully computationally reproducible research projects with tools from the software engineering community.
-Building upon this notion of fully automated reproducibility we present several application including the preregistration of research plans with code (Preregistration as Code, PAC).
+Building upon this notion of fully automated reproducibility we present several applications including the preregistration of research plans with code (Preregistration as Code, PAC).
 PAC eschews all ambiguity of traditional preregistration and offers several other advantages.
+Making technical advancements that serve reproducibility more widely accessible for researchers holds the potential to innovate the research process to become more productive, credible, and reliable.

### 2021-09-23 manuscript.Rmd

@@
 
+```{r, include=FALSE}
+options(opts)
+```
+
 These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
@@
 These results are initially based on simulated data with the same structure as the data they expect to obtain from their experiments.
-For guidance on how to simulate data, see @paxtonMonteCarloExperiments2001 and @skrondalDesignAnalysisMonte2000.
+For guidance on how to simulate data, see @morrisUsingSimulationStudies2019, @paxtonMonteCarloExperiments2001 and @skrondalDesignAnalysisMonte2000.
 Once the preregistration is submitted and real data have been collected or made available,
@@
 
+After the researcher is satisfied with their draft preregistration, they should create an archived and unchangeable version of the project that serves as the preregistration.
+zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides digital object identifiers (DOI) for these archives.
+While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
+Please note that you can only link public GitHub repositories to zenodo.
+You may log into zenodo.org through your GitHub account.
+
+> Log into zenodo.org → Account → GitHub[^zenodo-github]
+[^zenodo-github]: https://zenodo.org/account/settings/github/
+
+After you have linked a GitHub repository you trigger the archivation by creating a GitHub release.
+To create GitHub release, navigate to GitHub:
+
+```{r}
+usethis::browse_github()
+```
+
+And click on Releases → Draft a new release.
+Here you can add all relevant binary files but at least a rendered version of the manuscript and the docker image.
+
 We realize that PACs are challenging.

### 2021-09-24 manuscript.Rmd
not all decision can be translated to code
Update manuscript.Rmd
- I removed the paragraph on vagueness because we need to shorten the whole part and make a more concise story here
- I don't like the term "source of truth"
- some language editing
- I don't like the phrase "one man's trash" in this context (even though it's a good image in many other contexts of psychological research methodology)

@@
 Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
+Decisions that do not lend themselves to formulation in code, e.g.
+visual inspection, must still be described verbally or be treated as noted in the section [Deviating from the preregistration and exploration].
 With PAC researchers can precisely calibrate how much they want to explore.
@@
 
-The convenient comparison of preregistration and publication with PAC is particularly noteworthy.
-Because there is a single "source of truth", the dynamic document, including code, under version control, changes can be inspected by readers and authors alike.
-To make the comparison more accessible, we highly recommend that the authors review all changes to summarize and justify them for their reader, e.g. in the discussion [^summarize-changes].
-The reader is also in a position to inspects the changes and can therefore make an informed judgment about accepting or rejecting the justifications.
+We would like to note that in the PAC paradigm, the direct comparison of the preregistration and the final publication is particularly easy to accomplish.
+Because there is a single unambiguous reference point, that is the dynamic document, including code, under version control, changes can be inspected by authors and readers alike.
+To make the comparison even more accessible, we highly recommend that the authors review all changes made to the preregistered plan to summarize and justify them for their reader, e.g.
+in the discussion of the final manuscript [^summarize-changes].
+The reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.
 
 Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
-We, however, oppose this view.
-The problem is not deviating from the preregistration; it is not making these changes transparent.
-In fact, we expect that most PACs will require changes after the researcher collects data.
+However, we share the common view that a deviation from a preregistration is not a problem; it is a problem if these changes are not made transparent.
+In fact, we expect that most PACs will require changes after data have been collected.
 Often deviations provide an opportunity to learn from the unexpected.
@@
 The authors decide that for their purpose, these variables are interchangeable and remove one.
-For them, this change simply does not affect their theory's core but was an auxiliary assumption, and they might be right.
+For them, this change simply does not affect the core of their theory but was an auxiliary assumption, and they might be right.
 Another researcher, however, who, e.g.
 hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
-In other words, "one man's trash is another man's treasure".
+With transparent reporting, it is left to the reader to make up their own judgement.
 
@@
 
-Transparency of deviations is also preferable to avoiding them by vagueness in the first place.
-A vague preregistration must have reduced diagnosticity for demarcating exploration and conformation.
-A reader may falsely conclude that a result is explorative because the publication is more specific than the preregistration, therefore committing an alpha error (believing there is a deviation when, in fact, there is none).
-At the same time, a researcher may overlook a deviation---the goal of intentional vagueness---therefore committing a beta error.
-At its most extreme, the preregistration "We will do something with some data." has lost all ability to differentiate between exploration and conformation, which was what we set out to do with preregistrations.
-
 Another common misunderstanding is that preregistration, and therefore PAC, preclude exploratory analyses.
-We must differentiate here between two kinds of exploration, and neither of them is inhibited.
+We must differentiate here between two kinds of exploration, and neither of them is limited by PAC.
 The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
-Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results.
+Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results. <!-- citation? The Nosek paper?-->
 The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning.
+<!--citation?
+reproducibility in machine learning research?
+maybe cite the Brandmaier (in press) chapter, where I suggest to use PAC for these kinds of models?
+-->
 These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
-We recommend PAC for those exploratory statistical models without any irony.
+We specifically recommend PAC for those exploratory statistical models.
 The merit of preregistration in such cases is to communicate precisely how much exploration was done, a piece of information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.

### 2021-09-27 manuscript.Rmd
recomend changelog

@@
 Because there is a single unambiguous reference point, that is the dynamic document, including code, under version control, changes can be inspected by authors and readers alike.
-To make the comparison even more accessible, we highly recommend that the authors review all changes made to the preregistered plan to summarize and justify them for their reader, e.g.
-in the discussion of the final manuscript [^summarize-changes].
+Authors should at least review all changes made to the preregistered plan to summarize and justify them for their reader, e.g. in the discussion of the final manuscript [^summarize-changes].
+Such retrospective summary already serves to put changes into context, however, a change log that is kept continually and where changes are explained when they are made, provides more immediate and detailed information.
+Each entry in the change log should contain the commit id to the respective changes.
 The reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.

### 2021-09-27 manuscript.Rmd

--- 2021-09-24 manuscript.Rmd
@@
 the document can be reproduced with a single command, thus updating the Results to their final version.
+Reproducibility is of utmost importance at this stage since the preregistration must produce valid results at two points in time.
+We can only recommend PAC when the dangers to reproducibility we described above are eliminated and a state of automated reproducibility is reached.
 
@@
 Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
-Decisions that do not lend themselves to formulation in code, e.g.
-visual inspection, must still be described verbally or be treated as noted in the section [Deviating from the preregistration and exploration].
 With PAC researchers can precisely calibrate how much they want to explore.
 When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
-Fifth, deviations from the preregistration are clearly documented because they are reflected in changes to the code, which are managed and tracked with version control.
-
-## Deviating from the preregistration and exploration
-
-We would like to note that in the PAC paradigm, the direct comparison of the preregistration and the final publication is particularly easy to accomplish.
-Because there is a single unambiguous reference point, that is the dynamic document, including code, under version control, changes can be inspected by authors and readers alike.
-To make the comparison even more accessible, we highly recommend that the authors review all changes made to the preregistered plan to summarize and justify them for their reader, e.g.
-in the discussion of the final manuscript [^summarize-changes].
-The reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.
-
-Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
-However, we share the common view that a deviation from a preregistration is not a problem; it is a problem if these changes are not made transparent.
-In fact, we expect that most PACs will require changes after data have been collected.
-Often deviations provide an opportunity to learn from the unexpected.
-Transparency is essential in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].
-
-Take the following example: A preregistration intended to include "working memory" and "fluid intelligence" as predictors for reaction time in an experiment.
-However, after fitting the model on the actual data, both constructs exhibit high collinearity.
-The authors decide that for their purpose, these variables are interchangeable and remove one.
-For them, this change simply does not affect the core of their theory but was an auxiliary assumption, and they might be right.
-Another researcher, however, who, e.g.
-hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
-With transparent reporting, it is left to the reader to make up their own judgement.
-
-[^summarize-changes]: In section [Preregistration as Code --- a Tutorial] we have conducted an actual PAC and summarize the changes we made to the preregistered code in the discussion.
-
-Another common misunderstanding is that preregistration, and therefore PAC, preclude exploratory analyses.
-We must differentiate here between two kinds of exploration, and neither of them is limited by PAC.
-The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
-Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results. <!-- citation? The Nosek paper?-->
-The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning.
-<!--citation?
-reproducibility in machine learning research?
-maybe cite the Brandmaier (in press) chapter, where I suggest to use PAC for these kinds of models?
--->
-These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
-We specifically recommend PAC for those exploratory statistical models.
-The merit of preregistration in such cases is to communicate precisely how much exploration was done, a piece of information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.
+Fifth, deviations from the preregistration are clearly documented, because they are reflected in changes to the code which are managed and tracked with version control.
+One could even make a commit labeled "Necessary deviations from the preregistration after data collection",
+and reference it in the manuscript.
+
+Note that preregistration, or PAC, does not preclude exploratory analyses.
+Such analyses should merely be explicitly declared as such in the manuscript.
 
@@
 Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
-Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
-To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything there is a single "source of truth"---dynamic manuscript including code---under version control.
-
+Because the same dynamic document serves as preregistration and final manuscript, researchers, reviewers, and editors only have to revise one document.
+To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything is under version control.
 The likelihood of those deviations, while never unavoidable, is decreased by using simulated data to verify the suitability of the proposed analysis plan.

### 2021-09-27 manuscript.Rmd
suggest organization for multiple tests (#249)
* suggest organization for multiple tests

* move to dedicated section

* note that these are not repro functions

@@
 We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
-We capture these distinct tasks in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
-The simulate and report function can of course be reused, if appropriate, across multiple hypotheses.
+For each study these tasks have to specifically coded.
+While capture the tasks here in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
+The functions shown here are empty (they do not contain executable code) and the interested reader may find function filled with some life in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.
+
 At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` from which they later derive the method section.

### 2021-09-27 manuscript.Rmd

--- 2021-09-27 manuscript.Rmd
@@
 
+This example preregistration covers a single study with a single hypothesis.
+To organize studies with multi hypotheses we suggest multiple (possibly numbered in accord with the hypotheses, e.g. 1.2, 2.3 etc.) `planned_analysis()` and `report_analysis()` functions.
+Steps that apply to multiple analyses should be extracted into their own function e.g. `discard_outliers()` or `extract_scores()`.
+Preregistrations that cover multiple distinct data sources may employ multiple `simulate_data()` functions.
+These are merely suggestions, researcher who are more comfortable with another programming style will surely find an appropriate system.
+
 Of course it is likely that the preregistered code has to be adapted to the actual data.

### 2021-09-27 manuscript.Rmd
pilot is an alternative to simulation

@@
 
+Another approach that can replace but also complement simulated data is to conduct a pilot study [@thabaneTutorialPilotStudies2010] and use the resulting data to preregister.
+A pilot study has obvious advantages for study planning, since it lets researcher evaluate the feasibility of many assumptions.
+However, we must warn our readers, while piloting is more traditional than our approach of obfuscating the data before preregistration it may not be innocuous to use the resulting data and care must be taken to avoid errors of sequential testing [@whiteheadBiasMaximumLikelihood1986].
+
 When the researchers are satisfied with the function `planned_analysis()`, they can think about how they would like to report the analysis results via tables, plots, and text.

### 2021-09-29 manuscript.Rmd
more clearly define scope and readership

--- 2021-09-27 manuscript.Rmd
@@
 Second, we discuss opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated, but also how research itself is conducted.
-With regard to the first goal, we present a tutorial that is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
-In this tutorial, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
-The proposed workflow builds on several technical solutions originating from software engineering to automate reproduction. 
-We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they interact to form a coherent workflow.
-We further demonstrate how the R package `repro` supports researchers in setting up the required software and in adopting this workflow.
+With regard to the first goal, we discuss how to combat four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
+Then we present a tutorial on how to employ these tools to achieve automated reproducibility.
+This tutorial can not introduce the reader comprehensively into each tool but intents to help readers who have picked up bits and pieces to build a coherent workflow.
+This tutorial is aimed at researchers who regularly write code to analyze their data and who are willing to make relevant code, data, and materials available, either publicly or on request.
+Ideally, our reader already has created a dynamic document (e.g. with RMarkdown or Jupyter) and has tried their hand once at version control (e.g. with Git). 
+The R package `repro` supports researchers in setting up the required software and in adopting this workflow.
+We present the automated reproducibility as best practice, a goal that is not always archived because of limited resources, technical restrictions or practical considerations but is worth striving for.
+
 In pursuit of the second aim we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).

### 2021-09-29 manuscript.Rmd
suggest appendix for wordcount #228

@@
 Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
-If the PAC is submitted as a registered report [an article format in which the preregistration is reviewed and the article accepted "in principle", @nosekRegisteredReports2014; @chambersWhatNextRegistered2019] a journal article might be more familiar to the reviewers that ensure a comprehensive study plan.
+One factor that reduces comprehensiveness of such manuscript style but that is absent in standard preregistrations are strict word count limits that some journals impose.
+In such case we suggest using appendix materials to keep the manuscript concise, yet the preregistration comprehensive.
 
@@
 Writing incomplete code is technically impossible, or at least researchers would notice if something is omitted (e.g. if they forgot to aggregate the data).
+
+Peer review of preregistration is a method that also can ensure comprehensiveness e.g.
+a registered report [an article format in which the preregistration is reviewed and the article accepted "in principle", @nosekRegisteredReports2014; @chambersWhatNextRegistered2019].
+In such registered report peer reviewing a manuscript styled PAC might be more familiar to the reviewers that ensure a comprehensive study plan.
 

### 2021-09-29 manuscript.Rmd
readd sentance about vagueness

@@
 This tutorial is aimed at researchers who regularly write code to analyze their data and who are willing to make relevant code, data, and materials available, either publicly or on request.
-Ideally, our reader already has created a dynamic document (e.g. with RMarkdown or Jupyter) and has tried their hand once at version control (e.g. with Git). 
+Ideally, our reader already has created a dynamic document at some point (e.g. with RMarkdown or Jupyter) and has tried their hand once at version control (e.g. with Git). 
 The R package `repro` supports researchers in setting up the required software and in adopting this workflow.
-We present the automated reproducibility as best practice, a goal that is not always archived because of limited resources, technical restrictions or practical considerations but is worth striving for.
+We present automated reproducibility as best practice, a goal that is not always archived because of limited resources, technical restrictions or practical considerations but is worth striving for.
+If, due to whatever reason, some parts of the workflow proof to be unfeasible, we encourage the reader to find their own suitable solutions and to concentrate on the feasible parts.
 
@@
 We, however, prefer the second way of building images automatically from a textual description called `Dockerfile`.
-Because the `Dockerfile` clearly describes which software is installed how, users can quickly change the software environment e.g.
-update to another R or package version without being "stuck" with an unmaintainable because manually build `Docker` image.
+Because the `Dockerfile` clearly describes which software is installed how the installation process can be repeated automatically.
+Users can therefore quickly change the software environment e.g. update to another R or package version without being "stuck" with an `Docker` image they do not know how to rebuild.
 
@@
 The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing `Docker`.
+Unfortunately, Docker requires administrative rights to run.
+We recommend [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@R-renv] in cases where no administrative rights can be obtained but can not detail its use in this document.
 Start RStudio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
@@
 Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
+
+After you have installed the necessary software we suggest that you set up a secure connection to GitHub:
+
+```{r, include=FALSE}
+opts <- options()
+```
+
+```{r}
+repro::check_github()
+```
+
+If you know what SSH is and want to use it, you may alternatively use:
+
+```{r}
+# only an alternative: DO NOT USE if you are unsure what SSH means
+repro::check_github(auth_method = "ssh")
+```
+
+```{r, include=FALSE}
+options(opts)
+```
+
 If necessary, follow any instructions presented until all checks pass.
@@
 ```
+
 Now the file `private.md` will not be committed to `Git` and will not be made public.
+Please consider carefully if you can include data in the repository without violating privacy rights.
+If you can not share your data publicly add it to the `.gitignore` file and share it on request.
 
@@
 specific journal requirements.
+Noteworthy is the package output format supplied by the [`papaja`](https://github.com/crsh/papaja) package [@papaja] which produces APA compliant manuscripts [@apa7].
 Another advantage is that, because `R Markdown` files are in plain text, they are more suitable for version control using `Git`.
@@
 Together, these files allow users (including your future self) to reproduce every step in the analysis automatically.
+Please run `repro::automate()` in your project:
+
+```{r, eval=FALSE}
+repro::automate()
+```
+
 It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new `R Markdown` file to the project, to keep the `Makefile` and `Dockerfile` up to date.
@@
 This task is difficult because it impossible to decide without additional information if the analysis actually was carried out differently or just described differently.
+The comparison is made impossible if the authors of the preregistration were intentionally vague.
 Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare to different texts which may be worded differently or describe the same thing in varying level of detail.
@@
 For each study these tasks have to specifically coded.
-While capture the tasks here in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
+We capture the tasks here in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
 The functions shown here are empty (they do not contain executable code) and the interested reader may find function filled with some life in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.
@@
 And click on Releases → Draft a new release.
-Here you can add all relevant binary files but at least a rendered version of the manuscript and the docker image.
+Here you can add all relevant binary files but at least a rendered version of the manuscript and the Docker image.
 

### 2021-09-29 manuscript.Rmd
readd sentance about vagueness
change wording

--- 2021-09-29 manuscript.Rmd
@@
 Such retrospective summary already serves to put changes into context, however, a change log that is kept continually and where changes are explained when they are made, provides more immediate and detailed information.
-Each entry in the change log should contain the commit id to the respective changes.
-The reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.
+Each entry in the change log should explain the reasoning of the changes and link to the changes made by including the the commit id that applied the changes.
+The interested reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.
 

### 2021-09-29 manuscript.Rmd
add references

--- 2021-09-29 manuscript.Rmd
@@
 
-Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
-However, we share the common view that a deviation from a preregistration is not a problem; it is a problem if these changes are not made transparent.
+Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study [@szollosi_is_2020].
+However, we share the common view that a deviation from a preregistration is not a problem [@nosekPreregistrationHardWorthwhile2019]; it is a problem if these changes are not made transparent.
 In fact, we expect that most PACs will require changes after data have been collected.
@@
 Another researcher, however, who, e.g.
-hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
+hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because the change to the preregistration affects their theory directly.
 With transparent reporting, it is left to the reader to make up their own judgement.
@@
 The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
-Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results. <!-- citation? The Nosek paper?-->
-The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning.
-<!--citation?
-reproducibility in machine learning research?
-maybe cite the Brandmaier (in press) chapter, where I suggest to use PAC for these kinds of models?
--->
+Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
+The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning [@brandmaier2021].
 These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.

### 2021-09-29 manuscript.Rmd
typo

--- 2021-09-29 manuscript.Rmd
@@
 This task is difficult because it impossible to decide without additional information if the analysis actually was carried out differently or just described differently.
-The comparison is made impossible if the authors of the preregistration were intentionally vague.
 Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare to different texts which may be worded differently or describe the same thing in varying level of detail.
@@
 Such retrospective summary already serves to put changes into context, however, a change log that is kept continually and where changes are explained when they are made, provides more immediate and detailed information.
-Each entry in the change log should explain the reasoning of the changes and link to the changes made by including the the commit id that applied the changes.
-The interested reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.
-
-Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study [@szollosi_is_2020].
-However, we share the common view that a deviation from a preregistration is not a problem [@nosekPreregistrationHardWorthwhile2019]; it is a problem if these changes are not made transparent.
+Each entry in the change log should contain the commit id to the respective changes.
+The reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.
+
+Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
+However, we share the common view that a deviation from a preregistration is not a problem; it is a problem if these changes are not made transparent.
 In fact, we expect that most PACs will require changes after data have been collected.
@@
 Another researcher, however, who, e.g.
-hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because the change to the preregistration affects their theory directly.
+hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because of the change to the preregistrations.
 With transparent reporting, it is left to the reader to make up their own judgement.
@@
 The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
-Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
-The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning [@brandmaier2021].
+Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results. <!-- citation? The Nosek paper?-->
+The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning.
+<!--citation?
+reproducibility in machine learning research?
+maybe cite the Brandmaier (in press) chapter, where I suggest to use PAC for these kinds of models?
+-->
 These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.

### 2021-09-29 manuscript.Rmd
suggest to run automate

@@
 ```
+
 Now the file `private.md` will not be committed to `Git` and will not be made public.

### 2021-09-29 manuscript.Rmd
add correct brandmaier reference

--- 2021-09-29 manuscript.Rmd
@@
 Such retrospective summary already serves to put changes into context, however, a change log that is kept continually and where changes are explained when they are made, provides more immediate and detailed information.
-Each entry in the change log should contain the commit id to the respective changes.
-The reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.
-
-Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study.
-However, we share the common view that a deviation from a preregistration is not a problem; it is a problem if these changes are not made transparent.
+Each entry in the change log should explain the reasoning of the changes and link to the changes made by including the the commit id that applied the changes.
+The interested reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.
+
+Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study [@szollosi_is_2020].
+However, we share the common view that a deviation from a preregistration is not a problem [@nosekPreregistrationHardWorthwhile2019]; it is a problem if these changes are not made transparent.
 In fact, we expect that most PACs will require changes after data have been collected.
@@
 The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
-Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results. <!-- citation? The Nosek paper?-->
-The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning.
-<!--citation?
-reproducibility in machine learning research?
-maybe cite the Brandmaier (in press) chapter, where I suggest to use PAC for these kinds of models?
--->
+Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
+The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning [@brandmaier22mlsem].
 These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
@@
 And click on Releases → Draft a new release.
-Here you can add all relevant binary files but at least a rendered version of the manuscript and the Docker image.
+Here you can add all relevant binary files but at least a rendered version of the manuscript and the docker image.
 

### 2021-09-30 manuscript.Rmd
note privacy concerns closes #217
rephrase

@@
 Now the file `private.md` will not be committed to `Git` and will not be made public.
+Please consider carefully if you can include data in the repository without violating privacy rights.
+If you can not share your data publicly add it to the `.gitignore` file and share it on request.
 

### 2021-10-04 manuscript.Rmd
pilot is an alternative to simulation (#252)

@@
 members of group "A" have a greater mean in variable "Z").
+
+Another approach that can replace but also complement simulated data is to conduct a pilot study [@thabaneTutorialPilotStudies2010] and use the resulting data to preregister.
+A pilot study has obvious advantages for study planning, since it lets researcher evaluate the feasibility of many assumptions.
+However, we must warn our readers, while piloting is more traditional than our approach of obfuscating the data before preregistration it may not be innocuous to use the resulting data and care must be taken to avoid errors of sequential testing [@whiteheadBiasMaximumLikelihood1986].
 

### 2021-10-04 manuscript.Rmd
mention determinism and seeds

@@
 In essence, researchers provide a collection of *computational recipes*.
-A computational recipe describes how inputs are processed to create a specific output in a way that is automatically executable.
+A computational recipe describes how inputs are processed to deterministically create a specific output in a way that is automatically executable.
 The concept of computational recipes is central to our understanding of reproducibility because it enables a unified way to automatically reproduce a project.
@@
 A workflow that includes, e.g. renaming variable names by hand in a graphical spreadsheet application is therefore incompatible with automated reproducibility.
+Another property of a computational recipe is that the same inputs of should always result in the same outputs.
+Fo most computer code (given the same software is used) this property is fulfilled, however, one noteworthy exception are pseudo random numbers.
+Whenever random numbers are used in a computation, this computation is only reproducible when the same random numbers are used.
+To ensure identical random numbers, users may fix the state of the random number generate with a so called seed (e.g. in R with `set.seed()`).
 

### 2021-10-04 manuscript.Rmd
add clearer warning
be clear about what "shuffling" means

@@
 Of course researcher must not analyze the data before they write the preregistration.
-Nevertheless, they may obfuscate the real data by shuffling all variables or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
+Nevertheless, they may obfuscate the real data by shuffling all variables columnwise (independently from each other) or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
 Preregistration after data collection is common when reanalyzing data from another research group [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
@@
 members of group "A" have a greater mean in variable "Z").
+To be clear, any exploration of the data before preregistration renders the preregistration invalid and amounts to grave scientific (albeit sometime unintentional) misconduct.
 

### 2021-10-04 manuscript.Rmd
update repro
fake github access in docker

@@
     - gert
-    - aaronpeikert/repro@27e7d22
+    - aaronpeikert/repro@15a069
     - here
@@
 opts <- options()
+options(repro.github = TRUE,
+        repro.github.ssh = TRUE,
+        repro.ssh = TRUE,
+        repro.github.token = TRUE)
 ```
@@
 
+```{r, include=FALSE}
+opts <- options()
+options(repro.github = TRUE,
+        repro.github.ssh = TRUE,
+        repro.github.token = TRUE)
+```
+
 ```{r}
 repro::check_git()
+repro::check_github()
 repro::check_make()
 repro::check_docker()
+```
+
+```{r, include=FALSE}
+options(opts)
 ```

### 2021-10-05 manuscript.Rmd
add screenshot of rstudio git pane
remove double labels of figures
fix typo

@@
 
+\captionsetup{labelformat=empty}
+
 # Introduction
@@
 
+```{r modified, eval = TRUE, echo = FALSE, fig.cap="The Git pane in R Studio, showing `manuscript.Rmd` modified but unstaged and `modified.png` newly added and staged.  "}
+knitr::include_graphics(here::here("images", "modified.png"), auto_pdf = TRUE)
+```
+
 For `Git` to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
@@
 The comparison is made impossible if the authors of the preregistration were intentionally vague.
-Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare to different texts which may be worded differently or describe the same thing in varying level of detail.
+Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare two different texts which may be worded differently or describe the same thing in varying level of detail.
 A third limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.

### 2021-10-05 manuscript.Rmd
add section when pac is applicable

@@
 
+#### When is PAC applicable
+
+Every study that can be preregistered, can be preregistered with PAC.
+One significant difference is that PAC demands a comprehensive analysis plan.
+With traditional preregistration, however, the researcher gets away with deciding details later.
+Some researchers may not feel ready to decide every detail.
+In such case, we recommend to decide based on the best available knowledge and if these decisions prove infeasible, revise and explain them later as described in the section [Deviating from the preregistration and exploration].
+
+Two applications, however, lend themselves to PAC.
+First, registrations of clinical trials typically describe analyses in excruciating detail.
+Those registrations typically contain a detailed description of how results will be presented, including example tables and graphics.
+PAC may significantly reduce the required workload while maintaining (and exceeding) the required standards for registering a clinical trial.
+
+Second, preregistering explorative statistical models (i.e.
+machine learning) is hardly feasible with standard preregistrations since they are too complex to describe and depend strongly on their software implementation verbally.
+PAC, however, captures the precise algorithmic model, including its software implementation and is ideal for preregistering these models.
+
 # Preregistration as Code --- a Tutorial

### 2021-10-05 manuscript.Rmd
cite brandmaier
Update manuscript.Rmd
some rephrasing and a little bit more verbosity.

--- 2021-10-05 manuscript.Rmd
@@
 
+#### When is PAC applicable
+
+Every study that can be preregistered, can be preregistered with PAC.
+The only difference is that PAC demands a comprehensive and complete analysis plan.
+With traditional preregistration, researchers may omit details and get away with decision that are made only after having seen the data, thus, potentially biasing statistical results.
+Some researchers may not feel ready to decide every detail of the analysis plan before data collection.
+In such case, we recommend to make decisions based on the best available knowledge and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in the section [Deviating from the preregistration and exploration] of the final manuscript.
+
+Beyond standard preregistrations in psychological research, we believe that two applications specifically lend themselves to PAC.
+First, registrations of clinical trials typically describe analyses in exhaustive detail.
+Those registrations typically contain a detailed description of how results will be presented, including example tables and graphics. <!--citation needed-->
+PAC may significantly reduce the required workload while maintaining (and exceeding) the required standards for registering a clinical trial.
+
+Second, preregistering exploratory statistical models (i.e., those with large numbers of competing models or those inspired by machine learning) is hardly feasible with standard preregistrations since they are too complex to describe and depend strongly on their software implementation.
+PAC, however, captures the precise algorithmic model, including its software implementation and is ideal for preregistering these models [@brandmaier22mlsem].
+
 # Preregistration as Code --- a Tutorial

### 2021-10-06 manuscript.Rmd
cite E9

@@
 We argue that it is still an eligible preregistration.
+Guidelines for clinical trials also recommend analysis of blinded blinded to test the practicality of the preregistration [@ICH1998].
 Since the data is obfuscated, almost no results are revealed and the researchers can not employ their researcher degrees of freedom effectively to influence the results.
@@
 First, registrations of clinical trials typically describe analyses in exhaustive detail.
-Those registrations typically contain a detailed description of how results will be presented, including example tables and graphics. <!--citation needed-->
+Those registrations typically contain a detailed description of how results will be presented, including example tables and graphics [@ICH1998].
 PAC may significantly reduce the required workload while maintaining (and exceeding) the required standards for registering a clinical trial.

### 2021-10-06 manuscript.Rmd
added third affiliation to AB rewrote abstract

@@
   - name: Andreas M. Brandmaier
-    affil: 1, 4
+    affil: 1, 4, 5
     orcid: 0000-0001-8765-6982
@@
       Berlin, Germany and London, UK
+  - num: 5
+    address: |
+      MSB Medical School Berlin,
+      Berlin, Germany
 # firstnote to eighthnote
@@
   Reproducibility has long been considered integral to the scientific method.
-  Something is called reproducible when an independent person obtains the same results from the same data.
+  A scientific analysis is called reproducible when an independent person obtains the same results from the same data.
+<!--AB something seems to generic; we could generalize this to digital objects, though-->
   Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
   Technological advancements now enable scientists to achieve a more comprehensive standard;
-  one in which any individual can be granted access to a digital research repository, and reproduce the analyses from the raw data to the final report including all relevant statistical analyses with a single command.
+  a standard where anyone can access a digital research repository and reproduce all computational steps from raw data to final report including all relevant statistical analyses, with a single command.
   This method has far-reaching implications for scientific archiving,
@@
   scientific productivity,
-  and the credibility and reliability of scientific findings.
-  One obstacle preventing the widespread adoption of this method is that the underlying technological advancements
-  are complicated to use.
-  This paper introduces `repro`, an R-package, which guides researchers in the installation and use of the tools required for making a research project reproducible.
-  Finally, we suggest the use of the proposed tools for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
-  Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions of those analyses, which merely complement the PAC as a more readable summary.
-  PAC circumvents the shortcomings of ambiguous preregistrations that may give researchers undesired degrees of freedom.
-  Hence, reproducibility made convenient with automation has a wide range of applications to accelerate scientific progress.
+  and the credibility and reliability of scientific knowledge.
+  One obstacle to the widespread use of this method is that the underlying tools
+  are complex and not part of the regular training of most researchers.
+  This paper introduces `repro`, an R-package, which guides researchers in installing and using the tools needed to make a research project reproducible.
+  Finally, we suggest the use of the proposed workflow for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
+  Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions of those analyses, which then merely complement the PAC as a more readable summary.
+  PAC circumvents the shortcomings of ambiguous preregistrations that may give researchers undesired degrees of freedom in the execution of their analysis.
+  Reproducibility, facilitated by automation, therefore has a wide range of applications to accelerate scientific progress.
 abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"

### 2021-10-06 manuscript.Rmd
try to address caspars' critizism

--- 2021-10-05 manuscript.Rmd
@@
     - gert
-    - aaronpeikert/repro@15a069
+    - aaronpeikert/repro@27e7d22
     - here
@@
 
-\captionsetup{labelformat=empty}
-
 # Introduction
@@
 Second, we discuss opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated, but also how research itself is conducted.
-With regard to the first goal, we discuss how to combat four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
-Then we present a tutorial on how to employ these tools to achieve automated reproducibility.
-This tutorial can not introduce the reader comprehensively into each tool but intents to help readers who have picked up bits and pieces to build a coherent workflow.
-This tutorial is aimed at researchers who regularly write code to analyze their data and who are willing to make relevant code, data, and materials available, either publicly or on request.
-Ideally, our reader already has created a dynamic document at some point (e.g. with RMarkdown or Jupyter) and has tried their hand once at version control (e.g. with Git). 
-The R package `repro` supports researchers in setting up the required software and in adopting this workflow.
-We present automated reproducibility as best practice, a goal that is not always archived because of limited resources, technical restrictions or practical considerations but is worth striving for.
-If, due to whatever reason, some parts of the workflow proof to be unfeasible, we encourage the reader to find their own suitable solutions and to concentrate on the feasible parts.
-
+With regard to the first goal, we present a tutorial that is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
+In this tutorial, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.
+The proposed workflow builds on several technical solutions originating from software engineering to automate reproduction. 
+We discuss how these solutions address four common threats to reproducibility [see @Peikert2019 for details] and how they interact to form a coherent workflow.
+We further demonstrate how the R package `repro` supports researchers in setting up the required software and in adopting this workflow.
 In pursuit of the second aim we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
@@
 In essence, researchers provide a collection of *computational recipes*.
-A computational recipe describes how inputs are processed to deterministically create a specific output in a way that is automatically executable.
+A computational recipe describes how inputs are processed to create a specific output in a way that is automatically executable.
 The concept of computational recipes is central to our understanding of reproducibility because it enables a unified way to automatically reproduce a project.
@@
 A workflow that includes, e.g. renaming variable names by hand in a graphical spreadsheet application is therefore incompatible with automated reproducibility.
-Another property of a computational recipe is that the same inputs of should always result in the same outputs.
-Fo most computer code (given the same software is used) this property is fulfilled, however, one noteworthy exception are pseudo random numbers.
-Whenever random numbers are used in a computation, this computation is only reproducible when the same random numbers are used.
-To ensure identical random numbers, users may fix the state of the random number generate with a so called seed (e.g. in R with `set.seed()`).
 
@@
 We, however, prefer the second way of building images automatically from a textual description called `Dockerfile`.
-Because the `Dockerfile` clearly describes which software is installed how the installation process can be repeated automatically.
-Users can therefore quickly change the software environment e.g. update to another R or package version without being "stuck" with an `Docker` image they do not know how to rebuild.
+This enables users to quickly change the software environment e.g. update to another R or package version without being "stuck" with a unmaintainable because manually build `Docker` image.
 
@@
 The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing `Docker`.
-Unfortunately, Docker requires administrative rights to run.
-We recommend [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@R-renv] in cases where no administrative rights can be obtained but can not detail its use in this document.
 Start RStudio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
@@
 Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
-
-After you have installed the necessary software we suggest that you set up a secure connection to GitHub:
-
-```{r, include=FALSE}
-opts <- options()
-options(repro.github = TRUE,
-        repro.github.ssh = TRUE,
-        repro.ssh = TRUE,
-        repro.github.token = TRUE)
-```
-
-```{r}
-repro::check_github()
-```
-
-If you know what SSH is and want to use it, you may alternatively use:
-
-```{r}
-# only an alternative: DO NOT USE if you are unsure what SSH means
-repro::check_github(auth_method = "ssh")
-```
-
-```{r, include=FALSE}
-options(opts)
-```
-
 If necessary, follow any instructions presented until all checks pass.
@@
 ```
-
 Now the file `private.md` will not be committed to `Git` and will not be made public.
-Please consider carefully if you can include data in the repository without violating privacy rights.
-If you can not share your data publicly add it to the `.gitignore` file and share it on request.
 
@@
 usethis::use_git()
-```
-
-```{r modified, eval = TRUE, echo = FALSE, fig.cap="The Git pane in R Studio, showing `manuscript.Rmd` modified but unstaged and `modified.png` newly added and staged.  "}
-knitr::include_graphics(here::here("images", "modified.png"), auto_pdf = TRUE)
 ```
@@
 specific journal requirements.
-Noteworthy is the package output format supplied by the [`papaja`](https://github.com/crsh/papaja) package [@papaja] which produces APA compliant manuscripts [@apa7].
 Another advantage is that, because `R Markdown` files are in plain text, they are more suitable for version control using `Git`.
@@
 Together, these files allow users (including your future self) to reproduce every step in the analysis automatically.
-Please run `repro::automate()` in your project:
-
-```{r, eval=FALSE}
-repro::automate()
-```
-
 It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new `R Markdown` file to the project, to keep the `Makefile` and `Dockerfile` up to date.
@@
 
-```{r, include=FALSE}
-opts <- options()
-options(repro.github = TRUE,
-        repro.github.ssh = TRUE,
-        repro.github.token = TRUE)
-```
-
 ```{r}
 repro::check_git()
-repro::check_github()
 repro::check_make()
 repro::check_docker()
-```
-
-```{r, include=FALSE}
-options(opts)
 ```
@@
 However, templates cannot remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
-This ambiguity causes a second problem, that is comparing the preregistration and the resulting publication to decide if and how researcher deviated from the preregistration.
-This task is difficult because it impossible to decide without additional information if the analysis actually was carried out differently or just described differently.
-The comparison is made impossible if the authors of the preregistration were intentionally vague.
-Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare two different texts which may be worded differently or describe the same thing in varying level of detail.
-A third limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
+A second limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
 To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
@@
 These results are initially based on simulated data with the same structure as the data they expect to obtain from their experiments.
-For guidance on how to simulate data, see @morrisUsingSimulationStudies2019, @paxtonMonteCarloExperiments2001 and @skrondalDesignAnalysisMonte2000.
+For guidance on how to simulate data, see @paxtonMonteCarloExperiments2001 and @skrondalDesignAnalysisMonte2000.
 Once the preregistration is submitted and real data have been collected or made available,
 the document can be reproduced with a single command, thus updating the Results to their final version.
-Reproducibility is of utmost importance at this stage since the preregistration must produce valid results at two points in time.
-We can only recommend PAC when the dangers to reproducibility we described above are eliminated and a state of automated reproducibility is reached.
 
@@
 Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
-Decisions that do not lend themselves to formulation in code, e.g.
-visual inspection, must still be described verbally or be treated as noted in the section [Deviating from the preregistration and exploration].
 With PAC researchers can precisely calibrate how much they want to explore.
 When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
-Fifth, deviations from the preregistration are clearly documented because they are reflected in changes to the code, which are managed and tracked with version control.
-
-## Deviating from the preregistration and exploration
-
-We would like to note that in the PAC paradigm, the direct comparison of the preregistration and the final publication is particularly easy to accomplish.
-Because there is a single unambiguous reference point, that is the dynamic document, including code, under version control, changes can be inspected by authors and readers alike.
-Authors should at least review all changes made to the preregistered plan to summarize and justify them for their reader, e.g. in the discussion of the final manuscript [^summarize-changes].
-Such retrospective summary already serves to put changes into context, however, a change log that is kept continually and where changes are explained when they are made, provides more immediate and detailed information.
-Each entry in the change log should explain the reasoning of the changes and link to the changes made by including the the commit id that applied the changes.
-The interested reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.
-
-Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study [@szollosi_is_2020].
-However, we share the common view that a deviation from a preregistration is not a problem [@nosekPreregistrationHardWorthwhile2019]; it is a problem if these changes are not made transparent.
-In fact, we expect that most PACs will require changes after data have been collected.
-Often deviations provide an opportunity to learn from the unexpected.
-Transparency is essential in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].
-
-Take the following example: A preregistration intended to include "working memory" and "fluid intelligence" as predictors for reaction time in an experiment.
-However, after fitting the model on the actual data, both constructs exhibit high collinearity.
-The authors decide that for their purpose, these variables are interchangeable and remove one.
-For them, this change simply does not affect the core of their theory but was an auxiliary assumption, and they might be right.
-Another researcher, however, who, e.g.
-hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because the change to the preregistration affects their theory directly.
-With transparent reporting, it is left to the reader to make up their own judgement.
-
-[^summarize-changes]: In section [Preregistration as Code --- a Tutorial] we have conducted an actual PAC and summarize the changes we made to the preregistered code in the discussion.
-
-Another common misunderstanding is that preregistration, and therefore PAC, preclude exploratory analyses.
-We must differentiate here between two kinds of exploration, and neither of them is limited by PAC.
-The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
-Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
-The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning [@brandmaier22mlsem].
-These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
-We specifically recommend PAC for those exploratory statistical models.
-The merit of preregistration in such cases is to communicate precisely how much exploration was done, a piece of information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.
+Fifth, deviations from the preregistration are clearly documented, because they are reflected in changes to the code which are managed and tracked with version control.
+One could even make a commit labeled "Necessary deviations from the preregistration after data collection",
+and reference it in the manuscript.
+
+Note that preregistration, or PAC, does not preclude exploratory analyses.
+Such analyses should merely be explicitly declared as such in the manuscript.
 
@@
 Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
-One factor that reduces comprehensiveness of such manuscript style but that is absent in standard preregistrations are strict word count limits that some journals impose.
-In such case we suggest using appendix materials to keep the manuscript concise, yet the preregistration comprehensive.
+If the PAC is submitted as a registered report [an article format in which the preregistration is reviewed and the article accepted "in principle", @nosekRegisteredReports2014; @chambersWhatNextRegistered2019] a journal article might be more familiar to the reviewers that ensure a comprehensive study plan.
 
@@
 Writing incomplete code is technically impossible, or at least researchers would notice if something is omitted (e.g. if they forgot to aggregate the data).
-
-Peer review of preregistration is a method that also can ensure comprehensiveness e.g.
-a registered report [an article format in which the preregistration is reviewed and the article accepted "in principle", @nosekRegisteredReports2014; @chambersWhatNextRegistered2019].
-In such registered report peer reviewing a manuscript styled PAC might be more familiar to the reviewers that ensure a comprehensive study plan.
 
@@
 We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
-For each study these tasks have to specifically coded.
-We capture the tasks here in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
-The functions shown here are empty (they do not contain executable code) and the interested reader may find function filled with some life in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.
-
+We capture these distinct tasks in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
+The simulate and report function can of course be reused, if appropriate, across multiple hypotheses.
 At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` from which they later derive the method section.
@@
 If simulating data proves to be too challenging, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
-In order to do so researcher must walk a fine but utmost important line.
-Of course researcher must not analyze the data before they write the preregistration.
-Nevertheless, they may obfuscate the real data by shuffling all variables columnwise (independently from each other) or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
-Preregistration after data collection is common when reanalyzing data from another research group [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
-We argue that it is still an eligible preregistration.
-Since the data is obfuscated, almost no results are revealed and the researchers can not employ their researcher degrees of freedom effectively to influence the results.
-Shuffling leaves only univariate information, so this procedure is only applicable for bi-or multivariate hypotheses.
-Researcher still can access the information about means or proportions (e.g.
-how many participants belonging to group "A" are in the dataset) but are unbiased for all relations between variables (e.g.
-members of group "A" have a greater mean in variable "Z").
-To be clear, any exploration of the data before preregistration renders the preregistration invalid and amounts to grave scientific (albeit sometime unintentional) misconduct.
-
-Another approach that can replace but also complement simulated data is to conduct a pilot study [@thabaneTutorialPilotStudies2010] and use the resulting data to preregister.
-A pilot study has obvious advantages for study planning, since it lets researcher evaluate the feasibility of many assumptions.
-However, we must warn our readers, while piloting is more traditional than our approach of obfuscating the data before preregistration it may not be innocuous to use the resulting data and care must be taken to avoid errors of sequential testing [@whiteheadBiasMaximumLikelihood1986].
+Of course researcher must not analyze the data before they write the preregistration but they can obfuscate the real data by shuffling the dependent variable or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
 
@@
 After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data) and write the discussion.
-
-#### When is PAC applicable
-
-Every study that can be preregistered, can be preregistered with PAC.
-The only difference is that PAC demands a comprehensive and complete analysis plan.
-With traditional preregistration, researchers may omit details and get away with decision that are made only after having seen the data, thus, potentially biasing statistical results.
-Some researchers may not feel ready to decide every detail of the analysis plan before data collection.
-In such case, we recommend to make decisions based on the best available knowledge and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in the section [Deviating from the preregistration and exploration] of the final manuscript.
-
-Beyond standard preregistrations in psychological research, we believe that two applications specifically lend themselves to PAC.
-First, registrations of clinical trials typically describe analyses in exhaustive detail.
-Those registrations typically contain a detailed description of how results will be presented, including example tables and graphics. <!--citation needed-->
-PAC may significantly reduce the required workload while maintaining (and exceeding) the required standards for registering a clinical trial.
-
-Second, preregistering exploratory statistical models (i.e., those with large numbers of competing models or those inspired by machine learning) is hardly feasible with standard preregistrations since they are too complex to describe and depend strongly on their software implementation.
-PAC, however, captures the precise algorithmic model, including its software implementation and is ideal for preregistering these models [@brandmaier22mlsem].
 
@@
 
-This example preregistration covers a single study with a single hypothesis.
-To organize studies with multi hypotheses we suggest multiple (possibly numbered in accord with the hypotheses, e.g. 1.2, 2.3 etc.) `planned_analysis()` and `report_analysis()` functions.
-Steps that apply to multiple analyses should be extracted into their own function e.g. `discard_outliers()` or `extract_scores()`.
-Preregistrations that cover multiple distinct data sources may employ multiple `simulate_data()` functions.
-These are merely suggestions, researcher who are more comfortable with another programming style will surely find an appropriate system.
-
-Of course it is likely that the preregistered code has to be adapted to the actual data.
-Changing the preregistration must be discussed.
-This is our summary of what we changed:
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("discuss-changes")
-```
-
-But readers can inspect and judge the changes for themselves on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
-
-The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
-
-After the researcher is satisfied with their draft preregistration, they should create an archived and unchangeable version of the project that serves as the preregistration.
-zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides digital object identifiers (DOI) for these archives.
-While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
-Please note that you can only link public GitHub repositories to zenodo.
-You may log into zenodo.org through your GitHub account.
-
-> Log into zenodo.org → Account → GitHub[^zenodo-github]
-[^zenodo-github]: https://zenodo.org/account/settings/github/
-
-After you have linked a GitHub repository you trigger the archivation by creating a GitHub release.
-To create GitHub release, navigate to GitHub:
-
-```{r}
-usethis::browse_github()
-```
-
-And click on Releases → Draft a new release.
-Here you can add all relevant binary files but at least a rendered version of the manuscript and the Docker image.
+The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/releases/tag/compav0.0.2.0-results).
 
@@
 Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
-To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything there is a single "source of truth"---dynamic manuscript including code---under version control.
-
+To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything is under version control.
 The likelihood of those deviations, while never unavoidable, is decreased by using simulated data to verify the suitability of the proposed analysis plan.
@@
 
+We have to distinguish two kinds of limitations.
+First, some limitations are by design and are unlikely to be changed soon.
+Our workflow does inherit these from the software it relies on and the fundamental design principles these share with the workflow and `repro`.
+Second, there are other limitations that we hope development effort of the open-source community and, concerning `repro`, us will ease.
+
+As a principle, this workflow is incompatible with steps that cannot be automated.
+This principle is at odds with some ingrained habits of researchers to mix and match manual and automated steps in data analysis.
+To allow for automation, many researchers will have to search for alternative software.  
+The automation friendly software we present here has several technical but critical limitations.
+`Git` can track any files (files bigger than 1GB only with extensions) but can only compare text files (with endings like, `.txt`, `.csv`, `.R`,  or `.Rmd`), not binary files (with endings like `.docx`, `.exe` or `.zip`).
+With `R Markdown`, tables and graphics are dynamically generated from code but cannot be crafted by hand.
+`Make` can orchestrate any programmable software but fails to automate software that will only obey a user clicking at the right buttons.
+`Docker` can ship software that runs on Linux and can be automatically installed, but that requirement precludes many commercial or closed source software.
+
+This move away from software that has served researchers well for decades is understandably difficult and presents us with a conundrum.
+On the one hand, we firmly believe that automated reproducibility makes research more productive and collaboration easier.
+But, on the other hand, we expect researchers to invest considerable time in learning new tools and than must persuade their collaborators to do the same.
+Three arguments reconcile this apparent paradox.
+First, this change will not happen all at once.
+Automated reproducibility is an ideal that we believe has many advantages, but it is not an all or nothing decision.
+Researchers can pick up one skill at a time and then help their fellow collaborators do the same.
+Second, the upfront investment is one time (and efforts such as `repro` are underway to reduce it) and will pay dividends over many research projects.
+Third, the move towards open software for research offers several benefits beyond enabling automated reproducibility [@schaffnerFutureScientificJournals1994; @fitzgeraldTransformationOpenSource2006; @chaldecottHistoryScientificTechnical1965; @sonnenburgNeedOpenSource2007].
+
 `repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
 Despite our efforts to ensure this goal, the package is still in development and undoubtedly has limitations.
-The workflow we present here introduces several software solutions, which in itself present a threat to long-term reproducibility.
+The workflow we present here introduces several software solutions, which in itself present a threat to fully automated reproducibility.
 To benefit from automatic and convenient reproduction, researcher must have `Git`, `Make`, and `Docker`.

### 2021-10-06 manuscript.Rmd
suggested edits

--- 2021-10-06 manuscript.Rmd
@@
 Objectivity in this context means "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
-Only something reproducible can be tested and understood by anyone.
+Only a reproducible object can be tested and understood by anyone.
 Therefore, reproducibility has long been considered integral to empirical research.
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
-This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (e.g., missing software or ambiguous documentation of all steps taken), and there is a dire need for reproducibility standards.
-With the technological advancement, it is now possible to make all digital products related to a research project available so that a research project can be automatically reproduced with minimal effort.
-
-We pursue two aims.
+This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (e.g., missing software dependencies or ambiguous documentation of the exact computational steps taken), and there is a dire need for reproducibility standards.
+With the technological advancement, it is now possible to make all digital products related to a research project available so that a research project can be automatically reproduced with minimal effort.<!--AB I don't understand: why is availabbility based on technological advancement?
+aren't there still be legal restrictions?-->
+
+With this paper we pursue two aims.
 First, we want to introduce researchers to a notion of automated reproducibility that requires, apart from the setup of software, no manual steps. 
 Second, we discuss opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated, but also how research itself is conducted.
-With regard to the first goal, we discuss how to combat four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
+With regard to the first goal, we discuss how to address <!--AB please avoid military wordings--> four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
 Then we present a tutorial on how to employ these tools to achieve automated reproducibility.
-This tutorial can not introduce the reader comprehensively into each tool but intents to help readers who have picked up bits and pieces to build a coherent workflow.
+A single tutorial cannot comprehensively introduce the reader to the individual tools, but this tutorial is intended to help readers to build a coherent workflow.
 This tutorial is aimed at researchers who regularly write code to analyze their data and who are willing to make relevant code, data, and materials available, either publicly or on request.
-Ideally, our reader already has created a dynamic document at some point (e.g. with RMarkdown or Jupyter) and has tried their hand once at version control (e.g. with Git). 
+Ideally, our target reader already has created a dynamic document at some point in time (e.g. with RMarkdown or Jupyter) and knows the basic ideas behind version control (e.g. with Git). 
 The R package `repro` supports researchers in setting up the required software and in adopting this workflow.
@@
 
-In pursuit of the second aim we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
-PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including typical sections like Introduction, Methods, and Results.
+In pursuit of the second aim, we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
+PAC involves preregistering the intended analysis code and the major part of the final scientific report as a dynamic document, including typical sections like Introduction, Methods, and Results.
 The resulting dynamic document closely resembles the final manuscript, but temporarily uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
 With these placeholders, authors preregister the exact presentation of their results without having seen real data yet.
-Once the empirical data are analyzed, the Results are then updated automatically, and the Discussion can be written to finalize the report.
+Once the empirical data are available, they replace the simulated data, the results are then updated automatically, and the discussion can be written to finalize the report.
 
@@
 A workflow that includes, e.g. renaming variable names by hand in a graphical spreadsheet application is therefore incompatible with automated reproducibility.
-Another property of a computational recipe is that the same inputs of should always result in the same outputs.
+Another property of a computational recipe is that the same inputs <!--AB of what? I deleted 'of'--> should always result in the same outputs.
 Fo most computer code (given the same software is used) this property is fulfilled, however, one noteworthy exception are pseudo random numbers.
-Whenever random numbers are used in a computation, this computation is only reproducible when the same random numbers are used.
-To ensure identical random numbers, users may fix the state of the random number generate with a so called seed (e.g. in R with `set.seed()`).
+Whenever random numbers are used in a computation, this computation is only reproducible when the samerandom numbers are used.
+To ensure identical random numbers, users may fix the state of the random number generate with a so called seed (e.g.
+in R with `set.seed()`) but they also need to guarantee that the pseudo random number generator is unchanged (see @Peikert2019).
+<!--AB please check if citation is rendered correctly-->
 
@@
 Because the `Dockerfile` clearly describes which software is installed how the installation process can be repeated automatically.
-Users can therefore quickly change the software environment e.g. update to another R or package version without being "stuck" with an `Docker` image they do not know how to rebuild.
+Users can therefore quickly change the software environment, e.g. update to another R version or given package version.
 
@@
 The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing `Docker`.
-Unfortunately, Docker requires administrative rights to run.
-We recommend [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@R-renv] in cases where no administrative rights can be obtained but can not detail its use in this document.
+Unfortunately, Docker requires administrator rights to run, which may not be available to all researchers.
+We recommend [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@R-renv] in cases where no administrator rights can be obtained but can not detail its use in this document.
+<!-- add a half sentence about what R-env is-->
 Start RStudio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
@@
 
-After you have installed the necessary software we suggest that you set up a secure connection to GitHub:
+After you have installed the necessary software, we suggest that you set up a secure connection to GitHub:
 
@@
 Please consider carefully if you can include data in the repository without violating privacy rights.
-If you can not share your data publicly add it to the `.gitignore` file and share it on request.
+If you are not allowed to share your data publicly, add the data file(s) to the `.gitignore` file and only share them on request.
 
@@
 specific journal requirements.
-Noteworthy is the package output format supplied by the [`papaja`](https://github.com/crsh/papaja) package [@papaja] which produces APA compliant manuscripts [@apa7].
+Of particular note is the package output format supplied by the [`papaja`](https://github.com/crsh/papaja) package [@papaja] which produces APA compliant manuscripts [@apa7].
 Another advantage is that, because `R Markdown` files are in plain text, they are more suitable for version control using `Git`.
@@
 The comparison is made impossible if the authors of the preregistration were intentionally vague.
-Even when the researcher was faithful in each aspect to the preregistration the reader may reach opposite conclusions, because the reader has to compare two different texts which may be worded differently or describe the same thing in varying level of detail.
+Even when the researcher was faithful to the preregistration, the reader may reach opposite conclusions, because the reader has to compare two different texts, which may be worded differently or describe the same thing in varying levels of detail.
 A third limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
@@
 Once the preregistration is submitted and real data have been collected or made available,
-the document can be reproduced with a single command, thus updating the Results to their final version.
-Reproducibility is of utmost importance at this stage since the preregistration must produce valid results at two points in time.
-We can only recommend PAC when the dangers to reproducibility we described above are eliminated and a state of automated reproducibility is reached.
+the document can be reproduced with a single command, thus updating the Results section to the final version.
+Reproducibility is of utmost importance at this stage since the preregistration must produce valid results at two points in time, once before data collection and once after data collection.
+We can only recommend PAC when the dangers to reproducibility we described above are eliminated and a state of automated reproducibility is reached. <!--AB I do not understand this last sentence-->
 
@@
 Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
-One factor that reduces comprehensiveness of such manuscript style but that is absent in standard preregistrations are strict word count limits that some journals impose.
-In such case we suggest using appendix materials to keep the manuscript concise, yet the preregistration comprehensive.
+The comprehensiveness and completeness of published scientific reports is sometimes limited by strict word count limits that some journals impose.
+One may argue that PAC inherit this limitation if the same word count limit applies.
+However, in such case we suggest using reproducible appendix materials to keep the core manuscript concise, yet the preregistration comprehensive and exhaustive.
 
@@
 When they have have outlined the studies design, they should begin writing code.
-We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
-For each study these tasks have to specifically coded.
+If there is no other code structure available, we suggest writing a minimum of three functions for each planned hypothesis, one for the planned analysis, one for simulating the respective data and one for reporting the results.
 We capture the tasks here in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
@@
 
-If simulating data proves to be too challenging, researcher can delay their preregistration till they have gathered data and repurpose the actual data.
-In order to do so researcher must walk a fine but utmost important line.
-Of course researcher must not analyze the data before they write the preregistration.
-Nevertheless, they may obfuscate the real data by shuffling all variables columnwise (independently from each other) or otherwise create synthetic data that allows them to test the function `planned_analysis()` without actually revealing the research outcomes.
-Preregistration after data collection is common when reanalyzing data from another research group [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
+If simulating data proves to be too challenging, researcher can delay their preregistration until they have gathered data and repurpose the empirical data.
+In order to do so researcher must walk a fine but important line.
+Of course, researcher must not analyze the data before they write the preregistration.
+Nevertheless, they may obfuscate the real data (before actually looking at it) by shuffling all variables per each colum (independently from each other) in cross-sectional data <!--AB would not work well in longitudinal data--> or otherwise creating synthetic data that allows them to test the function `planned_analysis()` without actually revealing the observed research outcomes.
+Preregistration after data collection is common for secondary data analysis of data obtained by other research groups [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
 We argue that it is still an eligible preregistration.
-Guidelines for clinical trials also recommend analysis of blinded blinded to test the practicality of the preregistration [@ICH1998].
-Since the data is obfuscated, almost no results are revealed and the researchers can not employ their researcher degrees of freedom effectively to influence the results.
+Guidelines for clinical trials also recommend analysis of blinded data <!--AB I added data here?!--> to test the practicality of the preregistration [@ICH1998].
+Since the data is obfuscated, virtually no results<!--AB do you mean observed data? what are results here?--> are revealed and the researchers do not gain researcher degrees of freedom.
 Shuffling leaves only univariate information, so this procedure is only applicable for bi-or multivariate hypotheses.
@@
 members of group "A" have a greater mean in variable "Z").
-To be clear, any exploration of the data before preregistration renders the preregistration invalid and amounts to grave scientific (albeit sometime unintentional) misconduct.
+To be clear, any exploration of the data before preregistration renders the preregistration invalid and amounts to scientific (albeit sometime unintentional) misconduct.
 
@@
 A pilot study has obvious advantages for study planning, since it lets researcher evaluate the feasibility of many assumptions.
-However, we must warn our readers, while piloting is more traditional than our approach of obfuscating the data before preregistration it may not be innocuous to use the resulting data and care must be taken to avoid errors of sequential testing [@whiteheadBiasMaximumLikelihood1986].
+However, we must warn our readers, while piloting is more traditional than our approach of obfuscating the data before preregistration, it may not be innocuous to use the resulting data and care must be taken to avoid errors of sequential testing [@whiteheadBiasMaximumLikelihood1986].
+<!-- this may not be clear to everyone...
+do you mean that you cannot reuse the pilot data?-->
 
@@
 This example preregistration covers a single study with a single hypothesis.
-To organize studies with multi hypotheses we suggest multiple (possibly numbered in accord with the hypotheses, e.g. 1.2, 2.3 etc.) `planned_analysis()` and `report_analysis()` functions.
-Steps that apply to multiple analyses should be extracted into their own function e.g. `discard_outliers()` or `extract_scores()`.
+To organize studies with multiple hypotheses, we suggest multiple (possibly numbered in accordance with the hypotheses, e.g. 1.2, 2.3 etc.) `planned_analysis()` and `report_analysis()` functions.
+Computational steps that apply to multiple analyses should be extracted into their own function, for example, those that perform data cleaning or extract results from a statistical model in an identical way.
 Preregistrations that cover multiple distinct data sources may employ multiple `simulate_data()` functions.
-These are merely suggestions, researcher who are more comfortable with another programming style will surely find an appropriate system.
-
-Of course it is likely that the preregistered code has to be adapted to the actual data.
+These are merely suggestions and researchers are encouraged to find their own way of how to best organize their analysis code.
+
+Of course, it is likely that the preregistered code has to be adapted to the actual data.
 Changing the preregistration must be discussed.
@@
 
-The interested reader can find the example rendered as a pdf file with actual data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
-
-After the researcher is satisfied with their draft preregistration, they should create an archived and unchangeable version of the project that serves as the preregistration.
+The interested reader can find the example rendered as a PDF file with real data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
+
+After researchers are satisfied with their draft preregistration, they should create an archived and unchangeable version of the project that serves as the preregistration.
 zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides digital object identifiers (DOI) for these archives.
@@
 We recommend to archive the `Git` repository and the `Docker` image with zenodo.org, a non-profit long-term storage for scientific data.
-The necessary steps for archival on zenodo.org are described in at the end of section [Preregistration as Code --- a Tutorial].
+The necessary steps for archival on zenodo.org are described at the end of section [Preregistration as Code --- a Tutorial].
 The second vulnerability is that even if the existence of the `Docker` image and `Git` repository is guaranteed, future researchers still require software to run the image.

### 2021-10-06 manuscript.Rmd
address andreas' comments

@@
   Reproducibility has long been considered integral to the scientific method.
-  A scientific analysis is called reproducible when an independent person obtains the same results from the same data.
-<!--AB something seems to generic; we could generalize this to digital objects, though-->
+  An analysis is called reproducible when an independent person obtains the same results from the same data.
   Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
@@
 Objectivity in this context means "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
-Only a reproducible object can be tested and understood by anyone.
+Only a reproducible result can be tested and understood by anyone.
 Therefore, reproducibility has long been considered integral to empirical research.
@@
 This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (e.g., missing software dependencies or ambiguous documentation of the exact computational steps taken), and there is a dire need for reproducibility standards.
-With the technological advancement, it is now possible to make all digital products related to a research project available so that a research project can be automatically reproduced with minimal effort.<!--AB I don't understand: why is availabbility based on technological advancement?
-aren't there still be legal restrictions?-->
+With the technological advancement, it is now possible to make all digital products related to a research project available in a way that enables automatic reproduction of the research project with minimal effort.
 
@@
 Second, we discuss opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated, but also how research itself is conducted.
-With regard to the first goal, we discuss how to address <!--AB please avoid military wordings--> four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
+With regard to the first goal, we discuss how to address four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
 Then we present a tutorial on how to employ these tools to achieve automated reproducibility.
@@
 This tutorial is aimed at researchers who regularly write code to analyze their data and who are willing to make relevant code, data, and materials available, either publicly or on request.
-Ideally, our target reader already has created a dynamic document at some point in time (e.g. with RMarkdown or Jupyter) and knows the basic ideas behind version control (e.g. with Git). 
+Ideally, our reader already has created a dynamic document at some point in time (e.g. with RMarkdown or Jupyter) and attempted to track a file with version control (e.g. with Git). 
 The R package `repro` supports researchers in setting up the required software and in adopting this workflow.
@@
 A workflow that includes, e.g. renaming variable names by hand in a graphical spreadsheet application is therefore incompatible with automated reproducibility.
-Another property of a computational recipe is that the same inputs <!--AB of what? I deleted 'of'--> should always result in the same outputs.
+Another property of a computational recipe is that the same inputs should always result in the same outputs.
 Fo most computer code (given the same software is used) this property is fulfilled, however, one noteworthy exception are pseudo random numbers.
-Whenever random numbers are used in a computation, this computation is only reproducible when the samerandom numbers are used.
+Whenever random numbers are used in a computation, this computation is only reproducible when the same random numbers are used.
 To ensure identical random numbers, users may fix the state of the random number generate with a so called seed (e.g.
-in R with `set.seed()`) but they also need to guarantee that the pseudo random number generator is unchanged (see @Peikert2019).
-<!--AB please check if citation is rendered correctly-->
+in R with `set.seed()`) but they also need to guarantee that the pseudo random number generator is unchanged [see @Peikert2019].
 
@@
 The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing `Docker`.
+
 Unfortunately, Docker requires administrator rights to run, which may not be available to all researchers.
 We recommend [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@R-renv] in cases where no administrator rights can be obtained but can not detail its use in this document.
-<!-- add a half sentence about what R-env is-->
+`renv` tracks which R package is installed from which source in which version in a so called [`lockfile`](https://rstudio.github.io/renv/articles/lockfile.html).
+This lockfile is than used to reinstall the same packages on other computers or later in time.
+For a more thorogh discussion see @vanlissa2020worcs.
+
 Start RStudio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
@@
 Reproducibility is of utmost importance at this stage since the preregistration must produce valid results at two points in time, once before data collection and once after data collection.
-We can only recommend PAC when the dangers to reproducibility we described above are eliminated and a state of automated reproducibility is reached. <!--AB I do not understand this last sentence-->
+As outlined before, reproducibility builds upon four pillars (Version control, Dynamic document generation, Dependency tracking, Software management)
+We can therefore only recommend PAC when the dangers to reproducibility we described above are eliminated.
 
@@
 Of course, researcher must not analyze the data before they write the preregistration.
-Nevertheless, they may obfuscate the real data (before actually looking at it) by shuffling all variables per each colum (independently from each other) in cross-sectional data <!--AB would not work well in longitudinal data--> or otherwise creating synthetic data that allows them to test the function `planned_analysis()` without actually revealing the observed research outcomes.
+Nevertheless, they may obfuscate the real data (before actually looking at it) by shuffling all variables per each colum (independently from each other) in cross-sectional data or otherwise creating synthetic data that allows them to test the function `planned_analysis()` without actually revealing the observed research outcomes.
 Preregistration after data collection is common for secondary data analysis of data obtained by other research groups [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
 We argue that it is still an eligible preregistration.
-Guidelines for clinical trials also recommend analysis of blinded data <!--AB I added data here?!--> to test the practicality of the preregistration [@ICH1998].
-Since the data is obfuscated, virtually no results<!--AB do you mean observed data? what are results here?--> are revealed and the researchers do not gain researcher degrees of freedom.
+Guidelines for clinical trials also recommend analysis of blinded data  to test the practicality of the preregistration [@ICH1998].
+Since the data is obfuscated, virtually no information are revealed to the researchers and they can not employ their researcher degrees of freedom effectively to influence the results.
 Shuffling leaves only univariate information, so this procedure is only applicable for bi-or multivariate hypotheses.
@@
 members of group "A" have a greater mean in variable "Z").
-To be clear, any exploration of the data before preregistration renders the preregistration invalid and amounts to scientific (albeit sometime unintentional) misconduct.
+To be clear, any exploration of the data before preregistration (PAC or otherwise) renders the preregistration invalid and amounts to scientific (albeit sometime unintentional) misconduct.
 
@@
 A pilot study has obvious advantages for study planning, since it lets researcher evaluate the feasibility of many assumptions.
-However, we must warn our readers, while piloting is more traditional than our approach of obfuscating the data before preregistration, it may not be innocuous to use the resulting data and care must be taken to avoid errors of sequential testing [@whiteheadBiasMaximumLikelihood1986].
-<!-- this may not be clear to everyone...
-do you mean that you cannot reuse the pilot data?-->
+However, we must warn our readers, while piloting is more traditional than our approach of obfuscating the data before preregistration, it is not innocuous to use the data resulting from the pilot.
 

### 2021-10-07 manuscript.Rmd

@@
 
+We have to distinguish two kinds of limitations.
+First, some limitations are by design and are unlikely to be changed soon.
+Our workflow does inherit these from the software it relies on and the fundamental design principles these share with the workflow and `repro`.
+Second, there are other limitations that we hope development effort of the open-source community and, concerning `repro`, us will ease.
+
+As a principle, this workflow is incompatible with steps that cannot be automated.
+This principle is at odds with some ingrained habits of researchers to mix and match manual and automated steps in data analysis.
+To allow for automation, many researchers will have to search for alternative software.  
+The automation friendly software we present here has several technical but critical limitations.
+`Git` can track any files (files bigger than 1GB only with extensions) but can only compare text files (with endings like, `.txt`, `.csv`, `.R`,  or `.Rmd`), not binary files (with endings like `.docx`, `.exe` or `.zip`).
+With `R Markdown`, tables and graphics are dynamically generated from code but cannot be crafted by hand.
+`Make` can orchestrate any programmable software but fails to automate software that will only obey a user clicking at the right buttons.
+`Docker` can ship software that runs on Linux and can be automatically installed, but that requirement precludes many commercial or closed source software.
+
+This move away from software that has served researchers well for decades is understandably difficult and presents us with a conundrum.
+On the one hand, we firmly believe that automated reproducibility makes research more productive and collaboration easier.
+But, on the other hand, we expect researchers to invest considerable time in learning new tools and than must persuade their collaborators to do the same.
+Three arguments reconcile this apparent paradox.
+First, this change will not happen all at once.
+Automated reproducibility is an ideal that we believe has many advantages, but it is not an all or nothing decision.
+Researchers can pick up one skill at a time and then help their fellow collaborators do the same.
+Second, the upfront investment is one time (and efforts such as `repro` are underway to reduce it) and will pay dividends over many research projects.
+Third, the move towards open software for research offers several benefits beyond enabling automated reproducibility [@schaffnerFutureScientificJournals1994; @fitzgeraldTransformationOpenSource2006; @chaldecottHistoryScientificTechnical1965; @sonnenburgNeedOpenSource2007].
+
 `repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.

### 2021-10-07 manuscript.Rmd
grammerly on introduction

@@
 Scientists increasingly strive to make research data, materials, and analysis code openly available.
-Sharing these types of digital research products has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
-To some extent these potential benefits are contingent on whether these digital research products are reproducible. 
+Sharing these digital research products can increase scientific efficiency by enabling researchers to learn from each other, reuse materials, and increase scientific reliability by facilitating the review and replication of published research results.
+To some extent, these potential benefits are contingent on whether these digital research products are reproducible. 
 Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
@@
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
-This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges (e.g., missing software dependencies or ambiguous documentation of the exact computational steps taken), and there is a dire need for reproducibility standards.
-With the technological advancement, it is now possible to make all digital products related to a research project available in a way that enables automatic reproduction of the research project with minimal effort.
-
-With this paper we pursue two aims.
-First, we want to introduce researchers to a notion of automated reproducibility that requires, apart from the setup of software, no manual steps. 
-Second, we discuss opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated, but also how research itself is conducted.
+This is because making a research project with all code, data, and materials reproducible comes with various challenges (e.g., missing software dependencies or ambiguous documentation of the exact computational steps taken), and there is a dire need for reproducibility standards.
+With technological advancement, it is now possible to make all digital products related to a research project available in a manner that enables automatic reproduction of the research project with minimal effort.
+
+With this paper, we pursue two aims.
+First, we want to introduce researchers to a notion of automated reproducibility that requires, apart from the setup of the software, no manual steps. 
+Second, we discuss opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated but also how research itself is conducted.
 With regard to the first goal, we discuss how to address four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
@@
 We present automated reproducibility as best practice, a goal that is not always archived because of limited resources, technical restrictions or practical considerations but is worth striving for.
-If, due to whatever reason, some parts of the workflow proof to be unfeasible, we encourage the reader to find their own suitable solutions and to concentrate on the feasible parts.
-
-In pursuit of the second aim, we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon an implementation of this reproducible workflow, the so-called *preregistration as code* (PAC).
+If, due to whatever reason, some parts of the workflow prove to be unfeasible, we encourage the reader to find their own suitable solutions and to concentrate on the feasible parts.
+
+In pursuit of the second aim, we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon implementing this reproducible workflow, the so-called *preregistration as code* (PAC).
 PAC involves preregistering the intended analysis code and the major part of the final scientific report as a dynamic document, including typical sections like Introduction, Methods, and Results.
-The resulting dynamic document closely resembles the final manuscript, but temporarily uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
+The resulting dynamic document closely resembles the final manuscript but temporarily uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
 With these placeholders, authors preregister the exact presentation of their results without having seen real data yet.
@@
 This was illustrated by an attempt to reproduce results from open materials in the journal *Cognition* [@hardwicke2018]. 
-Out of 35 published articles with open code and data, results of 22 articles could be reproduced but in 11 of these cases, further assistance from the original authors was required.
+Out of 35 published articles with open code and data, results of 22 articles could be reproduced, but in 11 of these cases, further assistance from the original authors was required.
 For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
@@
 The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
-These failed attempts to reproduce highlight the need for widely accepted reproducibility standards, because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
+These failed attempts to reproduce highlight the need for widely accepted reproducibility standards because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
 If digital research products are available but not reproducible, their added value is limited.
@@
 
-1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during writing of a paper because a bug was removed at some point in time.
+1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during the writing of a paper because a bug was removed at some point in time.
 It may then be unclear which version of code and data was used to produce some reported set of results.
@@
 Version control allows a clear link between which results were generated by which version of code and data.
-This is important for reproducibility because only if it is obvious which version of data and code produced which results they can be said to be reproducible.
+This is important for reproducibility because only if it is obvious which version of data and code produced which results, those results can be said to be reproducible.
 We recommend using *Git* for version control, because of its widespread adoption in the R community.
@@
 This way, `Markdown` does not need any specialized software for editing.
-It is highly fool-proof [unlike, for example, `LaTeX` @lamportLATEXDocumentPreparation1994], works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using `Markdown` for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas, to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
+It is highly fool-proof [unlike, for example, `LaTeX` @lamportLATEXDocumentPreparation1994], works well with version control systems and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+We suggest using `Markdown` for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
 `R Markdown` extends regular `Markdown` by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a `Markdown` document.
@@
 
-One would hope that sharing a dynamic document would allow individuals (including one's future self) to simply download and reproduce a given analysis.
+One would hope that sharing a dynamic document would allow individuals (including one's future self) to download and reproduce a given analysis simply.
 While version control and dynamic document generation are becoming more common, we have argued that two more components are required and that each component alone is unlikely to guarantee reproducibility [@Peikert2019; @vanlissa2020worcs].
@@
 A computational recipe describes how inputs are processed to deterministically create a specific output in a way that is automatically executable.
-The concept of computational recipes is central to our understanding of reproducibility because it enables a unified way to automatically reproduce a project.
+The concept of computational recipes is central to our understanding of reproducibility because it enables a unified way to reproduce a project automatically.
 Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
@@
 Another property of a computational recipe is that the same inputs should always result in the same outputs.
-Fo most computer code (given the same software is used) this property is fulfilled, however, one noteworthy exception are pseudo random numbers.
+For most computer code (given the same software is used), this property is fulfilled.
+However, one noteworthy exception is pseudo-random numbers.
 Whenever random numbers are used in a computation, this computation is only reproducible when the same random numbers are used.
-To ensure identical random numbers, users may fix the state of the random number generate with a so called seed (e.g.
-in R with `set.seed()`) but they also need to guarantee that the pseudo random number generator is unchanged [see @Peikert2019].
+To ensure identical random numbers, users may fix the state of the random number generated with a so-called seed (e.g.
+in R with `set.seed()`), but they also need to guarantee that the pseudo-random number generator is unchanged [see @Peikert2019].
 
@@
 Here, if one were to build the final `manuscript.pdf` by rendering the `R Markdown` with the command shown in l.2, `Make` would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
-This ensures that the simulated data are present before the manuscript is built, and that the simulation is rerun and the manuscript is rebuilt if the simulation code was changed.
+This ensures that the simulated data are present before the manuscript is built and that the simulation is re-run, and the manuscript is rebuilt if the simulation code was changed.
 `Make` therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
-Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] does not contain this element of dependency tracking explicitly but its strict structure of only containing one definite `R Markdown` still makes dependencies between files unambiguous.
+Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] does not explicitly contain this dependency tracking element, but its strict structure of only containing one definite `R Markdown` still makes dependencies between files unambiguous.
 
@@
 
-A version controlled dynamic document with dependency tracking may still rely on external software.
-Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
+A version-controlled dynamic document with dependency tracking may still rely on external software.
+Troubleshooting issues specific to a particular programming language or dependent tool typically require considerable expertise and threatens reproducibility.
 One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via *software management*.
@@
 
-Note that the software environment of the Docker image is completely separate from software installed on your computer.
-This separation is excellent for reproducibility, but takes some getting used to.
+Note that the software environment of the Docker image is completely separate from the software installed on your computer.
+This separation is excellent for reproducibility but takes some getting used to.
 For example, it is important to realize that software available on your local computer will *not* be accessible within the confines of the `Docker` image.
@@
 There are two ways to build a `Docker` image.
-First users can manually install whatever software they like from within the virtual environment.
-Such manually build environment can still be ported to all computers which support `Docker`.
+First, users can manually install whatever software they like from within the virtual environment.
+Such a manually build environment can still be ported to all computers which support `Docker`.
 We, however, prefer the second way of building images automatically from a textual description called `Dockerfile`.
-Because the `Dockerfile` clearly describes which software is installed how the installation process can be repeated automatically.
-Users can therefore quickly change the software environment, e.g. update to another R version or given package version.
+Because the `Dockerfile` clearly describes which software is installed how, the installation process can be repeated automatically.
+Users can, therefore, quickly change the software environment, e.g. update to another R version or given package version.
 
@@
 The image for this project includes Ubuntu, R, RStudio, LaTeX as well as a variety of R packages like tidyverse [@tidyverse] and all its dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
-A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
+A second strategy is to save a container recipe, a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment instead of the software environment itself.
 `Dockerfiles` are tiny (the [`Dockerfile`](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
 However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
-For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the `Dockerfile`.
-A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
-Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] relies for software management on the R package `renv`[@R-renv] which is more lightweight and easier to use than `Docker` on the one hand, but not as comprehensive on the other hand because it only snapshots the R packages instead of all software used.
+For proper archival, we, therefore, recommend storing a complete image of the software environment, in addition to the `Dockerfile`.
+A more comprehensive overview of the use of containerization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
+Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] relies for software management on the R package `renv`[@R-renv], which is more lightweight and easier to use than `Docker` on the one hand, but not as comprehensive on the other hand because it only snapshots the R packages instead of all software used.
 
@@
 Users do not have to have `repro` installed to reproduce a project;
-in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
-In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.
+in fact, they do not even need to have R installed because the entire project can be rebuilt inside a container with R installed.
+In the remainder, we will walk you through creating a reproducible research project with the package `repro`.
 

### 2021-10-07 manuscript.Rmd
grammerly on tutorial

@@
 One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be quite time-intensive.
-To lower the threshold to adoption, the R-package `repro` introduces helper functions that wraps and hides the complexity of the necessary more complicated (and powerful) tools.
-The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user is supposed to do next.
+To lower the threshold to adoption, the R-package `repro` introduces helper functions that wrap and hide the complexity of the necessary more complicated (and powerful) tools.
+The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer does and suggest what the user is supposed to do next.
 The `repro` package extends the [`usethis`](https://usethis.r-lib.org) [@usethis] package.
@@
 We assume that you have installed R [@R-base, version `r with(R.version, paste0(major, ".", minor))`] and RStudio [@rstudio, version 1.4] already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert)[@R-repro].
-In case you either have not installed R and RStudio or are unsure if they are up to date you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers how to install all software necessary for this tutorial in three steps.
+In case you either have not installed R and RStudio or are unsure if they are up to date, you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers how to install all software necessary for this tutorial in three steps.
 The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing `Docker`.
@@
 We recommend [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@R-renv] in cases where no administrator rights can be obtained but can not detail its use in this document.
-`renv` tracks which R package is installed from which source in which version in a so called [`lockfile`](https://rstudio.github.io/renv/articles/lockfile.html).
-This lockfile is than used to reinstall the same packages on other computers or later in time.
-For a more thorogh discussion see @vanlissa2020worcs.
+`renv` tracks which R package is installed from which source in which version in a so-called [`lockfile`](https://rstudio.github.io/renv/articles/lockfile.html).
+This lockfile is then used to reinstall the same packages on other computers or later in time.
+For a more thorough discussion, see @vanlissa2020worcs.
 
@@
 
-An example of the assistance that `repro` provides are the check functions.
-These verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
+An example of the assistance that `repro` provides is the check functions.
+These verify that you indeed have installed and set up the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
 
@@
 These messages are adjusted to your specific operating system and installed dependencies.
-Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
+Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions and to prepare your system for creating reproducible projects.
 
@@
 `Git` tracks any files that you manually add to the repository.
-To make sure you do not accidentally add files that you do not wish to share (e.g., privacy sensitive data),
+To make sure you do not accidentally add files that you do not wish to share (e.g., privacy-sensitive data),
 you can list specific files that you do not want to track in the `.gitignore` file.
@@
 Click on the empty box next to the file you want to stage.
-A check mark appears to indicate that the file is staged.
+A checkmark appears to indicate that the file is staged.
 After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
@@
 While you can choose any `Git` service (like GitLab or BitBucket), we will use GitHub in this tutorial.
-Before you upload your project to `Git`, you need to decide if you would like the project publically accessable (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
+Before you upload your project to `Git`, you need to decide if you would like the project publically accessible (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
 To upload the project publically to GitHub use:
@@
 Depending on your computer configuration, this may ask to set up a secure connection to GitHub.
-In this case first follow the suggestions shown on the R console.
+In this case, first, follow the suggestions shown on the R console.
 
@@
 
-Now that you have created a version controlled project, we will proceed with dynamic document generation.
+Now that you have created a version-controlled project, we will proceed with dynamic document generation.
 A dynamic document has three elements:
@@
 specific journal requirements.
-Of particular note is the package output format supplied by the [`papaja`](https://github.com/crsh/papaja) package [@papaja] which produces APA compliant manuscripts [@apa7].
+Of particular note is the package output format supplied by the [`papaja`](https://github.com/crsh/papaja) package [@papaja], which produces APA compliant manuscripts [@apa7].
 Another advantage is that, because `R Markdown` files are in plain text, they are more suitable for version control using `Git`.
-Some users might find it easier to activate the "Visual Editor" of RStudio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the `R Markdown` document) which features more graphical elements like a traditional word processor but still creates an `R Markdown` underneath with all of its flexibility.
-The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
-and easing the generation of citations and references to Tables and Figures.
-
-Now that you are familiar with the basics of `Markdown` formatting, we turn our attention to including code into the text.
+Some users might find it easier to activate the "Visual Editor" of RStudio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the `R Markdown` document), which features more graphical elements like a traditional word processor but still creates an `R Markdown` underneath with all of its flexibility.
+The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a new line, which eases tracking changes across versions) and easing the generation of citations and references to Tables and Figures.
+
+Now that you are familiar with `Markdown` formatting basics, we turn our attention to including code and its results in the text.
 Code is separated by three backticks (and the programming language in curly braces) like here:
@@
 The results of code enclosed in such backticks will be dynamically inserted into the document (depending on specific settings).
-This means that, whenever you render the `R Markdown` to its intended output format, the code will be executed and the results updated.
-The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
+This means that whenever you render the `R Markdown` to its intended output format, the code will be executed and the results updated.
+The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g. on a preprint server.
 
@@
 This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
-To avoid ambiguity, we therefore recommend referencing the identifier of the `Git` commit at the time of rendering in the static document.
+To avoid ambiguity, we, therefore, recommend referencing the identifier of the `Git` commit at the time of rendering in the static document.
 Simply put, a static document should link to the version of the code that was used to create it.
@@
 
-This information makes explicit what dependencies (in form of files and R packages) a project relies on .
+This information clarifies what dependencies (in the form of files and R packages) a project relies on.
 `repro` uses this information to construct a `Makefile` for the dependencies on other files and a `Dockerfile` that includes all required packages.
@@
 
-It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new `R Markdown` file to the project, to keep the `Makefile` and `Dockerfile` up to date.
+It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new `R Markdown` file to the project to keep the `Makefile` and `Dockerfile` up to date.
 There is no harm in running it too often.

### 2021-10-07 manuscript.Rmd
grammerly on advanced sectction

@@
 This section is only for advanced users who want to overcome some limitations of `repro`.
-If you read this paper the first time you probably want to skip this section and continue reading at the section "Preregistration as Code".
+If you read this paper the first time, you probably want to skip this section and continue reading at the section "Preregistration as Code".
 As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.
@@
 Users who ask themselves either "How can I install software dependencies outside of R in the `Docker` image?" or "How can I express complex dependencies between files, e.g.
-hundreds of data files are preprocessed and combined?") need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
-Other users may safely skip this section, or return to it if they encounter such challenges.
+hundreds of data files are preprocessed and combined?") need to be aware of these restrictions and require a deeper understanding of the inner workings of `repro`.
+Other users may safely skip this section or return to it if they encounter such challenges.
 
@@
 The first limitation means that users must rely on software that is either already provided by the base `Dockerimage` "rocker/verse" or the R packages they list in the metadata.
-Other software that users might need like other programming languages, not already installed LaTeX packages etc. have to be added manually.
+Other software that users might need, like other programming languages, not already installed LaTeX packages, etc., must be added manually.
 We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g.
@@
 These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
-Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D, therefore leaving out important information or express the dependencies directly within the `Makefile`.
-
-The following section explains how to overcome these limitations while still relying on the automations afforded by `repro`.
+Users, therefore, have to either "flatten" the dependency structure by simply stating that A depends on B, C and D, therefore leaving out important information or express the dependencies directly within the `Makefile`.
+
+The following section explains how to overcome these limitations while still relying on the automation afforded by `repro`.
 Lifting these restrictions requires the user to interact more directly with `Make` or `Docker`.
-To satisfy more complicated requirements users need to understand how repro utilizes `Make` and `Docker` internally.
+Users need to understand how repro utilizes `Make` and `Docker` internally to satisfy more complicated requirements.
 
 Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two processing steps.
-First it recreates the virtual software environment (`Docker`) and then it executes computational recipes in the virtual software environment (`Make`).
+First, it recreates the virtual software environment (`Docker`), and then it executes computational recipes in the virtual software environment (`Make`).
 The first step is done by the command `make docker`.
@@
 The recipe for this target builds an image from the `Dockerfile` of the repository.
-The `&&` concatenates both commands and only runs the second command if the first was successful.
+The `&&` concatenates both commands and only runs the second command if the first is successful.
 Therefore, the computational steps are only executed when the software environment is set up.
@@
 Users can choose which R version they want to have inside the container by changing the version number in line 1 to the desired R version number.
-By default the R version corresponds to the locally installed version on which `repro::automate()` was called the first time.
-The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed, by default this date is set to the date on which `repro::automate()` was called the first time.
+By default, the R version corresponds to the locally installed version on which `repro::automate()` was called the first time.
+The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed.
+By default, this date is set to the date on which `repro::automate()` was called the first time. 
 This way, the call to the automate function virtually freezes the R environment to the state it was called the first time inside the container. 
@@
 Finally, there is one manually edited `Dockerfile`: `.repro/Dockerfile_manual`.
-It is blank by default, and can be used to add further dependencies outside of R, like system libraries or external software.
+It is blank by default and can be used to add further dependencies outside of R, like system libraries or external software.
 
@@
 
-With this approach, users of `repro` can not only build complex software environments but they can also implement complex file dependencies.
+With this approach, users of `repro` can build complex software environments and implement complex file dependencies.
 The standard `repro` metadata only makes sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
@@
 the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the `Makefile`.
-The `Makefile` that `repro` creates is only a template and you are free to change it.
-However, make sure to never remove the following two lines:
+The `Makefile` that `repro` creates is only a template, and you are free to change it.
+However, make sure never to remove the following two lines:
 
@@
 This file should not be altered manually.
-If you are not satisfied with the automatically generated target simply provide an alternative target in the main `Makefile`.
+If you are not satisfied with the automatically generated target, simply provide an alternative target in the main `Makefile`.
 Targets in the main `Makefile` take precedent.
@@
 Additionally, it provides facilities to execute commands within the container.
-If you write a computational recipe for a target it will be by default evaluated using the locally installed software.
+If you write a computational recipe for a target, it will be by default evaluated using the locally installed software.
 To instead evaluate commands inside the `Docker` image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first `Make` example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
@@
 
-To summarise, `repro` provides researchers with an easy to use solution for dependency tracking (in form of `Make`) and software management (using `Docker`) without the necessity to learn both tools.
-Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automations.
+To summarise, `repro` provides researchers with an easy to use solution for dependency tracking (in the form of `Make`) and software management (using `Docker`) without the necessity to learn both tools.
+Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automation.
 

### 2021-10-07 manuscript.Rmd
grammerly on discussion

@@
 Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
-The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
+The current best practice to ensure that the preregistrations are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
 These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
-However, templates cannot remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
-This ambiguity causes a second problem, that is comparing the preregistration and the resulting publication to decide if and how researcher deviated from the preregistration.
+However, templates cannot remove researchers degrees of freedom completely because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
+This ambiguity causes a second problem, that is, comparing the preregistration and the resulting publication to decide if and how researchers deviated from the preregistration.
 This task is difficult because it impossible to decide without additional information if the analysis actually was carried out differently or just described differently.
 The comparison is made impossible if the authors of the preregistration were intentionally vague.
-Even when the researcher was faithful to the preregistration, the reader may reach opposite conclusions, because the reader has to compare two different texts, which may be worded differently or describe the same thing in varying levels of detail.
+Even when the researcher was faithful to the preregistration, the reader may reach opposite conclusions because the reader has to compare two different texts, which may be worded differently or describe the same thing in varying levels of detail.
 A third limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
@@
 they sometimes inexplicably fail to reproduce their own results.
-For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
+For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistics than they reported initially (see Footnote 7).
 A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
-For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables, or more parameters than observed data.
-Conversely, a model may be mis-specified in a major way; for example by omitting direct effects when the research question is about mediation, thus leading to a model with unacceptable fit.
+For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables or more parameters than observed data.
+Conversely, a model may be misspecified in a major way; for example, by omitting direct effects when the research question is about mediation, thus leading to a model with an unacceptable fit.
 Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.
@@
 In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did include some form of script as part of their preregistration.
-But although the notion of preregistering code has been around for a while [cf. @steegenMeasuringCrowdAgain2014],
+Though the notion of preregistering code has been around for a while [cf. @steegenMeasuringCrowdAgain2014],
 it has not gained much traction---perhaps because, until now, this has constituted an extra non-standard step in the research process.
@@
 First, the PAC is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
-Relatedly, writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
+Relatedly, writing a preregistration in the form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
 By contrast, filling out a preregistration template is a novel type of scientific writing for many.
@@
 Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
-PAC therefore minimizes researchers degrees of freedom more effectively than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
-Fourth, despite its rigor, PAC does accommodate data-dependent decisions if these can be formulated as code.
+PAC, therefore, minimizes researchers degrees of freedom more effectively than standard preregistration [@bakkerEnsuringQualitySpecificity2020].
+Fourth, despite its rigour, PAC does accommodate data-dependent decisions if these can be formulated as code.
 Researchers can, for example, formulate conditions (e.g., in the form of if-else-blocks) under which they prefer one analysis type over the other.
 For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
-Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
+Another example of data-dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
 Decisions that do not lend themselves to formulation in code, e.g.
@@
 We would like to note that in the PAC paradigm, the direct comparison of the preregistration and the final publication is particularly easy to accomplish.
-Because there is a single unambiguous reference point, that is the dynamic document, including code, under version control, changes can be inspected by authors and readers alike.
+Because there is a single unambiguous reference point, that is, the dynamic document, including code, under version control, changes can be inspected by authors and readers alike.
 Authors should at least review all changes made to the preregistered plan to summarize and justify them for their reader, e.g. in the discussion of the final manuscript [^summarize-changes].
-Such retrospective summary already serves to put changes into context, however, a change log that is kept continually and where changes are explained when they are made, provides more immediate and detailed information.
-Each entry in the change log should explain the reasoning of the changes and link to the changes made by including the the commit id that applied the changes.
+Such retrospective summary already serves to put changes into context.
+However, a changelog that is kept continually and where changes are explained when they are made provides more immediate and detailed information.
+Each entry in the changelog should explain the reasoning of the changes and link to the changes made by including the commit id that applied the changes.
 The interested reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.
@@
 When planning a study and hence when writing a PAC, researchers ought to strive to meet three goals:
-First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
+First, the plan should be comprehensive.
+That is, researchers need to state the research question and describe the study design and analysis in detail.
 Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
@@
 The comprehensiveness and completeness of published scientific reports is sometimes limited by strict word count limits that some journals impose.
-One may argue that PAC inherit this limitation if the same word count limit applies.
-However, in such case we suggest using reproducible appendix materials to keep the core manuscript concise, yet the preregistration comprehensive and exhaustive.
+One may argue that PAC inherits this limitation if the same word count limit applies.
+However, in such case we suggest using reproducible appendix materials to keep the core manuscript concise yet the preregistration comprehensive and exhaustive.
 
@@
 Simulating data offers various benefits in the planning phase of a study.
-It forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
+It forces the researcher to think about and address data analytic problems on a conceptual level before they make an effort to acquire data [@axelrodAdvancingArtSimulation1997].
 It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
@@
 a registered report [an article format in which the preregistration is reviewed and the article accepted "in principle", @nosekRegisteredReports2014; @chambersWhatNextRegistered2019].
-In such registered report peer reviewing a manuscript styled PAC might be more familiar to the reviewers that ensure a comprehensive study plan.
+In such registered report peer-reviewing a manuscript styled PAC might be more familiar to the reviewers that ensure a comprehensive study plan.
 
@@
 it would be unreasonable to expect this analysis to work under realistic conditions.
-Because the simulation describes the researchers expectations about real world conditions,
+Because the simulation describes the researchers' expectations about real-world conditions,
 it is the best way to assess the effectiveness of the analysis without actually gathering the data.
 A simulation of data requires the researchers to translate their expectations about the underlying process into concrete computational terms, e.g.
-assumptions about which distribution does the dependent variable follow, or how do the independent variables influence each other and the dependent variable.
+assumptions about which distribution does the dependent variable follow or how do the independent variables influence each other and the dependent variable.
 If a researcher experiences difficulty in specifying such expectations, that may be an indication that further literature research, or exploratory pilot studies, are required.
@@
 In the simplest case, they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
-@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not increase the planned sample size on avaerage but could help to invest resources more efficiently.
+@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not increase the planned sample size on average but could help to invest resources more efficiently.
 Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
@@
 To write a PAC, researchers should start by drafting a manuscript in the form of an `R Markdown`.
-When they have have outlined the studies design, they should begin writing code.
+When they have outlined the studies design, they should begin writing code.
 If there is no other code structure available, we suggest writing a minimum of three functions for each planned hypothesis, one for the planned analysis, one for simulating the respective data and one for reporting the results.
-We capture the tasks here in the form of explicit R-functions but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
-The functions shown here are empty (they do not contain executable code) and the interested reader may find function filled with some life in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.
+We capture the tasks here in the form of explicit R-functions, but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
+The functions shown here are empty (they do not contain executable code), and the interested reader may find function filled with some life in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.
 
@@
 
-It is difficult to write code when it is not clear how exactly the input should look like, however, at the time of preregistration, researchers should remain insulated from real data.
-We therefore simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
-Using the simulated dataset as placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
+It is difficult to write code when it is not clear how exactly the input should look like.
+However, at the time of preregistration, researchers should remain insulated from real data.
+We, therefore, simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
+Using the simulated dataset as a placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
 
@@
 If simulating data proves to be too challenging, researcher can delay their preregistration until they have gathered data and repurpose the empirical data.
-In order to do so researcher must walk a fine but important line.
-Of course, researcher must not analyze the data before they write the preregistration.
-Nevertheless, they may obfuscate the real data (before actually looking at it) by shuffling all variables per each colum (independently from each other) in cross-sectional data or otherwise creating synthetic data that allows them to test the function `planned_analysis()` without actually revealing the observed research outcomes.
+In order to do so the researcher must walk a fine but important line.
+Of course, researchers must not analyze the data before they write the preregistration.
+Nevertheless, they may obfuscate the real data (before actually looking at it) by shuffling all variables per each column (independently from each other) in cross-sectional data or otherwise creating synthetic data that allows them to test the function `planned_analysis()` without actually revealing the observed research outcomes.
 Preregistration after data collection is common for secondary data analysis of data obtained by other research groups [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
@@
 Guidelines for clinical trials also recommend analysis of blinded data  to test the practicality of the preregistration [@ICH1998].
-Since the data is obfuscated, virtually no information are revealed to the researchers and they can not employ their researcher degrees of freedom effectively to influence the results.
+Since the data is obfuscated, virtually no information are revealed to the researchers, and they can not employ their researcher degrees of freedom effectively to influence the results.
 Shuffling leaves only univariate information, so this procedure is only applicable for bi-or multivariate hypotheses.
-Researcher still can access the information about means or proportions (e.g.
+The researcher still can access the information about means or proportions (e.g.
 how many participants belonging to group "A" are in the dataset) but are unbiased for all relations between variables (e.g.
@@
 Another approach that can replace but also complement simulated data is to conduct a pilot study [@thabaneTutorialPilotStudies2010] and use the resulting data to preregister.
-A pilot study has obvious advantages for study planning, since it lets researcher evaluate the feasibility of many assumptions.
+A pilot study has obvious advantages for study planning, since it lets the researcher evaluate the feasibility of many assumptions.
 However, we must warn our readers, while piloting is more traditional than our approach of obfuscating the data before preregistration, it is not innocuous to use the data resulting from the pilot.
@@
 The only difference is that PAC demands a comprehensive and complete analysis plan.
-With traditional preregistration, researchers may omit details and get away with decision that are made only after having seen the data, thus, potentially biasing statistical results.
+With traditional preregistration, researchers may omit details and get away with decisions that are made only after having seen the data, thus, potentially biasing statistical results.
 Some researchers may not feel ready to decide every detail of the analysis plan before data collection.
-In such case, we recommend to make decisions based on the best available knowledge and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in the section [Deviating from the preregistration and exploration] of the final manuscript.
+In such case, we recommend making decisions based on the best available knowledge, and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in the section [Deviating from the preregistration and exploration] of the final manuscript.
 
@@
 
-Again, we propose a preregistration format that closely resembles a classic journal article, but uses simulated data and dynamic document generation to create a document that starts out as a preregistration and eventually becomes the final report.
+Again, we propose a preregistration format that closely resembles a classic journal article but uses simulated data and dynamic document generation to create a document that starts out as a preregistration and eventually becomes the final report.
 The complete preregistration source is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
@@
 
-[^caution-mww]: The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
+[^caution-mww]: The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employed a rank based alternative in the New England Journal of Medicine in 2005].
 The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
@@
 One example would be using the function `wilcox.test` instead of the combinations of the functions `rank` and `t.test`.
-Either of them are valid implementations of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
+Either of them is a valid implementation of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
 In contrast, the second applies Welch's correction by default and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
@@
 
-We implemented a mechanism that only uses simulated data when the actual data is not yet available (in this example if the file `data/data.csv` does not exist) for the results section.
+We implemented a mechanism that only uses simulated data when the actual data is not yet available (in this example, if the file `data/data.csv` does not exist) for the results section.
 This mechanism also warns readers if these results are based on simulated data.
-The warning is colored red to avoid any confusion between mock and actual results.
+The warning is coloured red to avoid any confusion between mock and actual results.
 As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study.
@@
 Preregistrations that cover multiple distinct data sources may employ multiple `simulate_data()` functions.
-These are merely suggestions and researchers are encouraged to find their own way of how to best organize their analysis code.
+These are merely suggestions, and researchers are encouraged to find their own way of how to best organize their analysis code.
 
@@
 After researchers are satisfied with their draft preregistration, they should create an archived and unchangeable version of the project that serves as the preregistration.
-zenodo.org [@zenodo] is a publicly funded service provider that archives digital artifacts for research and provides digital object identifiers (DOI) for these archives.
+zenodo.org [@zenodo] is a publicly funded service provider that archives digital artefacts for research and provides digital object identifiers (DOI) for these archives.
 While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
-Please note that you can only link public GitHub repositories to zenodo.
+Please note that you can only link public GitHub repositories to zenodo.org.
 You may log into zenodo.org through your GitHub account.
@@
 
-After you have linked a GitHub repository you trigger the archivation by creating a GitHub release.
+After you have linked a GitHub repository, you trigger the archivation by creating a GitHub release.
 To create GitHub release, navigate to GitHub:
@@
 
-Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
+Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits nicely together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
 Automated research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
@@
 
-This paper demonstrated how the R package `repro` supports researchers in creating reproducible research projects including reproducible manuscripts.
+This paper demonstrated how the R package `repro` supports researchers in creating reproducible research projects, including reproducible manuscripts.
 These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
@@
 We first demonstrated how to create a reproducible research project.
-Then, we illustrated how such a project can be reproduced - either by the original author and/or collaborators, or by a third party.
+Then, we illustrated how such a project could be reproduced---either by the original author and/or collaborators or by a third party.
 We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations,
@@
 Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
-To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything there is a single "source of truth"---dynamic manuscript including code---under version control.
+To editors, reviewers, and readers, it is obvious how the researchers deviated from the preregistration because everything there is a single "source of truth"---dynamic manuscript including code---under version control.
 
 The likelihood of those deviations, while never unavoidable, is decreased by using simulated data to verify the suitability of the proposed analysis plan.
-This analysis plan may contain data-dependent decisions to increase flexibility of research plans.
+This analysis plan may contain data-dependent decisions to increase the flexibility of research plans.
 
@@
 On the one hand, we firmly believe that automated reproducibility makes research more productive and collaboration easier.
-But, on the other hand, we expect researchers to invest considerable time in learning new tools and than must persuade their collaborators to do the same.
+But, on the other hand, we expect researchers to invest considerable time in learning new tools and then must persuade their collaborators to do the same.
 Three arguments reconcile this apparent paradox.
@@
 The workflow we present here introduces several software solutions, which in itself present a threat to long-term reproducibility.
-To benefit from automatic and convenient reproduction, researcher must have `Git`, `Make`, and `Docker`.
+To benefit from automatic and convenient reproduction, researchers must use `Git`, `Make`, and `Docker`.
 However, `Git` and `Make` are itself included in the `Docker` image created by `repro`.
 Researchers can therefore employ the `Docker` image manually to download the `Git` repository and execute `Make` for full reproduction.
-In other words the only hard requirement for reproduction and therefore the Achilles' heel for this workflow is `Docker`.
+In other words, the only hard requirement for reproduction and therefore the Achilles' heel for this workflow is `Docker`.
 The `Docker` approach has two vulnerabilities.
-First, and more important, the `Docker` image for the project and the `Git` repository have to remain available.
-The `Dockerfile` (the plain text description to build a `Docker` image) as opposed to the image is insufficient because it relies on too many service providers (e.g.
+First, and more importantly, the `Docker` image for the project and the `Git` repository have to remain available.
+The `Dockerfile` (the plain text description to build a `Docker` image), as opposed to the image, is insufficient because it relies on too many service providers (e.g.
 Microsoft R Application Network, Ubuntu Package Archive).
-We recommend to archive the `Git` repository and the `Docker` image with zenodo.org, a non-profit long-term storage for scientific data.
+We recommend archiving the `Git` repository and the `Docker` image with zenodo.org, a non-profit long-term storage for scientific data.
 The necessary steps for archival on zenodo.org are described at the end of section [Preregistration as Code --- a Tutorial].
 The second vulnerability is that even if the existence of the `Docker` image and `Git` repository is guaranteed, future researchers still require software to run the image.
-To that end they can either rely on `Docker` itself or `Docker`-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
+To that end, they can either rely on `Docker` itself or `Docker`-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
 The only way to remove the reliance on such external software is to turn the `Docker` image into a full operating system that subsequently can be installed and run on almost any modern computer.
@@
 However, this process requires much technical knowledge and is currently not facilitated by `repro`.
-In any case and without such precautions do the `R Markdown`, the `Makefile` and the `Dockerfile` provide information that allow researchers to manually trace the computational steps and recreate the computational environment.
+In any case and without such precautions, do the `R Markdown`, the `Makefile` and the `Dockerfile` provide information that allows researchers to trace the computational steps and recreate the computational environment manually.
 The `Makefile`, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are, for whatever reason, unable to run `Make`.
@@
 Open science practices are a continually evolving field where technical innovations foster changes in research practice.
-Open data is much more widespread thanks to online storage facilities; preregistration are possible because there are preregistration platforms and so forth.
+Open data is much more widespread thanks to online storage facilities; preregistration is possible because there are preregistration platforms and so forth.
 Similarly, we hope that fully automatic reproduction, e.g. with `repro` as technical innovation, increase scientific rigour and efficiency.
@@
 `repro`'s modular structure allows such alternative workflow, though they have not been implemented.
-Depending on the demand by users, we will implement some of these alternative workflows in `repro` and hope for a broad adoption of computational reproducibility in the near future.
+Depending on the demand by users, we will implement some of these alternative workflows in `repro` and hope for broad adoption of computational reproducibility in the near future.
 

### 2021-10-07 manuscript.Rmd

@@
 Beyond standard preregistrations in psychological research, we believe that two applications specifically lend themselves to PAC.
-First, registrations of clinical trials typically describe analyses in exhaustive detail.
-Those registrations typically contain a detailed description of how results will be presented, including example tables and graphics [@ICH1998].
-PAC may significantly reduce the required workload while maintaining (and exceeding) the required standards for registering a clinical trial.
+First, pregistrations of clinical trials (there called statistical analysis plan, @ICH1998) typically describe analyses in exhaustive detail.
+Those preregistrations typically contain a detailed description of how results will be presented, including shells of tables and graphics [@yuanGuideStatisticalAnalysis2019].
+PAC may significantly reduce the required workload while maintaining (and exceeding) the required standards for preregistering a clinical trial.
 

### 2021-10-07 manuscript.Rmd
cite r packages for simulating data closes #258

@@
 These results are initially based on simulated data with the same structure as the data they expect to obtain from their experiments.
-For guidance on how to simulate data, see @morrisUsingSimulationStudies2019, @paxtonMonteCarloExperiments2001 and @skrondalDesignAnalysisMonte2000.
+For guidance on how to simulate data, see @morrisUsingSimulationStudies2019, @paxtonMonteCarloExperiments2001 and @skrondalDesignAnalysisMonte2000 as well as the R packages [`simstudy`](https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html)[@simstudy] and [`psych`](https://personality-project.org/r/psych/help/sim.html)[@psych].
+
 Once the preregistration is submitted and real data have been collected or made available,

### 2021-10-08 manuscript.Rmd
note again that we diviated from the preregistration

@@
 
-The last thing we need to do is to report our results with the combination of the functions `planned_analysis()` and `report_analysis()`.
+Naturally we did not access the data when we wrote this code initially.
+We therefore did not know is the exact format the data would have.
+So we did need to change our preregistration after accessing the data to include i.a. the recoding of gender (lines 17-22) and the items (lines 23-26).
+We invite the reader to evaluate the changes we made to the preregistratered code on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
+
+The last thing we need to preregister is the reporting of our results with the combination of the functions `planned_analysis()` and `report_analysis()`.
 
@@
 This is an example of how the results could be reported (based on simulated data):
+
+```{r, eval=FALSE, echo=-1}
+# this is a fake chunk, the faked output comes the chunk below
+report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)
+```
 

### 2021-10-08 manuscript.Rmd
fix link

@@
 The first step is to install the required software.
-We assume that you have installed R [@R-base, version `r with(R.version, paste0(major, ".", minor))`] and RStudio [@rstudio, version 1.4] already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert)[@R-repro].
+We assume that you have installed R [@R-base, version `r with(R.version, paste0(major, ".", minor))`] and RStudio [@rstudio, version 1.4] already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert/repro)[@R-repro].
 In case you either have not installed R and RStudio or are unsure if they are up to date, you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers how to install all software necessary for this tutorial in three steps.

### 2021-10-08 manuscript.Rmd

@@
 You may log into zenodo.org through your GitHub account.
+To log in with your GitHub account:
+
+> Navigate to https://zenodo.org/login/ →  Log in with GitHub
+
+\noindent{}To link zenodo.org and GitHub
 
 > Log into zenodo.org → Account → GitHub[^zenodo-github]
+
+\noindent{}Or:
+
+> Navigate to https://zenodo.org/account/settings/github/
+
 [^zenodo-github]: https://zenodo.org/account/settings/github/
@@
 
-```{r}
+```{r, eval=FALSE}
 usethis::browse_github()

### 2021-10-08 manuscript.Rmd
bump release number

@@
 
-The interested reader can find the example rendered as a PDF file with real data in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.2.0-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
+The interested reader can find the example rendered as a [PDF file with real data](https://github.com/aaronpeikert/repro-tutorial/files/7309455/preregistration.pdf) in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.3.1-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
 

### 2021-11-03 manuscript.Rmd
apply changes suggested by julia
correct p. 2

correct p. 3

correct p. 4

correct p. 5

correct p. 6

correct p. 7

correct p. 8

correct p. 9

correct p. 10

correct p. 11

correct p. 12

correct p. 13

correct p. 14

correct p. 16

correct p. 16

correct p. 17

correct p. 18

correct p. 19

finish up in text corrections

@@
 ---
-title: "Reproducible Research in R: A tutorial on how to do the same thing more than once"
+title: "Reproducible Research in R: A Tutorial on how to Do the Same Thing More Than Once"
 author:
@@
     orcid: 0000-0001-7813-818X
-  - name: Caspar J. Van Lissa
+  - name: Caspar J. van Lissa
     affil: 2, 3
@@
     address: |
-      Center for Lifespan Psychology---Max Planck Institute for Human Development
-      Lentzeallee 94, 14195 Berlin, Germany
+      Center for Lifespan Psychology---Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany
   - num: 2
     address: |
-      Department of Methodology & Statistics---Utrecht University faculty of Social and Behavioral Sciences, Utrecht, Netherlands
+      Department of Methodology & Statistics---Utrecht University, Faculty of Social and Behavioral Sciences, Utrecht, Netherlands
   - num: 3
@@
       Max Planck UCL Centre for Computational Psychiatry and Ageing Research
-      Berlin, Germany and London, UK
+      Berlin, Germany, and London, UK
   - num: 5
@@
   Reproducibility has long been considered integral to the scientific method.
-  An analysis is called reproducible when an independent person obtains the same results from the same data.
+  An analysis is considered reproducible if an independent person can obtain the same results from the same data.
   Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
   Technological advancements now enable scientists to achieve a more comprehensive standard;
-  a standard where anyone can access a digital research repository and reproduce all computational steps from raw data to final report including all relevant statistical analyses, with a single command.
+  a standard where that allows anyone to access a digital research repository and reproduce all computational steps from raw data to final report including all relevant statistical analyses, with a single command.
   This method has far-reaching implications for scientific archiving,
@@
   One obstacle to the widespread use of this method is that the underlying tools
-  are complex and not part of the regular training of most researchers.
-  This paper introduces `repro`, an R-package, which guides researchers in installing and using the tools needed to make a research project reproducible.
-  Finally, we suggest the use of the proposed workflow for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
+  are complex and not part of most researchers' basic training.
+  This paper introduces `repro`, an R package that guides researchers through installation and use of the tools required to make a research project reproducible.
+  We also suggest the utilization of the proposed workflow for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
   Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions of those analyses, which then merely complement the PAC as a more readable summary.
   PAC circumvents the shortcomings of ambiguous preregistrations that may give researchers undesired degrees of freedom in the execution of their analysis.
-  Reproducibility, facilitated by automation, therefore has a wide range of applications to accelerate scientific progress.
+  Reproducibility facilitated by automation therefore has a wide range of applications to accelerate scientific progress.
 abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
 keywords: |
-  open science; compuational reproducibility; preregistration; R; R Markdown; Make; GitHub; Docker
+  open science; computational reproducibility; preregistration; R; R Markdown; Make; GitHub; Docker
 acknowledgement: |
-  The authors received no financial support for the research, authorship, and/or publication of this article.
+  We would like to thank Maximilian Stefan Ernst for his contributions to the code for the simulation study. 
 authorcontributions: |
@@
   Andreas Brandmaier and Caspar J. Van Lissa contributed further ideas, critical feedback, and revisions of the original manuscript.
-  Furthermore, we would like to thank Maximilian Stefan Ernst (not an author) for his contributions to the code for the simulation study.
+
 conflictsofinterest: |
     The authors declare no conflict of interest.
+    We have received no financial support for the research, authorship, and/or publication of this article.
 abbreviations:
@@
 The credibility of empirical research results hinges on their objectivity.
-Objectivity in this context means "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
+Objectivity in this context means, "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
 Only a reproducible result can be tested and understood by anyone.
@@
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
-This is because making a research project with all code, data, and materials reproducible comes with various challenges (e.g., missing software dependencies or ambiguous documentation of the exact computational steps taken), and there is a dire need for reproducibility standards.
-With technological advancement, it is now possible to make all digital products related to a research project available in a manner that enables automatic reproduction of the research project with minimal effort.
+This is because making a research project reproducible with all code, data, and materials is accompanied by various challenges (e.g., missing software dependencies or ambiguous documentation of the exact computational steps taken), and there is a dire need for reproducibility standards.
+With technological advancement, however, it is now possible to make all digital products related to a research project available in a manner that enables automatic reproduction of the research project with minimal effort.
 
 With this paper, we pursue two aims.
-First, we want to introduce researchers to a notion of automated reproducibility that requires, apart from the setup of the software, no manual steps. 
-Second, we discuss opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated but also how research itself is conducted.
-With regard to the first goal, we discuss how to address four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
+First, we want to introduce researchers to a notion of automated reproducibility that requires no manual steps, apart from the setup of the software. 
+Second, we present the opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated but also how research itself is conducted.
+With regard to the first goal, we show how to address four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
 Then we present a tutorial on how to employ these tools to achieve automated reproducibility.
 A single tutorial cannot comprehensively introduce the reader to the individual tools, but this tutorial is intended to help readers to build a coherent workflow.
-This tutorial is aimed at researchers who regularly write code to analyze their data and who are willing to make relevant code, data, and materials available, either publicly or on request.
-Ideally, our reader already has created a dynamic document at some point in time (e.g. with RMarkdown or Jupyter) and attempted to track a file with version control (e.g. with Git). 
+It is aimed at researchers who regularly write code to analyze their data and are willing to make relevant code, data, and materials available, either publicly or on request.
+Ideally, our reader has already created a dynamic document at some point in time (e.g., with R Markdown or Jupyter) and attempted to track a file with version control (e.g., with Git). 
 The R package `repro` supports researchers in setting up the required software and in adopting this workflow.
-We present automated reproducibility as best practice, a goal that is not always archived because of limited resources, technical restrictions or practical considerations but is worth striving for.
-If, due to whatever reason, some parts of the workflow prove to be unfeasible, we encourage the reader to find their own suitable solutions and to concentrate on the feasible parts.
-
-In pursuit of the second aim, we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon implementing this reproducible workflow, the so-called *preregistration as code* (PAC).
-PAC involves preregistering the intended analysis code and the major part of the final scientific report as a dynamic document, including typical sections like Introduction, Methods, and Results.
-The resulting dynamic document closely resembles the final manuscript but temporarily uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
+We present automated reproducibility as best practice, a goal that is not always achieved due to limited resources, technical restrictions, or practical considerations, but is very much worth striving for.
+If some parts of the workflow prove to be unfeasible, we encourage the reader to find their own solutions and to concentrate on the feasible parts.
+
+In pursuit of the second aim, we present a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon implementing this reproducible workflow, the so-called *Preregistration as Code* (PAC).
+PAC involves preregistering the intended analysis code and the major part of the final scientific report as a dynamic document, including typical sections like introduction, methods, and results.
+The resulting dynamic document closely resembles the final manuscript but temporarily uses simulated data to generate placeholder results (e.g., figures, tables, and statistics).
 With these placeholders, authors preregister the exact presentation of their results without having seen real data yet.
@@
 
-[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept which he later generalized to inter-subjective criticism (Footnote *1, p.
+[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept that he later generalized to inter-subjective criticism (Footnote *1, p.
 22).
-Somewhat confusingly, @popperLogicScientificDiscovery2002 uses the term "Reproducibility" for what we would call "Replication".
+Somewhat confusingly, @popperLogicScientificDiscovery2002 uses the term "Reproducibility" for what we would call "Replication."
 
@@
 Out of 35 published articles with open code and data, results of 22 articles could be reproduced, but in 11 of these cases, further assistance from the original authors was required.
-For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
+For 13 articles, at least 1 outcome could not be reproduced—--even with the original authors' assistance.
 Another study of 62 Registered Reports found that only 41 had data available, and 37 had analysis scripts available [@obels2020].
 The authors could execute only 31 of the scripts without error and reproduce the results of only 21 articles (within a reasonable time).
-These failed attempts to reproduce highlight the need for widely accepted reproducibility standards because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
+These failed attempts to reproduce findings highlight the need for widely accepted reproducibility standards because open repositories do not routinely provide sufficient information to reproduce relevant computational and statistical results.
 If digital research products are available but not reproducible, their added value is limited.
@@
 An easy-to-use standard, on the other hand, is more likely to be adopted.
-Our approach is to promote broad adoption of easy-to-use practices, i.e., a "low threshold", and ensure that these practices are compatible with more complex rigorous solutions, i.e., a "high ceiling".
-As researchers become more proficient with the tools involved, they can thus improve the reproducibility of their work.
-
-We have structured the tutorial with a *learning by doing* approach in mind, such that readers can follow along on their own computers.
+Our approach is to promote broad adoption of such practices, by implementing a "low threshold," and to ensure that they are compatible with more complex rigorous solutions, i.e., a "high ceiling."
+As researchers become more proficient in using the tools involved, they can thus improve the reproducibility of their work.
+
+We have structured the tutorial with a *learning-by-doing* approach in mind, such that readers can follow along on their own computers.
 We explicitly encourage readers to try out all R commands for themselves.
@@
 
-# Threats to reproducibility and appropriate remedies
-
-From our own experience with various research projects, we identified the following common threats to reproducibility:
-
-1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code could have been modified during the writing of a paper because a bug was removed at some point in time.
+# Threats to Reproducibility and Appropriate Remedies
+
+From our own experience with various research projects, we have identified the following common threats to reproducibility:
+
+1.	*Multiple inconsistent versions of code, data, or both*; for example, the data set may have changed over time because outliers were removed at a later stage or an item was later recoded; or, the analysis code may have been modified during the writing of a paper because a bug was removed at some point in time.
 It may then be unclear which version of code and data was used to produce some reported set of results.
 2.	*Copy-and-paste errors*; for example, results are often manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result.
-3.	*Undocumented or ambiguous order of computation*; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually instead of in a script; copy-pasting results from one program to another).
-4.	*Ambiguous software dependencies*; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.
+3.	*Undocumented or ambiguous order of computation*; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually without documentation (e.g., executing a command manually rather than in a script; copy-and-pasting results from one program to another).
+4.	*Ambiguous software dependencies*; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer, or no longer exist at all; or a different version of the same software may produce different results.
 
@@
 Version control allows a clear link between which results were generated by which version of code and data.
-This is important for reproducibility because only if it is obvious which version of data and code produced which results, those results can be said to be reproducible.
-We recommend using *Git* for version control, because of its widespread adoption in the R community.
+This is important for reproducibility because only if it is clear which version of data and code produced which results, those results can be said to be reproducible.
+We recommend using `Git` for version control, because of its widespread adoption in the R community.
 
@@
 `Git` is built around snapshots that represent the project state at a given point in time.
-Those snapshots are called *commits* and work like a "save" action.
+These snapshots are called *commits* and work like a "save" action.
 Ideally, each commit has a message that succinctly describes these changes.
-It is good practice to make commits for concrete milestones (e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3").
+It is good practice to make commits for concrete milestones (e.g., "Commented on Introduction," "Added SES as a covariate," "Address Reviewer 2's comment 3").
 This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
 Each commit refers back to its ancestor, and they form a timeline together.
-The entirety of commits (i.e. the version-controlled project) is called a repository.
-In `Git`, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
+The entirety of commits (i.e., the version-controlled project) is called a repository.
+In `Git`, specific snapshots of a repository can be tagged, such that the user can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
 `Git` has additional features beyond basic version control, such as branches to facilitate collaboration.
-@vuorreCuratingResearchAssets2018 provide a more extensive treatment how `Git` functions and how to use `Git` for research.
+@vuorreCuratingResearchAssets2018 provide a more extensive treatment of how `Git` functions and how to use `Git` for research.
 @bryanExcuseMeYou2018 provides additional information on how to track `R Markdown` documents.
 Collaborating via `Git` requires the repository to be uploaded somewhere.
-We recommend GitHub to host `Git` repositories because of its popularity among R users.
-GitHub has many tools that ease project management and collaboration, and these tools are valuable in scientists everyday work, but they are not central to achieving reproducibility.
+We recommend GitHub as a host for `Git` repositories because of its popularity among R users.
+GitHub has many tools that ease project management and collaboration, and these tools are valuable in scientists' everyday work, but they are not central to achieving reproducibility.
 
@@
 Through dynamic document generation, code becomes an integral, although usually hidden, part of the manuscript, complementing the verbal description and allowing interested readers to gain a deeper understanding of the contents [@knuthCWEBSystemStructured; @claerboutElectronicDocumentsGive1992].
-*R Markdown* uses `Markdown` for formatting text and R (or other programming languages) for writing the statistical analysis.
+`R Markdown` uses `Markdown` for text formatting and R (or other programming languages) for writing the statistical analysis.
 `Markdown` is a lightweight text format in plain text with a minimal set of reserved symbols for formatting instructions.
 This way, `Markdown` does not need any specialized software for editing.
-It is highly fool-proof [unlike, for example, `LaTeX` @lamportLATEXDocumentPreparation1994], works well with version control systems and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+It is fool-proof (unlike, for example, `LaTeX` [@lamportLATEXDocumentPreparation1994]), works well with version control systems, and can be exported to various document formats, such as HTML websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
 We suggest using `Markdown` for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
-`R Markdown` extends regular `Markdown` by allowing users to include R code chunks [in fact, arbitrary computer code @riedererChapter15Other] into a `Markdown` document.
+`R Markdown` extends regular `Markdown` by allowing users to include R code chunks (in fact, arbitrary computer code [@xieMarkdownCookbook2020, Chapter 15, Chapter 15, Other Languages]) into a `Markdown` document.
 Upon rendering the document, the code blocks are executed, and their output is dynamically inserted into the document.
-This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).
+This allows the creation of (conditionally) formatted text, statistical results, and figures that are guaranteed to be up-to-date because they are created anew every time the document is rendered to its output format (e.g., presentation slides or a journal article).
 @xieMarkdownCookbook2020 provides an extensive yet practical introduction to most features of `R Markdown`.
@@
 While version control and dynamic document generation are becoming more common, we have argued that two more components are required and that each component alone is unlikely to guarantee reproducibility [@Peikert2019; @vanlissa2020worcs].
-In practice, dependencies between project files (for example, the information what script uses which data file and what script needs to be run first) or on external software (such as system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
+In practice, dependencies between project files (e.g., information on what script uses which data file and what script needs to be run first) or on external software (e.g., system libraries or components of the programming language, such as other R packages) are frequently unmentioned or not exhaustively and unambiguously documented.
 
@@
 One noteworthy property of computational recipes is that they require a textual representation and the ability of a program to make sense of this textual representation.
-A workflow that includes, e.g. renaming variable names by hand in a graphical spreadsheet application is therefore incompatible with automated reproducibility.
+A workflow that involves renaming variable names by hand in a graphical spreadsheet application, for example, is therefore incompatible with automated reproducibility.
 Another property of a computational recipe is that the same inputs should always result in the same outputs.
 For most computer code (given the same software is used), this property is fulfilled.
-However, one noteworthy exception is pseudo-random numbers.
-Whenever random numbers are used in a computation, this computation is only reproducible when the same random numbers are used.
+However, one noteworthy exception is the generation of pseudo-random numbers.
+Whenever random numbers are used in a computation, it is only reproducible if the random number generator generates the same numbers.
 To ensure identical random numbers, users may fix the state of the random number generated with a so-called seed (e.g.
-in R with `set.seed()`), but they also need to guarantee that the pseudo-random number generator is unchanged [see @Peikert2019].
-
-We recommend using *Make* for dependency tracking because it is language independent.
+`set.seed()` in R), but they also need to guarantee that the pseudo-random number generator is unchanged [see @Peikert2019].
+
+We recommend using `Make` for dependency tracking because it is language independent.
 The following hypothetical example illustrates the utility of `Make` and a suitable `Makefile`.
@@
 
-There are two targets, the final rendered report (`manuscript.pdf`, l.1) and the simulation results (`simulation_results.csv`, l.4).
+There are two targets, the final rendered report (`manuscript.pdf`, l. 1) and the simulation results (`simulation_results.csv`, l. 4).
 Each target is followed by a colon and a list of requirements.
 If a requirement is newer than the target, the recipe will be executed to rebuild the target.
-If a requirement does not exist, `Make` uses a recipe to build the requirement first before building the target.
-Here, if one were to build the final `manuscript.pdf` by rendering the `R Markdown` with the command shown in l.2, `Make` would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation first before rendering the manuscript.
-This ensures that the simulated data are present before the manuscript is built and that the simulation is re-run, and the manuscript is rebuilt if the simulation code was changed.
+If a requirement does not exist, `Make` uses a recipe to build the requirement before building the target.
+Here, if one were to build the final `manuscript.pdf` by rendering the `R Markdown` with the command shown in l.2, `Make` would check whether the file `simulation_results.csv` exists; if not, it would issue the command in l.5 to run the simulation before rendering the manuscript.
+This ensures that the simulated data are present before the manuscript is built, and that the simulation is re-run and the manuscript is rebuilt if the simulation code was changed.
 `Make` therefore offers a standardized process to reproduce projects, regardless of the complexity or configuration of the project.
@@
 A version-controlled dynamic document with dependency tracking may still rely on external software.
-Troubleshooting issues specific to a particular programming language or dependent tool typically require considerable expertise and threatens reproducibility.
+Troubleshooting issues specific to a particular programming language or dependent tool typically require considerable expertise and threaten reproducibility.
 One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via *software management*.
@@
 The central idea is that by
-"packaging the key elements of the computational environment needed to run the desired software the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p.
+"[..] packaging the key elements of the computational environment needed to run the desired software [makes] the software much easier to use, and the results easier to reproduce [...]" [@silverSoftwareSimplified2017, p.
 174].
@@
 It manages software dependencies by constructing a virtual software environment independent of the host software environment.
-These so-called "`Docker` images" function like a virtual computer  (i.e., a sort of "sand box").
-A `Docker` image does contain *all* software dependencies used in an analysis---not just R-packages, but also R and Rstudio, and even the operating system.
-All differences in computational results that could be caused by software are hence eliminated.
-This is important because some of its functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.
+These so-called "`Docker` images" function like a virtual computer (i.e., a "sand box" a computational environment seperated from the host).
+A `Docker` image contains *all* software dependencies used in an analysis---not just R packages, but also R and Rstudio, and even the operating system.
+This is important because low level functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.
+All of the differences in computational results that could be caused by software are hence eliminated.
 
@@
 First, users can manually install whatever software they like from within the virtual environment.
-Such a manually build environment can still be ported to all computers which support `Docker`.
-We, however, prefer the second way of building images automatically from a textual description called `Dockerfile`.
-Because the `Dockerfile` clearly describes which software is installed how, the installation process can be repeated automatically.
-Users can, therefore, quickly change the software environment, e.g. update to another R version or given package version.
-
+Such a manually build environment can still be ported to all computers that support `Docker`.
+However, we prefer the second way of building images automatically from a textual description called `Dockerfile`.
+Because the `Dockerfile` clearly describes how which software is installed, the installation process can be repeated automatically.
+Users can therefore quickly change the software environment, for example, update to another R version or given package version. 
 Packaging all required software in such an image requires considerable amounts of storage space.
 Two major strategies help to keep the storage requirements reasonable.
-One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
+One is to rely on pre-made images, that are maintained by a community for particular purposes.
 For example, there are pre-made images that only include what is necessary for R, based on Ubuntu containers [@boettigerIntroductionRockerDocker2017].
 Users can then install whatever they need in addition to what is provided by these pre-compiled images.
-`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
+`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}GiB of disk space."))`
 The image for this project includes Ubuntu, R, RStudio, LaTeX as well as a variety of R packages like tidyverse [@tidyverse] and all its dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
+
 A second strategy is to save a container recipe, a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment instead of the software environment itself.
-`Dockerfiles` are tiny (the [`Dockerfile`](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
+`Dockerfiles` are tiny (the [`Dockerfile`](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024, 2)`KiB).
 However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
-For proper archival, we, therefore, recommend storing a complete image of the software environment, in addition to the `Dockerfile`.
+For proper arching, we therefore recommend storing a complete image of the software environment, in addition to the `Dockerfile`.
 A more comprehensive overview of the use of containerization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
-Note that the Workflow for Open Reproducible Code in Science (WORCS) we presented elsewhere [@vanlissa2020worcs] relies for software management on the R package `renv`[@R-renv], which is more lightweight and easier to use than `Docker` on the one hand, but not as comprehensive on the other hand because it only snapshots the R packages instead of all software used.
-
-To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)), dynamic document generation (using `R Markdown`), version control (using `Git`), dependency tracking (using `Make`), and Software management (using `Docker`).
-While `R Markdown` and `Git` are well integrated into the R environment through RStudio, `Make` and `Docker` require a level of expertise that is often beyond the training of scholars outside the field of information technology, which can represent a considerable obstacle to the acceptance and implementation of the workflow.
-To remove this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
+Note that WORCS, which we presented elsewhere [@vanlissa2020worcs] relies on the R package `renv` [@R-renv] for software management.
+`renv` is more lightweight and easier to use than `Docker` on the one hand, but not as comprehensive on the other because it only takes snapshots of the R packages instead of all software used.
+
+To summarize, the workflow by @Peikert2019 requires four components (see Figure \@ref(fig:schematic)) dynamic document generation (using `R Markdown`), version control (using `Git`), dependency tracking (using `Make`), and software management (using `Docker`).
+While `R Markdown` and `Git` are well integrated into the R environment through RStudio, `Make` and `Docker` require a level of expertise that is often beyond the training of scholars outside the field of information technology.
+Thus can represent a considerable obstacle to the acceptance and implementation of the workflow.
+To overcome this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
 Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
@@
 in fact, they do not even need to have R installed because the entire project can be rebuilt inside a container with R installed.
-In the remainder, we will walk you through creating a reproducible research project with the package `repro`.
-
-```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (`Git`), dependency tracking (`Make`), software management (`Docker`), and dynamic document generation (`R Markdown`).
+In the remainder, we will walk you through the creation of a reproducible research project with the package `repro`.
+
+```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed columns) central to the reproducible workflow: version control (`Git`), dependency tracking (`Make`), software management (`Docker`), and dynamic document generation (`R Markdown`).
 `Git` tracks changes to the project over time.
@@
 
-# Creating reproducible research projects
+# Creating Reproducible Research Projects
 
 One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be quite time-intensive.
-To lower the threshold to adoption, the R-package `repro` introduces helper functions that wrap and hide the complexity of the necessary more complicated (and powerful) tools.
-The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer does and suggest what the user is supposed to do next.
+To lower the threshold, the R package `repro` introduces helper functions that wrap and hide the complexity of the necessary more complicated (and powerful) tools.
+The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer is doing and suggest what the user should do next.
 The `repro` package extends the [`usethis`](https://usethis.r-lib.org) [@usethis] package.
-While the `usethis` package aims to simplifies the development of R packages, `repro` contains reproducibility specific utilities.
+While the `usethis` package aims to simplifys the development of R packages, `repro` contains reproducibility-specific utilities.
 We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
@@
 
-This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
+This tutorial assumes that the user is predominantly working in R, with the help of RStudio.
 It describes basic steps that we expect to be representative of a small-scale psychological research project. 
@@
 After completing the tutorial, you should be able to customize your workflow accordingly.
+
 The first step is to install the required software.
-We assume that you have installed R [@R-base, version `r with(R.version, paste0(major, ".", minor))`] and RStudio [@rstudio, version 1.4] already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert/repro)[@R-repro].
-In case you either have not installed R and RStudio or are unsure if they are up to date, you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers how to install all software necessary for this tutorial in three steps.
+We assume that you have installed R [@R-base, version `r with(R.version, paste0(major, ".", minor))`] and RStudio [@rstudio, version 1.4] already but the tutorial will guide you in detail through the installation of other necessary software with the help of the R package [`repro`](https://github.com/aaronpeikert/repro) [@R-repro].
+In case you have either not installed R and RStudio or are unsure if they are up-to-date, you might want to consult our installation advice in the [Online Supplementary Material](https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md) that covers the installation of all software necessary for this tutorial in three steps.
 The [installation advice]((https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)) may also help Windows users who have problems installing `Docker`.
@@
 
-An example of the assistance that `repro` provides is the check functions.
-These verify that you indeed have installed and set up the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :
+The the check functions represent an example of the assistance that `repro` provides is .
+They verify that you have indeed installed and set up the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to clarify which functions are native `repro` functions):
 
@@
 
-These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
-Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have `Git` installed:
+These functions check whether specific dependencies are available on the user's system, and if not, explain what further action is needed to obtain it.
+Sometimes they ask the user to do something, for example, the following happens if you are a Windows user who does not have `Git` installed:
 
@@
 The messages from `repro` try to take the user by the hand to help them solve problems.
-These messages are adjusted to your specific operating system and installed dependencies.
-Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions and to prepare your system for creating reproducible projects.
+They are adjusted to your specific operating system and installed dependencies.
+Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions and to prepare your system for the creation of reproducible projects.
 
@@
 
-If you know what SSH is and want to use it, you may alternatively use:
+If you know what Secure Shell (SSH) is and want to use it, you may alternatively use:
 
@@
 
-If necessary, follow any instructions presented until all checks pass.
-
-## Implementing version control
+If necessary, follow any instructions presented until all checks are passed.
+
+## Implementing Version Control
 
@@
 
-This creates a project containing an example analysis, but you may use any other template or existing R Project to turn into a reproducible research project using the functionality of `repro`.
+This creates a project containing a sample analysis, but you may turn any other template or existing R project into a reproducible research project using the functionality of `repro`.
 
@@
 Now the file `private.md` will not be committed to `Git` and will not be made public.
-Please consider carefully if you can include data in the repository without violating privacy rights.
+Please also consider carefully whether you can include data in the repository without violating privacy rights.
 If you are not allowed to share your data publicly, add the data file(s) to the `.gitignore` file and only share them on request.
 
-Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
+New users, in particular, are advised it is recommended to follow this procedure and exclude sensitive files before proceeding.
 When you are ready, you can begin tracking your files using `Git` by running:
@@
 Click on the empty box next to the file you want to stage.
-A checkmark appears to indicate that the file is staged.
-After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
+A checkmark then indicates that the file is staged.
+After you have staged all of the files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
 This stores a snapshot of the current state of the project.
 
-The files you created and the changes you made still have not left your computer.
+The files you created and the changes you made have not yet left your computer.
 All snapshots are stored in a local repository in your project folder.
@@
 While you can choose any `Git` service (like GitLab or BitBucket), we will use GitHub in this tutorial.
-Before you upload your project to `Git`, you need to decide if you would like the project publically accessible (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
-To upload the project publically to GitHub use:
+Before you upload your project to GitHub, you need to decide whether you would like the project to be publicly accessible (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
+To upload the project publicly to GitHub use:
 
@@
 
-Depending on your computer configuration, this may ask to set up a secure connection to GitHub.
+Depending on your computer's configuration, it may ask you to set up a secure connection to GitHub.
 In this case, first, follow the suggestions shown on the R console.
 
-## Using dynamic document generation
+## Using Dynamic Document Generation
 
@@
 `R Markdown` is a type of dynamic document well-suited to the RStudio user interface.
-The text of an `R Markdown` is formatted by `Markdown` (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
-The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
+The text of an `R Markdown` is formatted by `Markdown` (see [@xieMarkdownDefinitiveGuide2019] for technical details and [@xieMarkdownCookbook2020] for practical guidance).
+The code mostly consists of R code (although other programming languages are supported, like Python, C++, Fortran, etc).
 The following example serves to illustrate the `Markdown` syntax. 
-It shows how to create a heading, a word in bold font, a citation and a list of several items in `Markdown`:
+It shows how to create a heading, a word in bold font, a citation, and a list of several items in `Markdown`:
 
@@
 
-* do research
-* do more research
-* spend time with family and friends
+* Do research
+* Do more research
+* Spend time with family and friends
 ```
@@
 specific journal requirements.
-Of particular note is the package output format supplied by the [`papaja`](https://github.com/crsh/papaja) package [@papaja], which produces APA compliant manuscripts [@apa7].
-Another advantage is that, because `R Markdown` files are in plain text, they are more suitable for version control using `Git`.
-Some users might find it easier to activate the "Visual Editor" of RStudio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the `R Markdown` document), which features more graphical elements like a traditional word processor but still creates an `R Markdown` underneath with all of its flexibility.
-The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a new line, which eases tracking changes across versions) and easing the generation of citations and references to Tables and Figures.
+Of particular note is the package output format supplied by the [`papaja`](https://github.com/crsh/papaja) package [@papaja], which produces manuscripts that follow the American Psychological Association formatting requirements [@apa7].
+`R Markdown` files are in plain text, with the advantage that they are more suitable for version control using `Git`.
+Some users might find it easier to activate the "Visual Editor" of RStudio ([Ctrl] + [Shift] + [F4] or click on the icon that resembles drawing materials or a compass in the upper right corner of the `R Markdown` document), which features more graphical elements like a traditional word processor but still creates an `R Markdown` underneath with all of its flexibility.
+The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a new line, which makes it easier to track changes across versions) and improving the generation of citations and references to tables and figures.
 
 Now that you are familiar with `Markdown` formatting basics, we turn our attention to including code and its results in the text.
-Code is separated by three backticks (and the programming language in curly braces) like here:
+Code is separated by three backticks (and the programming language in curly brackets) like this:
 
@@
 This means that whenever you render the `R Markdown` to its intended output format, the code will be executed and the results updated.
-The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g. on a preprint server.
+The resulting output document will be static, e.g., a pdf document, and can be shared wherever you like, e.g., on a preprint server.
 
@@
 These metadata contain information about the document, like the title and the output format.
-Metadata are placed at the beginning of the document and are separated by three dashes from the document body.
-The following example is a full markdown document where the metadata (the "YAML front matter") is in lines 1--6.
+Metadata are placed at the beginning of the document and are separated from the document body by three dashes.
+The following example is a full markdown document where the metadata (the "YAML front matter") are in lines 1--6.
 
@@
 ---
-title: "A tutorial on how to do the same thing more than once"
-author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
-abstract: A Hitchhiker's Guide to Reproducible Research in R
+title: "A Tutorial on how to Do the Same Thing More Than Once"
+author: Aaron Peikert, Caspar J. van Lissa, and Andreas M. Brandmaier
+abstract: A hitchhiker's guide to reproducible research in R
 output: html_document
@@
 
-1. *Version Control*
-2. *Dynamic Document Creation*
-3. *Dependency Tracking*
-4. *Software Management*
+1. *Version control*
+2. *Dynamic document creation*
+3. *Dependency tracking*
+4. *Software management*
 
@@
 
-## Manage software and file dependencies
+## Manage Software and File Dependencies
 
 Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
-However, `R Markdown` has been extended to hundreds of document formats which are often highly configurable by their metadata.
+However, `R Markdown` has been extended to hundreds of document formats that are often highly configurable by their metadata.
 Some R packages add additional metadata fields.
@@
 `repro` uses this information to construct a `Makefile` for the dependencies on other files and a `Dockerfile` that includes all required packages.
-These two files together form the basis for consistency within a research project and consistency across different systems.
+Together, these two files form the basis for consistency within a research project and consistency across different systems.
 The function `repro::automate()` converts the metadata from all `R Markdown` files in the project (all files with the ending `.Rmd`) to a `Makefile` and a `Dockerfile`.
-Together, these files allow users (including your future self) to reproduce every step in the analysis automatically.
+These files allow users (including your future self) to reproduce every step in the analysis automatically.
 Please run `repro::automate()` in your project:
@@
 
-It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new `R Markdown` file to the project to keep the `Makefile` and `Dockerfile` up to date.
+It is important to re-run `repro::automate()` whenever you change the `repro` metadata, change the output format, or add a new `R Markdown` file to the project to keep the `Makefile` and `Dockerfile` up to date.
 There is no harm in running it too often.
-Other than the `Makefile` and the `Dockerfile`, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to `Git`.
+Other than the `Makefile` and the `Dockerfile`, which are created in the document root path, `repro` generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to `Git`.
 
@@
 
-When you or someone else who received your project want to reproduce it, they first have to install the required software, which is `Make`, and `Docker`.
+If someone (including you) wants to reproduce your project, they first have to install the required software, that is `Make`, and `Docker`.
 Remember, you can use the `check_*`-functions to test if these are installed:
+
+```{r, include=FALSE}
+opts <- options()
+options(repro.docker = TRUE,
+        repro.docker.running = FALSE)
+```
 
@@
 
-When they are setup, they can ask `repro` to explain how they should use `Make` and `Docker` to reproduce the project (or you could explain it to them):
+```{r, include=FALSE}
+options(opts)
+```
+
+When they are set up, they can ask `repro` to explain how they should use `Make` and `Docker` to reproduce the project (or you could explain it to them):
 
@@
 
-If you feel uncomfortable using the terminal directly, you can sent the command to the terminal from within R:
+If you feel uncomfortable using the terminal directly, you can send the command to the terminal from within R:
 
@@
 
-The only software required for reproducing a project is `Docker`, assuming users know how to build a `Docker` image and run `Make` within the container.
-However, if they have installed `Make` additionally to `Docker`, they do not need to know how to use `Docker` and can simply rely on the two `Make` commands `make docker` and `make -B DOCKER=TRUE`.
-The commands can be pasted into a terminal and will reproduce the project in many cases.
+The only "hard" software requirement for reproducing a project is `Docker`, assuming users know how to build a `Docker` image and run `Make` within the container.
+However, if they have installed `Make` in addition to `Docker`, they do not even need to know how to use `Docker` and can simply rely on the two `Make` commands "`make docker`" and "`make -B DOCKER=TRUE`".
 
@@
 
-3. Create an R project or use an existing one. Do not forget to add `repro` metadata (i.e. packages, scripts, data).
+3. Create an R project or use an existing one. Do not forget to add `repro` metadata (i.e., packages, scripts, data).
 
@@
 
-The example `repro` project already has this metadata:
+The sample `repro` project already has theese metadata:
 
@@
 
-This section is only for advanced users who want to overcome some limitations of `repro`.
-If you read this paper the first time, you probably want to skip this section and continue reading at the section "Preregistration as Code".
+This section is for advanced users who want to overcome some limitations of `repro`.
+If you read this paper the first time, you will probably want to skip this section and continue reading from the section "Preregistration as Code."
 As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.
 The simplification of `Docker` and `Make` are possible because of two restrictions.
-Users who ask themselves either "How can I install software dependencies outside of R in the `Docker` image?" or "How can I express complex dependencies between files, e.g.
-hundreds of data files are preprocessed and combined?") need to be aware of these restrictions and require a deeper understanding of the inner workings of `repro`.
+Users who ask themselves either, "How can I install software dependencies outside of R in the `Docker` image?" or "How can I express complex dependencies between files (e.g., hundreds of data files are preprocessed and combined)?" need to be aware of these restrictions and require a deeper understanding of the inner workings of `repro`.
 Other users may safely skip this section or return to it if they encounter such challenges.
@@
 The first limitation means that users must rely on software that is either already provided by the base `Dockerimage` "rocker/verse" or the R packages they list in the metadata.
-Other software that users might need, like other programming languages, not already installed LaTeX packages, etc., must be added manually.
-We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g.
-for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
-The second limitation is about dependencies.
-`Make` can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
+Other software that users might need, like other programming languages, not yet installed LaTeX packages, etc., must be added manually.
+We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()`, for example, for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), Python packages (via `pip` the python package manager).
+The second limitation is related to dependencies.
+`Make` can represent complex dependencies, for example: A depends on B, which in turn depends on C and D.
 If B is missing in this example, `Make` would know how to recreate it from C and D.
-These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
-Users, therefore, have to either "flatten" the dependency structure by simply stating that A depends on B, C and D, therefore leaving out important information or express the dependencies directly within the `Makefile`.
-
-The following section explains how to overcome these limitations while still relying on the automation afforded by `repro`.
+These dependencies, and how they should be resolved, are difficult to represent in the metadata.
+Users, therefore, have to either "flatten" the dependency structure by simply stating that A depends on B, C, and D, thereby leaving out important information or express the dependencies directly within the `Makefile`.
+
+The following section explains how to overcome these limitations despite reliance on the automation afforded by `repro`.
 Lifting these restrictions requires the user to interact more directly with `Make` or `Docker`.
@@
 
-Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; it consists of two processing steps.
+Let us have a closer look at the command for reproducing a repro project: `make docker && make -B DOCKER=TRUE`; which consists of two processing steps.
 First, it recreates the virtual software environment (`Docker`), and then it executes computational recipes in the virtual software environment (`Make`).
@@
 The command `make docker` will trigger `Make` to build the target called `docker`.
-The recipe for this target builds an image from the `Dockerfile` of the repository.
+The recipe for this target builds an image from the `Dockerfile` in the repository.
 The `&&` concatenates both commands and only runs the second command if the first is successful.
@@
 First, a call to `make` without any explicit target will build the `Make` target `all`.
-Second, the flag `-B` means that `Make` will consider all dependencies as outdated and hence will rebuild everything.
+Second, the flag `-B` means that `Make` will consider all dependencies as outdated and will hence rebuild everything.
 Third, repro constructs `Make` targets so that if you supply `DOCKER=TRUE` they are executed within the `Docker` image of the project.
@@
 This interplay between `Docker` and `Make` is relatively complicated because it resembles a chicken or egg problem.
-We have computational steps (`Make`) which depend on the software environment (`Docker`) for which we again have computational steps that create it.
+We have computational steps (`Make`) that depend on the software environment (`Docker`) for which we again have computational steps that create it.
 Users only require a deeper understanding of this interdependence when they either want to have more complex computational recipes than rendering an `R Markdown` or require other software than R packages.
@@
 By default we rely on the "verse" images provided by the Rocker project [@boettigerIntroductionRockerDocker2017].
-These contain (among other software) the packages `tidyverse`, `rmarkdown` and a complete LaTeX installation, which makes these images ideal for the creation of scientific manuscripts.
+These contain (among other software) the packages `tidyverse`, `rmarkdown`, and a complete LaTeX installation, which makes these images ideal for the creation of scientific manuscripts.
 Users can choose which R version they want to have inside the container by changing the version number in line 1 to the desired R version number.
 By default, the R version corresponds to the locally installed version on which `repro::automate()` was called the first time.
-The build date is used to install packages in the version which was available on CRAN on this specific date and can also be changed.
+The build date is used to install packages in the version that was available on the Comprehensive R Archive Network on this specific date and can also be changed.
 By default, this date is set to the date on which `repro::automate()` was called the first time. 
@@
 The second automatically generated `Dockerfile` is `.repro/Dockerfile_packages`.
-Whenever `repro::automate()` is called, repro gathers all R packages from all `.Rmd` files and figures out whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.
-
+Whenever `repro::automate()` is called, `repro` gathers all R packages from all `.Rmd` files and determines whether they should be installed from CRAN or GitHub fixed to the date specified in `Dockerfile_base`.
 Finally, there is one manually edited `Dockerfile`: `.repro/Dockerfile_manual`.
 It is blank by default and can be used to add further dependencies outside of R, like system libraries or external software.
-
 Using `Docker` may require you to install software on an operating system that may not be familiar to you.
@@
 The most convenient way to install software on Ubuntu is through its package manager `apt`.
-
-If the following snippet is added to `.repro/Dockerfile_manual` the `Docker` image will have, for example, Python installed.
-Other software is installed identically only the software name is exchanged.
+If the following snippet is added to `.repro/Dockerfile_manual`, the `Docker` image will have, for example, Python installed.
+Other software is installed identically, only the software name is exchanged.
 
@@
 `Docker` eventually requires a single `Dockerfile` to run, so `repro::automate()` simply concatenates the three `Dockerfiles` and saves the result into the main `Dockerfile` at the top level of the R project.
-
 With this approach, users of `repro` can build complex software environments and implement complex file dependencies.
-The standard `repro` metadata only makes sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
+The standard `repro` metadata only make sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
 If you can formulate the creation of dependencies in terms of computational steps, e.g.
@@
 The `Makefile` that `repro` creates is only a template, and you are free to change it.
-However, make sure never to remove the following two lines:
+However, make sure you never remove the following two lines:
 
@@
 
-The file `.repro/Makefile_Docker` does again contain a rather complicated template which you could but usually should not modify.
+The file `.repro/Makefile_Docker` does again contain a rather complicated template that you could, but should usually not modify.
 This `Makefile` coordinates the interplay between `Make` and `Docker` and contains targets for building (with `make docker`) and saving (with `make save-docker`) the `Docker` image.
 Additionally, it provides facilities to execute commands within the container.
-If you write a computational recipe for a target, it will be by default evaluated using the locally installed software.
-To instead evaluate commands inside the `Docker` image, you should wrap them in `$(RUN1) command $(RUN2)`, like in this example, which is identical to the first `Make` example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
+If you write a computational recipe for a target, it will be evaluated using the locally installed software by default.
+To evaluate commands inside the `Docker` image instead, you should wrap them in `$(RUN1) command $(RUN2)`, as done in this example, which is identical to the first `Make` example we gave above except for the addition of `$(RUN1)` and `$(RUN2)` in l.
 2:
+
 ```bash
@@
 ```
-It behaves exactly as the in the first `Make` example, the script `R/simulate.R` is run using the locally installed R.
+It behaves exactly as in the first `Make` example, the script `R/simulate.R` is run using the locally installed R.
 Because this translates simply to:
@@
 
-To summarise, `repro` provides researchers with an easy to use solution for dependency tracking (in the form of `Make`) and software management (using `Docker`) without the necessity to learn both tools.
+To summarize, `repro` provides researchers with an easy-to-use solution for dependency tracking (in the form of `Make`) and software management (using `Docker`) without the necessity to learn both tools.
 Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automation.
@@
 Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
-Preregistration seeks to reduce "researcher degrees of freedom" [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could bias the reported results.
-Preregistrations, however, have several shortcomings.
-One shortcoming is that written study plans are often multi-interpretable and often do not achieve the level of unambiguity that are required to effectively zero the researcher degrees of freedom.
-Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
+Preregistration seeks to reduce researchers' degrees of freedom [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, that could bias the reported results.
+Preregistration, however, have several shortcomings.
+One is that written study plans are often interpretable in multiple ways and often do not achieve the level of unambiguity that is required to effectively set the researcher degrees of freedom, to zero.
+Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, their results vary considerably [@silberzahnManyAnalystsOne2018].
 The current best practice to ensure that the preregistrations are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
 These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
-However, templates cannot remove researchers degrees of freedom completely because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
-This ambiguity causes a second problem, that is, comparing the preregistration and the resulting publication to decide if and how researchers deviated from the preregistration.
-This task is difficult because it impossible to decide without additional information if the analysis actually was carried out differently or just described differently.
+However, such templates cannot remove the degrees of freedom completely because it is impossible to verbally describe every detail of an analysis for any but the most straightforward analysis.
+This ambiguity causes a second problem, namely, comparing the initial plan and the resulting publication to decide if and how researchers deviated from the preregistration.
+This task is difficult because it is impossible to decide without additional information weather the analysis was actually carried out differently or just described differently.
 The comparison is made impossible if the authors of the preregistration were intentionally vague.
-Even when the researcher was faithful to the preregistration, the reader may reach opposite conclusions because the reader has to compare two different texts, which may be worded differently or describe the same thing in varying levels of detail.
-A third limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
+Even if the researchers were faithful to the preregistration, readers may reach opposite conclusions because they have to compare two different text that may be worded differently or describe the same thing in varying levels of detail.
+A third limitation is that preregistrations are susceptible to non-reproducibility, just like primary research.
 To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
-Even when researchers have gone to great length in preregistering an analysis script,
+Even when researchers have gone to great lengths in preregistering an analysis script,
 they sometimes inexplicably fail to reproduce their own results.
-For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistics than they reported initially (see Footnote 7).
+For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code resulted in different test statistics than they reported initially (see their Footnote 7).
 A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
-For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables or more parameters than observed data.
+For example, a verbal description of a statistical model may be unidentified, e.g., if it includes reciprocal paths between variables or more parameters than observed data.
 Conversely, a model may be misspecified in a major way; for example, by omitting direct effects when the research question is about mediation, thus leading to a model with an unacceptable fit.
@@
 Instead of describing the analysis in prose, researchers include the code required to conduct the analysis in the preregistration.
-We term this approach of writing and publishing code at the preregistration stage *preregistration as code (PAC)*.
+We term this approach of writing and publishing code at the preregistration stage *Preregistration as Code (PAC)*.
 PAC has the potential to eliminate researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
@@
 For a PAC, researchers can write a reproducible, dynamically generated draft of their intended manuscript at the preregistration stage.
-This already includes most typical sections, such as Introduction, Methods, and Results.
-These results are initially based on simulated data with the same structure as the data they expect to obtain from their experiments.
-For guidance on how to simulate data, see @morrisUsingSimulationStudies2019, @paxtonMonteCarloExperiments2001 and @skrondalDesignAnalysisMonte2000 as well as the R packages [`simstudy`](https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html)[@simstudy] and [`psych`](https://personality-project.org/r/psych/help/sim.html)[@psych].
+This already includes most of the typical sections, such as introduction, methods, and results.
+These results are initially based on simulated data with the same structure as the data the authors expect to obtain from their experiments.
+For guidance on how to simulate data, see @morrisUsingSimulationStudies2019, @paxtonMonteCarloExperiments2001, and @skrondalDesignAnalysisMonte2000, as well as the R packages [`simstudy`](https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html), [@simstudy] and [`psych`](https://personality-project.org/r/psych/help/sim.html), [@psych].
 
@@
 Reproducibility is of utmost importance at this stage since the preregistration must produce valid results at two points in time, once before data collection and once after data collection.
-As outlined before, reproducibility builds upon four pillars (Version control, Dynamic document generation, Dependency tracking, Software management)
-We can therefore only recommend PAC when the dangers to reproducibility we described above are eliminated.
+As outlined before, reproducibility builds upon four pillars (version control, dynamic document generation, dependency tracking, and software management).
+To use PAC the dangers to reproducibility we described must be  eliminated.
 
@@
 A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
-In an informal literature search (we skimmed the first 300 results with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we have found close to a dozen published articles which did include some form of script as part of their preregistration.
+In an informal literature search (we skimmed the first 300 results of Google Scholar with the keywords `("pre registration"|"pre-registration"|preregistration)&(code|script|matlab|python|"R")`) we only found close to a dozen published articles that did include some form of script as part of their preregistration.
 Though the notion of preregistering code has been around for a while [cf. @steegenMeasuringCrowdAgain2014],
-it has not gained much traction---perhaps because, until now, this has constituted an extra non-standard step in the research process.
+it has not gained much traction---perhaps because, to date, this has constituted an extra non-standard step in the research process.
 This tutorial integrates the preregistration of code into the reproducible research workflow
@@
 
-## Advantages of PAC over traditional preregistration
-
-We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
-First, the PAC is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
-Relatedly, writing a preregistration in the form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
+## Advantages of PAC Over Traditional Preregistration
+
+We believe that pairing PAC with the workflow presented above offers five advantages over classical preregistration.
+First, PAC is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
+Relatedly, writing the preregistration in the form of a research article has the advantage that researchers usually have great expertise in writing a scientific manuscript anyway.
 By contrast, filling out a preregistration template is a novel type of scientific writing for many.
-Second, the PAC is a tool for study planning.
-When all steps to be executed are documented clearly, the execution of the study will proceed more efficiently than when every step is planned in medias res.
-Third, a PAC removes any ambiguity regarding the translation of verbal analysis plans into code.
-A PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
+Second, PAC is a tool for study planning.
+If all steps to be executed are documented clearly, the carrying out of the study will proceed more efficiently than when every step is planned in medias res.
+Third, PAC removes any ambiguity regarding the translation of verbal analysis plans into code.
+PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
 Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
-PAC, therefore, minimizes researchers degrees of freedom more effectively than standard preregistration [@bakkerEnsuringQualitySpecificity2020].
-Fourth, despite its rigour, PAC does accommodate data-dependent decisions if these can be formulated as code.
+PAC, therefore, minimizes researchers degrees of freedom more effectively than standard preregistration does [@bakkerEnsuringQualitySpecificity2020].
+Fourth, despite its rigor, PAC accommodates data-dependent decisions if these can be formulated as code.
 Researchers can, for example, formulate conditions (e.g., in the form of if-else-blocks) under which they prefer one analysis type over the other.
@@
 Another example of data-dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
-Decisions that do not lend themselves to formulation in code, e.g.
-visual inspection, must still be described verbally or be treated as noted in the section [Deviating from the preregistration and exploration].
+Decisions that do not lend themselves to formulation in code, e.g. visual inspection, must still be described verbally or be treated as noted in the next section.
 With PAC researchers can precisely calibrate how much they want to explore.
@@
 
-## Deviating from the preregistration and exploration
+## Deviating from the Preregistration and Exploration
 
@@
 
-Deviations from the preregistration are sometimes demonized as if encountering unexpected aspects would invalidate a whole carefully crafted study [@szollosi_is_2020].
-However, we share the common view that a deviation from a preregistration is not a problem [@nosekPreregistrationHardWorthwhile2019]; it is a problem if these changes are not made transparent.
+Deviations from the preregistration are sometimes maligned as if encountering unexpected aspects would invalidate an entire carefully crafted study [@szollosi_is_2020].
+However, we share the common view that deviation from a preregistration is not a problem [@nosekPreregistrationHardWorthwhile2019]; it is a problem if these changes are not made transparent.
 In fact, we expect that most PACs will require changes after data have been collected.
 Often deviations provide an opportunity to learn from the unexpected.
-Transparency is essential in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumption [@meehlTheoreticalRisksTabular1978].
+Transparency is essential in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumptions [@meehlTheoreticalRisksTabular1978].
 
@@
 For them, this change simply does not affect the core of their theory but was an auxiliary assumption, and they might be right.
-Another researcher, however, who, e.g.
-hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because the change to the preregistration affects their theory directly.
-With transparent reporting, it is left to the reader to make up their own judgement.
-
-[^summarize-changes]: In section [Preregistration as Code --- a Tutorial] we have conducted an actual PAC and summarize the changes we made to the preregistered code in the discussion.
+Another researcher, however, who hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because the change to the preregistration affects their theory directly.
+With transparent reporting, it is left to the readers to make up their own judgement.
+
+[^summarize-changes]: In section [Preregistration as Code --- a Tutorial] we conducted an actual PAC and summarize the changes we make to the preregistered code in the discussion.
 
@@
 We must differentiate here between two kinds of exploration, and neither of them is limited by PAC.
-The first and more traditional meaning refers to ad hoc statistical procedures and post hoc explanations of the results.
-Such traditional exploratory results should merely be explicitly declared in the manuscript to allow appropriate judgment of the results [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
-The second kind of exploration is through procedurally well-defined exploration with exploratory statistical models, which are standard in machine learning [@brandmaier22mlsem].
-These highly-parameterized models, which are difficult to preregister traditionally, can be preregistered with PAC like any other model.
-We specifically recommend PAC for those exploratory statistical models.
-The merit of preregistration in such cases is to communicate precisely how much exploration was done, a piece of information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.
-
-## What you need to know before writing a PAC
-
-When planning a study and hence when writing a PAC, researchers ought to strive to meet three goals:
+The first and more traditional meaning of exploration involves to ad hoc statistical procedures and post hoc explanations of the results.
+Such traditional exploratory findings should merely be explicitly declared in the manuscript to allow their appropriate evaluation [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
+The second kind of exploration is through procedurally well defined exploration with exploratory statistical models are standard in machine learning [@brandmaier22mlsem].
+These highly-parameterized models, which are difficult to preregister traditionally, but can be preregistered with PAC like any other model.
+We specifically recommend PAC for such exploratory statistical models.
+The merit of preregistration in these cases is to communicate precisely how much exploration was done, a piece of information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.
+
+## What You Need to Know Before Writing a PAC
+
+When planning a study and writing a PAC, researchers ought to strive to meet three goals:
 First, the plan should be comprehensive.
 That is, researchers need to state the research question and describe the study design and analysis in detail.
-Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
-Third, the plan should be efficient; the plan should only require the necessary amount of resources.
+Second, the plan should be effective, meaning that the researchers should reasonably expect it to answer the research question at hand.
+Third, the plan should be efficient; i.e., it should only require the necessary amount of resources.
 
@@
 Standards for scientific writing are designed to result in understandable and comprehensive research articles and can serve the same purpose for writing a preregistration.
-Standards for scientific writing [e.g. @apa7] require that research articles are described in such detail that any researcher that is familiar with the field is able to repeat a given study.
-They provide, hence, an excellent vehicle for planning a study.
-One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
+Such standards [e.g., @apa7] require that research articles describe a study in such detail that any researcher that is familiar with the field is able to repeat it.
+These well-known standards can thus provide an excellent template for planning a preregistered study.
+One particular advantage is introduction's and theoretical background's embedding in existing literature where that provide the basis for the study's design.
 Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
-The comprehensiveness and completeness of published scientific reports is sometimes limited by strict word count limits that some journals impose.
+The comprehensiveness and completeness of published scientific reports is sometimes limited by the strict word count limits that some journals impose.
 One may argue that PAC inherits this limitation if the same word count limit applies.
-However, in such case we suggest using reproducible appendix materials to keep the core manuscript concise yet the preregistration comprehensive and exhaustive.
-
-Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
+However, in such cases, we suggest using reproducible appendix materials to keep the core manuscript concise, yet the preregistration comprehensive and exhaustive.
+
+Another factor that ensures a comprehensive plan is writing their analysis based on simulated data.
 Simulating data offers various benefits in the planning phase of a study.
-It forces the researcher to think about and address data analytic problems on a conceptual level before they make an effort to acquire data [@axelrodAdvancingArtSimulation1997].
-It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
-Writing incomplete code is technically impossible, or at least researchers would notice if something is omitted (e.g. if they forgot to aggregate the data).
-
-Peer review of preregistration is a method that also can ensure comprehensiveness e.g.
-a registered report [an article format in which the preregistration is reviewed and the article accepted "in principle", @nosekRegisteredReports2014; @chambersWhatNextRegistered2019].
-In such registered report peer-reviewing a manuscript styled PAC might be more familiar to the reviewers that ensure a comprehensive study plan.
+It forces the researcher to think about and address data-analytic problems on a conceptual level before they make an effort to acquire data [@axelrodAdvancingArtSimulation1997].
+It also affords researchers with a simple technical check whether the analysis code will run without failure [@braiekTestingMachineLearning2020] (e.g., does the code also work with missing values, extreme outliers, etc.).
+Writing incomplete code is technically impossible, or at least the researchers would notice at this early stage if they overlooked something (e.g. if they forgot to aggregate the data).
+
+Peer review of preregistration is a method that also can ensure comprehensivenes.
+For example, registered reports [an article format in which the preregistration is reviewed and the article accepted for publication "in principle," @nosekRegisteredReports2014; @chambersWhatNextRegistered2019].
+In such registered report peer-reviewing a manuscript-styled PAC may well make it easier for the reviewers that ascertain and evaluate a comprehensive study plan.
 
@@
 
-When the comprehensiveness criterion is met, researchers can check if the analysis is effective at recovering the true parameters of the simulation.
+If the comprehensiveness criterion is met, researchers can check weather the analysis is effective in recovering the true parameters of the simulation.
 If an analysis fails to recover the correct values in idealized (simulated) conditions,
@@
 Because the simulation describes the researchers' expectations about real-world conditions,
-it is the best way to assess the effectiveness of the analysis without actually gathering the data.
+it is the best way to assess the effectiveness of the analysis without actually collecting the data.
 A simulation of data requires the researchers to translate their expectations about the underlying process into concrete computational terms, e.g.
-assumptions about which distribution does the dependent variable follow or how do the independent variables influence each other and the dependent variable.
-If a researcher experiences difficulty in specifying such expectations, that may be an indication that further literature research, or exploratory pilot studies, are required.
+assumptions about the distribution  the dependent variable follows or how the independent variables influence each other and the dependent variable.
+If a researcher experiences difficulty in specifying such expectations, it may be an indication that further literature research or exploratory pilot studies are required.
 
@@
 In the simplest case, they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
-@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not increase the planned sample size on average but could help to invest resources more efficiently.
+@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not increase the planned sample size on average but can help to invest resources more efficiently.
 Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
 
-### The writing process
+### The Writing Process
 
@@
 When they have outlined the studies design, they should begin writing code.
-If there is no other code structure available, we suggest writing a minimum of three functions for each planned hypothesis, one for the planned analysis, one for simulating the respective data and one for reporting the results.
-We capture the tasks here in the form of explicit R-functions, but researcher may use any form (e.g. scripts) as long as it serves to simulate, analyze, and report.
-The functions shown here are empty (they do not contain executable code), and the interested reader may find function filled with some life in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.
+If there is no other code structure available, we suggest writing a minimum of three functions for each planned hypothesis, one for the planned analysis, one for simulation of the respective data, and one for reporting the results.
+We capture the tasks here in the form of explicit R functions, but researchers may use any form (e.g., scripts) as long as it serves to simulate, analyze, and report.
+The functions shown here are empty (they do not contain executable code), but the interested reader can find working functions in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.
 
@@
 We, therefore, simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
-Using the simulated dataset as a placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results, which can serve as PAC.
+Using the simulated dataset as a placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results that can serve as a PAC.
 
@@
 
-As soon as they have written `planned_anaylsis()` and `simulate_data()` they can iteratively improve both functions, e.g.
+As soon as we have written `planned_anaylsis()` and `simulate_data()` we can iteratively improve both functions, e.g.
 until `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
@@
 
-#### Alternatives to simulated data
-
-If simulating data proves to be too challenging, researcher can delay their preregistration until they have gathered data and repurpose the empirical data.
-In order to do so the researcher must walk a fine but important line.
-Of course, researchers must not analyze the data before they write the preregistration.
-Nevertheless, they may obfuscate the real data (before actually looking at it) by shuffling all variables per each column (independently from each other) in cross-sectional data or otherwise creating synthetic data that allows them to test the function `planned_analysis()` without actually revealing the observed research outcomes.
+### Alternatives to simulated data
+
+If simulating data proves to be too challenging, researchers can delay their preregistration until they have gathered data and repurpose the empirical data.
+In doing so, the researchers must walk a fine but important line.
+Of course, they must not analyze the data before they write the preregistration.
+Nevertheless, they may obfuscate the real data (before actually looking at them) by shuffling the order of rows for each variable (independently of each other) in cross-sectional data or otherwise by creating synthetic data that allow them to test the function `planned_analysis()` without actually revealing the observed research outcomes.
 Preregistration after data collection is common for secondary data analysis of data obtained by other research groups [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
 We argue that it is still an eligible preregistration.
-Guidelines for clinical trials also recommend analysis of blinded data  to test the practicality of the preregistration [@ICH1998].
-Since the data is obfuscated, virtually no information are revealed to the researchers, and they can not employ their researcher degrees of freedom effectively to influence the results.
-Shuffling leaves only univariate information, so this procedure is only applicable for bi-or multivariate hypotheses.
-The researcher still can access the information about means or proportions (e.g.
-how many participants belonging to group "A" are in the dataset) but are unbiased for all relations between variables (e.g.
-members of group "A" have a greater mean in variable "Z").
-To be clear, any exploration of the data before preregistration (PAC or otherwise) renders the preregistration invalid and amounts to scientific (albeit sometime unintentional) misconduct.
-
-Another approach that can replace but also complement simulated data is to conduct a pilot study [@thabaneTutorialPilotStudies2010] and use the resulting data to preregister.
+In fact, guidelines for clinical trials also recommend analysis of blinded data  to test the practicality of the preregistration [@ICH1998].
+Since the data is obfuscated, virtually no information is revealed to the researchers, and they can not employ their degrees of freedom effectively to influence the results (see section above).
+Shuffling leaves only univariate information, so this procedure is only applicable for bi- or multivariate hypotheses.
+The researcher can still access the information about means or proportions (e.g., the number of participants belonging to group "A" are in the dataset), but are unbiased for all relations between variables (e.g., members of group "A" have a greater mean in variable "Z").
+To be clear, any exploration of the data before preregistration (PAC or otherwise) renders the preregistration invalid and amounts to (albeit sometimes unintentional) scientific misconduct.
+
+Another approach that can replace but also complement simulated data is to conduct a pilot study [@thabaneTutorialPilotStudies2010] and use the collected data to preregister.
 A pilot study has obvious advantages for study planning, since it lets the researcher evaluate the feasibility of many assumptions.
-However, we must warn our readers, while piloting is more traditional than our approach of obfuscating the data before preregistration, it is not innocuous to use the data resulting from the pilot.
-
-When the researchers are satisfied with the function `planned_analysis()`, they can think about how they would like to report the analysis results via tables, plots, and text.
+However, we must warn our readers, that while piloting is more traditional than our approach of obfuscating the data before preregistration, it is not innocuous to use the data resulting from the pilot study.
+
+When the researchers are satisfied with the function `planned_analysis()`, they can think about the way the would they would like to report the analysis results via tables, plots, and text.
 The implementation of this reporting should be in the function `report_analysis()`.
@@
 The idea is to automatically generate the full results section from the analysis.
-This way, the preregistration specifies not only the computation but also how the computation's results are reported.
-Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R objects to help with this objective.
+This way, the preregistration not only specifies the computation but also how the its results are reported.
+Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R goal to help with this objective.
 Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
-The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it generates not only tables but also a straightforward interpretation of the effects as actual prose (e.g.
-it verbally quantifies the size of an effect).
+The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables but also a straightforward interpretation of the effects as actual prose (e.g., it verbally quantifies the size of an effect).
 
@@
 
-To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration.
-After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data) and write the discussion.
-
-#### When is PAC applicable
-
-Every study that can be preregistered, can be preregistered with PAC.
+To summarize, researchers need to write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration.
+After they gathered the actual data, they can replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data), and write the discussion.
+
+#### When Is PAC Applicable?
+
+Every study that can be preregistered can also be preregistered with code.
 The only difference is that PAC demands a comprehensive and complete analysis plan.
-With traditional preregistration, researchers may omit details and get away with decisions that are made only after having seen the data, thus, potentially biasing statistical results.
-Some researchers may not feel ready to decide every detail of the analysis plan before data collection.
-In such case, we recommend making decisions based on the best available knowledge, and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in the section [Deviating from the preregistration and exploration] of the final manuscript.
+With traditional preregistration, researchers are able to omit details and get away with decisions that are made only after having seen the data, thus, potentially biasing statistical results.
+Some researchers may not feel ready to decide on every detail of the analysis plan before data collection.
+In such cases, we recommend making decisions based on the best available knowledge, and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in Section [Deviating from the preregistration and exploration] in the final manuscript.
 
 Beyond standard preregistrations in psychological research, we believe that two applications specifically lend themselves to PAC.
-First, pregistrations of clinical trials (there called statistical analysis plan, @ICH1998) typically describe analyses in exhaustive detail.
-Those preregistrations typically contain a detailed description of how results will be presented, including shells of tables and graphics [@yuanGuideStatisticalAnalysis2019].
+First, pregistrations of clinical trials (called statistical analysis plans, @ICH1998) typically describe analyses in exhaustive detail and typically contain a detailed description of how results will be presented, including shells of tables and graphics [@yuanGuideStatisticalAnalysis2019].
 PAC may significantly reduce the required workload while maintaining (and exceeding) the required standards for preregistering a clinical trial.
@@
 
-# Preregistration as Code --- a Tutorial
-
-We argued that PAC has several advantages over classic preregistration and have outlined how to implement it.
-To illustrate how PAC works in practice and help researchers implementing PAC themselves, we have conducted a PAC that serves as a guiding example.
+# Preregistration as Code: Tutorial
+
+We ahve argued that PAC has several advantages over classic preregistration and have outlined its implementation.
+To illustrate how PAC works in practice and help researchers to implement PAC themselves, we have conducted one that serves as a guiding example.
 We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
@@
 
-As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary background information and literature to understand the context and purpose of the research question.
+As usual, the authors state why they are interested in their research question in the "Introduction" section and provide the necessary background information and literature to understand the context and purpose of the research question.
 This example is drastically shortened for illustration purposes:
@@
 
-[^caution-mww]: The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employed a rank based alternative in the New England Journal of Medicine in 2005].
+[^caution-mww]: The t-test and Mann-Whitney-Wilcoxon test are arguably the most often used hypothesis tests (according to [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005] reports that 26% of all studies employed a t-test and 27% employed a rank-based alternative in the *New England Journal of Medicine* in 2005).
 The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.
+The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann-Whitney-Wilcoxon overestimates effect size with increasing sample size and skewness.
 
@@
 This function illustrates two advantages of PAC.
-First, a PAC can easily include data-dependent decisions creating different analysis branches under different conditions.
+First, a PAC can easily include data-dependent decisions by creating different analysis branches under different conditions.
 Second, it highlights how difficult it is to describe a statistical analysis precisely.
@@
 Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
-Still, these details can make a difference and provide unwanted researchers' degrees of freedom.
-
-The function `planned_analysis()` together with the function `simulate_data()` (not shown here) can be used to justify the planned sample size.
-To that end, `simulate_data()` is repeatedly called with increased sample size and the proportion of significant results (power) is recorded.
-The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
+Still, these details can make a difference and unwanted researchers' degrees of freedom.
+
+Together with the function `simulate_data()` (not shown here), the function `planned_analysis()` can be used to justify the planned sample size.
+To that end, `simulate_data()` is repeatedly called with increased sample sizes and the proportion of significant results (power) is recorded.
+The results for such a Monte Carlo simulation for this example are visualized in Figure \@ref(fig:power).
 The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R).
-The next snippet shows how we included the results dynamically into the preregistration (the origin of the R-variables `minn`, `choosen_power`, and `choosen_d` is not shown).
+The next snippet shows how we integrated the results dynamically into the preregistration (the origin of the R-variables `minn`, `choosen_power`, and `choosen_d` is not shown).
 
@@
 
-Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
-
-We implemented a mechanism that only uses simulated data when the actual data is not yet available (in this example, if the file `data/data.csv` does not exist) for the results section.
+Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate further relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
+
+We implemented a mechanism that only uses simulated data when the actual data are not yet available (in this example, if the file `data/data.csv` does not exist) for the results section.
 This mechanism also warns readers if these results are based on simulated data.
-The warning is coloured red to avoid any confusion between mock and actual results.
-As soon as the actual data are available, the simulated data is no longer used, and the results represent the actual empirical results of the study.
+The warning is colored red to avoid any confusion between mock and actual results.
+As soon as the actual data are available, the simulated data are no longer used, and the results represent the actual empirical results of the study.
 
@@
 
-Naturally we did not access the data when we wrote this code initially.
-We therefore did not know is the exact format the data would have.
-So we did need to change our preregistration after accessing the data to include i.a. the recoding of gender (lines 17-22) and the items (lines 23-26).
-We invite the reader to evaluate the changes we made to the preregistratered code on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
+Naturally we did not access the data when we initially wrote this code.
+We therefore did not know the exact format the data would have.
+This means that we did need to change our preregistration after accessing the data to include i.e. the recoding of gender (lines 17-22) and the items (lines 23-26).
+We invite the reader to evaluate the changes we made to the preregistered code on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
 
@@
 
-This example preregistration covers a single study with a single hypothesis.
-To organize studies with multiple hypotheses, we suggest multiple (possibly numbered in accordance with the hypotheses, e.g. 1.2, 2.3 etc.) `planned_analysis()` and `report_analysis()` functions.
+This example of a preregistration covers a single study with a single hypothesis.
+To organize studies with multiple hypotheses, we suggest multiple `planned_analysis()` and `report_analysis()` functions (possibly numbered in accordance with the hypotheses, e.g. 1.2, 2.3 etc.).
 Computational steps that apply to multiple analyses should be extracted into their own function, for example, those that perform data cleaning or extract results from a statistical model in an identical way.
@@
 
-After you have linked a GitHub repository, you trigger the archivation by creating a GitHub release.
+After you have linked a GitHub repository, you trigger the archival by creating a GitHub release.
 To create GitHub release, navigate to GitHub:
@@
 
-And click on Releases → Draft a new release.
+Then click on Releases → Draft a new release.
 Here you can add all relevant binary files but at least a rendered version of the manuscript and the Docker image.
 
-We realize that PACs are challenging.
+Taken together, we realize that PACs are challenging.
 Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove impractical for a PAC.
-Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not-preregistered research but as a complementary approach.
-We are convinced that many reasons justify changing the preregistered code, but are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
+Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or non-preregistered research but as a rather complementary approach.
+We are convinced that many reasons can justify changing the preregistered code, but equally certain that readers should judge for themselves weather they find these deviations from the preregistration meaningful.
 PAC enables just that.
-Researcher can change their plans and possibly defend their decision, e.g. in the discussion, while readers of the manuscript can clearly see what was changed through version control.
+Researchers can change their plans and possibly defend their decision to do so, e.g., in the discussion, while readers of the manuscript can clearly see what was changed through version control.
 
@@
 
-Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits nicely together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
-Automated research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
-
-Central to these innovations is a property we call "reusability" which we hope this workflow promotes.
+Increased automation is increasingly recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits nicely together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
+Automated research projects promise a wide range of applications, among them PAC [potentially to be submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], direct replication [@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], executable research articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations such as the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].
+
+Central to these innovations is a property we call "reusability," fully promoted by the present workflow.
 Reusable code can run on different inputs from a similar context and produce valid outputs.
-This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
+This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _built-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
 The reproducible workflow we present here is heavily automated and hence promotes reusability.
@@
 These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
-The workflow we present here rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to all but guarantee reproducibility.
+The workflow we present here rests on four software solutions, (1) version control, (2) dynamic document generation, (3) dependency tracking and (4) software management to guarantee reproducibility.
 We first demonstrated how to create a reproducible research project.
@@
 We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations,
-such as the preregistration as code (PAC).
+such as preregistration as code (PAC).
 A PAC incorporates all planned analysis code into a reproducible manuscript that already renders the expected results.
@@
 Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
-Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
+Because the same dynamic document serves as preregistration and final manuscript, researchers, reviewers and, editors only have to revise one document.
 To editors, reviewers, and readers, it is obvious how the researchers deviated from the preregistration because everything there is a single "source of truth"---dynamic manuscript including code---under version control.
 
-The likelihood of those deviations, while never unavoidable, is decreased by using simulated data to verify the suitability of the proposed analysis plan.
+The likelihood of such deviations, remains unavoidable, but is decreased by using simulated data to verify the suitability of the proposed analysis plan.
 This analysis plan may contain data-dependent decisions to increase the flexibility of research plans.
@@
 First, some limitations are by design and are unlikely to be changed soon.
-Our workflow does inherit these from the software it relies on and the fundamental design principles these share with the workflow and `repro`.
-Second, there are other limitations that we hope development effort of the open-source community and, concerning `repro`, us will ease.
+Our workflow inherits these from the software it relies on and the fundamental design principles these share with the workflow and `repro`.
+Second, there are other limitations that we hope development efforts of the open-source community and, concerning `repro`, we will solve
 
@@
 To allow for automation, many researchers will have to search for alternative software.  
-The automation friendly software we present here has several technical but critical limitations.
-`Git` can track any files (files bigger than 1GB only with extensions) but can only compare text files (with endings like, `.txt`, `.csv`, `.R`,  or `.Rmd`), not binary files (with endings like `.docx`, `.exe` or `.zip`).
+The automation-friendly software we present here has several technical but critical limitations.
+`Git` can track any files (files larger than 1 GB only with extensions) but can only compare text files (with endings like, `.txt`, `.csv`, `.R`, `.py`,  or `.Rmd`), not binary files (with endings like `.docx`, `.exe`, or `.zip`).
 With `R Markdown`, tables and graphics are dynamically generated from code but cannot be crafted by hand.
-`Make` can orchestrate any programmable software but fails to automate software that will only obey a user clicking at the right buttons.
-`Docker` can ship software that runs on Linux and can be automatically installed, but that requirement precludes many commercial or closed source software.
+`Make` can orchestrate any programmable software but fails to automate software that will only obey a user clicking on the right buttons.
+`Docker` can ship software that runs on Linux and can be automatically installed, but that requirement precludes much commercial or closed-source software.
 
@@
 On the one hand, we firmly believe that automated reproducibility makes research more productive and collaboration easier.
-But, on the other hand, we expect researchers to invest considerable time in learning new tools and then must persuade their collaborators to do the same.
+But, on the other hand, we expect researchers to invest considerable time in learning new tools and to persuade their collaborators to do the same.
 Three arguments reconcile this apparent paradox.
 First, this change will not happen all at once.
-Automated reproducibility is an ideal that we believe has many advantages, but it is not an all or nothing decision.
-Researchers can pick up one skill at a time and then help their fellow collaborators do the same.
-Second, the upfront investment is one time (and efforts such as `repro` are underway to reduce it) and will pay dividends over many research projects.
+Automated reproducibility is an ideal that we believe has many advantages, but it is not an all-or-nothing decision.
+Researchers can pick up one skill at a time and then help their fellow collaborators to do the same.
+Second, the upfront investment is required once (and efforts such as `repro` are underway to reduce it) and will pay dividends over many research projects.
 Third, the move towards open software for research offers several benefits beyond enabling automated reproducibility [@schaffnerFutureScientificJournals1994; @fitzgeraldTransformationOpenSource2006; @chaldecottHistoryScientificTechnical1965; @sonnenburgNeedOpenSource2007].
@@
 `repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
-Despite our efforts to ensure this goal, the package is still in development and undoubtedly has limitations.
-The workflow we present here introduces several software solutions, which in itself present a threat to long-term reproducibility.
+Despite our best efforts to ensure this goal, the package is still in development and undoubtedly has limitations.
+The workflow we present here introduces several software solutions.
+This sheer number of dependencies represents in itself a threat to long-term reproducibility.
 To benefit from automatic and convenient reproduction, researchers must use `Git`, `Make`, and `Docker`.
-However, `Git` and `Make` are itself included in the `Docker` image created by `repro`.
+However, `Git` and `Make` are themselves included in the `Docker` image created by `repro`.
 Researchers can therefore employ the `Docker` image manually to download the `Git` repository and execute `Make` for full reproduction.
-In other words, the only hard requirement for reproduction and therefore the Achilles' heel for this workflow is `Docker`.
+In other words, the only hard requirement for reproduction and therefore its Achilles' heel, is `Docker`.
 The `Docker` approach has two vulnerabilities.
 First, and more importantly, the `Docker` image for the project and the `Git` repository have to remain available.
-The `Dockerfile` (the plain text description to build a `Docker` image), as opposed to the image, is insufficient because it relies on too many service providers (e.g.
-Microsoft R Application Network, Ubuntu Package Archive).
+The `Dockerfile` (the plain text description to build a `Docker` image), as opposed to the image, is insufficient because it relies on too many service providers (e.g., Microsoft R Application Network, Ubuntu Package Archive).
 We recommend archiving the `Git` repository and the `Docker` image with zenodo.org, a non-profit long-term storage for scientific data.
-The necessary steps for archival on zenodo.org are described at the end of section [Preregistration as Code --- a Tutorial].
+The necessary steps for archival on zenodo.org are described at the end of Section [Preregistration as Code --- a Tutorial].
 The second vulnerability is that even if the existence of the `Docker` image and `Git` repository is guaranteed, future researchers still require software to run the image.
-To that end, they can either rely on `Docker` itself or `Docker`-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
+To that end, they can either rely on `Docker` itself or `Docker`-compatible alternatives (e.g., `CoreOS rkt`, `Mesos Containerizer`, `Singularity`).
 The only way to remove the reliance on such external software is to turn the `Docker` image into a full operating system that subsequently can be installed and run on almost any modern computer.
@@
 In any case and without such precautions, do the `R Markdown`, the `Makefile` and the `Dockerfile` provide information that allows researchers to trace the computational steps and recreate the computational environment manually.
-The `Makefile`, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are, for whatever reason, unable to run `Make`.
-Hypothetically, even if `Docker` were unavailable one day, the `Dockerfile` still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
+The `Makefile`, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are unable to run `Make` for some reason.
+Hypothetically, even if `Docker` were to become unavailable one day, the `Dockerfile` still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
 
@@
 Open science practices are a continually evolving field where technical innovations foster changes in research practice.
-Open data is much more widespread thanks to online storage facilities; preregistration is possible because there are preregistration platforms and so forth.
-Similarly, we hope that fully automatic reproduction, e.g. with `repro` as technical innovation, increase scientific rigour and efficiency.
-
-This ideal of a fully automatic reproduction of research projects meets in practice a wide range of demands on how user-friendly and how powerful the software solutions are.
-Some may find `Make` too complicated or that `Docker` requires too much storage.
-Yet others may find that they require other programming languages or scale their computation across hundreds of computers, e.g. via high-performance computing clusters or cloud computing.
+Open data are much more widespread thanks to online storage facilities; preregistration is possible because there are preregistration platforms and so forth.
+Similarly, we hope that fully automatic reproduction, e.g., with `repro` as a technical innovation, increase scientific rigor and efficiency.
+
+In practice, this ideal of a fully automatic reproduction of research projects can conflict with the wide range of demands for more user-friendly and powerful software.
+Some may find `Make` too complicated or that `Docker` requires too much storage space.
+Yet others may find that they require other programming languages or want to scale their computation across hundreds of computers, e.g., via high-performance computing clusters or cloud computing.
 
@@
 However, there are alternatives for each of these elements that may fit better into an individual research project.
-`R Markdown` could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer`[@officer] or `officedown`[@officedown] to accommodate a wider arrange of journal submission standards.
+`R Markdown` could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer` [@officer] or `officedown` [@officedown] to accommodate a wider range of journal submission standards.
 Instead of using formal version control with `Git`, `repro` could automatically save snapshots for increasing user-friendliness.
@@
 Alternatively, `Docker` could be replaced by the more lightweight `renv` if no dependencies outside of R are considered crucial.
-`Docker` does not satisfy the requirements of many HPC environments, but Singularity was designed to avoid this limitation while still being compatible with `Docker` images.
-
-`repro`'s modular structure allows such alternative workflow, though they have not been implemented.
-Depending on the demand by users, we will implement some of these alternative workflows in `repro` and hope for broad adoption of computational reproducibility in the near future.
+`Docker` does not satisfy the requirements of many HPC environments, but `Singularity` was designed to avoid this limitation while still being compatible with `Docker` images.
+
+`repro`'s modular structure allows such alternative workflows, though they have not yet been implemented.
+Depending on the demand by users, we will implement some of them in `repro` and hope for broad adoption of computational reproducibility in the near future.
 

### 2021-11-15 abstract.Rmd

@@
 Computational reproducibility is the ability to obtain identical results from the *same* data with the *same* computer code.
-It is a building block for transparent and cumulative science because it enables the originator and other researchers, on other computers and later in time, to reproduce and thus understand how results came about while, at the same time, it avoids a variety of errors that may lead to erroneous reporting of statistical and computational results.
+It is a building block for transparent and cumulative science because it enables the originator and other researchers, on other computers and later in time, to reproduce and thus understand how results came about while avoiding a variety of errors that may lead to erroneous reporting of statistical and computational results.
 In this tutorial, we demonstrate how the R package `repro` supports researchers in creating fully computationally reproducible research projects with tools from the software engineering community.
 Building upon this notion of fully automated reproducibility we present several applications including the preregistration of research plans with code (Preregistration as Code, PAC).
-PAC eschews all ambiguity of traditional preregistration and offers several other advantages.
+PAC eschews all ambiguity of traditional preregistration and offers several more advantages.
 Making technical advancements that serve reproducibility more widely accessible for researchers holds the potential to innovate the research process to become more productive, credible, and reliable.

### 2021-11-15 manuscript.Rmd
remove footnote closes #265

@@
 acknowledgement: |
-  We would like to thank Maximilian Stefan Ernst for his contributions to the code for the simulation study. 
+  We would like to thank Maximilian Stefan Ernst for his contributions to the code for the simulation study. We are grateful to Julia Delius for her helpful assistance in language and style editing.
 authorcontributions: |
@@
 The credibility of empirical research results hinges on their objectivity.
-Objectivity in this context means, "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
+Objectivity in this context means, "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22].
 Only a reproducible result can be tested and understood by anyone.
@@
 Once the empirical data are available, they replace the simulated data, the results are then updated automatically, and the discussion can be written to finalize the report.
-
-[^popper]: In fact, @popperLogicScientificDiscovery2002's famous criterion of demarcation is built around "inter-subjective testing", a concept that he later generalized to inter-subjective criticism (Footnote *1, p.
-22).
-Somewhat confusingly, @popperLogicScientificDiscovery2002 uses the term "Reproducibility" for what we would call "Replication."
 

### 2021-11-15 manuscript.Rmd
actually run fake chunk to ensure correctness closes #275

@@
 
-```{r, eval=FALSE, echo=-1}
+```{r, eval=TRUE, echo=-1, results='hide'}
 # this is a fake chunk, the faked output comes the chunk below
-report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)
+report_analysis(planned_analysis(simulate_data(900, 8, 0.3, 10)))
 ```

### 2021-11-15 manuscript.Rmd
remove causal language closes #263

@@
 In the simplest case, they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
-@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis does not increase the planned sample size on average but can help to invest resources more efficiently.
+@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis may help researchers to invest resources more efficiently.
 Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].

### 2021-11-15 manuscript.Rmd
update affiliation closes #267

@@
   - name: Aaron Peikert
-    affil: 1, *
+    affil: 1, 2, *
     orcid: 0000-0001-7813-818X
   - name: Caspar J. van Lissa
-    affil: 2, 3
+    affil: 3, 4
     orcid: 0000-0002-0808-5024
   - name: Andreas M. Brandmaier
-    affil: 1, 4, 5
+    affil: 1, 5, 6
     orcid: 0000-0001-8765-6982
@@
     address: |
-      Department of Methodology & Statistics---Utrecht University, Faculty of Social and Behavioral Sciences, Utrecht, Netherlands
+      Humboldt-Universität zu Berlin, Berlin, Germany
   - num: 3
     address: |
+      Department of Methodology & Statistics---Utrecht University, Faculty of Social and Behavioral Sciences, Utrecht, Netherlands
+  - num: 4
+    address: |
       Open Science Community Utrecht, Utrecht, Netherlands
-  - num: 4
+  - num: 5
     address: |
@@
       Berlin, Germany, and London, UK
-  - num: 5
+  - num: 6
     address: |

### 2021-11-15 manuscript.Rmd
add inkscape

@@
     - references.bib
+  apt:
+    - inkscape
 output:

### 2021-11-15 manuscript.Rmd
use new version of image closes #269
Update manuscript.Rmd
Here is a proposal. I tried to change as little as possible to your original sentence.

@@
 With this paper, we pursue two aims.
-First, we want to introduce researchers to a notion of automated reproducibility that requires no manual steps, apart from the setup of the software. 
-Second, we present the opportunities such automated reproducibility affords to improve not only how research outputs are created and recreated but also how research itself is conducted.
+First, we want to introduce researchers to a notion of automated reproducibility that requires no manual steps, apart from the initial setup of the software environment. 
+Secondly, we discuss the possibilities that such automated reproducibility offers to improve not only the way research results are created and recovered, but also how this may change the general approach to a research project.
 With regard to the first goal, we show how to address four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
@@
 `Docker` provides a container in which the final report is built using dynamic document generation in `R Markdown`.
-Reproduced from @Peikert2019 licensed under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0).
+Adapted from @Peikert2019 licensed under [CC BY 4.0](https://creativecommons.org/licenses/by/4.0).
 "}

### 2021-11-16 manuscript.Rmd
consistent use of "bias"/"researcher degree of freedome"
closes #270
closes #271

@@
   Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions of those analyses, which then merely complement the PAC as a more readable summary.
-  PAC circumvents the shortcomings of ambiguous preregistrations that may give researchers undesired degrees of freedom in the execution of their analysis.
+  PAC circumvents the shortcomings of ambiguous preregistrations that may give researchers undisclosed degrees of freedom in the execution of their analysis.
   Reproducibility facilitated by automation therefore has a wide range of applications to accelerate scientific progress.
@@
 Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
-Preregistration seeks to reduce researchers' degrees of freedom [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, that could bias the reported results.
+In designing, executing, and analyzing a study, researchers are faced with a myriad of choices, often called researchers degrees of freedom.
+These may be used to reach statistical significance or inflate effect size estimates introducing opportunistic bias [@decosterOpportunisticBiasesTheir2015].
+Preregistration clarifies when and how researchers employ their degrees of freedom; otherwise, readers of a manuscript must presume what liberties the researchers took.
+The goal is transparency and not to restrict what researchers may do to gather or analyze their data.
+
 Preregistration, however, have several shortcomings.
-One is that written study plans are often interpretable in multiple ways and often do not achieve the level of unambiguity that is required to effectively set the researcher degrees of freedom, to zero.
+One is that written study plans are often interpretable in multiple ways.
 Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, their results vary considerably [@silberzahnManyAnalystsOne2018].
@@
 These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
-However, such templates cannot remove the degrees of freedom completely because it is impossible to verbally describe every detail of an analysis for any but the most straightforward analysis.
+However, such templates cannot ensure full transparency because it is impossible to verbally describe every detail of an analysis for any but the most straightforward analysis.
 This ambiguity causes a second problem, namely, comparing the initial plan and the resulting publication to decide if and how researchers deviated from the preregistration.
@@
 We term this approach of writing and publishing code at the preregistration stage *Preregistration as Code (PAC)*.
-PAC has the potential to eliminate researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
+PAC has the potential to eliminate undisclosed researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
 Moreover, it removes the necessity to write two separate documents:
@@
 
-The idea of submitting code as part of a preregistration is not new.
+The idea of submitting code as part of a preregistration is not new [e.g. @wichertsDegreesFreedomPlanning2016].
 A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
@@
 Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
-PAC, therefore, minimizes researchers degrees of freedom more effectively than standard preregistration does [@bakkerEnsuringQualitySpecificity2020].
+PAC, therefore, minimizes undisclosed researchers degrees of freedom more effectively than standard preregistration does [@bakkerEnsuringQualitySpecificity2020; @wichertsDegreesFreedomPlanning2016].
 Fourth, despite its rigor, PAC accommodates data-dependent decisions if these can be formulated as code.
@@
 Shuffling leaves only univariate information, so this procedure is only applicable for bi- or multivariate hypotheses.
-The researcher can still access the information about means or proportions (e.g., the number of participants belonging to group "A" are in the dataset), but are unbiased for all relations between variables (e.g., members of group "A" have a greater mean in variable "Z").
+The researcher can still access the information about means or proportions (e.g., the number of participants belonging to group "A" are in the dataset), but remain uninformed about relations between variables (e.g., members of group "A" have a greater mean in variable "Z").
 To be clear, any exploration of the data before preregistration (PAC or otherwise) renders the preregistration invalid and amounts to (albeit sometimes unintentional) scientific misconduct.
@@
 The only difference is that PAC demands a comprehensive and complete analysis plan.
-With traditional preregistration, researchers are able to omit details and get away with decisions that are made only after having seen the data, thus, potentially biasing statistical results.
+With traditional preregistration, researchers are able to omit details and get away with decisions that are made only after having seen the data, thus, potentially introducing opportunistic bias in statistical results.
 Some researchers may not feel ready to decide on every detail of the analysis plan before data collection.
@@
 Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
-Still, these details can make a difference and unwanted researchers' degrees of freedom.
+Still, these details can make a difference and provide undisclosed researchers' degrees of freedom.
 
@@
 The entire manuscript, including planned analyses, is preregistered.
-This way, all researcher's degrees of freedom are eliminated.
+This way, every use of a researchers' degree of freedom is disclosed.
 Once real data is gathered, the reproducible manuscript is (re-)created with the real data.

### 2021-11-16 manuscript.Rmd
move to footnote closes #273 closes #274

--- 2021-11-15 manuscript.Rmd
@@
 We therefore did not know the exact format the data would have.
-This means that we did need to change our preregistration after accessing the data to include i.e. the recoding of gender (lines 17-22) and the items (lines 23-26).
-We invite the reader to evaluate the changes we made to the preregistered code on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
+This means that we did need to change our preregistration after accessing the data to include i.e. the recoding of gender (lines 17-22) and the items (lines 23-26).[^prereg-changes]
+
+[^prereg-changes]: We invite the reader to evaluate the changes we made to the preregistered code.
+Either on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg...main) → "Files changed" or directly in Git with `git diff v0.0.1.1-prereg preregistration.Rmd`.
 

### 2021-11-16 manuscript.Rmd
refine tutorial scope closes #278

@@
 This tutorial assumes that the user is predominantly working in R, with the help of RStudio.
-It describes basic steps that we expect to be representative of a small-scale psychological research project. 
+It describes basic steps that we expect to be representative of a small-scale psychological research project that relies on no external software or multistage data processing (for those requirements see section [Advanced Features]).
 Of course, your specific situation might involve additional, more specialized steps. 

### 2021-11-16 manuscript.Rmd
where is the line? closes #272
the sentence before makes it clear already closes #264

@@
 We present automated reproducibility as best practice, a goal that is not always achieved due to limited resources, technical restrictions, or practical considerations, but is very much worth striving for.
-If some parts of the workflow prove to be unfeasible, we encourage the reader to find their own solutions and to concentrate on the feasible parts.
 
@@
 To be clear, any exploration of the data before preregistration (PAC or otherwise) renders the preregistration invalid and amounts to (albeit sometimes unintentional) scientific misconduct.
+The aim is transparency and the line that shall not be crossed is learning about the research outcomes, but what this means differs for each research project.
 

### 2021-11-16 manuscript.Rmd
added a few lines

--- 2021-11-16 manuscript.Rmd
@@
 
-Every study that can be preregistered can also be preregistered with code.
+Every study that can be preregistered and ultimately uses computer code for the statistical analysis, can also be preregistered with code.
 The only difference is that PAC demands a comprehensive and complete analysis plan.
-With traditional preregistration, researchers are able to omit details and get away with decisions that are made only after having seen the data, thus, potentially biasing statistical results.
-Some researchers may not feel ready to decide on every detail of the analysis plan before data collection.
+With traditional preregistration, researchers may (unintentionally) omit details and thus introduce decisions that need to be made only after having seen the data.
+We emphasize again that deviations from the preregistration are not only possible but actually likely.
+The advtange of PAC is that all deviations from the planned analysis as implemented in code are tracked, and the final analysis can be easily compared to the PAC.
+Importantly, we see PAC as a means to increase transparency not as a means to limit researchers in their statistical analyses.
+In practice, researchers may not feel ready to decide on every detail of the analysis plan before data collection.
 In such cases, we recommend making decisions based on the best available knowledge, and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in Section [Deviating from the preregistration and exploration] in the final manuscript.
@@
 Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
-Still, these details can make a difference and unwanted researchers' degrees of freedom.
+Still, these details can make a difference in the interpretation of statistical results and, thus, represent undisclosed researchers' degrees of freedom.
 
@@
 We therefore did not know the exact format the data would have.
-This means that we did need to change our preregistration after accessing the data to include i.e. the recoding of gender (lines 17-22) and the items (lines 23-26).[^prereg-changes]
-
-[^prereg-changes]: We invite the reader to evaluate the changes we made to the preregistered code.
-Either on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg...main) → "Files changed" or directly in Git with `git diff v0.0.1.1-prereg preregistration.Rmd`.
+This means that we did need to change our preregistration after accessing the data to include i.e. the recoding of gender (lines 17-22) and the items (lines 23-26).
+We invite the reader to evaluate the changes we made to the preregistered code on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
 

### 2021-11-17 manuscript.Rmd
slight rephrasing

@@
 The only difference is that PAC demands a comprehensive and complete analysis plan.
-With traditional preregistration, researchers may (unintentionally) omit details and thus introduce decisions that need to be made only after having seen the data.
-We emphasize again that deviations from the preregistration are not only possible but actually likely.
-The advtange of PAC is that all deviations from the planned analysis as implemented in code are tracked, and the final analysis can be easily compared to the PAC.
+With traditional preregistration, researchers may (unintentionally) omit details and thus introduce decisions that appear to be made after having seen the data.
+We emphasize again that deviations from the preregistration are not only possible but likely.
+The advantage of PAC is transparency.
+All deviations from the planned analysis as implemented in code are tracked, and the final analysis can be easily compared to the PAC.
 Importantly, we see PAC as a means to increase transparency not as a means to limit researchers in their statistical analyses.

### 2021-11-17 manuscript.Rmd
mention figure 3.1 closes #279
fix typo

--- 2021-11-16 manuscript.Rmd
@@
 For `Git` to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
-One way to do this is through the visual user interface in the RStudio `Git` pane.
+One way to do this is through the visual user interface in the RStudio `Git` pane (see Figure \@ref(fig:modified)).
 Click on the empty box next to the file you want to stage.
@@
 simulated_data.csv: R/simulate.R
-  $(RUN1) Rscript -e 'source("R/simulate.R")' $(RUN1)
+  $(RUN1) Rscript -e 'source("R/simulate.R")' $(RUN2)
 ```

### 2021-11-17 manuscript.Rmd
shorten discussion closes #261

@@
 Then, we illustrated how such a project could be reproduced---either by the original author and/or collaborators or by a third party.
-We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations,
-such as preregistration as code (PAC).
-A PAC incorporates all planned analysis code into a reproducible manuscript that already renders the expected results.
-These results are based on simulated data, so that there is no danger of peeking at the real data while writing the scripts.
-The entire manuscript, including planned analyses, is preregistered.
-This way, all researcher's degrees of freedom are eliminated.
+We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations, such as preregistration as code (PAC).
+In PAC the entire reproducible manuscript, including planned analyses adn results based on simulated data, is preregistered.
 Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
-Because the same dynamic document serves as preregistration and final manuscript, researchers, reviewers and, editors only have to revise one document.
-To editors, reviewers, and readers, it is obvious how the researchers deviated from the preregistration because everything there is a single "source of truth"---dynamic manuscript including code---under version control.
-
-The likelihood of such deviations, remains unavoidable, but is decreased by using simulated data to verify the suitability of the proposed analysis plan.
-This analysis plan may contain data-dependent decisions to increase the flexibility of research plans.
+PAC only becomes possible because reproducibility is ensured and leverages version control and dynamic document generation as key features of the workflow.
 

### 2021-11-18 manuscript.Rmd
mention that `repro` was developed as part of my thesis
mention that `repro` was developed as part of my thesis (#280)
* mention that `repro` was developed as part of my thesis

* corrected spelling mistake and removed "here presented"

Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>
First wave of changes

@@
   Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
-  Technological advancements now enable scientists to achieve a more comprehensive standard;
-  a standard where that allows anyone to access a digital research repository and reproduce all computational steps from raw data to final report including all relevant statistical analyses, with a single command.
+  Technological advancements now enable scientists to achieve a more comprehensive standard
+  that allows anyone to access a digital research repository and reproduce all computational steps from raw data to final report, including all relevant statistical analyses, with a single command.
   This method has far-reaching implications for scientific archiving,
@@
   This paper introduces `repro`, an R package that guides researchers through installation and use of the tools required to make a research project reproducible.
-  We also suggest the utilization of the proposed workflow for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
-  Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions of those analyses, which then merely complement the PAC as a more readable summary.
-  PAC circumvents the shortcomings of ambiguous preregistrations that may give researchers undesired degrees of freedom in the execution of their analysis.
-  Reproducibility facilitated by automation therefore has a wide range of applications to accelerate scientific progress.
+  We also suggest using the proposed workflow for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
+  Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions. 
+  PAC circumvents the shortcomings of ambiguous preregistrations that may result in undisclosed researcher degrees of freedom when adjustments to the planned analyses are necessary.
+  Reproducibility, facilitated by automation, has a wide range of applications and could potentially accelerate scientific progress.
 abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
@@
 acknowledgement: |
-  We would like to thank Maximilian Stefan Ernst for his contributions to the code for the simulation study. We are grateful to Julia Delius for her helpful assistance in language and style editing.
+  We would like to thank Maximilian Stefan Ernst for his contributions to the code for the simulation study.
+We are grateful to Julia Delius for her helpful assistance in language and style editing.
+The R package `repro` was developed as part of the [first author's master thesis](https://aaronpeikert.github.io/repro-thesis/) at the Humboldt-Universität zu Berlin.
 authorcontributions: |
@@
 Objectivity in this context means, "in principle it [the research finding] can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22].
-Only a reproducible result can be tested and understood by anyone.
+Only a reproducible result meets these requirements.
 Therefore, reproducibility has long been considered integral to empirical research.
 Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
-This is because making a research project reproducible with all code, data, and materials is accompanied by various challenges (e.g., missing software dependencies or ambiguous documentation of the exact computational steps taken), and there is a dire need for reproducibility standards.
+This is because there are various challenges to making a research project reproducible (e.g., missing software dependencies or ambiguous documentation of the exact computational steps taken), and there is a lack of best practices for overcoming these challenges [but see @vanlissa2020worcs].
 With technological advancement, however, it is now possible to make all digital products related to a research project available in a manner that enables automatic reproduction of the research project with minimal effort.
@@
 First, we want to introduce researchers to a notion of automated reproducibility that requires no manual steps, apart from the initial setup of the software environment. 
-Secondly, we discuss the possibilities that such automated reproducibility offers to improve not only the way research results are created and recovered, but also how this may change the general approach to a research project.
-With regard to the first goal, we show how to address four common threats to reproducibility with tools originating from software engineering [see @Peikert2019 for details].
-Then we present a tutorial on how to employ these tools to achieve automated reproducibility.
-A single tutorial cannot comprehensively introduce the reader to the individual tools, but this tutorial is intended to help readers to build a coherent workflow.
-It is aimed at researchers who regularly write code to analyze their data and are willing to make relevant code, data, and materials available, either publicly or on request.
-Ideally, our reader has already created a dynamic document at some point in time (e.g., with R Markdown or Jupyter) and attempted to track a file with version control (e.g., with Git). 
+Secondly, we discuss the implications of automated reproducibility for changing the general approach to research.
+With regard to the first goal, we discuss how to address four common threats to reproducibility, using tools originating from software engineering [see @Peikert2019 for details], and present a tutorial on how to employ these tools to achieve automated reproducibility.
+A single tutorial cannot comprehensively introduce the reader to the individual tools, but this tutorial is intended to help readers get started with a basic workflow.
+The tutorial is aimed at researchers who regularly write code to analyze their data and are willing to make relevant code, data, and materials available, either publicly or on request.
+Ideally, the reader has already created a dynamic document at some point in time (e.g., with R Markdown or Jupyter) and used some form of version control (e.g., Git). 
 The R package `repro` supports researchers in setting up the required software and in adopting this workflow.
-We present automated reproducibility as best practice, a goal that is not always achieved due to limited resources, technical restrictions, or practical considerations, but is very much worth striving for.
+We present automated reproducibility as a best practice; a goal that is not always fully achieved due to limited resources, technical restrictions, or practical considerations, but is worth striving for nonetheless.
 
@@
 PAC involves preregistering the intended analysis code and the major part of the final scientific report as a dynamic document, including typical sections like introduction, methods, and results.
-The resulting dynamic document closely resembles the final manuscript but temporarily uses simulated data to generate placeholder results (e.g., figures, tables, and statistics).
-With these placeholders, authors preregister the exact presentation of their results without having seen real data yet.
-Once the empirical data are available, they replace the simulated data, the results are then updated automatically, and the discussion can be written to finalize the report.
+The resulting dynamic document closely resembles the final manuscript but uses simulated data to generate placeholder results (e.g., figures, tables, and statistics).
+Simulated data allows authors to write their intended analysis script and preregister it<!-- CJ: Aaron: I don't think the key point is that users preregister the presentation of their results (tables, figures), but rather that they preregister the analysis!!!-->.
+Once the empirical data are available, these replace the simulated data; the results are then updated automatically, and the discussion can be written to finalize the report.
 
@@
 
-Therefore, this tutorial demonstrates how to make R-based research projects reproducible, while striking a balance between rigor and ease-of-use.
+This tutorial demonstrates how R users can make digital research products more reproducible, while striking a balance between rigor and ease-of-use.
 A rigorous standard increases the likelihood that a project will remain reproducible as long as possible.
 An easy-to-use standard, on the other hand, is more likely to be adopted.
-Our approach is to promote broad adoption of such practices, by implementing a "low threshold," and to ensure that they are compatible with more complex rigorous solutions, i.e., a "high ceiling."
-As researchers become more proficient in using the tools involved, they can thus improve the reproducibility of their work.
+Our approach is to promote broad adoption of such practices by ensuring a "low threshold", by making it easy to get started, while enabling a "high ceiling" by ensuring that they are compatible with more complex rigorous solutions.
+As researchers become more proficient in using the tools involved, they can thus further improve the reproducibility of their work.
 
@@
 
-*Version control* prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime (and in particular, in the development phase) of a research project.
+*Version control* prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the lifetime of a research project.
 Version control allows a clear link between which results were generated by which version of code and data.
-This is important for reproducibility because only if it is clear which version of data and code produced which results, those results can be said to be reproducible.
+This addresses the first threat to reproducibility, because results can only be said to be reproducible if it is clear which version of data and code produced them.
 We recommend using `Git` for version control, because of its widespread adoption in the R community.
@@
 At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
-A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
+Moreover, contributions (e.g., from collaborators) can be compared to, and incorporated in the master version of the project. 
+Version control thus reduces the risk of losing work and facilitates collaboration.
 `Git` is built around snapshots that represent the project state at a given point in time.
@@
 This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
-Each commit refers back to its ancestor, and they form a timeline together.
+Each commit refers back to its ancestor, and all commits are thus linked in a timeline.
 The entirety of commits (i.e., the version-controlled project) is called a repository.
 In `Git`, specific snapshots of a repository can be tagged, such that the user can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
-`Git` has additional features beyond basic version control, such as branches to facilitate collaboration.
+`Git` has additional features beyond basic version control, such as "branches" (parallel versions of a project that can later be merged again) to facilitate simultaneous collaboration.
 @vuorreCuratingResearchAssets2018 provide a more extensive treatment of how `Git` functions and how to use `Git` for research.
 @bryanExcuseMeYou2018 provides additional information on how to track `R Markdown` documents.
-Collaborating via `Git` requires the repository to be uploaded somewhere.
+Collaborating via `Git` is facilitated by uploading the repository to a cloud-based service.<!-- CJ This is not true-->
 We recommend GitHub as a host for `Git` repositories because of its popularity among R users.
-GitHub has many tools that ease project management and collaboration, and these tools are valuable in scientists' everyday work, but they are not central to achieving reproducibility.
+GitHub has many tools that facilitate working with `Git` - in particular project management and collaboration - but these are not central to achieving reproducibility.
 
@@
 The traditional way of writing a scientific report based on a statistical data analysis uses two separate steps conducted in two different programs.
-In a word processor, the researcher writes the text, and in another software, they conduct the analysis.
-Typically, results are manually copied and pasted from one software to another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
+The researcher writes text in a word processor, and conducts the analysis in another program.
+Results are then (manually) copied and pasted from one program to another, a process that often produces inconsistencies [@nuijtenPrevalenceStatisticalReporting2016].
 
@@
 This way, `Markdown` does not need any specialized software for editing.
-It is fool-proof (unlike, for example, `LaTeX` [@lamportLATEXDocumentPreparation1994]), works well with version control systems, and can be exported to various document formats, such as HTML websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
-We suggest using `Markdown` for all sorts of documents that are created in the academic context, starting from simple sketches of your ideas to your scientific manuscripts [@R-rticles] and presentations [@revealjs], or even your résumé [@vitae].
+It is fool-proof <!-- CJ: I don't understand what this means-->(unlike, for example, `LaTeX` [@lamportLATEXDocumentPreparation1994]), works well with version control systems, and can be exported to various document formats, such as HTML websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+`Markdown` can be used for all sorts of academic documents, ranging from simple sketches of ideas to scientific manuscripts [@R-rticles] and presentations [@revealjs], or even résumés [@vitae].
 `R Markdown` extends regular `Markdown` by allowing users to include R code chunks (in fact, arbitrary computer code [@xieMarkdownCookbook2020, Chapter 15, Chapter 15, Other Languages]) into a `Markdown` document.
@@
 
-One would hope that sharing a dynamic document would allow individuals (including one's future self) to download and reproduce a given analysis simply.
 While version control and dynamic document generation are becoming more common, we have argued that two more components are required and that each component alone is unlikely to guarantee reproducibility [@Peikert2019; @vanlissa2020worcs].
@@
 
-To automatically resolve dependencies between project files, we rely on *dependency tracking*.
+*Dependency tracking* helps automatically resolve dependencies between project files.
 In essence, researchers provide a collection of *computational recipes*.
@@
 Similar to a collection of cooking recipes, we can have multiple products (_targets_) with different ingredients (_requirements_) and different steps of preparation (_recipes_).
-In the context of scientific data analysis, the targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
-One noteworthy property of computational recipes is that they require a textual representation and the ability of a program to make sense of this textual representation.
+In the context of scientific data analysis, targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
+One noteworthy property of computational recipes is that they require a textual representation and the ability of a program to make sense of this textual representation.<!-- CJ I don't understand, and is it necessary to explain this?-->
 A workflow that involves renaming variable names by hand in a graphical spreadsheet application, for example, is therefore incompatible with automated reproducibility.
@@
 
-A version-controlled dynamic document with dependency tracking may still rely on external software.
+A version-controlled dynamic document with dependency tracking still relies on external software (even if just to be rendered to its final output format).
 Troubleshooting issues specific to a particular programming language or dependent tool typically require considerable expertise and threaten reproducibility.
-One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via *software management*.
+*Software management* refers to the act of providing records of, or access to, all software packages and system libraries a project depends on.
 One comprehensive approach to software management is containerization.
@@
 This is important because low level functionality may impact the workings of higher-order software like R, such as calls to random number generators or linear algebra libraries.
-All of the differences in computational results that could be caused by software are hence eliminated.
+All of the differences in computational results that could be caused by variation in the software used are hence eliminated.
 
@@
 Two major strategies help to keep the storage requirements reasonable.
-One is to rely on pre-made images, that are maintained by a community for particular purposes.
+One is to rely on pre-made images that are maintained by a community for particular purposes.
 For example, there are pre-made images that only include what is necessary for R, based on Ubuntu containers [@boettigerIntroductionRockerDocker2017].
@@
 
-A second strategy is to save a container recipe, a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment instead of the software environment itself.
+A second strategy is to save a so-called  `Dockerfile`, which contains only a textual description of all commands that need to be executed to recreate the software environment.
 `Dockerfiles` are tiny (the [`Dockerfile`](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024, 2)`KiB).
-However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
-For proper arching, we therefore recommend storing a complete image of the software environment, in addition to the `Dockerfile`.
+However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will continue to remain accessible and provide historic software versions.
+For proper archiving, we therefore recommend storing a complete image of the software environment, in addition to the `Dockerfile`.
 A more comprehensive overview of the use of containerization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
@@
 While `R Markdown` and `Git` are well integrated into the R environment through RStudio, `Make` and `Docker` require a level of expertise that is often beyond the training of scholars outside the field of information technology.
-Thus can represent a considerable obstacle to the acceptance and implementation of the workflow.
+This presents a considerable obstacle to the acceptance and implementation of the workflow.
 To overcome this obstacle, we have developed the R package `repro` that supports scholars in setting up, maintaining, and reproducing research projects in R.
@@
 These projects will remain reproducible irrespective of whether `repro` remains accessible in future.
-Users do not have to have `repro` installed to reproduce a project;
+Users do not need to have `repro` installed to reproduce a project;
 in fact, they do not even need to have R installed because the entire project can be rebuilt inside a container with R installed.

### 2021-11-18 manuscript.Rmd
Changes until line 940

--- 2021-11-17 manuscript.Rmd
@@
 One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be quite time-intensive.
-To lower the threshold, the R package `repro` introduces helper functions that wrap and hide the complexity of the necessary more complicated (and powerful) tools.
-The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer is doing and suggest what the user should do next.
-The `repro` package extends the [`usethis`](https://usethis.r-lib.org) [@usethis] package.
-While the `usethis` package aims to simplifys the development of R packages, `repro` contains reproducibility-specific utilities.
+To lower the threshold, the R package `repro` introduces helper functions that simplify the use of complicated and powerful tools.
+The `repro` package follows the format of the [`usethis`](https://usethis.r-lib.org) [@usethis] package, which provides helper functions to simplify the development of R packages.
+These helper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer is doing and suggest what the user should do next.
+The `repro` package provides similar helper functions, but focuses on reproducibility-specific utilities.
 We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
-These wrappers are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
-
-This tutorial assumes that the user is predominantly working in R, with the help of RStudio.
-It describes basic steps that we expect to be representative of a small-scale psychological research project that relies on no external software or multistage data processing (for those requirements see section [Advanced Features]).
+These wrappers are merely a support system; as users learn to use the underlying tools, they can rely less on `repro` and use these ttools directly to solve more complex problems.
+
+This tutorial assumes that the user will be working predominantly in R, with the help of RStudio.
+It describes basic steps that we expect to be relevant for small-scale psychological research projects that do not rely on external software or multistage data processing (for those requirements see section [Advanced Features]).
 Of course, your specific situation might involve additional, more specialized steps. 
@@
 
-The the check functions represent an example of the assistance that `repro` provides is .
-They verify that you have indeed installed and set up the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to clarify which functions are native `repro` functions):
+To verify that you have indeed installed and set up the required software for this workflow,
+you can use the "check functions".
+These also illlustrate how `repro` assists the user in setting up a reproducible workflow.
+In the example below, we use the double-colon operator to explicitly indicate which functions originate in the `repro` package.
+If the package is loaded (using `library(repro)`), it is not necessary to use this double-colon notation.
 
@@
 These functions check whether specific dependencies are available on the user's system, and if not, explain what further action is needed to obtain it.
-Sometimes they ask the user to do something, for example, the following happens if you are a Windows user who does not have `Git` installed:
+Sometimes they ask the user to take action; for example, the following happens if you are a Windows user who does not have `Git` installed:
 
@@
 
-The messages from `repro` try to take the user by the hand to help them solve problems.
+The messages from `repro` try to help the user solve problems.
 They are adjusted to your specific operating system and installed dependencies.
-Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions and to prepare your system for the creation of reproducible projects.
+Before you continue, we ask you to run the above commands to check `Git`, `Make`, and `Docker`---both to become familiar with the functionality of the `check_*()` functions and to make sure your system is prepared for the remainder of this tutorial.
 
@@
 ```{r}
-# only an alternative: DO NOT USE if you are unsure what SSH means
+# only an alternative: DO NOT USE if you are unsure what SSH means<!--CJ: Does the default method above use HTTPS? This is deprecated by GitHub-->
 repro::check_github(auth_method = "ssh")
@@
 
+## Creating an RStudio Project
+
+<!-- CJ Explain what the repro project template does -->
+We start by creating such a project folder with RStudio by clicking the menu item:
+
+> File → New Project... → New Directory → Example Repro Template
+
+This creates a project containing a sample analysis, but you may turn any other template or existing R project into a reproducible research project using the functionality of `repro`.
+
 ## Implementing Version Control
 
-Now that your system is set up for this tutorial, we will introduce you to version control with `Git`.
-`Git` tracks files within one project folder.
-We start by creating such a project folder with RStudio by clicking the menu item:
-
-> File → New Project... → New Directory → Example Repro Template
-
-This creates a project containing a sample analysis, but you may turn any other template or existing R project into a reproducible research project using the functionality of `repro`.
-
-`Git` tracks any files that you manually add to the repository.
+Now that your project is set up, we will introduce you to version control with `Git`.
+`Git` does not automatically track all files in your project folder;
+rather, you must manually add files to the `Git` repository.
 To make sure you do not accidentally add files that you do not wish to share (e.g., privacy-sensitive data),
 you can list specific files that you do not want to track in the `.gitignore` file.
+You can also block specific filetypes; for example, to prevent accidentally sharing raw data.
 You can add something to the `.gitignore` file directly or with this command:
@@
 
-Now the file `private.md` will not be committed to `Git` and will not be made public.
+Now the file `private.md` will not be added to the `Git` repository, and hence also not be made public if you push the repository to a remote service like GitHub.
 Please also consider carefully whether you can include data in the repository without violating privacy rights.
@@
 
-New users, in particular, are advised it is recommended to follow this procedure and exclude sensitive files before proceeding.
-When you are ready, you can begin tracking your files using `Git` by running:
+New users are advised to explicitly exclude any sensitive files before proceeding.
+When you are ready, you can begin tracking your remaining files using `Git` by running:
 
@@
 In this case, first, follow the suggestions shown on the R console.
+Note that the `worcs` R-package contains a convenience function `worcs::git_update("commit message")` which adds, commits, and pushes files to a remote repository,
+thus functioning as a kind of 'quick save' <!--CJ reference worcs package-->.
 
@@
 specific journal requirements.
-Of particular note is the package output format supplied by the [`papaja`](https://github.com/crsh/papaja) package [@papaja], which produces manuscripts that follow the American Psychological Association formatting requirements [@apa7].
-`R Markdown` files are in plain text, with the advantage that they are more suitable for version control using `Git`.
+For social scientists, the [`papaja`](https://github.com/crsh/papaja) package [@papaja] may be relevant, as it produces manuscripts that follow the American Psychological Association formatting requirements [@apa7].
+`R Markdown` files are plain text, which is more suitable for version control using `Git` than binary files generated by some word processors.
 Some users might find it easier to activate the "Visual Editor" of RStudio ([Ctrl] + [Shift] + [F4] or click on the icon that resembles drawing materials or a compass in the upper right corner of the `R Markdown` document), which features more graphical elements like a traditional word processor but still creates an `R Markdown` underneath with all of its flexibility.
@@
 The following example is a full markdown document where the metadata (the "YAML front matter") are in lines 1--6.
+Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
+Others are specific to certain output formats or R packages.
 
@@
 
-Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
-However, `R Markdown` has been extended to hundreds of document formats that are often highly configurable by their metadata.
-Some R packages add additional metadata fields.
-The `repro` package adds fields to the metadata to help make the research project reproducible.
-These fields are used to list all dependencies of the research project.
+The `repro` package adds fields to the metadata to list all dependencies of the research project.
 This includes R scripts, data files, and external packages.
@@
 
-When they are set up, they can ask `repro` to explain how they should use `Make` and `Docker` to reproduce the project (or you could explain it to them):
+When these are set up, they can ask `repro` to explain how they should use `Make` and `Docker` to reproduce the project (or you could explain it to them):
 
@@
 The only "hard" software requirement for reproducing a project is `Docker`, assuming users know how to build a `Docker` image and run `Make` within the container.
-However, if they have installed `Make` in addition to `Docker`, they do not even need to know how to use `Docker` and can simply rely on the two `Make` commands "`make docker`" and "`make -B DOCKER=TRUE`".
+However, if they have installed `Make` in addition to `Docker`, they do not even need to know how to use `Docker` and can simply rely on the two `Make` commands "`make docker`" and "`make -B DOCKER=TRUE`".<!-- CJ Is this information in the default readme?-->
 
@@
 If you read this paper the first time, you will probably want to skip this section and continue reading from the section "Preregistration as Code."
-As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.
-The simplification of `Docker` and `Make` are possible because of two restrictions.
+As explained above, `repro` is merely a simplified interface to the tools that enable reproducibility.
+This simplified interface imposes two restrictions.
 Users who ask themselves either, "How can I install software dependencies outside of R in the `Docker` image?" or "How can I express complex dependencies between files (e.g., hundreds of data files are preprocessed and combined)?" need to be aware of these restrictions and require a deeper understanding of the inner workings of `repro`.
@@
 
+The first restriction is that users must rely on software that is either already provided by the base `Dockerimage` "rocker/verse" or the R packages they list in the metadata.
 The metadata the `repro::automate()` function relies on can only express R packages as dependencies for the `Dockerfile` and only trivial dependencies (in the form of "file must exist") for the `Makefile`.
-The first limitation means that users must rely on software that is either already provided by the base `Dockerimage` "rocker/verse" or the R packages they list in the metadata.
 Other software that users might need, like other programming languages, not yet installed LaTeX packages, etc., must be added manually.
@@
 Second, the flag `-B` means that `Make` will consider all dependencies as outdated and will hence rebuild everything.
-Third, repro constructs `Make` targets so that if you supply `DOCKER=TRUE` they are executed within the `Docker` image of the project.
-
-This interplay between `Docker` and `Make` is relatively complicated because it resembles a chicken or egg problem.
+Third, repro constructs `Make` targets so that if you supply `DOCKER=TRUE` they are executed within the `Docker` image of the project.<!-- CJ This paragraph is VERY hard to follow without knowing how make and docker work (you've explained docker a bit, but make not so much).
+I think you can remove this, it's too technical and not clear why you're explaining this-->
+
+The interplay between `Docker` and `Make` resembles a chicken or egg problem.
 We have computational steps (`Make`) that depend on the software environment (`Docker`) for which we again have computational steps that create it.
@@
 
-To summarize, `repro` provides researchers with an easy-to-use solution for dependency tracking (in the form of `Make`) and software management (using `Docker`) without the necessity to learn both tools.
-Users with more advanced requirements can manually change all aspects of both programs while still profiting from `repro`'s automation.
+To summarize, `repro` automates dependency tracking (in the form of `Make`) and software management (using `Docker`) without the necessity to learn both tools,
+but users with advanced requirements can still customize all aspects of both programs.
 
@@
 Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
-Preregistration seeks to reduce researchers' degrees of freedom [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, that could bias the reported results.
-Preregistration, however, have several shortcomings.
-One is that written study plans are often interpretable in multiple ways and often do not achieve the level of unambiguity that is required to effectively set the researcher degrees of freedom, to zero.
+Preregistration seeks to reduce researchers' degrees of freedom [@NosekRevolution2018]<!--CJ this is not exactly true; the explicit goal of preregistration is to distinguish exploratory research from confirmatory research.
+Reducing researcher degrees of freedom is a side effect, not the goal--> in the execution of a planned study by limiting the number of decisions to be made after data has been collected, that could bias the reported results.
+Preregistrations have several shortcomings, however.
+One is that written study plans can be interpreted in multiple ways and often do not achieve the level of unambiguity that is required to effectively set the researcher degrees of freedom, to zero.<!--CJ I don't think you've defined researcher degrees of freedom?
+And why is it important to reduce these to zero?
+Because that blurs the line between exploratory and confirmatory!-->
 Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, their results vary considerably [@silberzahnManyAnalystsOne2018].
@@
 These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
-However, such templates cannot remove the degrees of freedom completely because it is impossible to verbally describe every detail of an analysis for any but the most straightforward analysis.
-This ambiguity causes a second problem, namely, comparing the initial plan and the resulting publication to decide if and how researchers deviated from the preregistration.
-This task is difficult because it is impossible to decide without additional information weather the analysis was actually carried out differently or just described differently.
-The comparison is made impossible if the authors of the preregistration were intentionally vague.
+However, the effectiveness of such templates is limited by the ambiguity of using language to describe analytical procedures.
+The use of language to describe analytical procedures also complicates comparing the initial plan with the resulting analyses to decide if and how researchers deviated from the preregistration.
 Even if the researchers were faithful to the preregistration, readers may reach opposite conclusions because they have to compare two different text that may be worded differently or describe the same thing in varying levels of detail.

### 2021-11-19 manuscript.Rmd

--- 2021-11-18 manuscript.Rmd
@@
   Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
-  Technological advancements now enable scientists to achieve a more comprehensive standard
-  that allows anyone to access a digital research repository and reproduce all computational steps from raw data to final report, including all relevant statistical analyses, with a single command.
+  Technological advancements now enable scientists to achieve a more comprehensive standard that allows anyone to access a digital research repository and reproduce all computational steps from raw data to final report, including all relevant statistical analyses, with a single command.
   This method has far-reaching implications for scientific archiving,
@@
   Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions. 
-  PAC circumvents the shortcomings of ambiguous preregistrations that may result in undisclosed researcher degrees of freedom when adjustments to the planned analyses are necessary.
+  PAC circumvents the shortcomings of ambiguous preregistrations that may result in undisclosed use of researcher degrees of freedom.
   Reproducibility, facilitated by automation, has a wide range of applications and could potentially accelerate scientific progress.
@@
 With regard to the first goal, we discuss how to address four common threats to reproducibility, using tools originating from software engineering [see @Peikert2019 for details], and present a tutorial on how to employ these tools to achieve automated reproducibility.
-A single tutorial cannot comprehensively introduce the reader to the individual tools, but this tutorial is intended to help readers get started with a basic workflow.
+A single tutorial cannot comprehensively introduce the reader to the detail of individual tools, but this tutorial is intended to help readers get started with a basic workflow.
 The tutorial is aimed at researchers who regularly write code to analyze their data and are willing to make relevant code, data, and materials available, either publicly or on request.
@@
 The resulting dynamic document closely resembles the final manuscript but uses simulated data to generate placeholder results (e.g., figures, tables, and statistics).
-Simulated data allows authors to write their intended analysis script and preregister it<!-- CJ: Aaron: I don't think the key point is that users preregister the presentation of their results (tables, figures), but rather that they preregister the analysis!!!-->.
+Simulated data serve two functions, they allow to test the code for the planned analyses and for preregistering the exact presentation of the results.
 Once the empirical data are available, these replace the simulated data; the results are then updated automatically, and the discussion can be written to finalize the report.
@@
 At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
-Moreover, contributions (e.g., from collaborators) can be compared to, and incorporated in the master version of the project. 
+Moreover, contributions (e.g., from collaborators) can be compared to, and incorporated in the main version of the project. 
 Version control thus reduces the risk of losing work and facilitates collaboration.
@@
 @bryanExcuseMeYou2018 provides additional information on how to track `R Markdown` documents.
-Collaborating via `Git` is facilitated by uploading the repository to a cloud-based service.<!-- CJ This is not true-->
+Collaborating via `Git` is facilitated by uploading the repository to a cloud-based service.
 We recommend GitHub as a host for `Git` repositories because of its popularity among R users.
-GitHub has many tools that facilitate working with `Git` - in particular project management and collaboration - but these are not central to achieving reproducibility.
+GitHub has many tools that facilitate working with `Git` --- in particular project management and collaboration --- but these are not central to achieving reproducibility.
 
@@
 This way, `Markdown` does not need any specialized software for editing.
-It is fool-proof <!-- CJ: I don't understand what this means-->(unlike, for example, `LaTeX` [@lamportLATEXDocumentPreparation1994]), works well with version control systems, and can be exported to various document formats, such as HTML websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
+It is userfriendly (unlike, for example, `LaTeX` [@lamportLATEXDocumentPreparation1994]), works well with version control systems, and can be exported to various document formats, such as HTML websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
 `Markdown` can be used for all sorts of academic documents, ranging from simple sketches of ideas to scientific manuscripts [@R-rticles] and presentations [@revealjs], or even résumés [@vitae].
@@
 In the context of scientific data analysis, targets are typically the final scientific report (e.g., the one to be submitted to a journal) and possibly intermediate results (such as preprocessed data files, simulation results, and analysis results).
-One noteworthy property of computational recipes is that they require a textual representation and the ability of a program to make sense of this textual representation.<!-- CJ I don't understand, and is it necessary to explain this?-->
 A workflow that involves renaming variable names by hand in a graphical spreadsheet application, for example, is therefore incompatible with automated reproducibility.
@@
 
-A version-controlled dynamic document with dependency tracking still relies on external software (even if just to be rendered to its final output format).
+A version-controlled dynamic document with dependency tracking still relies on external software.
 Troubleshooting issues specific to a particular programming language or dependent tool typically require considerable expertise and threaten reproducibility.

### 2021-11-19 manuscript.Rmd

@@
 One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be quite time-intensive.
-To lower the threshold, the R package `repro` introduces helper functions that wrap and hide the complexity of the necessary more complicated (and powerful) tools.
-The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer is doing and suggest what the user should do next.
-The `repro` package extends the [`usethis`](https://usethis.r-lib.org) [@usethis] package.
-While the `usethis` package aims to simplifys the development of R packages, `repro` contains reproducibility-specific utilities.
+To lower the threshold, the R package `repro` introduces helper functions that simplify the use of complicated and powerful tools.
+The `repro` package follows the format of the [`usethis`](https://usethis.r-lib.org) [@usethis] package, which provides helper functions to simplify the development of R packages.
+The `repro` package provides similar helper functions, but focuses on reproducibility-specific utilities.
+These helper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer is doing and suggest what the user should do next.
 We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
-These wrappers are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.
-
-This tutorial assumes that the user is predominantly working in R, with the help of RStudio.
-It describes basic steps that we expect to be representative of a small-scale psychological research project that relies on no external software or multistage data processing (for those requirements see section [Advanced Features]).
+These wrappers are merely a support system; as users learn to use the underlying tools, they can rely less on `repro` and use these ttools directly to solve more complex problems.
+
+This tutorial assumes that the user will be working predominantly in R, with the help of RStudio.
+It describes basic steps that we expect to be relevant for small-scale psychological research projects that do not rely on external software or multistage data processing (for those requirements see section [Advanced Features]).
 Of course, your specific situation might involve additional, more specialized steps. 
@@
 
-The the check functions represent an example of the assistance that `repro` provides is .
-They verify that you have indeed installed and set up the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to clarify which functions are native `repro` functions):
+To verify that you have indeed installed and set up the required software for this workflow,
+you can use the "check functions".
+These also illlustrate how `repro` assists the user in setting up a reproducible workflow.
+In the example below, we use the double-colon operator to explicitly indicate which functions originate in the `repro` package.
+If the package is loaded (using `library("repro")`), it is not necessary to use this double-colon notation.
 
@@
 
+## Creating an RStudio Project
+
+We start by creating a project folder with RStudio by clicking the menu item:
+
+> File → New Project... → New Directory → Example Repro Template
+
+This creates a project with a sample analysis.
+This sample analysis consists of a single R Markdown document and a single data file.
+The only special thing about the R Markdown document is the `repro` metadata that we will learn later about. 
+However, you may turn any other template or existing R project into a reproducible research project by adding those `repro` metadata there.
+
 ## Implementing Version Control
 
-Now that your system is set up for this tutorial, we will introduce you to version control with `Git`.
-`Git` tracks files within one project folder.
-We start by creating such a project folder with RStudio by clicking the menu item:
-
-> File → New Project... → New Directory → Example Repro Template
-
-This creates a project containing a sample analysis, but you may turn any other template or existing R project into a reproducible research project using the functionality of `repro`.
-
-`Git` tracks any files that you manually add to the repository.
+Now that your project is set up, we will introduce you to version control with `Git`.
+`Git` does not automatically track all files in your project folder;
+rather, you must manually add files to the `Git` repository.
 To make sure you do not accidentally add files that you do not wish to share (e.g., privacy-sensitive data),
 you can list specific files that you do not want to track in the `.gitignore` file.
+You can also block specific filetypes; for example, to prevent accidentally sharing raw data.
 You can add something to the `.gitignore` file directly or with this command:
@@
 
-This interplay between `Docker` and `Make` is relatively complicated because it resembles a chicken or egg problem.
+The interplay between `Docker` and `Make` resembles a chicken or egg problem.
 We have computational steps (`Make`) that depend on the software environment (`Docker`) for which we again have computational steps that create it.

### 2021-11-19 manuscript.Rmd
Comments until 1113

@@
 
-Defining the research questions and planning data analysis before observing the research outcomes is referred to as *preregistration* [@NosekRevolution2018].
-Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
-Preregistration seeks to reduce researchers' degrees of freedom [@NosekRevolution2018] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, that could bias the reported results.
-Preregistration, however, have several shortcomings.
-One is that written study plans are often interpretable in multiple ways and often do not achieve the level of unambiguity that is required to effectively set the researcher degrees of freedom, to zero.
-Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, their results vary considerably [@silberzahnManyAnalystsOne2018].
-The current best practice to ensure that the preregistrations are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
-These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
-However, such templates cannot remove the degrees of freedom completely because it is impossible to verbally describe every detail of an analysis for any but the most straightforward analysis.
+*Preregistration* refers to the practice of defining research questions and planning data analysis before observing the research outcomes [@NosekRevolution2018].
+It serves to separate a-priori planned and theory-driven (confirmatory) analyses from unplanned and post-hoc (exploratory) analyses.<!-- CJ I removed the bit about credibility, because that requiresfurther explanation of why separating these two types results in more credible findings-->
+Exploratory analyses refer to both those aimed at discovering patterns in data, and those aimed at overcoming shortcomings of confirmatory analyses that come to light after seeing the data (e.g., violations of assumptions, modifications to address poor model fit).
+Researchers are faced with a myriad of choices in designing, executing, and analyzing a study, often called researchers degrees of freedom.
+Undisclosed researcher degrees may be used to modify planned analyses until a key finding reaches statistical significance or to inflate effect size estimates, a phenomenon referred to as *opportunistic bias* [@decosterOpportunisticBiasesTheir2015].
+Preregistration increases transparency by clarifying when and how researchers employ their degrees of freedom.
+It expressly does not restrict what researchers may do to gather or analyze their data.
+
+There are still several shortcomings to preregistration.
+One is that written study plans are often multiinterpretable.
+Empirical research has shown that, even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, their results vary considerably [@silberzahnManyAnalystsOne2018].
+The current best practice to ensure comprehensive and specific preregistration is to impose structure by following preregistration templates [@bowmanOSFPreregTemplate2020, @bakkerEnsuringQualitySpecificity2020].
+However, such templates cannot ensure full transparency because it is impossible to verbally describe every detail of an analysis for any but the most straightforward analysis.
 This ambiguity causes a second problem, namely, comparing the initial plan and the resulting publication to decide if and how researchers deviated from the preregistration.
 This task is difficult because it is impossible to decide without additional information weather the analysis was actually carried out differently or just described differently.
-The comparison is made impossible if the authors of the preregistration were intentionally vague.
-Even if the researchers were faithful to the preregistration, readers may reach opposite conclusions because they have to compare two different text that may be worded differently or describe the same thing in varying levels of detail.
+Even if researchers were faithful to the preregistration, readers may reach opposite conclusions because they have to compare two different text that may be worded differently or describe the same thing in varying levels of detail.
 A third limitation is that preregistrations are susceptible to non-reproducibility, just like primary research.
@@
 We term this approach of writing and publishing code at the preregistration stage *Preregistration as Code (PAC)*.
-PAC has the potential to eliminate researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
-Moreover, it removes the necessity to write two separate documents:
-A preregistration *and* a manuscript.
-For a PAC, researchers can write a reproducible, dynamically generated draft of their intended manuscript at the preregistration stage.
+PAC has the potential to eliminate undisclosed researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
+Moreover, it reduces overhead by obviating the need to write a separate preregistration and manuscript.
+For PAC, researchers can write a reproducible, dynamically generated draft of their intended manuscript at the preregistration stage.
 This already includes most of the typical sections, such as introduction, methods, and results.
@@
 Reproducibility is of utmost importance at this stage since the preregistration must produce valid results at two points in time, once before data collection and once after data collection.
-As outlined before, reproducibility builds upon four pillars (version control, dynamic document generation, dependency tracking, and software management).
-To use PAC the dangers to reproducibility we described must be  eliminated.
-
-The idea of submitting code as part of a preregistration is not new.
+As outlined before, reproducibility builds upon four pillars (version control, dynamic document generation, dependency tracking, and software management).<!-- CJ I don't understand why you repeat this here.
+Remove?-->
+To use PAC the dangers to reproducibility we described must be  eliminated.<!-- CJ I don't understand why you repeat this here. Remove?-->
+
+The idea of submitting code as part of a preregistration is not new [e.g.
+@wichertsDegreesFreedomPlanning2016].<!--CJ: Please keep this comment in the text.
+If possible, I would like to reference the upcoming many analysts paper here, for which I used PAC in the final revision.-->
 A prominent preregistration platform, [The Open Science Framework](https://osf.io/), suggests submitting scripts alongside the preregistration of methods.
@@
 
-We believe that pairing PAC with the workflow presented above offers five advantages over classical preregistration.
-First, PAC is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
-Relatedly, writing the preregistration in the form of a research article has the advantage that researchers usually have great expertise in writing a scientific manuscript anyway.
-By contrast, filling out a preregistration template is a novel type of scientific writing for many.
+We believe that pairing PAC with the workflow presented above offers several advantages over classical preregistration.
+First, PAC is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating, two separate documents.
+Relatedly, writing the preregistration in the form of a research article has the advantage that researchers are usually familiar with this format.
+By contrast, a preregistration template is a novelty for many.
 Second, PAC is a tool for study planning.
-If all steps to be executed are documented clearly, the carrying out of the study will proceed more efficiently than when every step is planned in medias res.
-Third, PAC removes any ambiguity regarding the translation of verbal analysis plans into code.
+A study can be carried out more efficiently if all steps are documented clearly than when every step is planned ad hoc.
+Third, PAC removes ambiguity regarding the translation of verbal analysis plans into code.
 PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
 Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
-PAC, therefore, minimizes researchers degrees of freedom more effectively than standard preregistration does [@bakkerEnsuringQualitySpecificity2020].
-Fourth, despite its rigor, PAC accommodates data-dependent decisions if these can be formulated as code.
+PAC therefore minimizes undisclosed researchers degrees of freedom more effectively than standard preregistration does [@bakkerEnsuringQualitySpecificity2020; @wichertsDegreesFreedomPlanning2016].
+Fourth, despite its rigor, PAC accommodates partly exploratory data-dependent decisions.
+Some decisions can be anticipated and formulated as conditional code.
 Researchers can, for example, formulate conditions (e.g., in the form of if-else-blocks) under which they prefer one analysis type over the other.
 For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
+Some necessary deviations from the preregistration only become apparent after seeing the data.<!--CJ I merged point 5 here; it's part of point 4-->
+Version control makes it possible to commit specific changes to the planned analyses.
+It is recommended to documented all such changes by maintaining a changelog, and referencing specific commits.<!--CJ I do this myself, I think it's a good idea.-->
 Another example of data-dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
 Decisions that do not lend themselves to formulation in code, e.g. visual inspection, must still be described verbally or be treated as noted in the next section.
-With PAC researchers can precisely calibrate how much they want to explore.
-When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
-Fifth, deviations from the preregistration are clearly documented because they are reflected in changes to the code, which are managed and tracked with version control.
+With PAC researchers can precisely calibrate how much they want to explore.<!--CJ This is vague-->
+When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.<!-- CJ Not clear what this refers to; you haven't given an example or anything!
+Cut?-->
 
@@
 
-We would like to note that in the PAC paradigm, the direct comparison of the preregistration and the final publication is particularly easy to accomplish.
-Because there is a single unambiguous reference point, that is, the dynamic document, including code, under version control, changes can be inspected by authors and readers alike.
-Authors should at least review all changes made to the preregistered plan to summarize and justify them for their reader, e.g. in the discussion of the final manuscript [^summarize-changes].
-Such retrospective summary already serves to put changes into context.
-However, a changelog that is kept continually and where changes are explained when they are made provides more immediate and detailed information.
-Each entry in the changelog should explain the reasoning of the changes and link to the changes made by including the commit id that applied the changes.
-The interested reader is also in a position to inspects the changes one by one and can therefore make an informed judgment about their implications for the inferences drawn.
-
-Deviations from the preregistration are sometimes maligned as if encountering unexpected aspects would invalidate an entire carefully crafted study [@szollosi_is_2020].
-However, we share the common view that deviation from a preregistration is not a problem [@nosekPreregistrationHardWorthwhile2019]; it is a problem if these changes are not made transparent.
-In fact, we expect that most PACs will require changes after data have been collected.
-Often deviations provide an opportunity to learn from the unexpected.
-Transparency is essential in this regard because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumptions [@meehlTheoreticalRisksTabular1978].
-
-Take the following example: A preregistration intended to include "working memory" and "fluid intelligence" as predictors for reaction time in an experiment.
-However, after fitting the model on the actual data, both constructs exhibit high collinearity.
-The authors decide that for their purpose, these variables are interchangeable and remove one.
-For them, this change simply does not affect the core of their theory but was an auxiliary assumption, and they might be right.
-Another researcher, however, who hopes to incorporate exactly this result into a theory about intelligence, might disagree about the importance of this assumption and would accordingly view the finding as more explorative because the change to the preregistration affects their theory directly.
-With transparent reporting, it is left to the readers to make up their own judgement.
+We would like to note that PAC allows explicit comparison of the preregistration and the final publication.
+Authors should retrospectively summarize and justify any changes made to the preregistered plan, e.g. in the discussion of the final manuscript [^summarize-changes].
+During the analysis process, authors can additionally maintain a running changelog to explain changes in detail as they arise.
+Each entry in the changelog should explain the reasoning behind the changes and link to the commit id that applied the changes.
+This enables readers and reviewers to inspects individual changes and make an informed judgment about their validity and implications.
+
+Deviations from the preregistration are sometimes maligned, as if encountering unexpected challenges invalidates a carefully crafted study [@szollosi_is_2020].
+However, we share the common view that deviation from a preregistration is not a problem [@nosekPreregistrationHardWorthwhile2019];
+rather, a faillure to disclose such deviations is a problem.
+In fact, it is expected that most PACs will require some modification after empirical data becomes available.
+Often, deviations provide an opportunity to learn from the unexpected.
+Transparency is essential in this regard <!--CJ: it is not clear how the following argument justifies the need for transparency.
+Please fix--> because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumptions [@meehlTheoreticalRisksTabular1978].
+
+For example, imagine that authors preregistered their intention to include both "working memory" and "fluid intelligence" as covariates in an experimental study, examining the effect of task novelty on reaction time.
+When evaluating the planned analyses on the empirical data, these two covariates reveal high collinearity, thus biasing estimates and compromising statistical inference.
+The authors decide to use PCA to extract common variance related to "intelligence", and include this component as a covariate instead.
+This change pertains to an auxiliary assumption (that working memory and fluid intelligence are distinct constructs), but does not undermine the core theory (that task novelty affects reaction time).
+Now imagine that a different researcher is interested in the structure of intelligence.
+This change to the preregistration directly relates their theory of intelligence.
+That researcher might thus interpret the same result as an explorative finding, suggesting that these aspects of intelligence are unidimensional.
+<!--CJ not sure what this final line intended to argue. I removed it. Feel free to add a more satisfactory conclusion to this paragraph-->
 
@@
 
-Another common misunderstanding is that preregistration, and therefore PAC, preclude exploratory analyses.
-We must differentiate here between two kinds of exploration, and neither of them is limited by PAC.
-The first and more traditional meaning of exploration involves to ad hoc statistical procedures and post hoc explanations of the results.
-Such traditional exploratory findings should merely be explicitly declared in the manuscript to allow their appropriate evaluation [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
-The second kind of exploration is through procedurally well defined exploration with exploratory statistical models are standard in machine learning [@brandmaier22mlsem].
-These highly-parameterized models, which are difficult to preregister traditionally, but can be preregistered with PAC like any other model.
+Another common misunderstanding is that preregistration, including PAC, precludes exploratory analyses.
+We differentiate between two kinds of exploration, neither of which is limited by PAC.
+The first, more traditional kind of exploration involves ad hoc statistical decisions and post hoc explanations of the results.
+Such traditional exploratory findings should be explicitly declared in the manuscript to distinguish them from confirmatory findings [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
+The second kind involves rigorous exploration using methods that are standard in machine learning [@brandmaier22mlsem].
+These models often involve dozens, if not hundreds of predictors, which makes it difficult to describe them verbally.
+With PAC, such models can be preregistered clearly and in comprehensive detail.
 We specifically recommend PAC for such exploratory statistical models.
-The merit of preregistration in these cases is to communicate precisely how much exploration was done, a piece of information that is crucial for assessing the uncertainty of such now "preregisterable" but still exploratory results.
-
-## What You Need to Know Before Writing a PAC
-
-When planning a study and writing a PAC, researchers ought to strive to meet three goals:
+The merit of preregistration in these cases is to communicate precisely how much exploration was done;
+a piece of information that is crucial to assess e.g., whether the results might be overfit <!--CJ reference Hastie & Tibshirani-->.
+
+## How to Write a Good PAC
+
+When planning a study and writing a PAC, researchers ought to strive to meet three goals:<!--CJ: None of these goals relate explicitly to code.
+Is that a problem?
+Also: Shouldn't you mention the goals when you first introduce PAC?-->
 First, the plan should be comprehensive.
@@
 Standards for scientific writing are designed to result in understandable and comprehensive research articles and can serve the same purpose for writing a preregistration.
-Such standards [e.g., @apa7] require that research articles describe a study in such detail that any researcher that is familiar with the field is able to repeat it.
-These well-known standards can thus provide an excellent template for planning a preregistered study.
-One particular advantage is introduction's and theoretical background's embedding in existing literature where that provide the basis for the study's design.
-Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.
-The comprehensiveness and completeness of published scientific reports is sometimes limited by the strict word count limits that some journals impose.
-One may argue that PAC inherits this limitation if the same word count limit applies.
-However, in such cases, we suggest using reproducible appendix materials to keep the core manuscript concise, yet the preregistration comprehensive and exhaustive.
-
-Another factor that ensures a comprehensive plan is writing their analysis based on simulated data.
-Simulating data offers various benefits in the planning phase of a study.
-It forces the researcher to think about and address data-analytic problems on a conceptual level before they make an effort to acquire data [@axelrodAdvancingArtSimulation1997].
-It also affords researchers with a simple technical check whether the analysis code will run without failure [@braiekTestingMachineLearning2020] (e.g., does the code also work with missing values, extreme outliers, etc.).
-Writing incomplete code is technically impossible, or at least the researchers would notice at this early stage if they overlooked something (e.g. if they forgot to aggregate the data).
-
-Peer review of preregistration is a method that also can ensure comprehensivenes.
-For example, registered reports [an article format in which the preregistration is reviewed and the article accepted for publication "in principle," @nosekRegisteredReports2014; @chambersWhatNextRegistered2019].
+For example, the APA standards [@apa7] require research articles describe a study in such detail that any researcher that is familiar with the field is able to repeat it.
+These well-known standards thus provide a good starting point for the structure of a PAC.<!--CJ What are you arguing here?
+You've just convinced the reader that existing templates are not comprehensive enough.
+Now you're arguing that APA7 can be considered a preregistration template, and that it is comprehensive?
+I am a little confused here-->
+One particular advantage is introduction's and theoretical background's embedding in existing literature where that provide the basis for the study's design.<!--CJ This sentence has been missed by the language editor-->
+Further inspiration may be taken from preregistration templates relevant to the type of study at hand,
+because such templates address methodological rigor more comprehensively.
+One factor that limits the comprehensiveness of published scientific reports is the word count imposed by journals.
+If this is a concern, we suggest relegating the PAC to a reproducible appendix to keep the core manuscript concise.
+
+Simulating data offers further benefits in ensuring the comprehensiveness of a preregistration.
+It forces the researcher to consider conceptual data-analytic challenges before obtaining empirical data [@axelrodAdvancingArtSimulation1997].
+It also allows researchers to verify that planned analyses are feasible and likely to run without failure [@braiekTestingMachineLearning2020] (e.g., does the code also work with missing values, extreme outliers, etc.).
+Many errors and omissions are likely to come to light when evaluating the intended analysis code on simulated data.
+
+One final method to ensure comprehensivenes of the PAC is peer review.
+The value of PAC would be further enhanced by submitting it as a registered report,
+a format in which articles are accepted for publication "in principle" based on peer review of the preregistered study plans [@nosekRegisteredReports2014; @chambersWhatNextRegistered2019].
 In such registered report peer-reviewing a manuscript-styled PAC may well make it easier for the reviewers that ascertain and evaluate a comprehensive study plan.
+Moreover, reviewers can suggest specific improvements to the planned analysis code.
 
@@
 
-If the comprehensiveness criterion is met, researchers can check weather the analysis is effective in recovering the true parameters of the simulation.
-If an analysis fails to recover the correct values in idealized (simulated) conditions,
-it would be unreasonable to expect this analysis to work under realistic conditions.
+If the comprehensiveness criterion is met, researchers can further check whether the planned analyses are effective.
+Effectiveness can be defined in terms of ability to recover true population parameters, or power to detect a population effect.<!--CJ You have not said anything about a simulation yet, so this is kind of surprising-->.
+Effectiveness can be assessed by means of a simulation study.
+Once code has been written to simulate synthetic data, it is often a minor step to perform a simulation study.
+It is sufficient to repeat the planned analyses many times, and plot the distribution of estimated parameters, or tabulate the proportion of replications in which a true effect was detected (power analysis).
+The R command `replicate()` can be used to this end.
+If an analysis fails to consistently recover the population parameters under simulated (idealized) conditions,
+it would be unreasonable to expect this analysis to be effective under realistic conditions.
 Because the simulation describes the researchers' expectations about real-world conditions,
-it is the best way to assess the effectiveness of the analysis without actually collecting the data.
-A simulation of data requires the researchers to translate their expectations about the underlying process into concrete computational terms, e.g.
-assumptions about the distribution  the dependent variable follows or how the independent variables influence each other and the dependent variable.
-If a researcher experiences difficulty in specifying such expectations, it may be an indication that further literature research or exploratory pilot studies are required.
-
-### Efficiency
+it is the best way to assess the effectiveness of the analysis prior to collecting empirical data.
+Simulating data requires researchers to explicate their assumptions about the data-generating mechanism in concrete computational terms; e.g.
+assumptions about the distributions of variables and how they relate to one another.
+If a researcher experiences difficulty in specifying such assumptions, this may indicate that further literature research or exploratory pilot studies are required.
+
+### Efficiency<!-- CJ I do not understand how 'efficiency' is different from effectiveness above.
+You cannot assess effectiveness by means of a single iteration.
+Moreover, efficiency has a specific statistical definition (pertaining to the amount of information required to estimate a parameter).
+Using this term can be highly confusing to our (probably statistically inclined) readership.-->
 
@@
 
-### The Writing Process
-
-To write a PAC, researchers should start by drafting a manuscript in the form of an `R Markdown`.
-When they have outlined the studies design, they should begin writing code.
-If there is no other code structure available, we suggest writing a minimum of three functions for each planned hypothesis, one for the planned analysis, one for simulation of the respective data, and one for reporting the results.
-We capture the tasks here in the form of explicit R functions, but researchers may use any form (e.g., scripts) as long as it serves to simulate, analyze, and report.
-The functions shown here are empty (they do not contain executable code), but the interested reader can find working functions in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.
-
-At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` from which they later derive the method section.
+### Planned Analyses as Functions
+
+Although researchers may use any form to preregister their planned analyses (e.g., scripts),
+we suggest writing three functions for each planned hypothesis: one to conduct the planned analysis, one to simulate the expected data, and to report the results.
+Using functions makes the analysis more portable (i.e., it can easily be used for other datasets), and facilitates repeated evaluation, as is the case in a simulation study. 
+The functions shown here do not contain executable code, but the interested reader can find working functions in the [online supplementary materials](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R) that power the example below.
+
+It is difficult to write analysis code when it is not clear what the expected data will look like.
+We therefore recommend first simulating a dataset that resembles the expected structure of the empirical data that will be used for the final analysis.
+Dedicated packages to simulate data for specific analyses exist <!--CJ Reference here-->.
+For linear models, simulating data is extremely simple:
+
+```{r}
+simulate_data <- function(n, effect_size){
+  # Draw n samples from a normal distribution for predictor X
+  x <- rnorm(n)
+  # Calculate dependent variable Y as a function of population effect size and residual error
+  y <- effect_size * x + rnorm(n)
+  # Return a data.frame
+  data.frame(x = x, y = y)
+}
+simulate_data(5, .2)
+```
+
+The general format of a simulation function might be as follows:
+
+```{r, eval=FALSE}
+simulate_data <- function(n, effect_size){
+  # 1. warn users that the results are "fake"
+  # 2. draw `n` samples with `effect_size`
+  # 3. format and return in expected data format
+}
+```
+
+Next, write a function to conduct the planned analysis.
 This function should receive the data and compute all relevant results from it.
+In the simplest case, an analysis function might already exist in R.
+For the linear model above, the analysis function might be:
+
+```{r}
+lm(y ~ x, data = simulate_data(5, .2))
+```
+
+The general format of an analysis function might be:
 
@@
 
-It is difficult to write code when it is not clear how exactly the input should look like.
-However, at the time of preregistration, researchers should remain insulated from real data.
-We, therefore, simulate a dataset first that resembles the expected structure of the empirical data that we will use for the final analysis.
-Using the simulated dataset as a placeholder, we can now write a scientific report including all analysis code blocks and the presentation of results that can serve as a PAC.
-
-```{r, eval=FALSE}
-simulate_data <- function(n, effect_size){
-  # 1. warn users that the results are "fake"
-  # 2. draw `n` samples with `effect_size`
-  # 3. format and return in expected data format
-}
-```
-
 As soon as we have written `planned_anaylsis()` and `simulate_data()` we can iteratively improve both functions, e.g.
 until `planned_analysis()` runs without error and recovers the correct parameters from `simulate_data()`.
-The goal is that the output of `simulate_data()` works as input to the function `planned_analysis()`.
+The goal is to ensure that the output of `simulate_data()` works as input to the function `planned_analysis()`.
+Then, using the simulated dataset as a placeholder, one can write a scientific report including all analysis code blocks and the presentation of results.
 
@@
 
-Every study that can be preregistered can also be preregistered with code.
+Every study that can be preregistered and ultimately uses computer code for the statistical analysis, can also be preregistered with code.
 The only difference is that PAC demands a comprehensive and complete analysis plan.
-With traditional preregistration, researchers are able to omit details and get away with decisions that are made only after having seen the data, thus, potentially biasing statistical results.
-Some researchers may not feel ready to decide on every detail of the analysis plan before data collection.
+With traditional preregistration, researchers may (unintentionally) omit details and thus introduce decisions that appear to be made after having seen the data.
+We emphasize again that deviations from the preregistration are not only possible but likely.
+The advantage of PAC is transparency.
+All deviations from the planned analysis as implemented in code are tracked, and the final analysis can be easily compared to the PAC.
+Importantly, we see PAC as a means to increase transparency not as a means to limit researchers in their statistical analyses.
+In practice, researchers may not feel ready to decide on every detail of the analysis plan before data collection.
 In such cases, we recommend making decisions based on the best available knowledge, and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in Section [Deviating from the preregistration and exploration] in the final manuscript.
@@
 Then, we illustrated how such a project could be reproduced---either by the original author and/or collaborators or by a third party.
+
 We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations, such as preregistration as code (PAC).
-In PAC the entire reproducible manuscript, including planned analyses adn results based on simulated data, is preregistered.
+In PAC the entire reproducible manuscript, including planned analyses and results based on simulated data, is preregistered.
+This way, every use of a researchers' degree of freedom is disclosed.
 Once real data is gathered, the reproducible manuscript is (re-)created with the real data.

### 2021-11-19 manuscript.Rmd
Final changes

@@
 The goal is to ensure that the output of `simulate_data()` works as input to the function `planned_analysis()`.
-Then, using the simulated dataset as a placeholder, one can write a scientific report including all analysis code blocks and the presentation of results.
-
-### Alternatives to simulated data
-
-If simulating data proves to be too challenging, researchers can delay their preregistration until they have gathered data and repurpose the empirical data.
-In doing so, the researchers must walk a fine but important line.
-Of course, they must not analyze the data before they write the preregistration.
-Nevertheless, they may obfuscate the real data (before actually looking at them) by shuffling the order of rows for each variable (independently of each other) in cross-sectional data or otherwise by creating synthetic data that allow them to test the function `planned_analysis()` without actually revealing the observed research outcomes.
-Preregistration after data collection is common for secondary data analysis of data obtained by other research groups [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
-We argue that it is still an eligible preregistration.
-In fact, guidelines for clinical trials also recommend analysis of blinded data  to test the practicality of the preregistration [@ICH1998].
-Since the data is obfuscated, virtually no information is revealed to the researchers, and they can not employ their degrees of freedom effectively to influence the results (see section above).
-Shuffling leaves only univariate information, so this procedure is only applicable for bi- or multivariate hypotheses.
-The researcher can still access the information about means or proportions (e.g., the number of participants belonging to group "A" are in the dataset), but remain uninformed about relations between variables (e.g., members of group "A" have a greater mean in variable "Z").
-To be clear, any exploration of the data before preregistration (PAC or otherwise) renders the preregistration invalid and amounts to (albeit sometimes unintentional) scientific misconduct.
-The aim is transparency and the line that shall not be crossed is learning about the research outcomes, but what this means differs for each research project.
-
-Another approach that can replace but also complement simulated data is to conduct a pilot study [@thabaneTutorialPilotStudies2010] and use the collected data to preregister.
-A pilot study has obvious advantages for study planning, since it lets the researcher evaluate the feasibility of many assumptions.
-However, we must warn our readers, that while piloting is more traditional than our approach of obfuscating the data before preregistration, it is not innocuous to use the data resulting from the pilot study.
 
@@
 
+### Alternatives to simulated data
+
+Simulating data may prove challenging to applied researchers.
+In the spirit of team science and collaboration, one feasible solution is to involve a statistical co-author.
+However, several easy alternatives exist. 
+The downside of these alternatives is that they all rely indirectly on the use of real data.
+This introduces a risk that the planned analyses may be cross-contaminated by any exploratory findings.
+It is crucial to disclose any exposure to the data in preparation of the preregistration (PAC or otherwise).
+This exposure to the data may decrease trust in the objectivity of the preregistration.
+Moreover, researchers should take rigorous measures to prevent exposure to exploratory findings that may unintentionally influence their decision making.
+
+The simplest method is to collect empirical data first, but set it aside and proceed with a copy of the data that is blinded by randomly shuffling the order of rows for each variable (independently of each other).
+Shuffling removes any associations between variables, while retaining information about the level of measurement and marginal distribution of each variable.
+If the hypotheses pertain to associations between variables, this treatment should thus be sufficient to revent cross-contamination.
+The researcher can still access the information about means or proportions (e.g., the number of participants belonging to group "A" are in the dataset), but remain uninformed about relations between variables (e.g., members of group "A" have a greater mean in variable "Z").
+<!-- CJ: This is really a different issue.
+You should not bring it up in relation to alternatives to simulated data.
+Preregistration after data collection is common for secondary data analysis of data obtained by other research groups [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
+We argue that it is still an eligible preregistration.-->
+Guidelines for clinical trials already recommend analysis of blinded data  to test the feasibility of a preregistration [@ICH1998].
+
+Another alternative to simulated data is to conduct a pilot study [@thabaneTutorialPilotStudies2010] and use the pilot data to develop the preregistration.
+A pilot study has obvious advantages for study planning, since it lets the researcher evaluate the feasibility of many assumptions.
+However, we must warn our readers, that while piloting is more traditional than our approach of blinding the data before preregistration, it is not innocuous to use the data resulting from the pilot study.<!--CJ What do you mean: innocuous!?-->
+
 #### When Is PAC Applicable?
 
-Every study that can be preregistered and ultimately uses computer code for the statistical analysis, can also be preregistered with code.
+PAC is applicable to every study that can be preregistered and ultimately uses computer code for the statistical analysis.
+<!-- This is all repetition of the rest of the paper. Cut it.
 The only difference is that PAC demands a comprehensive and complete analysis plan.
@@
 In practice, researchers may not feel ready to decide on every detail of the analysis plan before data collection.
-In such cases, we recommend making decisions based on the best available knowledge, and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in Section [Deviating from the preregistration and exploration] in the final manuscript.
-
-Beyond standard preregistrations in psychological research, we believe that two applications specifically lend themselves to PAC.
+In such cases, we recommend making decisions based on the best available knowledge, and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in Section [Deviating from the preregistration and exploration] in the final manuscript.-->
+Two types of preregistrations are particularly amenable to PAC.
 First, pregistrations of clinical trials (called statistical analysis plans, @ICH1998) typically describe analyses in exhaustive detail and typically contain a detailed description of how results will be presented, including shells of tables and graphics [@yuanGuideStatisticalAnalysis2019].
 PAC may significantly reduce the required workload while maintaining (and exceeding) the required standards for preregistering a clinical trial.
-
 Second, preregistering exploratory statistical models (i.e., those with large numbers of competing models or those inspired by machine learning) is hardly feasible with standard preregistrations since they are too complex to describe and depend strongly on their software implementation.
@@
 We ahve argued that PAC has several advantages over classic preregistration and have outlined its implementation.
-To illustrate how PAC works in practice and help researchers to implement PAC themselves, we have conducted one that serves as a guiding example.
-We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:
+To illustrate how PAC works in practice and help researchers to implement PAC themselves, we provide a worked example.
+We will use an exemplary research question that was based on openly available data:
 
@@
 [^caution-mww]: The t-test and Mann-Whitney-Wilcoxon test are arguably the most often used hypothesis tests (according to [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005] reports that 26% of all studies employed a t-test and 27% employed a rank-based alternative in the *New England Journal of Medicine* in 2005).
-The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
-The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann-Whitney-Wilcoxon overestimates effect size with increasing sample size and skewness.
+The analytical strategy presented here is, in fact, suboptimal in several respects (the assumption of measurement invariance is untested [@putnickMeasurementInvarianceConventions2016], the effect size is underestimated in the presence of measurement error [@frostCorrectingRegressionDilution2000], the effect size is overestimated for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).
+The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann-Whitney-Wilcoxon overestimates effect sizes with increasing sample size and skewness.
 
@@
 Second, it highlights how difficult it is to describe a statistical analysis precisely.
-The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and their expectations and beliefs.
+The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and assumptions.
 One example would be using the function `wilcox.test` instead of the combinations of the functions `rank` and `t.test`.
@@
 
-Naturally we did not access the data when we initially wrote this code.
+Following the recommendations outlined in this paper, we did not access the data when we initially wrote this code.
 We therefore did not know the exact format the data would have.
 This means that we did need to change our preregistration after accessing the data to include i.e. the recoding of gender (lines 17-22) and the items (lines 23-26).[^prereg-changes]
+This is our summary of what we changed:
+
+```{r, echo=FALSE, results='asis'}
+show_prereg("discuss-changes")
+```
+
+Readers can inspect and judge the changes for themselves on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
 
@@
 To organize studies with multiple hypotheses, we suggest multiple `planned_analysis()` and `report_analysis()` functions (possibly numbered in accordance with the hypotheses, e.g. 1.2, 2.3 etc.).
-Computational steps that apply to multiple analyses should be extracted into their own function, for example, those that perform data cleaning or extract results from a statistical model in an identical way.
+<!--CJ I would skip this.
+It is a recommendation rooted in computer science, but without explaining it further, resears will not understand the motivation.
+Computational steps that apply to multiple analyses should be extracted into their own function, for example, those that perform data cleaning or extract results from a statistical model in an identical way.-->
 Preregistrations that cover multiple distinct data sources may employ multiple `simulate_data()` functions.
@@
 
-Of course, it is likely that the preregistered code has to be adapted to the actual data.
-Changing the preregistration must be discussed.
-This is our summary of what we changed:
-
-```{r, echo=FALSE, results='asis'}
-show_prereg("discuss-changes")
-```
-
-But readers can inspect and judge the changes for themselves on [GitHub](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
-
-The interested reader can find the example rendered as a [PDF file with real data](https://github.com/aaronpeikert/repro-tutorial/files/7309455/preregistration.pdf) in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.3.1-results) and can view the changes we made since preregistering it on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
-
-After researchers are satisfied with their draft preregistration, they should create an archived and unchangeable version of the project that serves as the preregistration.
+The example rendered as a [PDF file with real data](https://github.com/aaronpeikert/repro-tutorial/files/7309455/preregistration.pdf) is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.3.1-results).
+The changes we made since preregistering it can be inspected on this [GitHub page](https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b).
+
+After researchers are satisfied with their draft preregistration, they should archive a time-stamped and indelible version of the project that serves as the preregistration.
 zenodo.org [@zenodo] is a publicly funded service provider that archives digital artefacts for research and provides digital object identifiers (DOI) for these archives.
@@
 Here you can add all relevant binary files but at least a rendered version of the manuscript and the Docker image.
-
-Taken together, we realize that PACs are challenging.
-Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove impractical for a PAC.
-Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or non-preregistered research but as a rather complementary approach.
-We are convinced that many reasons can justify changing the preregistered code, but equally certain that readers should judge for themselves weather they find these deviations from the preregistration meaningful.
-PAC enables just that.
-Researchers can change their plans and possibly defend their decision to do so, e.g., in the discussion, while readers of the manuscript can clearly see what was changed through version control.
 
@@
 
-We have to distinguish two kinds of limitations.
-First, some limitations are by design and are unlikely to be changed soon.
+We realize that the workflow outlined in this paper, and its application in PAC, remains challenging despite our efforts to simplify the procedure by means of the `repro` package.
+This paper should be considered as a starting point for those seeking to improve the reproducibility of their research.
+Two kinds of limitations can be distinguished.
+The first kind are limitations by design which are unlikely to change.
 Our workflow inherits these from the software it relies on and the fundamental design principles these share with the workflow and `repro`.
-Second, there are other limitations that we hope development efforts of the open-source community and, concerning `repro`, we will solve
-
-As a principle, this workflow is incompatible with steps that cannot be automated.
-This principle is at odds with some ingrained habits of researchers to mix and match manual and automated steps in data analysis.
+The second kind are limitations in `repro` and its dependencies that may be overcome by our future efforts and those of the open-source development community.
+
+With regard to limitations by design,
+the workflow outlined in this paper is fundamentally incompatible with steps that cannot be automated.
+This principle may be at odds with some ingrained habits of researchers to mix and match manual and automated steps in data analysis.
 To allow for automation, many researchers will have to search for alternative software.  
 The automation-friendly software we present here has several technical but critical limitations.
-`Git` can track any files (files larger than 1 GB only with extensions) but can only compare text files (with endings like, `.txt`, `.csv`, `.R`, `.py`,  or `.Rmd`), not binary files (with endings like `.docx`, `.exe`, or `.zip`).
-With `R Markdown`, tables and graphics are dynamically generated from code but cannot be crafted by hand.
-`Make` can orchestrate any programmable software but fails to automate software that will only obey a user clicking on the right buttons.
-`Docker` can ship software that runs on Linux and can be automatically installed, but that requirement precludes much commercial or closed-source software.
+For example, `Git` can track any filetype, but tracked changes are only meaningful for text files (with endings like, `.txt`, `.csv`, `.R`, `.py`,  or `.Rmd`), not for binary files (with endings like `.docx`, `.exe`, or `.zip`).
+Furthermore, tables and graphics dynamically generated from code are difficult to edit by hand.
+`Make` can automate any programmable software, but not software that is exclusively controlled through a point-and-click user interface.
+Finally, `Docker` can ship software that runs on Linux and can be automatically installed, which precludes much commercial or closed-source software.
 
@@
 Second, the upfront investment is required once (and efforts such as `repro` are underway to reduce it) and will pay dividends over many research projects.
-Third, the move towards open software for research offers several benefits beyond enabling automated reproducibility [@schaffnerFutureScientificJournals1994; @fitzgeraldTransformationOpenSource2006; @chaldecottHistoryScientificTechnical1965; @sonnenburgNeedOpenSource2007].
-
-`repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
-Despite our best efforts to ensure this goal, the package is still in development and undoubtedly has limitations.
-The workflow we present here introduces several software solutions.
-This sheer number of dependencies represents in itself a threat to long-term reproducibility.
-To benefit from automatic and convenient reproduction, researchers must use `Git`, `Make`, and `Docker`.
+Third, the move towards open software for research offers several benefits beyond enabling automated reproducibility [@schaffnerFutureScientificJournals1994; @fitzgeraldTransformationOpenSource2006; @chaldecottHistoryScientificTechnical1965; @sonnenburgNeedOpenSource2007].<!--CJ Summarize these.-->
+
+With regard to surmountable limitations, we acknowledge that the `repro` package is still in development.
+One limitation is that `repro` relies on several software dependencies, which represents a threat to long-term reproducibility in itself.
+For example, to benefit from automatic and convenient reproduction, researchers must use `Git`, `Make`, and `Docker`.
 However, `Git` and `Make` are themselves included in the `Docker` image created by `repro`.
@@
 The `Dockerfile` (the plain text description to build a `Docker` image), as opposed to the image, is insufficient because it relies on too many service providers (e.g., Microsoft R Application Network, Ubuntu Package Archive).
-We recommend archiving the `Git` repository and the `Docker` image with zenodo.org, a non-profit long-term storage for scientific data.
+To overcome this limitation, we recommend archiving the `Git` repository and the `Docker` image with zenodo.org, a non-profit long-term storage for scientific data.
 The necessary steps for archival on zenodo.org are described at the end of Section [Preregistration as Code --- a Tutorial].
+
 The second vulnerability is that even if the existence of the `Docker` image and `Git` repository is guaranteed, future researchers still require software to run the image.
@@
 However, this process requires much technical knowledge and is currently not facilitated by `repro`.
-In any case and without such precautions, do the `R Markdown`, the `Makefile` and the `Dockerfile` provide information that allows researchers to trace the computational steps and recreate the computational environment manually.
+With regard to this vulnerability, it is worthwhile to note that the `R Markdown`, `Makefile`, and `Dockerfile` do provide information that allows researchers to trace the computational steps and recreate the computational environment manually.
 The `Makefile`, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are unable to run `Make` for some reason.
-Hypothetically, even if `Docker` were to become unavailable one day, the `Dockerfile` still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
+Thus hypothetically, even if `Docker` were to become unavailable one day, the `Dockerfile` still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.
 
@@
 Open science practices are a continually evolving field where technical innovations foster changes in research practice.
-Open data are much more widespread thanks to online storage facilities; preregistration is possible because there are preregistration platforms and so forth.
-Similarly, we hope that fully automatic reproduction, e.g., with `repro` as a technical innovation, increase scientific rigor and efficiency.
+Open data are much more widespread today thanks to online storage facilities; preregistration is possible because there are preregistration platforms and so forth.
+Similarly, we hope that fully automatic reproduction, e.g., with `repro` as a technical innovation, will promote increased scientific rigor, efficiency, and productivity.
 
 In practice, this ideal of a fully automatic reproduction of research projects can conflict with the wide range of demands for more user-friendly and powerful software.
-Some may find `Make` too complicated or that `Docker` requires too much storage space.
+Some may find that `Make` is too complicated or that `Docker` requires too much storage space.
 Yet others may find that they require other programming languages or want to scale their computation across hundreds of computers, e.g., via high-performance computing clusters or cloud computing.

### 2021-11-22 manuscript.Rmd

--- 2021-11-19 manuscript.Rmd
@@
 *Preregistration* refers to the practice of defining research questions and planning data analysis before observing the research outcomes [@NosekRevolution2018].
-It serves to separate a-priori planned and theory-driven (confirmatory) analyses from unplanned and post-hoc (exploratory) analyses.<!-- CJ I removed the bit about credibility, because that requiresfurther explanation of why separating these two types results in more credible findings-->
+It serves to separate a-priori planned and theory-driven (confirmatory) analyses from unplanned and post-hoc (exploratory) analyses.
 Exploratory analyses refer to both those aimed at discovering patterns in data, and those aimed at overcoming shortcomings of confirmatory analyses that come to light after seeing the data (e.g., violations of assumptions, modifications to address poor model fit).
@@
 There are still several shortcomings to preregistration.
-One is that written study plans are often multiinterpretable.
+One is that written study plans are often interpretable in multiple ways.
 Empirical research has shown that, even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, their results vary considerably [@silberzahnManyAnalystsOne2018].
@@
 PAC has the potential to eliminate undisclosed researchers degrees of freedom to a much greater extent than, e.g., preregistration templates.
-Moreover, it reduces overhead by obviating the need to write a separate preregistration and manuscript.
+Moreover, it reduces overhead by removing the need to write a separate preregistration and manuscript.
 For PAC, researchers can write a reproducible, dynamically generated draft of their intended manuscript at the preregistration stage.
@@
 Reproducibility is of utmost importance at this stage since the preregistration must produce valid results at two points in time, once before data collection and once after data collection.
-As outlined before, reproducibility builds upon four pillars (version control, dynamic document generation, dependency tracking, and software management).<!-- CJ I don't understand why you repeat this here.
-Remove?-->
-To use PAC the dangers to reproducibility we described must be  eliminated.<!-- CJ I don't understand why you repeat this here. Remove?-->
+As outlined before, reproducibility builds upon four pillars (version control, dynamic document generation, dependency tracking, and software management).
+To use PAC the dangers to reproducibility we described must be  eliminated.
 
@@
 
-We believe that pairing PAC with the workflow presented above offers several advantages over classical preregistration.
+We believe that pairing PAC with the workflow presented above offers five advantages over classical preregistration.
 First, PAC is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating, two separate documents.
@@
 Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
-PAC therefore minimizes undisclosed researchers degrees of freedom more effectively than standard preregistration does [@bakkerEnsuringQualitySpecificity2020; @wichertsDegreesFreedomPlanning2016].
-Fourth, despite its rigor, PAC accommodates partly exploratory data-dependent decisions.
-Some decisions can be anticipated and formulated as conditional code.
+PAC, therefore, minimizes undisclosed researchers degrees of freedom more effectively than standard preregistration does [@bakkerEnsuringQualitySpecificity2020; @wichertsDegreesFreedomPlanning2016].
+Fourth, despite its rigor, PAC accommodates data-dependent decisions if these can be formulated as code.
 Researchers can, for example, formulate conditions (e.g., in the form of if-else-blocks) under which they prefer one analysis type over the other.
 For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
-Some necessary deviations from the preregistration only become apparent after seeing the data.<!--CJ I merged point 5 here; it's part of point 4-->
-Version control makes it possible to commit specific changes to the planned analyses.
-It is recommended to documented all such changes by maintaining a changelog, and referencing specific commits.<!--CJ I do this myself, I think it's a good idea.-->
 Another example of data-dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
 Decisions that do not lend themselves to formulation in code, e.g. visual inspection, must still be described verbally or be treated as noted in the next section.
-With PAC researchers can precisely calibrate how much they want to explore.<!--CJ This is vague-->
-When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.<!-- CJ Not clear what this refers to; you haven't given an example or anything!
-Cut?-->
+Fifth, deviations from the preregistration are clearly documented because they are reflected in changes to the code, which are managed and tracked with version control.
 
@@
 Often, deviations provide an opportunity to learn from the unexpected.
-Transparency is essential in this regard <!--CJ: it is not clear how the following argument justifies the need for transparency.
-Please fix--> because an empirical test of a theory can not test the theory alone but tests it in conjunction with auxiliary assumptions [@meehlTheoreticalRisksTabular1978].
 
 For example, imagine that authors preregistered their intention to include both "working memory" and "fluid intelligence" as covariates in an experimental study, examining the effect of task novelty on reaction time.
-When evaluating the planned analyses on the empirical data, these two covariates reveal high collinearity, thus biasing estimates and compromising statistical inference.
+When evaluating the planned analyses on the empirical data, these two covariates reveal high collinearity, thus compromising statistical inference.
 The authors decide to use PCA to extract common variance related to "intelligence", and include this component as a covariate instead.
@@
 Now imagine that a different researcher is interested in the structure of intelligence.
-This change to the preregistration directly relates their theory of intelligence.
+This change to the preregistration directly relates to their theory of intelligence.
 That researcher might thus interpret the same result as an explorative finding, suggesting that these aspects of intelligence are unidimensional.
-<!--CJ not sure what this final line intended to argue. I removed it. Feel free to add a more satisfactory conclusion to this paragraph-->
+A deviation from preregistration thus requires a judgement about what changes affect the test of the theory [@meehlTheoreticalRisksTabular1978].
+Only transparent reporting enables such judgment.
 
@@
 Such traditional exploratory findings should be explicitly declared in the manuscript to distinguish them from confirmatory findings [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
-The second kind involves rigorous exploration using methods that are standard in machine learning [@brandmaier22mlsem].
+The second kind of exploration is through procedurally well defined exploration with exploratory statistical models are standard in machine learning [@brandmaier22mlsem].
 These models often involve dozens, if not hundreds of predictors, which makes it difficult to describe them verbally.
-With PAC, such models can be preregistered clearly and in comprehensive detail.
+With PAC, such models can be preregistered clearly and in comprehensive detail and researcher can therefore precisely calibrate how much they want to explore.
 We specifically recommend PAC for such exploratory statistical models.
 The merit of preregistration in these cases is to communicate precisely how much exploration was done;
-a piece of information that is crucial to assess e.g., whether the results might be overfit <!--CJ reference Hastie & Tibshirani-->.
-
-## How to Write a Good PAC
-
-When planning a study and writing a PAC, researchers ought to strive to meet three goals:<!--CJ: None of these goals relate explicitly to code.
-Is that a problem?
-Also: Shouldn't you mention the goals when you first introduce PAC?-->
-First, the plan should be comprehensive.
-That is, researchers need to state the research question and describe the study design and analysis in detail.
-Second, the plan should be effective, meaning that the researchers should reasonably expect it to answer the research question at hand.
-Third, the plan should be efficient; i.e., it should only require the necessary amount of resources.
-
-### Comprehensiveness
-
-Standards for scientific writing are designed to result in understandable and comprehensive research articles and can serve the same purpose for writing a preregistration.
-For example, the APA standards [@apa7] require research articles describe a study in such detail that any researcher that is familiar with the field is able to repeat it.
-These well-known standards thus provide a good starting point for the structure of a PAC.<!--CJ What are you arguing here?
-You've just convinced the reader that existing templates are not comprehensive enough.
-Now you're arguing that APA7 can be considered a preregistration template, and that it is comprehensive?
-I am a little confused here-->
-One particular advantage is introduction's and theoretical background's embedding in existing literature where that provide the basis for the study's design.<!--CJ This sentence has been missed by the language editor-->
-Further inspiration may be taken from preregistration templates relevant to the type of study at hand,
-because such templates address methodological rigor more comprehensively.
-One factor that limits the comprehensiveness of published scientific reports is the word count imposed by journals.
-If this is a concern, we suggest relegating the PAC to a reproducible appendix to keep the core manuscript concise.
-
-Simulating data offers further benefits in ensuring the comprehensiveness of a preregistration.
-It forces the researcher to consider conceptual data-analytic challenges before obtaining empirical data [@axelrodAdvancingArtSimulation1997].
-It also allows researchers to verify that planned analyses are feasible and likely to run without failure [@braiekTestingMachineLearning2020] (e.g., does the code also work with missing values, extreme outliers, etc.).
-Many errors and omissions are likely to come to light when evaluating the intended analysis code on simulated data.
-
-One final method to ensure comprehensivenes of the PAC is peer review.
-The value of PAC would be further enhanced by submitting it as a registered report,
-a format in which articles are accepted for publication "in principle" based on peer review of the preregistered study plans [@nosekRegisteredReports2014; @chambersWhatNextRegistered2019].
-In such registered report peer-reviewing a manuscript-styled PAC may well make it easier for the reviewers that ascertain and evaluate a comprehensive study plan.
-Moreover, reviewers can suggest specific improvements to the planned analysis code.
-
-### Effectiveness
-
-If the comprehensiveness criterion is met, researchers can further check whether the planned analyses are effective.
-Effectiveness can be defined in terms of ability to recover true population parameters, or power to detect a population effect.<!--CJ You have not said anything about a simulation yet, so this is kind of surprising-->.
-Effectiveness can be assessed by means of a simulation study.
-Once code has been written to simulate synthetic data, it is often a minor step to perform a simulation study.
-It is sufficient to repeat the planned analyses many times, and plot the distribution of estimated parameters, or tabulate the proportion of replications in which a true effect was detected (power analysis).
-The R command `replicate()` can be used to this end.
-If an analysis fails to consistently recover the population parameters under simulated (idealized) conditions,
-it would be unreasonable to expect this analysis to be effective under realistic conditions.
-Because the simulation describes the researchers' expectations about real-world conditions,
-it is the best way to assess the effectiveness of the analysis prior to collecting empirical data.
-Simulating data requires researchers to explicate their assumptions about the data-generating mechanism in concrete computational terms; e.g.
-assumptions about the distributions of variables and how they relate to one another.
-If a researcher experiences difficulty in specifying such assumptions, this may indicate that further literature research or exploratory pilot studies are required.
-
-### Efficiency<!-- CJ I do not understand how 'efficiency' is different from effectiveness above.
-You cannot assess effectiveness by means of a single iteration.
-Moreover, efficiency has a specific statistical definition (pertaining to the amount of information required to estimate a parameter).
-Using this term can be highly confusing to our (probably statistically inclined) readership.-->
-
-A single simulated dataset already helps to ensure that the plan is viable under certain circumstances.
-However, repeated simulation with varying parameters, a so-called Monte Carlo simulation, lets researchers play through several scenarios.
-Monte Carlo simulation lets researchers navigate several trade-offs regarding their resources.
-In the simplest case, they can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study [@brandmaier2018precision].
-@bakkerRecommendationsPreregistrationsInternal2020 found that a formal power analysis may help researchers to invest resources more efficiently.
-Researchers may also evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
-
-### Planned Analyses as Functions
+a piece of information that is crucial to assess e.g., whether the results might be overfit [@hastieElementsStatisticalLearning2017, p. 220f.].
+
+## Planned Analyses as Functions
 
@@
 Dedicated packages to simulate data for specific analyses exist <!--CJ Reference here-->.
-For linear models, simulating data is extremely simple:
-
-```{r}
-simulate_data <- function(n, effect_size){
-  # Draw n samples from a normal distribution for predictor X
-  x <- rnorm(n)
-  # Calculate dependent variable Y as a function of population effect size and residual error
-  y <- effect_size * x + rnorm(n)
-  # Return a data.frame
-  data.frame(x = x, y = y)
-}
-simulate_data(5, .2)
-```
 
@@
 
+For linear models, simulating data is extremely simple:
+
+```{r, eval=FALSE}
+simulate_data <- function(n, effect_size){
+  # Draw n samples from a normal distribution for predictor X
+  x <- rnorm(n)
+  # Calculate dependent variable Y
+  # as a function of population effect size and residual error
+  y <- effect_size * x + rnorm(n)
+  # Return a data.frame
+  data.frame(x = x, y = y)
+}
+```
+
 Next, write a function to conduct the planned analysis.
 This function should receive the data and compute all relevant results from it.
-In the simplest case, an analysis function might already exist in R.
-For the linear model above, the analysis function might be:
-
-```{r}
-lm(y ~ x, data = simulate_data(5, .2))
-```
-
 The general format of an analysis function might be:
 
-```{r}
+```{r, eval=FALSE}
 planned_analyis <- function(data){
@@
   # 3. `return(results)`
+}
+```
+
+In the simplest case, an analysis function might already exist in R.
+For the linear model above, the analysis function might be:
+
+```{r, eval=FALSE}
+planned_analyis <- function(data){
+    lm(y ~ x, data = data)
 }
@@
 Shuffling removes any associations between variables, while retaining information about the level of measurement and marginal distribution of each variable.
-If the hypotheses pertain to associations between variables, this treatment should thus be sufficient to revent cross-contamination.
+If the hypotheses pertain to associations between variables, this treatment should thus be sufficient to prevent cross-contamination.
 The researcher can still access the information about means or proportions (e.g., the number of participants belonging to group "A" are in the dataset), but remain uninformed about relations between variables (e.g., members of group "A" have a greater mean in variable "Z").
-<!-- CJ: This is really a different issue.
-You should not bring it up in relation to alternatives to simulated data.
 Preregistration after data collection is common for secondary data analysis of data obtained by other research groups [@westonRecommendationsIncreasingTransparency2019] but not so much within the same research project.
-We argue that it is still an eligible preregistration.-->
+We argue that it is still an eligible preregistration.
 Guidelines for clinical trials already recommend analysis of blinded data  to test the feasibility of a preregistration [@ICH1998].
@@
 A pilot study has obvious advantages for study planning, since it lets the researcher evaluate the feasibility of many assumptions.
-However, we must warn our readers, that while piloting is more traditional than our approach of blinding the data before preregistration, it is not innocuous to use the data resulting from the pilot study.<!--CJ What do you mean: innocuous!?-->
+However, we must warn our readers, that while piloting is more traditional than our approach of blinding the data before preregistration, the data from the pilot study must not enter the analysis data set.
 
@@
 PAC is applicable to every study that can be preregistered and ultimately uses computer code for the statistical analysis.
-<!-- This is all repetition of the rest of the paper. Cut it.
-The only difference is that PAC demands a comprehensive and complete analysis plan.
-With traditional preregistration, researchers may (unintentionally) omit details and thus introduce decisions that appear to be made after having seen the data.
-We emphasize again that deviations from the preregistration are not only possible but likely.
-The advantage of PAC is transparency.
-All deviations from the planned analysis as implemented in code are tracked, and the final analysis can be easily compared to the PAC.
-Importantly, we see PAC as a means to increase transparency not as a means to limit researchers in their statistical analyses.
-In practice, researchers may not feel ready to decide on every detail of the analysis plan before data collection.
-In such cases, we recommend making decisions based on the best available knowledge, and if these decisions still prove infeasible, make simple assumptions and revise and explain them later as described in Section [Deviating from the preregistration and exploration] in the final manuscript.-->
 Two types of preregistrations are particularly amenable to PAC.
@@
 PAC may significantly reduce the required workload while maintaining (and exceeding) the required standards for preregistering a clinical trial.
+
 Second, preregistering exploratory statistical models (i.e., those with large numbers of competing models or those inspired by machine learning) is hardly feasible with standard preregistrations since they are too complex to describe and depend strongly on their software implementation.
@@
 To organize studies with multiple hypotheses, we suggest multiple `planned_analysis()` and `report_analysis()` functions (possibly numbered in accordance with the hypotheses, e.g. 1.2, 2.3 etc.).
-<!--CJ I would skip this.
-It is a recommendation rooted in computer science, but without explaining it further, resears will not understand the motivation.
-Computational steps that apply to multiple analyses should be extracted into their own function, for example, those that perform data cleaning or extract results from a statistical model in an identical way.-->
 Preregistrations that cover multiple distinct data sources may employ multiple `simulate_data()` functions.
@@
 
-After researchers are satisfied with their draft preregistration, they should archive a time-stamped and indelible version of the project that serves as the preregistration.
+After researchers are satisfied with their draft preregistration, they should archive a time-stamped and uneditable version of the project that serves as the preregistration.
 zenodo.org [@zenodo] is a publicly funded service provider that archives digital artefacts for research and provides digital object identifiers (DOI) for these archives.
@@
 Second, the upfront investment is required once (and efforts such as `repro` are underway to reduce it) and will pay dividends over many research projects.
-Third, the move towards open software for research offers several benefits beyond enabling automated reproducibility [@schaffnerFutureScientificJournals1994; @fitzgeraldTransformationOpenSource2006; @chaldecottHistoryScientificTechnical1965; @sonnenburgNeedOpenSource2007].<!--CJ Summarize these.-->
+Third, the move towards open software for research offers several benefits beyond enabling automated reproducibility [@schaffnerFutureScientificJournals1994; @fitzgeraldTransformationOpenSource2006; @chaldecottHistoryScientificTechnical1965; @sonnenburgNeedOpenSource2007].
 

### 2021-11-22 manuscript.Rmd
preregistration as code -> consistent capitalization

@@
   This paper introduces `repro`, an R package that guides researchers through installation and use of the tools required to make a research project reproducible.
-  We also suggest using the proposed workflow for the preregistration of study plans as reproducible computer code (preregistration as code; PAC).
+  We also suggest using the proposed workflow for the preregistration of study plans as reproducible computer code (Preregistration as Code; PAC).
   Since computer code represents the planned analyses exactly as they will be executed, it is more precise than natural language descriptions. 
@@
 
-We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations, such as preregistration as code (PAC).
+We finally presented an example of how the rigorous and automated reproducibility workflow introduced by `repro` may enable other innovations, such as Preregistration as Code (PAC).
 In PAC the entire reproducible manuscript, including planned analyses and results based on simulated data, is preregistered.

### 2021-11-22 manuscript.Rmd
correct citations

@@
 We explicitly encourage readers to try out all R commands for themselves.
-Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).
+Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R [@R-base, tested with version `r with(R.version, paste0(major, ".", minor))`].
 
@@
 Furthermore, using `Docker` may require you to install software on an operating system that may not be familiar to you.
-The images supplied by @boettigerIntroductionRockerDocker2017, for example, are based on Linux.
+The images supplied by the rocker project [@boettigerIntroductionRockerDocker2017], for example, are based on Linux.
 
@@
 We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
-These wrappers are merely a support system; as users learn to use the underlying tools, they can rely less on `repro` and use these ttools directly to solve more complex problems.
+These wrappers are merely a support system; as users learn to use the underlying tools, they can rely less on `repro` and use these tools directly to solve more complex problems.
 
@@
 
-Once the `R Markdown` file has been rendered to a static document, the resulting file is decoupled from the `R Markdown` and the code that created it.
+Once the `R Markdown` file has been rendered to a static document (the output, e.g., PDF), the resulting file is decoupled from the `R Markdown` and the code that created it.
 This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
@@
 Empirical research has shown that, even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, their results vary considerably [@silberzahnManyAnalystsOne2018].
-The current best practice to ensure comprehensive and specific preregistration is to impose structure by following preregistration templates [@bowmanOSFPreregTemplate2020, @bakkerEnsuringQualitySpecificity2020].
+The current best practice to ensure comprehensive and specific preregistration is to impose structure by following preregistration templates [@bowmanOSFPreregTemplate2020; @bakkerEnsuringQualitySpecificity2020].
 However, such templates cannot ensure full transparency because it is impossible to verbally describe every detail of an analysis for any but the most straightforward analysis.

### 2021-11-22 manuscript.Rmd
this is no binary choice

@@
 That researcher might thus interpret the same result as an explorative finding, suggesting that these aspects of intelligence are unidimensional.
-A deviation from preregistration thus requires a judgement about what changes affect the test of the theory [@meehlTheoreticalRisksTabular1978].
+A deviation from preregistration thus requires a judgement about what changes affect the test of the theory to what extend [@meehlTheoreticalRisksTabular1978].
 Only transparent reporting enables such judgment.
@@
 These models often involve dozens, if not hundreds of predictors, which makes it difficult to describe them verbally.
-With PAC, such models can be preregistered clearly and in comprehensive detail and researcher can therefore precisely calibrate how much they want to explore.
+With PAC, such models can be preregistered clearly and in comprehensive detail and researcher can precisely define a priori how much they want to explore.
 We specifically recommend PAC for such exploratory statistical models.

### 2021-11-22 manuscript.Rmd
add warning

@@
 simulate_data <- function(n, effect_size){
+  warning("This manuscript contains mock results based on simulated data.")
   # Draw n samples from a normal distribution for predictor X
   x <- rnorm(n)
-  # Calculate dependent variable Y
-  # as a function of population effect size and residual error
+  # Calculate dependent variable Y..
+  #.. as a function of population effect size and residual error
   y <- effect_size * x + rnorm(n)

### 2021-11-23 manuscript.Rmd
align headin levels
force build for examples even when up to date

@@
 cat("```bash",
-  stringi::stri_wrap(system2("make", " --dry-run data/simulation_results.csv", stdout = TRUE)),
+  stringi::stri_wrap(system2("make", "-B --dry-run data/simulation_results.csv", stdout = TRUE)),
   "```", sep = "\n")
@@
 cat("```bash",
-  stringi::stri_wrap(system2("make", " --dry-run DOCKER=TRUE simulated_data.csv", stdout = TRUE)),
+  stringi::stri_wrap(system2("make", "-B --dry-run DOCKER=TRUE simulated_data.csv", stdout = TRUE)),
   "```", sep = "\n")
@@
 
-### Alternatives to simulated data
+## Alternatives to simulated data
 
@@
 
-#### When Is PAC Applicable?
+## When Is PAC Applicable?
 
@@
 
-We ahve argued that PAC has several advantages over classic preregistration and have outlined its implementation.
+We have argued that PAC has several advantages over classic preregistration and have outlined its implementation.
 To illustrate how PAC works in practice and help researchers to implement PAC themselves, we provide a worked example.

### 2021-11-23 manuscript.Rmd
remove false dichotomy most conformatory analyses are aimed at discovering patterns -> weak theories
missing word

@@
 It serves to separate a-priori planned and theory-driven (confirmatory) analyses from unplanned and post-hoc (exploratory) analyses.
-Exploratory analyses refer to both those aimed at discovering patterns in data, and those aimed at overcoming shortcomings of confirmatory analyses that come to light after seeing the data (e.g., violations of assumptions, modifications to address poor model fit).
 Researchers are faced with a myriad of choices in designing, executing, and analyzing a study, often called researchers degrees of freedom.
@@
 Such traditional exploratory findings should be explicitly declared in the manuscript to distinguish them from confirmatory findings [@NosekRevolution2018; @nosekPreregistrationHardWorthwhile2019].
-The second kind of exploration is through procedurally well defined exploration with exploratory statistical models are standard in machine learning [@brandmaier22mlsem].
+The second kind of exploration is through procedurally well defined exploration with exploratory statistical models that are standard in machine learning [@brandmaier22mlsem].
 These models often involve dozens, if not hundreds of predictors, which makes it difficult to describe them verbally.

### 2021-11-23 manuscript.Rmd
move zenodo section

@@
 
-To summarize, researchers need to write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration.
+### Turning a Dynamic Document into a Preregistration
+
+After researchers are satisfied with their draft preregistration, they should archive a time-stamped and uneditable version of the project that serves as the preregistration.
+zenodo.org [@zenodo] is a publicly funded service provider that archives digital artefacts for research and provides digital object identifiers (DOI) for these archives.
+While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
+Please note that you can only link public GitHub repositories to zenodo.org.
+You may log into zenodo.org through your GitHub account.
+To log in with your GitHub account:
+
+> Navigate to https://zenodo.org/login/ →  Log in with GitHub
+
+\noindent{}To link zenodo.org and GitHub
+
+> Log into zenodo.org → Account → GitHub[^zenodo-github]
+
+\noindent{}Or:
+
+> Navigate to https://zenodo.org/account/settings/github/
+
+[^zenodo-github]: https://zenodo.org/account/settings/github/
+
+After you have linked a GitHub repository, you trigger the archival by creating a GitHub release.
+To create GitHub release, navigate to GitHub:
+
+```{r, eval=FALSE}
+usethis::browse_github()
+```
+
+Then click on Releases → Draft a new release.
+Here you can add all relevant binary files but at least a rendered version of the manuscript and the Docker image.
+
+To summarize, researchers need to write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration in an uneditable online repository.
 After they gathered the actual data, they can replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data), and write the discussion.
@@
 
-# Preregistration as Code: Tutorial
+## Preregistration as Code: Tutorial
 
@@
 
-After researchers are satisfied with their draft preregistration, they should archive a time-stamped and uneditable version of the project that serves as the preregistration.
-zenodo.org [@zenodo] is a publicly funded service provider that archives digital artefacts for research and provides digital object identifiers (DOI) for these archives.
-While the service is independent of GitHub---in terms of storage facilities and financing---you can link GitHub and zenodo.org.
-Please note that you can only link public GitHub repositories to zenodo.org.
-You may log into zenodo.org through your GitHub account.
-To log in with your GitHub account:
-
-> Navigate to https://zenodo.org/login/ →  Log in with GitHub
-
-\noindent{}To link zenodo.org and GitHub
-
-> Log into zenodo.org → Account → GitHub[^zenodo-github]
-
-\noindent{}Or:
-
-> Navigate to https://zenodo.org/account/settings/github/
-
-[^zenodo-github]: https://zenodo.org/account/settings/github/
-
-After you have linked a GitHub repository, you trigger the archival by creating a GitHub release.
-To create GitHub release, navigate to GitHub:
-
-```{r, eval=FALSE}
-usethis::browse_github()
-```
-
-Then click on Releases → Draft a new release.
-Here you can add all relevant binary files but at least a rendered version of the manuscript and the Docker image.
-
 # Discussion

### 2021-11-24 manuscript.Rmd
fix typo

@@
 ```{r, include=FALSE}
-fig_caption <- str_c("Results of simulation for the power analysis. The cross indicates the sample size that archives ", choosen_power*100, "% assuming a Cohen's D of ", choosen_d, ".")
+fig_caption <- str_c("Results of simulation for the power analysis. The cross indicates the sample size that archives ", choosen_power*100, "% assuming a Cohen's d of ", choosen_d, ".")
 ```
@@
        x = "Sample size",
-       color = "Cohen's D") +
+       color = "Cohen's d") +
   NULL

### 2021-12-01 manuscript.tex
commit original .tex from submission

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[psych,tutorial,submit,moreauthors,pdftex]{mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{4pt}}
\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}

\usepackage{longtable}

% see https://stackoverflow.com/a/47122900
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2019}
\copyrightyear{2019}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Reproducible Research in R: A Tutorial on how to Do the Same
Thing More Than Once}

% Authors, for the paper (add full first names)
\Author{Aaron Peikert$^{1, 2,
*}$\href{https://orcid.org/0000-0001-7813-818X}{\orcidicon}, Caspar J.
van Lissa$^{3,
4}$\href{https://orcid.org/0000-0002-0808-5024}{\orcidicon}, Andreas M.
Brandmaier$^{1, 5,
6}$\href{https://orcid.org/0000-0001-8765-6982}{\orcidicon}}

% Authors, for metadata in PDF
\AuthorNames{Aaron Peikert, Caspar J. van Lissa, Andreas M. Brandmaier}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Center for Lifespan Psychology---Max Planck Institute for
Human Development, Lentzeallee 94, 14195 Berlin, Germany; \\
$^{2}$ \quad Humboldt-Universität zu Berlin, Berlin, Germany; \\
$^{3}$ \quad Department of Methodology \& Statistics---Utrecht
University, Faculty of Social and Behavioral Sciences, Utrecht,
Netherlands; \\
$^{4}$ \quad Open Science Community Utrecht, Utrecht, Netherlands; \\
$^{5}$ \quad Max Planck UCL Centre for Computational Psychiatry and
Ageing Research Berlin, Germany, and London, UK; \\
$^{6}$ \quad MSB Medical School Berlin, Berlin, Germany; \\
}
% Contact information of the corresponding author
\corres{Correspondence: \href{mailto:peikert@mpib-berlin.mpg.de}{\nolinkurl{peikert@mpib-berlin.mpg.de}}}

% Current address and/or shared authorship

% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary
\simplesumm{Reproducibility has long been considered integral to the
scientific method. An analysis is considered reproducible if an
independent person can obtain the same results from the same data. Until
recently, detailed descriptions of methods and analyses were the primary
instrument for ensuring scientific reproducibility. Technological
advancements now enable scientists to achieve a more comprehensive
standard that allows anyone to access a digital research repository and
reproduce all computational steps from raw data to final report,
including all relevant statistical analyses, with a single command. This
method has far-reaching implications for scientific archiving,
reproducibility and replication, scientific productivity, and the
credibility and reliability of scientific knowledge. One obstacle to the
widespread use of this method is that the underlying tools are complex
and not part of most researchers' basic training. This paper introduces
\texttt{repro}, an R package that guides researchers through
installation and use of the tools required to make a research project
reproducible. We also suggest using the proposed workflow for the
preregistration of study plans as reproducible computer code
(Preregistration as Code; PAC). Since computer code represents the
planned analyses exactly as they will be executed, it is more precise
than natural language descriptions. PAC circumvents the shortcomings of
ambiguous preregistrations that may result in undisclosed use of
researcher degrees of freedom. Reproducibility, facilitated by
automation, has a wide range of applications and could potentially
accelerate scientific progress.}

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{Computational reproducibility is the ability to obtain
identical results from the \emph{same} data with the \emph{same}
computer code. It is a building block for transparent and cumulative
science because it enables the originator and other researchers, on
other computers and later in time, to reproduce and thus understand how
results came about while avoiding a variety of errors that may lead to
erroneous reporting of statistical and computational results. In this
tutorial, we demonstrate how the R package \texttt{repro} supports
researchers in creating fully computationally reproducible research
projects with tools from the software engineering community. Building
upon this notion of fully automated reproducibility we present several
applications including the preregistration of research plans with code
(Preregistration as Code, PAC). PAC eschews all ambiguity of traditional
preregistration and offers several more advantages. Making technical
advancements that serve reproducibility more widely accessible for
researchers holds the potential to innovate the research process to
become more productive, credible, and reliable.}

% Keywords
\keyword{open science; computational reproducibility; preregistration;
R; R Markdown; Make; GitHub; Docker}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Pandoc citation processing

\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, numbers=left}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\captionsetup{labelformat=empty}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Scientists increasingly strive to make research data, materials, and
analysis code openly available. Sharing these digital research products
can increase scientific efficiency by enabling researchers to learn from
each other, reuse materials, and increase scientific reliability by
facilitating the review and replication of published research results.
To some extent, these potential benefits are contingent on whether these
digital research products are reproducible. Reproducibility can be
defined as the ability of anyone to obtain identical results from the
\emph{same} data with the \emph{same} computer code \citep[see][ for
details]{Peikert2019}. The credibility of empirical research results
hinges on their objectivity. Objectivity in this context means, ``in
principle it {[}the research finding{]} can be tested and understood by
anybody.'' \citep[p.~22]{popperLogicScientificDiscovery2002}. Only a
reproducible result meets these requirements. Therefore, reproducibility
has long been considered integral to empirical research. Unfortunately,
despite increasing commitment to open science practices and good will,
many projects can not yet be reproduced by other research teams
\citep{obels2020}. This is because there are various challenges to
making a research project reproducible (e.g., missing software
dependencies or ambiguous documentation of the exact computational steps
taken), and there is a lack of best practices for overcoming these
challenges \citep[but see][]{vanlissa2020worcs}. With technological
advancement, however, it is now possible to make all digital products
related to a research project available in a manner that enables
automatic reproduction of the research project with minimal effort.

With this paper, we pursue two aims. First, we want to introduce
researchers to a notion of automated reproducibility that requires no
manual steps, apart from the initial setup of the software environment.
Secondly, we discuss the implications of automated reproducibility for
changing the general approach to research. With regard to the first
goal, we discuss how to address four common threats to reproducibility,
using tools originating from software engineering \citep[see][ for
details]{Peikert2019}, and present a tutorial on how to employ these
tools to achieve automated reproducibility. A single tutorial cannot
comprehensively introduce the reader to the detail of individual tools,
but this tutorial is intended to help readers get started with a basic
workflow. The tutorial is aimed at researchers who regularly write code
to analyze their data and are willing to make relevant code, data, and
materials available, either publicly or on request. Ideally, the reader
has already created a dynamic document at some point in time (e.g., with
R Markdown or Jupyter) and used some form of version control (e.g.,
Git). The R package \texttt{repro} supports researchers in setting up
the required software and in adopting this workflow. We present
automated reproducibility as a best practice; a goal that is not always
fully achieved due to limited resources, technical restrictions, or
practical considerations, but is worth striving for nonetheless.

In pursuit of the second aim, we present a strictly reproducible and
unambiguous form of preregistration \citep{NosekRevolution2018} that
builds upon implementing this reproducible workflow, the so-called
\emph{Preregistration as Code} (PAC). PAC involves preregistering the
intended analysis code and the major part of the final scientific report
as a dynamic document, including typical sections like introduction,
methods, and results. The resulting dynamic document closely resembles
the final manuscript but uses simulated data to generate placeholder
results (e.g., figures, tables, and statistics). Simulated data serve
two functions, they allow to test the code for the planned analyses and
for preregistering the exact presentation of the results. Once the
empirical data are available, these replace the simulated data; the
results are then updated automatically, and the discussion can be
written to finalize the report.

Scientific organizations and funding bodies increasingly demand
transparent sharing of digital research products, and researchers are
increasingly willing to do so. However, although the sharing of such
digital research products is a necessary condition for reproducibility,
it is not a sufficient one. This was illustrated by an attempt to
reproduce results from open materials in the journal \emph{Cognition}
\citep{hardwicke2018}. Out of 35 published articles with open code and
data, results of 22 articles could be reproduced, but in 11 of these
cases, further assistance from the original authors was required. For 13
articles, at least 1 outcome could not be reproduced-----even with the
original authors' assistance. Another study of 62 Registered Reports
found that only 41 had data available, and 37 had analysis scripts
available \citep{obels2020}. The authors could execute only 31 of the
scripts without error and reproduce the results of only 21 articles
(within a reasonable time). These failed attempts to reproduce findings
highlight the need for widely accepted reproducibility standards because
open repositories do not routinely provide sufficient information to
reproduce relevant computational and statistical results. If digital
research products are available but not reproducible, their added value
is limited.

This tutorial demonstrates how R users can make digital research
products more reproducible, while striking a balance between rigor and
ease-of-use. A rigorous standard increases the likelihood that a project
will remain reproducible as long as possible. An easy-to-use standard,
on the other hand, is more likely to be adopted. Our approach is to
promote broad adoption of such practices by ensuring a ``low
threshold'', by making it easy to get started, while enabling a ``high
ceiling'' by ensuring that they are compatible with more complex
rigorous solutions. As researchers become more proficient in using the
tools involved, they can thus further improve the reproducibility of
their work.

We have structured the tutorial with a \emph{learning-by-doing} approach
in mind, such that readers can follow along on their own computers. We
explicitly encourage readers to try out all R commands for themselves.
Unless stated otherwise, all code blocks are meant to be run in the
statistical programming language R \citep[tested with version
4.0.4]{R-base}.

\hypertarget{threats-to-reproducibility-and-appropriate-remedies}{%
\section{Threats to Reproducibility and Appropriate
Remedies}\label{threats-to-reproducibility-and-appropriate-remedies}}

From our own experience with various research projects, we have
identified the following common threats to reproducibility:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \emph{Multiple inconsistent versions of code, data, or both}; for
  example, the data set may have changed over time because outliers were
  removed at a later stage or an item was later recoded; or, the
  analysis code may have been modified during the writing of a paper
  because a bug was removed at some point in time. It may then be
  unclear which version of code and data was used to produce some
  reported set of results.
\item
  \emph{Copy-and-paste errors}; for example, results are often manually
  copied from a statistical computing language into a text processor; if
  a given analysis is re-run and results are manually updated in the
  text processor, this may inadvertently lead to inconsistencies between
  the reported result and the reproduced result.
\item
  \emph{Undocumented or ambiguous order of computation}; for example,
  with multiple data and code files, it may be unclear which scripts
  should be executed in what order; or, some of the computational steps
  are documented (e.g., final analysis), but other steps were conducted
  manually without documentation (e.g., executing a command manually
  rather than in a script; copy-and-pasting results from one program to
  another).
\item
  \emph{Ambiguous software dependencies}; for example, a given analysis
  may depend on a specific version of a specific software package, or
  rely on software that might not be available on a different computer,
  or no longer exist at all; or a different version of the same software
  may produce different results.
\end{enumerate}

We have developed a workflow that achieves long-term and cross-platform
computational reproducibility of scientific data analyses. It leverages
established tools and practices from software engineering and rests on
four pillars that address the aforementioned causes of
non-reproducibility \citep{Peikert2019}:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Version control
\item
  Dynamic document generation
\item
  Dependency tracking
\item
  Software management
\end{enumerate}

The remainder of this section briefly explains why each of these four
building blocks is needed and details their role in ensuring
reproducibility. A more extensive treatment of these tools is given in
\citet{Peikert2019}.

\emph{Version control} prevents the ambiguity that arises when multiple
versions of code and data are created in parallel during the lifetime of
a research project. Version control allows a clear link between which
results were generated by which version of code and data. This addresses
the first threat to reproducibility, because results can only be said to
be reproducible if it is clear which version of data and code produced
them. We recommend using \texttt{Git} for version control, because of
its widespread adoption in the R community.

\texttt{Git} tracks changes to all project-related files (e.g.,
materials, data, and code) over time. At any stage, individual files or
the entire project can be compared to, or reverted to, an earlier
version. Moreover, contributions (e.g., from collaborators) can be
compared to, and incorporated in the main version of the project.
Version control thus reduces the risk of losing work and facilitates
collaboration. \texttt{Git} is built around snapshots that represent the
project state at a given point in time. These snapshots are called
\emph{commits} and work like a ``save'' action. Ideally, each commit has
a message that succinctly describes these changes. It is good practice
to make commits for concrete milestones (e.g., ``Commented on
Introduction,'' ``Added SES as a covariate,'' ``Address Reviewer 2's
comment 3''). This makes it easier to revert specific changes than when
multiple milestones are joined in one commit, e.g., ``Changes made on
19/07/2021''. Each commit refers back to its ancestor, and all commits
are thus linked in a timeline. The entirety of commits (i.e., the
version-controlled project) is called a repository. In \texttt{Git},
specific snapshots of a repository can be tagged, such that the user can
clearly label which version of the project was used to create a
preregistration, preprint, or final version of the manuscript as
accepted by a journal. \texttt{Git} has additional features beyond basic
version control, such as ``branches'' (parallel versions of a project
that can later be merged again) to facilitate simultaneous
collaboration. \citet{vuorreCuratingResearchAssets2018} provide a more
extensive treatment of how \texttt{Git} functions and how to use
\texttt{Git} for research. \citet{bryanExcuseMeYou2018} provides
additional information on how to track \texttt{R\ Markdown} documents.
Collaborating via \texttt{Git} is facilitated by uploading the
repository to a cloud-based service. We recommend GitHub as a host for
\texttt{Git} repositories because of its popularity among R users.
GitHub has many tools that facilitate working with \texttt{Git} --- in
particular project management and collaboration --- but these are not
central to achieving reproducibility.

Second, we rely on \emph{dynamic document generation}. The traditional
way of writing a scientific report based on a statistical data analysis
uses two separate steps conducted in two different programs. The
researcher writes text in a word processor, and conducts the analysis in
another program. Results are then (manually) copied and pasted from one
program to another, a process that often produces inconsistencies
\citep{nuijtenPrevalenceStatisticalReporting2016}.

Dynamic document generation integrates both steps. Through dynamic
document generation, code becomes an integral, although usually hidden,
part of the manuscript, complementing the verbal description and
allowing interested readers to gain a deeper understanding of the
contents
\citep{knuthCWEBSystemStructured, claerboutElectronicDocumentsGive1992}.
\texttt{R\ Markdown} uses \texttt{Markdown} for text formatting and R
(or other programming languages) for writing the statistical analysis.
\texttt{Markdown} is a lightweight text format in plain text with a
minimal set of reserved symbols for formatting instructions. This way,
\texttt{Markdown} does not need any specialized software for editing. It
is userfriendly (unlike, for example, \texttt{LaTeX}
\citep{lamportLATEXDocumentPreparation1994}), works well with version
control systems, and can be exported to various document formats, such
as HTML websites, a Microsoft Word document, a typeset PDF file (for
example, via LaTeX journal templates), or a Powerpoint presentation.
\texttt{Markdown} can be used for all sorts of academic documents,
ranging from simple sketches of ideas to scientific manuscripts
\citep{R-rticles} and presentations \citep{revealjs}, or even résumés
\citep{vitae}. \texttt{R\ Markdown} extends regular \texttt{Markdown} by
allowing users to include R code chunks (in fact, arbitrary computer
code \citep[Chapter 15, Chapter 15, Other
Languages]{xieMarkdownCookbook2020}) into a \texttt{Markdown} document.
Upon rendering the document, the code blocks are executed, and their
output is dynamically inserted into the document. This allows the
creation of (conditionally) formatted text, statistical results, and
figures that are guaranteed to be up-to-date because they are created
anew every time the document is rendered to its output format (e.g.,
presentation slides or a journal article).
\citet{xieMarkdownCookbook2020} provides an extensive yet practical
introduction to most features of \texttt{R\ Markdown}.

While version control and dynamic document generation are becoming more
common, we have argued that two more components are required and that
each component alone is unlikely to guarantee reproducibility
\citep{Peikert2019, vanlissa2020worcs}. In practice, dependencies
between project files (e.g., information on what script uses which data
file and what script needs to be run first) or on external software
(e.g., system libraries or components of the programming language, such
as other R packages) are frequently unmentioned or not exhaustively and
unambiguously documented.

\emph{Dependency tracking} helps automatically resolve dependencies
between project files. In essence, researchers provide a collection of
\emph{computational recipes}. A computational recipe describes how
inputs are processed to deterministically create a specific output in a
way that is automatically executable. The concept of computational
recipes is central to our understanding of reproducibility because it
enables a unified way to reproduce a project automatically. Similar to a
collection of cooking recipes, we can have multiple products
(\emph{targets}) with different ingredients (\emph{requirements}) and
different steps of preparation (\emph{recipes}). In the context of
scientific data analysis, targets are typically the final scientific
report (e.g., the one to be submitted to a journal) and possibly
intermediate results (such as preprocessed data files, simulation
results, and analysis results). A workflow that involves renaming
variable names by hand in a graphical spreadsheet application, for
example, is therefore incompatible with automated reproducibility.
Another property of a computational recipe is that the same inputs
should always result in the same outputs. For most computer code (given
the same software is used), this property is fulfilled. However, one
noteworthy exception is the generation of pseudo-random numbers.
Whenever random numbers are used in a computation, it is only
reproducible if the random number generator generates the same numbers.
To ensure identical random numbers, users may fix the state of the
random number generated with a so-called seed (e.g.~\texttt{set.seed()}
in R), but they also need to guarantee that the pseudo-random number
generator is unchanged \citep[see][]{Peikert2019}.

We recommend using \texttt{Make} for dependency tracking because it is
language independent. The following hypothetical example illustrates the
utility of \texttt{Make} and a suitable \texttt{Makefile}. Consider a
research project that contains a script to simulate data
(\texttt{simulate.R}) and a scientific report of the simulation results
written in \texttt{R\ Markdown} (\texttt{manuscript.Rmd}). A
\texttt{Makefile} for this project could look like this:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{manuscript.pdf:}\NormalTok{ manuscript.Rmd simulated\_data.csv}
  \ExtensionTok{Rscript} \AttributeTok{{-}e} \StringTok{\textquotesingle{}rmarkdown::render("manuscript.Rmd")\textquotesingle{}}

\ExtensionTok{simulated\_data.csv:}\NormalTok{ simulate.R}
  \ExtensionTok{Rscript} \AttributeTok{{-}e} \StringTok{\textquotesingle{}source("simulate.R")\textquotesingle{}}
\end{Highlighting}
\end{Shaded}

There are two targets, the final rendered report
(\texttt{manuscript.pdf}, l. 1) and the simulation results
(\texttt{simulation\_results.csv}, l. 4). Each target is followed by a
colon and a list of requirements. If a requirement is newer than the
target, the recipe will be executed to rebuild the target. If a
requirement does not exist, \texttt{Make} uses a recipe to build the
requirement before building the target. Here, if one were to build the
final \texttt{manuscript.pdf} by rendering the \texttt{R\ Markdown} with
the command shown in l.2, \texttt{Make} would check whether the file
\texttt{simulation\_results.csv} exists; if not, it would issue the
command in l.5 to run the simulation before rendering the manuscript.
This ensures that the simulated data are present before the manuscript
is built, and that the simulation is re-run and the manuscript is
rebuilt if the simulation code was changed. \texttt{Make} therefore
offers a standardized process to reproduce projects, regardless of the
complexity or configuration of the project. Note that the Workflow for
Open Reproducible Code in Science (WORCS) we presented elsewhere
\citep{vanlissa2020worcs} does not explicitly contain this dependency
tracking element, but its strict structure of only containing one
definite \texttt{R\ Markdown} still makes dependencies between files
unambiguous.

A version-controlled dynamic document with dependency tracking still
relies on external software. Troubleshooting issues specific to a
particular programming language or dependent tool typically require
considerable expertise and threaten reproducibility. \emph{Software
management} refers to the act of providing records of, or access to, all
software packages and system libraries a project depends on. One
comprehensive approach to software management is containerization. The
central idea is that by ``{[}..{]} packaging the key elements of the
computational environment needed to run the desired software {[}makes{]}
the software much easier to use, and the results easier to reproduce
{[}\ldots{]}'' \citep[p.~174]{silverSoftwareSimplified2017}.

\emph{\texttt{Docker}} is a popular tool for containerization. It
manages software dependencies by constructing a virtual software
environment independent of the host software environment. These
so-called ``\texttt{Docker} images'' function like a virtual computer
(i.e., a ``sand box'' a computational environment seperated from the
host). A \texttt{Docker} image contains \emph{all} software dependencies
used in an analysis---not just R packages, but also R and Rstudio, and
even the operating system. This is important because low level
functionality may impact the workings of higher-order software like R,
such as calls to random number generators or linear algebra libraries.
All of the differences in computational results that could be caused by
variation in the software used are hence eliminated.

Note that the software environment of the Docker image is completely
separate from the software installed on your computer. This separation
is excellent for reproducibility but takes some getting used to. For
example, it is important to realize that software available on your
local computer will \emph{not} be accessible within the confines of the
\texttt{Docker} image. Each dependency that you want to use within the
\texttt{Docker} image must be explicitly added as a dependency.
Furthermore, using \texttt{Docker} may require you to install software
on an operating system that may not be familiar to you. The images
supplied by the rocker project
\citep{boettigerIntroductionRockerDocker2017}, for example, are based on
Linux.

There are two ways to build a \texttt{Docker} image. First, users can
manually install whatever software they like from within the virtual
environment. Such a manually build environment can still be ported to
all computers that support \texttt{Docker}. However, we prefer the
second way of building images automatically from a textual description
called \texttt{Dockerfile}. Because the \texttt{Dockerfile} clearly
describes how which software is installed, the installation process can
be repeated automatically. Users can therefore quickly change the
software environment, for example, update to another R version or given
package version. Packaging all required software in such an image
requires considerable amounts of storage space. Two major strategies
help to keep the storage requirements reasonable. One is to rely on
pre-made images that are maintained by a community for particular
purposes. For example, there are pre-made images that only include what
is necessary for R, based on Ubuntu containers
\citep{boettigerIntroductionRockerDocker2017}. Users can then install
whatever they need in addition to what is provided by these pre-compiled
images. The image that was used for this article uses 1.35GiB of disk
space. The image for this project includes Ubuntu, R, RStudio, LaTeX as
well as a variety of R packages like tidyverse \citep{tidyverse} and all
its dependent packages, amounting to 192 R packages.

A second strategy is to save a so-called \texttt{Dockerfile}, which
contains only a textual description of all commands that need to be
executed to recreate the software environment. \texttt{Dockerfiles} are
tiny (the
\href{https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile}{\texttt{Dockerfile}}
for this project has a size of only 1.55KiB). However, they rely on the
assumption that all software repositories that provide the dependent
operating systems, pieces of software, and R packages will continue to
remain accessible and provide historic software versions. For proper
archiving, we therefore recommend storing a complete image of the
software environment, in addition to the \texttt{Dockerfile}. A more
comprehensive overview of the use of containerization in research
projects is given by
\citet{wiebelsLeveragingContainersReproducible2021}. Note that WORCS,
which we presented elsewhere \citep{vanlissa2020worcs} relies on the R
package \texttt{renv} \citep{R-renv} for software management.
\texttt{renv} is more lightweight and easier to use than \texttt{Docker}
on the one hand, but not as comprehensive on the other because it only
takes snapshots of the R packages instead of all software used.

To summarize, the workflow by \citet{Peikert2019} requires four
components (see Figure 2.1) dynamic document generation (using
\texttt{R\ Markdown}), version control (using \texttt{Git}), dependency
tracking (using \texttt{Make}), and software management (using
\texttt{Docker}). While \texttt{R\ Markdown} and \texttt{Git} are well
integrated into the R environment through RStudio, \texttt{Make} and
\texttt{Docker} require a level of expertise that is often beyond the
training of scholars outside the field of information technology. This
presents a considerable obstacle to the acceptance and implementation of
the workflow. To overcome this obstacle, we have developed the R package
\texttt{repro} that supports scholars in setting up, maintaining, and
reproducing research projects in R. Importantly, a reproducible research
project created with \texttt{repro} does not have the \texttt{repro}
package itself as a dependency. These projects will remain reproducible
irrespective of whether \texttt{repro} remains accessible in future.
Users do not need to have \texttt{repro} installed to reproduce a
project; in fact, they do not even need to have R installed because the
entire project can be rebuilt inside a container with R installed. In
the remainder, we will walk you through the creation of a reproducible
research project with the package \texttt{repro}.

\begin{figure}
\centering
\includegraphics{/home/rstudio/images/nutshell.pdf}
\caption{Figure 2.1: Schematic illustration of the interplay of the four
components (in dashed columns) central to the reproducible workflow:
version control (\texttt{Git}), dependency tracking (\texttt{Make}),
software management (\texttt{Docker}), and dynamic document generation
(\texttt{R\ Markdown}). \texttt{Git} tracks changes to the project over
time. \texttt{Make} manages dependencies among the files.
\texttt{Docker} provides a container in which the final report is built
using dynamic document generation in \texttt{R\ Markdown}. Adapted from
\citet{Peikert2019} licensed under
\href{https://creativecommons.org/licenses/by/4.0}{CC BY 4.0}.}
\end{figure}

\hypertarget{creating-reproducible-research-projects}{%
\section{Creating Reproducible Research
Projects}\label{creating-reproducible-research-projects}}

One impediment to the widespread adoption of a standard for reproducible
research is that learning to use the required tools can be quite
time-intensive. To lower the threshold, the R package \texttt{repro}
introduces helper functions that simplify the use of complicated and
powerful tools. The \texttt{repro} package follows the format of the
\href{https://usethis.r-lib.org}{\texttt{usethis}} \citep{usethis}
package, which provides helper functions to simplify the development of
R packages. The \texttt{repro} package provides similar helper
functions, but focuses on reproducibility-specific utilities. These
helper functions guide end-users in the use of reproducibility tools,
provide feedback about what the computer is doing and suggest what the
user should do next. We hope this makes reproducibility tools more
accessible by enabling beginner-level users to detect their system's
state accurately and act correspondingly \citep[Chapter 8: ``Automation
and Situation Awareness'']{parasuramanAutomationHumanPerformance2018}.
These wrappers are merely a support system; as users learn to use the
underlying tools, they can rely less on \texttt{repro} and use these
tools directly to solve more complex problems.

This tutorial assumes that the user will be working predominantly in R,
with the help of RStudio. It describes basic steps that we expect to be
relevant for small-scale psychological research projects that do not
rely on external software or multistage data processing (for those
requirements see section \protect\hyperlink{advanced-features}{Advanced
Features}). Of course, your specific situation might involve additional,
more specialized steps. After completing the tutorial, you should be
able to customize your workflow accordingly.

The first step is to install the required software. We assume that you
have installed R \citep[version 4.0.4]{R-base} and RStudio
\citep[version 1.4]{rstudio} already but the tutorial will guide you in
detail through the installation of other necessary software with the
help of the R package
\href{https://github.com/aaronpeikert/repro}{\texttt{repro}}
\citep{R-repro}. In case you have either not installed R and RStudio or
are unsure if they are up-to-date, you might want to consult our
installation advice in the
\href{https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md}{Online
Supplementary Material} that covers the installation of all software
necessary for this tutorial in three steps. The
\href{(https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)}{installation
advice} may also help Windows users who have problems installing
\texttt{Docker}.

Unfortunately, Docker requires administrator rights to run, which may
not be available to all researchers. We recommend
\href{https://rstudio.github.io/renv/articles/renv.html}{\texttt{renv}}
\citep{R-renv} in cases where no administrator rights can be obtained
but can not detail its use in this document. \texttt{renv} tracks which
R package is installed from which source in which version in a so-called
\href{https://rstudio.github.io/renv/articles/lockfile.html}{\texttt{lockfile}}.
This lockfile is then used to reinstall the same packages on other
computers or later in time. For a more thorough discussion, see
\citet{vanlissa2020worcs}.

Start RStudio and install the package
\href{https://github.com/aaronpeikert}{\texttt{repro}}\citep{R-repro}.
It will assist you while you follow the tutorial.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# repro is not on CRAN yet}
\FunctionTok{options}\NormalTok{(}
  \AttributeTok{repos =} \FunctionTok{c}\NormalTok{(}\AttributeTok{aaronpeikert =} \StringTok{\textquotesingle{}https://aaronpeikert.r{-}universe.dev\textquotesingle{}}\NormalTok{,}
            \AttributeTok{CRAN =} \StringTok{\textquotesingle{}https://cloud.r{-}project.org\textquotesingle{}}\NormalTok{)}
\NormalTok{)}
\FunctionTok{install.packages}\NormalTok{(}\StringTok{\textquotesingle{}repro\textquotesingle{}}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

To verify that you have indeed installed and set up the required
software for this workflow, you can use the ``check functions''. These
also illlustrate how \texttt{repro} assists the user in setting up a
reproducible workflow. In the example below, we use the double-colon
operator to explicitly indicate which functions originate in the
\texttt{repro} package. If the package is loaded (using
\texttt{library("repro")}), it is not necessary to use this double-colon
notation.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# \textasciigrave{}package::function()\textasciigrave{} → use function from package without \textasciigrave{}library(package)\textasciigrave{}}
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_git}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## v Git is installed, don't worry.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_make}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## v Make is installed, don't worry.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_docker}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## v Docker is installed, don't worry.
\end{verbatim}

These functions check whether specific dependencies are available on the
user's system, and if not, explain what further action is needed to
obtain it. Sometimes they ask the user to take action; for example, the
following happens if you are a Windows user who does not have
\texttt{Git} installed:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_git}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## x Git is not installed.
\end{verbatim}

\begin{verbatim}
## i We recommend Chocolately for Windows users.
\end{verbatim}

\begin{verbatim}
## x Chocolately is not installed.
\end{verbatim}

\begin{verbatim}
## * To install it, follow directions on: 
##   'https://chocolatey.org/docs/installation'
\end{verbatim}

\begin{verbatim}
## i Use an administrator terminal to install chocolately.
\end{verbatim}

\begin{verbatim}
## * Restart your computer.
\end{verbatim}

\begin{verbatim}
## * Run `choco install -y git` in an admin terminal to install Git.
\end{verbatim}

The messages from \texttt{repro} try to help the user solve problems.
They are adjusted to your specific operating system and installed
dependencies. Before you continue, we ask you to run the above commands
to check \texttt{Git}, \texttt{Make}, and \texttt{Docker}---both to
become familiar with the functionality of the \texttt{check\_*()}
functions and to make sure your system is prepared for the remainder of
this tutorial.

After you have installed the necessary software, we suggest that you set
up a secure connection to GitHub:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_github}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## v You and GitHub are on good terms, don't worry.
\end{verbatim}

If you know what Secure Shell (SSH) is and want to use it, you may
alternatively use:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# only an alternative: DO NOT USE if you are unsure what SSH means}
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_github}\NormalTok{(}\AttributeTok{auth\_method =} \StringTok{"ssh"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## v You and GitHub are on good terms, don't worry.
\end{verbatim}

If necessary, follow any instructions presented until all checks are
passed.

\hypertarget{creating-an-rstudio-project}{%
\subsection{Creating an RStudio
Project}\label{creating-an-rstudio-project}}

We start by creating a project folder with RStudio by clicking the menu
item:

\begin{quote}
File → New Project\ldots{} → New Directory → Example Repro Template
\end{quote}

This creates a project with a sample analysis. This sample analysis
consists of a single R Markdown document and a single data file. The
only special thing about the R Markdown document is the \texttt{repro}
metadata that we will learn later about. However, you may turn any other
template or existing R project into a reproducible research project by
adding those \texttt{repro} metadata there.

\hypertarget{implementing-version-control}{%
\subsection{Implementing Version
Control}\label{implementing-version-control}}

Now that your project is set up, we will introduce you to version
control with \texttt{Git}. \texttt{Git} does not automatically track all
files in your project folder; rather, you must manually add files to the
\texttt{Git} repository. To make sure you do not accidentally add files
that you do not wish to share (e.g., privacy-sensitive data), you can
list specific files that you do not want to track in the
\texttt{.gitignore} file. You can also block specific filetypes; for
example, to prevent accidentally sharing raw data. You can add something
to the \texttt{.gitignore} file directly or with this command:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{usethis}\SpecialCharTok{::}\FunctionTok{use\_git\_ignore}\NormalTok{(}\StringTok{"private.md"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Now the file \texttt{private.md} will not be added to the \texttt{Git}
repository, and hence also not be made public if you push the repository
to a remote service like GitHub. Please also consider carefully whether
you can include data in the repository without violating privacy rights.
If you are not allowed to share your data publicly, add the data file(s)
to the \texttt{.gitignore} file and only share them on request.

New users are advised to explicitly exclude any sensitive files before
proceeding. When you are ready, you can begin tracking your remaining
files using \texttt{Git} by running:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{usethis}\SpecialCharTok{::}\FunctionTok{use\_git}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\centering
\includegraphics{/home/rstudio/images/modified.pdf}
\caption{Figure 3.1: The Git pane in R Studio, showing
\texttt{manuscript.Rmd} modified but unstaged and \texttt{modified.png}
newly added and staged.}
\end{figure}

For \texttt{Git} to recognize changes to a given file, you have to stage
and then commit these changes (this is the basic save action for a
project snapshot). One way to do this is through the visual user
interface in the RStudio \texttt{Git} pane (see Figure 3.1). Click on
the empty box next to the file you want to stage. A checkmark then
indicates that the file is staged. After you have staged all of the
files you want, click on the commit button, explain in the commit
message why you made those changes, and then click on commit. This
stores a snapshot of the current state of the project.

The files you created and the changes you made have not yet left your
computer. All snapshots are stored in a local repository in your project
folder. To back up and/or share the files online, you can push your
local repository to a remote repository. While you can choose any
\texttt{Git} service (like GitLab or BitBucket), we will use GitHub in
this tutorial. Before you upload your project to GitHub, you need to
decide whether you would like the project to be publicly accessible
(viewable by anyone, editable by selected collaborators) or if you want
to keep it private (only viewable and editable by selected
collaborators). To upload the project publicly to GitHub use:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{usethis}\SpecialCharTok{::}\FunctionTok{use\_github}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

To upload it privately:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{usethis}\SpecialCharTok{::}\FunctionTok{use\_github}\NormalTok{(}\AttributeTok{private =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Depending on your computer's configuration, it may ask you to set up a
secure connection to GitHub. In this case, first, follow the suggestions
shown on the R console.

\hypertarget{using-dynamic-document-generation}{%
\subsection{Using Dynamic Document
Generation}\label{using-dynamic-document-generation}}

Now that you have created a version-controlled project, we will proceed
with dynamic document generation. A dynamic document has three elements:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Text (prose; e.g., a scientific paper or presentation)
\item
  Executable code (e.g., analyses)
\item
  Metadata (e.g.~title, authors, document format)
\end{enumerate}

\texttt{R\ Markdown} is a type of dynamic document well-suited to the
RStudio user interface. The text of an \texttt{R\ Markdown} is formatted
by \texttt{Markdown} (see \citep{xieMarkdownDefinitiveGuide2019} for
technical details and \citep{xieMarkdownCookbook2020} for practical
guidance). The code mostly consists of R code (although other
programming languages are supported, like Python, C++, Fortran, etc).
The following example serves to illustrate the \texttt{Markdown} syntax.
It shows how to create a heading, a word in bold font, a citation, and a
list of several items in \texttt{Markdown}:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\textless{}!{-}{-}this is a Markdown file {-}{-}\textgreater{}}
\FunctionTok{\# Heading (level 1)}

\NormalTok{Normal text.}
\NormalTok{Important **word** in bold.}
\NormalTok{A citation: @einstein1935 did important research on this topic.}

\FunctionTok{\#\# Subheading (level 2)}

\NormalTok{To do list:}

\SpecialStringTok{* }\NormalTok{Do research}
\SpecialStringTok{* }\NormalTok{Do more research}
\SpecialStringTok{* }\NormalTok{Spend time with family and friends}
\end{Highlighting}
\end{Shaded}

One advantage of this type of markup for formatting is that it can be
rendered to many different output formats---both in terms of file types,
like \texttt{.docx}, \texttt{.html}, \texttt{.pdf}, and in terms of
style, e.g.~specific journal requirements. For social scientists, the
\href{https://github.com/crsh/papaja}{\texttt{papaja}} package
\citep{papaja} may be relevant, as it produces manuscripts that follow
the American Psychological Association formatting requirements
\citep{apa7}. \texttt{R\ Markdown} files are plain text, which is more
suitable for version control using \texttt{Git} than binary files
generated by some word processors. Some users might find it easier to
activate the ``Visual Editor'' of RStudio ({[}Ctrl{]} + {[}Shift{]} +
{[}F4{]} or click on the icon that resembles drawing materials or a
compass in the upper right corner of the \texttt{R\ Markdown} document),
which features more graphical elements like a traditional word processor
but still creates an \texttt{R\ Markdown} underneath with all of its
flexibility. The visual editor has some additional benefits, such as
promoting best practices (for example, each sentence should be written
on a new line, which makes it easier to track changes across versions)
and improving the generation of citations and references to tables and
figures.

Now that you are familiar with \texttt{Markdown} formatting basics, we
turn our attention to including code and its results in the text. Code
is separated by three backticks (and the programming language in curly
brackets) like this:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{This is normal text, written in Markdown.}

\InformationTok{\textasciigrave{}\textasciigrave{}\textasciigrave{}\{r\}}
\InformationTok{\# this is R code}
\InformationTok{1 + 1}
\InformationTok{\textasciigrave{}\textasciigrave{}\textasciigrave{}}
\end{Highlighting}
\end{Shaded}

The hotkey {[}Control{]}+{[}Alt{]}+{[}i{]} inserts a block of code in
the file. The results of code enclosed in such backticks will be
dynamically inserted into the document (depending on specific settings).
This means that whenever you render the \texttt{R\ Markdown} to its
intended output format, the code will be executed and the results
updated. The resulting output document will be static, e.g., a pdf
document, and can be shared wherever you like, e.g., on a preprint
server.

Once the \texttt{R\ Markdown} file has been rendered to a static
document (the output, e.g., PDF), the resulting file is decoupled from
the \texttt{R\ Markdown} and the code that created it. This introduces a
risk that multiple versions of the static document are disseminated,
each with slightly different results. To avoid ambiguity, we, therefore,
recommend referencing the identifier of the \texttt{Git} commit at the
time of rendering in the static document. Simply put, a static document
should link to the version of the code that was used to create it. The
\texttt{repro} package comes with the function
\texttt{repro::current\_hash()} for this purpose. This document was
created from the commit with the hash
\href{https://github.com/aaronpeikert/repro-tutorial/tree/a8bb0d4df4afa4f7dbb5cd2344ae16043cb8a99d}{\texttt{a8bb0d4}
(view on GitHub)}.

Now that you know how to write text and R code in an
\texttt{R\ Markdown}, you need to know about metadata (also called: YAML
front matter). These metadata contain information about the document,
like the title and the output format. Metadata are placed at the
beginning of the document and are separated from the document body by
three dashes. The following example is a full markdown document where
the metadata (the ``YAML front matter'') are in lines 1--6. Some
metadata fields are self-explanatory (like the author field), and exist
across all output formats (like the title field). Others are specific to
certain output formats or R packages.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{{-}{-}{-}}
\AnnotationTok{title:}\CommentTok{ "A Tutorial on how to Do the Same Thing More Than Once"}
\AnnotationTok{author:}\CommentTok{ Aaron Peikert, Caspar J. van Lissa, and Andreas M. Brandmaier}
\AnnotationTok{abstract:}\CommentTok{ A hitchhiker\textquotesingle{}s guide to reproducible research in R}
\AnnotationTok{output:}\CommentTok{ html\_document}
\CommentTok{{-}{-}{-}}

\FunctionTok{\# Introduction}

\NormalTok{Important for reproducibility:}

\SpecialStringTok{1. }\NormalTok{*Version control*}
\SpecialStringTok{2. }\NormalTok{*Dynamic document creation*}
\SpecialStringTok{3. }\NormalTok{*Dependency tracking*}
\SpecialStringTok{4. }\NormalTok{*Software management*}

\InformationTok{\textasciigrave{}\textasciigrave{}\textasciigrave{}\{r\}}
\InformationTok{\# this is R code}
\InformationTok{t.test(extra \textasciitilde{} group, data = sleep)}
\InformationTok{\textasciigrave{}\textasciigrave{}\textasciigrave{}}
\end{Highlighting}
\end{Shaded}

\hypertarget{manage-software-and-file-dependencies}{%
\subsection{Manage Software and File
Dependencies}\label{manage-software-and-file-dependencies}}

The \texttt{repro} package adds fields to the metadata to list all
dependencies of the research project. This includes R scripts, data
files, and external packages. The format is as follows (see everything
below the line \emph{repro:}):

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{{-}{-}{-}}
\AnnotationTok{title:}\CommentTok{ "A tutorial on how to do the same thing more than once"}
\AnnotationTok{author:}\CommentTok{ Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier}
\AnnotationTok{output:}\CommentTok{ html\_document}
\AnnotationTok{repro:}
\CommentTok{  scripts:}
\CommentTok{    {-} R/load.R}
\CommentTok{  data:}
\CommentTok{    {-} data/mtcars.csv}
\CommentTok{  packages:}
\CommentTok{    {-} tidyverse}
\CommentTok{    {-} usethis}
\CommentTok{    {-} gert}
\CommentTok{{-}{-}{-}}
\end{Highlighting}
\end{Shaded}

This information clarifies what dependencies (in the form of files and R
packages) a project relies on. \texttt{repro} uses this information to
construct a \texttt{Makefile} for the dependencies on other files and a
\texttt{Dockerfile} that includes all required packages. Together, these
two files form the basis for consistency within a research project and
consistency across different systems. The function
\texttt{repro::automate()} converts the metadata from all
\texttt{R\ Markdown} files in the project (all files with the ending
\texttt{.Rmd}) to a \texttt{Makefile} and a \texttt{Dockerfile}. These
files allow users (including your future self) to reproduce every step
in the analysis automatically. Please run \texttt{repro::automate()} in
your project:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{automate}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

It is important to re-run \texttt{repro::automate()} whenever you change
the \texttt{repro} metadata, change the output format, or add a new
\texttt{R\ Markdown} file to the project to keep the \texttt{Makefile}
and \texttt{Dockerfile} up to date. There is no harm in running it too
often. Other than the \texttt{Makefile} and the \texttt{Dockerfile},
which are created in the document root path, \texttt{repro} generates a
few more files in the \texttt{.repro} directory (which we will explain
in detail later), all of which you should add and commit to
\texttt{Git}.

\hypertarget{reproducing-a-project}{%
\subsection{Reproducing a project}\label{reproducing-a-project}}

If someone (including you) wants to reproduce your project, they first
have to install the required software, that is \texttt{Make}, and
\texttt{Docker}. Remember, you can use the \texttt{check\_*}-functions
to test if these are installed:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_make}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## v Make is installed, don't worry.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_docker}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## v Docker is installed, don't worry.
\end{verbatim}

When these are set up, they can ask \texttt{repro} to explain how they
should use \texttt{Make} and \texttt{Docker} to reproduce the project
(or you could explain it to them):

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{reproduce}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## * To reproduce this project, run the following code in a terminal:
\end{verbatim}

\begin{verbatim}
##   make docker &&
##   make -B DOCKER=TRUE
\end{verbatim}

If you feel uncomfortable using the terminal directly, you can send the
command to the terminal from within R:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{system}\NormalTok{(repro}\SpecialCharTok{::}\FunctionTok{reproduce}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

The only ``hard'' software requirement for reproducing a project is
\texttt{Docker}, assuming users know how to build a \texttt{Docker}
image and run \texttt{Make} within the container. However, if they have
installed \texttt{Make} in addition to \texttt{Docker}, they do not even
need to know how to use \texttt{Docker} and can simply rely on the two
\texttt{Make} commands ``\texttt{make\ docker}'' and
``\texttt{make\ -B\ DOCKER=TRUE}''.

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Install the \texttt{repro} package:
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{options}\NormalTok{(}
  \AttributeTok{repos =} \FunctionTok{c}\NormalTok{(}\AttributeTok{aaronpeikert =} \StringTok{\textquotesingle{}https://aaronpeikert.r{-}universe.dev\textquotesingle{}}\NormalTok{,}
            \AttributeTok{CRAN =} \StringTok{\textquotesingle{}https://cloud.r{-}project.org\textquotesingle{}}\NormalTok{)}
\NormalTok{)}
\FunctionTok{install.packages}\NormalTok{(}\StringTok{\textquotesingle{}repro\textquotesingle{}}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  Check the required software:
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_git}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## v Git is installed, don't worry.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_github}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## v You and GitHub are on good terms, don't worry.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_make}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## v Make is installed, don't worry.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{check\_docker}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## v You are inside a Docker container!
\end{verbatim}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Create an R project or use an existing one. Do not forget to add
  \texttt{repro} metadata (i.e., packages, scripts, data).
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{repro}\KeywordTok{:}
\AttributeTok{  }\FunctionTok{scripts}\KeywordTok{:}
\AttributeTok{    }\KeywordTok{{-}}\AttributeTok{ R/load.R}
\AttributeTok{  }\FunctionTok{data}\KeywordTok{:}
\AttributeTok{    }\KeywordTok{{-}}\AttributeTok{ data/mtcars.csv}
\AttributeTok{  }\FunctionTok{packages}\KeywordTok{:}
\AttributeTok{    }\KeywordTok{{-}}\AttributeTok{ tidyverse}
\end{Highlighting}
\end{Shaded}

The sample \texttt{repro} project already has theese metadata:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{use\_repro\_template}\NormalTok{(}\StringTok{"/some/folder"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Let \texttt{repro} generate \texttt{Docker}- and \texttt{Makefile}:
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{automate}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Enjoy automated reproducibility:
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{repro}\SpecialCharTok{::}\FunctionTok{reproduce}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## * To reproduce this project, run the following code in a terminal:
\end{verbatim}

\begin{verbatim}
##   make docker &&
##   make -B DOCKER=TRUE
\end{verbatim}

\hypertarget{advanced-features}{%
\section{Advanced Features}\label{advanced-features}}

This section is for advanced users who want to overcome some limitations
of \texttt{repro}. If you read this paper the first time, you will
probably want to skip this section and continue reading from the section
``Preregistration as Code.'' As explained above, \texttt{repro} is
merely a simplified interface to the tools that enable reproducibility.
This simplified interface imposes two restrictions. Users who ask
themselves either, ``How can I install software dependencies outside of
R in the \texttt{Docker} image?'' or ``How can I express complex
dependencies between files (e.g., hundreds of data files are
preprocessed and combined)?'' need to be aware of these restrictions and
require a deeper understanding of the inner workings of \texttt{repro}.
Other users may safely skip this section or return to it if they
encounter such challenges.

The first restriction is that users must rely on software that is either
already provided by the base \texttt{Dockerimage} ``rocker/verse'' or
the R packages they list in the metadata. The metadata the
\texttt{repro::automate()} function relies on can only express R
packages as dependencies for the \texttt{Dockerfile} and only trivial
dependencies (in the form of ``file must exist'') for the
\texttt{Makefile}. Other software that users might need, like other
programming languages, not yet installed LaTeX packages, etc., must be
added manually. We plan to add support for commonly used ways to install
software beyond R packages via the metadata and
\texttt{repro::automate()}, for example, for system libraries (via
\texttt{apt} the Ubuntu package manager), LaTeX packages (via
\texttt{tlmgr} the Tex Live package manager), Python packages (via
\texttt{pip} the python package manager). The second limitation is
related to dependencies. \texttt{Make} can represent complex
dependencies, for example: A depends on B, which in turn depends on C
and D. If B is missing in this example, \texttt{Make} would know how to
recreate it from C and D. These dependencies, and how they should be
resolved, are difficult to represent in the metadata. Users, therefore,
have to either ``flatten'' the dependency structure by simply stating
that A depends on B, C, and D, thereby leaving out important information
or express the dependencies directly within the \texttt{Makefile}.

The following section explains how to overcome these limitations despite
reliance on the automation afforded by \texttt{repro}. Lifting these
restrictions requires the user to interact more directly with
\texttt{Make} or \texttt{Docker}. Users need to understand how repro
utilizes \texttt{Make} and \texttt{Docker} internally to satisfy more
complicated requirements.

Let us have a closer look at the command for reproducing a repro
project: \texttt{make\ docker\ \&\&\ make\ -B\ DOCKER=TRUE}; which
consists of two processing steps. First, it recreates the virtual
software environment (\texttt{Docker}), and then it executes
computational recipes in the virtual software environment
(\texttt{Make}). The first step is done by the command
\texttt{make\ docker}. The command \texttt{make\ docker} will trigger
\texttt{Make} to build the target called \texttt{docker}. The recipe for
this target builds an image from the \texttt{Dockerfile} in the
repository. The \texttt{\&\&} concatenates both commands and only runs
the second command if the first is successful. Therefore, the
computational steps are only executed when the software environment is
set up. The second step executes the actual reproduction and is again a
call to \texttt{Make} in the form of \texttt{make\ -B\ DOCKER=TRUE} with
three noteworthy parts. First, a call to \texttt{make} without any
explicit target will build the \texttt{Make} target \texttt{all}.
Second, the flag \texttt{-B} means that \texttt{Make} will consider all
dependencies as outdated and will hence rebuild everything. Third, repro
constructs \texttt{Make} targets so that if you supply
\texttt{DOCKER=TRUE} they are executed within the \texttt{Docker} image
of the project.

The interplay between \texttt{Docker} and \texttt{Make} resembles a
chicken or egg problem. We have computational steps (\texttt{Make}) that
depend on the software environment (\texttt{Docker}) for which we again
have computational steps that create it. Users only require a deeper
understanding of this interdependence when they either want to have more
complex computational recipes than rendering an \texttt{R\ Markdown} or
require other software than R packages.

Users can have full control over the software installed within the image
of the project. \texttt{repro} creates three \texttt{Dockerfiles} inside
the \texttt{.repro} directory. Two \texttt{Dockerfiles} are
automatically generated. The first is \texttt{.repro/Dockerfile\_base}.
It contains information about the base image on which all the remaining
software is installed. By default we rely on the ``verse'' images
provided by the Rocker project
\citep{boettigerIntroductionRockerDocker2017}. These contain (among
other software) the packages \texttt{tidyverse}, \texttt{rmarkdown}, and
a complete LaTeX installation, which makes these images ideal for the
creation of scientific manuscripts. Users can choose which R version
they want to have inside the container by changing the version number in
line 1 to the desired R version number. By default, the R version
corresponds to the locally installed version on which
\texttt{repro::automate()} was called the first time. The build date is
used to install packages in the version that was available on the
Comprehensive R Archive Network on this specific date and can also be
changed. By default, this date is set to the date on which
\texttt{repro::automate()} was called the first time. This way, the call
to the automate function virtually freezes the R environment to the
state it was called the first time inside the container. Below, you see
the \texttt{Docker} base file we used to create this manuscript:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{FROM}\NormalTok{ rocker/verse:4.0.4}
\ExtensionTok{ARG}\NormalTok{ BUILD\_DATE=2021{-}05{-}06}
\ExtensionTok{WORKDIR}\NormalTok{ /home/rstudio}
\end{Highlighting}
\end{Shaded}

The second automatically generated \texttt{Dockerfile} is
\texttt{.repro/Dockerfile\_packages}. Whenever
\texttt{repro::automate()} is called, \texttt{repro} gathers all R
packages from all \texttt{.Rmd} files and determines whether they should
be installed from CRAN or GitHub fixed to the date specified in
\texttt{Dockerfile\_base}. Finally, there is one manually edited
\texttt{Dockerfile}: \texttt{.repro/Dockerfile\_manual}. It is blank by
default and can be used to add further dependencies outside of R, like
system libraries or external software. Using \texttt{Docker} may require
you to install software on an operating system that may not be familiar
to you. The images supplied by
\citep{boettigerIntroductionRockerDocker2017}, for example, are based on
the Ubuntu operating system. The most convenient way to install software
on Ubuntu is through its package manager \texttt{apt}. If the following
snippet is added to \texttt{.repro/Dockerfile\_manual}, the
\texttt{Docker} image will have, for example, Python installed. Other
software is installed identically, only the software name is exchanged.

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{RUN}\NormalTok{ apt{-}get update }\KeywordTok{\&\&} \ExtensionTok{apt{-}get}\NormalTok{ install }\AttributeTok{{-}y}\NormalTok{ python3}
\end{Highlighting}
\end{Shaded}

\texttt{Docker} eventually requires a single \texttt{Dockerfile} to run,
so \texttt{repro::automate()} simply concatenates the three
\texttt{Dockerfiles} and saves the result into the main
\texttt{Dockerfile} at the top level of the R project. With this
approach, users of \texttt{repro} can build complex software
environments and implement complex file dependencies. The standard
\texttt{repro} metadata only make sure that all dependencies are
available but does not allow you to specify custom recipes for them in
the metadata. If you can formulate the creation of dependencies in terms
of computational steps, e.g.~the file \texttt{data/clean.csv} is created
from \texttt{data/raw.csv} by script \texttt{R/preprocess.R}, you should
include these in the \texttt{Makefile}. The \texttt{Makefile} that
\texttt{repro} creates is only a template, and you are free to change
it. However, make sure you never remove the following two lines:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{include}\NormalTok{ .repro/Makefile\_Rmds}
\ExtensionTok{include}\NormalTok{ .repro/Makefile\_Docker}
\end{Highlighting}
\end{Shaded}

The file \texttt{.repro/Makefile\_Rmds} contains the automatically
generated targets from \texttt{repro::automate()} for the
\texttt{R\ Markdown} files. This file should not be altered manually. If
you are not satisfied with the automatically generated target, simply
provide an alternative target in the main \texttt{Makefile}. Targets in
the main \texttt{Makefile} take precedent.

The file \texttt{.repro/Makefile\_Docker} does again contain a rather
complicated template that you could, but should usually not modify. This
\texttt{Makefile} coordinates the interplay between \texttt{Make} and
\texttt{Docker} and contains targets for building (with
\texttt{make\ docker}) and saving (with \texttt{make\ save-docker}) the
\texttt{Docker} image. Additionally, it provides facilities to execute
commands within the container. If you write a computational recipe for a
target, it will be evaluated using the locally installed software by
default. To evaluate commands inside the \texttt{Docker} image instead,
you should wrap them in \texttt{\$(RUN1)\ command\ \$(RUN2)}, as done in
this example, which is identical to the first \texttt{Make} example we
gave above except for the addition of \texttt{\$(RUN1)} and
\texttt{\$(RUN2)} in l. 2:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{simulated\_data.csv:}\NormalTok{ R/simulate.R}
  \VariableTok{$(}\ExtensionTok{RUN1}\VariableTok{)}\NormalTok{ Rscript }\AttributeTok{{-}e} \StringTok{\textquotesingle{}source("R/simulate.R")\textquotesingle{}} \VariableTok{$(}\ExtensionTok{RUN2}\VariableTok{)}
\end{Highlighting}
\end{Shaded}

If users execute this in the terminal:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{make}\NormalTok{ data/simulation\_results.csv}
\end{Highlighting}
\end{Shaded}

It behaves exactly as in the first \texttt{Make} example, the script
\texttt{R/simulate.R} is run using the locally installed R. Because this
translates simply to:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{Rscript} \AttributeTok{{-}e} \StringTok{\textquotesingle{}source("R/simulation.R")\textquotesingle{}}
\end{Highlighting}
\end{Shaded}

But if users use

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{make}\NormalTok{ DOCKER=TRUE data/simulation\_results.csv}
\end{Highlighting}
\end{Shaded}

it is evaluated within the \texttt{Docker} container using the software
within it and not the locally installed R version:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{docker}\NormalTok{ run }\AttributeTok{{-}{-}rm} \AttributeTok{{-}{-}user}\NormalTok{ 1000 }\AttributeTok{{-}v} \StringTok{"/home/rstudio"}\NormalTok{:}\StringTok{"/home/rstudio/"}
\ExtensionTok{reprotutorial}\NormalTok{ Rscript }\AttributeTok{{-}e} \StringTok{\textquotesingle{}source("R/simulate.R")\textquotesingle{}}
\end{Highlighting}
\end{Shaded}

To summarize, \texttt{repro} automates dependency tracking (in the form
of \texttt{Make}) and software management (using \texttt{Docker})
without the necessity to learn both tools, but users with advanced
requirements can still customize all aspects of both programs.

\hypertarget{preregistration-as-code}{%
\section{Preregistration as Code}\label{preregistration-as-code}}

\emph{Preregistration} refers to the practice of defining research
questions and planning data analysis before observing the research
outcomes \citep{NosekRevolution2018}. It serves to separate a-priori
planned and theory-driven (confirmatory) analyses from unplanned and
post-hoc (exploratory) analyses. Researchers are faced with a myriad of
choices in designing, executing, and analyzing a study, often called
researchers degrees of freedom. Undisclosed researcher degrees may be
used to modify planned analyses until a key finding reaches statistical
significance or to inflate effect size estimates, a phenomenon referred
to as \emph{opportunistic bias}
\citep{decosterOpportunisticBiasesTheir2015}. Preregistration increases
transparency by clarifying when and how researchers employ their degrees
of freedom. It expressly does not restrict what researchers may do to
gather or analyze their data.

There are still several shortcomings to preregistration. One is that
written study plans are often interpretable in multiple ways. Empirical
research has shown that, even when several researchers describe their
analysis with the same terms, use the same data, and investigate the
same hypothesis, their results vary considerably
\citep{silberzahnManyAnalystsOne2018}. The current best practice to
ensure comprehensive and specific preregistration is to impose structure
by following preregistration templates
\citep{bowmanOSFPreregTemplate2020, bakkerEnsuringQualitySpecificity2020}.
However, such templates cannot ensure full transparency because it is
impossible to verbally describe every detail of an analysis for any but
the most straightforward analysis. This ambiguity causes a second
problem, namely, comparing the initial plan and the resulting
publication to decide if and how researchers deviated from the
preregistration. This task is difficult because it is impossible to
decide without additional information weather the analysis was actually
carried out differently or just described differently. Even if
researchers were faithful to the preregistration, readers may reach
opposite conclusions because they have to compare two different text
that may be worded differently or describe the same thing in varying
levels of detail. A third limitation is that preregistrations are
susceptible to non-reproducibility, just like primary research. To
illustrate, a review of 210 preregistrations found that, even though 174
(67\%) included a formal power analysis, only 34 (20\%) of these could
be reproduced \citep{bakkerRecommendationsPreregistrationsInternal2020}.
Even when researchers have gone to great lengths in preregistering an
analysis script, they sometimes inexplicably fail to reproduce their own
results. For example, \citet{steegenMeasuringCrowdAgain2014} realized
after publication that part of their preregistered code resulted in
different test statistics than they reported initially (see their
Footnote 7). A final limitation is that written plans may turn out to be
unfeasible once data are obtained and analyzed. For example, a verbal
description of a statistical model may be unidentified, e.g., if it
includes reciprocal paths between variables or more parameters than
observed data. Conversely, a model may be misspecified in a major way;
for example, by omitting direct effects when the research question is
about mediation, thus leading to a model with an unacceptable fit. Many
researchers would only realize that such a model cannot be estimated
once the data are obtained, thus necessitating a deviation from the
preregistered plans.

The workflow described in this paper facilitates a rigorous solution to
this problem: Instead of describing the analysis in prose, researchers
include the code required to conduct the analysis in the
preregistration. We term this approach of writing and publishing code at
the preregistration stage \emph{Preregistration as Code (PAC)}. PAC has
the potential to eliminate undisclosed researchers degrees of freedom to
a much greater extent than, e.g., preregistration templates. Moreover,
it reduces overhead by removing the need to write a separate
preregistration and manuscript. For PAC, researchers can write a
reproducible, dynamically generated draft of their intended manuscript
at the preregistration stage. This already includes most of the typical
sections, such as introduction, methods, and results. These results are
initially based on simulated data with the same structure as the data
the authors expect to obtain from their experiments. For guidance on how
to simulate data, see \citet{morrisUsingSimulationStudies2019},
\citet{paxtonMonteCarloExperiments2001}, and
\citet{skrondalDesignAnalysisMonte2000}, as well as the R packages
\href{https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html}{\texttt{simstudy}},
\citep{simstudy} and
\href{https://personality-project.org/r/psych/help/sim.html}{\texttt{psych}},
\citep{psych}.

Once the preregistration is submitted and real data have been collected
or made available, the document can be reproduced with a single command,
thus updating the Results section to the final version. Reproducibility
is of utmost importance at this stage since the preregistration must
produce valid results at two points in time, once before data collection
and once after data collection. As outlined before, reproducibility
builds upon four pillars (version control, dynamic document generation,
dependency tracking, and software management). To use PAC the dangers to
reproducibility we described must be eliminated.

The idea of submitting code as part of a preregistration is not new
\citep[e.g.][]{wichertsDegreesFreedomPlanning2016}. A prominent
preregistration platform, \href{https://osf.io/}{The Open Science
Framework}, suggests submitting scripts alongside the preregistration of
methods. In an informal literature search (we skimmed the first 300
results of Google Scholar with the keywords
\texttt{("pre\ registration"\textbar{}"pre-registration"\textbar{}preregistration)\&(code\textbar{}script\textbar{}matlab\textbar{}python\textbar{}"R")})
we only found close to a dozen published articles that did include some
form of script as part of their preregistration. Though the notion of
preregistering code has been around for a while
\citep[cf.][]{steegenMeasuringCrowdAgain2014}, it has not gained much
traction---perhaps because, to date, this has constituted an extra
non-standard step in the research process. This tutorial integrates the
preregistration of code into the reproducible research workflow by
encouraging researchers to preregister the whole manuscript as a dynamic
document.

\hypertarget{advantages-of-pac-over-traditional-preregistration}{%
\subsection{Advantages of PAC Over Traditional
Preregistration}\label{advantages-of-pac-over-traditional-preregistration}}

We believe that pairing PAC with the workflow presented above offers
five advantages over classical preregistration. First, PAC is merely an
intermediate stage of the final manuscript, thus sparing authors from
writing, and editors and reviewers from evaluating, two separate
documents. Relatedly, writing the preregistration in the form of a
research article has the advantage that researchers are usually familiar
with this format. By contrast, a preregistration template is a novelty
for many. Second, PAC is a tool for study planning. A study can be
carried out more efficiently if all steps are documented clearly than
when every step is planned ad hoc. Third, PAC removes ambiguity
regarding the translation of verbal analysis plans into code. PAC is
more comprehensive by design because its completeness can be empirically
checked with simulated data. Evaluating the intended analysis code on
simulated data will help identify missing steps or ambiguous decisions.
PAC, therefore, minimizes undisclosed researchers degrees of freedom
more effectively than standard preregistration does
\citep{bakkerEnsuringQualitySpecificity2020, wichertsDegreesFreedomPlanning2016}.
Fourth, despite its rigor, PAC accommodates data-dependent decisions if
these can be formulated as code. Researchers can, for example, formulate
conditions (e.g., in the form of if-else-blocks) under which they prefer
one analysis type over the other. For example, if distributional
assumptions are not met, the code may branch out to employ robust
methods; or, an analysis may perform automated variable selection
mechanisms before running the final model. Another example of
data-dependent decisions are more explorative analyses, i.e.~explorative
factor analysis or machine learning. Decisions that do not lend
themselves to formulation in code, e.g.~visual inspection, must still be
described verbally or be treated as noted in the next section. Fifth,
deviations from the preregistration are clearly documented because they
are reflected in changes to the code, which are managed and tracked with
version control.

\hypertarget{deviating-from-the-preregistration-and-exploration}{%
\subsection{Deviating from the Preregistration and
Exploration}\label{deviating-from-the-preregistration-and-exploration}}

We would like to note that PAC allows explicit comparison of the
preregistration and the final publication. Authors should
retrospectively summarize and justify any changes made to the
preregistered plan, e.g.~in the discussion of the final manuscript
\footnote{In section {[}Preregistration as Code --- a Tutorial{]} we
  conducted an actual PAC and summarize the changes we make to the
  preregistered code in the discussion.}. During the analysis process,
authors can additionally maintain a running changelog to explain changes
in detail as they arise. Each entry in the changelog should explain the
reasoning behind the changes and link to the commit id that applied the
changes. This enables readers and reviewers to inspects individual
changes and make an informed judgment about their validity and
implications.

Deviations from the preregistration are sometimes maligned, as if
encountering unexpected challenges invalidates a carefully crafted study
\citep{szollosi_is_2020}. However, we share the common view that
deviation from a preregistration is not a problem
\citep{nosekPreregistrationHardWorthwhile2019}; rather, a faillure to
disclose such deviations is a problem. In fact, it is expected that most
PACs will require some modification after empirical data becomes
available. Often, deviations provide an opportunity to learn from the
unexpected.

For example, imagine that authors preregistered their intention to
include both ``working memory'' and ``fluid intelligence'' as covariates
in an experimental study, examining the effect of task novelty on
reaction time. When evaluating the planned analyses on the empirical
data, these two covariates reveal high collinearity, thus compromising
statistical inference. The authors decide to use PCA to extract common
variance related to ``intelligence'', and include this component as a
covariate instead. This change pertains to an auxiliary assumption (that
working memory and fluid intelligence are distinct constructs), but does
not undermine the core theory (that task novelty affects reaction time).
Now imagine that a different researcher is interested in the structure
of intelligence. This change to the preregistration directly relates to
their theory of intelligence. That researcher might thus interpret the
same result as an explorative finding, suggesting that these aspects of
intelligence are unidimensional. A deviation from preregistration thus
requires a judgement about what changes affect the test of the theory to
what extend \citep{meehlTheoreticalRisksTabular1978}. Only transparent
reporting enables such judgment.

Another common misunderstanding is that preregistration, including PAC,
precludes exploratory analyses. We differentiate between two kinds of
exploration, neither of which is limited by PAC. The first, more
traditional kind of exploration involves ad hoc statistical decisions
and post hoc explanations of the results. Such traditional exploratory
findings should be explicitly declared in the manuscript to distinguish
them from confirmatory findings
\citep{NosekRevolution2018, nosekPreregistrationHardWorthwhile2019}. The
second kind of exploration is through procedurally well defined
exploration with exploratory statistical models that are standard in
machine learning \citep{brandmaier22mlsem}. These models often involve
dozens, if not hundreds of predictors, which makes it difficult to
describe them verbally. With PAC, such models can be preregistered
clearly and in comprehensive detail and researcher can precisely define
a priori how much they want to explore. We specifically recommend PAC
for such exploratory statistical models. The merit of preregistration in
these cases is to communicate precisely how much exploration was done; a
piece of information that is crucial to assess e.g., whether the results
might be overfit
\citep[p.~220f.]{hastieElementsStatisticalLearning2017}.

\hypertarget{planned-analyses-as-functions}{%
\subsection{Planned Analyses as
Functions}\label{planned-analyses-as-functions}}

Although researchers may use any form to preregister their planned
analyses (e.g., scripts), we suggest writing three functions for each
planned hypothesis: one to conduct the planned analysis, one to simulate
the expected data, and to report the results. Using functions makes the
analysis more portable (i.e., it can easily be used for other datasets),
and facilitates repeated evaluation, as is the case in a simulation
study. The functions shown here do not contain executable code, but the
interested reader can find working functions in the
\href{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R}{online
supplementary materials} that power the example below.

It is difficult to write analysis code when it is not clear what the
expected data will look like. We therefore recommend first simulating a
dataset that resembles the expected structure of the empirical data that
will be used for the final analysis. Dedicated packages to simulate data
for specific analyses exist .

The general format of a simulation function might be as follows:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{simulate\_data }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(n, effect\_size)\{}
  \CommentTok{\# 1. warn users that the results are "fake"}
  \CommentTok{\# 2. draw \textasciigrave{}n\textasciigrave{} samples with \textasciigrave{}effect\_size\textasciigrave{}}
  \CommentTok{\# 3. format and return in expected data format}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

For linear models, simulating data is extremely simple:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{simulate\_data }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(n, effect\_size)\{}
  \FunctionTok{warning}\NormalTok{(}\StringTok{"This manuscript contains mock results based on simulated data."}\NormalTok{)}
  \CommentTok{\# Draw n samples from a normal distribution for predictor X}
\NormalTok{  x }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(n)}
  \CommentTok{\# Calculate dependent variable Y..}
  \CommentTok{\#.. as a function of population effect size and residual error}
\NormalTok{  y }\OtherTok{\textless{}{-}}\NormalTok{ effect\_size }\SpecialCharTok{*}\NormalTok{ x }\SpecialCharTok{+} \FunctionTok{rnorm}\NormalTok{(n)}
  \CommentTok{\# Return a data.frame}
  \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{x =}\NormalTok{ x, }\AttributeTok{y =}\NormalTok{ y)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Next, write a function to conduct the planned analysis. This function
should receive the data and compute all relevant results from it. The
general format of an analysis function might be:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{planned\_analyis }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data)\{}
  \CommentTok{\# 1. preprocess e.g. with \textasciigrave{}rowMeans(data)\textasciigrave{}}
  \CommentTok{\# 2. conduct analysis e.g. with \textasciigrave{}t.test()\textasciigrave{}}
  \CommentTok{\# 3. \textasciigrave{}return(results)\textasciigrave{}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

In the simplest case, an analysis function might already exist in R. For
the linear model above, the analysis function might be:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{planned\_analyis }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data)\{}
    \FunctionTok{lm}\NormalTok{(y }\SpecialCharTok{\textasciitilde{}}\NormalTok{ x, }\AttributeTok{data =}\NormalTok{ data)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

As soon as we have written \texttt{planned\_anaylsis()} and
\texttt{simulate\_data()} we can iteratively improve both functions,
e.g.~until \texttt{planned\_analysis()} runs without error and recovers
the correct parameters from \texttt{simulate\_data()}. The goal is to
ensure that the output of \texttt{simulate\_data()} works as input to
the function \texttt{planned\_analysis()}.

When the researchers are satisfied with the function
\texttt{planned\_analysis()}, they can think about the way the would
they would like to report the analysis results via tables, plots, and
text. The implementation of this reporting should be in the function
\texttt{report\_analysis()}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{report\_analysis }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(results)\{}
  \CommentTok{\# 1. create markdown tables from results}
  \CommentTok{\# 2. conditionally interpret results e.g. if(p \textless{} .025)"Result is significant."}
  \CommentTok{\# (optional) visualize results}
  \CommentTok{\# 3. return results section formatted in markdown}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

This function should again accept the output of
\texttt{planned\_analysis()} as input. The output of this function
should be formatted in Markdown. The idea is to automatically generate
the full results section from the analysis. This way, the
preregistration not only specifies the computation but also how the its
results are reported. Various packages automatically generate
well-formatted Markdown outputs of statistical reports or even entire
tables of estimates or figures directly from R goal to help with this
objective. Packages like
\href{https://github.com/Rapporter/pander}{\texttt{pander}}
\citep{pander},
\href{https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf}{\texttt{stargazer}}
\citep{stargazer},
\href{https://dstanley4.github.io/apaTables/articles/apaTables.html}{\texttt{apaTables}}
\citep{apaTables} and
\href{http://frederikaust.com/papaja_man/}{\texttt{papaja}}
\citep{papaja} help you to create dynamically generated professional
looking results. The package
\href{https://github.com/easystats/report}{\texttt{report}}
\citep{report} is particularly noteworthy because it not only generates
tables but also a straightforward interpretation of the effects as
actual prose (e.g., it verbally quantifies the size of an effect).

Ideally, these three functions can be composed to create a ``fake''
results section, e.g.~when composed to
\texttt{report\_analysis(planned\_analysis(simulate\_data()))} or
\texttt{simulate\_data()\ \%\textgreater{}\%\ planned\_analysis()\ \%\textgreater{}\%\ report\_analysis()}
outputs a results section.

\hypertarget{turning-a-dynamic-document-into-a-preregistration}{%
\subsubsection{Turning a Dynamic Document into a
Preregistration}\label{turning-a-dynamic-document-into-a-preregistration}}

After researchers are satisfied with their draft preregistration, they
should archive a time-stamped and uneditable version of the project that
serves as the preregistration. zenodo.org \citep{zenodo} is a publicly
funded service provider that archives digital artefacts for research and
provides digital object identifiers (DOI) for these archives. While the
service is independent of GitHub---in terms of storage facilities and
financing---you can link GitHub and zenodo.org. Please note that you can
only link public GitHub repositories to zenodo.org. You may log into
zenodo.org through your GitHub account. To log in with your GitHub
account:

\begin{quote}
Navigate to \url{https://zenodo.org/login/} → Log in with GitHub
\end{quote}

\noindent{}To link zenodo.org and GitHub

\begin{quote}
Log into zenodo.org → Account → GitHub\footnote{\url{https://zenodo.org/account/settings/github/}}
\end{quote}

\noindent{}Or:

\begin{quote}
Navigate to \url{https://zenodo.org/account/settings/github/}
\end{quote}

After you have linked a GitHub repository, you trigger the archival by
creating a GitHub release. To create GitHub release, navigate to GitHub:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{usethis}\SpecialCharTok{::}\FunctionTok{browse\_github}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

Then click on Releases → Draft a new release. Here you can add all
relevant binary files but at least a rendered version of the manuscript
and the Docker image.

To summarize, researchers need to write three functions,
\texttt{planned\_analysis()}, \texttt{simulate\_data()}, and
\texttt{report\_analysis()} and embed these into a manuscript that
serves as a preregistration in an uneditable online repository. After
they gathered the actual data, they can replace the simulated data,
render the dynamic manuscript (therefore run
\texttt{planned\_analysis()} on the actual data), and write the
discussion.

\hypertarget{alternatives-to-simulated-data}{%
\subsection{Alternatives to simulated
data}\label{alternatives-to-simulated-data}}

Simulating data may prove challenging to applied researchers. In the
spirit of team science and collaboration, one feasible solution is to
involve a statistical co-author. However, several easy alternatives
exist. The downside of these alternatives is that they all rely
indirectly on the use of real data. This introduces a risk that the
planned analyses may be cross-contaminated by any exploratory findings.
It is crucial to disclose any exposure to the data in preparation of the
preregistration (PAC or otherwise). This exposure to the data may
decrease trust in the objectivity of the preregistration. Moreover,
researchers should take rigorous measures to prevent exposure to
exploratory findings that may unintentionally influence their decision
making.

The simplest method is to collect empirical data first, but set it aside
and proceed with a copy of the data that is blinded by randomly
shuffling the order of rows for each variable (independently of each
other). Shuffling removes any associations between variables, while
retaining information about the level of measurement and marginal
distribution of each variable. If the hypotheses pertain to associations
between variables, this treatment should thus be sufficient to prevent
cross-contamination. The researcher can still access the information
about means or proportions (e.g., the number of participants belonging
to group ``A'' are in the dataset), but remain uninformed about
relations between variables (e.g., members of group ``A'' have a greater
mean in variable ``Z''). Preregistration after data collection is common
for secondary data analysis of data obtained by other research groups
\citep{westonRecommendationsIncreasingTransparency2019} but not so much
within the same research project. We argue that it is still an eligible
preregistration. Guidelines for clinical trials already recommend
analysis of blinded data to test the feasibility of a preregistration
\citep{ICH1998}.

Another alternative to simulated data is to conduct a pilot study
\citep{thabaneTutorialPilotStudies2010} and use the pilot data to
develop the preregistration. A pilot study has obvious advantages for
study planning, since it lets the researcher evaluate the feasibility of
many assumptions. However, we must warn our readers, that while piloting
is more traditional than our approach of blinding the data before
preregistration, the data from the pilot study must not enter the
analysis data set.

\hypertarget{when-is-pac-applicable}{%
\subsection{When Is PAC Applicable?}\label{when-is-pac-applicable}}

PAC is applicable to every study that can be preregistered and
ultimately uses computer code for the statistical analysis. Two types of
preregistrations are particularly amenable to PAC. First, pregistrations
of clinical trials (called statistical analysis plans, \citet{ICH1998})
typically describe analyses in exhaustive detail and typically contain a
detailed description of how results will be presented, including shells
of tables and graphics \citep{yuanGuideStatisticalAnalysis2019}. PAC may
significantly reduce the required workload while maintaining (and
exceeding) the required standards for preregistering a clinical trial.

Second, preregistering exploratory statistical models (i.e., those with
large numbers of competing models or those inspired by machine learning)
is hardly feasible with standard preregistrations since they are too
complex to describe and depend strongly on their software
implementation. PAC, however, captures the precise algorithmic model,
including its software implementation and is ideal for preregistering
these models \citep{brandmaier22mlsem}.

\hypertarget{preregistration-as-code-tutorial}{%
\subsection{Preregistration as Code:
Tutorial}\label{preregistration-as-code-tutorial}}

We have argued that PAC has several advantages over classic
preregistration and have outlined its implementation. To illustrate how
PAC works in practice and help researchers to implement PAC themselves,
we provide a worked example. We will use an exemplary research question
that was based on openly available data:

\begin{quote}
``Is there a mean difference in the personality trait `Machiavellism'
between self-identified females and males?''
\end{quote}

Again, we propose a preregistration format that closely resembles a
classic journal article but uses simulated data and dynamic document
generation to create a document that starts out as a preregistration and
eventually becomes the final report. The complete preregistration source
is available in the
\href{https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd}{online
supplementary material}. In this section, we show code excerpts of this
preregistration (formatted in monospace) and explain the rationale
behind them.

As usual, the authors state why they are interested in their research
question in the ``Introduction'' section and provide the necessary
background information and literature to understand the context and
purpose of the research question. This example is drastically shortened
for illustration purposes:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{\# Theoretical Background}

\NormalTok{Machiavellianism describes a personality dimension characterized by a}
\NormalTok{cynical disregard of morals in the pursuit of one\textquotesingle{}s own interest, e.g.}
\NormalTok{through manipulation }\CommentTok{[}\OtherTok{@christie1970}\CommentTok{]}\NormalTok{. There is extensive literature reporting}
\NormalTok{differences in the dark triad (narcissism, machiavellianism, and psychopathy)}
\NormalTok{between self{-}identified males and females }\CommentTok{[}\OtherTok{@muris2017}\CommentTok{]}\NormalTok{ but only few studies}
\NormalTok{focus solely on machiavellianism. We aim to replicate the finding that males}
\NormalTok{tend to have higher machiavellianism scores }\CommentTok{[}\OtherTok{@muris2017}\CommentTok{]}\NormalTok{.}
\end{Highlighting}
\end{Shaded}

After researchers have provided the research question, they typically
proceed to explain how they want to study it. For simplicity, we will
use already published data that we have not yet analyzed:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{\# Method}

\NormalTok{We report how we determined our sample size, all data exclusions (if any), all}
\NormalTok{manipulations, and all measures in the study }\CommentTok{[}\OtherTok{cf. @simmons2012}\CommentTok{]}\NormalTok{. We use data}
\NormalTok{available from }\CommentTok{[}\OtherTok{openpsychometrics.org}\CommentTok{](https://openpsychometrics.org/\_rawdata/)}
\NormalTok{from the online version of the MACH{-}IV}\CommentTok{[}\OtherTok{@christie1970}\CommentTok{]}\NormalTok{ and included participants}
\NormalTok{that have responded to at least one machiavellianism item and reported their}
\NormalTok{gender as either "male" or "female".}
\end{Highlighting}
\end{Shaded}

We choose the following statistical procedure because many researchers
are familiar with it\footnote{The t-test and Mann-Whitney-Wilcoxon test
  are arguably the most often used hypothesis tests (according to
  \citep{fagerlandTtestsNonparametricTests2012, hortonStatisticalMethodsJournal2005}
  reports that 26\% of all studies employed a t-test and 27\% employed a
  rank-based alternative in the \emph{New England Journal of Medicine}
  in 2005). The analytical strategy presented here is, in fact,
  suboptimal in several respects (the assumption of measurement
  invariance is untested
  \citep{putnickMeasurementInvarianceConventions2016}, the effect size
  is underestimated in the presence of measurement error
  \citep{frostCorrectingRegressionDilution2000}, the effect size is
  overestimated for highly skewed distributions
  \citep{stonehouseRobustnessTestsCombined1998}). The interested reader
  can use the
  \href{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R}{provided
  code for the simulation} to verify that the t-test provides unbiased
  effect sizes but the Mann-Whitney-Wilcoxon overestimates effect sizes
  with increasing sample size and skewness.}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{We conduct a Student\textquotesingle{}s t{-}test }\CommentTok{[}\OtherTok{@studentProbableErrorMean1908}\CommentTok{]}\NormalTok{ with Welch\textquotesingle{}s}
\NormalTok{correction }\CommentTok{[}\OtherTok{@welchGeneralizationStudentProblem1947}\CommentTok{]}\NormalTok{ of the average of}
\NormalTok{machiavellianism items between the binary{-}coded gender groups. If the skew of}
\NormalTok{this average is greater than 1.0 we conduct a supposedly more robust Mann{-}{-}}
\NormalTok{Whitney{-}{-}Wilcoxon test }\CommentTok{[}\OtherTok{@Wilcoxon1945}\CommentTok{]}\NormalTok{ instead.}
\end{Highlighting}
\end{Shaded}

The methods section is the translation of the following
\texttt{planned\_analysis()} function:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{planned\_analysis }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data, }\AttributeTok{use\_rank =} \StringTok{"skew"}\NormalTok{, }\AttributeTok{skew\_cutoff =} \DecValTok{1}\NormalTok{)\{}
  \CommentTok{\# average over all variable supplied, except gender}
\NormalTok{  machiavellianism }\OtherTok{\textless{}{-}} \FunctionTok{rowMeans}\NormalTok{(data[}\StringTok{"gender"} \SpecialCharTok{!=} \FunctionTok{names}\NormalTok{(data)], }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{)}
  \CommentTok{\# discard rows that only contain NAs}
\NormalTok{  data }\OtherTok{\textless{}{-}}\NormalTok{ data[}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(machiavellianism),]}
\NormalTok{  machiavellianism }\OtherTok{\textless{}{-}}\NormalTok{ machiavellianism[}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(machiavellianism)]}
  \CommentTok{\# assure gender is factor}
\NormalTok{  gender }\OtherTok{\textless{}{-}} \FunctionTok{as.factor}\NormalTok{(data}\SpecialCharTok{$}\NormalTok{gender)}
  \CommentTok{\# note skewness and decide t.test vs wilcox based on it}
\NormalTok{  skew }\OtherTok{\textless{}{-}}\NormalTok{ moments}\SpecialCharTok{::}\FunctionTok{skewness}\NormalTok{(machiavellianism)}
  \CommentTok{\# skewness cutoff}
  \ControlFlowTok{if}\NormalTok{(use\_rank }\SpecialCharTok{==} \StringTok{"skew"}\NormalTok{)use\_rank }\OtherTok{\textless{}{-}} \FunctionTok{abs}\NormalTok{(skew) }\SpecialCharTok{\textgreater{}}\NormalTok{ skew\_cutoff}
  \ControlFlowTok{if}\NormalTok{(use\_rank)\{}
    \CommentTok{\# t.test + rank = wilcox test}
\NormalTok{    machiavellianism }\OtherTok{\textless{}{-}} \FunctionTok{rank}\NormalTok{(machiavellianism)}
\NormalTok{  \}}
\NormalTok{  test }\OtherTok{\textless{}{-}} \FunctionTok{t.test}\NormalTok{(machiavellianism }\SpecialCharTok{\textasciitilde{}}\NormalTok{ gender)}
  \CommentTok{\# return a bunch of information}
  \FunctionTok{list}\NormalTok{(}\AttributeTok{test =}\NormalTok{ test, }\AttributeTok{skew =}\NormalTok{ skew, }\AttributeTok{use\_rank =}\NormalTok{ use\_rank, }\AttributeTok{n =} \FunctionTok{length}\NormalTok{(gender))}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

This function illustrates two advantages of PAC. First, a PAC can easily
include data-dependent decisions by creating different analysis branches
under different conditions. Second, it highlights how difficult it is to
describe a statistical analysis precisely. The same verbal descriptions
may be implemented differently by different persons depending on their
statistical and programming knowledge and assumptions. One example would
be using the function \texttt{wilcox.test} instead of the combinations
of the functions \texttt{rank} and \texttt{t.test}. Either of them is a
valid implementation of the Mann--Whitney--Wilcoxon test, but the first
assumes equal variance. In contrast, the second applies Welch's
correction by default and hence is robust even with unequal variances
across groups \citep{zimmermanRankTransformationsPower1993}. Mentioning
every such minute implementation detail is almost impossible and would
result in overly verbose preregistrations. Still, these details can make
a difference in the interpretation of statistical results and, thus,
represent undisclosed researchers' degrees of freedom.

Together with the function \texttt{simulate\_data()} (not shown here),
the function \texttt{planned\_analysis()} can be used to justify the
planned sample size. To that end, \texttt{simulate\_data()} is
repeatedly called with increased sample sizes and the proportion of
significant results (power) is recorded. The results for such a Monte
Carlo simulation for this example are visualized in Figure 5.1. The code
for this power analysis can be found in the
\href{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R}{online
supplementary material}. The next snippet shows how we integrated the
results dynamically into the preregistration (the origin of the
R-variables \texttt{minn}, \texttt{choosen\_power}, and
\texttt{choosen\_d} is not shown).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/power-1.pdf}
\caption{Figure 5.1: Results of simulation for the power analysis. The
cross indicates the sample size that archives 80\% assuming a Cohen's d
of 0.2.}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{A simulation we conducted indicated that with a sample size of }\InformationTok{\textasciigrave{}r minn\textasciigrave{}}\NormalTok{ for}
\NormalTok{an alpha of .05 (two{-}sided) we achieve at least }\InformationTok{\textasciigrave{}r choosen\_power*100\textasciigrave{}}\NormalTok{\% power}
\NormalTok{assuming a standardized effect size of d=}\InformationTok{\textasciigrave{}r choosen\_d\textasciigrave{}}\NormalTok{.}
\end{Highlighting}
\end{Shaded}

Monte Carlo simulations are, of course, not only applicable for this
analysis method and also allow researchers to investigate further
relevant properties of their analysis method beyond power
\citep{brandmaier2018precision, harrisonIntroductionMonteCarlo2010, raychaudhuriIntroductionMonteCarlo2008}.

We implemented a mechanism that only uses simulated data when the actual
data are not yet available (in this example, if the file
\texttt{data/data.csv} does not exist) for the results section. This
mechanism also warns readers if these results are based on simulated
data. The warning is colored red to avoid any confusion between mock and
actual results. As soon as the actual data are available, the simulated
data are no longer used, and the results represent the actual empirical
results of the study.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{\# Results}

\InformationTok{\textasciigrave{}\textasciigrave{}\textasciigrave{}\{r, echo=FALSE, results=\textquotesingle{}asis\textquotesingle{}, warning=FALSE, message=FALSE\}}
\InformationTok{real\_data \textless{}{-} here::here("data", "data.csv")}
\InformationTok{simulated \textless{}{-} !fs::file\_exists(real\_data)}
\InformationTok{if(simulated)\{}
\InformationTok{  cat("\textbackslash{}\textbackslash{}textcolor\{red\}\{The results are based on simulated data and must not be}
\InformationTok{      interpreted. They only serve to illustrate the result of the preregistered}
\InformationTok{      code.\}")}
\InformationTok{  set.seed(1234)}
\InformationTok{  mach \textless{}{-} simulate\_data(900, 8, 0.3, 10)}
\InformationTok{\} else \{}
\InformationTok{  mach \textless{}{-} readr::read\_delim(real\_data, delim = "\textbackslash{}t", na = c("", "NA", "NULL"))}
\InformationTok{  \# only keep MACH items + gender}
\InformationTok{  mach \textless{}{-} dplyr::select(mach, dplyr::matches("\^{}Q\textbackslash{}\textbackslash{}d+A$"), gender)}
\InformationTok{  \# code gender according to codebook (3 would be other)}
\InformationTok{  mach \textless{}{-}}
\InformationTok{    dplyr::mutate(mach, gender = factor(}
\InformationTok{      gender,}
\InformationTok{      levels = 1:2,}
\InformationTok{      labels = c("male", "female")}
\InformationTok{    ))}
\InformationTok{  \# some items are reversed, see https://core.ac.uk/download/pdf/38810542.pdf}
\InformationTok{  reversed\_nr \textless{}{-} c(1, 15, 2, 12, 4, 11, 14, 19)}
\InformationTok{  reversed \textless{}{-} stringr::str\_c("Q", reversed\_nr, "A")}
\InformationTok{  mach \textless{}{-} dplyr::mutate(mach, dplyr::across(one\_of(reversed), \textasciitilde{} 6 {-} .x))}
\InformationTok{\}}
\InformationTok{\textasciigrave{}\textasciigrave{}\textasciigrave{}}
\end{Highlighting}
\end{Shaded}

Following the recommendations outlined in this paper, we did not access
the data when we initially wrote this code. We therefore did not know
the exact format the data would have. This means that we did need to
change our preregistration after accessing the data to include i.e.~the
recoding of gender (lines 17-22) and the items (lines 23-26).\footnote{We
  invite the reader to evaluate the changes we made to the preregistered
  code. Either on
  \href{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg...main}{GitHub}
  → ``Files changed'' or directly in Git with
  \texttt{git\ diff\ v0.0.1.1-prereg\ preregistration.Rmd}.} This is our
summary of what we changed:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{\# Discussion}

\NormalTok{This document only serves to illustrate Preregistration as Code. We, therefore,}
\NormalTok{do not discuss the results. After we have acquired the data, we realized that}
\NormalTok{we had to change the code for reading the data, including recoding gender,}
\NormalTok{missing values and reversed items (see commit }\CommentTok{[}\OtherTok{6556a93}\CommentTok{]}\NormalTok{(https://github.com/}
\NormalTok{aaronpeikert/repro{-}tutorial/commit/6556a9395fcdd600b5b0c5358f92a2c6635ae360)}
\NormalTok{and commit }\CommentTok{[}\OtherTok{9f7ab21}\CommentTok{]}\NormalTok{(https://github.com/aaronpeikert/repro{-}tutorial/commit/}
\NormalTok{9f7ab212dfaf84a0398752a4b80cf14c71000d00)). We do not believe that these changes}
\NormalTok{influence the results substantively.}
\end{Highlighting}
\end{Shaded}

Readers can inspect and judge the changes for themselves on
\href{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main\#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b}{GitHub}.

The last thing we need to preregister is the reporting of our results
with the combination of the functions \texttt{planned\_analysis()} and
\texttt{report\_analysis()}.

\begin{Shaded}
\begin{Highlighting}[]
\InformationTok{\textasciigrave{}\textasciigrave{}\textasciigrave{}\{r, echo=FALSE, results=\textquotesingle{}asis\textquotesingle{}\}}
\InformationTok{report\_analysis(planned\_analysis(mach))}
\InformationTok{\textasciigrave{}\textasciigrave{}\textasciigrave{}}
\end{Highlighting}
\end{Shaded}

This is an example of how the results could be reported (based on
simulated data):

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{report\_analysis}\NormalTok{(}\FunctionTok{planned\_analysis}\NormalTok{(}\FunctionTok{simulate\_data}\NormalTok{(}\DecValTok{900}\NormalTok{, }\DecValTok{8}\NormalTok{, }\FloatTok{0.3}\NormalTok{, }\DecValTok{10}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
The Welch Two Sample t-test testing the difference of machiavellianism by gender
(mean in group male = 0.96, mean in group female = 0.79) suggests that the
effect is - negative, statistically significant, and small (difference = -0.17,
95% CI [0.12, 0.22], t(887.46) = 6.38, p < .001; Cohen's d = 0.43, 95% CI [0.30,
0.56])
\end{verbatim}

This example of a preregistration covers a single study with a single
hypothesis. To organize studies with multiple hypotheses, we suggest
multiple \texttt{planned\_analysis()} and \texttt{report\_analysis()}
functions (possibly numbered in accordance with the hypotheses,
e.g.~1.2, 2.3 etc.). Preregistrations that cover multiple distinct data
sources may employ multiple \texttt{simulate\_data()} functions. These
are merely suggestions, and researchers are encouraged to find their own
way of how to best organize their analysis code.

The example rendered as a
\href{https://github.com/aaronpeikert/repro-tutorial/files/7309455/preregistration.pdf}{PDF
file with real data} is available in the
\href{https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.3.1-results}{online
supplementary material}. The changes we made since preregistering it can
be inspected on this
\href{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main\#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b}{GitHub
page}.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Increased automation is increasingly recognized as a means to improve
the research process \citep{rouderMinimizingMistakesPsychological2019},
and therefore this workflow fits nicely together with other innovations
that employ automation, like machine-readable hypothesis tests
\citep{lakensImprovingTransparencyFalsifiability2021} or automated data
documentation \citep{arslanHowAutomaticallyDocument2019}. Automated
research projects promise a wide range of applications, among them PAC
\citep[potentially to be submitted as a registered
report][]{nosekRegisteredReports2014, chambersWhatNextRegistered2019},
direct replication \citep{simonsValueDirectReplication2014}, fully
automated living metanalysis \citep{elliottLivingSystematicReviews2014},
executable research articles
\citep{elifesciencespublicationsELifeLaunchesExecutable2020}, and other
innovations such as the live analysis of born open data
\citep{rouderWhatWhyHow2016, kekecsRaisingValueResearch2019}.

Central to these innovations is a property we call ``reusability,''
fully promoted by the present workflow. Reusable code can run on
different inputs from a similar context and produce valid outputs. This
property is based on reproducibility but requires the researcher to more
carefully write the software
\citep{lanerganSoftwareEngineeringReusable1989} such that it is
\emph{built-for-reuse}
\citep{al-badareenReusableSoftwareComponents2010}. The reproducible
workflow we present here is heavily automated and hence promotes
reusability. Furthermore, adhering to principles of reusability
typically removes errors in the code and thus increases the likelihood
that the statistical analysis is correct. Therefore reproducibility
facilitates traditional good scientific practices and provides the
foundation for promising innovations.

\hypertarget{summary-1}{%
\subsection{Summary}\label{summary-1}}

This paper demonstrated how the R package \texttt{repro} supports
researchers in creating reproducible research projects, including
reproducible manuscripts. These are important building blocks for
transparent and cumulative science because they enable others to
reproduce statistical and computational results and reports later in
time and on different computers. The workflow we present here rests on
four software solutions, (1) version control, (2) dynamic document
generation, (3) dependency tracking and (4) software management to
guarantee reproducibility. We first demonstrated how to create a
reproducible research project. Then, we illustrated how such a project
could be reproduced---either by the original author and/or collaborators
or by a third party.

We finally presented an example of how the rigorous and automated
reproducibility workflow introduced by \texttt{repro} may enable other
innovations, such as Preregistration as Code (PAC). In PAC the entire
reproducible manuscript, including planned analyses and results based on
simulated data, is preregistered. This way, every use of a researchers'
degree of freedom is disclosed. Once real data is gathered, the
reproducible manuscript is (re-)created with the real data. PAC only
becomes possible because reproducibility is ensured and leverages
version control and dynamic document generation as key features of the
workflow.

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

We realize that the workflow outlined in this paper, and its application
in PAC, remains challenging despite our efforts to simplify the
procedure by means of the \texttt{repro} package. This paper should be
considered as a starting point for those seeking to improve the
reproducibility of their research. Two kinds of limitations can be
distinguished. The first kind are limitations by design which are
unlikely to change. Our workflow inherits these from the software it
relies on and the fundamental design principles these share with the
workflow and \texttt{repro}. The second kind are limitations in
\texttt{repro} and its dependencies that may be overcome by our future
efforts and those of the open-source development community.

With regard to limitations by design, the workflow outlined in this
paper is fundamentally incompatible with steps that cannot be automated.
This principle may be at odds with some ingrained habits of researchers
to mix and match manual and automated steps in data analysis. To allow
for automation, many researchers will have to search for alternative
software.\\
The automation-friendly software we present here has several technical
but critical limitations. For example, \texttt{Git} can track any
filetype, but tracked changes are only meaningful for text files (with
endings like, \texttt{.txt}, \texttt{.csv}, \texttt{.R}, \texttt{.py},
or \texttt{.Rmd}), not for binary files (with endings like
\texttt{.docx}, \texttt{.exe}, or \texttt{.zip}). Furthermore, tables
and graphics dynamically generated from code are difficult to edit by
hand. \texttt{Make} can automate any programmable software, but not
software that is exclusively controlled through a point-and-click user
interface. Finally, \texttt{Docker} can ship software that runs on Linux
and can be automatically installed, which precludes much commercial or
closed-source software.

This move away from software that has served researchers well for
decades is understandably difficult and presents us with a conundrum. On
the one hand, we firmly believe that automated reproducibility makes
research more productive and collaboration easier. But, on the other
hand, we expect researchers to invest considerable time in learning new
tools and to persuade their collaborators to do the same. Three
arguments reconcile this apparent paradox. First, this change will not
happen all at once. Automated reproducibility is an ideal that we
believe has many advantages, but it is not an all-or-nothing decision.
Researchers can pick up one skill at a time and then help their fellow
collaborators to do the same. Second, the upfront investment is required
once (and efforts such as \texttt{repro} are underway to reduce it) and
will pay dividends over many research projects. Third, the move towards
open software for research offers several benefits beyond enabling
automated reproducibility
\citep{schaffnerFutureScientificJournals1994, fitzgeraldTransformationOpenSource2006, chaldecottHistoryScientificTechnical1965, sonnenburgNeedOpenSource2007}.

With regard to surmountable limitations, we acknowledge that the
\texttt{repro} package is still in development. One limitation is that
\texttt{repro} relies on several software dependencies, which represents
a threat to long-term reproducibility in itself. For example, to benefit
from automatic and convenient reproduction, researchers must use
\texttt{Git}, \texttt{Make}, and \texttt{Docker}. However, \texttt{Git}
and \texttt{Make} are themselves included in the \texttt{Docker} image
created by \texttt{repro}. Researchers can therefore employ the
\texttt{Docker} image manually to download the \texttt{Git} repository
and execute \texttt{Make} for full reproduction. In other words, the
only hard requirement for reproduction and therefore its Achilles' heel,
is \texttt{Docker}. The \texttt{Docker} approach has two
vulnerabilities. First, and more importantly, the \texttt{Docker} image
for the project and the \texttt{Git} repository have to remain
available. The \texttt{Dockerfile} (the plain text description to build
a \texttt{Docker} image), as opposed to the image, is insufficient
because it relies on too many service providers (e.g., Microsoft R
Application Network, Ubuntu Package Archive). To overcome this
limitation, we recommend archiving the \texttt{Git} repository and the
\texttt{Docker} image with zenodo.org, a non-profit long-term storage
for scientific data. The necessary steps for archival on zenodo.org are
described at the end of Section {[}Preregistration as Code --- a
Tutorial{]}.

The second vulnerability is that even if the existence of the
\texttt{Docker} image and \texttt{Git} repository is guaranteed, future
researchers still require software to run the image. To that end, they
can either rely on \texttt{Docker} itself or \texttt{Docker}-compatible
alternatives (e.g., \texttt{CoreOS\ rkt}, \texttt{Mesos\ Containerizer},
\texttt{Singularity}). The only way to remove the reliance on such
external software is to turn the \texttt{Docker} image into a full
operating system that subsequently can be installed and run on almost
any modern computer. This process is technically possible and would
guarantee reproducibility for decades without any software dependency,
assuming hardware that conforms to the x86 instruction set architecture
continues to be available. However, this process requires much technical
knowledge and is currently not facilitated by \texttt{repro}. With
regard to this vulnerability, it is worthwhile to note that the
\texttt{R\ Markdown}, \texttt{Makefile}, and \texttt{Dockerfile} do
provide information that allows researchers to trace the computational
steps and recreate the computational environment manually. The
\texttt{Makefile}, for example, is written in a way that researchers can
manually trace the dependencies and execute commands in the right order,
in case they are unable to run \texttt{Make} for some reason. Thus
hypothetically, even if \texttt{Docker} were to become unavailable one
day, the \texttt{Dockerfile} still serves as unambiguous documentation
of how the original system was set up and may help future users to
create a software environment that closely resembles the original.

\hypertarget{outlook}{%
\subsection{Outlook}\label{outlook}}

Open science practices are a continually evolving field where technical
innovations foster changes in research practice. Open data are much more
widespread today thanks to online storage facilities; preregistration is
possible because there are preregistration platforms and so forth.
Similarly, we hope that fully automatic reproduction, e.g., with
\texttt{repro} as a technical innovation, will promote increased
scientific rigor, efficiency, and productivity.

In practice, this ideal of a fully automatic reproduction of research
projects can conflict with the wide range of demands for more
user-friendly and powerful software. Some may find that \texttt{Make} is
too complicated or that \texttt{Docker} requires too much storage space.
Yet others may find that they require other programming languages or
want to scale their computation across hundreds of computers, e.g., via
high-performance computing clusters or cloud computing.

\texttt{repro} was designed modularly to meet many such demands. At the
moment, \texttt{repro} only supports the combination of
\texttt{R\ Markdown}, \texttt{Git}, \texttt{Make}, and \texttt{Docker}.
However, there are alternatives for each of these elements that may fit
better into an individual research project. \texttt{R\ Markdown} could
be complemented or replaced by a dynamic Microsoft Word document with
the help of \texttt{officer} \citep{officer} or \texttt{officedown}
\citep{officedown} to accommodate a wider range of journal submission
standards. Instead of using formal version control with \texttt{Git},
\texttt{repro} could automatically save snapshots for increasing
user-friendliness. \texttt{Make} could be replaced by the more
R-centered alternative \texttt{targets} for more convenience.
\texttt{Docker} could be combined with \texttt{renv} \citep{R-renv} to
control the package versions precisely (our approach fixes the date,
\texttt{renv} the exact package version). Alternatively, \texttt{Docker}
could be replaced by the more lightweight \texttt{renv} if no
dependencies outside of R are considered crucial. \texttt{Docker} does
not satisfy the requirements of many HPC environments, but
\texttt{Singularity} was designed to avoid this limitation while still
being compatible with \texttt{Docker} images.

\texttt{repro}'s modular structure allows such alternative workflows,
though they have not yet been implemented. Depending on the demand by
users, we will implement some of them in \texttt{repro} and hope for
broad adoption of computational reproducibility in the near future.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %% optional
% \supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
%
% % Only for the journal Methods and Protocols:
% % If you wish to submit a video article, please do so with any other supplementary material.
% % \supplementary{The following are available at www.mdpi.com/link: Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{We would like to thank Maximilian Stefan Ernst for his
contributions to the code for the simulation study. We are grateful to
Julia Delius for her helpful assistance in language and style editing.
The R package \texttt{repro} was developed as part of the
\href{https://aaronpeikert.github.io/repro-thesis/}{first author's
master thesis} at the Humboldt-Universität zu Berlin.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{Aaron Peikert took the lead in writing and provided
the initial draft of the manuscript. Andreas Brandmaier and Caspar J.
Van Lissa contributed further ideas, critical feedback, and revisions of
the original manuscript.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{The authors declare no conflict of interest. We
have received no financial support for the research, authorship, and/or
publication of this article.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent
\begin{tabular}{@{}ll}
PAC & Preregistration as Code \\
Gb & Gigabyte \\
Kb & Kilobyte \\
CRAN & Comprehensive R Archive Network \\
WORCS & Workflow for Open Reproducible Code in Science \\
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: internal bibliography
%=====================================
%\reftitle{References}
%\begin{thebibliography}{999}
% Reference 1
%\bibitem[Author1(year)]{ref-journal}
%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
%\bibitem[Author2(year)]{ref-book}
%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
%\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%=====================================
% References, variant B: external bibliography
%=====================================
\reftitle{References}
\externalbibliography{yes}
\bibliography{temp.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2021-12-01 manuscript.tex
add mdpi proof

@@
 %=================================================================
-\documentclass[psych,tutorial,submit,moreauthors,pdftex]{mdpi}
+\documentclass[psych,tutorial,accept,moreauthors,pdftex]{Definitions/mdpi}
 
@@
 %% Some pieces required from the pandoc template
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{4pt}}
-\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
-\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}
+%\providecommand{%\tightlist}{%
+ % \setlength{\itemsep}{0pt}\setlength{\parskip}{4pt}}
+%\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
+%\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}
 
@@
 \makeatother
-\pubvolume{xx}
+\pubvolume{1}
 \issuenum{1}
@@
 \copyrightyear{2019}
-%\externaleditor{Academic Editor: name}
-\history{Received: date; Accepted: date; Published: date}
-\updates{yes} % If there is an update available, un-comment this line
+\externaleditor{Academic Editor: {Alexander Robitzsch}} %MDPI: Please add the academic editor.
+\datereceived{} 
+\dateaccepted{} 
+\datepublished{}
+\hreflink{https://doi.org/} % If needed use \linebreak
 
@@
 % Full title of the paper (Capitalized)
-\Title{Reproducible Research in R: A Tutorial on how to Do the Same
+\Title{Reproducible Research in R: A Tutorial on How to Do the Same
+Thing More \highlighting{Than Once}} %MDPI: the article type in the document is different with that in Susy, please confirm.
+
+\TitleCitation{Reproducible Research in R: A Tutorial on How to Do the Same
 Thing More Than Once}
@@
 % Authors, for the paper (add full first names)
-\Author{Aaron Peikert$^{1, 2,
-*}$\href{https://orcid.org/0000-0001-7813-818X}{\orcidicon}, Caspar J.
-van Lissa$^{3,
-4}$\href{https://orcid.org/0000-0002-0808-5024}{\orcidicon}, Andreas M.
-Brandmaier$^{1, 5,
-6}$\href{https://orcid.org/0000-0001-8765-6982}{\orcidicon}}
+\Author{\highlighting{Aaron Peikert~$^{1,2,*}$\href{https://orcid.org/0000-0001-7813-818X}{\orcidicon}} %Please carefully check the accuracy of names and affiliations.
+Changes will not be possible after proofreading.
+, \highlighting{Caspar J. van Lissa~$^{3,4}$\href{https://orcid.org/0000-0002-0808-5024}{\orcidicon}} %MDPI: This author is different from susy. Please check.
+ and Andreas M. Brandmaier~$^{1,5,6}$\href{https://orcid.org/0000-0001-8765-6982}{\orcidicon}}
 
 % Authors, for metadata in PDF
-\AuthorNames{Aaron Peikert, Caspar J. van Lissa, Andreas M. Brandmaier}
-
+\AuthorNames{Aaron Peikert, Caspar J. van Lissa and Andreas M. Brandmaier}
+
+\AuthorCitation{Peikert, A.;  van Lissa, J.; Brandmaier, A.M.}
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
 \address{%
-$^{1}$ \quad Center for Lifespan Psychology---Max Planck Institute for
-Human Development, Lentzeallee 94, 14195 Berlin, Germany; \\
-$^{2}$ \quad Humboldt-Universität zu Berlin, Berlin, Germany; \\
-$^{3}$ \quad Department of Methodology \& Statistics---Utrecht
-University, Faculty of Social and Behavioral Sciences, Utrecht,
-Netherlands; \\
-$^{4}$ \quad Open Science Community Utrecht, Utrecht, Netherlands; \\
-$^{5}$ \quad Max Planck UCL Centre for Computational Psychiatry and
-Ageing Research Berlin, Germany, and London, UK; \\
-$^{6}$ \quad MSB Medical School Berlin, Berlin, Germany; \\
-}
+$^{1}$ \quad Center for Lifespan Psychology---Max Planck Institute for Human Development, Lentzeallee 94, \mbox{14195 Berlin, Germany;} brandmaier@mpib-berlin.mpg.de\\
+$^{2}$ \quad Humboldt-Universität zu Berlin, \hl{Berlin}, Germany\\%MDPI: Please add the post code.
+$^{3}$ \quad \hl{Department} of Methodology \& Statistics---Utrecht \hl{University}, \hl{Faculty} of Social and Behavioral Sciences%MDPI: Please check the address order.
+(order guide: Center / Laboratory / Group / Programme / Unit < Department / Division / Faculty / Campus / Institute / School < College / University / Universiti)
+, \hl{Utrecht}, The Netherlands; C.J.vanLissa@uu.nl\\
+$^{4}$ \quad Open Science Community Utrecht, \hl{Utrecht}, The Netherlands\\
+$^{5}$ \quad Max Planck UCL Centre for Computational Psychiatry and Ageing Research, \mbox{\hl{Berlin, Germany, and London, UK}} %MDPI: 1.
+Please check if there are two separator affiliations.
+If so, please split them.
+2.
+Please add the post code.
+\\
+$^{6}$ \quad MSB Medical School Berlin, \hl{Berlin}, Germany}
+
 % Contact information of the corresponding author
-\corres{Correspondence: \href{mailto:peikert@mpib-berlin.mpg.de}{\nolinkurl{peikert@mpib-berlin.mpg.de}}}
+\corres{Correspondence: peikert@mpib-berlin.mpg.de}
 
@@
 other computers and later in time, to reproduce and thus understand how
-results came about while avoiding a variety of errors that may lead to
+results came about, while avoiding a variety of errors that may lead to
 erroneous reporting of statistical and computational results. In this
@@
 projects with tools from the software engineering community. Building
-upon this notion of fully automated reproducibility we present several
+upon this notion of fully automated reproducibility, we present several
 applications including the preregistration of research plans with code
@@
 advancements that serve reproducibility more widely accessible for
-researchers holds the potential to innovate the research process to
-become more productive, credible, and reliable.}
+researchers holds the potential to innovate the research process and to
+help it become more productive, credible, and reliable.}
 
@@
 \keyword{open science; computational reproducibility; preregistration;
-R; R Markdown; Make; GitHub; Docker}
+R; R Markdown; make; GitHub; docker}%MDPI: 1. The upper case has been removed. Please confirm. 2. Please check if R is duplicated here.
 
@@
 
-\captionsetup{labelformat=empty}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+%\captionsetup{labelformat=empty}
+
+%\hypertarget{introduction}
+\section{Introduction}\label{introduction}
 
@@
 defined as the ability of anyone to obtain identical results from the
-\emph{same} data with the \emph{same} computer code \citep[see][ for
-details]{Peikert2019}. The credibility of empirical research results
+\emph{same} data with the \hl{\emph{same}} %MDPI: Is the italic necessary? Can it be deleted?
+ computer code (see~\citep{Peikert2019} \hl{for
+details).} %MDPI: The structure of reference citation has been changed. So as the follows. Please confirm.
+ The credibility of empirical research results
 hinges on their objectivity. Objectivity in this context means, ``in
 principle it {[}the research finding{]} can be tested and understood by
-anybody.'' \citep[p.~22]{popperLogicScientificDiscovery2002}. Only a
+anybody.'' (\citep{popperLogicScientificDiscovery2002} p.~22). Only a
 reproducible result meets these requirements. Therefore, reproducibility
@@
 taken), and there is a lack of best practices for overcoming these
-challenges \citep[but see][]{vanlissa2020worcs}. With technological
+challenges (but see~\citep{vanlissa2020worcs}). With technological
 advancement, however, it is now possible to make all digital products
@@
 goal, we discuss how to address four common threats to reproducibility,
-using tools originating from software engineering \citep[see][ for
-details]{Peikert2019}, and present a tutorial on how to employ these
+using tools originating from software engineering (see~\citep{Peikert2019} for
+details), and present a tutorial on how to employ these
 tools to achieve automated reproducibility. A single tutorial cannot
@@
 In pursuit of the second aim, we present a strictly reproducible and
-unambiguous form of preregistration \citep{NosekRevolution2018} that
+unambiguous form of preregistration~\citep{NosekRevolution2018} that
 builds upon implementing this reproducible workflow, the so-called
@@
 \citep{hardwicke2018}. Out of 35 published articles with open code and
-data, results of 22 articles could be reproduced, but in 11 of these
-cases, further assistance from the original authors was required. For 13
-articles, at least 1 outcome could not be reproduced-----even with the
-original authors' assistance. Another study of 62 Registered Reports
+data, the results of 22 articles could be reproduced, but further assistance from the original authors was required in 11 of these cases. For 13
+articles, at least one outcome could not be reproduced---even with the
+original authors' assistance. Another study of 62 registered reports
 found that only 41 had data available, and 37 had analysis scripts
-available \citep{obels2020}. The authors could execute only 31 of the
+available~\citep{obels2020}. The authors could execute only 31 of the
 scripts without error and reproduce the results of only 21 articles
@@
 research products are available but not reproducible, their added value
-is limited.
+is~limited.
 
@@
 Unless stated otherwise, all code blocks are meant to be run in the
-statistical programming language R \citep[tested with version
-4.0.4]{R-base}.
-
-\hypertarget{threats-to-reproducibility-and-appropriate-remedies}{%
+statistical programming language R (\citep{R-base} tested with version
+4.0.4).
+
+%\hypertarget{threats-to-reproducibility-and-appropriate-remedies}
 \section{Threats to Reproducibility and Appropriate
-Remedies}\label{threats-to-reproducibility-and-appropriate-remedies}}
+Remedies}\label{threats-to-reproducibility-and-appropriate-remedies}
 
@@
 \def\labelenumi{\arabic{enumi}.}
-\tightlist
+%\tightlist
 \item
-  \emph{Multiple inconsistent versions of code, data, or both}; for
+  \highlighting{\emph{Multiple inconsistent versions of code, data, or both};} %MDPI: Is the italic necessary? Can it be deleted?
+ for
   example, the data set may have changed over time because outliers were
@@
 \item
-  \emph{Copy-and-paste errors}; for example, results are often manually
+  \highlighting{\emph{Copy-and-paste errors};} for example, results are often manually
   copied from a statistical computing language into a text processor; if
@@
 \item
-  \emph{Undocumented or ambiguous order of computation}; for example,
+  \highlighting{\emph{Undocumented or ambiguous order of computation};} for example,
   with multiple data and code files, it may be unclear which scripts
@@
 \item
-  \emph{Ambiguous software dependencies}; for example, a given analysis
+  \highlighting{\emph{Ambiguous software dependencies};} for example, a given analysis
   may depend on a specific version of a specific software package, or
@@
 four pillars that address the aforementioned causes of
-non-reproducibility \citep{Peikert2019}:
+non-reproducibility~\citep{Peikert2019}:
 
 \begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Version control
-\item
-  Dynamic document generation
-\item
-  Dependency tracking
-\item
-  Software management
+%\def\labelenumi{\arabic{enumi}.}
+%%\tightlist
+\item  Version control
+\item  Dynamic document generation
+\item  Dependency tracking
+\item  Software management
 \end{enumerate}
@@
 reproducibility. A more extensive treatment of these tools is given in
-\citet{Peikert2019}.
+Peikert and Brandmaier~\citep{Peikert2019}.
 
@@
 version. Moreover, contributions (e.g., from collaborators) can be
-compared to, and incorporated in the main version of the project.
+compared to and incorporated in the main version of the project.
 Version control thus reduces the risk of losing work and facilitates
@@
 that can later be merged again) to facilitate simultaneous
-collaboration. \citet{vuorreCuratingResearchAssets2018} provide a more
+collaboration. Vuorre and Curley~\citep{vuorreCuratingResearchAssets2018} provide a more
 extensive treatment of how \texttt{Git} functions and how to use
-\texttt{Git} for research. \citet{bryanExcuseMeYou2018} provides
+\texttt{Git} for research. Bryan~\citep{bryanExcuseMeYou2018} provides
 additional information on how to track \texttt{R\ Markdown} documents.
@@
 \texttt{Git} repositories because of its popularity among R users.
-GitHub has many tools that facilitate working with \texttt{Git} --- in
-particular project management and collaboration --- but these are not
+GitHub has many tools that facilitate working with \texttt{Git}---
+particularly in project management and collaboration---but these are not
 central to achieving reproducibility.
@@
 ranging from simple sketches of ideas to scientific manuscripts
-\citep{R-rticles} and presentations \citep{revealjs}, or even résumés
+\citep{R-rticles} and presentations~\citep{revealjs}, or even résumés
 \citep{vitae}. \texttt{R\ Markdown} extends regular \texttt{Markdown} by
@@
 presentation slides or a journal article).
-\citet{xieMarkdownCookbook2020} provides an extensive yet practical
+Xie et al.~\citep{xieMarkdownCookbook2020} provides an extensive yet practical
 introduction to most features of \texttt{R\ Markdown}.
@@
 scientific data analysis, targets are typically the final scientific
-report (e.g., the one to be submitted to a journal) and possibly
-intermediate results (such as preprocessed data files, simulation
+report (e.g., the one to be submitted to a journal) and produce possibly
+intermediate %pls check (such as preprocessed data files, simulation
 results, and analysis results). A workflow that involves renaming
@@
 in R), but they also need to guarantee that the pseudo-random number
-generator is unchanged \citep[see][]{Peikert2019}.
+generator is unchanged~(see~\citep{Peikert2019}).
 
@@
 
+\begin{adjustwidth}{0.5cm}{0cm} 
 \begin{Shaded}
@@
 \end{Shaded}
+\end{adjustwidth}
+
+%%\begin{figure*}[H]
+%\begin{Shaded}
+%%\begin{Highlighting}[]
+%\noindent\ExtensionTok{\texttt{manuscript.pdf:}}\NormalTok{\texttt{ manuscript.Rmd simulated\_data.csv}}\\[-2ex]
+%
+%\ExtensionTok{\texttt{Rscript}} \AttributeTok{\texttt{{-}e}} \StringTok{\texttt{\textquotesingle{}rmarkdown::render("manuscript.Rmd")\textquotesingle{}}}\\
+%
+%\noindent\ExtensionTok{\texttt{simulated\_data.csv:}}\NormalTok{\texttt{ simulate.R}}\\[-2ex]
+%
+%\ExtensionTok{\texttt{Rscript}} \AttributeTok{\texttt{{-}e}} \StringTok{\texttt{\textquotesingle{}source("simulate.R")\textquotesingle{}}}
+%%\end{Highlighting}
+%\end{Shaded}
+%%\end{figure*}
 
@@
 final \texttt{manuscript.pdf} by rendering the \texttt{R\ Markdown} with
-the command shown in l.2, \texttt{Make} would check whether the file
+the command shown in l.~2, \texttt{Make} would check whether the file
 \texttt{simulation\_results.csv} exists; if not, it would issue the
-command in l.5 to run the simulation before rendering the manuscript.
+command in l.~5 to run the simulation before rendering the manuscript.
 This ensures that the simulated data are present before the manuscript
@@
 relies on external software. Troubleshooting issues specific to a
-particular programming language or dependent tool typically require
-considerable expertise and threaten reproducibility. \emph{Software
+particular programming language or dependent tool typically requires
+considerable expertise and threatens reproducibility. \emph{Software
 management} refers to the act of providing records of, or access to, all
@@
 the software much easier to use, and the results easier to reproduce
-{[}\ldots{]}'' \citep[p.~174]{silverSoftwareSimplified2017}.
+{[}\ldots{]}'' (\citep{silverSoftwareSimplified2017} p.~174).
 
@@
 All of the differences in computational results that could be caused by
-variation in the software used are hence eliminated.
+variations in the software used are hence eliminated.
 
@@
 purposes. For example, there are pre-made images that only include what
-is necessary for R, based on Ubuntu containers
-\citep{boettigerIntroductionRockerDocker2017}. Users can then install
+is necessary for R, based on Ubuntu containers~\citep{boettigerIntroductionRockerDocker2017}. Users can then install
 whatever they need in addition to what is provided by these pre-compiled
-images. The image that was used for this article uses 1.35GiB of disk
+images. The image that was used for this article uses 1.35~GiB of disk
 space. The image for this project includes Ubuntu, R, RStudio, LaTeX as
-well as a variety of R packages like tidyverse \citep{tidyverse} and all
+well as a variety of R packages like tidyverse~\citep{tidyverse} and all
 its dependent packages, amounting to 192 R packages.
@@
 executed to recreate the software environment. \texttt{Dockerfiles} are
-tiny (the
-\href{https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile}{\texttt{Dockerfile}}
-for this project has a size of only 1.55KiB). However, they rely on the
+tiny (the \hl{\texttt{Dockerfile}} %MDPI: Please add the accessed date.
+ (\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile})
+for this project has a size of only 1.55~KiB). However, they rely on the
 assumption that all software repositories that provide the dependent
@@
 comprehensive overview of the use of containerization in research
-projects is given by
-\citet{wiebelsLeveragingContainersReproducible2021}. Note that WORCS,
-which we presented elsewhere \citep{vanlissa2020worcs} relies on the R
-package \texttt{renv} \citep{R-renv} for software management.
-\texttt{renv} is more lightweight and easier to use than \texttt{Docker}
-on the one hand, but not as comprehensive on the other because it only
+projects is given by Wiebels and Moreau~\citep{wiebelsLeveragingContainersReproducible2021}. Note that WORCS,
+which we presented elsewhere~\citep{vanlissa2020worcs} relies on the R
+package \texttt{renv}~\citep{R-renv} for software management. Although
+\texttt{renv} is more lightweight and easier to use than \texttt{Docker}, it is not as comprehensive because it only
 takes snapshots of the R packages instead of all software used.
 
-To summarize, the workflow by \citet{Peikert2019} requires four
-components (see Figure 2.1) dynamic document generation (using
+To summarize, the workflow by Peikert and Brandmaier~\citep{Peikert2019} requires four
+components (see \highlighting{Figure~\ref{figure1})} %MDPI: Figure 2.1 has been changed into Figure 1. Please confirm.
+ dynamic document generation (using
 \texttt{R\ Markdown}), version control (using \texttt{Git}), dependency
@@
 the remainder, we will walk you through the creation of a reproducible
-research project with the package \texttt{repro}.
-
-\begin{figure}
-\centering
-\includegraphics{/home/rstudio/images/nutshell.pdf}
-\caption{Figure 2.1: Schematic illustration of the interplay of the four
+research project with the package~\texttt{repro}.
+
+\begin{figure}[H]
+%\centering
+\includegraphics[width=13.5cm]{Figures/nutshell.pdf}
+\caption{Schematic illustration of the interplay of the four
 components (in dashed columns) central to the reproducible workflow:
@@
 using dynamic document generation in \texttt{R\ Markdown}. Adapted from
-\citet{Peikert2019} licensed under
-\href{https://creativecommons.org/licenses/by/4.0}{CC BY 4.0}.}
+Peikert and Brandmaier~\citep{Peikert2019} licensed under CC BY 4.0 
+(\url{https://creativecommons.org/licenses/by/4.0}).}
+\label{figure1}
 \end{figure}
 
-\hypertarget{creating-reproducible-research-projects}{%
+%\hypertarget{creating-reproducible-research-projects}
 \section{Creating Reproducible Research
-Projects}\label{creating-reproducible-research-projects}}
+Projects}\label{creating-reproducible-research-projects}
 
@@
 introduces helper functions that simplify the use of complicated and
-powerful tools. The \texttt{repro} package follows the format of the
-\href{https://usethis.r-lib.org}{\texttt{usethis}} \citep{usethis}
+powerful tools. The \texttt{repro} package follows the format of the{\texttt{usethis}} 
+ \mbox{(\url{https://usethis.r-lib.org})}~\citep{usethis}
 package, which provides helper functions to simplify the development of
@@
 accessible by enabling beginner-level users to detect their system's
-state accurately and act correspondingly \citep[Chapter 8: ``Automation
-and Situation Awareness'']{parasuramanAutomationHumanPerformance2018}.
+state accurately and act correspondingly (\citep{parasuramanAutomationHumanPerformance2018} Chapter 8: ``Automation
+and Situation Awareness'').
 These wrappers are merely a support system; as users learn to use the
@@
 
-This tutorial assumes that the user will be working predominantly in R,
+This tutorial assumes that the user will be working predominantly in R
 with the help of RStudio. It describes basic steps that we expect to be
@@
 rely on external software or multistage data processing (for those
-requirements see section \protect\hyperlink{advanced-features}{Advanced
-Features}). Of course, your specific situation might involve additional,
+requirements see Section~\ref{advanced-features}). Of course, your specific situation might involve additional,
 more specialized steps. After completing the tutorial, you should be
@@
 The first step is to install the required software. We assume that you
-have installed R \citep[version 4.0.4]{R-base} and RStudio
-\citep[version 1.4]{rstudio} already but the tutorial will guide you in
+have installed R (\citep{R-base} version 4.0.4) and RStudio
+(\citep{rstudio} version 1.4) already but the tutorial will guide you in
 detail through the installation of other necessary software with the
-help of the R package
-\href{https://github.com/aaronpeikert/repro}{\texttt{repro}}
-\citep{R-repro}. In case you have either not installed R and RStudio or
+help of the R package {\texttt{repro}} 
+(\url{https://github.com/aaronpeikert/repro})~\citep{R-repro}. In case you have either not installed R and RStudio or
 are unsure if they are up-to-date, you might want to consult our
-installation advice in the
-\href{https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md}{Online
-Supplementary Material} that covers the installation of all software
+installation advice in the Online Supplementary Material
+(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md}) that covers the installation of all software
 necessary for this tutorial in three steps. The
-\href{(https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md)}{installation
-advice} may also help Windows users who have problems installing
+\hl{installation advice} %MDPI: Please add the accessed date.
+(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md}) may also help Windows users who have problems installing
 \texttt{Docker}.
@@
 Unfortunately, Docker requires administrator rights to run, which may
-not be available to all researchers. We recommend
-\href{https://rstudio.github.io/renv/articles/renv.html}{\texttt{renv}}
+not be available to all researchers. We recommend {\texttt{renv}}
+(\url{https://rstudio.github.io/renv/articles/renv.html})
 \citep{R-renv} in cases where no administrator rights can be obtained
-but can not detail its use in this document. \texttt{renv} tracks which
-R package is installed from which source in which version in a so-called
-\href{https://rstudio.github.io/renv/articles/lockfile.html}{\texttt{lockfile}}.
+but we can not detail its use in this document. \texttt{renv} tracks which
+R package is installed from which source in which version in a so-called \hl{\texttt{lockfile}}
+(\url{https://rstudio.github.io/renv/articles/lockfile.html}).
 This lockfile is then used to reinstall the same packages on other
 computers or later in time. For a more thorough discussion, see
-\citet{vanlissa2020worcs}.
-
-Start RStudio and install the package
-\href{https://github.com/aaronpeikert}{\texttt{repro}}\citep{R-repro}.
+Lissa~et~al.~\citep{vanlissa2020worcs}.
+
+Start RStudio and install the package \hl{\texttt{repro}}
+(\url{https://github.com/aaronpeikert})~\citep{R-repro}.
 It will assist you while you follow the tutorial.
 
+\begin{adjustwidth}{0.5cm}{0cm} 
 \begin{Shaded}
@@
 \end{Shaded}
+\end{adjustwidth}
 
@@
 notation.
-
+\vspace{+12pt}
+% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{adjustwidth}{0.5cm}{0cm} 
 \begin{Shaded}
@@
 \end{verbatim}
+\end{adjustwidth}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
 
@@
 \end{verbatim}
+\end{adjustwidth}
 
@@
 
-\hypertarget{creating-an-rstudio-project}{%
+%\hypertarget{creating-an-rstudio-project}{%
 \subsection{Creating an RStudio
-Project}\label{creating-an-rstudio-project}}
+Project}\label{creating-an-rstudio-project}
 
@@
 only special thing about the R Markdown document is the \texttt{repro}
-metadata that we will learn later about. However, you may turn any other
+metadata that we will learn about later. However, you may turn any other
 template or existing R project into a reproducible research project by
@@
 
-\hypertarget{implementing-version-control}{%
+%\hypertarget{implementing-version-control}{%
 \subsection{Implementing Version
-Control}\label{implementing-version-control}}
+Control}\label{implementing-version-control}
 
@@
 \end{Shaded}
-
-\begin{figure}
-\centering
-\includegraphics{/home/rstudio/images/modified.pdf}
-\caption{Figure 3.1: The Git pane in R Studio, showing
-\texttt{manuscript.Rmd} modified but unstaged and \texttt{modified.png}
-newly added and staged.}
-\end{figure}
+\end{adjustwidth}
 
@@
 project snapshot). One way to do this is through the visual user
-interface in the RStudio \texttt{Git} pane (see Figure 3.1). Click on
+interface in the RStudio \texttt{Git} pane (see \highlighting{Figure~\ref{figure2}).} %MDPI: Figure 3.1 has been changed into Figure 2. Please confirm.
+ Click on
 the empty box next to the file you want to stage. A checkmark then
@@
 stores a snapshot of the current state of the project.
+
+\begin{figure}[H]
+%\centering
+\includegraphics{Figures/modified.pdf}
+\caption{The Git pane in R Studio, showing
+\texttt{manuscript.Rmd} modified but unstaged and \texttt{modified.png}
+newly added and staged.}
+\label{figure2}
+\end{figure}
 
@@
 
-\hypertarget{using-dynamic-document-generation}{%
+%\hypertarget{using-dynamic-document-generation}{%
 \subsection{Using Dynamic Document
-Generation}\label{using-dynamic-document-generation}}
+Generation}\label{using-dynamic-document-generation}
 
@@
 \def\labelenumi{\arabic{enumi}.}
-\tightlist
+%\tightlist
 \item
@@
 RStudio user interface. The text of an \texttt{R\ Markdown} is formatted
-by \texttt{Markdown} (see \citep{xieMarkdownDefinitiveGuide2019} for
-technical details and \citep{xieMarkdownCookbook2020} for practical
+by \texttt{Markdown} (see~\citep{xieMarkdownDefinitiveGuide2019} for
+technical details and~\citep{xieMarkdownCookbook2020} for practical
 guidance). The code mostly consists of R code (although other
@@
 \end{Shaded}
-
-\hypertarget{manage-software-and-file-dependencies}{%
+\end{adjustwidth}
+
+%\hypertarget{manage-software-and-file-dependencies}{%
 \subsection{Manage Software and File
-Dependencies}\label{manage-software-and-file-dependencies}}
+Dependencies}\label{manage-software-and-file-dependencies}
 
@@
 
-\hypertarget{reproducing-a-project}{%
-\subsection{Reproducing a project}\label{reproducing-a-project}}
+%\hypertarget{reproducing-a-project}{%
+\subsection{Reproducing a Project}\label{reproducing-a-project}
 
@@
 
-\hypertarget{summary}{%
-\subsection{Summary}\label{summary}}
+%\hypertarget{summary}{%
+\subsection{Summary}\label{summary}
 
@@
 \setcounter{enumi}{1}
-\tightlist
+%\tightlist
+\newpage
 \item
@@
 \setcounter{enumi}{2}
-\tightlist
+%\tightlist
 \item
@@
 \setcounter{enumi}{3}
-\tightlist
+%\tightlist
 \item
@@
 \setcounter{enumi}{4}
-\tightlist
+%\tightlist
 \item
@@
 \end{verbatim}
-
-\hypertarget{advanced-features}{%
-\section{Advanced Features}\label{advanced-features}}
+\end{adjustwidth}
+
+%\hypertarget{advanced-features}
+\section{Advanced Features}\label{advanced-features}
 
@@
 complex computational recipes than rendering an \texttt{R\ Markdown} or
-require other software than R packages.
+require software other than R packages.
 
@@
 
-\hypertarget{preregistration-as-code}{%
-\section{Preregistration as Code}\label{preregistration-as-code}}
+%\hypertarget{preregistration-as-code}
+\section{Preregistration as Code}\label{preregistration-as-code}
 
@@
 questions and planning data analysis before observing the research
-outcomes \citep{NosekRevolution2018}. It serves to separate a-priori
+outcomes~\citep{NosekRevolution2018}. It serves to separate a-priori
 planned and theory-driven (confirmatory) analyses from unplanned and
@@
 (67\%) included a formal power analysis, only 34 (20\%) of these could
-be reproduced \citep{bakkerRecommendationsPreregistrationsInternal2020}.
+be reproduced~\citep{bakkerRecommendationsPreregistrationsInternal2020}.
 Even when researchers have gone to great lengths in preregistering an
 analysis script, they sometimes inexplicably fail to reproduce their own
-results. For example, \citet{steegenMeasuringCrowdAgain2014} realized
+results. For example, Steegen~et~al.~\citep{steegenMeasuringCrowdAgain2014} realized
 after publication that part of their preregistered code resulted in
@@
 the authors expect to obtain from their experiments. For guidance on how
-to simulate data, see \citet{morrisUsingSimulationStudies2019},
-\citet{paxtonMonteCarloExperiments2001}, and
-\citet{skrondalDesignAnalysisMonte2000}, as well as the R packages
-\href{https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html}{\texttt{simstudy}},
-\citep{simstudy} and
-\href{https://personality-project.org/r/psych/help/sim.html}{\texttt{psych}},
+to simulate data, see Morris~et~al.~\citep{morrisUsingSimulationStudies2019},
+Paxton~et~al.~\citep{paxtonMonteCarloExperiments2001}, and
+Skrondal~\citep{skrondalDesignAnalysisMonte2000}, as well as the R packages
+{\texttt{simstudy}} 
+\mbox{ (\url{https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html})}
+~\citep{simstudy} and
+{\texttt{psych}} (\url{https://personality-project.org/r/psych/help/sim.html})
 \citep{psych}.
@@
 builds upon four pillars (version control, dynamic document generation,
-dependency tracking, and software management). To use PAC the dangers to
+dependency tracking, and software management). To use PAC, the dangers to
 reproducibility we described must be eliminated.
 
-The idea of submitting code as part of a preregistration is not new
-\citep[e.g.][]{wichertsDegreesFreedomPlanning2016}. A prominent
-preregistration platform, \href{https://osf.io/}{The Open Science
-Framework}, suggests submitting scripts alongside the preregistration of
-methods. In an informal literature search (we skimmed the first 300
-results of Google Scholar with the keywords
-\texttt{("pre\ registration"\textbar{}"pre-registration"\textbar{}preregistration)\&(code\textbar{}script\textbar{}matlab\textbar{}python\textbar{}"R")})
-we only found close to a dozen published articles that did include some
-form of script as part of their preregistration. Though the notion of
-preregistering code has been around for a while
-\citep[cf.][]{steegenMeasuringCrowdAgain2014}, it has not gained much
-traction---perhaps because, to date, this has constituted an extra
-non-standard step in the research process. This tutorial integrates the
-preregistration of code into the reproducible research workflow by
-encouraging researchers to preregister the whole manuscript as a dynamic
-document.
-
-\hypertarget{advantages-of-pac-over-traditional-preregistration}{%
-\subsection{Advantages of PAC Over Traditional
-Preregistration}\label{advantages-of-pac-over-traditional-preregistration}}
+\textls[+15]{The idea of submitting code as part of a preregistration is not new (e.g.~\citep{wichertsDegreesFreedomPlanning2016}).
+A prominent preregistration platform, \hl{The Open Science Framework} %MDPI: Please add the accessed date.
+(\url{https://osf.io/}), suggests submitting scripts alongside the preregistration of methods.
+In an informal literature search (we skimmed the first 300 results of Google Scholar with the keywords \texttt{\textls[+35]{("pre\ registration"\textbar{}"pre-registration"\textbar{}preregistration)\&(code\textbar{}script\textbar{}matlab\textbar{}}\linebreak python\textbar{}"R")}) we only found close to a dozen published articles that did include some form of script as part of their preregistration.
+Though the notion of preregistering code has been around for a while (cf.~\citep{steegenMeasuringCrowdAgain2014}), it has not gained much traction---perhaps because, to date, this has constituted an extra non-standard step in the research process.
+This tutorial integrates the preregistration of code into the reproducible research workflow by encouraging researchers to preregister the whole manuscript as a dynamic document.}
+
+%\hypertarget{advantages-of-pac-over-traditional-preregistration}{%
+\subsection{Advantages of PAC over Traditional
+Preregistration}\label{advantages-of-pac-over-traditional-preregistration}
 
@@
 assumptions are not met, the code may branch out to employ robust
-methods; or, an analysis may perform automated variable selection
+methods; or an analysis may perform automated variable selection
 mechanisms before running the final model. Another example of
-data-dependent decisions are more explorative analyses, i.e.~explorative
+data-dependent decisions are more explorative analyses, i.e.,~explorative
 factor analysis or machine learning. Decisions that do not lend
@@
 
-\hypertarget{deviating-from-the-preregistration-and-exploration}{%
+%\hypertarget{deviating-from-the-preregistration-and-exploration}{%
 \subsection{Deviating from the Preregistration and
-Exploration}\label{deviating-from-the-preregistration-and-exploration}}
+Exploration}\label{deviating-from-the-preregistration-and-exploration}
 
@@
 retrospectively summarize and justify any changes made to the
-preregistered plan, e.g.~in the discussion of the final manuscript
-\footnote{In section {[}Preregistration as Code --- a Tutorial{]} we
+preregistered plan, e.g.~in the discussion of the \hl{final manuscript} %MDPI: Footnote is not allowed in this journal. The content has been moved here. Please confirm.
+({In section {[}Preregistration as Code---a Tutorial{]} we
   conducted an actual PAC and summarize the changes we make to the
-  preregistered code in the discussion.}. During the analysis process,
+  preregistered code in the discussion}). During the analysis process,
 authors can additionally maintain a running changelog to explain changes
@@
 reasoning behind the changes and link to the commit id that applied the
-changes. This enables readers and reviewers to inspects individual
+changes. This enables readers and reviewers to inspect individual
 changes and make an informed judgment about their validity and
@@
 deviation from a preregistration is not a problem
-\citep{nosekPreregistrationHardWorthwhile2019}; rather, a faillure to
+\citep{nosekPreregistrationHardWorthwhile2019}; rather, a failure to
 disclose such deviations is a problem. In fact, it is expected that most
@@
 requires a judgement about what changes affect the test of the theory to
-what extend \citep{meehlTheoreticalRisksTabular1978}. Only transparent
+what extend~\citep{meehlTheoreticalRisksTabular1978}. Only transparent
 reporting enables such judgment.
@@
 exploration with exploratory statistical models that are standard in
-machine learning \citep{brandmaier22mlsem}. These models often involve
+machine learning~\citep{brandmaier22mlsem}. These models often involve
 dozens, if not hundreds of predictors, which makes it difficult to
 describe them verbally. With PAC, such models can be preregistered
-clearly and in comprehensive detail and researcher can precisely define
+clearly and in comprehensive detail, and the researcher can precisely define
 a priori how much they want to explore. We specifically recommend PAC
@@
 might be overfit
-\citep[p.~220f.]{hastieElementsStatisticalLearning2017}.
-
-\hypertarget{planned-analyses-as-functions}{%
+(\citep{hastieElementsStatisticalLearning2017}, p.~220f.).
+
+%\hypertarget{planned-analyses-as-functions}{%
 \subsection{Planned Analyses as
-Functions}\label{planned-analyses-as-functions}}
+Functions}\label{planned-analyses-as-functions}
 
@@
 planned hypothesis: one to conduct the planned analysis, one to simulate
-the expected data, and to report the results. Using functions makes the
+the expected data, and one to report the results. Using functions makes the
 analysis more portable (i.e., it can easily be used for other datasets),
@@
 study. The functions shown here do not contain executable code, but the
-interested reader can find working functions in the
-\href{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R}{online
-supplementary materials} that power the example below.
+interested reader can find working functions in the \hl{online
+supplementary materials}% MDPI: Please add the accessed date.
+~(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R}) that power the example below.
 
@@
 
+\vspace{+6pt}
+% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{adjustwidth}{0.5cm}{0cm} 
 \begin{Shaded}
@@
 \end{Shaded}
+\end{adjustwidth}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
 
@@
 
+\newpage
+% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{adjustwidth}{0.5cm}{0cm} 
 \begin{Shaded}
@@
 \end{Shaded}
-
-This function should again accept the output of
+\end{adjustwidth}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+
+\textls[-5]{This function should again accept the output of
 \texttt{planned\_analysis()} as input. The output of this function
@@
 tables of estimates or figures directly from R goal to help with this
-objective. Packages like
-\href{https://github.com/Rapporter/pander}{\texttt{pander}}
-\citep{pander},
-\href{https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf}{\texttt{stargazer}}
-\citep{stargazer},
-\href{https://dstanley4.github.io/apaTables/articles/apaTables.html}{\texttt{apaTables}}
-\citep{apaTables} and
-\href{http://frederikaust.com/papaja_man/}{\texttt{papaja}}
+objective. Packages like {\texttt{pander}}
+(\url{https://github.com/Rapporter/pander})
+\citep{pander}, {\texttt{stargazer}}
+\mbox{(\url{https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf})}
+\citep{stargazer}, {\texttt{apaTables}}
+(\url{https://dstanley4.github.io/apaTables/articles/apaTables.html})
+\citep{apaTables} and {\texttt{papaja}}
+\mbox{(\url{http://frederikaust.com/papaja_man/})}
 \citep{papaja} help you to create dynamically generated professional
-looking results. The package
-\href{https://github.com/easystats/report}{\texttt{report}}
-\citep{report} is particularly noteworthy because it not only generates
+looking results. The package {\texttt{report}}
+(\url{https://github.com/easystats/report})~\citep{report} is particularly noteworthy because it not only generates
 tables but also a straightforward interpretation of the effects as
-actual prose (e.g., it verbally quantifies the size of an effect).
+actual prose (e.g., it verbally quantifies the size of an effect).}
 
@@
 
-\hypertarget{turning-a-dynamic-document-into-a-preregistration}{%
-\subsubsection{Turning a Dynamic Document into a
-Preregistration}\label{turning-a-dynamic-document-into-a-preregistration}}
+%\hypertarget{turning-a-dynamic-document-into-a-preregistration}{%
+\subsubsection*{Turning a Dynamic Document into a
+Preregistration}\label{turning-a-dynamic-document-into-a-preregistration}
 
@@
 should archive a time-stamped and uneditable version of the project that
-serves as the preregistration. zenodo.org \citep{zenodo} is a publicly
+serves as the preregistration. zenodo.org~\citep{zenodo} is a publicly
 funded service provider that archives digital artefacts for research and
@@
 \begin{quote}
-Navigate to \url{https://zenodo.org/login/} → Log in with GitHub
+\hl{Navigate to} %MDPI: Please add the accessed date.
+~\url{https://zenodo.org/login/} → Log in with GitHub
 \end{quote}
@@
 \begin{quote}
-Log into zenodo.org → Account → GitHub\footnote{\url{https://zenodo.org/account/settings/github/}}
+Log into zenodo.org → Account → \hl{GitHub} %MDPI: 1. Footnote is not allowed in this journal. The content has been moved here. Please confirm. 2. Please add the accessed date
+~({\url{https://zenodo.org/account/settings/github/}})
 \end{quote}
@@
 \begin{quote}
-Navigate to \url{https://zenodo.org/account/settings/github/}
+\hl{Navigate to} %MDPI: Please add the accessed date.
+ \url{https://zenodo.org/account/settings/github/}
 \end{quote}
@@
 
-\hypertarget{alternatives-to-simulated-data}{%
-\subsection{Alternatives to simulated
-data}\label{alternatives-to-simulated-data}}
+%\hypertarget{alternatives-to-simulated-data}{%
+\subsection{Alternatives to Simulated
+Data}\label{alternatives-to-simulated-data}
 
@@
 mean in variable ``Z''). Preregistration after data collection is common
-for secondary data analysis of data obtained by other research groups
-\citep{westonRecommendationsIncreasingTransparency2019} but not so much
+for secondary data analysis of data obtained by other research groups~\citep{westonRecommendationsIncreasingTransparency2019} but not so much
 within the same research project. We argue that it is still an eligible
@@
 study planning, since it lets the researcher evaluate the feasibility of
-many assumptions. However, we must warn our readers, that while piloting
+many assumptions. However, we must warn our readers that while piloting
 is more traditional than our approach of blinding the data before
@@
 
-\hypertarget{when-is-pac-applicable}{%
-\subsection{When Is PAC Applicable?}\label{when-is-pac-applicable}}
+%\hypertarget{when-is-pac-applicable}{%
+\subsection{When Is PAC Applicable?}\label{when-is-pac-applicable}
 
@@
 preregistrations are particularly amenable to PAC. First, pregistrations
-of clinical trials (called statistical analysis plans, \citet{ICH1998})
+of clinical trials (called statistical analysis plans, International Council for Harmonisation of Technical Requirements for Registration of
+Pharmaceuticals for Human Use~\citep{ICH1998})
 typically describe analyses in exhaustive detail and typically contain a
 detailed description of how results will be presented, including shells
-of tables and graphics \citep{yuanGuideStatisticalAnalysis2019}. PAC may
+of tables and graphics~\citep{yuanGuideStatisticalAnalysis2019}. PAC may
 significantly reduce the required workload while maintaining (and
@@
 implementation. PAC, however, captures the precise algorithmic model,
-including its software implementation and is ideal for preregistering
-these models \citep{brandmaier22mlsem}.
-
-\hypertarget{preregistration-as-code-tutorial}{%
+including its software implementation, and is ideal for preregistering
+these models~\citep{brandmaier22mlsem}.
+
+%\hypertarget{preregistration-as-code-tutorial}{%
 \subsection{Preregistration as Code:
-Tutorial}\label{preregistration-as-code-tutorial}}
+Tutorial}\label{preregistration-as-code-tutorial}
 
@@
 preregistration and have outlined its implementation. To illustrate how
-PAC works in practice and help researchers to implement PAC themselves,
+PAC works in practice and to help researchers implement PAC themselves,
 we provide a worked example. We will use an exemplary research question
@@
 eventually becomes the final report. The complete preregistration source
-is available in the
-\href{https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd}{online
-supplementary material}. In this section, we show code excerpts of this
+is available in the \hl{online supplementary material}%MDPI: Please add the accessed date.
+~(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd}). In this section, we show code excerpts of this
 preregistration (formatted in monospace) and explain the rationale
@@
 \end{Shaded}
+\end{adjustwidth}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
 
 We choose the following statistical procedure because many researchers
-are familiar with it\footnote{The t-test and Mann-Whitney-Wilcoxon test
+are familiar \hl{with it} %MDPI: Footnote is not allowed. The content has been moved here. Please confirm.
+~(The \emph{t}-test and Mann-Whitney-Wilcoxon test
   are arguably the most often used hypothesis tests (according to
-  \citep{fagerlandTtestsNonparametricTests2012, hortonStatisticalMethodsJournal2005}
-  reports that 26\% of all studies employed a t-test and 27\% employed a
+ ~\citep{fagerlandTtestsNonparametricTests2012, hortonStatisticalMethodsJournal2005}
+  reports that 26\% of all studies employed a \emph{t}-test and 27\% employed a
   rank-based alternative in the \emph{New England Journal of Medicine}
@@
   invariance is untested
-  \citep{putnickMeasurementInvarianceConventions2016}, the effect size
+ ~\citep{putnickMeasurementInvarianceConventions2016}, the effect size
   is underestimated in the presence of measurement error
-  \citep{frostCorrectingRegressionDilution2000}, the effect size is
+ ~\citep{frostCorrectingRegressionDilution2000}, the effect size is
   overestimated for highly skewed distributions
-  \citep{stonehouseRobustnessTestsCombined1998}). The interested reader
-  can use the
-  \href{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R}{provided
-  code for the simulation} to verify that the t-test provides unbiased
+ ~\citep{stonehouseRobustnessTestsCombined1998}). The interested reader
+  can use the \hl{provided code for the simulation}
+~\mbox{(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R})} to verify that the \emph{t}-test provides unbiased
   effect sizes but the Mann-Whitney-Wilcoxon overestimates effect sizes
-  with increasing sample size and skewness.}:
-
+  with increasing sample size and skewness.):
+
+\vspace{+6pt}
+% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{adjustwidth}{0.5cm}{0cm} 
 \begin{Shaded}
@@
 correction by default and hence is robust even with unequal variances
-across groups \citep{zimmermanRankTransformationsPower1993}. Mentioning
+across groups~\citep{zimmermanRankTransformationsPower1993}. Mentioning
 every such minute implementation detail is almost impossible and would
@@
 Together with the function \texttt{simulate\_data()} (not shown here),
-the function \texttt{planned\_analysis()} can be used to justify the
+the function \texttt{planned\_\linebreak analysis()} can be used to justify the
 planned sample size. To that end, \texttt{simulate\_data()} is
@@
 significant results (power) is recorded. The results for such a Monte
-Carlo simulation for this example are visualized in Figure 5.1. The code
-for this power analysis can be found in the
-\href{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R}{online
-supplementary material}. The next snippet shows how we integrated the
+Carlo simulation for this example are visualized in \highlighting{Figure~\ref{figure3}.} %MDPI:  Figure 5.1 has been changed into Figure 3. Please confirm.
+ The code
+for this power analysis can be found in the \hl{online supplementary material}
+(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R}). The next snippet shows how we integrated the
 results dynamically into the preregistration (the origin of the
@@
 
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/power-1.pdf}
-\caption{Figure 5.1: Results of simulation for the power analysis. The
-cross indicates the sample size that archives 80\% assuming a Cohen's d
-of 0.2.}
-\end{figure}
-
+\vspace{+6pt}
+% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{adjustwidth}{0.5cm}{0cm} 
 \begin{Shaded}
@@
 
+\begin{figure}[H]
+%\widefigure
+\includegraphics[width=12.5cm]{Figures/power-1.pdf}
+\caption{Results of simulation for the power analysis. The
+cross indicates the sample size that archives 80\% assuming a Cohen's d
+of 0.2.}
+\label{figure3}
+\end{figure}
+
+\vspace{+6pt}
+% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{adjustwidth}{0.5cm}{0cm} 
 \begin{Shaded}
@@
 the exact format the data would have. This means that we did need to
-change our preregistration after accessing the data to include i.e.~the
-recoding of gender (lines 17-22) and the items (lines 23-26).\footnote{We
+change our preregistration after accessing the data to include i.e.,~the
+recoding of gender (lines 17--22) and the items\hl{ (lines 23--26).} %MDPI: Footnote is not allowed. The content has been moved here. Please confirm.
+ ({We
   invite the reader to evaluate the changes we made to the preregistered
-  code. Either on
-  \href{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg...main}{GitHub}
+  code. Either on \hl{GitHub}
+  (\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg...main})
   → ``Files changed'' or directly in Git with
-  \texttt{git\ diff\ v0.0.1.1-prereg\ preregistration.Rmd}.} This is our
+  \texttt{git\ diff\ v0.0.1.1-prereg\ preregistration.Rmd}.}) This is our
 summary of what we changed:
 
+\vspace{+6pt}
+% start a new page without indent 4.6cm
+%\clearpage
+\end{paracol}
+\nointerlineskip
+\begin{adjustwidth}{0.5cm}{0cm} 
 \begin{Shaded}
@@
 \end{Shaded}
-
-Readers can inspect and judge the changes for themselves on
-\href{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main\#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b}{GitHub}.
+\end{adjustwidth}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
+
+Readers can inspect and judge the changes for themselves on \hl{GitHub}
+\textls[-25]{(\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main\#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b})}.
 
@@
 effect is - negative, statistically significant, and small (difference = -0.17,
-95% CI [0.12, 0.22], t(887.46) = 6.38, p < .001; Cohen's d = 0.43, 95% CI [0.30,
+95% CI [0.12, 0.22], t(887.46) = 6.38, p < 0.001; Cohen's d = 0.43, 95% CI [0.30,
 0.56])
 \end{verbatim}
+\end{adjustwidth}
+\begin{paracol}{2}
+%\linenumbers
+\switchcolumn
 
@@
 
-The example rendered as a
-\href{https://github.com/aaronpeikert/repro-tutorial/files/7309455/preregistration.pdf}{PDF
-file with real data} is available in the
-\href{https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.3.1-results}{online
-supplementary material}. The changes we made since preregistering it can
-be inspected on this
-\href{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main\#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b}{GitHub
-page}.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+The example rendered as a \hl{PDF file with real data}
+(\url{https://github.com/aaronpeikert/repro-tutorial/files/7309455/preregistration.pdf}) is available in the \hl{online supplementary material}
+(\url{https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.3.1-results}). The changes we made since preregistering it can
+be inspected on this \hl{GitHub page}
+(\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main\#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b}).
+
+%\hypertarget{discussion}
+\section{Discussion}\label{discussion}
 
 Increased automation is increasingly recognized as a means to improve
-the research process \citep{rouderMinimizingMistakesPsychological2019},
+the research process~\citep{rouderMinimizingMistakesPsychological2019},
 and therefore this workflow fits nicely together with other innovations
@@
 \citep{lakensImprovingTransparencyFalsifiability2021} or automated data
-documentation \citep{arslanHowAutomaticallyDocument2019}. Automated
+documentation~\citep{arslanHowAutomaticallyDocument2019}. Automated
 research projects promise a wide range of applications, among them PAC
@@
 report][]{nosekRegisteredReports2014, chambersWhatNextRegistered2019},
-direct replication \citep{simonsValueDirectReplication2014}, fully
-automated living metanalysis \citep{elliottLivingSystematicReviews2014},
+direct replication~\citep{simonsValueDirectReplication2014}, fully
+automated living metanalysis~\citep{elliottLivingSystematicReviews2014},
 executable research articles
@@
 
-\hypertarget{summary-1}{%
-\subsection{Summary}\label{summary-1}}
+%\hypertarget{summary-1}{%
+\subsection{Summary}\label{summary-1}
 
@@
 four software solutions, (1) version control, (2) dynamic document
-generation, (3) dependency tracking and (4) software management to
+generation, (3) dependency tracking, and (4) software management to
 guarantee reproducibility. We first demonstrated how to create a
@@
 
-\hypertarget{limitations}{%
-\subsection{Limitations}\label{limitations}}
+%\hypertarget{limitations}{%
+\subsection{Limitations}\label{limitations}
 
@@
 reproducibility of their research. Two kinds of limitations can be
-distinguished. The first kind are limitations by design which are
+distinguished. The first kind are limitations by design, which are
 unlikely to change. Our workflow inherits these from the software it
@@
 for automation, many researchers will have to search for alternative
-software.\\
+software.
+
 The automation-friendly software we present here has several technical
@@
 for scientific data. The necessary steps for archival on zenodo.org are
-described at the end of Section {[}Preregistration as Code --- a
+described at the end of Section {[}Preregistration as Code---a
 Tutorial{]}.
@@
 manually trace the dependencies and execute commands in the right order,
-in case they are unable to run \texttt{Make} for some reason. Thus
+in case they are unable to run \texttt{Make} for some reason. Thus,
 hypothetically, even if \texttt{Docker} were to become unavailable one
 day, the \texttt{Dockerfile} still serves as unambiguous documentation
-of how the original system was set up and may help future users to
+of how the original system was set up, and may help future users to
 create a software environment that closely resembles the original.
 
-\hypertarget{outlook}{%
-\subsection{Outlook}\label{outlook}}
+%\hypertarget{outlook}{%
+\subsection{Outlook}\label{outlook}
 
@@
 be complemented or replaced by a dynamic Microsoft Word document with
-the help of \texttt{officer} \citep{officer} or \texttt{officedown}
+the help of \texttt{officer}~\citep{officer} or \texttt{officedown}
 \citep{officedown} to accommodate a wider range of journal submission
@@
 R-centered alternative \texttt{targets} for more convenience.
-\texttt{Docker} could be combined with \texttt{renv} \citep{R-renv} to
+\texttt{Docker} could be combined with \texttt{renv}~\citep{R-renv} to
 control the package versions precisely (our approach fixes the date,
@@
 
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% %% optional
-% \supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
-%
-% % Only for the journal Methods and Protocols:
-% % If you wish to submit a video article, please do so with any other supplementary material.
-% % \supplementary{The following are available at www.mdpi.com/link: Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
-
 \vspace{6pt}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Supplementary Materials{\hl{\quad \quad}}%The following are available online at www.mdpi.com/xxx/s1, Figure S1: title, Table S1: title, Video S1: title.
+
+\authorcontributions{A.P. took the lead in writing and provided
+the initial draft of the manuscript.
+A.M.B.
+and C.J.v.L.
+contributed further ideas, critical feedback, and revisions of the original manuscript.
+All authors have read and agreed to the published version of the manuscript.}
+
+\funding{\hl{\quad \quad}} %This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
+Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
+
+\institutionalreview{\hl{\quad \quad}} %In this section, please add the Institutional Review Board Statement and approval number for studies involving humans or animals.
+Please note that the Editorial Office might ask you for further information.
+Please add ``The study was conducted according to the guidelines of the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).'' OR ``Ethical review and approval were waived for this study, due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans or animals.
+You might also choose to exclude this statement if the study did not involve humans or animals.
+
+\informedconsent{\hl{\quad \quad}} %Any research article describing a study involving humans should contain this statement.
+Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
+You might also choose to exclude this statement if the study did not involve humans.
+Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
+Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.
+
+\dataavailability{\hl{\quad \quad}} %In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
+Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.
+You might choose to exclude this statement if the study did not report any data.
 
@@
 Julia Delius for her helpful assistance in language and style editing.
-The R package \texttt{repro} was developed as part of the
-\href{https://aaronpeikert.github.io/repro-thesis/}{first author's
-master thesis} at the Humboldt-Universität zu Berlin.}
+The R package \texttt{repro} was developed as part of the \hl{first author's
+master thesis}
+\mbox{(\url{https://aaronpeikert.github.io/repro-thesis/})} at the Humboldt-Universität zu Berlin.}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Aaron Peikert took the lead in writing and provided
-the initial draft of the manuscript. Andreas Brandmaier and Caspar J.
-Van Lissa contributed further ideas, critical feedback, and revisions of
-the original manuscript.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{The authors declare no conflict of interest. We
+\conflictsofinterest{The authors declare no conflicts of interest. We
 have received no financial support for the research, authorship, and/or
@@
 %% optional
-\abbreviations{The following abbreviations are used in this manuscript:\\
+\abbreviations{Abbreviations}{The following abbreviations are used in this manuscript:\\
 
@@
 \end{tabular}}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
-
-%=====================================
-% References, variant A: internal bibliography
-%=====================================
-%\reftitle{References}
-%\begin{thebibliography}{999}
+\end{paracol}
+\reftitle{References}
+%\externalbibliography{yes}
+%\bibliography{temp.bib}
+
+\begin{thebibliography}{999}
 % Reference 1
-%\bibitem[Author1(year)]{ref-journal}
-%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
+\bibitem[Peikert and Brandmaier(2021)]{Peikert2019}
+Peikert, A.; Brandmaier, A.M.
+A Reproducible Data Analysis Workflow with R Markdown, Git, Make, and Docker.
+\emph{Quant.
+Comput.
+Methods Behav.
+Sci.} \textbf{2021}, \emph{1}, e3763.
+https://doi.org/10.5964/qcmb.3763.
+
 % Reference 2
-%\bibitem[Author2(year)]{ref-book}
-%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
-%\end{thebibliography}
-
-% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
-For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-%=====================================
-% References, variant B: external bibliography
-%=====================================
-\reftitle{References}
-\externalbibliography{yes}
-\bibliography{temp.bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibitem[Popper(2002)]{popperLogicScientificDiscovery2002}
+Popper, K.R. \emph{The Logic of Scientific Discovery}; Routledge: London, UK, 2002.
+
+% Reference 3
+\bibitem[Obels \em{et~al.}(2020)Obels, Lakens, Coles, Gottfried, and
+  Green]{obels2020}
+Obels, P.; Lakens, D.; Coles, N.A.; Gottfried, J.; Green, S.A.
+Analysis of Open Data and Computational Reproducibility in Registered Reports in Psychology.
+\emph{Adv.
+Methods Pract.
+Psychol.
+Sci.} \textbf{2020}, \emph{3}, 229--237.
+
+% Reference 4
+\bibitem[van Lissa \em{et~al.}(2021)van Lissa, Brandmaier, Brinkman, Lamprecht,
+  Peikert, Struiksma, and Vreede]{vanlissa2020worcs}
+van Lissa, C.J.; Brandmaier, A.M.; Brinkman, L.; Lamprecht, A.L.; Peikert, A.; Struiksma, M.E.; Vreede, B.M.
+WORCS: A Workflow for Open Reproducible Code in Science.
+\emph{Data Sci.} \textbf{2021}, \emph{4}, 29--49.
+https://doi.org/10.3233/DS-210031.
+
+% Reference 5
+\bibitem[Nosek \em{et~al.}(2018)Nosek, Ebersole, DeHaven, and
+  Mellor]{NosekRevolution2018}
+Nosek, B.A.; Ebersole, C.R.; DeHaven, A.C.; Mellor, D.T.
+The Preregistration Revolution.
+\emph{Proc.
+Natl.
+Acad.
+Sci.
+USA} \textbf{2018}, \emph{115}, 2600--2606.
+https://doi.org/10.1073/pnas.1708274114.
+
+% Reference 6
+\bibitem[Hardwicke \em{et~al.}(2018)Hardwicke, Mathur, MacDonald, Nilsonne,
+  Banks, Kidwell, Hofelich~Mohr, Clayton, Yoon, Henry~Tessler, Lenne, Altman,
+  Long, and Frank]{hardwicke2018}
+Hardwicke, T.E.; Mathur, M.B.; MacDonald, K.; Nilsonne, G.; Banks, G.C.; Kidwell, M.C.; Hofelich Mohr, A.; Clayton, E.; Yoon, E.J.; Henry Tessler, M.; et al.
+Data Availability, Reusability, and Analytic Reproducibility: Evaluating the Impact of a Mandatory Open Data Policy at the Journal Cognition.
+\emph{R.
+Soc.
+Open Sci.} \textbf{2018}, \emph{5}, 180448.
+https://doi.org/10.1098/rsos.180448.
+
+% Reference 7
+\bibitem[{R Core Team}(2021)]{R-base}
+R Core Team. \emph{R: A Language and Environment for Statistical Computing}; R Foundation for Statistical Computing: Vienna, Austria, 2021.
+
+% Reference 8
+\bibitem[Vuorre and Curley(2018)]{vuorreCuratingResearchAssets2018}
+Vuorre, M.; Curley, J.P.
+Curating Research Assets: A Tutorial on the Git Version Control System.
+\emph{Adv.
+Methods Pract.
+Psychol.
+Sci.} \textbf{2018}, \emph{1}, 219--236.
+https://doi.org/10.1177/2515245918754826.
+
+% Reference 9
+\bibitem[Bryan(2018)]{bryanExcuseMeYou2018}
+Bryan, J. Excuse Me, Do You Have a Moment to Talk About Version Control? \emph{Am. Stat.} \textbf{2018}, \emph{72}, 20--27. https://doi.org/10.1080/\linebreak00031305.2017.1399928.
+
+% Reference 10
+\bibitem[Nuijten \em{et~al.}(2016)Nuijten, Hartgerink, {van Assen}, Epskamp,
+  and Wicherts]{nuijtenPrevalenceStatisticalReporting2016}
+Nuijten, M.B.; Hartgerink, C.H.J.; van Assen, M.A.L.M.; Epskamp, S.; Wicherts, J.M.
+The Prevalence of Statistical Reporting Errors in Psychology (1985--2013).
+\emph{Behav.
+Res.
+Methods} \textbf{2016}, \emph{48}, 1205--1226.
+https://doi.org/10.3758/s13428-015-0664-2.
+
+% Reference 11
+\bibitem[Knuth and Levy(1994)]{knuthCWEBSystemStructured}
+Knuth, D.E.; Levy, S. \emph{The CWEB System of Structured Documentation}; Addison-Wesley Longman: \hl{Boston, USA, 1994.}%MDPI: Please add the publisher’s location.
+
+% Reference 12
+\bibitem[Claerbout and Karrenbach(1992)]{claerboutElectronicDocumentsGive1992}
+Claerbout, J.F.; Karrenbach, M. Electronic Documents Give Reproducible Research a New Meaning. \emph{SEG Tech. Program Expand. Abstr.} \textbf{1992}, 601--604. https://doi.org/10.1190/1.1822162.
+
+% Reference 13
+\bibitem[Lamport(1994)]{lamportLATEXDocumentPreparation1994}
+Lamport, L. \emph{LATEX: A Document Preparation System: User’s Guide and Reference Manual}, 2nd ed.; Addison-Wesley: Reading, MA, USA, 1994.
+
+% Reference 14
+\bibitem[Allaire \em{et~al.}(2021)Allaire, Xie, {R Foundation}, Wickham,
+  {Journal of Statistical Software}, Vaidyanathan, {Association for Computing
+  Machinery}, Boettiger, {Elsevier}, Broman, Mueller, Quast, Pruim, Marwick,
+  Wickham, Keyes, Yu, Emaasit, Onkelinx, Gasparini, Desautels, Leutnant,
+  {MDPI}, {Taylor and Francis}, Öğreden, Hance, Nüst, Uvesten, Campitelli,
+  Muschelli, Hayes, Kamvar, Ross, Cannoodt, Luguern, Kaplan, Kreutzer, Wang,
+  Hesselberth, and Dervieux]{R-rticles}
+Allaire, J.; Xie, Y.; R Foundation; Wickham, H.; Journal of Statistical Software.; Vaidyanathan, R.; Association for Computing Machinery; Boettiger, C.; Elsevier; Broman, K.; et al.
+\emph{Articles: Article Formats for R Markdown}; R Package Version 0.19; \hl{2021.}%MDPI: Please add the publisher and its location.
+
+% Reference 15
+\bibitem[El~Hattab and Allaire(2017)]{revealjs}
+El Hattab, H.; Allaire, J. Revealjs: R Markdown Format for ’Reveal, Js’ Presentations, \hl{2017.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 16
+\bibitem[{O'Hara-Wild} and Hyndman(2021)]{vitae}
+O’Hara-Wild, M.; Hyndman, R. Vitae: Curriculum Vitae for r Markdown, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 17
+\bibitem[Xie \em{et~al.}(2020)Xie, Dervieux, and
+  Riederer]{xieMarkdownCookbook2020}
+Xie, Y.; Dervieux, C.; Riederer, E. \emph{R Markdown Cookbook}, 1st ed.; The R Series; Taylor and Francis, CRC Press: Boca Raton, FL, USA, 2020.
+
+% Reference 18
+\bibitem[Silver(2017)]{silverSoftwareSimplified2017}
+Silver, A. Software Simplified. \emph{Nature} \textbf{2017}, \emph{546}, 173--174.
+
+% Reference 19
+\bibitem[Boettiger and
+  Eddelbuettel(2017)]{boettigerIntroductionRockerDocker2017}
+Boettiger, C.; Eddelbuettel, D. An Introduction to Rocker: Docker Containers for R. \emph{R J.} \textbf{2017}, \emph{9}, 527. https://doi.org/10.32614/RJ-2017-065.
+
+% Reference 20
+\bibitem[Wickham \em{et~al.}(2019)Wickham, Averick, Bryan, Chang, McGowan,
+  Fran{\c c}ois, Grolemund, Hayes, Henry, Hester, Kuhn, Pedersen, Miller,
+  Bache, M{\"u}ller, Ooms, Robinson, Seidel, Spinu, Takahashi, Vaughan, Wilke,
+  Woo, and Yutani]{tidyverse}
+Wickham, H.; Averick, M.; Bryan, J.; Chang, W.; McGowan, L.D.; François, R.; Grolemund, G.; Hayes, A.; Henry, L.; Hester, J.; et al.
+Welcome to the tidyverse.
+\emph{J.
+Open Source Softw.} \textbf{2019}, \emph{4}, 1686.
+https://doi.org/10.21105/joss.01686.
+
+% Reference 21
+\bibitem[Wiebels and Moreau(2021)]{wiebelsLeveragingContainersReproducible2021}
+Wiebels, K.; Moreau, D.
+Leveraging Containers for Reproducible Psychological Research.
+Advances in Methods and Practices in Psychological.
+\emph{Science} \textbf{2021}, \emph{4}, https://doi.org/10.1177/25152459211017853.
+
+% Reference 22
+\bibitem[Ushey(2021)]{R-renv}
+Ushey, K. Renv: Project Environments, 2021. \hl{R Package Version 0.13.2.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 23
+\bibitem[Wickham and Bryan(2021)]{usethis}
+Wickham, H.; Bryan, J. Usethis: Automate Package and Project Setup, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 24
+\bibitem[Parasuraman and
+  Mouloua(2019)]{parasuramanAutomationHumanPerformance2018}
+Parasuraman, R.; Mouloua, M. \emph{Automation and Human Performance: Theory and Applications}, 1st ed.; CRC Press: Boca Raton, FL, USA, 2019.
+
+% Reference 25
+\bibitem[{RStudio Team}(2021)]{rstudio}
+RStudio Team. \emph{RStudio: Integrated Development Environment for r. RStudio}; PBC: Boston, MA, USA, 2021.
+
+% Reference 26
+\bibitem[Peikert \em{et~al.}(2021)Peikert, Brandmaier, and {van
+  Lissa}]{R-repro}
+Peikert, A.; Brandmaier, A.M.; van Lissa, C.J.
+Repro: Automated Setup of Reproducible Workflows and Their Dependencies, 2021.
+\hl{R Package Version 0.1.0.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 27
+\bibitem[Xie \em{et~al.}(2019)Xie, Allaire, and
+  Grolemund]{xieMarkdownDefinitiveGuide2019}
+Xie, Y.; Allaire, J.J.; Grolemund, G. \emph{R Markdown: The Definitive Guide}; CRC Press: Boca Raton, FL, USA, 2019.
+
+% Reference 28
+\bibitem[Aust and Barth(2020)]{papaja}
+Aust, F.; Barth, M. Papaja: Create APA Manuscripts with R Markdown, \hl{2020.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 29
+\bibitem[Association(2019)]{apa7}
+Association, A.P. (Ed.) \emph{Publication Manual of the American Psychological Association}, 7th ed.; American Psychological Association: Washington, DC, USA, 2019.
+
+% Reference 30
+\bibitem[DeCoster \em{et~al.}(2015)DeCoster, Sparks, Sparks, Sparks, and
+  Sparks]{decosterOpportunisticBiasesTheir2015}
+DeCoster, J.; Sparks, E.A.; Sparks, J.C.; Sparks, G.G.; Sparks, C.W.
+Opportunistic Biases: Their Origins, Effects, and an Integrated Solution.
+\emph{Am.
+Psychol.} \textbf{2015}, \emph{70}, 499–514.
+https://doi.org/10.1037/a0039191.
+
+% Reference 31
+\bibitem[Silberzahn \em{et~al.}(2018)Silberzahn, Uhlmann, Martin, Anselmi,
+  Aust, Awtrey, Bahnik, Bai, Bannard, Bonnier, Carlsson, Cheung, Christensen,
+  Clay, Craig, Dalla~Rosa, Dam, Evans, Flores~Cervantes, Fong, {Gamez-Djokic},
+  Glenz, {Gordon-McKeon}, Heaton, Hederos, Heene, Hofelich~Mohr, H{\"o}gden,
+  Hui, Johannesson, Kalodimos, Kaszubowski, Kennedy, Lei, Lindsay, Liverani,
+  Madan, Molden, Molleman, Morey, Mulder, Nijstad, Pope, Pope, Prenoveau, Rink,
+  Robusto, Roderique, Sandberg, Schl{\"u}ter, Sch{\"o}nbrodt, Sherman, Sommer,
+  Sotak, Spain, Sp{\"o}rlein, Stafford, Stefanutti, Tauber, Ullrich, Vianello,
+  Wagenmakers, Witkowiak, Yoon, and Nosek]{silberzahnManyAnalystsOne2018}
+Silberzahn, R.; Uhlmann, E.L.; Martin, D.P.; Anselmi, P.; Aust, F.; Awtrey, E.; Bahnik, S.; Bai, F.; Bannard, C.; Bonnier, E.; et al.
+Many Analysts, One Data Set: Making Transparent How Variations in Analytic Choices Affect Results.
+\emph{Adv.
+Methods Pract.
+Psychol.
+Sci.} \textbf{2018}, \emph{1}, 337--356.
+https://doi.org/10.1177/2515245917747646.
+
+% Reference 32
+\bibitem[Bowman \em{et~al.}(2020)Bowman, DeHaven, Errington, Hardwicke, Mellor,
+  Nosek, and Soderberg]{bowmanOSFPreregTemplate2020}
+Bowman, S.; DeHaven, A.; Errington, T.; Hardwicke, T.E.; Mellor, D.T.; Nosek, B.A.; Soderberg, C.K. OSF Prereg Template 2020. \emph{MetaArXiv}, \textbf{2020}, https://doi.org/10.31222/osf.io/epgjd.
+
+% Reference 33
+\bibitem[Bakker \em{et~al.}(2020{\natexlab{a}})Bakker, Veldkamp, {van Assen},
+  Crompvoets, Ong, Nosek, Soderberg, Mellor, and
+  Wicherts]{bakkerEnsuringQualitySpecificity2020}
+Bakker, M.; Veldkamp, C.L.S.; van Assen, M.A.L.M.; Crompvoets, E.A.V.; Ong, H.H.; Nosek, B.A.; Soderberg, C.K.; Mellor, D.; Wicherts, J.M.
+Ensuring the Quality and Specificity of Preregistrations.
+\emph{PLoS Biol.} \textbf{2020}, \emph{18}, e3000937.
+https://doi.org/10.1371/journal.pbio.3000937.
+
+% Reference 34
+\bibitem[Bakker \em{et~al.}(2020{\natexlab{b}})Bakker, Veldkamp, {van den
+  Akker}, {van Assen}, Crompvoets, Ong, and
+  Wicherts]{bakkerRecommendationsPreregistrationsInternal2020}
+Bakker, M.; Veldkamp, C.L.S.; van den Akker, O.R.; van Assen, M.A.L.M.; Crompvoets, E.; Ong, H.H.; Wicherts, J.M.
+Recommendations in Pre-Registrations and Internal Review Board Proposals Promote Formal Power Analyses but Do Not Increase Sample Size.
+\emph{PLoS ONE} \textbf{2020}, \emph{15}, e0236079.
+https://doi.org/10.1371/journal.pone.0236079.
+
+% Reference 35
+\bibitem[Steegen \em{et~al.}(2014)Steegen, Dewitte, Tuerlinckx, and
+  Vanpaemel]{steegenMeasuringCrowdAgain2014}
+Steegen, S.; Dewitte, L.; Tuerlinckx, F.; Vanpaemel, W.
+Measuring the Crowd within Again: A Pre-Registered Replication Study.
+\emph{Front.
+Psychol.} \textbf{2014}, \emph{5}.
+https://doi.org/10.3389/fpsyg.2014.00786.
+
+% Reference 36
+\bibitem[Morris \em{et~al.}(2019)Morris, White, and
+  Crowther]{morrisUsingSimulationStudies2019}
+Morris, T.P.; White, I.R.; Crowther, M.J. Using Simulation Studies to Evaluate Statistical Methods. \emph{Stat. Med.} \textbf{2019}, \emph{38}, 2074--2102. https://doi.org/10.1002/sim.8086.
+
+% Reference 37
+\bibitem[Paxton \em{et~al.}(2001)Paxton, Curran, Bollen, Kirby, and
+  Chen]{paxtonMonteCarloExperiments2001}
+Paxton, P.; Curran, P.J.; Bollen, K.A.; Kirby, J.; Chen, F.
+Monte Carlo Experiments: Design and Implementation.
+\emph{Struct.
+Equ.
+Modeling: A Multidiscip.
+J.} \textbf{2001}, \emph{8}, 287--312.
+https://doi.org/10.1207/S15328007SEM0802\_7.
+
+% Reference 38
+\bibitem[Skrondal(2000)]{skrondalDesignAnalysisMonte2000}
+Skrondal, A.
+Design and Analysis of Monte Carlo Experiments: Attacking the Conventional Wisdom.
+\emph{Multivar.
+Behav.
+Res.} \textbf{2000}, \emph{35}, 137--167.
+https://doi.org/10.1207/S15327906MBR3502\_1.
+
+% Reference 39
+\bibitem[Goldfeld and {Wujciak-Jens}(2020)]{simstudy}
+Goldfeld, K.; Wujciak-Jens, J. Simstudy: Illuminating Research Methods through Data Generation. \emph{J. Open Source Softw.} \textbf{2020}, \emph{5}, 2763. https://doi.org/10.21105/joss.02763.
+
+% Reference 40
+\bibitem[Revelle(2021)]{psych}
+Revelle, W. \emph{Psych: Procedures for Psychological, Psychometric, and Personality Research}; Northwestern University: Evanston, IL, USA, 2021.
+
+% Reference 41
+\bibitem[Wicherts \em{et~al.}(2016)Wicherts, Veldkamp, Augusteijn, Bakker, {van
+  Aert}, and {van Assen}]{wichertsDegreesFreedomPlanning2016}
+Wicherts, J.M.; Veldkamp, C.L.S.; Augusteijn, H.E.M.; Bakker, M.; van Aert, R.C.M.; van Assen, M.A.L.M.
+Degrees of Freedom in Planning, Running, Analyzing, and Reporting Psychological Studies: A Checklist to Avoid p-Hacking.
+\emph{Front.
+Psychol.} \textbf{2016}, \emph{7}, 1832.
+https://doi.org/10.3389/fpsyg.2016.01832.
+
+% Reference 42
+\bibitem[Szollosi \em{et~al.}(2020)Szollosi, Kellen, Navarro, Shiffrin, {van
+  Rooij}, Van~Zandt, and Donkin]{szollosi_is_2020}
+Szollosi, A.; Kellen, D.; Navarro, D.J.; Shiffrin, R.; van Rooij, I.; Van Zandt, T.; Donkin, C.
+Is Preregistration Worthwhile?
+\emph{Trends Cogn.
+Sci.} \textbf{2020}, \emph{24}, 94--95.
+https://doi.org/10.1016/j.tics.2019.11.009.
+
+% Reference 43
+\bibitem[Nosek \em{et~al.}(2019)Nosek, Beck, Campbell, Flake, Hardwicke,
+  Mellor, {van 't Veer}, and Vazire]{nosekPreregistrationHardWorthwhile2019}
+Nosek, B.A.; Beck, E.D.; Campbell, L.; Flake, J.K.; Hardwicke, T.E.; Mellor, D.T.; van ’t Veer, A.E.; Vazire, S.
+Preregistration Is Hard, And Worthwhile.
+\emph{Trends Cogn.
+Sci.} \textbf{2019}, \emph{23}, 815--818.
+https://doi.org/10.1016/j.tics.2019.07.009.
+
+% Reference 44
+\bibitem[Meehl(1978)]{meehlTheoreticalRisksTabular1978}
+Meehl, P.E.
+Theoretical Risks and Tabular Asterisks: Sir Karl, Sir Ronald, and the Slow Progress of Soft Psychology.
+\emph{J.
+Consult.
+Clin.
+Psychol.} \textbf{1978}, \emph{46}, 806--834.
+https://doi.org/10.1037/0022-006X.46.4.806.
+
+% Reference 45
+\bibitem[Brandmaier and Jacobucci(in press)]{brandmaier22mlsem}
+Brandmaier, A.M.; Jacobucci, R.
+Machine-Learning Approaches to Structural Equation Modeling.
+In \emph{Handbook of Structural Equation Modeling}, 2nd ed.; Hoyle, R.H., Ed.; \hl{Guilford Press, New York, NY, USA}, in press.%MDPI: Please add the location of the publisher and publication year.
+
+% Reference 46
+\bibitem[Hastie \em{et~al.}(2017)Hastie, Tibshirani, and
+  Friedman]{hastieElementsStatisticalLearning2017}
+Hastie, T.; Tibshirani, R.; Friedman, J.H.
+\emph{The Elements of Statistical Learning: Data Mining, Inference, and Prediction, Second Edition, Corrected at 12th Printing 2017 ed.}; Springer Series in Statisticsl Springer: New York, NY, USA, 2017.
+
+% Reference 47
+\bibitem[Dar{\'o}czi and Tsegelskyi(2021)]{pander}
+Daróczi, G.; Tsegelskyi, R. Pander: An R ‘pandoc’ Writer, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 48
+\bibitem[Hlavac(2018)]{stargazer}
+Hlavac, M. \emph{Stargazer: Well-Formatted Regression and Summary Statistics Tables}; Central European Labour Studies Institute (CELSI): Bratislava, Slovakia, 2018.
+
+% Reference 49
+\bibitem[Stanley(2021)]{apaTables}
+Stanley, D. apaTables: Create American Psychological Association (APA) Style Tables, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 50
+\bibitem[Makowski \em{et~al.}(2021)Makowski, {Ben-Shachar}, Patil, and
+  L{\"u}decke]{report}
+Makowski, D.; Ben-Shachar, M.S.; Patil, I.; Lüdecke, D.
+Automated Results Reporting as a Practical Tool to Improve Reproducibility and Methodological Best Practices Adoption.
+\hl{CRAN 2021.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 51
+\bibitem[{European Organization For Nuclear Research} and
+  {OpenAIRE}(2013)]{zenodo}
+European Organization For Nuclear Research. \emph{OpenAIRE}; \hl{Zenodo}, 2013. https://doi.org/10.25495/7GXK-RD71.%MDPI: Please add the location of the publisher.
+
+% Reference 52
+\bibitem[Weston \em{et~al.}(2019)Weston, Ritchie, Rohrer, and
+  Przybylski]{westonRecommendationsIncreasingTransparency2019}
+Weston, S.J.; Ritchie, S.J.; Rohrer, J.M.; Przybylski, A.K.
+Recommendations for Increasing the Transparency of Analysis of Preexisting Data Sets.
+\emph{Adv.
+Methods Pract.
+Psychol.
+Sci.} \textbf{2019}, \emph{2}, 214--227.
+https://doi.org/10.1177/2515245919848684.
+
+% Reference 53
+\bibitem[{International Council for Harmonisation of Technical Requirements for
+  Registration of Pharmaceuticals for Human Use}(1998)]{ICH1998}
+International Council for Harmonisation of Technical Requirements for Registration of Pharmaceuticals for Human Use.
+E 9 Statistical Principles for Clinical Trials, \hl{1998.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 54
+\bibitem[Thabane \em{et~al.}(2010)Thabane, Ma, Chu, Cheng, Ismaila, Rios,
+  Robson, Thabane, Giangregorio, and
+  Goldsmith]{thabaneTutorialPilotStudies2010}
+Thabane, L.; Ma, J.; Chu, R.; Cheng, J.; Ismaila, A.; Rios, L.P.; Robson, R.; Thabane, M.; Giangregorio, L.; Goldsmith, C.H.
+A Tutorial on Pilot Studies: The What, Why and How.
+\emph{BMC Med.
+Res.
+Methodol.} \textbf{2010}, \emph{10}, 1.
+https://doi.org/10.1186/1471-2288-10-1.
+
+% Reference 55
+\bibitem[Yuan \em{et~al.}(2019)Yuan, Topjian, Kurth, Kirschen, Ward, Zhang, and
+  Mensinger]{yuanGuideStatisticalAnalysis2019}
+Yuan, I.; Topjian, A.A.; Kurth, C.D.; Kirschen, M.P.; Ward, C.G.; Zhang, B.; Mensinger, J.L.
+Guide to the Statistical Analysis Plan.
+\emph{Pediatric Anesth.} \textbf{2019}, \emph{29}, 237--242.
+https://doi.org/10.1111/pan.13576.
+
+% Reference 56
+\bibitem[Fagerland(2012)]{fagerlandTtestsNonparametricTests2012}
+Fagerland, M.W. T-Tests, Non-Parametric Tests, and Large Studies—A Paradox of Statistical Practice? \emph{BMC Med Res. Methodol.} \textbf{2012}, \emph{12}, 78. https://doi.org/10.1186/1471-2288-12-78.
+
+% Reference 57
+\bibitem[Horton and Switzer(2005)]{hortonStatisticalMethodsJournal2005}
+Horton, N.J.; Switzer, S.S. Statistical Methods in the Journal. \emph{New Engl. J. Med.} \textbf{2005}, \emph{353}, 1977--1979. https://doi.org/10.1056/\linebreak NEJM200511033531823.
+
+% Reference 58
+\bibitem[Putnick and
+  Bornstein(2016)]{putnickMeasurementInvarianceConventions2016}
+Putnick, D.L.; Bornstein, M.H.
+Measurement Invariance Conventions and Reporting: The State of the Art and Future Directions for Psychological Research.
+\emph{Dev.
+Rev.} \textbf{2016}, \emph{41}, 71--90, https://doi.org/10.1016/j.dr.2016.06.004.
+
+% Reference 59
+\bibitem[Frost and Thompson(2000)]{frostCorrectingRegressionDilution2000}
+Frost, C.; Thompson, S.G.
+Correcting for Regression Dilution Bias: Comparison of Methods for a Single Predictor Variable.
+\emph{J.
+R.
+Stat.
+Soc.
+Ser.
+A} \textbf{2000}, \emph{163}, 173--189.
+https://doi.org/10.1111/1467-985x.00164.
+
+% Reference 60
+\bibitem[Stonehouse and Forrester(1998)]{stonehouseRobustnessTestsCombined1998}
+Stonehouse, J.M.; Forrester, G.J.
+Robustness of the \emph{t} and U Tests under Combined Assumption Violations.
+\emph{J.
+Appl.
+Stat.} \textbf{1998}, \emph{25}, 63--74.
+https://doi.org/10.1080/02664769823304.
+
+% Reference 61
+\bibitem[Zimmerman and Zumbo(1993)]{zimmermanRankTransformationsPower1993}
+Zimmerman, D.W.; Zumbo, B.D.
+Rank Transformations and the Power of the Student \emph{t} Test and Welch \emph{t} Test for Non-Normal Populations with Unequal Variances.
+\emph{Can.
+J.
+Exp.
+Psychol.
+/Rev.
+Can.
+De Psychol.
+Expérimentale} \textbf{1993}, \emph{47}, 523--539.
+https://doi.org/10.1037/h0078850.
+
+% Reference 62
+\bibitem[Brandmaier \em{et~al.}(2018)Brandmaier, {von Oertzen}, Ghisletta,
+  Lindenberger, and Hertzog]{brandmaier2018precision}
+Brandmaier, A.M.; von Oertzen, T.; Ghisletta, P.; Lindenberger, U.; Hertzog, C.
+Precision, Reliability, and Effect Size of Slope Variance in Latent Growth Curve Models: Implications for Statistical Power Analysis.
+\emph{Front.
+Psychol.} \textbf{2018}, \emph{9}, 294.
+https://doi.org/10.3389/fpsyg.2018.00294.
+
+% Reference 63
+\bibitem[Harrison(2010)]{harrisonIntroductionMonteCarlo2010}
+Harrison, R.L. Introduction to Monte Carlo Simulation. \emph{AIP Conf. Proc.} \textbf{2010}, \emph{1204}, 17--21. https://doi.org/10.1063/1.3295638.
+
+% Reference 64
+\bibitem[Raychaudhuri(2008)]{raychaudhuriIntroductionMonteCarlo2008}
+Raychaudhuri, S.
+Introduction to Monte Carlo Simulation.
+In Proceedings of the 2008 Winter Simulation Conference, Miami, FL, USA, 7--10 December 2008; pp.
+91--100.
+https://doi.org/10.1109/WSC.2008.4736059.
+
+% Reference 65
+\bibitem[Rouder \em{et~al.}(2019)Rouder, Haaf, and Snyder]{rouderMinimizingMistakesPsychological2019}
+Rouder, J.N.; Haaf, J.M.; Snyder, H.K. Minimizing Mistakes in Psychological Science. \emph{Adv. Methods Pract. Psychol. Sci.} \textbf{2019}, \emph{2}, 3--11. https://doi.org/10.1177/2515245918801915.
+
+% Reference 66
+\bibitem[Lakens and DeBruine(2021)]{lakensImprovingTransparencyFalsifiability2021}
+Lakens, D.; DeBruine, L.M.
+Improving Transparency, Falsifiability, and Rigor by Making Hypothesis Tests Machine-Readable.
+\emph{Adv.
+Methods Pract.
+Psychol.
+Sci.} \textbf{2021}, \emph{4}, https://doi.org/10.1177/2515245920970949.
+
+% Reference 67
+\bibitem[Arslan(2019)]{arslanHowAutomaticallyDocument2019}
+Arslan, R.C.
+How to Automatically Document Data With the Codebook Package to Facilitate Data Reuse.
+Advances in Methods and Practices in Psychological.
+\emph{Science} \textbf{2019}, \emph{2}, 169--187.
+https://doi.org/10.1177/2515245919838783.
+
+% Reference 68
+\bibitem[Nosek and Lakens(2014)]{nosekRegisteredReports2014}
+Nosek, B.A.; Lakens, D. Registered Reports. \emph{Soc. Psychol.} \textbf{2014}, \emph{45}, 137--141. https://doi.org/10.1027/1864-9335/a000192.
+
+% Reference 69
+\bibitem[Chambers(2019)]{chambersWhatNextRegistered2019}
+Chambers, C. What’s next for Registered Reports? \emph{Nature} \textbf{2019}, \emph{573}, 187--189. https://doi.org/10.1038/d41586-019-02674-6.
+
+% Reference 70
+\bibitem[Simons(2014)]{simonsValueDirectReplication2014}
+Simons, D.J. The Value of Direct Replication. \emph{Perspect. Psychol. Sci.} \textbf{2014}, \emph{9}, 76--80. https://doi.org/10.1177/1745691613514755.
+
+% Reference 71
+\bibitem[Elliott \em{et~al.}(2014)Elliott, Turner, Clavisi, Thomas, Higgins,
+  Mavergames, and Gruen]{elliottLivingSystematicReviews2014}
+Elliott, J.H.; Turner, T.; Clavisi, O.; Thomas, J.; Higgins, J.P.T.; Mavergames, C.; Gruen, R.L.
+Living Systematic Reviews: An Emerging Opportunity to Narrow the Evidence-Practice Gap.
+\emph{PLoS Med.} \textbf{2014}, \emph{11}, e1001603.
+https://doi.org/10.1371/journal.pmed.\linebreak1001603.
+
+% Reference 72
+\bibitem[{eLife Sciences
+  Publications}(2020)]{elifesciencespublicationsELifeLaunchesExecutable2020}
+eLife Sciences Publications. eLife Launches Executable Research Articles for Publishing Computationally Reproducible Results \hl{2020.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 73
+\bibitem[Rouder(2016)]{rouderWhatWhyHow2016}
+Rouder, J.N. The What, Why, and How of Born-Open Data. \emph{Behav. Res. Methods} \textbf{2016}, \emph{48}, 1062--1069. https://doi.org/10.3758/\linebreak13428-015-0630-z.
+
+% Reference 74
+\bibitem[Kekecs \em{et~al.}(2020)Kekecs, Aczel, Palfi, Szaszi, Szecsi, Zrubka,
+  Kovacs, Bakos, Cousineau, Tressoldi, Grassi, Arnold, Evans, Yamada, Miller,
+  Liu, Yonemitsu, and Dubrov]{kekecsRaisingValueResearch2019}
+Kekecs, Z.; Aczel, B.; Palfi, B.; Szaszi, B.; Szecsi, P.; Zrubka, M.; Kovacs, M.; Bakos, B.E.; Cousineau, D.; Tressoldi, P.; et al.
+Raising the Value of Research Studies in Psychological Science by Increasing the Credibility of Research Reports: The Transparent Psi Project—Preprint.
+\emph{PsyArXiv} \textbf{2020}.
+https://doi.org/10.31234/osf.io/uwk7y.
+
+% Reference 75
+\bibitem[Lanergan and Grasso(1989)]{lanerganSoftwareEngineeringReusable1989}
+Lanergan, R.G.; Grasso, C.A.
+Software Engineering with Reusable Designs and Code.
+In \emph{Software Reusability: Vol.
+2, Applications and Experience}; Association for Computing Machinery: New York, NY, USA, 1989; pp.
+187--195.
+
+% Reference 76
+\bibitem[{Al-Badareen} \em{et~al.}(2010){Al-Badareen}, Selamat, Jabar, Din, and
+  Turaev]{al-badareenReusableSoftwareComponents2010}
+Al-Badareen, A.B.; Selamat, M.H.; Jabar, M.A.; Din, J.; Turaev, S.
+Reusable Software Components Framework.
+In Proceedings of the European Conference of Systems, and European Conference of Circuits Technology and Devices, and European Conference of Communications, and European Conference on Computer Science, Kuantan, Pahang, Malaysia, 27--29 June 2011; World Scientific and Engineering Academy and Society (WSEAS): Stevens Point, WI, 2010; ECS’10/ECCTD’10/ECCOM’10/ECCS’10, pp.
+126--130.
+
+% Reference 77
+\bibitem[Schaffner(1994)]{schaffnerFutureScientificJournals1994}
+Schaffner, A.C. The Future of Scientific Journals: Lessons from the Past. \emph{Inf. Technol. Libr.} \textbf{1994}, \emph{13}, 239--247.
+
+% Reference 78
+\bibitem[Fitzgerald(2006)]{fitzgeraldTransformationOpenSource2006}
+Fitzgerald, B. The Transformation of Open Source Software. \emph{MIS Q.} \textbf{2006}, \emph{30}, 587--598. https://doi.org/10.2307/25148740.
+
+% Reference 79
+\bibitem[Chaldecott(1965)]{chaldecottHistoryScientificTechnical1965}
+Chaldecott, J.A.
+A History of Scientific and Technical Periodicals: The Origins and Development of the Scientific and Technological Press.
+\emph{Br.
+J.
+Hist.
+Sci.} \textbf{1965}, \emph{2}, 360--361.
+https://doi.org/10.1017/S0007087400002557.
+
+% Reference 80
+\bibitem[Sonnenburg \em{et~al.}(2007)Sonnenburg, Braun, Ong, Bengio, Bottou,
+  Holmes, LeCun, M\"{u}ller, Pereira, Rasmussen, R\"{a}tsch, Sch\"{o}lkopf,
+  Smola, Vincent, Weston, and Williamson]{sonnenburgNeedOpenSource2007}
+Sonnenburg, S.; Braun, M.L.; Ong, C.S.; Bengio, S.; Bottou, L.; Holmes, G.; LeCun, Y.; Müller, K.R.; Pereira, F.; Rasmussen, C.E.;~et~al.
+The Need for Open Source Software in Machine Learning.
+\emph{J.
+Mach.
+Learn.
+Res.} \textbf{2007}, \emph{8}, 2443--2466.
+
+% Reference 81
+\bibitem[Gohel(2021)]{officer}
+Gohel, D. Officer: Manipulation of Microsoft Word and PowerPoint Documents, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+
+% Reference 82
+\bibitem[Gohel and Ross(2021)]{officedown}
+Gohel, D.; Ross, N. Officedown: Enhanced ‘R Markdown’ Format for ‘Word’ and ‘PowerPoint’, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+\end{thebibliography}
+
 \end{document}

### 2021-12-03 manuscript.tex
reviewed but does not compile

@@
 \Title{Reproducible Research in R: A Tutorial on How to Do the Same
-Thing More \highlighting{Than Once}} %MDPI: the article type in the document is different with that in Susy, please confirm.
+Thing More \highlighting{Than Once}}
 
@@
 % Authors, for the paper (add full first names)
-\Author{\highlighting{Aaron Peikert~$^{1,2,*}$\href{https://orcid.org/0000-0001-7813-818X}{\orcidicon}} %Please carefully check the accuracy of names and affiliations.
-Changes will not be possible after proofreading.
-, \highlighting{Caspar J. van Lissa~$^{3,4}$\href{https://orcid.org/0000-0002-0808-5024}{\orcidicon}} %MDPI: This author is different from susy. Please check.
+\Author{\highlighting{Aaron Peikert~$^{1,2,*}$\href{https://orcid.org/0000-0001-7813-818X}{\orcidicon}}
+, \highlighting{Caspar J. van Lissa~$^{3,4}$\href{https://orcid.org/0000-0002-0808-5024}{\orcidicon}}
  and Andreas M. Brandmaier~$^{1,5,6}$\href{https://orcid.org/0000-0001-8765-6982}{\orcidicon}}
@@
 $^{1}$ \quad Center for Lifespan Psychology---Max Planck Institute for Human Development, Lentzeallee 94, \mbox{14195 Berlin, Germany;} brandmaier@mpib-berlin.mpg.de\\
-$^{2}$ \quad Humboldt-Universität zu Berlin, \hl{Berlin}, Germany\\%MDPI: Please add the post code.
-$^{3}$ \quad \hl{Department} of Methodology \& Statistics---Utrecht \hl{University}, \hl{Faculty} of Social and Behavioral Sciences%MDPI: Please check the address order.
-(order guide: Center / Laboratory / Group / Programme / Unit < Department / Division / Faculty / Campus / Institute / School < College / University / Universiti)
+$^{2}$ \quad Humboldt-Universität zu Berlin, Unter den Linden 6, \mbox{10117 Berlin, Germany}
+$^{3}$ \quad \hl{Department} of Methodology \& Statistics---Utrecht \hl{University}, \hl{Faculty} of Social and Behavioral Sciences
 , \hl{Utrecht}, The Netherlands; C.J.vanLissa@uu.nl\\
 $^{4}$ \quad Open Science Community Utrecht, \hl{Utrecht}, The Netherlands\\
-$^{5}$ \quad Max Planck UCL Centre for Computational Psychiatry and Ageing Research, \mbox{\hl{Berlin, Germany, and London, UK}} %MDPI: 1.
-Please check if there are two separator affiliations.
-If so, please split them.
-2.
-Please add the post code.
-\\
+$^{5}$ \quad Max Planck UCL Centre for Computational Psychiatry and Ageing Research, \mbox{\hl{Berlin, Germany, and London, UK}}\\
 $^{6}$ \quad MSB Medical School Berlin, \hl{Berlin}, Germany}
@@
 \keyword{open science; computational reproducibility; preregistration;
-R; R Markdown; make; GitHub; docker}%MDPI: 1. The upper case has been removed. Please confirm. 2. Please check if R is duplicated here.
+R; R Markdown; Make; GitHub; Docker}
 
@@
 defined as the ability of anyone to obtain identical results from the
-\emph{same} data with the \hl{\emph{same}} %MDPI: Is the italic necessary? Can it be deleted?
+\emph{same} data with the \hl{\emph{same}}
  computer code (see~\citep{Peikert2019} \hl{for
-details).} %MDPI: The structure of reference citation has been changed. So as the follows. Please confirm.
+details).}
  The credibility of empirical research results
@@
 \item
-  \highlighting{\emph{Multiple inconsistent versions of code, data, or both};} %MDPI: Is the italic necessary? Can it be deleted?
+  \highlighting{\emph{Multiple inconsistent versions of code, data, or both};}
  for
@@
 scientific data analysis, targets are typically the final scientific
-report (e.g., the one to be submitted to a journal) and produce possibly
-intermediate %pls check (such as preprocessed data files, simulation
+report (e.g., the one to be submitted to a journal) and possibly
+intermediate results (such as preprocessed data files, simulation
 results, and analysis results). A workflow that involves renaming
@@
 executed to recreate the software environment. \texttt{Dockerfiles} are
-tiny (the \hl{\texttt{Dockerfile}} %MDPI: Please add the accessed date.
+tiny (the \hl{\texttt{Dockerfile}}
  (\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile})
@@
 To summarize, the workflow by Peikert and Brandmaier~\citep{Peikert2019} requires four
-components (see \highlighting{Figure~\ref{figure1})} %MDPI: Figure 2.1 has been changed into Figure 1. Please confirm.
+components (see \highlighting{Figure~\ref{figure1})}
  dynamic document generation (using
@@
 necessary for this tutorial in three steps. The
-\hl{installation advice} %MDPI: Please add the accessed date.
+\hl{installation advice}
 (\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md}) may also help Windows users who have problems installing
@@
 project snapshot). One way to do this is through the visual user
-interface in the RStudio \texttt{Git} pane (see \highlighting{Figure~\ref{figure2}).} %MDPI: Figure 3.1 has been changed into Figure 2. Please confirm.
+interface in the RStudio \texttt{Git} pane (see \highlighting{Figure~\ref{figure2}).}
  Click on
@@
 \textls[+15]{The idea of submitting code as part of a preregistration is not new (e.g.~\citep{wichertsDegreesFreedomPlanning2016}).
-A prominent preregistration platform, \hl{The Open Science Framework} %MDPI: Please add the accessed date.
-(\url{https://osf.io/}), suggests submitting scripts alongside the preregistration of methods.
+A prominent preregistration platform, \href{https://osf.io/}{The Open Science
+Framework}, suggests submitting scripts alongside the preregistration of methods.
 In an informal literature search (we skimmed the first 300 results of Google Scholar with the keywords \texttt{\textls[+35]{("pre\ registration"\textbar{}"pre-registration"\textbar{}preregistration)\&(code\textbar{}script\textbar{}matlab\textbar{}}\linebreak python\textbar{}"R")}) we only found close to a dozen published articles that did include some form of script as part of their preregistration.
@@
 retrospectively summarize and justify any changes made to the
-preregistered plan, e.g.~in the discussion of the \hl{final manuscript} %MDPI: Footnote is not allowed in this journal. The content has been moved here. Please confirm.
+preregistered plan, e.g.~in the discussion of the \hl{final manuscript}
 ({In section {[}Preregistration as Code---a Tutorial{]} we
@@
 interested reader can find working functions in the \hl{online
-supplementary materials}% MDPI: Please add the accessed date.
+supplementary materials}
 ~(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R}) that power the example below.
@@
 tables of estimates or figures directly from R goal to help with this
-objective. Packages like {\texttt{pander}}
-(\url{https://github.com/Rapporter/pander})
-\citep{pander}, {\texttt{stargazer}}
-\mbox{(\url{https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf})}
-\citep{stargazer}, {\texttt{apaTables}}
-(\url{https://dstanley4.github.io/apaTables/articles/apaTables.html})
-\citep{apaTables} and {\texttt{papaja}}
-\mbox{(\url{http://frederikaust.com/papaja_man/})}
+objective. Packages like
+\href{https://github.com/Rapporter/pander}{\texttt{pander}}
+\citep{pander},
+\href{https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf}{\texttt{stargazer}}
+\citep{stargazer},
+\href{https://dstanley4.github.io/apaTables/articles/apaTables.html}{\texttt{apaTables}}
+\citep{apaTables} and
+\href{http://frederikaust.com/papaja_man/}{\texttt{papaja}}
 \citep{papaja} help you to create dynamically generated professional
-looking results. The package {\texttt{report}}
-(\url{https://github.com/easystats/report})~\citep{report} is particularly noteworthy because it not only generates
+looking results. The package
+\href{https://github.com/easystats/report}{\texttt{report}}~\citep{report} is particularly noteworthy because it not only generates
 tables but also a straightforward interpretation of the effects as
@@
 \begin{quote}
-\hl{Navigate to} %MDPI: Please add the accessed date.
+\hl{Navigate to}
 ~\url{https://zenodo.org/login/} → Log in with GitHub
@@
 \begin{quote}
-Log into zenodo.org → Account → \hl{GitHub} %MDPI: 1. Footnote is not allowed in this journal. The content has been moved here. Please confirm. 2. Please add the accessed date
-~({\url{https://zenodo.org/account/settings/github/}})
+Log into zenodo.org → Account → \hl{GitHub}~({\url{https://zenodo.org/account/settings/github/}})
 \end{quote}
@@
 eventually becomes the final report. The complete preregistration source
-is available in the \hl{online supplementary material}%MDPI: Please add the accessed date.
+is available in the \hl{online supplementary material}
 ~(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd}). In this section, we show code excerpts of this
@@
 We choose the following statistical procedure because many researchers
-are familiar \hl{with it} %MDPI: Footnote is not allowed. The content has been moved here. Please confirm.
+are familiar \hl{with it}
 ~(The \emph{t}-test and Mann-Whitney-Wilcoxon test
@@
 significant results (power) is recorded. The results for such a Monte
-Carlo simulation for this example are visualized in \highlighting{Figure~\ref{figure3}.} %MDPI:  Figure 5.1 has been changed into Figure 3. Please confirm.
+Carlo simulation for this example are visualized in \highlighting{Figure~\ref{figure3}.}
  The code
@@
 change our preregistration after accessing the data to include i.e.,~the
-recoding of gender (lines 17--22) and the items\hl{ (lines 23--26).} %MDPI: Footnote is not allowed. The content has been moved here. Please confirm.
- ({We
-  invite the reader to evaluate the changes we made to the preregistered
-  code. Either on \hl{GitHub}
-  (\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg...main})
+recoding of gender (lines 17--22) and the items\hl{ (lines 23--26).}
+We invite the reader to evaluate the changes we made to the preregistered code. Either on \hl{GitHub} (\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg...main})
   → ``Files changed'' or directly in Git with
-  \texttt{git\ diff\ v0.0.1.1-prereg\ preregistration.Rmd}.}) This is our
+  \texttt{git\ diff\ v0.0.1.1-prereg\ preregistration.Rmd}.) This is our
 summary of what we changed:
@@
 effect is - negative, statistically significant, and small (difference = -0.17,
-95% CI [0.12, 0.22], t(887.46) = 6.38, p < 0.001; Cohen's d = 0.43, 95% CI [0.30,
+95% CI [0.12, 0.22], t(887.46) = 6.38, p < .001; Cohen's d = 0.43, 95% CI [0.30,
 0.56])
@@
 
-Supplementary Materials{\hl{\quad \quad}}%The following are available online at www.mdpi.com/xxx/s1, Figure S1: title, Table S1: title, Video S1: title.
+Supplementary Materials{All materials (i.e., the source code, all figures, and the data) that are neccesary for reproducing the submitted version of this article are availible at \url{https://github.com/aaronpeikert/repro-tutorial} and archived under \url{https://doi.org/10.5281/zenodo.5724454}.
 
@@
 
-\funding{\hl{\quad \quad}} %This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
-
-\institutionalreview{\hl{\quad \quad}} %In this section, please add the Institutional Review Board Statement and approval number for studies involving humans or animals.
-Please note that the Editorial Office might ask you for further information.
-Please add ``The study was conducted according to the guidelines of the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).'' OR ``Ethical review and approval were waived for this study, due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans or animals.
-You might also choose to exclude this statement if the study did not involve humans or animals.
-
-\informedconsent{\hl{\quad \quad}} %Any research article describing a study involving humans should contain this statement.
-Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
-You might also choose to exclude this statement if the study did not involve humans.
-Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
-Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.
-
-\dataavailability{\hl{\quad \quad}} %In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
-Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.
-You might choose to exclude this statement if the study did not report any data.
+\funding{This research received no external funding}
+
+\institutionalreview{Not applicable.}
+
+\informedconsent{Not applicable.}
 

### 2021-12-04 manuscript.tex
wrap things up

@@
 \copyrightyear{2019}
-\externaleditor{Academic Editor: {Alexander Robitzsch}} %MDPI: Please add the academic editor.
+\externaleditor{Academic Editor: {Alexander Robitzsch}}
 \datereceived{} 
@@
 Unfortunately, Docker requires administrator rights to run, which may
-not be available to all researchers. We recommend {\texttt{renv}}
-(\url{https://rstudio.github.io/renv/articles/renv.html})
+not be available to all researchers. We recommend
+\href{https://rstudio.github.io/renv/articles/renv.html}{\texttt{renv}}
 \citep{R-renv} in cases where no administrator rights can be obtained
-but we can not detail its use in this document. \texttt{renv} tracks which
-R package is installed from which source in which version in a so-called \hl{\texttt{lockfile}}
-(\url{https://rstudio.github.io/renv/articles/lockfile.html}).
+but can not detail its use in this document. \texttt{renv} tracks which
+R package is installed from which source in which version in a so-called
+\href{https://rstudio.github.io/renv/articles/lockfile.html}{\texttt{lockfile}}.
 This lockfile is then used to reinstall the same packages on other
@@
 
-Start RStudio and install the package \hl{\texttt{repro}}
-(\url{https://github.com/aaronpeikert})~\citep{R-repro}.
+Start RStudio and install the package
+\href{https://github.com/aaronpeikert}{\texttt{repro}}\citep{R-repro}.
 It will assist you while you follow the tutorial.
@@
 Skrondal~\citep{skrondalDesignAnalysisMonte2000}, as well as the R packages
-{\texttt{simstudy}} 
-\mbox{ (\url{https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html})}
-~\citep{simstudy} and
-{\texttt{psych}} (\url{https://personality-project.org/r/psych/help/sim.html})
+\href{https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html}{\texttt{simstudy}},
+\citep{simstudy} and
+\href{https://personality-project.org/r/psych/help/sim.html}{\texttt{psych}},
 \citep{psych}.
@@
 \begin{quote}
-\hl{Navigate to} %MDPI: Please add the accessed date.
+\hl{Navigate to}
  \url{https://zenodo.org/account/settings/github/}
@@
 
-Supplementary Materials{All materials (i.e., the source code, all figures, and the data) that are neccesary for reproducing the submitted version of this article are availible at \url{https://github.com/aaronpeikert/repro-tutorial} and archived under \url{https://doi.org/10.5281/zenodo.5724454}.
+Supplementary Materials{All materials (i.e., the source code, all figures, and the data) that are neccesary for reproducing the submitted version of this article are availible at \url{https://github.com/aaronpeikert/repro-tutorial} and archived under \url{https://doi.org/10.5281/zenodo.5724454}}.
 

### 2021-12-09 manuscript.tex
add another proof

@@
 \Title{Reproducible Research in R: A Tutorial on How to Do the Same
-Thing More \highlighting{Than Once}}
+Thing More {Than Once}}
 
@@
 % Authors, for the paper (add full first names)
-\Author{\highlighting{Aaron Peikert~$^{1,2,*}$\href{https://orcid.org/0000-0001-7813-818X}{\orcidicon}}
-, \highlighting{Caspar J. van Lissa~$^{3,4}$\href{https://orcid.org/0000-0002-0808-5024}{\orcidicon}}
+\Author{{Aaron Peikert~$^{1,2,*}$\href{https://orcid.org/0000-0001-7813-818X}{\orcidicon}}
+, {Caspar J. van Lissa~$^{3,4}$\href{https://orcid.org/0000-0002-0808-5024}{\orcidicon}}
  and Andreas M. Brandmaier~$^{1,5,6}$\href{https://orcid.org/0000-0001-8765-6982}{\orcidicon}}
@@
 $^{2}$ \quad Humboldt-Universität zu Berlin, Unter den Linden 6, \mbox{10117 Berlin, Germany}
-$^{3}$ \quad \hl{Department} of Methodology \& Statistics---Utrecht \hl{University}, \hl{Faculty} of Social and Behavioral Sciences
-, \hl{Utrecht}, The Netherlands; C.J.vanLissa@uu.nl\\
-$^{4}$ \quad Open Science Community Utrecht, \hl{Utrecht}, The Netherlands\\
-$^{5}$ \quad Max Planck UCL Centre for Computational Psychiatry and Ageing Research, \mbox{\hl{Berlin, Germany, and London, UK}}\\
-$^{6}$ \quad MSB Medical School Berlin, \hl{Berlin}, Germany}
+
+$^{3}$ \quad \hl{Department} of Methodology \& Statistics---Utrecht {University}, {Faculty} of Social and Behavioral Sciences
+, Heidelberglaan 8, 3584 CS Utrecht, The Netherlands; C.J.vanLissa@uu.nl\\
+$^{4}$ \quad Open Science Community Utrecht, Heidelberglaan 8, 3584 CS Utrecht, The Netherlands\\
+$^{5}$ \quad Max Planck UCL Centre for Computational Psychiatry and Ageing Research, \mbox{\hl{Lentzeallee 94
+14195 Berlin, Germany, and London, WC1B 5EH, UK}}\\
+$^{6}$ \quad MSB Medical School Berlin, Rüdesheimer Str. 50, 14197 Berlin, Germany}
 
@@
 defined as the ability of anyone to obtain identical results from the
-\emph{same} data with the \hl{\emph{same}}
- computer code (see~\citep{Peikert2019} \hl{for
+\emph{same} data with the {\emph{same}}
+ computer code (see~\citep{Peikert2019} {for
 details).}
@@
 \item
-  \highlighting{\emph{Multiple inconsistent versions of code, data, or both};}
+  {\emph{Multiple inconsistent versions of code, data, or both};}
  for
@@
 \item
-  \highlighting{\emph{Copy-and-paste errors};} for example, results are often manually
+  {\emph{Copy-and-paste errors};} for example, results are often manually
   copied from a statistical computing language into a text processor; if
@@
 \item
-  \highlighting{\emph{Undocumented or ambiguous order of computation};} for example,
+  {\emph{Undocumented or ambiguous order of computation};} for example,
   with multiple data and code files, it may be unclear which scripts
@@
 \item
-  \highlighting{\emph{Ambiguous software dependencies};} for example, a given analysis
+  {\emph{Ambiguous software dependencies};} for example, a given analysis
   may depend on a specific version of a specific software package, or
@@
 executed to recreate the software environment. \texttt{Dockerfiles} are
-tiny (the \hl{\texttt{Dockerfile}}
- (\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile})
+tiny (the {\texttt{Dockerfile}}
+ (\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile} accessed on 11 October 2021)
 for this project has a size of only 1.55~KiB). However, they rely on the
@@
 To summarize, the workflow by Peikert and Brandmaier~\citep{Peikert2019} requires four
-components (see \highlighting{Figure~\ref{figure1})}
+components (see {Figure~\ref{figure1})}
  dynamic document generation (using
@@
 necessary for this tutorial in three steps. The
-\hl{installation advice}
-(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md}) may also help Windows users who have problems installing
+{installation advice}
+(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md} accessed on 11 October 2021) may also help Windows users who have problems installing
 \texttt{Docker}.
@@
 project snapshot). One way to do this is through the visual user
-interface in the RStudio \texttt{Git} pane (see \highlighting{Figure~\ref{figure2}).}
+interface in the RStudio \texttt{Git} pane (see {Figure~\ref{figure2}).}
  Click on
@@
 retrospectively summarize and justify any changes made to the
-preregistered plan, e.g.~in the discussion of the \hl{final manuscript}
+preregistered plan, e.g.~in the discussion of the {final manuscript}
 ({In section {[}Preregistration as Code---a Tutorial{]} we
@@
 study. The functions shown here do not contain executable code, but the
-interested reader can find working functions in the \hl{online
+interested reader can find working functions in the {online
 supplementary materials}
-~(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R}) that power the example below.
+~(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R} accessed on 11 October 2021) that power the example below.
 
@@
 \begin{quote}
-\hl{Navigate to}
-~\url{https://zenodo.org/login/} → Log in with GitHub
+{Navigate to}
+~\url{https://zenodo.org/login/ accessed on 11 October 2021} → Log in with GitHub
 \end{quote}
@@
 \begin{quote}
-Log into zenodo.org → Account → \hl{GitHub}~({\url{https://zenodo.org/account/settings/github/}})
+Log into zenodo.org → Account → {GitHub}~({\url{https://zenodo.org/account/settings/github/}})
 \end{quote}
@@
 \begin{quote}
-\hl{Navigate to}
+{Navigate to}
  \url{https://zenodo.org/account/settings/github/}
@@
 eventually becomes the final report. The complete preregistration source
-is available in the \hl{online supplementary material}
-~(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd}). In this section, we show code excerpts of this
+is available in the {online supplementary material}
+~(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd} accessed on 11 October 2021). In this section, we show code excerpts of this
 preregistration (formatted in monospace) and explain the rationale
@@
 We choose the following statistical procedure because many researchers
-are familiar \hl{with it}
+are familiar {with it}
 ~(The \emph{t}-test and Mann-Whitney-Wilcoxon test
@@
  ~\citep{stonehouseRobustnessTestsCombined1998}). The interested reader
-  can use the \hl{provided code for the simulation}
+  can use the {provided code for the simulation}
 ~\mbox{(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R})} to verify that the \emph{t}-test provides unbiased
@@
 significant results (power) is recorded. The results for such a Monte
-Carlo simulation for this example are visualized in \highlighting{Figure~\ref{figure3}.}
+Carlo simulation for this example are visualized in {Figure~\ref{figure3}.}
  The code
-for this power analysis can be found in the \hl{online supplementary material}
-(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R}). The next snippet shows how we integrated the
+for this power analysis can be found in the {online supplementary material}
+(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R} accessed on 11 October 2021). The next snippet shows how we integrated the
 results dynamically into the preregistration (the origin of the
@@
 change our preregistration after accessing the data to include i.e.,~the
-recoding of gender (lines 17--22) and the items\hl{ (lines 23--26).}
-We invite the reader to evaluate the changes we made to the preregistered code. Either on \hl{GitHub} (\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg...main})
+recoding of gender (lines 17--22) and the items{ (lines 23--26).}
+We invite the reader to evaluate the changes we made to the preregistered code.
+Either on {GitHub} (\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg...main} accessed on 11 October 2021)
   → ``Files changed'' or directly in Git with
@@
 
-Readers can inspect and judge the changes for themselves on \hl{GitHub}
+Readers can inspect and judge the changes for themselves on {GitHub}
 \textls[-25]{(\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main\#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b})}.
@@
 
-The example rendered as a \hl{PDF file with real data}
-(\url{https://github.com/aaronpeikert/repro-tutorial/files/7309455/preregistration.pdf}) is available in the \hl{online supplementary material}
-(\url{https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.3.1-results}). The changes we made since preregistering it can
-be inspected on this \hl{GitHub page}
-(\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main\#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b}).
+The example rendered as a {PDF file with real data}
+(\url{https://github.com/aaronpeikert/repro-tutorial/files/7309455/preregistration.pdf} accessed on 11 October 2021) is available in the {online supplementary material}
+(\url{https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.3.1-results} accessed on 11 October 2021). The changes we made since preregistering it can
+be inspected on this {GitHub page}
+(\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main\#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b} accessed on 11 October 2021).
 
@@
 
-Supplementary Materials{All materials (i.e., the source code, all figures, and the data) that are neccesary for reproducing the submitted version of this article are availible at \url{https://github.com/aaronpeikert/repro-tutorial} and archived under \url{https://doi.org/10.5281/zenodo.5724454}}.
+Supplementary Materials {All materials (i.e., the source code, all figures, and the data) that are neccesary for reproducing the submitted version of this article are availible at \url{https://github.com/aaronpeikert/repro-tutorial} and archived under \url{https://doi.org/10.5281/zenodo.5724454}}.
 
@@
 Julia Delius for her helpful assistance in language and style editing.
-The R package \texttt{repro} was developed as part of the \hl{first author's
+The R package \texttt{repro} was developed as part of the {first author's
 master thesis}
-\mbox{(\url{https://aaronpeikert.github.io/repro-thesis/})} at the Humboldt-Universität zu Berlin.}
+\mbox{(\url{https://aaronpeikert.github.io/repro-thesis/} accessed on 11 October 2021)} at the Humboldt-Universität zu Berlin.}
 
@@
 \bibitem[Knuth and Levy(1994)]{knuthCWEBSystemStructured}
-Knuth, D.E.; Levy, S. \emph{The CWEB System of Structured Documentation}; Addison-Wesley Longman: \hl{Boston, USA, 1994.}%MDPI: Please add the publisher’s location.
+Knuth, D.E.; Levy, S. \emph{The CWEB System of Structured Documentation}; Addison-Wesley Longman: \hl{Boston, USA, 1994.}
 
@@
 Machine-Learning Approaches to Structural Equation Modeling.
-In \emph{Handbook of Structural Equation Modeling}, 2nd ed.; Hoyle, R.H., Ed.; \hl{Guilford Press, New York, NY, USA}, in press.%MDPI: Please add the location of the publisher and publication year.
+In \emph{Handbook of Structural Equation Modeling}, 2nd ed.; Hoyle, R.H., Ed.; \hl{Guilford Press, New York, NY, USA}, in press.
 

### 2022-04-04 manuscript.tex

@@
 \articlenumber{5}
-\pubyear{2019}
-\copyrightyear{2019}
+\pubyear{2021}
+\copyrightyear{2021}
 \externaleditor{Academic Editor: {Alexander Robitzsch}}
-\datereceived{} 
-\dateaccepted{} 
-\datepublished{}
+\datereceived{11 October 2021} 
+\dateaccepted{25 November 2021} 
+\datepublished{7 December 2021}
 \hreflink{https://doi.org/} % If needed use \linebreak
@@
 $^{1}$ \quad Center for Lifespan Psychology---Max Planck Institute for Human Development, Lentzeallee 94, \mbox{14195 Berlin, Germany;} brandmaier@mpib-berlin.mpg.de\\
-$^{2}$ \quad Humboldt-Universität zu Berlin, Unter den Linden 6, \mbox{10117 Berlin, Germany}
-
-$^{3}$ \quad \hl{Department} of Methodology \& Statistics---Utrecht {University}, {Faculty} of Social and Behavioral Sciences
+$^{2}$ \quad Humboldt-Universität zu Berlin, Unter den Linden 6, \mbox{10117 Berlin, Germany}\\
+$^{3}$ \quad {Department} of Methodology \& Statistics---Utrecht {University}, {Faculty} of Social and Behavioral Sciences
 , Heidelberglaan 8, 3584 CS Utrecht, The Netherlands; C.J.vanLissa@uu.nl\\
 $^{4}$ \quad Open Science Community Utrecht, Heidelberglaan 8, 3584 CS Utrecht, The Netherlands\\
-$^{5}$ \quad Max Planck UCL Centre for Computational Psychiatry and Ageing Research, \mbox{\hl{Lentzeallee 94
-14195 Berlin, Germany, and London, WC1B 5EH, UK}}\\
+$^{5}$ \quad Max Planck UCL Centre for Computational Psychiatry and Ageing Research, \mbox{{Lentzeallee 94,
+14195 Berlin, Germany, and London WC1B 5EH, UK}}\\
 $^{6}$ \quad MSB Medical School Berlin, Rüdesheimer Str. 50, 14197 Berlin, Germany}
@@
 allowing users to include R code chunks (in fact, arbitrary computer
-code \citep[Chapter 15, Chapter 15, Other
-Languages]{xieMarkdownCookbook2020}) into a \texttt{Markdown} document.
+code (\citep{xieMarkdownCookbook2020} Chapter 15, Chapter 15, Other
+Languages)) into a \texttt{Markdown} document.
 Upon rendering the document, the code blocks are executed, and their
@@
 Peikert and Brandmaier~\citep{Peikert2019} licensed under CC BY 4.0 
-(\url{https://creativecommons.org/licenses/by/4.0}).}
+(\url{https://creativecommons.org/licenses/by/4.0} accessed on 4 December 2021).}
 \label{figure1}
@@
 The first step is to install the required software. We assume that you
-have installed R (\citep{R-base} version 4.0.4) and RStudio
+have installed R~(\citep{R-base} version 4.0.4) and RStudio
 (\citep{rstudio} version 1.4) already but the tutorial will guide you in
@@
 help of the R package {\texttt{repro}} 
-(\url{https://github.com/aaronpeikert/repro})~\citep{R-repro}. In case you have either not installed R and RStudio or
+(\url{https://github.com/aaronpeikert/repro} accessed on 4 December 2021)~\citep{R-repro}. In case you have either not installed R and RStudio or
 are unsure if they are up-to-date, you might want to consult our
 installation advice in the Online Supplementary Material
-(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md}) that covers the installation of all software
+(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md} accessed on 4 December 2021) that covers the installation of all software
 necessary for this tutorial in three steps. The
@@
 \item
-  Metadata (e.g.~title, authors, document format)
+  Metadata (e.g.,~title, authors, document format)
 \end{enumerate}
@@
 style, e.g.~specific journal requirements. For social scientists, the
-\href{https://github.com/crsh/papaja}{\texttt{papaja}} package
-\citep{papaja} may be relevant, as it produces manuscripts that follow
+\href{https://github.com/crsh/papaja}{\texttt{papaja}} package~\citep{papaja} may be relevant, as it produces manuscripts that follow
 the American Psychological Association formatting requirements
@@
 
-The hotkey {[}Control{]}+{[}Alt{]}+{[}i{]} inserts a block of code in
+The hotkey {[}Control{]} + {[}Alt{]} + {[}i{]} inserts a block of code in
 the file. The results of code enclosed in such backticks will be
@@
 Skrondal~\citep{skrondalDesignAnalysisMonte2000}, as well as the R packages
-\href{https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html}{\texttt{simstudy}},
-\citep{simstudy} and
+\href{https://cran.r-project.org/web/packages/simstudy/vignettes/simstudy.html}{\texttt{simstudy}},~\citep{simstudy} and
 \href{https://personality-project.org/r/psych/help/sim.html}{\texttt{psych}},
@@
 
-\textls[+15]{The idea of submitting code as part of a preregistration is not new (e.g.~\citep{wichertsDegreesFreedomPlanning2016}).
+\textls[+15]{The idea of submitting code as part of a preregistration is not new (e.g.,~\citep{wichertsDegreesFreedomPlanning2016}).
 A prominent preregistration platform, \href{https://osf.io/}{The Open Science
@@
 factor analysis or machine learning. Decisions that do not lend
-themselves to formulation in code, e.g.~visual inspection, must still be
+themselves to formulation in code, e.g.,~visual inspection, must still be
 described verbally or be treated as noted in the next section. Fifth,
@@
 interested reader can find working functions in the {online
-supplementary materials}
+Supplementary Materials}
 ~(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R} accessed on 11 October 2021) that power the example below.
@@
 Ideally, these three functions can be composed to create a ``fake''
-results section, e.g.~when composed to
-\texttt{report\_analysis(planned\_analysis(simulate\_data()))} or
+results section, e.g.,~when composed to
+\texttt{report\_analysis(planned\_analysis(simulate\_data()))} or\\
 \texttt{simulate\_data()\ \%\textgreater{}\%\ planned\_analysis()\ \%\textgreater{}\%\ report\_analysis()}
@@
 {Navigate to}
-~\url{https://zenodo.org/login/ accessed on 11 October 2021} → Log in with GitHub
+~\url{https://zenodo.org/login/} accessed on 11 October 2021 → Log in with GitHub
 \end{quote}
@@
 \begin{quote}
-Log into zenodo.org → Account → {GitHub}~({\url{https://zenodo.org/account/settings/github/}})
+Log into zenodo.org → Account → {GitHub}~({\url{https://zenodo.org/account/settings/github/} accessed on 4 December 2021 })
 \end{quote}
@@
 {Navigate to}
- \url{https://zenodo.org/account/settings/github/}
+ \url{https://zenodo.org/account/settings/github/} accessed on 4 December 2021
 \end{quote}
@@
 To summarize, researchers need to write three functions,
-\texttt{planned\_analysis()}, \texttt{simulate\_data()}, and
+\texttt{planned\_analysis()}, \\ \texttt{simulate\_data()}, and
 \texttt{report\_analysis()} and embed these into a manuscript that
@@
 eventually becomes the final report. The complete preregistration source
-is available in the {online supplementary material}
+is available in the {online Supplementary Material}
 ~(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd} accessed on 11 October 2021). In this section, we show code excerpts of this
@@
   can use the {provided code for the simulation}
-~\mbox{(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R})} to verify that the \emph{t}-test provides unbiased
+~{(\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R} accessed on 4 December 2021)} to verify that the \emph{t}-test provides unbiased
   effect sizes but the Mann-Whitney-Wilcoxon overestimates effect sizes
@@
  The code
-for this power analysis can be found in the {online supplementary material}
+for this power analysis can be found in the {online Supplementary Material}
 (\url{https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R} accessed on 11 October 2021). The next snippet shows how we integrated the
@@
 Readers can inspect and judge the changes for themselves on {GitHub}
-\textls[-25]{(\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main\#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b})}.
+\textls[-25]{(\url{https://github.com/aaronpeikert/repro-tutorial/compare/v0.0.1.1-prereg..main\#diff-e21a8fa2e44b297dfefef329a6ef56d283488d467c4b4ffe2a014111e52a170b} accessed on 4 December 2021)}.
 
@@
 The example rendered as a {PDF file with real data}
-(\url{https://github.com/aaronpeikert/repro-tutorial/files/7309455/preregistration.pdf} accessed on 11 October 2021) is available in the {online supplementary material}
+(\url{https://github.com/aaronpeikert/repro-tutorial/files/7309455/preregistration.pdf} accessed on 11 October 2021) is available in the {online Supplementary Material}
 (\url{https://github.com/aaronpeikert/repro-tutorial/releases/tag/v0.0.3.1-results} accessed on 11 October 2021). The changes we made since preregistering it can
@@
 research projects promise a wide range of applications, among them PAC
-\citep[potentially to be submitted as a registered
-report][]{nosekRegisteredReports2014, chambersWhatNextRegistered2019},
+(\citep{nosekRegisteredReports2014, chambersWhatNextRegistered2019} potentially to be submitted as a registered
+report),
 direct replication~\citep{simonsValueDirectReplication2014}, fully
@@
 users, we will implement some of them in \texttt{repro} and hope for
-broad adoption of computational reproducibility in the near future.
+broad adoption of computational reproducibility in the near~future.
 
@@
 
-Supplementary Materials {All materials (i.e., the source code, all figures, and the data) that are neccesary for reproducing the submitted version of this article are availible at \url{https://github.com/aaronpeikert/repro-tutorial} and archived under \url{https://doi.org/10.5281/zenodo.5724454}}.
+\supplementary{All materials (i.e., the source code, all figures, and the data) that are neccesary for reproducing the submitted version of this article are availible at \url{https://github.com/aaronpeikert/repro-tutorial} and archived under \url{https://doi.org/10.5281/zenodo.5724454} (accessed on 4 December 2021).}
 
@@
 
-\funding{This research received no external funding}
+\funding{This research received no external funding.}
 
@@
 \bibitem[Knuth and Levy(1994)]{knuthCWEBSystemStructured}
-Knuth, D.E.; Levy, S. \emph{The CWEB System of Structured Documentation}; Addison-Wesley Longman: \hl{Boston, USA, 1994.}
+Knuth, D.E.; Levy, S. \emph{The CWEB System of Structured Documentation}; Addison-Wesley Longman: {Boston, MA, USA, 1994.}
 
@@
   Hesselberth, and Dervieux]{R-rticles}
-Allaire, J.; Xie, Y.; R Foundation; Wickham, H.; Journal of Statistical Software.; Vaidyanathan, R.; Association for Computing Machinery; Boettiger, C.; Elsevier; Broman, K.; et al.
-\emph{Articles: Article Formats for R Markdown}; R Package Version 0.19; \hl{2021.}%MDPI: Please add the publisher and its location.
+Allaire, J.; Xie, Y.; R Foundation; Wickham, H.; Journal of Statistical Software; Vaidyanathan, R.; Association for Computing Machinery; Boettiger, C.; Elsevier; Broman, K.; et al.
+\emph{Articles: Article Formats for R Markdown}; R Package Version 0.19; {2021.}  Available online: \url{https://pkgs.rstudio.com/rticles/} (accessed on 4 December 2021).
+% access date: 04-12-21%MDPI: Please add the publisher and its location.
 
@@
 \bibitem[El~Hattab and Allaire(2017)]{revealjs}
-El Hattab, H.; Allaire, J. Revealjs: R Markdown Format for ’Reveal, Js’ Presentations, \hl{2017.}%MDPI: Please provide an available website and its accessed date.
+El Hattab, H.; Allaire, J. Revealjs: R Markdown Format for ’Reveal, Js’ Presentations. {2017.} Available online: \url{https://bookdown.org/yihui/rmarkdown/revealjs.html} (accessed on 4 December 2021).
+% access date: 04-12-21 %MDPI: Please provide an available website and its accessed date.
 
@@
 \bibitem[{O'Hara-Wild} and Hyndman(2021)]{vitae}
-O’Hara-Wild, M.; Hyndman, R. Vitae: Curriculum Vitae for r Markdown, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+O’Hara-Wild, M.; Hyndman, R.
+Vitae: Curriculum Vitae for r Markdown.
+{2021.} Available online: \url{https://cran.r-project.org/web/packages/vitae/vignettes/vitae.html} (accessed on 4 December 2021).%MDPI: Please provide an available website and its accessed date.
 
@@
 \bibitem[Ushey(2021)]{R-renv}
-Ushey, K. Renv: Project Environments, 2021. \hl{R Package Version 0.13.2.}%MDPI: Please provide an available website and its accessed date.
+Ushey, K.
+Renv: Project Environments, 2021.
+{R Package Version 0.13.2.} Available online: \url{https://rstudio.github.io/renv/articles/renv.html} (accessed on 4 December 2021).%MDPI: Please provide an available website and its accessed date.
 
@@
 \bibitem[Wickham and Bryan(2021)]{usethis}
-Wickham, H.; Bryan, J. Usethis: Automate Package and Project Setup, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+Wickham, H.; Bryan, J.
+Usethis: Automate Package and Project Setup.
+{2021.} Available online: \url{https://usethis.r-lib.org} (accessed on 4 December 2021).%MDPI: Please provide an available website and its accessed date.
 
@@
 Repro: Automated Setup of Reproducible Workflows and Their Dependencies, 2021.
-\hl{R Package Version 0.1.0.}%MDPI: Please provide an available website and its accessed date.
+{R Package Version 0.1.0.} Available online: \url{https://github.com/aaronpeikert/repro} (accessed on 4 December 2021).%MDPI: Please provide an available website and its accessed date.
 
@@
 \bibitem[Aust and Barth(2020)]{papaja}
-Aust, F.; Barth, M. Papaja: Create APA Manuscripts with R Markdown, \hl{2020.}%MDPI: Please provide an available website and its accessed date.
+Aust, F.; Barth, M.
+Papaja: Create APA Manuscripts with R Markdown.
+{2020.}  Available online: \url{http://frederikaust.com/papaja_man/} (accessed on 4 December 2021).%MDPI: Please provide an available website and its accessed date.
 
@@
   Nosek, and Soderberg]{bowmanOSFPreregTemplate2020}
-Bowman, S.; DeHaven, A.; Errington, T.; Hardwicke, T.E.; Mellor, D.T.; Nosek, B.A.; Soderberg, C.K. OSF Prereg Template 2020. \emph{MetaArXiv}, \textbf{2020}, https://doi.org/10.31222/osf.io/epgjd.
+Bowman, S.; DeHaven, A.; Errington, T.; Hardwicke, T.E.; Mellor, D.T.; Nosek, B.A.; Soderberg, C.K. OSF Prereg Template 2020. \emph{MetaArXiv} \textbf{2020}, https://doi.org/10.31222/osf.io/epgjd.
 
@@
 Machine-Learning Approaches to Structural Equation Modeling.
-In \emph{Handbook of Structural Equation Modeling}, 2nd ed.; Hoyle, R.H., Ed.; \hl{Guilford Press, New York, NY, USA}, in press.
+In \emph{Handbook of Structural Equation Modeling}, 2nd ed.; Hoyle, R.H., Ed.; {Guilford Press: New York, NY, USA}, in press.
 
@@
 Hastie, T.; Tibshirani, R.; Friedman, J.H.
-\emph{The Elements of Statistical Learning: Data Mining, Inference, and Prediction, Second Edition, Corrected at 12th Printing 2017 ed.}; Springer Series in Statisticsl Springer: New York, NY, USA, 2017.
+\emph{The Elements of Statistical Learning: Data Mining, Inference, and Prediction, Second Edition, Corrected at 12th Printing 2017 ed.}; Springer Series in Statisticsl;  Springer: New York, NY, USA, 2017.
 
@@
 \bibitem[Dar{\'o}czi and Tsegelskyi(2021)]{pander}
-Daróczi, G.; Tsegelskyi, R. Pander: An R ‘pandoc’ Writer, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+Daróczi, G.; Tsegelskyi, R.
+Pander: An R ‘pandoc’ Writer.
+{2021.} Available online: \url{https://www.r-project.org/nosvn/pandoc/pander.html} (accessed on 4 December 2021).%MDPI: Please provide an available website and its accessed date.
 
@@
 \bibitem[Stanley(2021)]{apaTables}
-Stanley, D. apaTables: Create American Psychological Association (APA) Style Tables, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+Stanley, D.
+apaTables: Create American Psychological Association (APA) Style Tables, {2021.} Available online: \url{https://dstanley4.github.io/apaTables/articles/apaTables.html} (accessed on 4 December 2021).
+%MDPI: Please provide an available website and its accessed date.
 
@@
 Automated Results Reporting as a Practical Tool to Improve Reproducibility and Methodological Best Practices Adoption.
-\hl{CRAN 2021.}%MDPI: Please provide an available website and its accessed date.
+{CRAN 2021.} Available online: \url{https://easystats.github.io/report/} (accessed on 4 December 2021).
+%MDPI: Please provide an available website and its accessed date.
 
@@
   {OpenAIRE}(2013)]{zenodo}
-European Organization For Nuclear Research. \emph{OpenAIRE}; \hl{Zenodo}, 2013. https://doi.org/10.25495/7GXK-RD71.%MDPI: Please add the location of the publisher.
+European Organization For Nuclear Research. \emph{OpenAIRE}; {Zenodo}: online, 2013. https://doi.org/10.25495/7GXK-RD71.%MDPI: Please add the location of the publisher.
 
@@
 International Council for Harmonisation of Technical Requirements for Registration of Pharmaceuticals for Human Use.
-E 9 Statistical Principles for Clinical Trials, \hl{1998.}%MDPI: Please provide an available website and its accessed date.
+E 9 Statistical Principles for Clinical Trials, {1998.} Available online: \url{https://www.ema.europa.eu/en/ich-e9-statistical-principles-clinical-trials} (accessed on 4 December 2021).
+%MDPI: Please provide an available website and its accessed date.
 
@@
   Publications}(2020)]{elifesciencespublicationsELifeLaunchesExecutable2020}
-eLife Sciences Publications. eLife Launches Executable Research Articles for Publishing Computationally Reproducible Results \hl{2020.}%MDPI: Please provide an available website and its accessed date.
+eLife Sciences Publications.
+eLife Launches Executable Research Articles for Publishing Computationally Reproducible Results.
+{2020.} Available online: \url{https://elifesciences.org/for-the-press/eb096af1/elife-launches-executable-research-articles-for-publishing-computationally-reproducible-results}  (accessed on 4 December 2021).
+%MDPI: Please provide an available website and its accessed date.
 
@@
 \bibitem[Gohel(2021)]{officer}
-Gohel, D. Officer: Manipulation of Microsoft Word and PowerPoint Documents, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+Gohel, D.
+Officer: Manipulation of Microsoft Word and PowerPoint Documents.
+{2021.} Available online: \url{https://davidgohel.github.io/officer/} (accessed on 4 December 2021).
+%MDPI: Please provide an available website and its accessed date.
 
@@
 \bibitem[Gohel and Ross(2021)]{officedown}
-Gohel, D.; Ross, N. Officedown: Enhanced ‘R Markdown’ Format for ‘Word’ and ‘PowerPoint’, \hl{2021.}%MDPI: Please provide an available website and its accessed date.
+Gohel, D.; Ross, N.
+Officedown: Enhanced ‘R Markdown’ Format for ‘Word’ and ‘PowerPoint’.
+{2021.} Available online: \url{https://davidgohel.github.io/officedown/} (accessed on 4 December 2021).
+%MDPI: Please provide an available website and its accessed date.
 \end{thebibliography}
